### 2020-03-09 Scripts/Initial_Analysis.Rmd
Initial analysis development

---
title: Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for
  Policy Response Under Climate Change
author: "Juliano Palacios Abrantes"
date: "09/03/2020"
output:
  word_document: default
  html_document: default
subtitle: Initital analysis
editor_options: 
  chunk_output_type: console
---

```{r Ch3_setup, eval = T, echo=F, warning=F,message=F, results='hide'}

#### READ ME !!! ####
# Run this chunk before knit so you make sure you have all pkgs installed in R

ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE,repos = "http://cran.us.r-project.org")
  sapply(pkg, require, character.only = TRUE)
}

#### Library ####
packages <- c(
  "tidyverse",
  "here", # for dbem_import `here()`
  "data.table", #dbem_import `fread()`
  "readxl", # for reading excell files
  "janitor", # for clearing names
  "geosphere", # estimate distances between points `distm()`
  "ggrepel",
  "zoo"
)

ipak(packages)

# Set paths depending on machine Beast (jepa88), "carmelia" or Hall1000
if(Sys.info()[7] == "jepa88"){
  data_path <- "Z:/DATA/DBEM/"
}
if(Sys.info()[7] == "carmelia"){
  data_path <- paste(here(),"/Temporal_Data/",sep = "")
}
if(Sys.info()[7] == "hall1000"){
  data_path <- "/Volumes/DATA/DBEM/"
}

```

# Methods

### Fun `GetSppDist`

This function reads the DBEM data for each year and ensamble (ESM model) from 102 20 111

```{r GetSppDist, eval = T, echo = T}

#### The function
dbem_import <- function(taxon_key,
                        iyr,
                        fyr,
                        ensamble_list,
                        data_type,
                        Path = NA){

  # Creat a sequence of the years to import  
  yr <- seq(iyr,
            fyr,
            by=1) #<- change this setting if you have bi/tri/etc annual data

  #### Set the Path####
  #First, set the path of the files under a loop for each year

  for(c in 1:length(ensamble_list)){

    D_Path <- paste(Path,"/mpa0F1ENS",ensamble_list[c],"/", #<- drobo path for 8.5
                    taxon_key,"/",taxon_key,data_type,yr,".txt",
                    sep="")

    total <- NULL

    #### Importing data ####
    #Now we import the data using data.table::fread and the path generated above

    for (i in 1:length(yr)){ #<-The years you want to have

      possibleError <-  tryCatch({ # <- This will solve for situations where the species is not found in the DBEM
        # Imports dataset 
        cur <- fread(D_Path[i],
                     na.strings = "NA",
                     col.names = c("index",
                                   yr[i]) #<- Calls colunmns by year
        )

      },
      error=function(e){ # Part of the "no species in DBEM
        e
      }
      )
      if(!inherits(possibleError, "error")){

        # possibleError_I <-  tryCatch({
        #Puts them together in one nice dataset 
        if(i == 1){
          total<- copy(cur) # <- copies the previouse data.table
        }else{ 
          setkey(total, # <- sets the data.table as "reference" ?setkey
                 index); setkey(cur,
                                index); 
          total <- merge(total, #<- Merges all the data in one single file.
                         cur,
                         all = TRUE) # <- This allows for the NA's

        } #Close datatable loop

      }

    } #First loop END (Path Creation)

    # Tidy version
    total <- total %>%
      gather(year,value,-index) %>%
      mutate(taxon_key = taxon_key,
             ensamble=ensamble_list[c])

    world_data <- world_grid %>%
      full_join(total,
                by = "index")
    # Messages
    if(inherits(possibleError, "error")){

      print(paste("Oops! --> No information for",taxon_key,yr[i], sep = " ")) # <- Display error message

    }else{

      print(paste("Analysis done for",taxon_key, "ensamble" ,ensamble_list[c]))
    }

    if(c == 1){
      partial_ensamble <- world_data
    }else{
      spp_ensamble <- bind_rows(partial_ensamble,world_data)
    }

  }
  return(spp_ensamble)

} #Function end

# Test function (works in Carmelia!)

# dbem_import(
#   taxon_key <- 600004,
#   iyr <- 1951,
#   fyr <- 1952,
#   data_type <- "Abd",
#   ensamble_list = c(102,103),
#   Path <- paste(data_path,"DBEM/",sep="")
# )

# unique(x$ensamble)
# head(x)

# Test multiple ensambles and spp

# ensamble_list <- c(102,103)
# taxon_key_list <- c(600004,600005)

# x <- bind_rows(
#   lapply(
#     # taxon_key_list,
#     ensamble_list,
#     FUN= dbem_import,
#     taxon_key = 600004,
#     # ensamble = 102,
#     iyr = 1951,
#     fyr = 1952,
#     data_type = "Abd",
#     Path = paste(data_path,"DBEM/",sep="")
#   )
# )

# OK!

```

### Fun `EstRangeIndex`

```{r load_data_per_ensamble, eval = T, echo = T}

#### Data needed (send to Control pannel latter)

# All ensambles
ensamble_list <- c(102,103)

# for dbem_impo_ensamble So we don't miss INDEX in different ensambles
world_grid <- tibble(index = seq(1,259200,1))
taxon_key <- 600005 #601477

# SAU relations between INDEX and Country's EEZs
EEZIDs_List <- read_excel(paste(data_path,"Spatial/V_Lam/Updated_EEZList_17June2016.xlsx",sep="")) %>% 
  clean_names()
EEZ_CellID <- read_excel(paste(data_path,"Spatial/V_Lam/EEZ_CellID.xlsx",sep="")) %>% 
  clean_names()
colnames(EEZ_CellID) <- c("eezid","index")

Lon_Lat_DBEM <- read.csv("~/GitHub/TransEmergence/Temporal_Data/Spatial/Lon_Lat_DBEM.txt", header=FALSE)
colnames(Lon_Lat_DBEM) <- c("index","lon","lat")

index_code <- EEZIDs_List %>% 
  left_join(EEZ_CellID) %>% 
  rename(eez_name = name) %>% 
  left_join(Lon_Lat_DBEM,
            by = "index")

EstRangeIndex <- function(taxon_key,iyr,fyr,ensamble_list){

  spp_data <- dbem_import(
    taxon_key = taxon_key,
    iyr <- 1951,
    fyr <- 2100,
    data_type <- "Abd",
    ensamble_list = ensamble_list,
    Path <- paste(data_path,"DBEM/",sep="")
  ) %>% 
    # Remove places where species is not destributed and set to cero differences between ensambles
    filter(!is.na(value)) %>% 
    spread(ensamble,value) %>% 
    gather(ensamble,value, -1:-3) %>% 
    mutate(value = ifelse(is.na(value),0,value)) %>% 
  # Include eez information
    left_join(index_code,
              by = "index") %>% 
    filter(!is.na(eezid))

  # Estimate mean eez centroid
  # centroid_data_eez <-  spp_data %>% 
  #   ungroup() %>% 
  #   group_by(taxon_key,ensamble,year,eezid,eez_name) %>% 
  #   top_n(1,value) %>% 
  #   group_by(year,taxon_key,ensamble,eezid,eez_name) %>% 
  #   summarise_at(vars(lon,lat),mean) %>%  # solves for multiple top 1
  #   rename(eez_centroid_lon = lon,
  #          eez_centroid_lat = lat) %>% 
  #   filter(!is.na(eezid))

  centroid_data_eez <-  spp_data %>% 
    ungroup() %>% 
    group_by(taxon_key,eezid,eez_name) %>%
    summarise_at(vars(lon,lat),mean) %>%  # solves for multiple top 1
    rename(eez_centroid_lon = lon,
           eez_centroid_lat = lat)

  # Visualization of  eez centroids
  # ggplot(spp_data) +
  #   geom_tile(
  #     aes(
  #     x = lon,
  #     y = lat,
  #     fill= eez_name
  #     )
  #   ) +
  #   geom_point(data = centroid_data_eez,
  #              aes(
  #                x = eez_centroid_lon,
  #                y = eez_centroid_lat,
  #                size = 1
  #              )
  #   )

# Estimate species distribution centroid
  centroid_spp <-  spp_data %>% 
    ungroup() %>% 
    group_by(taxon_key,ensamble,year) %>% 
    top_n(1,value) %>% 
    group_by(taxon_key,ensamble,year) %>% 
    summarise_at(vars(lon,lat),mean) %>%  # solves for multiple top 1
    rename(spp_centroid_lon = lon,
           spp_centroid_lat = lat)

  ggplot(centroid_spp) +
    geom_point(
      aes(
        x = spp_centroid_lon,
        y = spp_centroid_lat,
        size = as.numeric(year)#,
        # shape = eez_name
      )
    ) +
    geom_point(data = centroid_data_eez,
               aes(
                 x = eez_centroid_lon,
                 y = eez_centroid_lat,
                 color = eez_name
               )
    ) +
    facet_wrap(~ensamble)

# Estimate Transboundary Index
  trans_index_data <- centroid_data_eez %>% 
    left_join(centroid_spp,
              by = c("taxon_key") # removed year
              ) %>% 
    ungroup() %>% 
    # select(year,eez_name,ensamble,spp_centroid_lon,spp_centroid_lat,eez_centroid_lon, eez_centroid_lat) %>%
    rowwise() %>%
    mutate(d_eez = distm(c(eez_centroid_lon, eez_centroid_lat),c(spp_centroid_lon, spp_centroid_lat), fun = distHaversine))

trans_index_sd_data <- trans_index_data %>% 
  ungroup() %>% 
  filter(year <2000) %>% 
  group_by(eez_name,ensamble) %>% 
  summarise(
    s_sd = sd(d_eez)
  )

final_trans_index_data <- trans_index_data %>% 
  left_join(trans_index_sd_data,
            by = c("eez_name","ensamble")
            ) %>% 
  mutate(da_dsd = d_eez/s_sd) %>% 
  filter(eez_name != "Colombia (Pacific)") %>% 
  select(year,eez_name,ensamble,da_dsd) %>% 
  spread(eez_name,da_dsd) %>% 
  mutate(Trans_Index = `Mexico (Atlantic)`-`USA (Gulf of Mexico)`/2) %>% 
  ungroup() %>% 
  group_by(ensamble) %>% # Running mean as Kleisner & Pauly 2012
  mutate(RMean = rollmean(x = Trans_Index,
                            5,
                            align = "right",
                            fill = Trans_Index)
    )

ggplot(final_trans_index_data) +
  geom_line(
    aes(
      x = as.numeric(year),
      y = RMean,
      color = ensamble
    )
  ) +
  geom_line(data = final_trans_index_data,
    aes(
      x = as.numeric(year),
      y = Trans_Index,
      color = ensamble
    ),
    linetype = "dashed",
    alpha = 0.5
  ) +
  facet_wrap(~ensamble,
             ncol= 1)

  # %>% 
  #   filter(ensamble == 102) %>% 
  #   select(year,ensamble,eez_name,d_eez) %>% 
  #   spread(eez_name,d_eez) %>% 
  #   mutate()

  Year_MCP <-  spp_data %>% 
    # Aggregate MCP per EEZ
    group_by(taxon_key,eezid,eez_name,ensamble,year) %>% 
    summarise(eez_mcp = sum(value,na.rm=T)) %>% 
    # Ensamble mean
    group_by(taxon_key,eezid,eez_name,year) %>% 
    summarise(ensamble_mean = mean(eez_mcp,na.rm=T),
              ensamble_sd = sd(eez_mcp,na.rm=T)
              )

  # Test 
  # ggplot(spp_data) +
  #   geom_line(
  #     aes(
  #       x = as.numeric(year),
  #       y = ensamble_mean
  #     )
  #   ) +
  #   geom_point(
  #     aes(
  #       x = as.numeric(year),
  #       y = ensamble_mean+ensamble_sd
  #     )
  #   ) +
  #   geom_point(
  #     aes(
  #       x = as.numeric(year),
  #       y = ensamble_mean-ensamble_sd
  #     )
  #   ) +
  #   facet_wrap(~eez_name)

  } # close function

```

### 2020-03-09 Scripts/Manuscript.Rmd
Initial analysis development

---
title: "Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change"
subtitle: "[MANUSCRIPT DRAFT DO NOT SHARE WITHOUT PERMISSION]"
author: "Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, U. Rashid Sumaila^1^,  Vicky Lam^1^, David VanderZwaag^3^, and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern. ^3^Schulich School of Law, Dalhousie University"
output: pdf_document
---

* corresponding author: j.palacios@oceans.ubc.ca

# Abstract

# Introduction

# Methods

# Results

# Discussion

### 2020-03-11 Scripts/Initial_Analysis.Rmd
Future proportion change by mid and end of the century

@@
     }
+  }
 
-  }
   return(spp_ensamble)
@@
 # OK!
+
+```
+
+## Proportion approach
+
+### Fun `EstPropChange`
+
+```{r EstPropChange}
+
+EstPropChange <- function(Spp,iyr,fyr,data_type,ensamble_list,Path,Neighbours){
+
+  # Get model data from spps
+  SppDist <- dbem_import(
+    taxon_key = Spp,
+    iyr <- iyr,
+    fyr <- fyr,
+    data_type <- "Abd",
+    ensamble_list = ensamble_list,
+    Path <- paste(data_path,"DBEM/",sep="")
+  )
+
+  # Result 1. Number of Countries that share the species
+
+  #____________ Selecting only the transboundary nature of the species _________ #
+  Trans_Spp <- SppDist %>%
+    left_join(Neighbours,
+              by= "index") %>% 
+    filter(!is.na(value)) %>% 
+    semi_join(Transboundary_spp, # filter only transboundary cases
+              by = c("taxon_key","eez_name","eez_neighbour")
+    )
+
+  #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
+  # The number of species' cells present within each country's EEZ
+
+  #Step 1.  Get EEZ id and Neighbour
+  Neighbours_List <- Neighbours %>% 
+    group_by(eez_name,eez_neighbour) %>% 
+    summarise(n=n()) %>% 
+    ungroup() %>% 
+    select(-n)
+
+  # Step 2. Determines the amount of grids present in each country
+  Spp_Grid <- Trans_Spp %>% 
+    group_by(taxon_key,
+             eez_name,
+             year,
+             ensamble) %>% 
+    summarise(n_spp_eez = length(unique(index))) %>% 
+    left_join(Neighbours_List,
+              by = "eez_name") %>% 
+    filter(eez_name %in% Trans_Spp$eez_name, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
+           eez_neighbour %in% Trans_Spp$eez_name)
+
+  # Step 3. Sum total grids per Neighbours
+
+  # Split dataframes to merge latter
+  Territory_T <- Spp_Grid %>% 
+    ungroup() %>% 
+    select(
+      taxon_key,
+      Name=eez_name,
+      n_spp_eez,
+      year,
+      ensamble
+    )
+
+  Neighbour_T <- Spp_Grid %>% 
+    ungroup() %>% 
+    select(
+      taxon_key,
+      n_spp_eez,
+      Name=eez_neighbour,
+      eez_name,
+      year,
+      ensamble
+    )
+
+  per_change_d <- full_join(Territory_T,
+                            Neighbour_T, 
+                            by = c("Name","taxon_key","year","ensamble")
+  ) %>%
+    rowwise() %>%
+    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
+    distinct() %>% # Removes false duplicates from `full_join()`
+    rename(eez_name = Name,
+           eez_neighbour =eez_name,
+           n_spp_Country = n_spp_eez.x,
+           n_spp_Neighbour = n_spp_eez.y) %>% 
+    mutate(area_index = n_spp_Country/Spp_Total) %>% 
+    select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
+    mutate(
+      time_step = ifelse(year < 2000,"historical",
+                         ifelse(year >= 2045 & year <= 2064, "mid",
+                                ifelse(year >= 2080, "n_end","NA")
+                         )
+      )
+    ) %>%
+    filter(time_step != "NA") %>% 
+    # Time mean and sd
+    ungroup() %>% 
+    group_by(taxon_key,ensamble,eez_name,eez_neighbour,time_step) %>% 
+    summarise_at(vars("area_index"),
+                 c(temporal_mean=mean,
+                   temporal_sd=sd)
+    ) %>% 
+    # Ensamble mean and sd
+    group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% 
+    summarise_at(vars("temporal_mean"),
+                 c(ensamble_mean=mean,
+                   ensamble_sd=sd)
+    ) %>% 
+    mutate(ensamble_mean = ifelse(is.na(ensamble_mean),0,ensamble_mean)) %>% 
+    select(-ensamble_sd) %>% 
+    spread(time_step,ensamble_mean) %>% 
+    mutate_at(vars(mid,n_end),
+              funs(ifelse(historical == 0 & . == 0, 0, # If for no change than 0
+                          ifelse(historical == 0 & . > 0, 100, # If changes from 0 to anything then 100
+                                 ((.-historical)/historical)*100 # Estimate percentage
+                                 # ((.-historical)/((.historical)/2))*100 # Estimate percentage
+                          )
+              )
+              )
+    )
+
+  #### Statistical test ###
+
+  stat_change_d <- full_join(Territory_T,
+                             Neighbour_T, 
+                             by = c("Name","taxon_key","year","ensamble")
+  ) %>%
+    rowwise() %>%
+    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
+    distinct() %>% # Removes false duplicates from `full_join()`
+    rename(eez_name = Name,
+           eez_neighbour =eez_name,
+           n_spp_Country = n_spp_eez.x,
+           n_spp_Neighbour = n_spp_eez.y) %>% 
+    mutate(area_index = n_spp_Country/Spp_Total) %>% 
+    select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
+    mutate(
+      time_step = ifelse(year < 2000,"historical",
+                         ifelse(year >= 2045 & year <= 2064, "mid",
+                                ifelse(year >= 2080, "n_end","NA")
+                         )
+      )
+    )
+
+  return(per_change_d)
+
+}
+
+```
+
+### Fun `EstTransIndex`
+
+```{r funEstTransIndex, eval= T ,echo = T,warning = F,message = F}
+
+# Function to determine whether or not a species is transboundary
+
+# Varibales needed
+# Spp: Species to be analized, data with presence/absence per gridcell
+# Index_EEZ: Reference list of all INDEX that fall within EEZs
+# EEZ_Size: Number of grid-cells that each EEZ has
+# Neighbours: Reference list of neighbouring countries
+
+# Needs getSppDist
+
+#### For testing 
+Spp <- 600004  # Anchovie Equator, Chile and Peru
+# Spp <- 600107 # Tuna Many RFMOs
+# Spp <- 600004 # Argument 1 must have names (RFMO)
+# Spp <- 600464 # Discrete species 
+# Spp <- 600051 # One of Pierrs massive (wrong) discrete spp
+# x <- c(600464, # Discreat spp in one nation
+# 600059, # nowhere to be found spp
+# 601352, # Discrete spp in two nations
+# 600004) # transboundary spp
+# Model <- "All"
+# Coord <- CoorG
+# 
+# # For EEZ
+# Neighbours <- Neighbours_Data
+# Index_Code <- Index_Code
+
+# __________________________________ #
+
+# The function
+EstTransIndex <- function(Spp,iyr,fyr,data_type,ensamble_list,Path,Neighbours){
+
+  # Get model data from spps
+  SppDist <- dbem_import(
+    taxon_key = Spp,
+    iyr <- iyr,
+    fyr <- fyr,
+    data_type <- "Abd",
+    ensamble_list = ensamble_list,
+    Path <- paste(data_path,"DBEM/",sep="")
+  )
+
+  # Result 1. Number of Countries that share the species
+
+  #____________ Selecting only the transboundary nature of the species _________ #
+  Trans_Spp <- SppDist %>%
+    left_join(Neighbours,
+              by= "index") %>% 
+    filter(!is.na(value)) %>% 
+    semi_join(Transboundary_spp, # filter only transboundary cases
+              by = c("taxon_key","eez_name","eez_neighbour")
+    )
+
+  #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
+  # The number of species' cells present within each country's EEZ
+
+  #Step 1.  Get EEZ id and Neighbour
+  Neighbours_List <- Neighbours %>% 
+    group_by(eez_name,eez_neighbour) %>% 
+    summarise(n=n()) %>% 
+    ungroup() %>% 
+    select(-n)
+
+  # Step 2. Determines the amount of grids present in each country
+  Spp_Grid <- Trans_Spp %>% 
+    group_by(taxon_key,
+             eez_name,
+             year,
+             ensamble) %>% 
+    summarise(n_spp_eez = length(unique(index))) %>% 
+    left_join(Neighbours_List,
+              by = "eez_name") %>% 
+    filter(eez_name %in% Trans_Spp$eez_name, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
+           eez_neighbour %in% Trans_Spp$eez_name)
+
+  # Step 3. Sum total grids per Neighbours
+
+      # Split dataframes to merge latter
+    Territory_T <- Spp_Grid %>% 
+      ungroup() %>% 
+      select(
+        taxon_key,
+        Name=eez_name,
+        n_spp_eez,
+        year,
+        ensamble
+      )
+
+    Neighbour_T <- Spp_Grid %>% 
+      ungroup() %>% 
+      select(
+        taxon_key,
+        n_spp_eez,
+        Name=eez_neighbour,
+        eez_name,
+        year,
+        ensamble
+      )
+
+    # Merge dataframes to get totals per Neighbourds
+    area_index_d <- full_join(Territory_T,
+                              Neighbour_T, 
+                              by = c("Name","taxon_key","year","ensamble")
+    ) %>%
+      rowwise() %>%
+      mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
+      distinct() %>% # Removes false duplicates from `full_join()`
+      rename(eez_name = Name,
+             eez_neighbour =eez_name,
+             n_spp_Country = n_spp_eez.x,
+             n_spp_Neighbour = n_spp_eez.y) %>% 
+      mutate(area_index = n_spp_Country/Spp_Total) %>% 
+      select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
+      mutate(
+        time_step = ifelse(year < 2000,"historical",
+                           ifelse(year >= 2045 & year <= 2064, "mid",
+                                  ifelse(year >= 2080, "n_end","NA")
+                           )
+        )
+      ) %>%
+      # Ensamble mean and sd
+      ungroup() %>% 
+      group_by(taxon_key,year,eez_name,eez_neighbour,time_step) %>% 
+      summarise_at(vars("area_index"),
+                   c(ensamble_mean=mean,
+                     ensamble_sd=sd)
+                   ) %>% 
+      group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% #Running mean
+      mutate(RMean = rollmean(x = ensamble_mean,
+                            5,
+                            align = "right",
+                            fill = ensamble_mean)
+    )
+
+    # Determine if changes
+    natural_variation_d <- area_index_d %>% 
+      ungroup() %>% #remove rowise
+      filter(time_step == "historical") %>% 
+      # Temporal mean
+      group_by(taxon_key,eez_name,eez_neighbour) %>% 
+      summarise_at(vars("ensamble_mean"),
+                   c(hist_mean_index=mean,
+                     hist_sd_index=sd)
+      )
+
+    ### Variation treshold
+    variation_d <- area_index_d %>% 
+      ungroup() %>% #remove rowise
+      # Temporal mean
+      group_by(taxon_key,time_step,eez_name,eez_neighbour) %>% 
+      summarise_at(vars("RMean"),
+                   c(mean_index=mean,
+                     sd_index=sd)
+      ) %>% 
+      gather("variable","value",mean_index,sd_index) %>% 
+      spread(time_step,value) %>% 
+      mutate(larger_variation = ifelse(future > historical,"yes","no"))
+
+    # William's index methodology
+    trans_index_data <- area_index_d %>% 
+      left_join(natural_variation_d,
+                by = c("taxon_key","eez_name","eez_neighbour")) %>% 
+      mutate(d_sd = ensamble_mean/sd_index) %>% 
+      gather("variable","value",ensamble_mean:d_sd) %>% 
+      ungroup() %>% 
+      select(-time_step,-eez_neighbour) %>% 
+      rowid_to_column() %>% 
+      # spread(time_step,value) %>% 
+      spread(eez_name,value) %>% 
+      group_by(taxon_key,year,variable) %>% 
+      summarise_at(vars("Chile","Ecuador","Peru"),
+                   mean,na.rm=T)
+
+    ggplot(area_index_d) +
+      geom_ribbon(data=area_index_d,
+                  aes(
+                    x = as.numeric(year),
+                    ymin=RMean-ensamble_sd,
+                    ymax=RMean+ensamble_sd
+                  ),
+                  fill = "red",
+                  alpha = 0.5
+                  )+
+      geom_line(
+        aes(
+          x = as.numeric(year),
+          y = RMean
+        )
+      )  +
+      facet_wrap(~eez_name + eez_neighbour,
+                 scales = "free")
+
+} # closes function
+
+# __________________________________
+# Test function
+# __________________________________
 
@@
 ```
+
+## Control Panel
+
+```{r control_panel}
+
+Spp_list <- c(600004,600005)
+iyr <- 1951
+fyr <- 2099
+data_type <- "Catch"
+ensamble_list = c(102,103)
+Path <- paste(data_path,"DBEM/",sep="")
+world_grid <- tibble(index = seq(1,259200,1))
+Neighbours <- read_excel("Temporal_Data/Spatial/EEZ_Neighbour_List.xlsx") %>% 
+  clean_names()
+Transboundary_spp <- read.csv("~/GitHub/TransEmergence/Temporal_Data/Species/Transboundary_spp.csv") %>% clean_names()
+
+```
+
+## Main Routine
+
+```{r main_routine, eval = T, echo = F}
+
+####_______________________ ####
+# Estimating proportion change
+####_______________________ ####
+
+suppressWarnings(
+  r_proportion_change <- bind_rows(
+    lapply(
+      Spp_list, #c(600004,600005),
+      EstPropChange,
+      iyr <- iyr,
+      fyr <- fyr,
+      data_type <- "Abd",
+      ensamble_list = ensamble_list,
+      Path <- paste(data_path,"DBEM/",sep=""),
+      Neighbours = Neighbours
+    )
+  )
+)
+```

### 2020-03-13 Scripts/Initial_Analysis.Rmd
control panel updates

@@
 
-```{r EstPropChange}
+```{r EstPropChange, eval = T, echo = F}
 
@@
 
-```{r funEstTransIndex, eval= T ,echo = T,warning = F,message = F}
+```{r funEstTransIndex, eval= F ,echo = T,warning = F,message = F}
 
@@
 
-```{r load_data_per_ensamble, eval = T, echo = T}
+```{r load_data_per_ensamble, eval = F, echo = T}
 
@@
 
-Spp_list <- c(600004,600005)
+Spp_list <- read.csv(paste(data_path,"Species/Species_list.csv",sep="")) %>% clean_names()
 iyr <- 1951
@@
 data_type <- "Catch"
-ensamble_list = c(102,103)
+ensamble_list = seq(102,111,1)
 Path <- paste(data_path,"DBEM/",sep="")
 world_grid <- tibble(index = seq(1,259200,1))
-Neighbours <- read_excel("Temporal_Data/Spatial/EEZ_Neighbour_List.xlsx") %>% 
+Neighbours <- read_excel(paste(data_path,"Spatial/EEZ_Neighbour_List.xlsx",sep="")) %>% 
   clean_names()
-Transboundary_spp <- read.csv("~/GitHub/TransEmergence/Temporal_Data/Species/Transboundary_spp.csv") %>% clean_names()
+
+Transboundary_spp <- read.csv(paste(data_path,"Species/Transboundary_spp.csv",sep="")) %>% clean_names()
 

### 2020-03-13 Scripts/Initial_Analysis.Rmd
fic bugs

@@
 if(Sys.info()[7] == "hall1000"){
-  data_path <- "/Volumes/DATA/DBEM/"
+  data_path <- "/Volumes/DATA/JULIANO_NEYMAR/TransEmergence/Data/"
 }
@@
 
-    D_Path <- paste(Path,"/mpa0F1ENS",ensamble_list[c],"/", #<- drobo path for 8.5
+    D_Path <- paste(Path,"mpa0F1ENS",ensamble_list[c],"/", #<- drobo path for 8.5
                     taxon_key,"/",taxon_key,data_type,yr,".txt",
@@
 
-# dbem_import(
+dbem_import(
 #   taxon_key <- 600004,
@@
 #   ensamble_list = c(102,103),
-#   Path <- paste(data_path,"DBEM/",sep="")
+#   Path <- DBEM_path
 # )
@@
     fyr <- fyr,
-    data_type <- "Abd",
+    data_type <- data_type,
     ensamble_list = ensamble_list,
-    Path <- paste(data_path,"DBEM/",sep="")
+    Path <- Path
   )
@@
 
-Spp_list <- read.csv(paste(data_path,"Species/Species_list.csv",sep="")) %>% clean_names()
 iyr <- 1951
-fyr <- 2099
+fyr <- 1953
 data_type <- "Catch"
-ensamble_list = seq(102,111,1)
-Path <- paste(data_path,"DBEM/",sep="")
+ensamble_list = seq(102,103,1)
+DBEM_path <- "/Volumes/DATA/DATA/DBEM/"
 world_grid <- tibble(index = seq(1,259200,1))
@@
 Transboundary_spp <- read.csv(paste(data_path,"Species/Transboundary_spp.csv",sep="")) %>% clean_names()
+
+Spp_list <- unique(Transboundary_spp$taxon_key)
 
@@
     lapply(
-      Spp_list, #c(600004,600005),
+      Spp_list[1:2], #c(600004,600005),
       EstPropChange,
@@
       fyr <- fyr,
-      data_type <- "Abd",
+      data_type <- "Catch",
       ensamble_list = ensamble_list,
-      Path <- paste(data_path,"DBEM/",sep=""),
+      Path =DBEM_path,
       Neighbours = Neighbours
@@
 )
+
 ```

### 2020-03-16 Scripts/Initial_Analysis.Rmd
Changes to make it run on Hall

@@
 
-dbem_import(
+# dbem_import(
 #   taxon_key <- 600004,
@@
 iyr <- 1951
-fyr <- 1953
+fyr <- 2099
 data_type <- "Catch"
-ensamble_list = seq(102,103,1)
+ensamble_list <- seq(102,111,1)
 DBEM_path <- "/Volumes/DATA/DATA/DBEM/"
@@
 
-Spp_list <- unique(Transboundary_spp$taxon_key)
+Spp_list <- unique(Transboundary_spp$taxon_key)[3:4]
 
@@
 
-suppressWarnings(
+# suppressWarnings(
   r_proportion_change <- bind_rows(
     lapply(
-      Spp_list[1:2], #c(600004,600005),
+      Spp_list, #c(600004,600005),
       EstPropChange,
@@
   )
-)
+# )
+
+write_csv(r_proportion_change,
+          "r_proportion_change.csv")
 

### 2020-03-16 Scripts/Initial_Analysis.Rmd
small test changes

@@
 
-    total <- NULL
+     # total <- NULL
 
@@
 
-# dbem_import(
-#   taxon_key <- 600004,
-#   iyr <- 1951,
-#   fyr <- 1952,
-#   data_type <- "Abd",
+x <- dbem_import(
+  taxon_key <- 600004,
+  iyr <- 1951,
+  fyr <- 1952,
+  data_type <- "Catch",
   # ensamble_list = c(102,103),
-#   Path <- DBEM_path
-# )
+  ensamble_list =ensamble_list,
+  Path <- DBEM_path
+)
 
-# unique(x$ensamble)
+unique(x$ensamble)
 # head(x)
@@
 #### For testing 
-Spp <- 600004  # Anchovie Equator, Chile and Peru
+# Spp <- 600004  # Anchovie Equator, Chile and Peru
 # Spp <- 600107 # Tuna Many RFMOs
@@
 
-Spp_list <- unique(Transboundary_spp$taxon_key)[3:4]
+Spp_list <- unique(Transboundary_spp$taxon_key)[4:5]
 
@@
 
+# Problema con 600006
+
 ####_______________________ ####
@@
 ####_______________________ ####
-
+r_proportion_change <- NULL
 # suppressWarnings(

### 2020-03-16 Scripts/Initial_Analysis.Rmd
fixed ensamble labeling problem

@@
   data_path <- paste(here(),"/Temporal_Data/",sep = "")
+  DBEM_path <- paste(data_path,"DBEM/",sep="")
 }
@@
   data_path <- "/Volumes/DATA/JULIANO_NEYMAR/TransEmergence/Data/"
+  DBEM_path <- "/Volumes/DATA/DATA/DBEM/"
 }
@@
           total<- copy(cur)  # <- copies the previouse data.table
+
         }else{ 
@@
                 by = "index")
+
+    if(c == 1){
+      partial_ensamble <- world_data
+
+    }else{
+      spp_ensamble <- bind_rows(spp_ensamble,world_data)
+
+    }
+
   # Messages
@@
   }else{
-
     print(paste("Analysis done for",taxon_key, "ensamble" ,ensamble_list[c]))
@@
 
-    if(c == 1){
-      partial_ensamble <- world_data
-    }else{
-      spp_ensamble <- bind_rows(partial_ensamble,world_data)
-    }
   }
@@
   data_type <- "Catch",
-  # ensamble_list = c(102,103),
-  ensamble_list =ensamble_list,
+  ensamble_list = c(102,103,104),
+  # ensamble_list =ensamble_list,
   Path <- DBEM_path
@@
 ensamble_list <- seq(102,111,1)
-DBEM_path <- "/Volumes/DATA/DATA/DBEM/"
 world_grid <- tibble(index = seq(1,259200,1))

### 2020-03-25 Scripts/Initial_Analysis.Rmd
Analysis for estimating the proportion change

--- 2020-03-16 Scripts/Initial_Analysis.Rmd
@@
   data_path <- paste(here(),"/Temporal_Data/",sep = "")
+  DBEM_path <- paste(data_path,"DBEM/",sep="")
+  result_path <- paste(here(),"/Temporal_Data/Results/",sep="")
 }
@@
   data_path <- "/Volumes/DATA/JULIANO_NEYMAR/TransEmergence/Data/"
+  DBEM_path <- "/Volumes/DATA/DATA/DBEM/"
+  result_path <- paste(here(),"/Results/",sep="")
 }
@@
 
-#### The function
-dbem_import <- function(taxon_key,
-                        iyr,
-                        fyr, 
-                        ensamble_list,
-                        data_type,
-                        Path = NA){
+dbem_import <- function(taxon_key, #Except taxon_key 
+                        year, #Expects sequence of years 
+                        ensamble, #Expects a number of the ensamble
+                        data_type, #Expects either "Abd" or "Catch"
+                        path = "NA"){
 
-  # Creat a sequence of the years to import  
-  yr <- seq(iyr,
-            fyr,
-            by=1) #<- change this setting if you have bi/tri/etc annual data
+  years <- year_df %>% pull(year)
 
-  #### Set the Path####
-  #First, set the path of the files under a loop for each year
-
-  for(c in 1:length(ensamble_list)){
-
-    D_Path <- paste(Path,"mpa0F1ENS",ensamble_list[c],"/", #<- drobo path for 8.5
-                    taxon_key,"/",taxon_key,data_type,yr,".txt",
-                    sep="")
-
-     # total <- NULL
+  D_Path <- paste(path,"mpa0F1ENS",ensamble,"/",taxon_key,"/",taxon_key,data_type,years,".txt",sep="")
 
   #### Importing data ####
-    #Now we import the data using data.table::fread and the path generated above
+  cur <- lapply(D_Path, FUN=fread, na.strings="NA")
 
-    for (i in 1:length(yr)){ #<-The years you want to have
-
-      possibleError <-  tryCatch({ # <- This will solve for situations where the species is not found in the DBEM
-        # Imports dataset 
-        cur <- fread(D_Path[i],
-                     na.strings = "NA",
-                     col.names = c("index",
-                                   yr[i]) #<- Calls colunmns by year
+  if(length(cur)>0){
+    cur <- cur[sapply(cur, function(d) nrow(d) >= 1)] 
+    colnames <- c("index", "value")
+    cur <- lapply(cur, setNames, colnames)
+    df <- bind_rows(cur, .id = "column_label")
+    if(nrow(df)>0){
+      frame_key <- tibble(column_label = seq(1,length(unlist(years)),1), 
+                          "year"=years) %>% 
+        mutate(column_label=as.character(column_label))
+      df <- left_join(df, frame_key, by="column_label") %>% select(-column_label)
+      df <- df %>% mutate(data_type=data_type,
+                          taxon_key = taxon_key,
+                          ensamble = ensamble
       )
-
-      },
-      error=function(e){ # Part of the "no species in DBEM
-        e
     }
-      )
-      if(!inherits(possibleError, "error")){
-
-        # possibleError_I <-  tryCatch({
-        #Puts them together in one nice dataset 
-        if(i == 1){
-          total<- copy(cur) # <- copies the previouse data.table
-        }else{ 
-          setkey(total, # <- sets the data.table as "reference" ?setkey
-                 index); setkey(cur,
-                                index); 
-          total <- merge(total, #<- Merges all the data in one single file.
-                         cur,
-                         all = TRUE) # <- This allows for the NA's
-
-        } #Close datatable loop
-
+  } else {
+    df <- tibble()
   }
-
-    } #First loop END (Path Creation)
-
-    # Tidy version
-    total <- total %>%
-      gather(year,value,-index) %>%
-      mutate(taxon_key = taxon_key,
-             ensamble=ensamble_list[c])
-
-    world_data <- world_grid %>%
-      full_join(total,
-                by = "index")
-    # Messages
-    if(inherits(possibleError, "error")){
-
-      print(paste("Oops! --> No information for",taxon_key,yr[i], sep = " ")) # <- Display error message
-
-    }else{
-
-      print(paste("Analysis done for",taxon_key, "ensamble" ,ensamble_list[c]))
-    }
-
-    if(c == 1){
-      partial_ensamble <- world_data
-    }else{
-      spp_ensamble <- bind_rows(partial_ensamble,world_data)
-    }
-  }
-
-  return(spp_ensamble)
-
+    return(df)
 } #Function end
 
-# Test function (works in Carmelia!)
-
-x <- dbem_import(
-  taxon_key <- 600004,
-  iyr <- 1951,
-  fyr <- 1952,
-  data_type <- "Catch",
-  # ensamble_list = c(102,103),
-  ensamble_list =ensamble_list,
-  Path <- DBEM_path
-)
-
-unique(x$ensamble)
-# head(x)
-
-# Test multiple ensambles and spp
-
-# ensamble_list <- c(102,103)
-# taxon_key_list <- c(600004,600005)
-
-# x <- bind_rows(
-#   lapply(
-#     # taxon_key_list,
-#     ensamble_list,
-#     FUN= dbem_import,
-#     taxon_key = 600004,
-#     # ensamble = 102,
-#     iyr = 1951,
-#     fyr = 1952,
-#     data_type = "Abd",
-#     Path = paste(data_path,"DBEM/",sep="")
-#   )
+# x <- dbem_import(
+#   taxon_key <- 600004,
+#   Year <- year_df,
+#   data_type <- "Catch",
+#   ensamble = 102,
+#   # ensamble_list =ensamble,
+#   path <- DBEM_path
 # )
 
-# OK!
+# attach(x)
+# head(x)
+# unique(year)
+# unique(ensamble_list)
+# detach(x)
 
@@
 
-EstPropChange <- function(Spp,iyr,fyr,data_type,ensamble_list,Path,Neighbours){
+EstPropChange <- function(taxon_key,
+                          year,
+                          data_type,
+                          ensamble_list, # expects a list of ensambles to averrage by
+                          path,
+                          neighbours){
 
-  # Get model data from spps
-  SppDist <- dbem_import(
-    taxon_key = Spp,
-    iyr <- iyr,
-    fyr <- fyr,
-    data_type <- data_type,
-    ensamble_list = ensamble_list,
-    Path <- Path
+  # Loads species data for all ensambles
+  spp_dist <- bind_rows(
+    lapply(
+      ensamble_list,
+      dbem_import,
+      taxon_key = taxon_key,
+      year = year,
+      data_type = data_type,
+      path = path
     )
+  )
+
+  # head(spp_dist)
 
@@
 
+  suppressWarnings(
+    trans_EEZ <- Transboundary_spp %>% 
+      filter(taxon_key %in% spp_dist$taxon_key) %>% 
+      left_join(Neighbours,
+                by = c("eez_name","eez_neighbour")
+      )
+  )
+
+  # List of indexes per EEZ
+  index_id <- trans_EEZ %>% 
+    group_by(eez_name,index) %>% 
+    summarise(n=n()) %>% 
+    select(-n)
+
+  # --------------- #
+  # Warning messages:
+  # 1: Column `eez_name` joining factor and character vector, coercing into character vector 
+  # 2: Column `eez_neighbour` joining factor and character vector, coercing into character vector 
+  # --------------- #
+
   #____________ Selecting only the transboundary nature of the species _________ #
-  Trans_Spp <- SppDist %>%
-    left_join(Neighbours,
-              by= "index") %>% 
-    filter(!is.na(value)) %>% 
-    semi_join(Transboundary_spp, # filter only transboundary cases
-              by = c("taxon_key","eez_name","eez_neighbour")
-    )
+  trans_spp <- spp_dist %>%
+    filter(index %in% trans_EEZ$index) %>% 
+    left_join(index_id,
+              by = "index")
 
@@
     ungroup() %>% 
-    select(-n)
+    select(-n) %>% 
+    semi_join(trans_EEZ,
+              by = c("eez_name","eez_neighbour")
+    )
 
   # Step 2. Determines the amount of grids present in each country
-  Spp_Grid <- Trans_Spp %>% 
+  spp_grid <- trans_spp %>% 
     group_by(taxon_key,
@@
     left_join(Neighbours_List,
-              by = "eez_name") %>% 
-    filter(eez_name %in% Trans_Spp$eez_name, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
-           eez_neighbour %in% Trans_Spp$eez_name)
+              by = "eez_name")
 
@@
   # Split dataframes to merge latter
-  Territory_T <- Spp_Grid %>% 
+  Territory_T <- spp_grid %>% 
     ungroup() %>% 
@@
 
-  Neighbour_T <- Spp_Grid %>% 
+  Neighbour_T <- spp_grid %>% 
     ungroup() %>% 
@@
     mutate(area_index = n_spp_Country/Spp_Total) %>% 
-    select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
+    # select(taxon_key,year,ensamble_list,eez_name,eez_neighbour,area_index) %>% 
     mutate(
@@
                  c(ensamble_mean=mean,
-                   ensamble_sd=sd)
-    ) %>% 
-    mutate(ensamble_mean = ifelse(is.na(ensamble_mean),0,ensamble_mean)) %>% 
-    select(-ensamble_sd) %>% 
-    spread(time_step,ensamble_mean) %>% 
-    mutate_at(vars(mid,n_end),
-              funs(ifelse(historical == 0 & . == 0, 0, # If for no change than 0
-                          ifelse(historical == 0 & . > 0, 100, # If changes from 0 to anything then 100
-                                 ((.-historical)/historical)*100 # Estimate percentage
-                                 # ((.-historical)/((.historical)/2))*100 # Estimate percentage
-                          )
-              )
-              )
+                   ensamble_sd=sd),
+                 na.rm=T
     )
 
-  #### Statistical test ###
+  # --------------- #
+  # End function
+  # --------------- #
 
-  stat_change_d <- full_join(Territory_T,
-                             Neighbour_T, 
-                             by = c("Name","taxon_key","year","ensamble")
-  ) %>%
-    rowwise() %>%
-    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
-    distinct() %>% # Removes false duplicates from `full_join()`
-    rename(eez_name = Name,
-           eez_neighbour =eez_name,
-           n_spp_Country = n_spp_eez.x,
-           n_spp_Neighbour = n_spp_eez.y) %>% 
-    mutate(area_index = n_spp_Country/Spp_Total) %>% 
-    select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
-    mutate(
-      time_step = ifelse(year < 2000,"historical",
-                         ifelse(year >= 2045 & year <= 2064, "mid",
-                                ifelse(year >= 2080, "n_end","NA")
-                         )
-      )
-    )
+  if(nrow(per_change_d) > 0){
 
-  return(per_change_d)
+  # Save Result
+  name <- paste("proportion_",taxon_key,".csv",sep="")
+  path_name <- paste(result_path,"Proportion/",name, sep="")
 
+  write_csv(per_change_d,
+            path_name)
+
+  return(print(paste("analysis done",taxon_key)))
+
+  }else{
+    print(paste("no result for",taxon_key))
+  }
+
+  # Or return a df
+  # return(per_change_d)
 }
@@
 
-iyr <- 1951
-fyr <- 2099
+# Data needed for estimating Proportion change
+
+year_df <- tibble(year =  c(seq(1995,2014,1), seq(2041,2060,1), seq(2080,2099,1)),
+                  time_period = c(rep("Today", 20), rep("Mid Century", 20), rep("End of Century", 20))
+                  )
+
 data_type <- "Catch"
-ensamble_list <- seq(102,111,1)
-DBEM_path <- "/Volumes/DATA/DATA/DBEM/"
+ensamble_list <- seq(102,103,1)
 world_grid <- tibble(index = seq(1,259200,1))
@@
 
-Spp_list <- unique(Transboundary_spp$taxon_key)[4:5]
+spp_list <- unique(Transboundary_spp$taxon_key)[1:3]
 
@@
 
-## Main Routine
+## Proportion Routine
 
-```{r main_routine, eval = T, echo = F}
-
-# Problema con 600006
-
+```{r Proportion_routine, eval = T, echo = F}
 ####_______________________ ####
@@
 ####_______________________ ####
-r_proportion_change <- NULL
-# suppressWarnings(
-  r_proportion_change <- bind_rows(
+
 lapply(
-      Spp_list, #c(600004,600005),
+  spp_list, #c(600004,600005),
   EstPropChange,
-      iyr <- iyr,
-      fyr <- fyr,
+  year <- year_df,
   data_type <- "Catch",
   ensamble_list = ensamble_list,
-      Path =DBEM_path,
-      Neighbours = Neighbours
+  path = DBEM_path,
+  neighbours = Neighbours
 )
-  )
-# )
-
-write_csv(r_proportion_change,
-          "r_proportion_change.csv")
 

### 2020-03-25 Scripts/Initial_Analysis.Rmd
add time just for controling
added parallelization

@@
   "ggrepel",
-  "zoo"
+  "zoo",
+  "parallel" # for mclapply
 )
@@
 
-lapply(
+t_start <- Sys.time()
+
+mclapply(
   spp_list, #c(600004,600005),
@@
 
+t_end <- Sys.time()-t_start;t_end
+
 ```

### 2020-03-25 Scripts/Initial_Analysis.Rmd
Fixed results paths for hall 1000

@@
   DBEM_path <- "/Volumes/DATA/DATA/DBEM/"
-  result_path <- paste(here(),"/Results/",sep="")
+  result_path <- "/Volumes/DATA/JULIANO_NEYMAR/TransEmergence/Results/"
 }

### 2020-03-25 Scripts/Initial_Analysis.Rmd
changes for testing

@@
 data_type <- "Catch"
-ensamble_list <- seq(102,103,1)
+ensamble_list <- seq(102,111,1)
 world_grid <- tibble(index = seq(1,259200,1))
@@
 
-spp_list <- unique(Transboundary_spp$taxon_key)[1:3]
+spp_list <- unique(Transboundary_spp$taxon_key)
 
@@
 ```{r Proportion_routine, eval = T, echo = F}
+
 ####_______________________ ####
@@
 mclapply(
-  spp_list, #c(600004,600005),
+  spp_list[1],
   EstPropChange,
@@
   data_type <- "Catch",
-  ensamble_list = ensamble_list,
+  ensamble_list = ensamble_list[1],
   path = DBEM_path,

### 2020-03-26 Scripts/Initial_Analysis.Rmd
Code for first analisis done

@@
 mclapply(
-  spp_list[1],
+  spp_list,
   EstPropChange,
@@
   data_type <- "Catch",
-  ensamble_list = ensamble_list[1],
+  ensamble_list = ensamble_list,
   path = DBEM_path,
@@
 
+# Hall 1000 ran it in 5.33 hours
+
 ```
+
+# Results
+
+## Functions and Data needed
+
+### Fun `SummaryProp`
+
+This function estimates the total change in MCP proportion of each species for each EEZ. It looses the connection with neighbours as you only know for x species houw much is the proportion changing.
+
+```{r fun_SummaryProp, eval = T, echo =F}
+
+  SummaryProp <- function(taxon_key){
+
+    # Reads taxon data
+    proportion_data <- fread(paste(result_path,"Proportion/","proportion_",taxon_key,".csv", sep=""))
+
+    # summarizes by country
+    country_summary <- proportion_data %>% 
+      group_by(taxon_key,eez_name,time_step) %>% 
+      summarise(
+        mean_spp = mean(ensamble_mean,na.rm=T) # average spp change from all countries sharing
+      )
+
+    return(country_summary)
+  }
+
+```
+
+### Fun `EEZProp`
+
+The next step is to summarize all species per eez
+
+```{r fun_EEZProp, eval = T, echo = F}
+
+EEZProp <- function(taxon_list){
+
+ spp <- bind_rows(
+   lapply(taxon_list,SummaryProp)
+   )
+
+ eez_summary <- spp %>% 
+   group_by(eez_name,time_step) %>% 
+   summarise(
+      min_spp = min(mean_spp),
+     max_spp = max(mean_spp),
+     sd_spp = sd(mean_spp,na.rm=T), # for some weird reason it wont work if the order is fliped
+     mean_spp = mean(mean_spp,na.rm=T)
+   ) %>% 
+   select(eez_name:max_spp,mean_spp,sd_spp)
+
+ return(eez_summary)
+
+}
+
+taxon_list <- c(600015,600017)
+taxons <- c(600004,600004,600015,600017)
+x <- NULL
+x <- PropChange(taxons)
+
+```
+
+### Fun `PropChange`
+
+Finally, we estimate the average proportion change of all species within each eez
+
+```{r fun_PropChange, eval = T, echo =F}
+
+PropChange <- function(taxon_list){
+
+ eez <- EEZProp(taxon_list)
+
+ proportion_change <- eez %>% 
+   gather("variable","value",min_spp:sd_spp) %>% 
+   spread(time_step,value) %>% 
+   mutate_at(vars(
+     "per_mid" = mid,
+     "per_end" = n_end),
+            funs(round((((.-historical)/.))*100,2))
+   )
+
+ return(eez_summary)
+
+}
+
+```
+
+## Run Results Analysis
+
+```{r fun_PropChange, eval = T, echo =F}
+
+# Get taxon list
+taxon_list <- list.files(paste(result_path,"/Proportion/",sep=""))
+
+# clean names for function
+taxon_list <- gsub(".*_","",taxon_list) 
+taxon_list <- gsub("\\..*","",species_list)
+
+Restuls <- PropChange(taxon_list)
+
+```

### 2020-03-26 Scripts/Initial_Analysis.Rmd
save data for plotting
fixed bug

@@
 
-taxon_list <- c(600015,600017)
-taxons <- c(600004,600004,600015,600017)
-x <- NULL
-x <- PropChange(taxons)
-
 ```
@@
 
- return(eez_summary)
+ return(proportion_change)
 
@@
 taxon_list <- gsub(".*_","",taxon_list) 
-taxon_list <- gsub("\\..*","",species_list)
+taxon_list <- gsub("\\..*","",taxon_list)
 
-Restuls <- PropChange(taxon_list)
+Restuls <- PropChange(taxon_list[1:2])
+
+write.csv(Results,
+          "Clean_Results.csv")
 

### 2020-03-26 Scripts/Initial_Analysis.Rmd
fixing nan bugs

@@
    summarise(
-      min_spp = min(mean_spp),
-     max_spp = max(mean_spp),
+     n_spp = n(),
+     min_spp = min(mean_spp,na.rm=T),
+     max_spp = max(mean_spp,na.rm=T),
      sd_spp = sd(mean_spp,na.rm=T), # for some weird reason it wont work if the order is fliped
@@
  proportion_change <- eez %>% 
-   gather("variable","value",min_spp:sd_spp) %>% 
+   gather("variable","value",n_spp:sd_spp) %>% 
    spread(time_step,value) %>% 
@@
 
-Restuls <- PropChange(taxon_list[1:2])
+results <- PropChange(taxon_list)
 
-write.csv(Results,
-          "Clean_Results.csv")
+# write_csv(Results,
+#           "Clean_Results.csv")
+
+clean_results <- fread(paste(result_path,"Clean_Results.csv",sep=""))
+View(clean_results)
 
 ```
+
+## Figure 1. Change Map
+
+```{r figure_one, eval = T, echo =F}
+
+# Load SAU shapefile name and paths
+
+# The path
+path_world <- paste(data_path,"Spatial_Data/SAU_Shapefile/",sep="")
+
+# The File
+fnam_world <- "SAUEEZ_July2015.shp"
+
+# Load it!
+World_sf <- st_read(dsn = path_world,
+                    layer =file_path_sans_ext(fnam_world)) %>% 
+  rename(eez_name = Name)
+
+# View(unique(World_sf$eez_name))
+
+```

### 2020-03-27 Scripts/Initial_Analysis.Rmd
Included figure with connections

@@
   "zoo",
-  "parallel" # for mclapply
+  "parallel", # for mclapply,
+  "sf", # for mapping
+  "st", # for mapping
+  "rgdal", #Spatial analysis
+  "tools", #Spatial analysis 
+  "zeallot" # for Juanito's map
 )
@@
   result_path <- "/Volumes/DATA/JULIANO_NEYMAR/TransEmergence/Results/"
+}
+
+ggtheme_map <- function(base_size = 9, Region = "NA") {
+
+  theme(text             = element_text(#family = "Helvetica",
+    color = "gray30", size = base_size),
+    plot.title       = element_text(size = rel(1.25), hjust = 0, face = "bold"),
+    panel.background = element_blank(),
+    panel.border     = element_blank(),
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_line(color = "transparent"),
+    strip.background =element_rect(fill = "transparent"),
+    strip.text.x = element_text(size = 18, colour = "black",face= "bold.italic", angle = 0),
+    axis.line        = element_blank(), # element_line(colour = "grey30", size = .5))
+    axis.ticks       = element_blank(),
+    axis.text        = element_blank(),
+    axis.title       = element_blank(),
+    legend.key       = element_rect(colour = NA, fill = NA, size = 4),
+    legend.position = "bottom",
+    legend.key.width =unit(5,"line")
+  )
 }
@@
 
-## Functions and Data needed
+## Percentage Change
 
-### Fun `SummaryProp`
+This results only shows the average percentage change in proportion of all speecies within an EEZ without relationship with neighbours
+
+### Functions and Data needed
+
+#### Fun `SummaryProp`
 
@@
 
-  SummaryProp <- function(taxon_key){
+SummaryProp <- function(taxon_key,links="NA"){
 
@@
 
+  if(links == "y"){
+    # summarizes by country
+    country_summary <- proportion_data %>% 
+      group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% 
+      summarise(
+        mean_spp = mean(ensamble_mean,na.rm=T) # average spp change from all countries sharing
+      )
+
+  }else{
     # summarizes by country
@@
       )
+    }
 
@@
 
-### Fun `EEZProp`
+#### Fun `EEZProp`
 
@@
 
-EEZProp <- function(taxon_list){
+EEZProp <- function(taxon_list,links="NA"){
 
  spp <- bind_rows(
-   lapply(taxon_list,SummaryProp)
+   lapply(taxon_list,SummaryProp,links)
    )
+
+ if(links == "y"){
+
+   eez_summary <- spp %>% 
+     group_by(eez_name,eez_neighbour,time_step) %>% 
+     summarise(
+       n_spp = n(),
+       min_spp = min(mean_spp,na.rm=T),
+       max_spp = max(mean_spp,na.rm=T),
+       sd_spp = sd(mean_spp,na.rm=T), # for some weird reason it wont work if the order is fliped
+       mean_spp = mean(mean_spp,na.rm=T)
+     ) %>% 
+     select(eez_name:max_spp,mean_spp,sd_spp)
+
+ }else{
 
@@
 
+ }
+
  return(eez_summary)
@@
 
-### Fun `PropChange`
+#### Fun `PropChange`
 
@@
 
-PropChange <- function(taxon_list){
+PropChange <- function(taxon_list,links="NA"){
 
- eez <- EEZProp(taxon_list)
+ eez <- EEZProp(taxon_list,links)
 
@@
 
-## Run Results Analysis
+### Run Results Analysis
 
@@
 
-results <- PropChange(taxon_list)
+# Run results
+results <- PropChange(taxon_list, links="y") %>% 
+  mutate(per_mid =ifelse(per_mid >= 25, 25,
+                           ifelse(per_mid <= -25,-25,per_mid)
+                           )
+         )
 
-# write_csv(Results,
-#           "Clean_Results.csv")
+# View(results)
 
-clean_results <- fread(paste(result_path,"Clean_Results.csv",sep=""))
-View(clean_results)
-
-```
-
-## Figure 1. Change Map
-
-```{r figure_one, eval = T, echo =F}
+### Load Data
 
@@
 # The path
-path_world <- paste(data_path,"Spatial_Data/SAU_Shapefile/",sep="")
+path_world <- paste(data_path,"Spatial/SAU_Shapefile/",sep="")
 
@@
 # Load it!
-World_sf <- st_read(dsn = path_world,
+World_eez_sf <- st_read(dsn = path_world,
                     layer =file_path_sans_ext(fnam_world)) %>% 
-  rename(eez_name = Name)
+  rename(eez_name = Name) %>% 
+  st_transform(crs = 4326) %>% # 4326
+  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)
 
-# View(unique(World_sf$eez_name))
+# Call world map from natural earth
+world_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf"))
 
 ```
+
+### Fig.Percentage Change Map
+
+```{r fig_map_per, eval = F, echo =F}
+
+World_sf %>% 
+  left_join(results,
+            by = "eez_name") %>% 
+  filter(variable %in% c("mean_spp")#,
+         # eez_name %in% c("Peru","Chile","Ecuador","Argentina")
+         ) %>% 
+  ggplot() +
+  geom_sf(
+    aes(
+      fill = per_mid
+    )
+  ) + 
+  geom_sf(data=country10_sf) +
+  scale_fill_gradient2("Average Change in\nStock-Share Ratio (%)\nBy Mid Century") +
+  ggtheme_map() +
+  ggsave("figure_one.png",
+         width = 10,
+         height = 6,
+         units = "in"
+  )
+
+```
+
+### Fig. Netowrk
+
+```{r fig_map_Juan, eval = F, echo =F}
+
+# Figure made thanks to Juanito!
+# http://jsmayorga.com/post/mapping-the-global-network-of-transnational-fisheries/
+
+# Get the center poligon of each EEZ for map source/target
+coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
+  ungroup() %>% 
+  select(eez_name,geometry) %>% 
+  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,")
+
+# Remove "c()"
+coords$longitude <- gsub("\\(","",coords$longitude)
+coords$longitude <- gsub("c","",coords$longitude)
+coords$latitude <- gsub(")","",coords$latitude)
+
+coords <- coords %>% 
+  mutate_at(vars(longitude,latitude),as.numeric)
+
+# Prepare data from results
+Data <- results %>% 
+  # get coords for sources
+  left_join(coords,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  left_join(coords,
+            by = "eez_name") %>% 
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  mutate(
+    per_mid = ifelse(is.na(per_mid),0,per_mid)
+  ) %>% 
+  drop_na() %>% 
+  filter(per_mid > 0,
+         variable == "mean_spp"#,
+         # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
+  )
+
+# Convert to data to spatial
+network_data_sf <- Data %>%
+  select(start_lon, start_lat, end_lon, end_lat) %>% 
+  purrr::transpose() %>% 
+  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
+  purrr::map(st_linestring) %>%
+  st_sfc(crs = 4326) %>%
+  st_sf(geometry = .)%>%
+  bind_cols(Data) %>%
+  select(everything(), geometry) %>% 
+  # Set circles ('cus the world is not flat!)
+  st_segmentize(units::set_units(100, km)) %>% 
+  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
+  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
+  sf::st_sf(crs = 4326)
+
+#  unpack the coordinates of each feature 
+coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
+
+c(network_data_sf$line_id, network_data_sf$coords) %<-% (
+  coord_data %>% 
+    rename(subline_id = L1) %>%   
+    nest(-L2)
+)
+
+# Create data path for line connections
+paths <- network_data_sf %>% 
+  sf::st_set_geometry(NULL) %>% 
+  unnest() %>% 
+  mutate_at(vars(X,Y), round,2)
+
+# Plot it!
+
+x <- ggplot() +
+  geom_sf(data = world_map, size = 0.1, fill = "gray90", col = "gray90") +
+  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray70") +
+  # geom_point(data = Data,
+  #            aes(start_lon,
+  #                start_lat,
+  #                col = per_mid),
+  #            size = 0.1,
+  #            col = "black",
+  #            # shape = 21,
+  #            stroke = 1.5) +
+  geom_path(data = paths,
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = per_mid
+            ),
+            size = 0.3,
+            arrow = arrow(length = unit(0.009, "npc")),
+            show.legend = NA) +
+  # geom_sf_text(data =World_eez_sf, aes(label =eez_name),
+  #              size = 2) +
+  # geom_text_repel(data = paths,
+  #                  aes(
+  #                    x = X,
+  #                    y = Y,
+  #                    label =source),
+  #                  force = 1
+  # )+
+  ggtheme_map() +
+  scale_colour_gradientn("Gain of Stock-Share Ratio (%)",
+                         colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
+                         limits = c(0,25),
+                         breaks = seq(0,25,5)
+  )+ 
+  ggsave("figure_connections.png",
+       width = 10,
+       height = 6,
+       units = "in"
+)
+
+```

### 2020-04-06 Scripts/Manuscript.Rmd
Drafted methods

@@
 
+The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, established political boundaries across the distribution of many fish species, creating shared species between nations.
+Thus, transboundary fish stocks are defined as those populations that cross the EEZ of two or more neighboring nations.
+Despite not implicit, UNCLOS incentivizes nations to cooperate in the management of transboundary fisheries.
+Today, most management plans emplaced are based on historical distribution of catches.
+Climate change represents a challenge for the joint management of fish stocks.
+For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures, quota allocation, and fishery newcomers.
+Recent events suggests that even changes in species that contribute a relatively small proportion of the overall catch can lead to important international conflict.
+This is the case of the conflict between the EU, Norway, Iceland, and the Faroe Islands over Atlantic mackerel (*Scomber scombrus*) which represents 8% of the catch of involved nations.
+Thus, in this chapter I explore the climate-driven changes in stock-share-proportion (e.g.
+the proportion of the stock that each EEZ has) of the previously identified (Chapter One) transboundary species.
+Moreover, I determine the *time of emergence* of each fish stock, defined as the year in which the stock-share-proportion will change from its historical average.
+Preliminary results suggest that many neighboring EEZs will see changes in stock-share ratio between 1 and 10% with some exceeding 25%.
+The results of this chapter can inform fisheries management as it not only identifies at the EEZ level stocks that will be likely to create international conflict, but also estimates the time of emergence that can be used to prioritize management plans in the face of a changing world.
+
 # Introduction
 
+### Link paper un terms of SDGs
+
+- UnitedNations2018 Blanchard:2017hl Singh:2017ds
+
+### Collaborative management of share stocks
+
+- Theat points
+- Concept of treaty estabillity
+- Oremus:2020id
+
 # Methods
+
+## Projecting species distributions under cliamte change
+
+We projected the distribution of marine species using a dynamic bioclimatic envelope model (hearafter called DBEM).
+The DBEM algorythm includes species physiology, habitat suitabillity, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
+For pelagic species, the model uses environmental variables at the surface (e.g.
+sea surface temperature) whereas demersal species distribution are driven by ocean-bottom variables (e.g.
+bottom temperature).
+This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (See [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
+
+We coupled the DBEM with the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project species distribution from 1951 to 2100 under cliamte change [@Dunne:2012bo; @Dunne:2013db].
+We run the DBEM simulating biomass and MCP from 1950 to 2100 under a “worst-case” climate change scenario accroding to the Representative Concentration Pathway 8.5 (RCP 8.5) used by the Intergovernmental Panel on Climate Change (IPCC) [@IPCC:2014bz; Riahi:2011dk].
+According to the GFDL-ESM2M, under the RCP 8.5 scenario atmospheric temperature will increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial.
+We run the DBEM under ten different GFDL-ESM2M ensamble members [@Roggers:2015jq].
+While all ensamble members follow the same RCP, they all constitute different initial Earth conditions.
+As a consequence, for each GFDL-ESM2M-ensamble the Earth system is infinitesimally perturbed resulting in an unique atmosphere and ocean state at each point in time.
+Thus, each ensamble has a different state of internal variability representing a plausible future under RCP 8.5 [@Roggers:2015jq].
+
+## Data analysis
+
+The Food and Agricultural Organization (FAO) of the UN define transboundary species as those that freely swim within the EEZ of neighbouring nations [@REFFAO].
+Accordingly, we included in this analysis 633 marine exploited transboundary species representing 80% of the catch taken from the world's EEZs [@Palacios-Abrantes:xxxx].
+We consider each species as a single stock unit within an EEZ due to lack of more detailed information [@Teh:2015gd; @Palacios-Abrantes:xxxx].
+We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org).
+
+### Changes in Stocks Share Ratio
+
+In game theory, a “threat point” is known as the minimum payoff that a player is willing to receive in order to cooperate with other players [@Sumaila:E&S].
+We estimate each stock's stock share ratio (e.g.
+a players threat point) defined as the minimum proportion of the catch that each country will acept if they were to cooperate in the management of transboundary stocks [@Palacios-Abrantes:E&S].
+For each stock and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighbouring EEZs, and then calculated the proportion that each EEZ had [@Palacios-Abrantes:xxxx].
+We then average the results in three time periods in order tho reduce climate variabillity within the projections.
+The first time period ($t_h$) comprises (1951 - 2005) and is assumed as the historic distribution of each species, thus representing each neighbouring nation's threat point.
+This time frame matches the historic climate data that the GFDL uses to project climate change [@Dunne:2012bo; @Dunne:2013db], and has previouslly been used as a reference point in fisheries studies [@REF].
+We then selected two future periods; early $21^st$ century as the average of 2021 to 2040 ($t_e$), and mid $21^st$ century as the average of 2041 to 2060 ($t_m$).
+We choose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing) and  SDG 1.2 (poverty reduction), aimed to be achieved by 2020 and 2030, respectively [@Singh:2017ds].
+We then estimate the percentage change in stock share ratio of each future time period relative to the historic value.
+The analysis was repplicated for each of the ten ensamble members and ensamble variation captured by averaging results among ensamble members (mean $\pm$ standar deviation).
+
+### Time of Emergence Index
+
+An important question is estimating the time in which stock share ratio is going to change from a historical variablillty in order for policy to cope with the change.
+
+For that we developed an index of transboundary range shift (*I*), fisheries and climate change based on a species distribution centroid:
+
+$$I =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h)}})^2$$
+
+where $D_A$ and $D_B$ represent the distribution centroid of a species in EEZs A and B, respectiveley for each time step from 2000 to 2100 (*t*) and *sd* is the standar deviation of the historical (1951 - 2000) centroid distribution for $D_A$ and $D_B$.
+
+In cases where the species is distributed in more than 2 neighbouring EEZs (e.g. *xxx*), the index is as follows:
+
+$$I =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h})}) |  \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{C,y}}{sd(D_{B,h})}) | \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{Cy}}{sd(D_{B,h})})/N$$
+
+This way, for two neighbouring EEZs sharing a stock, if the centroid shifts towards EEZ A (*DA* decreases) but away from EEZ A B (*DB* increases), EEZ A may catch a bigger share over B and the risk of dispute increases.
+Consequently, if the centroid shifts are within historical variability, risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability, the risk of dispute is higher.
+Thus, risk is positively related to the frequency of centroid shifts beyond historical variability (Fig.
+1).
+
+![Fig 1.
+Theorical representation of transboundary Index.
+(A) Case with two neighbouring EEZs.
+(B) Case with three neigbouring EEZs.
+Grey dotted boxes represent each EEZ boundaries, t1 and t2 represent time periods one and two, respectiveley, and green and red lines represent an increase and decrease in D, respectiveley](~/GitHub/TransEmergence/Figures/Fig1.png)
+
+### Statistical analysis
+
+All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *wesanderson* [@Packagewesanderson:2018vn], *rfishbase* [@PackagerfishbaseR:2019th], *R.matlab* [@PackageRmatlabRe:2018un], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], and *zoo* [@PackagezooSInfr:2019tm].
+All code is available at https://github.com/jepa/TransEmergence
+
+Things to consider:
+
+- Disputed territories (e.g. China)
 

### 2020-04-06 Scripts/Initial_Analysis.Rmd
Fixed bugs in connection figure

@@
 
-```{r GetSppDist, eval = T, echo = T}
+```{r GetSppDist, eval = F, echo = T}
 
@@
 
-## Proportion approach
+## Changes in Proportion
 
@@
 
-```{r EstPropChange, eval = T, echo = F}
+```{r EstPropChange, eval = F, echo = F}
 
@@
 
-```{r control_panel}
+```{r control_panel, eval = F}
 
@@
 world_grid <- tibble(index = seq(1,259200,1))
-Neighbours <- read_excel(paste(data_path,"Spatial/EEZ_Neighbour_List.xlsx",sep="")) %>% 
-  clean_names()
 
@@
 
-```{r Proportion_routine, eval = T, echo = F}
+```{r Proportion_routine, eval = F, echo = F}
 
@@
   select(eez_name,geometry) %>% 
-  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,")
+  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
+  # Manually fix issues
+  mutate(
+    longitude = ifelse(
+      eez_name == "Fiji", 178.065033,
+      ifelse(eez_name == "Tuvalu",177.6493,
+             ifelse( eez_name == "New Zealand",174.8860,
+                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,
+                             ifelse( eez_name == "Brazil",-36.937,longitude
+                             )
+                     )
+             )
+      )
+    )
+  )
 
@@
   geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray70") +
-  # geom_point(data = Data,
-  #            aes(start_lon,
-  #                start_lat,
-  #                col = per_mid),
-  #            size = 0.1,
-  #            col = "black",
-  #            # shape = 21,
-  #            stroke = 1.5) +
+  geom_point(data = Data,
+             aes(start_lon,
+                 start_lat,
+                 col = per_mid),
+             size = 0.05,
+             col = "black",
+             shape = 20,
+             stroke = 1.5) +
   geom_path(data = paths,
@@
 ```
+
+## Troubleshooting
+
+### Weird connections from Australia
+
+```{r fig_map_Juan, eval = F, echo =F}
+
+##_----------------------- #
+# Checking on China and Indonesia
+# Probably due to dispute territories
+
+Neighbours <- read_excel(paste(data_path,"Spatial/EEZ_Neighbour_List.xlsx",sep="")) %>% 
+  clean_names()
+
+# 
+# Neighbours %>% 
+#   filter(eez_name == "China") %>% 
+#   pull(eez_neighbour) %>% 
+#   unique()
+
+##_----------------------- #
+
+# Get the center poligon of each EEZ for map source/target
+coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
+  ungroup() %>% 
+  select(eez_name,geometry) %>% 
+  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
+  # Manually fix issues
+  mutate(
+    longitude = ifelse(
+      eez_name == "Fiji", 178.065033,
+      ifelse(eez_name == "Tuvalu",177.6493,
+             ifelse( eez_name == "New Zealand",174.8860,
+                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
+                     )
+             )
+      )
+    )
+  )
+
+# Remove "c()"
+coords$longitude <- gsub("\\(","",coords$longitude)
+coords$longitude <- gsub("c","",coords$longitude)
+coords$latitude <- gsub(")","",coords$latitude)
+
+coords <- coords %>% 
+  mutate_at(vars(longitude,latitude),as.numeric)
+
+# Prepare data from results
+Test_Data <- results %>% 
+  # get coords for sources
+  left_join(coords,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  left_join(coords,
+            by = "eez_name") %>% 
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  mutate(
+    per_mid = ifelse(is.na(per_mid),0,per_mid)
+  ) %>% 
+  drop_na() %>% 
+  filter(
+    # per_mid >= 20,
+    variable == "mean_spp",
+  )
+
+# Convert to data to spatial
+network_data_sf <- Test_Data %>%
+  select(start_lon, start_lat, end_lon, end_lat) %>% 
+  purrr::transpose() %>% 
+  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
+  purrr::map(st_linestring) %>%
+  st_sfc(crs = 4326) %>%
+  st_sf(geometry = .)%>%
+  bind_cols(Test_Data) %>%
+  select(everything(), geometry) %>% 
+  # Set circles ('cus the world is not flat!)
+  st_segmentize(units::set_units(100, km)) %>% 
+  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
+  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
+  sf::st_sf(crs = 4326)
+
+#  unpack the coordinates of each feature 
+coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
+
+c(network_data_sf$line_id, network_data_sf$coords) %<-% (
+  coord_data %>% 
+    rename(subline_id = L1) %>%   
+    nest(-L2)
+)
+
+# Create data path for line connections
+paths <- network_data_sf %>% 
+  sf::st_set_geometry(NULL) %>% 
+  unnest() %>% 
+  mutate_at(vars(X,Y), round,2)
+
+# Plot it!
+
+country <- c("Fiji","Solomon Isl.","Tuvalu")
+
+ggplot() +
+  geom_sf(
+    # data = subset(world_map,name %in% country),
+    data = world_map,
+    size = 0.1, fill = "gray90", col = "gray90") +
+  geom_sf(
+    # data = subset(World_eez_sf, eez_name %in% country),
+    data = World_eez_sf,
+    size = 0.1, fill = "white", col = "gray70") +
+  geom_point(data = Test_Data,
+             # data = subset(Test_Data, source %in% country & target %in% source),
+             aes(start_lon,
+                 start_lat,
+                 col = per_mid),
+             size = 0.1,
+             col = "black",
+             # shape = 21,
+             stroke = 1.5) +
+  geom_path(data = paths,
+            # data = subset(paths, source %in% country),
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = per_mid
+            ),
+            size = 0.3,
+            arrow = arrow(length = unit(0.009, "npc")),
+            show.legend = NA) +
+  # geom_sf_text(data = subset(World_eez_sf, eez_name %in% country),
+  #              aes(label =eez_name),
+  #              size = 2) +
+  # geom_text(data = subset(Test_Data, source %in% country),
+  #              aes(start_lon,
+  #                start_lat,
+  #                label = paste(source,"to",target)
+  #              ),
+  #              size = 2) +
+  # geom_text_repel(data = Test_Data,
+  #   aes(start_lon,
+# start_lat,
+# label = paste(source,"to",target)
+# )
+# )+
+ggtheme_map() +
+  scale_colour_gradientn("Gain of Stock-Share Ratio (%)",
+                         colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
+                         limits = c(0,25),
+                         breaks = seq(0,25,5)
+  )
+
+```

### 2020-04-15 Scripts/Initial_Analysis.Rmd
Working emerging function

@@
 
-```{r Ch3_setup, eval = T, echo=F, warning=F,message=F, results='hide'}
+```{r setup, eval = T, echo=F, warning=F,message=F, results='hide'}
 
@@
   "tools", #Spatial analysis 
-  "zeallot" # for Juanito's map
+  "zeallot", # for Juanito's map
+  "gmt" # for estimating distances between points
 )
@@
 
-### Fun `GetSppDist`
+### Fun `dbem_import`
 
@@
 
-```{r GetSppDist, eval = F, echo = T}
+```{r dbem_import, eval = F, echo = T}
 
@@
 
-## Changes in Proportion
-
-### Fun `EstPropChange`
+## Determine Coords centroids and Neighbouring ids
+
+```{r, get_eez_centroids, eval = F}
+
+# Load SAU shapefile name and paths
+
+# The path
+path_world <- paste(data_path,"Spatial/SAU_Shapefile/",sep="")
+
+# The File
+fnam_world <- "SAUEEZ_July2015.shp"
+
+# Load it!
+World_eez_sf <- st_read(dsn = path_world,
+                    layer =file_path_sans_ext(fnam_world)) %>% 
+  rename(eez_name = Name) %>% 
+  st_transform(crs = 4326) %>% # 4326
+  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)
+
+# Get the center poligon of each EEZ for map source/target with st_centroid
+coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
+  ungroup() %>% 
+  select(eez_name,geometry) %>% 
+  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
+  # Manually fix issues
+  mutate(
+    longitude = ifelse(
+      eez_name == "Fiji", 178.065033,
+      ifelse(eez_name == "Tuvalu",177.6493,
+             ifelse( eez_name == "New Zealand",174.8860,
+                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,
+                             ifelse( eez_name == "Brazil",-36.937,longitude
+                             )
+                     )
+             )
+      )
+    )
+  )
+
+# Remove "c()"
+coords$longitude <- gsub("\\(","",coords$longitude)
+coords$longitude <- gsub("c","",coords$longitude)
+coords$latitude <- gsub(")","",coords$latitude)
+
+coords <- coords %>% 
+  mutate_at(vars(longitude,latitude),as.numeric)
+
+write_csv(coords,
+          "eez_centroids.csv")
+
+# Get id of neighbouring pairs
+Neighbours_spp <- Transboundary_spp %>% 
+  group_by(eez_name,eez_neighbour) %>% 
+  summarise(n=n()) %>% 
+  ungroup() %>% 
+  select(-n) %>% 
+  rownames_to_column("neighbour_id")
+
+#  set ids 
+df1 <- Neighbours_spp
+df2 <- Neighbours_spp
+colnames(df2) <- c("neighbour_id", "eez_neighbour", "eez_name")
+
+df_full <- bind_rows(df1, df2) %>% 
+  group_by(eez_neighbour,eez_name) %>% 
+  summarise(neighbour_id = min(neighbour_id))
+
+# write_csv(df_full, "Neighbours_eez_id.csv")
+
+```
+
+## Estimating Proportion Change
+
+#### Fun `EstPropChange`
 
@@
 
-### Fun `EstTransIndex`
-
-```{r funEstTransIndex, eval= F ,echo = T,warning = F,message = F}
-
-# Function to determine whether or not a species is transboundary
-
-# Varibales needed
-# Spp: Species to be analized, data with presence/absence per gridcell
-# Index_EEZ: Reference list of all INDEX that fall within EEZs
-# EEZ_Size: Number of grid-cells that each EEZ has
-# Neighbours: Reference list of neighbouring countries
-
-# Needs getSppDist
-
-#### For testing 
-# Spp <- 600004  # Anchovie Equator, Chile and Peru
-# Spp <- 600107 # Tuna Many RFMOs
-# Spp <- 600004 # Argument 1 must have names (RFMO)
-# Spp <- 600464 # Discrete species 
-# Spp <- 600051 # One of Pierrs massive (wrong) discrete spp
-# x <- c(600464, # Discreat spp in one nation
-# 600059, # nowhere to be found spp
-# 601352, # Discrete spp in two nations
-# 600004) # transboundary spp
-# Model <- "All"
-# Coord <- CoorG
-# 
-# # For EEZ
-# Neighbours <- Neighbours_Data
-# Index_Code <- Index_Code
-
-# __________________________________ #
-
-# The function
-EstTransIndex <- function(Spp,iyr,fyr,data_type,ensamble_list,Path,Neighbours){
-
-  # Get model data from spps
-  SppDist <- dbem_import(
-    taxon_key = Spp,
-    iyr <- iyr,
-    fyr <- fyr,
-    data_type <- "Abd",
-    ensamble_list = ensamble_list,
-    Path <- paste(data_path,"DBEM/",sep="")
-  )
-
-  # Result 1. Number of Countries that share the species
-
-  #____________ Selecting only the transboundary nature of the species _________ #
-  Trans_Spp <- SppDist %>%
-    left_join(Neighbours,
-              by= "index") %>% 
-    filter(!is.na(value)) %>% 
-    semi_join(Transboundary_spp, # filter only transboundary cases
-              by = c("taxon_key","eez_name","eez_neighbour")
-    )
-
-  #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
-  # The number of species' cells present within each country's EEZ
-
-  #Step 1.  Get EEZ id and Neighbour
-  Neighbours_List <- Neighbours %>% 
-    group_by(eez_name,eez_neighbour) %>% 
-    summarise(n=n()) %>% 
-    ungroup() %>% 
-    select(-n)
-
-  # Step 2. Determines the amount of grids present in each country
-  Spp_Grid <- Trans_Spp %>% 
-    group_by(taxon_key,
-             eez_name,
-             year,
-             ensamble) %>% 
-    summarise(n_spp_eez = length(unique(index))) %>% 
-    left_join(Neighbours_List,
-              by = "eez_name") %>% 
-    filter(eez_name %in% Trans_Spp$eez_name, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
-           eez_neighbour %in% Trans_Spp$eez_name)
-
-  # Step 3. Sum total grids per Neighbours
-
-      # Split dataframes to merge latter
-    Territory_T <- Spp_Grid %>% 
-      ungroup() %>% 
-      select(
-        taxon_key,
-        Name=eez_name,
-        n_spp_eez,
-        year,
-        ensamble
-      )
-
-    Neighbour_T <- Spp_Grid %>% 
-      ungroup() %>% 
-      select(
-        taxon_key,
-        n_spp_eez,
-        Name=eez_neighbour,
-        eez_name,
-        year,
-        ensamble
-      )
-
-    # Merge dataframes to get totals per Neighbourds
-    area_index_d <- full_join(Territory_T,
-                              Neighbour_T, 
-                              by = c("Name","taxon_key","year","ensamble")
-    ) %>%
-      rowwise() %>%
-      mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
-      distinct() %>% # Removes false duplicates from `full_join()`
-      rename(eez_name = Name,
-             eez_neighbour =eez_name,
-             n_spp_Country = n_spp_eez.x,
-             n_spp_Neighbour = n_spp_eez.y) %>% 
-      mutate(area_index = n_spp_Country/Spp_Total) %>% 
-      select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
-      mutate(
-        time_step = ifelse(year < 2000,"historical",
-                           ifelse(year >= 2045 & year <= 2064, "mid",
-                                  ifelse(year >= 2080, "n_end","NA")
-                           )
-        )
-      ) %>%
-      # Ensamble mean and sd
-      ungroup() %>% 
-      group_by(taxon_key,year,eez_name,eez_neighbour,time_step) %>% 
-      summarise_at(vars("area_index"),
-                   c(ensamble_mean=mean,
-                     ensamble_sd=sd)
-                   ) %>% 
-      group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% #Running mean
-      mutate(RMean = rollmean(x = ensamble_mean,
-                            5,
-                            align = "right",
-                            fill = ensamble_mean)
-    )
-
-    # Determine if changes
-    natural_variation_d <- area_index_d %>% 
-      ungroup() %>% #remove rowise
-      filter(time_step == "historical") %>% 
-      # Temporal mean
-      group_by(taxon_key,eez_name,eez_neighbour) %>% 
-      summarise_at(vars("ensamble_mean"),
-                   c(hist_mean_index=mean,
-                     hist_sd_index=sd)
-      )
-
-    ### Variation treshold
-    variation_d <- area_index_d %>% 
-      ungroup() %>% #remove rowise
-      # Temporal mean
-      group_by(taxon_key,time_step,eez_name,eez_neighbour) %>% 
-      summarise_at(vars("RMean"),
-                   c(mean_index=mean,
-                     sd_index=sd)
-      ) %>% 
-      gather("variable","value",mean_index,sd_index) %>% 
-      spread(time_step,value) %>% 
-      mutate(larger_variation = ifelse(future > historical,"yes","no"))
-
-    # William's index methodology
-    trans_index_data <- area_index_d %>% 
-      left_join(natural_variation_d,
-                by = c("taxon_key","eez_name","eez_neighbour")) %>% 
-      mutate(d_sd = ensamble_mean/sd_index) %>% 
-      gather("variable","value",ensamble_mean:d_sd) %>% 
-      ungroup() %>% 
-      select(-time_step,-eez_neighbour) %>% 
-      rowid_to_column() %>% 
-      # spread(time_step,value) %>% 
-      spread(eez_name,value) %>% 
-      group_by(taxon_key,year,variable) %>% 
-      summarise_at(vars("Chile","Ecuador","Peru"),
-                   mean,na.rm=T)
-
-    ggplot(area_index_d) +
-      geom_ribbon(data=area_index_d,
-                  aes(
-                    x = as.numeric(year),
-                    ymin=RMean-ensamble_sd,
-                    ymax=RMean+ensamble_sd
-                  ),
-                  fill = "red",
-                  alpha = 0.5
-                  )+
-      geom_line(
-        aes(
-          x = as.numeric(year),
-          y = RMean
-        )
-      )  +
-      facet_wrap(~eez_name + eez_neighbour,
-                 scales = "free")
-
-} # closes function
-
-# __________________________________
-# Test function
-# __________________________________
-
-```
-
-### Fun `EstRangeIndex`
-
-```{r load_data_per_ensamble, eval = F, echo = T}
-
-#### Data needed (send to Control pannel latter)
-
-# All ensambles
-ensamble_list <- c(102,103)
-
-# for dbem_impo_ensamble So we don't miss INDEX in different ensambles
-world_grid <- tibble(index = seq(1,259200,1))
-taxon_key <- 600005 #601477
-
-# SAU relations between INDEX and Country's EEZs
-EEZIDs_List <- read_excel(paste(data_path,"Spatial/V_Lam/Updated_EEZList_17June2016.xlsx",sep="")) %>% 
-  clean_names()
-EEZ_CellID <- read_excel(paste(data_path,"Spatial/V_Lam/EEZ_CellID.xlsx",sep="")) %>% 
-  clean_names()
-colnames(EEZ_CellID) <- c("eezid","index")
-
-Lon_Lat_DBEM <- read.csv("~/GitHub/TransEmergence/Temporal_Data/Spatial/Lon_Lat_DBEM.txt", header=FALSE)
-colnames(Lon_Lat_DBEM) <- c("index","lon","lat")
-
-index_code <- EEZIDs_List %>% 
-  left_join(EEZ_CellID) %>% 
-  rename(eez_name = name) %>% 
-  left_join(Lon_Lat_DBEM,
-            by = "index")
-
-EstRangeIndex <- function(taxon_key,iyr,fyr,ensamble_list){
-
-  spp_data <- dbem_import(
-    taxon_key = taxon_key,
-    iyr <- 1951,
-    fyr <- 2100,
-    data_type <- "Abd",
-    ensamble_list = ensamble_list,
-    Path <- paste(data_path,"DBEM/",sep="")
-  ) %>% 
-    # Remove places where species is not destributed and set to cero differences between ensambles
-    filter(!is.na(value)) %>% 
-    spread(ensamble,value) %>% 
-    gather(ensamble,value, -1:-3) %>% 
-    mutate(value = ifelse(is.na(value),0,value)) %>% 
-  # Include eez information
-    left_join(index_code,
-              by = "index") %>% 
-    filter(!is.na(eezid))
-
-  # Estimate mean eez centroid
-  # centroid_data_eez <-  spp_data %>% 
-  #   ungroup() %>% 
-  #   group_by(taxon_key,ensamble,year,eezid,eez_name) %>% 
-  #   top_n(1,value) %>% 
-  #   group_by(year,taxon_key,ensamble,eezid,eez_name) %>% 
-  #   summarise_at(vars(lon,lat),mean) %>%  # solves for multiple top 1
-  #   rename(eez_centroid_lon = lon,
-  #          eez_centroid_lat = lat) %>% 
-  #   filter(!is.na(eezid))
-
-  centroid_data_eez <-  spp_data %>% 
-    ungroup() %>% 
-    group_by(taxon_key,eezid,eez_name) %>%
-    summarise_at(vars(lon,lat),mean) %>%  # solves for multiple top 1
-    rename(eez_centroid_lon = lon,
-           eez_centroid_lat = lat)
-
-  # Visualization of  eez centroids
-  # ggplot(spp_data) +
-  #   geom_tile(
-  #     aes(
-  #     x = lon,
-  #     y = lat,
-  #     fill= eez_name
-  #     )
-  #   ) +
-  #   geom_point(data = centroid_data_eez,
-  #              aes(
-  #                x = eez_centroid_lon,
-  #                y = eez_centroid_lat,
-  #                size = 1
-  #              )
-  #   )
-
-# Estimate species distribution centroid
-  centroid_spp <-  spp_data %>% 
-    ungroup() %>% 
-    group_by(taxon_key,ensamble,year) %>% 
-    top_n(1,value) %>% 
-    group_by(taxon_key,ensamble,year) %>% 
-    summarise_at(vars(lon,lat),mean) %>%  # solves for multiple top 1
-    rename(spp_centroid_lon = lon,
-           spp_centroid_lat = lat)
-
-  ggplot(centroid_spp) +
-    geom_point(
-      aes(
-        x = spp_centroid_lon,
-        y = spp_centroid_lat,
-        size = as.numeric(year)#,
-        # shape = eez_name
-      )
-    ) +
-    geom_point(data = centroid_data_eez,
-               aes(
-                 x = eez_centroid_lon,
-                 y = eez_centroid_lat,
-                 color = eez_name
-               )
-    ) +
-    facet_wrap(~ensamble)
-
-# Estimate Transboundary Index
-  trans_index_data <- centroid_data_eez %>% 
-    left_join(centroid_spp,
-              by = c("taxon_key") # removed year
-              ) %>% 
-    ungroup() %>% 
-    # select(year,eez_name,ensamble,spp_centroid_lon,spp_centroid_lat,eez_centroid_lon, eez_centroid_lat) %>%
-    rowwise() %>%
-    mutate(d_eez = distm(c(eez_centroid_lon, eez_centroid_lat),c(spp_centroid_lon, spp_centroid_lat), fun = distHaversine))
-
-trans_index_sd_data <- trans_index_data %>% 
-  ungroup() %>% 
-  filter(year <2000) %>% 
-  group_by(eez_name,ensamble) %>% 
-  summarise(
-    s_sd = sd(d_eez)
-  )
-
-final_trans_index_data <- trans_index_data %>% 
-  left_join(trans_index_sd_data,
-            by = c("eez_name","ensamble")
-            ) %>% 
-  mutate(da_dsd = d_eez/s_sd) %>% 
-  filter(eez_name != "Colombia (Pacific)") %>% 
-  select(year,eez_name,ensamble,da_dsd) %>% 
-  spread(eez_name,da_dsd) %>% 
-  mutate(Trans_Index = `Mexico (Atlantic)`-`USA (Gulf of Mexico)`/2) %>% 
-  ungroup() %>% 
-  group_by(ensamble) %>% # Running mean as Kleisner & Pauly 2012
-  mutate(RMean = rollmean(x = Trans_Index,
-                            5,
-                            align = "right",
-                            fill = Trans_Index)
-    )
-
-ggplot(final_trans_index_data) +
-  geom_line(
-    aes(
-      x = as.numeric(year),
-      y = RMean,
-      color = ensamble
-    )
-  ) +
-  geom_line(data = final_trans_index_data,
-    aes(
-      x = as.numeric(year),
-      y = Trans_Index,
-      color = ensamble
-    ),
-    linetype = "dashed",
-    alpha = 0.5
-  ) +
-  facet_wrap(~ensamble,
-             ncol= 1)
-
-  # %>% 
-  #   filter(ensamble == 102) %>% 
-  #   select(year,ensamble,eez_name,d_eez) %>% 
-  #   spread(eez_name,d_eez) %>% 
-  #   mutate()
-
-  Year_MCP <-  spp_data %>% 
-    # Aggregate MCP per EEZ
-    group_by(taxon_key,eezid,eez_name,ensamble,year) %>% 
-    summarise(eez_mcp = sum(value,na.rm=T)) %>% 
-    # Ensamble mean
-    group_by(taxon_key,eezid,eez_name,year) %>% 
-    summarise(ensamble_mean = mean(eez_mcp,na.rm=T),
-              ensamble_sd = sd(eez_mcp,na.rm=T)
-              )
-
-  # Test 
-  # ggplot(spp_data) +
-  #   geom_line(
-  #     aes(
-  #       x = as.numeric(year),
-  #       y = ensamble_mean
-  #     )
-  #   ) +
-  #   geom_point(
-  #     aes(
-  #       x = as.numeric(year),
-  #       y = ensamble_mean+ensamble_sd
-  #     )
-  #   ) +
-  #   geom_point(
-  #     aes(
-  #       x = as.numeric(year),
-  #       y = ensamble_mean-ensamble_sd
-  #     )
-  #   ) +
-  #   facet_wrap(~eez_name)
-
-  } # close function
-
-```
-
-## Control Panel
-
-```{r control_panel, eval = F}
-
-# Data needed for estimating Proportion change
-
-year_df <- tibble(year =  c(seq(1995,2014,1), seq(2041,2060,1), seq(2080,2099,1)),
-                  time_period = c(rep("Today", 20), rep("Mid Century", 20), rep("End of Century", 20))
-                  )
-
-data_type <- "Catch"
-ensamble_list <- seq(102,111,1)
-world_grid <- tibble(index = seq(1,259200,1))
-
-Transboundary_spp <- read.csv(paste(data_path,"Species/Transboundary_spp.csv",sep="")) %>% clean_names()
-
-spp_list <- unique(Transboundary_spp$taxon_key)
-
-```
-
-## Proportion Routine
+#### Proportion routine
 
@@
 
+## Estimating emerging year
+
+#### Fun `TransIndex`
+
+```{r TransIndex, eval = F, echo = T}
+
+TransIndex <- function(Spp,year, data_type = "Catch",ensamble_list,path,Neighbours,eez_centroid){
+
+  # ----------------- #
+  # Get Species Centrodis
+  #  ---------------- #
+
+  # Get model data from spps
+  Spp_Dist <- bind_rows(
+    lapply(
+      ensamble_list,
+      dbem_import,
+      taxon_key = Spp,
+      year = year,
+      data_type = data_type,
+      path = path
+    )
+  )
+
+  print(Spp)
+
+  #____________ Selecting only the transboundary nature of the species _________ #
+  Trans_Spp <- Spp_Dist %>%
+    left_join(index_code,
+              by= "index") %>% 
+    filter(!is.na(value)) %>% 
+    semi_join(Transboundary_spp, # filter only transboundary cases
+              by = c("taxon_key","eez_name")
+    )
+
+  # Step
+  nr <- nrow(Trans_Spp)
+  if(nr > 0){
+
+    # Filter eez by those transboundary
+  unique_eez <- unique(Trans_Spp$eez_name)
+
+  Neighbours_combo <- Neighbours %>% 
+    filter(eez_name %in% unique_eez,
+           eez_neighbour %in% unique_eez)
+
+    # Get the centroid of each country
+    EEZ_centroid <- eez_centroid %>% 
+      filter(eez_name %in% Trans_Spp$eez_name) %>% 
+      mutate(taxon_key = Spp)
+
+    Ids <- unique(Neighbours_combo$neighbour_id)
+
+    for(i in 1:length(Ids)){
+
+      EEZ_ids <- Neighbours_combo %>% 
+        filter(neighbour_id == Ids[i])
+
+      Sub_Trans_Spp <- Trans_Spp %>% 
+        filter(eez_name %in% EEZ_ids$eez_name)
+
+      Sub_EEZ_centroid <- EEZ_centroid %>% 
+        filter(eez_name %in% EEZ_ids$eez_name)
+
+      # Centroid is defined as the tile with the higest species value
+      Centroids <- Sub_Trans_Spp %>% 
+        left_join(EEZ_ids, by = "eez_name") %>% 
+        group_by(taxon_key,year,ensamble,neighbour_id) %>%  # for whole distribution
+        top_n(1,value) %>% 
+        distinct(value,year, .keep_all = TRUE) %>%  # removes duplicates when tile is in between EEZs
+        left_join(Sub_EEZ_centroid, by = "taxon_key") %>%
+        select(eez_name =eez_name.y, everything() ,-eez_name.x)
+
+      # Estimate distances between spp centroid and EEZ centroid
+      Distance_Data <- Centroids %>%
+        rename(
+          catch_lon = lon,
+          catch_lat = lat,
+          eez_lon = longitude,
+          eez_lat= latitude
+        ) %>% 
+        # Estimate the distance between centriods
+        mutate(
+          distance = geodist(eez_lon, eez_lat, catch_lon, catch_lat, units="km")
+        ) %>% 
+        select(taxon_key,eez_name,neighbour_id,ensamble,year,distance)
+
+      # Historical mean and sd of distance
+      historical <- Distance_Data %>%
+        filter(year <= 2000) %>%
+        ungroup() %>%
+        group_by(taxon_key,ensamble,eez_name,neighbour_id) %>%
+        summarise_at(vars("distance"),
+                     c(hd_temp_mean=mean,
+                       hd_temp_sd=sd)
+        )
+
+      # Future results
+      future <- Distance_Data %>% 
+        filter(year > 2000)
+
+      # Transboundary Index
+      Trans_index <- Distance_Data %>% 
+        left_join(historical,
+                  by = c("taxon_key","eez_name","ensamble","neighbour_id")
+        ) %>%
+        # Re arrenging table for mutate
+        gather("variable","value",distance:hd_temp_sd) %>%
+        ungroup()
+
+      Partial <- Trans_index %>% 
+        mutate(
+          names = ifelse(eez_name ==Sub_EEZ_centroid$eez_name[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_"))
+        ) %>%
+        select(-eez_name,-variable) %>%
+        spread(names,value) %>% 
+        # mutate index
+        mutate(
+          trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
+        ) %>% 
+        group_by(taxon_key,ensamble,neighbour_id) %>% 
+        mutate(RMean = rollmean(x = trans_index, 
+                                10, 
+                                align = "right", 
+                                fill = trans_index)
+        ) %>% 
+        filter(year > 1960) %>% 
+        select(taxon_key,neighbour_id,year,ensamble,trans_index,RMean)
+
+      # Average of 10 years mean of ensamble members results
+      mean_sd_index <- Partial %>% 
+        group_by(taxon_key,year,neighbour_id) %>% 
+        summarise_at(vars("RMean"),
+                     c(trend=mean,
+                       ensamble_sd=sd)
+        )
+
+      # Estimate noice from ensamble members
+      historic_sd <- mean_sd_index %>% 
+        filter(year < 2000) %>% 
+        group_by(taxon_key) %>% 
+        summarise_at(vars("trend"),
+                     c(hist_mean=mean,
+                       noise=sd)
+        )
+
+      # Plot the result, bruh
+      # ggplot() +
+      #   geom_line(data = mean_sd_index,
+      #             aes(
+      #               x = year,
+      #               y = trend
+      #             )
+      #   ) +
+      #   geom_ribbon(data = mean_sd_index,
+      #               aes(x=year,
+      #                   ymin=historic_sd$hist_mean-historic_sd$noise,
+      #                   ymax=historic_sd$hist_mean+historic_sd$noise),
+      #               alpha = 0.5,
+      #               fill = "grey50"
+      #   ) +
+      #   geom_line(data = mean_sd_index,
+      #             aes(
+      #               x = year,
+      #               y = historic_sd$hist_mean
+      #             ),
+      #             colour = "grey50"
+      #   ) +
+      #   geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
+      #              aes(
+      #                x = year,
+      #                y = trend,
+      #                colour = "red",
+      #              ),
+      #              shape = 1,
+      #              size = 3
+      #   ) +
+      #   geom_vline(xintercept = 2000,colour="blue",alpha = 0.5) +
+      #   theme_classic()
+
+      # #     
+      Final_Result <- mean_sd_index %>%
+        filter(year > 2000,
+               trend > (historic_sd$hist_mean+historic_sd$noise) |
+                 trend < (historic_sd$hist_mean-historic_sd$noise)
+        ) %>%
+        group_by(taxon_key,neighbour_id) %>%
+        summarise(emerging_yr= min(year)) %>%
+        left_join(EEZ_ids) %>%
+        ungroup() %>%
+        select(taxon_key,eez_name,eez_neighbour,emerging_yr)
+
+      if(i == 1){
+        f_df <- Final_Result
+      }else{
+        f_df <- bind_rows(f_df,Final_Result)
+      }
+    }
+
+    return(f_df)
+
+  }else{
+
+    print("no share data")
+
+    f_df <- tibble(
+      taxon_key = Spp,
+      eez_name = NA,
+      emerging_yr = NA
+    )
+
+  }
+
+  return(f_df)
+
+}
+
+#  #One spp works 
+# suppressMessages(
+# x <-  TransIndex(Spp = 600245,
+#              year,
+#              data_type = "Catch",
+#              ensamble_list,
+#              path,
+#              Neighbours = Neighbours,
+#              eez_centroid)
+# )
+# Multiple spp
+
+# spp_list <- c(600004,
+#               600005,
+#               600245,
+#               600006
+#               )
+# 
+# suppressMessages(
+#   suppressWarnings(
+#     Test <- bind_rows(
+#       lapply(spp_list,
+#              TransIndex,
+#              year,
+#              data_type = "Catch",
+#              ensamble_list,
+#              path,
+#              Neighbours,
+#              eez_centroid
+#       )
+#     )
+#   )
+# )
+
+```
+
+#### Emerging Control Panel
+
+```{r Emerging_control_panel, eval = F}
+
+# Data needed for estimating Proportion change
+
+# SAU relations between INDEX and Country's EEZs
+EEZIDs_List <- read_excel(paste(data_path,"Spatial/V_Lam/Updated_EEZList_17June2016.xlsx",sep="")) %>% 
+  clean_names()
+EEZ_CellID <- read_excel(paste(data_path,"Spatial/V_Lam/EEZ_CellID.xlsx",sep="")) %>% 
+  clean_names()
+colnames(EEZ_CellID) <- c("eezid","index")
+
+Lon_Lat_DBEM <- read.csv("~/GitHub/TransEmergence/Temporal_Data/Spatial/Lon_Lat_DBEM.txt", header=FALSE)
+colnames(Lon_Lat_DBEM) <- c("index","lon","lat")
+
+index_code <- EEZIDs_List %>% 
+  left_join(EEZ_CellID) %>% 
+  rename(eez_name = name) %>% 
+  left_join(Lon_Lat_DBEM,
+            by = "index")
+
+year_df <- tibble(year =  c(seq(1951,2000,1), seq(2001,2099,1)),
+                  time_period = c(rep("historic", 50), rep("future", 99))
+                  )
+
+data_type <- "catch"
+ensamble_list <- seq(102,103,1)
+world_grid <- tibble(index = seq(1,259200,1))
+path = DBEM_path
+
+eez_centroid <- read.csv("~/GitHub/TransEmergence/Temporal_Data/Spatial/eez_centroids.csv", header=TRUE)
+
+Neighbours <- read.csv(paste(data_path,"Spatial/Neighbours_eez_id.csv",sep="")) %>% clean_names()
+
+Transboundary_spp <- read.csv(paste(data_path,"Species/Transboundary_spp.csv",sep="")) %>% 
+  clean_names()
+
+```
+
+#### Emerging Routine
+```{r Emerging_routine, eval = F, echo = F}
+
+####_______________________ ####
+# Estimating year of eergence
+####_______________________ ####
+
+spp_list <- c(600004,
+              600005,
+              600245,
+              600006
+              )
+
+t_start <- Sys.time()
+
+suppressMessages(
+  suppressWarnings(
+    Test <- bind_rows(
+      lapply(spp_list,
+             TransIndex,
+             year,
+             data_type = "Catch",
+             ensamble_list,
+             path,
+             Neighbours,
+             eez_centroid
+      )
+    )
+  )
+)
+
+t_end <- Sys.time()-t_start;t_end
+
+```
+
 # Results
@@
 
+coords <- read.csv(coords)
+
 ```
@@
 # http://jsmayorga.com/post/mapping-the-global-network-of-transnational-fisheries/
-
-# Get the center poligon of each EEZ for map source/target
-coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
-  ungroup() %>% 
-  select(eez_name,geometry) %>% 
-  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
-  # Manually fix issues
-  mutate(
-    longitude = ifelse(
-      eez_name == "Fiji", 178.065033,
-      ifelse(eez_name == "Tuvalu",177.6493,
-             ifelse( eez_name == "New Zealand",174.8860,
-                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,
-                             ifelse( eez_name == "Brazil",-36.937,longitude
-                             )
-                     )
-             )
-      )
-    )
-  )
-
-# Remove "c()"
-coords$longitude <- gsub("\\(","",coords$longitude)
-coords$longitude <- gsub("c","",coords$longitude)
-coords$latitude <- gsub(")","",coords$latitude)
-
-coords <- coords %>% 
-  mutate_at(vars(longitude,latitude),as.numeric)
 

### 2020-04-15 Scripts/Initial_Analysis.Rmd
Function working (mechanically) in carmelia

@@
 
-    for(i in 1:length(Ids)){
+  corefx <- function(Ids){
 
       EEZ_ids <- Neighbours_combo %>% 
-        filter(neighbour_id == Ids[i])
+        filter(neighbour_id == Ids)
 
@@
 
-      if(i == 1){
-        f_df <- Final_Result
+      # if(i == 1){
+      #   f_df <- Final_Result
+      # }else{
+      #   f_df <- bind_rows(f_df,Final_Result)
+      # }
+    return(Final_Result)
+      }
+
+  x <- bind_rows(
+    mclapply(Ids,corefx)
+  )
+    # return(f_df)
+
   }else{
-        f_df <- bind_rows(f_df,Final_Result)
-      }
-    }
-
-    return(f_df)
-
-  }else{
 
@@
 
-    f_df <- tibble(
+    x <- tibble(
       taxon_key = Spp,
@@
 
-  return(f_df)
+  return(x)
 
@@
 # suppressMessages(
-# x <-  TransIndex(Spp = 600245,
+# TransIndex(Spp = 600006,
 #              year,
@@
 #              eez_centroid)
-# )
 # Multiple spp
@@
               600005,
-              600245,
-              600006
+              600245#,
+              # 600006
               )
@@
     Test <- bind_rows(
-      lapply(spp_list,
+      mclapply(spp_list,
              TransIndex,
@@
 t_end <- Sys.time()-t_start;t_end
+
+# Version 1 (embeded loop) <- 3.01
+# Version 2 (embeded lapply) <- 2.71
+# Version 3 (embeded mclapply) <- 1.597311 mins
+# Version 4 (embeded mclapply) + mclapply in spp <- 1.597311 mins
 

### 2020-04-15 Scripts/Initial_Analysis.Rmd
fixed paths on emergence cont.pan

@@
 
-Lon_Lat_DBEM <- read.csv("~/GitHub/TransEmergence/Temporal_Data/Spatial/Lon_Lat_DBEM.txt", header=FALSE)
+Lon_Lat_DBEM <- read.csv(paste(data_path,"Spatial/Lon_Lat_DBEM.txt",sep=""), header=FALSE)
 colnames(Lon_Lat_DBEM) <- c("index","lon","lat")
@@
 
-eez_centroid <- read.csv("~/GitHub/TransEmergence/Temporal_Data/Spatial/eez_centroids.csv", header=TRUE)
+eez_centroid <- read.csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""), header=TRUE)
 

### 2020-04-15 Scripts/Initial_Analysis.Rmd
Included reporting to fix specief species errors

@@
 
-  print(Spp)
+  #### Control for Spp id so I can fix problems
+
+if(file.exists("~/Desktop/spp_df.csv") == FALSE){
+  spp_df <- tibble() %>% 
+    write_csv(.,
+            "~/Desktop/spp_df.csv")
+}
+
+  suppressMessages(
+    spp_df <- read_csv("~/Desktop/spp_df.csv")
+  )
+  spp_df <- tibble(Spp) %>% 
+    bind_rows(spp_df) %>% 
+    write_csv(.,
+              "~/Desktop/spp_df.csv")
 
@@
 
+spp_list <- unique(Transboundary_spp$taxon_key)
+
 ```
@@
               600005,
-              600245#,
-              # 600006
+              600245,
+              600006
               )

### 2020-04-16 Scripts/Initial_Analysis.Rmd
Modified function to see if it uses less memory

@@
 
-#### Fun `TransIndex`
-
-```{r TransIndex, eval = F, echo = T}
-
-TransIndex <- function(Spp,year, data_type = "Catch",ensamble_list,path,Neighbours,eez_centroid){
+#### Fun `GetCentroids`
+
+```{r GetCentroids, eval = F, echo = T}
+
+GetCentroids <- function(taxon_key,year, data_type = "Catch", ensamble, path, Neighbours, eez_centroid){
 
@@
   # Get model data from spps
-  Spp_Dist <- bind_rows(
-    lapply(
-      ensamble_list,
-      dbem_import,
-      taxon_key = Spp,
+  Spp_Dist <- dbem_import(
+    taxon_key = taxon_key,
     year = year,
-      data_type = data_type,
+    data_type = "Catch",
+    ensamble = ensamble,
     path = path
-    )
   )
@@
   )
-  spp_df <- tibble(Spp) %>% 
+  spp_df <- tibble(taxon_key,
+                   ensamble) %>% 
     bind_rows(spp_df) %>% 
@@
       filter(eez_name %in% Trans_Spp$eez_name) %>% 
-      mutate(taxon_key = Spp)
-
-    Ids <- unique(Neighbours_combo$neighbour_id)
-
-  corefx <- function(Ids){
+      mutate(taxon_key = taxon_key)
+
+    CentroidsFx <- function(Ids){
 
@@
 
+      return(Distance_Data)
+
+    }
+
+    Ids <- unique(Neighbours_combo$neighbour_id)
+
+    Centroids_Data <- bind_rows(
+      lapply(Ids,CentroidsFx)
+    )
+
+  }else{
+
+    Centroids_Data <- tibble()
+
+  }
+
+  return(Centroids_Data)
+
+}
+
+```
+
+#### Fun `TransIndex`
+
+```{r TransIndex, eval = F, echo = T}
+
+TransIndex <- function(Centroids,Ids){
+
+  Centroids_id <- Centroids %>% 
+    filter(neighbour_id %in% Ids)
+
+  EEZ_ids <- Neighbours %>% 
+    filter(neighbour_id %in% Ids)
+
   # Historical mean and sd of distance
-      historical <- Distance_Data %>%
+  historical <- Centroids_id %>%
     filter(year <= 2000) %>% 
@@
   # Future results
-      future <- Distance_Data %>% 
+  future <- Centroids_id %>% 
     filter(year > 2000)
@@
   # Transboundary Index
-      Trans_index <- Distance_Data %>% 
+  Trans_index <- Centroids_id %>% 
     left_join(historical,
@@
 
+  Unique_EEZs <- unique(Trans_index$eez_name)
+
   Partial <- Trans_index %>% 
     mutate(
-          names = ifelse(eez_name ==Sub_EEZ_centroid$eez_name[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_"))
+      names = ifelse(eez_name ==Unique_EEZs[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_"))
     ) %>%
@@
 
-      # if(i == 1){
-      #   f_df <- Final_Result
-      # }else{
-      #   f_df <- bind_rows(f_df,Final_Result)
-      # }
   return(Final_Result)
@@
 
-  x <- bind_rows(
-    mclapply(Ids,corefx)
-  )
-    # return(f_df)
-
-  }else{
-
-    print("no share data")
-
-    x <- tibble(
-      taxon_key = Spp,
-      eez_name = NA,
-      emerging_yr = NA
-    )
-
-  }
-
-  return(x)
-
-}
-
-#  #One spp works 
-# suppressMessages(
-# TransIndex(Spp = 600006,
-#              year,
+# test <- bind_rows(
+#   lapply(ensamble_list,
+#          GetCentroids,
+#          taxon_key = 600245,
+#          year = year,
 #          data_type = "Catch",
-#              ensamble_list,
-#              path,
+#          path = path,
 #          Neighbours = Neighbours,
-#              eez_centroid)
-# Multiple spp
-
-# spp_list <- c(600004,
-#               600005,
-#               600245,
-#               600006
+#          eez_centroid = eez_centroid)
 # )
 # 
-# suppressMessages(
-#   suppressWarnings(
-#     Test <- bind_rows(
+# 
+#   TransIndex(test,819)
+#   
+#   bind_rows(
+#   lapply(Ids,
+#          TransIndex,
+#          Centroid = test)
+#   )
+
+```
+
+#### Fun `GetTransIndex`
+
+```{r GetTransIndex, eval = F, echo = T}
+
+GetTransIndex <- function(taxon_key,
+                            year,
+                            data_type = "Catch",
+                            path,
+                            ensamble_list,
+                            Neighbours,
+                            eez_centroid){
+
+    # First step
+    # Call Get centroids fucntion for centroids data per ensamble
+    Step_one <- bind_rows(
+      lapply(ensamble_list,
+             GetCentroids,
+             taxon_key = taxon_key,
+             year = year,
+             data_type = "Catch",
+             path = path,
+             Neighbours = Neighbours,
+             eez_centroid = eez_centroid)
+    )
+
+    Ids <- unique(Step_one$neighbour_id)
+
+    # Second Step
+    # Get the transboudary index based on all centroids
+    Step_two <- bind_rows(
+      lapply(Ids,
+             TransIndex,
+             Centroid = Step_one)
+    )
+
+    return(Step_two)
+  }
+
+  # GetTransIndex(
+  #   taxon_key = 600004,
+  #   ensamble_list = ensamble_list,
+  #   year = year,
+  #   data_type = "Catch",
+  #   path = path,
+  #   Neighbours = Neighbours,
+  #   eez_centroid = eez_centroid
+  # )
+  # 
+  # 
+  # spp_list <- c(600004, 600245)
+  # 
+  #  test <- bind_rows(
   #   lapply(spp_list,
-#              TransIndex,
-#              year,
+  #          GetTransIndex,
+  #          year = year,
   #          data_type = "Catch",
-#              ensamble_list,
-#              path,
-#              Neighbours,
-#              eez_centroid
+  #          path = path,
+  #          Neighbours = Neighbours,
+  #          eez_centroid = eez_centroid,
+  #          ensamble_list = ensamble_list
   #   )
   # )
-#   )
-# )
+  #
 
@@
 data_type <- "catch"
-ensamble_list <- seq(102,103,1)
+# ensamble_list <- seq(102,103,1)
+ensamble_list <- seq(102,111,1)
+
 world_grid <- tibble(index = seq(1,259200,1))
@@
 
-spp_list <- unique(Transboundary_spp$taxon_key)
+spp_list <- unique(Transboundary_spp$taxon_key)[1:2]
 
@@
 #### Emerging Routine
+
 ```{r Emerging_routine, eval = F, echo = F}
@@
 ####_______________________ ####
-
-spp_list <- c(600004,
-              600005,
-              600245,
-              600006
-              )
 
@@
       mclapply(spp_list,
-             TransIndex,
-             year,
+               GetTransIndex,
+               year = year,
                data_type = "Catch",
-             ensamble_list,
-             path,
-             Neighbours,
-             eez_centroid
+               ensamble_list = ensamble_list,
+               path = path,
+               Neighbours = Neighbours,
+               eez_centroid = eez_centroid
       )
@@
 t_end <- Sys.time()-t_start;t_end
+
+write_csv(Test,
+          "~/Desktop/Test.csv")
 
@@
 ```
+
+# Old Code
+
+```{r TransIndex, eval = F, echo = T}
+
+TransIndex <- function(Spp,year, data_type = "Catch",ensamble_list,path,Neighbours,eez_centroid){
+
+  # ----------------- #
+  # Get Species Centrodis
+  #  ---------------- #
+
+  # Get model data from spps
+  Spp_Dist <- bind_rows(
+    lapply(
+      ensamble_list,
+      dbem_import,
+      taxon_key = Spp,
+      year = year,
+      data_type = data_type,
+      path = path
+    )
+  )
+
+  #### Control for Spp id so I can fix problems
+
+if(file.exists("~/Desktop/spp_df.csv") == FALSE){
+  spp_df <- tibble() %>% 
+    write_csv(.,
+            "~/Desktop/spp_df.csv")
+}
+
+  suppressMessages(
+    spp_df <- read_csv("~/Desktop/spp_df.csv")
+  )
+  spp_df <- tibble(Spp) %>% 
+    bind_rows(spp_df) %>% 
+    write_csv(.,
+              "~/Desktop/spp_df.csv")
+
+  #____________ Selecting only the transboundary nature of the species _________ #
+  Trans_Spp <- Spp_Dist %>%
+    left_join(index_code,
+              by= "index") %>% 
+    filter(!is.na(value)) %>% 
+    semi_join(Transboundary_spp, # filter only transboundary cases
+              by = c("taxon_key","eez_name")
+    )
+
+  # Step
+  nr <- nrow(Trans_Spp)
+  if(nr > 0){
+
+    # Filter eez by those transboundary
+  unique_eez <- unique(Trans_Spp$eez_name)
+
+  Neighbours_combo <- Neighbours %>% 
+    filter(eez_name %in% unique_eez,
+           eez_neighbour %in% unique_eez)
+
+    # Get the centroid of each country
+    EEZ_centroid <- eez_centroid %>% 
+      filter(eez_name %in% Trans_Spp$eez_name) %>% 
+      mutate(taxon_key = Spp)
+
+    Ids <- unique(Neighbours_combo$neighbour_id)
+
+  corefx <- function(Ids){
+
+      EEZ_ids <- Neighbours_combo %>% 
+        filter(neighbour_id == Ids)
+
+      Sub_Trans_Spp <- Trans_Spp %>% 
+        filter(eez_name %in% EEZ_ids$eez_name)
+
+      Sub_EEZ_centroid <- EEZ_centroid %>% 
+        filter(eez_name %in% EEZ_ids$eez_name)
+
+      # Centroid is defined as the tile with the higest species value
+      Centroids <- Sub_Trans_Spp %>% 
+        left_join(EEZ_ids, by = "eez_name") %>% 
+        group_by(taxon_key,year,ensamble,neighbour_id) %>%  # for whole distribution
+        top_n(1,value) %>% 
+        distinct(value,year, .keep_all = TRUE) %>%  # removes duplicates when tile is in between EEZs
+        left_join(Sub_EEZ_centroid, by = "taxon_key") %>%
+        select(eez_name =eez_name.y, everything() ,-eez_name.x)
+
+      # Estimate distances between spp centroid and EEZ centroid
+      Distance_Data <- Centroids %>%
+        rename(
+          catch_lon = lon,
+          catch_lat = lat,
+          eez_lon = longitude,
+          eez_lat= latitude
+        ) %>% 
+        # Estimate the distance between centriods
+        mutate(
+          distance = geodist(eez_lon, eez_lat, catch_lon, catch_lat, units="km")
+        ) %>% 
+        select(taxon_key,eez_name,neighbour_id,ensamble,year,distance)
+
+      # Historical mean and sd of distance
+      historical <- Distance_Data %>%
+        filter(year <= 2000) %>%
+        ungroup() %>%
+        group_by(taxon_key,ensamble,eez_name,neighbour_id) %>%
+        summarise_at(vars("distance"),
+                     c(hd_temp_mean=mean,
+                       hd_temp_sd=sd)
+        )
+
+      # Future results
+      future <- Distance_Data %>% 
+        filter(year > 2000)
+
+      # Transboundary Index
+      Trans_index <- Distance_Data %>% 
+        left_join(historical,
+                  by = c("taxon_key","eez_name","ensamble","neighbour_id")
+        ) %>%
+        # Re arrenging table for mutate
+        gather("variable","value",distance:hd_temp_sd) %>%
+        ungroup()
+
+      Partial <- Trans_index %>% 
+        mutate(
+          names = ifelse(eez_name ==Sub_EEZ_centroid$eez_name[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_"))
+        ) %>%
+        select(-eez_name,-variable) %>%
+        spread(names,value) %>% 
+        # mutate index
+        mutate(
+          trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
+        ) %>% 
+        group_by(taxon_key,ensamble,neighbour_id) %>% 
+        mutate(RMean = rollmean(x = trans_index, 
+                                10, 
+                                align = "right", 
+                                fill = trans_index)
+        ) %>% 
+        filter(year > 1960) %>% 
+        select(taxon_key,neighbour_id,year,ensamble,trans_index,RMean)
+
+      # Average of 10 years mean of ensamble members results
+      mean_sd_index <- Partial %>% 
+        group_by(taxon_key,year,neighbour_id) %>% 
+        summarise_at(vars("RMean"),
+                     c(trend=mean,
+                       ensamble_sd=sd)
+        )
+
+      # Estimate noice from ensamble members
+      historic_sd <- mean_sd_index %>% 
+        filter(year < 2000) %>% 
+        group_by(taxon_key) %>% 
+        summarise_at(vars("trend"),
+                     c(hist_mean=mean,
+                       noise=sd)
+        )
+
+      # Plot the result, bruh
+      # ggplot() +
+      #   geom_line(data = mean_sd_index,
+      #             aes(
+      #               x = year,
+      #               y = trend
+      #             )
+      #   ) +
+      #   geom_ribbon(data = mean_sd_index,
+      #               aes(x=year,
+      #                   ymin=historic_sd$hist_mean-historic_sd$noise,
+      #                   ymax=historic_sd$hist_mean+historic_sd$noise),
+      #               alpha = 0.5,
+      #               fill = "grey50"
+      #   ) +
+      #   geom_line(data = mean_sd_index,
+      #             aes(
+      #               x = year,
+      #               y = historic_sd$hist_mean
+      #             ),
+      #             colour = "grey50"
+      #   ) +
+      #   geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
+      #              aes(
+      #                x = year,
+      #                y = trend,
+      #                colour = "red",
+      #              ),
+      #              shape = 1,
+      #              size = 3
+      #   ) +
+      #   geom_vline(xintercept = 2000,colour="blue",alpha = 0.5) +
+      #   theme_classic()
+
+      # #     
+      Final_Result <- mean_sd_index %>%
+        filter(year > 2000,
+               trend > (historic_sd$hist_mean+historic_sd$noise) |
+                 trend < (historic_sd$hist_mean-historic_sd$noise)
+        ) %>%
+        group_by(taxon_key,neighbour_id) %>%
+        summarise(emerging_yr= min(year)) %>%
+        left_join(EEZ_ids) %>%
+        ungroup() %>%
+        select(taxon_key,eez_name,eez_neighbour,emerging_yr)
+
+      # if(i == 1){
+      #   f_df <- Final_Result
+      # }else{
+      #   f_df <- bind_rows(f_df,Final_Result)
+      # }
+    return(Final_Result)
+      }
+
+  x <- bind_rows(
+    mclapply(Ids,corefx)
+  )
+    # return(f_df)
+
+  }else{
+
+    print("no share data")
+
+    x <- tibble(
+      taxon_key = Spp,
+      eez_name = NA,
+      emerging_yr = NA
+    )
+
+  }
+
+  return(x)
+
+}
+
+#  #One spp works 
+# suppressMessages(
+# TransIndex(Spp = 600006,
+#              year,
+#              data_type = "Catch",
+#              ensamble_list,
+#              path,
+#              Neighbours = Neighbours,
+#              eez_centroid)
+# Multiple spp
+
+# spp_list <- c(600004,
+#               600005,
+#               600245,
+#               600006
+#               )
+# 
+# suppressMessages(
+#   suppressWarnings(
+#     Test <- bind_rows(
+#       lapply(spp_list,
+#              TransIndex,
+#              year,
+#              data_type = "Catch",
+#              ensamble_list,
+#              path,
+#              Neighbours,
+#              eez_centroid
+#       )
+#     )
+#   )
+# )
+
+```

### 2020-04-20 Scripts/Initial_Analysis.Rmd
Small changes to the code

@@
 
+This function estimates the distance between the centroid of each neighbouring EEZ and the distribution of the species along both EEZs.
+We define species centroid as the tile with the highest value of maimum catch potential from all of the species distribution in each year.
+
 ```{r GetCentroids, eval = F, echo = T}
@@
   # ----------------- #
-  # Get Species Centrodis
+  # Get model data from spps for one ensamble
   #  ---------------- #
 
-  # Get model data from spps
   Spp_Dist <- dbem_import(
@@
 
-  #### Control for Spp id so I can fix problems
-
+  # ----------------- #
+  # Extra step 
+  # ----------------- #
+
+  #### Creates a data to control computational time
   if(file.exists("~/Desktop/spp_df.csv") == FALSE){
@@
 
-  #____________ Selecting only the transboundary nature of the species _________ #
+  # ----------------- #
+  # Estimate distance between centroids
+  # ----------------- #
+
+  # Filter only the transboundary nature of the species 
   Trans_Spp <- Spp_Dist %>%
@@
 
-  # Step
+  # ----------------- #
+  # Extra step 
+  # ----------------- #
+  # Controls for cases with no-data
   nr <- nrow(Trans_Spp)
@@
 
+    # ----------------- #
+    # Centroid function
+    # ----------------- #
+
+    # This function estimates the distance of each neighbouring EEZ
+    # It uses Ids wich is a pre-determine identifier of neighbouring
+    # EEZs. Fo example Peru and Chile have Id = 167, and viceversa
+
     CentroidsFx <- function(Ids){
 
+      # Filter data by Ids
       EEZ_ids <- Neighbours_combo %>% 
@@
         left_join(Sub_EEZ_centroid, by = "taxon_key") %>%
-        select(eez_name =eez_name.y, everything() ,-eez_name.x)
-
+        select(eez_name =eez_name.y, everything() ,-eez_name.x) %>% 
         # Estimate distances between spp centroid and EEZ centroid
-      Distance_Data <- Centroids %>%
-        rename(
-          catch_lon = lon,
-          catch_lat = lat,
-          eez_lon = longitude,
-          eez_lat= latitude
-        ) %>% 
-        # Estimate the distance between centriods
         mutate(
-          distance = geodist(eez_lon, eez_lat, catch_lon, catch_lat, units="km")
+          distance = geodist(longitude, # eez longitude
+                             latitude,  # eez latitude
+                             lon,  # species longitude
+                             lat, # species latitude
+                             units="km")
         ) %>% 
@@
 
-      return(Distance_Data)
-
-    }
-
+      return(Centroids)
+
+    }
+
+    # Determine the ids
     Ids <- unique(Neighbours_combo$neighbour_id)
 
+    # Run function
     Centroids_Data <- bind_rows(
@@
 
+  # Filter data by neighbouring ID
   Centroids_id <- Centroids %>% 

### 2020-04-20 Scripts/Initial_Analysis.Rmd
Updates to code

@@
 output:
-  word_document: default
+  word_document: 
+    keep_md: yes
   html_document: default
@@
 
-```{r setup, eval = T, echo=F, warning=F,message=F, results='hide'}
+```{r setup, eval = T, echo = F, warning = F, message = F, results = 'hide'}
 
@@
 
-#### Library ####
+#### Project's Library
 packages <- c(
@@
   "ggrepel",
-  "zoo",
+  "zoo", # for average mean
   "parallel", # for mclapply,
@@
 
+# standarize maps
 ggtheme_map <- function(base_size = 9, Region = "NA") {
@@
 
+See methods section of the paper (e.g. Manusript.Rmd)
+
+## Global functions
+
+Here we just run functions that will be used multiple times douring the analysis
+
 ### Fun `dbem_import`
@@
 
-```{r dbem_import, eval = F, echo = T}
+```{r dbem_import, eval = T, echo = T}
 
@@
 
-## Determine Coords centroids and Neighbouring ids
-
-```{r, get_eez_centroids, eval = F}
+## Creating Data
+
+Here we re-arrange existing data for the analysis. We have three steps:
+
+1- Set the SAU relations between index and EEZs and include lat long data for each index
+
+2.- Deteremine the centroid of each EEZ and asigned an id to each neighbouring interaction (e.g. Chile and Peru or Canada West Coast and US West Coast)
+
+3.- Set Neigbouring IDs based on the interactions between neighbouring EEZs for all of those sharing transboundary species
+
+```{r pre_data_creation, eval = F, echo = F}
+
+## ---------------##
+#### 1. SAU relations between INDEX and Country's EEZs
+## ---------------##
+
+EEZIDs_List <- read_excel(paste(data_path,"Spatial/V_Lam/Updated_EEZList_17June2016.xlsx",sep="")) %>% 
+  clean_names()
+EEZ_CellID <- read_excel(paste(data_path,"Spatial/V_Lam/EEZ_CellID.xlsx",sep="")) %>% 
+  clean_names()
+colnames(EEZ_CellID) <- c("eezid","index")
+
+Lon_Lat_DBEM <- read.csv(paste(data_path,"Spatial/Lon_Lat_DBEM.txt",sep=""), header=FALSE)
+colnames(Lon_Lat_DBEM) <- c("index","lon","lat")
+
+index_code <- EEZIDs_List %>% 
+  left_join(EEZ_CellID) %>% 
+  rename(eez_name = name) %>% 
+  left_join(Lon_Lat_DBEM,
+            by = "index")
+
+write_csv(index_code,
+          "sau_index_code.csv")
+
+## ---------------##
+#### 2. entroids 
+## ---------------##
 
@@
 
-write_csv(coords,
-          "eez_centroids.csv")
+# Save dataframe so we don't need to run this ewvery time
+# write_csv(coords,
+#           "eez_centroids.csv")
+
+## ---------------##
+#### 3. Set Neigbouring IDs
+## ---------------##
+
+# Read transboundary species database from Palacios-Abrantes et al 2020 (FishForVisa)
+Transboundary_spp <- read.csv(paste(data_path,"Species/Transboundary_spp.csv",sep="")) %>% 
+  clean_names()
 
@@
 
-#  set ids 
+# Split dataframes to merge neighbouring ids
 df1 <- Neighbours_spp
@@
 
+# Merge neighbouring ids
 df_full <- bind_rows(df1, df2) %>% 
@@
 
-# write_csv(df_full, "Neighbours_eez_id.csv")
+# Save dataframe so we don't need to run this every time
+# write_csv(df_full, 
+#           "Neighbours_eez_id.csv")
 
@@
 
-#### Fun `EstPropChange`
+This part of the analysis estimates the Stock Share Ratio and its change under cliamte change under two time frames; early (2030) and mid (2050).
+Note that 2020 is the average of 2021 to 2040 and mid is the average of 2041 to 2060.
+
+### Fun `EstPropChange`
 
@@
                           year,
-                          data_type,
+                          hist_y,
+                          data_type = "Catch",
                           ensamble_list, # expects a list of ensambles to averrage by
@@
   # Loads species data for all ensambles
-  spp_dist <- bind_rows(
+  Spp_Dist <- bind_rows(
     lapply(
@@
 
-  # head(spp_dist)
-
-  # Result 1. Number of Countries that share the species
+  # Filter Species distribution to that of the neighbouring EEZs
+  # ------------ #
+  # Warning message:
+  # Column `eez_name` joining factor and character vector, coercing into character vector 
+# ------------ #
 
   suppressWarnings(
-    trans_EEZ <- Transboundary_spp %>% 
-      filter(taxon_key %in% spp_dist$taxon_key) %>% 
-      left_join(Neighbours,
+    Trans_EEZ <- Transboundary_spp %>% 
+      filter(taxon_key %in% Spp_Dist$taxon_key) %>% 
+      left_join(index_code,
+                by = c("eez_name")
+      )
+    )
+
+  # Get neighbouring id
+  Neighbours_List <- Neighbours %>% 
+    semi_join(Trans_EEZ,
               by = c("eez_name","eez_neighbour")
     )
-  )
-
-  # List of indexes per EEZ
-  index_id <- trans_EEZ %>% 
-    group_by(eez_name,index) %>% 
-    summarise(n=n()) %>% 
-    select(-n)
-
-  # --------------- #
-  # Warning messages:
-  # 1: Column `eez_name` joining factor and character vector, coercing into character vector 
-  # 2: Column `eez_neighbour` joining factor and character vector, coercing into character vector 
-  # --------------- #
-
-  #____________ Selecting only the transboundary nature of the species _________ #
-  trans_spp <- spp_dist %>%
-    filter(index %in% trans_EEZ$index) %>% 
-    left_join(index_id,
-              by = "index")
-
-  #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
+
+  # Selecting only the transboundary nature of the species
+  Trans_Spp <- Spp_Dist %>%
+    filter(index %in% Trans_EEZ$index) %>% 
+    left_join(index_code,
+              by = "index") %>% 
+    filter(eez_name %in% Neighbours_List$eez_name)
+
+  # -------------------- #
+  # Estimate area index
   # The number of species' cells present within each country's EEZ
-
-  #Step 1.  Get EEZ id and Neighbour
-  Neighbours_List <- Neighbours %>% 
-    group_by(eez_name,eez_neighbour) %>% 
-    summarise(n=n()) %>% 
-    ungroup() %>% 
-    select(-n) %>% 
-    semi_join(trans_EEZ,
-              by = c("eez_name","eez_neighbour")
-    )
-
-  # Step 2. Determines the amount of grids present in each country
-  spp_grid <- trans_spp %>% 
+  # -------------------- #
+
+  #Determines the amount of grids present in each country
+  Spp_Grid <- Trans_Spp %>% 
     group_by(taxon_key,
@@
 
-  # Step 3. Sum total grids per Neighbours
-
   # Split dataframes to merge latter
-  Territory_T <- spp_grid %>% 
+  Territory_T <- Spp_Grid %>% 
     ungroup() %>% 
@@
 
-  Neighbour_T <- spp_grid %>% 
+  Neighbour_T <- Spp_Grid %>% 
     ungroup() %>% 
@@
 
-  per_change_d <- full_join(Territory_T,
+  # Estimate Area index and % change
+  Per_Ghange_D <- full_join(Territory_T,
                             Neighbour_T, 
@@
     mutate(area_index = n_spp_Country/Spp_Total) %>% 
-    # select(taxon_key,year,ensamble_list,eez_name,eez_neighbour,area_index) %>% 
+    # Determine the three time frames for comparrison
     mutate(
-      time_step = ifelse(year < 2000,"historical",
-                         ifelse(year >= 2045 & year <= 2064, "mid",
-                                ifelse(year >= 2080, "n_end","NA")
+      time_step = ifelse(year < hist_y,"historical",
+                         ifelse(year >= 2021 & year <= 2040, "early",
+                                ifelse(year >= 2041 & year <= 2060, "mid","NA")
                          )
@@
 
-  if(nrow(per_change_d) > 0){
+  if(nrow(Per_Ghange_D) > 0){
 
@@
 
-  write_csv(per_change_d,
+  write_csv(Per_Ghange_D,
             path_name)
@@
   return(print(paste("analysis done",taxon_key)))
+  # return(Per_Ghange_D)
 
@@
   # Or return a df
-  # return(per_change_d)
-}
-
-```
-
-#### Proportion routine
+  return(Per_Ghange_D)
+}
+
+# EstPropChange(taxon_key = 600004,
+#               year = year_df,
+#               hist_y = 2000,
+#               data_type = "Catch",
+#               ensamble_list = ensamble_list,
+#               path = path,
+#               neighbours = neighbours)
+
+```
+
+### Proportion Control Panel
+
+```{r Emerging_control_panel, eval = F}
+
+## ----------------- ##
+# Data needed for estimating year of emergence
+## ----------------- ##
+
+# SAU relations between INDEX and Country's EEZs (see chunk 3)
+index_code <- read_csv(paste(data_path,"Spatial/sau_index_code.csv",sep=""))
+
+# Neighbour list and their respective id (see chunk 3)
+Neighbours <- read_csv(paste(data_path,"Spatial/Neighbours_eez_id.csv",sep="")) %>% clean_names()
+
+# Transboundary species and their sharing EEZs (see Palacios-Abrantes et al; FishForVisa)
+Transboundary_spp <- read_csv(paste(data_path,"Species/Transboundary_spp.csv",sep="")) %>% 
+  clean_names() %>% 
+  select(-x1)
+
+## ----------------- ##
+# Global variables
+## ----------------- ##
+
+# Year timeframe
+year_df <- tibble(year =  c(seq(1951,2000,1), seq(2001,2099,1)),
+                  time_period = c(rep("historic", 50), rep("future", 99))
+                  )
+
+# Data path
+path = DBEM_path
+
+# List of ensamble members
+ensamble_list <- seq(102,103,1) # for testing
+# ensamble_list <- seq(102,111,1)
+
+#  List os species
+spp_list <- unique(Transboundary_spp$taxon_key)[1:2]
+
+```
+
+### Proportion routine
 
@@
 
-#### Fun `GetCentroids`
+This rutine estimates the time (e.g.
+year) in which the changes in stock share ratio is more likely to happen based on historical variations (e.g.
+model noise) and future trends (e.g signal).
+This result uses teh average of the 10 ensamble members
+
+### Fun `GetCentroids`
 
@@
 
-```{r GetCentroids, eval = F, echo = T}
-
-GetCentroids <- function(taxon_key,year, data_type = "Catch", ensamble, path, Neighbours, eez_centroid){
+```{r GetCentroids, eval = T, echo = F}
+
+GetCentroids <- function(taxon_key,
+                         year, 
+                         data_type = "Catch",
+                         ensamble, 
+                         path,
+                         Neighbours,
+                         eez_centroid){
 
@@
     year = year,
-    data_type = "Catch",
+    data_type = data_type,
     ensamble = ensamble,
@@
 
-    # Run function
+    # Run sub-function
     Centroids_Data <- bind_rows(
@@
   }else{
-
+    # In case there's no data
     Centroids_Data <- tibble()
@@
 
-```
-
-#### Fun `TransIndex`
-
-```{r TransIndex, eval = F, echo = T}
-
-TransIndex <- function(Centroids,Ids){
+GetCentroids(taxon_key = 600004,
+             year = year_df,
+             # data_type = data,
+             ensamble = 102,
+             path = path,
+             Neighbours = Neighbours,
+             eez_centroid = eez_centroid
+             )
+
+```
+
+### Fun `TransIndex`
+
+This functions takes on `GetCentroids` function and estimates the TransIndex (*TI*) for each pair of neighbouring countries. Presents a 10 years average of the results
+
+$$TI =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h)}})^2$$
+
+The year of emergece is defined as the first year where the trend is larger than the 95% CI of the historic variation.
+
+```{r TransIndex, eval = T, echo = T}
+
+TransIndex <- function(Centroids, # Expecting the results from the GetCentroids function
+                       Ids, # A numeric value representing a neighbouring interaction
+                       Hist_y = 2000){ # The LAST year of Historical data
 
@@
   historical <- Centroids_id %>%
-    filter(year <= 2000) %>% 
+    filter(year <= Hist_y) %>% 
     ungroup() %>%
@@
   future <- Centroids_id %>% 
-    filter(year > 2000)
+    filter(year > Hist_y)
 
@@
 
+  # Get unique eezs
   Unique_EEZs <- unique(Trans_index$eez_name)
@@
     ) %>% 
+    # Ten years average of results
     group_by(taxon_key,ensamble,neighbour_id) %>% 
@@
   historic_sd <- mean_sd_index %>% 
-    filter(year < 2000) %>% 
+    filter(year < Hist_y) %>% 
     group_by(taxon_key) %>% 
@@
 
-  # #     
+  # Identifies the year of emergence
   Final_Result <- mean_sd_index %>%
-    filter(year > 2000,
+    filter(year > Hist_y,
            trend > (historic_sd$hist_mean+historic_sd$noise) |
@@
 
-# test <- bind_rows(
-#   lapply(ensamble_list,
-#          GetCentroids,
-#          taxon_key = 600245,
-#          year = year,
-#          data_type = "Catch",
-#          path = path,
-#          Neighbours = Neighbours,
-#          eez_centroid = eez_centroid)
-# )
-# 
-# 
-#   TransIndex(test,819)
-#   
-#   bind_rows(
-#   lapply(Ids,
-#          TransIndex,
-#          Centroid = test)
-#   )
-
-```
-
-#### Fun `GetTransIndex`
+```
+
+### Fun `GetTransIndex`
+
+This is the last function of the emerging year evaluation.
+It was made so we could run the analysis per each species.
+The function function simply runs the functions `GetCentroids` by ensabmble member and then `TransIndex` by pairing EEZ.
+Thus, with this function we run the whole set by species * paired EEZ * ten ensamble members.
 
@@
                             year,
+                          Hist_y,
                             data_type = "Catch",
@@
              year = year,
-             data_type = "Catch",
+             data_type = data_type,
              path = path,
@@
 
+    # Set the different neighbouring pairs
     Ids <- unique(Step_one$neighbour_id)
@@
              TransIndex,
+             Hist_y = Hist_y,
              Centroid = Step_one)
@@
   #   taxon_key = 600004,
-  #   ensamble_list = ensamble_list,
+  #   ensamble_list = ensamble_list[1:2],
   #   year = year,
-  #   data_type = "Catch",
+  #   Hist_y = 2000,
+  #   data_type = "Abd",
   #   path = path,
@@
   # )
-  # 
-  # 
+
   # spp_list <- c(600004, 600245)
@@
 
-#### Emerging Control Panel
+### Emerging Control Panel
 
@@
 
-# Data needed for estimating Proportion change
-
-# SAU relations between INDEX and Country's EEZs
-EEZIDs_List <- read_excel(paste(data_path,"Spatial/V_Lam/Updated_EEZList_17June2016.xlsx",sep="")) %>% 
-  clean_names()
-EEZ_CellID <- read_excel(paste(data_path,"Spatial/V_Lam/EEZ_CellID.xlsx",sep="")) %>% 
-  clean_names()
-colnames(EEZ_CellID) <- c("eezid","index")
-
-Lon_Lat_DBEM <- read.csv(paste(data_path,"Spatial/Lon_Lat_DBEM.txt",sep=""), header=FALSE)
-colnames(Lon_Lat_DBEM) <- c("index","lon","lat")
-
-index_code <- EEZIDs_List %>% 
-  left_join(EEZ_CellID) %>% 
-  rename(eez_name = name) %>% 
-  left_join(Lon_Lat_DBEM,
-            by = "index")
-
-year_df <- tibble(year =  c(seq(1951,2000,1), seq(2001,2099,1)),
-                  time_period = c(rep("historic", 50), rep("future", 99))
-                  )
-
-data_type <- "catch"
-# ensamble_list <- seq(102,103,1)
-ensamble_list <- seq(102,111,1)
-
-world_grid <- tibble(index = seq(1,259200,1))
-path = DBEM_path
-
+## ----------------- ##
+# Data needed for estimating year of emergence
+## ----------------- ##
+
+# SAU relations between INDEX and Country's EEZs (see chunk 3)
+index_code <- read_csv(paste(data_path,"Spatial/sau_index_code.csv",sep=""))
+
+# EEZ list and their centroid of distribution (see chunk 3)
 eez_centroid <- read.csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""), header=TRUE)
 
+# Neighbour list and their respective id (see chunk 3)
 Neighbours <- read.csv(paste(data_path,"Spatial/Neighbours_eez_id.csv",sep="")) %>% clean_names()
 
+# Transboundary species and their sharing EEZs (see Palacios-Abrantes et al; FishForVisa)
 Transboundary_spp <- read.csv(paste(data_path,"Species/Transboundary_spp.csv",sep="")) %>% 
@@
 
+## ----------------- ##
+# Global variables
+## ----------------- ##
+
+# Year timeframe
+year_df <- tibble(year =  seq(1951,2099,1))
+
+# Last year of "historic" data
+Historic_year = 2000
+
+# Data path
+path = DBEM_path
+
+# List of ensamble members
+ensamble_list <- seq(102,103,1) # for testing
+# ensamble_list <- seq(102,111,1)
+
+#  List os species
 spp_list <- unique(Transboundary_spp$taxon_key)[1:2]
@@
 
-#### Emerging Routine
+### Emerging Routine
 
@@
   suppressWarnings(
-    Test <- bind_rows(
-      mclapply(spp_list,
+    Emerging_result <- bind_rows(
+      mclapply(spp_list[1],
                GetTransIndex,
-               year = year,
+               year = year_df,
+               Hist_y = Historic_year,
                data_type = "Catch",
@@
 
-write_csv(Test,
+write_csv(Emerging_result,
           "~/Desktop/Test.csv")
@@
   if(links == "y"){
-    # summarizes by country
+    # summarizes by country and neighbour
     country_summary <- proportion_data %>% 
@@
 
-The next step is to summarize all species per eez
+The next step is to summarize all species per eez so our results shows average change of all species within an EEZ towards a neighbour
 
@@
 # Get taxon list
-taxon_list <- list.files(paste(result_path,"/Proportion/",sep=""))
+Taxon_List <- list.files(paste(result_path,"/Proportion/",sep=""))
 
 # clean names for function
-taxon_list <- gsub(".*_","",taxon_list) 
-taxon_list <- gsub("\\..*","",taxon_list)
+Taxon_List <- gsub(".*_","",Taxon_List) 
+Taxon_List <- gsub("\\..*","",Taxon_List)
 
 # Run results
-results <- PropChange(taxon_list, links="y") %>% 
+Proportion_Results <- PropChange(Taxon_List, links="y") %>% 
   mutate(per_mid =ifelse(per_mid >= 25, 25,
@@
 # Call world map from natural earth
-world_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf"))
-
-coords <- read.csv(coords)
+World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf"))
+
+eez_centroid <- read_csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""))
 
@@
   ggtheme_map() +
-  ggsave("figure_one.png",
+  ggsave("per_mid.png",
          width = 10,
@@
 ```{r fig_map_Juan, eval = F, echo =F}
-
 # Figure made thanks to Juanito!
@@
   # get coords for sources
-  left_join(coords,
+  left_join(eez_centroid,
             by = "eez_name") %>%
@@
   # get coords for targets
-  left_join(coords,
+  left_join(eez_centroid,
             by = "eez_name") %>% 
@@
   drop_na() %>%
-  filter(per_mid > 0,
+  filter(
     variable == "mean_spp"#,
@@
 
+gc()# clean ram
 x <- ggplot() +
-  geom_sf(data = world_map, size = 0.1, fill = "gray90", col = "gray90") +
+  geom_sf(data = subset(World_map, sovereignt != "Antarctica"), size = 0.1, fill = "gray90", col = "gray90") +
   geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray70") +
-  geom_point(data = Data,
+  geom_point(data = subset(Data, per_mid < 0),
              aes(start_lon,
@@
              size = 0.05,
-             col = "black",
+             col = "gray30",
              shape = 20,
              stroke = 1.5) +
-  geom_path(data = paths,
+  geom_path(data = subset(paths, per_mid > 0),
             aes(X, Y,
@@
             size = 0.3,
-            arrow = arrow(length = unit(0.009, "npc")),
+            arrow = arrow(length = unit(0.009, "npc"),
+                          ends = "first"),
             show.legend = NA) +
-  # geom_sf_text(data =World_eez_sf, aes(label =eez_name),
-  #              size = 2) +
-  # geom_text_repel(data = paths,
-  #                  aes(
-  #                    x = X,
-  #                    y = Y,
-  #                    label =source),
-  #                  force = 1
-  # )+
   ggtheme_map() +
-  scale_colour_gradientn("Gain of Stock-Share Ratio (%)",
-                         colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
+  scale_colour_viridis(
+    name="Average Gain of\nStock-Share Ratio (%)",
     limits = c(0,25),
-                         breaks = seq(0,25,5)
-  )+ 
-  ggsave("figure_connections.png",
+    breaks = seq(0,25,5),
+    # option="magma",
+    direction = 1
+  ) +
+  ggsave("figure_connectionsc.png",
          width = 10,

### 2020-05-01 Scripts/Manuscript.Rmd
Drafted intro

@@
 
-## Projecting species distributions under cliamte change
+## Projecting species distributions under climate change
 
@@
 According to the GFDL-ESM2M, under the RCP 8.5 scenario atmospheric temperature will increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial.
-We run the DBEM under ten different GFDL-ESM2M ensamble members [@Roggers:2015jq].
+We run the DBEM under ten different GFDL-ESM2M ensamble members [@Rodgers:2015jq].
 While all ensamble members follow the same RCP, they all constitute different initial Earth conditions.
@@
 An important question is estimating the time in which stock share ratio is going to change from a historical variablillty in order for policy to cope with the change.
+In here we call this event, time of emergence defined as the moment in time when the a secular trend (e.g.
+future variation in SSR) becomes larger than the historical trend variabillity [@Roggers:2015jq].
+Thus we have to understand not only the impact of future climate change on the variabillity of the species distribution, but also the historical variabillity and its difference in the hopes to identify changes between both variabillities.
+We do this by following an ensemble member approach [@Roggers:2015jq; @Frolicher:2009ge].
+Identifying a secular trend for the natural variabillity (e.g.
+backround noise) with a coupled cliamte change model (e.g.
+GFDL) requires multiple initial-condition ensemble simulations in order to rembove natural variability by averaging over the ensemble members [@Roggers:2015jq].
 
-For that we developed an index of transboundary range shift (*I*), fisheries and climate change based on a species distribution centroid:
+Thus, rely on ten GFDL ensable members described above, and developed an index of transboundary range shift, fisheries and climate change (*I*).
+The index is based on the variation of a species distribution centroid as follows:
 
@@
 
-where $D_A$ and $D_B$ represent the distribution centroid of a species in EEZs A and B, respectiveley for each time step from 2000 to 2100 (*t*) and *sd* is the standar deviation of the historical (1951 - 2000) centroid distribution for $D_A$ and $D_B$.
+where $D_A$ and $D_B$ represent the mean (among ensamble members) distribution centroid of a species in EEZs A and B, respectiveley for each time step from 2000 to 2100 (*t*) and *sd* is the standar deviation of the historical (1951 - 2000) centroid distribution for $D_A$ and $D_B$.
 

### 2020-05-13 Scripts/Initial_Analysis.Rmd
All ready to re-run proportion analysis in Hall 1000
fied return function fun proportion
updated paths

@@
 
-ipak <- function(pkg){
-  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
-  if (length(new.pkg)) 
-    install.packages(new.pkg, dependencies = TRUE,repos = "http://cran.us.r-project.org")
-  sapply(pkg, require, character.only = TRUE)
-}
+library(MyFunctions)
 
@@
 
-ipak(packages)
-
-# Set paths depending on machine Beast (jepa88), "carmelia" or Hall1000
-if(Sys.info()[7] == "jepa88"){
-  data_path <- "Z:/DATA/DBEM/"
-}
-if(Sys.info()[7] == "carmelia"){
-  data_path <- paste(here(),"/Temporal_Data/",sep = "")
-  DBEM_path <- paste(data_path,"DBEM/",sep="")
-  result_path <- paste(here(),"/Temporal_Data/Results/",sep="")
-}
-if(Sys.info()[7] == "hall1000"){
-  data_path <- "/Volumes/DATA/JULIANO_NEYMAR/TransEmergence/Data/"
-  DBEM_path <- "/Volumes/DATA/DATA/DBEM/"
-  result_path <- "/Volumes/DATA/JULIANO_NEYMAR/TransEmergence/Results/"
-}
-
-# standarize maps
-ggtheme_map <- function(base_size = 9, Region = "NA") {
-
-  theme(text             = element_text(#family = "Helvetica",
-    color = "gray30", size = base_size),
-    plot.title       = element_text(size = rel(1.25), hjust = 0, face = "bold"),
-    panel.background = element_blank(),
-    panel.border     = element_blank(),
-    panel.grid.minor = element_blank(),
-    panel.grid.major = element_line(color = "transparent"),
-    strip.background =element_rect(fill = "transparent"),
-    strip.text.x = element_text(size = 18, colour = "black",face= "bold.italic", angle = 0),
-    axis.line        = element_blank(), # element_line(colour = "grey30", size = .5))
-    axis.ticks       = element_blank(),
-    axis.text        = element_blank(),
-    axis.title       = element_blank(),
-    legend.key       = element_rect(colour = NA, fill = NA, size = 4),
-    legend.position = "bottom",
-    legend.key.width =unit(5,"line")
-  )
-}
+my_lib(packages)
 
@@
 See methods section of the paper (e.g. Manusript.Rmd)
-
-## Global functions
-
-Here we just run functions that will be used multiple times douring the analysis
-
-### Fun `dbem_import`
-
-This function reads the DBEM data for each year and ensamble (ESM model) from 102 20 111
-
-```{r dbem_import, eval = T, echo = T}
-
-dbem_import <- function(taxon_key, #Except taxon_key 
-                        year, #Expects sequence of years 
-                        ensamble, #Expects a number of the ensamble
-                        data_type, #Expects either "Abd" or "Catch"
-                        path = "NA"){
-
-  years <- year_df %>% pull(year)
-
-  D_Path <- paste(path,"mpa0F1ENS",ensamble,"/",taxon_key,"/",taxon_key,data_type,years,".txt",sep="")
-
-  #### Importing data ####
-  cur <- lapply(D_Path, FUN=fread, na.strings="NA")
-
-  if(length(cur)>0){
-    cur <- cur[sapply(cur, function(d) nrow(d) >= 1)] 
-    colnames <- c("index", "value")
-    cur <- lapply(cur, setNames, colnames)
-    df <- bind_rows(cur, .id = "column_label")
-    if(nrow(df)>0){
-      frame_key <- tibble(column_label = seq(1,length(unlist(years)),1), 
-                          "year"=years) %>% 
-        mutate(column_label=as.character(column_label))
-      df <- left_join(df, frame_key, by="column_label") %>% select(-column_label)
-      df <- df %>% mutate(data_type=data_type,
-                          taxon_key = taxon_key,
-                          ensamble = ensamble
-      )
-    }
-  } else {
-    df <- tibble()
-  }
-    return(df)
-} #Function end
-
-# x <- dbem_import(
-#   taxon_key <- 600004,
-#   Year <- year_df,
-#   data_type <- "Catch",
-#   ensamble = 102,
-#   # ensamble_list =ensamble,
-#   path <- DBEM_path
-# )
-
-# attach(x)
-# head(x)
-# unique(year)
-# unique(ensamble_list)
-# detach(x)
-
-```
 
@@
 
-EEZIDs_List <- read_excel(paste(data_path,"Spatial/V_Lam/Updated_EEZList_17June2016.xlsx",sep="")) %>% 
+EEZIDs_List <- my_path("G", extra_path = "Spatial/DBEM", name="Updated_EEZList_17June2016.xlsx", read = TRUE) %>% 
   clean_names()
-EEZ_CellID <- read_excel(paste(data_path,"Spatial/V_Lam/EEZ_CellID.xlsx",sep="")) %>% 
+
+EEZ_CellID <- my_path("G", extra_path = "Spatial/DBEM", name="EEZ_CellID.xlsx", read = TRUE) %>% 
   clean_names()
+
 colnames(EEZ_CellID) <- c("eezid","index")
 
-Lon_Lat_DBEM <- read.csv(paste(data_path,"Spatial/Lon_Lat_DBEM.txt",sep=""), header=FALSE)
+Lon_Lat_DBEM <- my_path("G", extra_path = "Spatial/DBEM", name="Lon_Lat_DBEM.txt", read = TRUE, header = FALSE)
 colnames(Lon_Lat_DBEM) <- c("index","lon","lat")
 
-index_code <- EEZIDs_List %>% 
+Index_Code <- EEZIDs_List %>% 
   left_join(EEZ_CellID) %>% 
@@
 
+# Save Index data for future 
 write_csv(index_code,
-          "sau_index_code.csv")
+          paste(my_path("G",extra_path = "Spatial/DBEM"),"sau_index_code.csv",sep="")
+          )
 
@@
 # The path
-path_world <- paste(data_path,"Spatial/SAU_Shapefile/",sep="")
+Path_SAU <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile")
 
 # The File
-fnam_world <- "SAUEEZ_July2015.shp"
+File_Name <- "SAUEEZ_July2015.shp"
 
 # Load it!
-World_eez_sf <- st_read(dsn = path_world,
-                        layer =file_path_sans_ext(fnam_world)) %>% 
+SAU_eez_sf <- st_read(dsn = Path_SAU,
+                        layer =file_path_sans_ext(File_Name)
+                      ) %>% 
   rename(eez_name = Name) %>% 
@@
 
-```{r EstPropChange, eval = F, echo = F}
+```{r EstPropChange, eval = T, echo = F}
 
@@
 
-```{r Emerging_control_panel, eval = F}
+```{r Emerging_control_panel, eval = T, echo = F}
 
@@
 
-```{r Proportion_routine, eval = F, echo = F}
+```{r Proportion_routine, eval = T, echo = F}
 
@@
 
-```{r GetTransIndex, eval = F, echo = T}
+```{r GetTransIndex, eval = T, echo = T}
 
@@
 
-```{r Emerging_control_panel, eval = F}
+```{r Emerging_control_panel, eval = T, echo = T}
 
@@
 
-```{r fun_SummaryProp, eval = T, echo =F}
+```{r fun_SummaryProp, eval = F, echo =F}
 
@@
 
-```{r fun_EEZProp, eval = T, echo = F}
+```{r fun_EEZProp, eval = F, echo = F}
 
@@
 
-```{r fun_PropChange, eval = T, echo =F}
+```{r fun_PropChange, eval = F, echo =F}
 
@@
 
-```{r TransIndex, eval = F, echo = T}
+```{r TransIndex, eval = F, echo = F}
 

### 2020-05-21 Scripts/Initial_Analysis.Rmd
updated analysis with correct paths and functions

@@
 packages <- c(
+  "dbemImport",
   "tidyverse",
@@
   clean_names()
-
 colnames(EEZ_CellID) <- c("eezid","index")
@@
 # Read transboundary species database from Palacios-Abrantes et al 2020 (FishForVisa)
-Transboundary_spp <- read.csv(paste(data_path,"Species/Transboundary_spp.csv",sep="")) %>% 
-  clean_names()
+Transboundary_spp <- my_path("D", extra_path = "Species/", name="Transboundary_spp.csv", read = TRUE)
 
@@
                           data_type = "Catch",
-                          ensamble_list, # expects a list of ensambles to averrage by
+                          ensemble_list, # expects a list of ensembles to averrage by
                           path,
@@
 
-  # Loads species data for all ensambles
+  # Loads species data for all ensembles
   Spp_Dist <- bind_rows(
     lapply(
-      ensamble_list,
+      ensemble_list,
       dbem_import,
@@
              year,
-             ensamble) %>% 
+             ensemble) %>% 
     summarise(n_spp_eez = length(unique(index))) %>% 
@@
       year,
-      ensamble
+      ensemble
     )
@@
                             Neighbour_T, 
-                            by = c("Name","taxon_key","year","ensamble")
+                            by = c("Name","taxon_key","year","ensemble")
   ) %>%
@@
     ungroup() %>% 
-    group_by(taxon_key,ensamble,eez_name,eez_neighbour,time_step) %>% 
+    group_by(taxon_key,ensemble,eez_name,eez_neighbour,time_step) %>% 
     summarise_at(vars("area_index"),
@@
     ) %>% 
-    # Ensamble mean and sd
+    # ensemble mean and sd
     group_by(taxon_key,eez_name,eez_neighbour,time_step) %>%
     summarise_at(vars("temporal_mean"),
-                 c(ensamble_mean=mean,
-                   ensamble_sd=sd),
+                 c(ensemble_mean=mean,
+                   ensemble_sd=sd),
                  na.rm=T
@@
   # Or return a df
-  return(Per_Ghange_D)
+  # return(Per_Ghange_D)
 }
@@
 #               data_type = "Catch",
-#               ensamble_list = ensamble_list,
+#               ensemble_list = ensemble_list,
 #               path = path,
@@
 
-```{r Emerging_control_panel, eval = T, echo = F}
+```{r Proportion_control_panel, eval = T, echo = F}
 
@@
 # SAU relations between INDEX and Country's EEZs (see chunk 3)
-index_code <- read_csv(paste(data_path,"Spatial/sau_index_code.csv",sep=""))
+index_code <- my_path("G", extra_path = "Spatial/SAU/", name="sau_index_code.csv", read = TRUE)
 
 # Neighbour list and their respective id (see chunk 3)
-Neighbours <- read_csv(paste(data_path,"Spatial/Neighbours_eez_id.csv",sep="")) %>% clean_names()
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)
 
 # Transboundary species and their sharing EEZs (see Palacios-Abrantes et al; FishForVisa)
-Transboundary_spp <- read_csv(paste(data_path,"Species/Transboundary_spp.csv",sep="")) %>% 
+Transboundary_spp <- my_path("D", extra_path = "Species/", name="Transboundary_spp.csv", read = TRUE) %>% 
   clean_names() %>% 
@@
 
-# List of ensamble members
-ensamble_list <- seq(102,103,1) # for testing
-# ensamble_list <- seq(102,111,1)
+# List of ensemble members
+ensemble_list <- seq(102,103,1) # for testing
+# ensemble_list <- seq(102,111,1)
 
 #  List os species
-spp_list <- unique(Transboundary_spp$taxon_key)[1:2]
+spp_list <- unique(Transboundary_spp$taxon_key)
 
@@
   data_type <- "Catch",
-  ensamble_list = ensamble_list,
+  ensemble_list = ensemble_list,
   path = DBEM_path,
@@
 model noise) and future trends (e.g signal).
-This result uses teh average of the 10 ensamble members
+This result uses teh average of the 10 ensemble members
 
@@
                          data_type = "Catch",
-                         ensamble, 
-                         path,
+                         ensemble, 
+                         # path,
                          Neighbours,
@@
   # ----------------- #
-  # Get model data from spps for one ensamble
+  # Get model data from spps for one ensemble
   #  ---------------- #
 
-  Spp_Dist <- dbem_import(
+  Spp_Dist <- read_dbem_ens(
     taxon_key = taxon_key,
@@
     data_type = data_type,
-    ensamble = ensamble,
-    path = path
+    ensemble = ensemble,
+    path = my_path("D",extra_path = "/DBEM")
   )
@@
 
-  #### Creates a data to control computational time
-  if(file.exists("~/Desktop/spp_df.csv") == FALSE){
+  #### Creates a data to control computational time (lol!)
+    if(file.exists(paste(my_path("R"),"spp_df.csv",sep="")) == FALSE){
     spp_df <- tibble() %>% 
       write_csv(.,
-                "~/Desktop/spp_df.csv")
+                paste(my_path("R"),"spp_df.csv",sep=""))
   }
@@
   suppressMessages(
-    spp_df <- read_csv("~/Desktop/spp_df.csv")
+    spp_df <- my_path("R",name="spp_df.csv",read = TRUE)
   )
   spp_df <- tibble(taxon_key,
-                   ensamble) %>% 
+                   ensemble) %>% 
     bind_rows(spp_df) %>% 
     write_csv(.,
-              "~/Desktop/spp_df.csv")
+              paste(my_path("R"),"spp_df.csv",sep=""))
 
@@
         left_join(EEZ_ids, by = "eez_name") %>% 
-        group_by(taxon_key,year,ensamble,neighbour_id) %>%  # for whole distribution
+        group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
         top_n(1,value) %>% 
@@
         ) %>% 
-        select(taxon_key,eez_name,neighbour_id,ensamble,year,distance)
+        select(taxon_key,eez_name,neighbour_id,ensemble,year,distance)
 
@@
 
-GetCentroids(taxon_key = 600004,
-             year = year_df,
-             # data_type = data,
-             ensamble = 102,
-             path = path,
-             Neighbours = Neighbours,
-             eez_centroid = eez_centroid
-             )
+# Test function
+
+# Test_centroids <- GetCentroids(taxon_key = 600004,
+#              year = year_df$year,
+#              # data_type = data,
+#              ensemble = 102,
+#              # path = path,
+#              Neighbours = Neighbours,
+#              eez_centroid = eez_centroid
+#              )
+# head(Test_centroids)
+# unique(Test_centroids$ensemble)
+# unique(Test_centroids$neigh)
+# 
+# ggplot(Test_centroids) +
+#   geom_line(
+#     aes(
+#       x = year,
+#       y = distance,
+#       color = eez_name
+#     )
+#   )
 
@@
     ungroup() %>%
-    group_by(taxon_key,ensamble,eez_name,neighbour_id) %>%
+    group_by(taxon_key,ensemble,eez_name,neighbour_id) %>%
     summarise_at(vars("distance"),
@@
     left_join(historical,
-              by = c("taxon_key","eez_name","ensamble","neighbour_id")
+              by = c("taxon_key","eez_name","ensemble","neighbour_id")
     ) %>%
@@
     # Ten years average of results
-    group_by(taxon_key,ensamble,neighbour_id) %>% 
+    group_by(taxon_key,ensemble,neighbour_id) %>% 
     mutate(RMean = rollmean(x = trans_index, 
@@
     filter(year > 1960) %>% 
-    select(taxon_key,neighbour_id,year,ensamble,trans_index,RMean)
-
-  # Average of 10 years mean of ensamble members results
+    select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)
+
+  # Average of 10 years mean of ensemble members results
   mean_sd_index <- Partial %>% 
@@
                  c(trend=mean,
-                   ensamble_sd=sd)
-    )
-
-  # Estimate noice from ensamble members
-  historic_sd <- mean_sd_index %>% 
+                   ensemble_sd=sd)
+    )
+
+  # Estimate noice from ensemble members
+  historic_sd <- Partial %>% 
     filter(year < Hist_y) %>% 
     group_by(taxon_key) %>% 
-    summarise_at(vars("trend"),
+    summarise_at(vars("RMean"),
                  c(hist_mean=mean,
@@
   # Plot the result, bruh
-  # ggplot() +
-  #   geom_line(data = mean_sd_index,
-  #             aes(
-  #               x = year,
-  #               y = trend
-  #             )
-  #   ) +
-  #   geom_ribbon(data = mean_sd_index,
-  #               aes(x=year,
-  #                   ymin=historic_sd$hist_mean-historic_sd$noise,
-  #                   ymax=historic_sd$hist_mean+historic_sd$noise),
-  #               alpha = 0.5,
-  #               fill = "grey50"
-  #   ) +
-  #   geom_line(data = mean_sd_index,
-  #             aes(
-  #               x = year,
-  #               y = historic_sd$hist_mean
-  #             ),
-  #             colour = "grey50"
-  #   ) +
-  #   geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
-  #              aes(
-  #                x = year,
-  #                y = trend,
-  #                colour = "red",
-  #              ),
-  #              shape = 1,
-  #              size = 3
-  #   ) +
-  #   geom_vline(xintercept = 2000,colour="blue",alpha = 0.5) +
-  #   theme_classic()
+  # For testing
+  ggplot() +
+    geom_line(data = mean_sd_index,
+              aes(
+                x = year,
+                y = trend
+              )
+    ) +
+    geom_ribbon(data = mean_sd_index,
+                aes(x=year,
+                    ymin=historic_sd$hist_mean-historic_sd$noise,
+                    ymax=historic_sd$hist_mean+historic_sd$noise),
+                alpha = 0.5,
+                fill = "grey50"
+    ) +
+    geom_line(data = mean_sd_index,
+              aes(
+                x = year,
+                y = historic_sd$hist_mean
+              ),
+              colour = "grey50"
+    ) +
+    geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
+               aes(
+                 x = year,
+                 y = trend,
+                 colour = "red",
+               ),
+               shape = 1,
+               size = 3
+    ) +
+    geom_vline(xintercept = 2000,colour="blue",alpha = 0.5) +
+    scale_x_continuous(breaks = seq(1950,2100,5)) +
+    theme_classic() +
+    ggsave("plot_example.jpg",
+       plot = last_plot(),
+       width = 10,
+       height = 8,
+       units = "in",
+       path = my_path("R")
+       )
 
@@
 The function function simply runs the functions `GetCentroids` by ensabmble member and then `TransIndex` by pairing EEZ.
-Thus, with this function we run the whole set by species * paired EEZ * ten ensamble members.
+Thus, with this function we run the whole set by species * paired EEZ * ten ensemble members.
 
@@
                             path,
-                            ensamble_list,
+                            ensemble_list,
                             Neighbours,
@@
     # First step
-    # Call Get centroids fucntion for centroids data per ensamble
+    # Call Get centroids fucntion for centroids data per ensemble
     Step_one <- bind_rows(
-      lapply(ensamble_list,
+      lapply(ensemble_list,
              GetCentroids,
@@
              data_type = data_type,
-             path = path,
              Neighbours = Neighbours,
@@
   #   taxon_key = 600004,
-  #   ensamble_list = ensamble_list[1:2],
+  #   ensemble_list = ensemble_list[1:2],
   #   year = year,
@@
   #   data_type = "Abd",
-  #   path = path,
   #   Neighbours = Neighbours,
@@
   #          eez_centroid = eez_centroid,
-  #          ensamble_list = ensamble_list
+  #          ensemble_list = ensemble_list
   #   )
@@
 # SAU relations between INDEX and Country's EEZs (see chunk 3)
-index_code <- read_csv(paste(data_path,"Spatial/sau_index_code.csv",sep=""))
+index_code <- my_path("G", extra_path = "Spatial/SAU/", name="sau_index_code.csv", read = TRUE)
 
 # EEZ list and their centroid of distribution (see chunk 3)
-eez_centroid <- read.csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""), header=TRUE)
+eez_centroid <- my_path("D", extra_path = "Spatial/", name="eez_centroids.csv", read = TRUE)
 
 # Neighbour list and their respective id (see chunk 3)
-Neighbours <- read.csv(paste(data_path,"Spatial/Neighbours_eez_id.csv",sep="")) %>% clean_names()
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)
 
 # Transboundary species and their sharing EEZs (see Palacios-Abrantes et al; FishForVisa)
-Transboundary_spp <- read.csv(paste(data_path,"Species/Transboundary_spp.csv",sep="")) %>% 
+Transboundary_spp <- my_path("D", extra_path = "Species/", name="Transboundary_spp.csv", read = TRUE) %>% 
   clean_names()
@@
 # Data path
-path = DBEM_path
-
-# List of ensamble members
-ensamble_list <- seq(102,103,1) # for testing
-# ensamble_list <- seq(102,111,1)
+# path = DBEM_path
+
+# List of ensemble members
+ensemble_list <- seq(102,103,1) # for testing
+# ensemble_list <- seq(102,111,1)
 
@@
                GetTransIndex,
-               year = year_df,
+               year = year_df$year,
                Hist_y = Historic_year,
                data_type = "Catch",
-               ensamble_list = ensamble_list,
-               path = path,
+               ensemble_list = ensemble_list,
+               # path = path,
                Neighbours = Neighbours,
@@
       summarise(
-        mean_spp = mean(ensamble_mean,na.rm=T) # average spp change from all countries sharing
+        mean_spp = mean(ensemble_mean,na.rm=T) # average spp change from all countries sharing
       )
@@
 
-Neighbours <- read_excel(paste(data_path,"Spatial/EEZ_Neighbour_List.xlsx",sep="")) %>% 
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
   clean_names()
@@
 
-TransIndex <- function(Spp,year, data_type = "Catch",ensamble_list,path,Neighbours,eez_centroid){
+TransIndex <- function(Spp,year, data_type = "Catch",ensemble_list,path,Neighbours,eez_centroid){
 
@@
     lapply(
-      ensamble_list,
+      ensemble_list,
       dbem_import,
@@
         ungroup() %>%
-        group_by(taxon_key,ensamble,eez_name,neighbour_id) %>%
+        group_by(taxon_key,ensemble,eez_name,neighbour_id) %>%
         summarise_at(vars("distance"),
@@
         left_join(historical,
-                  by = c("taxon_key","eez_name","ensamble","neighbour_id")
+                  by = c("taxon_key","eez_name","ensemble","neighbour_id")
         ) %>%
@@
         ) %>% 
-        group_by(taxon_key,ensamble,neighbour_id) %>% 
+        group_by(taxon_key,ensemble,neighbour_id) %>% 
         mutate(RMean = rollmean(x = trans_index, 
@@
         filter(year > 1960) %>% 
-        select(taxon_key,neighbour_id,year,ensamble,trans_index,RMean)
-
-      # Average of 10 years mean of ensamble members results
+        select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)
+
+      # Average of 10 years mean of ensemble members results
       mean_sd_index <- Partial %>% 
@@
                      c(trend=mean,
-                       ensamble_sd=sd)
-        )
-
-      # Estimate noice from ensamble members
+                       ensemble_sd=sd)
+        )
+
+      # Estimate noice from ensemble members
       historic_sd <- mean_sd_index %>% 
@@
 #              data_type = "Catch",
-#              ensamble_list,
+#              ensemble_list,
 #              path,

### 2020-05-21 Scripts/Initial_Analysis.Rmd
updates on functions tp work in all systems

@@
                          ensemble, 
-                         # path,
+                         path,
                          Neighbours,
-                         eez_centroid){
+                         eez_centroid
+                         ){
+
+  if(Sys.info()[7] == "jepa88"){
+      Sys <-FALSE
+    }else{
+      Sys <- TRUE
+    }
 
@@
     ensemble = ensemble,
-    path = my_path("D",extra_path = "/DBEM")
+    path = path,
+    my_path = Sys
   )
@@
   #### Creates a data to control computational time (lol!)
-    if(file.exists(paste(my_path("R"),"spp_df.csv",sep="")) == FALSE){
+    if(file.exists(paste(my_path("R",repo=Sys),"spp_df.csv",sep="")) == FALSE){
     spp_df <- tibble() %>% 
       write_csv(.,
-                paste(my_path("R"),"spp_df.csv",sep=""))
+                paste(my_path("R", repo =Sys),"spp_df.csv",sep=""))
   }
@@
   suppressMessages(
-    spp_df <- my_path("R",name="spp_df.csv",read = TRUE)
+    spp_df <- my_path("R",name="spp_df.csv",read = TRUE, repo =Sys)
   )
@@
     write_csv(.,
-              paste(my_path("R"),"spp_df.csv",sep=""))
+              paste(my_path("R", repo =Sys),"spp_df.csv",sep=""))
 
@@
 
-# Test_centroids <- GetCentroids(taxon_key = 600004,
+# GetCentroids(taxon_key = 600004,
 #              year = year_df$year,
@@
 #              ensemble = 102,
-#              # path = path,
 #              Neighbours = Neighbours,
@@
     theme_classic() +
-    ggsave("plot_example.jpg",
+    facet_wrap(~neighbour_id)+
+    ggsave(paste(taxon_key,"plot_example.jpg",sep="_"),
        plot = last_plot(),
@@
        units = "in",
-       path = my_path("R")
+       path = my_path("R", repo = Sys)
        )
@@
 
-  # GetTransIndex(
-  #   taxon_key = 600004,
-  #   ensemble_list = ensemble_list[1:2],
-  #   year = year,
-  #   Hist_y = 2000,
-  #   data_type = "Abd",
-  #   Neighbours = Neighbours,
-  #   eez_centroid = eez_centroid
-  # )
+  GetTransIndex(
+    taxon_key = 600006,
+    ensemble_list = ensemble_list,
+    year = year_df$year,
+    Hist_y = 2000,
+    data_type = "Abd",
+    Neighbours = Neighbours,
+    eez_centroid = eez_centroid
+  )
 
@@
 # Data needed for estimating year of emergence
+# Note: Use repo = F when running on the Beast or Hall1000
 ## ----------------- ##
@@
 # SAU relations between INDEX and Country's EEZs (see chunk 3)
-index_code <- my_path("G", extra_path = "Spatial/SAU/", name="sau_index_code.csv", read = TRUE)
+index_code <- my_path("G", extra_path = "Spatial/SAU/", name="sau_index_code.csv", read = TRUE, repo = F)
 
 # EEZ list and their centroid of distribution (see chunk 3)
-eez_centroid <- my_path("D", extra_path = "Spatial/", name="eez_centroids.csv", read = TRUE)
+eez_centroid <- my_path("D", extra_path = "Spatial/", name="eez_centroids.csv", read = TRUE, repo=F )
 
 # Neighbour list and their respective id (see chunk 3)
-Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE, repo = F)
 
 # Transboundary species and their sharing EEZs (see Palacios-Abrantes et al; FishForVisa)
-Transboundary_spp <- my_path("D", extra_path = "Species/", name="Transboundary_spp.csv", read = TRUE) %>% 
+Transboundary_spp <- my_path("D", extra_path = "Species/", name="Transboundary_spp.csv", read = TRUE, repo = F) %>% 
   clean_names()
@@
 # List of ensemble members
-ensemble_list <- seq(102,103,1) # for testing
-# ensemble_list <- seq(102,111,1)
+# ensemble_list <- seq(102,103,1) # for testing
+ensemble_list <- seq(102,111,1)
 

### 2020-05-22 Scripts/Initial_Analysis.Rmd
Should be ready to run in Beasty Beast

--- 2020-05-21 Scripts/Initial_Analysis.Rmd
@@
                          ensemble, 
-                         # path,
+                         path,
                          Neighbours,
-                         eez_centroid){
+                         eez_centroid
+                         ){
 
@@
     data_type = data_type,
-    ensemble = ensemble,
-    path = my_path("D",extra_path = "/DBEM")
+    ensemble = ensemble
   )
@@
 
-# Test_centroids <- GetCentroids(taxon_key = 600004,
-#              year = year_df$year,
-#              # data_type = data,
+# GetCentroids(taxon_key = 600006,
+#              year = year_df$year[1:5],
+#              data_type = "Catch",
 #              ensemble = 102,
-#              # path = path,
 #              Neighbours = Neighbours,
@@
 #              )
+
 # head(Test_centroids)
@@
 TransIndex <- function(Centroids, # Expecting the results from the GetCentroids function
-                       Ids, # A numeric value representing a neighbouring interaction
-                       Hist_y = 2000){ # The LAST year of Historical data
+                       Id, # A numeric value representing a neighbouring interaction (ONE!)
+                       Hist_y = 2000,
+                       Plot = F){ # The LAST year of Historical data
 
@@
   Centroids_id <- Centroids %>% 
-    filter(neighbour_id %in% Ids)
+    filter(neighbour_id %in% Id)
 
   EEZ_ids <- Neighbours %>% 
-    filter(neighbour_id %in% Ids)
+    filter(neighbour_id %in% Id)
 
@@
 
+  # For testing
+  if(Plot == T){
   # Plot the result, bruh
-  # For testing
+
+  # Step to determine if data exists
+  Plot_Test <- nrow(subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)))>0
+
   ggplot() +
@@
     ) +
-    geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
+    # Conditional, will only plot circles if the data exists
+    {if(Plot_Test)geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
                aes(
@@
                size = 3
-    ) +
+    )
+      } +
     geom_vline(xintercept = 2000,colour="blue",alpha = 0.5) +
@@
     theme_classic() +
-    ggsave("plot_example.jpg",
+    ggsave(paste(taxon_key,Id,"plot_example.jpg",sep="_"),
        plot = last_plot(),
@@
        )
+
+  }
 
@@
     Step_one <- bind_rows(
-      lapply(ensemble_list,
+      lapply(ensemble_list[1:2],
              GetCentroids,
              taxon_key = taxon_key,
-             year = year,
+             year = year_df$year,
              data_type = data_type,
              Neighbours = Neighbours,
-             eez_centroid = eez_centroid)
+             eez_centroid = eez_centroid
+             )
     )
@@
   # GetTransIndex(
-  #   taxon_key = 600004,
-  #   ensemble_list = ensemble_list[1:2],
-  #   year = year,
+  #   taxon_key = 600006,
+  #   ensemble_list = ensemble_list,
+  #   year = year_df$year,
   #   Hist_y = 2000,

### 2020-05-25 Scripts/Initial_Analysis.Rmd
Fixing Nas issues

@@
   # Future results
-  future <- Centroids_id %>% 
+  future <- Centroids_id %>% # FIX ME CUS I PRODUCE A BUNCH OF NAs WITH SSP 600076 ENS 102 AND 104
     filter(year > Hist_y)
@@
     ) %>%
+    # Filter out ensembels that have no data
+    filter(!is.na(value)) %>% 
     select(-eez_name,-variable) %>%
@@
     Step_one <- bind_rows(
-      lapply(ensemble_list[1:2],
+      lapply(ensemble_list,
              GetCentroids,
@@
     Step_two <- bind_rows(
-      lapply(Ids,
+      lapply(Ids[7],
              TransIndex,
@@
 #  List os species
-spp_list <- unique(Transboundary_spp$taxon_key)[1:2]
+spp_list <- unique(Transboundary_spp$taxon_key)
 
@@
     Emerging_result <- bind_rows(
-      mclapply(spp_list[1],
+      mclapply(spp_list,
                GetTransIndex,
@@
                ensemble_list = ensemble_list,
-               # path = path,
                Neighbours = Neighbours,
@@
 write_csv(Emerging_result,
-          "~/Desktop/Test.csv")
+          paste(my_path("R"),"Emerging_Results.csv",sep="")
+)
 

### 2020-06-15 Scripts/Initial_Analysis.Rmd
Fixed code and made first result run

--- 2020-05-22 Scripts/Initial_Analysis.Rmd
@@
   #  ---------------- #
+ if(ensemble == 102){
+   print(taxon_key)
+ }
 
@@
   #### Creates a data to control computational time (lol!)
-    if(file.exists(paste(my_path("R"),"spp_df.csv",sep="")) == FALSE){
-    spp_df <- tibble() %>% 
-      write_csv(.,
-                paste(my_path("R"),"spp_df.csv",sep=""))
-  }
-
-  suppressMessages(
-    spp_df <- my_path("R",name="spp_df.csv",read = TRUE)
-  )
-  spp_df <- tibble(taxon_key,
-                   ensemble) %>% 
-    bind_rows(spp_df) %>% 
-    write_csv(.,
-              paste(my_path("R"),"spp_df.csv",sep=""))
+  #   if(file.exists(paste(my_path("R"),"spp_df.csv",sep="")) == FALSE){
+  #   spp_df <- tibble() %>% 
+  #     write_csv(.,
+  #               paste(my_path("R"),"spp_df.csv",sep=""))
+  # }
+  # 
+  # suppressMessages(
+  #   spp_df <- my_path("R",name="spp_df.csv",read = TRUE)
+  # )
+  # spp_df <- tibble(taxon_key,
+  #                  ensemble) %>% 
+  #   bind_rows(spp_df) %>% 
+  #   write_csv(.,
+  #             paste(my_path("R"),"spp_df.csv",sep=""))
 
@@
 
-# GetCentroids(taxon_key = 600006,
+# GetCentroids(taxon_key = 600076,
 #              year = year_df$year[1:5],
@@
 
+Note: If no historic data available then emerging year is the first year of data in an EEZ
+
 ```{r TransIndex, eval = T, echo = T}
@@
                  c(hd_temp_mean=mean,
-                   hd_temp_sd=sd)
+                   hd_temp_sd=sd,
+                   n_yrs =~n())
                  )
@@
     filter(year > Hist_y)
+
+   # Get unique eezs
+  Unique_EEZs <- unique(EEZ_ids$eez_name)
 
@@
     gather("variable","value",distance:hd_temp_sd) %>%
-    ungroup()
-
-  # Get unique eezs
-  Unique_EEZs <- unique(Trans_index$eez_name)
-
-  Partial <- Trans_index %>% 
+    ungroup() %>% 
+    mutate(value = ifelse(is.na(value),0,value)) %>% 
     mutate(
-      names = ifelse(eez_name ==Unique_EEZs[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_"))
+      names = ifelse(eez_name ==Unique_EEZs[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_")
+    )
     ) %>%
@@
       trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
+    # )
     ) %>% 
@@
   # Average of 10 years mean of ensemble members results
-  mean_sd_index <- Partial %>% 
+  mean_sd_index <- Trans_index %>% 
     group_by(taxon_key,year,neighbour_id) %>% 
@@
                  c(trend=mean,
-                   ensemble_sd=sd)
+                   ensemble_sd=sd,
+                   n_ens = ~n())
     )
@@
   # Estimate noice from ensemble members
-  historic_sd <- Partial %>% 
+  historic_sd <- Trans_index %>% 
     filter(year < Hist_y) %>% 
@@
                  c(hist_mean=mean,
-                   noise=sd)
+                   noise=sd,
+                   n_ens = ~n())
     )
@@
 
-  # Identifies the year of emergence
+  # If no historic data then emergin year is the first year of data
+  if(nrow(historic_sd) == 0){
+
+    Final_Result <- mean_sd_index %>%
+      group_by(taxon_key,neighbour_id,n_ens) %>%
+      summarise(emerging_yr= min(year)) %>%
+      left_join(EEZ_ids,
+                by = "neighbour_id") %>%
+      ungroup() %>%
+      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr)
+
+  return(Final_Result)
+
+  }else{
+
+  # Identifies the year of emergence (e.g. trend > noise)
   Final_Result <- mean_sd_index %>%
@@
     ) %>%
-    group_by(taxon_key,neighbour_id) %>%
+    group_by(taxon_key,neighbour_id,n_ens) %>%
     summarise(emerging_yr= min(year)) %>%
-    left_join(EEZ_ids) %>%
+    left_join(EEZ_ids,
+              by = "neighbour_id") %>%
     ungroup() %>%
-    select(taxon_key,eez_name,eez_neighbour,emerging_yr)
+    select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr)
 
   return(Final_Result)
+  }
 }
@@
                           eez_centroid){
-
+  t_start <- Sys.time()
+  print("Starting Step One")
     # First step
@@
     Ids <- unique(Step_one$neighbour_id)
+
+    print("Starting Step Two")
 
@@
 
-    return(Step_two)
+    name <- paste(my_path("R",extra_path = "/Emergence/"),taxon_key,"_emergence.txt",sep="")
+
+    write_csv(Step_two,
+              name)
+
+    time_dif <- Sys.time()-t_start
+
+    message <- paste("Analysis done for", taxon_key, "in",time_dif, "minutes")
+    return(message)
   }
@@
   # GetTransIndex(
-  #   taxon_key = 600006,
+  #   taxon_key = 600076,
   #   ensemble_list = ensemble_list,
@@
   #   Hist_y = 2000,
-  #   data_type = "Abd",
+  #   data_type = "Catch",
   #   Neighbours = Neighbours,
@@
 #  List os species
-spp_list <- unique(Transboundary_spp$taxon_key)[1:2]
+spp_list <- unique(Transboundary_spp$taxon_key)
+
+# Already modeled species
+spp_done <- list.files(my_path("R", extra_path = "Emergence"))
+
+spp_done <- gsub("\\_.*","",spp_done)
+
+spp_reman <- spp_list[!spp_list %in% spp_done]
 
@@
 ####_______________________ ####
-
+n_cores <- 24
 t_start <- Sys.time()
 
-suppressMessages(
-  suppressWarnings(
     Emerging_result <- bind_rows(
-      mclapply(spp_list[1],
+      lapply(spp_reman[63],
                GetTransIndex,
@@
     )
-  )
-)
+
+library(doParallel)
+library(doSNOW)
+
+cl <- makeCluster(12)
+registerDoSNOW(cl)
+
+# To show processing times
+repetitions <- length(spp_reman)
+pb <- txtProgressBar(max = repetitions, style = 3)
+
+progress <- function(n) setTxtProgressBar(pb, n)
+opts <- list(progress = progress)
+
+results <- foreach(i=1:repetitions,
+                    .combine = rbind,
+                    .packages = c("MyFunctions",packages),
+                    .options.snow = opts,
+                   .verbose = T) %dopar% 
+  GetTransIndex(
+    taxon_key =spp_reman[i],
+    ensemble_list = ensemble_list,
+    year = year_df$year,
+    Hist_y = 2000,
+    data_type = "Catch",
+    Neighbours = Neighbours,
+    eez_centroid = eez_centroid
+  )
+
+close(pb)
+stopCluster(cl)
 

### 2020-06-15 Scripts/Initial_Analysis.Rmd
Juanitos map for emerging stocks

@@
 
+```
+
+```{r Emerging_routine_DoPar, eval = F, echo = F}
+
 library(doParallel)
@@
 
-results <- foreach(i=1:repetitions,
+foreach(i=1:repetitions,
         .combine = rbind,
@@
 
-t_end <- Sys.time()-t_start;t_end
-
-write_csv(Emerging_result,
-          paste(my_path("R"),"Emerging_Results.csv",sep="")
-)
-
-# Version 1 (embeded loop) <- 3.01
-# Version 2 (embeded lapply) <- 2.71
-# Version 3 (embeded mclapply) <- 1.597311 mins
-# Version 4 (embeded mclapply) + mclapply in spp <- 1.597311 mins
-
 ```
@@
 
-```{r fun_PropChange, eval = F, echo =F}
+####  Prooportion Analysis
+
+```{r PropChange, eval = F, echo =F}
 
 # Get taxon list
-Taxon_List <- list.files(paste(result_path,"/Proportion/",sep=""))
+Taxon_List <- list.files(paste(my_path("R"),"Propotion",sep=""))
 
@@
 
-### Fig.Percentage Change Map
+##### Fig.Percentage Change Map
 
@@
 
-### Fig. Netowrk
+##### Fig. Netowrk
 
@@
   )
+
+```
+
+####  Year of Eergence Analysis
+
+```{r PropChange, eval = F, echo =F}
+
+# Get taxon list
+Taxon_List <- list.files(my_path("R",extra_path = "Emergence_yr"))
+
+# clean names for function
+Taxon_List <- gsub("_.*","",Taxon_List)
+
+Paths <- paste(my_path("R",extra_path = "Emergence_yr"),Taxon_List,"_emergence.txt",sep="")
+
+# Load all sp in noe table
+Eme_Res <- bind_rows(lapply(Paths, fread))
+# warnings() for empty data species
+
+View(Eme_Res)
+
+### Load Data
+
+# Load SAU shapefile name and paths
+
+# The path
+path_world <- paste(data_path,"Spatial/SAU_Shapefile/",sep="")
+
+# The File
+fnam_world <- "SAUEEZ_July2015.shp"
+
+# Load it!
+World_eez_sf <- st_read(dsn = path_world,
+                    layer =file_path_sans_ext(fnam_world)) %>% 
+  rename(eez_name = Name) %>% 
+  st_transform(crs = 4326) %>% # 4326
+  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)
+
+# Call world map from natural earth
+World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf"))
+
+eez_centroid <- read_csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""))
+
+```
+
+##### Histogram of Data
+
+```{r histo,  eval = T, echo = T}
+
+Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(eez_name,emerging_yr) %>% 
+  summarise(
+    n_tax = length(unique(taxon_key))
+  ) %>% 
+  ggplot() +
+  geom_histogram(
+    aes(
+      x = emerging_yr
+    ),
+    stat = "count"
+  ) # +
+  # facet_wrap(~n_ens)
+
+Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr) %>% 
+  summarise(
+    n_tax = length(unique(taxon_key))
+  ) %>% 
+  group_by() %>%
+  mutate(RMean = rollmean(x = n_tax, 
+                          10,
+                          align = "right",
+                          fill = n_tax)
+  ) %>%
+  filter(emerging_yr >= 2020) %>% 
+  ggplot() +
+  geom_line(
+    aes(
+      x = emerging_yr,
+      y = RMean
+    )
+  )
+
+```
+
+##### Mapping the time of emergence
+
+```{r time_emergence_map,  eval = T, echo = T}
+
+results <- Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # emerging_yr >= 2020
+  ) %>% 
+  group_by(eez_name,eez_neighbour,emerging_yr) %>% 
+  summarise(
+    n_tax = length(unique(taxon_key))
+  ) %>% 
+  mutate(
+    time_step = ifelse(emerging_yr <= 2050, "Mid_Century","Then_End_Century")
+  ) %>% 
+  # filter(
+  #   emerging_yr >= 2030
+  # ) %>% 
+  group_by(eez_name,eez_neighbour,time_step) %>% 
+  summarise(
+    n_tax_emer = sum(n_tax)
+  )
+
+##_----------------------- #
+# Checking on China and Indonesia
+# Probably due to dispute territories
+
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
+  clean_names()
+
+# 
+# Neighbours %>% 
+#   filter(eez_name == "China") %>% 
+#   pull(eez_neighbour) %>% 
+#   unique()
+
+##_----------------------- #
+
+# Get the center poligon of each EEZ for map source/target
+coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
+  ungroup() %>% 
+  select(eez_name,geometry) %>% 
+  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
+  # Manually fix issues
+  mutate(
+    longitude = ifelse(
+      eez_name == "Fiji", 178.065033,
+      ifelse(eez_name == "Tuvalu",177.6493,
+             ifelse( eez_name == "New Zealand",174.8860,
+                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
+                     )
+             )
+      )
+    )
+  )
+
+# Remove "c()"
+coords$longitude <- gsub("\\(","",coords$longitude)
+coords$longitude <- gsub("c","",coords$longitude)
+coords$latitude <- gsub(")","",coords$latitude)
+
+coords <- coords %>% 
+  mutate_at(vars(longitude,latitude),as.numeric)
+
+# Prepare data from results
+Test_Data <- results %>% 
+  # get coords for sources
+  left_join(coords,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  left_join(coords,
+            by = "eez_name") %>% 
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  # mutate(
+  #   per_mid = ifelse(is.na(per_mid),0,per_mid)
+  # ) %>% 
+  drop_na() #%>% 
+  # filter(
+    # per_mid >= 20,
+    # variable == "mean_spp",
+  # )
+
+# Convert to data to spatial
+network_data_sf <- Test_Data %>%
+  select(start_lon, start_lat, end_lon, end_lat) %>% 
+  purrr::transpose() %>% 
+  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
+  purrr::map(st_linestring) %>%
+  st_sfc(crs = 4326) %>%
+  st_sf(geometry = .)%>%
+  bind_cols(Test_Data) %>%
+  select(everything(), geometry) %>% 
+  # Set circles ('cus the world is not flat!)
+  st_segmentize(units::set_units(100, km)) %>% 
+  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
+  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
+  sf::st_sf(crs = 4326)
+
+#  unpack the coordinates of each feature 
+coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
+
+c(network_data_sf$line_id, network_data_sf$coords) %<-% (
+  coord_data %>% 
+    rename(subline_id = L1) %>%   
+    nest(-L2)
+)
+
+# Create data path for line connections
+paths <- network_data_sf %>% 
+  sf::st_set_geometry(NULL) %>% 
+  unnest() %>% 
+  mutate_at(vars(X,Y), round,2)
+
+# Plot it!
+
+country <- c("Fiji","Solomon Isl.","Tuvalu")
+
+ggplot() +
+  geom_sf(
+    # data = subset(world_map,name %in% country),
+    data = World_map,
+    size = 0.1, fill = "gray90", col = "gray90") +
+  geom_sf(
+    # data = subset(World_eez_sf, eez_name %in% country),
+    data = World_eez_sf,
+    size = 0.1, fill = "white", col = "gray70") +
+  geom_point(data = Test_Data,
+             # data = subset(Test_Data, source %in% country & target %in% source),
+             aes(start_lon,
+                 start_lat,
+                 col = n_tax_emer),
+             size = 0.1,
+             col = "black",
+             # shape = 21,
+             stroke = 1.5) +
+  geom_path(data = paths,
+            # data = subset(paths, source %in% country),
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = n_tax_emer
+            ),
+            size = 0.3,
+            arrow = arrow(length = unit(0.009, "npc")),
+            show.legend = NA) +
+  # geom_sf_text(data = subset(World_eez_sf, eez_name %in% country),
+  #              aes(label =eez_name),
+  #              size = 2) +
+  # geom_text(data = subset(Test_Data, source %in% country),
+  #              aes(start_lon,
+  #                start_lat,
+  #                label = paste(source,"to",target)
+  #              ),
+  #              size = 2) +
+  # geom_text_repel(data = Test_Data,
+  #   aes(start_lon,
+# start_lat,
+# label = paste(source,"to",target)
+# )
+# )+
+ggtheme_map() +
+  scale_colour_gradientn("Number of emerging stocks",
+                         colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
+                         limits = c(0,35),
+                         breaks = seq(0,35,5)
+  ) +
+  facet_wrap(~time_step,
+             ncol = 1)
+
+```
 

### 2020-06-15 Scripts/Initial_Analysis.Rmd
Included plots and code to get climate data

--- 2020-06-15 Scripts/Initial_Analysis.Rmd
@@
   "zeallot", # for Juanito's map
-  "gmt" # for estimating distances between points
+  "gmt", # for estimating distances between points
+  "viridis"
 )
@@
   ggsave("figure_connectionsc.png",
+         width = 10,
+         height = 6,
+         units = "in"
+  )
+
+```
+
+####  Year of Eergence Analysis
+
+```{r PropChange, eval = F, echo =F}
+
+# Get taxon list
+Taxon_List <- list.files(my_path("R",extra_path = "Emergence_yr"))
+
+# clean names for function
+Taxon_List <- gsub("_.*","",Taxon_List)
+
+Paths <- paste(my_path("R",extra_path = "Emergence_yr"),Taxon_List,"_emergence.txt",sep="")
+
+# Load all sp in noe table
+Eme_Res <- bind_rows(lapply(Paths, fread))
+# warnings() for empty data species
+
+# View(Eme_Res)
+
+### Load Data
+
+# Load SAU shapefile name and paths
+
+# The path
+path_world <- paste(data_path,"Spatial/SAU_Shapefile/",sep="")
+
+# The File
+fnam_world <- "SAUEEZ_July2015.shp"
+
+# Load it!
+World_eez_sf <- st_read(dsn = path_world,
+                    layer =file_path_sans_ext(fnam_world)) %>% 
+  rename(eez_name = Name) %>% 
+  st_transform(crs = 4326) %>% # 4326
+  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)
+
+# Call world map from natural earth
+World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf"))
+
+eez_centroid <- read_csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""))
+
+```
+
+##### Histogram of Data
+
+```{r histo,  eval = T, echo = T}
+
+# Per Stock
+stock_d_hist <- Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr,taxon_key) %>%
+  # summarise(n_tax = length(unique(eez_name))) %>%
+  summarise(n()) %>% 
+  mutate(
+    level = "stock"
+  ) %>% 
+  select(1,3,4)
+  # ggplot() +
+  # geom_histogram(
+  #   aes(
+  #     x = emerging_yr
+  #   ),
+  #   stat = "count"
+  # )
+
+# Per EEZs
+eez_d_hist <- Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr,eez_name) %>% 
+  # summarise(n_tax = length(unique(taxon_key))) %>% 
+  summarise(n()) %>% 
+  mutate(
+    level = "EEZ"
+  ) %>% 
+  select(1,3,4)
+
+hist_data <- stock_d_hist %>% 
+  bind_rows(eez_d_hist)
+
+ggplot() +
+  geom_density(data = subset(hist_data, level == "EEZ"),
+    aes(
+      x = emerging_yr,
+      fill = level
+    ),
+    stat = "count",
+    alpha= 0.8,
+    size=0.5,
+  ) +
+  geom_density(data = subset(hist_data, level == "stock"),
+    aes(
+      x = emerging_yr,
+      fill = level
+    ),
+    stat = "count",
+    alpha= 0.8,
+    size=0.5,
+  ) +
+    scale_y_continuous("Frequency",
+                       breaks = seq(0,90,10),
+                       limits = c(0,100)
+    ) +
+    scale_x_continuous("Year of emergence",
+                       breaks = seq(2000,2100,10)
+    ) +
+    my_ggtheme_p() +
+    scale_fill_viridis(discrete = T)
+
+```
+
+##### Scatter plots
+
+```{r Scatter_plot,  eval = T, echo = T}
+
+ data_stock <- Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr,eez_name) %>% 
+  summarise(
+    tax_n = length(unique(taxon_key))
+    ) %>% 
+  group_by(emerging_yr) %>% 
+  summarise(emerging_tax = sum(tax_n))
+
+data_eez <- Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr,taxon_key) %>% 
+  summarise(
+    eez_n = length(unique(eez_name))
+    ) %>% 
+    group_by(emerging_yr) %>% 
+  summarise(emerging_eez = sum(eez_n))
+
+data %>% 
+  left_join(data_eez,
+            by = "emerging_yr") %>% 
+  ggplot() +
+  geom_point(
+    aes(
+      x = emerging_eez,
+      y = emerging_tax,
+      fill = emerging_yr,
+      colour = emerging_yr
+    )
+  )
+
+```
+
+##### Temperature gradient
+
+##### Fun `GetEnvData`
+
+```{r get_env_data, eval = T, echo = T}
+
+getEnvVar=function(yr,ModelRCP,RCP,Variable){
+
+  if(ModelRCP == "IPSL"){
+    ModelRCP = "IPSL-CM5A-MR"
+    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
+  }
+
+  if(ModelRCP == "MPI"){
+    ModelRCP = "MPI-ESM-MR"
+    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
+  }
+
+  if(ModelRCP == "GFDL"){
+
+    if(RCP == "rcp26"){
+      ModelRCP <- "GFDL26"
+    }else{
+      ModelRCP <- "GFDL85"
+    }
+    Filepath <- paste("/Volumes/DATA/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")
+  }
+
+  if(Variable == "Ice"){
+
+    Bottom <- paste("IceExt_",yr,".txt",sep="")
+
+  }
+
+  if(Variable == "Temperature"){
+
+    Bottom = paste("SST_",yr,".txt",sep="")
+
+  }
+
+  if(Variable == "Salinity"){
+    # Bottom
+    Bottom = paste("Salinity_btm_",yr,".txt",sep="")
+
+  }
+
+  Read_me <- paste(Filepath,Bottom,sep="")[1]
+  Bottom_Data <- read.csv(Read_me,header=F)
+
+  # Retrun
+
+  getTemp <- Bottom_Data 
+}
+
+Years <-seq(2020,2100,1)
+
+for(y in 1:length(Years)){
+
+        Data <- getEnvVar(yr = Years[y],
+                          Model = "GFDL",
+                          Variable = "Temperature",
+                          RCP = "rcp85"
+        ) %>% 
+          mutate(INDEX = all_Grid) %>% 
+          filter(INDEX %in% Grid$INDEX)
+
+        if(y == 1){
+
+          F_BData <- Data[1]
+
+        }else{
+
+          F_BData <- cbind(Data[1],F_BData)
+        }
+      }
+
+```
+
+```{r Scatter_plot,  eval = T, echo = T}
+
+ data_stock <- Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr,eez_name) %>% 
+  summarise(
+    tax_n = length(unique(taxon_key))
+    ) %>% 
+  group_by(emerging_yr) %>% 
+  summarise(emerging_tax = sum(tax_n))
+
+data_eez <- Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr,taxon_key) %>% 
+  summarise(
+    eez_n = length(unique(eez_name))
+    ) %>% 
+    group_by(emerging_yr) %>% 
+  summarise(emerging_eez = sum(eez_n))
+
+data %>% 
+  left_join(data_eez,
+            by = "emerging_yr") %>% 
+  ggplot() +
+  geom_point(
+    aes(
+      x = emerging_eez,
+      y = emerging_tax,
+      fill = emerging_yr,
+      colour = emerging_yr
+    )
+  )
+
+```
+
+##### Mapping the time of emergence
+
+```{r time_emergence_map,  eval = T, echo = T}
+
+##_----------------------- #
+# Checking on China and Indonesia
+# Probably due to dispute territories
+
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
+  clean_names()
+
+# 
+# Neighbours %>% 
+#   filter(eez_name == "China") %>% 
+#   pull(eez_neighbour) %>% 
+#   unique()
+
+##_----------------------- #
+
+# Get the center poligon of each EEZ for map source/target
+coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
+  ungroup() %>% 
+  select(eez_name,geometry) %>% 
+  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
+  # Manually fix issues
+  mutate(
+    longitude = ifelse(
+      eez_name == "Fiji", 178.065033,
+      ifelse(eez_name == "Tuvalu",177.6493,
+             ifelse( eez_name == "New Zealand",174.8860,
+                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
+                     )
+             )
+      )
+    )
+  )
+
+# Remove "c()"
+coords$longitude <- gsub("\\(","",coords$longitude)
+coords$longitude <- gsub("c","",coords$longitude)
+coords$latitude <- gsub(")","",coords$latitude)
+
+coords <- coords %>% 
+  mutate_at(vars(longitude,latitude),as.numeric)
+
+# Prepare data from results
+Test_Data <- Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # emerging_yr >= 2020
+  ) %>% 
+  group_by(eez_name,eez_neighbour,emerging_yr) %>% 
+  summarise(
+    n_tax = length(unique(taxon_key))
+  ) %>% 
+  filter(
+    emerging_yr <= 2050
+  ) %>%
+  group_by(eez_name,eez_neighbour) %>% 
+  summarise(
+    n_tax_emer = sum(n_tax)
+  ) %>% 
+  # get coords for sources
+  left_join(coords,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  left_join(coords,
+            by = "eez_name") %>% 
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  # mutate(
+  #   per_mid = ifelse(is.na(per_mid),0,per_mid)
+  # ) %>% 
+  drop_na() #%>% 
+  # filter(
+    # per_mid >= 20,
+    # variable == "mean_spp",
+  # )
+
+# Convert to data to spatial
+network_data_sf <- Test_Data %>%
+  select(start_lon, start_lat, end_lon, end_lat) %>% 
+  purrr::transpose() %>% 
+  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
+  purrr::map(st_linestring) %>%
+  st_sfc(crs = 4326) %>%
+  st_sf(geometry = .)%>%
+  bind_cols(Test_Data) %>%
+  select(everything(), geometry) %>% 
+  # Set circles ('cus the world is not flat!)
+  st_segmentize(units::set_units(100, km)) %>% 
+  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
+  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
+  sf::st_sf(crs = 4326)
+
+#  unpack the coordinates of each feature 
+coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
+
+c(network_data_sf$line_id, network_data_sf$coords) %<-% (
+  coord_data %>% 
+    rename(subline_id = L1) %>%   
+    nest(-L2)
+)
+
+# Create data path for line connections
+paths <- network_data_sf %>% 
+  sf::st_set_geometry(NULL) %>% 
+  unnest() %>% 
+  mutate_at(vars(X,Y), round,2)
+
+# Plot it!
+
+country <- c("Fiji","Solomon Isl.","Tuvalu")
+
+ggplot() +
+  geom_sf(
+    # data = subset(world_map,name %in% country),
+    data = World_map,
+    size = 0.1, fill = "gray90", col = "gray90") +
+  geom_sf(
+    # data = subset(World_eez_sf, eez_name %in% country),
+    data = World_eez_sf,
+    size = 0.1, fill = "white", col = "gray70") +
+  geom_point(data = Test_Data,
+             # data = subset(Test_Data, source %in% country & target %in% source),
+             aes(start_lon,
+                 start_lat,
+                 col = n_tax_emer
+                 ),
+             size = 0.1,
+             col = "grey50",
+             # shape = 21,
+             stroke = 1.5) +
+  geom_path(data = paths,
+            # data = subset(paths, source %in% country),
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = n_tax_emer
+            ),
+            size = 0.3,
+            # arrow = arrow(length = unit(0.009, "npc")),
+            show.legend = NA) +
+  ggtheme_map() +
+  scale_colour_viridis(
+    name="Number of emerging stocks\nby 2050",
+    limits = c(0,40),
+    breaks = seq(0,40,10),
+    # option="magma",
+    direction = 1
+  ) +
+  ggsave("./Figures/figure_emerging_2050.png",
          width = 10,

### 2020-06-15 Scripts/Initial_Analysis.Rmd
Get SST data from GFDL85

@@
     }
-    Filepath <- paste("/Volumes/DATA/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")
+    Filepath <- paste("Z:/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")
   }
@@
 for(y in 1:length(Years)){
-
         Data <- getEnvVar(yr = Years[y],
@@
                           RCP = "rcp85"
-        ) %>% 
-          mutate(INDEX = all_Grid) %>% 
-          filter(INDEX %in% Grid$INDEX)
+        )
 
@@
 }
+
+colnames(F_BData) <- Years
+
+F_BData <- F_BData %>% 
+          rowid_to_column("index") %>% 
+  gather("year","temp",-index) %>% 
+  write_csv(.,
+            "SST_data.csv")
 

### 2020-06-15 Scripts/Initial_Analysis.Rmd
Changes in climate function

@@
           F_BData <- Data[1]
+          colnames(F_BData) <- 2020
 
@@
           F_BData <- cbind(Data[1],F_BData)
+          colnames(F_BData)[y] <- Years[y]
         }
@@
           rowid_to_column("index") %>% 
+  %>% 
+  mutate(temp = ifelse(temp == -9999,NA,temp)) %>% 
+  group_by(year) %>% 
+  summarise(
+    mean_temp = mean(temp,na.rm=T)
+  ) %>% 
   gather("year","temp",-index) %>% 
@@
             "SST_data.csv")
+
+```
+
+```{r temp_plot, eval = T, echo = T}
+
+sst_data <- my_path("D",T,"Climate","SST_data.csv",read = T)
+
+ggplot(sst_data) +
+  geom_line(
+    aes(
+      x = year,
+      y = mean_temp
+    )
+  )
 

### 2020-06-15 Scripts/Initial_Analysis.Rmd
Fixed cliamte data

--- 2020-06-15 Scripts/Initial_Analysis.Rmd
@@
 
-Years <-seq(2020,2100,1)
+Years <-seq(,2100,1)
 
@@
 
-          F_BData <- cbind(Data[1],F_BData)
-        }
-}
-
-colnames(F_BData) <- Years
+          F_BData <- cbind(F_BData,Data[1])
+          colnames(F_BData)[y] <- Years[y]
+        }
+}
 
@@
   gather("year","temp",-index) %>% 
+   mutate(temp = ifelse(temp == -9999,NA,temp)) %>% 
+  group_by(year) %>% 
+  summarise(
+    mean_temp = mean(temp,na.rm=T)
+  ) %>% 
   write_csv(.,
             "SST_data.csv")
+
+```
+
+```{r temp_plot, eval = T, echo = T}
+
+sst_data <- my_path("D",T,"Climate","SST_data.csv",read = T)
+
+ggplot(F_BData) +
+  geom_line(
+    aes(
+      x = as.numeric(year),
+      y = mean_temp
+    )
+  )
 

### 2020-06-15 Scripts/Initial_Analysis.Rmd
Produced first set of potential figures

@@
 # Load all sp in noe table
-Eme_Res <- bind_rows(lapply(Paths, fread))
+Eme_Res <- bind_rows(lapply(Paths, fread)) %>% 
+  filter(
+    n_ens == 10
+  )
 # warnings() for empty data species
@@
 stock_d_hist <- Eme_Res %>% 
-  filter(
-    n_ens == 10,
-    emerging_yr >= 2020
-  ) %>% 
   group_by(emerging_yr,taxon_key) %>%
@@
 eez_d_hist <- Eme_Res %>% 
-  filter(
-    n_ens == 10,
-    emerging_yr >= 2020
-  ) %>% 
   group_by(emerging_yr,eez_name) %>% 
@@
 
+##### Time Lapse plot
+
+```{r Time_lapse_plot,  eval = T, echo = T}
+
+# Per Stock
+stock_d <-  Eme_Res %>% 
+  left_join(Neighbours) %>% 
+    distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
+    group_by(taxon_key) %>% 
+    summarise(
+      taxon_emer = min(emerging_yr)
+    ) %>% 
+    group_by(taxon_emer) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "Stock"
+  ) %>% 
+    select(1,3,4)
+
+  #  EEZ data
+eez_d <-  Eme_Res %>% 
+  left_join(Neighbours) %>% 
+  filter(n_ens == 10) %>% 
+    distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
+    group_by(eez_name) %>% 
+    summarise(
+      taxon_emer = min(emerging_yr)
+    ) %>% 
+    group_by(taxon_emer) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "EEZ"
+  ) %>% 
+    select(1,3,4)
+
+# Plot Data
+plot_data <- stock_d %>% 
+  bind_rows(eez_d)
+
+  # filter(emerging_yr >= 2020) %>% 
+  ggplot() +
+    geom_area(data = subset(plot_data, level == "Stock"),
+              aes(
+                x = as.numeric(taxon_emer),
+                y = com_sum,
+                fill = level
+              ),
+              alpha= 0.8,
+              size=0.5,
+              colour="black"
+    )  +
+    geom_area(data = subset(plot_data, level == "EEZ"),
+              aes(
+                x = taxon_emer,
+                y = com_sum,
+                fill = level
+              ),
+              alpha= 0.8,
+              size=0.5,
+              colour="black"
+    ) +
+    # Comment out for option A (Change plot name!!)
+    geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
+  geom_text(aes(label="Total number of EEZs", x = 2010, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
+    geom_hline(yintercept = length(unique(Eme_Res$taxon_key)), color = "grey",linetype = "dashed") +
+  geom_text(aes(label="Total number of Taxa", x = 2010, y = length(unique(Eme_Res$taxon_key))+10), color = "grey70",size =3) +
+    scale_y_continuous("Cumulatuve Curve",
+                       breaks = seq(0,500,50),
+                       limits = c(0,500)
+    ) +
+    scale_x_continuous("Year of emergence",
+                       breaks = seq(2000,2100,10)
+    ) +
+    my_ggtheme_p() +
+    scale_fill_viridis(discrete = T) +
+  ggsave("./Figures/figure_accumulative_curve_B.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+```
+
 ##### Scatter plots
@@
 
-Years <-seq(,2100,1)
+Years <-seq(1951,2100,1)
 
@@
           F_BData <- Data[1]
-          colnames(F_BData) <- 2020
+          colnames(F_BData) <- Years[1]
 
@@
 
+### NEED TO RUN TIME LAPSE CHUNK FIRST TO GET ACCUMULATIVE DATA
+
+# Get sst_data
 sst_data <- my_path("D",T,"Climate","SST_data.csv",read = T)
 
-ggplot(F_BData) +
+sst_hist <- sst_data %>% 
+  filter(year < 2000) %>% 
+  group_by() %>% 
+  summarise(hist_sst =mean(mean_temp)) %>% 
+  pull()
+
+sst_change <- sst_data %>%
+  filter(year > 2020) %>% 
+  mutate(chng = round(mean_temp-sst_hist,2))
+
+  # Plot Data
+stock_d %>% 
+  bind_rows(eez_d) %>% 
+  filter(taxon_emer >= 2020) %>% 
+  rename(year = taxon_emer) %>% 
+  left_join(sst_change) %>% 
+  ggplot() +
   geom_line(
     aes(
-      x = as.numeric(year),
-      y = mean_temp
-    )
+      x = chng,
+      y = com_sum,
+      colour = level
+    )
+  ) +
+  geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
+  geom_text(aes(label="Total number of EEZs", x = 0.5, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
+    scale_y_continuous("Cumulatuve Curve",
+                       breaks = seq(0,450,50),
+                       limits = c(0,500)
+    ) +
+    scale_x_continuous("Change in SST (C)",
+                       breaks = seq(0,2,0.2),
+                       limits = c(0.3,2.3)
+    ) +
+    my_ggtheme_p() +
+    scale_colour_viridis(discrete = T) +
+  ggsave("./Figures/figure_temp.png",
+         width = 6,
+         height = 6,
+         units = "in"
   )
@@
             by = "emerging_yr") %>% 
+  rename(year = emerging_yr) %>% 
+  left_join(sst_data) %>% 
   ggplot() +
@@
     aes(
-      x = emerging_eez,
-      y = emerging_tax,
-      fill = emerging_yr,
-      colour = emerging_yr
+      x = mean_temp,
+      y = emerging_eez,
+      fill = year,
+      colour = year
     )

### 2020-06-17 Scripts/Manuscript.Rmd
First take at the intro

@@
 
-### Link paper un terms of SDGs
+The world is moving towards sustainable development and ocean-based activities plays a mayor role in achieveing international policies.
 
+One of the main goals of the United Nations Sustainable Development Goal No.14 (SDG 14-Life below water) is to achieve sustainable development of fisheries [@SDG14].
+(https://sustainabledevelopment.un.org/sdg14).
+
+Climate change is expected to reduce the production from global fisheries in about XX% to xx% [@Cheung; @Barange] compromisisng our capacity to reach SDG 14.
+
+Recent reseach show that the impacts of climate change on the productivity of marine wild fisheries can be offset with the proper reform of current fisheries management [@Gaines; @CheungClimateReadyFish].
+While bringing fisheries that are poorly managed to a sustainable level will likeley offset the negative impacts of climate change to fisheries production [@Gaines], it will not necessarilly solve the spatial consequences of species moving from one mangement area to another.
+
+Shifts in species distribution does not necessarilly mean a reduction in production (e.g. reduction in stock biomass) but rather a spatial re-distribution of the stock's biomass [@RefLiam?].
+
+Documented cases of stock shifts have resulted in international conflict and usustainable fishing rates by parties involved in the conflict.
+
+This is of specific concern to a set of 633 fished species that cross international borders, also known as transbounday species [@FishForVisa].
+
+With the implementation of arbritary spatial management boundaries, at least 633 species became transboudary representing an average catch (2005 - 2010) of 47.6 million tons and generating USD 65,483 million in fishing revenue [@FishorVisa].
+
+Transboudary fish stocks refere to those species that cross international jurisdictions known as Economic Exclusive Zones (EEZs) of neighbouring nations [@FAO].
+These stocks are responsable for an average catch (2005 - 2010) of 47.6 million tons generating USD 65,483 million in fishing revenue [@FishorVisa].
+
+Historically, environmentally-driven shits in shared stocks have cause problems and threathen stock's sustainabillity.
+**In early 1980's Canada and the US blablabla.** Many fish stocks managed by the European Union ICES regultaroy areas have been shifting since the 1970's [@EuropeShift].
+Such shifts have already result in confrontations between nations as evidenced by the expansion of the Atlantic mackerel (*Scomber scombrus*) distribution reaching the EEZ of Iceland [@Spijkers:2017ij].
+More recently, Chile and Peru have declare a cooperation strategy to mange the transboundary stock of Peruvean Anchovera [@Peru].
+Furthermore, as we continue to enter the $21^{st}$ century, commercial marine species are expected to continue shifting (mostly poleward) potentilly compromising shared stocks management.
+By the end of the century, all EEZs of the world are proyected to have at least one new transboundary species with some of them having over 10 new stocks [@Pinsky].
+Consequently, as species occupy new waters they leave historic areas and so, xx% of species are expected to exit EEZs by the end of che century [@Oremus].
+This entering and exicting national jurisdicions will complicate fisheries management potentially creating confllict between neighbouring nations [@Pinksy @Oremus:2020id].
+
+ So, preparing policy to species on the move is key to achieve sustainable development goals and proper governance of the world fisheries [@Pinksy].
+
+Undrstanding the degree by which shared stocks are going to change and the time of emergence is key for creating climate-ready fisheries policy.
+
+<!-- The concepts of transboundary stocks and EEZs were introduced in the early 1980's by the United Nations Law of the Sea (UCLOS) [@UnitedNations:1986tl].
+Under UNCLOS each costal nation has the right (and responsabillity) to manage wild fish stocks up to 200 nautical mailes of the coast.
+Exceptions exist when territorial waters are in dispute (e.g.
+South China sea[@Ref]) or when menagement agreements exist between nations [@Ref].
+In fact, UNCLOS incentives actions to cooperate on the management of shared stocks [Article 63, @UnitedNations:1986tl].
+Climate change is shifting the distribution of marine species to the extreem that some spatial policies implemented to manage fisheries are becoming obsolete [@Pololanza; @Jessica].
+-->
+
+## Link paper un terms of SDGs
 - UnitedNations2018 Blanchard:2017hl Singh:2017ds
 
-### Collaborative management of share stocks
-
+## Collaborative management of share stocks
 - Theat points
 - Concept of treaty estabillity
-- Oremus:2020id
 

### 2020-07-03 Scripts/Initial_Analysis.Rmd
re-ran untill 2005

@@
                           year,
-                          hist_y,
+                          hist_y = 2005,
                           data_type = "Catch",
@@
       ensemble_list,
-      dbem_import,
+      read_dbem_ens,
       taxon_key = taxon_key,
@@
       data_type = data_type,
-      path = path
+       model = "GFDL",
+      rcp = 85,
+      path = "",
     )
@@
   name <- paste("proportion_",taxon_key,".csv",sep="")
-  path_name <- paste(result_path,"Proportion/",name, sep="")
+  path_name <- my_path("R",extra_path = "Proportion")
 
@@
   clean_names() %>% 
-  select(-x1)
+  select(-v1)
 
@@
 # Year timeframe
-year_df <- tibble(year =  c(seq(1951,2000,1), seq(2001,2099,1)),
-                  time_period = c(rep("historic", 50), rep("future", 99))
-                  )
-
-# Data path
-path = DBEM_path
+year_df <- tibble(year =  c(seq(1951,2005,1), seq(2006,2099,1)),
+                  time_period = c(rep("historic", 55), rep("future", 94))
+                  )
 
 # List of ensemble members
-ensemble_list <- seq(102,103,1) # for testing
-# ensemble_list <- seq(102,111,1)
+# ensemble_list <- seq(102,103,1) # for testing
+ensemble_list <- seq(102,111,1)
 

### 2020-07-03 Scripts/Initial_Analysis.Rmd
re run model for speices that didnt work

@@
        model = "GFDL",
-      rcp = 85,
-      path = "",
+      rcp = "85",
+      path = "NA",
+      my_path = F
     )
@@
   name <- paste("proportion_",taxon_key,".csv",sep="")
-  path_name <- my_path("R",extra_path = "Proportion")
+  path_name <- paste(my_path("R",extra_path = "Proportion_n"),name,sep= "")
 
@@
 # EstPropChange(taxon_key = 600004,
-#               year = year_df,
+#               year = year_df$year,
 #               hist_y = 2000,
 #               data_type = "Catch",
-#               ensemble_list = ensemble_list,
+#               ensemble_list = 102,
 #               path = path,
 #               neighbours = neighbours)
+#
 
@@
 
+modeled_spp <- list.files(my_path("R", extra_path = "Proportion_n"))
+
+modeled_spp <- gsub("proportion_","",modeled_spp)
+modeled_spp <- gsub(".csv","",modeled_spp)
+
+spp_list <- spp_list[!spp_list %in% modeled_spp]
+
 ```
@@
 ####_______________________ ####
-
+gc()
 t_start <- Sys.time()
@@
   EstPropChange,
-  year <- year_df,
+  year <- year_df$year,
   data_type <- "Catch",

### 2020-07-03 Scripts/Initial_Analysis.Rmd
Ezstimat change in results

--- 2020-06-15 Scripts/Initial_Analysis.Rmd
@@
   # Reads taxon data
-  proportion_data <- fread(paste(result_path,"Proportion/","proportion_",taxon_key,".csv", sep=""))
+  proportion_data <- fread(paste(my_path("R",extra_path = "Proportion/"),"proportion_",taxon_key,".csv", sep=""))
 
@@
       summarise(
-        mean_spp = mean(ensemble_mean,na.rm=T) # average spp change from all countries sharing
+        mean_spp = mean(ensamble_mean,na.rm=T) # average spp change from all countries sharing
       )
@@
      "per_end" = n_end),
-            funs(round((((.-historical)/.))*100,2))
+            funs(round((((.-historical)/historical))*100,2))
    )
@@
 # Get taxon list
-Taxon_List <- list.files(paste(my_path("R"),"Propotion",sep=""))
+Taxon_List <- list.files(paste(my_path("R"),"Proportion",sep=""))
 
@@
 # Run results
-Proportion_Results <- PropChange(Taxon_List, links="y") %>% 
-  mutate(per_mid =ifelse(per_mid >= 25, 25,
-                           ifelse(per_mid <= -25,-25,per_mid)
+results <- PropChange(Taxon_List, links="y") %>% 
+  mutate(per_mid_plot =ifelse(per_mid >= 50, 50,
+                           ifelse(per_mid <= -50,-50,per_mid)
                            )
@@
 # The path
-path_world <- paste(data_path,"Spatial/SAU_Shapefile/",sep="")
+path_world <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile")
 
@@
 
-eez_centroid <- read_csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""))
+eez_centroid <- my_path("D",extra_path = "Spatial",name = "eez_centroids.csv",read = T)
 
@@
   mutate(
-    per_mid = ifelse(is.na(per_mid),0,per_mid)
+    per_mid = ifelse(is.na(per_mid_plot),0,per_mid_plot)
   ) %>% 
@@
 gc()# clean ram
+
 x <- ggplot() +
@@
     name="Average Gain of\nStock-Share Ratio (%)",
-    limits = c(0,25),
-    breaks = seq(0,25,5),
+    limits = c(0,max(Data$per_mid)),
+    breaks = seq(0,max(Data$per_mid),5),
     # option="magma",
@@
   ) +
-  ggsave("figure_connectionsc.png",
+  ggsave("figure_connections_n50.png",
          width = 10,

### 2020-07-03 Scripts/Initial_Analysis.Rmd

@@
       data_type = data_type,
-      path = path
+       model = "GFDL",
+      rcp = "85",
+      path = "NA",
+      my_path = F
     )
@@
   name <- paste("proportion_",taxon_key,".csv",sep="")
-  path_name <- paste(result_path,"Proportion/",name, sep="")
+  path_name <- paste(my_path("R",extra_path = "Proportion_n"),name,sep= "")
 
@@
 spp_list <- unique(Transboundary_spp$taxon_key)
+
+modeled_spp <- list.files(my_path("R", extra_path = "Proportion_n"))
+
+modeled_spp <- gsub("proportion_","",modeled_spp)
+modeled_spp <- gsub(".csv","",modeled_spp)
+
+spp_list <- spp_list[!spp_list %in% modeled_spp]
 

### 2020-07-06 Scripts/Initial_Analysis.Rmd
small changes

@@
   # Reads taxon data
-  proportion_data <- fread(paste(my_path("R",extra_path = "Proportion/"),"proportion_",taxon_key,".csv", sep=""))
+  proportion_data <- fread(paste(my_path("R",extra_path = "Proportion_2005/"),"proportion_",taxon_key,".csv", sep=""))
 
@@
       summarise(
-        mean_spp = mean(ensamble_mean,na.rm=T) # average spp change from all countries sharing
-      )
+        mean_spp = mean(ensemble_mean, na.rm = T) # average spp change from all countries sharing
+      ) %>% 
+      filter(!is.na(mean_spp))
 
@@
         mean_spp = mean(ensemble_mean,na.rm=T) # average spp change from all countries sharing
-      )
+      ) %>% 
+      filter(!is.na(mean_spp))
     }
@@
        n_spp = n(),
-       min_spp = min(mean_spp,na.rm=T),
-       max_spp = max(mean_spp,na.rm=T),
-       sd_spp = sd(mean_spp,na.rm=T), # for some weird reason it wont work if the order is fliped
-       mean_spp = mean(mean_spp,na.rm=T)
-     ) %>% 
-     select(eez_name:max_spp,mean_spp,sd_spp)
+       min_eez = min(mean_spp, na.rm=T),
+       max_eez = max(mean_spp, na.rm=T),
+       sd_eez = sd(mean_spp, na.rm=T), # for some weird reason it wont work if the order is fliped
+       mean_eez = mean(mean_spp,na.rm=T)
+     ) %>% 
+     select(eez_name:max_eez,mean_eez,sd_eez)
 
@@
        n_spp = n(),
-     min_spp = min(mean_spp,na.rm=T),
-     max_spp = max(mean_spp,na.rm=T),
-     sd_spp = sd(mean_spp,na.rm=T), # for some weird reason it wont work if the order is fliped
-     mean_spp = mean(mean_spp,na.rm=T)
-   ) %>% 
-   select(eez_name:max_spp,mean_spp,sd_spp)
+       min_eez = min(mean_spp, na.rm=T),
+       max_eez = max(mean_spp, na.rm=T),
+       sd_eez = sd(mean_spp, na.rm=T), # for some weird reason it wont work if the order is fliped
+       mean_eez = mean(mean_spp,na.rm=T)
+     ) %>% 
+     select(eez_name:max_eez,mean_eez,sd_eez)
 
@@
  proportion_change <- eez %>% 
-   gather("variable","value",n_spp:sd_spp) %>% 
+   gather("variable","value",n_spp:sd_eez) %>% 
    spread(time_step,value) %>% 
    mutate_at(vars(
-     "per_mid" = mid,
-     "per_end" = n_end),
+     "per_early" = early,
+     "per_mid" = mid
+     ),
             funs(round((((.-historical)/historical))*100,2))
@@
 # Get taxon list
-Taxon_List <- list.files(paste(my_path("R"),"Proportion",sep=""))
+Taxon_List <- list.files(paste(my_path("R"),"Proportion_2005",sep=""))
 
@@
                            ifelse(per_mid <= -50,-50,per_mid)
+                           ),
+         per_early_plot =ifelse(per_early >= 50, 50,
+                           ifelse(per_early <= -50,-50,per_early)
                            )
@@
   mutate(
+    per_early = ifelse(is.na(per_early_plot),0,per_early_plot),
     per_mid = ifelse(is.na(per_mid_plot),0,per_mid_plot)
@@
   filter(
-    variable == "mean_spp"#,
+    variable == "mean_eez"#,
     # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
@@
   unnest() %>% 
-  mutate_at(vars(X,Y), round,2)
+  mutate_at(vars(X,Y), round,2) %>% 
+  gather("time_step","per_chng",per_mid_plot:per_early_plot)
 
@@
 
+Data <- Data %>% 
+  gather("time_step","per_chng",per_mid_plot:per_early_plot)
+
 gc()# clean ram
+
+# time <- "per_mid_plot"
+time <- "per_early_plot"
 
@@
   geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray70") +
-  geom_point(data = subset(Data, per_mid < 0),
+  geom_point(data = subset(Data, time_step == time & per_chng < 0),
              aes(start_lon,
                  start_lat,
-                 col = per_mid),
+                 col = per_chng),
              size = 0.05,
@@
              stroke = 1.5) +
-  geom_path(data = subset(paths, per_mid > 0),
+  geom_path(data = subset(paths, time_step == time & per_chng > 0),
             aes(X, Y,
                 group = interaction(line_id,subline_id),
-                col = per_mid
+                col = per_chng
             ),
@@
             show.legend = NA) +
-  ggtheme_map() +
+  my_ggtheme_m() +
   scale_colour_viridis(
     name="Average Gain of\nStock-Share Ratio (%)",
-    limits = c(0,max(Data$per_mid)),
-    breaks = seq(0,max(Data$per_mid),5),
+    limits = c(0,max(Data$per_chng)),
+    breaks = seq(0,max(Data$per_chng),5),
     # option="magma",
@@
   ) +
-  ggsave("figure_connections_n50.png",
+  ggsave("./Figures/figure_connections_early_n50.png",
          width = 10,
@@
 
-# Load all sp in noe table
+# Load all sp in one table
 Eme_Res <- bind_rows(lapply(Paths, fread)) %>% 
@@
 eez_d_hist <- Eme_Res %>% 
-  group_by(emerging_yr,eez_name) %>% 
+  group_by(eez_name,emerging_yr) %>% 
   # summarise(n_tax = length(unique(taxon_key))) %>% 

### 2020-07-06 Scripts/Manuscript.Rmd
expanded intro

--- 2020-04-06 Scripts/Manuscript.Rmd
@@
 title: "Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change"
-subtitle: "[MANUSCRIPT DRAFT DO NOT SHARE WITHOUT PERMISSION]"
+subtitle: "Climate-ready transboundary fisheries management under Climate Change"
 author: "Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, U. Rashid Sumaila^1^,  Vicky Lam^1^, David VanderZwaag^3^, and William W.L. Cheung^1^"
@@
 ^3^Schulich School of Law, Dalhousie University"
-output: pdf_document
+# output: pdf_document
+output: word_document
 ---
 
-* corresponding author: j.palacios@oceans.ubc.ca
+```{r setup, eval = T, echo = F, warning = F, message = F, results = 'hide'}
+
+library(MyFunctions)
+
+#### Project's Library
+packages <- c(
+  "tidyverse" 
+  # "here", # for dbem_import `here()`
+  # "data.table", #dbem_import `fread()`
+  # "readxl", # for reading excell files
+  # "janitor", # for clearing names
+  # "geosphere", # estimate distances between points `distm()`
+  # "ggrepel",
+  # "zoo", # for average mean
+  # "parallel", # for mclapply,
+  # "sf", # for mapping
+  # "st", # for mapping
+  # "rgdal", #Spatial analysis
+  # "tools", #Spatial analysis 
+  # "zeallot", # for Juanito's map
+  # "gmt", # for estimating distances between points
+  # "viridis"
+)
+
+my_lib(packages)
+
+```
+
+<!-- * corresponding author: j.palacios@oceans.ubc.ca -->
 
@@
 
-The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, established political boundaries across the distribution of many fish species, creating shared species between nations.
-Thus, transboundary fish stocks are defined as those populations that cross the EEZ of two or more neighboring nations.
-Despite not implicit, UNCLOS incentivizes nations to cooperate in the management of transboundary fisheries.
-Today, most management plans emplaced are based on historical distribution of catches.
+The delineation of Economic Exclusive Zones (EEZs) by the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, established political boundaries across the distribution of many fish species, creating shared species between nations.
+Despite not implicit, UNCLOS incentivizes nations to cooperate in the management of transboundary fisheries and most management plans emplaced are based on historical distribution of catches.
 Climate change represents a challenge for the joint management of fish stocks.
 For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures, quota allocation, and fishery newcomers.
-Recent events suggests that even changes in species that contribute a relatively small proportion of the overall catch can lead to important international conflict.
-This is the case of the conflict between the EU, Norway, Iceland, and the Faroe Islands over Atlantic mackerel (*Scomber scombrus*) which represents 8% of the catch of involved nations.
-Thus, in this chapter I explore the climate-driven changes in stock-share-proportion (e.g.
-the proportion of the stock that each EEZ has) of the previously identified (Chapter One) transboundary species.
-Moreover, I determine the *time of emergence* of each fish stock, defined as the year in which the stock-share-proportion will change from its historical average.
-Preliminary results suggest that many neighboring EEZs will see changes in stock-share ratio between 1 and 10% with some exceeding 25%.
-The results of this chapter can inform fisheries management as it not only identifies at the EEZ level stocks that will be likely to create international conflict, but also estimates the time of emergence that can be used to prioritize management plans in the face of a changing world.
+Recent events suggests that even changes in species that contribute a relatively small proportion of the overall catch can lead to important international conflict when no legal framework is available.
+Thus, it is fundamental to prepare for cliamte change-driven impacts such as identifying the stocks that are going to change historical proportions within neighbouring EEZs, the degree by which they will cahnge, and the time when such change is going to happen.
+Only by adapting management plans and polocy that antecipates to these effects will international fisheries treaties be climate-ready.
+Here, we build on previous work to identify the transboundary stocks that will see changes in their stock-share-proportion (e.g.
+the proportion of the stock that each EEZ has) around the world.
+Moreover, we determine the time of emergence of each fish stock, defined as the year in which the stock-share-proportion will change from its historical range.
+Our results suggest that by the mid of the 21st century EEZs will see changes in stock-share ratio between 1 and 10% with some exceeding 25%.
+Most of these changes will hapen before 2050 urging policy to be re-formulated to account for the possible implications of these changes.
+By the end of the 21st century virtually all EEZs and transboundary sepeices will have experienced changes in stock-share-proportion relative to historical distributions.
+The results of this research has the potential to improve climate-ready fisheries if policy is modify to respond to the consequences presented here.
 
-# Introduction
+<!-- # Introduction -->
 
-### Link paper un terms of SDGs
+<!-- The world is moving towards sustainable development and ocean-based activities plays a mayor role in achieveing international policies. -->
 
-- UnitedNations2018 Blanchard:2017hl Singh:2017ds
+<!-- One of the main goals of the United Nations Sustainable Development Goal No.14 (SDG 14-Life below water) is to achieve sustainable development of fisheries [@SDG14]. -->
+<!-- (https://sustainabledevelopment.un.org/sdg14). -->
 
-### Collaborative management of share stocks
+<!-- Climate change is expected to reduce the production from global fisheries in about XX% to xx% [@Cheung; @Barange] compromisisng our capacity to reach SDG 14. -->
 
-- Theat points
-- Concept of treaty estabillity
-- Oremus:2020id
+<!-- Recent reseach show that the impacts of climate change on the productivity of marine wild fisheries can be offset with the proper reform of current fisheries management [@Gaines; @CheungClimateReadyFish].
+While bringing fisheries that are poorly managed to a sustainable level will likeley offset the negative impacts of climate change to fisheries production [@Gaines], it will not necessarilly solve the spatial consequences of species moving from one mangement area to another.
+-->
+
+<!-- Shifts in species distribution does not necessarilly mean a reduction in production (e.g. reduction in stock biomass) but rather a spatial re-distribution of the stock's biomass [@RefLiam?].  -->
+
+<!-- Documented cases of stock shifts have resulted in international conflict and usustainable fishing rates by parties involved in the conflict. -->
+
+<!-- This is of specific concern to a set of 633 fished species that cross international borders, also known as transbounday species [@FishForVisa].  -->
+
+<!-- With the implementation of arbritary spatial management boundaries, at least 633 species became transboudary representing an average catch (2005 - 2010) of 47.6 million tons and generating USD 65,483 million in fishing revenue [@FishorVisa].
+-->
+
+<!-- Transboudary fish stocks refere to those species that cross international jurisdictions known as Economic Exclusive Zones (EEZs) of neighbouring nations [@FAO].
+These stocks are responsable for an average catch (2005 - 2010) of 47.6 million tons generating USD 65,483 million in fishing revenue [@FishorVisa].
+-->
+
+<!-- Historically, environmentally-driven shits in shared stocks have cause problems and threathen stock's sustainabillity.
+**In early 1980's Canada and the US blablabla.** Many fish stocks managed by the European Union ICES regultaroy areas have been shifting since the 1970's [@EuropeShift].
+Such shifts have already result in confrontations between nations as evidenced by the expansion of the Atlantic mackerel (*Scomber scombrus*) distribution reaching the EEZ of Iceland [@Spijkers:2017ij].
+More recently, Chile and Peru have declare a cooperation strategy to mange the transboundary stock of Peruvean Anchovera [@Peru].
+Furthermore, as we continue to enter the $21^{st}$ century, commercial marine species are expected to continue shifting (mostly poleward) potentilly compromising shared stocks management.
+By the end of the century, all EEZs of the world are proyected to have at least one new transboundary species with some of them having over 10 new stocks [@Pinsky].
+Consequently, as species occupy new waters they leave historic areas and so, xx% of species are expected to exit EEZs by the end of che century [@Oremus].
+This entering and exicting national jurisdicions will complicate fisheries management potentially creating confllict between neighbouring nations [@Pinksy @Oremus:2020id].
+-->
+
+<!--  So, preparing policy to species on the move is key to achieve sustainable development goals and proper governance of the world fisheries [@Pinksy]. -->
+
+<!-- Undrstanding the degree by which shared stocks are going to change and the time of emergence is key for creating climate-ready fisheries policy. -->
+
+<!-- <!-- The concepts of transboundary stocks and EEZs were introduced in the early 1980's by the United Nations Law of the Sea (UCLOS) [@UnitedNations:1986tl].
+Under UNCLOS each costal nation has the right (and responsabillity) to manage wild fish stocks up to 200 nautical mailes of the coast.
+Exceptions exist when territorial waters are in dispute (e.g.
+South China sea[@Ref]) or when menagement agreements exist between nations [@Ref].
+In fact, UNCLOS incentives actions to cooperate on the management of shared stocks [Article 63, @UnitedNations:1986tl].
+Climate change is shifting the distribution of marine species to the extreem that some spatial policies implemented to manage fisheries are becoming obsolete [@Pololanza; @Jessica].
+--> -->
+
+<!-- ## Link paper un terms of SDGs -->
+<!-- - UnitedNations2018 Blanchard:2017hl Singh:2017ds -->
+
+<!-- ## Collaborative management of share stocks -->
+<!-- - Theat points -->
+<!-- - Concept of treaty estabillity -->
 
@@
 
-We coupled the DBEM with the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project species distribution from 1951 to 2100 under cliamte change [@Dunne:2012bo; @Dunne:2013db].
+We coupled the DBEM with the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project species distribution from 1951 to 2100 under climate change [@Dunne:2012bo; @Dunne:2013db].
 We run the DBEM simulating biomass and MCP from 1950 to 2100 under a “worst-case” climate change scenario accroding to the Representative Concentration Pathway 8.5 (RCP 8.5) used by the Intergovernmental Panel on Climate Change (IPCC) [@IPCC:2014bz; Riahi:2011dk].
-According to the GFDL-ESM2M, under the RCP 8.5 scenario atmospheric temperature will increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial.
-We run the DBEM under ten different GFDL-ESM2M ensamble members [@Roggers:2015jq].
-While all ensamble members follow the same RCP, they all constitute different initial Earth conditions.
-As a consequence, for each GFDL-ESM2M-ensamble the Earth system is infinitesimally perturbed resulting in an unique atmosphere and ocean state at each point in time.
-Thus, each ensamble has a different state of internal variability representing a plausible future under RCP 8.5 [@Roggers:2015jq].
+According to the GFDL-ESM2M, under the RCP 8.5 scenario atmospheric temperature will increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a species distribution, we have to understand both the historical as well as the future climate-driven distribution variabillity, in order to infer the differences.
+We do this by following an ensemble member approach where each member constitutes a different initial Earth system condition [@Roggers:2015jq; @Frolicher:2009ge].
+Thus, for our experiment, in each of the ten GFDL-ESM2M-ensambles the Earth system initial condition is infinitesimally different resulting in an unique atmosphere and ocean state at each point in time.
+This results in a different state of internal variability for each ensamble representing a plausible future under RCP 8.5 [@Roggers:2015jq].
+Following this approach allows us to estimate a secular trend for the natural variabillity (e.g.
+backround noise).
+At the same time, by having multiple initial-condition ensemble simulations of the GFDL-ESM2M we were able to remove the natural variability of a stock's distributio by averaging the secular trend over all ensemble members [@Roggers:2015jq].
 
@@
 
-The Food and Agricultural Organization (FAO) of the UN define transboundary species as those that freely swim within the EEZ of neighbouring nations [@REFFAO].
+The Food and Agricultural Organization (FAO) of the UN define transboundary stocks as those that freely swim across the EEZ of neighbouring nations [@REFFAO].
 Accordingly, we included in this analysis 633 marine exploited transboundary species representing 80% of the catch taken from the world's EEZs [@Palacios-Abrantes:xxxx].
@@
 
-In game theory, a “threat point” is known as the minimum payoff that a player is willing to receive in order to cooperate with other players [@Sumaila:E&S].
-We estimate each stock's stock share ratio (e.g.
-a players threat point) defined as the minimum proportion of the catch that each country will acept if they were to cooperate in the management of transboundary stocks [@Palacios-Abrantes:E&S].
+In game theory, a threat point is known as the minimum payoff that a player is willing to receive in order to cooperate with other players.
+Thus, if a player's payoff is lower than the theat point, such player will liekly break off any collaboration agreement and proceed to act solo [@Sumaila:E&S].
+Here, we adopted the concept of threat point as the minimum proportion of the catch that each country will acept if they were to cooperate in the management of a transboundary stock [@Palacios-Abrantes:E&S] hereafter referred as stock share ratio (SSR).
 For each stock and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighbouring EEZs, and then calculated the proportion that each EEZ had [@Palacios-Abrantes:xxxx].
-We then average the results in three time periods in order tho reduce climate variabillity within the projections.
+We then average the results in three time periods in order to reduce climate variabillity within the projections.
 The first time period ($t_h$) comprises (1951 - 2005) and is assumed as the historic distribution of each species, thus representing each neighbouring nation's threat point.
-This time frame matches the historic climate data that the GFDL uses to project climate change [@Dunne:2012bo; @Dunne:2013db], and has previouslly been used as a reference point in fisheries studies [@REF].
+This time frame matches the historic climate data that the GFDL uses to project climate change [@Dunne:2012bo; @Dunne:2013db].
 We then selected two future periods; early $21^st$ century as the average of 2021 to 2040 ($t_e$), and mid $21^st$ century as the average of 2041 to 2060 ($t_m$).
 We choose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing) and SDG 1.2 (poverty reduction), aimed to be achieved by 2020 and 2030, respectively [@Singh:2017ds].
-We then estimate the percentage change in stock share ratio of each future time period relative to the historic value.
-The analysis was repplicated for each of the ten ensamble members and ensamble variation captured by averaging results among ensamble members (mean $\pm$ standar deviation).
+We then estimate the percentage change in sthe SSR of each future time period relative to the historic time period.
+To address the natural spatal variabillity of the stock we only considered a SSR chgange when the future SSR was higher or lower than the average historic SSR $\pm$ 1 s.d.
+The analysis was repplicated for each of the ten ensamble members capturing variation by averaging results among ensamble members (mean $\pm$ standar deviation).
 
@@
 
-An important question is estimating the time in which stock share ratio is going to change from a historical variablillty in order for policy to cope with the change.
-
-For that we developed an index of transboundary range shift (*I*), fisheries and climate change based on a species distribution centroid:
+Knowing the time (e.g.
+year) in whitch SSR is going to diverge from a historical variablillty can be key for policy to prepare to such change.
+In here we call this event time of emergence and defined it as the moment in time when the future variation in SSR becomes larger than the historical trend variabillity [@Roggers:2015jq].
+We emplot the ten GFDL ensable members described above, and developed an index of transboundary range shift, fisheries and climate change (*I*).
+The index is based on the variation of a species distribution centroid in realtion to the EEZs it ocuppys as follows:
 
@@
 
-where $D_A$ and $D_B$ represent the distribution centroid of a species in EEZs A and B, respectiveley for each time step from 2000 to 2100 (*t*) and *sd* is the standar deviation of the historical (1951 - 2000) centroid distribution for $D_A$ and $D_B$.
+where $D_A$ and $D_B$ represent the mean (among ensamble members) distribution centroid of a species in EEZs A and B, respectiveley for each future time step from 2006 to 2100 (*t*) and *sd* is the standar deviation of the historical (*h*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
 
-In cases where the species is distributed in more than 2 neighbouring EEZs (e.g. *xxx*), the index is as follows:
+<!-- In cases where the species is distributed in more than 2 neighbouring EEZs (e.g. *xxx*), the index is as follows: -->
 
-$$I =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h})}) |  \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{C,y}}{sd(D_{B,h})}) | \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{Cy}}{sd(D_{B,h})})/N$$
+<!-- $$I =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h})}) |  \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{C,y}}{sd(D_{B,h})}) | \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{Cy}}{sd(D_{B,h})})/N$$ -->
 
-This way, for two neighbouring EEZs sharing a stock, if the centroid shifts towards EEZ A (*DA* decreases) but away from EEZ A B (*DB* increases), EEZ A may catch a bigger share over B and the risk of dispute increases.
+This way, for two neighbouring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases.
 Consequently, if the centroid shifts are within historical variability, risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability, the risk of dispute is higher.
@@
 
-### Statistical analysis
-
 All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *wesanderson* [@Packagewesanderson:2018vn], *rfishbase* [@PackagerfishbaseR:2019th], *R.matlab* [@PackageRmatlabRe:2018un], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], and *zoo* [@PackagezooSInfr:2019tm].
@@
 
-Things to consider:
-
-- Disputed territories (e.g. China)
-
 # Results
 
+```{r Load_Results, eval = TRUE,  echo = FALSE, include = FALSE}
+
+prop_path <- "/Volumes/HALL2000/Data/TransEmergence/Results/Proportion_2005/"
+prop_names  <- list.files(prop_path)
+
+prop_read <- paste(prop_path,prop_names,sep="")
+
+results <- bind_rows(lapply(prop_read, data.table::fread))
+
+```
+
+## Changes in Stock Share Ratio
+
+```{r Load_Results, eval = TRUE,  echo = FALSE, include = FALSE}
+
+# Mid century
+n_eez <- prop_r %>%
+  filter(variable == "mean_spp" & 
+           !is.na(per_mid)
+         ) %>% 
+  pull(eez_name) %>% 
+  unique() %>% 
+  length()
+
+n_eez_per <- n_eez/275*100
+
+### Max values
+max_mid <- prop_r %>% 
+  filter(variable == "max_spp" & 
+           !is.na(per_mid)
+         )
+
+### Average
+average_mid <- prop_r %>% 
+  filter(variable == "mean_spp" & 
+           !is.na(per_mid)
+         ) %>% 
+  pull(per_mid) %>% 
+  mean()
+
+```
+
+Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by mid $21^{st}$ century (Fig.
+1).
+The change in SSR will range from gains of over 100% (e.g.
+Honduras' SSR of *species* shared with Mexico is projected to go from 28% to 60%) to loses of up to -80% of SSR (e.g.
+Nauru' SSR of *species* shared with the Marshall Isl.
+is projected to decline from 14% to 3%).
+By the end of the century, SSR changes will be more exrteme with 13 countries gaining over 100% SSR and Palau loosing -93.61% of the SSR of *spp* to the	Philippines (from 20% to 1.2%).
+
+```{r SSR_gain, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Average transboundary species SSR gain of neighbouring EEZs by 2030 (2021-2040) relative to 1951-2005 under RCP 8.5.
+Arrows go from the loosing EEZ (point) to the winner (arrowhead)"}
+
+img <- readPNG("../Figures/figure_connections_n50.png")
+grid.raster(img)
+
+```
+
+![Average transboundary species SSR gain of neighbouring EEZs by 2030 (2021-2040) relative to 1951-2005 under RCP 8.5.
+Arrows go from the loosing EEZ (point) to the winner (arrowhead)](../Figures/figure_connections_n50.png)
+
+## Time of Emergence Index
 # Discussion

### 2020-07-06 Scripts/Initial_Analysis.Rmd
changes to functions to work on hall

--- 2020-07-03 Scripts/Initial_Analysis.Rmd
@@
       taxon_key = taxon_key,
-      year = year,
+      year = year_df$year,
       data_type = data_type,
@@
       rcp = "85",
-      path = "NA",
+      # path = "/Volumes/DATA/DATA",
       my_path = F
@@
     group_by(taxon_key,eez_name,eez_neighbour,time_step) %>%
-    summarise_at(vars("temporal_mean"),
-                 c(ensemble_mean=mean,
-                   ensemble_sd=sd),
-                 na.rm=T
+    summarise(
+      ensemble_mean = mean(temporal_mean, na.rm = T),
+      ensemble_sd = sd(temporal_mean, na.rm = T),
+      ensemble_n = n()
     )
@@
   name <- paste("proportion_",taxon_key,".csv",sep="")
-  path_name <- paste(my_path("R",extra_path = "Proportion_n"),name,sep= "")
+  path_name <- paste(my_path("R",extra_path = "Proportion_2005"),name,sep= "")
 
@@
 # Last year of "historic" data
-Historic_year = 2000
+Historic_year = 2005
 
@@
   # Reads taxon data
-  proportion_data <- fread(paste(my_path("R",extra_path = "Proportion/"),"proportion_",taxon_key,".csv", sep=""))
+  proportion_data <- fread(paste(result_path,"Proportion/","proportion_",taxon_key,".csv", sep=""))
 
@@
      "per_end" = n_end),
-            funs(round((((.-historical)/historical))*100,2))
+            funs(round((((.-historical)/.))*100,2))
    )
@@
 # Get taxon list
-Taxon_List <- list.files(paste(my_path("R"),"Proportion",sep=""))
+Taxon_List <- list.files(paste(my_path("R"),"Propotion",sep=""))
 
@@
 # Run results
-results <- PropChange(Taxon_List, links="y") %>% 
-  mutate(per_mid_plot =ifelse(per_mid >= 50, 50,
-                           ifelse(per_mid <= -50,-50,per_mid)
+Proportion_Results <- PropChange(Taxon_List, links="y") %>% 
+  mutate(per_mid =ifelse(per_mid >= 25, 25,
+                           ifelse(per_mid <= -25,-25,per_mid)
                            )
@@
 # The path
-path_world <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile")
+path_world <- paste(data_path,"Spatial/SAU_Shapefile/",sep="")
 
@@
 
-eez_centroid <- my_path("D",extra_path = "Spatial",name = "eez_centroids.csv",read = T)
+eez_centroid <- read_csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""))
 
@@
   mutate(
-    per_mid = ifelse(is.na(per_mid_plot),0,per_mid_plot)
+    per_mid = ifelse(is.na(per_mid),0,per_mid)
   ) %>% 
@@
 gc()# clean ram
-
 x <- ggplot() +
@@
     name="Average Gain of\nStock-Share Ratio (%)",
-    limits = c(0,max(Data$per_mid)),
-    breaks = seq(0,max(Data$per_mid),5),
+    limits = c(0,25),
+    breaks = seq(0,25,5),
     # option="magma",
@@
   ) +
-  ggsave("figure_connections_n50.png",
+  ggsave("figure_connectionsc.png",
          width = 10,

### 2020-07-07 Scripts/Initial_Analysis.Rmd
Created new figures for 2005 year of proportion

--- 2020-07-06 Scripts/Initial_Analysis.Rmd
@@
 results <- PropChange(Taxon_List, links="y") %>% 
+  filter(!is.na(eez_neighbour)) %>% 
   mutate(per_mid_plot =ifelse(per_mid >= 50, 50,
@@
          )
+
+# For dat availabillity
+ # write_csv(results,
+          # "proportion_change_2005.csv")
 
@@
     # option="magma",
-    direction = 1
+    direction = -1
   ) +
@@
 
+##### Histogram of change
+
+```{r hist_change}
+
+ results %>% 
+  gather("time_step","change",per_early,per_mid) %>% 
+  filter(variable == "mean_eez" & 
+           !is.na(change)
+         ) %>%
+ggplot() +
+  geom_density(
+               aes(
+                 x = ifelse(change >= 100,100,change),
+                 # x = change,
+                 fill = ifelse(time_step == "per_early","q",time_step)
+                 # fill = time_step,
+
+               ),
+               # stat = "count",
+               # alpha= 0.8,
+               size=0.1,
+               # position = "stack"
+  ) +
+  scale_y_continuous("Density",
+                     # breaks = seq(0,90,10),
+                     # limits = c(0,100)
+  ) +
+  scale_x_continuous("Percentage Change",
+                     breaks = seq(-75,500,25)
+  ) +
+  my_ggtheme_p() +
+  theme(legend.position = "right") +
+  scale_fill_viridis(discrete = T, 
+                     option = "plasma",
+                     alpha = 0.5,
+                     name = "Time Step", labels = c("Middle", "Early")
+                     ) +
+  coord_flip() +
+  ggsave("./Figures/hist_change.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
+```
+
 ####  Year of Eergence Analysis
@@
 # Get taxon list
-Taxon_List <- list.files(my_path("R",extra_path = "Emergence_yr"))
+Taxon_List <- list.files(my_path("R",extra_path = "Emergence_2005"))
 
@@
 
-Paths <- paste(my_path("R",extra_path = "Emergence_yr"),Taxon_List,"_emergence.txt",sep="")
+# Neighbour list and their respective id (see chunk 3)
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)
+
+Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List,"_emergence.txt",sep="")
 
@@
 stock_d_hist <- Eme_Res %>% 
-  group_by(emerging_yr,taxon_key) %>%
+  group_by(emerging_yr,taxon_key,Tresh) %>%
   # summarise(n_tax = length(unique(eez_name))) %>%
@@
     level = "stock"
-  ) %>% 
-  select(1,3,4)
+  ) 
   # ggplot() +
@@
 eez_d_hist <- Eme_Res %>% 
-  group_by(eez_name,emerging_yr) %>% 
+  group_by(eez_name,emerging_yr,Tresh) %>% 
   # summarise(n_tax = length(unique(taxon_key))) %>% 
@@
   ) %>% 
+  ungroup() %>% 
   select(1,3,4)
@@
 hist_data <- stock_d_hist %>% 
-  bind_rows(eez_d_hist)
+  bind_rows(eez_d_hist) %>% 
+  filter(Tresh != "NA")
 
@@
     my_ggtheme_p() +
-    scale_fill_viridis(discrete = T)
+    scale_fill_viridis(discrete = T) +
+  facet_wrap(~Tresh)
 
@@
     distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
-    group_by(taxon_key) %>% 
+    group_by(taxon_key,Tresh) %>% 
     summarise(
@@
     ) %>% 
-    group_by(taxon_emer) %>% 
+    group_by(taxon_emer,Tresh) %>% 
   summarise(
@@
     distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
-    group_by(eez_name) %>% 
+    group_by(eez_name,Tresh) %>% 
     summarise(
@@
 plot_data <- stock_d %>% 
-  bind_rows(eez_d)
+  bind_rows(eez_d) %>% 
+  filter(Tresh != "NA")
 
@@
     scale_fill_viridis(discrete = T) +
-  ggsave("./Figures/figure_accumulative_curve_B.png",
-         width = 6,
-         height = 6,
-         units = "in"
-  )
+    facet_wrap(~Tresh)
+  # ggsave("./Figures/figure_accumulative_curve_B.png",
+  #        width = 6,
+  #        height = 6,
+  #        units = "in"
+  # )
 ```

### 2020-07-26 Scripts/Initial_Analysis.Rmd
Changed code so that the spp's centroid is the average of the girds >0.95 percentile

@@
 ```{r setup, eval = T, echo = F, warning = F, message = F, results = 'hide'}
-
-#### READ ME !!! ####
-# Run this chunk before knit so you make sure you have all pkgs installed in R
 
@@
                          Neighbours,
-                         eez_centroid
+                         eez_centroid,
+                         p = 0.95 # percentile of data for central distribution
                          ){
@@
     data_type = data_type,
-    ensemble = ensemble
+    ensemble = ensemble,
+    path = my_path("G","DBEM")
   )
@@
 
-      # Centroid is defined as the tile with the higest species value
+      # Function to get percentiles
+      # https://www.tidyverse.org/blog/2020/03/dplyr-1-0-0-summarise/
+      quibble <- function(x, q = p) {
+        tibble(x = quantile(x, q), q = q)
+      }
+
+      # Get the percentile treshold per year, ensemble and neighbour id
+      percentile <- Sub_Trans_Spp %>% 
+        left_join(EEZ_ids, by = "eez_name") %>% 
+        group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
+        do(quibble(.$value, p))
+
        Centroids <- Sub_Trans_Spp %>% 
          left_join(EEZ_ids, by = "eez_name") %>% 
+         left_join(percentile,
+                   by = c("taxon_key","year","ensemble","neighbour_id")
+                   ) %>% 
+         mutate(to_filter = ifelse(value < x,"out","keep")) %>% 
+         filter(to_filter == "keep") %>% 
          group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
-        top_n(1,value) %>% 
-        distinct(value,year, .keep_all = TRUE) %>%  # removes duplicates when tile is in between EEZs
+         summarise_at(vars(lon,lat),mean) %>% 
          left_join(Sub_EEZ_centroid, by = "taxon_key") %>%
-        select(eez_name =eez_name.y, everything() ,-eez_name.x) %>% 
+         # select(eez_name =eez_name.y, everything() ,-eez_name.x) %>% 
          # Estimate distances between spp centroid and EEZ centroid
@@
     }
-
     # Determine the ids
@@
     filter(year < Hist_y) %>% 
-    group_by(taxon_key) %>% 
-    summarise_at(vars("RMean"),
+    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per ensemble
+    summarise_at(vars("trend"),
+                 c(trend= mean)
+                 ) %>% 
+    group_by(taxon_key,neighbour_id) %>% #Capture ensemble variation
+    summarise_at(vars("trend"),
                  c(hist_mean=mean,
@@
     n_ens == 10
-  )
+  ) %>% 
+  left_join(Neighbours,
+            by = c("eez_name","eez_neighbour"))
 # warnings() for empty data species
@@
   st_transform(crs = 4326) %>% # 4326
-  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)
+  st_simplify(preserveTopology = TRUE, dTolerance = 100)
 
@@
 
-# Per Stock
+# First time a Species becomes emergent
+Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # Tresh == "Tresh_one"
+  ) %>% 
+  group_by(Tresh,taxon_key,emerging_yr) %>% 
+  summarise(
+    n_trans = length(unique(eez_name))
+  ) %>% 
+  group_by(Tresh,taxon_key) %>% 
+  summarise(
+    ntrans = min(emerging_yr)
+  ) %>% 
+  group_by(Tresh,ntrans) %>% 
+  summarise(
+    n_taxa = length(unique(taxon_key))
+  ) %>% 
+  arrange(-n_taxa) %>% 
+  # View()
+  # group_by(Tresh) %>% 
+  # summarise(sum(n_taxa))
+  ggplot() +
+  geom_line(
+    aes(
+      x = ntrans,
+      y = n_taxa,
+      color = Tresh
+    )
+  )
+
+# EEZ, first time an EEZ has a emergent stock
+Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # Tresh == "Tresh_one"
+  ) %>% 
+  group_by(Tresh,eez_name,emerging_yr) %>% 
+  summarise(
+    n_trans = length(unique(taxon_key))
+  ) %>% 
+  group_by(Tresh,eez_name) %>% 
+  summarise(
+    ntrans = min(emerging_yr)
+  ) %>% 
+  group_by(Tresh,ntrans) %>% 
+  summarise(
+    n_taxa = length(unique(eez_name))
+  ) %>% 
+  arrange(-n_taxa) %>% 
+  View()
+  # group_by(Tresh) %>%
+  # summarise(sum(n_taxa))
+  # ggplot() +
+  # geom_line(
+  #   aes(
+  #     x = ntrans,
+  #     y = n_taxa,
+  #     color = Tresh
+  #   )
+  # )
+
+# Numer of EEZs where stocks will emerge
 stock_d_hist <- Eme_Res %>% 
+  distinct(taxon_key,neighbour_id,Tresh, .keep_all = TRUE) %>%
   group_by(emerging_yr,taxon_key,Tresh) %>%
-  # summarise(n_tax = length(unique(eez_name))) %>%
-  summarise(n()) %>% 
+  summarise(n_tax = length(unique(eez_name))) %>%
+  # summarise(n()) %>% 
   mutate(
-    level = "stock"
-  ) 
-  # ggplot() +
-  # geom_histogram(
-  #   aes(
-  #     x = emerging_yr
-  #   ),
-  #   stat = "count"
-  # )
+    level = "EEZs per Spp"
+  )  %>% 
+  ungroup() %>% 
+  select(-taxon_key)
 
@@
 eez_d_hist <- Eme_Res %>% 
-  group_by(eez_name,emerging_yr,Tresh) %>% 
+  distinct(taxon_key,neighbour_id,Tresh, .keep_all = TRUE) %>%
+  group_by(emerging_yr,eez_name,Tresh,neighbour_id) %>%  
   # summarise(n_tax = length(unique(taxon_key))) %>% 
@@
   mutate(
-    level = "EEZ"
+    level = "Spp per EEZ"
   ) %>% 
   ungroup() %>% 
-  select(1,3,4)
-
-hist_data <- stock_d_hist %>% 
+  select(-eez_name)
+
+stock_d_hist %>% 
   bind_rows(eez_d_hist) %>% 
-  filter(Tresh != "NA")
-
+  filter(Tresh != "NA") %>% 
   ggplot() +
-  geom_density(data = subset(hist_data, level == "EEZ"),
-    aes(
-      x = emerging_yr,
-      fill = level
-    ),
-    stat = "count",
-    alpha= 0.8,
-    size=0.5,
-  ) +
-  geom_density(data = subset(hist_data, level == "stock"),
+  geom_density(
     aes(
@@
   scale_y_continuous("Frequency",
-                       breaks = seq(0,90,10),
-                       limits = c(0,100)
+                     # breaks = seq(0,300,50),
+                     # limits = c(0,300)
   ) +
   scale_x_continuous("Year of emergence",
-                       breaks = seq(2000,2100,10)
+                     # breaks = seq(2006,2100,25),
+                     # limits = c(2000,2100)
   ) +
@@
   scale_fill_viridis(discrete = T) +
-  facet_wrap(~Tresh)
-
-```
-
-##### Time Lapse plot
+  facet_wrap(~Tresh) +
+  ggsave("./Figures/hist_freq.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
+```
+
+##### Accumulative plots
 
@@
   left_join(Neighbours) %>% 
-    distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
+    # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
     group_by(taxon_key,Tresh) %>% 
@@
   ) %>% 
+  group_by(Tresh) %>% 
   mutate(
@@
     level = "Stock"
-  ) %>% 
-    select(1,3,4)
+  )
 
@@
   filter(n_ens == 10) %>% 
-    distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
+    # distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
     group_by(eez_name,Tresh) %>% 
@@
     level = "EEZ"
-  ) %>% 
-    select(1,3,4)
+  )
 
@@
   bind_rows(eez_d) %>% 
-  filter(Tresh != "NA")
+  filter(Tresh != "NA") %>% 
+  mutate(
+    total = ifelse(level == "Stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$eez_name))),
+    per = com_sum/total*100
+  ) ;plot_data
 
@@
     ) +
+    facet_wrap(~Tresh) + 
     # Comment out for option A (Change plot name!!)
-    geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
-  geom_text(aes(label="Total number of EEZs", x = 2010, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
-    geom_hline(yintercept = length(unique(Eme_Res$taxon_key)), color = "grey",linetype = "dashed") +
-  geom_text(aes(label="Total number of Taxa", x = 2010, y = length(unique(Eme_Res$taxon_key))+10), color = "grey70",size =3) +
+  #   geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
+  # geom_text(aes(label="Total number of EEZs", x = 2010, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
+  #   geom_hline(yintercept = length(unique(Eme_Res$taxon_key)), color = "grey",linetype = "dashed") +
+  # geom_text(aes(label="Total number of Taxa", x = 2010, y = length(unique(Eme_Res$taxon_key))+10), color = "grey70",size =3) +
     scale_y_continuous("Cumulatuve Curve",
-                       breaks = seq(0,500,50),
-                       limits = c(0,500)
+                       breaks = seq(0,600,50),
+                       limits = c(0,600)
     ) +
     scale_x_continuous("Year of emergence",
-                       breaks = seq(2000,2100,10)
+                       breaks = seq(2005,2100,10),
+                       limits = c(2005,2100)
     ) +
@@
     scale_fill_viridis(discrete = T) +
-    facet_wrap(~Tresh)
-  # ggsave("./Figures/figure_accumulative_curve_B.png",
-  #        width = 6,
-  #        height = 6,
-  #        units = "in"
-  # )
+  ggsave("./Figures/figure_accumulative_curve_ab.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
+  ### Percentage accumulation
+
+  # filter(emerging_yr >= 2020) %>% 
+  ggplot() +
+    geom_area(data = subset(plot_data, level == "Stock"),
+              aes(
+                x = as.numeric(taxon_emer),
+                y = per,
+                fill = level
+              ),
+              alpha= 0.8,
+              size=0.5,
+              colour="black"
+    )  +
+    geom_area(data = subset(plot_data, level == "EEZ"),
+              aes(
+                x = taxon_emer,
+                y = per,
+                fill = level
+              ),
+              alpha= 0.8,
+              size=0.5,
+              colour="black"
+    ) +
+    scale_y_continuous("Cumulatuve Curve",
+                       breaks = seq(0,100,25),
+                       limits = c(0,100)
+    ) +
+    scale_x_continuous("Year of emergence",
+                       breaks = seq(2005,2100,10),
+                       limits = c(2005,2100)
+    ) +
+    my_ggtheme_p() +
+    scale_fill_viridis(discrete = T) +
+    facet_wrap(~Tresh) + 
+  ggsave("./Figures/figure_accumulative_curve_per.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
 ```
@@
 ##### Mapping the time of emergence
+
+###### Just mean time
+
+```{r emergence_map, eval = T}
+
+# Per EEZs
+eez_d_hist <- Eme_Res %>% 
+  group_by(eez_name,Tresh) %>% 
+  # summarise(n_tax = length(unique(taxon_key))) %>% 
+  summarise(
+    mean_yr = mean(emerging_yr,na.rm=T),
+    sd_yr = sd(emerging_yr,na.rm=T)) %>% 
+  gather("metric","value",mean_yr:sd_yr)
+
+### Random plot
+
+ggplot(eez_d_hist) +
+  geom_density(
+    aes(
+      x = value,
+      fill = Tresh
+    ),
+    bins = 10,
+    alpha = 0.8
+  ) +
+  facet_wrap(~metric,
+             scales = "free") +
+  scale_fill_viridis(discrete = T) +
+  my_ggtheme_p() +
+  ggsave("./Figures/hist_mean.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
+map <- World_eez_sf %>% 
+  left_join(eez_d_hist,
+            by = "eez_name") %>% 
+  # filter(Tresh == "Tresh_one") %>%
+  ggplot() +
+  geom_sf(
+    aes(
+      fill = mean_yr
+    )
+  ) +
+  geom_sf(data =World_map, aes(), size = 0.1, fill = "gray90", col = "gray90") +
+  scale_fill_viridis() +
+  my_ggtheme_m() +
+  facet_wrap(~Tresh,
+             ncol = 1) +
+  ggsave("./Figures/emer_map_T12.png",
+         width = 12,
+         height = 10,
+         units = "in"
+  )
+
+```
+
+###### With relation to neighbours
 

### 2020-07-26 Scripts/Initial_Analysis.Rmd
Changed trans_index function to account for noise in two steps, time variation and model variarion

@@
                        Id, # A numeric value representing a neighbouring interaction (ONE!)
-                       Hist_y = 2000,
+                       Hist_y = 2005,
                        Plot = F){ # The LAST year of Historical data
@@
   historic_sd <- Trans_index %>% 
-    filter(year < Hist_y) %>% 
+    filter(year <= Hist_y) %>% 
     group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per ensemble
-    summarise_at(vars("trend"),
+    summarise_at(vars("trans_index"),
                  c(trend= mean)

### 2020-07-26 Scripts/Initial_Analysis.Rmd
Changes to functions

--- 2020-07-07 Scripts/Initial_Analysis.Rmd
@@
   EstPropChange,
-  year <- year_df$year,
-  data_type <- "Catch",
+  year = year_df$year,
+  data_type = "Catch",
   ensemble_list = ensemble_list,
@@
 
-# Hall 1000 ran it in 5.33 hours
+# Hall 1000 ran it in 5.33 hours past
+# Hall 1000 ran it in  hours (July 2020)
 
@@
         mutate(
-          distance = geodist(longitude, # eez longitude
-                             latitude,  # eez latitude
+          distance = geodist(latitude,  # eez latitude
+                             longitude, # eez longitude
+                             lat, # species latitude
                              lon,  # species longitude
-                             lat, # species latitude
                              units="km")
@@
     ) %>% 
-    filter(year > 1960) %>%
+    # filter(year > 1960) %>%
     select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)
@@
                    ensemble_sd=sd,
-                   n_ens = ~n())
-    )
+                   n_ens = ~n()
+    )
+    )
+
+    # Valid ensembles
+  # Only perform analysis on ensembles containing all years between 1951 and 2040
+    n_years <- mean_sd_index %>% 
+      group_by(n_ens) %>% 
+      summarise(
+        n_year = n()
+      ) %>% 
+      filter(
+        n_year == 149
+      )
+
+    mean_sd_index <- mean_sd_index %>% 
+      filter(n_ens %in% n_years$n_ens)
 
   # Estimate noice from ensemble members
-  historic_sd <- Trans_index %>% 
-    filter(year < Hist_y) %>% 
-    group_by(taxon_key) %>% 
-    summarise_at(vars("RMean"),
+  historic_sd <- mean_sd_index %>% 
+    filter(year <= Hist_y,
+           n_ens %in% n_years$n_ens) %>% 
+    group_by(taxon_key,neighbour_id) %>% 
+    summarise_at(vars("trend"),
                  c(hist_mean=mean,
@@
   # For testing
-  if(Plot == T){
+  if(Plot == T & length(historic_sd$hist_mean) > 1){
   # Plot the result, bruh
@@
                 fill = "grey50"
+    ) +
+    geom_ribbon(data = mean_sd_index,
+                aes(x=year,
+                    ymin=historic_sd$hist_mean-(historic_sd$noise*2),
+                    ymax=historic_sd$hist_mean+(historic_sd$noise)*2)
+                ,
+                alpha = 0.5,
+                fill = "grey30"
     ) +
@@
       } +
-    geom_vline(xintercept = 2000,colour="blue",alpha = 0.5) +
+    geom_vline(xintercept = Hist_y,colour="blue",alpha = 0.5) +
     scale_x_continuous(breaks = seq(1950,2100,5)) +
     theme_classic() +
-    ggsave(paste(taxon_key,Id,"plot_example.jpg",sep="_"),
+
+    tk <- unique(mean_sd_index$taxon_key)
+
+    ggsave(paste(tk,Id,"plot_example.jpg",sep="_"),
        plot = last_plot(),
@@
       ungroup() %>%
-      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr)
+      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr) %>% 
+      mutate(Tresh= NA)
 
@@
 
-  # Identifies the year of emergence (e.g. trend > noise)
   Final_Result <- mean_sd_index %>%
-    filter(year > Hist_y,
-           trend > (historic_sd$hist_mean+historic_sd$noise) |
-             trend < (historic_sd$hist_mean-historic_sd$noise)
-    ) %>%
-    group_by(taxon_key,neighbour_id,n_ens) %>%
+    filter(year > Hist_y) %>% 
+    mutate(
+           Tresh_one = ifelse(trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise),"One",NA),
+           Tresh_two = ifelse(trend > (historic_sd$hist_mean+(2*historic_sd$noise)) | trend < (historic_sd$hist_mean-(2*historic_sd$noise)),"Two",NA)
+    ) %>%
+    gather("Tresh","YesNo",Tresh_one,Tresh_two) %>% 
+    filter(!is.na(YesNo)) %>% 
+    group_by(taxon_key,neighbour_id,n_ens,Tresh) %>%
     summarise(emerging_yr= min(year)) %>%
@@
     ungroup() %>%
-    select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr)
+    select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr,Tresh)
 
@@
                           Neighbours,
-                          eez_centroid){
+                          eez_centroid,
+                          Plot = F){
   t_start <- Sys.time()
@@
 
-    name <- paste(my_path("R",extra_path = "/Emergence/"),taxon_key,"_emergence.txt",sep="")
+    name <- paste(my_path("R",extra_path = "/Emergence_2005/"),taxon_key,"_emergence.txt",sep="")
 
@@
 spp_reman <- spp_list[!spp_list %in% spp_done]
+
+rand_plot <- sample(spp_reman,5)
 
@@
 
+gc()
+
     Emerging_result <- bind_rows(
-      lapply(spp_reman[63],
+      mclapply(spp_list,
                GetTransIndex,
@@
                Hist_y = Historic_year,
-               data_type = "Catch",
+               data_type = "Abd",
                ensemble_list = ensemble_list,
                Neighbours = Neighbours,
-               eez_centroid = eez_centroid
+               eez_centroid = eez_centroid,
+               Plot = F
       )
@@
   # Reads taxon data
-  proportion_data <- fread(paste(my_path("R",extra_path = "Proportion_2005/"),"proportion_",taxon_key,".csv", sep=""))
+  proportion_data <- fread(paste(result_path,"Proportion/","proportion_",taxon_key,".csv", sep=""))
 
@@
       summarise(
-        mean_spp = mean(ensemble_mean, na.rm = T) # average spp change from all countries sharing
-      ) %>% 
-      filter(!is.na(mean_spp))
+        mean_spp = mean(ensemble_mean,na.rm=T) # average spp change from all countries sharing
+      )
 
@@
         mean_spp = mean(ensemble_mean,na.rm=T) # average spp change from all countries sharing
-      ) %>% 
-      filter(!is.na(mean_spp))
+      )
     }
@@
        n_spp = n(),
-       min_eez = min(mean_spp, na.rm=T),
-       max_eez = max(mean_spp, na.rm=T),
-       sd_eez = sd(mean_spp, na.rm=T), # for some weird reason it wont work if the order is fliped
-       mean_eez = mean(mean_spp,na.rm=T)
-     ) %>% 
-     select(eez_name:max_eez,mean_eez,sd_eez)
+       min_spp = min(mean_spp,na.rm=T),
+       max_spp = max(mean_spp,na.rm=T),
+       sd_spp = sd(mean_spp,na.rm=T), # for some weird reason it wont work if the order is fliped
+       mean_spp = mean(mean_spp,na.rm=T)
+     ) %>% 
+     select(eez_name:max_spp,mean_spp,sd_spp)
 
@@
      n_spp = n(),
-       min_eez = min(mean_spp, na.rm=T),
-       max_eez = max(mean_spp, na.rm=T),
-       sd_eez = sd(mean_spp, na.rm=T), # for some weird reason it wont work if the order is fliped
-       mean_eez = mean(mean_spp,na.rm=T)
-     ) %>% 
-     select(eez_name:max_eez,mean_eez,sd_eez)
+     min_spp = min(mean_spp,na.rm=T),
+     max_spp = max(mean_spp,na.rm=T),
+     sd_spp = sd(mean_spp,na.rm=T), # for some weird reason it wont work if the order is fliped
+     mean_spp = mean(mean_spp,na.rm=T)
+   ) %>% 
+   select(eez_name:max_spp,mean_spp,sd_spp)
 
@@
  proportion_change <- eez %>% 
-   gather("variable","value",n_spp:sd_eez) %>% 
+   gather("variable","value",n_spp:sd_spp) %>% 
    spread(time_step,value) %>% 
    mutate_at(vars(
-     "per_early" = early,
-     "per_mid" = mid
-     ),
-            funs(round((((.-historical)/historical))*100,2))
+     "per_mid" = mid,
+     "per_end" = n_end),
+            funs(round((((.-historical)/.))*100,2))
    )
@@
 # Get taxon list
-Taxon_List <- list.files(paste(my_path("R"),"Proportion_2005",sep=""))
+Taxon_List <- list.files(paste(my_path("R"),"Propotion",sep=""))
 
@@
 # Run results
-results <- PropChange(Taxon_List, links="y") %>% 
-  filter(!is.na(eez_neighbour)) %>% 
-  mutate(per_mid_plot =ifelse(per_mid >= 50, 50,
-                           ifelse(per_mid <= -50,-50,per_mid)
-                           ),
-         per_early_plot =ifelse(per_early >= 50, 50,
-                           ifelse(per_early <= -50,-50,per_early)
-                           )
-         )
-
-# For dat availabillity
- # write_csv(results,
-          # "proportion_change_2005.csv")
+Proportion_Results <- PropChange(Taxon_List, links="y") %>% 
+  mutate(per_mid =ifelse(per_mid >= 25, 25,
+                           ifelse(per_mid <= -25,-25,per_mid)
+                           )
+         )
 
@@
   mutate(
-    per_early = ifelse(is.na(per_early_plot),0,per_early_plot),
-    per_mid = ifelse(is.na(per_mid_plot),0,per_mid_plot)
+    per_mid = ifelse(is.na(per_mid),0,per_mid)
   ) %>% 
@@
   filter(
-    variable == "mean_eez"#,
+    variable == "mean_spp"#,
     # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
@@
   unnest() %>% 
-  mutate_at(vars(X,Y), round,2) %>% 
-  gather("time_step","per_chng",per_mid_plot:per_early_plot)
+  mutate_at(vars(X,Y), round,2)
 
@@
 
-Data <- Data %>% 
-  gather("time_step","per_chng",per_mid_plot:per_early_plot)
-
 gc()# clean ram
-
-# time <- "per_mid_plot"
-time <- "per_early_plot"
-
 x <- ggplot() +
@@
   geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray70") +
-  geom_point(data = subset(Data, time_step == time & per_chng < 0),
+  geom_point(data = subset(Data, per_mid < 0),
              aes(start_lon,
                  start_lat,
-                 col = per_chng),
+                 col = per_mid),
              size = 0.05,
@@
              stroke = 1.5) +
-  geom_path(data = subset(paths, time_step == time & per_chng > 0),
+  geom_path(data = subset(paths, per_mid > 0),
             aes(X, Y,
                 group = interaction(line_id,subline_id),
-                col = per_chng
+                col = per_mid
             ),
@@
             show.legend = NA) +
-  my_ggtheme_m() +
+  ggtheme_map() +
   scale_colour_viridis(
     name="Average Gain of\nStock-Share Ratio (%)",
-    limits = c(0,max(Data$per_chng)),
-    breaks = seq(0,max(Data$per_chng),5),
+    limits = c(0,25),
+    breaks = seq(0,25,5),
     # option="magma",
-    direction = -1
+    direction = 1
   ) +
-  ggsave("./Figures/figure_connections_early_n50.png",
+  ggsave("figure_connectionsc.png",
          width = 10,
@@
 
-##### Histogram of change
-
-```{r hist_change}
-
- results %>% 
-  gather("time_step","change",per_early,per_mid) %>% 
-  filter(variable == "mean_eez" & 
-           !is.na(change)
-         ) %>%
-ggplot() +
-  geom_density(
-               aes(
-                 x = ifelse(change >= 100,100,change),
-                 # x = change,
-                 fill = ifelse(time_step == "per_early","q",time_step)
-                 # fill = time_step,
-
-               ),
-               # stat = "count",
-               # alpha= 0.8,
-               size=0.1,
-               # position = "stack"
-  ) +
-  scale_y_continuous("Density",
-                     # breaks = seq(0,90,10),
-                     # limits = c(0,100)
-  ) +
-  scale_x_continuous("Percentage Change",
-                     breaks = seq(-75,500,25)
-  ) +
-  my_ggtheme_p() +
-  theme(legend.position = "right") +
-  scale_fill_viridis(discrete = T, 
-                     option = "plasma",
-                     alpha = 0.5,
-                     name = "Time Step", labels = c("Middle", "Early")
-                     ) +
-  coord_flip() +
-  ggsave("./Figures/hist_change.png",
-         width = 6,
-         height = 6,
-         units = "in"
-  )
-
-```
-
 ####  Year of Eergence Analysis
@@
 # Get taxon list
-Taxon_List <- list.files(my_path("R",extra_path = "Emergence_2005"))
+Taxon_List <- list.files(my_path("R",extra_path = "Emergence_yr"))
 
@@
 
-# Neighbour list and their respective id (see chunk 3)
-Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)
-
-Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List,"_emergence.txt",sep="")
-
-# Load all sp in one table
+Paths <- paste(my_path("R",extra_path = "Emergence_yr"),Taxon_List,"_emergence.txt",sep="")
+
+# Load all sp in noe table
 Eme_Res <- bind_rows(lapply(Paths, fread)) %>% 
@@
 stock_d_hist <- Eme_Res %>% 
-  group_by(emerging_yr,taxon_key,Tresh) %>%
+  group_by(emerging_yr,taxon_key) %>%
   # summarise(n_tax = length(unique(eez_name))) %>%
@@
     level = "stock"
-  ) 
+  ) %>% 
+  select(1,3,4)
   # ggplot() +
@@
 eez_d_hist <- Eme_Res %>% 
-  group_by(eez_name,emerging_yr,Tresh) %>% 
+  group_by(emerging_yr,eez_name) %>% 
   # summarise(n_tax = length(unique(taxon_key))) %>% 
@@
   ) %>% 
-  ungroup() %>% 
   select(1,3,4)
@@
 hist_data <- stock_d_hist %>% 
-  bind_rows(eez_d_hist) %>% 
-  filter(Tresh != "NA")
+  bind_rows(eez_d_hist)
 
@@
     my_ggtheme_p() +
-    scale_fill_viridis(discrete = T) +
-  facet_wrap(~Tresh)
+    scale_fill_viridis(discrete = T)
 
@@
     distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
-    group_by(taxon_key,Tresh) %>% 
+    group_by(taxon_key) %>% 
     summarise(
@@
     ) %>% 
-    group_by(taxon_emer,Tresh) %>% 
+    group_by(taxon_emer) %>% 
   summarise(
@@
     distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
-    group_by(eez_name,Tresh) %>% 
+    group_by(eez_name) %>% 
     summarise(
@@
 plot_data <- stock_d %>% 
-  bind_rows(eez_d) %>% 
-  filter(Tresh != "NA")
+  bind_rows(eez_d)
 
@@
     scale_fill_viridis(discrete = T) +
-    facet_wrap(~Tresh)
-  # ggsave("./Figures/figure_accumulative_curve_B.png",
-  #        width = 6,
-  #        height = 6,
-  #        units = "in"
-  # )
+  ggsave("./Figures/figure_accumulative_curve_B.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
 ```

### 2020-07-26 Scripts/Initial_Analysis.Rmd

--- 2020-07-26 Scripts/Initial_Analysis.Rmd
@@
                    ensemble_sd=sd,
-                   n_ens = ~n())
-    )
+                   n_ens = ~n()
+    )
+    )
+
+    # Valid ensembles
+  # Only perform analysis on ensembles containing all years between 1951 and 2040
+    n_years <- mean_sd_index %>% 
+      group_by(n_ens) %>% 
+      summarise(
+        n_year = n()
+      ) %>% 
+      filter(
+        n_year == 149
+      )
+
+    mean_sd_index <- mean_sd_index %>% 
+      filter(n_ens %in% n_years$n_ens)
 
@@
 
+rand_plot <- sample(spp_reman,5)
+
 ```

### 2020-07-26 Scripts/Manuscript.Rmd

@@
 Because the main approach of this paper relies on understanding the spatial and temporal variation of a species distribution, we have to understand both the historical as well as the future climate-driven distribution variabillity, in order to infer the differences.
-We do this by following an ensemble member approach where each member constitutes a different initial Earth system condition [@Roggers:2015jq; @Frolicher:2009ge].
+We do this by following an ensemble member approach where each member constitutes a different initial Earth system condition [@Rodggers:2015jq; @Frolicher:2009ge].
 Thus, for our experiment, in each of the ten GFDL-ESM2M-ensambles the Earth system initial condition is infinitesimally different resulting in an unique atmosphere and ocean state at each point in time.
-This results in a different state of internal variability for each ensamble representing a plausible future under RCP 8.5 [@Roggers:2015jq].
+This results in a different state of internal variability for each ensamble representing a plausible future under RCP 8.5 [@Rodggers:2015jq].
 Following this approach allows us to estimate a secular trend for the natural variabillity (e.g.
 backround noise).
-At the same time, by having multiple initial-condition ensemble simulations of the GFDL-ESM2M we were able to remove the natural variability of a stock's distributio by averaging the secular trend over all ensemble members [@Roggers:2015jq].
+At the same time, by having multiple initial-condition ensemble simulations of the GFDL-ESM2M we were able to remove the natural variability of a stock's distributio by averaging the secular trend over all ensemble members [@Rodggers:2015jq].
 
@@
 In here we call this event time of emergence and defined it as the moment in time when the future variation in SSR becomes larger than the historical trend variabillity [@Roggers:2015jq].
-We emplot the ten GFDL ensable members described above, and developed an index of transboundary range shift, fisheries and climate change (*I*).
-The index is based on the variation of a species distribution centroid in realtion to the EEZs it ocuppys as follows:
+We emplot the ten GFDL ensable members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
+The index is based on the variation of a species distribution centroid in realtion to the centroid of the neighbouring EEZs it ocuppys.
+We define the species centroid as the gridcell within neighbouring EEZs with the highest abundance in each year.
+First, for each ensamble member, neighbouring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring elevation differences:
 
-$$I =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h)}})^2$$
+$$D_{ens} = acos(sin(lat_{eez})*sin(lat_{spp}) + cos(lat_{eez})*cos(lat_{spp})*cos(lon_{eez}-lon_{spp}))$$
+where $lat_{eez}$ and $lat_{spp}$ are the latitudes of the EEZ and specie centroid, respectiveley and $lon_{eez}$ and $lon_{spp}$ are the respectevly longitudes.
 
-where $D_A$ and $D_B$ represent the mean (among ensamble members) distribution centroid of a species in EEZs A and B, respectiveley for each future time step from 2006 to 2100 (*t*) and *sd* is the standar deviation of the historical (*h*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+Then, for each year we calculated the transboundary index as follows:
+
+$$TI =(\frac{D_{A,t}}{sd(D_{A,h})} - \frac{D_{B,t}}{sd(D_{B,h)}})^2$$
+
+where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectiveley for each future time step from 2006 to 2100 (*t*); *sd* is the standar deviation of the historical (*h*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+The index was then smoothed to a 10 years average.
+Finally, to determine the time of emergence we estimated a signal to noise ratio (SNR) and set an arbritary treshold as follows [@Rodggers:2015jq]:
+
+$$SNR = TREND/NOISE$$
+
+where *TREND* is the yearly mean *TI* ($\mu$) of all ten ensemble members and *NOISE* is the standar deviation of the TREND ($\sigma$) between 1951 and 2005.
+Note that in some cases not all ensemble members show the presence of a species in a certain EEZ.
+We presented the results only for those cases wher all 10 ensemble members had results.
+We determine a $\mu \pm \sigma$ (confidence interval of 68%) to show the first year when the "signal" of the transboundary index overpasses the historic "noise".
+Since this was an arbritary decission, we have included a more conservative estimation ($\mu \pm 2*\sigma$ for a 95% confidence interval).
 
@@
 
-This way, for two neighbouring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases.
-Consequently, if the centroid shifts are within historical variability, risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability, the risk of dispute is higher.
+In other words, for two neighbouring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases (see *TI* equation).
+Consequently, if the centroid shifts are within historical variability (*SNR* < 1), risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability (*SNR* > 1), the risk of dispute is higher.
 Thus, risk is positively related to the frequency of centroid shifts beyond historical variability (Fig.
@@
 
-All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *wesanderson* [@Packagewesanderson:2018vn], *rfishbase* [@PackagerfishbaseR:2019th], *R.matlab* [@PackageRmatlabRe:2018un], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], and *zoo* [@PackagezooSInfr:2019tm].
+All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *wesanderson* [@Packagewesanderson:2018vn], *rfishbase* [@PackagerfishbaseR:2019th], *R.matlab* [@PackageRmatlabRe:2018un], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@gmt] and *zoo* [@PackagezooSInfr:2019tm].
 All code is available at https://github.com/jepa/TransEmergence
@@
 
-prop_path <- "/Volumes/HALL2000/Data/TransEmergence/Results/Proportion_2005/"
-prop_names  <- list.files(prop_path)
-
-prop_read <- paste(prop_path,prop_names,sep="")
-
-results <- bind_rows(lapply(prop_read, data.table::fread))
+prop_r <- read_csv("/Volumes/HALL2000/Data/TransEmergence/Results/proportion_change_2005.csv") %>% 
+  gather("time_step","change",per_early,per_mid)
 
@@
 
-# Mid century
+# Early century
 n_eez <- prop_r %>%
-  filter(variable == "mean_spp" & 
-           !is.na(per_mid)
+  filter(variable == "mean_eez" & 
+           !is.na(change),
+         time_step == "per_early"
          ) %>% 
@@
 
-n_eez_per <- n_eez/275*100
+n_eez_per <- round(n_eez/275*100,2)
 
-### Max values
-max_mid <- prop_r %>% 
-  filter(variable == "max_spp" & 
-           !is.na(per_mid)
+### Max values early
+max_values <- prop_r %>% 
+  filter(variable == "max_eez" & 
+           !is.na(change),
+         # time_step == "per_early"
          )
@@
 average_mid <- prop_r %>% 
-  filter(variable == "mean_spp" & 
-           !is.na(per_mid)
-         ) %>% 
-  pull(per_mid) %>% 
-  mean()
-
+  filter(variable == "mean_eez" & 
+           !is.na(change)
+         ) 
 ```
 
-Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by mid $21^{st}$ century (Fig.
+Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by early $21^{st}$ century (Fig.
 1).
-The change in SSR will range from gains of over 100% (e.g.
-Honduras' SSR of *species* shared with Mexico is projected to go from 28% to 60%) to loses of up to -80% of SSR (e.g.
-Nauru' SSR of *species* shared with the Marshall Isl.
-is projected to decline from 14% to 3%).
-By the end of the century, SSR changes will be more exrteme with 13 countries gaining over 100% SSR and Palau loosing -93.61% of the SSR of *spp* to the	Philippines (from 20% to 1.2%).
+The change in SSR will range from gains of over 100% of a sertain shared stock (e.g.
+Guatemala' SSR of *species* shared with Mexico is projected to go from 12% to 26%) to loses of up to -71% of SSR (e.g.
+Honduras SSR of *species* shared with the Nicaragua in the Pacific coast is projected to decline from 50% to 14%).
+Despite no substantial changes in SSR per EEZ can be seen by the mid $21^st$ century (**Supplements Fig 1 (map)**), changes in SSR are expected to increase in intensity (**Supplements Fig 2 (histogram)**).
 

### 2020-08-13 Scripts/Initial_Analysis.Rmd
misc changes

@@
 
-```{r PropChange, eval = F, echo =F}
+```{r Emergence-analysis, eval = F, echo =F}
 
@@
   left_join(Neighbours,
-            by = c("eez_name","eez_neighbour"))
+            by = c("eez_name","eez_neighbour")
+            )
 # warnings() for empty data species
@@
   st_transform(crs = 4326) %>% # 4326
-  st_simplify(preserveTopology = TRUE, dTolerance = 100)
+  # st_transform(crs = "+proj=eck4") %>% 
+  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)
 
@@
 
-data %>% 
+data_stock %>% 
   left_join(data_eez,
@@
             by = "eez_name") %>% 
-  # filter(Tresh == "Tresh_one") %>%
+  filter(Tresh == "Tresh_two",
+         metric == "mean_yr") %>%
   ggplot() +
@@
     aes(
-      fill = mean_yr
+      fill = value
     )
@@
   my_ggtheme_m() +
-  facet_wrap(~Tresh,
-             ncol = 1) +
-  ggsave("./Figures/emer_map_T12.png",
+  # facet_wrap(~Tresh,
+             # ncol = 1) +
+  ggsave("./Figures/emer_map_T2a.png",
          width = 12,

### 2020-08-13 Scripts/Initial_Analysis.Rmd
testing coords

@@
 #### 1. SAU relations between INDEX and Country's EEZs
+# Re-ran with new names data
 ## ---------------##
 
-EEZIDs_List <- my_path("G", extra_path = "Spatial/DBEM", name="Updated_EEZList_17June2016.xlsx", read = TRUE) %>% 
-  clean_names()
-
-EEZ_CellID <- my_path("G", extra_path = "Spatial/DBEM", name="EEZ_CellID.xlsx", read = TRUE) %>% 
+# EEZIDs_List <- my_path("G", extra_path = "Spatial/V_Lam", name="Updated_EEZList_17June2016.xlsx", read = TRUE) %>%
+  # clean_names()
+
+EEZ_CellID <- my_path("G", extra_path = "Spatial/V_Lam", name="EEZ_CellID.xlsx", read = TRUE) %>%
   clean_names()
@@
 
-Lon_Lat_DBEM <- my_path("G", extra_path = "Spatial/DBEM", name="Lon_Lat_DBEM.txt", read = TRUE, header = FALSE)
+# SAU name relatioal table (Fish For Visa)
+sau_clean_names <- my_path("G", extra_path = "Spatial/SAU/", name="SAU_matching_names.csv", read = TRUE)
+
+Lon_Lat_DBEM <- my_path("G", extra_path = "Spatial", name="Lon_Lat_DBEM.txt", read = TRUE, header = FALSE)
 colnames(Lon_Lat_DBEM) <- c("index","lon","lat")
 
-Index_Code <- EEZIDs_List %>% 
-  left_join(EEZ_CellID) %>% 
-  rename(eez_name = name) %>% 
+# Index_Code <- EEZIDs_List %>% 
+#   left_join(EEZ_CellID) %>% 
+#   rename(eez_name = name) %>% 
+#   left_join(Lon_Lat_DBEM,
+#             by = "index")
+
+Index_Code <-  sau_clean_names %>% 
+  select(eezid = sf_eezid) %>% 
+  left_join(EEZ_CellID,
+            by = "eezid") %>% 
   left_join(Lon_Lat_DBEM,
@@
 # Save Index data for future 
-write_csv(index_code,
-          paste(my_path("G",extra_path = "Spatial/DBEM"),"sau_index_code.csv",sep="")
-          )
+# write_csv(index_code,
+          # paste(my_path("G",extra_path = "Spatial/DBEM"),"sau_index_code.csv",sep="")
+          # )
 
@@
 # Get the center poligon of each EEZ for map source/target with st_centroid
-coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
+coords <- as.data.frame(st_centroid(SAU_eez_sf)) %>% 
   ungroup() %>% 
@@
   mutate_at(vars(longitude,latitude),as.numeric)
+
+# Test coords
+
+maptest <- ggplot() +
+  geom_sf(data = SAU_eez_sf, aes(), fill = NA) +
+  geom_point(data = coords,
+             aes(
+               x = longitude,
+               y = latitude
+             )
+  )
+
+ggsave("maptest.png",
+       plot = maptest,
+       width = 10,
+       height = 8,
+       units = "in"
+       )
 
@@
 # SAU relations between INDEX and Country's EEZs (see chunk 3)
-index_code <- my_path("G", extra_path = "Spatial/SAU/", name="sau_index_code.csv", read = TRUE)
+index_code <- my_path("G", extra_path = "Spatial/SAU/", name="SAU_matching_names.csv", read = TRUE)
 

### 2020-08-13 Scripts/Initial_Analysis.Rmd
double checked coordenates

@@
 
-# Test coords
-
-maptest <- ggplot() +
+# Double check coords
+# Note that Australia and Ggreenland have points in the middle of the country because of the EEZ shape
+
+coords_map <- ggplot() +
   geom_sf(data = SAU_eez_sf, aes(), fill = NA) +
-  geom_point(data = coords,
+  geom_point(data = eez_centroid,
              aes(
@@
 
-ggsave("maptest.png",
-       plot = maptest,
+ggsave("coords_map.png",
+       plot = coords_map,
        width = 10,

### 2020-08-14 Scripts/Initial_Analysis.Rmd
small updates

@@
     data_type = data_type,
-    ensemble = ensemble,
+    ensemble = ensemble, # change me to ensemble
     path = my_path("G","DBEM")
@@
 
+       ### Code to produce graphs for Brazil and ruguay
+
+       # name <- paste(taxon_key, "distrbution p = 90%") 
+       # Centroids %>% 
+       #   # filter(year %in% c(1951,2000,2010)) %>% 
+       #   ggplot() +
+       #   geom_point(
+       #     aes(
+       #       x = lon,
+       #       y = lat,
+       #       size = distance,
+       #       color = year,
+       #       shape= eez_name
+       #     )
+       #   ) +
+       #   geom_point(
+       #     aes(
+       #       x = longitude,
+       #       y = latitude,
+       #     ),
+       #     fill = "red"
+       #   ) +
+       #   geom_sf(data = braur_sf,
+       #           aes()
+       #   )  +
+       #   ggtitle(name)
+       # 
+       # ggsave(paste(taxon_key,"distribution.jpg",sep="_"),
+       # plot = last_plot(),
+       # width = 10,
+       # height = 8,
+       # units = "in",
+       # path = my_path("D")
+       # )
+
       return(Centroids)
@@
     }
+
     # Determine the ids
@@
 #
-# ggplot(Test_centroids) +
-#   geom_line(
+
+# braur_sf <- World_ %>% 
+#   filter(name %in% c("Brazil","Uruguay"))
+
+# ggplot(braur_sf) +
+#   geom_sf() +
+  # geom_point(data = subset(Centroids, year %in% seq(1951,2100,20)),
+  # geom_point(data = subset(Centroids,eez_name =="Brazil"),
+  # geom_point(data = Centroids,
   #            aes(
-#       x = year,
-#       y = distance,
-#       color = eez_name
+  #              x = lon,
+  #              y = lat,
+  #              color = year#,
+  #              # size = value
   #            )
-#   )
+  # ) #+
+  # facet_wrap(~to_filter+eez_name)
 
@@
                             align = "right", 
-                            fill = trans_index)
+                            fill = trans_index,
+                            na.rm=T)
     ) %>% 
@@
                    n_ens = ~n()
-    )
+
+    ),
+    na.rm=T
     )
@@
     filter(year <= Hist_y) %>% 
-    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per ensemble
+    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
     summarise_at(vars("trans_index"),
-                 c(trend= mean)
+                 c(trend= mean),
+                 na.rm=T
                  ) %>% 
@@
     scale_x_continuous(breaks = seq(1950,2100,5)) +
-    theme_classic() +
+    theme_classic()
 
@@
     # Call Get centroids fucntion for centroids data per ensemble
+
   Step_one <- bind_rows(
@@
              GetCentroids,
-             taxon_key = taxon_key,
+             taxon_key = taxon_key, # for testing
              year = year_df$year,
@@
 # SAU relations between INDEX and Country's EEZs (see chunk 3)
-index_code <- my_path("G", extra_path = "Spatial/SAU/", name="SAU_matching_names.csv", read = TRUE)
+index_code <- my_path("G", extra_path = "Spatial/SAU/", name="sau_index_code.csv", read = TRUE)
 

### 2020-08-18 Scripts/Initial_Analysis.Rmd
Ready for results

@@
 
+results <- my_path("R",name="proportion_change_2005.csv", read=T)
+
 # View(results)
@@
   st_transform(crs = 4326) %>% # 4326
-  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)
+  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
+  clean_names()
 
 # Call world map from natural earth
-World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf"))
+World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf")) %>% 
+  rename(fishing_entity = name_long) %>% 
+  clean_names()
 
@@
 
+# Clean names to match data
+matching_names <- my_path("G","Spatial/SAU", "sau_matching_names.csv", read = T)
+
 # Load all sp in one table
@@
             by = c("eez_name","eez_neighbour")
-            )
+  ) %>% 
+  rename(sf_eez_name=eez_name) %>% 
+  left_join(matching_names,
+            by=  "sf_eez_name") %>% 
+  clean_names()
 # warnings() for empty data species
@@
   rename(eez_name = Name) %>% 
-  st_transform(crs = 4326) %>% # 4326
-  # st_transform(crs = "+proj=eck4") %>% 
-  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)
+  # st_transform(crs = 4326) %>% # 4326
+  st_transform(crs = "+proj=eck4") %>% # for tbular plot
+  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
+  clean_names()
 
 # Call world map from natural earth
-World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf"))
-
-eez_centroid <- read_csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""))
+World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf")) %>% 
+  rename(fishing_entity = name_long) %>%
+  left_join(matching_names,
+            by = "fishing_entity") %>% 
+  distinct(fishing_entity,.keep_all = TRUE) %>% 
+  st_transform(crs = "+proj=eck4") %>% # for tbular plot
+
+# eez_centroid <- read_csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""))
 
@@
 eez_d_hist <- Eme_Res %>% 
-  group_by(eez_name,Tresh) %>% 
+  group_by(eezid  = sf_eezid,fishing_entity,tresh) %>% 
   # summarise(n_tax = length(unique(taxon_key))) %>% 
@@
     mean_yr = mean(emerging_yr,na.rm=T),
-    sd_yr = sd(emerging_yr,na.rm=T)) %>% 
-  gather("metric","value",mean_yr:sd_yr)
+    sd_yr = sd(emerging_yr,na.rm=T),
+    n_taxa = length(unique(taxon_key))
+    ) %>% 
+  gather("metric","value",mean_yr:n_taxa)
+
+# Per fishing entity
+
+fish_d_hist <- Eme_Res %>% 
+  group_by(fishing_entity,tresh) %>% 
+  # summarise(n_tax = length(unique(taxon_key))) %>% 
+  summarise(
+    value = length(unique(taxon_key))
+    ) %>% 
+  filter(!is.na(fishing_entity))
 
@@
 
+World_map_nemer <- World_map %>%
+  mutate(fishing_entity = ifelse(fishing_entity == "United States","USA", fishing_entity)) %>% 
+  left_join(fish_d_hist,
+            by = "fishing_entity")
+
+for(i in 1:2){
+
+  Tresh_list <- c("Tresh_one","Tresh_two")
+  name_save <- paste("./Figures/emer_map_",Tresh_list[i],".png")
+
   map <- World_eez_sf %>% 
     left_join(eez_d_hist,
-            by = "eez_name") %>% 
-  filter(Tresh == "Tresh_two",
+              by = "eezid") %>% 
+    filter(tresh == Tresh_list[i],
            metric == "mean_yr") %>% 
@@
         fill = value
-    )
-  ) +
-  geom_sf(data =World_map, aes(), size = 0.1, fill = "gray90", col = "gray90") +
-  scale_fill_viridis() +
+      ),
+      color =  "white",
+      size = 0.1
+    ) +
+    geom_sf(data = World_map_nemer, aes(fill = value +2000   # to match scales
+                                        ),
+      color =  "white",
+      size = 0.1
+      ) +
+    annotate("text", 
+             x =(c(-85.5,-30,25,81,136)+5),
+             y = -93.5,
+             # x =(c(-8500000,-3000000,2500000,7500000,12800000)),
+             # y = -9050000,
+             label = c(0,25,50,75,100), 
+             color="black", 
+             size=3, 
+             angle=0, 
+             # fontface="bold",
+             hjust=0) +
+    scale_fill_viridis(
+      limits = c(2000,2100),
+      breaks = seq(2000,2100,25)
+    ) +
     my_ggtheme_m() +
-  # facet_wrap(~Tresh,
-             # ncol = 1) +
-  ggsave("./Figures/emer_map_T2a.png",
+    ggsave(name_save,
            width = 12,
@@
     )
+
+}
 

### 2020-08-18 Scripts/Manuscript.Rmd
updated methods

@@
 We consider each species as a single stock unit within an EEZ due to lack of more detailed information [@Teh:2015gd; @Palacios-Abrantes:xxxx].
-We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org).
+We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) **noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic) including islands territories**, and determined the intersections between polygons using the R package *sf* [@PackagesfSimpleF:2018vp].
 
@@
 In here we call this event time of emergence and defined it as the moment in time when the future variation in SSR becomes larger than the historical trend variabillity [@Roggers:2015jq].
-We emplot the ten GFDL ensable members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
+We employed the ten GFDL ensable members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
 The index is based on the variation of a species distribution centroid in realtion to the centroid of the neighbouring EEZs it ocuppys.
-We define the species centroid as the gridcell within neighbouring EEZs with the highest abundance in each year.
-First, for each ensamble member, neighbouring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring elevation differences:
+Each species centroid was determined by the average latitude and longitude of the grid cells containing the top 10% abundance percentile of each year.
+The EEZ centroid was estimated using the *st* package of *R* statistical software.
+For each ensamble member, neighbouring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring elevation differences:
 
@@
 
-where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectiveley for each future time step from 2006 to 2100 (*t*); *sd* is the standar deviation of the historical (*h*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectiveley for each future time step from 2006 to 2100 (*t*); *sd* is the standard deviation of the historical (*h*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
 The index was then smoothed to a 10 years average.
-Finally, to determine the time of emergence we estimated a signal to noise ratio (SNR) and set an arbritary treshold as follows [@Rodggers:2015jq]:
+Finally, to determine the time of emergence we estimated a signal to noise ratio (SNR) and set an arbitrary threshold as follows [@Rodggers:2015jq]:
 
@@
 where *TREND* is the yearly mean *TI* ($\mu$) of all ten ensemble members and *NOISE* is the standar deviation of the TREND ($\sigma$) between 1951 and 2005.
-Note that in some cases not all ensemble members show the presence of a species in a certain EEZ.
-We presented the results only for those cases wher all 10 ensemble members had results.
+Note that in some cases not all ensemble members show the presence of a species in a certain EEZ throughout the study's time frame.
+We presented the results only for those cases where all 10 ensemble members had results for each year between 1951 and 2100 (n = 150).
 We determine a $\mu \pm \sigma$ (confidence interval of 68%) to show the first year when the "signal" of the transboundary index overpasses the historic "noise".

### 2020-08-18 Scripts/Initial_Analysis.Rmd
Re-organized results

@@
 
-####  Prooportion Analysis
-
-```{r PropChange, eval = F, echo =F}
+#### Data Needed for Analysis
+
+```{r General_Data, eval = F, echo =F}
+
+# --------------- #
+# General Data needed
+# --------------- #
 
 # Get taxon list
-Taxon_List <- list.files(paste(my_path("R"),"Proportion_2005",sep=""))
+Taxon_List_emer <- list.files(my_path("R",extra_path = "Emergence_2005"))
 
 # clean names for function
-Taxon_List <- gsub(".*_","",Taxon_List) 
-Taxon_List <- gsub("\\..*","",Taxon_List)
-
-# Run results
-results <- PropChange(Taxon_List, links="y") %>% 
-  filter(!is.na(eez_neighbour)) %>% 
-  mutate(per_mid_plot =ifelse(per_mid >= 50, 50,
-                           ifelse(per_mid <= -50,-50,per_mid)
-                           ),
-         per_early_plot =ifelse(per_early >= 50, 50,
-                           ifelse(per_early <= -50,-50,per_early)
-                           )
-         )
+Taxon_List_emer <- gsub("_.*","",Taxon_List_emer)
+
+# Neighbour list and their respective id (see chunk 3)
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)
+
+Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List,"_emergence.txt",sep="")
+
+# Clean names to match data
+matching_names <- my_path("G","Spatial/SAU", "sau_matching_names.csv", read = T)
+
+# --------------- #
+# Load Proportion results
+# --------------- #
+
+# Read proportion results
+# prop_results <- PropChange(Taxon_List, links="y") %>% 
+  # filter(!is.na(eez_neighbour)) %>% 
+  # mutate(per_mid_plot =ifelse(per_mid >= 50, 50,
+  #                          ifelse(per_mid <= -50,-50,per_mid)
+  #                          ),
+  #        per_early_plot =ifelse(per_early >= 50, 50,
+  #                          ifelse(per_early <= -50,-50,per_early)
+  #                          )
+  #        )
 
 # For dat availabillity
- # write_csv(results,
+ # write_csv(prop_results,
           # "proportion_change_2005.csv")
 
-results <- my_path("R",name="proportion_change_2005.csv", read=T)
-
-# View(results)
-
-### Load Data
+prop_results <- my_path("R",name="proportion_change_2005.csv", read=T)
+
+# --------------- #
+# Load year of emergence results
+# --------------- #
+
+# Load all sp in one table
+Eme_Res <- bind_rows(lapply(Paths, fread)) %>% 
+  filter(
+    n_ens == 10
+  ) %>% 
+  left_join(Neighbours,
+            by = c("eez_name","eez_neighbour")
+  ) %>% 
+  rename(sf_eez_name=eez_name) %>% 
+  left_join(matching_names,
+            by=  "sf_eez_name") %>% 
+  clean_names()
+# warnings() for empty data species
+
+# View(Eme_Res)
+
+# --------------- #
+# Load shapefiles for figures
+# --------------- #
 
@@
   rename(eez_name = Name) %>% 
-  st_transform(crs = 4326) %>% # 4326
+  # st_transform(crs = 4326) %>% # 4326
+  st_transform(crs = "+proj=eck4") %>% # for tbular plot
   st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
@@
   rename(fishing_entity = name_long) %>%
-  clean_names()
-
-eez_centroid <- my_path("D",extra_path = "Spatial",name = "eez_centroids.csv",read = T)
-
-```
+  left_join(matching_names,
+            by = "fishing_entity") %>% 
+  distinct(fishing_entity,.keep_all = TRUE) %>% 
+  st_transform(crs = "+proj=eck4") %>% # for tbular plot
+
+# eez_centroid <- read_csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""))
+
+```
+
+####  Prooportion Analysis
 
@@
 
-####  Year of Eergence Analysis
-
-```{r Emergence-analysis, eval = F, echo =F}
-
-# Get taxon list
-Taxon_List <- list.files(my_path("R",extra_path = "Emergence_2005"))
-
-# clean names for function
-Taxon_List <- gsub("_.*","",Taxon_List)
-
-# Neighbour list and their respective id (see chunk 3)
-Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)
-
-Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List,"_emergence.txt",sep="")
-
-# Clean names to match data
-matching_names <- my_path("G","Spatial/SAU", "sau_matching_names.csv", read = T)
-
-# Load all sp in one table
-Eme_Res <- bind_rows(lapply(Paths, fread)) %>% 
-  filter(
-    n_ens == 10
-  ) %>% 
-  left_join(Neighbours,
-            by = c("eez_name","eez_neighbour")
-  ) %>% 
-  rename(sf_eez_name=eez_name) %>% 
-  left_join(matching_names,
-            by=  "sf_eez_name") %>% 
-  clean_names()
-# warnings() for empty data species
-
-# View(Eme_Res)
-
-### Load Data
-
-# Load SAU shapefile name and paths
-
-# The path
-path_world <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile")
-
-# The File
-fnam_world <- "SAUEEZ_July2015.shp"
-
-# Load it!
-World_eez_sf <- st_read(dsn = path_world,
-                    layer =file_path_sans_ext(fnam_world)) %>% 
-  rename(eez_name = Name) %>% 
-  # st_transform(crs = 4326) %>% # 4326
-  st_transform(crs = "+proj=eck4") %>% # for tbular plot
-  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
-  clean_names()
-
-# Call world map from natural earth
-World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf")) %>% 
-  rename(fishing_entity = name_long) %>%
-  left_join(matching_names,
-            by = "fishing_entity") %>% 
-  distinct(fishing_entity,.keep_all = TRUE) %>% 
-  st_transform(crs = "+proj=eck4") %>% # for tbular plot
-
-# eez_centroid <- read_csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""))
-
-```
+####  Year of Eergence Analysisç
 

### 2020-08-18 Scripts/Initial_Analysis.Rmd
Produced figure 2 with regional zooms

@@
   "gmt", # for estimating distances between points
-  "viridis"
+  "viridis",
+  "cowplot"
 )
@@
 
-Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List,"_emergence.txt",sep="")
+# Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List_emer,"_emergence.txt",sep="")
 
 # Clean names to match data
-matching_names <- my_path("G","Spatial/SAU", "sau_matching_names.csv", read = T)
+Matching_names <- my_path("G","Spatial/SAU", "matching_names.csv", read = T)
 
@@
 # Read proportion results
-# prop_results <- PropChange(Taxon_List, links="y") %>% 
+# prop_results <- PropChange(Taxon_List_emer, links="y") %>% 
   # filter(!is.na(eez_neighbour)) %>% 
@@
 
-prop_results <- my_path("R",name="proportion_change_2005.csv", read=T)
+Prop_results <- my_path("R",name="proportion_change_2005.csv", read=T)
 
@@
   rename(sf_eez_name=eez_name) %>% 
-  left_join(matching_names,
+  left_join(Matching_names,
             by=  "sf_eez_name") %>% 
   clean_names()
+
 # warnings() for empty data species
@@
   rename(eez_name = Name) %>% 
-  # st_transform(crs = 4326) %>% # 4326
-  st_transform(crs = "+proj=eck4") %>% # for tbular plot
+  st_transform(crs = 4326) %>% # 4326
+  # st_transform(crs = "+proj=eck4") %>% # for tbular plot
   st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
@@
 # Call world map from natural earth
+
+World_map_dat <- Matching_names %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,iso_a3) %>% 
+  summarise(n=n()) %>% 
+  select(-n)
+
 World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf")) %>%
-  rename(fishing_entity = name_long) %>%
-  left_join(matching_names,
-            by = "fishing_entity") %>% 
-  distinct(fishing_entity,.keep_all = TRUE) %>% 
-  st_transform(crs = "+proj=eck4") %>% # for tbular plot
-
-# eez_centroid <- read_csv(paste(data_path,"Spatial/eez_centroids.csv", sep = ""))
+  # st_transform(crs = "+proj=eck4") %>%
+  st_transform(crs = 4326) %>% # 4326
+  left_join(World_map_dat,
+            by = "iso_a3")
+
+# Load eez_centroid data
+Eez_centroid <- my_path("D", "Spatial","eez_centroids.csv",read = T)
+
+# Tranform to tubular
+Eez_centroid_sf <-st_as_sf(Eez_centroid,
+            coords = c("longitude", "latitude")
+            ) %>% 
+  st_set_crs(4326) #%>% # to match shapefiles
+  # st_transform(crs = "+proj=eck4")
+
+# Transform sf to df
+Eez_centroid_eck <- as.data.frame(Eez_centroid_sf)
+
+Eez_centroid_eck <- do.call(rbind, st_geometry(Eez_centroid_sf)) %>% 
+  as_tibble() %>% 
+  setNames(c("lon","lat")) %>% 
+  bind_cols(Eez_centroid) %>% 
+  select(eez_name,longitude = lon,latitude = lat)
 
@@
 # Prepare data from results
-Data <- results %>% 
+Data <- Prop_results %>% 
   # get coords for sources
-  left_join(eez_centroid,
+  left_join(Eez_centroid_eck,
             by = "eez_name") %>%
@@
   # get coords for targets
-  left_join(eez_centroid,
+  left_join(Eez_centroid_eck,
             by = "eez_name") %>% 
@@
   purrr::map(st_linestring) %>%
-  st_sfc(crs = 4326) %>%
+  # st_sfc(crs = 4326) %>%
+  st_sfc(crs = "+proj=eck4") %>% # for tbular plot
   st_sf(geometry = .)%>%
@@
   mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>%
+  # mutate(geometry = (geometry + c(18000000,9000000)) %% c(360) - c(18000000,9000000)) %>%
   st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>%
-  sf::st_sf(crs = 4326)
+  st_wrap_dateline(options = c("WRAPDATELINE=YES"), quiet = TRUE) %>%
+  # sf::st_sf(crs = 4326)
+  sf::st_sf(crs = "+proj=eck4") # for tbular plot
 
@@
 
-x <- ggplot() +
-  geom_sf(data = subset(World_map, sovereignt != "Antarctica"), size = 0.1, fill = "gray90", col = "gray90") +
-  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray70") +
+## ------------------- ##
+# Global Plot
+## ------------------- ##
+
+Prop_map <- ggplot() +
+  geom_sf(data = subset(World_map, sovereignt != "Antarctica"), size = 0.1, fill = "gray50", col = "gray50") +
+  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
   geom_point(data = subset(Data, time_step == time & per_chng < 0),
@@
     # option="magma",
-    direction = -1
+    # direction = -1
   ) +
@@
   units = "in"
+  )
+
+## ------------------- ##
+# Regional Plots
+## ------------------- ##
+
+Lat_am_land <- World_map %>% 
+  filter(region_wb %in% c("Latin America & Caribbean") |
+          sau_fishing_entity %in% c("USA","Mexico")
+         )
+
+Lat_am_eez <- World_eez_sf %>% 
+  rename(sf_eez_name = eez_name) %>% 
+  left_join(Matching_names) %>% 
+  filter(
+    # sau_fishing_entity %in% Lat_am_land$sau_fishing_entity,
+         !str_detect(sf_eez_name,"Pacific"),
+         !sf_eez_name %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador")
+         )
+
+Lat_am_eez %>% filter(str_detect(sf_eez_name,"Tri"))
+
+Lat_am_data <- Data %>% 
+  filter(source %in% Lat_am_eez$sf_eez_name)
+
+Lat_am_paths <- paths %>% 
+  filter(source %in% Lat_am_eez$sf_eez_name,
+         target %in% Lat_am_eez$sf_eez_name)
+
+Lat_am_map <- ggplot() +
+  geom_sf(data = Lat_am_eez, size = 0.1, fill = "white", col = "gray30") +
+  geom_sf(data = Lat_am_land, size = 0.1, fill = "gray50", col = "gray40") +
+  geom_point(data = subset(Lat_am_data, time_step == time & per_chng < 0 ),
+             aes(start_lon,
+                 start_lat,
+                 col = per_chng),
+             size = 1,
+             col = "gray30",
+             shape = 20,
+             stroke = 1.5) +
+  geom_path(data = subset(Lat_am_paths, time_step == time & per_chng > 0),
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = per_chng
+            ),
+            size = 1,
+            arrow = arrow(length = unit(0.05, "npc"),
+                          ends = "first"),
+            show.legend = NA) +
+  coord_sf(xlim = c(-100, -55),ylim = c(9, 30)) + 
+  labs(x = "Lon", y = "Lat") +
+  scale_x_continuous(breaks = seq(-100,60,10)) +
+  scale_y_continuous(breaks = seq(10,30,5)) +
+  # my_ggtheme_m() +
+  theme(legend.position = "",
+        panel.background = element_rect(fill = "white")) +
+  scale_colour_viridis(
+    name="Average Gain of\nStock-Share Ratio (%)",
+    limits = c(0,max(Lat_am_data$per_chng)),
+    breaks = seq(0,max(Lat_am_data$per_chng),5),
+    # option="magma",
+    # direction = -1
+  );Lat_am_map
+
+## ------------------- ##
+# The Med
+## ------------------- ##
+
+# Land polygon
+Med_land <- World_map %>% 
+  filter(region_wb %in% c("Europe & Central Asia","Middle East & North Africa") |
+          sau_fishing_entity %in% c("Spain","France")
+         )
+
+# EEZ polygon
+Med_eez <- World_eez_sf %>% 
+  rename(sf_eez_name = eez_name) %>% 
+  left_join(Matching_names) %>% 
+  filter(sau_fishing_entity %in% Med_land$sau_fishing_entity,
+         sf_eez_name != "Spain (Northwest)")
+
+# Med data
+Med_data <- Data %>% 
+  filter(source %in% Med_eez$sf_eez_name)
+
+# Med paths
+Med_paths <- paths %>% 
+  filter(source %in% Med_eez$sf_eez_name,
+         target %in% Med_eez$sf_eez_name)
+
+# Med plot
+Med_map <- ggplot() +
+  geom_sf(data = Med_eez, size = 0.5, fill = "white", col = "gray30") +
+  geom_sf(data = Med_land, size = 0.5, fill = "gray50", col = "gray40") +
+  geom_point(data = subset(Med_data, time_step == time & per_chng < 0 ),
+             aes(start_lon,
+                 start_lat,
+                 col = per_chng),
+             size = 1,
+             col = "gray30",
+             shape = 20,
+             stroke = 1.5) +
+  geom_path(data = subset(Med_paths, time_step == time & per_chng > 0),
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = per_chng
+            ),
+            size = 1,
+            arrow = arrow(length = unit(0.05, "npc"),
+                          ends = "first"),
+            show.legend = NA) +
+  coord_sf(xlim = c(-3, 35),ylim = c(30, 45)) +
+  labs(x = "Lon", y = "Lat") +
+  scale_x_continuous(breaks = seq(-5,35,10)) +
+  scale_y_continuous(breaks = seq(30,50,5)) +
+  # my_ggtheme_m() +
+  theme(legend.position = "",
+        panel.background = element_rect(fill = "white")) +
+  scale_colour_viridis(
+    name="Average Gain of\nStock-Share Ratio (%)",
+    limits = c(0,max(Med_data$per_chng)),
+    breaks = seq(0,max(Med_data$per_chng),5),
+    # option="magma",
+    # direction = -1
+  );Med_map
+
+## ------------------- ##
+# Northern Europe
+## ------------------- ##
+
+Eu_land <- World_map %>% 
+  filter(
+    continent == "Europe",
+    sau_fishing_entity != "Iceland"
+    )
+
+# 
+# ggplot(N_Eu_land) +
+#   geom_sf()
+
+Eu_eez <- World_eez_sf %>% 
+  rename(sf_eez_name = eez_name) %>% 
+  left_join(Matching_names) %>% 
+  filter(
+    !sf_eez_name %in% c("Iceland", "Norway","Italy","Greece","Croatia","Bulgaria","Romania"),
+    !str_detect(sf_eez_name,"Spain")
+  )
+
+Eu_eez %>% filter(str_detect(sf_eez_name,"Romania"))
+
+Eu_data <- Data %>% 
+  filter(source %in% Eu_eez$sf_eez_name)
+
+Eu_paths <- paths %>% 
+  filter(source %in% Eu_eez$sf_eez_name,
+         target %in% Eu_eez$sf_eez_name)
+
+Eu_map <- ggplot() +
+  geom_sf(data = Eu_eez, size = 0.5, fill = "white", col = "gray30") +
+  geom_sf(data = Eu_land, size = 0.5, fill = "gray50", col = "gray40") +
+  geom_point(data = subset(Med_data, time_step == time & per_chng < 0 ),
+             aes(start_lon,
+                 start_lat,
+                 col = per_chng),
+             size = 1,
+             col = "gray30",
+             shape = 20,
+             stroke = 1.5) +
+  geom_path(data = subset(Eu_paths, time_step == time & per_chng > 0),
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = per_chng
+            ),
+            size = 1,
+            arrow = arrow(length = unit(0.05, "npc"),
+                          ends = "first"),
+            show.legend = NA) +
+  coord_sf(xlim = c(-15, 30), ylim = c(45, 65)) +
+  labs(x = "Lon", y = "Lat") +
+  scale_y_continuous(breaks = seq(45,65,10)) +
+  # scale_x_continuous(breaks = seq(45,65,10)) +
+  # my_ggtheme_m() +
+  theme(legend.position = "",
+        panel.background = element_rect(fill = "white", color = "white"),
+        panel.grid.major =  element_line(color = "white")
+        ) +
+  scale_colour_viridis(
+    name="Average Gain of\nStock-Share Ratio (%)",
+    limits = c(0,max(Eu_data$per_chng)),
+    breaks = seq(0,max(Eu_data$per_chng),5),
+    # option="magma",
+    # direction = -1
+  );Eu_map
+
+## ------------------- ##
+# Final Figure
+## ------------------- ##
+
+gc()
+plot_all <- ggdraw() +
+      # map
+      draw_plot(Prop_map, x = 0, y = 0.45, width = 1, height = 0.58) +
+      # Lat America reg plor
+      draw_plot(Lat_am_map, x = 0.1, y = 0.08, width = 0.3, height = 0.4) +
+      # Europe reg plot
+      draw_plot(Eu_map, x = 0.35, y = 0.08, width = 0.3, height = 0.4) +
+      # Med regional plot
+      draw_plot(Med_map, x = 0.65, y = 0.08, width = 0.3, height = 0.4) +
+      draw_plot_label(label = c("a", "b"), 
+                      size = 20,
+                      x = c(0, 0), 
+                      y = c(1, 0.45)
+      ) #+
+      # draw_plot_label(label = c("Revenue (USD) | Catch (Tonnes)","(Million)"), 
+                      # size = 8,
+                      # x = c(0.325,0.485), 
+                      # y = c(0.31,0.29)
+      )
+
+    save_plot(
+      # paste(Figures_Path,"100_25_Bins_Value_Region.png",sep=""),
+      "test.png",
+      plot_all,
+      base_height = 8,
+      base_width = 8
     )

### 2020-08-19 Scripts/Initial_Analysis.Rmd
Included value participation on figure 3

@@
       select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr) %>% 
-      mutate(Tresh= NA)
+      mutate(tresh= NA)
 
@@
       mutate(
-        Tresh_one = ifelse(trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise),"One",NA),
-        Tresh_two = ifelse(trend > (historic_sd$hist_mean+(2*historic_sd$noise)) | trend < (historic_sd$hist_mean-(2*historic_sd$noise)),"Two",NA)
-      ) %>%
-      gather("Tresh","YesNo",Tresh_one,Tresh_two) %>% 
+        tresh_one = ifelse(trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise),"One",NA),
+        tresh_two = ifelse(trend > (historic_sd$hist_mean+(2*historic_sd$noise)) | trend < (historic_sd$hist_mean-(2*historic_sd$noise)),"Two",NA)
+      ) %>%
+      gather("tresh","YesNo",tresh_one,tresh_two) %>% 
       filter(!is.na(YesNo)) %>% 
-      group_by(taxon_key,neighbour_id,n_ens,Tresh) %>%
+      group_by(taxon_key,neighbour_id,n_ens,tresh) %>%
       summarise(emerging_yr= min(year)) %>%
@@
       ungroup() %>%
-      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr,Tresh)
+      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr,tresh)
 
@@
 # --------------- #
+# Load SAU catch and revenue data
+# --------------- #
+
+Sau_data <-fread(paste(my_path("D"),"SAU/sau_catch_value_country_taxon_JEPA.csv", sep ="")) %>%
+  clean_names() %>% 
+  # Remove high seas and 207 and 208 that are unknown
+  filter(!fishing_entity_id %in% c(213,223)) %>%  # 207,208
+  # Allocate catch of islands to UK
+  mutate(fishing_entity_id = ifelse(fishing_entity_id %in% c(219,220), 183,fishing_entity_id)) %>% 
+  # Convert to 2019 real USD
+  mutate(values = 1.16 * value) %>%
+  # Average 10 years of data
+  group_by(fishing_entity_id,taxon_key) %>% 
+  summarise_at(c("catch","value"),
+               mean,na.rm=T) %>% 
+  gather("variable","value",catch:value)
+
+# --------------- #
 # Load shapefiles for figures
@@
 World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf")) %>%
+  clean_names() %>% 
   # st_transform(crs = "+proj=eck4") %>%
@@
 
-##### Fig. Netowrk
+##### Fig. 2 and S1. (Netowrk)
 
@@
 
-# time <- "per_mid_plot"
-time <- "per_early_plot"
+time <- "per_mid_plot"
+# time <- "per_early_plot"
 
@@
   ) +
-  ggsave("./Figures/figure_connections_early_n50.png",
+  ggsave("./Figures/figS1a.png",
   width = 10,
@@
 
-gc()
-plot_all <- ggdraw() +
-      # map
-      draw_plot(Prop_map, x = 0, y = 0.45, width = 1, height = 0.58) +
-      # Lat America reg plor
-      draw_plot(Lat_am_map, x = 0.1, y = 0.08, width = 0.3, height = 0.4) +
-      # Europe reg plot
-      draw_plot(Eu_map, x = 0.35, y = 0.08, width = 0.3, height = 0.4) +
-      # Med regional plot
-      draw_plot(Med_map, x = 0.65, y = 0.08, width = 0.3, height = 0.4) +
-      draw_plot_label(label = c("a", "b"), 
-                      size = 20,
-                      x = c(0, 0), 
-                      y = c(1, 0.45)
-      ) #+
-      # draw_plot_label(label = c("Revenue (USD) | Catch (Tonnes)","(Million)"), 
-                      # size = 8,
-                      # x = c(0.325,0.485), 
-                      # y = c(0.31,0.29)
-      )
-
-    save_plot(
-      # paste(Figures_Path,"100_25_Bins_Value_Region.png",sep=""),
-      "test.png",
-      plot_all,
-      base_height = 8,
-      base_width = 8
-    )
+# gc()
+# plot_all <- ggdraw() +
+#       # map
+#       draw_plot(Prop_map, x = 0, y = 0.45, width = 1, height = 0.58) +
+#       # Lat America reg plor
+#       draw_plot(Lat_am_map, x = 0.1, y = 0.08, width = 0.3, height = 0.4) +
+#       # Europe reg plot
+#       draw_plot(Eu_map, x = 0.35, y = 0.08, width = 0.3, height = 0.4) +
+#       # Med regional plot
+#       draw_plot(Med_map, x = 0.65, y = 0.08, width = 0.3, height = 0.4) +
+#       draw_plot_label(label = c("a", "b"), 
+#                       size = 20,
+#                       x = c(0, 0), 
+#                       y = c(1, 0.45)
+#       ) #+
+#       # draw_plot_label(label = c("Revenue (USD) | Catch (Tonnes)","(Million)"), 
+#                       # size = 8,
+#                       # x = c(0.325,0.485), 
+#                       # y = c(0.31,0.29)
+#       )
+#     
+#     
+#     save_plot(
+#       # paste(Figures_Path,"100_25_Bins_Value_Region.png",sep=""),
+#       "test.png",
+#       plot_all,
+#       base_height = 8,
+#       base_width = 8
+#     )
 
@@
 
-##### Histogram of change
+##### Fig S2 (Histogram of change)
 
@@
 
- results %>% 
+ Prop_results %>% 
   gather("time_step","change",per_early,per_mid) %>% 
@@
 
-####  Year of Eergence Analysisç
+####  Year of Eergence Analysis
 
@@
     n_ens == 10#,
-    # Tresh == "Tresh_one"
-  ) %>% 
-  group_by(Tresh,taxon_key,emerging_yr) %>% 
+    # tresh == "tresh_one"
+  ) %>% 
+  group_by(tresh,taxon_key,emerging_yr) %>% 
   summarise(
-    n_trans = length(unique(eez_name))
-  ) %>% 
-  group_by(Tresh,taxon_key) %>% 
+    n_trans = length(unique(sf_eez_name))
+  ) %>% 
+  group_by(tresh,taxon_key) %>% 
   summarise(
@@
   ) %>% 
-  group_by(Tresh,ntrans) %>% 
+  group_by(tresh,ntrans) %>% 
   summarise(
@@
   # View()
-  # group_by(Tresh) %>% 
+  # group_by(tresh) %>% 
   # summarise(sum(n_taxa))
@@
       y = n_taxa,
-      color = Tresh
+      color = tresh
     )
@@
     n_ens == 10#,
-    # Tresh == "Tresh_one"
-  ) %>% 
-  group_by(Tresh,eez_name,emerging_yr) %>% 
+    # tresh == "tresh_one"
+  ) %>% 
+  group_by(tresh,sf_eez_name,emerging_yr) %>% 
   summarise(
@@
   ) %>% 
-  group_by(Tresh,eez_name) %>% 
+  group_by(tresh,sf_eez_name) %>% 
   summarise(
@@
   ) %>% 
-  group_by(Tresh,ntrans) %>% 
+  group_by(tresh,ntrans) %>% 
   summarise(
-    n_taxa = length(unique(eez_name))
+    n_taxa = length(unique(sf_eez_name))
   ) %>% 
@@
   View()
-  # group_by(Tresh) %>%
+  # group_by(tresh) %>%
   # summarise(sum(n_taxa))
@@
   #     y = n_taxa,
-  #     color = Tresh
+  #     color = tresh
   #   )
@@
 stock_d_hist <- Eme_Res %>% 
-  distinct(taxon_key,neighbour_id,Tresh, .keep_all = TRUE) %>%
-  group_by(emerging_yr,taxon_key,Tresh) %>%
-  summarise(n_tax = length(unique(eez_name))) %>%
+  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
+  group_by(emerging_yr,taxon_key,tresh) %>%
+  summarise(n_tax = length(unique(sf_eez_name))) %>%
   # summarise(n()) %>% 
@@
 eez_d_hist <- Eme_Res %>% 
-  distinct(taxon_key,neighbour_id,Tresh, .keep_all = TRUE) %>%
-  group_by(emerging_yr,eez_name,Tresh,neighbour_id) %>%  
+  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
+  group_by(emerging_yr,sf_eez_name,tresh,neighbour_id) %>%  
   # summarise(n_tax = length(unique(taxon_key))) %>% 
@@
   ungroup() %>% 
-  select(-eez_name)
+  select(-sf_eez_name)
 
@@
   bind_rows(eez_d_hist) %>% 
-  filter(Tresh != "NA") %>% 
+  filter(tresh != "NA") %>% 
   ggplot() +
@@
   scale_fill_viridis(discrete = T) +
-  facet_wrap(~Tresh) +
+  facet_wrap(~tresh) +
   ggsave("./Figures/hist_freq.png",
@@
     # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
-    group_by(taxon_key,Tresh) %>% 
+    group_by(taxon_key,tresh) %>% 
     summarise(
@@
     ) %>% 
-    group_by(taxon_emer,Tresh) %>% 
+    group_by(taxon_emer,tresh) %>% 
   summarise(
@@
   ) %>% 
-  group_by(Tresh) %>% 
+  group_by(tresh) %>% 
   mutate(
@@
     # distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
-    group_by(eez_name,Tresh) %>% 
+    group_by(sf_eez_name,tresh) %>% 
     summarise(
@@
   bind_rows(eez_d) %>% 
-  filter(Tresh != "NA") %>% 
+  filter(tresh != "NA") %>% 
   mutate(
@@
     ) +
-    facet_wrap(~Tresh) + 
+    facet_wrap(~tresh) + 
     # Comment out for option A (Change plot name!!)
@@
     scale_x_continuous("Year of emergence",
-                       breaks = seq(2005,2100,10),
+                       breaks = seq(2005,2100,15),
                        limits = c(2005,2100)
@@
     scale_fill_viridis(discrete = T) +
-    facet_wrap(~Tresh) + 
-  ggsave("./Figures/figure_accumulative_curve_per.png",
-         width = 6,
-         height = 6,
+    facet_wrap(~tresh) + 
+  ggsave("./Figures/Fig4.png",
+         width = 8,
+         height = 8,
          units = "in"
@@
 
-##### Mapping the time of emergence
+##### Fig. 3 Mapping the time of emergence
 
@@
 # Per EEZs
-eez_d_hist <- Eme_Res %>% 
-  group_by(eezid  = sf_eezid,fishing_entity,tresh) %>% 
+Eez_d_hist <- Eme_Res %>% 
+  group_by(eezid= eez_id ,sau_fishing_entity,tresh) %>% 
   # summarise(n_tax = length(unique(taxon_key))) %>% 
@@
     ) %>% 
-  gather("metric","value",mean_yr:n_taxa)
+  gather("metric","value",mean_yr:n_taxa) %>% 
+  filter(metric != "n_taxa")
+
+# Estimate the value (catch and USD) of emerging species
+Eme_value_data <- Eme_Res %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,tresh,taxon_key) %>% 
+  summarise(n()) %>% 
+  left_join(Sau_data,
+            by = c("fishing_entity_id","taxon_key")) %>% 
+  filter(!is.na(variable)) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,tresh,variable) %>% 
+  summarise(
+   sum_value = sum(value,na.rm=T),
+    n_taxa = length(unique(taxon_key)) 
+  )
+
+# Total value of each fishing nation
+
+Total_value <- Sau_data %>% 
+  group_by(fishing_entity_id,variable) %>% 
+  summarise(tot_value= sum(value,na.rm=T))
+
+Value_map_data <- Eme_value_data %>% 
+  left_join(Total_value,
+            by = c("fishing_entity_id","variable")) %>% 
+  mutate(participation = (sum_value/tot_value)*100)
 
@@
 fish_d_hist <- Eme_Res %>% 
-  group_by(fishing_entity,tresh) %>% 
+  group_by(sau_fishing_entity,tresh) %>% 
   # summarise(n_tax = length(unique(taxon_key))) %>% 
@@
     ) %>% 
-  filter(!is.na(fishing_entity))
+  filter(!is.na(sau_fishing_entity))
 
@@
 
-ggplot(eez_d_hist) +
+ggplot(subset(eez_d_hist, tresh == "Tresh_one")) +
   geom_density(
@@
       x = value,
-      fill = Tresh
+      fill = tresh
     ),
@@
 World_map_nemer <- World_map %>%
-  mutate(fishing_entity = ifelse(fishing_entity == "United States","USA", fishing_entity)) %>% 
-  left_join(fish_d_hist,
-            by = "fishing_entity")
+  left_join(Value_map_data,
+            by = c("sau_fishing_entity","fishing_entity_id")
+  )
+
+grey_land <- World_map %>% 
+  filter(!fishing_entity_id %in% World_map_nemer$fishing_entity_id)
+
+grey_eez <- World_eez_sf %>% 
+  filter(!eezid %in% eez_d_hist$eezid)
 
@@
 
-  Tresh_list <- c("Tresh_one","Tresh_two")
-  name_save <- paste("./Figures/emer_map_",Tresh_list[i],".png")
+  tresh_list <- c("Tresh_one","Tresh_two")
+  name_save <- paste("./Figures/emer_map_par_",tresh_list[i],".png",sep = "")
+
+  World_map_nemer <- World_map_nemer %>% 
+    filter(
+      tresh ==  tresh_list[i], 
+      variable == "value"
+    )
 
@@
               by = "eezid") %>% 
-    filter(tresh == Tresh_list[i],
+    filter(tresh == tresh_list[i],
            metric == "mean_yr") %>% 
@@
     ) +
-    geom_sf(data = World_map_nemer, aes(fill = value +2000   # to match scales
-                                        ),
+    geom_sf(data = World_map_nemer,
+            aes(fill = participation+2000),
             color =  "white",
@@
             ) +
+    # grey maps
+    geom_sf(data = grey_land, aes(),fill = "grey50", color =  "white",size = 0.1) +
+    geom_sf(data = grey_eez, aes(),fill = "grey50", color =  "white",size = 0.1) +
     annotate("text", 

### 2020-08-19 Scripts/Initial_Analysis.Rmd
All figures created. Next: Stats

@@
 
-#### Data Needed for Analysis
+#### Read Data for Analysis
 
@@
 Matching_names <- my_path("G","Spatial/SAU", "matching_names.csv", read = T)
+
+# Species depth clasification
+
+Spp_depth <-fread(paste(my_path("D"),"SAU/SppTaxonName.csv", sep ="")) %>% 
+  clean_names()
 
@@
   mutate(
-    per_early = ifelse(is.na(per_early_plot),0,per_early_plot),
-    per_mid = ifelse(is.na(per_mid_plot),0,per_mid_plot)
+    per_early = ifelse(is.na(per_early),0,per_early),
+    per_mid = ifelse(is.na(per_mid),0,per_mid)
   ) %>% 
@@
   filter(
-    variable == "mean_eez"#,
+    variable %in% c("mean_eez","n_spp")#,
     # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
@@
   mutate_at(vars(X,Y), round,2) %>% 
-  gather("time_step","per_chng",per_mid_plot:per_early_plot)
+  gather("time_step","per_chng",per_early:per_mid) %>% 
+  mutate(per_plot = ifelse(per_chng > 100,100,per_chng))
 
@@
 
-Data <- Data %>% 
-  gather("time_step","per_chng",per_mid_plot:per_early_plot)
+# Proportion map data 
+Map_data <- Data %>% 
+  gather("time_step","per_chng",per_early:per_mid) %>% 
+  mutate(per_plot = ifelse(per_chng > 100,100,per_chng))
+
+# Average number of taxa to change per eez
+
+n_taxa_data <- Prop_results %>% 
+  filter(variable == "n_spp") %>% 
+  rename(sf_eez_name = eez_name) %>% 
+  left_join(Matching_names) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id) %>% 
+  summarise(total_spp = mean(early,na.rm=T))
+
+n_taxa_map <- World_map %>% 
+  left_join(n_taxa_data,
+            by = c("sau_fishing_entity","fishing_entity_id")
+            )
 
@@
 
-time <- "per_mid_plot"
-# time <- "per_early_plot"
+# time <- "per_mid_plot"
+time <- "per_early"
 
@@
 Prop_map <- ggplot() +
-  geom_sf(data = subset(World_map, sovereignt != "Antarctica"), size = 0.1, fill = "gray50", col = "gray50") +
+  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+          aes(fill = total_spp),
+          size = 0.1,col = "white") +
   geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
-  geom_point(data = subset(Data, time_step == time & per_chng < 0),
+  geom_point(data = subset(Map_data, time_step == time & per_plot < 0),
              aes(start_lon,
                  start_lat,
-                 col = per_chng),
+                 col = per_plot),
              size = 0.05,
@@
              stroke = 1.5) +
-  geom_path(data = subset(paths, time_step == time & per_chng > 0),
+  geom_path(data = subset(paths, time_step == time & per_plot > 0),
             aes(X, Y,
                 group = interaction(line_id,subline_id),
-                col = per_chng
+                col = per_plot
             ),
             size = 0.3,
-            arrow = arrow(length = unit(0.009, "npc"),
+            arrow = arrow(length = unit(0.008, "npc"),
                           ends = "first"),
@@
   scale_colour_viridis(
-    name="Average Gain of\nStock-Share Ratio (%)",
-    limits = c(0,max(Data$per_chng)),
-    breaks = seq(0,max(Data$per_chng),5),
+    name="Average Gain of Stock-Share Ratio (%)\nand Number of taxa (n)",
+    limits = c(0,max(Map_data$per_plot)),
+    breaks = seq(0,max(Map_data$per_plot),10),
     # option="magma",
@@
   ) +
-  ggsave("./Figures/figS1a.png",
-  width = 10,
-  height = 6,
+  scale_fill_viridis(
+    name="Average Gain of Stock-Share Ratio (%)\nand Number of taxa (n)",
+    limits = c(0,max(Map_data$per_plot)),
+    breaks = seq(0,max(Map_data$per_plot),10),
+    alpha = 0.8
+    # option="magma",
+    # direction = -1
+  ) +
+  ggsave("./Figures/fig2ax.png",
+  width = 12,
+  height = 8,
   units = "in"
@@
 
-Lat_am_eez %>% filter(str_detect(sf_eez_name,"Tri"))
-
-Lat_am_data <- Data %>% 
+# Lat_am_eez %>% filter(str_detect(sf_eez_name,"Tri"))
+
+Lat_am_data <- Map_data %>% 
   filter(source %in% Lat_am_eez$sf_eez_name)
@@
   geom_sf(data = Lat_am_eez, size = 0.1, fill = "white", col = "gray30") +
-  geom_sf(data = Lat_am_land, size = 0.1, fill = "gray50", col = "gray40") +
-  geom_point(data = subset(Lat_am_data, time_step == time & per_chng < 0 ),
+  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+          aes(fill = total_spp),
+          size = 0.1,col = "white") +
+  geom_point(data = subset(Lat_am_data, time_step == time & per_plot < 0 ),
              aes(start_lon,
                  start_lat,
-                 col = per_chng),
+                 col = per_plot),
              size = 1,
@@
              stroke = 1.5) +
-  geom_path(data = subset(Lat_am_paths, time_step == time & per_chng > 0),
+  geom_path(data = subset(Lat_am_paths, time_step == time & per_plot > 0),
             aes(X, Y,
                 group = interaction(line_id,subline_id),
-                col = per_chng
+                col = per_plot
             ),
@@
   theme(legend.position = "",
-        panel.background = element_rect(fill = "white")) +
+        panel.background = element_rect(fill = "white", color = "white"),
+        panel.grid.major =  element_line(color = "white")
+        ) +
   scale_colour_viridis(
     name="Average Gain of\nStock-Share Ratio (%)",
-    limits = c(0,max(Lat_am_data$per_chng)),
-    breaks = seq(0,max(Lat_am_data$per_chng),5),
+    limits = c(0,max(Lat_am_data$per_plot)),
+    breaks = seq(0,max(Lat_am_data$per_plot),10),
+    # option="magma",
+    # direction = -1
+  )+
+  scale_fill_viridis(
+    name="Average Gain of\nStock-Share Ratio (%)",
+    limits = c(0,max(Lat_am_data$per_plot)),
+    breaks = seq(0,max(Lat_am_data$per_plot),10),
     # option="magma",
@@
 # Med data
-Med_data <- Data %>% 
+Med_data <- Map_data %>% 
   filter(source %in% Med_eez$sf_eez_name)
@@
   geom_sf(data = Med_eez, size = 0.5, fill = "white", col = "gray30") +
-  geom_sf(data = Med_land, size = 0.5, fill = "gray50", col = "gray40") +
-  geom_point(data = subset(Med_data, time_step == time & per_chng < 0 ),
+  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+          aes(fill = total_spp),
+          size = 0.1,col = "white") +
+  geom_point(data = subset(Med_data, time_step == time & per_plot < 0 ),
              aes(start_lon,
                  start_lat,
-                 col = per_chng),
+                 col = per_plot),
              size = 1,
@@
              stroke = 1.5) +
-  geom_path(data = subset(Med_paths, time_step == time & per_chng > 0),
+  geom_path(data = subset(Med_paths, time_step == time & per_plot > 0),
             aes(X, Y,
                 group = interaction(line_id,subline_id),
-                col = per_chng
+                col = per_plot
             ),
@@
   theme(legend.position = "",
-        panel.background = element_rect(fill = "white")) +
+        panel.background = element_rect(fill = "white", color = "white"),
+        panel.grid.major =  element_line(color = "white")
+        ) +
   scale_colour_viridis(
     name="Average Gain of\nStock-Share Ratio (%)",
-    limits = c(0,max(Med_data$per_chng)),
-    breaks = seq(0,max(Med_data$per_chng),5),
+    limits = c(0,max(Med_data$per_plot)),
+    breaks = seq(0,max(Med_data$per_plot),5),
+    # option="magma",
+    # direction = -1
+  )+
+  scale_fill_viridis(
+    name="Average Gain of\nStock-Share Ratio (%)",
+    limits = c(0,max(Med_data$per_plot)),
+    breaks = seq(0,max(Med_data$per_plot),5),
     # option="magma",
@@
 
-Eu_eez %>% filter(str_detect(sf_eez_name,"Romania"))
-
-Eu_data <- Data %>% 
+# Eu_eez %>% filter(str_detect(sf_eez_name,"Romania"))
+
+Eu_data <- Map_data %>% 
   filter(source %in% Eu_eez$sf_eez_name)
@@
   geom_sf(data = Eu_eez, size = 0.5, fill = "white", col = "gray30") +
-  geom_sf(data = Eu_land, size = 0.5, fill = "gray50", col = "gray40") +
-  geom_point(data = subset(Med_data, time_step == time & per_chng < 0 ),
+  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+          aes(fill = total_spp),
+          size = 0.1,col = "white") +
+  geom_point(data = subset(Med_data, time_step == time & per_plot < 0 ),
              aes(start_lon,
                  start_lat,
-                 col = per_chng),
+                 col = per_plot),
              size = 1,
@@
              stroke = 1.5) +
-  geom_path(data = subset(Eu_paths, time_step == time & per_chng > 0),
+  geom_path(data = subset(Eu_paths, time_step == time & per_plot > 0),
             aes(X, Y,
                 group = interaction(line_id,subline_id),
-                col = per_chng
+                col = per_plot
             ),
@@
     name="Average Gain of\nStock-Share Ratio (%)",
-    limits = c(0,max(Eu_data$per_chng)),
-    breaks = seq(0,max(Eu_data$per_chng),5),
+    limits = c(0,max(Eu_data$per_plot)),
+    breaks = seq(0,max(Eu_data$per_plot),5),
+    # option="magma",
+    # direction = -1
+   )+
+  scale_fill_viridis(
+    name="Average Gain of\nStock-Share Ratio (%)",
+    limits = c(0,max(Med_data$per_plot)),
+    breaks = seq(0,max(Med_data$per_plot),5),
     # option="magma",
@@
   )
+
+```
+
+### Species level analysis
+
+#### Demersal Vs Pelagic
+
+```{r deme_pela, eval = F, echo =F}
+
+Dem_pel <- Eme_Res %>% 
+  filter(tresh == "Tresh_one",
+         !is.na(un_sub_region)) %>% 
+  left_join(Spp_depth)
+
+Dem_pel_mean <- Dem_pel %>% 
+  group_by(dem_pel) %>% 
+  summarise(mean  = mean(emerging_yr))
+
+ggplot(Dem_pel) +
+  geom_boxplot(
+    aes(
+      x = reorder(dem_pel,emerging_yr),
+      y = emerging_yr,
+      fill = un_sub_region
+    )
+  ) +
+  geom_line(data = Dem_pel_mean,
+    aes(
+      x = dem_pel,
+      y = mean
+    )
+)
+
+ggplot(Dem_pel) +
+  geom_boxplot(
+    aes(
+      x = reorder(dem_pel,emerging_yr),
+      y = emerging_yr
+    )
+  ) +
+  facet_wrap(~un_sub_region)
 

### 2020-08-19 Scripts/Manuscript.Rmd
Results draft

@@
 
+max_values$per_early
+
 ### Average
@@
 
-Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by early $21^{st}$ century (Fig.
-1).
-The change in SSR will range from gains of over 100% of a sertain shared stock (e.g.
-Guatemala' SSR of *species* shared with Mexico is projected to go from 12% to 26%) to loses of up to -71% of SSR (e.g.
-Honduras SSR of *species* shared with the Nicaragua in the Pacific coast is projected to decline from 50% to 14%).
-Despite no substantial changes in SSR per EEZ can be seen by the mid $21^st$ century (**Supplements Fig 1 (map)**), changes in SSR are expected to increase in intensity (**Supplements Fig 2 (histogram)**).
+We estimate the changes from today (1951-2005) to the early (2020-2040) and mid $21^{st}$ century (2040-2060) in transboundary Stock Share Ratio between neighbouring countries.
+Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
+1A).
+While most EEZs will see and average species change lower than 20%, a few EEZs are expected to have 50% change in SSR (Fig.
+2A).
+For example.
+Guatemala' SSR of *species* shared with Mexico is projected to go from 12% to 26% while Honduras SSR of *species* shared with the Nicaragua in the Pacific coast is projected to decline from 50% to 14%.
+Regions where EEZs are particularly small and/or have multiple borders (e.g.
+The Caribeean) will see a complex dynamic of changes in SSR (Fig.
+2B).
+Towards the mid $21^st$ century, changes in SSR per EEZ will no be substantial (**Supplements Fig 1 (map)**) but are still are expected to increase in intensity (**Supplements Fig 2 (histogram)**).
 
@@
 
-img <- readPNG("../Figures/figure_connections_n50.png")
+img <- readPNG("../Figures/Fig2.png")
 grid.raster(img)
@@
 ![Average transboundary species SSR gain of neighbouring EEZs by 2030 (2021-2040) relative to 1951-2005 under RCP 8.5.
-Arrows go from the loosing EEZ (point) to the winner (arrowhead)](../Figures/figure_connections_n50.png)
+Arrows go from the loosing EEZ (point) to the winner (arrowhead)](../Figures/Fig2.png)
 
-## Time of Emergence Index
+## Time of Emergence
+
+We estimated the time of emergence between neighbouring EEZs for 633 transbboundary stocks.
+On average, most of the world EEZ's time of emergence will be 2030 $\pm$ 25 yers (Fig S3-A), however, all EEZs have already seen changes for at least one species (Fig S3-B).
+
 # Discussion

### 2020-08-25 Scripts/Initial_Analysis.Rmd
Started code to get sst dat but can't figure out path

@@
 stopCluster(cl)
+
+```
+
+## SST time of emergence
+
+##### Fun `GetEnvData`
+
+```{r get_env_data, eval = T, echo = T}
+
+ModelRCP <- "GFDL85"
+
+getEnvVar=function(yr,ModelRCP,ensemble,Variable){
+
+  filepath <- paste("Z:/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")
+
+  if(Variable == "sst"){
+    sst = paste("SST_",yr,".txt",sep="")
+    read_me_sst <- paste(filepath,sst,sep="")[1]
+    data <- read.csv(read_me_sst,header=F)
+
+  }
+
+  if(Variable == "botom"){
+    botom = paste("botom_",yr,".txt",sep="")
+    read_me_botom <- paste(filepath,botom,sep="")[1]
+    data <- read.csv(read_me_botom,header=F)
+  }
+
+  # Retrun
+
+  getTemp <- data %>% 
+    mutate(ensemble = ensemble)
+}
+
+```
+
+```{r temp_emergence, eval = F, echo = F}
+
+TempEmer <- function(Hist_y = 2005){ # The LAST year of Historical data
+
+  # Historical mean and sd of distance
+  historical <- Centroids_id %>%
+    filter(year <= Hist_y) %>% 
+    ungroup() %>%
+    group_by(taxon_key,ensemble,eez_name,neighbour_id) %>%
+    summarise_at(vars("distance"),
+                 c(hd_temp_mean=mean,
+                   hd_temp_sd=sd,
+                   n_yrs =~n())
+                 )
+
+    # Future results
+  future <- Centroids_id %>% 
+    filter(year > Hist_y)
+
+   # Get unique eezs
+  Unique_EEZs <- unique(EEZ_ids$eez_name)
+
+  # Transboundary Index
+  Trans_index <- Centroids_id %>% 
+    left_join(historical,
+              by = c("taxon_key","eez_name","ensemble","neighbour_id")
+    ) %>%
+    # Re arrenging table for mutate
+    gather("variable","value",distance:hd_temp_sd) %>%
+    ungroup() %>% 
+    mutate(value = ifelse(is.na(value),0,value)) %>% 
+    mutate(
+      names = ifelse(eez_name ==Unique_EEZs[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_")
+    )
+    ) %>%
+    select(-eez_name,-variable) %>%
+    spread(names,value) %>% 
+    # mutate index
+    mutate(
+      trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
+    # )
+    ) %>% 
+    # Ten years average of results
+    group_by(taxon_key,ensemble,neighbour_id) %>% 
+    mutate(RMean = rollmean(x = trans_index, 
+                            10, 
+                            align = "right", 
+                            fill = trans_index,
+                            na.rm=T)
+    ) %>% 
+    # filter(year > 1960) %>%
+    select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)
+
+  # Average of 10 years mean of ensemble members results
+  mean_sd_index <- Trans_index %>% 
+    group_by(taxon_key,year,neighbour_id) %>% 
+    summarise_at(vars("RMean"),
+                 c(trend=mean,
+                   ensemble_sd=sd,
+                   n_ens = ~n()
+
+    ),
+    na.rm=T
+    )
+
+    # Valid ensembles
+  # Only perform analysis on ensembles containing all years between 1951 and 2040
+    n_years <- mean_sd_index %>% 
+      group_by(n_ens) %>% 
+      summarise(
+        n_year = n()
+      ) %>% 
+      filter(
+        n_year == 149
+      )
+
+    mean_sd_index <- mean_sd_index %>% 
+      filter(n_ens %in% n_years$n_ens)
+
+  # Estimate noice from ensemble members
+  historic_sd <- Trans_index %>% 
+    filter(year <= Hist_y) %>% 
+    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
+    summarise_at(vars("trans_index"),
+                 c(trend= mean),
+                 na.rm=T
+                 ) %>% 
+    group_by(taxon_key,neighbour_id) %>% #Capture ensemble variation
+    summarise_at(vars("trend"),
+                 c(hist_mean=mean,
+                   noise=sd,
+                   n_ens = ~n())
+                 )
+
+  # For testing
+  if(Plot == T & length(historic_sd$hist_mean) > 1){
+  # Plot the result, bruh
+
+  # Step to determine if data exists
+  Plot_Test <- nrow(subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)))>0
+
+  ggplot() +
+    geom_line(data = mean_sd_index,
+              aes(
+                x = year,
+                y = trend
+              )
+    ) +
+    geom_ribbon(data = mean_sd_index,
+                aes(x=year,
+                    ymin=historic_sd$hist_mean-historic_sd$noise,
+                    ymax=historic_sd$hist_mean+historic_sd$noise),
+                alpha = 0.5,
+                fill = "grey50"
+    ) +
+    geom_ribbon(data = mean_sd_index,
+                aes(x=year,
+                    ymin=historic_sd$hist_mean-(historic_sd$noise*2),
+                    ymax=historic_sd$hist_mean+(historic_sd$noise)*2)
+                ,
+                alpha = 0.5,
+                fill = "grey30"
+    ) +
+    geom_line(data = mean_sd_index,
+              aes(
+                x = year,
+                y = historic_sd$hist_mean
+              ),
+              colour = "grey50"
+    ) +
+    # Conditional, will only plot circles if the data exists
+    {if(Plot_Test)geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
+               aes(
+                 x = year,
+                 y = trend,
+                 colour = "red",
+               ),
+               shape = 1,
+               size = 3
+    )
+      } +
+    geom_vline(xintercept = Hist_y,colour="blue",alpha = 0.5) +
+    scale_x_continuous(breaks = seq(1950,2100,5)) +
+    theme_classic()
+
+    tk <- unique(mean_sd_index$taxon_key)
+
+    ggsave(paste(tk,Id,"plot_example.jpg",sep="_"),
+       plot = last_plot(),
+       width = 10,
+       height = 8,
+       units = "in",
+       path = my_path("R")
+       )
+
+  }
+
+  # If no historic data then emergin year is the first year of data
+  if(nrow(historic_sd) == 0){
+
+    Final_Result <- mean_sd_index %>%
+      group_by(taxon_key,neighbour_id,n_ens) %>%
+      summarise(emerging_yr= min(year)) %>%
+      left_join(EEZ_ids,
+                by = "neighbour_id") %>%
+      ungroup() %>%
+      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr) %>% 
+      mutate(tresh= NA)
+
+  return(Final_Result)
+
+  }else{
+
+    Final_Result <- mean_sd_index %>%
+      filter(year > Hist_y) %>%
+      mutate(
+        tresh_one = ifelse(trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise),"One",NA),
+        tresh_two = ifelse(trend > (historic_sd$hist_mean+(2*historic_sd$noise)) | trend < (historic_sd$hist_mean-(2*historic_sd$noise)),"Two",NA)
+      ) %>%
+      gather("tresh","YesNo",tresh_one,tresh_two) %>% 
+      filter(!is.na(YesNo)) %>% 
+      group_by(taxon_key,neighbour_id,n_ens,tresh) %>%
+      summarise(emerging_yr= min(year)) %>%
+      left_join(EEZ_ids,
+                by = "neighbour_id") %>%
+      ungroup() %>%
+      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr,tresh)
+
+  return(Final_Result)
+  }
+}
+
+```
+
+```{r william_code, eval = F, echo = F}
+
+datascale<-subset(testdata,!is.na(MHW) & !is.na(MeanChange))
+MHWBreaks<-c(0.2,0.4,0.5,0.6,0.8,2.1)
+MeanChangeBreaks<-c(-0.4,0.6,0.7,0.8,1.0,2.0)
+#testdata$MHW <- with(testdata, cut(MHW,
+#                                   breaks=quantile(MHW, probs=seq(0,1, by=0.2), na.rm=TRUE),
+#                                   include.lowest=TRUE))
+testdata$MHW <- with(testdata, cut(MHW,
+                                   breaks=MHWBreaks, na.rm=TRUE),
+                                   include.lowest=TRUE)
+testdata$MHW <- as.numeric(testdata$MHW)
+#testdata$MeanChange <- with(testdata, cut(MeanChange,
+#                                          breaks=quantile(MeanChange, probs=seq(0,1, by=0.2), na.rm=TRUE),
+#                                          include.lowest=TRUE))
+testdata$MeanChange <- with(testdata, cut(MeanChange,
+                                          breaks=MeanChangeBreaks, na.rm=TRUE),
+                                          include.lowest=TRUE)
+
+testdata$MeanChange <- as.numeric(testdata$MeanChange)
+
+library(ggplot2)
+d <- data.frame(testdata,fill_val = rgb(blue=0.5,red = testdata$MHW/5, green = testdata$MeanChange/5))
+
+world <- map_data("world")
+map<- ggplot(d, aes(x=lon, y=lat)) +
+  geom_tile(aes(fill = fill_val)) +
+  scale_fill_identity()+
+  geom_cartogram(data=world, map=world,
+                 aes(x=long, y=lat, map_id=region),fill="white",color="black")+
+  coord_quickmap(xlim=c(-180,180), ylim=c(-89, 90))+
+  ylab("")+
+  xlab("")+
+  scale_y_continuous(expand = c(0, -5))+
+  scale_x_continuous(expand = c(0, 0))+
+  labs(title=paste("(A) Sea Surface Temperature",sep=""))+
+  theme(axis.line = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        legend.position="bottom",
+        #panel.border = element_blank(),
+        panel.border = element_rect(colour = "black", fill=NA, size=1),
+        panel.background = element_blank(),
+        #axis.text.y=element_blank(),
+        axis.text=element_text(size=9,colour="black"),
+        plot.title=element_text(hjust=0,vjust=-0.4,face="bold",size=24)
+  )
+dlegend <- expand.grid(x = seq(0, 1, 0.25), y = seq(0, 1, 0.25)) %>%
+  mutate(mix2 = rgb(red = x, green = y, blue = 0.5))
+library(cowplot)
+xyLegend <- ggplot(dlegend, aes(x, y)) +
+  geom_tile(aes(fill = mix2)) +
+  scale_fill_identity() +
+  labs(x = "Higher MHW change->",
+       y = "Higher mean change->") +
+  theme_void() +
+  theme(axis.title = element_text(size = 8),
+        axis.title.y = element_text(angle = 90)) +
+  coord_fixed()
+p=ggdraw() +
+  draw_plot(map, 0, 0, 1, 1) +
+  draw_plot(xyLegend, 0.05, 0.1, 0.2, 0.2)
+#ggsave(p,filename="/Users/william/Google Drive/Projects/MHWs Global/MHW_SST_plot.png",height=9,width=8,dpi=300)
+ggsave(map,filename="/Users/william/Google Drive/Projects/MHWs Global/MHW_SST_plot_map.png",height=8,width=11,dpi=300) (edited)
 

### 2020-08-25 Scripts/Initial_Analysis.Rmd
reversed scale of Fig 3

@@
 
-# Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List_emer,"_emergence.txt",sep="")
+Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List_emer,"_emergence.txt",sep="")
 
@@
   gather("metric","value",mean_yr:n_taxa) %>% 
-  filter(metric != "n_taxa")
+  filter(metric != "n_taxa") %>% 
+  mutate(
+    reverse = abs(value-2100)  + 2000
+  )
 
@@
 
-ggplot(subset(eez_d_hist, tresh == "Tresh_one")) +
+ggplot(subset(Eez_d_hist, tresh == "Tresh_one")) +
   geom_density(
@@
 grey_eez <- World_eez_sf %>% 
-  filter(!eezid %in% eez_d_hist$eezid)
+  filter(!eezid %in% Eez_d_hist$eezid)
 
@@
   tresh_list <- c("Tresh_one","Tresh_two")
-  name_save <- paste("./Figures/emer_map_par_",tresh_list[i],".png",sep = "")
-
-  World_map_nemer <- World_map_nemer %>% 
+  name_save <- paste("./Figures/emer_map_rev_",tresh_list[i],".png",sep = "")
+
+  World_map_part <- World_map_nemer %>% 
     filter(
@@
   map <- World_eez_sf %>% 
-    left_join(eez_d_hist,
+    left_join(Eez_d_hist,
               by = "eezid") %>% 
@@
       aes(
-        fill = value
+        fill = reverse
       ),
@@
     ) +
-    geom_sf(data = World_map_nemer,
+    # Inlcude the participation of species
+    geom_sf(data = World_map_part,
             aes(fill = participation+2000),
@@
       limits = c(2000,2100),
-      breaks = seq(2000,2100,25)
+      breaks = seq(2000,2100,25),
+      labels = seq(2100,2000,-25),
+      direction = -1
     ) +

### 2020-08-27 Scripts/Initial_Analysis.Rmd
continue sst emerg. analysis
Working on the multilinealmodel

@@
   "viridis",
-  "cowplot"
+  "cowplot",
+  "rfishbase"
 )
@@
 
-EEZ_CellID <- my_path("G", extra_path = "Spatial/V_Lam", name="EEZ_CellID.xlsx", read = TRUE) %>%
+EEZ_CellID <- my_path("G", extra_path = "Spatial/DBEM", name="EEZ_CellID.xlsx", read = TRUE) %>%
   clean_names()
@@
 # SAU name relatioal table (Fish For Visa)
-sau_clean_names <- my_path("G", extra_path = "Spatial/SAU/", name="SAU_matching_names.csv", read = TRUE)
+sau_clean_names <- my_path("G", extra_path = "Spatial/SAU/", name="matching_names.csv", read = TRUE)
 
@@
 
-ModelRCP <- "GFDL85"
-
-getEnvVar=function(yr,ModelRCP,ensemble,Variable){
-
-  filepath <- paste("Z:/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")
+getEnvVar=function(yr,ensemble,Variable){
+
+  filepath <- paste("Z:/DATA/Environmental data/GFDL_CMIP5_Large_Ensemble/ens",ensemble,"",sep="/")
 
@@
 # Get taxon list
-Taxon_List_emer <- list.files(my_path("R",extra_path = "Emergence_2005"))
+Taxon_list_emer <- list.files(my_path("R",extra_path = "Emergence_2005"))
 
 # clean names for function
-Taxon_List_emer <- gsub("_.*","",Taxon_List_emer)
+Taxon_list_emer <- gsub("_.*","",Taxon_list_emer)
+
+# SAU exploited species list
+Exploited_spp_list <- my_path("G","SAU","exploited_species_list.csv",read = T) %>%
+  clean_names() %>% 
+  pull(taxon_name)
+
+# Get habitat preference from fishbase
+Spp_env_pref <- rbind(species(Exploited_spp_list, #species from rfishbase package
+                              fields = species_fields$habitat[4])
+                      ) %>% 
+  mutate(taxon_name = Exploited_spp_list) %>% 
+  left_join(my_path("G","SAU","exploited_species_list.csv", read = T) %>%
+              clean_names(),
+            by = "taxon_name")
 
@@
 
+# Clean names to match data
+Matching_names <- my_path("G","Spatial/SAU", "matching_names.csv", read = T) %>% 
+  rename(eezid = eez_id)
+
+Fishing_names <- Matching_names %>% 
+  group_by(sau_fishing_entity,fishing_entity_id) %>% 
+  summarise(n=n()) %>% 
+  select(-n)
+
 Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List_emer,"_emergence.txt",sep="")
 
-# Clean names to match data
-Matching_names <- my_path("G","Spatial/SAU", "matching_names.csv", read = T)
-
-# Species depth clasification
-
+# eez_centroids
+Eez_centroids <- my_path("D", extra_path = "Spatial/", name="eez_centroids.csv", read = TRUE) %>% 
+  rename(sf_eez_name = eez_name) %>% 
+  left_join(Matching_names) %>% 
+  select(eezid,latitude,longitude)
+
+# Species depth clasification (SAU-fishbase)
 Spp_depth <-fread(paste(my_path("D"),"SAU/SppTaxonName.csv", sep ="")) %>% 
   clean_names()
+
+# Time of Emergence of SST using methods here
+Toe_sst <-fread(paste(my_path("D"),"Climate/ToE_sst.csv", sep ="")) %>% 
+  clean_names()
+
+# Biogeographic regions (from Rey 2020)
+Biogeo_reg <-fread(paste(my_path("D"),"Climate/SPATIAL_PARTITION_OCEAN.csv", sep ="")) %>% 
+  clean_names() %>% 
+  rowid_to_column("index")
+
+# Get eez to index relational data
+EEZ_CellID <- my_path("G", extra_path = "Spatial/DBEM", name="EEZ_CellID.xlsx", read = TRUE) %>%
+  clean_names()
+colnames(EEZ_CellID) <- c("eezid","index")
+
+Eez_biogeo <- EEZ_CellID %>% 
+  left_join(Biogeo_reg,
+       by = "index") %>% 
+  # Deal with NaNs in data
+  mutate(biome = ifelse(biome == "NaN",0,biome)) %>% 
+  filter(biome > 0) %>% 
+  # Get frequency of biome per index
+  group_by(eezid,basin,biome) %>% 
+  summarise(nbiom=n(),
+            nbasin=n()) %>% 
+  # Select the more frequent biome per eez
+  group_by(eezid,basin) %>% 
+  top_n(1,nbiom) %>% 
+  group_by(eezid,biome) %>% 
+  top_n(1,nbasin) %>% 
+  ungroup() %>% 
+  select(-nbiom,-nbasin)
+
+# Species catch trend (from FishForVisa)
+
+Catch_trend <- my_path("D","Species","Catch_trend.csv", read = T) %>% 
+  clean_names() %>% 
+  filter(taxon_key %in% Taxon_list_emer) %>% 
+  rename(sau_fishing_entity = territory) %>% 
+  left_join(fishing_names,
+            by = "sau_fishing_entity") %>% 
+  filter(status != "No Trend") %>% 
+  mutate(category = ifelse(status %in% c("Developing","Rebuilding"),"Category A",
+                         ifelse(status == "Max Exploited","Category B",
+                                ifelse(status %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
+                         )
+         )
+  )
 
@@
             by=  "sf_eez_name") %>% 
-  clean_names()
+  clean_names() %>% 
+  rename(eezid = eez_id)
 
@@
 Eez_d_hist <- Eme_Res %>% 
-  group_by(eezid= eez_id ,sau_fishing_entity,tresh) %>% 
+  group_by(eezid ,sau_fishing_entity,tresh) %>% 
   # summarise(n_tax = length(unique(taxon_key))) %>% 
@@
   facet_wrap(~un_sub_region)
+
+```
+
+### Can we explain the ToE?
+
+```{r multiple_reg, eval = F, echo = F}
+
+# Possible explanatory variables
+
+## Habitat association (DemersPelag)
+## Dem Vs Pelagic (Dem_pel)
+## SST time of Emer (Toe_sst)
+## Stock Status [FishForVisa] (Catch_trend)
+## Biogeographic location (Biogeo_reg)
+
+# AIC value; model between predictive capacity and complexity (i.e., number of explanatory variables) is given by the model with lowest AIC value
+
+# -------------------------------- #
+#### Prepare data
+# -------------------------------- #
+
+Lm_data <- Eme_Res %>% 
+  filter(tresh == "Tresh_one") %>% 
+  left_join(Spp_depth) %>% 
+  select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel) %>% 
+  left_join(Toe_sst,
+            by = c("eezid","tresh")
+            ) %>% 
+  left_join(Catch_trend,
+            by = c("taxon_key","sau_fishing_entity","fishing_entity_id")
+            ) %>% 
+  left_join(Eez_biogeo,
+            by = "eezid") %>% 
+  left_join(Spp_env_pref,
+            by = "taxon_key") %>% 
+  left_join(Eez_centroids,
+            by = "eezid") %>% 
+  mutate(latitude = abs(latitude)) %>% 
+  filter(tresh == "Tresh_one") %>% 
+  distinct()
+
+# -------------------------------- #
+#### Build models
+# -------------------------------- #
+
+# Model 1; all variables
+M_all_var <- lm(emerging_yr ~ sst_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)
+
+all_sum <- summary(M_all_var)
+all_aic <- AIC(M_all_var)
+all_r <-  all_sum$adj.r.squared
+
+# Model 2; non demPel
+M2 <- lm(emerging_yr ~ sst_toe + status + biome + dem_pel, data = Lm_data)
+
+M2_sum <- summary(M2)
+M2_aic <- AIC(M2)
+M2_r <-  M2_sum$adj.r.squared
+
+# Model 3; non biome
+M3 <- lm(emerging_yr ~ sst_toe + status + dem_pel + DemersPelag, data = Lm_data)
+
+M3_sum <- summary(M3)
+M3_aic <- AIC(M3)
+M3_r <-  M3_sum$adj.r.squared
+
+# Model 4; non status
+M4 <- lm(emerging_yr ~ sst_toe + biome + dem_pel + DemersPelag, data = Lm_data)
+
+M4_sum <- summary(M4)
+M4_aic <- AIC(M4)
+M4_r <-  M4_sum$adj.r.squared
+
+# Model 5; non sst
+M5 <- lm(emerging_yr ~ status + biome + dem_pel + DemersPelag, data = Lm_data)
+
+M5_sum <- summary(M5)
+M5_aic <- AIC(M5)
+M5_r <-  M5_sum$adj.r.squared
+
+# Model 6; all + basin
+M6 <- lm(emerging_yr ~ sst_toe + status + biome + basin + dem_pel + DemersPelag, data = Lm_data)
+
+M6_sum <- summary(M6)
+M6_aic <- AIC(M6)
+M6_r <-  M6_sum$adj.r.squared
+
+# Model 7; Categories instead of status
+M7 <- lm(emerging_yr ~ sst_toe + category + biome + dem_pel + DemersPelag, data = Lm_data)
+
+M7_sum <- summary(M7)
+M7_aic <- AIC(M7)
+M7_r <-  M7_sum$adj.r.squared
+
+# Model 8; All variables + ccentroid lat
+M8 <- lm(emerging_yr ~ sst_toe + category + biome + dem_pel + DemersPelag + latitude, data = Lm_data)
+
+M8_sum <- summary(M8)
+M8_aic <- AIC(M8)
+M8_r <-  M8_sum$adj.r.squared
+
+tibble(
+  "model" = c("all variables", "no demPel","no biome","no status","no sst","all + basin","Categories","lat centroids"),
+  "AIC" = c(all_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic),
+  "rsquared" = c(all_r,M2_r,M3_r,M4_r,M5_r,M6_r,M7_r,M8_r)
+)
+
+20282659
 

### 2020-08-27 Scripts/Initial_Analysis.Rmd
Fixed figure 3 labels and LM

@@
   "cowplot",
-  "rfishbase"
+  "rfishbase",
+  "stargazer" # For nice lm tables
 )
@@
 
-Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List_emer,"_emergence.txt",sep="")
-
 # eez_centroids
@@
 
-# Time of Emergence of SST using methods here
-Toe_sst <-fread(paste(my_path("D"),"Climate/ToE_sst.csv", sep ="")) %>% 
+# Temperature Time of Emergence
+Toe_temp <- left_join(my_path("D","Climate", "ToE_sst.csv", read = T),
+                      my_path("D","Climate", "ToE_bottemp.csv", read = T)
+                      ) %>% 
   clean_names()
@@
 Eez_biogeo <- EEZ_CellID %>% 
+  # filter(eezid == 191) %>% 
   left_join(Biogeo_reg,
@@
   # Get frequency of biome per index
-  group_by(eezid,basin,biome) %>% 
-  summarise(nbiom=n(),
-            nbasin=n()) %>% 
+  group_by(eezid,biome) %>% 
+  summarise(nbiom=n()) %>% 
   # Select the more frequent biome per eez
-  group_by(eezid,basin) %>% 
+  group_by(eezid) %>% 
   top_n(1,nbiom) %>% 
-  group_by(eezid,biome) %>% 
-  top_n(1,nbasin) %>% 
   ungroup() %>% 
-  select(-nbiom,-nbasin)
+  select(-nbiom) %>% 
+  distinct(eezid, .keep_all = T) # Croatia is 1/2. We arbritary allocated the first one
 
@@
   rename(sau_fishing_entity = territory) %>% 
-  left_join(fishing_names,
+  left_join(Fishing_names,
             by = "sau_fishing_entity") %>% 
@@
 
+Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_list_emer,"_emergence.txt",sep="")
+
 # Load all sp in one table
@@
             by=  "sf_eez_name") %>% 
-  clean_names() %>% 
-  rename(eezid = eez_id)
+  clean_names()
 
@@
   # get coords for sources
-  left_join(Eez_centroid_eck,
+  left_join(Eez_centroid,
             by = "eez_name") %>%
@@
   # get coords for targets
-  left_join(Eez_centroid_eck,
+  left_join(Eez_centroid,
             by = "eez_name") %>% 
@@
   purrr::map(st_linestring) %>%
-  # st_sfc(crs = 4326) %>%
-  st_sfc(crs = "+proj=eck4") %>% # for tbular plot
+  st_sfc(crs = 4326) %>%
+  # st_sfc(crs = "+proj=eck4") %>% # for tbular plot
   st_sf(geometry = .)%>%
@@
   mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>%
-  # mutate(geometry = (geometry + c(18000000,9000000)) %% c(360) - c(18000000,9000000)) %>%
   st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>%
-  st_wrap_dateline(options = c("WRAPDATELINE=YES"), quiet = TRUE) %>%
-  # sf::st_sf(crs = 4326)
-  sf::st_sf(crs = "+proj=eck4") # for tbular plot
+  sf::st_sf(crs = 4326)
+  # sf::st_sf(crs = "+proj=eck4") # for tbular plot
 
@@
            size = 0.05,
-             col = "gray30",
+           col = "gray80",
            shape = 20,
@@
 geom_path(data = subset(paths, time_step == time & per_plot > 0),
-            aes(X, Y,
+          aes(X,
+              Y,
               group = interaction(line_id,subline_id),
@@
           size = 0.3,
+          lineend = "round",
           arrow = arrow(length = unit(0.008, "npc"),
@@
   limits = c(0,max(Map_data$per_plot)),
-    breaks = seq(0,max(Map_data$per_plot),10),
+  breaks = seq(0,max(Map_data$per_plot),10)#,
   # option="magma",
@@
   limits = c(0,max(Map_data$per_plot)),
-    breaks = seq(0,max(Map_data$per_plot),10),
-    alpha = 0.8
+  breaks = seq(0,max(Map_data$per_plot),10)#,
+  # alpha = 0.8
   # option="magma",
@@
     geom_sf(data = grey_eez, aes(),fill = "grey50", color =  "white",size = 0.1) +
+    # Set Participation labels
     annotate("text", 
-             x =(c(-85.5,-30,25,81,136)+5),
-             y = -93.5,
+             x =-130,
+             y = -69,
+             label = "Participation (USD)", 
+             color="black", 
+             size=4, 
+             angle=0, 
+             fontface="bold",
+             hjust=0) +
+    annotate("text", 
+             x = c(-73,-34,5,45,83),
+             y = -69,
              # x =(c(-8500000,-3000000,2500000,7500000,12800000)),
@@
              color="black", 
-             size=3, 
+             size=4, 
              angle=0, 
@@
              hjust=0) +
-    scale_fill_viridis(
+    # Set ToE labels
+    annotate("text", 
+             x =-130,
+             y = -85,
+             label = "Time of Emergence",
+             color="black", 
+             size=4, 
+             angle=0, 
+             fontface="bold",
+             hjust=0) +
+    annotate("text", 
+             x = c(-77,-37,3,42,82),
+             y = -85,
+             label = seq(2100,2000,-25), 
+             color="black", 
+             size=4, 
+             angle=0, 
+             # fontface="bold",
+             hjust=0) +
+    scale_fill_viridis("",
       limits = c(2000,2100),
       breaks = seq(2000,2100,25),
-      labels = seq(2100,2000,-25),
-      direction = -1
+      labels = c("","","","","")#,
+      # direction = -1
     ) +
     my_ggtheme_m() +
+    theme(
+      legend.box.margin = margin(t = -66, l = 250),
+      legend.text = element_blank(),
+      legend.key.width = unit(5, "line"),
+      axis.text = element_text(colour = "blue")
+    ) +
+    # guides(fill = guide_legend(nrow = 1))+
     ggsave(name_save,
            width = 12,
-           height = 10,
+           height = 6,
            units = "in"
@@
 
-### Can we explain the ToE?
+### Tab.1 Explain the ToE
 
@@
 
-Lm_data <- Eme_Res %>% 
-  filter(tresh == "Tresh_one") %>% 
+Lm_data <- Eme_Res %>% # 19728
+  filter(tresh == "Tresh_one") %>% # 11490
   left_join(Spp_depth) %>% 
   select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel) %>% 
-  left_join(Toe_sst,
+  left_join(Toe_temp,
             by = c("eezid","tresh")
@@
             by = c("taxon_key","sau_fishing_entity","fishing_entity_id")
-            ) %>% 
+            ) %>% # 11492
   left_join(Eez_biogeo,
@@
             by = "eezid") %>% 
-  mutate(latitude = abs(latitude)) %>% 
-  filter(tresh == "Tresh_one") %>% 
-  distinct()
+  mutate(latitude = abs(latitude))
 
@@
 # Model 1; all variables
-M_all_var <- lm(emerging_yr ~ sst_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)
+M_all_var <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)
 
@@
 # Model 2; non demPel
-M2 <- lm(emerging_yr ~ sst_toe + status + biome + dem_pel, data = Lm_data)
+M2 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel, data = Lm_data)
 
@@
 # Model 3; non biome
-M3 <- lm(emerging_yr ~ sst_toe + status + dem_pel + DemersPelag, data = Lm_data)
+M3 <- lm(emerging_yr ~ sst_toe + bot_toe + status + dem_pel + DemersPelag, data = Lm_data)
 
@@
 # Model 4; non status
-M4 <- lm(emerging_yr ~ sst_toe + biome + dem_pel + DemersPelag, data = Lm_data)
+M4 <- lm(emerging_yr ~ sst_toe + bot_toe + biome + dem_pel + DemersPelag, data = Lm_data)
 
@@
 # Model 5; non sst
-M5 <- lm(emerging_yr ~ status + biome + dem_pel + DemersPelag, data = Lm_data)
+M5 <- lm(emerging_yr ~ status + bot_toe + biome + dem_pel + DemersPelag, data = Lm_data)
 
@@
 # Model 6; all + basin
-M6 <- lm(emerging_yr ~ sst_toe + status + biome + basin + dem_pel + DemersPelag, data = Lm_data)
+M6 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)
 
@@
 # Model 7; Categories instead of status
-M7 <- lm(emerging_yr ~ sst_toe + category + biome + dem_pel + DemersPelag, data = Lm_data)
+M7 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + dem_pel + DemersPelag, data = Lm_data)
 
@@
 # Model 8; All variables + ccentroid lat
-M8 <- lm(emerging_yr ~ sst_toe + category + biome + dem_pel + DemersPelag + latitude, data = Lm_data)
+M8 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + DemersPelag + latitude, data = Lm_data)
 
@@
 tibble(
-  "model" = c("all variables", "no demPel","no biome","no status","no sst","all + basin","Categories","lat centroids"),
+  "model" = paste("M",seq(1,8,1),sep=""),
   "AIC" = c(all_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic),
@@
 
-20282659
+# UNIDAS 20282659
+
+# The Winner is M8!
+
+stargazer(M8, title = "Results", align=TRUE)
 

### 2020-08-27 Scripts/Initial_Analysis.Rmd
Finalized Figure 3 for draft

@@
 
-### Random plot
-
-ggplot(subset(Eez_d_hist, tresh == "Tresh_one")) +
-  geom_density(
-    aes(
-      x = value,
-      fill = tresh
-    ),
-    bins = 10,
-    alpha = 0.8
-  ) +
-  facet_wrap(~metric,
-             scales = "free") +
-  scale_fill_viridis(discrete = T) +
-  my_ggtheme_p() +
-  ggsave("./Figures/hist_mean.png",
-         width = 6,
-         height = 6,
-         units = "in"
-  )
-
 World_map_nemer <- World_map %>%
@@
             by = c("sau_fishing_entity","fishing_entity_id")
-  )
+  ) %>% 
+  filter(!is.na(participation))
 
 grey_land <- World_map %>% 
-  filter(!fishing_entity_id %in% World_map_nemer$fishing_entity_id)
+  filter(!fishing_entity_id %in% World_map_nemer$fishing_entity_id) %>% 
+  filter(sovereignt != "Antarctica")
 
@@
   tresh_list <- c("Tresh_one","Tresh_two")
-  name_save <- paste("./Figures/emer_map_rev_",tresh_list[i],".png",sep = "")
+  name_save <- paste("./Figures/emer_map_lat_",tresh_list[i],".png",sep = "")
+
+  ## --------------- ##
+  # Latitudinal plot
+  ## --------------- ##
+
+  ### Random plot
+
+Eez_hist <- Eme_Res %>% 
+  group_by(eezid,eez_name = sf_eez_name) %>% 
+  left_join(Eez_centroid) %>% 
+  filter(tresh == tresh_list[i]) %>% 
+  group_by(latitude) %>% 
+  summarise(
+    number = n()
+  ) %>% 
+  group_by() %>% 
+  mutate(RMean = rollmean(x = number, 
+                          10, 
+                          align = "center", 
+                          fill = number
+  )
+  ) %>% 
+  mutate(lab="lab") %>% 
+  ggplot() +
+  geom_area(
+    aes(
+      x = latitude,
+      y = RMean,
+      fill = lab,
+      color = lab
+    )
+  ) +
+  coord_flip() +
+  scale_x_continuous("",
+    limits = c(-50,80),
+    breaks = c(-50,0,50,80)
+  ) +
+  scale_y_continuous(
+    "Emerging stocks",
+    limits = c(0, 124),
+    breaks = c(0,62,124)
+
+  ) +
+  scale_fill_viridis(option="magma", discrete = T, begin = 0.7) +
+  scale_color_viridis(option="magma", discrete = T, begin = 0.7) +
+  theme_minimal() + 
+  theme(
+    panel.background = element_rect(fill = "white", colour = "white"),
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.text = element_text(size = 12, colour="black"),
+    axis.title.x = element_text(size = 12, colour="black", face = "bold"),
+    legend.position = ""
+  );Eez_hist
+
+  ## --------------- ##
+  # Map
+  ## --------------- ##
 
@@
     annotate("text", 
-             x =-130,
+             x =-135,
              y = -69,
-             label = "Participation (USD)", 
+             label = "Participation (%)", 
              color="black", 
@@
     annotate("text", 
-             x = c(-73,-34,5,45,83),
+             x = c(-76,-35,6,48,88),
              y = -69,
@@
     annotate("text", 
-             x =-130,
+             x =-135,
              y = -85,
@@
     annotate("text", 
-             x = c(-77,-37,3,42,82),
+             x = c(-79,-39,3,45,87),
              y = -85,
@@
     ) +
+    # geom_hline(yintercept = c(-50,0,50,80)) +
     my_ggtheme_m() +
     theme(
-      legend.box.margin = margin(t = -66, l = 250),
+      legend.box.margin = margin(t = -64, l = 229),
       legend.text = element_blank(),
-      legend.key.width = unit(5, "line"),
-      axis.text = element_text(colour = "blue")
-    ) +
-    # guides(fill = guide_legend(nrow = 1))+
-    ggsave(name_save,
-           width = 12,
-           height = 6,
-           units = "in"
+      legend.key.width = unit(5, "line")
+    )
+
+  ## --------------- ##
+  # Combine plots and save them
+  ## --------------- ##
+
+  p <- ggdraw() +
+    # ToE Latitudinal plot
+    draw_plot(Eez_hist, x = 0.873, y = 0.13, width = 0.124, height = 0.77) +
+    # Time of Emergence Map
+    draw_plot(map, x = 0, y = 0, width = 0.95, height = 1)
+
+  save_plot(
+    name_save,
+    p,
+    base_height = 6,
+    base_width = 12
   )
@@
   }
-
-```
-
-###### With relation to neighbours
-
-```{r time_emergence_map,  eval = T, echo = T}
-
-##_----------------------- #
-# Checking on China and Indonesia
-# Probably due to dispute territories
-
-Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
-  clean_names()
-
-# 
-# Neighbours %>% 
-#   filter(eez_name == "China") %>% 
-#   pull(eez_neighbour) %>% 
-#   unique()
-
-##_----------------------- #
-
-# Get the center poligon of each EEZ for map source/target
-coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
-  ungroup() %>% 
-  select(eez_name,geometry) %>% 
-  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
-  # Manually fix issues
-  mutate(
-    longitude = ifelse(
-      eez_name == "Fiji", 178.065033,
-      ifelse(eez_name == "Tuvalu",177.6493,
-             ifelse( eez_name == "New Zealand",174.8860,
-                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
-                     )
-             )
-      )
-    )
-  )
-
-# Remove "c()"
-coords$longitude <- gsub("\\(","",coords$longitude)
-coords$longitude <- gsub("c","",coords$longitude)
-coords$latitude <- gsub(")","",coords$latitude)
-
-coords <- coords %>% 
-  mutate_at(vars(longitude,latitude),as.numeric)
-
-# Prepare data from results
-Test_Data <- Eme_Res %>% 
-  filter(
-    n_ens == 10#,
-    # emerging_yr >= 2020
-  ) %>% 
-  group_by(eez_name,eez_neighbour,emerging_yr) %>% 
-  summarise(
-    n_tax = length(unique(taxon_key))
-  ) %>% 
-  filter(
-    emerging_yr <= 2050
-  ) %>%
-  group_by(eez_name,eez_neighbour) %>% 
-  summarise(
-    n_tax_emer = sum(n_tax)
-  ) %>% 
-  # get coords for sources
-  left_join(coords,
-            by = "eez_name") %>%
-  rename(source = eez_name,
-         start_lon = longitude,
-         start_lat = latitude,
-         eez_name=eez_neighbour) %>% 
-  # get coords for targets
-  left_join(coords,
-            by = "eez_name") %>% 
-  rename(target = eez_name,
-         end_lon = longitude,
-         end_lat = latitude) %>% 
-  ungroup() %>% 
-  # Plot miscs
-  # mutate(
-  #   per_mid = ifelse(is.na(per_mid),0,per_mid)
-  # ) %>% 
-  drop_na() #%>% 
-  # filter(
-    # per_mid >= 20,
-    # variable == "mean_spp",
-  # )
-
-# Convert to data to spatial
-network_data_sf <- Test_Data %>%
-  select(start_lon, start_lat, end_lon, end_lat) %>% 
-  purrr::transpose() %>% 
-  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
-  purrr::map(st_linestring) %>%
-  st_sfc(crs = 4326) %>%
-  st_sf(geometry = .)%>%
-  bind_cols(Test_Data) %>%
-  select(everything(), geometry) %>% 
-  # Set circles ('cus the world is not flat!)
-  st_segmentize(units::set_units(100, km)) %>% 
-  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
-  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
-  sf::st_sf(crs = 4326)
-
-#  unpack the coordinates of each feature 
-coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
-
-c(network_data_sf$line_id, network_data_sf$coords) %<-% (
-  coord_data %>% 
-    rename(subline_id = L1) %>%   
-    nest(-L2)
-)
-
-# Create data path for line connections
-paths <- network_data_sf %>% 
-  sf::st_set_geometry(NULL) %>% 
-  unnest() %>% 
-  mutate_at(vars(X,Y), round,2)
-
-# Plot it!
-
-country <- c("Fiji","Solomon Isl.","Tuvalu")
-
-ggplot() +
-  geom_sf(
-    # data = subset(world_map,name %in% country),
-    data = World_map,
-    size = 0.1, fill = "gray90", col = "gray90") +
-  geom_sf(
-    # data = subset(World_eez_sf, eez_name %in% country),
-    data = World_eez_sf,
-    size = 0.1, fill = "white", col = "gray70") +
-  geom_point(data = Test_Data,
-             # data = subset(Test_Data, source %in% country & target %in% source),
-             aes(start_lon,
-                 start_lat,
-                 col = n_tax_emer
-                 ),
-             size = 0.1,
-             col = "grey50",
-             # shape = 21,
-             stroke = 1.5) +
-  geom_path(data = paths,
-            # data = subset(paths, source %in% country),
-            aes(X, Y,
-                group = interaction(line_id,subline_id),
-                col = n_tax_emer
-            ),
-            size = 0.3,
-            # arrow = arrow(length = unit(0.009, "npc")),
-            show.legend = NA) +
-  ggtheme_map() +
-  scale_colour_viridis(
-    name="Number of emerging stocks\nby 2050",
-    limits = c(0,40),
-    breaks = seq(0,40,10),
-    # option="magma",
-    direction = 1
-  ) +
-  ggsave("./Figures/figure_emerging_2050.png",
-         width = 10,
-         height = 6,
-         units = "in"
-  )
 
@@
 ```
+
+###### With relation to neighbours
+
+```{r time_emergence_map,  eval = T, echo = T}
+
+##_----------------------- #
+# Checking on China and Indonesia
+# Probably due to dispute territories
+
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
+  clean_names()
+
+# 
+# Neighbours %>% 
+#   filter(eez_name == "China") %>% 
+#   pull(eez_neighbour) %>% 
+#   unique()
+
+##_----------------------- #
+
+# Get the center poligon of each EEZ for map source/target
+coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
+  ungroup() %>% 
+  select(eez_name,geometry) %>% 
+  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
+  # Manually fix issues
+  mutate(
+    longitude = ifelse(
+      eez_name == "Fiji", 178.065033,
+      ifelse(eez_name == "Tuvalu",177.6493,
+             ifelse( eez_name == "New Zealand",174.8860,
+                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
+                     )
+             )
+      )
+    )
+  )
+
+# Remove "c()"
+coords$longitude <- gsub("\\(","",coords$longitude)
+coords$longitude <- gsub("c","",coords$longitude)
+coords$latitude <- gsub(")","",coords$latitude)
+
+coords <- coords %>% 
+  mutate_at(vars(longitude,latitude),as.numeric)
+
+# Prepare data from results
+Test_Data <- Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # emerging_yr >= 2020
+  ) %>% 
+  group_by(eez_name,eez_neighbour,emerging_yr) %>% 
+  summarise(
+    n_tax = length(unique(taxon_key))
+  ) %>% 
+  filter(
+    emerging_yr <= 2050
+  ) %>%
+  group_by(eez_name,eez_neighbour) %>% 
+  summarise(
+    n_tax_emer = sum(n_tax)
+  ) %>% 
+  # get coords for sources
+  left_join(coords,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  left_join(coords,
+            by = "eez_name") %>% 
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  # mutate(
+  #   per_mid = ifelse(is.na(per_mid),0,per_mid)
+  # ) %>% 
+  drop_na() #%>% 
+  # filter(
+    # per_mid >= 20,
+    # variable == "mean_spp",
+  # )
+
+# Convert to data to spatial
+network_data_sf <- Test_Data %>%
+  select(start_lon, start_lat, end_lon, end_lat) %>% 
+  purrr::transpose() %>% 
+  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
+  purrr::map(st_linestring) %>%
+  st_sfc(crs = 4326) %>%
+  st_sf(geometry = .)%>%
+  bind_cols(Test_Data) %>%
+  select(everything(), geometry) %>% 
+  # Set circles ('cus the world is not flat!)
+  st_segmentize(units::set_units(100, km)) %>% 
+  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
+  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
+  sf::st_sf(crs = 4326)
+
+#  unpack the coordinates of each feature 
+coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
+
+c(network_data_sf$line_id, network_data_sf$coords) %<-% (
+  coord_data %>% 
+    rename(subline_id = L1) %>%   
+    nest(-L2)
+)
+
+# Create data path for line connections
+paths <- network_data_sf %>% 
+  sf::st_set_geometry(NULL) %>% 
+  unnest() %>% 
+  mutate_at(vars(X,Y), round,2)
+
+# Plot it!
+
+country <- c("Fiji","Solomon Isl.","Tuvalu")
+
+ggplot() +
+  geom_sf(
+    # data = subset(world_map,name %in% country),
+    data = World_map,
+    size = 0.1, fill = "gray90", col = "gray90") +
+  geom_sf(
+    # data = subset(World_eez_sf, eez_name %in% country),
+    data = World_eez_sf,
+    size = 0.1, fill = "white", col = "gray70") +
+  geom_point(data = Test_Data,
+             # data = subset(Test_Data, source %in% country & target %in% source),
+             aes(start_lon,
+                 start_lat,
+                 col = n_tax_emer
+                 ),
+             size = 0.1,
+             col = "grey50",
+             # shape = 21,
+             stroke = 1.5) +
+  geom_path(data = paths,
+            # data = subset(paths, source %in% country),
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = n_tax_emer
+            ),
+            size = 0.3,
+            # arrow = arrow(length = unit(0.009, "npc")),
+            show.legend = NA) +
+  ggtheme_map() +
+  scale_colour_viridis(
+    name="Number of emerging stocks\nby 2050",
+    limits = c(0,40),
+    breaks = seq(0,40,10),
+    # option="magma",
+    direction = 1
+  ) +
+  ggsave("./Figures/figure_emerging_2050.png",
+         width = 10,
+         height = 6,
+         units = "in"
+  )
+
+```

### 2020-08-27 Scripts/Manuscript.Rmd
Finalized Figure 3 for draft

@@
 
-```{r SSR_gain, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Average transboundary species SSR gain of neighbouring EEZs by 2030 (2021-2040) relative to 1951-2005 under RCP 8.5.
-Arrows go from the loosing EEZ (point) to the winner (arrowhead)"}
+```{r SSR_gain_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Average gain of transboundary stock share ratio (lines) and number of shifting stocks (polygon) by 2030 (2021-2040) relative to 1951-2005.
+Arrows go from the losing EEZ (point) to the winner.
+(arrowhead)"}
 
@@
 
-![Average transboundary species SSR gain of neighbouring EEZs by 2030 (2021-2040) relative to 1951-2005 under RCP 8.5.
-Arrows go from the loosing EEZ (point) to the winner (arrowhead)](../Figures/Fig2.png)
+![Average gain of transboundary stock share ratio (lines) and number of shifting stocks (polygon) by 2030 (2021-2040) relative to 1951-2005.
+Arrows go from the losing EEZ (point) to the winner.
+(arrowhead)](../Figures/Fig2.png)
 
@@
 
+```{r Toe_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}
+
+img <- readPNG("../Figures/Fig3.png")
+grid.raster(img)
+
+```
+
+![(Average time of emergence (ToE) per Exclusive Economic Zone (EEZ polygon) and participation of emerging stocks in total EEZ fishing revenue (land polygon).
+Darker tones represent low participation and late ToE; lighter colors represent large participation and earlier ToE.
+Density plot on right shows the number of emerging stocks across EEZ centroids (10^o running mean))](../Figures/Fig3.png)
+
 # Discussion

### 2020-08-28 Scripts/Initial_Analysis.Rmd
Included buble plot

@@
 
-##### Fig.Percentage Change Map
+##### Fig. Percentage Change Map
 
@@
 
-##### Fig. 2 and S1. (Netowrk)
+##### Fig. 1 and S1. (Netowrk)
 
@@
 
-##### Fig. 3 Mapping the time of emergence
+##### Fig. 2 Mapping the time of emergence
 
@@
   gather("metric","value",mean_yr:n_taxa) %>% 
-  filter(metric != "n_taxa") %>% 
+  # filter(metric != "n_taxa") %>% 
   mutate(
@@
 
+##### Fig. 3. Bubble plot
+
+```{r, buble_plot, eval = F, echo = F}
+
+Buble_data <- Eez_d_hist %>% 
+  filter(metric == "mean_yr") %>% 
+  group_by(sau_fishing_entity,tresh) %>% 
+  summarise(mean_yr = mean(value)) %>% 
+  left_join(Eme_value_data,
+            by = c("sau_fishing_entity","tresh")) %>% 
+  filter(tresh == tresh_list[i],
+         variable == "value") %>% 
+  mutate(sum_value_mi = sum_value/1000000) %>%   # Million usd
+  left_join(Matching_names,
+            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
+  distinct(sau_fishing_entity,.keep_all = TRUE) %>% 
+  filter(!is.na(un_geo_region)) %>% 
+  ungroup() %>% 
+  mutate(
+    sau_fishing_entity = ifelse(sum_value_mi >= 1000 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity)
+  )
+
+# Plot the bubbles!
+
+ggplot(Buble_data) +
+  geom_point(
+    aes(
+      x = mean_yr,
+      y = sum_value_mi,
+      size = n_taxa,
+      colour = un_geo_region
+    ),
+    alpha = 0.7
+  ) +
+  # scale_size(range = c(.1, 10), name="Fishing revenue (USD)")  +
+  geom_label_repel(data = subset(Buble_data, sum_value_mi >= 1000),
+                   aes(
+                     x = mean_yr,
+                     y = sum_value_mi,
+                     label = sau_fishing_entity
+                   ),
+                   arrow = arrow(length = unit(0.009, "npc"), 
+                                 type = "closed",
+                                 ends = "first"),
+                   force = 2,
+                   box.padding = 2,
+                   fill = "NA"
+
+  ) +
+  # scale_fill_viridis(discrete = F)
+  scale_color_viridis(name="Georaphic Region",
+                      discrete = T) +
+  scale_x_continuous("Average time of emergence",
+                     breaks = seq(2000,2080,20)
+  ) +
+  scale_y_continuous("Fishing revenue (million)",
+                     breaks = seq(1000,5000,1000),
+                     labels = scales::dollar
+  ) +
+  scale_size("Number of\nemerging taxa") +
+  ggsave("./Figures/bubble_plot.png",
+         width = 8,
+         height = 8,
+         units = "in"
+  )
+
+```
+
 ### Species level analysis

### 2020-08-28 Scripts/Initial_Analysis.Rmd
log transformed bubble plot

--- 2020-08-27 Scripts/Initial_Analysis.Rmd
@@
 
+##### Fig. 3. Bubble plot
+
+```{r, buble_plot, eval = F, echo = F}
+
+Buble_data <- Eez_d_hist %>% 
+  filter(metric == "mean_yr") %>% 
+  group_by(sau_fishing_entity,tresh) %>% 
+  summarise(mean_yr = mean(value)) %>% 
+  left_join(Eme_value_data,
+            by = c("sau_fishing_entity","tresh")) %>% 
+  filter(tresh == tresh_list[i],
+         variable == "value") %>% 
+  mutate(
+    weight_toe = sum_value/mean_yr,
+    sum_value_mi = sum_value/1000000000,
+         logvalue = log10(sum_value)
+         ) %>%   # Billion usd
+  left_join(Matching_names,
+            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
+  distinct(sau_fishing_entity,.keep_all = TRUE) %>% 
+  filter(!is.na(un_geo_region)) %>% 
+  ungroup() %>% 
+  mutate(
+    sau_fishing_entity = ifelse(sum_value_mi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity)
+  )
+
+# Labels Data
+
+Labels_data <- tibble(
+  x = c(2020,2040,2070),
+  y = c(0.3,0.3,9.1),
+  label = c("Early emergence", "Late emergence","Fisheries over the 75 percentile")
+)
+
+# Plot the bubbles!
+
+ggplot(Buble_data) +
+  geom_point(
+    aes(
+      x = mean_yr,
+      # y = sum_value_mi,
+      y = logvalue,
+      size = n_taxa,
+      colour = un_geo_region
+    ),
+    alpha = 0.7
+  ) +
+  geom_vline(xintercept = 2030, color = "grey90", size = 0.5) +
+  geom_hline(yintercept = 9, color = "grey90", size = 0.5)  +
+  geom_label_repel(data = subset(Buble_data, sum_value_mi >= 1),
+                   aes(
+                     x = mean_yr,
+                     # y = sum_value_mi,
+                     y = logvalue,
+                     label = sau_fishing_entity
+                   ),
+                   arrow = arrow(length = unit(0.009, "npc"),
+                                 type = "closed",
+                                 ends = "first"),
+                   force = 3,
+                   box.padding = 2,
+                   fill = "NA"
+
+  ) +
+  geom_text(data = Labels_data,
+    aes(
+      x = x,
+      y = y,
+      label = label
+    ),
+    size = 3,
+    color = "grey30"
+  ) +
+  # scale_fill_viridis(discrete = F)
+  scale_color_viridis(name="Region",
+                      discrete = T) +
+  scale_x_continuous("Average time of emergence",
+                     limits = c(2000,2080),
+                     breaks = seq(2000,2080,10)
+  ) +
+  scale_y_continuous("Fishing revenue (log10)",
+                     limits = c(3,12),
+                     breaks = seq(3,12,3),
+                     expand = c(0,0)
+  ) +
+  scale_size("Number of\nemerging taxa") +
+  ggsave("./Figures/bubble_plot.png",
+         width = 8,
+         height = 8,
+         units = "in"
+  )
+
+```
+
 ### Species level analysis

### 2020-08-31 Scripts/Initial_Analysis.Rmd
Created code for ToE tempr

--- 2020-08-27 Scripts/Initial_Analysis.Rmd
@@
   "viridis",
-  "cowplot",
-  "rfishbase",
-  "stargazer" # For nice lm tables
+  "cowplot"
 )
@@
 
-EEZ_CellID <- my_path("G", extra_path = "Spatial/DBEM", name="EEZ_CellID.xlsx", read = TRUE) %>%
+EEZ_CellID <- my_path("G", extra_path = "Spatial/V_Lam", name="EEZ_CellID.xlsx", read = TRUE) %>%
   clean_names()
@@
 # SAU name relatioal table (Fish For Visa)
-sau_clean_names <- my_path("G", extra_path = "Spatial/SAU/", name="matching_names.csv", read = TRUE)
+sau_clean_names <- my_path("G", extra_path = "Spatial/SAU/", name="SAU_matching_names.csv", read = TRUE)
 
@@
 
-getEnvVar=function(yr,ensemble,Variable){
-
-  filepath <- paste("Z:/DATA/Environmental data/GFDL_CMIP5_Large_Ensemble/ens",ensemble,"",sep="/")
-
-  if(Variable == "sst"){
-    sst = paste("SST_",yr,".txt",sep="")
-    read_me_sst <- paste(filepath,sst,sep="")[1]
-    data <- read.csv(read_me_sst,header=F)
-
+getEnvVar=function(yr,ensemble,variable){
+
+  filepath <- paste("/Volumes/DATA/DATA/Environmental data/GFDL_CMIP5_Large_Ensemble/All var/ens",ensemble,"/",sep="")
+
+    file = paste(variable,"_",yr,".txt",sep="")
+    read_me <- paste(filepath,file,sep="")
+
+    data <- bind_cols(lapply(read_me,FUN= data.table::fread, na.strings = "NA",header=F)
+      ) %>% 
+      mutate(index = seq(1,259200,1))
+
+    colnames(data) <- c(as.character(ensemble),"index")
+
+    data <- data %>% 
+      gather("ensemble","variable",1:length(ensemble)) %>%  # fix it has to be untill 10
+      mutate(variable = ifelse(variable == -9999.00000,NA,variable)) # Fixing NAs connotation
+
+  # Retrun
+
+  getTemp <- data %>% 
+    mutate(ensemble = ensemble,
+           year = yr)
+
+  return(getTemp)
 }
 
-  if(Variable == "botom"){
-    botom = paste("botom_",yr,".txt",sep="")
-    read_me_botom <- paste(filepath,botom,sep="")[1]
-    data <- read.csv(read_me_botom,header=F)
-  }
-
-  # Retrun
-
-  getTemp <- data %>% 
-    mutate(ensemble = ensemble)
-}
-
 ```
@@
 
-TempEmer <- function(Hist_y = 2005){ # The LAST year of Historical data
-
-  # Historical mean and sd of distance
-  historical <- Centroids_id %>%
-    filter(year <= Hist_y) %>% 
-    ungroup() %>%
-    group_by(taxon_key,ensemble,eez_name,neighbour_id) %>%
-    summarise_at(vars("distance"),
-                 c(hd_temp_mean=mean,
-                   hd_temp_sd=sd,
-                   n_yrs =~n())
-                 )
-
-    # Future results
-  future <- Centroids_id %>% 
-    filter(year > Hist_y)
-
-   # Get unique eezs
-  Unique_EEZs <- unique(EEZ_ids$eez_name)
-
-  # Transboundary Index
-  Trans_index <- Centroids_id %>% 
-    left_join(historical,
-              by = c("taxon_key","eez_name","ensemble","neighbour_id")
-    ) %>%
-    # Re arrenging table for mutate
-    gather("variable","value",distance:hd_temp_sd) %>%
-    ungroup() %>% 
-    mutate(value = ifelse(is.na(value),0,value)) %>% 
-    mutate(
-      names = ifelse(eez_name ==Unique_EEZs[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_")
-    )
-    ) %>%
-    select(-eez_name,-variable) %>%
-    spread(names,value) %>% 
-    # mutate index
-    mutate(
-      trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
-    # )
-    ) %>% 
-    # Ten years average of results
-    group_by(taxon_key,ensemble,neighbour_id) %>% 
-    mutate(RMean = rollmean(x = trans_index, 
+TempEmer <- function(yrs, hist_y = 2005, variable, ensemble){ # The LAST year of Historical data
+
+  # Load environmental data
+  sst_data <- bind_rows(
+    lapply(
+      yrs,
+      getEnvVar, 
+      ensemble = ensemble, 
+      variable = variable
+    )
+  ) %>% 
+    left_join(index_code,
+              by = "index") %>%  # join with EEZ indexes
+    select(-lat,-lon,-eez_name) %>% 
+    filter(!is.na(eezid)) %>% 
+    # Get average temperature per eez
+    group_by(ensemble,eezid,year) %>%
+    summarise_at(vars("variable"),
+                 c(hd_eez_mean=mean),
+                 na.rm=T
+    ) %>% 
+    # estimate 10 years running mean
+    group_by(ensemble,eezid) %>% 
+    mutate(Rmean = rollmean(x = hd_eez_mean, 
                             10, 
                             align = "right", 
-                            fill = trans_index,
-                            na.rm=T)
-    ) %>% 
-    # filter(year > 1960) %>%
-    select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)
+                            fill = hd_eez_mean)
+    )
 
   # Average of 10 years mean of ensemble members results
-  mean_sd_index <- Trans_index %>% 
-    group_by(taxon_key,year,neighbour_id) %>% 
-    summarise_at(vars("RMean"),
+  mean_sd_index <- sst_data %>% 
+    group_by(eezid,year) %>% 
+    summarise_at(vars("Rmean"),
                  c(trend=mean,
@@
 
-    # Valid ensembles
-  # Only perform analysis on ensembles containing all years between 1951 and 2040
-    n_years <- mean_sd_index %>% 
-      group_by(n_ens) %>% 
-      summarise(
-        n_year = n()
-      ) %>% 
-      filter(
-        n_year == 149
-      )
-
-    mean_sd_index <- mean_sd_index %>% 
-      filter(n_ens %in% n_years$n_ens)
-
-  # Estimate noice from ensemble members
-  historic_sd <- Trans_index %>% 
-    filter(year <= Hist_y) %>% 
-    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
-    summarise_at(vars("trans_index"),
-                 c(trend= mean),
-                 na.rm=T
-                 ) %>% 
-    group_by(taxon_key,neighbour_id) %>% #Capture ensemble variation
-    summarise_at(vars("trend"),
+  # Historic data
+  historical <- sst_data %>%
+    filter(year <= hist_y) %>%
+    ungroup() %>%
+    # get EEZ time average
+    group_by(ensemble,eezid) %>%
+    summarise_at(vars("Rmean"),
+                 c(hd_temp_mean=mean),
+                 na.rm= T
+    ) %>% 
+    # get ensemble member mean and noice
+    group_by(eezid) %>% 
+    summarise_at(vars("hd_temp_mean"),
                  c(hist_mean=mean,
-                   noise=sd,
-                   n_ens = ~n())
-                 )
-
-  # For testing
-  if(Plot == T & length(historic_sd$hist_mean) > 1){
-  # Plot the result, bruh
-
-  # Step to determine if data exists
-  Plot_Test <- nrow(subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)))>0
-
-  ggplot() +
-    geom_line(data = mean_sd_index,
-              aes(
-                x = year,
-                y = trend
-              )
-    ) +
-    geom_ribbon(data = mean_sd_index,
-                aes(x=year,
-                    ymin=historic_sd$hist_mean-historic_sd$noise,
-                    ymax=historic_sd$hist_mean+historic_sd$noise),
-                alpha = 0.5,
-                fill = "grey50"
-    ) +
-    geom_ribbon(data = mean_sd_index,
-                aes(x=year,
-                    ymin=historic_sd$hist_mean-(historic_sd$noise*2),
-                    ymax=historic_sd$hist_mean+(historic_sd$noise)*2)
-                ,
-                alpha = 0.5,
-                fill = "grey30"
-    ) +
-    geom_line(data = mean_sd_index,
-              aes(
-                x = year,
-                y = historic_sd$hist_mean
-              ),
-              colour = "grey50"
-    ) +
-    # Conditional, will only plot circles if the data exists
-    {if(Plot_Test)geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
-               aes(
-                 x = year,
-                 y = trend,
-                 colour = "red",
-               ),
-               shape = 1,
-               size = 3
-    )
-      } +
-    geom_vline(xintercept = Hist_y,colour="blue",alpha = 0.5) +
-    scale_x_continuous(breaks = seq(1950,2100,5)) +
-    theme_classic()
-
-    tk <- unique(mean_sd_index$taxon_key)
-
-    ggsave(paste(tk,Id,"plot_example.jpg",sep="_"),
-       plot = last_plot(),
-       width = 10,
-       height = 8,
-       units = "in",
-       path = my_path("R")
-       )
-
-  }
-
-  # If no historic data then emergin year is the first year of data
-  if(nrow(historic_sd) == 0){
-
-    Final_Result <- mean_sd_index %>%
-      group_by(taxon_key,neighbour_id,n_ens) %>%
-      summarise(emerging_yr= min(year)) %>%
-      left_join(EEZ_ids,
-                by = "neighbour_id") %>%
-      ungroup() %>%
-      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr) %>% 
-      mutate(tresh= NA)
-
-  return(Final_Result)
-
-  }else{
-
-    Final_Result <- mean_sd_index %>%
-      filter(year > Hist_y) %>%
+                   noise = sd),
+                 na.rm = T
+    )
+
+  # ----------------------- #
+  # Plot for data viz
+  # ----------------------- #
+
+  # samp <- sample(mean_sd_index$eezid,30,replace = F)
+
+  # mean_sd_index <- mean_sd_index %>% 
+    # filter(eezid %in% samp)
+
+    # plot_data <- mean_sd_index %>% 
+  # left_join(historical,
+  #           by = "eezid")
+
+  # ggplot() +
+  #   geom_line(data = plot_data,
+  #             aes(
+  #               x = year,
+  #               y = trend
+  #             )
+  #   ) +
+  #   geom_ribbon(data = plot_data,
+  #               aes(x=year,
+  #                   ymin=hist_mean-noise,
+  #                   ymax=hist_mean+noise
+  #                   ),
+  #               alpha = 0.5
+  #   ) +
+  #   geom_line(data = plot_data,
+  #             aes(
+  #               x = year,
+  #               y = hist_mean
+  #               ),
+  #             linetype = "dashed"
+  #   ) +
+  #   facet_wrap(~eezid,
+  #              scales = "free") +
+  #   # Conditional, will only plot circles if the data exists
+  #   geom_point(data = subset(plot_data, trend > (hist_mean+noise) | trend < (hist_mean-noise)),
+  #              aes(
+  #                x = year,
+  #                y = trend,
+  #                colour = "red",
+  #              ),
+  #              shape = 1,
+  #              size = 3
+  #   ) +
+  #   geom_vline(xintercept = hist_y,colour="black",alpha = 0.5) +
+  #   scale_x_continuous(breaks = seq(1950,2100,5)) +
+  #   theme_classic()
+
+  # Future results
+  future <- sst_data %>% 
+    filter(year > hist_y) %>% 
+    # get ensemble member mean and noice
+    group_by(eezid,year) %>% 
+    summarise_at(vars("Rmean"),
+                 c(trend=mean,
+                   f_noise = sd)
+    ) %>% 
+    group_by(eezid)
+
+  Final_Result <- future %>%
+    left_join(historical,
+              by = "eezid") %>% 
     mutate(
-        tresh_one = ifelse(trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise),"One",NA),
-        tresh_two = ifelse(trend > (historic_sd$hist_mean+(2*historic_sd$noise)) | trend < (historic_sd$hist_mean-(2*historic_sd$noise)),"Two",NA)
+      tresh_one = ifelse(trend > (hist_mean+noise) | trend < (hist_mean-noise),"One",NA),
+      tresh_two = ifelse(trend > (hist_mean+(2*noise)) | trend < (hist_mean-(2*noise)),"Two",NA)
     ) %>%
@@
     filter(!is.na(YesNo)) %>% 
-      group_by(taxon_key,neighbour_id,n_ens,tresh) %>%
-      summarise(emerging_yr= min(year)) %>%
-      left_join(EEZ_ids,
-                by = "neighbour_id") %>%
-      ungroup() %>%
-      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr,tresh)
+    group_by(eezid,tresh) %>% 
+    summarise(emerging_yr = min(year))
 
@@
 }
-}
+
+gc()
+
+sst <- TempEmer(
+  yrs = seq(1951,2070,1),
+  hist_y = 2005,
+  variable = "SST",
+  ensemble = seq(102,111,1)
+)
+
+write_csv(sst,
+          "ToE_sst.csv")
+
+botTemp <- TempEmer(
+  yrs = seq(1951,2070,1),
+  hist_y = 2005,
+  variable = "botTemp",
+  ensemble = seq(102,111,1)
+)
+
+write_csv(botTemp,
+          "ToE_bottemp.csv")
+
+test <- sst %>% 
+  filter(tresh == "tresh_one")
+
+hist(test$emerging_yr)
+
+gc()
+
+test_map <- World_eez_sf %>% 
+  left_join(test,
+            by = "eezid") %>% 
+  ggplot() +
+  geom_sf(
+    aes(
+      fill = emerging_yr
+    )
+  )+ scale_fill_viridis(
+    limits = c(2000,2070),
+    breaks = seq(2000,2070,10)#,
+    # direction = -1
+  ) +
+  my_ggtheme_m() +
+  ggsave("botTemp_ToE.png",
+         width = 12,
+         height = 10,
+         units = "in"
+  )
 
@@
 # Get taxon list
-Taxon_list_emer <- list.files(my_path("R",extra_path = "Emergence_2005"))
+Taxon_List_emer <- list.files(my_path("R",extra_path = "Emergence_2005"))
 
 # clean names for function
-Taxon_list_emer <- gsub("_.*","",Taxon_list_emer)
-
-# SAU exploited species list
-Exploited_spp_list <- my_path("G","SAU","exploited_species_list.csv",read = T) %>%
-  clean_names() %>% 
-  pull(taxon_name)
-
-# Get habitat preference from fishbase
-Spp_env_pref <- rbind(species(Exploited_spp_list, #species from rfishbase package
-                              fields = species_fields$habitat[4])
-                      ) %>% 
-  mutate(taxon_name = Exploited_spp_list) %>% 
-  left_join(my_path("G","SAU","exploited_species_list.csv", read = T) %>%
-              clean_names(),
-            by = "taxon_name")
+Taxon_List_emer <- gsub("_.*","",Taxon_List_emer)
 
@@
 
+Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_List_emer,"_emergence.txt",sep="")
+
 # Clean names to match data
-Matching_names <- my_path("G","Spatial/SAU", "matching_names.csv", read = T) %>% 
-  rename(eezid = eez_id)
-
-Fishing_names <- Matching_names %>% 
-  group_by(sau_fishing_entity,fishing_entity_id) %>% 
-  summarise(n=n()) %>% 
-  select(-n)
-
-# eez_centroids
-Eez_centroids <- my_path("D", extra_path = "Spatial/", name="eez_centroids.csv", read = TRUE) %>% 
-  rename(sf_eez_name = eez_name) %>% 
-  left_join(Matching_names) %>% 
-  select(eezid,latitude,longitude)
-
-# Species depth clasification (SAU-fishbase)
+Matching_names <- my_path("G","Spatial/SAU", "matching_names.csv", read = T)
+
+# Species depth clasification
+
 Spp_depth <-fread(paste(my_path("D"),"SAU/SppTaxonName.csv", sep ="")) %>% 
   clean_names()
-
-# Temperature Time of Emergence
-Toe_temp <- left_join(my_path("D","Climate", "ToE_sst.csv", read = T),
-                      my_path("D","Climate", "ToE_bottemp.csv", read = T)
-                      ) %>% 
-  clean_names()
-
-# Biogeographic regions (from Rey 2020)
-Biogeo_reg <-fread(paste(my_path("D"),"Climate/SPATIAL_PARTITION_OCEAN.csv", sep ="")) %>% 
-  clean_names() %>% 
-  rowid_to_column("index")
-
-# Get eez to index relational data
-EEZ_CellID <- my_path("G", extra_path = "Spatial/DBEM", name="EEZ_CellID.xlsx", read = TRUE) %>%
-  clean_names()
-colnames(EEZ_CellID) <- c("eezid","index")
-
-Eez_biogeo <- EEZ_CellID %>% 
-  # filter(eezid == 191) %>% 
-  left_join(Biogeo_reg,
-       by = "index") %>% 
-  # Deal with NaNs in data
-  mutate(biome = ifelse(biome == "NaN",0,biome)) %>% 
-  filter(biome > 0) %>% 
-  # Get frequency of biome per index
-  group_by(eezid,biome) %>% 
-  summarise(nbiom=n()) %>% 
-  # Select the more frequent biome per eez
-  group_by(eezid) %>% 
-  top_n(1,nbiom) %>% 
-  ungroup() %>% 
-  select(-nbiom) %>% 
-  distinct(eezid, .keep_all = T) # Croatia is 1/2. We arbritary allocated the first one
-
-# Species catch trend (from FishForVisa)
-
-Catch_trend <- my_path("D","Species","Catch_trend.csv", read = T) %>% 
-  clean_names() %>% 
-  filter(taxon_key %in% Taxon_list_emer) %>% 
-  rename(sau_fishing_entity = territory) %>% 
-  left_join(Fishing_names,
-            by = "sau_fishing_entity") %>% 
-  filter(status != "No Trend") %>% 
-  mutate(category = ifelse(status %in% c("Developing","Rebuilding"),"Category A",
-                         ifelse(status == "Max Exploited","Category B",
-                                ifelse(status %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
-                         )
-         )
-  )
 
@@
 # --------------- #
-
-Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_list_emer,"_emergence.txt",sep="")
 
@@
   # get coords for sources
-  left_join(Eez_centroid,
+  left_join(Eez_centroid_eck,
             by = "eez_name") %>%
@@
   # get coords for targets
-  left_join(Eez_centroid,
+  left_join(Eez_centroid_eck,
             by = "eez_name") %>% 
@@
   purrr::map(st_linestring) %>%
-  st_sfc(crs = 4326) %>%
-  # st_sfc(crs = "+proj=eck4") %>% # for tbular plot
+  # st_sfc(crs = 4326) %>%
+  st_sfc(crs = "+proj=eck4") %>% # for tbular plot
   st_sf(geometry = .)%>%
@@
   mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>%
+  # mutate(geometry = (geometry + c(18000000,9000000)) %% c(360) - c(18000000,9000000)) %>%
   st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>%
-  sf::st_sf(crs = 4326)
-  # sf::st_sf(crs = "+proj=eck4") # for tbular plot
+  st_wrap_dateline(options = c("WRAPDATELINE=YES"), quiet = TRUE) %>%
+  # sf::st_sf(crs = 4326)
+  sf::st_sf(crs = "+proj=eck4") # for tbular plot
 
@@
              size = 0.05,
-           col = "gray80",
+             col = "gray30",
              shape = 20,
@@
   geom_path(data = subset(paths, time_step == time & per_plot > 0),
-          aes(X,
-              Y,
+            aes(X, Y,
                 group = interaction(line_id,subline_id),
@@
             size = 0.3,
-          lineend = "round",
             arrow = arrow(length = unit(0.008, "npc"),
@@
     limits = c(0,max(Map_data$per_plot)),
-  breaks = seq(0,max(Map_data$per_plot),10)#,
+    breaks = seq(0,max(Map_data$per_plot),10),
     # option="magma",
@@
     limits = c(0,max(Map_data$per_plot)),
-  breaks = seq(0,max(Map_data$per_plot),10)#,
-  # alpha = 0.8
+    breaks = seq(0,max(Map_data$per_plot),10),
+    alpha = 0.8
     # option="magma",
@@
 Eez_d_hist <- Eme_Res %>% 
-  group_by(eezid ,sau_fishing_entity,tresh) %>% 
+  group_by(eezid= eez_id ,sau_fishing_entity,tresh) %>% 
   # summarise(n_tax = length(unique(taxon_key))) %>% 
@@
 
+### Random plot
+
+ggplot(subset(Eez_d_hist, tresh == "Tresh_one")) +
+  geom_density(
+    aes(
+      x = value,
+      fill = tresh
+    ),
+    bins = 10,
+    alpha = 0.8
+  ) +
+  facet_wrap(~metric,
+             scales = "free") +
+  scale_fill_viridis(discrete = T) +
+  my_ggtheme_p() +
+  ggsave("./Figures/hist_mean.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
 World_map_nemer <- World_map %>%
@@
             by = c("sau_fishing_entity","fishing_entity_id")
-  ) %>% 
-  filter(!is.na(participation))
+  )
 
 grey_land <- World_map %>% 
-  filter(!fishing_entity_id %in% World_map_nemer$fishing_entity_id) %>% 
-  filter(sovereignt != "Antarctica")
+  filter(!fishing_entity_id %in% World_map_nemer$fishing_entity_id)
 
@@
   tresh_list <- c("Tresh_one","Tresh_two")
-  name_save <- paste("./Figures/emer_map_lat_",tresh_list[i],".png",sep = "")
-
-  ## --------------- ##
-  # Latitudinal plot
-  ## --------------- ##
-
-  ### Random plot
-
-Eez_hist <- Eme_Res %>% 
-  group_by(eezid,eez_name = sf_eez_name) %>% 
-  left_join(Eez_centroid) %>% 
-  filter(tresh == tresh_list[i]) %>% 
-  group_by(latitude) %>% 
-  summarise(
-    number = n()
-  ) %>% 
-  group_by() %>% 
-  mutate(RMean = rollmean(x = number, 
-                          10, 
-                          align = "center", 
-                          fill = number
-  )
-  ) %>% 
-  mutate(lab="lab") %>% 
-  ggplot() +
-  geom_area(
-    aes(
-      x = latitude,
-      y = RMean,
-      fill = lab,
-      color = lab
-    )
-  ) +
-  coord_flip() +
-  scale_x_continuous("",
-    limits = c(-50,80),
-    breaks = c(-50,0,50,80)
-  ) +
-  scale_y_continuous(
-    "Emerging stocks",
-    limits = c(0, 124),
-    breaks = c(0,62,124)
-
-  ) +
-  scale_fill_viridis(option="magma", discrete = T, begin = 0.7) +
-  scale_color_viridis(option="magma", discrete = T, begin = 0.7) +
-  theme_minimal() + 
-  theme(
-    panel.background = element_rect(fill = "white", colour = "white"),
-    panel.grid.major.y = element_blank(),
-    panel.grid.minor.y = element_blank(),
-    panel.grid.major.x = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.text = element_text(size = 12, colour="black"),
-    axis.title.x = element_text(size = 12, colour="black", face = "bold"),
-    legend.position = ""
-  );Eez_hist
-
-  ## --------------- ##
-  # Map
-  ## --------------- ##
+  name_save <- paste("./Figures/emer_map_rev_",tresh_list[i],".png",sep = "")
 
@@
     geom_sf(data = grey_eez, aes(),fill = "grey50", color =  "white",size = 0.1) +
-    # Set Participation labels
     annotate("text", 
-             x =-135,
-             y = -69,
-             label = "Participation (%)", 
-             color="black", 
-             size=4, 
-             angle=0, 
-             fontface="bold",
-             hjust=0) +
-    annotate("text", 
-             x = c(-76,-35,6,48,88),
-             y = -69,
+             x =(c(-85.5,-30,25,81,136)+5),
+             y = -93.5,
              # x =(c(-8500000,-3000000,2500000,7500000,12800000)),
@@
              color="black", 
-             size=4, 
+             size=3, 
              angle=0, 
@@
              hjust=0) +
-    # Set ToE labels
-    annotate("text", 
-             x =-135,
-             y = -85,
-             label = "Time of Emergence",
-             color="black", 
-             size=4, 
-             angle=0, 
-             fontface="bold",
-             hjust=0) +
-    annotate("text", 
-             x = c(-79,-39,3,45,87),
-             y = -85,
-             label = seq(2100,2000,-25), 
-             color="black", 
-             size=4, 
-             angle=0, 
-             # fontface="bold",
-             hjust=0) +
-    scale_fill_viridis("",
+    scale_fill_viridis(
       limits = c(2000,2100),
       breaks = seq(2000,2100,25),
-      labels = c("","","","","")#,
-      # direction = -1
-    ) +
-    # geom_hline(yintercept = c(-50,0,50,80)) +
+      labels = seq(2100,2000,-25),
+      direction = -1
+    ) +
     my_ggtheme_m() +
-    theme(
-      legend.box.margin = margin(t = -64, l = 229),
-      legend.text = element_blank(),
-      legend.key.width = unit(5, "line")
-    )
-
-  ## --------------- ##
-  # Combine plots and save them
-  ## --------------- ##
-
-  p <- ggdraw() +
-    # ToE Latitudinal plot
-    draw_plot(Eez_hist, x = 0.873, y = 0.13, width = 0.124, height = 0.77) +
-    # Time of Emergence Map
-    draw_plot(map, x = 0, y = 0, width = 0.95, height = 1)
-
-  save_plot(
-    name_save,
-    p,
-    base_height = 6,
-    base_width = 12
+    ggsave(name_save,
+           width = 12,
+           height = 10,
+           units = "in"
     )
@@
 }
+
+```
+
+###### With relation to neighbours
+
+```{r time_emergence_map,  eval = T, echo = T}
+
+##_----------------------- #
+# Checking on China and Indonesia
+# Probably due to dispute territories
+
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
+  clean_names()
+
+# 
+# Neighbours %>% 
+#   filter(eez_name == "China") %>% 
+#   pull(eez_neighbour) %>% 
+#   unique()
+
+##_----------------------- #
+
+# Get the center poligon of each EEZ for map source/target
+coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
+  ungroup() %>% 
+  select(eez_name,geometry) %>% 
+  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
+  # Manually fix issues
+  mutate(
+    longitude = ifelse(
+      eez_name == "Fiji", 178.065033,
+      ifelse(eez_name == "Tuvalu",177.6493,
+             ifelse( eez_name == "New Zealand",174.8860,
+                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
+                     )
+             )
+      )
+    )
+  )
+
+# Remove "c()"
+coords$longitude <- gsub("\\(","",coords$longitude)
+coords$longitude <- gsub("c","",coords$longitude)
+coords$latitude <- gsub(")","",coords$latitude)
+
+coords <- coords %>% 
+  mutate_at(vars(longitude,latitude),as.numeric)
+
+# Prepare data from results
+Test_Data <- Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # emerging_yr >= 2020
+  ) %>% 
+  group_by(eez_name,eez_neighbour,emerging_yr) %>% 
+  summarise(
+    n_tax = length(unique(taxon_key))
+  ) %>% 
+  filter(
+    emerging_yr <= 2050
+  ) %>%
+  group_by(eez_name,eez_neighbour) %>% 
+  summarise(
+    n_tax_emer = sum(n_tax)
+  ) %>% 
+  # get coords for sources
+  left_join(coords,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  left_join(coords,
+            by = "eez_name") %>% 
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  # mutate(
+  #   per_mid = ifelse(is.na(per_mid),0,per_mid)
+  # ) %>% 
+  drop_na() #%>% 
+  # filter(
+    # per_mid >= 20,
+    # variable == "mean_spp",
+  # )
+
+# Convert to data to spatial
+network_data_sf <- Test_Data %>%
+  select(start_lon, start_lat, end_lon, end_lat) %>% 
+  purrr::transpose() %>% 
+  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
+  purrr::map(st_linestring) %>%
+  st_sfc(crs = 4326) %>%
+  st_sf(geometry = .)%>%
+  bind_cols(Test_Data) %>%
+  select(everything(), geometry) %>% 
+  # Set circles ('cus the world is not flat!)
+  st_segmentize(units::set_units(100, km)) %>% 
+  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
+  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
+  sf::st_sf(crs = 4326)
+
+#  unpack the coordinates of each feature 
+coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
+
+c(network_data_sf$line_id, network_data_sf$coords) %<-% (
+  coord_data %>% 
+    rename(subline_id = L1) %>%   
+    nest(-L2)
+)
+
+# Create data path for line connections
+paths <- network_data_sf %>% 
+  sf::st_set_geometry(NULL) %>% 
+  unnest() %>% 
+  mutate_at(vars(X,Y), round,2)
+
+# Plot it!
+
+country <- c("Fiji","Solomon Isl.","Tuvalu")
+
+ggplot() +
+  geom_sf(
+    # data = subset(world_map,name %in% country),
+    data = World_map,
+    size = 0.1, fill = "gray90", col = "gray90") +
+  geom_sf(
+    # data = subset(World_eez_sf, eez_name %in% country),
+    data = World_eez_sf,
+    size = 0.1, fill = "white", col = "gray70") +
+  geom_point(data = Test_Data,
+             # data = subset(Test_Data, source %in% country & target %in% source),
+             aes(start_lon,
+                 start_lat,
+                 col = n_tax_emer
+                 ),
+             size = 0.1,
+             col = "grey50",
+             # shape = 21,
+             stroke = 1.5) +
+  geom_path(data = paths,
+            # data = subset(paths, source %in% country),
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = n_tax_emer
+            ),
+            size = 0.3,
+            # arrow = arrow(length = unit(0.009, "npc")),
+            show.legend = NA) +
+  ggtheme_map() +
+  scale_colour_viridis(
+    name="Number of emerging stocks\nby 2050",
+    limits = c(0,40),
+    breaks = seq(0,40,10),
+    # option="magma",
+    direction = 1
+  ) +
+  ggsave("./Figures/figure_emerging_2050.png",
+         width = 10,
+         height = 6,
+         units = "in"
+  )
 
@@
   facet_wrap(~un_sub_region)
-
-```
-
-### Tab.1 Explain the ToE
-
-```{r multiple_reg, eval = F, echo = F}
-
-# Possible explanatory variables
-
-## Habitat association (DemersPelag)
-## Dem Vs Pelagic (Dem_pel)
-## SST time of Emer (Toe_sst)
-## Stock Status [FishForVisa] (Catch_trend)
-## Biogeographic location (Biogeo_reg)
-
-# AIC value; model between predictive capacity and complexity (i.e., number of explanatory variables) is given by the model with lowest AIC value
-
-# -------------------------------- #
-#### Prepare data
-# -------------------------------- #
-
-Lm_data <- Eme_Res %>% # 19728
-  filter(tresh == "Tresh_one") %>% # 11490
-  left_join(Spp_depth) %>% 
-  select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel) %>% 
-  left_join(Toe_temp,
-            by = c("eezid","tresh")
-            ) %>% 
-  left_join(Catch_trend,
-            by = c("taxon_key","sau_fishing_entity","fishing_entity_id")
-            ) %>% # 11492
-  left_join(Eez_biogeo,
-            by = "eezid") %>% 
-  left_join(Spp_env_pref,
-            by = "taxon_key") %>% 
-  left_join(Eez_centroids,
-            by = "eezid") %>% 
-  mutate(latitude = abs(latitude))
-
-# -------------------------------- #
-#### Build models
-# -------------------------------- #
-
-# Model 1; all variables
-M_all_var <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)
-
-all_sum <- summary(M_all_var)
-all_aic <- AIC(M_all_var)
-all_r <-  all_sum$adj.r.squared
-
-# Model 2; non demPel
-M2 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel, data = Lm_data)
-
-M2_sum <- summary(M2)
-M2_aic <- AIC(M2)
-M2_r <-  M2_sum$adj.r.squared
-
-# Model 3; non biome
-M3 <- lm(emerging_yr ~ sst_toe + bot_toe + status + dem_pel + DemersPelag, data = Lm_data)
-
-M3_sum <- summary(M3)
-M3_aic <- AIC(M3)
-M3_r <-  M3_sum$adj.r.squared
-
-# Model 4; non status
-M4 <- lm(emerging_yr ~ sst_toe + bot_toe + biome + dem_pel + DemersPelag, data = Lm_data)
-
-M4_sum <- summary(M4)
-M4_aic <- AIC(M4)
-M4_r <-  M4_sum$adj.r.squared
-
-# Model 5; non sst
-M5 <- lm(emerging_yr ~ status + bot_toe + biome + dem_pel + DemersPelag, data = Lm_data)
-
-M5_sum <- summary(M5)
-M5_aic <- AIC(M5)
-M5_r <-  M5_sum$adj.r.squared
-
-# Model 6; all + basin
-M6 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)
-
-M6_sum <- summary(M6)
-M6_aic <- AIC(M6)
-M6_r <-  M6_sum$adj.r.squared
-
-# Model 7; Categories instead of status
-M7 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + dem_pel + DemersPelag, data = Lm_data)
-
-M7_sum <- summary(M7)
-M7_aic <- AIC(M7)
-M7_r <-  M7_sum$adj.r.squared
-
-# Model 8; All variables + ccentroid lat
-M8 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + DemersPelag + latitude, data = Lm_data)
-
-M8_sum <- summary(M8)
-M8_aic <- AIC(M8)
-M8_r <-  M8_sum$adj.r.squared
-
-tibble(
-  "model" = paste("M",seq(1,8,1),sep=""),
-  "AIC" = c(all_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic),
-  "rsquared" = c(all_r,M2_r,M3_r,M4_r,M5_r,M6_r,M7_r,M8_r)
-)
-
-# UNIDAS 20282659
-
-# The Winner is M8!
-
-stargazer(M8, title = "Results", align=TRUE)
 
@@
 ```
-
-###### With relation to neighbours
-
-```{r time_emergence_map,  eval = T, echo = T}
-
-##_----------------------- #
-# Checking on China and Indonesia
-# Probably due to dispute territories
-
-Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
-  clean_names()
-
-# 
-# Neighbours %>% 
-#   filter(eez_name == "China") %>% 
-#   pull(eez_neighbour) %>% 
-#   unique()
-
-##_----------------------- #
-
-# Get the center poligon of each EEZ for map source/target
-coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
-  ungroup() %>% 
-  select(eez_name,geometry) %>% 
-  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
-  # Manually fix issues
-  mutate(
-    longitude = ifelse(
-      eez_name == "Fiji", 178.065033,
-      ifelse(eez_name == "Tuvalu",177.6493,
-             ifelse( eez_name == "New Zealand",174.8860,
-                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
-                     )
-             )
-      )
-    )
-  )
-
-# Remove "c()"
-coords$longitude <- gsub("\\(","",coords$longitude)
-coords$longitude <- gsub("c","",coords$longitude)
-coords$latitude <- gsub(")","",coords$latitude)
-
-coords <- coords %>% 
-  mutate_at(vars(longitude,latitude),as.numeric)
-
-# Prepare data from results
-Test_Data <- Eme_Res %>% 
-  filter(
-    n_ens == 10#,
-    # emerging_yr >= 2020
-  ) %>% 
-  group_by(eez_name,eez_neighbour,emerging_yr) %>% 
-  summarise(
-    n_tax = length(unique(taxon_key))
-  ) %>% 
-  filter(
-    emerging_yr <= 2050
-  ) %>%
-  group_by(eez_name,eez_neighbour) %>% 
-  summarise(
-    n_tax_emer = sum(n_tax)
-  ) %>% 
-  # get coords for sources
-  left_join(coords,
-            by = "eez_name") %>%
-  rename(source = eez_name,
-         start_lon = longitude,
-         start_lat = latitude,
-         eez_name=eez_neighbour) %>% 
-  # get coords for targets
-  left_join(coords,
-            by = "eez_name") %>% 
-  rename(target = eez_name,
-         end_lon = longitude,
-         end_lat = latitude) %>% 
-  ungroup() %>% 
-  # Plot miscs
-  # mutate(
-  #   per_mid = ifelse(is.na(per_mid),0,per_mid)
-  # ) %>% 
-  drop_na() #%>% 
-  # filter(
-    # per_mid >= 20,
-    # variable == "mean_spp",
-  # )
-
-# Convert to data to spatial
-network_data_sf <- Test_Data %>%
-  select(start_lon, start_lat, end_lon, end_lat) %>% 
-  purrr::transpose() %>% 
-  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
-  purrr::map(st_linestring) %>%
-  st_sfc(crs = 4326) %>%
-  st_sf(geometry = .)%>%
-  bind_cols(Test_Data) %>%
-  select(everything(), geometry) %>% 
-  # Set circles ('cus the world is not flat!)
-  st_segmentize(units::set_units(100, km)) %>% 
-  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
-  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
-  sf::st_sf(crs = 4326)
-
-#  unpack the coordinates of each feature 
-coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
-
-c(network_data_sf$line_id, network_data_sf$coords) %<-% (
-  coord_data %>% 
-    rename(subline_id = L1) %>%   
-    nest(-L2)
-)
-
-# Create data path for line connections
-paths <- network_data_sf %>% 
-  sf::st_set_geometry(NULL) %>% 
-  unnest() %>% 
-  mutate_at(vars(X,Y), round,2)
-
-# Plot it!
-
-country <- c("Fiji","Solomon Isl.","Tuvalu")
-
-ggplot() +
-  geom_sf(
-    # data = subset(world_map,name %in% country),
-    data = World_map,
-    size = 0.1, fill = "gray90", col = "gray90") +
-  geom_sf(
-    # data = subset(World_eez_sf, eez_name %in% country),
-    data = World_eez_sf,
-    size = 0.1, fill = "white", col = "gray70") +
-  geom_point(data = Test_Data,
-             # data = subset(Test_Data, source %in% country & target %in% source),
-             aes(start_lon,
-                 start_lat,
-                 col = n_tax_emer
-                 ),
-             size = 0.1,
-             col = "grey50",
-             # shape = 21,
-             stroke = 1.5) +
-  geom_path(data = paths,
-            # data = subset(paths, source %in% country),
-            aes(X, Y,
-                group = interaction(line_id,subline_id),
-                col = n_tax_emer
-            ),
-            size = 0.3,
-            # arrow = arrow(length = unit(0.009, "npc")),
-            show.legend = NA) +
-  ggtheme_map() +
-  scale_colour_viridis(
-    name="Number of emerging stocks\nby 2050",
-    limits = c(0,40),
-    breaks = seq(0,40,10),
-    # option="magma",
-    direction = 1
-  ) +
-  ggsave("./Figures/figure_emerging_2050.png",
-         width = 10,
-         height = 6,
-         units = "in"
-  )
-
-```

### 2020-08-31 Scripts/Initial_Analysis.Rmd

--- 2020-08-28 Scripts/Initial_Analysis.Rmd
@@
 
-##### Fig. Percentage Change Map
+##### Fig.Percentage Change Map
 
@@
 
-##### Fig. 1 and S1. (Netowrk)
+##### Fig. 2 and S1. (Netowrk)
 
@@
 
-##### Fig. 2 Mapping the time of emergence
+##### Fig. 3 Mapping the time of emergence
 
@@
   gather("metric","value",mean_yr:n_taxa) %>% 
-  # filter(metric != "n_taxa") %>% 
+  filter(metric != "n_taxa") %>% 
   mutate(
@@
 
-##### Fig. 3. Bubble plot
-
-```{r, buble_plot, eval = F, echo = F}
-
-Buble_data <- Eez_d_hist %>% 
-  filter(metric == "mean_yr") %>% 
-  group_by(sau_fishing_entity,tresh) %>% 
-  summarise(mean_yr = mean(value)) %>% 
-  left_join(Eme_value_data,
-            by = c("sau_fishing_entity","tresh")) %>% 
-  filter(tresh == tresh_list[i],
-         variable == "value") %>% 
-  mutate(sum_value_mi = sum_value/1000000) %>%   # Million usd
-  left_join(Matching_names,
-            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
-  distinct(sau_fishing_entity,.keep_all = TRUE) %>% 
-  filter(!is.na(un_geo_region)) %>% 
-  ungroup() %>% 
-  mutate(
-    sau_fishing_entity = ifelse(sum_value_mi >= 1000 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity)
-  )
-
-# Plot the bubbles!
-
-ggplot(Buble_data) +
-  geom_point(
-    aes(
-      x = mean_yr,
-      y = sum_value_mi,
-      size = n_taxa,
-      colour = un_geo_region
-    ),
-    alpha = 0.7
-  ) +
-  # scale_size(range = c(.1, 10), name="Fishing revenue (USD)")  +
-  geom_label_repel(data = subset(Buble_data, sum_value_mi >= 1000),
-                   aes(
-                     x = mean_yr,
-                     y = sum_value_mi,
-                     label = sau_fishing_entity
-                   ),
-                   arrow = arrow(length = unit(0.009, "npc"), 
-                                 type = "closed",
-                                 ends = "first"),
-                   force = 2,
-                   box.padding = 2,
-                   fill = "NA"
-
-  ) +
-  # scale_fill_viridis(discrete = F)
-  scale_color_viridis(name="Georaphic Region",
-                      discrete = T) +
-  scale_x_continuous("Average time of emergence",
-                     breaks = seq(2000,2080,20)
-  ) +
-  scale_y_continuous("Fishing revenue (million)",
-                     breaks = seq(1000,5000,1000),
-                     labels = scales::dollar
-  ) +
-  scale_size("Number of\nemerging taxa") +
-  ggsave("./Figures/bubble_plot.png",
-         width = 8,
-         height = 8,
-         units = "in"
-  )
-
-```
-
 ### Species level analysis

### 2020-08-31 Scripts/Initial_Analysis.Rmd
finalized bubble plot

--- 2020-08-28 Scripts/Initial_Analysis.Rmd
@@
     weight_toe = sum_value/mean_yr,
-    sum_value_mi = sum_value/1000000000,
-         logvalue = log10(sum_value)
-         ) %>%   # Billion usd
+    sum_value_bi = sum_value/1000000000) %>%   # Billion usd
   left_join(Matching_names,
@@
   mutate(
-    sau_fishing_entity = ifelse(sum_value_mi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity)
-  )
+    sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity)
+  )
+
+### Fucntion for estimating percentile
+
+quibble <- function(x, q = p) {
+        tibble(x = quantile(x, q), q = q)
+}
+
+Percentile_value <- Buble_data %>% 
+  group_by() %>%  # for whole distribution
+        do(quibble(.$sum_value_bi, 0.75))
 
@@
 Labels_data <- tibble(
-  x = c(2020,2040,2070),
-  y = c(0.3,0.3,9.1),
-  label = c("Early emergence", "Late emergence","Fisheries over the 75 percentile")
+  x = c(2020,2040,2007),
+  y = c(3.2,3.2,Percentile_value$x+.1),
+  label = c("Early emergence", "Late emergence","Fisheries > 75th percentile")
 )
@@
       x = mean_yr,
-      # y = sum_value_mi,
-      y = logvalue,
+      y = sum_value_bi,
       size = n_taxa,
@@
   ) +
+  # Divisory lines
   geom_vline(xintercept = 2030, color = "grey90", size = 0.5) +
-  geom_hline(yintercept = 9, color = "grey90", size = 0.5)  +
-  geom_label_repel(data = subset(Buble_data, sum_value_mi >= 1),
+  geom_hline(data = Percentile_value,
+             aes(yintercept = x),
+             color = "grey90"
+             )  +
+  geom_label_repel(data = subset(Buble_data, sum_value_bi >= 1),
                    aes(
                      x = mean_yr,
-                     # y = sum_value_mi,
-                     y = logvalue,
+                     y = sum_value_bi,
                      label = sau_fishing_entity
@@
                                  ends = "first"),
-                   force = 3,
-                   box.padding = 2,
+                   force = 2,
+                   box.padding = 1,
                    fill = "NA"
@@
   ) +
-  # scale_fill_viridis(discrete = F)
   scale_color_viridis(name="Region",
@@
   ) +
-  scale_y_continuous("Fishing revenue (log10)",
-                     limits = c(3,12),
-                     breaks = seq(3,12,3),
+  scale_y_continuous("Fishing revenue (billion USD)",
+                     trans = "log10",
+                     # limits = c(3,12),
+                     # breaks = seq(3,12,3),
                      expand = c(0,0)

### 2020-08-31 Scripts/Initial_Analysis.Rmd

--- 2020-08-31 Scripts/Initial_Analysis.Rmd
@@
 
+##### Fig. 3. Bubble plot
+
+```{r, buble_plot, eval = F, echo = F}
+
+Buble_data <- Eez_d_hist %>% 
+  filter(metric == "mean_yr") %>% 
+  group_by(sau_fishing_entity,tresh) %>% 
+  summarise(mean_yr = mean(value)) %>% 
+  left_join(Eme_value_data,
+            by = c("sau_fishing_entity","tresh")) %>% 
+  filter(tresh == tresh_list[i],
+         variable == "value") %>% 
+  mutate(
+    weight_toe = sum_value/mean_yr,
+    sum_value_bi = sum_value/1000000000) %>%   # Billion usd
+  left_join(Matching_names,
+            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
+  distinct(sau_fishing_entity,.keep_all = TRUE) %>% 
+  filter(!is.na(un_geo_region)) %>% 
+  ungroup() %>% 
+  mutate(
+    sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity)
+  )
+
+### Fucntion for estimating percentile
+
+quibble <- function(x, q = p) {
+        tibble(x = quantile(x, q), q = q)
+}
+
+Percentile_value <- Buble_data %>% 
+  group_by() %>%  # for whole distribution
+        do(quibble(.$sum_value_bi, 0.75))
+
+# Labels Data
+
+Labels_data <- tibble(
+  x = c(2020,2040,2007),
+  y = c(3.2,3.2,Percentile_value$x+.1),
+  label = c("Early emergence", "Late emergence","Fisheries > 75th percentile")
+)
+
+# Plot the bubbles!
+
+ggplot(Buble_data) +
+  geom_point(
+    aes(
+      x = mean_yr,
+      y = sum_value_bi,
+      size = n_taxa,
+      colour = un_geo_region
+    ),
+    alpha = 0.7
+  ) +
+  # Divisory lines
+  geom_vline(xintercept = 2030, color = "grey90", size = 0.5) +
+  geom_hline(data = Percentile_value,
+             aes(yintercept = x),
+             color = "grey90"
+             )  +
+  geom_label_repel(data = subset(Buble_data, sum_value_bi >= 1),
+                   aes(
+                     x = mean_yr,
+                     y = sum_value_bi,
+                     label = sau_fishing_entity
+                   ),
+                   arrow = arrow(length = unit(0.009, "npc"),
+                                 type = "closed",
+                                 ends = "first"),
+                   force = 2,
+                   box.padding = 1,
+                   fill = "NA"
+
+  ) +
+  geom_text(data = Labels_data,
+    aes(
+      x = x,
+      y = y,
+      label = label
+    ),
+    size = 3,
+    color = "grey30"
+  ) +
+  scale_color_viridis(name="Region",
+                      discrete = T) +
+  scale_x_continuous("Average time of emergence",
+                     limits = c(2000,2080),
+                     breaks = seq(2000,2080,10)
+  ) +
+  scale_y_continuous("Fishing revenue (billion USD)",
+                     trans = "log10",
+                     # limits = c(3,12),
+                     # breaks = seq(3,12,3),
+                     expand = c(0,0)
+  ) +
+  scale_size("Number of\nemerging taxa") +
+  ggsave("./Figures/bubble_plot.png",
+         width = 8,
+         height = 8,
+         units = "in"
+  )
+
+ ```
+
 ### Species level analysis

### 2020-08-31 Scripts/Initial_Analysis.Rmd
Completed first draft of results

--- 2020-08-31 Scripts/Initial_Analysis.Rmd
@@
   "rfishbase",
-  "stargazer" # For nice lm tables
+  "stargazer", # For nice lm tables
+  "tidytext" # for boxplot order
 )
@@
 
-##### Fig.Percentage Change Map
-
-```{r fig_map_per, eval = F, echo =F}
-
-World_sf %>% 
-  left_join(results,
-            by = "eez_name") %>% 
-  filter(variable %in% c("mean_spp")#,
-         # eez_name %in% c("Peru","Chile","Ecuador","Argentina")
-         ) %>% 
-  ggplot() +
-  geom_sf(
-    aes(
-      fill = per_mid
-    )
-  ) + 
-  geom_sf(data=country10_sf) +
-  scale_fill_gradient2("Average Change in\nStock-Share Ratio (%)\nBy Mid Century") +
-  ggtheme_map() +
-  ggsave("per_mid.png",
-         width = 10,
-         height = 6,
-         units = "in"
-  )
-
-```
-
-##### Fig. 2 and S1. (Netowrk)
+##### Fig. 1 and S1. (Netowrk)
 
@@
 
-##### Histogram of Data
-
-```{r histo,  eval = T, echo = T}
-
-# First time a Species becomes emergent
-Eme_Res %>% 
-  filter(
-    n_ens == 10#,
-    # tresh == "tresh_one"
-  ) %>% 
-  group_by(tresh,taxon_key,emerging_yr) %>% 
-  summarise(
-    n_trans = length(unique(sf_eez_name))
-  ) %>% 
-  group_by(tresh,taxon_key) %>% 
-  summarise(
-    ntrans = min(emerging_yr)
-  ) %>% 
-  group_by(tresh,ntrans) %>% 
-  summarise(
-    n_taxa = length(unique(taxon_key))
-  ) %>% 
-  arrange(-n_taxa) %>% 
-  # View()
-  # group_by(tresh) %>% 
-  # summarise(sum(n_taxa))
-  ggplot() +
-  geom_line(
-    aes(
-      x = ntrans,
-      y = n_taxa,
-      color = tresh
-    )
-  )
-
-# EEZ, first time an EEZ has a emergent stock
-Eme_Res %>% 
-  filter(
-    n_ens == 10#,
-    # tresh == "tresh_one"
-  ) %>% 
-  group_by(tresh,sf_eez_name,emerging_yr) %>% 
-  summarise(
-    n_trans = length(unique(taxon_key))
-  ) %>% 
-  group_by(tresh,sf_eez_name) %>% 
-  summarise(
-    ntrans = min(emerging_yr)
-  ) %>% 
-  group_by(tresh,ntrans) %>% 
-  summarise(
-    n_taxa = length(unique(sf_eez_name))
-  ) %>% 
-  arrange(-n_taxa) %>% 
-  View()
-  # group_by(tresh) %>%
-  # summarise(sum(n_taxa))
-  # ggplot() +
-  # geom_line(
-  #   aes(
-  #     x = ntrans,
-  #     y = n_taxa,
-  #     color = tresh
-  #   )
-  # )
-
-# Numer of EEZs where stocks will emerge
-stock_d_hist <- Eme_Res %>% 
-  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
-  group_by(emerging_yr,taxon_key,tresh) %>%
-  summarise(n_tax = length(unique(sf_eez_name))) %>%
-  # summarise(n()) %>% 
-  mutate(
-    level = "EEZs per Spp"
-  )  %>% 
-  ungroup() %>% 
-  select(-taxon_key)
-
-# Per EEZs
-eez_d_hist <- Eme_Res %>% 
-  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
-  group_by(emerging_yr,sf_eez_name,tresh,neighbour_id) %>%  
-  # summarise(n_tax = length(unique(taxon_key))) %>% 
-  summarise(n()) %>% 
-  mutate(
-    level = "Spp per EEZ"
-  ) %>% 
-  ungroup() %>% 
-  select(-sf_eez_name)
-
-stock_d_hist %>% 
-  bind_rows(eez_d_hist) %>% 
-  filter(tresh != "NA") %>% 
-  ggplot() +
-  geom_density(
-    aes(
-      x = emerging_yr,
-      fill = level
-    ),
-    stat = "count",
-    alpha= 0.8,
-    size=0.5,
-  ) +
-  scale_y_continuous("Frequency",
-                     # breaks = seq(0,300,50),
-                     # limits = c(0,300)
-  ) +
-  scale_x_continuous("Year of emergence",
-                     # breaks = seq(2006,2100,25),
-                     # limits = c(2000,2100)
-  ) +
-  my_ggtheme_p() +
-  scale_fill_viridis(discrete = T) +
-  facet_wrap(~tresh) +
-  ggsave("./Figures/hist_freq.png",
-         width = 6,
-         height = 6,
-         units = "in"
-  )
-
-```
-
-##### Accumulative plots
+##### Fig. 2. Accumulative plots
 
@@
     com_sum = cumsum(n_tax),
-    level = "Stock"
+    level = "stock"
   )
@@
     com_sum = cumsum(n_tax),
-    level = "EEZ"
+    level = "eez"
   )
@@
   mutate(
-    total = ifelse(level == "Stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$eez_name))),
+    total = ifelse(level == "stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$sf_eez_name))),
     per = com_sum/total*100
-  ) ;plot_data
-
-  # filter(emerging_yr >= 2020) %>% 
+  ) %>% 
+  gather("metric","value",com_sum,per) %>% 
+  filter(tresh == "Tresh_one") %>% 
+  mutate(
+    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum"),
+    label = ifelse(total == 478,"Total number of stocks","Total number of EEZs")
+  );plot_data
+
+#### Both plots for tresh one
+
 ggplot() +
-    geom_area(data = subset(plot_data, level == "Stock"),
+  geom_area(data = subset(plot_data, level == "stock"),
             aes(
               x = as.numeric(taxon_emer),
-                y = com_sum,
+              y = value,
               fill = level
@@
   )  +
-    geom_area(data = subset(plot_data, level == "EEZ"),
+  geom_area(data = subset(plot_data, level == "eez"),
             aes(
-                x = taxon_emer,
-                y = com_sum,
+              x = as.numeric(taxon_emer),
+              y = value,
               fill = level
@@
   )  +
-    facet_wrap(~tresh) + 
-    # Comment out for option A (Change plot name!!)
-  #   geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
-  # geom_text(aes(label="Total number of EEZs", x = 2010, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
-  #   geom_hline(yintercept = length(unique(Eme_Res$taxon_key)), color = "grey",linetype = "dashed") +
-  # geom_text(aes(label="Total number of Taxa", x = 2010, y = length(unique(Eme_Res$taxon_key))+10), color = "grey70",size =3) +
-    scale_y_continuous("Cumulatuve Curve",
-                       breaks = seq(0,600,50),
-                       limits = c(0,600)
-    ) +
-    scale_x_continuous("Year of emergence",
-                       breaks = seq(2005,2100,10),
-                       limits = c(2005,2100)
-    ) +
-    my_ggtheme_p() +
-    scale_fill_viridis(discrete = T) +
-  ggsave("./Figures/figure_accumulative_curve_ab.png",
-         width = 6,
-         height = 6,
-         units = "in"
-  )
-
-  ### Percentage accumulation
-
-  # filter(emerging_yr >= 2020) %>% 
-  ggplot() +
-    geom_area(data = subset(plot_data, level == "Stock"),
+  geom_line(data = subset(plot_data, metric == "com_sum"),
             aes(
               x = as.numeric(taxon_emer),
-                y = per,
+              y = total,
               fill = level
             ),
-              alpha= 0.8,
-              size=0.5,
-              colour="black"
-    )  +
-    geom_area(data = subset(plot_data, level == "EEZ"),
+            colour="grey30",
+  )  +
+  geom_text(data = subset(plot_data, metric == "com_sum"),
             aes(
-                x = taxon_emer,
-                y = per,
-                fill = level
+              x = 2037,
+              y = total+5,
+              label = label
             ),
-              alpha= 0.8,
-              size=0.5,
-              colour="black"
-    ) +
-    scale_y_continuous("Cumulatuve Curve",
-                       breaks = seq(0,100,25),
-                       limits = c(0,100)
-    ) +
-    scale_x_continuous("Year of emergence",
-                       breaks = seq(2005,2100,15),
-                       limits = c(2005,2100)
+            size = 3,
+            colour="grey30",
+            hjust=0.6,
+  )  +
+  facet_wrap(~my_facet,
+             scales = "free") + 
+  scale_y_continuous("",expand = expand_scale(mult = c(0, .05))) +
+  scale_x_continuous("Time of emergence",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     expand = c(0, 0)
   ) +
   my_ggtheme_p() +
-    scale_fill_viridis(discrete = T) +
-    facet_wrap(~tresh) + 
-  ggsave("./Figures/Fig4.png",
-         width = 8,
-         height = 8,
+  theme(
+    legend.position = "right"
+  ) +
+  guides(fill=guide_legend(title= "Level")) +
+  scale_fill_viridis(discrete = T,
+                     labels = c("EEZs","Stocks")
+  ) +
+  ggsave("./Figures/cumu_per_freq.png",
+         width = 7,
+         height = 7,
          units = "in"
-  )
-
-```
-
-##### Scatter plots
-
-```{r Scatter_plot,  eval = T, echo = T}
-
- data_stock <- Eme_Res %>% 
-  filter(
-    n_ens == 10,
-    emerging_yr >= 2020
-  ) %>% 
-  group_by(emerging_yr,eez_name) %>% 
-  summarise(
-    tax_n = length(unique(taxon_key))
-    ) %>% 
-  group_by(emerging_yr) %>% 
-  summarise(emerging_tax = sum(tax_n))
-
-data_eez <- Eme_Res %>% 
-  filter(
-    n_ens == 10,
-    emerging_yr >= 2020
-  ) %>% 
-  group_by(emerging_yr,taxon_key) %>% 
-  summarise(
-    eez_n = length(unique(eez_name))
-    ) %>% 
-    group_by(emerging_yr) %>% 
-  summarise(emerging_eez = sum(eez_n))
-
-data_stock %>% 
-  left_join(data_eez,
-            by = "emerging_yr") %>% 
-  ggplot() +
-  geom_point(
-    aes(
-      x = emerging_eez,
-      y = emerging_tax,
-      fill = emerging_yr,
-      colour = emerging_yr
-    )
-  )
-
-```
-
-##### Temperature gradient
-
-##### Fun `GetEnvData`
-
-```{r get_env_data, eval = T, echo = T}
-
-getEnvVar=function(yr,ModelRCP,RCP,Variable){
-
-  if(ModelRCP == "IPSL"){
-    ModelRCP = "IPSL-CM5A-MR"
-    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
-  }
-
-  if(ModelRCP == "MPI"){
-    ModelRCP = "MPI-ESM-MR"
-    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
-  }
-
-  if(ModelRCP == "GFDL"){
-
-    if(RCP == "rcp26"){
-      ModelRCP <- "GFDL26"
-    }else{
-      ModelRCP <- "GFDL85"
-    }
-    Filepath <- paste("Z:/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")
-  }
-
-  if(Variable == "Ice"){
-
-    Bottom <- paste("IceExt_",yr,".txt",sep="")
-
-  }
-
-  if(Variable == "Temperature"){
-
-    Bottom = paste("SST_",yr,".txt",sep="")
-
-  }
-
-  if(Variable == "Salinity"){
-    # Bottom
-    Bottom = paste("Salinity_btm_",yr,".txt",sep="")
-
-  }
-
-  Read_me <- paste(Filepath,Bottom,sep="")[1]
-  Bottom_Data <- read.csv(Read_me,header=F)
-
-  # Retrun
-
-  getTemp <- Bottom_Data 
-}
-
-Years <-seq(1951,2100,1)
-
-for(y in 1:length(Years)){
-        Data <- getEnvVar(yr = Years[y],
-                          Model = "GFDL",
-                          Variable = "Temperature",
-                          RCP = "rcp85"
-        )
-
-        if(y == 1){
-
-          F_BData <- Data[1]
-          colnames(F_BData) <- Years[1]
-
-        }else{
-
-          F_BData <- cbind(F_BData,Data[1])
-          colnames(F_BData)[y] <- Years[y]
-        }
-}
-
-F_BData <- F_BData %>% 
-          rowid_to_column("index") %>% 
-  gather("year","temp",-index) %>% 
-   mutate(temp = ifelse(temp == -9999,NA,temp)) %>% 
-  group_by(year) %>% 
-  summarise(
-    mean_temp = mean(temp,na.rm=T)
-  ) %>% 
-  write_csv(.,
-            "SST_data.csv")
-
-```
-
-```{r temp_plot, eval = T, echo = T}
-
-### NEED TO RUN TIME LAPSE CHUNK FIRST TO GET ACCUMULATIVE DATA
-
-# Get sst_data
-sst_data <- my_path("D",T,"Climate","SST_data.csv",read = T)
-
-sst_hist <- sst_data %>% 
-  filter(year < 2000) %>% 
-  group_by() %>% 
-  summarise(hist_sst =mean(mean_temp)) %>% 
-  pull()
-
-sst_change <- sst_data %>%
-  filter(year > 2020) %>% 
-  mutate(chng = round(mean_temp-sst_hist,2))
-
-  # Plot Data
-stock_d %>% 
-  bind_rows(eez_d) %>% 
-  filter(taxon_emer >= 2020) %>% 
-  rename(year = taxon_emer) %>% 
-  left_join(sst_change) %>% 
-  ggplot() +
-  geom_line(
-    aes(
-      x = chng,
-      y = com_sum,
-      colour = level
-    )
-  ) +
-  geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
-  geom_text(aes(label="Total number of EEZs", x = 0.5, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
-    scale_y_continuous("Cumulatuve Curve",
-                       breaks = seq(0,450,50),
-                       limits = c(0,500)
-    ) +
-    scale_x_continuous("Change in SST (C)",
-                       breaks = seq(0,2,0.2),
-                       limits = c(0.3,2.3)
-    ) +
-    my_ggtheme_p() +
-    scale_colour_viridis(discrete = T) +
-  ggsave("./Figures/figure_temp.png",
-         width = 6,
-         height = 6,
-         units = "in"
-  )
-
-```
-
-```{r Scatter_plot,  eval = T, echo = T}
-
- data_stock <- Eme_Res %>% 
-  filter(
-    n_ens == 10,
-    emerging_yr >= 2020
-  ) %>% 
-  group_by(emerging_yr,eez_name) %>% 
-  summarise(
-    tax_n = length(unique(taxon_key))
-    ) %>% 
-  group_by(emerging_yr) %>% 
-  summarise(emerging_tax = sum(tax_n))
-
-data_eez <- Eme_Res %>% 
-  filter(
-    n_ens == 10,
-    emerging_yr >= 2020
-  ) %>% 
-  group_by(emerging_yr,taxon_key) %>% 
-  summarise(
-    eez_n = length(unique(eez_name))
-    ) %>% 
-    group_by(emerging_yr) %>% 
-  summarise(emerging_eez = sum(eez_n))
-
-data_stock %>% 
-  left_join(data_eez,
-            by = "emerging_yr") %>% 
-  rename(year = emerging_yr) %>% 
-  left_join(sst_data) %>% 
-  ggplot() +
-  geom_point(
-    aes(
-      x = mean_temp,
-      y = emerging_eez,
-      fill = year,
-      colour = year
-    )
   )
@@
 
+##### Fig. 4. Bubble plot
+
+```{r, buble_plot, eval = F, echo = F}
+
+Buble_data <- Eez_d_hist %>% 
+  filter(metric == "mean_yr") %>% 
+  group_by(sau_fishing_entity,tresh) %>% 
+  summarise(mean_yr = mean(value)) %>% 
+  left_join(Eme_value_data,
+            by = c("sau_fishing_entity","tresh")) %>% 
+  filter(tresh == "Tresh_one",
+         variable == "value") %>% 
+  mutate(
+    weight_toe = sum_value/mean_yr,
+    sum_value_bi = sum_value/1000000000,
+    log_value = log10(sum_value_bi)
+    ) %>%   # Billion usd
+  left_join(Matching_names,
+            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
+  distinct(sau_fishing_entity,.keep_all = TRUE) %>% 
+  filter(!is.na(un_geo_region)) %>% 
+  ungroup() %>% 
+  mutate(
+    sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity))
+
+### Fucntion for estimating percentile
+
+quibble <- function(x, q = p) {
+        tibble(x = quantile(x, q), q = q)
+}
+
+Percentile_value <- Buble_data %>% 
+  group_by() %>%  # for whole distribution
+        do(quibble(.$log_value, 0.75))
+
+# Labels Data
+
+Labels_data <- tibble(
+  x = c(2020,2040,2007),
+  y = c(-5.8,-5.8,Percentile_value$x+.1),
+  label = c("Early emergence", "Late emergence","Fisheries > 75th percentile")
+)
+
+# Plot the bubbles!
+ggplot(Buble_data) +
+  geom_point(
+    aes(
+      x = mean_yr,
+      y = log_value,
+      size = n_taxa,
+      colour = un_geo_region
+    ),
+    alpha = 0.7
+  ) +
+  # Divisory lines
+  geom_vline(xintercept = 2030, color = "grey90", size = 0.5) +
+  geom_hline(data = Percentile_value,
+             aes(yintercept = x),
+             color = "grey90"
+  )  +
+  geom_label_repel(data = subset(Buble_data, sum_value_bi >= 1),
+                   aes(
+                     x = mean_yr,
+                     y = log_value,
+                     label = sau_fishing_entity
+                   ),
+                   arrow = arrow(length = unit(0.009, "npc"),
+                                 type = "closed",
+                                 ends = "first"),
+                   force = 2,
+                   box.padding = 1,
+                   fill = "NA"
+
+  ) +
+  geom_text(data = Labels_data,
+            aes(
+              x = x,
+              y = y,
+              label = label
+            ),
+            size = 3,
+            color = "grey30"
+  ) +
+  scale_color_viridis(name="Region",
+                      discrete = T) +
+  scale_x_continuous("Average time of emergence",
+                     limits = c(2000,2080),
+                     breaks = seq(2000,2080,10)
+  ) +
+  scale_y_continuous("Fishing revenue (log10)",
+                     limits = c(-6,2),
+                     breaks = seq(-6,2,2),
+                     expand = c(0,0)
+  ) +
+  scale_size("Number of\nemerging taxa") +
+  ggsave("./Figures/bubble_plot.png",
+         width = 8,
+         height = 8,
+         units = "in"
+  )
+
+```
+
 ### Species level analysis
 
-#### Demersal Vs Pelagic
+#### Fig. 5 Demersal Vs Pelagic
 
@@
          !is.na(un_sub_region)) %>% 
-  left_join(Spp_depth)
-
-Dem_pel_mean <- Dem_pel %>% 
-  group_by(dem_pel) %>% 
-  summarise(mean  = mean(emerging_yr))
-
+  left_join(Spp_env_pref) %>% 
+  mutate(DemersPelag = ifelse(is.na(DemersPelag),"other",
+                              ifelse(DemersPelag %in% c("bathypelagic","benthopelagic"), "bathy-bentho-pelagic",paste(DemersPelag)
+                              )
+  ),
+    un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and the Car.",paste(un_sub_region)))
+
+## Averages
+
+Average_data <- Dem_pel %>% 
+  group_by(un_sub_region) %>% 
+  summarise(mean_toe = mean(emerging_yr,na.rm=T)) %>% 
+  mutate(regional_average="average")
+
+# Dem Pelagic boxplot
 ggplot(Dem_pel) +
+  geom_hline(data = Average_data,
+             aes(
+               yintercept = mean_toe,
+               colour = regional_average
+             ),
+             linetype = "dashed"
+             ) +
   geom_boxplot(
     aes(
-      x = reorder(dem_pel,emerging_yr),
+      x = reorder_within(DemersPelag,emerging_yr,un_sub_region),
       y = emerging_yr,
-      fill = un_sub_region
-    )
-  ) +
-  geom_line(data = Dem_pel_mean,
-    aes(
-      x = dem_pel,
-      y = mean
-    )
-)
-
-ggplot(Dem_pel) +
-  geom_boxplot(
-    aes(
-      x = reorder(dem_pel,emerging_yr),
-      y = emerging_yr
-    )
-  ) +
-  facet_wrap(~un_sub_region)
+      fill = DemersPelag
+    ),
+    alpha = 0.8
+  ) +
+  facet_wrap(~un_sub_region,scales = "free") +
+  scale_fill_viridis("Habitat preference",
+                     discrete = T) +
+  scale_color_grey("Regional average") +
+  xlab("") +
+  ylab("Time of emergence") +
+  my_ggtheme_p() +
+  theme(
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    legend.position = "right"
+  ) +
+  scale_x_reordered() +
+  ggsave("./Figures/regional_hab_boxplot.png",
+         width = 14,
+         height = 8,
+         units = "in"
+  )
+```
+
+### Statistics
+
+## Regional differences
+
+```{r multiple_reg, eval = F, echo = F}
+
+reg_diff_data <- Eme_Res %>% 
+  filter(tresh == "Tresh_one",
+         !is.na(un_sub_region)
+         )
+
+Reg_aov <- aov(emerging_yr~un_sub_region, data = reg_diff_data)
+Reg_aov
+
+summary(Reg_aov)
+
+TukeyHSD(Reg_aov) %>% 
+  tidy() %>% 
+   filter(adj.p.value < 0.05) %>%
+  View()
+
+```
+
+## Habitat preference differences
+
+```{r multiple_reg, eval = F, echo = F}
+
+hab_diff_data <- Eme_Res %>% 
+  filter(tresh == "Tresh_one",
+         !is.na(un_sub_region)
+         ) %>% 
+  left_join(Spp_env_pref,
+            by = "taxon_key")
+
+Hab_aov <- aov(emerging_yr~DemersPelag, data = hab_diff_data)
+Hab_aov
+
+summary(Hab_aov)
+
+TukeyHSD(Hab_aov) %>% 
+  tidy() %>% 
+   filter(adj.p.value < 0.05) %>%
+  View()
 
@@
 # Model 8; All variables + ccentroid lat
-M8 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + DemersPelag + latitude, data = Lm_data)
+M8 <- lm(emerging_yr ~ sst_toe + bot_toe + category + as.factor(biome) + DemersPelag + latitude, data = Lm_data)
 
@@
 ```
+
+### Non selected Results
+
+```{r, olda_acumulative}
+
+# filter(emerging_yr >= 2020) %>% 
+  ggplot() +
+    geom_area(data = subset(plot_data, level == "stock"),
+              aes(
+                x = as.numeric(taxon_emer),
+                y = com_sum,
+                fill = level
+              ),
+              alpha= 0.8,
+              size=0.5,
+              colour="black"
+    )  +
+    geom_area(data = subset(plot_data, level == "eez"),
+              aes(
+                x = taxon_emer,
+                y = com_sum,
+                fill = level
+              ),
+              alpha= 0.8,
+              size=0.5,
+              colour="black"
+    ) +
+    facet_wrap(~tresh) + 
+    # Comment out for option A (Change plot name!!)
+  #   geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
+  # geom_text(aes(label="Total number of EEZs", x = 2010, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
+  #   geom_hline(yintercept = length(unique(Eme_Res$taxon_key)), color = "grey",linetype = "dashed") +
+  # geom_text(aes(label="Total number of Taxa", x = 2010, y = length(unique(Eme_Res$taxon_key))+10), color = "grey70",size =3) +
+    scale_y_continuous("Cumulatuve Curve",
+                       breaks = seq(0,600,50),
+                       limits = c(0,600)
+    ) +
+    scale_x_continuous("Year of emergence",
+                       breaks = seq(2005,2100,10),
+                       limits = c(2005,2100)
+    ) +
+    my_ggtheme_p() +
+    scale_fill_viridis(discrete = T) +
+  ggsave("./Figures/figure_accumulative_curve_ab.png",
+         width = 6,
+         height = 6,
+         units = "in" 
+  )
+
+  ### Percentage accumulation
+
+  ggplot() +
+    geom_area(data = subset(plot_data, level == "stock"),
+              aes(
+                x = as.numeric(taxon_emer),
+                y = per,
+                fill = level
+              ),
+              alpha= 0.8,
+              size=0.5,
+              colour="black"
+    )  +
+    geom_area(data = subset(plot_data, level == "eez"),
+              aes(
+                x = taxon_emer,
+                y = per,
+                fill = level
+              ),
+              alpha= 0.8,
+              size=0.5,
+              colour="black"
+    ) +
+    scale_y_continuous("Cumulatuve Curve",
+                       breaks = seq(0,100,25),
+                       limits = c(0,100)
+    ) +
+    scale_x_continuous("Year of emergence",
+                       breaks = seq(2005,2100,15),
+                       limits = c(2005,2100)
+    ) +
+    my_ggtheme_p() +
+    scale_fill_viridis(discrete = T) +
+    # facet_wrap(~tresh) + 
+  ggsave("./Figures/Fig4.png",
+         width = 8,
+         height = 8,
+         units = "in"
+  )
+
+```
+
+```{r histo,  eval = T, echo = T}
+
+# First time a Species becomes emergent
+Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # tresh == "tresh_one"
+  ) %>% 
+  group_by(tresh,taxon_key,emerging_yr) %>% 
+  summarise(
+    n_trans = length(unique(sf_eez_name))
+  ) %>% 
+  group_by(tresh,taxon_key) %>% 
+  summarise(
+    ntrans = min(emerging_yr)
+  ) %>% 
+  group_by(tresh,ntrans) %>% 
+  summarise(
+    n_taxa = length(unique(taxon_key))
+  ) %>% 
+  arrange(-n_taxa) %>% 
+  # View()
+  # group_by(tresh) %>% 
+  # summarise(sum(n_taxa))
+  ggplot() +
+  geom_line(
+    aes(
+      x = ntrans,
+      y = n_taxa,
+      color = tresh
+    )
+  )
+
+# EEZ, first time an EEZ has a emergent stock
+Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # tresh == "tresh_one"
+  ) %>% 
+  group_by(tresh,sf_eez_name,emerging_yr) %>% 
+  summarise(
+    n_trans = length(unique(taxon_key))
+  ) %>% 
+  group_by(tresh,sf_eez_name) %>% 
+  summarise(
+    ntrans = mean(emerging_yr)
+  ) %>% 
+  group_by(tresh,ntrans) %>% 
+  summarise(
+    n_taxa = length(unique(sf_eez_name))
+  ) %>% 
+  arrange(-n_taxa) %>% 
+  View()
+  # group_by(tresh) %>%
+  # summarise(sum(n_taxa))
+  # ggplot() +
+  # geom_line(
+  #   aes(
+  #     x = ntrans,
+  #     y = n_taxa,
+  #     color = tresh
+  #   )
+  # )
+
+# Numer of EEZs where stocks will emerge
+stock_d_hist <- Eme_Res %>% 
+  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
+  group_by(emerging_yr,taxon_key,tresh) %>%
+  summarise(n_tax = length(unique(sf_eez_name))) %>%
+  # summarise(n()) %>% 
+  mutate(
+    level = "EEZs per Spp"
+  )  %>% 
+  ungroup() %>% 
+  select(-taxon_key)
+
+# Per EEZs
+eez_d_hist <- Eme_Res %>% 
+  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
+  group_by(emerging_yr,sf_eez_name,tresh,neighbour_id) %>%  
+  # summarise(n_tax = length(unique(taxon_key))) %>% 
+  summarise(n()) %>% 
+  mutate(
+    level = "Spp per EEZ"
+  ) %>% 
+  ungroup() %>% 
+  select(-sf_eez_name)
+
+stock_d_hist %>% 
+  bind_rows(eez_d_hist) %>% 
+  filter(tresh != "NA") %>% 
+  ggplot() +
+  geom_density(
+    aes(
+      x = emerging_yr,
+      fill = level
+    ),
+    stat = "count",
+    alpha= 0.8,
+    size=0.5,
+  ) +
+  scale_y_continuous("Frequency",
+                     # breaks = seq(0,300,50),
+                     # limits = c(0,300)
+  ) +
+  scale_x_continuous("Year of emergence",
+                     # breaks = seq(2006,2100,25),
+                     # limits = c(2000,2100)
+  ) +
+  my_ggtheme_p() +
+  scale_fill_viridis(discrete = T) +
+  facet_wrap(~tresh) +
+  ggsave("./Figures/hist_freq.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
+```
+
+##### Temperature gradient
+
+##### Fun `GetEnvData`
+
+```{r get_env_data, eval = T, echo = T}
+
+getEnvVar=function(yr,ModelRCP,RCP,Variable){
+
+  if(ModelRCP == "IPSL"){
+    ModelRCP = "IPSL-CM5A-MR"
+    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
+  }
+
+  if(ModelRCP == "MPI"){
+    ModelRCP = "MPI-ESM-MR"
+    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
+  }
+
+  if(ModelRCP == "GFDL"){
+
+    if(RCP == "rcp26"){
+      ModelRCP <- "GFDL26"
+    }else{
+      ModelRCP <- "GFDL85"
+    }
+    Filepath <- paste("Z:/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")
+  }
+
+  if(Variable == "Ice"){
+
+    Bottom <- paste("IceExt_",yr,".txt",sep="")
+
+  }
+
+  if(Variable == "Temperature"){
+
+    Bottom = paste("SST_",yr,".txt",sep="")
+
+  }
+
+  if(Variable == "Salinity"){
+    # Bottom
+    Bottom = paste("Salinity_btm_",yr,".txt",sep="")
+
+  }
+
+  Read_me <- paste(Filepath,Bottom,sep="")[1]
+  Bottom_Data <- read.csv(Read_me,header=F)
+
+  # Retrun
+
+  getTemp <- Bottom_Data 
+}
+
+Years <-seq(1951,2100,1)
+
+for(y in 1:length(Years)){
+        Data <- getEnvVar(yr = Years[y],
+                          Model = "GFDL",
+                          Variable = "Temperature",
+                          RCP = "rcp85"
+        )
+
+        if(y == 1){
+
+          F_BData <- Data[1]
+          colnames(F_BData) <- Years[1]
+
+        }else{
+
+          F_BData <- cbind(F_BData,Data[1])
+          colnames(F_BData)[y] <- Years[y]
+        }
+}
+
+F_BData <- F_BData %>% 
+          rowid_to_column("index") %>% 
+  gather("year","temp",-index) %>% 
+   mutate(temp = ifelse(temp == -9999,NA,temp)) %>% 
+  group_by(year) %>% 
+  summarise(
+    mean_temp = mean(temp,na.rm=T)
+  ) %>% 
+  write_csv(.,
+            "SST_data.csv")
+
+```
+
+```{r temp_plot, eval = T, echo = T}
+
+### NEED TO RUN TIME LAPSE CHUNK FIRST TO GET ACCUMULATIVE DATA
+
+# Get sst_data
+sst_data <- my_path("D",T,"Climate","SST_data.csv",read = T)
+
+sst_hist <- sst_data %>% 
+  filter(year < 2000) %>% 
+  group_by() %>% 
+  summarise(hist_sst =mean(mean_temp)) %>% 
+  pull()
+
+sst_change <- sst_data %>%
+  filter(year > 2020) %>% 
+  mutate(chng = round(mean_temp-sst_hist,2))
+
+  # Plot Data
+stock_d %>% 
+  bind_rows(eez_d) %>% 
+  filter(taxon_emer >= 2020) %>% 
+  rename(year = taxon_emer) %>% 
+  left_join(sst_change) %>% 
+  ggplot() +
+  geom_line(
+    aes(
+      x = chng,
+      y = com_sum,
+      colour = level
+    )
+  ) +
+  geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
+  geom_text(aes(label="Total number of EEZs", x = 0.5, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
+    scale_y_continuous("Cumulatuve Curve",
+                       breaks = seq(0,450,50),
+                       limits = c(0,500)
+    ) +
+    scale_x_continuous("Change in SST (C)",
+                       breaks = seq(0,2,0.2),
+                       limits = c(0.3,2.3)
+    ) +
+    my_ggtheme_p() +
+    scale_colour_viridis(discrete = T) +
+  ggsave("./Figures/figure_temp.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
+```
+
+```{r Scatter_plot,  eval = T, echo = T}
+
+ data_stock <- Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr,eez_name) %>% 
+  summarise(
+    tax_n = length(unique(taxon_key))
+    ) %>% 
+  group_by(emerging_yr) %>% 
+  summarise(emerging_tax = sum(tax_n))
+
+data_eez <- Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr,taxon_key) %>% 
+  summarise(
+    eez_n = length(unique(eez_name))
+    ) %>% 
+    group_by(emerging_yr) %>% 
+  summarise(emerging_eez = sum(eez_n))
+
+data_stock %>% 
+  left_join(data_eez,
+            by = "emerging_yr") %>% 
+  rename(year = emerging_yr) %>% 
+  left_join(sst_data) %>% 
+  ggplot() +
+  geom_point(
+    aes(
+      x = mean_temp,
+      y = emerging_eez,
+      fill = year,
+      colour = year
+    )
+  )
+
+```

### 2020-08-31 Scripts/Manuscript.Rmd
log transformed bubble plot
Completed first draft of results

@@
 We consider each species as a single stock unit within an EEZ due to lack of more detailed information [@Teh:2015gd; @Palacios-Abrantes:xxxx].
-We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) **noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic) including islands territories**, and determined the intersections between polygons using the R package *sf* [@PackagesfSimpleF:2018vp].
+We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic) including islands territories, and determined the intersections between polygons using the R package *sf* (**Supplements Fig 1 (SuppM)**) [@PackagesfSimpleF:2018vp].
 
@@
 Each species centroid was determined by the average latitude and longitude of the grid cells containing the top 10% abundance percentile of each year.
-The EEZ centroid was estimated using the *st* package of *R* statistical software.
+The EEZ centroid was estimated using the *st* package of *R* statistical software (**Supplements Fig 1 (SuppM)**).
 For each ensamble member, neighbouring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring elevation differences:
@@
 Since this was an arbritary decission, we have included a more conservative estimation ($\mu \pm 2*\sigma$ for a 95% confidence interval).
+The SNR for both bottom and surface sea temperature was also estimated following the previos equation.
 
@@
 Consequently, if the centroid shifts are within historical variability (*SNR* < 1), risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability (*SNR* > 1), the risk of dispute is higher.
-Thus, risk is positively related to the frequency of centroid shifts beyond historical variability (Fig.
-1).
+Thus, risk is positively related to the frequency of centroid shifts beyond historical variability.
 
-![Fig 1.
-Theorical representation of transboundary Index.
-(A) Case with two neighbouring EEZs.
-(B) Case with three neigbouring EEZs.
-Grey dotted boxes represent each EEZ boundaries, t1 and t2 represent time periods one and two, respectiveley, and green and red lines represent an increase and decrease in D, respectiveley](~/GitHub/TransEmergence/Figures/Fig1.png)
+We performed a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
+Specifically, each species within an EEZ was classified based on the habitat preference [@REF] and catch trend [@FishForVisa].
+Moreover, each EEZ was classified based on the continental [@REFUN] and biome [@Gabs] location.
+Finally, we also include the ToE of both bottom and sea surface temperature.
+We hypothesize that all of these factors would have significant influence in the time of emergence of transboundary stocks.
+Finally, we ran a series of ANOVAS to test continental and habitat preference ToE differences betwwn transboundary species.
 
@@
 
-## Changes in Stock Share Ratio
+## Changes in Stock Share Ratio (SSR)
 
@@
 
-We estimate the changes from today (1951-2005) to the early (2020-2040) and mid $21^{st}$ century (2040-2060) in transboundary Stock Share Ratio between neighbouring countries.
+We estimated changes in the Stock Share Ratio of transboundary species shared by neighbouring EEZs by early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
 Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
 1A).
-While most EEZs will see and average species change lower than 20%, a few EEZs are expected to have 50% change in SSR (Fig.
-2A).
-For example.
-Guatemala' SSR of *species* shared with Mexico is projected to go from 12% to 26% while Honduras SSR of *species* shared with the Nicaragua in the Pacific coast is projected to decline from 50% to 14%.
-Regions where EEZs are particularly small and/or have multiple borders (e.g.
-The Caribeean) will see a complex dynamic of changes in SSR (Fig.
+While most EEZs will experience an average change > 20%, a few EEZs are expected to have larger changes in SSR (Fig.
+1A).
+For example, Guatemala' average SSR of *species* shared with Mexico is projected to go from 12% to 26% while Honduras SSR of *species* shared with the Nicaragua in the Pacific coast is projected to decline from 50% to 14%.
+The number of stocks per EEZ that will see a change in SSR by early $21^{st}$ century is relativelly low (*XX%*), however, some regions like West Africa and the Mediterranean will see changes in over 50 species.
+These regions, where EEZs are relativelly small and have multiple borders will see a complex dynamic of changes in SSR with no clear patterns (Fig.
 2B).
-Towards the mid $21^st$ century, changes in SSR per EEZ will no be substantial (**Supplements Fig 1 (map)**) but are still are expected to increase in intensity (**Supplements Fig 2 (histogram)**).
+Towards the mid $21^st$ century, changes in SSR per EEZ will no be substantial (**Supplements Fig 1 (map)**) relative to early century, but are expected to increase in intensity (**Supplements Fig 2 (histogram)**).
 
 ```{r SSR_gain_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Average gain of transboundary stock share ratio (lines) and number of shifting stocks (polygon) by 2030 (2021-2040) relative to 1951-2005.
-Arrows go from the losing EEZ (point) to the winner.
-(arrowhead)"}
+Arrows go from the losing EEZ (point) to the winner (arrowhead).
+Pannel B zooms in to specific areas"}
 
@@
 ![Average gain of transboundary stock share ratio (lines) and number of shifting stocks (polygon) by 2030 (2021-2040) relative to 1951-2005.
-Arrows go from the losing EEZ (point) to the winner.
-(arrowhead)](../Figures/Fig2.png)
+Arrows go from the losing EEZ (point) to the winner (arrowhead).
+Pannel B zooms in to specific areas ](../Figures/Fig1.png)
 
-## Time of Emergence
+## Time of Emergence (ToE)
 
 We estimated the time of emergence between neighbouring EEZs for 633 transbboundary stocks.
-On average, most of the world EEZ's time of emergence will be 2030 $\pm$ 25 yers (Fig S3-A), however, all EEZs have already seen changes for at least one species (Fig S3-B).
+Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least one neighbouring EEZ (e.g.
+Peruvean anchoveta emerged in 2010 between Chile and Peru).
+The cumulative sum of emerging stocks will have a steeper increase untill 2040 where the increment wil be slightly less intense.
+Similarly, for the number of EEZs, the increment curve seem to flattern by 2020.
+By 2081, all of the EEZs analyzed here will have at least one emerging stock while 2097 will be the year where all 633 stocks have emerged in at least one neighbouring EEZ (Fig.
+2).
+
+```{r SSR_gain_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Cumulative increase in time of emergence by EEZ and stock"}
+
+img <- readPNG("../Figures/Fig2.png")
+grid.raster(img)
+
+```
+
+![Cumulative increase in time of emergence by EEZ and stock](../Figures/Fig2.png)
+
+On average, most of the world EEZ's will see a stock ToE by 2030 with a higher frfequency of number of emerging stocks between latitudes 0$^o$ and 50$^o$N (Fig 2).
+In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50%, however, in some cases it represents 85% of total fishing revenue (Peru) or more, like the case of island territories such as Marshall Islands (92%) or Tokelau- New Zealand (100%).
+In few cases like Brazil's EEZ and Alaska's polar EEZ, no stock will emerge.
 
@@
 
+The regional patterns of ToE suggest that most stocks will emerge by early 21st with many stocks in the Americas and Oceania already emerging.
+For the larger fisheries of the world in terms of revenue representig 75 percentile, ony the USA will see an average ToE before 2030 while fisheries in all other top fishing countries will be emerging within the following 10 years.
+
+```{r Toe_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}
+
+img <- readPNG("../Figures/Fig4.png")
+grid.raster(img)
+
+```
+
+![(Average time of emergence (ToE) per fishing entity.
+Fishing revenue in logarithmic sacle.
+Showing names for the top 75th percentile).
+Countries marked with a asterisk have at least one border](../Figures/Fig4.png)
+
+The mean ToE of transbounudary species varied significantly according to the geographic region of the neighbouring EEZs (one-way ANOVA, *DF*(15,5478) = 18.64, *p* < 0.001, $\alpha$ = 0.05) (Fig.
+5).
+Specifically, for species in EEZs of Easter Asia and Wetern Europe who have the latest average ToE across all habitat types (2041) and Polynesia who has the earliest everage ToE across all habitats (2018).
+Other regions like Wester Europe (regional average 2035) and Latin America and the Caribbean (regional average 2021) also significantly differe from other regions (*See Table Sup X for full statistics*).
+When looking specifically at habitat association, we also found significant differences on ToE (one-way ANOVA; *DF*(7,5041) = 12.68, *p* < 0.001, $\alpha$ = 0.05).
+Specifically, the ToE of pelagic-oceanic species is significantly different (*p* < 0.001) to that of bathydemersal, benthopelagic, demersal, and pelagic-neritic while the ToE of reef associated species is significantly different (*p* < 0.001) than bathydemersal, bethopelagic and demersal species (*See Table Sup XX for full statistics*).
+
+```{r Toe_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}
+
+img <- readPNG("../Figures/Fig5.png")
+grid.raster(img)
+
+```
+
+![Regional time of emergence per habitat preference.
+Regions acording to the United Nations and habitat preference to FishBase.
+Whiskers represent 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Line represents median, and black points represent outliers (outside of 1.5*IQR.)](../Figures/Fig5.png)
+
+We hypothesize that species environmental preferences (e.g.
+presence in sertain biomes and habitat), climate change effects on water temperature, human drivers (e.g.
+catch rate), and latitudinal gradient would have significant influence in the time of emergence of transboundary stocks (Table 1).
+Results from the multiple linear regression analysis show a significant relation between ToE of transbounuadry stocks and latitudinal gradient (*p* > 0.001), **biome2** (*p* = 0.006) and speicies habitat association (Table 1).
+
 # Discussion
+
+The combination of the time of emergence and the revenue participation of the species could shed a light in the importance of the time of emergence in terms of management.
+For example, we identified that in Peru 12 species will emerge by 2038 in average, representing 92% of the countriy's fishing value.
+Morevoer, our results show that Peruvean anchoveta, the second largest fishery of the world [@SOFIA] and the main fishery in Peru [@REF] have already emerged (2010), with its stock share ratio benefititng Chile (Fig.
+1).
+
+Form all siginificantly differnt variables, only latitudinal gradient postpone the ToE suggesting that the ToE will happen first at equatorian latitudes and later towards the poles.

### 2020-09-01 Scripts/Initial_Analysis.Rmd
Fixed Gabs coordinate system and selected LM

@@
   "stargazer", # For nice lm tables
-  "tidytext" # for boxplot order
+  "tidytext", # for boxplot order
+  "moments", # for skewness and kurto
+  "pgirmess" # for non parametric post hoc test
 )
@@
 Biogeo_reg <-fread(paste(my_path("D"),"Climate/SPATIAL_PARTITION_OCEAN.csv", sep ="")) %>% 
-  clean_names() %>% 
-  rowid_to_column("index")
+  clean_names()
+
+# Fix coordinate system and re-name biomes
+Biogeo_reg <- suppressWarnings(Biogeo_reg[order(Biogeo_reg$lat, rev(Biogeo_reg$lon),decreasing=TRUE), ]) %>% 
+  mutate(index = seq(1,259200,1)) %>% 
+  mutate(
+    # biome = ifelse(biome2 == 1, "c_tropical",
+    #                ifelse(biome2 == 2,"c_temperate",
+    #                       ifelse(biome2 == 3, "c_polar",
+    #                              ifelse(biome2 == 4, "o_tropical",
+    #                                     ifelse(biome2 == 5, "o_temperate",
+    #                                            ifelse(biome2 == 6, "o_polar",
+    #                                                   biome2
+    #                                            )
+    #                                     )
+    #                              ) 
+    #                       )
+    #                )
+    # ),
+    biome = ifelse(biome2 %in% c(1,4), "tropical",
+                   ifelse(biome2  %in% c(2,5),"temperate",
+                          ifelse(biome2  %in% c(3,6), "polar",
+                                 biome2)
+                   )
+    )
+    ) %>% 
+  filter(biome2 != "NaN")
+
+# biome2 1 = Coastal tropical 
+# biome2 2 = Coastal sub-tropical
+# biome2 3 = Coastal polar
+# biome2 4 = Oceanic tropical
+# biome2 5 = Oceanic sub-tropical
+# biome2 6 = Oceanic polar
+
+# Biogeo_reg %>%
+  # gather("biome_type","value",biome:biome2) %>%
+# ggplot() +
+# geom_tile(
+#   aes(
+#     x = lon,
+#     y = lat,
+#     fill = biome
+#   )
+# )
+#   facet_wrap(~biome_type,
+#              ncol=1)
 
@@
 Eez_biogeo <- EEZ_CellID %>% 
-  # filter(eezid == 191) %>% 
+  # filter(eezid == 8) %>%
   left_join(Biogeo_reg,
@@
 # Per Stock
-stock_d <-  Eme_Res %>% 
+stock_da <-  Eme_Res %>% 
   left_join(Neighbours) %>% 
@@
 
-#### Fig. 5 Demersal Vs Pelagic
+#### Fig. 5 Boxplot
 
@@
 
+#### Test for normallity
+
+```{r, normality_test, eval = F, echo = F}
+# Test for normality
+
+Stats_data <- Eme_Res %>% 
+  filter(tresh == "Tresh_one",
+         n_ens == 10)
+
+# ------------------- #
+# Test normal asumptions 
+# ------------------- #
+
+hist(Stats_data$emerging_yr) # Positevely skewed
+qqnorm(Stats_data$emerging_yr) # not normally distributed
+skewness(Stats_data$emerging_yr) #1.02
+kurtosis(Stats_data$emerging_yr) #2.82
+
+# ------------------- #
+
+# (NOT NORMAL DISTRIRBUTION)
+
+```
+
 ## Regional differences
@@
 
-reg_diff_data <- Eme_Res %>% 
+Reg_diff_data <- Eme_Res %>% 
   filter(tresh == "Tresh_one",
@@
 
-Reg_aov <- aov(emerging_yr~un_sub_region, data = reg_diff_data)
+# ------------------- #
+# Run non-parametric Krustal
+# ------------------- #
+
+Reg_kw <- kruskal.test(un_sub_region~emerging_yr, data = Reg_diff_data)
+
+Reg_kw
+# Kruskal-Wallis rank sum test
+# 
+# data:  un_sub_region by emerging_yr
+# Kruskal-Wallis chi-squared = 287.23, df = 93, p-value < 2.2e-16
+
+kruskalmc(emerging_yr~un_sub_region, data = Reg_diff_data)
+
+# Multiple differences but Eastern Asia, Lat. Ame. Car. and Polinesia are the nes who differe the most (e.g. differe with more revions).
+
+# ------------------- #
+# Just to double check ANOVA results (give the same)
+# ------------------- #
+Reg_aov <- aov(emerging_yr~un_sub_region, data = Reg_diff_data)
 Reg_aov
@@
 
-hab_diff_data <- Eme_Res %>% 
+Hab_diff_data <- Eme_Res %>% 
   filter(tresh == "Tresh_one",
@@
             by = "taxon_key")
+
+# ------------------- #
+# Run non-parametric Krustal
+# ------------------- #
+
+Hab_kw <- kruskal.test(DemersPelag~emerging_yr, data = Hab_diff_data)
+
+Hab_kw
+# Kruskal-Wallis rank sum test
+
+# data:  DemersPelag by emerging_yr
+# Kruskal-Wallis chi-squared = 286.48, df = 93, p-value < 2.2e-16
+
+kruskalmc(emerging_yr~DemersPelag, data = Hab_diff_data)
+
+# Multiple differences but main differences for pelagic-ocenaic and reefs
+
+# ------------------- #
+# Just to double check ANOVA results (give the same)
+# ------------------- #
 
@@
 
-### Tab.1 Explain the ToE
+## Table.1 Multy linear model
 
@@
   left_join(Spp_depth) %>% 
-  select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel) %>% 
+  select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel,un_region= un_geo_region) %>% 
   left_join(Toe_temp,
@@
 # Model 1; all variables
-M_all_var <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)
-
-all_sum <- summary(M_all_var)
-all_aic <- AIC(M_all_var)
-all_r <-  all_sum$adj.r.squared
+M1 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)
+
+M1_sum <- summary(M1)
+M1_aic <- AIC(M1)
+M1_r <-  M1_sum$adj.r.squared
 
@@
 
+# Model 9; Include UN regions
+M9 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + DemersPelag + latitude + un_region, data = Lm_data)
+
+M9_sum <- summary(M9)
+M9_aic <- AIC(M9)
+M9_r <-  M9_sum$adj.r.squared
+
 tibble(
-  "model" = paste("M",seq(1,8,1),sep=""),
-  "AIC" = c(all_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic),
-  "rsquared" = c(all_r,M2_r,M3_r,M4_r,M5_r,M6_r,M7_r,M8_r)
-)
-
-# UNIDAS 20282659
-
-# The Winner is M8!
-
-stargazer(M8, title = "Results", align=TRUE)
-
+  "model" = paste("M",seq(1,9,1),sep=""),
+  "AIC" = c(M1_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic, M9_aic),
+  "rsquared" = c(M1_r,M2_r,M3_r,M4_r,M5_r,M6_r,M7_r,M8_r,M9_r)
+)
+
+# The Winner is M9    49283.   0.0765
+
+#Table for paper
+M9_sum %>% 
+  tidy() %>% 
+  filter(p.value < 0.05) %>% 
+  mutate(p.value = ifelse(p.value < 0.001,0.001,round(p.value,3))) %>% 
+  # stargazer(title = "Multiple regression results for time of emergence of Transboundary stocks.
+Only showing explanatory variables.
+The model significantly predicts time of emergencec (F(31,5270) = 15.38, p < 0.001, R2 = 0.06).",
+            # align=TRUE)
+  # knitr::kable() %>% 
+  write_csv(.,
+            "./Tables/mlr_result.csv")
+
+#Table for supplements
+M9_sum %>% 
+  tidy() %>% 
+  # filter(p.value < 0.05) %>% 
+  # mutate(p.value = ifelse(p.value < 0.001,0.001,round(p.value,3))) %>% 
+  # stargazer(title = "Multiple regression results for time of emergence of Transboundary stocks.
+Only showing explanatory variables.
+The model significantly predicts time of emergencec (F(31,5270) = 15.38, p < 0.001, R2 = 0.06).",
+            # align=TRUE)
+  # knitr::kable() %>% 
+  write_csv(.,
+            "./Tables/mlr_result_sup.csv")
 ```
 
-## Troubleshooting
+# END PAPER
+
+# Troubleshooting
 

### 2020-09-01 Scripts/Initial_Analysis.Rmd
Analysis done

@@
   mutate(
-    # biome = ifelse(biome2 == 1, "c_tropical",
-    #                ifelse(biome2 == 2,"c_temperate",
-    #                       ifelse(biome2 == 3, "c_polar",
-    #                              ifelse(biome2 == 4, "o_tropical",
-    #                                     ifelse(biome2 == 5, "o_temperate",
-    #                                            ifelse(biome2 == 6, "o_polar",
-    #                                                   biome2
-    #                                            )
-    #                                     )
-    #                              ) 
-    #                       )
-    #                )
-    # ),
     biome = ifelse(biome2 %in% c(1,4), "tropical",
@@
 Only showing explanatory variables.
-The model significantly predicts time of emergencec (F(31,5270) = 15.38, p < 0.001, R2 = 0.06).",
+The model significantly predicts time of emergencec (F(18,6155) = 20.68, p < 0.001, R2 = 0.05).",
             # align=TRUE)

### 2020-09-01 Scripts/Manuscript.Rmd
Methods completed for first draft

@@
 Recent events suggests that even changes in species that contribute a relatively small proportion of the overall catch can lead to important international conflict when no legal framework is available.
-Thus, it is fundamental to prepare for cliamte change-driven impacts such as identifying the stocks that are going to change historical proportions within neighbouring EEZs, the degree by which they will cahnge, and the time when such change is going to happen.
-Only by adapting management plans and polocy that antecipates to these effects will international fisheries treaties be climate-ready.
+Thus, it is fundamental to prepare for climate change-driven impacts such as identifying the stocks that are going to change historical proportions within neighboring EEZs, the degree by which they will change, and the time when such change is going to happen.
+Only by adapting management plans and policy that anticipates to these effects will international fisheries treaties be climate-ready.
 Here, we build on previous work to identify the transboundary stocks that will see changes in their stock-share-proportion (e.g.
@@
 Our results suggest that by the mid of the 21st century EEZs will see changes in stock-share ratio between 1 and 10% with some exceeding 25%.
-Most of these changes will hapen before 2050 urging policy to be re-formulated to account for the possible implications of these changes.
-By the end of the 21st century virtually all EEZs and transboundary sepeices will have experienced changes in stock-share-proportion relative to historical distributions.
+Most of these changes will happen before 2050 urging policy to be re-formulated to account for the possible implications of these changes.
+By the end of the 21st century virtually all EEZs and transboundary species will have experienced changes in stock-share-proportion relative to historical distributions.
 The results of this research has the potential to improve climate-ready fisheries if policy is modify to respond to the consequences presented here.
@@
 
-<!-- The world is moving towards sustainable development and ocean-based activities plays a mayor role in achieveing international policies. -->
+The world is moving towards sustainable development and ocean-based activities plays a mayor role in achieving international policies.
 
-<!-- One of the main goals of the United Nations Sustainable Development Goal No.14 (SDG 14-Life below water) is to achieve sustainable development of fisheries [@SDG14]. -->
+One of the main goals of the United Nations Sustainable Development Goal No.14 (SDG 14-Life below water) is to achieve sustainable development of fisheries [@SDG14].
 <!-- (https://sustainabledevelopment.un.org/sdg14). -->
 
-<!-- Climate change is expected to reduce the production from global fisheries in about XX% to xx% [@Cheung; @Barange] compromisisng our capacity to reach SDG 14. -->
+Climate change is expected to reduce the production from global fisheries in about XX% to xx% [@Cheung; @Barange] compromising our capacity to reach SDG 14.
 
-<!-- Recent reseach show that the impacts of climate change on the productivity of marine wild fisheries can be offset with the proper reform of current fisheries management [@Gaines; @CheungClimateReadyFish].
-While bringing fisheries that are poorly managed to a sustainable level will likeley offset the negative impacts of climate change to fisheries production [@Gaines], it will not necessarilly solve the spatial consequences of species moving from one mangement area to another.
--->
+Recent research show that the impacts of climate change on the productivity of marine wild fisheries can be offset with the proper reform of current fisheries management [@Gaines; @CheungClimateReadyFish].
+While bringing fisheries that are poorly managed to a sustainable level will likely offset the negative impacts of climate change to fisheries production [@Gaines], it will not necessarily solve the spatial consequences of species moving from one management area to another.
 
-<!-- Shifts in species distribution does not necessarilly mean a reduction in production (e.g. reduction in stock biomass) but rather a spatial re-distribution of the stock's biomass [@RefLiam?].  -->
+Shifts in species distribution does not necessarily mean a reduction in production (e.g. reduction in stock biomass) but rather a spatial re-distribution of the stock's biomass [@RefLiam?].
 
-<!-- Documented cases of stock shifts have resulted in international conflict and usustainable fishing rates by parties involved in the conflict. -->
+Documented cases of stock shifts have resulted in international conflict and sustainable fishing rates by parties involved in the conflict.
 
-<!-- This is of specific concern to a set of 633 fished species that cross international borders, also known as transbounday species [@FishForVisa].  -->
+This is of specific concern to a set of 633 fished species that cross international borders, also known as transponder species [@FishForVisa].
 
-<!-- With the implementation of arbritary spatial management boundaries, at least 633 species became transboudary representing an average catch (2005 - 2010) of 47.6 million tons and generating USD 65,483 million in fishing revenue [@FishorVisa].
--->
+With the implementation of arbitrary spatial management boundaries, at least 633 species became transboundary representing an average catch (2005 - 2010) of 47.6 million tons and generating USD 65,483 million in fishing revenue [@FishorVisa].
 
-<!-- Transboudary fish stocks refere to those species that cross international jurisdictions known as Economic Exclusive Zones (EEZs) of neighbouring nations [@FAO].
-These stocks are responsable for an average catch (2005 - 2010) of 47.6 million tons generating USD 65,483 million in fishing revenue [@FishorVisa].
--->
+Transboundary fish stocks refer to those species that cross international jurisdictions known as Economic Exclusive Zones (EEZs) of neighboring nations [@FAO].
+These stocks are responsible for an average catch (2005 - 2010) of 47.6 million tons generating USD 65,483 million in fishing revenue [@FishorVisa].
 
-<!-- Historically, environmentally-driven shits in shared stocks have cause problems and threathen stock's sustainabillity.
+Historically, environmentally-driven shits in shared stocks have cause problems and threaten stock's sustainability.
 **In early 1980's Canada and the US blablabla.** Many fish stocks managed by the European Union ICES regultaroy areas have been shifting since the 1970's [@EuropeShift].
@@
 This entering and exicting national jurisdicions will complicate fisheries management potentially creating confllict between neighbouring nations [@Pinksy @Oremus:2020id].
--->
 
-<!--  So, preparing policy to species on the move is key to achieve sustainable development goals and proper governance of the world fisheries [@Pinksy]. -->
+ So, preparing policy to species on the move is key to achieve sustainable development goals and proper governance of the world fisheries [@Pinksy].
 
-<!-- Undrstanding the degree by which shared stocks are going to change and the time of emergence is key for creating climate-ready fisheries policy. -->
+Undrstanding the degree by which shared stocks are going to change and the time of emergence is key for creating climate-ready fisheries policy.
 
-<!-- <!-- The concepts of transboundary stocks and EEZs were introduced in the early 1980's by the United Nations Law of the Sea (UCLOS) [@UnitedNations:1986tl].
+The concepts of transboundary stocks and EEZs were introduced in the early 1980's by the United Nations Law of the Sea (UCLOS) [@UnitedNations:1986tl].
 Under UNCLOS each costal nation has the right (and responsabillity) to manage wild fish stocks up to 200 nautical mailes of the coast.
@@
 Climate change is shifting the distribution of marine species to the extreem that some spatial policies implemented to manage fisheries are becoming obsolete [@Pololanza; @Jessica].
---> -->
 
@@
 
+## Databases and species selection
+
+The Food and Agricultural Organization (FAO) of the UN define transboundary stocks as those that freely swim across the EEZ of neighboring nations [@REFFAO].
+Accordingly, we included in this analysis 633 marine exploited transboundary species representing 80% of the catch taken from the world's EEZs [@Palacios-Abrantes:xxxx].
+We consider each species as a single stock unit within an EEZ due to a lack of more detailed spatial information [@Teh:2015gd; @Palacios-Abrantes:xxxx].
+We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the *Sea Around Us* shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic) including islands territories, and determined the intersections between polygons using the R package *sf* (**Supplements Fig 1 (SuppM)**) [@PackagesfSimpleF:2018vp].
+For each species and EEZ we used the *Sea Around Us* catch data to estimated the average catch and fishing revenue (USD 2019).
+We only used data from fishing activities within the global EEZs reported as the average from the last decade (2005-2014) of available spatial data [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+Moreover, we catalogue catch trends as increasing, constant and decreasing, following Palacios-Abrates *et al.* (*hopefully in press*).
+Finally, each EEZ was attributed to a geopolitical region [@UN] and biome [@Gabriel], and species were attributed a habitat preference [@FishBase].
+For the present study we first projected the 633 species distributions under climate change and then estimated two main metrics; a stock's share ratio (SSR) and time of emergence (ToE) as described in the following sections.
+
 ## Projecting species distributions under climate change
 
-We projected the distribution of marine species using a dynamic bioclimatic envelope model (hearafter called DBEM).
-The DBEM algorythm includes species physiology, habitat suitabillity, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
+We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM).
+The DBEM algorithm includes species physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
 For pelagic species, the model uses environmental variables at the surface (e.g.
@@
 We coupled the DBEM with the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project species distribution from 1951 to 2100 under climate change [@Dunne:2012bo; @Dunne:2013db].
-We run the DBEM simulating biomass and MCP from 1950 to 2100 under a “worst-case” climate change scenario accroding to the Representative Concentration Pathway 8.5 (RCP 8.5) used by the Intergovernmental Panel on Climate Change (IPCC) [@IPCC:2014bz; Riahi:2011dk].
+We run the DBEM simulating biomass and MCP from 1950 to 2100 under a “worst-case” climate change scenario according to the Representative Concentration Pathway 8.5 (RCP 8.5) used by the Intergovernmental Panel on Climate Change (IPCC) [@IPCC:2014bz; Riahi:2011dk].
 According to the GFDL-ESM2M, under the RCP 8.5 scenario atmospheric temperature will increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a species distribution, we have to understand both the historical as well as the future climate-driven distribution variabillity, in order to infer the differences.
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a species distribution, we have to understand both the historical as well as the future climate-driven distribution variability, in order to infer the differences.
 We do this by following an ensemble member approach where each member constitutes a different initial Earth system condition [@Rodggers:2015jq; @Frolicher:2009ge].
-Thus, for our experiment, in each of the ten GFDL-ESM2M-ensambles the Earth system initial condition is infinitesimally different resulting in an unique atmosphere and ocean state at each point in time.
-This results in a different state of internal variability for each ensamble representing a plausible future under RCP 8.5 [@Rodggers:2015jq].
-Following this approach allows us to estimate a secular trend for the natural variabillity (e.g.
-backround noise).
-At the same time, by having multiple initial-condition ensemble simulations of the GFDL-ESM2M we were able to remove the natural variability of a stock's distributio by averaging the secular trend over all ensemble members [@Rodggers:2015jq].
+Thus, for our experiment, in each of the ten GFDL-ESM2M-ensembles the Earth system initial condition is infinitesimally different resulting in an unique atmosphere and ocean state at each point in time.
+This results in a different state of internal variability for each ensemble representing a plausible future under RCP 8.5 [@Rodggers:2015jq].
+Following this approach allows us to estimate a secular trend for the natural variability (e.g.
+background noise).
+At the same time, by having multiple initial-condition ensemble simulations of the GFDL-ESM2M we were able to remove the natural variability of a stock's distribution by averaging the secular trend over all ensemble members [@Rodggers:2015jq].
 
-## Data analysis
-
-The Food and Agricultural Organization (FAO) of the UN define transboundary stocks as those that freely swim across the EEZ of neighbouring nations [@REFFAO].
-Accordingly, we included in this analysis 633 marine exploited transboundary species representing 80% of the catch taken from the world's EEZs [@Palacios-Abrantes:xxxx].
-We consider each species as a single stock unit within an EEZ due to lack of more detailed information [@Teh:2015gd; @Palacios-Abrantes:xxxx].
-We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic) including islands territories, and determined the intersections between polygons using the R package *sf* (**Supplements Fig 1 (SuppM)**) [@PackagesfSimpleF:2018vp].
-
-### Changes in Stocks Share Ratio
+## Changes in Stocks Share Ratio
 
 In game theory, a threat point is known as the minimum payoff that a player is willing to receive in order to cooperate with other players.
-Thus, if a player's payoff is lower than the theat point, such player will liekly break off any collaboration agreement and proceed to act solo [@Sumaila:E&S].
-Here, we adopted the concept of threat point as the minimum proportion of the catch that each country will acept if they were to cooperate in the management of a transboundary stock [@Palacios-Abrantes:E&S] hereafter referred as stock share ratio (SSR).
-For each stock and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighbouring EEZs, and then calculated the proportion that each EEZ had [@Palacios-Abrantes:xxxx].
-We then average the results in three time periods in order to reduce climate variabillity within the projections.
-The first time period ($t_h$) comprises (1951 - 2005) and is assumed as the historic distribution of each species, thus representing each neighbouring nation's threat point.
+Thus, if a player's payoff is lower than the thread point, such player will likely break off any collaboration agreement and proceed to act solo [@Sumaila:E&S].
+Here, we adopted the concept of threat point as the minimum proportion of the catch that each country will accept if they were to cooperate in the management of a transboundary stock [@Palacios-Abrantes:E&S] hereafter referred as stock share ratio (SSR).
+For each stock and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighboring EEZs, and then calculated the proportion that each EEZ had [@Palacios-Abrantes:xxxx].
+We then average the results in three time periods in order to reduce climate variability within the projections.
+The first time period ($t_h$) comprises (1951 - 2005) and is assumed as the historic distribution of each species, thus representing each neighboring nation's threat point.
 This time frame matches the historic climate data that the GFDL uses to project climate change [@Dunne:2012bo; @Dunne:2013db].
@@
 We choose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing) and SDG 1.2 (poverty reduction), aimed to be achieved by 2020 and 2030, respectively [@Singh:2017ds].
-We then estimate the percentage change in sthe SSR of each future time period relative to the historic time period.
-To address the natural spatal variabillity of the stock we only considered a SSR chgange when the future SSR was higher or lower than the average historic SSR $\pm$ 1 s.d.
-The analysis was repplicated for each of the ten ensamble members capturing variation by averaging results among ensamble members (mean $\pm$ standar deviation).
+We then estimate the percentage change in the SSR of each future time period relative to the historic time period.
+To address the natural spatial variability of the stock we only considered a SSR change when the future SSR was higher or lower than the average historic SSR $\pm$ 1 s.d.
+The analysis was replicated for each of the ten ensemble members capturing variation by averaging results among ensemble members (mean $\pm$ standard deviation).
 
-### Time of Emergence Index
+## Time of Emergence Index
 
 Knowing the time (e.g.
-year) in whitch SSR is going to diverge from a historical variablillty can be key for policy to prepare to such change.
-In here we call this event time of emergence and defined it as the moment in time when the future variation in SSR becomes larger than the historical trend variabillity [@Roggers:2015jq].
-We employed the ten GFDL ensable members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
-The index is based on the variation of a species distribution centroid in realtion to the centroid of the neighbouring EEZs it ocuppys.
+year) in which SSR is going to diverge from a historical variability can be key for policy to prepare to such change.
+In here we call this event time of emergence and defined it as the moment in time when the future variation in SSR becomes larger than the historical trend variability [@Roggers:2015jq].
+We employed the ten GFDL ensemble members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
+The index is based on the variation of a species distribution centroid in relation to the centroid of the neighboring EEZs it occupy.
 Each species centroid was determined by the average latitude and longitude of the grid cells containing the top 10% abundance percentile of each year.
 The EEZ centroid was estimated using the *st* package of *R* statistical software (**Supplements Fig 1 (SuppM)**).
-For each ensamble member, neighbouring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring elevation differences:
+For each ensemble member, neighboring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring elevation differences using the *geosphere* package of *R* statistical software;
 
 $$D_{ens} = acos(sin(lat_{eez})*sin(lat_{spp}) + cos(lat_{eez})*cos(lat_{spp})*cos(lon_{eez}-lon_{spp}))$$
-where $lat_{eez}$ and $lat_{spp}$ are the latitudes of the EEZ and specie centroid, respectiveley and $lon_{eez}$ and $lon_{spp}$ are the respectevly longitudes.
+where $lat_{eez}$ and $lat_{spp}$ are the latitudes of the EEZ and specie centroid, respectively and $lon_{eez}$ and $lon_{spp}$ are the respectively longitudes.
 
@@
 
-where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectiveley for each future time step from 2006 to 2100 (*t*); *sd* is the standard deviation of the historical (*h*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectively for each future time step from 2006 to 2100 (*t*); *sd* is the standard deviation of the historical (*h*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
 The index was then smoothed to a 10 years average.
@@
 
-where *TREND* is the yearly mean *TI* ($\mu$) of all ten ensemble members and *NOISE* is the standar deviation of the TREND ($\sigma$) between 1951 and 2005.
+where *TREND* is the yearly mean *TI* ($\mu$) of all ten ensemble members and *NOISE* is the standard deviation of the TREND ($\sigma$) between 1951 and 2005.
 Note that in some cases not all ensemble members show the presence of a species in a certain EEZ throughout the study's time frame.
@@
 We determine a $\mu \pm \sigma$ (confidence interval of 68%) to show the first year when the "signal" of the transboundary index overpasses the historic "noise".
-Since this was an arbritary decission, we have included a more conservative estimation ($\mu \pm 2*\sigma$ for a 95% confidence interval).
-The SNR for both bottom and surface sea temperature was also estimated following the previos equation.
+Since this was an arbitrary decision, we have included a more conservative estimation ($\mu \pm 2*\sigma$ for a 95% confidence interval).
+The SNR for both bottom and surface sea temperature was also estimated following the previous equation.
 
-<!-- In cases where the species is distributed in more than 2 neighbouring EEZs (e.g. *xxx*), the index is as follows: -->
+<!-- In cases where the species is distributed in more than 2 neighboring EEZs (e.g. *xxx*), the index is as follows: -->
 
@@
 
-In other words, for two neighbouring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases (see *TI* equation).
+In other words, for two neighboring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases (see *TI* equation).
 Consequently, if the centroid shifts are within historical variability (*SNR* < 1), risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability (*SNR* > 1), the risk of dispute is higher.
@@
 
-We performed a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
-Specifically, each species within an EEZ was classified based on the habitat preference [@REF] and catch trend [@FishForVisa].
-Moreover, each EEZ was classified based on the continental [@REFUN] and biome [@Gabs] location.
-Finally, we also include the ToE of both bottom and sea surface temperature.
+## Statistical analysis
+
+We tested data for normallity (e.g.
+skewness, kurtosis) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the time of emergence of transboundary stocks.
+Moreover, we implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
+Specifically, we lookd at habitat preferences, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
 We hypothesize that all of these factors would have significant influence in the time of emergence of transboundary stocks.
-Finally, we ran a series of ANOVAS to test continental and habitat preference ToE differences betwwn transboundary species.
-
-All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *wesanderson* [@Packagewesanderson:2018vn], *rfishbase* [@PackagerfishbaseR:2019th], *R.matlab* [@PackageRmatlabRe:2018un], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@gmt] and *zoo* [@PackagezooSInfr:2019tm].
+All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@gmt] and *zoo* [@PackagezooSInfr:2019tm], *readxl* [@REF], *ggrepel* [@REF], *zeallot* [@REF], *viridis* [@], *cowplot* [@], *moments* [@REF] and *pgirmess* [@REF].
 All code is available at https://github.com/jepa/TransEmergence
@@
 
-We estimated changes in the Stock Share Ratio of transboundary species shared by neighbouring EEZs by early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
+We estimated changes in the Stock Share Ratio of transboundary species shared by neighboring EEZs by early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
 Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
@@
 For example, Guatemala' average SSR of *species* shared with Mexico is projected to go from 12% to 26% while Honduras SSR of *species* shared with the Nicaragua in the Pacific coast is projected to decline from 50% to 14%.
-The number of stocks per EEZ that will see a change in SSR by early $21^{st}$ century is relativelly low (*XX%*), however, some regions like West Africa and the Mediterranean will see changes in over 50 species.
-These regions, where EEZs are relativelly small and have multiple borders will see a complex dynamic of changes in SSR with no clear patterns (Fig.
+The number of stocks per EEZ that will see a change in SSR by early $21^{st}$ century is relatively low (*XX%*), however, some regions like West Africa and the Mediterranean will see changes in over 50 species.
+These regions, where EEZs are relatively small and have multiple borders will see a complex dynamic of changes in SSR with no clear patterns (Fig.
 2B).
@@
 
-We estimated the time of emergence between neighbouring EEZs for 633 transbboundary stocks.
-Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least one neighbouring EEZ (e.g.
+We estimated the time of emergence between neighboring EEZs for 633 transboundary stocks.
+Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least one neighboring EEZ (e.g.
 Peruvean anchoveta emerged in 2010 between Chile and Peru).
-The cumulative sum of emerging stocks will have a steeper increase untill 2040 where the increment wil be slightly less intense.
-Similarly, for the number of EEZs, the increment curve seem to flattern by 2020.
-By 2081, all of the EEZs analyzed here will have at least one emerging stock while 2097 will be the year where all 633 stocks have emerged in at least one neighbouring EEZ (Fig.
+The cumulative sum of emerging stocks will have a steeper increase until 2040 where the increment will be slightly less intense.
+Similarly, for the number of EEZs, the increment curve seem to flatten by 2020.
+By 2081, all of the EEZs analyzed here will have at least one emerging stock while 2097 will be the year where all 633 stocks have emerged in at least one neighboring EEZ (Fig.
 2).
@@
 
-On average, most of the world EEZ's will see a stock ToE by 2030 with a higher frfequency of number of emerging stocks between latitudes 0$^o$ and 50$^o$N (Fig 2).
+On average, most of the world EEZ's will see a stock ToE by 2030 with a higher frequency of number of emerging stocks between latitudes 0$^o$ and 50$^o$N (Fig 2).
 In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50%, however, in some cases it represents 85% of total fishing revenue (Peru) or more, like the case of island territories such as Marshall Islands (92%) or Tokelau- New Zealand (100%).
@@
 The regional patterns of ToE suggest that most stocks will emerge by early 21st with many stocks in the Americas and Oceania already emerging.
-For the larger fisheries of the world in terms of revenue representig 75 percentile, ony the USA will see an average ToE before 2030 while fisheries in all other top fishing countries will be emerging within the following 10 years.
+For the larger fisheries of the world in terms of revenue representing 75 percentile, only the USA will see an average ToE before 2030 while fisheries in all other top fishing countries will be emerging within the following 10 years.
 
@@
 
-The mean ToE of transbounudary species varied significantly according to the geographic region of the neighbouring EEZs (one-way ANOVA, *DF*(15,5478) = 18.64, *p* < 0.001, $\alpha$ = 0.05) (Fig.
+The mean ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001) (Fig.
 5).
-Specifically, for species in EEZs of Easter Asia and Wetern Europe who have the latest average ToE across all habitat types (2041) and Polynesia who has the earliest everage ToE across all habitats (2018).
-Other regions like Wester Europe (regional average 2035) and Latin America and the Caribbean (regional average 2021) also significantly differe from other regions (*See Table Sup X for full statistics*).
-When looking specifically at habitat association, we also found significant differences on ToE (one-way ANOVA; *DF*(7,5041) = 12.68, *p* < 0.001, $\alpha$ = 0.05).
-Specifically, the ToE of pelagic-oceanic species is significantly different (*p* < 0.001) to that of bathydemersal, benthopelagic, demersal, and pelagic-neritic while the ToE of reef associated species is significantly different (*p* < 0.001) than bathydemersal, bethopelagic and demersal species (*See Table Sup XX for full statistics*).
+Post-hoc tests revealed that EEZs in Easter Asia who have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ with almost all other regions.
+Other regions also significantly differ from each other but in a less frequent manner (*See Table Sup X for full statistics*).
+When looking specifically at habitat association, we also found significant differences on ToE (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
+Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats is significantly different than other for other habitat preferences (*See Table Sup XX for full statistics*).
 
@@
 We hypothesize that species environmental preferences (e.g.
-presence in sertain biomes and habitat), climate change effects on water temperature, human drivers (e.g.
-catch rate), and latitudinal gradient would have significant influence in the time of emergence of transboundary stocks (Table 1).
-Results from the multiple linear regression analysis show a significant relation between ToE of transbounuadry stocks and latitudinal gradient (*p* > 0.001), **biome2** (*p* = 0.006) and speicies habitat association (Table 1).
+presence in certain biomes and habitat preference), climate change effects on water temperature (e.g.
+ToE of SST), human drivers (e.g.
+catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary stocks (*Table Sup 3*).
+Results from the multiple linear regression analysis show a significant relation between ToE of transboundary stocks and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (Table 1).
+
+Table 2.
+Multiple regression results for time of emergence of transboundary stocks.
+Only showing significant variables.
+The model significantly predicts time of emergence (F[18,6155] = 20.68, p < 0.001, R2 = 0.05).
+
+```{r table1, eval = T,  echo = F"}
+
+read_csv("../Tables/mlr_result.csv") %>% 
+  knirt::kable()
+
+```
 
@@
 The combination of the time of emergence and the revenue participation of the species could shed a light in the importance of the time of emergence in terms of management.
-For example, we identified that in Peru 12 species will emerge by 2038 in average, representing 92% of the countriy's fishing value.
-Morevoer, our results show that Peruvean anchoveta, the second largest fishery of the world [@SOFIA] and the main fishery in Peru [@REF] have already emerged (2010), with its stock share ratio benefititng Chile (Fig.
+For example, we identified that in Peru 12 species will emerge by 2038 in average, representing 92% of the country's fishing value.
+Moreover, our results show that Peruvean anchoveta, the second largest fishery of the world [@SOFIA] and the main fishery in Peru [@REF] have already emerged (2010), with its stock share ratio benefiting Chile (Fig.
 1).
 
-Form all siginificantly differnt variables, only latitudinal gradient postpone the ToE suggesting that the ToE will happen first at equatorian latitudes and later towards the poles.
+Form all significantly different variables, only latitudinal gradient postpone the ToE suggesting that the ToE will happen first at equatorial latitudes and later towards the poles.

### 2020-09-01 Scripts/Manuscript.Rmd
Completed draft of results and methods

@@
 ---
-title: "Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change"
-subtitle: "Climate-ready transboundary fisheries management under Climate Change"
-author: "Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, U. Rashid Sumaila^1^,  Vicky Lam^1^, David VanderZwaag^3^, and William W.L. Cheung^1^"
-date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
-^2^Climate and Environmental Physics, Physics Institute, University of Bern.
-^3^Schulich School of Law, Dalhousie University"
-# output: pdf_document
-output: word_document
+title: Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for
+  Policy Response Under Climate Change
+author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, U. Rashid Sumaila^1^,  Vicky
+  Lam^1^, David VanderZwaag^3^, and William W.L. Cheung^1^'
+date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British
+  Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute,
+  University of Bern. ^3^Schulich School of Law, Dalhousie University'
+output:
+  word_document: default
+  pdf_document: default
+subtitle: Climate-ready transboundary fisheries management under Climate Change
 ---
@@
 
-```{r Load_Results, eval = TRUE,  echo = FALSE, include = FALSE}
+```{r analysis_results, eval = T,  echo = FALSE, include = FALSE}
 
@@
 
-We estimated changes in the Stock Share Ratio of transboundary species shared by neighboring EEZs by early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
+We estimated changes in the Stock Share Ratio of transboundary species shared by neighboring EEZs by early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
 Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
 1A).
-While most EEZs will experience an average change > 20%, a few EEZs are expected to have larger changes in SSR (Fig.
-1A).
-For example, Guatemala' average SSR of *species* shared with Mexico is projected to go from 12% to 26% while Honduras SSR of *species* shared with the Nicaragua in the Pacific coast is projected to decline from 50% to 14%.
-The number of stocks per EEZ that will see a change in SSR by early $21^{st}$ century is relatively low (*XX%*), however, some regions like West Africa and the Mediterranean will see changes in over 50 species.
+While most EEZs will experience an average change lower than 25%, a few EEZs are expected to have larger changes in SSR relative to early century.
+Such change is expected to slightly increase in intensity towards the  mid $21^{st}$ century (**Supplements Fig 1 (histogram)**).
+For example, Guatemala's SSR of *Panulirus gracilis* shared with Mexico is projected to go from 12% to 26%.
+The number of stocks per EEZ that will see a change in SSR by early $21^{st}$ century is relatively low (median of 13 spp per EEZ), however, some regions like West Africa and the Mediterranean will see changes in over 50 species (e.g.
+Senegal, 94).
 These regions, where EEZs are relatively small and have multiple borders will see a complex dynamic of changes in SSR with no clear patterns (Fig.
-2B).
-Towards the mid $21^st$ century, changes in SSR per EEZ will no be substantial (**Supplements Fig 1 (map)**) relative to early century, but are expected to increase in intensity (**Supplements Fig 2 (histogram)**).
+1B).
+Towards the mid $21^st$ century, changes in SSR per EEZ will no be substantial (**Supplements Fig 1 (map)**) relative to early century, but are expected to increase in intensity (**Supplements Fig 1 (histogram)**).
 
@@
 
-## Time of Emergence (ToE)
+## Time of Emergence
 
-We estimated the time of emergence between neighboring EEZs for 633 transboundary stocks.
-Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least one neighboring EEZ (e.g.
-Peruvean anchoveta emerged in 2010 between Chile and Peru).
-The cumulative sum of emerging stocks will have a steeper increase until 2040 where the increment will be slightly less intense.
+We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock's distrtibution overpasses the historical variation.
+Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g.
+*Engraulis ringens* shared by Chile and Peru emerged in 2010).
+The cumulative sum of emerging stocks will have a steeper increase until 2040 where the increment will be slightly less intense (Fig.
+2).
 Similarly, for the number of EEZs, the increment curve seem to flatten by 2020.
@@
 
-```{r SSR_gain_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Cumulative increase in time of emergence by EEZ and stock"}
+```{r freq_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Cumulative increase in time of emergence by EEZ and stock"}
 
@@
 
-On average, most of the world EEZ's will see a stock ToE by 2030 with a higher frequency of number of emerging stocks between latitudes 0$^o$ and 50$^o$N (Fig 2).
-In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50%, however, in some cases it represents 85% of total fishing revenue (Peru) or more, like the case of island territories such as Marshall Islands (92%) or Tokelau- New Zealand (100%).
-In few cases like Brazil's EEZ and Alaska's polar EEZ, no stock will emerge.
+On average, most of the world EEZ's will see a stock emerge by 2030 with a few EEZs having a relative late average ToE (e.g.
+Honduras average ToE is 2071) and 54 EEZs having already emergered (e.g.
+average ToE >= 2020).
+A higher number of emerging stocks will happen between latitudes 0$^o$ and 50$^o$N corresponding to regions of relatively small EEZs with multiple borders (e.g.
+The Carribeean and the Medditerranean) (Fig 2).
+In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50%.
+However, in some cases emerging stocks represent 85% of total fishing revenue (e.g.
+Peru) or more (e.g.
+Marshall Islands, 92%).
+Only in very few cases like Brazil and Alaska's polar EEZ, presented no emerging stock between 2006 and 2100.
 
@@
 
-The regional patterns of ToE suggest that most stocks will emerge by early 21st with many stocks in the Americas and Oceania already emerging.
-For the larger fisheries of the world in terms of revenue representing 75 percentile, only the USA will see an average ToE before 2030 while fisheries in all other top fishing countries will be emerging within the following 10 years.
+Most of the main fishing countries in terms of revenue from transbounudary fisheries (e.g.
+the top $75^{th}$ percentile) will have an average ToE across all species after 2030 (n = 28).
+The remaining twelve countries will have an average ToE before 2030, this includes the United States who is the $4^{th}$ place it total reveneue (Fig.
+4).
+These 47 countries are also among those with the largest number of emerging taxa.
+Regionally speaking, most of African, Asian and European EEZs will see a latter ToE, while the Americas and Oceania will see an early ToE (Fig.
+4).
+From the top ten countries with more revenue from transboundary species, nine of them share a border.
+Most of these countries (e.g.
+Spain and China) will also have a large number of emering stocks, however, some others like Peru and Norway will have a small number of emerging taxa that represents high revenue fisheries.
 
-```{r Toe_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}
+```{r bubble_plot, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}
 
@@
 
+We compared species ToE within EEZs by geogrpahic region and by species habitat preference.
 The mean ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001) (Fig.
@@
 Other regions also significantly differ from each other but in a less frequent manner (*See Table Sup X for full statistics*).
-When looking specifically at habitat association, we also found significant differences on ToE (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
-Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats is significantly different than other for other habitat preferences (*See Table Sup XX for full statistics*).
+When looking specifically at habitat association, we found significant differences on ToE of different habitat types (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
+Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other for other habitat preferences (*See Table Sup XX for full statistics*).
 
-```{r Toe_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}
+```{r Boxplot, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}
 
@@
 
-We hypothesize that species environmental preferences (e.g.
-presence in certain biomes and habitat preference), climate change effects on water temperature (e.g.
+We hypothesized that species environmental preferences (e.g.
+presence in certain biomes and habitat preferences), climate change effects on water temperature (e.g.
 ToE of SST), human drivers (e.g.
 catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary stocks (*Table Sup 3*).
-Results from the multiple linear regression analysis show a significant relation between ToE of transboundary stocks and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (Table 1).
+Results from the multiple linear regression analysis show a significant relation between ToE of transboundary stocks and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table 1).
 
-Table 2.
+Table 1.
 Multiple regression results for time of emergence of transboundary stocks.
@@
 
-```{r table1, eval = T,  echo = F"}
+```{r table1, eval = T,  echo = F}
 
 read_csv("../Tables/mlr_result.csv") %>% 
-  knirt::kable()
+  knitr::kable()
 
@@
 
-# Discussion
+<!-- # Discussion -->
 
-The combination of the time of emergence and the revenue participation of the species could shed a light in the importance of the time of emergence in terms of management.
+<!-- The combination of the time of emergence and the revenue participation of the species could shed a light in the importance of the time of emergence in terms of management.
 For example, we identified that in Peru 12 species will emerge by 2038 in average, representing 92% of the country's fishing value.
@@
 1).
+-->
 
-Form all significantly different variables, only latitudinal gradient postpone the ToE suggesting that the ToE will happen first at equatorial latitudes and later towards the poles.
+<!-- Form all significantly different variables, only latitudinal gradient postpone the ToE suggesting that the ToE will happen first at equatorial latitudes and later towards the poles. -->

### 2020-09-01 Scripts/Initial_Analysis.Rmd
sip fig 1

@@
 
+Main_data <- Eme_Res %>% # 19728
+  filter(tresh == "Tresh_one") %>% # 11490
+  left_join(Spp_depth) %>% 
+  select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel,un_region= un_geo_region) %>% 
+  left_join(Toe_temp,
+            by = c("eezid","tresh")
+            ) %>% 
+  left_join(Catch_trend,
+            by = c("taxon_key","sau_fishing_entity","fishing_entity_id")
+            ) %>% # 11492
+  left_join(Eez_biogeo,
+            by = "eezid") %>% 
+  left_join(Spp_env_pref,
+            by = "taxon_key") %>% 
+  left_join(Eez_centroids,
+            by = "eezid") %>% 
+  mutate(latitude = abs(latitude))
+
 ```
@@
 # http://jsmayorga.com/post/mapping-the-global-network-of-transnational-fisheries/
+
+### Median number
+
+Prop_results %>% 
+  filter(variable == "n_spp") %>% 
+  group_by() %>% 
+  summarise(
+    median(early,na.rm=T),
+    sd(early,na.rm=T)
+  )
 
@@
 
-# time <- "per_mid_plot"
-time <- "per_early"
+time <- "per_mid"
+# time <- "per_early"
 
@@
 ) +
-ggsave("./Figures/fig2ax.png",
+ggsave("./Figures/fig_sup_1.png",
        width = 12,
@@
 
+## Top fishing nations
+Buble_data %>% 
+  filter(log_value > Percentile_value$x) %>% 
+  View()
+
 # Labels Data
@@
 # AIC value; model between predictive capacity and complexity (i.e., number of explanatory variables) is given by the model with lowest AIC value
-
-# -------------------------------- #
-#### Prepare data
-# -------------------------------- #
-
-Lm_data <- Eme_Res %>% # 19728
-  filter(tresh == "Tresh_one") %>% # 11490
-  left_join(Spp_depth) %>% 
-  select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel,un_region= un_geo_region) %>% 
-  left_join(Toe_temp,
-            by = c("eezid","tresh")
-            ) %>% 
-  left_join(Catch_trend,
-            by = c("taxon_key","sau_fishing_entity","fishing_entity_id")
-            ) %>% # 11492
-  left_join(Eez_biogeo,
-            by = "eezid") %>% 
-  left_join(Spp_env_pref,
-            by = "taxon_key") %>% 
-  left_join(Eez_centroids,
-            by = "eezid") %>% 
-  mutate(latitude = abs(latitude))
 

### 2020-09-02 Scripts/Manuscript.Rmd
Drafted intrto, methods and results

@@
   Policy Response Under Climate Change
-author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, U. Rashid Sumaila^1^,  Vicky
-  Lam^1^, David VanderZwaag^3^, and William W.L. Cheung^1^'
+author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, U. Rashid Sumaila^1^,  Vicky Lam^1^, David VanderZwaag^3^, and William W.L. Cheung^1^'
 date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British
@@
   pdf_document: default
-subtitle: Climate-ready transboundary fisheries management under Climate Change
 ---
@@
 
-The delineation of Economic Exclusive Zones (EEZs) by the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, established political boundaries across the distribution of many fish species, creating shared species between nations.
+<!-- The delineation of Economic Exclusive Zones (EEZs) by the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, established political boundaries across the distribution of many fish species, creating shared species between nations.
 Despite not implicit, UNCLOS incentivizes nations to cooperate in the management of transboundary fisheries and most management plans emplaced are based on historical distribution of catches.
@@
 The results of this research has the potential to improve climate-ready fisheries if policy is modify to respond to the consequences presented here.
+-->
 
-<!-- # Introduction -->
+# Introduction
 
-The world is moving towards sustainable development and ocean-based activities plays a mayor role in achieving international policies.
+Over the last century human activities have fundamentally changed the physicochemical composition of the ocean, warming the waters, increasing the pH and reducing the oxygen concentration [@IPCC:2019to].
+As a consequence, many marine species are changing their distribution [@ Poloczanska:2016kk] and will continue to do so in the search of the apropiate environmental conditions needed for their subsistence [@Cheung:2010dt.
+Climate change is expected to impact fisheries production at different latitudes compromising our capacity to reach international sustainability goals such as the SDG 14 - life below water [@SDG14; @Cheung:2010dt; @Barange:2014kr].
+Such impacts can be offset with the proper reform of current fisheries management [@Gaines:2018sg; @CheungClimateReadyFish].
+However, such reforms will require comprehensive plans and institutions to manage species that cross international borders, also known as shared stocks [@Gaines:2018sg;@Pinsky:2018cb;].
 
-One of the main goals of the United Nations Sustainable Development Goal No.14 (SDG 14-Life below water) is to achieve sustainable development of fisheries [@SDG14].
-<!-- (https://sustainabledevelopment.un.org/sdg14). -->
+The concept of shared stocks gained attention in the early 1980's when the United Nations officialized the concept of Exclusive Economic Zone (EEZs) though the Law of the Sea (UCLOS) [@UnitedNations:1986tl].
+The implementation of arbitrary jurisdictions across the global ocean created a partition in the distribution of several stocks, thus, creating what we known today as shared stocks.
+As defined by the FAO, shared stocks can be classified as transboundary, those that cross neighboring EEZs; straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and highly migratory, mainly tunas and bill-fishes that cover broad oceanic regions [@FAO; @Munro].
+While counties are responsible for the management of stocks within their EEZs, UNCLOS incentives actions to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
+Today, an estimate of 633 transboundary fished stocks are found in different continents and across multiple habitats averaging 48.5 million tones and representing USD 78 billion in fishing revenue between 2005 and 2010 [@FishorVisa].
 
-Climate change is expected to reduce the production from global fisheries in about XX% to xx% [@Cheung; @Barange] compromising our capacity to reach SDG 14.
+The management, and thus sustainability, of fisheries that target shared stocks is jeopardized as marine species shift their known distribution to cope with a changing world [@Pinsky:2018cb].
+There is evidence that such shifts have already complicated shared marine resources around the globe.
+In the North-eastern Atlantic, evidence suggest that shared stocks managed within the European Union divisions have been shifting since the 1970's compromising the region's sustainability [@EuropeShift].
+In the North-eastern Pacific, sockeye salmon (*Oncorhynchus nerka*) shifted in the 1980's towards Alaska, creating a ten years conflict between Canada and the United States [@Song].
+In the South-eastern Pacific, the southern stock of Peruvean anchoveta (*Engraulis ringens*) shared by Chile and Peru has gradually shifted south (towards Chilean waters) adding to the over-exploitation of the resource [@Cashion:2018cg].
+Global shifts in the distribution of tunas [@ErauskinExtramiana:2019fw] have brought additional management challenges in both the Western Indian ocean and South-western Pacific [@Miller:2007kx].
+Moreover, as species continue shifting with climate change, current policies might become obsolete, and thus, require new management strategies that create equitable systems across all dependent fishing nations [@Bell:2013ga; @Pinsky:2018cb].
+We argue that two concepts will be essential in addressing the challenges that climate change will bring to the management of transboundary fisheries.
 
-Recent research show that the impacts of climate change on the productivity of marine wild fisheries can be offset with the proper reform of current fisheries management [@Gaines; @CheungClimateReadyFish].
-While bringing fisheries that are poorly managed to a sustainable level will likely offset the negative impacts of climate change to fisheries production [@Gaines], it will not necessarily solve the spatial consequences of species moving from one management area to another.
+The first concept alludes to the game theory concept of thread point, known as the minimum payoff a player is willing to receive in order to cooperate with others.
+Thus, if a player's future payoff drops below the thread point, such player will likely break off any collaboration agreement and proceed to act solo [@Sumaila:E&S].This can be seen as the proportion of a species distribution that each neighboring nation has (or the quota allocation that each one gets) at the time of signing off an agreement [@PalaciosAbrantes:E&S].
+Many shared fisheries operate this way [@EuropeShift; @GoMA].
+If not addressed, future changes in the thread point (hereafter called Stock Share Ratio -SSR) of jointly managed fisheries can have adverse consequences for shared fisheries [@PalaciosAbrantes:E&S; @Sumaila:E&S].
+The second concept refers to the Time of Emergence (ToE), defined as the moment when a future trend of a particular variable exceeds the historical variability [@IPCC:2019to].
+Such concept has been used to look at the emergence of diverse oceanic variables [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg] and ecological consequences of marine heatwaves [@Cheung:2020ja].
+Similarly, climate change is expected to drive biomass further away from the historic range limits of a species [@Alexa?].
+Assessing the ToE of shared stocks is as important as the change itself as it could reduce potential international conflicts [@Spijkers:2017ij] and prepare for emerging fisheries [@Pinsky:2018cb;  @Pinsky:2014by].
 
-Shifts in species distribution does not necessarily mean a reduction in production (e.g. reduction in stock biomass) but rather a spatial re-distribution of the stock's biomass [@RefLiam?].
-
-Documented cases of stock shifts have resulted in international conflict and sustainable fishing rates by parties involved in the conflict.
-
-This is of specific concern to a set of 633 fished species that cross international borders, also known as transponder species [@FishForVisa].
-
-With the implementation of arbitrary spatial management boundaries, at least 633 species became transboundary representing an average catch (2005 - 2010) of 47.6 million tons and generating USD 65,483 million in fishing revenue [@FishorVisa].
-
-Transboundary fish stocks refer to those species that cross international jurisdictions known as Economic Exclusive Zones (EEZs) of neighboring nations [@FAO].
-These stocks are responsible for an average catch (2005 - 2010) of 47.6 million tons generating USD 65,483 million in fishing revenue [@FishorVisa].
-
-Historically, environmentally-driven shits in shared stocks have cause problems and threaten stock's sustainability.
-**In early 1980's Canada and the US blablabla.** Many fish stocks managed by the European Union ICES regultaroy areas have been shifting since the 1970's [@EuropeShift].
-Such shifts have already result in confrontations between nations as evidenced by the expansion of the Atlantic mackerel (*Scomber scombrus*) distribution reaching the EEZ of Iceland [@Spijkers:2017ij].
-More recently, Chile and Peru have declare a cooperation strategy to mange the transboundary stock of Peruvean Anchovera [@Peru].
-Furthermore, as we continue to enter the $21^{st}$ century, commercial marine species are expected to continue shifting (mostly poleward) potentilly compromising shared stocks management.
-By the end of the century, all EEZs of the world are proyected to have at least one new transboundary species with some of them having over 10 new stocks [@Pinsky].
-Consequently, as species occupy new waters they leave historic areas and so, xx% of species are expected to exit EEZs by the end of che century [@Oremus].
-This entering and exicting national jurisdicions will complicate fisheries management potentially creating confllict between neighbouring nations [@Pinksy @Oremus:2020id].
-
- So, preparing policy to species on the move is key to achieve sustainable development goals and proper governance of the world fisheries [@Pinksy].
-
-Undrstanding the degree by which shared stocks are going to change and the time of emergence is key for creating climate-ready fisheries policy.
-
-The concepts of transboundary stocks and EEZs were introduced in the early 1980's by the United Nations Law of the Sea (UCLOS) [@UnitedNations:1986tl].
-Under UNCLOS each costal nation has the right (and responsabillity) to manage wild fish stocks up to 200 nautical mailes of the coast.
-Exceptions exist when territorial waters are in dispute (e.g.
-South China sea[@Ref]) or when menagement agreements exist between nations [@Ref].
-In fact, UNCLOS incentives actions to cooperate on the management of shared stocks [Article 63, @UnitedNations:1986tl].
-Climate change is shifting the distribution of marine species to the extreem that some spatial policies implemented to manage fisheries are becoming obsolete [@Pololanza; @Jessica].
-
-<!-- ## Link paper un terms of SDGs -->
-<!-- - UnitedNations2018 Blanchard:2017hl Singh:2017ds -->
-
-<!-- ## Collaborative management of share stocks -->
-<!-- - Theat points -->
-<!-- - Concept of treaty estabillity -->
+Here we estimate the changes in Stock Share Ratio and the Time of Emergence of 633 transboundary stocks shared by 280 EEZs of 198 coastal countries under climate change (See methods).
+We rely on a species distribution and Earth system model with ten ensemble members to project the distribution of the 663 selected fish stocks under a high greenhouse gas emission scenario.
+We treated each species in an EEZ as a single stock due to a lack of more detailed spatial data [@FishForVisa] and only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g.
+we did not include the high seas).
+Our results suggest that virtually all EEZs will have a change in SSR by 2030 and that 87% of them already had at least one emerging species since 2006, with significant regional and environmental differences.
+Towards the end of the $21^{st}$ century, all of the 280 EEZs will have at least one emerging stock and all 633 stocks would have emerged at least once.
+Our results have direct implications for marine governance of shared marine resources under a changing world.
 
@@
 We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the *Sea Around Us* shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic) including islands territories, and determined the intersections between polygons using the R package *sf* (**Supplements Fig 1 (SuppM)**) [@PackagesfSimpleF:2018vp].
-For each species and EEZ we used the *Sea Around Us* catch data to estimated the average catch and fishing revenue (USD 2019).
-We only used data from fishing activities within the global EEZs reported as the average from the last decade (2005-2014) of available spatial data [@Zeller2016; @Tai2017; @Sumaila:2015uc].
-Moreover, we catalogue catch trends as increasing, constant and decreasing, following Palacios-Abrates *et al.* (*hopefully in press*).
-Finally, each EEZ was attributed to a geopolitical region [@UN] and biome [@Gabriel], and species were attributed a habitat preference [@FishBase].
-For the present study we first projected the 633 species distributions under climate change and then estimated two main metrics; a stock's share ratio (SSR) and time of emergence (ToE) as described in the following sections.
+For each species and EEZ we used the Sea Around Us data to estimate the catch and fishing revenue (USD 2019 from fishing activities within the global EEZs reported as the average from the last decade (2005-2014) of available spatial data [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+Moreover, we classified catch trends as increasing, constant and decreasing, following Palacios-Abrantes *et al.* (*hopefully in press*).
+Finally, each EEZ was attributed to a geopolitical region [@UN] and biome [@Reygondeau:2019uh], and species were attributed a habitat preference [@FishBase].
 
@@
 We do this by following an ensemble member approach where each member constitutes a different initial Earth system condition [@Rodggers:2015jq; @Frolicher:2009ge].
-Thus, for our experiment, in each of the ten GFDL-ESM2M-ensembles the Earth system initial condition is infinitesimally different resulting in an unique atmosphere and ocean state at each point in time.
+Thus, for our experiment, in each of the ten GFDL-ESM2M-ensembles the Earth system initial condition is infinitesimally different resulting in a unique atmosphere and ocean state at each point in time.
 This results in a different state of internal variability for each ensemble representing a plausible future under RCP 8.5 [@Rodggers:2015jq].
@@
 
-In game theory, a threat point is known as the minimum payoff that a player is willing to receive in order to cooperate with other players.
-Thus, if a player's payoff is lower than the thread point, such player will likely break off any collaboration agreement and proceed to act solo [@Sumaila:E&S].
-Here, we adopted the concept of threat point as the minimum proportion of the catch that each country will accept if they were to cooperate in the management of a transboundary stock [@Palacios-Abrantes:E&S] hereafter referred as stock share ratio (SSR).
+We adopted the concept of threat point as the minimum proportion of the stocks’ joint distribution between neighboring EEZs that each country will accept if they were to cooperate in the management of the stock [@Palacios-Abrantes:E&S].
 For each stock and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighboring EEZs, and then calculated the proportion that each EEZ had [@Palacios-Abrantes:xxxx].
@@
 This time frame matches the historic climate data that the GFDL uses to project climate change [@Dunne:2012bo; @Dunne:2013db].
-We then selected two future periods; early $21^st$ century as the average of 2021 to 2040 ($t_e$), and mid $21^st$ century as the average of 2041 to 2060 ($t_m$).
+We then selected two future periods; early $21^{st}$ century as the average of 2021 to 2040 ($t_e$), and mid $21^{st}$ century as the average of 2041 to 2060 ($t_m$).
 We choose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing) and SDG 1.2 (poverty reduction), aimed to be achieved by 2020 and 2030, respectively [@Singh:2017ds].
 We then estimate the percentage change in the SSR of each future time period relative to the historic time period.
-To address the natural spatial variability of the stock we only considered a SSR change when the future SSR was higher or lower than the average historic SSR $\pm$ 1 s.d.
+To address the natural spatial variability of the stock we only considered an  SR change when the future SSR was higher or lower than the average historic SSR $\pm$ 1 s.d.
 The analysis was replicated for each of the ten ensemble members capturing variation by averaging results among ensemble members (mean $\pm$ standard deviation).
 
-## Time of Emergence Index
+## Time of Emergence
 
 Knowing the time (e.g.
-year) in which SSR is going to diverge from a historical variability can be key for policy to prepare to such change.
-In here we call this event time of emergence and defined it as the moment in time when the future variation in SSR becomes larger than the historical trend variability [@Roggers:2015jq].
+year) in which a shared stock is going to diverge from a historical variability can be key for preparing to shifting stocks.
 We employed the ten GFDL ensemble members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
@@
 Since this was an arbitrary decision, we have included a more conservative estimation ($\mu \pm 2*\sigma$ for a 95% confidence interval).
-The SNR for both bottom and surface sea temperature was also estimated following the previous equation.
+The SNR for both bottom and surface sea temperature was also estimated following the SNR method.
 
@@
 
-In other words, for two neighboring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases (see *TI* equation).
+<!-- In other words, for two neighboring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases (see *TI* equation).
 Consequently, if the centroid shifts are within historical variability (*SNR* < 1), risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability (*SNR* > 1), the risk of dispute is higher.
 Thus, risk is positively related to the frequency of centroid shifts beyond historical variability.
+-->
 
@@
 
-We tested data for normallity (e.g.
-skewness, kurtosis) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the time of emergence of transboundary stocks.
+We tested data for normality (e.g.
+skewness, kurtosis) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary stocks.
 Moreover, we implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
-Specifically, we lookd at habitat preferences, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
-We hypothesize that all of these factors would have significant influence in the time of emergence of transboundary stocks.
+Specifically, we looked at habitat preferences, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
+We hypothesize that all of these factors would have significant influence in the ToE of transboundary stocks.
 All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@gmt] and *zoo* [@PackagezooSInfr:2019tm], *readxl* [@REF], *ggrepel* [@REF], *zeallot* [@REF], *viridis* [@], *cowplot* [@], *moments* [@REF] and *pgirmess* [@REF].
@@
 
-We estimated changes in the Stock Share Ratio of transboundary species shared by neighboring EEZs by early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
+We estimated changes in the SSR of transboundary species shared by neighboring EEZs by early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
 Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
 1A).
-While most EEZs will experience an average change lower than 25%, a few EEZs are expected to have larger changes in SSR relative to early century.
+While most EEZs will experience an average change lower than 25%, a few EEZs are expected to have larger changes in SSR.
 Such change is expected to slightly increase in intensity towards the  mid $21^{st}$ century (**Supplements Fig 1 (histogram)**).
 For example, Guatemala's SSR of *Panulirus gracilis* shared with Mexico is projected to go from 12% to 26%.
-The number of stocks per EEZ that will see a change in SSR by early $21^{st}$ century is relatively low (median of 13 spp per EEZ), however, some regions like West Africa and the Mediterranean will see changes in over 50 species (e.g.
+The number of stocks per EEZ that will see a change in SSR by early $21^{st}$ century is relatively low (median of 13 spp.
+per EEZ), however, some regions like West Africa and the Mediterranean will see changes in over 50 species (e.g.
 Senegal, 94).
@@
 
-![Average gain of transboundary stock share ratio (lines) and number of shifting stocks (polygon) by 2030 (2021-2040) relative to 1951-2005.
+![Figure 1.
+Average gain of transboundary stock share ratio (lines) and number of shifting stocks (polygon) by 2030 (2021-2040) relative to 1951-2005.
 Arrows go from the losing EEZ (point) to the winner (arrowhead).
-Pannel B zooms in to specific areas ](../Figures/Fig1.png)
+Panel B zooms in to specific areas.](../Figures/Fig1.png)
 
@@
 
-We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock's distrtibution overpasses the historical variation.
+We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock's distribution overpasses the historical variation.
 Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g.
 *Engraulis ringens* shared by Chile and Peru emerged in 2010).
-The cumulative sum of emerging stocks will have a steeper increase until 2040 where the increment will be slightly less intense (Fig.
+The cumulative sum of emerging stocks will have a steeper increase until 2025 where the increment will be slightly less intense (Fig.
 2).
@@
 
-![Cumulative increase in time of emergence by EEZ and stock](../Figures/Fig2.png)
+![Figure 2 Cumulative increase in time of emergence by EEZ and stock](../Figures/Fig2.png)
 
-On average, most of the world EEZ's will see a stock emerge by 2030 with a few EEZs having a relative late average ToE (e.g.
-Honduras average ToE is 2071) and 54 EEZs having already emergered (e.g.
-average ToE >= 2020).
+Most of the world EEZ’s will see an average ToE across transboundary stocks by 2030 with a few EEZs having a relative late average ToE (e.g.
+Honduras average ToE is 2071) and 54 EEZs having already emerged (e.g.
+average ToE <= 2020).
 A higher number of emerging stocks will happen between latitudes 0$^o$ and 50$^o$N corresponding to regions of relatively small EEZs with multiple borders (e.g.
-The Carribeean and the Medditerranean) (Fig 2).
+The Caribbean and the Mediterranean) (Fig 2).
 In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50%.
@@
 Marshall Islands, 92%).
-Only in very few cases like Brazil and Alaska's polar EEZ, presented no emerging stock between 2006 and 2100.
+Only in very few cases like Brazil and Alaska's polar EEZ, no stock will emerge between 2006 and 2100.
 
@@
 
-![(Average time of emergence (ToE) per Exclusive Economic Zone (EEZ polygon) and participation of emerging stocks in total EEZ fishing revenue (land polygon).
+![Figure 3.
+Average time of emergence (ToE) per Exclusive Economic Zone (EEZ polygon) and participation of emerging stocks in total EEZ fishing revenue (land polygon).
 Darker tones represent low participation and late ToE; lighter colors represent large participation and earlier ToE.
-Density plot on right shows the number of emerging stocks across EEZ centroids (10^o running mean))](../Figures/Fig3.png)
+Density plot on right shows the number of emerging stocks across EEZ centroids (10^o running mean).](../Figures/Fig3.png)
 
-Most of the main fishing countries in terms of revenue from transbounudary fisheries (e.g.
-the top $75^{th}$ percentile) will have an average ToE across all species after 2030 (n = 28).
-The remaining twelve countries will have an average ToE before 2030, this includes the United States who is the $4^{th}$ place it total reveneue (Fig.
+Most fisheries within the EEZ of the main transboundary fishing countries in terms of revenue (e.g.
+the top $75^{th}$ percentile) will emerge after 2030 (n = 28).
+The remaining twelve countries will have an average ToE before 2030, this includes the United States who is the $4^{th}$ place in total revenue (Fig.
 4).
@@
 Most of these countries (e.g.
-Spain and China) will also have a large number of emering stocks, however, some others like Peru and Norway will have a small number of emerging taxa that represents high revenue fisheries.
+Spain and China) will also have a large number of emerging stocks, while Peru and Norway will have a small number of emerging taxa but representing high revenue fisheries.
 
@@
 
-![(Average time of emergence (ToE) per fishing entity.
-Fishing revenue in logarithmic sacle.
-Showing names for the top 75th percentile).
-Countries marked with a asterisk have at least one border](../Figures/Fig4.png)
+![Figure 4.
+Average time of emergence (ToE) per fishing entity.
+Fishing revenue in logarithmic scale Showing names for the top 75th percentile).
+Countries marked with an asterisk have at least one border.](../Figures/Fig4.png)
 
-We compared species ToE within EEZs by geogrpahic region and by species habitat preference.
+We compared species ToE within EEZs by geographic region and by species habitat preference.
 The mean ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001) (Fig.
 5).
-Post-hoc tests revealed that EEZs in Easter Asia who have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ with almost all other regions.
+Post-hoc tests revealed that EEZs in Easter Asia which have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ from (almost) all other regions.
 Other regions also significantly differ from each other but in a less frequent manner (*See Table Sup X for full statistics*).
 When looking specifically at habitat association, we found significant differences on ToE of different habitat types (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
-Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other for other habitat preferences (*See Table Sup XX for full statistics*).
+Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (*See Table Sup XX for full statistics*).
 
@@
 
-![Regional time of emergence per habitat preference.
+![Figure 5.
+Regional time of emergence per habitat preference.
 Regions acording to the United Nations and habitat preference to FishBase.
@@
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers (outside of 1.5*IQR.)](../Figures/Fig5.png)
+Line represents median, and black points represent outliers (outside of 1.5*IQR.](../Figures/Fig5.png)
 
@@
 <!-- Form all significantly different variables, only latitudinal gradient postpone the ToE suggesting that the ToE will happen first at equatorial latitudes and later towards the poles. -->
+
+<!-- Historically, environmentally-driven shits in shared stocks have cause problems and threaten stock's sustainability.
+**In early 1980's Canada and the US blablabla.** Many fish stocks managed by the European Union ICES regultaroy areas have been shifting since the 1970's [@EuropeShift].
+Such shifts have already result in confrontations between nations as evidenced by the expansion of the Atlantic mackerel (*Scomber scombrus*) distribution reaching the EEZ of Iceland [@Spijkers:2017ij].
+More recently, Chile and Peru have declare a cooperation strategy to mange the transboundary stock of Peruvean Anchovera [@Peru].
+Furthermore, as we continue to enter the $21^{st}$ century, commercial marine species are expected to continue shifting (mostly poleward) potentilly compromising shared stocks management.
+By the end of the century, all EEZs of the world are proyected to have at least one new transboundary species with some of them having over 10 new stocks [@Pinsky].
+Consequently, as species occupy new waters they leave historic areas and so, xx% of species are expected to exit EEZs by the end of che century [@Oremus].
+This entering and exicting national jurisdicions will complicate fisheries management potentially creating confllict between neighboring nations [@Pinksy @Oremus:2020id].
+-->
+
+<!-- Conclusion -->
+
+<!-- Preparing policy to species on the move is key to achieve sustainable development goals and proper governance of the world fisheries [@Pinksy]. -->
+
+# Recycling text
+
+Ignoring anthropocentric effects, a species biomass distribution follows a variation in time and space according to environmental preferences, species interactions and landscape features [@Huch].
+Thus, a neighboring nation will not necessarily have, at all times  (e.g.
+months or years), a fixed SSR but rather a series of SSRs [@GoMA].

### 2020-09-03 Scripts/Initial_Analysis.Rmd
Including 1 sd treshold step in proportion analysis

@@
 ```{r fun_SummaryProp, eval = F, echo =F}
+taxon_key <- 600006
 
@@
   # Reads taxon data
-  proportion_data <- fread(paste(my_path("R",extra_path = "Proportion_2005/"),"proportion_",taxon_key,".csv", sep=""))
+  proportion_data <- fread(paste(my_path("R",extra_path = "Proportion_2005/"),"proportion_",taxon_key,".csv", sep="")) %>% 
+    filter(!is.na(eez_neighbour))
+
+  # Identify SSR treshold for natural spatial variability of the stock
+  proportion_tresh <- proportion_data %>% 
+    filter(time_step == "historical") %>% 
+    mutate(
+      top_tresh = ensemble_mean+ensemble_sd,
+      low_tresh = ensemble_mean-ensemble_sd
+    ) %>% 
+    select(taxon_key:eez_neighbour,top_tresh:low_tresh) %>% 
+    left_join(proportion_data) %>% 
+    # Only keeps change if surpasses treshold
+    mutate(
+      over_tresh = ifelse(ensemble_mean > top_tresh | ensemble_mean < low_tresh,"keep",
+                          ifelse(time_step == "historical","keep","drop")
+      )
+    ) %>% 
+    filter(over_tresh == "keep") %>% 
+    select(-top_tresh,-low_tresh)
 
@@
     # summarizes by country and neighbour
-    country_summary <- proportion_data %>% 
+    country_summary <- proportion_tresh %>% 
       group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% 
@@
     # summarizes by country
-    country_summary <- proportion_data %>% 
-      group_by(taxon_key,eez_name,time_step) %>% 
+    country_summary <- proportion_tresh %>% 
+      group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% 
       summarise(
@@
 
-The next step is to summarize all species per eez so our results shows average change of all species within an EEZ towards a neighbour
+The next step is to summarize all species per eez so our results shows average change of all species within an EEZ towards a neighbour.
+
+**Note: Contains step to filter out changes less than +-1 s.d.**
 
@@
 ```{r fun_PropChange, eval = F, echo =F}
+taxon_list <- unique(Transboundary_spp$taxon_key)[1:15]
 

### 2020-09-03 Scripts/Manuscript.Rmd
Updated version of the manuscript

@@
 We then estimate the percentage change in the SSR of each future time period relative to the historic time period.
-To address the natural spatial variability of the stock we only considered an  SR change when the future SSR was higher or lower than the average historic SSR $\pm$ 1 s.d.
+To address the natural spatial variability of the stock we only considered a SSR change when the future SSR was higher or lower than the average historic SSR $\pm$ 1 s.d.
 The analysis was replicated for each of the ten ensemble members capturing variation by averaging results among ensemble members (mean $\pm$ standard deviation).

### 2020-09-04 Scripts/Initial_Analysis.Rmd
Sent first draft to W. up to results

@@
 ```{r fun_SummaryProp, eval = F, echo =F}
-taxon_key <- 600006
 
@@
 
+    # Number of SSR stocks
+    n_ssr <- spp %>% 
+      group_by(eez_name,time_step) %>% 
+      summarise(
+        n_spp = length(unique(taxon_key))
+      ) %>% 
+      spread(time_step,n_spp) %>% 
+      mutate_at(vars(early,mid),
+                .funs = ~(./historical)*100
+      ) %>% 
+      select(-historical) %>%
+      gather("time_step","n_per",early:mid)
+
+    # Min, max and median change per eez_name and neighbour
     eez_summary <- spp %>% 
@@
       summarise(
-       n_spp = n(),
+        min_eez = min(mean_spp, na.rm=T),
+        max_eez = max(mean_spp, na.rm=T),
+        sd_eez = sd(mean_spp, na.rm=T), # for some weird reason it wont work if the order is fliped
+        mean_eez = median(mean_spp,na.rm=T)
+        ) %>% 
+      select(eez_name:max_eez,mean_eez,sd_eez) %>% 
+      left_join(n_ssr)
+
+  }else{
+
+    eez_summary <- spp %>% 
+      group_by(eez_name,time_step) %>% 
+      summarise(
+        n_spp = length(unique(taxon_key)),
+        # n_spp = n(),
         min_eez = min(mean_spp, na.rm=T),
@@
 
- }else{
-
-   eez_summary <- spp %>% 
-     group_by(eez_name,time_step) %>% 
-     summarise(
-       n_spp = n(),
-       min_eez = min(mean_spp, na.rm=T),
-       max_eez = max(mean_spp, na.rm=T),
-       sd_eez = sd(mean_spp, na.rm=T), # for some weird reason it wont work if the order is fliped
-       mean_eez = mean(mean_spp,na.rm=T)
-     ) %>% 
-     select(eez_name:max_eez,mean_eez,sd_eez)
-
   }
@@
 ```{r fun_PropChange, eval = F, echo =F}
-taxon_list <- unique(Transboundary_spp$taxon_key)[1:15]
 
@@
   proportion_change <- eez %>% 
-   gather("variable","value",n_spp:sd_eez) %>% 
+    gather("variable","value",min_eez:n_per) %>% 
     spread(time_step,value) %>% 
@@
 Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)
+
+Neighbours_n <- Neighbours %>% 
+  group_by(eez_name) %>% 
+  summarise(
+    n = length(unique(eez_neighbour))
+  )
 
@@
 # --------------- #
+# Get taxon list
+Taxon_list <- list.files(my_path("R",extra_path = "Proportion_2005"))
+
+# clean names for function
+Taxon_list <- gsub(".*_","",Taxon_list)
+
+# clean names for function
+Taxon_list <- gsub(".csv","",Taxon_list)
 
 # Read proportion results
-# prop_results <- PropChange(Taxon_List_emer, links="y") %>% 
-  # filter(!is.na(eez_neighbour)) %>% 
-  # mutate(per_mid_plot =ifelse(per_mid >= 50, 50,
-  #                          ifelse(per_mid <= -50,-50,per_mid)
-  #                          ),
-  #        per_early_plot =ifelse(per_early >= 50, 50,
-  #                          ifelse(per_early <= -50,-50,per_early)
-  #                          )
-  #        )
+Prop_results <- PropChange(Taxon_list, links="y") %>%
+  filter(!is.na(eez_neighbour)) %>%
+  mutate(per_mid_plot =ifelse(per_mid >= 50, 50,
+                              ifelse(per_mid <= -50,-50,per_mid)
+  ),
+  per_early_plot =ifelse(per_early >= 50, 50,
+                         ifelse(per_early <= -50,-50,per_early)
+  )
+  )
 
@@
   clean_names() %>% 
+  select(iso_a3,sovereignt,sov_a3,name,name_long,region_wb,continent,geometry) %>% 
   # st_transform(crs = "+proj=eck4") %>%
   st_transform(crs = 4326) %>% # 4326
+  mutate(iso_a3 = ifelse(sovereignt == "Norway","NOR",
+                         ifelse(name == "France","FRA",
+                                iso_a3)
+                         )
+         ) %>% 
     left_join(World_map_dat,
@@
 Prop_results %>% 
-  filter(variable == "n_spp") %>% 
+  filter(variable == "n_per") %>% 
   group_by() %>% 
@@
   filter(
-    variable %in% c("mean_eez","n_spp")#,
+    variable %in% c("mean_eez","n_per")#,
     # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
@@
   gather("time_step","per_chng",per_early:per_mid) %>% 
-  mutate(per_plot = ifelse(per_chng > 100,100,per_chng))
+  mutate(per_plot = ifelse(per_chng > 100,100,per_chng),
+         direction = ifelse(per_chng > 0,"positive","negative")
+         )
 
@@
 n_taxa_data <- Prop_results %>% 
-  filter(variable == "n_spp") %>% 
+  filter(variable == "n_per") %>%
+  select(eez_name,early,mid) %>% 
   rename(sf_eez_name = eez_name) %>% 
@@
   group_by(sau_fishing_entity,fishing_entity_id) %>% 
-  summarise(total_spp = mean(early,na.rm=T))
+  summarise(avrg_spp_per = mean(early,na.rm=T))
 
@@
 
-gc()# clean ram
+mean(n_taxa_data$avrg_spp_per,na.rm=T)
+
+# Rectangles dataset
+# coord_sf(xlim = c(-3, 35),ylim = c(30, 45)) +
+Rectangles <- tibble(
+    x1=c(-100,-3,-14), #LatAm, Med, NE
+    x2=c(-55,35,23),
+    y1=c(9,31,50), 
+    y2=c(30,44,67)
+  )
+
+# gc()# clean ram
 
@@
   geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
-          aes(fill = total_spp),
+          aes(fill = avrg_spp_per),
           size = 0.1,col = "white") +
@@
              aes(start_lon,
-               start_lat,
-               col = per_plot),
+                 start_lat
+             ),
              size = 0.05,
@@
             show.legend = NA) +
+  # Closeup Rectangles 
+  geom_rect(data=Rectangles,
+            aes(xmin=x1, 
+                xmax=x2,
+                ymin=y1,
+                ymax=y2),
+            color="grey50", 
+            fill = NA,
+            linetype = "dashed"
+            ) +
   my_ggtheme_m() +
   scale_colour_viridis(
-  name="Average Gain of Stock-Share Ratio (%)\nand Number of taxa (n)",
+    name="Stock-Share Ratio Gain \nand taxa change (%)",
     limits = c(0,max(Map_data$per_plot)),
@@
   scale_fill_viridis(
-  name="Average Gain of Stock-Share Ratio (%)\nand Number of taxa (n)",
+    name="Stock-Share Ratio Gain \nand taxa change (%)",
     limits = c(0,max(Map_data$per_plot)),
@@
     # direction = -1
-) +
-ggsave("./Figures/fig_sup_1.png",
-       width = 12,
-       height = 8,
-       units = "in"
-)
+  ) #+
+  # ggsave("./Figures/fig_sup_1.png",
+  # width = 12,
+  # height = 8,
+  # units = "in"
+  # )
 
@@
   filter(
-    # sau_fishing_entity %in% Lat_am_land$sau_fishing_entity,
          !str_detect(sf_eez_name,"Pacific"),
@@
          )
-
-# Lat_am_eez %>% filter(str_detect(sf_eez_name,"Tri"))
 
@@
   coord_sf(xlim = c(-100, -55),ylim = c(9, 30)) + 
-  labs(x = "Lon", y = "Lat") +
+  labs(x = "", y = "") +
   scale_x_continuous(breaks = seq(-100,60,10)) +
-  scale_y_continuous(breaks = seq(10,30,5)) +
-  # my_ggtheme_m() +
+  scale_y_continuous(breaks = seq(10,30,10)) +
+  my_ggtheme_m() +
   theme(legend.position = "",
@@
   coord_sf(xlim = c(-3, 35),ylim = c(30, 45)) +
-  labs(x = "Lon", y = "Lat") +
+  labs(x = "", y = "") +
   scale_x_continuous(breaks = seq(-5,35,10)) +
   scale_y_continuous(breaks = seq(30,50,5)) +
-  # my_ggtheme_m() +
+  my_ggtheme_m() +
   theme(legend.position = "",
@@
     # direction = -1
-  );Med_map
+  )#;Med_map
 
@@
   coord_sf(xlim = c(-15, 30), ylim = c(45, 65)) +
-  labs(x = "Lon", y = "Lat") +
+  labs(x = "", y = "") +
   scale_y_continuous(breaks = seq(45,65,10)) +
-  # scale_x_continuous(breaks = seq(45,65,10)) +
-  # my_ggtheme_m() +
+  # scale_x_continuous(breaks = seq(45,65,5)) +
+  my_ggtheme_m() +
   theme(legend.position = "",
@@
     # direction = -1
-  );Eu_map
+  )#;Eu_map
+
+# Legend
+map_plot <- get_legend(Prop_map)
+
+Prop_map <- Prop_map + 
+  theme(legend.position = "")
 
 ## ------------------- ##
-# Final Figure
+# Render final plot
 ## ------------------- ##
 
-# gc()
-# plot_all <- ggdraw() +
-#       # map
-#       draw_plot(Prop_map, x = 0, y = 0.45, width = 1, height = 0.58) +
-#       # Lat America reg plor
-#       draw_plot(Lat_am_map, x = 0.1, y = 0.08, width = 0.3, height = 0.4) +
-#       # Europe reg plot
-#       draw_plot(Eu_map, x = 0.35, y = 0.08, width = 0.3, height = 0.4) +
-#       # Med regional plot
-#       draw_plot(Med_map, x = 0.65, y = 0.08, width = 0.3, height = 0.4) +
-#       draw_plot_label(label = c("a", "b"), 
-#                       size = 20,
-#                       x = c(0, 0), 
-#                       y = c(1, 0.45)
-#       ) #+
-#       # draw_plot_label(label = c("Revenue (USD) | Catch (Tonnes)","(Million)"), 
-#                       # size = 8,
-#                       # x = c(0.325,0.485), 
-#                       # y = c(0.31,0.29)
-#       )
-#     
-#     
-#     save_plot(
-#       # paste(Figures_Path,"100_25_Bins_Value_Region.png",sep=""),
-#       "test.png",
-#       plot_all,
-#       base_height = 8,
-#       base_width = 8
-#     )
+# Clean memory
+gc()
+
+# Draw figure
+plot_all <- ggdraw() +
+  # Main map
+  draw_plot(Prop_map, x = 0, y = 0.40, width = 1, height = 0.6) +
+  # Latin America and the Car.
+  draw_plot(Lat_am_map, x = 0.01, y = 0.02, width = 0.33, height = 0.45) +
+  # Europe
+  draw_plot(Eu_map, x = 0.33, y = 0.01, width = 0.33, height = 0.45) +
+  # The Med
+  draw_plot(Med_map, x = 0.66, y = 0.02 , width = 0.33, height = 0.45) +
+  # Legend
+  draw_plot(map_plot, x = 0.3, y = 0, width = 0.59, height = 0.1) +
+  # Labels
+  draw_plot_label(label = c("A", "B", "The Caribbean","North West Europe","The Mediterranean"),
+                  size = c(20,20,15,15,15),
+                  x = c(0.01,0.01,0.058,0.38,0.70),
+                  y = c(1,0.45,0.44,0.44,0.44)
+  )
+
+# Save plot
+save_plot(
+  "./Figures/figS3.png",
+  plot_all,
+  base_height = 10,
+  base_width = 12
+)
 
@@
 
+# Needs previous chunk
+
 Buble_data <- Eez_d_hist %>% 
@@
   summarise(mean_toe = mean(emerging_yr,na.rm=T)) %>% 
-  mutate(regional_average="average")
+  mutate(regional_average="regional average")
 
@@
   ) +
-  facet_wrap(~un_sub_region,scales = "free") +
+  facet_wrap(~un_sub_region,scales = "free_x") +
   scale_fill_viridis("Habitat preference",
                      discrete = T) +
-  scale_color_grey("Regional average") +
+  scale_color_grey("") +
   xlab("") +

### 2020-09-04 Scripts/Manuscript.Rmd
Sent first draft to W. up to results

@@
 Over the last century human activities have fundamentally changed the physicochemical composition of the ocean, warming the waters, increasing the pH and reducing the oxygen concentration [@IPCC:2019to].
-As a consequence, many marine species are changing their distribution [@ Poloczanska:2016kk] and will continue to do so in the search of the apropiate environmental conditions needed for their subsistence [@Cheung:2010dt.
+As a consequence, many marine species have changed their distribution [@Poloczanska:2016kk] and will continue to do so in the search of the appropriate environmental conditions needed for their subsistence [@Cheung:2010dt].
 Climate change is expected to impact fisheries production at different latitudes compromising our capacity to reach international sustainability goals such as the SDG 14 - life below water [@SDG14; @Cheung:2010dt; @Barange:2014kr].
@@
 
-The concept of shared stocks gained attention in the early 1980's when the United Nations officialized the concept of Exclusive Economic Zone (EEZs) though the Law of the Sea (UCLOS) [@UnitedNations:1986tl].
-The implementation of arbitrary jurisdictions across the global ocean created a partition in the distribution of several stocks, thus, creating what we known today as shared stocks.
-As defined by the FAO, shared stocks can be classified as transboundary, those that cross neighboring EEZs; straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and highly migratory, mainly tunas and bill-fishes that cover broad oceanic regions [@FAO; @Munro].
+In the early 1980's the United Nations officialized the concept of Exclusive Economic Zone (EEZs) though the Law of the Sea (UCLOS) [@UnitedNations:1986tl] partitioning the distribution of several species, thus, creating what we known today as shared stocks.
+As defined by the Food and Agriculture Organization, shared stocks can be classified as transboundary, those that cross neighboring EEZs; straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@FAO; @Munro].
 While counties are responsible for the management of stocks within their EEZs, UNCLOS incentives actions to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
@@
 There is evidence that such shifts have already complicated shared marine resources around the globe.
-In the North-eastern Atlantic, evidence suggest that shared stocks managed within the European Union divisions have been shifting since the 1970's compromising the region's sustainability [@EuropeShift].
+In the North-eastern Atlantic, evidence shows that shared stocks managed within the European Union divisions have been shifting since the 1970's compromising the region's sustainability [@EuropeShift].
 In the North-eastern Pacific, sockeye salmon (*Oncorhynchus nerka*) shifted in the 1980's towards Alaska, creating a ten years conflict between Canada and the United States [@Song].
-In the South-eastern Pacific, the southern stock of Peruvean anchoveta (*Engraulis ringens*) shared by Chile and Peru has gradually shifted south (towards Chilean waters) adding to the over-exploitation of the resource [@Cashion:2018cg].
+Further south, the southern stock of Peruvean anchoveta (*Engraulis ringens*) shared by Chile and Peru has gradually shifted towards Chilean waters adding to the over-exploitation of the resource [@Cashion:2018cg].
 Global shifts in the distribution of tunas [@ErauskinExtramiana:2019fw] have brought additional management challenges in both the Western Indian ocean and South-western Pacific [@Miller:2007kx].
 Moreover, as species continue shifting with climate change, current policies might become obsolete, and thus, require new management strategies that create equitable systems across all dependent fishing nations [@Bell:2013ga; @Pinsky:2018cb].
-We argue that two concepts will be essential in addressing the challenges that climate change will bring to the management of transboundary fisheries.
+The objective of the current paper is to explore two concepts that will be essential in addressing the challenges that climate change will bring to the management of transboundary fisheries [@Link:2010ei].
 
 The first concept alludes to the game theory concept of thread point, known as the minimum payoff a player is willing to receive in order to cooperate with others.
-Thus, if a player's future payoff drops below the thread point, such player will likely break off any collaboration agreement and proceed to act solo [@Sumaila:E&S].This can be seen as the proportion of a species distribution that each neighboring nation has (or the quota allocation that each one gets) at the time of signing off an agreement [@PalaciosAbrantes:E&S].
-Many shared fisheries operate this way [@EuropeShift; @GoMA].
+If a player's future payoff drops below the thread point, such player will likely break off any collaboration agreement and proceed to act solo [@Sumaila:E&S].
+A thread point can be translated to the proportion of a species distribution that each neighboring nation has (or the quota allocation that each one gets) at the time of signing off an agreement [@PalaciosAbrantes:E&S].
+Many shared fisheries operate this way [@EuropeShift; @GoMA; @NOAA].
 If not addressed, future changes in the thread point (hereafter called Stock Share Ratio -SSR) of jointly managed fisheries can have adverse consequences for shared fisheries [@PalaciosAbrantes:E&S; @Sumaila:E&S].
@@
 Such concept has been used to look at the emergence of diverse oceanic variables [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg] and ecological consequences of marine heatwaves [@Cheung:2020ja].
-Similarly, climate change is expected to drive biomass further away from the historic range limits of a species [@Alexa?].
-Assessing the ToE of shared stocks is as important as the change itself as it could reduce potential international conflicts [@Spijkers:2017ij] and prepare for emerging fisheries [@Pinsky:2018cb;  @Pinsky:2014by].
+Similarly, climate change is expected to drive species further away from their historic range limits [@Alexa?].
+Assessing the ToE of shared stocks is as important as the change itself, as it could reduce potential international conflicts [@Spijkers:2017ij] and aid the strengthen of marine governance [@Pinsky:2018cb;  @Pinsky:2014by; @Link:2010ei].
 
@@
 
-The Food and Agricultural Organization (FAO) of the UN define transboundary stocks as those that freely swim across the EEZ of neighboring nations [@REFFAO].
-Accordingly, we included in this analysis 633 marine exploited transboundary species representing 80% of the catch taken from the world's EEZs [@Palacios-Abrantes:xxxx].
-We consider each species as a single stock unit within an EEZ due to a lack of more detailed spatial information [@Teh:2015gd; @Palacios-Abrantes:xxxx].
+We included in this analysis 633 marine exploited transboundary species representing 80% of the catch taken from the world's EEZs between 2005 and 2014 [@Palacios-Abrantes:xxxx].
+Each species in within an EEZ was considered a single stock unit due to a lack of more detailed spatial information [@Teh:2015gd; @Palacios-Abrantes:xxxx].
 We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the *Sea Around Us* shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic) including islands territories, and determined the intersections between polygons using the R package *sf* (**Supplements Fig 1 (SuppM)**) [@PackagesfSimpleF:2018vp].
@@
 background noise).
-At the same time, by having multiple initial-condition ensemble simulations of the GFDL-ESM2M we were able to remove the natural variability of a stock's distribution by averaging the secular trend over all ensemble members [@Rodggers:2015jq].
-
+At the same time, by having multiple initial-condition ensemble simulations of the GFDL-ESM2M we were able to remove the natural variability of a stock's distribution by averaging the secular trend over all ensemble members (See Estimating the Time of Emergence).
 ## Changes in Stocks Share Ratio
@@
 We adopted the concept of threat point as the minimum proportion of the stocks’ joint distribution between neighboring EEZs that each country will accept if they were to cooperate in the management of the stock [@Palacios-Abrantes:E&S].
-For each stock and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighboring EEZs, and then calculated the proportion that each EEZ had [@Palacios-Abrantes:xxxx].
+For each stock and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighboring EEZs, and then calculated the proportion that each EEZ had [@FishForVisa].
 We then average the results in three time periods in order to reduce climate variability within the projections.
@@
 We choose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing) and SDG 1.2 (poverty reduction), aimed to be achieved by 2020 and 2030, respectively [@Singh:2017ds].
-We then estimate the percentage change in the SSR of each future time period relative to the historic time period.
-To address the natural spatial variability of the stock we only considered a SSR change when the future SSR was higher or lower than the average historic SSR $\pm$ 1 s.d.
-The analysis was replicated for each of the ten ensemble members capturing variation by averaging results among ensemble members (mean $\pm$ standard deviation).
+The analysis was replicated for each of the ten ensemble members capturing the variation by averaging results among ensemble members ($\mu \pm \sigma$).
+We then estimate the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_{f}$) relative to the historic time period ($SSR_h$) as follows,
 
-## Time of Emergence
+$$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$
+To address the natural spatial variability of the stock we only considered $\Delta{SSR_{e,m}}$ to be true when $SSR_f \geq (SSR_f + \sigma)$ or $SSR_f \leq (SSR_f - \sigma)$.
+
+## Estimating the Time of Emergence
 
 Knowing the time (e.g.
-year) in which a shared stock is going to diverge from a historical variability can be key for preparing to shifting stocks.
-We employed the ten GFDL ensemble members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
+year) in which a shared stock is going to diverge from a historical variability is key for preparing to shifting stocks [@Link:2010ei].
+We employed the DBEM and the ten GFDL ensemble members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
 The index is based on the variation of a species distribution centroid in relation to the centroid of the neighboring EEZs it occupy.
-Each species centroid was determined by the average latitude and longitude of the grid cells containing the top 10% abundance percentile of each year.
+Each species centroid was determined by the average latitude and longitude of the grid cells containing the top 10% abundance percentile of each year derived from the DBEM.
 The EEZ centroid was estimated using the *st* package of *R* statistical software (**Supplements Fig 1 (SuppM)**).
@@
 
-$$TI =(\frac{D_{A,t}}{sd(D_{A,h})} - \frac{D_{B,t}}{sd(D_{B,h)}})^2$$
+$$TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2$$
 
-where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectively for each future time step from 2006 to 2100 (*t*); *sd* is the standard deviation of the historical (*h*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectively for each future time step from 2006 to 2100 (*t*); *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
 The index was then smoothed to a 10 years average.
@@
 Since this was an arbitrary decision, we have included a more conservative estimation ($\mu \pm 2*\sigma$ for a 95% confidence interval).
-The SNR for both bottom and surface sea temperature was also estimated following the SNR method.
+The SNR for both bottom and surface sea temperature was also estimated for statistical analysis following the SNR method.
 
@@
 
-We tested data for normality (e.g.
+We tested the results for normality (e.g.
 skewness, kurtosis) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary stocks.
@@
 
-We estimated changes in the SSR of transboundary species shared by neighboring EEZs by early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
+We estimated the change in SSR (larger than 1 s.d) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
 Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
 1A).
-While most EEZs will experience an average change lower than 25%, a few EEZs are expected to have larger changes in SSR.
-Such change is expected to slightly increase in intensity towards the  mid $21^{st}$ century (**Supplements Fig 1 (histogram)**).
+Most EEZs will average a change lower than 25% which is expected to slightly increase in intensity towards the  mid $21^{st}$ century (**Supplements Fig 1 (histogram)**).
+However, this result masks larger changes in SSR for specific species.
 For example, Guatemala's SSR of *Panulirus gracilis* shared with Mexico is projected to go from 12% to 26%.
-The number of stocks per EEZ that will see a change in SSR by early $21^{st}$ century is relatively low (median of 13 spp.
-per EEZ), however, some regions like West Africa and the Mediterranean will see changes in over 50 species (e.g.
-Senegal, 94).
-These regions, where EEZs are relatively small and have multiple borders will see a complex dynamic of changes in SSR with no clear patterns (Fig.
+By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR >< 1 s.d.
+of historical distribution will be relatively large (global mean of 69% of species in an EEZ).
+For 17 EEZs, 100% of their analyzed transboundary stocks will shift, while and only six EEZs will have less than 25% shifting stocks (Fig.
+1A-Polygons).
+Regions where EEZs are relatively small and have multiple borders (e.g.
+The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Fig.
 1B).
-Towards the mid $21^st$ century, changes in SSR per EEZ will no be substantial (**Supplements Fig 1 (map)**) relative to early century, but are expected to increase in intensity (**Supplements Fig 1 (histogram)**).
+Towards the mid $21^{st}$ century, changes in SSR per EEZ will no be substantial (**Supplements Fig 1 (map)**) relative to the early $21^{st}$ century but are expected to increase in intensity (**Supplements Fig 1 (histogram)**).
 
@@
 ![Figure 1.
-Average gain of transboundary stock share ratio (lines) and number of shifting stocks (polygon) by 2030 (2021-2040) relative to 1951-2005.
-Arrows go from the losing EEZ (point) to the winner (arrowhead).
-Panel B zooms in to specific areas.](../Figures/Fig1.png)
+Changes in Stock Share Ration by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial stocks that will see a shift.
+Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Fig1.png)
 
@@
 
-We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock's distribution overpasses the historical variation.
+We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock's distribution (2006-2100) overpasses the historical variation (1951-2005).
 Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g.
 *Engraulis ringens* shared by Chile and Peru emerged in 2010).
-The cumulative sum of emerging stocks will have a steeper increase until 2025 where the increment will be slightly less intense (Fig.
+Stocks will continue to emerge steadily towards 2097 when all 633 stocks will have had emerged in at least one neighboring EEZ (Fig.
 2).
-Similarly, for the number of EEZs, the increment curve seem to flatten by 2020.
-By 2081, all of the EEZs analyzed here will have at least one emerging stock while 2097 will be the year where all 633 stocks have emerged in at least one neighboring EEZ (Fig.
+The number of EEZs with emerging stocks will increase steeply until 2020 when it seems to flatten approaching 2081 when all of the EEZs analyzed here will have at least one emerging stock (Fig.
 2).
@@
 
-![Figure 2 Cumulative increase in time of emergence by EEZ and stock](../Figures/Fig2.png)
+![Figure 2. Cumulative increase in time of emergence by Exclusive Economic Zone and transbounudary stock.](../Figures/Fig2.png)
 
-Most of the world EEZ’s will see an average ToE across transboundary stocks by 2030 with a few EEZs having a relative late average ToE (e.g.
-Honduras average ToE is 2071) and 54 EEZs having already emerged (e.g.
-average ToE <= 2020).
+Most of the world EEZ’s will see an average ToE across transboundary stocks before 2040 (Fig.
+3).
+In a few EEZs stocks will have a relative late average ToE (e.g.
+Honduras average ToE is 2071) while 54 of the 280 EEZs (19%) have already seen their stocks emerged (e.g.
+average ToE across all stocks <= 2020).
 A higher number of emerging stocks will happen between latitudes 0$^o$ and 50$^o$N corresponding to regions of relatively small EEZs with multiple borders (e.g.
 The Caribbean and the Mediterranean) (Fig 2).
-In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50%.
+In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50% (Fig.
+2 - Polygons).
 However, in some cases emerging stocks represent 85% of total fishing revenue (e.g.
@@
 ![Figure 3.
-Average time of emergence (ToE) per Exclusive Economic Zone (EEZ polygon) and participation of emerging stocks in total EEZ fishing revenue (land polygon).
-Darker tones represent low participation and late ToE; lighter colors represent large participation and earlier ToE.
-Density plot on right shows the number of emerging stocks across EEZ centroids (10^o running mean).](../Figures/Fig3.png)
+Average time of emergence of transboundary stock.
+Land polygon represents  the current participation of emerging stocks in the country's fishing revenue.
+Exclusive Economic Zone polygon represents the average time of emergence across all stocks.
+Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean).](../Figures/Fig3.png)
 
-Most fisheries within the EEZ of the main transboundary fishing countries in terms of revenue (e.g.
-the top $75^{th}$ percentile) will emerge after 2030 (n = 28).
-The remaining twelve countries will have an average ToE before 2030, this includes the United States who is the $4^{th}$ place in total revenue (Fig.
+Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transbondary stocks between 2005 and 2010 will see an average ToE before 2030 (Fig.
 4).
+This includes the United States who occupies the $4^{th}$ place.
 These 47 countries are also among those with the largest number of emerging taxa.
@@
 From the top ten countries with more revenue from transboundary species, nine of them share a border.
-Most of these countries (e.g.
-Spain and China) will also have a large number of emerging stocks, while Peru and Norway will have a small number of emerging taxa but representing high revenue fisheries.
+Most of these countries (e.g.China) will also have a large number of emerging stocks, while others like Peru and Norway will have a small number of emerging taxa but representing high revenue fisheries.
 
@@
 Average time of emergence (ToE) per fishing entity.
-Fishing revenue in logarithmic scale Showing names for the top 75th percentile).
-Countries marked with an asterisk have at least one border.](../Figures/Fig4.png)
+Fishing revenue in logarithmic scale.
+Showing names for the top 75th percentile.
+Countries marked with an asterisk have at least one neighbouring border.](../Figures/Fig4.png)
 
-We compared species ToE within EEZs by geographic region and by species habitat preference.
-The mean ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001) (Fig.
+We compared species ToE within EEZs by geographic region and by species habitat preference (Fig.
 5).
-Post-hoc tests revealed that EEZs in Easter Asia which have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ from (almost) all other regions.
+The median ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001).
+Post-hoc tests revealed regional patterns.
+EEZs in Easter Asia which have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ from (almost) all other regions.
 Other regions also significantly differ from each other but in a less frequent manner (*See Table Sup X for full statistics*).
-When looking specifically at habitat association, we found significant differences on ToE of different habitat types (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
+When looking specifically at habitat association, we found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
 Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (*See Table Sup XX for full statistics*).
@@
 We hypothesized that species environmental preferences (e.g.
-presence in certain biomes and habitat preferences), climate change effects on water temperature (e.g.
+presence in certain biomes and habitats), climate change effects on water temperature (e.g.
 ToE of SST), human drivers (e.g.

### 2020-09-18 Scripts/Initial_Analysis.Rmd
Drafted a part of the discussion

@@
   select(eez_name,longitude = lon,latitude = lat)
+
+# Join all data for analysis
 
@@
 # Model 1; all variables
-M1 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)
+M1 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Main_data)
 
@@
 # Model 2; non demPel
-M2 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel, data = Lm_data)
+M2 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel, data = Main_data)
 
@@
 # Model 3; non biome
-M3 <- lm(emerging_yr ~ sst_toe + bot_toe + status + dem_pel + DemersPelag, data = Lm_data)
+M3 <- lm(emerging_yr ~ sst_toe + bot_toe + status + dem_pel + DemersPelag, data = Main_data)
 
@@
 # Model 4; non status
-M4 <- lm(emerging_yr ~ sst_toe + bot_toe + biome + dem_pel + DemersPelag, data = Lm_data)
+M4 <- lm(emerging_yr ~ sst_toe + bot_toe + biome + dem_pel + DemersPelag, data = Main_data)
 
@@
 # Model 5; non sst
-M5 <- lm(emerging_yr ~ status + bot_toe + biome + dem_pel + DemersPelag, data = Lm_data)
+M5 <- lm(emerging_yr ~ status + bot_toe + biome + dem_pel + DemersPelag, data = Main_data)
 
@@
 # Model 6; all + basin
-M6 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)
+M6 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Main_data)
 
@@
 # Model 7; Categories instead of status
-M7 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + dem_pel + DemersPelag, data = Lm_data)
+M7 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + dem_pel + DemersPelag, data = Main_data)
 
@@
 # Model 8; All variables + ccentroid lat
-M8 <- lm(emerging_yr ~ sst_toe + bot_toe + category + as.factor(biome) + DemersPelag + latitude, data = Lm_data)
+M8 <- lm(emerging_yr ~ sst_toe + bot_toe + category + as.factor(biome) + DemersPelag + latitude, data = Main_data)
 
@@
 # Model 9; Include UN regions
-M9 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + DemersPelag + latitude + un_region, data = Lm_data)
+M9 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + DemersPelag + latitude + un_region, data = Main_data)
 

### 2020-09-18 Scripts/Manuscript.Rmd
Drafted a part of the discussion

@@
 ---
-title: Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for
-  Policy Response Under Climate Change
-author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, U. Rashid Sumaila^1^,  Vicky Lam^1^, David VanderZwaag^3^, and William W.L. Cheung^1^'
-date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British
-  Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute,
-  University of Bern. ^3^Schulich School of Law, Dalhousie University'
+title: Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change
+author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^'
+date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern.'
 output:
@@
 At the same time, by having multiple initial-condition ensemble simulations of the GFDL-ESM2M we were able to remove the natural variability of a stock's distribution by averaging the secular trend over all ensemble members (See Estimating the Time of Emergence).
+
 ## Changes in Stocks Share Ratio
@@
 
-We estimated the change in SSR (larger than 1 s.d) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
+We estimated the change in SSR (larger than 1 s.d) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) under a high emission climate change scenario.
 Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
@@
 
-Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transbondary stocks between 2005 and 2010 will see an average ToE before 2030 (Fig.
+Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see an average ToE before 2030 (Fig.
 4).
@@
 
-<!-- # Discussion -->
+# Discussion
 
-<!-- The combination of the time of emergence and the revenue participation of the species could shed a light in the importance of the time of emergence in terms of management.
-For example, we identified that in Peru 12 species will emerge by 2038 in average, representing 92% of the country's fishing value.
-Moreover, our results show that Peruvean anchoveta, the second largest fishery of the world [@SOFIA] and the main fishery in Peru [@REF] have already emerged (2010), with its stock share ratio benefiting Chile (Fig.
+We found no clear pattern in the direction of change of future SSR under a high emission climate change scenario by 2030 relative to present (Fig.
 1).
--->
+Previous modelling exercises predicted an overall poleward shift in species distributions towards the end of the $21^{st}$ century [@Cheung; @Poloczanska:2016kk; @ETC].
+Poleward shifts in SSR can be seen in some regions of the world with relative large EEZs and few boundaries like the North East Pacific.
+However, even here, Alaska's EEZ is expected to loose to Canada in the south.
+This is likely the effect of the geographical barriers [@Cheung:2015fm; @Kleisner:2016je] as fish cannot move north due to the land mass [@Cheung:2015fm].
+Contrarily, the Eastern Tropical Pacific will see an equatorial shift with EEZs from Mexico to Ecuador loosing SSR to their southern neighbors.
+This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will shift between 50 and 100 km southward of Mexico's EEZ towards Peru under a high emission climate change scenario [@TayClarkeChp2].
+Similarly, poleward shifts patterns are hard to identify in regions where EEZs are relatively small like the south west Atlantic.
+The influence of local processes such as upwelling systems [@REF; @TayClarkeChp2] and local currents [@REF; @TayClarkeChp2] could be influencing the re-distribution of marine species in these two regions.
+Finally, in regions where EEZs are both small and with multiple borders like the Caribbean or North East Atlantic, changes in SSR will have no unidirectional pattern but rather present a complex network of changes.
+Despite these regions been already highly connected by both fish adult migration [@REF] and larvae dispersal [@Ramesh:2019va], our results suggest that distributional changes will still happen influencing the SSR of local fisheries.
 
-<!-- Form all significantly different variables, only latitudinal gradient postpone the ToE suggesting that the ToE will happen first at equatorial latitudes and later towards the poles. -->
+Our results suggest that most of the EEZs of the world and more than half of the transboundary species already emerged since the early 2000s (Fig.
+2).
+Shifting species distribution  has been documented in several regions and for different species [@Poloczanska:2016kk].
+In the early 2000s Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range reaching Californian waters in the Eastern North Pacific [@Zeidberg:2007gs].
+In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, lead to years of unsettled international conflict between fishing nations [@Spijkers:2017ij].
+The large amount of transboundary species emerging since early 2000s aligns with the time of emergence of several environmental variables (e.g.
+temperature, oxygen, $CO_2$) that influence the distribution of marine species [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
+It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 22.6% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while $O_2$ will have doubled [@Rodggers:2015j].
 
-<!-- Historically, environmentally-driven shits in shared stocks have cause problems and threaten stock's sustainability.
-**In early 1980's Canada and the US blablabla.** Many fish stocks managed by the European Union ICES regultaroy areas have been shifting since the 1970's [@EuropeShift].
-Such shifts have already result in confrontations between nations as evidenced by the expansion of the Atlantic mackerel (*Scomber scombrus*) distribution reaching the EEZ of Iceland [@Spijkers:2017ij].
-More recently, Chile and Peru have declare a cooperation strategy to mange the transboundary stock of Peruvean Anchovera [@Peru].
-Furthermore, as we continue to enter the $21^{st}$ century, commercial marine species are expected to continue shifting (mostly poleward) potentilly compromising shared stocks management.
+Regional differences in ToE are likely the result of environmental factors and the characteristics of political boundaries.
+The tropics are the most species-rich biome of the global ocean [@Reygondeau:2019no] and showed the larger number of emerging species, specifically between latitudes $10^o$S and $45^o$N (Fig.
+3).
+Historical trends suggest that both tropical Atlantic and Pacific oceans have experienced the greatest levels of warming since 2006 [@IPCC:2019tn; @Roemmich:2015id].
+Similarly, equatorial regions have potentially experienced some of the largest declines in $O_2$ [@IPCC:2019tn].
+Consequently, species living in these habitats that cannot cope with changing oceans will have too move towards more suitable waters [@Poloczanska:2016kk].
+In addition to the environmental pressure from climate change, latitudes $10^o$S and $45^o$N comprise areas with a complex system of small EEZs with many borders (e.g.
+The Caribbean and the South China Sea).
+The time and number of emerging stocks decreases as we move away from the equator and into temperate-polar biomes with only a peak related to the Baltic and North seas.
+
+The combination of the ToE and the participation of the emerged species to fishing revenue could be seen as a metric to measure the vulnerability of transboundary fisheries to climate change (Fig.
+3).
+The mismatch between species distributional change across international jurisdictions and regulations emplaced [or lack of thereof] can contribute to future over-exploitation of transboundary species and initiate international conflict [@PalaciosAbrantesOC; @SumauilaOC; @Spijkers:2017ij].
+Thus, the sooner a transboundary species emerge, the more vulnerable its fisheries will be as policy might not be prepare to respond to the change [@Pisnky2018] nor move fast enough to cope with the emerging species [@Pinsky:2012kq; @EuropeShift].
+In European waters, quota is allocated following a fix historical share based on the distribution of the 1970's.
+However, as species change distributions, the quota system has become outdated, potentially generating economic losses to the continent.
+Moreover, if the species has a large participation on a country or region fisheries revenue or catch, the management complications could potentially lead to larger economic and food security impacts.
+Polynesia (2019) and Micronesia (2021) the regions with the earliest average time of emergence (Fig.
+5) are a case of early ToE and highly important fisheries.
+Countries in these regions are highly dependent of pelagic-oceanic species (e.g.
+tunas) and reef-associated species (e.g.
+*Speceies*) for both food security [@Cabral:2018hc] and profits [@].
+The prompt ToE of species with a high social value makes this region particularly vulnerable to the impacts of climate change to transbounudary species.
+
+<!-- Furthermore, as we continue to enter the $21^{st}$ century, commercial marine species are expected to continue shifting (mostly poleward) potentilly compromising shared stocks management.
 By the end of the century, all EEZs of the world are proyected to have at least one new transboundary species with some of them having over 10 new stocks [@Pinsky].
@@
 
+# Caveats
+
 <!-- Conclusion -->
@@
 <!-- Preparing policy to species on the move is key to achieve sustainable development goals and proper governance of the world fisheries [@Pinksy]. -->
-
-# Recycling text
-
-Ignoring anthropocentric effects, a species biomass distribution follows a variation in time and space according to environmental preferences, species interactions and landscape features [@Huch].
-Thus, a neighboring nation will not necessarily have, at all times  (e.g.
-months or years), a fixed SSR but rather a series of SSRs [@GoMA].

### 2020-09-22 Scripts/Initial_Analysis.Rmd
Vulnerabillity index

@@
 
+```{r vulnerabillity_index, eval = T}
+
+# Per EEZs
+Eez_d_hist <- Eme_Res %>% 
+  group_by(eezid ,sau_fishing_entity,fishing_entity_id,tresh) %>% 
+  # summarise(n_tax = length(unique(taxon_key))) %>% 
+  summarise(
+    mean_yr = mean(emerging_yr,na.rm=T),
+    sd_yr = sd(emerging_yr,na.rm=T),
+    n_taxa = length(unique(taxon_key))
+    ) %>% 
+  gather("metric","value",mean_yr:n_taxa) %>% 
+  # filter(metric != "n_taxa") %>% 
+  mutate(
+    reverse = abs(value-2100)  + 2000
+  ) %>% 
+  filter(
+    tresh == "Tresh_one",
+    metric == "mean_yr"
+  )
+
+Value_map_data <- Eme_value_data %>% 
+  left_join(Total_value,
+            by = c("fishing_entity_id","variable")) %>% 
+  mutate(participation = (sum_value/tot_value)*100) %>% 
+  filter(
+    tresh == "Tresh_one",
+    variable == "value"
+  )
+
+```
+
 ```{r emergence_map, eval = T}
@@
   mutate(participation = (sum_value/tot_value)*100)
+
+### ----------------- ###
+### Index alternative ###
+### ----------------- ###
+
+Index_data <- Eez_d_hist %>% 
+  left_join(Value_map_data) %>% 
+  mutate(
+    index = (participation/value),
+    index = ifelse(is.na(index),0,index)
+  ) %>% 
+  select(1:3,value,participation,index,everything())
+
+##### Normalize index
+# https://www.datanovia.com/en/blog/how-to-normalize-and-standardize-data-in-r-for-great-heatmap-visualization/
+
+# Get min and maximums
+min_index <- min(Index_data$index)
+max_index <- max(Index_data$index)
+
+Index_data <- Index_data %>% 
+  mutate(
+    index = 2000 + (index-min_index)*(2100-2000)/(max_index-min_index)
+  )
+
+Value_map_data <- Index_data %>% 
+  select(sau_fishing_entity,fishing_entity_id,tresh,variable,participation = indexb)
+
+### ----------------- ###
 
@@
       aes(
-        fill = reverse
+        fill = reverse #original map
+        # fill = value
       ),
@@
     geom_sf(data = World_map_part,
-            aes(fill = participation+2000),
+            # aes(fill = participation+2000), #original map
+            aes(fill = participation),
             color =  "white",
@@
              y = -69,
-             label = "Participation (%)", 
+             label = "Vulnerabillity index",
              color="black",
@@
              y = -85,
-             label = seq(2100,2000,-25), 
+             label = seq(2100,2000,-25), # original
+             # label = seq(2000,2100,25), 
              color="black", 

### 2020-09-22 Scripts/Manuscript.Rmd
Discussion draft

@@
 Similarly, climate change is expected to drive species further away from their historic range limits [@Alexa?].
-Assessing the ToE of shared stocks is as important as the change itself, as it could reduce potential international conflicts [@Spijkers:2017ij] and aid the strengthen of marine governance [@Pinsky:2018cb;  @Pinsky:2014by; @Link:2010ei].
+Assessing the ToE of shared stocks is as important as the change itself, as it could reduce potential international conflicts [@Spijkers:2017ij] and aid the strengthen of marine governance [@Pinsky:2018cb;  @Pinsky:2014by; @Link:2010ei; @Fogarty].
 
@@
 
-We estimated the change in SSR (larger than 1 s.d) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) under a high emission climate change scenario.
+We estimated the change in SSR (with a 63% of confidence) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) under a high emission climate change scenario.
 Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
@@
 
-We found no clear pattern in the direction of change of future SSR under a high emission climate change scenario by 2030 relative to present (Fig.
+In this paper we endeavour to look at the changes in the sharing proportion of transboundary stocks and identify the year by witch future stocks' distribution whill emerge from historical variabillity.
+Our findings here are noteworthy in several ways.
+First, we identified that the historical share proportion of transboundary stocks will change by as soon as 2030 in most neighbouring EEZs (Fig.
 1).
-Previous modelling exercises predicted an overall poleward shift in species distributions towards the end of the $21^{st}$ century [@Cheung; @Poloczanska:2016kk; @ETC].
-Poleward shifts in SSR can be seen in some regions of the world with relative large EEZs and few boundaries like the North East Pacific.
-However, even here, Alaska's EEZ is expected to loose to Canada in the south.
-This is likely the effect of the geographical barriers [@Cheung:2015fm; @Kleisner:2016je] as fish cannot move north due to the land mass [@Cheung:2015fm].
-Contrarily, the Eastern Tropical Pacific will see an equatorial shift with EEZs from Mexico to Ecuador loosing SSR to their southern neighbors.
-This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will shift between 50 and 100 km southward of Mexico's EEZ towards Peru under a high emission climate change scenario [@TayClarkeChp2].
-Similarly, poleward shifts patterns are hard to identify in regions where EEZs are relatively small like the south west Atlantic.
-The influence of local processes such as upwelling systems [@REF; @TayClarkeChp2] and local currents [@REF; @TayClarkeChp2] could be influencing the re-distribution of marine species in these two regions.
-Finally, in regions where EEZs are both small and with multiple borders like the Caribbean or North East Atlantic, changes in SSR will have no unidirectional pattern but rather present a complex network of changes.
-Despite these regions been already highly connected by both fish adult migration [@REF] and larvae dispersal [@Ramesh:2019va], our results suggest that distributional changes will still happen influencing the SSR of local fisheries.
+Second, we show that the current distribution of most transboundary stocks have already emerged from the historical record and that virtually all EEZs have seen at least one emerging stock since 2006 (Fig.
+2).
+Thirdly, we identify regional patterns of complex changes in SSR and potentially high vulnerabillity for transbounudary fisheries management under climate change.
 
+By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Fig.
+1).
+Poleward shifts in SSR will be seen in regions of the world with relative large EEZs and few boundaries like the North East Pacific.
+However, despite a general recognition that climate change is shifting species distributions poleward [@Cheung; @Poloczanska:2016kk; @IPCC], the Eastern Tropical Pacific will see an equatorial shift.
+Here, the EEZs from Mexico to Ecuador will loose SSR to their southern neighbors.
+This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico's EEZ towards Peru under a high emission climate change scenario [@TayClarkeChp2].
+Moreover, regions with presence of multiple frontiers like the Caribbean and the North East Atlantic will see a complex network of SSR change with no marked direction.
+These regions are known for beign highly connected by both fish adult migration [@REF] and larvae dispersal [@Ramesh:2019va] leading to joint management plans [@EUFisheries] and the call for new implemented shared policies [@McNananman].
+However, game theory suggest that the more number of parties negotiating, the harder it is for parties to reach an agreement [@Sumaila:2013], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.
+
+We set out to evaluate the emergence characteristics of transboundary stocks across the global EEZs.
 Our results suggest that most of the EEZs of the world and more than half of the transboundary species already emerged since the early 2000s (Fig.
 2).
-Shifting species distribution  has been documented in several regions and for different species [@Poloczanska:2016kk].
-In the early 2000s Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range reaching Californian waters in the Eastern North Pacific [@Zeidberg:2007gs].
+This is in alignment with previous research showing that the change in the composition of marine catches since 1970's is an evidence that climate change is already acting uppont marine fishes [@Cheung:2013gk].
+Indeed an expansion in the distribution of multiple marine species has been documented in several regions of the world [@Poloczanska:2016kk].
+For example, in the early 2000s Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range reaching Californian and latter Washington waters in the Eastern North Pacific [@Zeidberg:2007gs] where a new fishery rapidly developed [@Pinsky:2014by].
 In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, lead to years of unsettled international conflict between fishing nations [@Spijkers:2017ij].
-The large amount of transboundary species emerging since early 2000s aligns with the time of emergence of several environmental variables (e.g.
+The large amount of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g.
 temperature, oxygen, $CO_2$) that influence the distribution of marine species [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
@@
 
-Regional differences in ToE are likely the result of environmental factors and the characteristics of political boundaries.
-The tropics are the most species-rich biome of the global ocean [@Reygondeau:2019no] and showed the larger number of emerging species, specifically between latitudes $10^o$S and $45^o$N (Fig.
+Regional differences in ToE are likely the result of environmental factors, species vulnerabillity, and the characteristics of political boundaries.
+The tropics are the most species-rich biome of the global ocean [@Reygondeau:2019no] presenting a larger number of emerging species, specifically between latitudes $10^o$S and $25^o$N (Fig.
 3).
-Historical trends suggest that both tropical Atlantic and Pacific oceans have experienced the greatest levels of warming since 2006 [@IPCC:2019tn; @Roemmich:2015id].
-Similarly, equatorial regions have potentially experienced some of the largest declines in $O_2$ [@IPCC:2019tn].
-Consequently, species living in these habitats that cannot cope with changing oceans will have too move towards more suitable waters [@Poloczanska:2016kk].
-In addition to the environmental pressure from climate change, latitudes $10^o$S and $45^o$N comprise areas with a complex system of small EEZs with many borders (e.g.
-The Caribbean and the South China Sea).
-The time and number of emerging stocks decreases as we move away from the equator and into temperate-polar biomes with only a peak related to the Baltic and North seas.
+Equatorial regions have potentially experienced some of the largest declines in $O_2$ [@IPCC:2019tn] and historical trends suggest that both tropical Atlantic and Pacific oceans have experienced the greatest levels of warming since 2006 [@IPCC:2019tn; @Roemmich:2015id].
+Here, marine species live close to the limits of their termal tolerance, making them highly vulnerable to warming waters [@IPCC] and consequently having to move towards deeper or sub-tropical colder waters [@Poloczanska:2016kk].
+Thus, the high number of stocks emerging in the tropcis (e.g.
+between latitudes $10^o$S and $45^o$N) could be a consequence of tropical species expaning their distributions towards sup-tropical waters.
+This is supported by observational evidence showing that the majority of fish first sightings (e.g.
+the first time a fish was recorder outside its known distribution) were tropical and subtropical species in suptropical biomes [@Fogarty:2017eo].
+Similarly, catch-data analysis suggests a reduction in subtropical species cathes in tropical waters between 1970 and 2006 evidencing a polweard shift in subtropical species [@Cheung:2013gk].
+Moreover, the earlier emergece of tropical regions (e.g.
+Fig.
+5 - Micronesia, Polynesia and South America and the Caribbean) supports the premise that such effect could be alredy happening [@Fogarty:2017eo; @Cheung:2013gk].
+In addition to species natural traits and the environmental pressure from climate change, latitudes $10^o$S and $45^o$N comprise areas with a complex system of relative small EEZs with many borders that match peaks of emerging stocks.
+Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsable for substantial increases in emerging stocks (Fig.
+3).
+The time and number of emerging stocks decreases as we move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
+
+## Vulnerabillity index-nonIndex
 
@@
 
-<!-- Furthermore, as we continue to enter the $21^{st}$ century, commercial marine species are expected to continue shifting (mostly poleward) potentilly compromising shared stocks management.
-By the end of the century, all EEZs of the world are proyected to have at least one new transboundary species with some of them having over 10 new stocks [@Pinsky].
-Consequently, as species occupy new waters they leave historic areas and so, xx% of species are expected to exit EEZs by the end of che century [@Oremus].
-This entering and exicting national jurisdicions will complicate fisheries management potentially creating confllict between neighboring nations [@Pinksy @Oremus:2020id].
--->
+## Implications for management
+
+The early detection of fish moving to new regions is key for fisheries management [@Fogarty]
+
+## Potential solutions
 

### 2020-09-23 Scripts/Initial_Analysis.Rmd
First discussion draft

@@
 
-###### Just mean time
-
-```{r vulnerabillity_index, eval = T}
-
-# Per EEZs
-Eez_d_hist <- Eme_Res %>% 
-  group_by(eezid ,sau_fishing_entity,fishing_entity_id,tresh) %>% 
-  # summarise(n_tax = length(unique(taxon_key))) %>% 
-  summarise(
-    mean_yr = mean(emerging_yr,na.rm=T),
-    sd_yr = sd(emerging_yr,na.rm=T),
-    n_taxa = length(unique(taxon_key))
-    ) %>% 
-  gather("metric","value",mean_yr:n_taxa) %>% 
-  # filter(metric != "n_taxa") %>% 
-  mutate(
-    reverse = abs(value-2100)  + 2000
-  ) %>% 
-  filter(
-    tresh == "Tresh_one",
-    metric == "mean_yr"
-  )
-
-Value_map_data <- Eme_value_data %>% 
-  left_join(Total_value,
-            by = c("fishing_entity_id","variable")) %>% 
-  mutate(participation = (sum_value/tot_value)*100) %>% 
-  filter(
-    tresh == "Tresh_one",
-    variable == "value"
-  )
-
-```
+###### Mean ToE and vulnerabillity Index
 
@@
   left_join(Value_map_data) %>% 
+  filter(
+    tresh == "Tresh_one",
+    metric == "mean_yr"
+  ) %>% 
+  group_by(sau_fishing_entity,variable) %>% 
+  summarise(
+    toe = mean(value,na.rm=T),
+    # sum_value = sum(sum_value),
+    mean_part = mean(participation,na.rm=T)
+  ) %>% 
+  filter(!is.na(variable)) %>% 
+  # spread(variable,sum_value) %>% 
+  spread(variable,mean_part) %>% 
   mutate(
-    index = (participation/value),
-    index = ifelse(is.na(index),0,index)
-  ) %>% 
-  select(1:3,value,participation,index,everything())
+    imp = catch+value,
+    index = imp/toe
+  )
 
@@
   mutate(
-    index = 2000 + (index-min_index)*(2100-2000)/(max_index-min_index)
-  )
+    index_norm =(index-min_index)/(max_index-min_index),
+    index_2000 = 2000 + (index-min_index)*(2100-2000)/(max_index-min_index)
+  )
+
+### Explore index ####
+
+ggplot(data = Index_data) +
+  geom_point(
+    aes(
+      x = toe,
+      y = imp,
+      color = index_2000
+    ),
+    size = 4
+  ) +
+   scale_color_viridis(option="magma", discrete = F)
+
+#### ------------  ####
 
 Value_map_data <- Index_data %>% 
-  select(sau_fishing_entity,fishing_entity_id,tresh,variable,participation = indexb)
+  select(sau_fishing_entity,fishing_entity_id,tresh,variable,participation = index)
 

### 2020-09-23 Scripts/Manuscript.Rmd
First discussion draft

--- 2020-09-04 Scripts/Manuscript.Rmd
@@
 As a consequence, many marine species have changed their distribution [@Poloczanska:2016kk] and will continue to do so in the search of the appropriate environmental conditions needed for their subsistence [@Cheung:2010dt].
-Climate change is expected to impact fisheries production at different latitudes compromising our capacity to reach international sustainability goals such as the SDG 14 - life below water [@SDG14; @Cheung:2010dt; @Barange:2014kr].
-Such impacts can be offset with the proper reform of current fisheries management [@Gaines:2018sg; @CheungClimateReadyFish].
-However, such reforms will require comprehensive plans and institutions to manage species that cross international borders, also known as shared stocks [@Gaines:2018sg;@Pinsky:2018cb;].
+Climate change is expected to impact fisheries production at different latitudes compromising our capacity to reach international sustainability goals such as the SDG 14 - life below water [@UnitedNations2018; @Cheung:2010dt; @Barange:2014kr].
+Such impacts can be offset with the proper reform of current fisheries management [@Gaines:2018sg; @Cheung:2018dh].
+However, such reforms will require comprehensive plans and institutions to manage species that cross international borders, also known as shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
 
 In the early 1980's the United Nations officialized the concept of Exclusive Economic Zone (EEZs) though the Law of the Sea (UCLOS) [@UnitedNations:1986tl] partitioning the distribution of several species, thus, creating what we known today as shared stocks.
-As defined by the Food and Agriculture Organization, shared stocks can be classified as transboundary, those that cross neighboring EEZs; straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@FAO; @Munro].
+As defined by the Food and Agriculture Organization, shared stocks can be classified as transboundary, those that cross neighboring EEZs; straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@Munro:2004th].
 While counties are responsible for the management of stocks within their EEZs, UNCLOS incentives actions to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
@@
 There is evidence that such shifts have already complicated shared marine resources around the globe.
-In the North-eastern Atlantic, evidence shows that shared stocks managed within the European Union divisions have been shifting since the 1970's compromising the region's sustainability [@EuropeShift].
-In the North-eastern Pacific, sockeye salmon (*Oncorhynchus nerka*) shifted in the 1980's towards Alaska, creating a ten years conflict between Canada and the United States [@Song].
+In the North-eastern Atlantic, evidence shows that shared stocks managed within the European Union divisions have been shifting since the 1970's compromising the region's sustainability [@Baudron:2020dj].
+In the North-eastern Pacific, sockeye salmon (*Oncorhynchus nerka*) shifted in the 1980's towards Alaska, creating a ten years conflict between Canada and the United States [@Song:2017va].
 Further south, the southern stock of Peruvean anchoveta (*Engraulis ringens*) shared by Chile and Peru has gradually shifted towards Chilean waters adding to the over-exploitation of the resource [@Cashion:2018cg].
@@
 The first concept alludes to the game theory concept of thread point, known as the minimum payoff a player is willing to receive in order to cooperate with others.
-If a player's future payoff drops below the thread point, such player will likely break off any collaboration agreement and proceed to act solo [@Sumaila:E&S].
-A thread point can be translated to the proportion of a species distribution that each neighboring nation has (or the quota allocation that each one gets) at the time of signing off an agreement [@PalaciosAbrantes:E&S].
-Many shared fisheries operate this way [@EuropeShift; @GoMA; @NOAA].
-If not addressed, future changes in the thread point (hereafter called Stock Share Ratio -SSR) of jointly managed fisheries can have adverse consequences for shared fisheries [@PalaciosAbrantes:E&S; @Sumaila:E&S].
+If a player's future payoff drops below the thread point, such player will likely break off any collaboration agreement and proceed to act solo [@OCSumaila].
+A thread point can be translated to the proportion of a species distribution that each neighboring nation has (or the quota allocation that each one gets) at the time of signing off an agreement [@OCPalacios].
+Many shared fisheries operate this way having quota allocated based on some historical level [@Baudron:2020dj; @GoMA; @NOAA].
+If not addressed, future changes in the thread point (hereafter called Stock Share Ratio -SSR) of jointly managed fisheries can have adverse consequences for shared fisheries [@OCPalacios; @OCSumaila].
 The second concept refers to the Time of Emergence (ToE), defined as the moment when a future trend of a particular variable exceeds the historical variability [@IPCC:2019to].
 Such concept has been used to look at the emergence of diverse oceanic variables [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg] and ecological consequences of marine heatwaves [@Cheung:2020ja].
-Similarly, climate change is expected to drive species further away from their historic range limits [@Alexa?].
-Assessing the ToE of shared stocks is as important as the change itself, as it could reduce potential international conflicts [@Spijkers:2017ij] and aid the strengthen of marine governance [@Pinsky:2018cb;  @Pinsky:2014by; @Link:2010ei].
+Similarly, climate change is expected to drive species further away from their historic range limits.
+Assessing the ToE of shared stocks is as important as the change itself, as it could reduce potential international conflicts [@Spijkers:2017ij] and aid the strengthen of marine governance [@Pinsky:2018cb;  @Pinsky:2014by; @Link:2010ei; @Fogarty:2017eo].
 
@@
 
-We included in this analysis 633 marine exploited transboundary species representing 80% of the catch taken from the world's EEZs between 2005 and 2014 [@Palacios-Abrantes:xxxx].
-Each species in within an EEZ was considered a single stock unit due to a lack of more detailed spatial information [@Teh:2015gd; @Palacios-Abrantes:xxxx].
+We included in this analysis 633 marine exploited transboundary species representing 80% of the catch taken from the world's EEZs between 2005 and 2014 [@FishForVisa].
+Each species in within an EEZ was considered a single stock unit due to a lack of more detailed spatial information [@Teh:2015gd; @FishForVisa].
 We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the *Sea Around Us* shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic) including islands territories, and determined the intersections between polygons using the R package *sf* (**Supplements Fig 1 (SuppM)**) [@PackagesfSimpleF:2018vp].
@@
 Moreover, we classified catch trends as increasing, constant and decreasing, following Palacios-Abrantes *et al.* (*hopefully in press*).
-Finally, each EEZ was attributed to a geopolitical region [@UN] and biome [@Reygondeau:2019uh], and species were attributed a habitat preference [@FishBase].
+Finally, each EEZ was attributed to a geopolitical region according to the UN and biome [@Reygondeau:2019uh], and species were attributed a habitat preference [@FishBase].
 
@@
 bottom temperature).
-This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (See [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
+This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (See Cheung *et al.* [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
 
@@
 At the same time, by having multiple initial-condition ensemble simulations of the GFDL-ESM2M we were able to remove the natural variability of a stock's distribution by averaging the secular trend over all ensemble members (See Estimating the Time of Emergence).
+
 ## Changes in Stocks Share Ratio
 
-We adopted the concept of threat point as the minimum proportion of the stocks’ joint distribution between neighboring EEZs that each country will accept if they were to cooperate in the management of the stock [@Palacios-Abrantes:E&S].
+We adopted the concept of threat point as the minimum proportion of the stocks’ joint distribution between neighboring EEZs that each country will accept if they were to cooperate in the management of the stock [@OCPalacios].
 For each stock and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighboring EEZs, and then calculated the proportion that each EEZ had [@FishForVisa].
 We then average the results in three time periods in order to reduce climate variability within the projections.
-The first time period ($t_h$) comprises (1951 - 2005) and is assumed as the historic distribution of each species, thus representing each neighboring nation's threat point.
+The first time period ($t_h$) comprises (1951 - 2005) and is assumed as the historic distribution of each species, thus, representing each neighboring nation's threat point.
 This time frame matches the historic climate data that the GFDL uses to project climate change [@Dunne:2012bo; @Dunne:2013db].
@@
 
+## Vulnerability index
+
+We created a vulnerability index of fisheries management to shifts in transboundary species.
+The index is based on two factors; the importance of the shifting species to a country's total catch and revenue from transboundary species, and the time of emergence;
+
+$$ index' = (P_c + P_v)/ToE $$
+where *P_c* and *P_v* are the proportion that the shifting species' will have in the total catch and value generated from transboundary fisheries within a county's EEZs.
+Here, total catch and value are represented as the average of 2005 to 2014 according to the *Sea Around Us*.
+The index was then normalized to a 0 to 1 scale going from less vulnerable to more vulnerable:
+
+$$ index = \frac{(index' - N_{min})}{N_{max}-N_{min}} $$
+
+The premise of the index is that, from a management perspective, countries whose shifting transboundary stocks provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
+Similarly, the index is capable to capture different sources of societal benefits (e.g.
+food security trough catch or livelihood trough value).
+We assume that the management of fisheries that provide large societal benefits will have more pressure from stakeholders while earlier times of emergence will result in less time to act, thus resulting in higher vulnerability.
+
+## Caveats and uncertainties
+
+We have identified three main sources of uncertainty in our analysis and point out ways to move forward in order to produce more reliable results to inform policy dealing with international fisheries under climate change.
+First, in this analysis we relay in one Earth System Model (GFDL-ESM2M) and climate change scenario (RCP 8.5).
+While the use of multiple ESMs would allow to capture the structural uncertainty of the models [@Frolicher:2016dg], using the GFDL-ESM2M results in conservative estimates as its projections are the less extreme from the CIMIP5 models [@Bopp:2013fg].
+Addressing the uncertainty in future projections is key to advance marine governance for species on the move [@Pinsky:2018cb], thus, future efforts should look at the ensemble members of multiple ESM to capture the uncertainty embedded in the models.
+Secondly, we set an arbitrary threshold of $\pm$ 1 s.d.
+(67% confidence) to determine when a SSR and ToE would happen (see methods).
+We tested the sensitivity of our results to a larger threshold (e.g.
+$\pm$ 2 s.d for a 95% confidence).
+While the selected threshold is more relaxed than a 2 s.d., the distribution of results do not dramatically change between thresholds with a difference in ToE of about 4 years suggesting that this is a robust threshold (Fig.
+S4).
+Thirdly, in here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations [@FishForVisa].
+Thus, our analysis is carried at the species level, rather than a more detailed organizational level (e.g.
+fish stock) due to limited spatial-specific data on species sub populations (stocks).
+Reproducing our analysis with spatially-explicit stock data would allow to produce more certain results in changes in both SSR and ToE of transboundary stocks, and potentially identity different types of shared stock shifts [@Link:2010ei].
+However, many regions of the world lack these detailed data.
+Moreover, in many EEZs of the world fisheries are often managed at the species level [@FishForVisa].
+
 ## Statistical analysis
@@
 We hypothesize that all of these factors would have significant influence in the ToE of transboundary stocks.
-All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@gmt] and *zoo* [@PackagezooSInfr:2019tm], *readxl* [@REF], *ggrepel* [@REF], *zeallot* [@REF], *viridis* [@], *cowplot* [@], *moments* [@REF] and *pgirmess* [@REF].
+All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@gmt] and *zoo* [@PackagezooSInfr:2019tm], *readxl* [@REF], *ggrepel* [@REF], *zeallot* [@REF], *viridis* [@REF], *cowplot* [@REF], *moments* [@REF] and *pgirmess* [@REF].
 All code is available at https://github.com/jepa/TransEmergence
@@
 
-We estimated the change in SSR (larger than 1 s.d) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005).
+We estimated the change in SSR (with a 63% of confidence) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) under a high emission climate change scenario.
 Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
@@
 
-<!-- # Discussion -->
+# Discussion
 
-<!-- The combination of the time of emergence and the revenue participation of the species could shed a light in the importance of the time of emergence in terms of management.
-For example, we identified that in Peru 12 species will emerge by 2038 in average, representing 92% of the country's fishing value.
-Moreover, our results show that Peruvean anchoveta, the second largest fishery of the world [@SOFIA] and the main fishery in Peru [@REF] have already emerged (2010), with its stock share ratio benefiting Chile (Fig.
+In this paper we endeavor to look at the timing and intensity of changes in the sharing of transboundary species.
+Our findings here are noteworthy in several ways.
+First, we identified that the historical share proportion of transboundary species will change by as soon as 2030 in most neighboring EEZs (Fig.
 1).
+Second, we show that the current distribution of most transboundary species have already emerged from the historical record and that virtually all EEZs have seen at least one emerging species since 2006 (Fig.
+2).
+Thirdly, we identify regional patterns of complex changes in SSR and levels of vulnerability of transboundary fisheries management to climate change (Fig.
+3).
+
+By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Fig.
+1).
+Poleward shifts in SSR will be seen in regions of the world with relative large EEZs and few boundaries like the North East Pacific.
+However, despite a general recognition that climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn], the Eastern Tropical Pacific will see an equatorial shift with EEZs from Mexico to Ecuador loosing SSR to their southern neighbors.
+This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico's EEZ towards Peru under a high emission climate change scenario [@TayClarkeChp2].
+Moreover, regions with presence of multiple frontiers like the Caribbean and the North East Atlantic will see a complex network of changes in SSR with no marked direction (Fig.
+1B).
+These regions are known for being highly connected by both fish adult migration [@REF] and larvae dispersal [@Ramesh:2019va] leading to joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
+However, game theory suggest that the more number of parties negotiating, the harder it is for parties to reach an agreement [@Sumaila:2013], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.
+
+We set out to evaluate the emergence characteristics of transboundary stocks across the global EEZs.
+Our results suggest that most of the EEZs of the world and more than half of the transboundary species already emerged since the early 2000s (Fig.
+2).
+This is in alignment with previous research showing changes in the composition of marine catches since 1970's as evidence of climate change [@Cheung:2013gk].
+Indeed, an expansion in the distribution of multiple marine species has been documented in several regions of the world [@Poloczanska:2016kk].
+For example, in the early 2000s Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range reaching Washington waters where a new fishery rapidly developed [@Pinsky:2014by; @@Zeidberg:2007gs].
+In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, lead to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij].
+The large amount of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g.
+temperature, oxygen, $CO_2$) that influence the distribution of marine species [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
+It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 22.6% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while $O_2$ will have doubled [@Rodggers:2015j].
+
+Regional differences in ToE are likely the result of environmental factors, species vulnerability, and the characteristics of political boundaries.
+For example, the high number of stocks emerging early in the tropics suggest that tropical species are already on the move.
+The tropics are among the most species-rich biome [@Reygondeau:2019no] and have potentially experienced some of the largest declines in $O_2$ and greatest levels of warming since 2006 [@IPCC:2019tn; @Roemmich:2015id].
+Here, marine species live close to the thermal tolerance making them highly vulnerable to warming waters [@IPCC:2014bz] consequently having to move towards deeper or sub-tropical waters [@Poloczanska:2016kk].
+The premise of a shift already happening is supported by observational evidence of tropical species apearing in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis showing a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
+In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relative small EEZs and many borders that match peaks of emerging stocks.
+Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsible for substantial increases in emerging stocks (Fig.
+3).
+
+An early ToE and large fisheries benefits could indicate high vulnerability of transboundary fisheries management to climate change (Fig.
+3).
+For example, Polynesia and Micronesia are the only regions where the average ToE across species and EEZs already happened; 2019 and 2021, respectively (Fig.
+5).
+Countries in these regions are highly dependent of pelagic-oceanic (e.g.
+tunas) and reef-associated transboundary species (e.g.
+*Speceies*) for both food security [@REF] and profits [@REF].
+For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, in averaged, emerged since 2018 resulting in high vulnerability index (Fig.
+5).
+In this cases policy has to work against the clock and under potentially high societal pressure to address an issue that will impact important fisheries.
+However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to stocks sustainability.
+While it gives countries more time to prepare for substantial changes, species with relatively small overall participation can still be a source of conflict as portraited by the Atlantic mackerel case [@FishForVisa; @Spijkers:2017ij].
+
+# Conclusion
+
+Preparing policy to species on the move is key to achieve sustainable development goals and proper governance of the world fisheries [@Pinsky:2018cb].
+Here, we have addressed two fundamental steps to achieve sustainable international fisheries under a changing world [@Link:2010ei]; the degree of change that share species will have and the year in witch species will shift.
+Such results have direct implications for ocean governance.
+By identifying the degree location and time by which transboundary species will change from historical normal we provide critical information for regional management bodies and countries to anticipate the potential complications that climate change will bring to international fisheries management.
+While our results are not free of uncertainty and would benefit from more detailed analysis at localize levels, it represents a crucial step towards preparing ocean governance for species on the move.
+To address the issues raised in this paper, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
+
+# Supplements
+
+```{r fig_s1, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)"}
+
+img <- readPNG("../Figures/Supp1.png")
+grid.raster(img)
+
+```
+
+![Figure S1. World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points).](../Figures/Supp1.png)
+```{r fig_s2, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  " Distribution of Stock Share Ratio percentage change by the early (2020-2040) and middle (2040-2060) 21st centurury relative to today (1970-2005)"}
+
+img <- readPNG("../Figures/Supp2.png")
+grid.raster(img)
+
+```
+
+![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005))](../Figures/Supp2.png)
+
+```{r fig_s3, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  " Distribution of Stock Share Ratio percentage change by the early (2020-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
+
+img <- readPNG("../Figures/Supp2.png")
+grid.raster(img)
+
+```
+
+![Figure S3.
+Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
+Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial stocks that will see a shift.
+Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Supp3.png)
+
+# Recycled  text
+<!-- All of these results have direct implications for managing shared fisheries under climate change that can be partially addressed by the early implementation of joint-management international policies.
 -->
 
-<!-- Form all significantly different variables, only latitudinal gradient postpone the ToE suggesting that the ToE will happen first at equatorial latitudes and later towards the poles. -->
-
-<!-- Historically, environmentally-driven shits in shared stocks have cause problems and threaten stock's sustainability.
-**In early 1980's Canada and the US blablabla.** Many fish stocks managed by the European Union ICES regultaroy areas have been shifting since the 1970's [@EuropeShift].
-Such shifts have already result in confrontations between nations as evidenced by the expansion of the Atlantic mackerel (*Scomber scombrus*) distribution reaching the EEZ of Iceland [@Spijkers:2017ij].
-More recently, Chile and Peru have declare a cooperation strategy to mange the transboundary stock of Peruvean Anchovera [@Peru].
-Furthermore, as we continue to enter the $21^{st}$ century, commercial marine species are expected to continue shifting (mostly poleward) potentilly compromising shared stocks management.
-By the end of the century, all EEZs of the world are proyected to have at least one new transboundary species with some of them having over 10 new stocks [@Pinsky].
-Consequently, as species occupy new waters they leave historic areas and so, xx% of species are expected to exit EEZs by the end of che century [@Oremus].
-This entering and exicting national jurisdicions will complicate fisheries management potentially creating confllict between neighboring nations [@Pinksy @Oremus:2020id].
+<!-- The time and number of emerging stocks decreases as we move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
 -->
 
-<!-- Conclusion -->
-
-<!-- Preparing policy to species on the move is key to achieve sustainable development goals and proper governance of the world fisheries [@Pinksy]. -->
-
-# Recycling text
-
-Ignoring anthropocentric effects, a species biomass distribution follows a variation in time and space according to environmental preferences, species interactions and landscape features [@Huch].
-Thus, a neighboring nation will not necessarily have, at all times  (e.g.
-months or years), a fixed SSR but rather a series of SSRs [@GoMA].
+<!-- Changes in stock share ratio has contribute to past conflicts between nations with long history of collaboration including Canada and the United States over Pacific salmon and Norway and Iceland over spring spawning herring [@Miller:2013iv].
+In both cases, rules were emplaced and agreed by all parties, however, treaties were not prepared to address shifts in species distributions [@Miller:2013iv; @Song:2017va] and when the benefits felt below the desired level (e.g.
+threat point) the affected country broke off relationships.
+In the case of Canada and the US, the agreement was re-stated when a conservation fund was implemented [@Miller:2013iv].
+-->

### 2020-10-05 Scripts/Manuscript.Rmd
changes for thesis

@@
 
-# Abstract
+<!-- # Abstract -->
 
@@
 
-# Introduction
+## Introduction
 
@@
 
-# Methods
+## Materials and methods
 
-## Databases and species selection
+### Databases and species selection
 
@@
 
-## Projecting species distributions under climate change
+### Projecting species distributions under climate change
 
@@
 
-## Changes in Stocks Share Ratio
+### Changes in Stocks Share Ratio
 
@@
 
-## Estimating the Time of Emergence
+### Estimating the Time of Emergence
 
@@
 
-## Vulnerability index
+### Vulnerability index
 
@@
 
-## Caveats and uncertainties
+### Caveats and uncertainties
 
@@
 
-## Statistical analysis
+### Statistical analysis
 
@@
 
-# Results
+## Results
 
@@
 
-## Changes in Stock Share Ratio (SSR)
+### Changes in Stock Share Ratio (SSR)
 
@@
 We estimated the change in SSR (with a 63% of confidence) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) under a high emission climate change scenario.
-Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
-1A).
+Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Figure \ref{fig:Figure1}A).
 Most EEZs will average a change lower than 25% which is expected to slightly increase in intensity towards the  mid $21^{st}$ century (**Supplements Fig 1 (histogram)**).
@@
 For example, Guatemala's SSR of *Panulirus gracilis* shared with Mexico is projected to go from 12% to 26%.
-By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR >< 1 s.d.
+By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR $\pm$ 1 s.d.
 of historical distribution will be relatively large (global mean of 69% of species in an EEZ).
-For 17 EEZs, 100% of their analyzed transboundary stocks will shift, while and only six EEZs will have less than 25% shifting stocks (Fig.
-1A-Polygons).
+For 17 EEZs, 100% of their analyzed transboundary stocks will shift, while and only six EEZs will have less than 25% shifting stocks (Figure \ref{fig:Figure1}A-Polygons).
 Regions where EEZs are relatively small and have multiple borders (e.g.
-The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Fig.
-1B).
-Towards the mid $21^{st}$ century, changes in SSR per EEZ will no be substantial (**Supplements Fig 1 (map)**) relative to the early $21^{st}$ century but are expected to increase in intensity (**Supplements Fig 1 (histogram)**).
+The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Figure \ref{fig:Figure1}B).
+Towards the mid $21^{st}$ century, changes in SSR per EEZ will no be substantial (Figure \ref{fig:FigureS1}) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \ref{fig:FigureS2}).
 
-```{r SSR_gain_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Average gain of transboundary stock share ratio (lines) and number of shifting stocks (polygon) by 2030 (2021-2040) relative to 1951-2005.
-Arrows go from the losing EEZ (point) to the winner (arrowhead).
-Pannel B zooms in to specific areas"}
+```{r Figure1, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial stocks that will see a shift.
+Panel B zooms in to specific areas shown in grey boxes of A."}
+
+img <- readPNG("../Figures/Fig1.png")
+grid.raster(img)
+
+```
+
+<!-- ![Figure 1.
+Changes in Stock Share Ration by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial stocks that will see a shift.
+Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Fig1.png) -->
+
+### Time of Emergence
+
+We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock's distribution (2006-2100) overpasses the historical variation (1951-2005).
+Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g.
+*Engraulis ringens* shared by Chile and Peru emerged in 2010).
+Stocks will continue to emerge steadily towards 2097 when all 633 stocks will have had emerged in at least one neighboring EEZ (Figure \ref{fig:Figure2}).
+The number of EEZs with emerging stocks will increase steeply until 2020 when it seems to flatten approaching 2081 when all of the EEZs analyzed here will have at least one emerging stock (Figure \ref{fig:Figure2}).
+
+```{r Figure2, eval = T,  echo = F, fig.width = 5, fig.height=5, fig.cap =  "Cumulative increase in time of emergence by Exclusive Economic Zone and transbounudary stock."}
 
@@
 
-![Figure 1.
-Changes in Stock Share Ration by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial stocks that will see a shift.
-Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Fig1.png)
+<!-- ![Figure 2. Cumulative increase in time of emergence by Exclusive Economic Zone and transbounudary stock.](../Figures/Fig2.png) -->
 
-## Time of Emergence
-
-We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock's distribution (2006-2100) overpasses the historical variation (1951-2005).
-Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g.
-*Engraulis ringens* shared by Chile and Peru emerged in 2010).
-Stocks will continue to emerge steadily towards 2097 when all 633 stocks will have had emerged in at least one neighboring EEZ (Fig.
-2).
-The number of EEZs with emerging stocks will increase steeply until 2020 when it seems to flatten approaching 2081 when all of the EEZs analyzed here will have at least one emerging stock (Fig.
-2).
-
-```{r freq_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Cumulative increase in time of emergence by EEZ and stock"}
-
-img <- readPNG("../Figures/Fig2.png")
-grid.raster(img)
-
-```
-
-![Figure 2. Cumulative increase in time of emergence by Exclusive Economic Zone and transbounudary stock.](../Figures/Fig2.png)
-
-Most of the world EEZ’s will see an average ToE across transboundary stocks before 2040 (Fig.
-3).
+Most of the world EEZ’s will see an average ToE across transboundary stocks before 2040 (Figure \ref{fig:Figure3}).
 In a few EEZs stocks will have a relative late average ToE (e.g.
@@
 A higher number of emerging stocks will happen between latitudes 0$^o$ and 50$^o$N corresponding to regions of relatively small EEZs with multiple borders (e.g.
-The Caribbean and the Mediterranean) (Fig 2).
-In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50% (Fig.
-2 - Polygons).
+The Caribbean and the Mediterranean) (Figure \ref{fig:Figure2}).
+In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50% (Figure \ref{fig:Figure2} - Polygons).
 However, in some cases emerging stocks represent 85% of total fishing revenue (e.g.
@@
 
-```{r Toe_map, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}
+```{r Figure3, eval = T,  echo = F, fig.width  = 12, fig.height =10, fig.cap =  "Average time of emergence of transboundary stock.
+Land polygon represents  the current participation of emerging stocks in the country's fishing revenue.
+Exclusive Economic Zone polygon represents the average time of emergence across all stocks.
+Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean)."}
 
@@
 
-![Figure 3.
+<!-- ![Figure 3.
 Average time of emergence of transboundary stock.
@@
 Exclusive Economic Zone polygon represents the average time of emergence across all stocks.
-Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean).](../Figures/Fig3.png)
+Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean).](../Figures/Fig3.png) -->
 
-Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see an average ToE before 2030 (Fig.
-4).
+Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see an average ToE before 2030 (Figure \ref{fig:Figure4}).
 This includes the United States who occupies the $4^{th}$ place.
 These 47 countries are also among those with the largest number of emerging taxa.
-Regionally speaking, most of African, Asian and European EEZs will see a latter ToE, while the Americas and Oceania will see an early ToE (Fig.
-4).
+Regionally speaking, most of African, Asian and European EEZs will see a latter ToE, while the Americas and Oceania will see an early ToE (Figure \ref{fig:Figure4}).
 From the top ten countries with more revenue from transboundary species, nine of them share a border.
@@
 
-```{r bubble_plot, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}
+```{r Figure4, eval = T,  echo = F, fig.width=5, fig.height=6, fig.cap =  "Average time of emergence (ToE) per fishing entity.
+Fishing revenue in logarithmic scale.
+Showing names for the top 75th percentile.
+Countries marked with an asterisk have at least one neighbouring border."}
 
@@
 
-![Figure 4.
+<!-- ![Figure 4.
 Average time of emergence (ToE) per fishing entity.
@@
 Showing names for the top 75th percentile.
-Countries marked with an asterisk have at least one neighbouring border.](../Figures/Fig4.png)
+Countries marked with an asterisk have at least one neighbouring border.](../Figures/Fig4.png) -->
 
-We compared species ToE within EEZs by geographic region and by species habitat preference (Fig.
-5).
+We compared species ToE within EEZs by geographic region and by species habitat preference (Figure \ref{fig:Figure5}).
 The median ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001).
@@
 EEZs in Easter Asia which have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ from (almost) all other regions.
-Other regions also significantly differ from each other but in a less frequent manner (*See Table Sup X for full statistics*).
+Other regions also significantly differ from each other but in a less frequent manner (See Table \ref{tab:TabS1}) for full statistics).
 When looking specifically at habitat association, we found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
-Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (*See Table Sup XX for full statistics*).
+Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table \ref{tab:TabS2} for full statistics).
 
-```{r Boxplot, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  ""}
+```{r Figure5, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap =  " Regional time of emergence per habitat preference.
+Regions acording to the United Nations and habitat preference to FishBase.
+Whiskers represent 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Line represents median, and black points represent outliers (outside of 1.5*IQR."}
 
@@
 
-![Figure 5.
+<!-- ![Figure 5.
 Regional time of emergence per habitat preference.
@@
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers (outside of 1.5*IQR.](../Figures/Fig5.png)
+Line represents median, and black points represent outliers (outside of 1.5*IQR.](../Figures/Fig5.png) -->
 
@@
 ToE of SST), human drivers (e.g.
-catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary stocks (*Table Sup 3*).
-Results from the multiple linear regression analysis show a significant relation between ToE of transboundary stocks and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table 1).
+catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary stocks (Table \ref{tab:TableS3}).
+Results from the multiple linear regression analysis show a significant relation between ToE of transboundary stocks and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref{tab:Table1}).
 
-Table 1.
-Multiple regression results for time of emergence of transboundary stocks.
-Only showing significant variables.
-The model significantly predicts time of emergence (F[18,6155] = 20.68, p < 0.001, R2 = 0.05).
-
-```{r table1, eval = T,  echo = F}
+```{r Table1, eval = T,  echo = F, message = F, warning  = F}
 
 read_csv("../Tables/mlr_result.csv") %>% 
-  knitr::kable()
+  knitr::kable(
+    "latex",
+    booktabs = T,
+    captio= "Multiple regression results for time of emergence of transboundary stocks.
+Only showing significant variables.
+The model significantly predicts time of emergence (F[18,6155] = 20.68, p < 0.001, R2 = 0.05)"
+  )
 
@@
 
-# Discussion
+## Discussion
 
@@
 Our findings here are noteworthy in several ways.
-First, we identified that the historical share proportion of transboundary species will change by as soon as 2030 in most neighboring EEZs (Fig.
-1).
-Second, we show that the current distribution of most transboundary species have already emerged from the historical record and that virtually all EEZs have seen at least one emerging species since 2006 (Fig.
-2).
-Thirdly, we identify regional patterns of complex changes in SSR and levels of vulnerability of transboundary fisheries management to climate change (Fig.
-3).
+First, we identified that the historical share proportion of transboundary species will change by as soon as 2030 in most neighboring EEZs (Figure \ref{tab:Figure1}).
+Second, we show that the current distribution of most transboundary species have already emerged from the historical record and that virtually all EEZs have seen at least one emerging species since 2006 (Figure \ref{tab:Figure2}).
+Thirdly, we identify regional patterns of complex changes in SSR and levels of vulnerability of transboundary fisheries management to climate change (Figure \ref{tab:Figure3}).
 
-By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Fig.
-1).
+By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Figure \ref{tab:Figure1}).
 Poleward shifts in SSR will be seen in regions of the world with relative large EEZs and few boundaries like the North East Pacific.
@@
 We set out to evaluate the emergence characteristics of transboundary stocks across the global EEZs.
-Our results suggest that most of the EEZs of the world and more than half of the transboundary species already emerged since the early 2000s (Fig.
-2).
+Our results suggest that most of the EEZs of the world and more than half of the transboundary species already emerged since the early 2000s (Figure \ref{tab:Figure2}).
 This is in alignment with previous research showing changes in the composition of marine catches since 1970's as evidence of climate change [@Cheung:2013gk].
 Indeed, an expansion in the distribution of multiple marine species has been documented in several regions of the world [@Poloczanska:2016kk].
-For example, in the early 2000s Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range reaching Washington waters where a new fishery rapidly developed [@Pinsky:2014by; @@Zeidberg:2007gs].
+For example, in the early 2000s Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range reaching Washington waters where a new fishery rapidly developed [@Pinsky:2014by; @Zeidberg:2007gs].
 In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, lead to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij].
@@
 In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relative small EEZs and many borders that match peaks of emerging stocks.
-Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsible for substantial increases in emerging stocks (Fig.
-3).
+Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsible for substantial increases in emerging stocks (Figure \ref{tab:Figure3}).
 
-An early ToE and large fisheries benefits could indicate high vulnerability of transboundary fisheries management to climate change (Fig.
-3).
-For example, Polynesia and Micronesia are the only regions where the average ToE across species and EEZs already happened; 2019 and 2021, respectively (Fig.
-5).
+An early ToE and large fisheries benefits could indicate high vulnerability of transboundary fisheries management to climate change (Figure \ref{tab:Figure3})..
+For example, Polynesia and Micronesia are the only regions where the average ToE across species and EEZs already happened; 2019 and 2021, respectively (Figure \ref{tab:Figure5}).
 Countries in these regions are highly dependent of pelagic-oceanic (e.g.
@@
 *Speceies*) for both food security [@REF] and profits [@REF].
-For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, in averaged, emerged since 2018 resulting in high vulnerability index (Fig.
-5).
+For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, in averaged, emerged since 2018 resulting in high vulnerability index (Figure \ref{tab:Figure5}).
 In this cases policy has to work against the clock and under potentially high societal pressure to address an issue that will impact important fisheries.
@@
 
-# Conclusion
+## Conclusion
 
@@
 
-# Supplements
+## Supplements
 
-```{r fig_s1, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)"}
+```{r FigureS1, eval = T,  echo = F, fig.width=9, fig.height=6, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
 
@@
 
-![Figure S1. World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points).](../Figures/Supp1.png)
-```{r fig_s2, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  " Distribution of Stock Share Ratio percentage change by the early (2020-2040) and middle (2040-2060) 21st centurury relative to today (1970-2005)"}
+<!-- ![Figure S1. World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points).](../Figures/Supp1.png) -->
+
+```{r Figure2, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
 
@@
 
-![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005))](../Figures/Supp2.png)
+<!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005))](../Figures/Supp2.png) -->
 
-```{r fig_s3, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  " Distribution of Stock Share Ratio percentage change by the early (2020-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
+```{r FigureS3, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
+Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial stocks that will see a shift.
+Panel B zooms in to specific areas shown in grey boxes of A."}
 
-img <- readPNG("../Figures/Supp2.png")
+img <- readPNG("../Figures/Supp3.png")
 grid.raster(img)
@@
 
-![Figure S3.
-Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial stocks that will see a shift.
-Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Supp3.png)
+```{r FigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98 confidenc teshold (yellow"}
 
-# Recycled  text
+img <- readPNG("../Figures/Supp4.png")
+grid.raster(img)
+
+```
+
+<!-- # Recycled  text -->
 <!-- All of these results have direct implications for managing shared fisheries under climate change that can be partially addressed by the early implementation of joint-management international policies.

### 2020-10-14 Sci_Adv/cover_letter.Rmd
First draft to share with authors

---
output: pdf_document
---

Cover letter, containing:

<!-- The title of the paper and a statement of its main point -->

<!-- Any information needed to ensure a fair review process, including related manuscripts submitted to other journals -->

<!-- Names of colleagues who have reviewed the paper -->

<!-- A statement that none of the material has been published or is under consideration elsewhere, including the Internet. -->

<!-- Names, telephone, and e-mail addresses for all authors, including selection of one to be corresponding author. -->

Dr. Ali Shilatifard                        03-06-2020
Editor, *Science Advances*

We recommend our paper to be handled by Deputy Editor Dr. Jeremy Jackson and Associate Editor Dr. Jennifer Jacquet and suggest the following reviewers:

<!-- Names, affiliations, and e-mail addresses of five potential referees. -->
-
- Gretta Pecl, 
- Rebecca Selden, Wellesley College, Wellesley, MA, USA; bselden@wellesley.edu  
- 
-

This manuscript is not being considered for publication in any other journal. All data will be available upon request and the code used to analyze the data can be found at https://github.com/jepa/SharedToE All contributing authors and institutions have been identified, and the authors have no competing financial interests. This manuscript has been reviewed and approved by all authors.

Thank you for your consideration of this submission.
Sincerely (on behalf of all authors),

Juliano Palacios Abrantes  
The Institute for Oceans and Fisheries  
The University of British Columbia

### 2020-10-14 Scripts/Manuscript.Rmd
First draft to share with authors

@@
 ---
-title: Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change
+title: Matching the Time of Emergence of Transboundary Fish species to Lead Time for Policy Response Under Climate Change
+subtitle: Challenges to transboundary fisheries under climate change
 author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^'
@@
 output:
-  word_document: default
+  word_document: 
+    reference_docx: ../References/template.docx
   pdf_document: default
+# bibliography: ../References/FishForVisa_Ref.bib
+csl: ../References/science.csl
 ---
+
+<!-- # Instructions  -->
+
+<!-- Research articles should present a major advance and can be short or long (maximum 15,000 words) format.
+Submissions must include an abstract (up to 150 words), an introduction, and sections with brief informative subheadings.
+You may include up to ten figures and/or tables and about 60 references.
+Supplementary materials should be limited to information that is not essential for the general understanding of the research presented in the main text and can include data sets, figures, tables, videos, or audio files.
+You may include up to a total of ten figures and/or tables (combined) throughout the supplemental text.
+-->
 
@@
 packages <- c(
-  "tidyverse" 
+  "tidyverse" ,
+  "png",
+  "grid"
   # "here", # for dbem_import `here()`
@@
 
-<!-- # Abstract -->
+# Abstract (150  words)
 
-<!-- The delineation of Economic Exclusive Zones (EEZs) by the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, established political boundaries across the distribution of many fish species, creating shared species between nations.
-Despite not implicit, UNCLOS incentivizes nations to cooperate in the management of transboundary fisheries and most management plans emplaced are based on historical distribution of catches.
-Climate change represents a challenge for the joint management of fish stocks.
-For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures, quota allocation, and fishery newcomers.
-Recent events suggests that even changes in species that contribute a relatively small proportion of the overall catch can lead to important international conflict when no legal framework is available.
-Thus, it is fundamental to prepare for climate change-driven impacts such as identifying the stocks that are going to change historical proportions within neighboring EEZs, the degree by which they will change, and the time when such change is going to happen.
-Only by adapting management plans and policy that anticipates to these effects will international fisheries treaties be climate-ready.
-Here, we build on previous work to identify the transboundary stocks that will see changes in their stock-share-proportion (e.g.
-the proportion of the stock that each EEZ has) around the world.
-Moreover, we determine the time of emergence of each fish stock, defined as the year in which the stock-share-proportion will change from its historical range.
-Our results suggest that by the mid of the 21st century EEZs will see changes in stock-share ratio between 1 and 10% with some exceeding 25%.
-Most of these changes will happen before 2050 urging policy to be re-formulated to account for the possible implications of these changes.
-By the end of the 21st century virtually all EEZs and transboundary species will have experienced changes in stock-share-proportion relative to historical distributions.
-The results of this research has the potential to improve climate-ready fisheries if policy is modify to respond to the consequences presented here.
+<!-- An opening sentence that states the question/problem addressed by the research AND Enough background content to give context to the study AND A brief statement of primary results AND A short concluding sentence.
 -->
 
-## Introduction
+Climate change shifts in marine species distributions is affecting the current frameworks emplaced to manage species shared fisheries.
+Here, we rely on a modeling framework with multiple ensemble members to identify the time frame by which shared stocks distribution will change from historical ranges, and the change in the proportion of the distrirbution that each EEZ has under climate change.
+Our results suggest that 60% of shared species have emerged since 2006 in 87% of the world EEZs.
+By 2075 all EEZs will have an emerging species and by 2097 and all species will have emerged in at least one EEZ.
+Average changes in the ratio of species-share between neighboring EEZs will fluctuate between 1 and 10% with some exceeding 25% by 2030 relative to today’s ratio.
+Our results suggest that most changes will happen before 2030 urging policy to be re-formulated to account for the possible implications of these changes.
 
-Over the last century human activities have fundamentally changed the physicochemical composition of the ocean, warming the waters, increasing the pH and reducing the oxygen concentration [@IPCC:2019to].
+# Introduction
+
+Over the last century human activities have fundamentally changed the physicochemical composition of the ocean, warming the waters, decreasing the pH and reducing the oxygen concentration [@IPCC:2019to].
 As a consequence, many marine species have changed their distribution [@Poloczanska:2016kk] and will continue to do so in the search of the appropriate environmental conditions needed for their subsistence [@Cheung:2010dt].
-Climate change is expected to impact fisheries production at different latitudes compromising our capacity to reach international sustainability goals such as the SDG 14 - life below water [@UnitedNations2018; @Cheung:2010dt; @Barange:2014kr].
-Such impacts can be offset with the proper reform of current fisheries management [@Gaines:2018sg; @Cheung:2018dh].
-However, such reforms will require comprehensive plans and institutions to manage species that cross international borders, also known as shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
+Climate change is expected to impact fisheries production worldwide compromising our capacity to reach international sustainability goals such as the SDG 14 - life below water [@UnitedNations2018; @Cheung:2010dt; @Barange:2014kr].
+Such impacts can be partly offset with the proper reform of current fisheries management [@Gaines:2018sg; @Cheung:2018dh].
+However, these reforms will require comprehensive plans and institutions to manage species that cross international borders, also known as shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
 
-In the early 1980's the United Nations officialized the concept of Exclusive Economic Zone (EEZs) though the Law of the Sea (UCLOS) [@UnitedNations:1986tl] partitioning the distribution of several species, thus, creating what we known today as shared stocks.
-As defined by the Food and Agriculture Organization, shared stocks can be classified as transboundary, those that cross neighboring EEZs; straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@Munro:2004th].
-While counties are responsible for the management of stocks within their EEZs, UNCLOS incentives actions to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
-Today, an estimate of 633 transboundary fished stocks are found in different continents and across multiple habitats averaging 48.5 million tones and representing USD 78 billion in fishing revenue between 2005 and 2010 [@FishorVisa].
+In the early 1980’s the United Nations officialized the concept of Exclusive Economic Zone (EEZs) through the Law of the Sea (UCLOS) [@UnitedNations:1986tl], partitioning the distribution of several species, thus, creating what we know today as shared stocks.
+As defined by the Food and Agriculture Organization, shared stocks can be classified in three non-exclusive categories.
+(*i*) Transboundary, those that cross neighboring EEZs; (*ii*) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@Munro:2004th].
+While countries are responsible for the management of stocks within their EEZs, UNCLOS encourages actions to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
+Today, an estimate of 633 transboundary fish species are found in different continents and across multiple habitats averaging 48.5 million tones and representing USD 78 billion in fishing revenue between 2005 and 2010 (Chapter 1).
 
-The management, and thus sustainability, of fisheries that target shared stocks is jeopardized as marine species shift their known distribution to cope with a changing world [@Pinsky:2018cb].
-There is evidence that such shifts have already complicated shared marine resources around the globe.
-In the North-eastern Atlantic, evidence shows that shared stocks managed within the European Union divisions have been shifting since the 1970's compromising the region's sustainability [@Baudron:2020dj].
-In the North-eastern Pacific, sockeye salmon (*Oncorhynchus nerka*) shifted in the 1980's towards Alaska, creating a ten years conflict between Canada and the United States [@Song:2017va].
-Further south, the southern stock of Peruvean anchoveta (*Engraulis ringens*) shared by Chile and Peru has gradually shifted towards Chilean waters adding to the over-exploitation of the resource [@Cashion:2018cg].
-Global shifts in the distribution of tunas [@ErauskinExtramiana:2019fw] have brought additional management challenges in both the Western Indian ocean and South-western Pacific [@Miller:2007kx].
-Moreover, as species continue shifting with climate change, current policies might become obsolete, and thus, require new management strategies that create equitable systems across all dependent fishing nations [@Bell:2013ga; @Pinsky:2018cb].
-The objective of the current paper is to explore two concepts that will be essential in addressing the challenges that climate change will bring to the management of transboundary fisheries [@Link:2010ei].
+The management, and thus sustainability, of fisheries that target shared species is jeopardized as marine species shift their known distribution to cope with a changing world [@Pinsky:2018cb].
+Frequently, the rules placed to manage these species are set to historical trends [@Baudron:2020dj] and do not necessarily consider the biogeography of species [@Alexa], nor the effects of climate change [@FishForVisa].
+Misalignment between policies and distributional shifts have previously resulted in unsustainable harvest and international conflict [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij] and are expected to continue towards the end of the $21^{st}$ century.
+Thus, there is a need to understand the time when climate change will affect the sharing distribution of these species as well as the intensity in order to prepare ocean governance to species on the move [@Pinsky:2018cb; @FishForVisa; @Link:2010ei].
 
-The first concept alludes to the game theory concept of thread point, known as the minimum payoff a player is willing to receive in order to cooperate with others.
-If a player's future payoff drops below the thread point, such player will likely break off any collaboration agreement and proceed to act solo [@OCSumaila].
-A thread point can be translated to the proportion of a species distribution that each neighboring nation has (or the quota allocation that each one gets) at the time of signing off an agreement [@OCPalacios].
-Many shared fisheries operate this way having quota allocated based on some historical level [@Baudron:2020dj; @GoMA; @NOAA].
-If not addressed, future changes in the thread point (hereafter called Stock Share Ratio -SSR) of jointly managed fisheries can have adverse consequences for shared fisheries [@OCPalacios; @OCSumaila].
-The second concept refers to the Time of Emergence (ToE), defined as the moment when a future trend of a particular variable exceeds the historical variability [@IPCC:2019to].
-Such concept has been used to look at the emergence of diverse oceanic variables [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg] and ecological consequences of marine heatwaves [@Cheung:2020ja].
-Similarly, climate change is expected to drive species further away from their historic range limits.
-Assessing the ToE of shared stocks is as important as the change itself, as it could reduce potential international conflicts [@Spijkers:2017ij] and aid the strengthen of marine governance [@Pinsky:2018cb;  @Pinsky:2014by; @Link:2010ei; @Fogarty:2017eo].
-
-Here we estimate the changes in Stock Share Ratio and the Time of Emergence of 633 transboundary stocks shared by 280 EEZs of 198 coastal countries under climate change (See methods).
-We rely on a species distribution and Earth system model with ten ensemble members to project the distribution of the 663 selected fish stocks under a high greenhouse gas emission scenario.
+Here we endeavor to address this gap by estimating the Time of Emergence (ToE) [@Hawkins] and Stock Share Ratio (SSR) [@OCPalacios] of 633 transboundary species worldwide.
+We define ToE as the time when a change  of a particular variable exceeds the historical natural internal variability [@IPCC:2019to], a concept previously used to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja].
+Moreover, SSR is adopted from Game Theory as a proxy of thread point - the minimum payoff a player is willing to receive in order to cooperate with others [@OCSumaila; @Sumaila:2013].
+We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected shared species [@FishForVisa] by 280 EEZs of 198 coastal countries under a high greenhouse gas emission scenario RCP8.5 (See methods).
 We treated each species in an EEZ as a single stock due to a lack of more detailed spatial data [@FishForVisa] and only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g.
 we did not include the high seas).
-Our results suggest that virtually all EEZs will have a change in SSR by 2030 and that 87% of them already had at least one emerging species since 2006, with significant regional and environmental differences.
-Towards the end of the $21^{st}$ century, all of the 280 EEZs will have at least one emerging stock and all 633 stocks would have emerged at least once.
-Our results have direct implications for marine governance of shared marine resources under a changing world.
+Our results suggest that many shared species have already emerged and shifted their SSR in most neighboring EEZs posing a threat to the sustainability of shared stocks and the resilience of international treaties.
 
-## Materials and methods
+# Materials and methods
 
-### Databases and species selection
+## Databases and species selection
 
 We included in this analysis 633 marine exploited transboundary species representing 80% of the catch taken from the world's EEZs between 2005 and 2014 [@FishForVisa].
-Each species in within an EEZ was considered a single stock unit due to a lack of more detailed spatial information [@Teh:2015gd; @FishForVisa].
-We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the *Sea Around Us* shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic) including islands territories, and determined the intersections between polygons using the R package *sf* (**Supplements Fig 1 (SuppM)**) [@PackagesfSimpleF:2018vp].
-For each species and EEZ we used the Sea Around Us data to estimate the catch and fishing revenue (USD 2019 from fishing activities within the global EEZs reported as the average from the last decade (2005-2014) of available spatial data [@Zeller2016; @Tai2017; @Sumaila:2015uc].
-Moreover, we classified catch trends as increasing, constant and decreasing, following Palacios-Abrantes *et al.* (*hopefully in press*).
-Finally, each EEZ was attributed to a geopolitical region according to the UN and biome [@Reygondeau:2019uh], and species were attributed a habitat preference [@FishBase].
+Each species within an EEZ was considered a single stock unit due to a lack of more detailed spatial information [@Teh:2015gd; @FishForVisa].
+We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the *Sea Around Us* shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic; Fig.
+S1) including islands territories, and determined the intersections between polygons using the R package *sf* [@PackagesfSimpleF:2018vp].
+Each EEZ was attributed to a geopolitical region according to the UN and biome [@Reygondeau:2019uh], and species were attributed a habitat preference [@FishBase].
+For each species and EEZ we used the *Sea Around Us* data to estimate the catch and fishing revenue from fishing activities within the global EEZs.
+We report both catch and revenue the average from the last decade (2005-2014) of available spatial data [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+The monthly average US Consumer Price Index (CPI) according to the U.S.
+Bureau of Labor Statistic (https://www.bls.gov/cpi/) was used to convert the original 2010 real USD value to 2019 real USD.
 
-### Projecting species distributions under climate change
+## Projecting species distributions under climate change
 
@@
 Following this approach allows us to estimate a secular trend for the natural variability (e.g.
-background noise).
-At the same time, by having multiple initial-condition ensemble simulations of the GFDL-ESM2M we were able to remove the natural variability of a stock's distribution by averaging the secular trend over all ensemble members (See Estimating the Time of Emergence).
+background noise) and isolate the climate signal of a stock’s distribution by averaging the secular trend over all ensemble members.
 
-### Changes in Stocks Share Ratio
+## Estimating the Time of Emergence of transboundary species
 
-We adopted the concept of threat point as the minimum proportion of the stocks’ joint distribution between neighboring EEZs that each country will accept if they were to cooperate in the management of the stock [@OCPalacios].
-For each stock and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighboring EEZs, and then calculated the proportion that each EEZ had [@FishForVisa].
+Knowing the time (e.g., year) in which a shared stock is going to diverge from a historical variability is key for preparing to shifting species [@Link:2010ei; @FishForVisa].
+Here, we coupled the DBEM with the ten GFDL ensemble members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
+The index is based on the variation of a species distribution centroid in relation to the centroid of the neighboring EEZs it occupies.
+Each species centroid was determined by the average latitude and longitude of the grid cells containing the top 10% abundance percentile of each year derived from the DBEM.
+The EEZ centroid was estimated using the *st* package of *R* statistical software (**Supplements Fig 1 (SuppM)**).
+For each ensemble member, neighboring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring elevation differences using the *geosphere* package of *R* statistical software;
+
+$$D_{ens} = acos(sin(lat_{eez})*sin(lat_{spp}) + cos(lat_{eez})*cos(lat_{spp})*cos(lon_{eez}-lon_{spp}))$$
+where $lat_{eez}$ and $lat_{spp}$ are the latitudes of the EEZ and specie centroid, respectively and $lon_{eez}$ and $lon_{spp}$ are the respective longitudes.
+
+Then, for each year we calculated the transboundary index as follows:
+
+$$TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2$$
+
+where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectively for each future time step from 2006 to 2100 (*t*); *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+The index was then smoothed to a 10 years average.
+Finally, we determined the time of emergence setting an arbitrary threshold following Rodggers (2015):
+
+$$ToE = TREND/NOISE$$
+
+where *TREND* is the yearly mean *TI* ($\mu$) of all ten ensemble members and *NOISE* is the standard deviation of the TREND ($\sigma$) between 1951 and 2005. 
+<!-- Note that in some cases not all ensemble members show the presence of a species in a certain EEZ throughout the study's time frame.
+We presented the results only for those cases where all 10 ensemble members had results for each year between 1951 and 2100 (n = 150).
+-->
+We estimated the mean and one standar deviation ($\mu \pm \sigma$ for a confidence interval of 68%) to show the first year when the "signal" of the transboundary index overpasses the historic "noise".
+Since this was an arbitrary decision, we have included a more conservative estimation ($\mu \pm 2*\sigma$ for a 95% confidence interval).
+The SNR for both bottom and surface sea temperature was also estimated for statistical analysis following the SNR method.
+
+### Vulnerability index
+
+We created a vulnerability index of fisheries management to shifts in transboundary species.
+The index is based on two factors; the ToE and the importance of the shifting species to a country's total catch and revenue from transboundary species;
+
+$$ index' = (P_c + P_v)/ToE $$
+where *P_c* and *P_v* are the proportion that the shifting species' will have in the total catch and value generated from transboundary fisheries within a county's EEZs, respectively.
+Here, total catch and value are represented as the average of 2005 to 2014 within the global EEZs according to the *Sea Around Us*.
+The index was then normalized to a 0 to 1 scale going from less vulnerable to more vulnerable:
+
+$$ index = \frac{(index' - N_{min})}{N_{max}-N_{min}} $$
+
+## Changes in transboundary Species Share Ratio
+
+We adopted the concept of threat point as the minimum proportion of the species’ joint distribution between neighboring EEZs that each country will accept if they were to cooperate in the management of the stock [@OCPalacios].
+For each species and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighboring EEZs, and then calculated the proportion that each EEZ had [@FishForVisa].
 We then average the results in three time periods in order to reduce climate variability within the projections.
-The first time period ($t_h$) comprises (1951 - 2005) and is assumed as the historic distribution of each species, thus, representing each neighboring nation's threat point.
+The first time period ($t_h$) comprises (1951 - 2005) and is assumed as the historic share distribution of each species, thus, representing each neighboring nation's threat point.
 This time frame matches the historic climate data that the GFDL uses to project climate change [@Dunne:2012bo; @Dunne:2013db].
@@
 
-### Estimating the Time of Emergence
-
-Knowing the time (e.g.
-year) in which a shared stock is going to diverge from a historical variability is key for preparing to shifting stocks [@Link:2010ei].
-We employed the DBEM and the ten GFDL ensemble members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
-The index is based on the variation of a species distribution centroid in relation to the centroid of the neighboring EEZs it occupy.
-Each species centroid was determined by the average latitude and longitude of the grid cells containing the top 10% abundance percentile of each year derived from the DBEM.
-The EEZ centroid was estimated using the *st* package of *R* statistical software (**Supplements Fig 1 (SuppM)**).
-For each ensemble member, neighboring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring elevation differences using the *geosphere* package of *R* statistical software;
-
-$$D_{ens} = acos(sin(lat_{eez})*sin(lat_{spp}) + cos(lat_{eez})*cos(lat_{spp})*cos(lon_{eez}-lon_{spp}))$$
-where $lat_{eez}$ and $lat_{spp}$ are the latitudes of the EEZ and specie centroid, respectively and $lon_{eez}$ and $lon_{spp}$ are the respectively longitudes.
-
-Then, for each year we calculated the transboundary index as follows:
-
-$$TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2$$
-
-where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectively for each future time step from 2006 to 2100 (*t*); *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
-The index was then smoothed to a 10 years average.
-Finally, to determine the time of emergence we estimated a signal to noise ratio (SNR) and set an arbitrary threshold as follows [@Rodggers:2015jq]:
-
-$$SNR = TREND/NOISE$$
-
-where *TREND* is the yearly mean *TI* ($\mu$) of all ten ensemble members and *NOISE* is the standard deviation of the TREND ($\sigma$) between 1951 and 2005.
-Note that in some cases not all ensemble members show the presence of a species in a certain EEZ throughout the study's time frame.
-We presented the results only for those cases where all 10 ensemble members had results for each year between 1951 and 2100 (n = 150).
-We determine a $\mu \pm \sigma$ (confidence interval of 68%) to show the first year when the "signal" of the transboundary index overpasses the historic "noise".
-Since this was an arbitrary decision, we have included a more conservative estimation ($\mu \pm 2*\sigma$ for a 95% confidence interval).
-The SNR for both bottom and surface sea temperature was also estimated for statistical analysis following the SNR method.
-
-<!-- In cases where the species is distributed in more than 2 neighboring EEZs (e.g. *xxx*), the index is as follows: -->
-
-<!-- $$I =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h})}) |  \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{C,y}}{sd(D_{B,h})}) | \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{Cy}}{sd(D_{B,h})})/N$$ -->
-
-<!-- In other words, for two neighboring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases (see *TI* equation).
-Consequently, if the centroid shifts are within historical variability (*SNR* < 1), risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability (*SNR* > 1), the risk of dispute is higher.
-Thus, risk is positively related to the frequency of centroid shifts beyond historical variability.
--->
-
-### Vulnerability index
-
-We created a vulnerability index of fisheries management to shifts in transboundary species.
-The index is based on two factors; the importance of the shifting species to a country's total catch and revenue from transboundary species, and the time of emergence;
-
-$$ index' = (P_c + P_v)/ToE $$
-where *P_c* and *P_v* are the proportion that the shifting species' will have in the total catch and value generated from transboundary fisheries within a county's EEZs.
-Here, total catch and value are represented as the average of 2005 to 2014 according to the *Sea Around Us*.
-The index was then normalized to a 0 to 1 scale going from less vulnerable to more vulnerable:
-
-$$ index = \frac{(index' - N_{min})}{N_{max}-N_{min}} $$
-
-The premise of the index is that, from a management perspective, countries whose shifting transboundary stocks provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
-Similarly, the index is capable to capture different sources of societal benefits (e.g.
-food security trough catch or livelihood trough value).
-We assume that the management of fisheries that provide large societal benefits will have more pressure from stakeholders while earlier times of emergence will result in less time to act, thus resulting in higher vulnerability.
-
-### Caveats and uncertainties
-
-We have identified three main sources of uncertainty in our analysis and point out ways to move forward in order to produce more reliable results to inform policy dealing with international fisheries under climate change.
-First, in this analysis we relay in one Earth System Model (GFDL-ESM2M) and climate change scenario (RCP 8.5).
-While the use of multiple ESMs would allow to capture the structural uncertainty of the models [@Frolicher:2016dg], using the GFDL-ESM2M results in conservative estimates as its projections are the less extreme from the CIMIP5 models [@Bopp:2013fg].
-Addressing the uncertainty in future projections is key to advance marine governance for species on the move [@Pinsky:2018cb], thus, future efforts should look at the ensemble members of multiple ESM to capture the uncertainty embedded in the models.
-Secondly, we set an arbitrary threshold of $\pm$ 1 s.d.
-(67% confidence) to determine when a SSR and ToE would happen (see methods).
-We tested the sensitivity of our results to a larger threshold (e.g.
-$\pm$ 2 s.d for a 95% confidence).
-While the selected threshold is more relaxed than a 2 s.d., the distribution of results do not dramatically change between thresholds with a difference in ToE of about 4 years suggesting that this is a robust threshold (Fig.
-S4).
-Thirdly, in here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations [@FishForVisa].
-Thus, our analysis is carried at the species level, rather than a more detailed organizational level (e.g.
-fish stock) due to limited spatial-specific data on species sub populations (stocks).
-Reproducing our analysis with spatially-explicit stock data would allow to produce more certain results in changes in both SSR and ToE of transboundary stocks, and potentially identity different types of shared stock shifts [@Link:2010ei].
-However, many regions of the world lack these detailed data.
-Moreover, in many EEZs of the world fisheries are often managed at the species level [@FishForVisa].
-
-### Statistical analysis
+## Statistical analysis
 
 We tested the results for normality (e.g.
-skewness, kurtosis) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary stocks.
-Moreover, we implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
+skewness, kurtosis) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
+<!-- Moreover, we implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
 Specifically, we looked at habitat preferences, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
-We hypothesize that all of these factors would have significant influence in the ToE of transboundary stocks.
+We hypothesize that all of these factors would have significant influence in the ToE of transboundary species.
+-->
 All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@gmt] and *zoo* [@PackagezooSInfr:2019tm], *readxl* [@REF], *ggrepel* [@REF], *zeallot* [@REF], *viridis* [@REF], *cowplot* [@REF], *moments* [@REF] and *pgirmess* [@REF].
@@
 
-## Results
+# Results
 
@@
 
-### Changes in Stock Share Ratio (SSR)
+## Time of Emergence of transbounudary species
+
+We estimated the ToE between neighboring EEZs for 633 transboundary species, that is, the year where the future stock's distribution (2006-2100) overpasses the historical variation (1951-2005).
+Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g.
+*Engraulis ringens* shared by Chile and Peru emerged in 2010).
+species will continue to emerge steadily towards 2097 when all 633 species will have emerged in at least one neighboring EEZ (Fig.
+1).
+The number of EEZs with emerging species will increase steeply until 2020 when it seems to flatten approaching 2081 when all of the EEZs analyzed here will have at least one emerging stock (Fig.
+1).
+
+```{r Figure1, eval = T,  echo = F, fig.width = 5, fig.height=5, fig.cap =  "Cumulative increase in time of emergence by Exclusive Economic Zone and transboundary species in absolute numbers and percentage."}
+
+img <- readPNG("../Figures/Fig1.png")
+grid.raster(img)
+
+```
+
+<!-- ![Figure 1. Cumulative increase in time of emergence by Exclusive Economic Zone and transboundary species in absolute numbers and percentage.](../Figures/Fig2.png) -->
+
+Most of the world EEZ’s will see an average ToE across transboundary species before 2040 (Fig.
+2).
+In a few EEZs species will have a relative late average ToE (e.g.
+Honduras average ToE is 2071) while 54 of the 280 EEZs (19%) have already seen their species emerged (e.g.
+average ToE across all species <= 2020).
+A higher number of emerging species will happen between latitudes 0$^o$ and 50$^o$N corresponding to regions of relatively small EEZs with multiple borders (e.g.
+The Caribbean and the Mediterranean) (Fig.2).
+In most cases, the current participation of the emerging species in a country's total fishing revenue is less than 50% (Fig.
+2 - Polygons).
+However, in some cases emerging species represent 85% of current transboundary fishing revenue (e.g.
+Peru) or more (e.g.
+Marshall Islands, 92%).
+Only in very few cases like Brazil and Alaska's polar EEZ, no specie will emerge between 2006 and 2100.
+
+```{r Figure2, eval = T,  echo = F, fig.width  = 12, fig.height =10, fig.cap =  "Average time of emergence of transboundary specie.
+Land polygon shows the vulnerability index.
+Exclusive Economic Zone polygon represents the average time of emergence across all species.
+Density plot on the right shows the latitudinal number of emerging species across EEZ centroids (10^o running mean)."}
+
+img <- readPNG("../Figures/Fig2.png")
+grid.raster(img)
+
+```
+
+<!-- ![Figure 2.
+Average time of emergence of transboundary specie.
+Land polygon shows the vulnerability index.
+Exclusive Economic Zone polygon represents the average time of emergence across all species.
+Density plot on the right shows the latitudinal number of emerging species across EEZ centroids (10^o running mean).](../Figures/Fig3.png) -->
+
+Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transboundary species between 2005 and 2010 will see an average ToE before 2030 (Fig.
+3).
+This includes the United States who occupies the $4^{th}$ place in revenue generated from transboundary fisheries.
+These 47 countries are also among those with the largest number of emerging taxa.
+Regionally speaking, most of African, Asian and European EEZs will see a latter ToE, while the Americas and Oceania will see an early ToE (Fig.
+3).
+From the top ten countries with more revenue from transboundary species, nine of them share a border.
+Most of these countries (e.g.
+China) will also have a large number of emerging species, while others like Peru and Norway will have a small number of emerging taxa but representing high revenue fisheries.
+
+```{r Figure3, eval = T,  echo = F, fig.width=5, fig.height=6, fig.cap =  "Average time of emergence of transboundary species per fishing entity.
+Fishing revenue in logarithmic scale.
+Showing names for the top 75th percentile.
+Countries marked with an asterisk have at least one neighboring border in the group."}
+
+img <- readPNG("../Figures/Fig3.png")
+grid.raster(img)
+
+```
+
+<!-- ![Figure 3.
+Average time of emergence of transboundary species per fishing entity.
+Fishing revenue in logarithmic scale.
+Showing names for the top 75th percentile.
+Countries marked with an asterisk have at least one neighboring border in the group.](../Figures/Fig4.png) -->
+
+We compared species ToE within EEZs by geographic region and by species habitat preference (Fig.
+4).
+The median ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001).
+Post-hoc tests revealed regional patterns.
+EEZs in Easter Asia which have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ from (almost) all other regions.
+Other regions also significantly differ from each other but in a less frequent manner (See Table S1 for full statistics).
+When looking specifically at habitat association, we found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
+Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table S2 for full statistics).
+
+```{r Figure4, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap =  " Regional time of emergence per habitat preference.
+Regions according to the United Nations and habitat preference to FishBase.
+Whiskers represent 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Line represents median, and black points represent outliers (outside of 1.5*IQR."}
+
+img <- readPNG("../Figures/Fig4.png")
+grid.raster(img)
+
+```
+
+<!-- ![Figure 4.
+Regional time of emergence per habitat preference.
+Regions according to the United Nations and habitat preference to FishBase.
+Whiskers represent 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Line represents median, and black points represent outliers (outside of 1.5*IQR.](../Figures/Fig5.png) -->
+
+<!-- We hypothesized that species environmental preferences (e.g.
+presence in certain biomes and habitats), climate change effects on water temperature (e.g.
+ToE of SST), human drivers (e.g.
+catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary species (Table \ref{tab:TableS3}).
+Results from the multiple linear regression analysis show a significant relation between ToE of transboundary species and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref{tab:Table1}).
+-->
+
+```{r Table1, eval = F,  echo = F, message = F, warning  = F}
+
+read_csv("../Tables/mlr_result.csv") %>% 
+  knitr::kable(
+    "latex",
+    booktabs = T,
+    captio= "Multiple regression results for time of emergence of transboundary species.
+Only showing significant variables.
+The model significantly predicts time of emergence (F[18,6155] = 20.68, p < 0.001, R2 = 0.05)"
+  )
+
+```
+
+## Changes in specie Share Ratio (SSR)
 
@@
 We estimated the change in SSR (with a 63% of confidence) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) under a high emission climate change scenario.
-Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Figure \ref{fig:Figure1}A).
-Most EEZs will average a change lower than 25% which is expected to slightly increase in intensity towards the  mid $21^{st}$ century (**Supplements Fig 1 (histogram)**).
+Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
+5A).
+Most EEZs will average a change lower than 25% which is expected to slightly increase in intensity towards the  mid $21^{st}$ century (**Supplements Fig 5 (histogram)**).
 However, this result masks larger changes in SSR for specific species.
 For example, Guatemala's SSR of *Panulirus gracilis* shared with Mexico is projected to go from 12% to 26%.
-By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR $\pm$ 1 s.d.
+By early $21^{st}$ century, the percentage of transboundary species per EEZ that will see a change in SSR $\pm$ 1 s.d.
 of historical distribution will be relatively large (global mean of 69% of species in an EEZ).
-For 17 EEZs, 100% of their analyzed transboundary stocks will shift, while and only six EEZs will have less than 25% shifting stocks (Figure \ref{fig:Figure1}A-Polygons).
+For 17 EEZs, 100% of their analyzed transboundary species will shift, while only six EEZs will have less than 25% shifting species (Fig.
+5A-Polygons).
 Regions where EEZs are relatively small and have multiple borders (e.g.
-The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Figure \ref{fig:Figure1}B).
-Towards the mid $21^{st}$ century, changes in SSR per EEZ will no be substantial (Figure \ref{fig:FigureS1}) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \ref{fig:FigureS2}).
+The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Fig.
+5B).
+Towards the mid $21^{st}$ century, changes in SSR per EEZ will not be substantial (Fig.
+S1) relative to the early $21^{st}$ century but are expected to increase in intensity (Fig.
+S2).
 
-```{r Figure1, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial stocks that will see a shift.
+```{r Figure1, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average gain of transboundary specie share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial species that will see a shift.
 Panel B zooms in to specific areas shown in grey boxes of A."}
@@
 <!-- ![Figure 1.
-Changes in Stock Share Ration by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial stocks that will see a shift.
+Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average gain of transboundary specie share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial species that will see a shift.
 Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Fig1.png) -->
 
-### Time of Emergence
-
-We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock's distribution (2006-2100) overpasses the historical variation (1951-2005).
-Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g.
-*Engraulis ringens* shared by Chile and Peru emerged in 2010).
-Stocks will continue to emerge steadily towards 2097 when all 633 stocks will have had emerged in at least one neighboring EEZ (Figure \ref{fig:Figure2}).
-The number of EEZs with emerging stocks will increase steeply until 2020 when it seems to flatten approaching 2081 when all of the EEZs analyzed here will have at least one emerging stock (Figure \ref{fig:Figure2}).
-
-```{r Figure2, eval = T,  echo = F, fig.width = 5, fig.height=5, fig.cap =  "Cumulative increase in time of emergence by Exclusive Economic Zone and transbounudary stock."}
-
-img <- readPNG("../Figures/Fig2.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure 2. Cumulative increase in time of emergence by Exclusive Economic Zone and transbounudary stock.](../Figures/Fig2.png) -->
-
-Most of the world EEZ’s will see an average ToE across transboundary stocks before 2040 (Figure \ref{fig:Figure3}).
-In a few EEZs stocks will have a relative late average ToE (e.g.
-Honduras average ToE is 2071) while 54 of the 280 EEZs (19%) have already seen their stocks emerged (e.g.
-average ToE across all stocks <= 2020).
-A higher number of emerging stocks will happen between latitudes 0$^o$ and 50$^o$N corresponding to regions of relatively small EEZs with multiple borders (e.g.
-The Caribbean and the Mediterranean) (Figure \ref{fig:Figure2}).
-In most cases, the current participation of the emerging stocks in a country's total fishing revenue is less than 50% (Figure \ref{fig:Figure2} - Polygons).
-However, in some cases emerging stocks represent 85% of total fishing revenue (e.g.
-Peru) or more (e.g.
-Marshall Islands, 92%).
-Only in very few cases like Brazil and Alaska's polar EEZ, no stock will emerge between 2006 and 2100.
-
-```{r Figure3, eval = T,  echo = F, fig.width  = 12, fig.height =10, fig.cap =  "Average time of emergence of transboundary stock.
-Land polygon represents  the current participation of emerging stocks in the country's fishing revenue.
-Exclusive Economic Zone polygon represents the average time of emergence across all stocks.
-Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean)."}
-
-img <- readPNG("../Figures/Fig3.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure 3.
-Average time of emergence of transboundary stock.
-Land polygon represents  the current participation of emerging stocks in the country's fishing revenue.
-Exclusive Economic Zone polygon represents the average time of emergence across all stocks.
-Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean).](../Figures/Fig3.png) -->
-
-Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see an average ToE before 2030 (Figure \ref{fig:Figure4}).
-This includes the United States who occupies the $4^{th}$ place.
-These 47 countries are also among those with the largest number of emerging taxa.
-Regionally speaking, most of African, Asian and European EEZs will see a latter ToE, while the Americas and Oceania will see an early ToE (Figure \ref{fig:Figure4}).
-From the top ten countries with more revenue from transboundary species, nine of them share a border.
-Most of these countries (e.g.China) will also have a large number of emerging stocks, while others like Peru and Norway will have a small number of emerging taxa but representing high revenue fisheries.
-
-```{r Figure4, eval = T,  echo = F, fig.width=5, fig.height=6, fig.cap =  "Average time of emergence (ToE) per fishing entity.
-Fishing revenue in logarithmic scale.
-Showing names for the top 75th percentile.
-Countries marked with an asterisk have at least one neighbouring border."}
-
-img <- readPNG("../Figures/Fig4.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure 4.
-Average time of emergence (ToE) per fishing entity.
-Fishing revenue in logarithmic scale.
-Showing names for the top 75th percentile.
-Countries marked with an asterisk have at least one neighbouring border.](../Figures/Fig4.png) -->
-
-We compared species ToE within EEZs by geographic region and by species habitat preference (Figure \ref{fig:Figure5}).
-The median ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001).
-Post-hoc tests revealed regional patterns.
-EEZs in Easter Asia which have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ from (almost) all other regions.
-Other regions also significantly differ from each other but in a less frequent manner (See Table \ref{tab:TabS1}) for full statistics).
-When looking specifically at habitat association, we found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
-Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table \ref{tab:TabS2} for full statistics).
-
-```{r Figure5, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap =  " Regional time of emergence per habitat preference.
-Regions acording to the United Nations and habitat preference to FishBase.
-Whiskers represent 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers (outside of 1.5*IQR."}
-
-img <- readPNG("../Figures/Fig5.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure 5.
-Regional time of emergence per habitat preference.
-Regions acording to the United Nations and habitat preference to FishBase.
-Whiskers represent 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers (outside of 1.5*IQR.](../Figures/Fig5.png) -->
-
-We hypothesized that species environmental preferences (e.g.
-presence in certain biomes and habitats), climate change effects on water temperature (e.g.
-ToE of SST), human drivers (e.g.
-catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary stocks (Table \ref{tab:TableS3}).
-Results from the multiple linear regression analysis show a significant relation between ToE of transboundary stocks and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref{tab:Table1}).
-
-```{r Table1, eval = T,  echo = F, message = F, warning  = F}
-
-read_csv("../Tables/mlr_result.csv") %>% 
-  knitr::kable(
-    "latex",
-    booktabs = T,
-    captio= "Multiple regression results for time of emergence of transboundary stocks.
-Only showing significant variables.
-The model significantly predicts time of emergence (F[18,6155] = 20.68, p < 0.001, R2 = 0.05)"
-  )
-
-```
-
-## Discussion
+# Discussion
 
@@
 Our findings here are noteworthy in several ways.
-First, we identified that the historical share proportion of transboundary species will change by as soon as 2030 in most neighboring EEZs (Figure \ref{tab:Figure1}).
-Second, we show that the current distribution of most transboundary species have already emerged from the historical record and that virtually all EEZs have seen at least one emerging species since 2006 (Figure \ref{tab:Figure2}).
-Thirdly, we identify regional patterns of complex changes in SSR and levels of vulnerability of transboundary fisheries management to climate change (Figure \ref{tab:Figure3}).
+First, we show that the current distribution of most transboundary species have already emerged from the historical record and that virtually all EEZs have seen at least one emerging species since 2006 (Fig.
+5).
+Second, we identify regional patterns of complex changes in SSR and levels of vulnerability of transboundary fisheries management to climate change (Fig.
+3).
+Third, we identified that the historical share proportion of transboundary species will change by as soon as 2030 in most neighboring EEZs (Fig.
+5).
 
-By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Figure \ref{tab:Figure1}).
-Poleward shifts in SSR will be seen in regions of the world with relative large EEZs and few boundaries like the North East Pacific.
-However, despite a general recognition that climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn], the Eastern Tropical Pacific will see an equatorial shift with EEZs from Mexico to Ecuador loosing SSR to their southern neighbors.
-This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico's EEZ towards Peru under a high emission climate change scenario [@TayClarkeChp2].
-Moreover, regions with presence of multiple frontiers like the Caribbean and the North East Atlantic will see a complex network of changes in SSR with no marked direction (Fig.
-1B).
-These regions are known for being highly connected by both fish adult migration [@REF] and larvae dispersal [@Ramesh:2019va] leading to joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
-However, game theory suggest that the more number of parties negotiating, the harder it is for parties to reach an agreement [@Sumaila:2013], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.
+### Emergence characteristics of transboundary species
 
-We set out to evaluate the emergence characteristics of transboundary stocks across the global EEZs.
-Our results suggest that most of the EEZs of the world and more than half of the transboundary species already emerged since the early 2000s (Figure \ref{tab:Figure2}).
-This is in alignment with previous research showing changes in the composition of marine catches since 1970's as evidence of climate change [@Cheung:2013gk].
+We set out to evaluate the emergence characteristics of transboundary species across the global EEZs.
+Our results suggest that most of the EEZs of the world and more than half of the transboundary species already emerged since the early 2000s (Fig.
+2).
+This is in alignment with previous research showing changes in the composition of marine catches since the 1970's as evidence of climate change [@Cheung:2013gk].
 Indeed, an expansion in the distribution of multiple marine species has been documented in several regions of the world [@Poloczanska:2016kk].
 For example, in the early 2000s Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range reaching Washington waters where a new fishery rapidly developed [@Pinsky:2014by; @Zeidberg:2007gs].
-In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, lead to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij].
+In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, led to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij].
 The large amount of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g.
@@
 Regional differences in ToE are likely the result of environmental factors, species vulnerability, and the characteristics of political boundaries.
-For example, the high number of stocks emerging early in the tropics suggest that tropical species are already on the move.
+For example, the high number of species emerging early in the tropics suggest that tropical species are already on the move.
 The tropics are among the most species-rich biome [@Reygondeau:2019no] and have potentially experienced some of the largest declines in $O_2$ and greatest levels of warming since 2006 [@IPCC:2019tn; @Roemmich:2015id].
 Here, marine species live close to the thermal tolerance making them highly vulnerable to warming waters [@IPCC:2014bz] consequently having to move towards deeper or sub-tropical waters [@Poloczanska:2016kk].
-The premise of a shift already happening is supported by observational evidence of tropical species apearing in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis showing a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
-In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relative small EEZs and many borders that match peaks of emerging stocks.
-Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsible for substantial increases in emerging stocks (Figure \ref{tab:Figure3}).
+The premise of a shift already happening is supported by observational evidence of tropical species appearing in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis showing a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
+In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs and many borders that match peaks of emerging species.
+Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsible for substantial increases in emerging species (Fig.
+3).
 
-An early ToE and large fisheries benefits could indicate high vulnerability of transboundary fisheries management to climate change (Figure \ref{tab:Figure3})..
-For example, Polynesia and Micronesia are the only regions where the average ToE across species and EEZs already happened; 2019 and 2021, respectively (Figure \ref{tab:Figure5}).
+## Some fisheries will be more vulnerable to emerging species
+
+The premise of our index is that, from a management perspective, countries whose shifting transboundary species provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
+Similarly, the index is capable to capture different sources of societal benefits (e.g.
+food security trough catch or livelihood trough value).
+We assume that the management of fisheries that provide large societal benefits will have more pressure from stakeholders while earlier times of emergence will result in less time to act, thus resulting in higher vulnerability.
+Thus, an early ToE and large fisheries benefits could indicate high vulnerability of transboundary fisheries management to climate change (Fig.
+3).
+For example, Polynesia and Micronesia are the only regions where the average ToE across species and EEZs already happened; 2019 and 2021, respectively (Fig.
+5).
 Countries in these regions are highly dependent of pelagic-oceanic (e.g.
@@
 *Speceies*) for both food security [@REF] and profits [@REF].
-For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, in averaged, emerged since 2018 resulting in high vulnerability index (Figure \ref{tab:Figure5}).
-In this cases policy has to work against the clock and under potentially high societal pressure to address an issue that will impact important fisheries.
-However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to stocks sustainability.
-While it gives countries more time to prepare for substantial changes, species with relatively small overall participation can still be a source of conflict as portraited by the Atlantic mackerel case [@FishForVisa; @Spijkers:2017ij].
+For example, in the Marshall Islands 16 transboundary species responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, in averaged, emerged since 2018 resulting in high vulnerability index (Fig.
+5).
+In this cases policy has to work against the clock and under potentially high societal pressure to address an issue that could be impacting important fisheries.
+However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to species sustainability.
+While it gives countries more time to prepare for substantial changes, species with relatively small overall participation can still be a source of conflict as portrayed by the Atlantic mackerel case [@FishForVisa; @Spijkers:2017ij].
 
-## Conclusion
+## Changes in specie-share ratio will be everywhere
+
+By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Fig.
+1).
+Poleward shifts in SSR will be seen in regions of the world with relatively large EEZs and few boundaries like the North East Pacific.
+However, despite a general recognition that climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn], the Eastern Tropical Pacific will see an equatorial shift with EEZs from Mexico to Ecuador losing SSR to their southern neighbors.
+This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico's EEZ towards Peru under a high emission climate change scenario [@TayClarkeChp2].
+Moreover, regions with presence of multiple frontiers like the Caribbean and the North East Atlantic will see a complex network of changes in SSR with no marked direction (Fig.
+1B).
+These regions are known for being highly connected by both fish adult migration [@REF] and larvae dispersal [@Ramesh:2019va] leading to joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
+However, game theory suggests that the ggreater number of parties negotiating, the harder it is for parties to reach an agreement [@Sumaila:2013], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared species.
+
+# Caveats and uncertainties
+
+We have identified three main sources of uncertainty in our analysis and point out ways to move forward in order to produce more reliable results to inform policy dealing with international fisheries under climate change.
+First, in this analysis we relay in one Earth System Model (GFDL-ESM2M) and climate change scenario (RCP 8.5).
+While the use of multiple ESMs would allow to capture the structural uncertainty of the models [@Frolicher:2016dg], using the GFDL-ESM2M results in conservative estimates as its projections are overall less extreme than similar models [@Bopp:2013fg].
+Addressing the uncertainty in future projections is key to advance marine governance for species on the move [@Pinsky:2018cb], thus, future efforts should look at the ensemble members of multiple ESM to capture the uncertainty embedded in the models.
+Secondly, we set an arbitrary threshold of $\pm$ 1 s.d.
+(67% confidence) to determine when a ToE and SSR would happen (see methods).
+We tested the sensitivity of our results to a larger threshold (e.g.
+$\pm$ 2 s.d for a 95% confidence).
+While the selected threshold is more relaxed than a 2 s.d., the distribution of results do not dramatically change between thresholds with a difference in ToE of about 4 years suggesting that this is a robust threshold (Fig.
+S4).
+Thirdly, in here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations [@FishForVisa].
+Thus, our analysis is carried at the species level, rather than a more detailed organizational level (e.g.
+fish stock) due to limited spatial-specific data on species sub populations (species).
+Reproducing our analysis with spatially-explicit stock data would allow to produce more certain results in changes in both SSR and ToE of transboundary species, and potentially identity different types of shared stock shifts [@Link:2010ei].
+However, many regions of the world lack these detailed data.
+Moreover, in many EEZs of the world fisheries are often managed at the species level [@FishForVisa].
+
+# Conclusion
 
@@
 
-## Supplements
+# Supplements
 
@@
 
-```{r Figure2, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
+```{r FigureS2, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
 
@@
 Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial stocks that will see a shift.
+Polygons represent the percentage of initial species that will see a shift.
 Panel B zooms in to specific areas shown in grey boxes of A."}
@@
 
-<!-- The time and number of emerging stocks decreases as we move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
+<!-- The time and number of emerging species decreases as we move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
 -->
@@
 -->
+
+<!-- In cases where the species is distributed in more than 2 neighboring EEZs (e.g. *xxx*), the index is as follows: -->
+
+<!-- $$I =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h})}) |  \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{C,y}}{sd(D_{B,h})}) | \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{Cy}}{sd(D_{B,h})})/N$$ -->
+
+<!-- In other words, for two neighboring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases (see *TI* equation).
+Consequently, if the centroid shifts are within historical variability (*SNR* < 1), risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability (*SNR* > 1), the risk of dispute is higher.
+Thus, risk is positively related to the frequency of centroid shifts beyond historical variability.
+-->

### 2020-10-15 Scripts/Manuscript.Rmd
Updated figures orders and manuscriopt

@@
   pdf_document: default
-# bibliography: ../References/FishForVisa_Ref.bib
+bibliography: ../References/ToE_Ref.bib
 csl: ../References/science.csl
@@
 
-# Abstract (150  words)
+<!-- # Abstract (150  words) -->
 
@@
 
-Climate change shifts in marine species distributions is affecting the current frameworks emplaced to manage species shared fisheries.
+<!-- Climate change shifts in marine species distributions is affecting the current frameworks emplaced to manage species shared fisheries.
 Here, we rely on a modeling framework with multiple ensemble members to identify the time frame by which shared stocks distribution will change from historical ranges, and the change in the proportion of the distrirbution that each EEZ has under climate change.
@@
 Our results suggest that most changes will happen before 2030 urging policy to be re-formulated to account for the possible implications of these changes.
+-->
 
-# Introduction
+## Introduction
 
-Over the last century human activities have fundamentally changed the physicochemical composition of the ocean, warming the waters, decreasing the pH and reducing the oxygen concentration [@IPCC:2019to].
+Over the last century human activities have fundamentally changed the physicochemical composition of the ocean, warming the waters, decreasing the pH and reducing the oxygen concentration [@IPCC:2019tn].
 As a consequence, many marine species have changed their distribution [@Poloczanska:2016kk] and will continue to do so in the search of the appropriate environmental conditions needed for their subsistence [@Cheung:2010dt].
@@
 While countries are responsible for the management of stocks within their EEZs, UNCLOS encourages actions to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
-Today, an estimate of 633 transboundary fish species are found in different continents and across multiple habitats averaging 48.5 million tones and representing USD 78 billion in fishing revenue between 2005 and 2010 (Chapter 1).
+Today, an estimate of 633 transboundary fish species are found in different continents and across multiple habitats averaging 48.5 million tones and representing USD 78 billion in fishing revenue between 2005 and 2010 (Chapter Two).
 
 The management, and thus sustainability, of fisheries that target shared species is jeopardized as marine species shift their known distribution to cope with a changing world [@Pinsky:2018cb].
-Frequently, the rules placed to manage these species are set to historical trends [@Baudron:2020dj] and do not necessarily consider the biogeography of species [@Alexa], nor the effects of climate change [@FishForVisa].
+Frequently, the rules placed to manage these species are set to historical trends [@Baudron:2020dj] and do not necessarily consider the biogeography of species [@FredstonHermann:2018kp], nor the effects of climate change (Chpater One).
 Misalignment between policies and distributional shifts have previously resulted in unsustainable harvest and international conflict [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij] and are expected to continue towards the end of the $21^{st}$ century.
-Thus, there is a need to understand the time when climate change will affect the sharing distribution of these species as well as the intensity in order to prepare ocean governance to species on the move [@Pinsky:2018cb; @FishForVisa; @Link:2010ei].
+Thus, there is a need to understand the time when climate change will affect the sharing distribution of these species as well as the intensity in order to prepare ocean governance to species on the move [@Pinsky:2018cb; @Link:2010ei, Chpater One].
 
-Here we endeavor to address this gap by estimating the Time of Emergence (ToE) [@Hawkins] and Stock Share Ratio (SSR) [@OCPalacios] of 633 transboundary species worldwide.
-We define ToE as the time when a change  of a particular variable exceeds the historical natural internal variability [@IPCC:2019to], a concept previously used to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja].
-Moreover, SSR is adopted from Game Theory as a proxy of thread point - the minimum payoff a player is willing to receive in order to cooperate with others [@OCSumaila; @Sumaila:2013].
-We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected shared species [@FishForVisa] by 280 EEZs of 198 coastal countries under a high greenhouse gas emission scenario RCP8.5 (See methods).
-We treated each species in an EEZ as a single stock due to a lack of more detailed spatial data [@FishForVisa] and only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g.
+Here we endeavor to address this gap by estimating the Time of Emergence (ToE) [@Hawkins:2012hc] and Stock Share Ratio (SSR) (Chapter Four) of 633 transboundary species worldwide.
+We define ToE as the time when a change  of a particular variable exceeds the historical natural internal variability [@IPCC:2019tn], a concept previously used to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja].
+Moreover, SSR is adopted from Game Theory as a proxy of thread point - the minimum payoff a player is willing to receive in order to cooperate with others [@Sumaila:2020tp; @Sumaila:2013hv].
+We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected shared species (Chpater One) by 280 EEZs of 198 coastal countries under a high greenhouse gas emission scenario RCP8.5 (See methods).
+We treated each species in an EEZ as a single stock due to a lack of more detailed spatial data (Chpater One) and only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g.
 we did not include the high seas).
@@
 
-# Materials and methods
+## Materials and methods
 
-## Databases and species selection
+### Databases and species selection
 
-We included in this analysis 633 marine exploited transboundary species representing 80% of the catch taken from the world's EEZs between 2005 and 2014 [@FishForVisa].
-Each species within an EEZ was considered a single stock unit due to a lack of more detailed spatial information [@Teh:2015gd; @FishForVisa].
-We defined the boundaries of the world EEZs (e.g., Mexico Pacific EEZ and Mexico Atlantic EEZ) using the *Sea Around Us* shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic; Fig.
-S1) including islands territories, and determined the intersections between polygons using the R package *sf* [@PackagesfSimpleF:2018vp].
-Each EEZ was attributed to a geopolitical region according to the UN and biome [@Reygondeau:2019uh], and species were attributed a habitat preference [@FishBase].
+We included in this analysis 633 marine exploited transboundary species previously identified by representing 80% of the catch taken from the world’s EEZs between 2005 and 2014 (Chapter Two).
+Each species within an EEZ was considered a single stock unit due to a lack of more detailed spatial information per exploited population [@Teh:2015gd, Chpater One].
+We defined the boundaries of the world EEZs using the *Sea Around Us* spatial division (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided in Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:FigureS1)) including islands territories, and determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp].
+Each EEZ was attributed to a geopolitical region according to the UN and biome [@Reygondeau:2019uh], and species were attributed a habitat preference according to FishBase.
 For each species and EEZ we used the *Sea Around Us* data to estimate the catch and fishing revenue from fishing activities within the global EEZs.
@@
 
-## Projecting species distributions under climate change
+### Projecting species distributions under climate change
 
-We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM).
+We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
 The DBEM algorithm includes species physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
@@
 
-We coupled the DBEM with the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project species distribution from 1951 to 2100 under climate change [@Dunne:2012bo; @Dunne:2013db].
-We run the DBEM simulating biomass and MCP from 1950 to 2100 under a “worst-case” climate change scenario according to the Representative Concentration Pathway 8.5 (RCP 8.5) used by the Intergovernmental Panel on Climate Change (IPCC) [@IPCC:2014bz; Riahi:2011dk].
-According to the GFDL-ESM2M, under the RCP 8.5 scenario atmospheric temperature will increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
+We forced the DBEM with environmental data from a ten ensemble member simulation of  the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project species distribution from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db].
+The GFDL ESM2M is run under historical forcing till 2005 and follows the high ” climate change emission scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
+In GFDL-ESM2M, the atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
 Because the main approach of this paper relies on understanding the spatial and temporal variation of a species distribution, we have to understand both the historical as well as the future climate-driven distribution variability, in order to infer the differences.
-We do this by following an ensemble member approach where each member constitutes a different initial Earth system condition [@Rodggers:2015jq; @Frolicher:2009ge].
-Thus, for our experiment, in each of the ten GFDL-ESM2M-ensembles the Earth system initial condition is infinitesimally different resulting in a unique atmosphere and ocean state at each point in time.
-This results in a different state of internal variability for each ensemble representing a plausible future under RCP 8.5 [@Rodggers:2015jq].
-Following this approach allows us to estimate a secular trend for the natural variability (e.g.
-background noise) and isolate the climate signal of a stock’s distribution by averaging the secular trend over all ensemble members.
+We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
+Thus, for our experiment, each of the ten GFDL-ESM2M-ensembles were started from infinitesimally different Earth system initial conditions resulting in a unique atmosphere and ocean state at each point in time.
+Following this approach allows us to estimate the natural internal variability (e.g.
+background noise) and isolate the forced climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
 
-## Estimating the Time of Emergence of transboundary species
+### Estimating the Time of Emergence of transboundary species
 
-Knowing the time (e.g., year) in which a shared stock is going to diverge from a historical variability is key for preparing to shifting species [@Link:2010ei; @FishForVisa].
-Here, we coupled the DBEM with the ten GFDL ensemble members described above, and developed an index of transboundary range shift, fisheries and climate change (*TI*).
+Knowing the time (e.g.
+year) in which a shared stock is going to diverge from the natural internal variability is key for preparing to shifting stocks [@Link:2010ei].
+Here, we coupled the DBEM with the ten GFDL ensemble members described above, and developed an index of transboundary range shift, fisheries and climate change (TI).
 The index is based on the variation of a species distribution centroid in relation to the centroid of the neighboring EEZs it occupies.
 Each species centroid was determined by the average latitude and longitude of the grid cells containing the top 10% abundance percentile of each year derived from the DBEM.
-The EEZ centroid was estimated using the *st* package of *R* statistical software (**Supplements Fig 1 (SuppM)**).
-For each ensemble member, neighboring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring elevation differences using the *geosphere* package of *R* statistical software;
+The EEZ centroid was estimated using the st package of R statistical software (Figure \ref{fig:FigureS1}).
+For each ensemble member, neighboring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring geographic barriers using the geosphere package of R statistical software;
 
@@
 
-where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectively for each future time step from 2006 to 2100 (*t*); *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectively for each time step from 2006 to 2100 (t); sd is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
 The index was then smoothed to a 10 years average.
-Finally, we determined the time of emergence setting an arbitrary threshold following Rodggers (2015):
+Finally, we determined the time of emergence following Rodgers et al.
+(2015):
 
@@
 -->
-We estimated the mean and one standar deviation ($\mu \pm \sigma$ for a confidence interval of 68%) to show the first year when the "signal" of the transboundary index overpasses the historic "noise".
-Since this was an arbitrary decision, we have included a more conservative estimation ($\mu \pm 2*\sigma$ for a 95% confidence interval).
-The SNR for both bottom and surface sea temperature was also estimated for statistical analysis following the SNR method.
 
-### Vulnerability index
+We estimated the mean and one standard deviation ($\mu \pm \sigma$ for a confidence interval of 68%) to show the first year when the “signal” of the transboundary index overpasses the historic “noise”.
+Since this was an arbitrary decision, we have included also a more conservative estimation in the supplements ($\mu \pm 2*\sigma$ for a 95% confidence interval).
+The ToE for both bottom and surface sea temperature was also estimated for statistical analysis following the SNR method.
+
+#### Vulnerability index
 
@@
 
-## Changes in transboundary Species Share Ratio
+### Changes in transboundary Species Share Ratio
 
-We adopted the concept of threat point as the minimum proportion of the species’ joint distribution between neighboring EEZs that each country will accept if they were to cooperate in the management of the stock [@OCPalacios].
-For each species and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighboring EEZs, and then calculated the proportion that each EEZ had [@FishForVisa].
+We adopted the concept of threat point as the minimum proportion of the species’ joint distribution between neighboring EEZs that each country will accept if they were to cooperate in the management of the stock (Chapter Four).
+For each species and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighboring EEZs, and then calculated the proportion that each EEZ had Chpater One.
 We then average the results in three time periods in order to reduce climate variability within the projections.
@@
 
-## Time of Emergence of transbounudary species
+### Time of Emergence of transbounudary species
 
-We estimated the ToE between neighboring EEZs for 633 transboundary species, that is, the year where the future stock's distribution (2006-2100) overpasses the historical variation (1951-2005).
-Since 2006, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g.
-*Engraulis ringens* shared by Chile and Peru emerged in 2010).
-species will continue to emerge steadily towards 2097 when all 633 species will have emerged in at least one neighboring EEZ (Fig.
-1).
-The number of EEZs with emerging species will increase steeply until 2020 when it seems to flatten approaching 2081 when all of the EEZs analyzed here will have at least one emerging stock (Fig.
-1).
+We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock’s distribution (2006-2100) overpasses the historical variation (1951-2005).
+Since 2006 till 2020, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g., *Engraulis ringens* shared by Chile and Peru emerged in 2010).
+Stocks will continue to emerge steadily towards 2097 when all 633 species will have emerged in at least one neighboring EEZ (Figure \ref{fig:Figure1}).
+The number of EEZs with emerging stocks will increase steeply until 2025 when it seems to flatten approaching 2081 when all of the EEZs analyzed here will have at least one emerging stock (Figure \ref{fig:Figure1}).
 
-```{r Figure1, eval = T,  echo = F, fig.width = 5, fig.height=5, fig.cap =  "Cumulative increase in time of emergence by Exclusive Economic Zone and transboundary species in absolute numbers and percentage."}
+```{r Figure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.cap =  "Cumulative increase in time of emergence by Exclusive Economic Zone and transboundary species in absolute numbers and percentage."}
 
@@
 
-Most of the world EEZ’s will see an average ToE across transboundary species before 2040 (Fig.
-2).
-In a few EEZs species will have a relative late average ToE (e.g.
+Most of the world EEZs are projected to experience an average ToE across transboundary species before 2040 (Figure \ref{fig:Figure2}).
+In a few EEZs species are projected to have a relative late average ToE (e.g.
 Honduras average ToE is 2071) while 54 of the 280 EEZs (19%) have already seen their species emerged (e.g.
-average ToE across all species <= 2020).
+average ToE across all species $\leq$ 2020).
 A higher number of emerging species will happen between latitudes 0$^o$ and 50$^o$N corresponding to regions of relatively small EEZs with multiple borders (e.g.
-The Caribbean and the Mediterranean) (Fig.2).
-In most cases, the current participation of the emerging species in a country's total fishing revenue is less than 50% (Fig.
-2 - Polygons).
+The Caribbean and the Mediterranean) (Figure \ref{fig:Figure2}).
+In most cases, the current participation of the emerging species in a country’s total fishing revenue is less than 50% (Figure \ref{fig:Figure2}- Polygons).
 However, in some cases emerging species represent 85% of current transboundary fishing revenue (e.g.
-Peru) or more (e.g.
-Marshall Islands, 92%).
-Only in very few cases like Brazil and Alaska's polar EEZ, no specie will emerge between 2006 and 2100.
+Peru) or more (e.g., Marshall Islands, 92%).
+Only in very few cases like Brazil and Alaska’s polar EEZ, no stock will emerge between 2006 and 2100.
 
-```{r Figure2, eval = T,  echo = F, fig.width  = 12, fig.height =10, fig.cap =  "Average time of emergence of transboundary specie.
+```{r Figure2, eval = T,  echo = F, fig.width  = 13, fig.height =11, fig.cap =  "Average time of emergence of transboundary specie.
 Land polygon shows the vulnerability index.
@@
 
-Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transboundary species between 2005 and 2010 will see an average ToE before 2030 (Fig.
-3).
+Regionally, most of African, Asian and European EEZs are projected to have a later ToE, while the Americas and Oceania will see an early ToE (Figure \ref{fig:Figure3}).
+Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transboundary species between 2005 and 2010 will see an average ToE before 2030 (Figure \ref{fig:Figure3}).
 This includes the United States who occupies the $4^{th}$ place in revenue generated from transboundary fisheries.
-These 47 countries are also among those with the largest number of emerging taxa.
-Regionally speaking, most of African, Asian and European EEZs will see a latter ToE, while the Americas and Oceania will see an early ToE (Fig.
-3).
+These 47 countries are also among those with the largest number of emerging taxa (Figure \ref{fig:Figure3}).
 From the top ten countries with more revenue from transboundary species, nine of them share a border.
 Most of these countries (e.g.
-China) will also have a large number of emerging species, while others like Peru and Norway will have a small number of emerging taxa but representing high revenue fisheries.
+China) will also have a large number of emerging stocks, while others like Peru and Norway will have a small number of emerging taxa but representing high revenue fisheries.
 
-```{r Figure3, eval = T,  echo = F, fig.width=5, fig.height=6, fig.cap =  "Average time of emergence of transboundary species per fishing entity.
+```{r Figure3, eval = T,  echo = F, fig.width=4, fig.height=5, fig.cap =  "Average time of emergence of transboundary species per fishing entity.
 Fishing revenue in logarithmic scale.
@@
 
-We compared species ToE within EEZs by geographic region and by species habitat preference (Fig.
-4).
+We compared species ToE within EEZs by geographic region and by species habitat preference (Figure \ref{fig:Figure4}).
 The median ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001).
@@
 EEZs in Easter Asia which have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ from (almost) all other regions.
-Other regions also significantly differ from each other but in a less frequent manner (See Table S1 for full statistics).
+Other regions also significantly differ from each other but in a less frequent manner (see Table \ref{tab:TableS1}).
 When looking specifically at habitat association, we found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
-Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table S2 for full statistics).
+Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table \ref{tab:TableS2}).
 
-```{r Figure4, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap =  " Regional time of emergence per habitat preference.
+```{r Figure4, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap =  "Regional time of emergence per habitat preference.
 Regions according to the United Nations and habitat preference to FishBase.
@@
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers (outside of 1.5*IQR."}
+Line represents median, and black points represent outliers outside of 1.5*IQR."}
 
@@
 
-## Changes in specie Share Ratio (SSR)
+### Changes in specie Share Ratio (SSR)
 
-```{r analysis_results, eval = T,  echo = FALSE, include = FALSE}
+```{r analysis_results, eval = F,  echo = FALSE, include = FALSE}
 
@@
 
-We estimated the change in SSR (with a 63% of confidence) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) under a high emission climate change scenario.
-Our results suggest that `r n_eez_per`% (n = `r n_eez`) of the world EEZs will see a change in SSR by 2030 (Fig.
-5A).
-Most EEZs will average a change lower than 25% which is expected to slightly increase in intensity towards the  mid $21^{st}$ century (**Supplements Fig 5 (histogram)**).
+We estimated the change in SSR (with a 63% of confidence) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) (Figure \ref{fig:Figure5}).
+Our results suggest that 87% (n = 238) of the world EEZs will see a change in SSR by 2030 (Figure \ref{fig:Figure5}A).
+Most EEZs will average a change lower than 25% which is expected to slightly increase in intensity towards the mid $21^{st}$ century (Figure \ref{fig:FigureS2}).
 However, this result masks larger changes in SSR for specific species.
-For example, Guatemala's SSR of *Panulirus gracilis* shared with Mexico is projected to go from 12% to 26%.
-By early $21^{st}$ century, the percentage of transboundary species per EEZ that will see a change in SSR $\pm$ 1 s.d.
+For example, Guatemala’s SSR of *Panulirus gracilis* shared with Mexico is projected to go from 12% to 26%.
+By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR $\pm$ 1 s.d.
 of historical distribution will be relatively large (global mean of 69% of species in an EEZ).
-For 17 EEZs, 100% of their analyzed transboundary species will shift, while only six EEZs will have less than 25% shifting species (Fig.
-5A-Polygons).
+For 17 EEZs, 100% of their analyzed transboundary stocks will shift, while only six EEZs will have less than 25% shifting stocks (Figure \ref{fig:Figure5}A-Polygons).
 Regions where EEZs are relatively small and have multiple borders (e.g.
-The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Fig.
-5B).
-Towards the mid $21^{st}$ century, changes in SSR per EEZ will not be substantial (Fig.
-S1) relative to the early $21^{st}$ century but are expected to increase in intensity (Fig.
-S2).
+The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Figure \ref{fig:Figure1}B).
+Towards the mid $21^{st}$ century, changes in SSR per EEZ will not be substantial (Figure \ref{fig:FigureS3}) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \ref{fig:FigureS3}).
 
-```{r Figure1, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average gain of transboundary specie share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial species that will see a shift.
+```{r Figure5, eval = T,  echo = F, fig.width=12, fig.height=10, fig.cap =  "Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average gain of transboundary stock share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial stocks that will see a shift.
 Panel B zooms in to specific areas shown in grey boxes of A."}
 
-img <- readPNG("../Figures/Fig1.png")
+img <- readPNG("../Figures/Fig5.png")
 grid.raster(img)
@@
 Our findings here are noteworthy in several ways.
-First, we show that the current distribution of most transboundary species have already emerged from the historical record and that virtually all EEZs have seen at least one emerging species since 2006 (Fig.
-5).
-Second, we identify regional patterns of complex changes in SSR and levels of vulnerability of transboundary fisheries management to climate change (Fig.
-3).
-Third, we identified that the historical share proportion of transboundary species will change by as soon as 2030 in most neighboring EEZs (Fig.
-5).
+First, we show that the current distribution of most transboundary species have already emerged from the historical record and that virtually all EEZs have seen at least one emerging species since 2006 (Figure \ref{fig:Figure5}).
+Second, we identify regional patterns of complex changes in SSR and levels of vulnerability of transboundary fisheries management to climate change (Figure \ref{fig:Figure3}).
+Third, we identified that the historical share proportion of transboundary species will change by 2030 in most neighboring EEZs (Figure \ref{fig:Figure5}).
 
-### Emergence characteristics of transboundary species
+#### Emergence characteristics of transboundary species
 
-We set out to evaluate the emergence characteristics of transboundary species across the global EEZs.
-Our results suggest that most of the EEZs of the world and more than half of the transboundary species already emerged since the early 2000s (Fig.
-2).
+We set out to evaluate the emergence characteristics of transboundary stocks across the global EEZs.
+Our results suggest that in most of the EEZs of the world more than half of the transboundary species already emerged since the early 2000s (Figure \ref{fig:Figure2}).
 This is in alignment with previous research showing changes in the composition of marine catches since the 1970's as evidence of climate change [@Cheung:2013gk].
@@
 In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, led to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij].
-The large amount of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g.
-temperature, oxygen, $CO_2$) that influence the distribution of marine species [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
-It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 22.6% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while $O_2$ will have doubled [@Rodggers:2015j].
+The large amount of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
+It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 23% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while ($O_2$) will have doubled under high greenhouse gas emission scenario [@Rodgers:2015jq].
 
 Regional differences in ToE are likely the result of environmental factors, species vulnerability, and the characteristics of political boundaries.
-For example, the high number of species emerging early in the tropics suggest that tropical species are already on the move.
-The tropics are among the most species-rich biome [@Reygondeau:2019no] and have potentially experienced some of the largest declines in $O_2$ and greatest levels of warming since 2006 [@IPCC:2019tn; @Roemmich:2015id].
+For example, the high number of stocks emerging early in the tropics suggest that tropical species are already on the move.
+The tropics are among the most species-rich biome [@Reygondeau:2019uh] and have potentially experienced some of the greatest levels of warming relative to the natural internal variability  [@IPCC:2019tn; @Roemmich:2015id].
 Here, marine species live close to the thermal tolerance making them highly vulnerable to warming waters [@IPCC:2014bz] consequently having to move towards deeper or sub-tropical waters [@Poloczanska:2016kk].
 The premise of a shift already happening is supported by observational evidence of tropical species appearing in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis showing a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
-In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs and many borders that match peaks of emerging species.
-Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsible for substantial increases in emerging species (Fig.
-3).
+In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs and many borders that match peaks of emerging stocks.
+Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsible for substantial increases in emerging stocks (Figure \ref{fig:Figure3}).
 
-## Some fisheries will be more vulnerable to emerging species
+### Some fisheries will be more vulnerable to emerging species
 
-The premise of our index is that, from a management perspective, countries whose shifting transboundary species provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
-Similarly, the index is capable to capture different sources of societal benefits (e.g.
-food security trough catch or livelihood trough value).
+The premise of our vulnerability index is that, from a management perspective, countries whose shifting transboundary stocks provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
+Similarly, the index is capable of capturing different sources of societal benefits (e.g.
+food security through catch or livelihood through value).
 We assume that the management of fisheries that provide large societal benefits will have more pressure from stakeholders while earlier times of emergence will result in less time to act, thus resulting in higher vulnerability.
-Thus, an early ToE and large fisheries benefits could indicate high vulnerability of transboundary fisheries management to climate change (Fig.
-3).
-For example, Polynesia and Micronesia are the only regions where the average ToE across species and EEZs already happened; 2019 and 2021, respectively (Fig.
-5).
-Countries in these regions are highly dependent of pelagic-oceanic (e.g.
+Thus, an early ToE and large fisheries benefits could indicate high vulnerability of transboundary fisheries management to climate change (Figure \ref{fig:Figure3}).
+For example, Polynesia and Micronesia are the only regions where the average ToE across all species and EEZs already happened; 2019 or is about too 2021, respectively (Figure \ref{fig:Figure5}).
+Countries in these regions are highly dependent on pelagic-oceanic (e.g.
 tunas) and reef-associated transboundary species (e.g.
-*Speceies*) for both food security [@REF] and profits [@REF].
-For example, in the Marshall Islands 16 transboundary species responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, in averaged, emerged since 2018 resulting in high vulnerability index (Fig.
-5).
-In this cases policy has to work against the clock and under potentially high societal pressure to address an issue that could be impacting important fisheries.
-However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to species sustainability.
-While it gives countries more time to prepare for substantial changes, species with relatively small overall participation can still be a source of conflict as portrayed by the Atlantic mackerel case [@FishForVisa; @Spijkers:2017ij].
+Species) for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh; @Cabral:2018hc].
+For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \ref{fig:Figure5}).
+In this case, policy has to work against the clock and under potentially high societal pressure to address an issue that could be impacting important fisheries.
+However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to stocks sustainability.
+While it gives countries more time to prepare for substantial changes, species with relatively small overall participation can still be a source of conflict as portrayed by the Atlantic mackerel case (@Spijkers:2017ij, Chapter One].
 
-## Changes in specie-share ratio will be everywhere
+### Changes in specie-share ratio will be everywhere
 
-By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Fig.
-1).
+By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Figure \ref{fig:Figure5}).
 Poleward shifts in SSR will be seen in regions of the world with relatively large EEZs and few boundaries like the North East Pacific.
 However, despite a general recognition that climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn], the Eastern Tropical Pacific will see an equatorial shift with EEZs from Mexico to Ecuador losing SSR to their southern neighbors.
-This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico's EEZ towards Peru under a high emission climate change scenario [@TayClarkeChp2].
-Moreover, regions with presence of multiple frontiers like the Caribbean and the North East Atlantic will see a complex network of changes in SSR with no marked direction (Fig.
-1B).
-These regions are known for being highly connected by both fish adult migration [@REF] and larvae dispersal [@Ramesh:2019va] leading to joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
-However, game theory suggests that the ggreater number of parties negotiating, the harder it is for parties to reach an agreement [@Sumaila:2013], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared species.
+This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario (Clark *et al.,* n.d.).
+Moreover, regions with presence of multiple frontiers like the Caribbean and the North East Atlantic will see a complex network of changes in SSR with no marked direction (Figure \ref{fig:Figure5}B).
+These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] leading to joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
+However, game theory suggests that the greater number of parties negotiating, the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared species.
 
-# Caveats and uncertainties
+## Caveats and uncertainties
+
+We have identified three main sources of uncertainty in our analysis and point out ways to move forward in order to produce more reliable results to inform policy dealing with international fisheries under climate change.
+First, ouranalysis relies one Earth system model (GFDL-ESM2M) and climate change scenario (RCP 8.5).
+While the use of multiple ESMs would allow to capture the structural uncertainty of the models [@Frolicher:2016dg], using the GFDL-ESM2M results in conservative estimates as its projections, for example in sea surface temperature, are overall less extreme than in other CMIP5-type Earth system models [@Bopp:2013fg].
+Addressing the uncertainty in future projections is key to advance marine governance for species on the move [@Pinsky:2018cb], thus, future analysis should include  multiple ESMs.
+Secondly, we set an arbitrary threshold of $\pm$ 1 s.d.
+(67% confidence) for determining a ToE and SSR (see methods).
+We tested the sensitivity of our results using a larger threshold (e.g., $\pm$ 2 s.d for a 95% confidence).
+While the selected  1 s.d.
+threshold is more relaxed than a 2 s.d., the distribution of results do not dramatically change between thresholds with a difference in ToE of about 4 years (Figure \ref{fig:FigureS4}) Thirdly, here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations (Chpater One).
+Thus, our analysis is carried at the species level, rather than a more detailed organizational level (e.g.
+fish stock) due to limited spatial-specific data on species sub populations (stocks).
+Moreover, in many EEZs of the world fisheries are often managed at the species level (Chpater One) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
+However, reproducing our analysis at the local level, where spatially-explicit stock data is available would allow to produce more certain results in changes in both ToE and SSR of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level[@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
 
@@
 $\pm$ 2 s.d for a 95% confidence).
-While the selected threshold is more relaxed than a 2 s.d., the distribution of results do not dramatically change between thresholds with a difference in ToE of about 4 years suggesting that this is a robust threshold (Fig.
-S4).
-Thirdly, in here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations [@FishForVisa].
+While the selected threshold is more relaxed than a 2 s.d., the distribution of results do not dramatically change between thresholds with a difference in ToE of about 4 years suggesting that this is a robust threshold (Figure \ref{fig:FigureS4}).
+Thirdly, in here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations (Chpater One).
 Thus, our analysis is carried at the species level, rather than a more detailed organizational level (e.g.
@@
 However, many regions of the world lack these detailed data.
-Moreover, in many EEZs of the world fisheries are often managed at the species level [@FishForVisa].
+Moreover, in many EEZs of the world fisheries are often managed at the species level (Chpater One).
 
-# Conclusion
+## Conclusion
 
-Preparing policy to species on the move is key to achieve sustainable development goals and proper governance of the world fisheries [@Pinsky:2018cb].
-Here, we have addressed two fundamental steps to achieve sustainable international fisheries under a changing world [@Link:2010ei]; the degree of change that share species will have and the year in witch species will shift.
-Such results have direct implications for ocean governance.
-By identifying the degree location and time by which transboundary species will change from historical normal we provide critical information for regional management bodies and countries to anticipate the potential complications that climate change will bring to international fisheries management.
-While our results are not free of uncertainty and would benefit from more detailed analysis at localize levels, it represents a crucial step towards preparing ocean governance for species on the move.
+Global society has set the ambitious goal to manage all fisheries sustainably with no overfishing (SDG 14 – Live below water) [@UnitedNations2018] in the incoming years, which in turn will have direct repercussion in others [@Singh:2017ds].
+Preparing policy to species on the move is key to achieve these sustainable development goals and proper governance of the world fisheries [@Pinsky:2018cb].
+Here, we have addressed two fundamental steps to achieve sustainable international fisheries under a changing world [@Link:2010ei].
+We have identified the transboundary species that will see a shift in shared distribution, the year in which such distribution will shift, as well as the degree of change.
+Our results have direct implications for ocean governance.
+By identifying the time, degree, and location by which transboundary species will change from a historical normal we provide critical information for regional management bodies and countries to anticipate the potential complications that climate change will bring to international fisheries management.
+While our results are not free of uncertainty and would benefit from more detailed analysis at local levels, they represent a crucial approximation towards preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg, Chapter One].
 To address the issues raised in this paper, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
 
-# Supplements
+## Supplements
 
-```{r FigureS1, eval = T,  echo = F, fig.width=9, fig.height=6, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
+```{r FigureS1, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
 
@@
 
-```{r FigureS2, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
+```{r FigureS2, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
 
@@
 
-<!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005))](../Figures/Supp2.png) -->
+<!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005))](../Figures/Supp2.png) -->
 
@@
 
-```{r FigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98 confidenc teshold (yellow"}
+```{r FigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
 

### 2020-11-09 Scripts/Manuscript.Rmd
Addressed W. and C. comments untill Results

@@
 ---
-title: Matching the Time of Emergence of Transboundary Fish species to Lead Time for Policy Response Under Climate Change
+title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change
 subtitle: Challenges to transboundary fisheries under climate change
@@
 
-Over the last century human activities have fundamentally changed the physicochemical composition of the ocean, warming the waters, decreasing the pH and reducing the oxygen concentration [@IPCC:2019tn].
-As a consequence, many marine species have changed their distribution [@Poloczanska:2016kk] and will continue to do so in the search of the appropriate environmental conditions needed for their subsistence [@Cheung:2010dt].
-Climate change is expected to impact fisheries production worldwide compromising our capacity to reach international sustainability goals such as the SDG 14 - life below water [@UnitedNations2018; @Cheung:2010dt; @Barange:2014kr].
-Such impacts can be partly offset with the proper reform of current fisheries management [@Gaines:2018sg; @Cheung:2018dh].
-However, these reforms will require comprehensive plans and institutions to manage species that cross international borders, also known as shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
+Over the last century, human activities have fundamentally changed the physico-chemical composition of the ocean, substantially warming the waters, decreasing the pH and reducing oxygen concentration [@IPCC:2019to].
+Throughout the ocean, marine species are distributed according to preferences in given environmental parameters (e.g., salinity, temperature, oxygen) [@Hutchinson:1957nt; @Nelson:2016kk].
+As a result of climate change, many marine species have changed their distribution towards higher latitude, deeper water or following local environmental gradients [@Poloczanska:2016kk].
+Biogeography of marine species is projected to continue to shift as ocean conditions continue to change in the 21st century under climate change [@Cheung:2010dt].
+Climate change is expected to impact fisheries production, compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Cheung:2016Sci; @Barange:2014kr; @Singht:2019].
+The projected risks of impacts can be reduced through improving the effectiveness of current fisheries management [@Gaines:2018sg; @Cheung:2018dh], including managing fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
 
-In the early 1980’s the United Nations officialized the concept of Exclusive Economic Zone (EEZs) through the Law of the Sea (UCLOS) [@UnitedNations:1986tl], partitioning the distribution of several species, thus, creating what we know today as shared stocks.
-As defined by the Food and Agriculture Organization, shared stocks can be classified in three non-exclusive categories.
-(*i*) Transboundary, those that cross neighboring EEZs; (*ii*) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@Munro:2004th].
-While countries are responsible for the management of stocks within their EEZs, UNCLOS encourages actions to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
-Today, an estimate of 633 transboundary fish species are found in different continents and across multiple habitats averaging 48.5 million tones and representing USD 78 billion in fishing revenue between 2005 and 2010 (Chapter Two).
+The concept of shared stocks were developed following the ratification of  the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
+As defined by the Food and Agriculture Organization, shared stocks can be classified into three non-exclusive categories: (*i*) transboundary, those that cross neighboring EEZs; (*ii*) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@Munro:2004th].
+While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
+A recent study estimates that, globally, there are 633 transboundary fish species, yielding an average 48.5 million tonnes of catch and USD 78 billion in fishing revenue, per year, between 2005 and 2010 [@PalaciosAbrantes:2020tv].
 
-The management, and thus sustainability, of fisheries that target shared species is jeopardized as marine species shift their known distribution to cope with a changing world [@Pinsky:2018cb].
-Frequently, the rules placed to manage these species are set to historical trends [@Baudron:2020dj] and do not necessarily consider the biogeography of species [@FredstonHermann:2018kp], nor the effects of climate change (Chpater One).
-Misalignment between policies and distributional shifts have previously resulted in unsustainable harvest and international conflict [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij] and are expected to continue towards the end of the $21^{st}$ century.
-Thus, there is a need to understand the time when climate change will affect the sharing distribution of these species as well as the intensity in order to prepare ocean governance to species on the move [@Pinsky:2018cb; @Link:2010ei, Chpater One].
+The effectiveness of managing fisheries targeting shared stocks is challenged by the shifts in marine species distribution under climate change [@Pinsky:2018cb; @Pinsky:wq].
+In many cases, catch or fishing effort quota on shared stocks are based on historical records, [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change in the fish stocks and fisheries [@PalaciosAbrantes:2020tv].
+Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by worsening climate change [@Pinsky:2018cb].
+With future shifts in species distributions under climate change, the challenges posed on shared stocks fisheries management will continue to increase in the $21^{st}$ century.
+Thus, there is a need to understand the time when climate change will affect the sharing distribution of these species as well as the intensity in order to prepare ocean governance to species on the move [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
 
-Here we endeavor to address this gap by estimating the Time of Emergence (ToE) [@Hawkins:2012hc] and Stock Share Ratio (SSR) (Chapter Four) of 633 transboundary species worldwide.
-We define ToE as the time when a change  of a particular variable exceeds the historical natural internal variability [@IPCC:2019tn], a concept previously used to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja].
-Moreover, SSR is adopted from Game Theory as a proxy of thread point - the minimum payoff a player is willing to receive in order to cooperate with others [@Sumaila:2020tp; @Sumaila:2013hv].
-We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected shared species (Chpater One) by 280 EEZs of 198 coastal countries under a high greenhouse gas emission scenario RCP8.5 (See methods).
-We treated each species in an EEZ as a single stock due to a lack of more detailed spatial data (Chpater One) and only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g.
-we did not include the high seas).
-Our results suggest that many shared species have already emerged and shifted their SSR in most neighboring EEZs posing a threat to the sustainability of shared stocks and the resilience of international treaties.
+We propose that adapting the concepts of "Time of Emergence" (ToE) and “Thread Point” (TP) can help address some of the management challenges imposed by shifting species.
+Roughly, ToE is defined as the moment in time when an anthropogenic change signal (e.g., future trend) rises above the background noise of natural variability (e.g., historical variation) [@IPCC:2019tn].
+This concept has been applied to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
+The premise behind ToE is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background of natural climate variability [@Hawkins:2012hc].
+The concept of thread point (TP) comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
+Game theory has been widely used to investigate and manage shared stocks [@Sumaila:2013hv].
+Specifically, Sumaila et al., *in press* applied this concept to understand the impacts of climate change to the economics of shared species between Canada and the United States.
+
+Here, we estimate the ToE and TP of the share distribution of transboundary stocks worldwide.
+We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected shared species [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emission scenario RCP8.5 (see Methods).
+We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., we did not include the high seas).
+Our results suggest that many shared stocks have already ‘emerged’ and overpassed their share TP in most neighboring EEZs posing a threat to the sustainability of shared stocks and the resilience of international treaties.
 
@@
 
-We included in this analysis 633 marine exploited transboundary species previously identified by representing 80% of the catch taken from the world’s EEZs between 2005 and 2014 (Chapter Two).
-Each species within an EEZ was considered a single stock unit due to a lack of more detailed spatial information per exploited population [@Teh:2015gd, Chpater One].
-We defined the boundaries of the world EEZs using the *Sea Around Us* spatial division (updated 1 July 2015, available from http://www.seaaroundus.org) noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided in Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:FigureS1)) including islands territories, and determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp].
-Each EEZ was attributed to a geopolitical region according to the UN and biome [@Reygondeau:2019uh], and species were attributed a habitat preference according to FishBase.
-For each species and EEZ we used the *Sea Around Us* data to estimate the catch and fishing revenue from fishing activities within the global EEZs.
-We report both catch and revenue the average from the last decade (2005-2014) of available spatial data [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+This analysis includes 633 marine exploited transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
+Each species shared by a pair of neighbouring EEZs was considered a transboundary stock [@Teh:2015gd; @PalaciosAbrantes:2020tv].
+We defined the boundaries of the world’s EEZs using the *Sea Around Us* spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:FigureS1)) including island territories.
+We determined the intersections between polygons using the *R*-statistical software package *sf* [@PackagesfSimpleF:2018vp].
+Each EEZ was categorized by the geopolitical regions and natural biomes according to the United Nations and by natural biome [@Reygondeau:2019].
+Habitat preference of each species was determined based on information available from FishBase (http://www.fishbase.org).
+For each species and EEZ, we used the *Sea Around Us* data to estimate catch and fishing revenue from fishing activities within global EEZs.
+We report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
 The monthly average US Consumer Price Index (CPI) according to the U.S.
-Bureau of Labor Statistic (https://www.bls.gov/cpi/) was used to convert the original 2010 real USD value to 2019 real USD.
+Bureau of Labor Statistic (https://www.bls.gov/cpi/) was used to standardize all monetary values to 2019 real USD.
 
@@
 We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
-The DBEM algorithm includes species physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
-For pelagic species, the model uses environmental variables at the surface (e.g.
-sea surface temperature) whereas demersal species distribution are driven by ocean-bottom variables (e.g.
-bottom temperature).
-This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (See Cheung *et al.* [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
+The DBEM  represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
+For pelagic species, the model uses environmental variables at the surface (e.g., sea surface temperature) whereas demersal species distribution are driven by ocean bottom variables (e.g., bottom temperature).
+This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (see Cheung *et al.* [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
 
-We forced the DBEM with environmental data from a ten ensemble member simulation of  the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project species distribution from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db].
-The GFDL ESM2M is run under historical forcing till 2005 and follows the high ” climate change emission scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
-In GFDL-ESM2M, the atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a species distribution, we have to understand both the historical as well as the future climate-driven distribution variability, in order to infer the differences.
-We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
-Thus, for our experiment, each of the ten GFDL-ESM2M-ensembles were started from infinitesimally different Earth system initial conditions resulting in a unique atmosphere and ocean state at each point in time.
-Following this approach allows us to estimate the natural internal variability (e.g.
-background noise) and isolate the forced climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
+We forced the DBEM with environmental data from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+The GFDL-ESM2M is run under historical forcing until 2005 and follows the ‘high’ climate change emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [Riahi:2011dk].
+In GFDL-ESM2M, atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a species’ distribution, we have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
+We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodggers:2015jq; @Frolicher:2009ge].
+Thus, for our experiment, each of the ten GFDL-ESM2M-ensembles were started from infinitesimally small differences in Earth system initial conditions resulting in a unique atmosphere and ocean state at each point in time.
+This approach allows us to estimate the natural internal variability (e.g., background noise) and isolate the forced climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
 
-### Estimating the Time of Emergence of transboundary species
+### Calculating an index of transboundary range shift
 
-Knowing the time (e.g.
-year) in which a shared stock is going to diverge from the natural internal variability is key for preparing to shifting stocks [@Link:2010ei].
-Here, we coupled the DBEM with the ten GFDL ensemble members described above, and developed an index of transboundary range shift, fisheries and climate change (TI).
-The index is based on the variation of a species distribution centroid in relation to the centroid of the neighboring EEZs it occupies.
-Each species centroid was determined by the average latitude and longitude of the grid cells containing the top 10% abundance percentile of each year derived from the DBEM.
-The EEZ centroid was estimated using the st package of R statistical software (Figure \ref{fig:FigureS1}).
-For each ensemble member, neighboring EEZ and species, we computed the distance between centroids assuming a perfect sphere and ignoring geographic barriers using the geosphere package of R statistical software;
+We developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
+This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \ref{fig:FigureS2}).
+The centroid of a transboundary stock was determined by the average latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
+Therefore,
 
-$$D_{ens} = acos(sin(lat_{eez})*sin(lat_{spp}) + cos(lat_{eez})*cos(lat_{spp})*cos(lon_{eez}-lon_{spp}))$$
-where $lat_{eez}$ and $lat_{spp}$ are the latitudes of the EEZ and specie centroid, respectively and $lon_{eez}$ and $lon_{spp}$ are the respective longitudes.
+$lat_{ts} = \mu (lat_{per})$
+
+$lon_{ts} = \mu (lon_{per})$
+
+where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
+To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top $10^{th}$ percentile (e.g., *per* = 10%).
+We ran a sensitivity analysis (e.g.
+*per* = 20, 50, 90) to relax this parameter (Figure \ref{fig:FigureS2}).
+The centroid of a EEZ was estimated using the *st* package in *R* (version 3.2.1) (Figure \ref{fig:FigureS1}).
+For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in *R*;
+
+$$D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))$$
+where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively and $lon_{eez}$ and $lon_{ts}$ are the respective longitudes.
 
@@
 
-where $D_A$ and $D_B$ represent the distance between EEZ and distribution centroids of a species in EEZs A and B, respectively for each time step from 2006 to 2100 (t); sd is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
-The index was then smoothed to a 10 years average.
-Finally, we determined the time of emergence following Rodgers et al.
-(2015):
+where $D_A$ and $D_B$ represent the distance between the distribution centroids of a species and the centroid of the pair of neighbouring EEZs (A and B) sharing the species for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+Thus, a higher TI means that a shared species becomes more or less abundant relative to its neighbouring EEZs that is beyond what the fisheries experienced historically.
+From the perspective of a state’s fisheries, higher TI indicates an increase in the challenges to the effective management and sustainability of the shared species’ fisheries.
+In contrast, a smaller TI suggests that the sharing of the species between neighbouring EEZs is becoming relatively more stable; a condition that favours effective management and sustainability of the shared species’ fisheries.
+The index was smoothed to a 10 year average to xxx.
 
-$$ToE = TREND/NOISE$$
+#### Calculating the Time-of-Emergence of the transboundary index
 
-where *TREND* is the yearly mean *TI* ($\mu$) of all ten ensemble members and *NOISE* is the standard deviation of the TREND ($\sigma$) between 1951 and 2005. 
-<!-- Note that in some cases not all ensemble members show the presence of a species in a certain EEZ throughout the study's time frame.
-We presented the results only for those cases where all 10 ensemble members had results for each year between 1951 and 2100 (n = 150).
--->
+Knowing the point in time (e.g., year) at which the distribution of a shared stock will diverge from its natural internal variability is important in informing the lead-time to which climate adaptation in transboundary fisheries management needs to be implemented [@Link:2010ei].
+Thus, we define ToE as the time when the shared stock between neighbouring EEZs will exceed the historical natural internal variability as follows:
 
-We estimated the mean and one standard deviation ($\mu \pm \sigma$ for a confidence interval of 68%) to show the first year when the “signal” of the transboundary index overpasses the historic “noise”.
-Since this was an arbitrary decision, we have included also a more conservative estimation in the supplements ($\mu \pm 2*\sigma$ for a 95% confidence interval).
-The ToE for both bottom and surface sea temperature was also estimated for statistical analysis following the SNR method.
+$EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$
+$ToE = t when Emergence index[t] > 1$
 
-#### Vulnerability index
+where $\mu_{TI}$ is the yearly mean TI ($\mu$) across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005).
+We set a threshold of TI above which we consider a shared stock to have emerged from the historical internal variability of the TI.
+The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., eq.
+**x**, representing a probability of 68% that the transboundary index has emerged from historical variability).
+We tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
 
-We created a vulnerability index of fisheries management to shifts in transboundary species.
-The index is based on two factors; the ToE and the importance of the shifting species to a country's total catch and revenue from transboundary species;
+#### Computing a transboundary fisheries vulnerability index
+
+We created a country-specific index to examine the risk of impacts of species’ range shifts on transboundary fisheries management.
+The index reflects both the rate of change of shared stocks’ distribution (indicated by the time-of-emergence of the transboundary index) and the importance of the shifting species to a country’s total catch and revenue from the species.
+The index is calculated from:
 
 $$ index' = (P_c + P_v)/ToE $$
-where *P_c* and *P_v* are the proportion that the shifting species' will have in the total catch and value generated from transboundary fisheries within a county's EEZs, respectively.
-Here, total catch and value are represented as the average of 2005 to 2014 within the global EEZs according to the *Sea Around Us*.
-The index was then normalized to a 0 to 1 scale going from less vulnerable to more vulnerable:
 
-$$ index = \frac{(index' - N_{min})}{N_{max}-N_{min}} $$
+where *P_c* and *P_v* indicate the proportion accounted for by shared species in total catch and value generated from transboundary fisheries within a country’s EEZ, respectively.
+Here, the proportion of total catch and value are represented as the average of 2005 to 2014 within individual EEZs according to the *Sea Around Us*.
+The index was then normalized to a 0 (less vulnerable) to 1 (more vulnerable) scale.
 
-### Changes in transboundary Species Share Ratio
+### Estimating the threat point of transboundary stock share
 
-We adopted the concept of threat point as the minimum proportion of the species’ joint distribution between neighboring EEZs that each country will accept if they were to cooperate in the management of the stock (Chapter Four).
-For each species and time step (1951-2100) we first aggregate the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells where the species was present within neighboring EEZs, and then calculated the proportion that each EEZ had Chpater One.
-We then average the results in three time periods in order to reduce climate variability within the projections.
-The first time period ($t_h$) comprises (1951 - 2005) and is assumed as the historic share distribution of each species, thus, representing each neighboring nation's threat point.
-This time frame matches the historic climate data that the GFDL uses to project climate change [@Dunne:2012bo; @Dunne:2013db].
-We then selected two future periods; early $21^{st}$ century as the average of 2021 to 2040 ($t_e$), and mid $21^{st}$ century as the average of 2041 to 2060 ($t_m$).
-We choose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing) and SDG 1.2 (poverty reduction), aimed to be achieved by 2020 and 2030, respectively [@Singh:2017ds].
-The analysis was replicated for each of the ten ensemble members capturing the variation by averaging results among ensemble members ($\mu \pm \sigma$).
-We then estimate the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_{f}$) relative to the historic time period ($SSR_h$) as follows,
+In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than operating on a non-cooperative basis [@Munro:1980sd].
+However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., *Threat Point* (TP) [@Sumaila:2013].
+Here, we define TP as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbour [@OCPalacios; @OCSumaila].
+Any proportion below the defined threat point of a country would result in a deflection from the joint management of the shared stock.
+
+First we estimated the stock share ratio (SSR) for each species and year (1951-2100) by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the species is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
+Second, we average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
+The first time period ($t_h$) spans 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
+It is assumed as the historic baseline of shared distributions for each species.
+We then selected two future periods; early $21^{st}$ century as the average of 2021 to 2040 (*te*), and mid $21^{st}$ century as the average of 2041 to 2060 (*tm*).
+We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing) and SDG 1.2 (poverty reduction), to be achieved by 2020 and 2030, respectively [@Singh:2017ds].
+The analysis was replicated for projected species distributions from each of the ten ensemble members and results were averaged across ensemble members.
+Third, we estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_h$) following Palacios-Abrantes et al.
+(*in press*),
 
 $$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$
-To address the natural spatial variability of the stock we only considered $\Delta{SSR_{e,m}}$ to be true when $SSR_f \geq (SSR_f + \sigma)$ or $SSR_f \leq (SSR_f - \sigma)$.
+
+$TP = \Delta{SSR}_{e,m} < μ±σ$
+
+Finally, the TP of each EEZ was calculated when $\Delta{SSR_{e,m}}$ is beyond one standard deviation of the historical distribution i.e., when $SSR_f \geq (SSR_f + \sigma)$ or $SSR_f \leq (SSR_f - \sigma)$.
 
@@
 skewness, kurtosis) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
+Specifically, we tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of transboundary stocks.
+For both cases, our null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
 <!-- Moreover, we implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
@@
 
-Global society has set the ambitious goal to manage all fisheries sustainably with no overfishing (SDG 14 – Live below water) [@UnitedNations2018] in the incoming years, which in turn will have direct repercussion in others [@Singh:2017ds].
-Preparing policy to species on the move is key to achieve these sustainable development goals and proper governance of the world fisheries [@Pinsky:2018cb].
-Here, we have addressed two fundamental steps to achieve sustainable international fisheries under a changing world [@Link:2010ei].
-We have identified the transboundary species that will see a shift in shared distribution, the year in which such distribution will shift, as well as the degree of change.
-Our results have direct implications for ocean governance.
-By identifying the time, degree, and location by which transboundary species will change from a historical normal we provide critical information for regional management bodies and countries to anticipate the potential complications that climate change will bring to international fisheries management.
-While our results are not free of uncertainty and would benefit from more detailed analysis at local levels, they represent a crucial approximation towards preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg, Chapter One].
+Global society has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) [@UnitedNations2018] by 2030; achieving this goal would have clear effects on multiple  others [@Singh:2017ds].
+Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@Pinsky:2018cb].
+Here, we have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
+We have identified the transboundary species that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
+Our results have direct implications for ocean governance as the data provided  can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by  climate change to international fisheries management.
+While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our  results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg; @PalaciosAbrantes:2020tv].
 To address the issues raised in this paper, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.

### 2020-11-13 Scripts/Manuscript.Rmd
Introduction and methods comments addressed

@@
 csl: ../References/science.csl
+header-includes:
+- \usepackage{amsmath}
 ---
@@
 This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (see Cheung *et al.* [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
+Recent work has demonstrated that, at the global scale, DBEM is able to reproduce catch changes similar in magnitude to those detected in observed catch time series in the
+world’s EEZs (Cheung et al. in revision).
 
@@
 
-$lat_{ts} = \mu (lat_{per})$
+<!-- $lat_{ts} = \mu (lat_{per})$ -->
 
-$lon_{ts} = \mu (lon_{per})$
+<!-- $lon_{ts} = \mu (lon_{per})$ -->
+
+\begin{equation}
+\begin{split}
+\label{eq:one}
+lat_{ts} = \mu (lat_{per}) \\
+lon_{ts} = \mu (lon_{per})
+\end{split}
+\end{equation}
 
@@
 
-$$D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))$$
+<!-- $$D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))$$ -->
+
+\begin{equation}
+\label{eq:two}
+
+D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
+
+\end{equation}
+
 where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively and $lon_{eez}$ and $lon_{ts}$ are the respective longitudes.
@@
 
-$$TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2$$
+<!-- $$TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2$$ -->
+
+\begin{equation}
+\label{eq:three}
+TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
+\end{equation}
 
@@
 In contrast, a smaller TI suggests that the sharing of the species between neighbouring EEZs is becoming relatively more stable; a condition that favours effective management and sustainability of the shared species’ fisheries.
-The index was smoothed to a 10 year average to xxx.
+The index was smoothed to a 10 year average to reduce interanual variabillity.
 
@@
 
-$EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$
-$ToE = t when Emergence index[t] > 1$
+<!-- $EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$ -->
+<!-- $ToE = t when Emergence index[t] > 1$ -->
+
+\begin{equation}
+\begin{split}
+\label{eq:four}
+EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}} \\
+ToE = t when Emergence index[t] > 1
+\end{split}
+\end{equation}
 
@@
 We set a threshold of TI above which we consider a shared stock to have emerged from the historical internal variability of the TI.
-The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., eq.
-**x**, representing a probability of 68% that the transboundary index has emerged from historical variability).
+The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \ref{eq:four}, representing a probability of 68% that the transboundary index has emerged from historical variability).
 We tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
@@
 
-$$ index' = (P_c + P_v)/ToE $$
+<!-- $$ index' = (P_c + P_v)/ToE $$ -->
+
+\begin{equation}
+\label{eq:five}
+index' = (P_c + P_v)/ToE
+\end{equation}
 
@@
 We then selected two future periods; early $21^{st}$ century as the average of 2021 to 2040 (*te*), and mid $21^{st}$ century as the average of 2041 to 2060 (*tm*).
-We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing) and SDG 1.2 (poverty reduction), to be achieved by 2020 and 2030, respectively [@Singh:2017ds].
+We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
 The analysis was replicated for projected species distributions from each of the ten ensemble members and results were averaged across ensemble members.
@@
 
-$$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$
+<!-- $$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$ -->
+<!-- $TP = \Delta{SSR}_{e,m} < \mu \pm \sigma$ -->
 
-$TP = \Delta{SSR}_{e,m} < μ±σ$
+\begin{equation}
+\begin{split}
+\label{eq:six}
+\Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 \\
+TP = \Delta{SSR}_{e,m} < \mu \pm \sigma
+\end{split}
+\end{equation}
 
@@
 -->
-All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@gmt] and *zoo* [@PackagezooSInfr:2019tm], *readxl* [@REF], *ggrepel* [@REF], *zeallot* [@REF], *viridis* [@REF], *cowplot* [@REF], *moments* [@REF] and *pgirmess* [@REF].
+All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@PackagegmtInterfa:2017vh] and *zoo* [@PackagezooSInfr:2019tm], *ggrepel* [@PackageggrepelAut:2020to], *zeallot* [@PackagezeallotMul:2018wy], *viridis* [@PackageviridisDef:2018us], *cowplot* [@CowplotStreamlined:2019wt], *moments* [@PackagemomensMome:2015uh] and *pgirmess* [@PackagepgirmessSp:2018um].
 All code is available at https://github.com/jepa/TransEmergence
@@
 However, despite a general recognition that climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn], the Eastern Tropical Pacific will see an equatorial shift with EEZs from Mexico to Ecuador losing SSR to their southern neighbors.
-This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario (Clark *et al.,* n.d.).
+This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
 Moreover, regions with presence of multiple frontiers like the Caribbean and the North East Atlantic will see a complex network of changes in SSR with no marked direction (Figure \ref{fig:Figure5}B).

### 2020-11-14 Scripts/Initial_Analysis.Rmd
Addressed comments for first round

--- 2020-09-18 Scripts/Initial_Analysis.Rmd
@@
   clean_names()
+
+# ------------------------------------------- #
+### Estimate number of transboundary stocks ###
+# ------------------------------------------- #
+
+Transboundary_spp %>% 
+  left_join(Neighbours,
+            by = c("eez_name", "eez_neighbour")) %>% 
+  # filter(neighbour_id %in% c(1,105)) %>%
+  # filter(neighbour_id == 470) %>%
+  group_by(neighbour_id) %>% 
+  summarise(n_stock = length(unique(taxon_key))) %>% 
+  # View()
+  group_by() %>% 
+  summarise(t_trans_stocks = sum(n_stock)) # 9132
 
@@
 # Clean names to match data
-Matching_names <- my_path("G","Spatial/SAU", "matching_names.csv", read = T) %>% 
+Matching_names <- my_path("G","Spatial/SAU", "sau_matching_names.csv", read = T) %>% 
   rename(eezid = eez_id)
@@
   mutate(latitude = abs(latitude))
+
+```
+
+###  Year of Eergence Analysis
+
+```{r ToE_generals,  eval = T, echo = T}
+
+# Total number of emerging stocks
+
+# Global estimates
+Eme_Res %>% 
+  group_by(tresh,neighbour_id) %>% 
+  summarise(
+    emerging_stocks = length(unique(taxon_key))
+    ) %>% 
+  group_by(tresh) %>% 
+  summarise(
+    emerging_stocks = sum(emerging_stocks),
+    prop_total = (emerging_stocks/9132)*100,
+    )
+
+# Global average ToE
+Eme_Res %>% 
+  group_by(tresh) %>% 
+  summarise(
+    # emerging_stocks = length(unique(taxon_key)),
+    # prop_total = (emerging_stocks/9132)*100,
+    mean= mean(emerging_yr),
+    sd = sd(emerging_yr),
+    min = min(emerging_yr),
+    max = max(emerging_yr)
+           )
+
+```
+
+##### Fig. 1. Accumulative plots
+
+```{r Time_lapse_data,  eval = T, echo = T}
+
+# Per Stock
+stock_d <-  
+  Eme_Res %>% 
+  left_join(Neighbours) %>% 
+  # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
+  group_by(taxon_key,neighbour_id,tresh) %>% 
+  summarise(
+    taxon_emer = min(emerging_yr)
+  ) %>% 
+  group_by(taxon_emer,tresh) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  group_by(tresh) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "stock"
+  )
+
+max(stock_d$com_sum)
+
+#  EEZ data
+eez_d <-  Eme_Res %>% 
+  left_join(Neighbours) %>% 
+  filter(n_ens == 10) %>% 
+  # distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
+  group_by(sf_eez_name,tresh) %>% 
+  summarise(
+    taxon_emer = min(emerging_yr)
+  ) %>% 
+  group_by(taxon_emer,tresh) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  group_by(tresh) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "eez"
+  )
+
+max(eez_d$com_sum)
+
+# Plot Data
+plot_data <- stock_d %>% 
+  bind_rows(eez_d) %>% 
+  filter(tresh != "NA") %>% 
+  mutate(
+    # total = ifelse(level == "stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$sf_eez_name))),
+    total = ifelse(level == "stock",max(stock_d$com_sum),max(eez_d$com_sum)),
+    per = com_sum/total*100
+  ) %>% 
+  gather("metric","value",com_sum,per) %>% 
+  filter(tresh == "Tresh_one")
+
+# Manually include rest of EEZs
+
+cont_eez <- plot_data %>% 
+  filter(level == "eez") %>% 
+  slice(1:36) %>%
+  select(-1) %>% 
+  mutate(
+    value = ifelse(metric == "per",100,232),
+    taxon_emer = c(seq(2082,2099,1),seq(2082,2099,1))
+  )
+
+plot_data <- plot_data %>% 
+  bind_rows(cont_eez) %>% 
+  mutate(
+    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
+    label = ifelse(total == max(stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs")
+  )
+
+# How many between 2006 and 2020?
+plot_data %>% 
+  filter(taxon_emer <= 2020,
+         tresh == "Tresh_one") %>% 
+  group_by(level,metric) %>% 
+  summarise(
+    max(taxon_emer),
+    min(taxon_emer),
+    max(taxon_emer),
+    min(taxon_emer)
+  )
+
+# Percentage increase of EEZ after 2020
+lm_eez <- plot_data %>% 
+  filter(taxon_emer <= 2020,
+         tresh == "Tresh_one",
+         metric == "per")
+
+summary(lm(lm_eez$taxon_emer~lm_eez$value))
+
+```
+
+```{r Time_lapse_plot}
+#### Both plots for tresh one
+
+ggplot() +
+  geom_area(data = subset(plot_data, level == "eez"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            alpha= 0.5,
+            size=0.5,
+            colour="black"
+  )  +
+  geom_area(data = subset(plot_data, level == "stock"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            alpha= 0.5,
+            size=0.5,
+            colour="black"
+  )  +
+  # geom_line(data = subset(plot_data, metric == "com_sum"),
+  #           aes(
+  #             x = as.numeric(taxon_emer),
+  #             y = total,
+  #             fill = level
+  #           ),
+  #           colour="grey30",
+  # )  +
+  # geom_text(data = subset(plot_data, metric == "com_sum"),
+  #           aes(
+  #             x = 2037,
+  #             y = total+5,
+  #             label = label
+  #           ),
+  #           size = 3,
+  #           colour="grey30",
+  #           hjust=0.6,
+  # )  +
+  facet_wrap(~my_facet,
+             scales = "free") + 
+  scale_y_continuous("",expand = expand_scale(mult = c(0, .05))) +
+  scale_x_continuous("Time of emergence",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     expand = c(0, 0)
+  ) +
+  my_ggtheme_p() +
+  theme(
+    legend.position = "right"
+  ) +
+  guides(fill=guide_legend(title= "Level")) +
+  scale_fill_viridis(discrete = T,
+                     labels = c("EEZs","Stocks")
+  ) +
+  ggsave("./Figures/cumu_per_freq.png",
+  width = 12,
+  height = 7,
+  units = "in"
+  )
+
+```
+
+##### Fig. 2 Mapping the time of emergence
+
+###### Mean ToE and vulnerabillity Index
+
+```{r emergence_map, eval = T}
+
+# Per EEZs
+Eez_d_hist <- Eme_Res %>% 
+  group_by(eezid,sau_fishing_entity,tresh) %>% 
+  # summarise(n_tax = length(unique(taxon_key))) %>% 
+  summarise(
+    mean_yr = mean(emerging_yr,na.rm=T),
+    sd_yr = sd(emerging_yr,na.rm=T),
+    n_taxa = length(unique(taxon_key))
+    ) %>% 
+  gather("metric","value",mean_yr:n_taxa) %>% 
+  # filter(metric != "n_taxa") %>% 
+  mutate(
+    reverse = abs(value-2100)  + 2000
+  )
+
+# Estimate the value (catch and USD) of emerging species
+Eme_value_data <- Eme_Res %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,tresh,taxon_key) %>% 
+  summarise(n()) %>% 
+  left_join(Sau_data,
+            by = c("fishing_entity_id","taxon_key")) %>% 
+  filter(!is.na(variable)) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,tresh,variable) %>% 
+  summarise(
+   sum_value = sum(value,na.rm=T),
+    n_taxa = length(unique(taxon_key)) 
+  )
+
+# Total value of each fishing nation
+
+Total_value <- Sau_data %>% 
+  group_by(fishing_entity_id,variable) %>% 
+  summarise(tot_value= sum(value,na.rm=T))
+
+Value_map_data <- Eme_value_data %>% 
+  left_join(Total_value,
+            by = c("fishing_entity_id","variable")) %>% 
+  mutate(participation = (sum_value/tot_value)*100)
+
+### ----------------- ###
+### Index alternative ###
+### ----------------- ###
+
+Index_data <- Eez_d_hist %>% 
+  left_join(Value_map_data) %>% 
+  filter(
+    tresh == "Tresh_one",
+    metric == "mean_yr"
+  ) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,variable) %>% 
+  summarise(
+    toe = mean(value,na.rm=T),
+    # sum_value = sum(sum_value),
+    mean_part = mean(participation,na.rm=T)
+  ) %>% 
+  filter(!is.na(variable)) %>% 
+  # spread(variable,sum_value) %>% 
+  spread(variable,mean_part) %>% 
+  mutate(
+    imp = catch+value,
+    index = imp/toe
+  )
+
+##### Normalize index
+# https://www.datanovia.com/en/blog/how-to-normalize-and-standardize-data-in-r-for-great-heatmap-visualization/
+
+# Get min and maximums
+min_index <- min(Index_data$index)
+max_index <- max(Index_data$index)
+
+Index_data <- Index_data %>% 
+  mutate(
+    index_norm =(index-min_index)/(max_index-min_index),
+    index_2000 = 2000 + (index-min_index)*(2100-2000)/(max_index-min_index)
+  )
+
+### Explore index ####
+
+# ggplot(data = Index_data) +
+#   geom_point(
+#     aes(
+#       x = toe,
+#       y = imp,
+#       color = index_2000
+#     ),
+#     size = 4
+#   ) +
+#    scale_color_viridis(option="magma", discrete = F)
+
+#### ------------  ####
+
+Value_map_data <- Index_data %>%
+  select(sau_fishing_entity,fishing_entity_id,
+         # tresh,
+         # variable,
+         participation = index_2000)
+
+### ----------------- ###
+
+# Per fishing entity
+
+fish_d_hist <- Eme_Res %>% 
+  group_by(sau_fishing_entity,tresh) %>% 
+  # summarise(n_tax = length(unique(taxon_key))) %>% 
+  summarise(
+    value = length(unique(taxon_key))
+    ) %>% 
+  filter(!is.na(sau_fishing_entity))
+
+World_map_nemer <- World_map %>%
+  full_join(Value_map_data,
+            by = c("sau_fishing_entity","fishing_entity_id")
+  ) %>% 
+  filter(!is.na(participation))
+
+# These are regions without transboundary species
+grey_land <- World_map %>% 
+  filter(!fishing_entity_id %in% Value_map_data$fishing_entity_id) %>% 
+  filter(!sovereignt %in% c("Antarctica","Brazil"))
+
+grey_eez <- World_eez_sf %>% 
+  filter(!eezid %in% Eez_d_hist$eezid,
+         !eez_name %in% Neighbours$eez_name)
+
+Brazil_land <- World_map %>% 
+  filter(sovereignt == "Brazil")
+
+Brazil_eez <- World_eez_sf %>% 
+  filter(eez_name %in% c("Brazil","USA (Alaska, Arctic)"))
+
+for(i in 1:1){
+
+  tresh_list <- c("Tresh_one","Tresh_two")
+  name_save <- paste("./Figures/emer_map_lat_",tresh_list[i],".png",sep = "")
+
+  ## --------------- ##
+  # Latitudinal plot
+  ## --------------- ##
+
+  ### Random plot
+
+Eez_hist <- Eme_Res %>% 
+  group_by(eezid,eez_name = sf_eez_name) %>% 
+  left_join(Eez_centroid) %>% 
+  filter(tresh == tresh_list[i]) %>% 
+  group_by(latitude) %>% 
+  summarise(
+    number = n()
+  ) %>% 
+  group_by() %>% 
+  mutate(RMean = rollmean(x = number, 
+                          10, 
+                          align = "center", 
+                          fill = number
+  )
+  ) %>% 
+  mutate(lab="lab") %>% 
+  ggplot() +
+  geom_area(
+    aes(
+      x = latitude,
+      y = RMean,
+      fill = lab,
+      color = lab
+    )
+  ) +
+  coord_flip() +
+  scale_x_continuous("",
+    limits = c(-50,80),
+    breaks = c(-50,0,50,80)
+  ) +
+  scale_y_continuous(
+    "Emerging stocks",
+    limits = c(0, 124),
+    breaks = c(0,62,124)
+
+  ) +
+  scale_fill_viridis(option="magma", discrete = T, begin = 0.7) +
+  scale_color_viridis(option="magma", discrete = T, begin = 0.7) +
+  theme_minimal() + 
+  theme(
+    panel.background = element_rect(fill = "white", colour = "white"),
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.text = element_text(size = 12, colour="black"),
+    axis.title.x = element_text(size = 12, colour="black", face = "bold"),
+    legend.position = ""
+  )
+
+  ## --------------- ##
+  # Map
+  ## --------------- ##
+
+  World_map_part <- World_map_nemer# %>% 
+    # filter(
+    #   tresh ==  tresh_list[i], 
+    #   variable == "value"
+    # )
+
+  map <- World_eez_sf %>% 
+    left_join(Eez_d_hist,
+              by = "eezid") %>% 
+    filter(tresh == tresh_list[i],
+           metric == "mean_yr") %>% 
+    ggplot() +
+    geom_sf(
+      aes(
+        fill = reverse #original map
+        # fill = value
+      ),
+      color =  "white",
+      size = 0.1
+    ) +
+    # Inlcude the participation of species
+    geom_sf(data = World_map_part,
+            # aes(fill = participation+2000), #original map
+            aes(fill = participation),
+            color =  "white",
+            size = 0.1
+            ) +
+    # grey maps
+    # ggplot() +
+    geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
+    geom_sf(data = grey_eez, aes(), fill = "white", color =  "grey50",size = 0.1) +
+    geom_sf(data = Brazil_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    geom_sf(data = Brazil_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    # Set Participation labels
+    annotate("text",
+             x =-135,
+             y = -69,
+             label = "Vulnerabillity index",
+             color="black",
+             size=4,
+             angle=0,
+             fontface="bold",
+             hjust=0) +
+    annotate("text", 
+             x = c(-76,-35,6,48,88),
+             y = -69,
+             # x =(c(-8500000,-3000000,2500000,7500000,12800000)),
+             # y = -9050000,
+             label = c(0,25,50,75,100), 
+             color="black", 
+             size=4, 
+             angle=0, 
+             # fontface="bold",
+             hjust=0) +
+    # Set ToE labels
+    annotate("text", 
+             x =-135,
+             y = -85,
+             label = "Time of Emergence",
+             color="black", 
+             size=4, 
+             angle=0, 
+             fontface="bold",
+             hjust=0) +
+    annotate("text", 
+             x = c(-79,-39,3,45,87),
+             y = -85,
+             label = seq(2100,2000,-25), # original
+             # label = c(">2100",seq(2075,2000,-25)), # not inluding the > because we dont know
+             color="black", 
+             size=4, 
+             angle=0, 
+             # fontface="bold",
+             hjust=0) +
+    scale_fill_viridis("",
+      limits = c(2000,2100),
+      breaks = seq(2000,2100,25),
+      labels = c("","","","","")#,
+      # direction = -1
+    ) +
+    # scale_fill_gradientn("",
+    #                      limits = c(2000,2100),
+    #                      breaks = seq(2000,2100,25),
+    #                       labels = c("","","","",""),
+    #                      colours = pal
+    # ) +
+    # geom_hline(yintercept = c(-50,0,50,80)) +
+    my_ggtheme_m() +
+    theme(
+      legend.box.margin = margin(t = -64, l = 229),
+      legend.text = element_blank(),
+      legend.key.width = unit(5, "line")
+    )
+
+  ## --------------- ##
+  # Regional Boxplot
+  ## --------------- ##
+
+  boxplot <- Eme_Res %>% 
+    filter(tresh == "Tresh_one",
+           !is.na(un_sub_region)) %>%
+    group_by(un_sub_region) %>% 
+    summarise(order = median(emerging_yr)) %>% 
+    left_join(Eme_Res) %>% 
+    filter(tresh == "Tresh_one",
+           !is.na(un_sub_region)) %>% 
+    mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and the Car.",paste(un_sub_region))) %>% 
+    ggplot() +
+    geom_boxplot(
+      aes(
+        # x = un_sub_region,
+        x = reorder(un_sub_region,order),
+        y = emerging_yr#,
+        # fill = un_geo_region
+      ),
+      fill = "grey"
+    ) +
+    facet_wrap(~un_geo_region,
+               scales = "free_x",
+               nrow = 1) +
+    scale_fill_viridis("UN Sub region",
+                       discrete = T) +
+    scale_color_grey("") +
+    xlab("") +
+    ylab("Time of Emergence") +
+    my_ggtheme_p() +
+    theme(
+      axis.text.x = element_text(size = 10, angle = 45, face = "plain", hjust = 1), 
+      strip.text.x = element_text(size = 14, colour = "black"), 
+      # axis.ticks.x = element_blank(),
+      legend.position = "right"
+    )
+
+  ## --------------- ##
+  # Combine plots and save them
+  ## --------------- ##
+
+  p <- ggdraw() +
+    # ToE Latitudinal plot
+    draw_plot(Eez_hist, x = 0.873, y = 0.53, width = 0.124, height = 0.37) +
+    # Time of Emergence Map
+    draw_plot(map, x = 0, y = 0.4, width = 0.95, height = 0.6) +
+    # Regional boxplot
+    draw_plot(boxplot, x = 0, y = -0.03, width = 1, height = 0.44) +
+    # Labels
+    draw_plot_label(label = c("A", "B"), 
+                      size = 20,
+                      x = c(0, 0), 
+                      y = c(0.95, 0.46)
+      )
+
+  save_plot(
+    name_save,
+    p,
+    base_height = 11,
+    base_width = 12
+  )
+
+}
+
+```
+
+##### Fig. 4. Bubble plot
+
+```{r, buble_plot, eval = F, echo = F}
+
+# Needs previous chunk
+
+Buble_data <- Eez_d_hist %>% 
+  filter(metric == "mean_yr") %>% 
+  group_by(sau_fishing_entity,tresh) %>% 
+  summarise(mean_yr = mean(value)) %>% 
+  left_join(Eme_value_data,
+            by = c("sau_fishing_entity","tresh")) %>% 
+  filter(tresh == "Tresh_one",
+         variable == "value") %>% 
+  mutate(
+    weight_toe = sum_value/mean_yr,
+    sum_value_bi = sum_value/1000000000,
+    log_value = log10(sum_value_bi)
+    ) %>%   # Billion usd
+  left_join(Matching_names,
+            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
+  distinct(sau_fishing_entity,.keep_all = TRUE) %>% 
+  filter(!is.na(un_geo_region)) %>% 
+  ungroup() %>% 
+  mutate(
+    sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity))
+
+### Fucntion for estimating percentile
+
+quibble <- function(x, q = p) {
+        tibble(x = quantile(x, q), q = q)
+}
+
+Percentile_value <- Buble_data %>% 
+  group_by() %>%  # for whole distribution
+        do(quibble(.$log_value, 0.75))
+
+## Top fishing nations
+Buble_data %>% 
+  filter(log_value > Percentile_value$x) %>% 
+  View()
+
+# Labels Data
+
+Labels_data <- tibble(
+  x = c(2020,2040,2007),
+  y = c(-5.8,-5.8,Percentile_value$x+.1),
+  label = c("Early emergence", "Late emergence","Fisheries > 75th percentile")
+)
+
+# Plot the bubbles!
+ggplot(Buble_data) +
+  geom_point(
+    aes(
+      x = mean_yr,
+      y = log_value,
+      size = n_taxa,
+      # colour = un_geo_region,
+      shape = un_geo_region
+    ),
+    alpha = 0.7
+  ) +
+  # Divisory lines
+  geom_vline(xintercept = 2030, color = "grey90", size = 0.5) +
+  geom_hline(data = Percentile_value,
+             aes(yintercept = x),
+             color = "grey90"
+  )  +
+  geom_label_repel(data = subset(Buble_data, sum_value_bi >= 1),
+                   aes(
+                     x = mean_yr,
+                     y = log_value,
+                     label = sau_fishing_entity
+                   ),
+                   arrow = arrow(length = unit(0.009, "npc"),
+                                 type = "closed",
+                                 ends = "first"),
+                   force = 2,
+                   box.padding = 1,
+                   fill = "NA"
+
+  ) +
+  geom_text(data = Labels_data,
+            aes(
+              x = x,
+              y = y,
+              label = label
+            ),
+            size = 3,
+            color = "grey30"
+  ) +
+  scale_color_viridis(name="Region",
+                      discrete = T) +
+  scale_x_continuous("Average time of emergence",
+                     limits = c(2000,2080),
+                     breaks = seq(2000,2080,10)
+  ) +
+  scale_y_continuous("Fishing revenue (log10)",
+                     limits = c(-6,2),
+                     breaks = seq(-6,2,2),
+                     expand = c(0,0)
+  ) +
+  scale_size("Number of\nemerging taxa") +
+  ggsave("./Figures/bubble_plot.png",
+         width = 8,
+         height = 8,
+         units = "in"
+  )
+
+```
+
+### Species level analysis
+
+#### Fig. 5 Species Boxplot
+
+```{r deme_pela, eval = F, echo =F}
+
+# Dem Pelagic boxplot
+Eme_Res %>% 
+  left_join(Spp_env_pref) %>% 
+  mutate(DemersPelag = ifelse(is.na(DemersPelag),"other",DemersPelag)) %>% 
+  group_by(DemersPelag) %>% 
+  summarise(mean_toe = median(emerging_yr,na.rm=T),
+            n_spp = length(unique(taxon_key))) %>% 
+  left_join(Dem_pel) %>% 
+  mutate(
+    label = paste0(DemersPelag, "\n(n=",n_spp,")")
+  ) %>% 
+  filter(tresh == "Tresh_one",
+         DemersPelag != "pelagic") %>% # just one spp.
+  ggplot() +
+  geom_boxplot(
+    aes(x = reorder(label,mean_toe),
+        # aes(x = DemersPelag,
+        y = emerging_yr,
+        # fill = DemersPelag
+    ),
+    alpha = 0.8,
+    fill = "grey"
+  ) +
+  # scale_fill_viridis("Habitat preference",
+                     # discrete = T) +
+  xlab("") +
+  ylab("Time of emergence") +
+  my_ggtheme_p() +
+  theme(
+    axis.text.x = element_text(size = 10, angle = 45, face = "plain", hjust = 1), 
+    # axis.ticks.x = element_blank(),
+    legend.position = "right"
+  ) +
+  scale_x_reordered() +
+  ggsave("./Figures/regional_hab_boxplot.png",
+         width = 10,
+         height = 5,
+         units = "in"
+  )
+
+```
+
+### Statistics
+
+#### Test for normallity
+
+```{r, normality_test, eval = F, echo = F}
+# Test for normality
+
+Stats_data <- Eme_Res %>% 
+  filter(tresh == "Tresh_one",
+         n_ens == 10)
+
+# ------------------- #
+# Test normal asumptions 
+# ------------------- #
+
+hist(Stats_data$emerging_yr) # Positevely skewed
+qqnorm(Stats_data$emerging_yr) # not normally distributed
+skewness(Stats_data$emerging_yr) #1.02
+kurtosis(Stats_data$emerging_yr) #2.82
+
+# ------------------- #
+
+# (NOT NORMAL DISTRIRBUTION)
+
+```
+
+## Regional differences
+
+```{r multiple_reg, eval = F, echo = F}
+
+Reg_diff_data <- Eme_Res %>% 
+  filter(tresh == "Tresh_one",
+         !is.na(un_sub_region)
+         )
+
+# ------------------- #
+# Run non-parametric Krustal
+# ------------------- #
+
+Reg_kw <- kruskal.test(un_sub_region~emerging_yr, data = Reg_diff_data)
+
+Reg_kw
+# Kruskal-Wallis rank sum test
+# 
+# data:  un_sub_region by emerging_yr
+# Kruskal-Wallis chi-squared = 287.23, df = 93, p-value < 2.2e-16
+
+kruskalmc(emerging_yr~un_sub_region, data = Reg_diff_data)
+
+# Multiple differences but Eastern Asia, Lat. Ame. Car. and Polinesia are the nes who differe the most (e.g. differe with more revions).
+
+# ------------------- #
+# Just to double check ANOVA results (give the same)
+# ------------------- #
+Reg_aov <- aov(emerging_yr~un_sub_region, data = Reg_diff_data)
+Reg_aov
+
+summary(Reg_aov)
+
+TukeyHSD(Reg_aov) %>% 
+  tidy() %>% 
+   filter(adj.p.value < 0.05) %>%
+  View()
+
+```
+
+## Habitat preference differences
+
+```{r multiple_reg, eval = F, echo = F}
+
+Hab_diff_data <- Eme_Res %>% 
+  filter(tresh == "Tresh_one",
+         !is.na(un_sub_region)
+         ) %>% 
+  left_join(Spp_env_pref,
+            by = "taxon_key")
+
+# ------------------- #
+# Run non-parametric Krustal
+# ------------------- #
+
+Hab_kw <- kruskal.test(DemersPelag~emerging_yr, data = Hab_diff_data)
+
+Hab_kw
+# Kruskal-Wallis rank sum test
+
+# data:  DemersPelag by emerging_yr
+# Kruskal-Wallis chi-squared = 286.48, df = 93, p-value < 2.2e-16
+
+kruskalmc(emerging_yr~DemersPelag, data = Hab_diff_data)
+
+# Multiple differences but main differences for pelagic-ocenaic and reefs
+
+# ------------------- #
+# Just to double check ANOVA results (give the same)
+# ------------------- #
+
+Hab_aov <- aov(emerging_yr~DemersPelag, data = hab_diff_data)
+Hab_aov
+
+summary(Hab_aov)
+
+TukeyHSD(Hab_aov) %>% 
+  tidy() %>% 
+   filter(adj.p.value < 0.05) %>%
+  View()
 
@@
 
-```
-
-####  Year of Eergence Analysis
-
-##### Fig. 2. Accumulative plots
-
-```{r Time_lapse_plot,  eval = T, echo = T}
-
-# Per Stock
-stock_da <-  Eme_Res %>% 
-  left_join(Neighbours) %>% 
-  # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
-  group_by(taxon_key,tresh) %>% 
-  summarise(
-    taxon_emer = min(emerging_yr)
-  ) %>% 
-  group_by(taxon_emer,tresh) %>% 
-  summarise(
-    n_tax = n()
-  ) %>% 
-  group_by(tresh) %>% 
-  mutate(
-    com_sum = cumsum(n_tax),
-    level = "stock"
-  )
-
-#  EEZ data
-eez_d <-  Eme_Res %>% 
-  left_join(Neighbours) %>% 
-  filter(n_ens == 10) %>% 
-  # distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
-  group_by(sf_eez_name,tresh) %>% 
-  summarise(
-    taxon_emer = min(emerging_yr)
-  ) %>% 
-  group_by(taxon_emer,tresh) %>% 
-  summarise(
-    n_tax = n()
-  ) %>% 
-  group_by(tresh) %>% 
-  mutate(
-    com_sum = cumsum(n_tax),
-    level = "eez"
-  )
-
-# Plot Data
-plot_data <- stock_d %>% 
-  bind_rows(eez_d) %>% 
-  filter(tresh != "NA") %>% 
-  mutate(
-    total = ifelse(level == "stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$sf_eez_name))),
-    per = com_sum/total*100
-  ) %>% 
-  gather("metric","value",com_sum,per) %>% 
-  filter(tresh == "Tresh_one") %>% 
-  mutate(
-    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum"),
-    label = ifelse(total == 478,"Total number of stocks","Total number of EEZs")
-  );plot_data
-
-#### Both plots for tresh one
-
-ggplot() +
-  geom_area(data = subset(plot_data, level == "stock"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = value,
-              fill = level
-            ),
-            alpha= 0.8,
-            size=0.5,
-            colour="black"
-  )  +
-  geom_area(data = subset(plot_data, level == "eez"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = value,
-              fill = level
-            ),
-            alpha= 0.8,
-            size=0.5,
-            colour="black"
-  )  +
-  geom_line(data = subset(plot_data, metric == "com_sum"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = total,
-              fill = level
-            ),
-            colour="grey30",
-  )  +
-  geom_text(data = subset(plot_data, metric == "com_sum"),
-            aes(
-              x = 2037,
-              y = total+5,
-              label = label
-            ),
-            size = 3,
-            colour="grey30",
-            hjust=0.6,
-  )  +
-  facet_wrap(~my_facet,
-             scales = "free") + 
-  scale_y_continuous("",expand = expand_scale(mult = c(0, .05))) +
-  scale_x_continuous("Time of emergence",
-                     breaks = seq(2000,2100,25),
-                     limits = c(2005,2100),
-                     expand = c(0, 0)
-  ) +
-  my_ggtheme_p() +
-  theme(
-    legend.position = "right"
-  ) +
-  guides(fill=guide_legend(title= "Level")) +
-  scale_fill_viridis(discrete = T,
-                     labels = c("EEZs","Stocks")
-  ) +
-  ggsave("./Figures/cumu_per_freq.png",
-         width = 7,
-         height = 7,
-         units = "in"
-  )
-
-```
-
-##### Fig. 3 Mapping the time of emergence
-
-###### Just mean time
-
-```{r emergence_map, eval = T}
-
-# Per EEZs
-Eez_d_hist <- Eme_Res %>% 
-  group_by(eezid ,sau_fishing_entity,tresh) %>% 
-  # summarise(n_tax = length(unique(taxon_key))) %>% 
-  summarise(
-    mean_yr = mean(emerging_yr,na.rm=T),
-    sd_yr = sd(emerging_yr,na.rm=T),
-    n_taxa = length(unique(taxon_key))
-    ) %>% 
-  gather("metric","value",mean_yr:n_taxa) %>% 
-  # filter(metric != "n_taxa") %>% 
-  mutate(
-    reverse = abs(value-2100)  + 2000
-  )
-
-# Estimate the value (catch and USD) of emerging species
-Eme_value_data <- Eme_Res %>% 
-  group_by(sau_fishing_entity,fishing_entity_id,tresh,taxon_key) %>% 
-  summarise(n()) %>% 
-  left_join(Sau_data,
-            by = c("fishing_entity_id","taxon_key")) %>% 
-  filter(!is.na(variable)) %>% 
-  group_by(sau_fishing_entity,fishing_entity_id,tresh,variable) %>% 
-  summarise(
-   sum_value = sum(value,na.rm=T),
-    n_taxa = length(unique(taxon_key)) 
-  )
-
-# Total value of each fishing nation
-
-Total_value <- Sau_data %>% 
-  group_by(fishing_entity_id,variable) %>% 
-  summarise(tot_value= sum(value,na.rm=T))
-
-Value_map_data <- Eme_value_data %>% 
-  left_join(Total_value,
-            by = c("fishing_entity_id","variable")) %>% 
-  mutate(participation = (sum_value/tot_value)*100)
-
-# Per fishing entity
-
-fish_d_hist <- Eme_Res %>% 
-  group_by(sau_fishing_entity,tresh) %>% 
-  # summarise(n_tax = length(unique(taxon_key))) %>% 
-  summarise(
-    value = length(unique(taxon_key))
-    ) %>% 
-  filter(!is.na(sau_fishing_entity))
-
-World_map_nemer <- World_map %>%
-  left_join(Value_map_data,
-            by = c("sau_fishing_entity","fishing_entity_id")
-  ) %>% 
-  filter(!is.na(participation))
-
-grey_land <- World_map %>% 
-  filter(!fishing_entity_id %in% World_map_nemer$fishing_entity_id) %>% 
-  filter(sovereignt != "Antarctica")
-
-grey_eez <- World_eez_sf %>% 
-  filter(!eezid %in% Eez_d_hist$eezid)
-
-for(i in 1:2){
-
-  tresh_list <- c("Tresh_one","Tresh_two")
-  name_save <- paste("./Figures/emer_map_lat_",tresh_list[i],".png",sep = "")
-
-  ## --------------- ##
-  # Latitudinal plot
-  ## --------------- ##
-
-  ### Random plot
-
-Eez_hist <- Eme_Res %>% 
-  group_by(eezid,eez_name = sf_eez_name) %>% 
-  left_join(Eez_centroid) %>% 
-  filter(tresh == tresh_list[i]) %>% 
-  group_by(latitude) %>% 
-  summarise(
-    number = n()
-  ) %>% 
-  group_by() %>% 
-  mutate(RMean = rollmean(x = number, 
-                          10, 
-                          align = "center", 
-                          fill = number
-  )
-  ) %>% 
-  mutate(lab="lab") %>% 
-  ggplot() +
-  geom_area(
-    aes(
-      x = latitude,
-      y = RMean,
-      fill = lab,
-      color = lab
-    )
-  ) +
-  coord_flip() +
-  scale_x_continuous("",
-    limits = c(-50,80),
-    breaks = c(-50,0,50,80)
-  ) +
-  scale_y_continuous(
-    "Emerging stocks",
-    limits = c(0, 124),
-    breaks = c(0,62,124)
-
-  ) +
-  scale_fill_viridis(option="magma", discrete = T, begin = 0.7) +
-  scale_color_viridis(option="magma", discrete = T, begin = 0.7) +
-  theme_minimal() + 
-  theme(
-    panel.background = element_rect(fill = "white", colour = "white"),
-    panel.grid.major.y = element_blank(),
-    panel.grid.minor.y = element_blank(),
-    panel.grid.major.x = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.text = element_text(size = 12, colour="black"),
-    axis.title.x = element_text(size = 12, colour="black", face = "bold"),
-    legend.position = ""
-  );Eez_hist
-
-  ## --------------- ##
-  # Map
-  ## --------------- ##
-
-  World_map_part <- World_map_nemer %>% 
-    filter(
-      tresh ==  tresh_list[i], 
-      variable == "value"
-    )
-
-  map <- World_eez_sf %>% 
-    left_join(Eez_d_hist,
-              by = "eezid") %>% 
-    filter(tresh == tresh_list[i],
-           metric == "mean_yr") %>% 
-    ggplot() +
-    geom_sf(
-      aes(
-        fill = reverse
-      ),
-      color =  "white",
-      size = 0.1
-    ) +
-    # Inlcude the participation of species
-    geom_sf(data = World_map_part,
-            aes(fill = participation+2000),
-            color =  "white",
-            size = 0.1
-            ) +
-    # grey maps
-    geom_sf(data = grey_land, aes(),fill = "grey50", color =  "white",size = 0.1) +
-    geom_sf(data = grey_eez, aes(),fill = "grey50", color =  "white",size = 0.1) +
-    # Set Participation labels
-    annotate("text", 
-             x =-135,
-             y = -69,
-             label = "Participation (%)", 
-             color="black", 
-             size=4, 
-             angle=0, 
-             fontface="bold",
-             hjust=0) +
-    annotate("text", 
-             x = c(-76,-35,6,48,88),
-             y = -69,
-             # x =(c(-8500000,-3000000,2500000,7500000,12800000)),
-             # y = -9050000,
-             label = c(0,25,50,75,100), 
-             color="black", 
-             size=4, 
-             angle=0, 
-             # fontface="bold",
-             hjust=0) +
-    # Set ToE labels
-    annotate("text", 
-             x =-135,
-             y = -85,
-             label = "Time of Emergence",
-             color="black", 
-             size=4, 
-             angle=0, 
-             fontface="bold",
-             hjust=0) +
-    annotate("text", 
-             x = c(-79,-39,3,45,87),
-             y = -85,
-             label = seq(2100,2000,-25), 
-             color="black", 
-             size=4, 
-             angle=0, 
-             # fontface="bold",
-             hjust=0) +
-    scale_fill_viridis("",
-      limits = c(2000,2100),
-      breaks = seq(2000,2100,25),
-      labels = c("","","","","")#,
-      # direction = -1
-    ) +
-    # geom_hline(yintercept = c(-50,0,50,80)) +
-    my_ggtheme_m() +
-    theme(
-      legend.box.margin = margin(t = -64, l = 229),
-      legend.text = element_blank(),
-      legend.key.width = unit(5, "line")
-    )
-
-  ## --------------- ##
-  # Combine plots and save them
-  ## --------------- ##
-
-  p <- ggdraw() +
-    # ToE Latitudinal plot
-    draw_plot(Eez_hist, x = 0.873, y = 0.13, width = 0.124, height = 0.77) +
-    # Time of Emergence Map
-    draw_plot(map, x = 0, y = 0, width = 0.95, height = 1)
-
-  save_plot(
-    name_save,
-    p,
-    base_height = 6,
-    base_width = 12
-  )
-
-}
-
-```
-
-##### Fig. 4. Bubble plot
-
-```{r, buble_plot, eval = F, echo = F}
-
-# Needs previous chunk
-
-Buble_data <- Eez_d_hist %>% 
-  filter(metric == "mean_yr") %>% 
-  group_by(sau_fishing_entity,tresh) %>% 
-  summarise(mean_yr = mean(value)) %>% 
-  left_join(Eme_value_data,
-            by = c("sau_fishing_entity","tresh")) %>% 
-  filter(tresh == "Tresh_one",
-         variable == "value") %>% 
-  mutate(
-    weight_toe = sum_value/mean_yr,
-    sum_value_bi = sum_value/1000000000,
-    log_value = log10(sum_value_bi)
-    ) %>%   # Billion usd
-  left_join(Matching_names,
-            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
-  distinct(sau_fishing_entity,.keep_all = TRUE) %>% 
-  filter(!is.na(un_geo_region)) %>% 
-  ungroup() %>% 
-  mutate(
-    sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity))
-
-### Fucntion for estimating percentile
-
-quibble <- function(x, q = p) {
-        tibble(x = quantile(x, q), q = q)
-}
-
-Percentile_value <- Buble_data %>% 
-  group_by() %>%  # for whole distribution
-        do(quibble(.$log_value, 0.75))
-
-## Top fishing nations
-Buble_data %>% 
-  filter(log_value > Percentile_value$x) %>% 
-  View()
-
-# Labels Data
-
-Labels_data <- tibble(
-  x = c(2020,2040,2007),
-  y = c(-5.8,-5.8,Percentile_value$x+.1),
-  label = c("Early emergence", "Late emergence","Fisheries > 75th percentile")
-)
-
-# Plot the bubbles!
-ggplot(Buble_data) +
-  geom_point(
-    aes(
-      x = mean_yr,
-      y = log_value,
-      size = n_taxa,
-      colour = un_geo_region
-    ),
-    alpha = 0.7
-  ) +
-  # Divisory lines
-  geom_vline(xintercept = 2030, color = "grey90", size = 0.5) +
-  geom_hline(data = Percentile_value,
-             aes(yintercept = x),
-             color = "grey90"
-  )  +
-  geom_label_repel(data = subset(Buble_data, sum_value_bi >= 1),
-                   aes(
-                     x = mean_yr,
-                     y = log_value,
-                     label = sau_fishing_entity
-                   ),
-                   arrow = arrow(length = unit(0.009, "npc"),
-                                 type = "closed",
-                                 ends = "first"),
-                   force = 2,
-                   box.padding = 1,
-                   fill = "NA"
-
-  ) +
-  geom_text(data = Labels_data,
-            aes(
-              x = x,
-              y = y,
-              label = label
-            ),
-            size = 3,
-            color = "grey30"
-  ) +
-  scale_color_viridis(name="Region",
-                      discrete = T) +
-  scale_x_continuous("Average time of emergence",
-                     limits = c(2000,2080),
-                     breaks = seq(2000,2080,10)
-  ) +
-  scale_y_continuous("Fishing revenue (log10)",
-                     limits = c(-6,2),
-                     breaks = seq(-6,2,2),
-                     expand = c(0,0)
-  ) +
-  scale_size("Number of\nemerging taxa") +
-  ggsave("./Figures/bubble_plot.png",
-         width = 8,
-         height = 8,
-         units = "in"
-  )
-
-```
-
-### Species level analysis
-
-#### Fig. 5 Boxplot
-
-```{r deme_pela, eval = F, echo =F}
-
-Dem_pel <- Eme_Res %>% 
-  filter(tresh == "Tresh_one",
-         !is.na(un_sub_region)) %>% 
-  left_join(Spp_env_pref) %>% 
-  mutate(DemersPelag = ifelse(is.na(DemersPelag),"other",
-                              ifelse(DemersPelag %in% c("bathypelagic","benthopelagic"), "bathy-bentho-pelagic",paste(DemersPelag)
-                              )
-  ),
-    un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and the Car.",paste(un_sub_region)))
-
-## Averages
-
-Average_data <- Dem_pel %>% 
-  group_by(un_sub_region) %>% 
-  summarise(mean_toe = mean(emerging_yr,na.rm=T)) %>% 
-  mutate(regional_average="regional average")
-
-# Dem Pelagic boxplot
-ggplot(Dem_pel) +
-  geom_hline(data = Average_data,
-             aes(
-               yintercept = mean_toe,
-               colour = regional_average
-             ),
-             linetype = "dashed"
-             ) +
-  geom_boxplot(
-    aes(
-      x = reorder_within(DemersPelag,emerging_yr,un_sub_region),
-      y = emerging_yr,
-      fill = DemersPelag
-    ),
-    alpha = 0.8
-  ) +
-  facet_wrap(~un_sub_region,scales = "free_x") +
-  scale_fill_viridis("Habitat preference",
-                     discrete = T) +
-  scale_color_grey("") +
-  xlab("") +
-  ylab("Time of emergence") +
-  my_ggtheme_p() +
-  theme(
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    legend.position = "right"
-  ) +
-  scale_x_reordered() +
-  ggsave("./Figures/regional_hab_boxplot.png",
-         width = 14,
-         height = 8,
-         units = "in"
-  )
-```
-
-### Statistics
-
-#### Test for normallity
-
-```{r, normality_test, eval = F, echo = F}
-# Test for normality
-
-Stats_data <- Eme_Res %>% 
-  filter(tresh == "Tresh_one",
-         n_ens == 10)
-
-# ------------------- #
-# Test normal asumptions 
-# ------------------- #
-
-hist(Stats_data$emerging_yr) # Positevely skewed
-qqnorm(Stats_data$emerging_yr) # not normally distributed
-skewness(Stats_data$emerging_yr) #1.02
-kurtosis(Stats_data$emerging_yr) #2.82
-
-# ------------------- #
-
-# (NOT NORMAL DISTRIRBUTION)
-
-```
-
-## Regional differences
-
-```{r multiple_reg, eval = F, echo = F}
-
-Reg_diff_data <- Eme_Res %>% 
-  filter(tresh == "Tresh_one",
-         !is.na(un_sub_region)
-         )
-
-# ------------------- #
-# Run non-parametric Krustal
-# ------------------- #
-
-Reg_kw <- kruskal.test(un_sub_region~emerging_yr, data = Reg_diff_data)
-
-Reg_kw
-# Kruskal-Wallis rank sum test
-# 
-# data:  un_sub_region by emerging_yr
-# Kruskal-Wallis chi-squared = 287.23, df = 93, p-value < 2.2e-16
-
-kruskalmc(emerging_yr~un_sub_region, data = Reg_diff_data)
-
-# Multiple differences but Eastern Asia, Lat. Ame. Car. and Polinesia are the nes who differe the most (e.g. differe with more revions).
-
-# ------------------- #
-# Just to double check ANOVA results (give the same)
-# ------------------- #
-Reg_aov <- aov(emerging_yr~un_sub_region, data = Reg_diff_data)
-Reg_aov
-
-summary(Reg_aov)
-
-TukeyHSD(Reg_aov) %>% 
-  tidy() %>% 
-   filter(adj.p.value < 0.05) %>%
-  View()
-
-```
-
-## Habitat preference differences
-
-```{r multiple_reg, eval = F, echo = F}
-
-Hab_diff_data <- Eme_Res %>% 
-  filter(tresh == "Tresh_one",
-         !is.na(un_sub_region)
-         ) %>% 
-  left_join(Spp_env_pref,
-            by = "taxon_key")
-
-# ------------------- #
-# Run non-parametric Krustal
-# ------------------- #
-
-Hab_kw <- kruskal.test(DemersPelag~emerging_yr, data = Hab_diff_data)
-
-Hab_kw
-# Kruskal-Wallis rank sum test
-
-# data:  DemersPelag by emerging_yr
-# Kruskal-Wallis chi-squared = 286.48, df = 93, p-value < 2.2e-16
-
-kruskalmc(emerging_yr~DemersPelag, data = Hab_diff_data)
-
-# Multiple differences but main differences for pelagic-ocenaic and reefs
-
-# ------------------- #
-# Just to double check ANOVA results (give the same)
-# ------------------- #
-
-Hab_aov <- aov(emerging_yr~DemersPelag, data = hab_diff_data)
-Hab_aov
-
-summary(Hab_aov)
-
-TukeyHSD(Hab_aov) %>% 
-  tidy() %>% 
-   filter(adj.p.value < 0.05) %>%
-  View()
-
-```
-
-## Table.1 Multy linear model
-
-```{r multiple_reg, eval = F, echo = F}
-
-# Possible explanatory variables
-
-## Habitat association (DemersPelag)
-## Dem Vs Pelagic (Dem_pel)
-## SST time of Emer (Toe_sst)
-## Stock Status [FishForVisa] (Catch_trend)
-## Biogeographic location (Biogeo_reg)
-
-# AIC value; model between predictive capacity and complexity (i.e., number of explanatory variables) is given by the model with lowest AIC value
-
-# -------------------------------- #
-#### Build models
-# -------------------------------- #
-
-# Model 1; all variables
-M1 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Main_data)
-
-M1_sum <- summary(M1)
-M1_aic <- AIC(M1)
-M1_r <-  M1_sum$adj.r.squared
-
-# Model 2; non demPel
-M2 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel, data = Main_data)
-
-M2_sum <- summary(M2)
-M2_aic <- AIC(M2)
-M2_r <-  M2_sum$adj.r.squared
-
-# Model 3; non biome
-M3 <- lm(emerging_yr ~ sst_toe + bot_toe + status + dem_pel + DemersPelag, data = Main_data)
-
-M3_sum <- summary(M3)
-M3_aic <- AIC(M3)
-M3_r <-  M3_sum$adj.r.squared
-
-# Model 4; non status
-M4 <- lm(emerging_yr ~ sst_toe + bot_toe + biome + dem_pel + DemersPelag, data = Main_data)
-
-M4_sum <- summary(M4)
-M4_aic <- AIC(M4)
-M4_r <-  M4_sum$adj.r.squared
-
-# Model 5; non sst
-M5 <- lm(emerging_yr ~ status + bot_toe + biome + dem_pel + DemersPelag, data = Main_data)
-
-M5_sum <- summary(M5)
-M5_aic <- AIC(M5)
-M5_r <-  M5_sum$adj.r.squared
-
-# Model 6; all + basin
-M6 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Main_data)
-
-M6_sum <- summary(M6)
-M6_aic <- AIC(M6)
-M6_r <-  M6_sum$adj.r.squared
-
-# Model 7; Categories instead of status
-M7 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + dem_pel + DemersPelag, data = Main_data)
-
-M7_sum <- summary(M7)
-M7_aic <- AIC(M7)
-M7_r <-  M7_sum$adj.r.squared
-
-# Model 8; All variables + ccentroid lat
-M8 <- lm(emerging_yr ~ sst_toe + bot_toe + category + as.factor(biome) + DemersPelag + latitude, data = Main_data)
-
-M8_sum <- summary(M8)
-M8_aic <- AIC(M8)
-M8_r <-  M8_sum$adj.r.squared
-
-# Model 9; Include UN regions
-M9 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + DemersPelag + latitude + un_region, data = Main_data)
-
-M9_sum <- summary(M9)
-M9_aic <- AIC(M9)
-M9_r <-  M9_sum$adj.r.squared
-
-tibble(
-  "model" = paste("M",seq(1,9,1),sep=""),
-  "AIC" = c(M1_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic, M9_aic),
-  "rsquared" = c(M1_r,M2_r,M3_r,M4_r,M5_r,M6_r,M7_r,M8_r,M9_r)
-)
-
-# The Winner is M9    49283.   0.0765
-
-#Table for paper
-M9_sum %>% 
-  tidy() %>% 
-  filter(p.value < 0.05) %>% 
-  mutate(p.value = ifelse(p.value < 0.001,0.001,round(p.value,3))) %>% 
-  # stargazer(title = "Multiple regression results for time of emergence of Transboundary stocks.
-Only showing explanatory variables.
-The model significantly predicts time of emergencec (F(31,5270) = 15.38, p < 0.001, R2 = 0.06).",
-            # align=TRUE)
-  # knitr::kable() %>% 
-  write_csv(.,
-            "./Tables/mlr_result.csv")
-
-#Table for supplements
-M9_sum %>% 
-  tidy() %>% 
-  # filter(p.value < 0.05) %>% 
-  # mutate(p.value = ifelse(p.value < 0.001,0.001,round(p.value,3))) %>% 
-  # stargazer(title = "Multiple regression results for time of emergence of Transboundary stocks.
-Only showing explanatory variables.
-The model significantly predicts time of emergencec (F(18,6155) = 20.68, p < 0.001, R2 = 0.05).",
-            # align=TRUE)
-  # knitr::kable() %>% 
-  write_csv(.,
-            "./Tables/mlr_result_sup.csv")
 ```
@@
 
+###### Table.1 Multy linear model
+
+```{r multiple_reg, eval = F, echo = F}
+
+# Possible explanatory variables
+
+## Habitat association (DemersPelag)
+## Dem Vs Pelagic (Dem_pel)
+## SST time of Emer (Toe_sst)
+## Stock Status [FishForVisa] (Catch_trend)
+## Biogeographic location (Biogeo_reg)
+
+# AIC value; model between predictive capacity and complexity (i.e., number of explanatory variables) is given by the model with lowest AIC value
+
+# -------------------------------- #
+#### Build models
+# -------------------------------- #
+
+# Model 1; all variables
+M1 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Main_data)
+
+M1_sum <- summary(M1)
+M1_aic <- AIC(M1)
+M1_r <-  M1_sum$adj.r.squared
+
+# Model 2; non demPel
+M2 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel, data = Main_data)
+
+M2_sum <- summary(M2)
+M2_aic <- AIC(M2)
+M2_r <-  M2_sum$adj.r.squared
+
+# Model 3; non biome
+M3 <- lm(emerging_yr ~ sst_toe + bot_toe + status + dem_pel + DemersPelag, data = Main_data)
+
+M3_sum <- summary(M3)
+M3_aic <- AIC(M3)
+M3_r <-  M3_sum$adj.r.squared
+
+# Model 4; non status
+M4 <- lm(emerging_yr ~ sst_toe + bot_toe + biome + dem_pel + DemersPelag, data = Main_data)
+
+M4_sum <- summary(M4)
+M4_aic <- AIC(M4)
+M4_r <-  M4_sum$adj.r.squared
+
+# Model 5; non sst
+M5 <- lm(emerging_yr ~ status + bot_toe + biome + dem_pel + DemersPelag, data = Main_data)
+
+M5_sum <- summary(M5)
+M5_aic <- AIC(M5)
+M5_r <-  M5_sum$adj.r.squared
+
+# Model 6; all + basin
+M6 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Main_data)
+
+M6_sum <- summary(M6)
+M6_aic <- AIC(M6)
+M6_r <-  M6_sum$adj.r.squared
+
+# Model 7; Categories instead of status
+M7 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + dem_pel + DemersPelag, data = Main_data)
+
+M7_sum <- summary(M7)
+M7_aic <- AIC(M7)
+M7_r <-  M7_sum$adj.r.squared
+
+# Model 8; All variables + ccentroid lat
+M8 <- lm(emerging_yr ~ sst_toe + bot_toe + category + as.factor(biome) + DemersPelag + latitude, data = Main_data)
+
+M8_sum <- summary(M8)
+M8_aic <- AIC(M8)
+M8_r <-  M8_sum$adj.r.squared
+
+# Model 9; Include UN regions
+M9 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + DemersPelag + latitude + un_region, data = Main_data)
+
+M9_sum <- summary(M9)
+M9_aic <- AIC(M9)
+M9_r <-  M9_sum$adj.r.squared
+
+tibble(
+  "model" = paste("M",seq(1,9,1),sep=""),
+  "AIC" = c(M1_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic, M9_aic),
+  "rsquared" = c(M1_r,M2_r,M3_r,M4_r,M5_r,M6_r,M7_r,M8_r,M9_r)
+)
+
+# The Winner is M9    49283.   0.0765
+
+#Table for paper
+M9_sum %>% 
+  tidy() %>% 
+  filter(p.value < 0.05) %>% 
+  mutate(p.value = ifelse(p.value < 0.001,0.001,round(p.value,3))) %>% 
+  # stargazer(title = "Multiple regression results for time of emergence of Transboundary stocks.
+Only showing explanatory variables.
+The model significantly predicts time of emergencec (F(31,5270) = 15.38, p < 0.001, R2 = 0.06).",
+            # align=TRUE)
+  # knitr::kable() %>% 
+  write_csv(.,
+            "./Tables/mlr_result.csv")
+
+#Table for supplements
+M9_sum %>% 
+  tidy() %>% 
+  # filter(p.value < 0.05) %>% 
+  # mutate(p.value = ifelse(p.value < 0.001,0.001,round(p.value,3))) %>% 
+  # stargazer(title = "Multiple regression results for time of emergence of Transboundary stocks.
+Only showing explanatory variables.
+The model significantly predicts time of emergencec (F(18,6155) = 20.68, p < 0.001, R2 = 0.05).",
+            # align=TRUE)
+  # knitr::kable() %>% 
+  write_csv(.,
+            "./Tables/mlr_result_sup.csv")
+```
+
 ```{r TransIndex, eval = F, echo = F}

### 2020-11-14 Scripts/Manuscript.Rmd
Addressed comments for first round

@@
 output:
-  word_document: 
-    reference_docx: ../References/template.docx
+  # word_document: 
+    # reference_docx: ../References/template.docx
   pdf_document: default
@@
 - \usepackage{amsmath}
+- \usepackage{float} # Use the 'float' package
+- \floatplacement{figure}{H} # Make every figure with caption = h
 ---
@@
 
-Over the last century, human activities have fundamentally changed the physico-chemical composition of the ocean, substantially warming the waters, decreasing the pH and reducing oxygen concentration [@IPCC:2019to].
+Over the last century, human activities have fundamentally changed the physico-chemical composition of the ocean, substantially warming the waters, decreasing the pH and reducing oxygen concentration [@IPCC:2019tn].
 Throughout the ocean, marine species are distributed according to preferences in given environmental parameters (e.g., salinity, temperature, oxygen) [@Hutchinson:1957nt; @Nelson:2016kk].
@@
 Biogeography of marine species is projected to continue to shift as ocean conditions continue to change in the 21st century under climate change [@Cheung:2010dt].
-Climate change is expected to impact fisheries production, compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Cheung:2016Sci; @Barange:2014kr; @Singht:2019].
+Climate change is expected to impact fisheries production, compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].
 The projected risks of impacts can be reduced through improving the effectiveness of current fisheries management [@Gaines:2018sg; @Cheung:2018dh], including managing fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
@@
 Roughly, ToE is defined as the moment in time when an anthropogenic change signal (e.g., future trend) rises above the background noise of natural variability (e.g., historical variation) [@IPCC:2019tn].
-This concept has been applied to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Roggers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
+This concept has been applied to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Rodgers:2015jqq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
 The premise behind ToE is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background of natural climate variability [@Hawkins:2012hc].
@@
 This analysis includes 633 marine exploited transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-Each species shared by a pair of neighbouring EEZs was considered a transboundary stock [@Teh:2015gd; @PalaciosAbrantes:2020tv].
+Each time a species was shared by a pair of neighbouring EEZs it considered a transboundary stock [@Teh:2015gd; @PalaciosAbrantes:2020tv] for a total of 9,132 transboundary stocks.
 We defined the boundaries of the world’s EEZs using the *Sea Around Us* spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:FigureS1)) including island territories.
 We determined the intersections between polygons using the *R*-statistical software package *sf* [@PackagesfSimpleF:2018vp].
-Each EEZ was categorized by the geopolitical regions and natural biomes according to the United Nations and by natural biome [@Reygondeau:2019].
+Each EEZ was categorized by the geopolitical regions and natural biomes according to the United Nations and by natural biome [@Reygondeau:2019uh].
 Habitat preference of each species was determined based on information available from FishBase (http://www.fishbase.org).
@@
 
-We forced the DBEM with environmental data from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+We forced the DBEM with environmental data from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jqq].
 The GFDL-ESM2M is run under historical forcing until 2005 and follows the ‘high’ climate change emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [Riahi:2011dk].
@@
 Because the main approach of this paper relies on understanding the spatial and temporal variation of a species’ distribution, we have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
-We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodggers:2015jq; @Frolicher:2009ge].
+We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jqq; @Frolicher:2009ge].
 Thus, for our experiment, each of the ten GFDL-ESM2M-ensembles were started from infinitesimally small differences in Earth system initial conditions resulting in a unique atmosphere and ocean state at each point in time.
@@
 We developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
-This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \ref{fig:FigureS2}).
+This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref{fig:FigureS2}).
 The centroid of a transboundary stock was determined by the average latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
@@
 We ran a sensitivity analysis (e.g.
-*per* = 20, 50, 90) to relax this parameter (Figure \ref{fig:FigureS2}).
-The centroid of a EEZ was estimated using the *st* package in *R* (version 3.2.1) (Figure \ref{fig:FigureS1}).
+*per* = 20, 50, 90) to relax this parameter (Figure \@ref{fig:FigureS2}).
+The centroid of a EEZ was estimated using the *st* package in *R* (version 3.2.1) (Figure \@ref{fig:FigureS1}).
 For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in *R*;
@@
 \label{eq:two}
-
 D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
-
 \end{equation}
@@
 From the perspective of a state’s fisheries, higher TI indicates an increase in the challenges to the effective management and sustainability of the shared species’ fisheries.
-In contrast, a smaller TI suggests that the sharing of the species between neighbouring EEZs is becoming relatively more stable; a condition that favours effective management and sustainability of the shared species’ fisheries.
-The index was smoothed to a 10 year average to reduce interanual variabillity.
+In contrast, a smaller TI suggests that the sharing of the species between neighbouring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the shared species’ fisheries.
+The index was smoothed to a 10 year average to reduce interanual variability.
 
@@
 
-In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than operating on a non-cooperative basis [@Munro:1980sd].
-However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., *Threat Point* (TP) [@Sumaila:2013].
-Here, we define TP as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbour [@OCPalacios; @OCSumaila].
+In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than operating on a non-cooperative basis [@Munro:1979dg].
+However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., *Threat Point* (TP) [@Sumaila:2013hv].
+Here, we define TP as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
 Any proportion below the defined threat point of a country would result in a deflection from the joint management of the shared stock.
 
-First we estimated the stock share ratio (SSR) for each species and year (1951-2100) by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the species is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
+First we estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
+We did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
 Second, we average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
 The first time period ($t_h$) spans 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
-It is assumed as the historic baseline of shared distributions for each species.
+It is assumed as the historic baseline of shared distributions for each stock.
 We then selected two future periods; early $21^{st}$ century as the average of 2021 to 2040 (*te*), and mid $21^{st}$ century as the average of 2041 to 2060 (*tm*).
 We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
-The analysis was replicated for projected species distributions from each of the ten ensemble members and results were averaged across ensemble members.
+The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members.
 Third, we estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_h$) following Palacios-Abrantes et al.
@@
 
-Finally, the TP of each EEZ was calculated when $\Delta{SSR_{e,m}}$ is beyond one standard deviation of the historical distribution i.e., when $SSR_f \geq (SSR_f + \sigma)$ or $SSR_f \leq (SSR_f - \sigma)$.
+Finally, a substantial change in SSR of each EEZ was determined when $\Delta{SSR_{e,m}}$ was beyond one standard deviation of the historical distribution i.e., when $SSR_f \geq (SSR_f + \sigma)$ or $SSR_f \leq (SSR_f - \sigma)$.
+In other words, when the future SSR is below or above a neighbouring EEZ TP.
 
@@
 We tested the results for normality (e.g.
-skewness, kurtosis) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
+skewness = 1.02, kurtosis = 2.83) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
 Specifically, we tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of transboundary stocks.
@@
 
-### Time of Emergence of transbounudary species
+### Time of Emergence of transboundary species
 
-We estimated the ToE between neighboring EEZs for 633 transboundary stocks, that is, the first year where the future stock’s distribution (2006-2100) overpasses the historical variation (1951-2005).
-Since 2006 till 2020, 87% of the world EEZs have already seen at least one emerging stock and 60% of the transboundary species analyzed have emerged in at least a set of neighboring EEZ (e.g., *Engraulis ringens* shared by Chile and Peru emerged in 2010).
-Stocks will continue to emerge steadily towards 2097 when all 633 species will have emerged in at least one neighboring EEZ (Figure \ref{fig:Figure1}).
-The number of EEZs with emerging stocks will increase steeply until 2025 when it seems to flatten approaching 2081 when all of the EEZs analyzed here will have at least one emerging stock (Figure \ref{fig:Figure1}).
+The average ToE of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years with the first stocks emerging in 2006 and until 2100 (Figure \@ref(fig:Figure1).
+In total, 5,745 stocks will emerge between 2006 and 2100, that is 63% of the total transboundary stocks considered in this study (9,132).
+About 55% (n = 3,154) of those stocks are projected to have their shifts in distributions emerged between 2006 and 2020.
+Also, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 (e.g.
+*Engraulis ringens* shared by Chile and Peru emerged in 2010).
+Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century and are not expected to stop there (Figure \@ref(fig:Figure1)).
+The number of EEZs with emerging stocks is projected to increase steeply from 87% to 96% during the period of 2006 to 2020.
+After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.74% per year.
+By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:cumulative)).
 
-```{r Figure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.cap =  "Cumulative increase in time of emergence by Exclusive Economic Zone and transboundary species in absolute numbers and percentage."}
+```{r cumulative, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Cumulative increase in time of emergence by Exclusive Economic Zone and transboundary species in absolute numbers and percentage."}
 
@@
 
-Most of the world EEZs are projected to experience an average ToE across transboundary species before 2040 (Figure \ref{fig:Figure2}).
-In a few EEZs species are projected to have a relative late average ToE (e.g.
-Honduras average ToE is 2071) while 54 of the 280 EEZs (19%) have already seen their species emerged (e.g.
-average ToE across all species $\leq$ 2020).
-A higher number of emerging species will happen between latitudes 0$^o$ and 50$^o$N corresponding to regions of relatively small EEZs with multiple borders (e.g.
-The Caribbean and the Mediterranean) (Figure \ref{fig:Figure2}).
-In most cases, the current participation of the emerging species in a country’s total fishing revenue is less than 50% (Figure \ref{fig:Figure2}- Polygons).
-However, in some cases emerging species represent 85% of current transboundary fishing revenue (e.g.
-Peru) or more (e.g., Marshall Islands, 92%).
-Only in very few cases like Brazil and Alaska’s polar EEZ, no stock will emerge between 2006 and 2100.
+The median ToE of transboundary stocks varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, X2 = 287.23, DF = 93, p < 0.001).
+Around one fifth (19%) of the EEZs have an average time of emergence across species prior to 2020 (Figure \@ref{fig:Figure2}).
+Notable exceptions where the ToE is relative late include Honduras (2071), New Zealand (2055), Ireland (2067) and Pitcairn (2070).
+The EEZs in Eastern Asia, Latin America and the Caribbean as well as Polynesia differed significantly (*p* < 0.05) from (almost) all other regions (see Table S1 for full statistics).
+Tropical latitudes also hold the majority of countries with high vulnerability to shifts in transboundary stocks due to climate change (Figure \@ref{fig:Figure2}).
+Only in very few cases like Brazil and Alaska’s and Russia’s polar EEZs, no stock will emerge between 2006 and 2100.
+From the top 10 more vulnerable countries, seven are located in tropical waters (Tokelau 100, the	Marshall Islands 92, Peru 88, Kiribati 80, Tuvalu 80, Papua New Guinea 78, and the Bahamas 74) while only three are outside the tropics (Estonia 86, Finland 82 and Sweden 78).
+We set an arbitrary threshold of  1 s.d.
+(67% confidence) to determine the ToE (see Methods).
+We tested the sensitivity of our results using a larger threshold (i.e.,  2 s.d for a 95% confidence).
+While the selected 1 s.d.
+threshold is less conservative than a 2 s.d., the distribution of results do not dramatically change between thresholds, with a difference in ToE of about 4 years (Figure S\@ref{fig:S2}).
 
-```{r Figure2, eval = T,  echo = F, fig.width  = 13, fig.height =11, fig.cap =  "Average time of emergence of transboundary specie.
-Land polygon shows the vulnerability index.
-Exclusive Economic Zone polygon represents the average time of emergence across all species.
-Density plot on the right shows the latitudinal number of emerging species across EEZ centroids (10^o running mean)."}
+```{r Figure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of transboundary species.
+A) Land polygon shows the vulnerability index of individual country’s fisheries management to shifts in transboundary species.
+The warmer the colour the greater the vulnerability.
+Exclusive Economic Zone polygon represents the average time of emergence across all species, with the cooler the colour the later the average time of emergence.
+The Density plot on the right shows the latitudinal number of emerging species across EEZ centroids (10 degrees latitude running mean).
+B) Time of emergence according to the United Nations sub regions.
+Whiskers represent 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Line represents median, and black points represent points outside of 1.5*IQR."}
 
@@
 
-Regionally, most of African, Asian and European EEZs are projected to have a later ToE, while the Americas and Oceania will see an early ToE (Figure \ref{fig:Figure3}).
-Twelve of the 47 countries responsible for 75% of the yearly revenue generated from transboundary species between 2005 and 2010 will see an average ToE before 2030 (Figure \ref{fig:Figure3}).
-This includes the United States who occupies the $4^{th}$ place in revenue generated from transboundary fisheries.
-These 47 countries are also among those with the largest number of emerging taxa (Figure \ref{fig:Figure3}).
-From the top ten countries with more revenue from transboundary species, nine of them share a border.
-Most of these countries (e.g.
-China) will also have a large number of emerging stocks, while others like Peru and Norway will have a small number of emerging taxa but representing high revenue fisheries.
+Our analysis suggests that most countries responsible for 75% (n = 47) of the yearly revenue generated from transboundary species between 2005 and 2010 will see their stocks emerge in average before 2040 (\@ref{fig:Figure3}).
+In some cases like Indonesia, Thailand and the United States, stocks will emerge ass soon as 2030.
+These 47 countries are also among those with the largest number of emerging stocks.
+Notably, nine of the top ten countries earning the most revenue from transboundary species, share a border.
 
-```{r Figure3, eval = T,  echo = F, fig.width=4, fig.height=5, fig.cap =  "Average time of emergence of transboundary species per fishing entity.
-Fishing revenue in logarithmic scale.
-Showing names for the top 75th percentile.
+```{r Figure3, eval = T,  echo = F, fig.width=4, fig.height=5,fig.pos= "H", fig.cap =  "Average time of emergence of transboundary species per fishing entity, aggregated by colour according to region.
+Fishing revenue on a logarithmic scale.
+Showing country names for the top 90th percentile.
 Countries marked with an asterisk have at least one neighboring border in the group."}
@@
 
-We compared species ToE within EEZs by geographic region and by species habitat preference (Figure \ref{fig:Figure4}).
-The median ToE of transboundary species varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, *DF* = 93, *p* < 0.001).
-Post-hoc tests revealed regional patterns.
-EEZs in Easter Asia which have the latest average ToE across all habitat types (2041), and Latin America and the Caribbean and Polynesia who have the earliest average ToE across all habitats (2021 and 2018, respectively), significantly differ from (almost) all other regions.
-Other regions also significantly differ from each other but in a less frequent manner (see Table \ref{tab:TableS1}).
-When looking specifically at habitat association, we found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
+We compared species ToE within EEZs by species habitat preference according to FishBase's classification (Figure \@ref{fig:Figure4}).
+We found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
 Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table \ref{tab:TableS2}).
+In average, species associated with pelagic oceanic such as tunas (*Thunnus sp.*) and bathy-pelagic habitats like blue whiting (*Micromesistius poutassou*) as well as species associated to corals (e.g., Bigeye snapper	*Lutjanus	lutjanus*) will have an earliest time of emergence (Figure \@ref{fig:Figure4}).
+Only species associated to bathydemersal habitats such as Alaska plaice	*Pleuronectes	quadrituberculatus* and other habitats will have an average ToE later than 2025 (Figure \@ref{fig:Figure4}).
 
-```{r Figure4, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap =  "Regional time of emergence per habitat preference.
+```{r Figure4, eval = T,  echo = F, fig.width=14, fig.height=14,fig.pos= "H", fig.cap =  "Regional time of emergence per habitat preference.
 Regions according to the United Nations and habitat preference to FishBase.
@@
 
-### Changes in specie Share Ratio (SSR)
+### Changes in the threat point of transboundary stocks
 
@@
 
-We estimated the change in SSR (with a 63% of confidence) of transboundary species shared by neighboring EEZs by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) (Figure \ref{fig:Figure5}).
-Our results suggest that 87% (n = 238) of the world EEZs will see a change in SSR by 2030 (Figure \ref{fig:Figure5}A).
-Most EEZs will average a change lower than 25% which is expected to slightly increase in intensity towards the mid $21^{st}$ century (Figure \ref{fig:FigureS2}).
-However, this result masks larger changes in SSR for specific species.
-For example, Guatemala’s SSR of *Panulirus gracilis* shared with Mexico is projected to go from 12% to 26%.
-By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR $\pm$ 1 s.d.
-of historical distribution will be relatively large (global mean of 69% of species in an EEZ).
-For 17 EEZs, 100% of their analyzed transboundary stocks will shift, while only six EEZs will have less than 25% shifting stocks (Figure \ref{fig:Figure5}A-Polygons).
-Regions where EEZs are relatively small and have multiple borders (e.g.
-The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Figure \ref{fig:Figure1}B).
-Towards the mid $21^{st}$ century, changes in SSR per EEZ will not be substantial (Figure \ref{fig:FigureS3}) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \ref{fig:FigureS3}).
+We estimated the changes in the SSR of transboundary stocks by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) highlighting changes below a neighbouring country's TP.
+Our results suggest that 87% (n = 238) of the world EEZs will see a change in SSR below the country's TP by 2030 (Figure \@ref{fig:Figure5}A).
+Globally, most EEZs will average a change lower than 25% by the early $21^{st}$ century which is expected to slightly increase in intensity towards the mid $21^{st}$ century (Figures \@ref{fig:FigureS4} and S5).
+However, this result masks larger changes in SSR for specific stocks and EEZs.
+For example, Guatemala’s SSR of *Panulirus gracilis* shared with Mexico is projected to go from a historical 12% to 26%, a change of over 50%.
+By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR above the country's TP historical distribution (e.g., $\Delta{SSR} > \mu \pm \sigma$) will be relatively large (global mean of 69% of stocks in an EEZ).
+For 17 EEZs, 100% of their analyzed transboundary stocks will shift below the, while only six EEZs will have less than 25% stocks shifting below the TP (Figure \@ref{fig:Figure5}A-Polygons).
+In terms of direction of change, regions where EEZs are relatively small and have multiple borders (e.g.
+The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Figure \@ref{fig:Figure1}B).
+Towards the mid $21^{st}$ century, changes in SSR per EEZ will not be substantial (Figure \@ref{fig:FigureS5}) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref{fig:FigureS5}).
 
-```{r Figure5, eval = T,  echo = F, fig.width=12, fig.height=10, fig.cap =  "Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
+```{r Figure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
 Lines represent the average gain of transboundary stock share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
@@
 In this paper we endeavor to look at the timing and intensity of changes in the sharing of transboundary species.
-Our findings here are noteworthy in several ways.
-First, we show that the current distribution of most transboundary species have already emerged from the historical record and that virtually all EEZs have seen at least one emerging species since 2006 (Figure \ref{fig:Figure5}).
-Second, we identify regional patterns of complex changes in SSR and levels of vulnerability of transboundary fisheries management to climate change (Figure \ref{fig:Figure3}).
-Third, we identified that the historical share proportion of transboundary species will change by 2030 in most neighboring EEZs (Figure \ref{fig:Figure5}).
+Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world indicating climate change is already affecting international fisheries management (Figure \@ref{fig:Figure1}).
+Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require fast responses from transboundary fisheries management (Figure \@ref{fig:Figure2} and \@ref{fig:Figure3}).
+Finally, we show that in many instances around the world, the incoming years will see many transboundary stocks shifting their shared proportion below acceptable levels for sharing nations potentially becoming future sources of conflict (Figure \@ref{fig:Figure5})
 
-#### Emergence characteristics of transboundary species
+#### High present-day climate risk of on transboundary fisheries management
 
-We set out to evaluate the emergence characteristics of transboundary stocks across the global EEZs.
-Our results suggest that in most of the EEZs of the world more than half of the transboundary species already emerged since the early 2000s (Figure \ref{fig:Figure2}).
-This is in alignment with previous research showing changes in the composition of marine catches since the 1970's as evidence of climate change [@Cheung:2013gk].
-Indeed, an expansion in the distribution of multiple marine species has been documented in several regions of the world [@Poloczanska:2016kk].
-For example, in the early 2000s Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range reaching Washington waters where a new fishery rapidly developed [@Pinsky:2014by; @Zeidberg:2007gs].
+Our results suggest that in most EEZs, more than half of transboundary species emerged in the early 2000s (Figure \@ref{fig:Figure1}).
+This is in alignment with previous research pointing to changes in marine catch  composition since the 1970s as evidence of climate change [@Cheung:2013gk, @Poloczanska:2016kk].
+For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range, reaching Washington state waters where a new fishery rapidly developed [@Pinsky:2014by; @Zeidberg:2007gs].
 In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, led to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij].
-The large amount of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
-It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 23% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while ($O_2$) will have doubled under high greenhouse gas emission scenario [@Rodgers:2015jq].
+The large number of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jqq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
+It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 23% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while $O_2$ will have doubled under a high greenhouse gas emission scenario [@Rodgers:2015jq].
 
 Regional differences in ToE are likely the result of environmental factors, species vulnerability, and the characteristics of political boundaries.
-For example, the high number of stocks emerging early in the tropics suggest that tropical species are already on the move.
-The tropics are among the most species-rich biome [@Reygondeau:2019uh] and have potentially experienced some of the greatest levels of warming relative to the natural internal variability  [@IPCC:2019tn; @Roemmich:2015id].
-Here, marine species live close to the thermal tolerance making them highly vulnerable to warming waters [@IPCC:2014bz] consequently having to move towards deeper or sub-tropical waters [@Poloczanska:2016kk].
+For example, our findings underscore previous findings suggesting that tropical species are already on the move.
+The tropics are among the most species-rich biome [@Reygondeau:2019uhno] and have potentially experienced some of the greatest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
+In the tropics, marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz].
+Consequently, as a result of climate change, species have to move towards deeper or sub-tropical waters [@Poloczanska:2016kk] although recent studies suggest that in some regions species are shifting towards the equator and to shallow waters [@Clarke:2020ep].
 The premise of a shift already happening is supported by observational evidence of tropical species appearing in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis showing a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
 In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs and many borders that match peaks of emerging stocks.
-Specifically, it seems like South East Asia, The Caribbean, and the Mediterranean are responsible for substantial increases in emerging stocks (Figure \ref{fig:Figure3}).
+Specifically, it seems like South East Asia, the Caribbean, and the Mediterranean harbour substantial increases in emerging stocks (Figure \@ref{fig:Figure2}).
 
-### Some fisheries will be more vulnerable to emerging species
+### Hotspot of climate risk on transboundary fisheries management
 
+We have identified hotspot regions of climate risk for transboundary fisheries management (Figure \@ref{fig:Figure2}).
 The premise of our vulnerability index is that, from a management perspective, countries whose shifting transboundary stocks provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
-Similarly, the index is capable of capturing different sources of societal benefits (e.g.
-food security through catch or livelihood through value).
-We assume that the management of fisheries that provide large societal benefits will have more pressure from stakeholders while earlier times of emergence will result in less time to act, thus resulting in higher vulnerability.
-Thus, an early ToE and large fisheries benefits could indicate high vulnerability of transboundary fisheries management to climate change (Figure \ref{fig:Figure3}).
-For example, Polynesia and Micronesia are the only regions where the average ToE across all species and EEZs already happened; 2019 or is about too 2021, respectively (Figure \ref{fig:Figure5}).
-Countries in these regions are highly dependent on pelagic-oceanic (e.g.
-tunas) and reef-associated transboundary species (e.g.
-Species) for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh; @Cabral:2018hc].
-For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \ref{fig:Figure5}).
-In this case, policy has to work against the clock and under potentially high societal pressure to address an issue that could be impacting important fisheries.
+Thus, climate risk hotspots for transboundary management happen where EEZs will see an early ToE and where transboundary stocks are important in terms of fisheries catches and revenues (Figure \@ref{fig:Figure2}).
+For example, Polynesia and Micronesia are the only regions where the average ToE across all stocks and EEZs already happened; 2019 or is about too 2021, respectively (Figure \@ref{fig:Figure5}).
+Countries in these regions are highly dependent on pelagic-oceanic (e.g.,tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
+For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \@ref{fig:Figure5}).
+In this case, policy has to work against the clock and under potentially high societal pressure to address an issue that could be impacting important fisheries [@Burden:2019ij].
+Currently, these fisheries operate under the Vessel Day Scheme - which takes into account shifts in stocks and catches making it a good example where a management system has been set up that is somewhat anticipatory [@Aqorau:2018bh].
 However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to stocks sustainability.
-While it gives countries more time to prepare for substantial changes, species with relatively small overall participation can still be a source of conflict as portrayed by the Atlantic mackerel case (@Spijkers:2017ij, Chapter One].
+While it gives countries more time to prepare for substantial changes, stocks with relatively small overall participation on catches or revenue can still be a source of conflict as portrayed by the Atlantic mackerel case [@Spijkers:2017ij, Chapter 1].
 
-### Changes in specie-share ratio will be everywhere
+### Changes in species-share ratio will be ubiquitous
 
-By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Figure \ref{fig:Figure5}).
-Poleward shifts in SSR will be seen in regions of the world with relatively large EEZs and few boundaries like the North East Pacific.
-However, despite a general recognition that climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn], the Eastern Tropical Pacific will see an equatorial shift with EEZs from Mexico to Ecuador losing SSR to their southern neighbors.
-This shift is consistent with regional modeling studies suggesting that by mid $21^{st}$ century marine species will displace between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
-Moreover, regions with presence of multiple frontiers like the Caribbean and the North East Atlantic will see a complex network of changes in SSR with no marked direction (Figure \ref{fig:Figure5}B).
-These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] leading to joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
-However, game theory suggests that the greater number of parties negotiating, the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared species.
+By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Figure \@ref{fig:Figure5}).
+Poleward shifts in SSR are expected in regions of the world with relatively large EEZs and few boundaries, like the North East Pacific.
+However, despite observations and modelling work indicating that, generally, climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], the Eastern Tropical Pacific will see EEZs from Mexico to Ecuador losing SSR to their southern neighbors.
+Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems.
+They are consistent with regional modeling studies suggesting that by mid $21^{st}$ century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
+Moreover, regions characterized by multiple political boundaries, like the Caribbean and the North East Atlantic, will see a complex network of changes in SSR with no clear pattern (Figure \@ref{fig:Figure5}B).
+These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] supporting the implementation of joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
+However, game theory suggests that the greater the number of negotiating parties , the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared species.
 
@@
 
-We have identified three main sources of uncertainty in our analysis and point out ways to move forward in order to produce more reliable results to inform policy dealing with international fisheries under climate change.
-First, ouranalysis relies one Earth system model (GFDL-ESM2M) and climate change scenario (RCP 8.5).
-While the use of multiple ESMs would allow to capture the structural uncertainty of the models [@Frolicher:2016dg], using the GFDL-ESM2M results in conservative estimates as its projections, for example in sea surface temperature, are overall less extreme than in other CMIP5-type Earth system models [@Bopp:2013fg].
-Addressing the uncertainty in future projections is key to advance marine governance for species on the move [@Pinsky:2018cb], thus, future analysis should include  multiple ESMs.
-Secondly, we set an arbitrary threshold of $\pm$ 1 s.d.
-(67% confidence) for determining a ToE and SSR (see methods).
-We tested the sensitivity of our results using a larger threshold (e.g., $\pm$ 2 s.d for a 95% confidence).
-While the selected  1 s.d.
-threshold is more relaxed than a 2 s.d., the distribution of results do not dramatically change between thresholds with a difference in ToE of about 4 years (Figure \ref{fig:FigureS4}) Thirdly, here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations (Chpater One).
-Thus, our analysis is carried at the species level, rather than a more detailed organizational level (e.g.
-fish stock) due to limited spatial-specific data on species sub populations (stocks).
-Moreover, in many EEZs of the world fisheries are often managed at the species level (Chpater One) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
-However, reproducing our analysis at the local level, where spatially-explicit stock data is available would allow to produce more certain results in changes in both ToE and SSR of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level[@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
-
-We have identified three main sources of uncertainty in our analysis and point out ways to move forward in order to produce more reliable results to inform policy dealing with international fisheries under climate change.
-First, in this analysis we relay in one Earth System Model (GFDL-ESM2M) and climate change scenario (RCP 8.5).
-While the use of multiple ESMs would allow to capture the structural uncertainty of the models [@Frolicher:2016dg], using the GFDL-ESM2M results in conservative estimates as its projections are overall less extreme than similar models [@Bopp:2013fg].
-Addressing the uncertainty in future projections is key to advance marine governance for species on the move [@Pinsky:2018cb], thus, future efforts should look at the ensemble members of multiple ESM to capture the uncertainty embedded in the models.
-Secondly, we set an arbitrary threshold of $\pm$ 1 s.d.
-(67% confidence) to determine when a ToE and SSR would happen (see methods).
-We tested the sensitivity of our results to a larger threshold (e.g.
-$\pm$ 2 s.d for a 95% confidence).
-While the selected threshold is more relaxed than a 2 s.d., the distribution of results do not dramatically change between thresholds with a difference in ToE of about 4 years suggesting that this is a robust threshold (Figure \ref{fig:FigureS4}).
-Thirdly, in here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations (Chpater One).
-Thus, our analysis is carried at the species level, rather than a more detailed organizational level (e.g.
-fish stock) due to limited spatial-specific data on species sub populations (species).
-Reproducing our analysis with spatially-explicit stock data would allow to produce more certain results in changes in both SSR and ToE of transboundary species, and potentially identity different types of shared stock shifts [@Link:2010ei].
-However, many regions of the world lack these detailed data.
-Moreover, in many EEZs of the world fisheries are often managed at the species level (Chpater One).
+There are two main sources of uncertainty in our analysis that can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
+First, our analysis relies on the combination of a single Earth system model (GFDL-ESM2M) and climate change scenario (RCP 8.5).
+The use of multiple ESMs can capture the structural uncertainty of the models [@Frolicher:2016dg; @Lotze:2019ce], however, exercises looking at changes in fish biomass with multiple models show an overall agreement between models in terms of direction of change but variable magnitude.
+Using the GFDL-ESM2M results in conservative estimates as its projections, for example in sea surface temperature, are overall less extreme than in other CMIP5-type Earth system models [@Bopp:2013fg].
+Addressing the uncertainty in future projections is key to advance marine governance for species on the move [@Pinsky:2018cb], thus, further research that includes multiple ESMs could provide more accurate results in terms of magnitude but will likely follow the same trend showed here [@Lotze:2019ce].
+Secondly, here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations (Chapter One).
+We consider that each species shared by a set of neighbouring EEZs is a shared stock for a total of 9,132 shared stocks.
+Thus, our analysis does not consider a more detailed organizational level such as biologically-meaningful sub-populations within an EEZ due to limited biological and spatial-specific data on stocks sub-populations.
+Moreover, in many EEZs of the world fisheries are often managed at the species level (Chapter One) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
+However, reproducing our analysis regionally, where spatially-explicit stock data is available would allow to produce more certain results in changes in both ToE and SSR of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
 
@@
 
-<!-- ![Figure S1. World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points).](../Figures/Supp1.png) -->
-
-```{r FigureS2, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
+```{r FigureS2, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel).
+Time of emergence is defined as the first year when the TI overpasses historical values (Bottom panel)."}
 
@@
 
-<!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005))](../Figures/Supp2.png) -->
-
-```{r FigureS3, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial species that will see a shift.
-Panel B zooms in to specific areas shown in grey boxes of A."}
+```{r FigureS3, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
 
@@
 
-```{r FigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
+```{r FigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
 
 img <- readPNG("../Figures/Supp4.png")
+grid.raster(img)
+
+```
+
+<!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005))](../Figures/Supp2.png) -->
+
+```{r FigureS5, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
+Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial species that will see a shift.
+Panel B zooms in to specific areas shown in grey boxes of A."}
+
+img <- readPNG("../Figures/Supp5.png")
 grid.raster(img)

### 2020-11-16 Scripts/Initial_Analysis.Rmd
updates to figure 1

@@
 ```{r Time_lapse_plot}
+
 #### Both plots for tresh one
 
-ggplot() +
+plot <- ggplot() +
   geom_area(data = subset(plot_data, level == "eez"),
@@
             ),
-            alpha= 0.5,
+            alpha= 0.8,
             size=0.5,
@@
              scales = "free") + 
-  scale_y_continuous("",expand = expand_scale(mult = c(0, .05))) +
+  # scale_y_continuous("Cumulative number",expand = expand_scale(mult = c(0, .05))
+  #                    ) +
   scale_x_continuous("Time of emergence",
@@
   theme(
-    legend.position = "right"
+    legend.position = "right",
+    strip.text.x = element_blank()
   ) +
@@
   scale_fill_viridis(discrete = T,
+                     begin = 0.1, 
+                     end = 0.7,
                      labels = c("EEZs","Stocks")

### 2020-11-16 Scripts/Manuscript.Rmd
updates to figure 1

@@
 
-The concept of shared stocks were developed following the ratification of  the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
-As defined by the Food and Agriculture Organization, shared stocks can be classified into three non-exclusive categories: (*i*) transboundary, those that cross neighboring EEZs; (*ii*) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@Munro:2004th].
+The concept of shared stocks was developed following the ratification of  the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
+As defined by the Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (*i*) transboundary, those that cross neighboring EEZs; (*ii*) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@Munro:2004th].
 While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
@@
 Roughly, ToE is defined as the moment in time when an anthropogenic change signal (e.g., future trend) rises above the background noise of natural variability (e.g., historical variation) [@IPCC:2019tn].
-This concept has been applied to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Rodgers:2015jqq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
+This concept has been applied to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
 The premise behind ToE is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background of natural climate variability [@Hawkins:2012hc].
@@
 This analysis includes 633 marine exploited transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-Each time a species was shared by a pair of neighbouring EEZs it considered a transboundary stock [@Teh:2015gd; @PalaciosAbrantes:2020tv] for a total of 9,132 transboundary stocks.
+Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd; @PalaciosAbrantes:2020tv] for a total of 9,132 transboundary stocks.
 We defined the boundaries of the world’s EEZs using the *Sea Around Us* spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:FigureS1)) including island territories.
 We determined the intersections between polygons using the *R*-statistical software package *sf* [@PackagesfSimpleF:2018vp].
-Each EEZ was categorized by the geopolitical regions and natural biomes according to the United Nations and by natural biome [@Reygondeau:2019uh].
-Habitat preference of each species was determined based on information available from FishBase (http://www.fishbase.org).
+Each EEZ was categorized by the geopolitical regions according to the United Nations and by natural biome [@Reygondeau:2019uh].
+The habitat preference of each species was determined based on information available from FishBase (http://www.fishbase.org).
 For each species and EEZ, we used the *Sea Around Us* data to estimate catch and fishing revenue from fishing activities within global EEZs.
@@
 
-We forced the DBEM with environmental data from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jqq].
-The GFDL-ESM2M is run under historical forcing until 2005 and follows the ‘high’ climate change emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [Riahi:2011dk].
+We forced the DBEM with environmental data from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+The GFDL-ESM2M is run under historical forcing until 2005 and follows the ‘high’ climate change emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
 In GFDL-ESM2M, atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
 Because the main approach of this paper relies on understanding the spatial and temporal variation of a species’ distribution, we have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
-We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jqq; @Frolicher:2009ge].
+We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
 Thus, for our experiment, each of the ten GFDL-ESM2M-ensembles were started from infinitesimally small differences in Earth system initial conditions resulting in a unique atmosphere and ocean state at each point in time.
@@
 In contrast, a smaller TI suggests that the sharing of the species between neighbouring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the shared species’ fisheries.
-The index was smoothed to a 10 year average to reduce interanual variability.
+The index was smoothed to a 10 year average to reduce interannual variability.
 
@@
 Knowing the point in time (e.g., year) at which the distribution of a shared stock will diverge from its natural internal variability is important in informing the lead-time to which climate adaptation in transboundary fisheries management needs to be implemented [@Link:2010ei].
-Thus, we define ToE as the time when the shared stock between neighbouring EEZs will exceed the historical natural internal variability as follows:
+Thus, we define ToE as the time when the stock shared between neighbouring EEZs will exceed the historical natural internal variability as follows:
 
@@
 In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, led to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij].
-The large number of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jqq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
+The large number of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
 It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 23% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while $O_2$ will have doubled under a high greenhouse gas emission scenario [@Rodgers:2015jq].
@@
 For example, our findings underscore previous findings suggesting that tropical species are already on the move.
-The tropics are among the most species-rich biome [@Reygondeau:2019uhno] and have potentially experienced some of the greatest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
+The tropics are among the most species-rich biome [@Reygondeau:2019uh] and have potentially experienced some of the greatest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
 In the tropics, marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz].

### 2020-11-16 Scripts/Manuscript.Rmd
Updates to chapter

@@
 ---
-title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change
-subtitle: Challenges to transboundary fisheries under climate change
-author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^'
-date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern.'
 output:
-  # word_document: 
-    # reference_docx: ../References/template.docx
-  pdf_document: default
+  bookdown::pdf_book:
+    toc : no
+    keep_tex: no
+    number_sections: yes
+geometry: margin=1in
+editor_options:
+  chunk_output_type: inline
 bibliography: ../References/ToE_Ref.bib
-csl: ../References/science.csl
 header-includes:
@@
 ---
+
+<!-- --- -->
+<!-- title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change -->
+<!-- subtitle: Challenges to transboundary fisheries under climate change -->
+<!-- author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^' -->
+<!-- date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
+^2^Climate and Environmental Physics, Physics Institute, University of Bern.' -->
+<!-- output: -->
+<!--   # word_document:  -->
+<!--     # reference_docx: ../References/template.docx -->
+<!--   pdf_document: default -->
+<!-- bibliography: ../References/ToE_Ref.bib -->
+<!-- csl: ../References/science.csl -->
+<!-- header-includes: -->
+<!-- - \usepackage{amsmath} -->
+<!-- - \usepackage{float} # Use the 'float' package -->
+<!-- - \floatplacement{figure}{H} # Make every figure with caption = h -->
+<!-- --- -->
 
@@
 
-The effectiveness of managing fisheries targeting shared stocks is challenged by the shifts in marine species distribution under climate change [@Pinsky:2018cb; @Pinsky:wq].
+The effectiveness of managing fisheries targeting shared stocks is challenged by the shifts in marine species distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
 In many cases, catch or fishing effort quota on shared stocks are based on historical records, [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change in the fish stocks and fisheries [@PalaciosAbrantes:2020tv].
@@
 We developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
-This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref{fig:FigureS2}).
+This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:FigureS2)).
 The centroid of a transboundary stock was determined by the average latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
@@
 We ran a sensitivity analysis (e.g.
-*per* = 20, 50, 90) to relax this parameter (Figure \@ref{fig:FigureS2}).
-The centroid of a EEZ was estimated using the *st* package in *R* (version 3.2.1) (Figure \@ref{fig:FigureS1}).
+*per* = 20, 50, 90) to relax this parameter (Figure \@ref{fig:FigureS2)).
+The centroid of a EEZ was estimated using the *st* package in *R* (version 3.2.1) (Figure \@ref{fig:FigureS1)).
 For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in *R*;
@@
 \label{eq:five}
-index' = (P_c + P_v)/ToE
+index' = \frac{(P_c + P_v)}{ToE}
 \end{equation}
 
-where *P_c* and *P_v* indicate the proportion accounted for by shared species in total catch and value generated from transboundary fisheries within a country’s EEZ, respectively.
+where $P_c$ and $P_v$ indicate the proportion accounted for by shared species in total catch and value generated from transboundary fisheries within a country’s EEZ, respectively.
 Here, the proportion of total catch and value are represented as the average of 2005 to 2014 within individual EEZs according to the *Sea Around Us*.
@@
 In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than operating on a non-cooperative basis [@Munro:1979dg].
-However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., *Threat Point* (TP) [@Sumaila:2013hv].
+However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., "Threat Point" (TP) [@Sumaila:2013hv].
 Here, we define TP as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
@@
 -->
-All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@PackagegmtInterfa:2017vh] and *zoo* [@PackagezooSInfr:2019tm], *ggrepel* [@PackageggrepelAut:2020to], *zeallot* [@PackagezeallotMul:2018wy], *viridis* [@PackageviridisDef:2018us], *cowplot* [@CowplotStreamlined:2019wt], *moments* [@PackagemomensMome:2015uh] and *pgirmess* [@PackagepgirmessSp:2018um].
+All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@PackagegmtInterfa:2017vh] and *zoo* [@PackagezooSInfr:2019tm], *ggrepel* [@PackageggrepelAut:2020to], *zeallot* [@PackagezeallotMul:2018wy], *viridis* [@PackageviridisDef:2018us], *cowplot* [@CowplotStreamlined:2019wt], *moments* [@PackagemomensMome:2015uh] and *pgirmess* [@PackagepgirmessSp:2018um].
 All code is available at https://github.com/jepa/TransEmergence
@@
 The median ToE of transboundary stocks varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, X2 = 287.23, DF = 93, p < 0.001).
-Around one fifth (19%) of the EEZs have an average time of emergence across species prior to 2020 (Figure \@ref{fig:Figure2}).
+Around one fifth (19%) of the EEZs have an average time of emergence across species prior to 2020 (Figure \@ref(fig:Figure2).
 Notable exceptions where the ToE is relative late include Honduras (2071), New Zealand (2055), Ireland (2067) and Pitcairn (2070).
 The EEZs in Eastern Asia, Latin America and the Caribbean as well as Polynesia differed significantly (*p* < 0.05) from (almost) all other regions (see Table S1 for full statistics).
-Tropical latitudes also hold the majority of countries with high vulnerability to shifts in transboundary stocks due to climate change (Figure \@ref{fig:Figure2}).
+Tropical latitudes also hold the majority of countries with high vulnerability to shifts in transboundary stocks due to climate change (Figure \@ref(fig:Figure2).
 Only in very few cases like Brazil and Alaska’s and Russia’s polar EEZs, no stock will emerge between 2006 and 2100.
@@
 While the selected 1 s.d.
-threshold is less conservative than a 2 s.d., the distribution of results do not dramatically change between thresholds, with a difference in ToE of about 4 years (Figure S\@ref{fig:S2}).
+threshold is less conservative than a 2 s.d., the distribution of results do not dramatically change between thresholds, with a difference in ToE of about 4 years (Figure \@ref(fig:FigureS2).
 
@@
 The Density plot on the right shows the latitudinal number of emerging species across EEZ centroids (10 degrees latitude running mean).
+Aqua color = no emerging stocks.
 B) Time of emergence according to the United Nations sub regions.
@@
 
-Our analysis suggests that most countries responsible for 75% (n = 47) of the yearly revenue generated from transboundary species between 2005 and 2010 will see their stocks emerge in average before 2040 (\@ref{fig:Figure3}).
+Our analysis suggests that most countries responsible for 75% (n = 47) of the yearly revenue generated from transboundary species between 2005 and 2010 will see their stocks emerge in average before 2040 (Figure \@ref(fig:Figure3).
 In some cases like Indonesia, Thailand and the United States, stocks will emerge ass soon as 2030.
@@
 
-```{r Figure3, eval = T,  echo = F, fig.width=4, fig.height=5,fig.pos= "H", fig.cap =  "Average time of emergence of transboundary species per fishing entity, aggregated by colour according to region.
+```{r Figure3, eval = T,  echo = F, fig.width=5, fig.height=6,fig.pos= "H", fig.cap =  "Average time of emergence of transboundary species per fishing entity, aggregated by colour according to region.
 Fishing revenue on a logarithmic scale.
@@
 
-We compared species ToE within EEZs by species habitat preference according to FishBase's classification (Figure \@ref{fig:Figure4}).
+We compared species ToE within EEZs by species habitat preference according to FishBase's classification (Figure \@ref(fig:Figure4).
 We found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
-Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table \ref{tab:TableS2}).
-In average, species associated with pelagic oceanic such as tunas (*Thunnus sp.*) and bathy-pelagic habitats like blue whiting (*Micromesistius poutassou*) as well as species associated to corals (e.g., Bigeye snapper	*Lutjanus	lutjanus*) will have an earliest time of emergence (Figure \@ref{fig:Figure4}).
-Only species associated to bathydemersal habitats such as Alaska plaice	*Pleuronectes	quadrituberculatus* and other habitats will have an average ToE later than 2025 (Figure \@ref{fig:Figure4}).
+Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table @\ref(tab:TableS2).
+In average, species associated with pelagic oceanic such as tunas (*Thunnus sp.*) and bathy-pelagic habitats like blue whiting (*Micromesistius poutassou*) as well as species associated to corals (e.g., Bigeye snapper	*Lutjanus	lutjanus*) will have an earliest time of emergence (Figure \@ref(fig:Figure4)).
+Only species associated to bathydemersal habitats such as Alaska plaice	*Pleuronectes	quadrituberculatus* and other habitats will have an average ToE later than 2025 (Figure \@ref{fig:Figure4)).
 
@@
 ToE of SST), human drivers (e.g.
-catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary species (Table \ref{tab:TableS3}).
-Results from the multiple linear regression analysis show a significant relation between ToE of transboundary species and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref{tab:Table1}).
+catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary species (Table \ref{tab:TableS3)).
+Results from the multiple linear regression analysis show a significant relation between ToE of transboundary species and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref{tab:Table1)).
 -->
@@
 The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Figure \@ref{fig:Figure1}B).
-Towards the mid $21^{st}$ century, changes in SSR per EEZ will not be substantial (Figure \@ref{fig:FigureS5}) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref{fig:FigureS5}).
+Towards the mid $21^{st}$ century, changes in SSR per EEZ will not be substantial (Figure \@ref{fig:FigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref{fig:FigureS5)).
 
@@
 In this paper we endeavor to look at the timing and intensity of changes in the sharing of transboundary species.
-Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world indicating climate change is already affecting international fisheries management (Figure \@ref{fig:Figure1}).
-Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require fast responses from transboundary fisheries management (Figure \@ref{fig:Figure2} and \@ref{fig:Figure3}).
-Finally, we show that in many instances around the world, the incoming years will see many transboundary stocks shifting their shared proportion below acceptable levels for sharing nations potentially becoming future sources of conflict (Figure \@ref{fig:Figure5})
+Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world indicating climate change is already affecting international fisheries management (Figure \@ref{fig:Figure1)).
+Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require fast responses from transboundary fisheries management (Figure \@ref{fig:Figure2} and \@ref{fig:Figure3)).
+Finally, we show that in many instances around the world, the incoming years will see many transboundary stocks shifting their shared proportion below acceptable levels for sharing nations potentially becoming future sources of conflict (Figure \@ref{fig:Figure5))
 
@@
 
-Our results suggest that in most EEZs, more than half of transboundary species emerged in the early 2000s (Figure \@ref{fig:Figure1}).
+Our results suggest that in most EEZs, more than half of transboundary species emerged in the early 2000s (Figure \@ref{fig:Figure1)).
 This is in alignment with previous research pointing to changes in marine catch  composition since the 1970s as evidence of climate change [@Cheung:2013gk, @Poloczanska:2016kk].
@@
 In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs and many borders that match peaks of emerging stocks.
-Specifically, it seems like South East Asia, the Caribbean, and the Mediterranean harbour substantial increases in emerging stocks (Figure \@ref{fig:Figure2}).
+Specifically, it seems like South East Asia, the Caribbean, and the Mediterranean harbour substantial increases in emerging stocks (Figure \@ref{fig:Figure2)).
 
@@
 
-We have identified hotspot regions of climate risk for transboundary fisheries management (Figure \@ref{fig:Figure2}).
+We have identified hotspot regions of climate risk for transboundary fisheries management (Figure \@ref{fig:Figure2)).
 The premise of our vulnerability index is that, from a management perspective, countries whose shifting transboundary stocks provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
-Thus, climate risk hotspots for transboundary management happen where EEZs will see an early ToE and where transboundary stocks are important in terms of fisheries catches and revenues (Figure \@ref{fig:Figure2}).
-For example, Polynesia and Micronesia are the only regions where the average ToE across all stocks and EEZs already happened; 2019 or is about too 2021, respectively (Figure \@ref{fig:Figure5}).
+Thus, climate risk hotspots for transboundary management happen where EEZs will see an early ToE and where transboundary stocks are important in terms of fisheries catches and revenues (Figure \@ref{fig:Figure2)).
+For example, Polynesia and Micronesia are the only regions where the average ToE across all stocks and EEZs already happened; 2019 or is about too 2021, respectively (Figure \@ref{fig:Figure5)).
 Countries in these regions are highly dependent on pelagic-oceanic (e.g.,tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
-For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \@ref{fig:Figure5}).
+For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \@ref{fig:Figure5)).
 In this case, policy has to work against the clock and under potentially high societal pressure to address an issue that could be impacting important fisheries [@Burden:2019ij].
@@
 
-By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Figure \@ref{fig:Figure5}).
+By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Figure \@ref{fig:Figure5)).
 Poleward shifts in SSR are expected in regions of the world with relatively large EEZs and few boundaries, like the North East Pacific.
@@
 
-<!-- $$I =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h})}) |  \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{C,y}}{sd(D_{B,h})}) | \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{Cy}}{sd(D_{B,h})})/N$$ -->
+<!-- $$I =(\frac{D_{A,y}}{sd(D_{A,h))} - \frac{D_{B,y}}{sd(D_{B,h)))) |  \\  +(\frac{D_{A,y}}{sd(D_{A,h))} - \frac{D_{C,y}}{sd(D_{B,h)))) | \\  +(\frac{D_{A,y}}{sd(D_{A,h))} - \frac{D_{Cy}}{sd(D_{B,h))))/N$$ -->
 

### 2020-11-20 Scripts/Initial_Analysis.Rmd
updates on figures and stock from species

@@
              scales = "free") + 
-  # scale_y_continuous("Cumulative number",expand = expand_scale(mult = c(0, .05))
-  #                    ) +
+  scale_y_continuous("",
+                     expand = c(0,0)
+                     ) +
   scale_x_continuous("Time of emergence",
@@
                      limits = c(2005,2100),
+                     label = c(seq(2000,2075,25),">2100"),
                      expand = c(0, 0)
@@
             by = c("fishing_entity_id","variable")) %>% 
-  mutate(participation = (sum_value/tot_value)*100)
+  mutate(participation = (sum_value/tot_value)*100+2000)
 
@@
 
-Index_data <- Eez_d_hist %>% 
-  left_join(Value_map_data) %>% 
-  filter(
-    tresh == "Tresh_one",
-    metric == "mean_yr"
-  ) %>% 
-  group_by(sau_fishing_entity,fishing_entity_id,variable) %>% 
-  summarise(
-    toe = mean(value,na.rm=T),
-    # sum_value = sum(sum_value),
-    mean_part = mean(participation,na.rm=T)
-  ) %>% 
-  filter(!is.na(variable)) %>% 
-  # spread(variable,sum_value) %>% 
-  spread(variable,mean_part) %>% 
-  mutate(
-    imp = catch+value,
-    index = imp/toe
-  )
+# Index_data <- Eez_d_hist %>% 
+#   left_join(Value_map_data) %>% 
+#   filter(
+#     tresh == "Tresh_one",
+#     metric == "mean_yr"
+#   ) %>% 
+#   group_by(sau_fishing_entity,fishing_entity_id,variable) %>% 
+#   summarise(
+#     toe = mean(value,na.rm=T),
+#     # sum_value = sum(sum_value),
+#     mean_part = mean(participation,na.rm=T)
+#   ) %>% 
+#   filter(!is.na(variable)) %>% 
+#   # spread(variable,sum_value) %>% 
+#   spread(variable,mean_part) %>% 
+#   mutate(
+#     imp = catch+value,
+#     index = imp/toe
+#   )
 
@@
 # Get min and maximums
-min_index <- min(Index_data$index)
-max_index <- max(Index_data$index)
-
-Index_data <- Index_data %>% 
-  mutate(
-    index_norm =(index-min_index)/(max_index-min_index),
-    index_2000 = 2000 + (index-min_index)*(2100-2000)/(max_index-min_index)
-  )
+# min_index <- min(Index_data$index)
+# max_index <- max(Index_data$index)
+
+# Index_data <- Index_data %>% 
+#   mutate(
+    # index_norm =(index-min_index)/(max_index-min_index),
+  #   index_2000 = 2000 + (index-min_index)*(2100-2000)/(max_index-min_index)
+  # )
 
@@
 
-Value_map_data <- Index_data %>%
-  select(sau_fishing_entity,fishing_entity_id,
+# Value_map_data <- Index_data %>%
+  # select(sau_fishing_entity,fishing_entity_id,
          # tresh,
          # variable,
-         participation = index_2000)
+         # participation = index_2000)
 
@@
 
-for(i in 1:1){
+for(i in 1:2){
 
@@
              y = -69,
-             label = "Vulnerabillity index",
+             label = "Participation (%)",
              color="black",
@@
 
-kruskalmc(emerging_yr~un_sub_region, data = Reg_diff_data)
+kruskalmc(emerging_yr~un_sub_region, data = Reg_diff_data) %>% 
+  as.data.frame() %>% 
+  rownames_to_column("comparrison") %>% 
+  # head()
+  write_csv("kw_regional.csv")
 
@@
 
-kruskalmc(emerging_yr~DemersPelag, data = Hab_diff_data)
+kruskalmc(emerging_yr~DemersPelag, data = Hab_diff_data) %>% 
+  as.data.frame() %>% 
+  rownames_to_column("comparrison") %>% 
+  # head()
+  write_csv("kw_habitat.csv")
 
@@
 n_taxa_data <- Prop_results %>% 
-  filter(variable == "n_per") %>%
+  filter(variable == "n_spp") %>%
   select(eez_name,early,mid) %>% 

### 2020-11-20 Scripts/Manuscript.Rmd
updates to manuscript and supplements

@@
   "png",
-  "grid"
+  "grid",
+  "knitr",
+ "kableExtra"
   # "here", # for dbem_import `here()`
@@
 
-Over the last century, human activities have fundamentally changed the physico-chemical composition of the ocean, substantially warming the waters, decreasing the pH and reducing oxygen concentration [@IPCC:2019tn].
-Throughout the ocean, marine species are distributed according to preferences in given environmental parameters (e.g., salinity, temperature, oxygen) [@Hutchinson:1957nt; @Nelson:2016kk].
+Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming the waters, decreasing the pH and reducing oxygen content [@IPCC:2019tn].
+Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk].
 As a result of climate change, many marine species have changed their distribution towards higher latitude, deeper water or following local environmental gradients [@Poloczanska:2016kk].
-Biogeography of marine species is projected to continue to shift as ocean conditions continue to change in the 21st century under climate change [@Cheung:2010dt].
+Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt].
 Climate change is expected to impact fisheries production, compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].
@@
 The concept of shared stocks was developed following the ratification of  the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
-As defined by the Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (*i*) transboundary, those that cross neighboring EEZs; (*ii*) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@Munro:2004th].
+As defined by the United Nation's Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (i) transboundary, those that cross neighboring EEZs; (ii) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (iii) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions including both the high seas and EEZs [@Munro:2004th].
 While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
@@
 With future shifts in species distributions under climate change, the challenges posed on shared stocks fisheries management will continue to increase in the $21^{st}$ century.
-Thus, there is a need to understand the time when climate change will affect the sharing distribution of these species as well as the intensity in order to prepare ocean governance to species on the move [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
+Thus, there is a need to understand the time when climate change will affect the sharing distribution of these species as well as the intensity intensity of the impact in order to prepare ocean governance to species on the move [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
 
-We propose that adapting the concepts of "Time of Emergence" (ToE) and “Thread Point” (TP) can help address some of the management challenges imposed by shifting species.
+The timing and intensity of climate impacts on managing transboundary fish stocks can be quantitative examined through the concepts of "Time of Emergence" (ToE) and “Thread Point” (TP) can help address some of the management challenges imposed by shifting species.
 Roughly, ToE is defined as the moment in time when an anthropogenic change signal (e.g., future trend) rises above the background noise of natural variability (e.g., historical variation) [@IPCC:2019tn].
@@
 Game theory has been widely used to investigate and manage shared stocks [@Sumaila:2013hv].
-Specifically, Sumaila et al., *in press* applied this concept to understand the impacts of climate change to the economics of shared species between Canada and the United States.
+Specifically, Sumaila et al.
+(*in press*) applied this concept to understand the impacts of climate change to the economics of shared species between Canada and the United States.
 
 Here, we estimate the ToE and TP of the share distribution of transboundary stocks worldwide.
-We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected shared species [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emission scenario RCP8.5 (see Methods).
+We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emission scenario RCP8.5 (see Methods).
 We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., we did not include the high seas).
-Our results suggest that many shared stocks have already ‘emerged’ and overpassed their share TP in most neighboring EEZs posing a threat to the sustainability of shared stocks and the resilience of international treaties.
+Our results suggest that many shared stocks have already ‘emerged’ from historical natural variability and overpassed their share TP in most neighboring EEZs posing a threat to the sustainability of shared stocks and the resilience of international treaties.
 
@@
 
-This analysis includes 633 marine exploited transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
+This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
 Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd; @PalaciosAbrantes:2020tv] for a total of 9,132 transboundary stocks.
-We defined the boundaries of the world’s EEZs using the *Sea Around Us* spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:FigureS1)) including island territories.
-We determined the intersections between polygons using the *R*-statistical software package *sf* [@PackagesfSimpleF:2018vp].
+We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:FigureS1)) including island territories.
+We determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
 Each EEZ was categorized by the geopolitical regions according to the United Nations and by natural biome [@Reygondeau:2019uh].
-The habitat preference of each species was determined based on information available from FishBase (http://www.fishbase.org).
-For each species and EEZ, we used the *Sea Around Us* data to estimate catch and fishing revenue from fishing activities within global EEZs.
+The habitat association of each species was determined based on information available from FishBase (http://www.fishbase.org).
+For each species and EEZ, we used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
 We report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
@@
 This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (see Cheung *et al.* [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
-Recent work has demonstrated that, at the global scale, DBEM is able to reproduce catch changes similar in magnitude to those detected in observed catch time series in the
-world’s EEZs (Cheung et al. in revision).
+The DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd], Cheung et al.
+*in revision*].
 
-We forced the DBEM with environmental data from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+We forced the DBEM with simulated ocean conditions from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
 The GFDL-ESM2M is run under historical forcing until 2005 and follows the ‘high’ climate change emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
@@
 where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
-To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top $10^{th}$ percentile (e.g., *per* = 10%).
+To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top $10^{th}$ percentile (e.g., per = 10%).
 We ran a sensitivity analysis (e.g.
-*per* = 20, 50, 90) to relax this parameter (Figure \@ref{fig:FigureS2)).
-The centroid of a EEZ was estimated using the *st* package in *R* (version 3.2.1) (Figure \@ref{fig:FigureS1)).
-For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in *R*;
+per = $20^{th}$, $50^{th}$, $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:FigureS2)).
+The centroid of a EEZ was estimated using the st package in R (version 3.2.1) (Figure \@ref(fig:FigureS1)).
+For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
 
@@
 Thus, a higher TI means that a shared species becomes more or less abundant relative to its neighbouring EEZs that is beyond what the fisheries experienced historically.
-From the perspective of a state’s fisheries, higher TI indicates an increase in the challenges to the effective management and sustainability of the shared species’ fisheries.
+From the perspective of a state’s fisheries, higher TI indicates an increase in the challenges to effective management and sustainability of the shared species’ fisheries.
 In contrast, a smaller TI suggests that the sharing of the species between neighbouring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the shared species’ fisheries.
@@
 We set a threshold of TI above which we consider a shared stock to have emerged from the historical internal variability of the TI.
-The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \ref{eq:four}, representing a probability of 68% that the transboundary index has emerged from historical variability).
+The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four), representing a probability of 68% that the transboundary index has emerged from historical variability).
 We tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
 
-#### Computing a transboundary fisheries vulnerability index
+<!-- #### Computing a transboundary fisheries vulnerability index (VI) -->
 
-We created a country-specific index to examine the risk of impacts of species’ range shifts on transboundary fisheries management.
-The index reflects both the rate of change of shared stocks’ distribution (indicated by the time-of-emergence of the transboundary index) and the importance of the shifting species to a country’s total catch and revenue from the species.
-The index is calculated from:
+<!-- We created a country-specific vulnerability index (VI) to examine the risk of impacts of species’ range shifts on transboundary fisheries management.
+The VI reflects both the rate of change of shared stocks’ distribution (indicated by the time-of-emergence of the transboundary index) and the importance of the shifting species to a country’s total catch and revenue from the species.
+The VI is calculated from: -->
 
@@
 
-\begin{equation}
-\label{eq:five}
-index' = \frac{(P_c + P_v)}{ToE}
-\end{equation}
+<!-- \begin{equation} -->
+<!-- \label{eq:five} -->
+<!-- VI' = \frac{(P_c + P_v)}{ToE} -->
+<!-- \end{equation} -->
 
-where $P_c$ and $P_v$ indicate the proportion accounted for by shared species in total catch and value generated from transboundary fisheries within a country’s EEZ, respectively.
-Here, the proportion of total catch and value are represented as the average of 2005 to 2014 within individual EEZs according to the *Sea Around Us*.
-The index was then normalized to a 0 (less vulnerable) to 1 (more vulnerable) scale.
+<!-- where $P_c$ and $P_v$ indicate the proportion accounted for by shared species in total catch and value generated from transboundary fisheries within a country’s EEZ, respectively.
+Here, the proportion of total catch and value are represented as the average of 2005 to 2014 within individual EEZs according to the Sea Around Us.
+The VI was then normalized to a 0 (less vulnerable) to 100 (more vulnerable) scale.
+-->
 
-### Estimating the threat point of transboundary stock share
+### Estimating the threat point (TP) of transboundary stock share
 
 In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than operating on a non-cooperative basis [@Munro:1979dg].
-However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., "Threat Point" (TP) [@Sumaila:2013hv].
+However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., "threat point" (TP) [@Sumaila:2013hv].
 Here, we define TP as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
@@
 Second, we average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
-The first time period ($t_h$) spans 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
+The first time period ($t_h$) spans from 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
 It is assumed as the historic baseline of shared distributions for each stock.
-We then selected two future periods; early $21^{st}$ century as the average of 2021 to 2040 (*te*), and mid $21^{st}$ century as the average of 2041 to 2060 (*tm*).
+We then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 (*te*), and the mid $21^{st}$ century as the average of 2041 to 2060 (*tm*).
 We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
@@
 \label{eq:six}
-\Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 \\
-TP = \Delta{SSR}_{e,m} < \mu \pm \sigma
+\Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 \\
+TP = \Delta{SSR}_{e,m} < \Delta{SSR}_{th} \pm \sigma
 \end{split}
@@
 
+where *e* is EEZ and *s* is stock.
 Finally, a substantial change in SSR of each EEZ was determined when $\Delta{SSR_{e,m}}$ was beyond one standard deviation of the historical distribution i.e., when $SSR_f \geq (SSR_f + \sigma)$ or $SSR_f \leq (SSR_f - \sigma)$.
@@
 <!-- Moreover, we implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
-Specifically, we looked at habitat preferences, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
+Specifically, we looked at habitat associations, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
 We hypothesize that all of these factors would have significant influence in the ToE of transboundary species.
 -->
-All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@PackagegmtInterfa:2017vh] and *zoo* [@PackagezooSInfr:2019tm], *ggrepel* [@PackageggrepelAut:2020to], *zeallot* [@PackagezeallotMul:2018wy], *viridis* [@PackageviridisDef:2018us], *cowplot* [@CowplotStreamlined:2019wt], *moments* [@PackagemomensMome:2015uh] and *pgirmess* [@PackagepgirmessSp:2018um].
+All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages data.table [@Packagedatatable:2019uh], janitor [@PackagejanitorSim:2018ur], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], gmt [@PackagegmtInterfa:2017vh] and zoo [@PackagezooSInfr:2019tm], ggrepel [@PackageggrepelAut:2020to], zeallot [@PackagezeallotMul:2018wy], viridis [@PackageviridisDef:2018us], cowplot [@CowplotStreamlined:2019wt], moments [@PackagemomensMome:2015uh] and pgirmess [@PackagepgirmessSp:2018um].
 All code is available at https://github.com/jepa/TransEmergence
@@
 
-The average ToE of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years with the first stocks emerging in 2006 and until 2100 (Figure \@ref(fig:Figure1).
-In total, 5,745 stocks will emerge between 2006 and 2100, that is 63% of the total transboundary stocks considered in this study (9,132).
+The average ToE of the transboundary index (TI) across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:Figure1).
+Fish stocks with TI that first emerged from their historical variability in each EEZ was projected to be as early as 2006.
+In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100.
 About 55% (n = 3,154) of those stocks are projected to have their shifts in distributions emerged between 2006 and 2020.
@@
 After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.74% per year.
-By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:cumulative)).
+By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:Figure1)).
 
-```{r cumulative, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Cumulative increase in time of emergence by Exclusive Economic Zone and transboundary species in absolute numbers and percentage."}
+```{r Figure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone and transboundary stock.
+A) Showing results in absolute numbers.
+B) Showing the result as a percentage of the total stocks and EEZ."}
 
@@
 
-The median ToE of transboundary stocks varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, X2 = 287.23, DF = 93, p < 0.001).
-Around one fifth (19%) of the EEZs have an average time of emergence across species prior to 2020 (Figure \@ref(fig:Figure2).
-Notable exceptions where the ToE is relative late include Honduras (2071), New Zealand (2055), Ireland (2067) and Pitcairn (2070).
-The EEZs in Eastern Asia, Latin America and the Caribbean as well as Polynesia differed significantly (*p* < 0.05) from (almost) all other regions (see Table S1 for full statistics).
-Tropical latitudes also hold the majority of countries with high vulnerability to shifts in transboundary stocks due to climate change (Figure \@ref(fig:Figure2).
+The median ToE of the transboundary index of stocks varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, X2 = 287.23, DF = 93, p < 0.001).
+Around one fifth (19%) of the EEZs have an average time of emergence across stocks prior to 2020 (Figure 2).
+Notable exceptions where the ToE is relatively late include Honduras (ToE = 2071), New Zealand (ToE = 2055), Ireland (ToE = 2067) and Pitcairn (ToE = 2070).
+The EEZs in Eastern Asia, Latin America, the Caribbean and Polynesia regions differed significantly (p < 0.05) from (almost) all other regions (see Table S1 for test statistics).
+(Figure \@ref(tab:TableS1).
+
+Tropical latitudes also hold the majority of countries with the largest number of shifts in transboundary stocks due to climate change (Figure \@ref(fig:Figure2).
 Only in very few cases like Brazil and Alaska’s and Russia’s polar EEZs, no stock will emerge between 2006 and 2100.
@@
 While the selected 1 s.d.
-threshold is less conservative than a 2 s.d., the distribution of results do not dramatically change between thresholds, with a difference in ToE of about 4 years (Figure \@ref(fig:FigureS2).
+threshold is less conservative than a 2 s.d., the distribution of results do not dramatically change between thresholds, with a difference in ToE of about 4 years (Figure \@ref(fig:FigureS2)).
 
-```{r Figure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of transboundary species.
-A) Land polygon shows the vulnerability index of individual country’s fisheries management to shifts in transboundary species.
+```{r Figure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of transboundary stocks.
+A) Land polygon shows the vulnerability index of individual country’s fisheries management to shifts in transboundary stocks.
 The warmer the colour the greater the vulnerability.
-Exclusive Economic Zone polygon represents the average time of emergence across all species, with the cooler the colour the later the average time of emergence.
-The Density plot on the right shows the latitudinal number of emerging species across EEZ centroids (10 degrees latitude running mean).
+Exclusive Economic Zone polygon represents the average time of emergence across all stocks, with the cooler the colour the later the average time of emergence.
+The Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10 degrees latitude running mean).
 Aqua color = no emerging stocks.
@@
 Land polygon shows the vulnerability index.
-Exclusive Economic Zone polygon represents the average time of emergence across all species.
-Density plot on the right shows the latitudinal number of emerging species across EEZ centroids (10^o running mean).](../Figures/Fig3.png) -->
+Exclusive Economic Zone polygon represents the average time of emergence across all stocks.
+Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean).](../Figures/Fig3.png) -->
 
-Our analysis suggests that most countries responsible for 75% (n = 47) of the yearly revenue generated from transboundary species between 2005 and 2010 will see their stocks emerge in average before 2040 (Figure \@ref(fig:Figure3).
-In some cases like Indonesia, Thailand and the United States, stocks will emerge ass soon as 2030.
-These 47 countries are also among those with the largest number of emerging stocks.
-Notably, nine of the top ten countries earning the most revenue from transboundary species, share a border.
+Our analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary species between 2005 and 2010 will see their distribution range shifts beyond historical variabilities before 2040 (Figure \@ref(fig:Figure3)).
+Particularly, nine of the top ten countries earning the most revenue from transboundary species, share a border.
+In some cases, like Indonesia, Thailand and the United States, stocks’ range shifts swill emerge by around 2030.
+The 47 countries that make up the bulk of the global fisheries revenues also have more stocks with range shifts emerging from historical variabilities.
 
-```{r Figure3, eval = T,  echo = F, fig.width=5, fig.height=6,fig.pos= "H", fig.cap =  "Average time of emergence of transboundary species per fishing entity, aggregated by colour according to region.
+```{r Figure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average time of emergence of transboundary species per fishing entity, aggregated by colour according to region.
 Fishing revenue on a logarithmic scale.
-Showing country names for the top 90th percentile.
-Countries marked with an asterisk have at least one neighboring border in the group."}
+Showing country names for the top 90th percentile of annual fishing revenue."}
 
@@
 
-We compared species ToE within EEZs by species habitat preference according to FishBase's classification (Figure \@ref(fig:Figure4).
-We found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
-Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table @\ref(tab:TableS2).
-In average, species associated with pelagic oceanic such as tunas (*Thunnus sp.*) and bathy-pelagic habitats like blue whiting (*Micromesistius poutassou*) as well as species associated to corals (e.g., Bigeye snapper	*Lutjanus	lutjanus*) will have an earliest time of emergence (Figure \@ref(fig:Figure4)).
-Only species associated to bathydemersal habitats such as Alaska plaice	*Pleuronectes	quadrituberculatus* and other habitats will have an average ToE later than 2025 (Figure \@ref{fig:Figure4)).
+We compared species time of emergence within EEZs by species’ habitat association (Figure \@ref(fig:Figure4)).
+Each species is assigned to be associated to one of the xx habitat types based on information available from FishBase (for fish, ref) or SeaLifeBase (for invertebrates, ref).
+We found significant differences in the time of emergence of stocks; range shifts by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
+Specifically, time of emergence of range shifts of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:TableS2)).
+On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (Figure \@ref(fig:Figure4)).
+Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:Figure4)).
 
-```{r Figure4, eval = T,  echo = F, fig.width=14, fig.height=14,fig.pos= "H", fig.cap =  "Regional time of emergence per habitat preference.
-Regions according to the United Nations and habitat preference to FishBase.
+```{r Figure4, eval = T,  echo = F, fig.width=14, fig.height=14,fig.pos= "H", fig.cap =  "Regional time of emergence per habitat association.
+Regions according to the United Nations and habitat association to FishBase.
 Whiskers represent 1.5* interquartile range.
@@
 <!-- ![Figure 4.
-Regional time of emergence per habitat preference.
-Regions according to the United Nations and habitat preference to FishBase.
+Regional time of emergence per habitat association.
+Regions according to the United Nations and habitat association to FishBase.
 Whiskers represent 1.5* interquartile range.
@@
 ToE of SST), human drivers (e.g.
-catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary species (Table \ref{tab:TableS3)).
-Results from the multiple linear regression analysis show a significant relation between ToE of transboundary species and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref{tab:Table1)).
+catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary species (Table \ref(tab:TableS3)).
+Results from the multiple linear regression analysis show a significant relation between ToE of transboundary species and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref(tab:Table1)).
 -->
@@
 
-We estimated the changes in the SSR of transboundary stocks by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) highlighting changes below a neighbouring country's TP.
-Our results suggest that 87% (n = 238) of the world EEZs will see a change in SSR below the country's TP by 2030 (Figure \@ref{fig:Figure5}A).
-Globally, most EEZs will average a change lower than 25% by the early $21^{st}$ century which is expected to slightly increase in intensity towards the mid $21^{st}$ century (Figures \@ref{fig:FigureS4} and S5).
-However, this result masks larger changes in SSR for specific stocks and EEZs.
-For example, Guatemala’s SSR of *Panulirus gracilis* shared with Mexico is projected to go from a historical 12% to 26%, a change of over 50%.
-By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR above the country's TP historical distribution (e.g., $\Delta{SSR} > \mu \pm \sigma$) will be relatively large (global mean of 69% of stocks in an EEZ).
-For 17 EEZs, 100% of their analyzed transboundary stocks will shift below the, while only six EEZs will have less than 25% stocks shifting below the TP (Figure \@ref{fig:Figure5}A-Polygons).
-In terms of direction of change, regions where EEZs are relatively small and have multiple borders (e.g.
-The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Figure \@ref{fig:Figure1}B).
-Towards the mid $21^{st}$ century, changes in SSR per EEZ will not be substantial (Figure \@ref{fig:FigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref{fig:FigureS5)).
+We estimated the changes in the SSR of transboundary stocks by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) and highlighting changes below a neighbouring country’s TP.
+Our results suggest that 87% (n = 238) of the world EEZs will see a change in SSR below the country’s TP by 2030 (Figure \@ref(fig:Figure5)A).
+Globally, average changes in SSR of transboundary stocks in most EEZs are projected to be lower than 25% by the early $21^{st}$ century.
+The changes in SSR is expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:FigureS4) and \@ref(fig:FigureS5)).
+However, changes in SST for specific stocks and in some EEZs are projected to be exceptionally high.
+For example, Guatemala is sharing *Panulirus gracilis* stock with Mexico, and Guatemala’s share is projected to increase by >50% from 12% to 26% by the early $21^{st}$ century.
+By the early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR above the country’s TP historical distribution  (e.g., $\Delta{SSR} > \mu \pm \sigma$) will be relatively large (global mean of 69% of stocks in an EEZ).
 
-```{r Figure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial stocks that will see a shift.
-Panel B zooms in to specific areas shown in grey boxes of A."}
+```{r Figure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in Specie-Share Ratio by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average gain of transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those that are gaining shares (arrowhead).
+Land polygons represent the percentage of initial stocks that will see a shift.
+Panel B zooms in to specific areas shown within the grey dashed line boundaries in A."}
 
@@
 In this paper we endeavor to look at the timing and intensity of changes in the sharing of transboundary species.
-Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world indicating climate change is already affecting international fisheries management (Figure \@ref{fig:Figure1)).
-Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require fast responses from transboundary fisheries management (Figure \@ref{fig:Figure2} and \@ref{fig:Figure3)).
-Finally, we show that in many instances around the world, the incoming years will see many transboundary stocks shifting their shared proportion below acceptable levels for sharing nations potentially becoming future sources of conflict (Figure \@ref{fig:Figure5))
+Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
+Such early emergences of range shifts indicate that the immediate challenges climate change is posing on international fisheries management (Figure \@ref(fig:Figure1)).
+Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require fast responses from transboundary fisheries management \@ref(fig:Figure2) and \@ref(fig:Figure3)).
+Finally, we show that in many instances around the world, the coming years will see that the shares of many transboundary stocks by countries will drop below level that conduce cooperation between countries in managing the shared stocks (Figure \@ref(fig:Figure5)).
 
@@
 
-Our results suggest that in most EEZs, more than half of transboundary species emerged in the early 2000s (Figure \@ref{fig:Figure1)).
-This is in alignment with previous research pointing to changes in marine catch  composition since the 1970s as evidence of climate change [@Cheung:2013gk, @Poloczanska:2016kk].
-For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range, reaching Washington state waters where a new fishery rapidly developed [@Pinsky:2014by; @Zeidberg:2007gs].
-In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, led to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij].
-The large number of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
+Our results suggest that in most EEZs, more than half of their transboundary stocks emerged in the early 2000s (Figure \@ref(fig:Figure1)).
+Our projections in the emergence of distribution shifts of transboundary stocks corroborate with previous studies that have detected and attributed changes in marine catch composition since the 1970s [@Cheung:2013gk; @Poloczanska:2016kk].
+For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (USA).
+A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs].
+In the North East Atlantic, the range expansion of Atlantic mackerel (*Scomber scombrus*) into Icelandic waters in 2010 was followed by the increase in fishery quota **xxx** without consultation with **xxxx.** Such changes in quota resulted in disputes Iceland and Faroes Islands with multiple fishing nations co-managing the Atlantic mackerel stocks [@Spijkers:2017ij].
+
+The projected range shifts of large number of transboundary species emerging since around the 2000s period  can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine stocks [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
 It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 23% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while $O_2$ will have doubled under a high greenhouse gas emission scenario [@Rodgers:2015jq].
 
-Regional differences in ToE are likely the result of environmental factors, species vulnerability, and the characteristics of political boundaries.
+Regional differences in time of emergence are likely the result of environmental factors, species vulnerability, and the characteristics of political boundaries.
 For example, our findings underscore previous findings suggesting that tropical species are already on the move.
-The tropics are among the most species-rich biome [@Reygondeau:2019uh] and have potentially experienced some of the greatest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
+The tropics are among the most species-rich biome [@Reygondeau:2019uh] and have potentially experienced some of the highest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
 In the tropics, marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz].
-Consequently, as a result of climate change, species have to move towards deeper or sub-tropical waters [@Poloczanska:2016kk] although recent studies suggest that in some regions species are shifting towards the equator and to shallow waters [@Clarke:2020ep].
-The premise of a shift already happening is supported by observational evidence of tropical species appearing in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis showing a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
+Consequently, as a result of climate change, species have to move towards deeper or sub-tropical waters [@Poloczanska:2016kk] although recent studies suggest that in some regions species are shifting towards the equator and to shallow waters to follow local temperature gradients [@Clarke:2020ep].
+Moreover, tropical species are expanding their range in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis shows a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
 In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs and many borders that match peaks of emerging stocks.
-Specifically, it seems like South East Asia, the Caribbean, and the Mediterranean harbour substantial increases in emerging stocks (Figure \@ref{fig:Figure2)).
 
@@
 
-We have identified hotspot regions of climate risk for transboundary fisheries management (Figure \@ref{fig:Figure2)).
+Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management (Figure \@ref(fig:Figure2)).
 The premise of our vulnerability index is that, from a management perspective, countries whose shifting transboundary stocks provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
-Thus, climate risk hotspots for transboundary management happen where EEZs will see an early ToE and where transboundary stocks are important in terms of fisheries catches and revenues (Figure \@ref{fig:Figure2)).
-For example, Polynesia and Micronesia are the only regions where the average ToE across all stocks and EEZs already happened; 2019 or is about too 2021, respectively (Figure \@ref{fig:Figure5)).
+Thus, climate risk hotspots for transboundary management happen where EEZs will see an early ToE and where transboundary stocks are important in terms of fisheries catches and revenues (Figure \@ref(fig:Figure2)).
+For example, Polynesia and Micronesia are the only regions where the average ToE across all stocks and EEZs already happened; 2019 or is about too 2021, respectively (Figure \@ref(fig:Figure5)).
 Countries in these regions are highly dependent on pelagic-oceanic (e.g.,tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
-For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \@ref{fig:Figure5)).
+For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \@ref(fig:Figure5)).
 In this case, policy has to work against the clock and under potentially high societal pressure to address an issue that could be impacting important fisheries [@Burden:2019ij].
-Currently, these fisheries operate under the Vessel Day Scheme - which takes into account shifts in stocks and catches making it a good example where a management system has been set up that is somewhat anticipatory [@Aqorau:2018bh].
+Currently, fisheries targeting shared stocks like tunas in the South Pacific Islands region operate under the Vessel Day Scheme - which takes into account shifts in stocks and catches making it a good example where a management system has been set up that is somewhat anticipatory [@Aqorau:2018bh].
 However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to stocks sustainability.
-While it gives countries more time to prepare for substantial changes, stocks with relatively small overall participation on catches or revenue can still be a source of conflict as portrayed by the Atlantic mackerel case [@Spijkers:2017ij, Chapter 1].
+While it gives countries more time to prepare for substantial changes, stocks with relatively small overall participation on catches or revenue can destabilize cooperative shared stock management agreement as shown by the Atlantic mackerel example [@Spijkers:2017ij, Chapter 1].
 
-### Changes in species-share ratio will be ubiquitous
+### Changes in stocks-share ratio will be ubiquitous
 
-By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Figure \@ref{fig:Figure5)).
+Changes in SSR will be seen across the world, suggesting that climate change impacts on transboundary stock management is a global problem (Figure \@ref(fig:Figure5)).
 Poleward shifts in SSR are expected in regions of the world with relatively large EEZs and few boundaries, like the North East Pacific.
-However, despite observations and modelling work indicating that, generally, climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], the Eastern Tropical Pacific will see EEZs from Mexico to Ecuador losing SSR to their southern neighbors.
+However, despite observations and modelling work indicating that, generally, climate change is shifting stocks distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], the Eastern Tropical Pacific will see EEZs from Mexico to Ecuador losing SSR to their southern neighbors.
 Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems.
-They are consistent with regional modeling studies suggesting that by mid $21^{st}$ century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
-Moreover, regions characterized by multiple political boundaries, like the Caribbean and the North East Atlantic, will see a complex network of changes in SSR with no clear pattern (Figure \@ref{fig:Figure5}B).
+They are consistent with regional modeling studies suggesting that by mid $21^{st}$ century, marine stocks will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
+Moreover, regions characterized by multiple political boundaries, like the Caribbean and the North East Atlantic, will see a complex network of changes in SSR with no clear pattern (Figure \@ref(fig:Figure5)B).
 These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] supporting the implementation of joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
-However, game theory suggests that the greater the number of negotiating parties , the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared species.
+However, game theory suggests that the greater the number of negotiating parties , the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.
 
@@
 The use of multiple ESMs can capture the structural uncertainty of the models [@Frolicher:2016dg; @Lotze:2019ce], however, exercises looking at changes in fish biomass with multiple models show an overall agreement between models in terms of direction of change but variable magnitude.
-Using the GFDL-ESM2M results in conservative estimates as its projections, for example in sea surface temperature, are overall less extreme than in other CMIP5-type Earth system models [@Bopp:2013fg].
-Addressing the uncertainty in future projections is key to advance marine governance for species on the move [@Pinsky:2018cb], thus, further research that includes multiple ESMs could provide more accurate results in terms of magnitude but will likely follow the same trend showed here [@Lotze:2019ce].
-Secondly, here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations (Chapter One).
-We consider that each species shared by a set of neighbouring EEZs is a shared stock for a total of 9,132 shared stocks.
-Thus, our analysis does not consider a more detailed organizational level such as biologically-meaningful sub-populations within an EEZ due to limited biological and spatial-specific data on stocks sub-populations.
-Moreover, in many EEZs of the world fisheries are often managed at the species level (Chapter One) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
-However, reproducing our analysis regionally, where spatially-explicit stock data is available would allow to produce more certain results in changes in both ToE and SSR of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
+Further research that includes multiple ESMs and species distribution models could elucidate a broader range structural uncertainty of the climate, fish and fisheries models [@Lotze:2019ce].
+Secondly, due to limited biological and spatial-specific data on stocks sub-populations, our analysis uses political boundaries (EEZ) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
+On the other hand, in many EEZs, fisheries are often managed at the species level (Chapter One) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
+However, reproducing our analysis regionally, where spatially-explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
 
@@
 Here, we have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
-We have identified the transboundary species that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
+We have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
 Our results have direct implications for ocean governance as the data provided  can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by  climate change to international fisheries management.
@@
 
-```{r FigureS1, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
+```{r FigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
 
@@
 
-```{r FigureS2, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+```{r FigureS2, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
 The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel).
@@
 
-```{r FigureS3, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
+```{r FigureS3, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
 
@@
 
-```{r FigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
+```{r FigureS4, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
 
@@
 
-```{r FigureS5, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
+```{r FigureS5, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
 Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial species that will see a shift.
+Polygons represent the percentage of initial stocks that will see a shift.
 Panel B zooms in to specific areas shown in grey boxes of A."}
@@
 grid.raster(img)
+
+```
+
+```{r TableS1, eval = F, echo = F}
+
+read.csv("~/GitHub/TransEmergence/Tables/kw_regional.csv") %>% 
+  select(-statistic,
+         "habitat association comparrison" = comparrison,
+         "Significan level" = signif.level,
+         "Observation difference" = dif.com.obs.dif,
+         "Critical difference" = dif.com.critical.dif,
+          "K-W Difference" = dif.com.difference) %>% 
+  kable("latex",
+        booktabs = T,
+        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  ) %>% 
+  kable_styling(full_width = F,
+                position = "center")
+
+```
+
+```{r TableS2, eval = F, echo = F}
+
+read.csv("~/GitHub/TransEmergence/Tables/kw_habitat.csv") %>% 
+  select(-statistic,
+         "Sub region comparrison" = comparrison,
+         "Significan level" = signif.level,
+         "Observation difference" = dif.com.obs.dif,
+         "Critical difference" = dif.com.critical.dif,
+          "K-W Difference" = dif.com.difference) %>% 
+  kable("latex",
+        booktabs = T,
+        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  ) %>% 
+  kable_styling(full_width = F,
+                position = "center")
 

### 2020-11-20 Scripts/Supplements.Rmd
updates to manuscript and supplements

--- 2020-11-16 Scripts/Manuscript.Rmd
@@
 ---
-title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change
-subtitle: Challenges to transboundary fisheries under climate change
-author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^'
-date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern.'
 output:
-  # word_document: 
-    # reference_docx: ../References/template.docx
-  pdf_document: default
-bibliography: ../References/ToE_Ref.bib
-csl: ../References/science.csl
-header-includes:
-- \usepackage{amsmath}
-- \usepackage{float} # Use the 'float' package
-- \floatplacement{figure}{H} # Make every figure with caption = h
+  bookdown::pdf_book:
+    toc : no
+    keep_tex: no
+    number_sections: yes
+header-includes: #allows you to add in your own Latex packages
+- \usepackage{float} #use the 'float' package
+- \floatplacement{figure}{H} #make every figure with caption = h
+- \floatplacement{table}{H}
+- \usepackage{caption}
+- \captionsetup[figure]{name = Figure A3}
+- \captionsetup[table]{name = Table A3}
 ---
 
-<!-- # Instructions  -->
-
-<!-- Research articles should present a major advance and can be short or long (maximum 15,000 words) format.
-Submissions must include an abstract (up to 150 words), an introduction, and sections with brief informative subheadings.
-You may include up to ten figures and/or tables and about 60 references.
-Supplementary materials should be limited to information that is not essential for the general understanding of the research presented in the main text and can include data sets, figures, tables, videos, or audio files.
-You may include up to a total of ten figures and/or tables (combined) throughout the supplemental text.
--->
-
-```{r setup, eval = T, echo = F, warning = F, message = F, results = 'hide'}
-
+```{r setup, eval = T, echo=F, warning=F,message=F, results='hide'}
 library(MyFunctions)
-
-#### Project's Library
+#### Library ####
 packages <- c(
-  "tidyverse" ,
+  "dplyr", # Data manipulation
+  "knitr",
+  "kableExtra",
   "png",
-  "grid"
-  # "here", # for dbem_import `here()`
-  # "data.table", #dbem_import `fread()`
-  # "readxl", # for reading excell files
-  # "janitor", # for clearing names
-  # "geosphere", # estimate distances between points `distm()`
-  # "ggrepel",
-  # "zoo", # for average mean
-  # "parallel", # for mclapply,
-  # "sf", # for mapping
-  # "st", # for mapping
-  # "rgdal", #Spatial analysis
-  # "tools", #Spatial analysis 
-  # "zeallot", # for Juanito's map
-  # "gmt", # for estimating distances between points
-  # "viridis"
+  "grid",
+  "janitor"
 )
-
 my_lib(packages)
-
 ```
-
-<!-- * corresponding author: j.palacios@oceans.ubc.ca -->
-
-<!-- # Abstract (150  words) -->
-
-<!-- An opening sentence that states the question/problem addressed by the research AND Enough background content to give context to the study AND A brief statement of primary results AND A short concluding sentence.
--->
-
-<!-- Climate change shifts in marine species distributions is affecting the current frameworks emplaced to manage species shared fisheries.
-Here, we rely on a modeling framework with multiple ensemble members to identify the time frame by which shared stocks distribution will change from historical ranges, and the change in the proportion of the distrirbution that each EEZ has under climate change.
-Our results suggest that 60% of shared species have emerged since 2006 in 87% of the world EEZs.
-By 2075 all EEZs will have an emerging species and by 2097 and all species will have emerged in at least one EEZ.
-Average changes in the ratio of species-share between neighboring EEZs will fluctuate between 1 and 10% with some exceeding 25% by 2030 relative to today’s ratio.
-Our results suggest that most changes will happen before 2030 urging policy to be re-formulated to account for the possible implications of these changes.
--->
-
-## Introduction
-
-Over the last century, human activities have fundamentally changed the physico-chemical composition of the ocean, substantially warming the waters, decreasing the pH and reducing oxygen concentration [@IPCC:2019tn].
-Throughout the ocean, marine species are distributed according to preferences in given environmental parameters (e.g., salinity, temperature, oxygen) [@Hutchinson:1957nt; @Nelson:2016kk].
-As a result of climate change, many marine species have changed their distribution towards higher latitude, deeper water or following local environmental gradients [@Poloczanska:2016kk].
-Biogeography of marine species is projected to continue to shift as ocean conditions continue to change in the 21st century under climate change [@Cheung:2010dt].
-Climate change is expected to impact fisheries production, compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].
-The projected risks of impacts can be reduced through improving the effectiveness of current fisheries management [@Gaines:2018sg; @Cheung:2018dh], including managing fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
-
-The concept of shared stocks was developed following the ratification of  the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
-As defined by the Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (*i*) transboundary, those that cross neighboring EEZs; (*ii*) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions [@Munro:2004th].
-While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species, yielding an average 48.5 million tonnes of catch and USD 78 billion in fishing revenue, per year, between 2005 and 2010 [@PalaciosAbrantes:2020tv].
-
-The effectiveness of managing fisheries targeting shared stocks is challenged by the shifts in marine species distribution under climate change [@Pinsky:2018cb; @Pinsky:wq].
-In many cases, catch or fishing effort quota on shared stocks are based on historical records, [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change in the fish stocks and fisheries [@PalaciosAbrantes:2020tv].
-Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by worsening climate change [@Pinsky:2018cb].
-With future shifts in species distributions under climate change, the challenges posed on shared stocks fisheries management will continue to increase in the $21^{st}$ century.
-Thus, there is a need to understand the time when climate change will affect the sharing distribution of these species as well as the intensity in order to prepare ocean governance to species on the move [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
-
-We propose that adapting the concepts of "Time of Emergence" (ToE) and “Thread Point” (TP) can help address some of the management challenges imposed by shifting species.
-Roughly, ToE is defined as the moment in time when an anthropogenic change signal (e.g., future trend) rises above the background noise of natural variability (e.g., historical variation) [@IPCC:2019tn].
-This concept has been applied to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
-The premise behind ToE is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background of natural climate variability [@Hawkins:2012hc].
-The concept of thread point (TP) comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
-Game theory has been widely used to investigate and manage shared stocks [@Sumaila:2013hv].
-Specifically, Sumaila et al., *in press* applied this concept to understand the impacts of climate change to the economics of shared species between Canada and the United States.
-
-Here, we estimate the ToE and TP of the share distribution of transboundary stocks worldwide.
-We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected shared species [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emission scenario RCP8.5 (see Methods).
-We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., we did not include the high seas).
-Our results suggest that many shared stocks have already ‘emerged’ and overpassed their share TP in most neighboring EEZs posing a threat to the sustainability of shared stocks and the resilience of international treaties.
-
-## Materials and methods
-
-### Databases and species selection
-
-This analysis includes 633 marine exploited transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd; @PalaciosAbrantes:2020tv] for a total of 9,132 transboundary stocks.
-We defined the boundaries of the world’s EEZs using the *Sea Around Us* spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:FigureS1)) including island territories.
-We determined the intersections between polygons using the *R*-statistical software package *sf* [@PackagesfSimpleF:2018vp].
-Each EEZ was categorized by the geopolitical regions according to the United Nations and by natural biome [@Reygondeau:2019uh].
-The habitat preference of each species was determined based on information available from FishBase (http://www.fishbase.org).
-For each species and EEZ, we used the *Sea Around Us* data to estimate catch and fishing revenue from fishing activities within global EEZs.
-We report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
-The monthly average US Consumer Price Index (CPI) according to the U.S.
-Bureau of Labor Statistic (https://www.bls.gov/cpi/) was used to standardize all monetary values to 2019 real USD.
-
-### Projecting species distributions under climate change
-
-We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
-The DBEM  represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
-For pelagic species, the model uses environmental variables at the surface (e.g., sea surface temperature) whereas demersal species distribution are driven by ocean bottom variables (e.g., bottom temperature).
-This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (see Cheung *et al.* [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
-Recent work has demonstrated that, at the global scale, DBEM is able to reproduce catch changes similar in magnitude to those detected in observed catch time series in the
-world’s EEZs (Cheung et al. in revision).
-
-We forced the DBEM with environmental data from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
-The GFDL-ESM2M is run under historical forcing until 2005 and follows the ‘high’ climate change emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
-In GFDL-ESM2M, atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a species’ distribution, we have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
-We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
-Thus, for our experiment, each of the ten GFDL-ESM2M-ensembles were started from infinitesimally small differences in Earth system initial conditions resulting in a unique atmosphere and ocean state at each point in time.
-This approach allows us to estimate the natural internal variability (e.g., background noise) and isolate the forced climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
-
-### Calculating an index of transboundary range shift
-
-We developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
-This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref{fig:FigureS2}).
-The centroid of a transboundary stock was determined by the average latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
-Therefore,
-
-<!-- $lat_{ts} = \mu (lat_{per})$ -->
-
-<!-- $lon_{ts} = \mu (lon_{per})$ -->
-
-\begin{equation}
-\begin{split}
-\label{eq:one}
-lat_{ts} = \mu (lat_{per}) \\
-lon_{ts} = \mu (lon_{per})
-\end{split}
-\end{equation}
-
-where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
-To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top $10^{th}$ percentile (e.g., *per* = 10%).
-We ran a sensitivity analysis (e.g.
-*per* = 20, 50, 90) to relax this parameter (Figure \@ref{fig:FigureS2}).
-The centroid of a EEZ was estimated using the *st* package in *R* (version 3.2.1) (Figure \@ref{fig:FigureS1}).
-For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in *R*;
-
-<!-- $$D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))$$ -->
-
-\begin{equation}
-\label{eq:two}
-D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
-\end{equation}
-
-where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively and $lon_{eez}$ and $lon_{ts}$ are the respective longitudes.
-
-Then, for each year we calculated the transboundary index as follows:
-
-<!-- $$TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2$$ -->
-
-\begin{equation}
-\label{eq:three}
-TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
-\end{equation}
-
-where $D_A$ and $D_B$ represent the distance between the distribution centroids of a species and the centroid of the pair of neighbouring EEZs (A and B) sharing the species for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
-Thus, a higher TI means that a shared species becomes more or less abundant relative to its neighbouring EEZs that is beyond what the fisheries experienced historically.
-From the perspective of a state’s fisheries, higher TI indicates an increase in the challenges to the effective management and sustainability of the shared species’ fisheries.
-In contrast, a smaller TI suggests that the sharing of the species between neighbouring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the shared species’ fisheries.
-The index was smoothed to a 10 year average to reduce interannual variability.
-
-#### Calculating the Time-of-Emergence of the transboundary index
-
-Knowing the point in time (e.g., year) at which the distribution of a shared stock will diverge from its natural internal variability is important in informing the lead-time to which climate adaptation in transboundary fisheries management needs to be implemented [@Link:2010ei].
-Thus, we define ToE as the time when the stock shared between neighbouring EEZs will exceed the historical natural internal variability as follows:
-
-<!-- $EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$ -->
-<!-- $ToE = t when Emergence index[t] > 1$ -->
-
-\begin{equation}
-\begin{split}
-\label{eq:four}
-EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}} \\
-ToE = t when Emergence index[t] > 1
-\end{split}
-\end{equation}
-
-where $\mu_{TI}$ is the yearly mean TI ($\mu$) across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005).
-We set a threshold of TI above which we consider a shared stock to have emerged from the historical internal variability of the TI.
-The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \ref{eq:four}, representing a probability of 68% that the transboundary index has emerged from historical variability).
-We tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
-
-#### Computing a transboundary fisheries vulnerability index
-
-We created a country-specific index to examine the risk of impacts of species’ range shifts on transboundary fisheries management.
-The index reflects both the rate of change of shared stocks’ distribution (indicated by the time-of-emergence of the transboundary index) and the importance of the shifting species to a country’s total catch and revenue from the species.
-The index is calculated from:
-
-<!-- $$ index' = (P_c + P_v)/ToE $$ -->
-
-\begin{equation}
-\label{eq:five}
-index' = (P_c + P_v)/ToE
-\end{equation}
-
-where *P_c* and *P_v* indicate the proportion accounted for by shared species in total catch and value generated from transboundary fisheries within a country’s EEZ, respectively.
-Here, the proportion of total catch and value are represented as the average of 2005 to 2014 within individual EEZs according to the *Sea Around Us*.
-The index was then normalized to a 0 (less vulnerable) to 1 (more vulnerable) scale.
-
-### Estimating the threat point of transboundary stock share
-
-In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than operating on a non-cooperative basis [@Munro:1979dg].
-However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., *Threat Point* (TP) [@Sumaila:2013hv].
-Here, we define TP as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-Any proportion below the defined threat point of a country would result in a deflection from the joint management of the shared stock.
-
-First we estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
-We did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
-Second, we average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
-The first time period ($t_h$) spans 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
-It is assumed as the historic baseline of shared distributions for each stock.
-We then selected two future periods; early $21^{st}$ century as the average of 2021 to 2040 (*te*), and mid $21^{st}$ century as the average of 2041 to 2060 (*tm*).
-We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
-The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members.
-Third, we estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_h$) following Palacios-Abrantes et al.
-(*in press*),
-
-<!-- $$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$ -->
-<!-- $TP = \Delta{SSR}_{e,m} < \mu \pm \sigma$ -->
-
-\begin{equation}
-\begin{split}
-\label{eq:six}
-\Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 \\
-TP = \Delta{SSR}_{e,m} < \mu \pm \sigma
-\end{split}
-\end{equation}
-
-Finally, a substantial change in SSR of each EEZ was determined when $\Delta{SSR_{e,m}}$ was beyond one standard deviation of the historical distribution i.e., when $SSR_f \geq (SSR_f + \sigma)$ or $SSR_f \leq (SSR_f - \sigma)$.
-In other words, when the future SSR is below or above a neighbouring EEZ TP.
-
-### Statistical analysis
-
-We tested the results for normality (e.g.
-skewness = 1.02, kurtosis = 2.83) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
-Specifically, we tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of transboundary stocks.
-For both cases, our null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
-<!-- Moreover, we implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
-Specifically, we looked at habitat preferences, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
-We hypothesize that all of these factors would have significant influence in the ToE of transboundary species.
--->
-All analyses were run using the statistical software *R version 3.5.2 (2018-12-20; Eggshell Igloo)* with the packages *data.table* [@Packagedatatable:2019uh], *janitor* [@PackagejanitorSim:2018ur], *rfishbase* [@PackagerfishbaseR:2019th], *sf* [@PackagesfSimpleF:2018vp], *sp* [@PackagespClasses:2019tn], *tidiverse* [@PackagetidyverseE:2017vq], *tidytext* [@PackagetidytextTe:2019vk], *gmt* [@PackagegmtInterfa:2017vh] and *zoo* [@PackagezooSInfr:2019tm], *ggrepel* [@PackageggrepelAut:2020to], *zeallot* [@PackagezeallotMul:2018wy], *viridis* [@PackageviridisDef:2018us], *cowplot* [@CowplotStreamlined:2019wt], *moments* [@PackagemomensMome:2015uh] and *pgirmess* [@PackagepgirmessSp:2018um].
-All code is available at https://github.com/jepa/TransEmergence
-
-## Results
-
-```{r Load_Results, eval = TRUE,  echo = FALSE, include = FALSE}
-
-prop_r <- read_csv("/Volumes/HALL2000/Data/TransEmergence/Results/proportion_change_2005.csv") %>% 
-  gather("time_step","change",per_early,per_mid)
-
-```
-
-### Time of Emergence of transboundary species
-
-The average ToE of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years with the first stocks emerging in 2006 and until 2100 (Figure \@ref(fig:Figure1).
-In total, 5,745 stocks will emerge between 2006 and 2100, that is 63% of the total transboundary stocks considered in this study (9,132).
-About 55% (n = 3,154) of those stocks are projected to have their shifts in distributions emerged between 2006 and 2020.
-Also, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 (e.g.
-*Engraulis ringens* shared by Chile and Peru emerged in 2010).
-Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century and are not expected to stop there (Figure \@ref(fig:Figure1)).
-The number of EEZs with emerging stocks is projected to increase steeply from 87% to 96% during the period of 2006 to 2020.
-After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.74% per year.
-By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:cumulative)).
-
-```{r cumulative, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Cumulative increase in time of emergence by Exclusive Economic Zone and transboundary species in absolute numbers and percentage."}
-
-img <- readPNG("../Figures/Fig1.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure 1. Cumulative increase in time of emergence by Exclusive Economic Zone and transboundary species in absolute numbers and percentage.](../Figures/Fig2.png) -->
-
-The median ToE of transboundary stocks varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, X2 = 287.23, DF = 93, p < 0.001).
-Around one fifth (19%) of the EEZs have an average time of emergence across species prior to 2020 (Figure \@ref{fig:Figure2}).
-Notable exceptions where the ToE is relative late include Honduras (2071), New Zealand (2055), Ireland (2067) and Pitcairn (2070).
-The EEZs in Eastern Asia, Latin America and the Caribbean as well as Polynesia differed significantly (*p* < 0.05) from (almost) all other regions (see Table S1 for full statistics).
-Tropical latitudes also hold the majority of countries with high vulnerability to shifts in transboundary stocks due to climate change (Figure \@ref{fig:Figure2}).
-Only in very few cases like Brazil and Alaska’s and Russia’s polar EEZs, no stock will emerge between 2006 and 2100.
-From the top 10 more vulnerable countries, seven are located in tropical waters (Tokelau 100, the	Marshall Islands 92, Peru 88, Kiribati 80, Tuvalu 80, Papua New Guinea 78, and the Bahamas 74) while only three are outside the tropics (Estonia 86, Finland 82 and Sweden 78).
-We set an arbitrary threshold of  1 s.d.
-(67% confidence) to determine the ToE (see Methods).
-We tested the sensitivity of our results using a larger threshold (i.e.,  2 s.d for a 95% confidence).
-While the selected 1 s.d.
-threshold is less conservative than a 2 s.d., the distribution of results do not dramatically change between thresholds, with a difference in ToE of about 4 years (Figure S\@ref{fig:S2}).
-
-```{r Figure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of transboundary species.
-A) Land polygon shows the vulnerability index of individual country’s fisheries management to shifts in transboundary species.
-The warmer the colour the greater the vulnerability.
-Exclusive Economic Zone polygon represents the average time of emergence across all species, with the cooler the colour the later the average time of emergence.
-The Density plot on the right shows the latitudinal number of emerging species across EEZ centroids (10 degrees latitude running mean).
-B) Time of emergence according to the United Nations sub regions.
-Whiskers represent 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent points outside of 1.5*IQR."}
-
-img <- readPNG("../Figures/Fig2.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure 2.
-Average time of emergence of transboundary specie.
-Land polygon shows the vulnerability index.
-Exclusive Economic Zone polygon represents the average time of emergence across all species.
-Density plot on the right shows the latitudinal number of emerging species across EEZ centroids (10^o running mean).](../Figures/Fig3.png) -->
-
-Our analysis suggests that most countries responsible for 75% (n = 47) of the yearly revenue generated from transboundary species between 2005 and 2010 will see their stocks emerge in average before 2040 (\@ref{fig:Figure3}).
-In some cases like Indonesia, Thailand and the United States, stocks will emerge ass soon as 2030.
-These 47 countries are also among those with the largest number of emerging stocks.
-Notably, nine of the top ten countries earning the most revenue from transboundary species, share a border.
-
-```{r Figure3, eval = T,  echo = F, fig.width=4, fig.height=5,fig.pos= "H", fig.cap =  "Average time of emergence of transboundary species per fishing entity, aggregated by colour according to region.
-Fishing revenue on a logarithmic scale.
-Showing country names for the top 90th percentile.
-Countries marked with an asterisk have at least one neighboring border in the group."}
-
-img <- readPNG("../Figures/Fig3.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure 3.
-Average time of emergence of transboundary species per fishing entity.
-Fishing revenue in logarithmic scale.
-Showing names for the top 75th percentile.
-Countries marked with an asterisk have at least one neighboring border in the group.](../Figures/Fig4.png) -->
-
-We compared species ToE within EEZs by species habitat preference according to FishBase's classification (Figure \@ref{fig:Figure4}).
-We found significant differences in ToE by habitat preference (Kruskal-Wallis, $X^2$ = 286.48, *DF* = 93, *p* < 0.001).
-Specifically, the ToE of species with preferences for pelagic-oceanic and reef habitats was significantly different than other habitat preferences (See Table \ref{tab:TableS2}).
-In average, species associated with pelagic oceanic such as tunas (*Thunnus sp.*) and bathy-pelagic habitats like blue whiting (*Micromesistius poutassou*) as well as species associated to corals (e.g., Bigeye snapper	*Lutjanus	lutjanus*) will have an earliest time of emergence (Figure \@ref{fig:Figure4}).
-Only species associated to bathydemersal habitats such as Alaska plaice	*Pleuronectes	quadrituberculatus* and other habitats will have an average ToE later than 2025 (Figure \@ref{fig:Figure4}).
-
-```{r Figure4, eval = T,  echo = F, fig.width=14, fig.height=14,fig.pos= "H", fig.cap =  "Regional time of emergence per habitat preference.
-Regions according to the United Nations and habitat preference to FishBase.
-Whiskers represent 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers outside of 1.5*IQR."}
-
-img <- readPNG("../Figures/Fig4.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure 4.
-Regional time of emergence per habitat preference.
-Regions according to the United Nations and habitat preference to FishBase.
-Whiskers represent 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers (outside of 1.5*IQR.](../Figures/Fig5.png) -->
-
-<!-- We hypothesized that species environmental preferences (e.g.
-presence in certain biomes and habitats), climate change effects on water temperature (e.g.
-ToE of SST), human drivers (e.g.
-catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary species (Table \ref{tab:TableS3}).
-Results from the multiple linear regression analysis show a significant relation between ToE of transboundary species and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref{tab:Table1}).
--->
-
-```{r Table1, eval = F,  echo = F, message = F, warning  = F}
-
-read_csv("../Tables/mlr_result.csv") %>% 
-  knitr::kable(
-    "latex",
-    booktabs = T,
-    captio= "Multiple regression results for time of emergence of transboundary species.
-Only showing significant variables.
-The model significantly predicts time of emergence (F[18,6155] = 20.68, p < 0.001, R2 = 0.05)"
-  )
-
-```
-
-### Changes in the threat point of transboundary stocks
-
-```{r analysis_results, eval = F,  echo = FALSE, include = FALSE}
-
-# Early century
-n_eez <- prop_r %>%
-  filter(variable == "mean_eez" & 
-           !is.na(change),
-         time_step == "per_early"
-         ) %>% 
-  pull(eez_name) %>% 
-  unique() %>% 
-  length()
-
-n_eez_per <- round(n_eez/275*100,2)
-
-### Max values early
-max_values <- prop_r %>% 
-  filter(variable == "max_eez" & 
-           !is.na(change),
-         # time_step == "per_early"
-         )
-
-max_values$per_early
-
-### Average
-average_mid <- prop_r %>% 
-  filter(variable == "mean_eez" & 
-           !is.na(change)
-         )
-```
-
-We estimated the changes in the SSR of transboundary stocks by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to present (1951-2005) highlighting changes below a neighbouring country's TP.
-Our results suggest that 87% (n = 238) of the world EEZs will see a change in SSR below the country's TP by 2030 (Figure \@ref{fig:Figure5}A).
-Globally, most EEZs will average a change lower than 25% by the early $21^{st}$ century which is expected to slightly increase in intensity towards the mid $21^{st}$ century (Figures \@ref{fig:FigureS4} and S5).
-However, this result masks larger changes in SSR for specific stocks and EEZs.
-For example, Guatemala’s SSR of *Panulirus gracilis* shared with Mexico is projected to go from a historical 12% to 26%, a change of over 50%.
-By early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR above the country's TP historical distribution (e.g., $\Delta{SSR} > \mu \pm \sigma$) will be relatively large (global mean of 69% of stocks in an EEZ).
-For 17 EEZs, 100% of their analyzed transboundary stocks will shift below the, while only six EEZs will have less than 25% stocks shifting below the TP (Figure \@ref{fig:Figure5}A-Polygons).
-In terms of direction of change, regions where EEZs are relatively small and have multiple borders (e.g.
-The Caribbean) will see a complex dynamic of changes in SSR with no clear patterns (Figure \@ref{fig:Figure1}B).
-Towards the mid $21^{st}$ century, changes in SSR per EEZ will not be substantial (Figure \@ref{fig:FigureS5}) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref{fig:FigureS5}).
-
-```{r Figure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial stocks that will see a shift.
-Panel B zooms in to specific areas shown in grey boxes of A."}
-
-img <- readPNG("../Figures/Fig5.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure 1.
-Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average gain of transboundary specie share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial species that will see a shift.
-Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Fig1.png) -->
-
-## Discussion
-
-In this paper we endeavor to look at the timing and intensity of changes in the sharing of transboundary species.
-Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world indicating climate change is already affecting international fisheries management (Figure \@ref{fig:Figure1}).
-Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require fast responses from transboundary fisheries management (Figure \@ref{fig:Figure2} and \@ref{fig:Figure3}).
-Finally, we show that in many instances around the world, the incoming years will see many transboundary stocks shifting their shared proportion below acceptable levels for sharing nations potentially becoming future sources of conflict (Figure \@ref{fig:Figure5})
-
-#### High present-day climate risk of on transboundary fisheries management
-
-Our results suggest that in most EEZs, more than half of transboundary species emerged in the early 2000s (Figure \@ref{fig:Figure1}).
-This is in alignment with previous research pointing to changes in marine catch  composition since the 1970s as evidence of climate change [@Cheung:2013gk, @Poloczanska:2016kk].
-For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range, reaching Washington state waters where a new fishery rapidly developed [@Pinsky:2014by; @Zeidberg:2007gs].
-In the North East Atlantic, the emergence of Atlantic mackerel (*Scomber scombrus*) in Icelandic waters in 2010, led to years of unsettled international conflict between multiple fishing nations [@Spijkers:2017ij].
-The large number of transboundary species emerging since early 2000s identified in this study can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
-It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 23% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while $O_2$ will have doubled under a high greenhouse gas emission scenario [@Rodgers:2015jq].
-
-Regional differences in ToE are likely the result of environmental factors, species vulnerability, and the characteristics of political boundaries.
-For example, our findings underscore previous findings suggesting that tropical species are already on the move.
-The tropics are among the most species-rich biome [@Reygondeau:2019uh] and have potentially experienced some of the greatest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
-In the tropics, marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz].
-Consequently, as a result of climate change, species have to move towards deeper or sub-tropical waters [@Poloczanska:2016kk] although recent studies suggest that in some regions species are shifting towards the equator and to shallow waters [@Clarke:2020ep].
-The premise of a shift already happening is supported by observational evidence of tropical species appearing in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis showing a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
-In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs and many borders that match peaks of emerging stocks.
-Specifically, it seems like South East Asia, the Caribbean, and the Mediterranean harbour substantial increases in emerging stocks (Figure \@ref{fig:Figure2}).
-
-### Hotspot of climate risk on transboundary fisheries management
-
-We have identified hotspot regions of climate risk for transboundary fisheries management (Figure \@ref{fig:Figure2}).
-The premise of our vulnerability index is that, from a management perspective, countries whose shifting transboundary stocks provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
-Thus, climate risk hotspots for transboundary management happen where EEZs will see an early ToE and where transboundary stocks are important in terms of fisheries catches and revenues (Figure \@ref{fig:Figure2}).
-For example, Polynesia and Micronesia are the only regions where the average ToE across all stocks and EEZs already happened; 2019 or is about too 2021, respectively (Figure \@ref{fig:Figure5}).
-Countries in these regions are highly dependent on pelagic-oceanic (e.g.,tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
-For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \@ref{fig:Figure5}).
-In this case, policy has to work against the clock and under potentially high societal pressure to address an issue that could be impacting important fisheries [@Burden:2019ij].
-Currently, these fisheries operate under the Vessel Day Scheme - which takes into account shifts in stocks and catches making it a good example where a management system has been set up that is somewhat anticipatory [@Aqorau:2018bh].
-However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to stocks sustainability.
-While it gives countries more time to prepare for substantial changes, stocks with relatively small overall participation on catches or revenue can still be a source of conflict as portrayed by the Atlantic mackerel case [@Spijkers:2017ij, Chapter 1].
-
-### Changes in species-share ratio will be ubiquitous
-
-By 2030, changes in SSR will be seen across the globe, but with no clear pattern in the direction of change (Figure \@ref{fig:Figure5}).
-Poleward shifts in SSR are expected in regions of the world with relatively large EEZs and few boundaries, like the North East Pacific.
-However, despite observations and modelling work indicating that, generally, climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], the Eastern Tropical Pacific will see EEZs from Mexico to Ecuador losing SSR to their southern neighbors.
-Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems.
-They are consistent with regional modeling studies suggesting that by mid $21^{st}$ century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
-Moreover, regions characterized by multiple political boundaries, like the Caribbean and the North East Atlantic, will see a complex network of changes in SSR with no clear pattern (Figure \@ref{fig:Figure5}B).
-These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] supporting the implementation of joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
-However, game theory suggests that the greater the number of negotiating parties , the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared species.
-
-## Caveats and uncertainties
-
-There are two main sources of uncertainty in our analysis that can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
-First, our analysis relies on the combination of a single Earth system model (GFDL-ESM2M) and climate change scenario (RCP 8.5).
-The use of multiple ESMs can capture the structural uncertainty of the models [@Frolicher:2016dg; @Lotze:2019ce], however, exercises looking at changes in fish biomass with multiple models show an overall agreement between models in terms of direction of change but variable magnitude.
-Using the GFDL-ESM2M results in conservative estimates as its projections, for example in sea surface temperature, are overall less extreme than in other CMIP5-type Earth system models [@Bopp:2013fg].
-Addressing the uncertainty in future projections is key to advance marine governance for species on the move [@Pinsky:2018cb], thus, further research that includes multiple ESMs could provide more accurate results in terms of magnitude but will likely follow the same trend showed here [@Lotze:2019ce].
-Secondly, here we present the result based on 633 species that were previously identified as shared by multiple neighboring nations (Chapter One).
-We consider that each species shared by a set of neighbouring EEZs is a shared stock for a total of 9,132 shared stocks.
-Thus, our analysis does not consider a more detailed organizational level such as biologically-meaningful sub-populations within an EEZ due to limited biological and spatial-specific data on stocks sub-populations.
-Moreover, in many EEZs of the world fisheries are often managed at the species level (Chapter One) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
-However, reproducing our analysis regionally, where spatially-explicit stock data is available would allow to produce more certain results in changes in both ToE and SSR of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
-
-## Conclusion
-
-Global society has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) [@UnitedNations2018] by 2030; achieving this goal would have clear effects on multiple  others [@Singh:2017ds].
-Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@Pinsky:2018cb].
-Here, we have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
-We have identified the transboundary species that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
-Our results have direct implications for ocean governance as the data provided  can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by  climate change to international fisheries management.
-While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our  results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg; @PalaciosAbrantes:2020tv].
-To address the issues raised in this paper, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
-
-## Supplements
 
@@
 
-```{r FigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
+```{r FigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
 
@@
 
-<!-- # Recycled  text -->
-<!-- All of these results have direct implications for managing shared fisheries under climate change that can be partially addressed by the early implementation of joint-management international policies.
--->
+```{r TableS1, eval = T, echo = F}
 
-<!-- The time and number of emerging species decreases as we move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
--->
+read.csv("~/GitHub/TransEmergence/Tables/kw_regional.csv") %>% 
+  select(-statistic,
+         "Habitat preference comparrison" = comparrison,
+         "Significan level" = signif.level,
+         "Observation difference" = dif.com.obs.dif,
+         "Critical difference" = dif.com.critical.dif,
+          "K-W Difference" = dif.com.difference) %>% 
+  kable("latex",
+        booktabs = T,
+        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  ) %>% 
+  kable_styling(full_width = F,
+                position = "center")
 
-<!-- Changes in stock share ratio has contribute to past conflicts between nations with long history of collaboration including Canada and the United States over Pacific salmon and Norway and Iceland over spring spawning herring [@Miller:2013iv].
-In both cases, rules were emplaced and agreed by all parties, however, treaties were not prepared to address shifts in species distributions [@Miller:2013iv; @Song:2017va] and when the benefits felt below the desired level (e.g.
-threat point) the affected country broke off relationships.
-In the case of Canada and the US, the agreement was re-stated when a conservation fund was implemented [@Miller:2013iv].
--->
+```
 
-<!-- In cases where the species is distributed in more than 2 neighboring EEZs (e.g. *xxx*), the index is as follows: -->
+```{r TableS2, eval = T, echo = F}
 
-<!-- $$I =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h})}) |  \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{C,y}}{sd(D_{B,h})}) | \\  +(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{Cy}}{sd(D_{B,h})})/N$$ -->
+read.csv("~/GitHub/TransEmergence/Tables/kw_habitat.csv") %>% 
+  select(-statistic,
+         "Sub region comparrison" = comparrison,
+         "Significan level" = signif.level,
+         "Observation difference" = dif.com.obs.dif,
+         "Critical difference" = dif.com.critical.dif,
+          "K-W Difference" = dif.com.difference) %>% 
+  kable("latex",
+        booktabs = T,
+        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  ) %>% 
+  kable_styling(full_width = F,
+                position = "center")
 
-<!-- In other words, for two neighboring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases (see *TI* equation).
-Consequently, if the centroid shifts are within historical variability (*SNR* < 1), risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability (*SNR* > 1), the risk of dispute is higher.
-Thus, risk is positively related to the frequency of centroid shifts beyond historical variability.
--->
+```

### 2020-11-24 Scripts/Initial_Analysis.Rmd
William's comments and changes proportion estimations

@@
 
+World_eez_sf %>% 
+  filter(str_detect(eez_name,"Russia")) %>% 
+  ggplot() +
+  geom_sf() +
+  aes(fill = eez_name)
+
 # Call world map from natural earth
@@
 #### Both plots for tresh one
-
-plot <- ggplot() +
+#  Embarrasing coding Juliano!
+
+ggplot() +
+  geom_area(data = subset(plot_data, level == "stock"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            # alpha= 1,
+            size=0.5,
+            colour="black"
+  )  +
   geom_area(data = subset(plot_data, level == "eez"),
@@
             ),
-            alpha= 0.8,
+            # alpha= 0.8,
             size=0.5,
@@
   )  +
-  geom_area(data = subset(plot_data, level == "stock"),
+  # Shameslly plot it again to fix colors
+  geom_area(data = subset(plot_data, level == "stock" & my_facet == "Percentage (%)"),
             aes(
               x = as.numeric(taxon_emer),
+              # x = reorder(as.numeric(taxon_emer),level),
               y = value,
@@
             ),
-            alpha= 0.8,
+            # alpha= 1,
             size=0.5,
@@
     legend.position = "right",
-    strip.text.x = element_blank()
-  ) +
-  guides(fill=guide_legend(title= "Level")) +
+    strip.text.x = element_blank(),
+    panel.spacing = unit(3, "lines") # space between plots
+  ) +
+  guides(fill=guide_legend(title= "Level", reverse=FALSE)) +
   scale_fill_viridis(discrete = T,
@@
   ggsave("./Figures/cumu_per_freq.png",
-  width = 12,
-  height = 7,
+  width = 10,
+  height = 5,
   units = "in"
@@
 
-Eez_hist <- Eme_Res %>% 
-  group_by(eezid,eez_name = sf_eez_name) %>% 
-  left_join(Eez_centroid) %>% 
-  filter(tresh == tresh_list[i]) %>% 
-  group_by(latitude) %>% 
-  summarise(
-    number = n()
-  ) %>% 
-  group_by() %>% 
-  mutate(RMean = rollmean(x = number, 
-                          10, 
-                          align = "center", 
-                          fill = number
-  )
-  ) %>% 
-  mutate(lab="lab") %>% 
-  ggplot() +
-  geom_area(
-    aes(
-      x = latitude,
-      y = RMean,
-      fill = lab,
-      color = lab
-    )
-  ) +
-  coord_flip() +
-  scale_x_continuous("",
-    limits = c(-50,80),
-    breaks = c(-50,0,50,80)
-  ) +
-  scale_y_continuous(
-    "Emerging stocks",
-    limits = c(0, 124),
-    breaks = c(0,62,124)
-
-  ) +
-  scale_fill_viridis(option="magma", discrete = T, begin = 0.7) +
-  scale_color_viridis(option="magma", discrete = T, begin = 0.7) +
-  theme_minimal() + 
-  theme(
-    panel.background = element_rect(fill = "white", colour = "white"),
-    panel.grid.major.y = element_blank(),
-    panel.grid.minor.y = element_blank(),
-    panel.grid.major.x = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.text = element_text(size = 12, colour="black"),
-    axis.title.x = element_text(size = 12, colour="black", face = "bold"),
-    legend.position = ""
-  )
-
+# Eez_hist <- Eme_Res %>% 
+#   group_by(eezid,eez_name = sf_eez_name) %>% 
+#   left_join(Eez_centroid) %>% 
+#   filter(tresh == tresh_list[i]) %>% 
+#   group_by(latitude) %>% 
+#   summarise(
+#     number = n()
+#   ) %>% 
+#   group_by() %>% 
+#   mutate(RMean = rollmean(x = number, 
+#                           10, 
+#                           align = "center", 
+#                           fill = number
+#   )
+#   ) %>% 
+#   mutate(lab="lab") %>% 
+#   ggplot() +
+#   geom_area(
+#     aes(
+#       x = latitude,
+#       y = RMean,
+#       fill = lab,
+#       color = lab
+#     )
+#   ) +
+#   coord_flip() +
+#   scale_x_continuous("",
+#     limits = c(-50,80),
+#     breaks = c(-50,0,50,80)
+#   ) +
+#   scale_y_continuous(
+#     "Emerging stocks",
+#     limits = c(0, 124),
+#     breaks = c(0,62,124)
+#     
+#   ) +
+#   scale_fill_viridis(option="magma", discrete = T, begin = 0.7) +
+#   scale_color_viridis(option="magma", discrete = T, begin = 0.7) +
+#   theme_minimal() + 
+#   theme(
+#     panel.background = element_rect(fill = "white", colour = "white"),
+#     panel.grid.major.y = element_blank(),
+#     panel.grid.minor.y = element_blank(),
+#     panel.grid.major.x = element_blank(),
+#     panel.grid.minor.x = element_blank(),
+#     axis.text.y = element_blank(),
+#     axis.ticks.y = element_blank(),
+#     axis.text = element_text(size = 12, colour="black"),
+#     axis.title.x = element_text(size = 12, colour="black", face = "bold"),
+#     legend.position = ""
+#   )
+#   
   ## --------------- ##
@@
     annotate("text", 
-             x = c(-76,-35,6,48,88),
+             x = c(-72,-35,5,46,82),
              y = -69,
@@
     annotate("text", 
-             x = c(-79,-39,3,45,87),
+             x = c(-79,-39,2,42,81),
              y = -85,
@@
     theme(
-      legend.box.margin = margin(t = -64, l = 229),
+      legend.box.margin = margin(t = -64, l = 249),
       legend.text = element_blank(),
@@
     # ToE Latitudinal plot
-    draw_plot(Eez_hist, x = 0.873, y = 0.53, width = 0.124, height = 0.37) +
+    # draw_plot(Eez_hist, x = 0.873, y = 0.53, width = 0.124, height = 0.37) +
     # Time of Emergence Map
-    draw_plot(map, x = 0, y = 0.4, width = 0.95, height = 0.6) +
+    draw_plot(map, x = 0, y = 0.4, width = 1, height = 0.6) +
     # Regional boxplot
@@
   filter(!is.na(un_geo_region)) %>% 
-  ungroup() %>% 
-  mutate(
-    sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity))
+  ungroup()# %>%  # no more asterisk
+  # mutate(
+  #   sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity))
 
@@
       size = n_taxa,
-      # colour = un_geo_region,
+      colour = un_geo_region,
       shape = un_geo_region
@@
   )  +
-  geom_label_repel(data = subset(Buble_data, sum_value_bi >= 1),
+  # right side labels
+  geom_label_repel(data = subset(Buble_data, mean_yr > 2035 & sum_value_bi >= 1),
                   aes(
@@
                   force = 2,
-                   box.padding = 1,
-                   fill = "NA"
-
+                  box.padding = 0.2,
+                  fill = "NA",
+                  nudge_x = 2050 + subset(Buble_data, mean_yr > 2035 & sum_value_bi >= 1)$mean_yr,
+                  segment.size  = 0.2,
+                  segment.color = "grey50",
+                  direction     = "y",
+                  hjust         = 1
+  ) +
+  # right side labels
+  geom_label_repel(data = subset(Buble_data, mean_yr < 2035 & sum_value_bi >= 1),
+                  aes(
+                    x = mean_yr,
+                    y = log_value,
+                    label = sau_fishing_entity
+                  ),
+                  arrow = arrow(length = unit(0.009, "npc"),
+                                type = "closed",
+                                ends = "first"),
+                  force = 1,
+                  box.padding = 0.2,
+                  fill = "NA",
+                  nudge_x = 2010 - subset(Buble_data, mean_yr < 2035 & sum_value_bi >= 1)$mean_yr,
+                  segment.size  = 0.2,
+                  segment.color = "grey50",
+                  direction     = "y",
+                  hjust         = 1
   ) +
@@
   ) +
+  scale_shape("Region") +
   scale_color_viridis(name="Region",
@@
   ) +
-  scale_size("Number of\nemerging taxa") +
+  scale_size("Number of\nemerging stock") +
   ggsave("./Figures/bubble_plot.png",
@@
 Prop_results %>% 
-  filter(variable == "n_per") %>% 
+  filter(variable == "n_spp") %>% 
   group_by() %>% 
@@
   drop_na() %>%
+  # gather("variable","value",per_early:per_early_plot) %>% 
   filter(
-    variable %in% c("mean_eez","n_per")#,
+    variable %in% c("mean_eez","n_spp")#,
     # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
@@
 
+# Total number of transboundary stocks per EEZ
+n_trans_eez <- Transboundary_spp %>% 
+  group_by(eez_name) %>% 
+  summarise(
+    n_trans = unique(length(taxon_key))
+  )
+
+# Estimate the percentage changing
 n_taxa_data <- Prop_results %>% 
@@
   select(eez_name,early,mid) %>% 
+  left_join(n_trans_eez) %>% 
+  # mutate(per = (early/n_trans)*100) %>% 
+  mutate_at(vars(early,mid),
+            .funs = ~(./n_trans)*100) %>% 
   rename(sf_eez_name = eez_name) %>% 
   left_join(Matching_names) %>% 
-  group_by(sau_fishing_entity,fishing_entity_id) %>% 
-  summarise(avrg_spp_per = mean(early,na.rm=T))
-
+  select(sau_fishing_entity,fishing_entity_id,early,mid) %>% 
+  gather("time_step","avrg_spp_per",early:mid) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,time_step) %>%
+  summarise(avrg_spp_per = mean(avrg_spp_per,na.rm=T)) %>% 
+  mutate(time_step = ifelse(time_step == "early","per_early","per_mid"))
+
+# Paper numbers
+mean(n_taxa_data$avrg_spp_per,na.rm=T)
+sd(n_taxa_data$avrg_spp_per,na.rm=T)
+
+# Included in map
 n_taxa_map <- World_map %>% 
-  left_join(n_taxa_data,
+  full_join(n_taxa_data,
             by = c("sau_fishing_entity","fishing_entity_id")
             )
-
-mean(n_taxa_data$avrg_spp_per,na.rm=T)
 
@@
 
-# gc()# clean ram
-
-time <- "per_mid"
-# time <- "per_early"
+time_sets <- c("per_early","per_mid")
+# i = 1
+for(i in 1:2){
+  gc()# clean ram
+
+  time <- time_sets[i]
 
@@
   Prop_map <- ggplot() +
-  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+    geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
             aes(fill = avrg_spp_per),
-          size = 0.1,col = "white") +
+            size = 0.1,
+            col = "white") +
     geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
+    geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
     geom_point(data = subset(Map_data, time_step == time & per_plot < 0),
@@
     scale_colour_viridis(
-    name="Stock-Share Ratio Gain \nand taxa change (%)",
+      name="Stock share ratio gain \nand changing stocks (%)",
       limits = c(0,max(Map_data$per_plot)),
@@
     scale_fill_viridis(
-    name="Stock-Share Ratio Gain \nand taxa change (%)",
+      name="Stock share ratio gain \nand changing stocks (%)",
       limits = c(0,max(Map_data$per_plot)),
-    breaks = seq(0,max(Map_data$per_plot),10)#,
+      breaks = seq(0,max(Map_data$per_plot),10),
       # alpha = 0.8
@@
     geom_sf(data = Lat_am_eez, size = 0.1, fill = "white", col = "gray30") +
-  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+    geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
+    geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
             aes(fill = avrg_spp_per),
@@
       # direction = -1
-  );Lat_am_map
+    )
 
@@
     geom_sf(data = Med_eez, size = 0.5, fill = "white", col = "gray30") +
-  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+    geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
+    geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
             aes(fill = avrg_spp_per),
@@
     geom_sf(data = Eu_eez, size = 0.5, fill = "white", col = "gray30") +
-  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+    geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
+    geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
             aes(fill = avrg_spp_per),
@@
     # Labels
-  draw_plot_label(label = c("A", "B", "The Caribbean","North West Europe","The Mediterranean"),
+    draw_plot_label(label = c("A", "B", "The Caribbean","Northwestern Europe","The Mediterranean"),
                     size = c(20,20,15,15,15),
-                  x = c(0.01,0.01,0.058,0.38,0.70),
+                    x = c(0.01,0.01,0.058,0.33,0.68),
                     y = c(1,0.45,0.44,0.44,0.44)
     )
+
+  name <- paste0("./Figures/prop_",time,".png")
 
@@
   save_plot(
-  "./Figures/figS3.png",
+    name,
     plot_all,
@@
 
+}
 ```
@@
   )
+
+```
+
+## Sensitivity analysis
+
+Testing different Per values for the centroids
+
+```{r data_loading, eval = T, echo = F}
+
+# Get taxon list
+Taxon_list_sens <- list.files(my_path("R",extra_path = "Toe_Sens_per/Emergence_2005_p20"))
+
+# clean names for function
+Taxon_list_sens <- gsub("_.*","",Taxon_list_sens)
+
+# SAU exploited species list
+Exploited_spp_list <- my_path("G","SAU","exploited_species_list.csv",read = T) %>%
+  clean_names() %>% 
+  pull(taxon_name)
 
@@
 
+### Arctic and Brasil missing species
+
+```{r arctic, eval = F, echo =F}
+
 # Old Code
@@
 
-# Possible explanatory variables
-
-## Habitat association (DemersPelag)
-## Dem Vs Pelagic (Dem_pel)
-## SST time of Emer (Toe_sst)
-## Stock Status [FishForVisa] (Catch_trend)
-## Biogeographic location (Biogeo_reg)
-
-# AIC value; model between predictive capacity and complexity (i.e., number of explanatory variables) is given by the model with lowest AIC value
-
-# -------------------------------- #
-#### Build models
-# -------------------------------- #
-
-# Model 1; all variables
-M1 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Main_data)
-
-M1_sum <- summary(M1)
-M1_aic <- AIC(M1)
-M1_r <-  M1_sum$adj.r.squared
-
-# Model 2; non demPel
-M2 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel, data = Main_data)
-
-M2_sum <- summary(M2)
-M2_aic <- AIC(M2)
-M2_r <-  M2_sum$adj.r.squared
-
-# Model 3; non biome
-M3 <- lm(emerging_yr ~ sst_toe + bot_toe + status + dem_pel + DemersPelag, data = Main_data)
-
-M3_sum <- summary(M3)
-M3_aic <- AIC(M3)
-M3_r <-  M3_sum$adj.r.squared
-
-# Model 4; non status
-M4 <- lm(emerging_yr ~ sst_toe + bot_toe + biome + dem_pel + DemersPelag, data = Main_data)
-
-M4_sum <- summary(M4)
-M4_aic <- AIC(M4)
-M4_r <-  M4_sum$adj.r.squared
-
-# Model 5; non sst
-M5 <- lm(emerging_yr ~ status + bot_toe + biome + dem_pel + DemersPelag, data = Main_data)
-
-M5_sum <- summary(M5)
-M5_aic <- AIC(M5)
-M5_r <-  M5_sum$adj.r.squared
-
-# Model 6; all + basin
-M6 <- lm(emerging_yr ~ sst_toe + bot_toe + status + biome + dem_pel + DemersPelag, data = Main_data)
-
-M6_sum <- summary(M6)
-M6_aic <- AIC(M6)
-M6_r <-  M6_sum$adj.r.squared
-
-# Model 7; Categories instead of status
-M7 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + dem_pel + DemersPelag, data = Main_data)
-
-M7_sum <- summary(M7)
-M7_aic <- AIC(M7)
-M7_r <-  M7_sum$adj.r.squared
-
-# Model 8; All variables + ccentroid lat
-M8 <- lm(emerging_yr ~ sst_toe + bot_toe + category + as.factor(biome) + DemersPelag + latitude, data = Main_data)
-
-M8_sum <- summary(M8)
-M8_aic <- AIC(M8)
-M8_r <-  M8_sum$adj.r.squared
-
-# Model 9; Include UN regions
-M9 <- lm(emerging_yr ~ sst_toe + bot_toe + category + biome + DemersPelag + latitude + un_region, data = Main_data)
-
-M9_sum <- summary(M9)
-M9_aic <- AIC(M9)
-M9_r <-  M9_sum$adj.r.squared
-
-tibble(
-  "model" = paste("M",seq(1,9,1),sep=""),
-  "AIC" = c(M1_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic, M9_aic),
-  "rsquared" = c(M1_r,M2_r,M3_r,M4_r,M5_r,M6_r,M7_r,M8_r,M9_r)
-)
-
-# The Winner is M9    49283.   0.0765
-
-#Table for paper
-M9_sum %>% 
-  tidy() %>% 
-  filter(p.value < 0.05) %>% 
-  mutate(p.value = ifelse(p.value < 0.001,0.001,round(p.value,3))) %>% 
-  # stargazer(title = "Multiple regression results for time of emergence of Transboundary stocks.
-Only showing explanatory variables.
-The model significantly predicts time of emergencec (F(31,5270) = 15.38, p < 0.001, R2 = 0.06).",
-            # align=TRUE)
-  # knitr::kable() %>% 
-  write_csv(.,
-            "./Tables/mlr_result.csv")
-
-#Table for supplements
-M9_sum %>% 
-  tidy() %>% 
-  # filter(p.value < 0.05) %>% 
-  # mutate(p.value = ifelse(p.value < 0.001,0.001,round(p.value,3))) %>% 
-  # stargazer(title = "Multiple regression results for time of emergence of Transboundary stocks.
-Only showing explanatory variables.
-The model significantly predicts time of emergencec (F(18,6155) = 20.68, p < 0.001, R2 = 0.05).",
-            # align=TRUE)
-  # knitr::kable() %>% 
-  write_csv(.,
-            "./Tables/mlr_result_sup.csv")
+Arctic <- c("Canada (Arctic)","USA (Alaska, Arctic)", "Russia (Laptev to Chukchi Sea)")
+
+can_us_arctic <- Transboundary_spp %>% 
+  filter(eez_name == "Canada (Arctic)",
+         eez_neighbour == "USA (Alaska, Arctic)")
+
+arctic_spp <- Transboundary_spp %>% 
+  filter(eez_name %in% Arctic,
+         eez_neighbour %in% Arctic)
+
+arctic_spp <- Spp_depth %>% 
+  filter(taxon_key %in% can_us_arctic$taxon_key)
+
+# Transbounudary species that will see no change
+# 600252 Mallotus villosus  (Capelin)
+# 600315 Eleginus gracilis (Safforn cod)
+# 601520 Clupea pallasii  (Pacific herring)
+
+Eme_Res %>% 
+  filter(taxon_key %in% can_us_arctic$taxon_key) %>% 
+  # filter(sau_eez_name %in% arctic_spp$eez_name) %>% 
+  View(.)
+
+```
+
+```{r Brasa, eval = F, echo =F}
+
+brasil <- Transboundary_spp %>% 
+  filter(eez_name == "Brazil")
+
+brasil_spp <- Spp_depth %>% 
+  filter(taxon_key %in% brasil$taxon_key)
+
+# Transbounudary species that will see no change
+# 601590  Brevoortia pectinata  
+# 690022  Pleoticus muelleri 
+# 690023  Illex argentinus  
+# 690338  Penaeus paulensis
+
+Eme_Res %>% 
+  filter(taxon_key %in% brasil_spp$taxon_key) %>% 
+  View(.)
+
 ```

### 2020-11-24 Scripts/Manuscript.Rmd
misc

@@
 Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd; @PalaciosAbrantes:2020tv] for a total of 9,132 transboundary stocks.
-We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:FigureS1)) including island territories.
+We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:toeFigureS1)) including island territories.
 We determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
@@
 We developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
-This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:FigureS2)).
+This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2)).
 The centroid of a transboundary stock was determined by the average latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
@@
 We ran a sensitivity analysis (e.g.
-per = $20^{th}$, $50^{th}$, $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:FigureS2)).
-The centroid of a EEZ was estimated using the st package in R (version 3.2.1) (Figure \@ref(fig:FigureS1)).
+per = $20^{th}$, $50^{th}$, $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS2)).
+The centroid of a EEZ was estimated using the st package in R (version 3.2.1) (Figure \@ref(fig:toeFigureS1)).
 For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
@@
 
-The average ToE of the transboundary index (TI) across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:Figure1).
+The average ToE of the transboundary index (TI) across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:toeFigure1).
 Fish stocks with TI that first emerged from their historical variability in each EEZ was projected to be as early as 2006.
@@
 *Engraulis ringens* shared by Chile and Peru emerged in 2010).
-Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century and are not expected to stop there (Figure \@ref(fig:Figure1)).
+Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century and are not expected to stop there (Figure \@ref(fig:toeFigure1)).
 The number of EEZs with emerging stocks is projected to increase steeply from 87% to 96% during the period of 2006 to 2020.
 After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.74% per year.
-By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:Figure1)).
+By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:toeFigure1)).
 
-```{r Figure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone and transboundary stock.
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone and transboundary stock.
 A) Showing results in absolute numbers.
@@
 The median ToE of the transboundary index of stocks varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, X2 = 287.23, DF = 93, p < 0.001).
-Around one fifth (19%) of the EEZs have an average time of emergence across stocks prior to 2020 (Figure 2).
+Around one fifth (19%) of the EEZs have an average time of emergence across stocks prior to 2020 (Figure \@ref(fig:toeFigure2).
 Notable exceptions where the ToE is relatively late include Honduras (ToE = 2071), New Zealand (ToE = 2055), Ireland (ToE = 2067) and Pitcairn (ToE = 2070).
 The EEZs in Eastern Asia, Latin America, the Caribbean and Polynesia regions differed significantly (p < 0.05) from (almost) all other regions (see Table S1 for test statistics).
-(Figure \@ref(tab:TableS1).
+(Figure \@ref(tab:toeTableS1).
 
-Tropical latitudes also hold the majority of countries with the largest number of shifts in transboundary stocks due to climate change (Figure \@ref(fig:Figure2).
+Tropical latitudes also hold the majority of countries with the largest number of shifts in transboundary stocks due to climate change (Figure \@ref(fig:toeFigure2).
 Only in very few cases like Brazil and Alaska’s and Russia’s polar EEZs, no stock will emerge between 2006 and 2100.
@@
 While the selected 1 s.d.
-threshold is less conservative than a 2 s.d., the distribution of results do not dramatically change between thresholds, with a difference in ToE of about 4 years (Figure \@ref(fig:FigureS2)).
+threshold is less conservative than a 2 s.d., the distribution of results do not dramatically change between thresholds, with a difference in ToE of about 4 years (toeFigure \@ref(fig:toeFigureS2)).
 
-```{r Figure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of transboundary stocks.
+```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of transboundary stocks.
 A) Land polygon shows the vulnerability index of individual country’s fisheries management to shifts in transboundary stocks.
@@
 
-<!-- ![Figure 2.
+<!-- ![toeFigure 2.
 Average time of emergence of transboundary specie.
@@
 
-Our analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary species between 2005 and 2010 will see their distribution range shifts beyond historical variabilities before 2040 (Figure \@ref(fig:Figure3)).
+Our analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary species between 2005 and 2010 will see their distribution range shifts beyond historical variabilities before 2040 (toeFigure \@ref(fig:toeFigure3)).
 Particularly, nine of the top ten countries earning the most revenue from transboundary species, share a border.
@@
 
-```{r Figure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average time of emergence of transboundary species per fishing entity, aggregated by colour according to region.
+```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average time of emergence of transboundary species per fishing entity, aggregated by colour according to region.
 Fishing revenue on a logarithmic scale.
@@
 
-<!-- ![Figure 3.
+<!-- ![toeFigure 3.
 Average time of emergence of transboundary species per fishing entity.
@@
 
-We compared species time of emergence within EEZs by species’ habitat association (Figure \@ref(fig:Figure4)).
+We compared species time of emergence within EEZs by species’ habitat association (toeFigure \@ref(fig:toeFigure4)).
 Each species is assigned to be associated to one of the xx habitat types based on information available from FishBase (for fish, ref) or SeaLifeBase (for invertebrates, ref).
 We found significant differences in the time of emergence of stocks; range shifts by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
-Specifically, time of emergence of range shifts of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:TableS2)).
-On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (Figure \@ref(fig:Figure4)).
-Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:Figure4)).
+Specifically, time of emergence of range shifts of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
+On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (toeFigure \@ref(fig:toeFigure4)).
+Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (toeFigure \@ref(fig:toeFigure4)).
 
-```{r Figure4, eval = T,  echo = F, fig.width=14, fig.height=14,fig.pos= "H", fig.cap =  "Regional time of emergence per habitat association.
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Regional time of emergence per habitat association.
 Regions according to the United Nations and habitat association to FishBase.
@@
 
-<!-- ![Figure 4.
+<!-- ![toeFigure 4.
 Regional time of emergence per habitat association.
@@
 ToE of SST), human drivers (e.g.
-catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary species (Table \ref(tab:TableS3)).
-Results from the multiple linear regression analysis show a significant relation between ToE of transboundary species and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref(tab:Table1)).
+catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary species (Table \ref(tab:toeTableS3)).
+Results from the multiple linear regression analysis show a significant relation between ToE of transboundary species and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref(tab:toeTable1)).
 -->
@@
 We estimated the changes in the SSR of transboundary stocks by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) and highlighting changes below a neighbouring country’s TP.
-Our results suggest that 87% (n = 238) of the world EEZs will see a change in SSR below the country’s TP by 2030 (Figure \@ref(fig:Figure5)A).
+Our results suggest that 87% (n = 238) of the world EEZs will see a change in SSR below the country’s TP by 2030 (toeFigure \@ref(fig:toeFigure5)A).
 Globally, average changes in SSR of transboundary stocks in most EEZs are projected to be lower than 25% by the early $21^{st}$ century.
-The changes in SSR is expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:FigureS4) and \@ref(fig:FigureS5)).
+The changes in SSR is expected to increase slightly towards the mid $21^{st}$ century (toeFigures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)).
 However, changes in SST for specific stocks and in some EEZs are projected to be exceptionally high.
@@
 
-```{r Figure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in Specie-Share Ratio by 2030 (2021-2040) relative to 1951-2005.
+```{r toeFigure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in Specie-Share Ratio by 2030 (2021-2040) relative to 1951-2005.
 Lines represent the average gain of transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those that are gaining shares (arrowhead).
@@
 Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
-Such early emergences of range shifts indicate that the immediate challenges climate change is posing on international fisheries management (Figure \@ref(fig:Figure1)).
-Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require fast responses from transboundary fisheries management \@ref(fig:Figure2) and \@ref(fig:Figure3)).
-Finally, we show that in many instances around the world, the coming years will see that the shares of many transboundary stocks by countries will drop below level that conduce cooperation between countries in managing the shared stocks (Figure \@ref(fig:Figure5)).
+Such early emergences of range shifts indicate that the immediate challenges climate change is posing on international fisheries management (Figure \@ref(fig:toeFigure1)).
+Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require fast responses from transboundary fisheries management \@ref(fig:toeFigure2) and \@ref(fig:toeFigure3)).
+Finally, we show that in many instances around the world, the coming years will see that the shares of many transboundary stocks by countries will drop below level that conduce cooperation between countries in managing the shared stocks (Figure \@ref(fig:toeFigure5)).
 
@@
 
-Our results suggest that in most EEZs, more than half of their transboundary stocks emerged in the early 2000s (Figure \@ref(fig:Figure1)).
+Our results suggest that in most EEZs, more than half of their transboundary stocks emerged in the early 2000s (Figure \@ref(fig:toeFigure1)).
 Our projections in the emergence of distribution shifts of transboundary stocks corroborate with previous studies that have detected and attributed changes in marine catch composition since the 1970s [@Cheung:2013gk; @Poloczanska:2016kk].
@@
 
-Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management (Figure \@ref(fig:Figure2)).
+Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management (Figure \@ref(fig:toeFigure2)).
 The premise of our vulnerability index is that, from a management perspective, countries whose shifting transboundary stocks provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
-Thus, climate risk hotspots for transboundary management happen where EEZs will see an early ToE and where transboundary stocks are important in terms of fisheries catches and revenues (Figure \@ref(fig:Figure2)).
-For example, Polynesia and Micronesia are the only regions where the average ToE across all stocks and EEZs already happened; 2019 or is about too 2021, respectively (Figure \@ref(fig:Figure5)).
+Thus, climate risk hotspots for transboundary management happen where EEZs will see an early ToE and where transboundary stocks are important in terms of fisheries catches and revenues (Figure \@ref(fig:toeFigure2)).
+For example, Polynesia and Micronesia are the only regions where the average ToE across all stocks and EEZs already happened; 2019 or is about too 2021, respectively (Figure \@ref(fig:toeFigure5)).
 Countries in these regions are highly dependent on pelagic-oceanic (e.g.,tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
-For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \@ref(fig:Figure5)).
+For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \@ref(fig:toeFigure5)).
 In this case, policy has to work against the clock and under potentially high societal pressure to address an issue that could be impacting important fisheries [@Burden:2019ij].
@@
 
-Changes in SSR will be seen across the world, suggesting that climate change impacts on transboundary stock management is a global problem (Figure \@ref(fig:Figure5)).
+Changes in SSR will be seen across the world, suggesting that climate change impacts on transboundary stock management is a global problem (Figure \@ref(fig:toeFigure5)).
 Poleward shifts in SSR are expected in regions of the world with relatively large EEZs and few boundaries, like the North East Pacific.
@@
 They are consistent with regional modeling studies suggesting that by mid $21^{st}$ century, marine stocks will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
-Moreover, regions characterized by multiple political boundaries, like the Caribbean and the North East Atlantic, will see a complex network of changes in SSR with no clear pattern (Figure \@ref(fig:Figure5)B).
+Moreover, regions characterized by multiple political boundaries, like the Caribbean and the North East Atlantic, will see a complex network of changes in SSR with no clear pattern (Figure \@ref(fig:toeFigure5)B).
 These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] supporting the implementation of joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
@@
 
-## Supplements
+<!-- ## Supplements -->
 
-```{r FigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
+```{r toeFigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
 
@@
 
-```{r FigureS2, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+```{r toeFigureS2, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
 The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel).
@@
 
-```{r FigureS3, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
+```{r toeFigureS3, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
 
@@
 
-```{r FigureS4, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
+```{r toeFigureS4, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
 

### 2020-11-24 Scripts/Supplements.Rmd
misc

@@
 
-```{r FigureS1, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
+```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
 
@@
 
-```{r FigureS2, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+```{r toeFigureS2, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
 The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel).
@@
 
-```{r FigureS3, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
+```{r toeFigureS3, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
 
@@
 
-```{r FigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
+```{r toeFigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
 
@@
 
-```{r FigureS5, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
+```{r toeFigureS5, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
 Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
@@
 
-```{r TableS1, eval = T, echo = F}
+```{r toeTableS1, eval = T, echo = F}
 
@@
 
-```{r TableS2, eval = T, echo = F}
+```{r toeTableS2, eval = T, echo = F}
 

### 2020-11-24 Scripts/Initial_Analysis.Rmd
included er gid sensitivity

@@
 
+# Load all sp in one table
+sen_list <- c(20,50,90,95)
+
+read_sen <- function(sen){
+
+  if(sen != 95){
+
+  level <- paste0("Emergence_2005_p",sen)
+  extra_path <- paste0("Toe_Sens_per/",level)
+
+  # Get taxon list
+  Taxon_list_sens <- list.files(my_path("R",extra_path = extra_path))
+
+  Paths_sen <- paste(my_path("R",extra_path = extra_path),Taxon_list_sens,sep="")
+
+  }else{
+
+  Taxon_list_sens <- list.files(my_path("R",extra_path = "Toe_Sens_per/Emergence_2005_p20"))
+
+  Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_list_emer,"_emergence.txt",sep="")
+
+  }
+
+  Eme_sen <- bind_rows(lapply(Paths_sen, fread)) %>% 
+    mutate(sen = sen) %>% 
+    filter(
+      n_ens == 10
+    ) %>% 
+    left_join(Neighbours,
+              by = c("eez_name","eez_neighbour")
+    ) %>% 
+    rename(sf_eez_name=eez_name) %>% 
+    left_join(Matching_names,
+              by=  "sf_eez_name") %>% 
+    clean_names()
+
+  return(Eme_sen)
+
+}
+
+Sen_data <- bind_rows(
+  lapply(sen_list,read_sen)
+) %>% 
+  filter(
+    tresh== "tresh_one"
+  )
+
+```
+
+```{r data_analysis, eval = T, echo = F}
+
+# Per Stock
+sen_stock_d <-  
+  Sen_data %>% 
+  left_join(Neighbours) %>%
+  group_by(taxon_key,neighbour_id,sen) %>% 
+  summarise(
+    taxon_emer = min(emerging_yr)
+  ) %>% 
+  group_by(taxon_emer,sen) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  group_by(sen) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "stock"
+  )
+
+#  EEZ data
+sen_eez_d <-  Sen_data %>% 
+  left_join(Neighbours) %>% 
+  filter(n_ens == 10) %>% 
+  group_by(sf_eez_name,sen) %>% 
+  summarise(
+    taxon_emer = min(emerging_yr)
+  ) %>% 
+  group_by(taxon_emer,sen) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  group_by(sen) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "eez"
+  )
+
+# Plot Data
+sen_plot_data <- sen_stock_d %>% 
+  bind_rows(sen_eez_d) %>% 
+  mutate(
+    # total = ifelse(level == "stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$sf_eez_name))),
+    total = ifelse(level == "stock",max(stock_d$com_sum),max(eez_d$com_sum)),
+    per = com_sum/total*100
+  ) %>% 
+  gather("metric","value",com_sum,per) %>% 
+  ungroup() %>% 
+  mutate(
+    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
+    label = ifelse(total == max(stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs"),
+  sen = paste("Percentile above",sen,"%") 
+  )
+
+# Plot me
+ggplot() +
+  geom_area(data = subset(sen_plot_data, level == "stock"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  )  +
+  geom_area(data = subset(sen_plot_data, level == "eez"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  )  +
+  # Shameslly plot it again to fix colors
+  geom_area(data = subset(sen_plot_data, level == "stock" & my_facet == "Percentage (%)"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+            ) +
+  facet_wrap(~my_facet + sen,
+             scales = "free",
+             nrow=2) + 
+  scale_y_continuous("",
+                     expand = c(0,0)
+                     ) +
+  scale_x_continuous("Time of emergence",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     label = c(seq(2000,2075,25),">2100"),
+                     expand = c(0,0)
+  ) +
+  my_ggtheme_p() +
+  theme(
+    legend.position = "right",
+    # strip.text.x = element_blank(),
+    panel.spacing = unit(3, "lines") # space between plots
+  ) +
+  guides(fill=guide_legend(title= "Level", reverse=FALSE)) +
+  scale_fill_viridis(discrete = T,
+                     begin = 0.1, 
+                     end = 0.7,
+                     labels = c("EEZs","Stocks")
+  ) +
+  ggsave("./Figures/toe_sen.png",
+  width = 14,
+  height = 10,
+  units = "in"
+  )
+
 ```

### 2020-11-24 Scripts/Manuscript.Rmd
William comments 2nd round

@@
 
-Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming the waters, decreasing the pH and reducing oxygen content [@IPCC:2019tn].
+Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn].
 Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk].
-As a result of climate change, many marine species have changed their distribution towards higher latitude, deeper water or following local environmental gradients [@Poloczanska:2016kk].
-Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt].
-Climate change is expected to impact fisheries production, compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].
-The projected risks of impacts can be reduced through improving the effectiveness of current fisheries management [@Gaines:2018sg; @Cheung:2018dh], including managing fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
+).
+As a result of climate change, many marine species have changed their distribution towards higher latitude, deeper water or following local temperature gradients [@Poloczanska:2016kk].
+Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt] impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].The projected risks of impacts can be reduced through improving the effectiveness of current fisheries managemen [@Gaines:2018sg; @Cheung:2018dh], including managing fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
 
 The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
-As defined by the United Nation's Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (i) transboundary, those that cross neighboring EEZs; (ii) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (iii) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions including both the high seas and EEZs [@Munro:2004th].
-While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [Article 63, @UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species, yielding an average 48.5 million tonnes of catch and USD 78 billion in fishing revenue, per year, between 2005 and 2010 [@PalaciosAbrantes:2020tv].
+As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (*i*) transboundary, those that cross neighboring EEZs; (*ii*) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions including both the high seas and EEZs [@Munro:2004th].
+While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
+A recent study estimates that, globally, there are 633 transboundary fish species, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
 
-The effectiveness of managing fisheries targeting shared stocks is challenged by the shifts in marine species distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
-In many cases, catch or fishing effort quota on shared stocks are based on historical records, [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change in the fish stocks and fisheries [@PalaciosAbrantes:2020tv].
+The effectiveness of managing fisheries targeting shared stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
+In many cases, catch or fishing effort quota on shared stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change in the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
 Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by worsening climate change [@Pinsky:2018cb].
 With future shifts in species distributions under climate change, the challenges posed on shared stocks fisheries management will continue to increase in the $21^{st}$ century.
-Thus, there is a need to understand the time when climate change will affect the sharing distribution of these species as well as the intensity intensity of the impact in order to prepare ocean governance to species on the move [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
+To prepare ocean governance to meet the challenges in resource management arising from shifting shared stocks, it is important to understand the time when climate change will affect the sharing distribution of these stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
 
-The timing and intensity of climate impacts on managing transboundary fish stocks can be quantitative examined through the concepts of "Time of Emergence" (ToE) and “Thread Point” (TP) can help address some of the management challenges imposed by shifting species.
-Roughly, ToE is defined as the moment in time when an anthropogenic change signal (e.g., future trend) rises above the background noise of natural variability (e.g., historical variation) [@IPCC:2019tn].
+The timing and intensity of climate impacts on managing transboundary fish stocks can be quantitatively examined through the concepts of ‘time of emergence’ (ToE) and ‘threat point’ (TP).
+Time of emergence is defined as the moment in time when an anthropogenic change signal (e.g., future trend) rises above the background noise of natural variability (e.g., historical variation) [@IPCC:2019tn].
 This concept has been applied to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
-The premise behind ToE is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background of natural climate variability [@Hawkins:2012hc].
-The concept of thread point (TP) comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
+The premise behind the time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
+The concept of threat point comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
 Game theory has been widely used to investigate and manage shared stocks [@Sumaila:2013hv].
@@
 
-Here, we estimate the ToE and TP of the share distribution of transboundary stocks worldwide.
-We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emission scenario RCP8.5 (see Methods).
-We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., we did not include the high seas).
-Our results suggest that many shared stocks have already ‘emerged’ from historical natural variability and overpassed their share TP in most neighboring EEZs posing a threat to the sustainability of shared stocks and the resilience of international treaties.
+Here, we estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
+We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
+We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks, excluding the high seas in this study).
+Our results suggest that many transboundary stocks have already ‘emerged’ from historical natural variability and overpassed their share threat point in most neighboring EEZs, posing a threat to the sustainability of shared stocks and the resilience of international treaties.
 
@@
 This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd; @PalaciosAbrantes:2020tv] for a total of 9,132 transboundary stocks.
-We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:toeFigureS1)) including island territories.
+Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv]; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks in this analysis.
+Using the transboundary species Atlantic cod (Gadus morhua) as an example, based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of G.
+morhua, but the United States and Greenland do not share a stock of this transboundary species.
+We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; \@ref(fig:toeFigureS1))) including island territories.
 We determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
-Each EEZ was categorized by the geopolitical regions according to the United Nations and by natural biome [@Reygondeau:2019uh].
-The habitat association of each species was determined based on information available from FishBase (http://www.fishbase.org).
-For each species and EEZ, we used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
+Each EEZ was categorized by the geopolitical regions according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and by natural biome [@Reygondeau:2019uh].
+The habitat association of each species was determined based on information available from FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org).
+For each stock and EEZ, we used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
 We report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
@@
 The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
-For pelagic species, the model uses environmental variables at the surface (e.g., sea surface temperature) whereas demersal species distribution are driven by ocean bottom variables (e.g., bottom temperature).
-This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (see Cheung *et al.* [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
-The DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd], Cheung et al.
+For pelagic species, the model uses environmental variables at the surface (e.g., sea surface temperature) whereas demersal species distribution is driven by ocean bottom variables (e.g., bottom temperature).
+This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (see Cheung et al.
+[-@Cheung:2010dt; -@Cheung:2016jd] for model details).
+DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd], Cheung et al.
 *in revision*].
 
-We forced the DBEM with simulated ocean conditions from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+We forced the DBEM with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
 The GFDL-ESM2M is run under historical forcing until 2005 and follows the ‘high’ climate change emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
-In GFDL-ESM2M, atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a species’ distribution, we have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
+In the GFDL-ESM2M, atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
 We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
@@
 We developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
-This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2)).
-The centroid of a transboundary stock was determined by the average latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
+This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2).
+The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
 Therefore,
@@
 where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
-To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top $10^{th}$ percentile (e.g., per = 10%).
-We ran a sensitivity analysis (e.g.
-per = $20^{th}$, $50^{th}$, $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS2)).
-The centroid of a EEZ was estimated using the st package in R (version 3.2.1) (Figure \@ref(fig:toeFigureS1)).
+To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top 95th percentile (e.g., per = 95%).
+We ran a sensitivity analysis using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
+The centroid of each EEZ was estimated using the st package in R (Figure \@ref(fig:toeFigureS1)).
 For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
@@
 where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively and $lon_{eez}$ and $lon_{ts}$ are the respective longitudes.
-
 Then, for each year we calculated the transboundary index as follows:
@@
 where $D_A$ and $D_B$ represent the distance between the distribution centroids of a species and the centroid of the pair of neighbouring EEZs (A and B) sharing the species for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
-Thus, a higher TI means that a shared species becomes more or less abundant relative to its neighbouring EEZs that is beyond what the fisheries experienced historically.
-From the perspective of a state’s fisheries, higher TI indicates an increase in the challenges to effective management and sustainability of the shared species’ fisheries.
-In contrast, a smaller TI suggests that the sharing of the species between neighbouring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the shared species’ fisheries.
-The index was smoothed to a 10 year average to reduce interannual variability.
+Thus, a higher TI means that a shared stock becomes more or less abundant relative to its neighbouring EEZs that is beyond what the fisheries experienced historically.
+From the perspective of a state’s fisheries, higher TI indicates an increase in the challenges to effective management and sustainability of the shared stock’ fisheries.
+In contrast, a smaller TI suggests that the sharing of the stocks between neighbouring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the shared stocks’ fisheries.
+The index was smoothed to a 10-year average to reduce interannual variability.
 
-#### Calculating the Time-of-Emergence of the transboundary index
+#### Calculating the time of emergence of the transboundary index
 
-Knowing the point in time (e.g., year) at which the distribution of a shared stock will diverge from its natural internal variability is important in informing the lead-time to which climate adaptation in transboundary fisheries management needs to be implemented [@Link:2010ei].
-Thus, we define ToE as the time when the stock shared between neighbouring EEZs will exceed the historical natural internal variability as follows:
+Knowing the point in time (e.g., year) at which the distribution of a shared stock will diverge from its natural internal variability is important in informing the lead-time to which climate adaptation needs to be implemented [@Link:2010ei].
+Here, we define time of emergence as the time when the stock shared between neighbouring EEZs will exceed the historical natural internal variability as follows:
 
@@
 EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}} \\
-ToE = t when Emergence index[t] > 1
+ToE = t when Emergence index[t] > 1 s.d.
 \end{split}
@@
 
-### Estimating the threat point (TP) of transboundary stock share
+Estimating the threat point of transboundary stock share
 
 In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than operating on a non-cooperative basis [@Munro:1979dg].
-However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., "threat point" (TP) [@Sumaila:2013hv].
-Here, we define TP as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., "threat point" [@Sumaila:2013hv].
+Here, we define threat point as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
 Any proportion below the defined threat point of a country would result in a deflection from the joint management of the shared stock.
@@
 Second, we average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
-The first time period ($t_h$) spans from 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
+The first time period ($SSR_{th}$) spans from 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
 It is assumed as the historic baseline of shared distributions for each stock.
-We then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 (*te*), and the mid $21^{st}$ century as the average of 2041 to 2060 (*tm*).
-We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs [@UnitedNations2018] such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
+We then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$).
+We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
 The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members.
-Third, we estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_h$) following Palacios-Abrantes et al.
+Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$.
+This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{et} \geqq (SSR_{th} + \sigma)$ or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
+Finally, we estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al.
 (*in press*),
@@
 <!-- $$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$ -->
-<!-- $TP = \Delta{SSR}_{e,m} < \mu \pm \sigma$ -->
 
 \begin{equation}
-\begin{split}
-\label{eq:six}
-\Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 \\
-TP = \Delta{SSR}_{e,m} < \Delta{SSR}_{th} \pm \sigma
-\end{split}
+\label{eq:five}
+\Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100
 \end{equation}
-
-where *e* is EEZ and *s* is stock.
-Finally, a substantial change in SSR of each EEZ was determined when $\Delta{SSR_{e,m}}$ was beyond one standard deviation of the historical distribution i.e., when $SSR_f \geq (SSR_f + \sigma)$ or $SSR_f \leq (SSR_f - \sigma)$.
-In other words, when the future SSR is below or above a neighbouring EEZ TP.
 
@@
 
-### Time of Emergence of transboundary species
+Time of emergence of transboundary stocks
 
-The average ToE of the transboundary index (TI) across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:toeFigure1).
-Fish stocks with TI that first emerged from their historical variability in each EEZ was projected to be as early as 2006.
+The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 \pm 26 years (Figure \@ref(fig:toeFigure1)).
 In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100.
-About 55% (n = 3,154) of those stocks are projected to have their shifts in distributions emerged between 2006 and 2020.
-Also, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 (e.g.
-*Engraulis ringens* shared by Chile and Peru emerged in 2010).
-Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century and are not expected to stop there (Figure \@ref(fig:toeFigure1)).
-The number of EEZs with emerging stocks is projected to increase steeply from 87% to 96% during the period of 2006 to 2020.
-After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.74% per year.
-By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:toeFigure1)).
+Fish stocks with transboundary index that first emerged from their historical variability in each EEZ was projected to be as early as 2006 and the last one in 2100.
+About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g.
+Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010).
+The number of EEZs with emerging stocks is projected to increase steeply from 87% in 2006 to 96% in 2020.
+After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.7% per year.
+By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure Figure \@ref(fig:toeFigure1)).
+Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020.
+Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century and are not expected to stop there (Figure Figure \@ref(fig:toeFigure1)).
+We based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
+Results from the sensitivity analysis suggest that our method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g.
+the top 90%) of the stock (Figure Figure \@ref(fig:toeFigures·3).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone and transboundary stock.
-A) Showing results in absolute numbers.
-B) Showing the result as a percentage of the total stocks and EEZ."}
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone (EEZ) and transboundary stock: A) results in absolute numbers (green).
+B) results expressed as a percentage of the total emerging stocks and EEZ with emerging stock (purple)."}
 
@@
 
-The median ToE of the transboundary index of stocks varied significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, X2 = 287.23, DF = 93, p < 0.001).
-Around one fifth (19%) of the EEZs have an average time of emergence across stocks prior to 2020 (Figure \@ref(fig:toeFigure2).
-Notable exceptions where the ToE is relatively late include Honduras (ToE = 2071), New Zealand (ToE = 2055), Ireland (ToE = 2067) and Pitcairn (ToE = 2070).
-The EEZs in Eastern Asia, Latin America, the Caribbean and Polynesia regions differed significantly (p < 0.05) from (almost) all other regions (see Table S1 for test statistics).
-(Figure \@ref(tab:toeTableS1).
+The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A).
+Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1)) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B).
+In contrast, EEZs located in template regions like northern Europe and eastern Asia will have a significantly later time of emergence (p < 0.05; see Table \@ref(tab:toeTableS1 for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058).
+However, notable exceptions of tropical EEZs with a later average time of emergence across stocks include Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070).
+Moreover, despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) will have relative early average time of emergence (Figure \@ref(fig:toeFigure2)A).
+Only in a few cases, the transboundary index of a stock will not emerge between 2006 and 2100.
+In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
+In the case of Brazil, we included four transboundary stocks, all shared with Uruguay [@PalaciosAbrantes:2020]: Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
+In all cases, our study suggest that the transboundary index of these stocks will not emerge before 2100 relative to the historic (1951-2005) distribution \@ref(fig:toeFigure2)A)
 
-Tropical latitudes also hold the majority of countries with the largest number of shifts in transboundary stocks due to climate change (Figure \@ref(fig:toeFigure2).
-Only in very few cases like Brazil and Alaska’s and Russia’s polar EEZs, no stock will emerge between 2006 and 2100.
-From the top 10 more vulnerable countries, seven are located in tropical waters (Tokelau 100, the	Marshall Islands 92, Peru 88, Kiribati 80, Tuvalu 80, Papua New Guinea 78, and the Bahamas 74) while only three are outside the tropics (Estonia 86, Finland 82 and Sweden 78).
-We set an arbitrary threshold of  1 s.d.
-(67% confidence) to determine the ToE (see Methods).
-We tested the sensitivity of our results using a larger threshold (i.e.,  2 s.d for a 95% confidence).
-While the selected 1 s.d.
-threshold is less conservative than a 2 s.d., the distribution of results do not dramatically change between thresholds, with a difference in ToE of about 4 years (toeFigure \@ref(fig:toeFigureS2)).
-
-```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of transboundary stocks.
-A) Land polygon shows the vulnerability index of individual country’s fisheries management to shifts in transboundary stocks.
-The warmer the colour the greater the vulnerability.
-Exclusive Economic Zone polygon represents the average time of emergence across all stocks, with the cooler the colour the later the average time of emergence.
-The Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10 degrees latitude running mean).
-Aqua color = no emerging stocks.
-B) Time of emergence according to the United Nations sub regions.
+```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of stocks’ transboundary index.
+A) Land polygon shows the participation of the emerging stocks in the country’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygon represents the average time of emergence of the transboundary index across all stocks.
+Aqua color represents EEZs with no time of emergence between 2006 and 2100.
+B) Time of emergence of the transboundary index according to the United Nations sub-regions.
 Whiskers represent 1.5* interquartile range.
@@
 <!-- ![toeFigure 2.
-Average time of emergence of transboundary specie.
-Land polygon shows the vulnerability index.
-Exclusive Economic Zone polygon represents the average time of emergence across all stocks.
-Density plot on the right shows the latitudinal number of emerging stocks across EEZ centroids (10^o running mean).](../Figures/Fig3.png) -->
+Average time of emergence of stocks’ transboundary index.
+A) Land polygon shows the participation of the emerging stocks in the country’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygon represents the average time of emergence of the transboundary index across all stocks.
+Aqua color represents EEZs with no time of emergence between 2006 and 2100.
+B) Time of emergence of the transboundary index according to the United Nations sub-regions.
+Whiskers represent 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Line represents median, and black points represent points outside of 1.5*IQR.](../Figures/Fig3.png) -->
 
-Our analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary species between 2005 and 2010 will see their distribution range shifts beyond historical variabilities before 2040 (toeFigure \@ref(fig:toeFigure3)).
-Particularly, nine of the top ten countries earning the most revenue from transboundary species, share a border.
-In some cases, like Indonesia, Thailand and the United States, stocks’ range shifts swill emerge by around 2030.
-The 47 countries that make up the bulk of the global fisheries revenues also have more stocks with range shifts emerging from historical variabilities.
+Our analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see their distribution range shifts beyond historical variabilities between 2025 and 2040 (Figure \@ref(fig:toeFigure3)).
+In some cases, like Indonesia (ranked $8^{th}$ in terms of revenue from transboundary stocks), Thailand (ranked $7^{th}$ in terms of revenue from transboundary stocks) and the United States (ranked $4^{th}$ in terms of fishing revenue from transboundary stocks), stocks’ range shifts swill emerge by around 2030.
+The 47 countries that make up the bulk of the global fisheries revenues also have more stocks with range shifts emerging from historical variabilities, being Spain (n = 122), France (n = 85), China and Portugal (both n = 73) and Senegal (n = 66) the top 5 leading countries in number of emerging stocks.
 
-```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average time of emergence of transboundary species per fishing entity, aggregated by colour according to region.
+On average, stocks that will see their transboundary index emerge by 2100 represent 27% ± 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010.
+However, large variation exists in the proportion that stocks with emerging transboundary index have in the current revenue participation across EEZs.
+In some cases like El Salvador and New Zealand, stock’s with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries  in their respective EEZs between 2004 and 2014.
+On the contrary, stocks with emerging transboundary index from the Marshal Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A).
+In some cases, a few emerging stocks represent large proportion of revenue from transboundary stocks such as Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively).
+On the contrary, Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the participation of such stocks on fishing revenue is less than 20%.
+
+```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average time of emergence of stocks’ transboundary index per fishing entity, aggregated by colour and shape according to region.
 Fishing revenue on a logarithmic scale.
@@
 
-<!-- ![toeFigure 3.
-Average time of emergence of transboundary species per fishing entity.
-Fishing revenue in logarithmic scale.
-Showing names for the top 75th percentile.
-Countries marked with an asterisk have at least one neighboring border in the group.](../Figures/Fig4.png) -->
+We set an arbitrary threshold of 1 s.d.
+(67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods).
+We tested the sensitivity of our methods by considering a larger threshold for the time of emergence results (i.e., 2 s.d for a 95% confidence).
+The global average time of emergence of the transboundary index considering a 2 s.d.
+threshold was 2036 \pm 28, a difference of about 7 years relative to the 1 s.d.
+threshold (Figure \@ref(fig:toeFigureS2)).
+In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric, that is, 40% (n = 1626) fewer emerging stocks than using a 1.s.d threshold.
 
-We compared species time of emergence within EEZs by species’ habitat association (toeFigure \@ref(fig:toeFigure4)).
-Each species is assigned to be associated to one of the xx habitat types based on information available from FishBase (for fish, ref) or SeaLifeBase (for invertebrates, ref).
-We found significant differences in the time of emergence of stocks; range shifts by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
+We compared species time of emergence within EEZs by species’ habitat association (Figure \@ref(fig:toeFigure4)).
+Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods).
+We found significant differences in the time of emergence of stocks range shifts, indicated by the transbounudary index, by their specie's habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
 Specifically, time of emergence of range shifts of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
 On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (toeFigure \@ref(fig:toeFigure4)).
-Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (toeFigure \@ref(fig:toeFigure4)).
+Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:toeFigure4)).
 
-```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Regional time of emergence per habitat association.
-Regions according to the United Nations and habitat association to FishBase.
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association.
+Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Methods).
+The number of species included in this analysis for each habitat type is noted in parenthesis.
 Whiskers represent 1.5* interquartile range.
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers outside of 1.5*IQR."}
+Line represents median, and black points represent outliers outside of 1.5 times the interquartile range."}
 
@@
 
-### Changes in the threat point of transboundary stocks
+### Changes in the stock share ratio of transboundary stocks
 
-```{r analysis_results, eval = F,  echo = FALSE, include = FALSE}
+We have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods - equation \@ref(eq:five)).
+By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% \pm 15% (Figure \@ref(fig:toeFigure5)A).
+The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)).
+However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high.
+For example, Guatemala shares Panulirus gracilis stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century.
+Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)).
+Our results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
+These EEZs will see one third (33% \pm 21%) of their transboundary stocks changing the share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A – Land).
+Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin will see changes in all of the transboundary stocks analyzed while Venezuela (6%), Bonaire (9%), China (10%) and Viet Nam (10%) will be the only countries with less than 10% of stocks changing beyond the threat point.
 
-# Early century
-n_eez <- prop_r %>%
-  filter(variable == "mean_eez" & 
-           !is.na(change),
-         time_step == "per_early"
-         ) %>% 
-  pull(eez_name) %>% 
-  unique() %>% 
-  length()
-
-n_eez_per <- round(n_eez/275*100,2)
-
-### Max values early
-max_values <- prop_r %>% 
-  filter(variable == "max_eez" & 
-           !is.na(change),
-         # time_step == "per_early"
-         )
-
-max_values$per_early
-
-### Average
-average_mid <- prop_r %>% 
-  filter(variable == "mean_eez" & 
-           !is.na(change)
-         )
-```
-
-We estimated the changes in the SSR of transboundary stocks by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) and highlighting changes below a neighbouring country’s TP.
-Our results suggest that 87% (n = 238) of the world EEZs will see a change in SSR below the country’s TP by 2030 (toeFigure \@ref(fig:toeFigure5)A).
-Globally, average changes in SSR of transboundary stocks in most EEZs are projected to be lower than 25% by the early $21^{st}$ century.
-The changes in SSR is expected to increase slightly towards the mid $21^{st}$ century (toeFigures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)).
-However, changes in SST for specific stocks and in some EEZs are projected to be exceptionally high.
-For example, Guatemala is sharing *Panulirus gracilis* stock with Mexico, and Guatemala’s share is projected to increase by >50% from 12% to 26% by the early $21^{st}$ century.
-By the early $21^{st}$ century, the percentage of transboundary stocks per EEZ that will see a change in SSR above the country’s TP historical distribution  (e.g., $\Delta{SSR} > \mu \pm \sigma$) will be relatively large (global mean of 69% of stocks in an EEZ).
-
-```{r toeFigure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in Specie-Share Ratio by 2030 (2021-2040) relative to 1951-2005.
+```{r toeFigure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.
 Lines represent the average gain of transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those that are gaining shares (arrowhead).
-Land polygons represent the percentage of initial stocks that will see a shift.
-Panel B zooms in to specific areas shown within the grey dashed line boundaries in A."}
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point.
+Panel B zooms in to specific areas shown within the grey dashed line boundaries in A.s"}
 
@@
 
+Our results offer no clear global pattern in the change of stock share ratio (Figure \@ref(fig:toeFigure5)).
+In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, stock share ratio is expected to follow a poleward shift.
+However, central America Pacific and West African coasts are expected to follow an equatorial direction.
+Moreover, some regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no stablished pattern, this is the case of the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).
+
 ## Discussion
 
-In this paper we endeavor to look at the timing and intensity of changes in the sharing of transboundary species.
+In this paper, we investigate the timing and intensity of changes in the sharing of transboundary stocks between countries.
 Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
-Such early emergences of range shifts indicate that the immediate challenges climate change is posing on international fisheries management (Figure \@ref(fig:toeFigure1)).
-Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require fast responses from transboundary fisheries management \@ref(fig:toeFigure2) and \@ref(fig:toeFigure3)).
-Finally, we show that in many instances around the world, the coming years will see that the shares of many transboundary stocks by countries will drop below level that conduce cooperation between countries in managing the shared stocks (Figure \@ref(fig:toeFigure5)).
+Such early emergences of range shifts indicate the urgency of the challenges that climate change is posing on international fisheries management.
+Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management.
+Finally, we show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to level below that conduce cooperative resource management between countries.
+Adaptation measures are needed to reduce the impacts of species range shifts on current international transboundary fisheries treaties.
 
@@
 
-Our results suggest that in most EEZs, more than half of their transboundary stocks emerged in the early 2000s (Figure \@ref(fig:toeFigure1)).
-Our projections in the emergence of distribution shifts of transboundary stocks corroborate with previous studies that have detected and attributed changes in marine catch composition since the 1970s [@Cheung:2013gk; @Poloczanska:2016kk].
-For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (USA).
+Our projections in the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century corroborate with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk].
+For example, in the early 2000s, Humboldt squid (Dosidicus gigas) substantially expanded its geographic range poleward, reaching the coast of Washington state (US).
 A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs].
-In the North East Atlantic, the range expansion of Atlantic mackerel (*Scomber scombrus*) into Icelandic waters in 2010 was followed by the increase in fishery quota **xxx** without consultation with **xxxx.** Such changes in quota resulted in disputes Iceland and Faroes Islands with multiple fishing nations co-managing the Atlantic mackerel stocks [@Spijkers:2017ij].
+In the northeast Atlantic, Atlantic mackerel (Scomber scombrus) is multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
+However, a range expansion of Atlantic mackerel (Scomber scombrus) into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008 without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij].
+Such changes in quota resulted in disputes between Iceland and Faroes Islands, and the NEAFC member states [@Spijkers:2017ij].
+As transboundary stocks continue to emerge in the incoming years, imal-adapted fisheries management and international disputes like that of A.
+mackerel are likely to increase in frequency [@Pinksy2018]
 
-The projected range shifts of large number of transboundary species emerging since around the 2000s period  can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine stocks [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
-It has been estimated that since the early 2000s, SST and oxygen ($O_2$) have emerged in 79% and 23% of the global ocean, respectively, and by 2080 SST will have emerged in 90% of the global ocean while $O_2$ will have doubled under a high greenhouse gas emission scenario [@Rodgers:2015jq].
+The projected range shifts of large number of transboundary stocks emerging since the early 2000s can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
+Marine organisms have a thermal tolerance that limits the temperature range they could live in [@IPCC:2019tn].
+This thermal tolerance explains the sensitivity of the animal to warming and is a main driver of species’ distributional change [@IPCC:2019tn].
+Oxygen plays a similar role to that of temperature [Clarke:2020ep], especially on tuna species that require large amounts of dissolve oxygen for they metabolism [@IPCC2014].
+It has been estimated that since the early 2000s, increase in sea surface temperature (SST) and decrease in oxygen content have emerged in 79% and 23% of the global ocean, respectively, and by 2080, SST is projected to emerge in 90% of the global ocean while O2 will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq].
+Moreover, before 2100, surface acidification, nitrate content and net primary production, also important variables for species distributions, will have emerged from the historical variability of 1861-1900 in 30% and 60% of the ocean under a low (RCP 2.6) and strong (RCP 8.5) climate change scenario respectively [@IPCC:2019tn].
+These oceanographic changes are important drivers of change in species distributions as represented in the DBEM [@Cheung:2016jd; @Cheung:0ur].
 
-Regional differences in time of emergence are likely the result of environmental factors, species vulnerability, and the characteristics of political boundaries.
-For example, our findings underscore previous findings suggesting that tropical species are already on the move.
+Regional differences in time of emergence are attributed to the variations in environmental conditions, species’ climate vulnerability, and the characteristics of political boundaries.
+Specifically, our findings are consistent with results from previous studies that show tropical species are already shifting their distributions in the last few decades [@Poloczanska:2016kk].
 The tropics are among the most species-rich biome [@Reygondeau:2019uh] and have potentially experienced some of the highest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
 In the tropics, marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz].
-Consequently, as a result of climate change, species have to move towards deeper or sub-tropical waters [@Poloczanska:2016kk] although recent studies suggest that in some regions species are shifting towards the equator and to shallow waters to follow local temperature gradients [@Clarke:2020ep].
+Consequently, as a result of climate change, species have to move towards deeper or sub-tropical waters [@Poloczanska:2016kk] although recent studies suggest that in some regions species are shifting towards the equator and to shallow waters to follow local temperature gradients [@Clarke:2020ep.
 Moreover, tropical species are expanding their range in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis shows a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
-In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs and many borders that match peaks of emerging stocks.
+In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs that border with multiple countries (e.g., The Caribbean).
+These tropical areas are projected to have the highest number of transboundary stocks’ distribution shifts emerging in the incoming years.
 
@@
 
-Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management (Figure \@ref(fig:toeFigure2)).
-The premise of our vulnerability index is that, from a management perspective, countries whose shifting transboundary stocks provide larger societal benefits and have an early ToE will be more vulnerable to shifts in SSR than their counterpart.
-Thus, climate risk hotspots for transboundary management happen where EEZs will see an early ToE and where transboundary stocks are important in terms of fisheries catches and revenues (Figure \@ref(fig:toeFigure2)).
-For example, Polynesia and Micronesia are the only regions where the average ToE across all stocks and EEZs already happened; 2019 or is about too 2021, respectively (Figure \@ref(fig:toeFigure5)).
-Countries in these regions are highly dependent on pelagic-oceanic (e.g.,tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
-For example, in the Marshall Islands 16 transboundary stocks responsible for 92% and 94% of the transboundary fisheries revenue and catch, respectively, have, on average, emerged since 2018 resulting in high vulnerability index (Figure \@ref(fig:toeFigure5)).
-In this case, policy has to work against the clock and under potentially high societal pressure to address an issue that could be impacting important fisheries [@Burden:2019ij].
-Currently, fisheries targeting shared stocks like tunas in the South Pacific Islands region operate under the Vessel Day Scheme - which takes into account shifts in stocks and catches making it a good example where a management system has been set up that is somewhat anticipatory [@Aqorau:2018bh].
-However, a low vulnerability index does not necessarily translate to no-risk of international conflict or threat to stocks sustainability.
-While it gives countries more time to prepare for substantial changes, stocks with relatively small overall participation on catches or revenue can destabilize cooperative shared stock management agreement as shown by the Atlantic mackerel example [@Spijkers:2017ij, Chapter 1].
+Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)).
+For example, Polynesia and Micronesia are the only regions where the average time of emergence of the distribution of where all transboundary stocks across all EEZs emerged before 2022 (Figure \@ref(fig:toeFigure5)).
+Moreover, countries in these hotspot regions are highly dependent on pelagic-oceanic (e.g., tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
+On the other hand, “coldspots” of climate risk for transboundary fisheries like the Northeast Atlantic and Northwest Pacific where stocks are expected to emerge later in the 21st century.
+Assuming that these countries already have adequate transboundary management to deal with historical and current variations in fish stocks’ distribution e.g., the Pacific Salmon Treaties [@Miller:2004fpa], management agencies are likely to have more time to prepare for emerging stock shifts.
+In cases where stocks are emerging or have already emerged, adaptation policies are urgently needed to address an issue that could be impacting important fisheries [@Burden:2019ij].
+Currently, fisheries targeting tunas shared stocks in the South Pacific Islands region operate under the Vessel Day Scheme - which takes into account shifts in stocks and catches making it a good example where a management system has been set up that is somewhat anticipatory [@Aqorau:2018bh].
+However, even strategies that take into consideration the shift of the species within the Party’s area will likely require to deal with the newcomer issue that NEAFC is currently facing over Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+As species have been emerging from their historical distribution since 2006, regional management bodies require to revise their management actions as today most of them are not equipped with strategies to cope with shifting species distributions [@Oremus:2020id; @Sumby:2021gw].
+This is of specific concern for those management bodies located in the emerging “hotspots” here identified.
 
-### Changes in stocks-share ratio will be ubiquitous
+The lack of Arctic stocks with emerging transboundary index is likely a combination of the few species currently fished in this region.
+Historically, there are 59 commercial fish stocks fished the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas with limited knowledge about the species biology [@Christiansen:2014gg].
+However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@REF].
+As a result, the biomass of Arctic marine animals [@BryndumBuchholz:2019kh] and catch and revenue of associated fisheries [@Tai:2019cq; @Lam:2014fq] are expected to increase in the incoming years, although substantial uncertainty exists on the magnitude of change [@BryndumBuchholz:2019kh].
+Moreover, projections suggest a larger increment on transboundary stocks in the incoming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further North [@Morley:2018fn; @Frainer:2017ji].
+For Brazil and Uruguay, the current and future distribution of the transboundary stocks in combination with the particular characteristics of these EEZs could be influencing the results.
+First, both Argentine menhaden and Argentine shortfin squid have distributions with larger biomass concentration further south, with Brazil serving as the northern limit of the distributional range [@AquaMaps:2019ui; @AquaMaps:2019wt].
+Second, Brazil’s EEZ is substantially larger than Uruguay’s resulting in a long distance between Brazil’s and the stocks’ centroids.
+Third, both stocks are expected to shift further south in the incoming years [@AquaMaps:2019ui; @AquaMaps:2019w].
+Thus, the further southern shift of Argentine menhaden and shortfin squid in combination with the reduced latitudinal gradient of Uruguay’s EEZ results in a small effect in the already long latitudinal distance between Brazil’s EEZ and the stocks’ centroids.
+In the case of the two shrimp species, their destitution range spams along the cost of Brazil and is not expected to have a strong shift in the incoming years, also resulting in a stable transboundary index [@AquaMaps:2019wn; @AquaMaps:2019wj].
 
-Changes in SSR will be seen across the world, suggesting that climate change impacts on transboundary stock management is a global problem (Figure \@ref(fig:toeFigure5)).
-Poleward shifts in SSR are expected in regions of the world with relatively large EEZs and few boundaries, like the North East Pacific.
-However, despite observations and modelling work indicating that, generally, climate change is shifting stocks distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], the Eastern Tropical Pacific will see EEZs from Mexico to Ecuador losing SSR to their southern neighbors.
-Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems.
-They are consistent with regional modeling studies suggesting that by mid $21^{st}$ century, marine stocks will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
-Moreover, regions characterized by multiple political boundaries, like the Caribbean and the North East Atlantic, will see a complex network of changes in SSR with no clear pattern (Figure \@ref(fig:toeFigure5)B).
-These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] supporting the implementation of joint management plans [@Baudron:2020dj] and the call for new implemented shared policies [@Levin:2018in].
-However, game theory suggests that the greater the number of negotiating parties , the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.
+### Changes in stock share ratio will be ubiquitous
+
+Changes in stock share ratio beyond a country’s threat point will be seen across the world, suggesting that climate change impacts on transboundary stock management is a global problem (Figure \@ref(fig:toeFigure5)).
+Poleward shifts in stock share ratio are expected in regions of the world with relatively long EEZs and few boundaries, like the northeast Pacific and the southeast Atlantic.
+However, despite observations and modelling work indicating that, generally, climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], some regions like the eastern tropical Pacific and eastern tropical Atlantic Oceans will experience a shift in distribution towards the equator.
+Specifically, in the eastern tropical Pacific, the direction of shift is expected to be from Mexico to Ecuador with EEZs losing stock share ratio to their southern neighbors.
+Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems [@Clarke:2020ep].
+They are consistent with regional modeling studies suggesting that by mid 21st century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
+Moreover, regions characterized by multiple political boundaries, like the Caribbean and the northeast Atlantic, will see a complex network of changes in SSR with no clear pattern (\@ref(fig:toeFigure5)B).
+These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] supporting the implementation of joint management plans [@Burden:2019ij] and the call for new implemented shared policies [@Levin:2018in].
+However, game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.
 
@@
 
-There are two main sources of uncertainty in our analysis that can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
+There are two main sources of uncertainty in our analysis.
 First, our analysis relies on the combination of a single Earth system model (GFDL-ESM2M) and climate change scenario (RCP 8.5).
-The use of multiple ESMs can capture the structural uncertainty of the models [@Frolicher:2016dg; @Lotze:2019ce], however, exercises looking at changes in fish biomass with multiple models show an overall agreement between models in terms of direction of change but variable magnitude.
+The use of multiple ESMs can capture the structural uncertainty of the models [@Frolicher:2016dg; @Lotze:2019ce], however, exercises looking at changes in fish biomass with multiple models show an overall agreement between models in terms of direction of change but variable in magnitude.
 Further research that includes multiple ESMs and species distribution models could elucidate a broader range structural uncertainty of the climate, fish and fisheries models [@Lotze:2019ce].
 Secondly, due to limited biological and spatial-specific data on stocks sub-populations, our analysis uses political boundaries (EEZ) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
-On the other hand, in many EEZs, fisheries are often managed at the species level (Chapter One) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
-However, reproducing our analysis regionally, where spatially-explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
+On the other hand, in many EEZs, fisheries are often managed at the species level (Chapter 1) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
+However, reproducing our analysis regionally, where spatially explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
+Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
 
@@
 
-Global society has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) [@UnitedNations2018] by 2030; achieving this goal would have clear effects on multiple  others [@Singh:2017ds].
-Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@Pinsky:2018cb].
+Global society has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear effects on multiple others [@UnitedNations2018; @Singh:2017ds].
+Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@@Pinsky:2018cb].
 Here, we have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].

### 2020-11-24 Scripts/Supplements.Rmd
William comments 2nd round

@@
 
-```{r toeFigureS3, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
+```{r toeFigureS3, eval = T,  echo = F, fig.width=14, fig.height=7, fig.cap ="Sensitivity analysis of the number of grid cells with projected stock abundance within the neighbouring EEZs sharing the stock.
+The level used in this article was the top 95th percentile (e.g., far right column)."}
 
@@
 
-```{r toeFigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
+```{r toeFigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
 
@@
 
+```{r toeFigureS5, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
+
+img <- readPNG("../Figures/Supp5.png")
+grid.raster(img)
+
+```
+
 <!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005))](../Figures/Supp2.png) -->
 
-```{r toeFigureS5, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
+```{r toeFigureS6, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
 Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
@@
 
-img <- readPNG("../Figures/Supp5.png")
+img <- readPNG("../Figures/Supp6.png")
 grid.raster(img)

### 2020-11-25 Scripts/Initial_Analysis.Rmd
Final edits before sending  to comm.

@@
 
-```{r Time_lapse_plot}
-
-#### Both plots for tresh one
-#  Embarrasing coding Juliano!
-
-ggplot() +
-  geom_area(data = subset(plot_data, level == "stock"),
+```{r Time_lapse_plot, eval = T, echo = F}
+
+### Three plots
+
+count_eez_plot <- ggplot() +
+  geom_area(data = subset(plot_data, metric == "com_sum" & level == "eez"),
             aes(
@@
             ),
-            # alpha= 1,
             size=0.5,
@@
   )  +
-  geom_area(data = subset(plot_data, level == "eez"),
+  scale_y_continuous("Cumulative number of EEZs (n)",
+                     expand = c(0,0)
+                     ) +
+  scale_x_continuous("",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     # label = c(seq(2000,2075,25),">2100"),
+                     expand = c(0, 0)
+  ) +
+  my_ggtheme_p() +
+  theme(legend.position = "") +
+  guides(fill=guide_legend(title= "Level", reverse=FALSE)) +
+  scale_fill_viridis(discrete = T,
+                     begin = 0.1, 
+                     end = 0.1
+  )
+
+#### Count stock plot 
+count_stock_plot <- ggplot() +
+  geom_area(data = subset(plot_data, metric == "com_sum" & level == "stock"),
             aes(
@@
             ),
-            # alpha= 0.8,
             size=0.5,
@@
   )  +
-  # Shameslly plot it again to fix colors
+  scale_y_continuous("Cumulative number of stocks (n)",
+                     expand = c(0,0)
+                     ) +
+  scale_x_continuous("Time of emergence of transboundary index",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     # label = c(seq(2000,2075,25),">2100"),
+                     expand = c(0, 0)
+  ) +
+  my_ggtheme_p() +
+  theme(legend.position = "") +
+  guides(fill=guide_legend(title= "Level", reverse=FALSE)) +
+  scale_fill_viridis(discrete = T,
+                     begin = 0.7, 
+                     end = 0.7
+  )
+
+# Percentage plot
+percentage_plot <- ggplot() +
+  geom_area(data = subset(plot_data, level == "eez" & my_facet == "Percentage (%)"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  ) + 
   geom_area(data = subset(plot_data, level == "stock" & my_facet == "Percentage (%)"),
@@
               x = as.numeric(taxon_emer),
-              # x = reorder(as.numeric(taxon_emer),level),
               y = value,
@@
   ) + 
-  # geom_line(data = subset(plot_data, metric == "com_sum"),
-  #           aes(
-  #             x = as.numeric(taxon_emer),
-  #             y = total,
-  #             fill = level
-  #           ),
-  #           colour="grey30",
-  # )  +
-  # geom_text(data = subset(plot_data, metric == "com_sum"),
-  #           aes(
-  #             x = 2037,
-  #             y = total+5,
-  #             label = label
-  #           ),
-  #           size = 3,
-  #           colour="grey30",
-  #           hjust=0.6,
-  # )  +
-  facet_wrap(~my_facet,
-             scales = "free") + 
-  scale_y_continuous("",
+  scale_y_continuous("Cumulative proportion of the total (%)",
                      expand = c(0,0)
   ) +
-  scale_x_continuous("Time of emergence",
+  scale_x_continuous("",
                      breaks = seq(2000,2100,25),
                      limits = c(2005,2100),
-                     label = c(seq(2000,2075,25),">2100"),
+                     # label = c(seq(2000,2075,25),">2100"),
                      expand = c(0, 0)
@@
     legend.position = "right",
-    strip.text.x = element_blank(),
-    panel.spacing = unit(3, "lines") # space between plots
+    strip.text.x = element_blank()
   ) +
@@
                      labels = c("EEZs","Stocks")
-  ) +
-  ggsave("./Figures/cumu_per_freq.png",
-  width = 10,
-  height = 5,
-  units = "in"
+  )
+
+# Legend
+  legend <- get_legend(percentage_plot)
+
+  percentage_plot <- percentage_plot + 
+    theme(legend.position = "")
+
+# Draw figure
+  plot_all <- ggdraw() +
+    # Latin America and the Car.
+    draw_plot(count_eez_plot, x = 0.01, y = 0, width = 0.28, height = 1) +
+    # Europe
+    draw_plot(count_stock_plot, x = 0.28, y = 0, width = 0.28, height = 1) +
+    # Percentage plot
+    draw_plot(percentage_plot, x = 0.56, y = 0 , width = 0.28, height = 1) +
+    # legend
+    draw_plot(legend, x = 0.84, y = 0 , width = 0.28, height = 1)
+
+  # Save plot
+  save_plot(
+    "./Figures/cumu_per_freq.png",
+    plot_all,
+    base_height = 5,
+    base_width = 12
   )
@@
   ggsave("./Figures/bubble_plot.png",
-         width = 8,
+         width = 12,
          height = 8,

### 2020-11-25 Scripts/Manuscript.Rmd

@@
 ---
+title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change
+subtitle: Challenges to transboundary fisheries under climate change
+author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^'
+date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern.'
 output:
-  bookdown::pdf_book:
-    toc : no
-    keep_tex: no
-    number_sections: yes
-geometry: margin=1in
-editor_options:
-  chunk_output_type: inline
+  # word_document:
+    # reference_docx: ../References/template.docx
+  pdf_document: default
 bibliography: ../References/ToE_Ref.bib
+csl: ../References/science.csl
 header-includes:
@@
 ---
-
-<!-- --- -->
-<!-- title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change -->
-<!-- subtitle: Challenges to transboundary fisheries under climate change -->
-<!-- author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^' -->
-<!-- date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
-^2^Climate and Environmental Physics, Physics Institute, University of Bern.' -->
-<!-- output: -->
-<!--   # word_document:  -->
-<!--     # reference_docx: ../References/template.docx -->
-<!--   pdf_document: default -->
-<!-- bibliography: ../References/ToE_Ref.bib -->
-<!-- csl: ../References/science.csl -->
-<!-- header-includes: -->
-<!-- - \usepackage{amsmath} -->
-<!-- - \usepackage{float} # Use the 'float' package -->
-<!-- - \floatplacement{figure}{H} # Make every figure with caption = h -->
-<!-- --- -->
 
@@
 Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn].
-Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk].
-).
-As a result of climate change, many marine species have changed their distribution towards higher latitude, deeper water or following local temperature gradients [@Poloczanska:2016kk].
-Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt] impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].The projected risks of impacts can be reduced through improving the effectiveness of current fisheries managemen [@Gaines:2018sg; @Cheung:2018dh], including managing fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
+Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk]).
+As a result of climate change, many marine species have changed their distributions towards higher latitude, deeper water or following local temperature gradients [@Poloczanska:2016kk].
+Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].The projected risks of impacts can be reduced through improving the effectiveness of current fisheries managemen [@Gaines:2018sg; @Cheung:2018dh], including the management of fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
 
 The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
-As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (*i*) transboundary, those that cross neighboring EEZs; (*ii*) straddling, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory, mainly tunas and bill-fishes that cover vast oceanic regions including both the high seas and EEZs [@Munro:2004th].
+As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (*i*) transboundary stocks, those that cross neighboring EEZs; (*ii*) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs [@Munro:2004th].
 While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
@@
 Here, we estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
-We rely on a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
+We employ a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
 We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks, excluding the high seas in this study).
@@
 After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.7% per year.
-By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure Figure \@ref(fig:toeFigure1)).
+By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:toeFigure1)).
 Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020.
-Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century and are not expected to stop there (Figure Figure \@ref(fig:toeFigure1)).
+Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure Figure \@ref(fig:toeFigure1)).
 We based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
 Results from the sensitivity analysis suggest that our method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g.
-the top 90%) of the stock (Figure Figure \@ref(fig:toeFigures·3).
+the top 90%) of the stock (Figure \@ref(fig:toeFigureS3).
 
@@
 However, notable exceptions of tropical EEZs with a later average time of emergence across stocks include Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070).
-Moreover, despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) will have relative early average time of emergence (Figure \@ref(fig:toeFigure2)A).
+Moreover, despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) will have relatively early average time of emergence (Figure \@ref(fig:toeFigure2)A).
 Only in a few cases, the transboundary index of a stock will not emerge between 2006 and 2100.
 In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
-In the case of Brazil, we included four transboundary stocks, all shared with Uruguay [@PalaciosAbrantes:2020]: Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
-In all cases, our study suggest that the transboundary index of these stocks will not emerge before 2100 relative to the historic (1951-2005) distribution \@ref(fig:toeFigure2)A)
+In the case of Brazil, we included four transboundary stocks, all shared with Uruguay [@PalaciosAbrantes:2020tv]: Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
+In all cases, our study suggest that the transboundary index of these stocks will not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A)
 
@@
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent points outside of 1.5*IQR."}
+Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range."}
 
@@
 
-On average, stocks that will see their transboundary index emerge by 2100 represent 27% ± 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010.
+On average, stocks that will see their transboundary index emerge by 2100 represent 27% \pm 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010.
 However, large variation exists in the proportion that stocks with emerging transboundary index have in the current revenue participation across EEZs.
-In some cases like El Salvador and New Zealand, stock’s with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries  in their respective EEZs between 2004 and 2014.
+In some cases e.g., El Salvador and New Zealand, stock’s with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries  in their respective EEZs between 2004 and 2014.
 On the contrary, stocks with emerging transboundary index from the Marshal Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A).
-In some cases, a few emerging stocks represent large proportion of revenue from transboundary stocks such as Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively).
-On the contrary, Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the participation of such stocks on fishing revenue is less than 20%.
+In addition, some EEZs are projectede to have a few emerging stocks that represent large proportion of revenue from transboundary stocks e.g., in Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively).
+Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks on fishing revenue is less than 20%.
 
@@
 (67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods).
-We tested the sensitivity of our methods by considering a larger threshold for the time of emergence results (i.e., 2 s.d for a 95% confidence).
-The global average time of emergence of the transboundary index considering a 2 s.d.
-threshold was 2036 \pm 28, a difference of about 7 years relative to the 1 s.d.
+We tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence).
+The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 \pm 28, a difference of about 7 years relative to the 1 s.d.
 threshold (Figure \@ref(fig:toeFigureS2)).
@@
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers outside of 1.5 times the interquartile range."}
+Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range."}
 
@@
 Our results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
-These EEZs will see one third (33% \pm 21%) of their transboundary stocks changing the share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A – Land).
-Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin will see changes in all of the transboundary stocks analyzed while Venezuela (6%), Bonaire (9%), China (10%) and Viet Nam (10%) will be the only countries with less than 10% of stocks changing beyond the threat point.
+These EEZs will see one third (33% \pm 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A).
+Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin will see changes in all of the transboundary stocks analyzed while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) will be the only countries with less than 10% of stocks changing beyond the threat point.
 
@@
 
-Our results offer no clear global pattern in the change of stock share ratio (Figure \@ref(fig:toeFigure5)).
-In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, stock share ratio is expected to follow a poleward shift.
-However, central America Pacific and West African coasts are expected to follow an equatorial direction.
+Changes of stock share ratio are largely related to regional changes in biogeography and the geometry of the EEZs (Figure \@ref(fig:toeFigure5)).
+In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa stock share ratio is expected to follow the projected poleward shift of transboundary species under climate change.
+However, central America Pacific and West African coasts are expected to follow an equatorial direction as range shifts of the transboundary species are following local oceanographic gradients with relatively lower temperature waters in lower latitude [@Clarke:2020ep].
 Moreover, some regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no stablished pattern, this is the case of the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).
@@
 A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs].
-In the northeast Atlantic, Atlantic mackerel (Scomber scombrus) is multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
+In the northeast Atlantic, the Atlantic mackerel (Scomber scombrus) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
 However, a range expansion of Atlantic mackerel (Scomber scombrus) into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008 without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij].
 Such changes in quota resulted in disputes between Iceland and Faroes Islands, and the NEAFC member states [@Spijkers:2017ij].
-As transboundary stocks continue to emerge in the incoming years, imal-adapted fisheries management and international disputes like that of A.
-mackerel are likely to increase in frequency [@Pinksy2018]
+As transboundary stocks continue to emerge in the incoming years, mal-adapted fisheries management and international disputes like that of Atlantic mackerel are likely to increase in frequency [@Pinsky:2018cb]
 
@@
 This thermal tolerance explains the sensitivity of the animal to warming and is a main driver of species’ distributional change [@IPCC:2019tn].
-Oxygen plays a similar role to that of temperature [Clarke:2020ep], especially on tuna species that require large amounts of dissolve oxygen for they metabolism [@IPCC2014].
+Oxygen plays a similar role to that of temperature [@Clarke:2020ep], especially on tuna species that require large amounts of dissolve oxygen for their metabolism [@IPCC:2014bz].
 It has been estimated that since the early 2000s, increase in sea surface temperature (SST) and decrease in oxygen content have emerged in 79% and 23% of the global ocean, respectively, and by 2080, SST is projected to emerge in 90% of the global ocean while O2 will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq].
-Moreover, before 2100, surface acidification, nitrate content and net primary production, also important variables for species distributions, will have emerged from the historical variability of 1861-1900 in 30% and 60% of the ocean under a low (RCP 2.6) and strong (RCP 8.5) climate change scenario respectively [@IPCC:2019tn].
+Moreover, before 2100, surface acidification, nitrate content and net primary production, also important variables for species distributions, will have emerged from the historical variability of 1861-1900 in 30% and 60% of the ocean under a low (RCP 2.6) and strong (RCP 8.5) climate change scenario, respectively [@IPCC:2019tn].
 These oceanographic changes are important drivers of change in species distributions as represented in the DBEM [@Cheung:2016jd; @Cheung:0ur].
@@
 
-The lack of Arctic stocks with emerging transboundary index is likely a combination of the few species currently fished in this region.
-Historically, there are 59 commercial fish stocks fished the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas with limited knowledge about the species biology [@Christiansen:2014gg].
+The presentce of EEZs lacking stocks with emerging transboundary index is a combination of EEZ shape, climate projections, and a reduced number of transbounudary stocks.
+For example, there are 59 commercial fish stocks fished the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas with limited knowledge about the species biology [@Christiansen:2014gg].
 However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@REF].
@@
 Moreover, projections suggest a larger increment on transboundary stocks in the incoming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further North [@Morley:2018fn; @Frainer:2017ji].
-For Brazil and Uruguay, the current and future distribution of the transboundary stocks in combination with the particular characteristics of these EEZs could be influencing the results.
+In the Case of Brazil, in addition to having few transbounudary stocks (only 4 shared with Uruguay in this study), the large diffrence in the EEZs lenghts could also be influencing the results.
+
+<!-- For Brazil and Uruguay, the current and future distribution of the transboundary stocks in combination with the particular characteristics of these EEZs could be influencing the results.
 First, both Argentine menhaden and Argentine shortfin squid have distributions with larger biomass concentration further south, with Brazil serving as the northern limit of the distributional range [@AquaMaps:2019ui; @AquaMaps:2019wt].
@@
 In the case of the two shrimp species, their destitution range spams along the cost of Brazil and is not expected to have a strong shift in the incoming years, also resulting in a stable transboundary index [@AquaMaps:2019wn; @AquaMaps:2019wj].
+-->
 
@@
 
+Changes in stock share ratio beyond the threat point of countries sharing a common resource could instigate international conflict and hinder bi-lateral agreements (@Miller:2013iv; @Sumaila:2020tp).
+Lessons learned from historic distributional shifts of shared stocks suggest that treaties that are not prepare to respond to such shifts will be less resilient to changes in transboundary fish stocks share ratio due to climate change [@Miller:2013iv; @Sumaila:2020tp].
+For example, disputes over Pacific salmon between Canada and the USA arose in the 1990s when a climate-related shift in stock abundance favoured Alaska’s salmon fisheries over Canada’s [@Miller:2013iv; @Munro:2002uf; @Song:2017va].
+The conflict lasted over ten years until Canada and the USA agreed on a mutual conservation fund where the US would contribute the larger proportion to support scientific research, habitat restoration and enhancement of wild stock production in transboundary rivers [@Munro:2002uf].
+Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Chapter5; @Pinsky2018].
+Side payments are mechanisms that provide a solution to the inequality produced by shifts in the stock’s distribution [@Tunca:2019if; @Miller:2004fpa].
+These can be monetary (e.g., the stock-“winning” State compensates the stock-“loosing” state for the proportional shift) or other methods like the Pacific salmon case (e.g., a conservation fund payed mainly by one Party) [@Miller:2004fpa; @Miller:2013iv].
+Strengthening current international cooperation will be required as stock distribution will move to neighboring jurisdictions, requiring combine efforts to generate information and set proper management rules [@Miller:2013iv].
+Finally, flexible management rules might be better fitted to capture the stock’s distributional shift.
+For example, quota allocation based on the stock’s current distribution like that for Pacific halibut (*Hippoglossus stenolepis*) managed by Canada and the US [@Sumaila:2020tp; @PalaciosAbrantes:2020vt] might be better able to capture the stock’s shifts than a quota based on a fix-historical proportion like the case of the EU [@Baudron:2020dj].
+
 ## Caveats and uncertainties
@@
 There are two main sources of uncertainty in our analysis.
-First, our analysis relies on the combination of a single Earth system model (GFDL-ESM2M) and climate change scenario (RCP 8.5).
+First, our analysis relies on the combination of a single Earth system model (GFDL-ESM2M), fish and fisheries model, and climate change scenario (RCP 8.5).
 The use of multiple ESMs can capture the structural uncertainty of the models [@Frolicher:2016dg; @Lotze:2019ce], however, exercises looking at changes in fish biomass with multiple models show an overall agreement between models in terms of direction of change but variable in magnitude.
 Further research that includes multiple ESMs and species distribution models could elucidate a broader range structural uncertainty of the climate, fish and fisheries models [@Lotze:2019ce].
-Secondly, due to limited biological and spatial-specific data on stocks sub-populations, our analysis uses political boundaries (EEZ) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
-On the other hand, in many EEZs, fisheries are often managed at the species level (Chapter 1) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
+Secondly, due to limited biological and spatially-specific data on stocks sub-population structure of transboundary species, our analysis uses political boundaries (EEZ) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
+On the other hand, in many EEZs, fisheries are often managed at the species level [@PalaciosAbrantes:2020tv] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
 However, reproducing our analysis regionally, where spatially explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].

### 2020-11-26 Scripts/Manuscript.Rmd
small details

@@
 
-Here, we estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
-We employ a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
-We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks, excluding the high seas in this study).
+Here, I estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
+I employ a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
+I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks, excluding the high seas in this study).
 Our results suggest that many transboundary stocks have already ‘emerged’ from historical natural variability and overpassed their share threat point in most neighboring EEZs, posing a threat to the sustainability of shared stocks and the resilience of international treaties.
@@
 This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv]; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks in this analysis.
+Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks in this analysis.
 Using the transboundary species Atlantic cod (Gadus morhua) as an example, based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of G.
 morhua, but the United States and Greenland do not share a stock of this transboundary species.
-We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; \@ref(fig:toeFigureS1))) including island territories.
-We determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
+The boundaries of the world’s EEZs were defined using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:toeFigureS1))) including island territories.
+The intersections between polygons were defined using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
 Each EEZ was categorized by the geopolitical regions according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and by natural biome [@Reygondeau:2019uh].
 The habitat association of each species was determined based on information available from FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org).
-For each stock and EEZ, we used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
-We report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+For each stock and EEZ, I used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
+I report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
 The monthly average US Consumer Price Index (CPI) according to the U.S.
@@
 
-We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
+The distribution of marine species were proejcted using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
 The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
@@
 [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
-DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd], Cheung et al.
+DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
 *in revision*].
 
-We forced the DBEM with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+The DBEM was forced with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
 The GFDL-ESM2M is run under historical forcing until 2005 and follows the ‘high’ climate change emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
 In the GFDL-ESM2M, atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
-We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, I have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
+I do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
 Thus, for our experiment, each of the ten GFDL-ESM2M-ensembles were started from infinitesimally small differences in Earth system initial conditions resulting in a unique atmosphere and ocean state at each point in time.
@@
 
-We developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
-This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2).
+I developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
+This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2)).
 The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
@@
 where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
-To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top 95th percentile (e.g., per = 95%).
-We ran a sensitivity analysis using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
+To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, I included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top 95th percentile (e.g., per = 95%).
+I ran a sensitivity analysis using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
 The centroid of each EEZ was estimated using the st package in R (Figure \@ref(fig:toeFigureS1)).
-For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
+For each ensemble member, neighboring EEZs and transboundary stock, the distance between centroids was computed assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
 
@@
 where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively and $lon_{eez}$ and $lon_{ts}$ are the respective longitudes.
-Then, for each year we calculated the transboundary index as follows:
+Then, for each year I calculated the transboundary index as follows:
 
@@
 Knowing the point in time (e.g., year) at which the distribution of a shared stock will diverge from its natural internal variability is important in informing the lead-time to which climate adaptation needs to be implemented [@Link:2010ei].
-Here, we define time of emergence as the time when the stock shared between neighbouring EEZs will exceed the historical natural internal variability as follows:
+Here, I define time of emergence as the time when the stock shared between neighbouring EEZs will exceed the historical natural internal variability as follows:
 
@@
 where $\mu_{TI}$ is the yearly mean TI ($\mu$) across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005).
-We set a threshold of TI above which we consider a shared stock to have emerged from the historical internal variability of the TI.
+I set a threshold of TI above which it was consider a shared stock to have emerged from the historical internal variability of the TI.
 The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four), representing a probability of 68% that the transboundary index has emerged from historical variability).
-We tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
+I tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
 
@@
 However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., "threat point" [@Sumaila:2013hv].
-Here, we define threat point as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Here, I define threat point as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
 Any proportion below the defined threat point of a country would result in a deflection from the joint management of the shared stock.
 
-First we estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
-We did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
-Second, we average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
+First I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
+I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
+Second, I average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
 The first time period ($SSR_{th}$) spans from 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
 It is assumed as the historic baseline of shared distributions for each stock.
-We then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$).
-We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
+I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$).
+I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
 The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members.
-Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$.
+Third, I defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$.
 This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{et} \geqq (SSR_{th} + \sigma)$ or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
-Finally, we estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al.
+Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al.
 (*in press*),
@@
 
-We tested the results for normality (e.g.
+The time of emergence results were tested for normality (e.g.
 skewness = 1.02, kurtosis = 2.83) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
-Specifically, we tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of transboundary stocks.
+Specifically, I tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of the transboundary index of transboundary stocks.
 For both cases, our null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
@@
 
-The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 \pm 26 years (Figure \@ref(fig:toeFigure1)).
+The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:toeFigure1)).
 In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100.
@@
 Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020.
-Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure Figure \@ref(fig:toeFigure1)).
-We based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
+Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure \@ref(fig:toeFigure1)).
+I based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
 Results from the sensitivity analysis suggest that our method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g.
-the top 90%) of the stock (Figure \@ref(fig:toeFigureS3).
+the top 90%) of the stock (Figure \@ref(fig:toeFigureS3)).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone (EEZ) and transboundary stock: A) results in absolute numbers (green).
-B) results expressed as a percentage of the total emerging stocks and EEZ with emerging stock (purple)."}
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone (EEZ) and transboundary stock."}
 
@@
 In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
-In the case of Brazil, we included four transboundary stocks, all shared with Uruguay [@PalaciosAbrantes:2020tv]: Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
+In the case of Brazil, there were four transboundary stocks included, all shared with Uruguay [@PalaciosAbrantes:2020tv]: Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
 In all cases, our study suggest that the transboundary index of these stocks will not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A)
@@
 
-On average, stocks that will see their transboundary index emerge by 2100 represent 27% \pm 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010.
+On average, stocks that will see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010.
 However, large variation exists in the proportion that stocks with emerging transboundary index have in the current revenue participation across EEZs.
@@
 
-We set an arbitrary threshold of 1 s.d.
+I set an arbitrary threshold of 1 s.d.
 (67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods).
-We tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence).
-The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 \pm 28, a difference of about 7 years relative to the 1 s.d.
+I tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence).
+The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
 threshold (Figure \@ref(fig:toeFigureS2)).
@@
 
-We compared species time of emergence within EEZs by species’ habitat association (Figure \@ref(fig:toeFigure4)).
+I compared species time of emergence within EEZs by species’ habitat association (Figure \@ref(fig:toeFigure4)).
 Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods).
-We found significant differences in the time of emergence of stocks range shifts, indicated by the transbounudary index, by their specie's habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
+I found significant differences in the time of emergence of stocks range shifts, indicated by the transbounudary index, by their specie's habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
 Specifically, time of emergence of range shifts of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
@@
 
-We have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods - equation \@ref(eq:five)).
-By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% \pm 15% (Figure \@ref(fig:toeFigure5)A).
-The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)).
+I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods - equation \@ref(eq:five)).
+By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A).
+The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4); \@ref(fig:toeFigureS5)).
 However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high.
@@
 Our results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
-These EEZs will see one third (33% \pm 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A).
+These EEZs will see one third (33% $\pm$ 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A).
 Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin will see changes in all of the transboundary stocks analyzed while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) will be the only countries with less than 10% of stocks changing beyond the threat point.
@@
 Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point.
-Panel B zooms in to specific areas shown within the grey dashed line boundaries in A.s"}
+Panel B zooms in to specific areas shown within the grey dashed line boundaries in A."}
 
@@
 
-In this paper, we investigate the timing and intensity of changes in the sharing of transboundary stocks between countries.
+In this paper, I investigate the timing and intensity of changes in the sharing of transboundary stocks between countries.
 Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
 Such early emergences of range shifts indicate the urgency of the challenges that climate change is posing on international fisheries management.
-Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management.
-Finally, we show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to level below that conduce cooperative resource management between countries.
+Moreover, I identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management.
+Finally, I show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to level below that conduce cooperative resource management between countries.
 Adaptation measures are needed to reduce the impacts of species range shifts on current international transboundary fisheries treaties.
@@
 For example, there are 59 commercial fish stocks fished the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas with limited knowledge about the species biology [@Christiansen:2014gg].
-However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@REF].
+However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@IPCC:2019tn].
 As a result, the biomass of Arctic marine animals [@BryndumBuchholz:2019kh] and catch and revenue of associated fisheries [@Tai:2019cq; @Lam:2014fq] are expected to increase in the incoming years, although substantial uncertainty exists on the magnitude of change [@BryndumBuchholz:2019kh].
@@
 The conflict lasted over ten years until Canada and the USA agreed on a mutual conservation fund where the US would contribute the larger proportion to support scientific research, habitat restoration and enhancement of wild stock production in transboundary rivers [@Munro:2002uf].
-Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Chapter5; @Pinsky2018].
+Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Pinsky:2018cd].
 Side payments are mechanisms that provide a solution to the inequality produced by shifts in the stock’s distribution [@Tunca:2019if; @Miller:2004fpa].
@@
 Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@@Pinsky:2018cb].
-Here, we have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
-We have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
+Here, I have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
+I have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
 Our results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management.
@@
 
-```{r toeFigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
+```{r FigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
 
@@
 
-```{r toeFigureS2, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+```{r FigureS2, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
 The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel).
@@
 
-```{r toeFigureS3, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
+```{r FigureS3, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
 
@@
 
-```{r toeFigureS4, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
+```{r FigureS4, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
 

### 2020-11-26 Scripts/Manuscript.Rmd
Changed we for I and Palacios for CHapter

@@
 ---
-title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change
-subtitle: Challenges to transboundary fisheries under climate change
-author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^'
-date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern.'
 output:
-  # word_document:
-    # reference_docx: ../References/template.docx
-  pdf_document: default
+  bookdown::pdf_book:
+    toc : no
+    keep_tex: no
+    number_sections: yes
+geometry: margin=1in
+editor_options:
+  chunk_output_type: inline
 bibliography: ../References/ToE_Ref.bib
-csl: ../References/science.csl
 header-includes:
@@
   "knitr",
- "kableExtra"
+ # "kableExtra"
   # "here", # for dbem_import `here()`
@@
 While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
+A recent study estimates that, globally, there are 633 transboundary fish species, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 (Chapter 2).
 
 The effectiveness of managing fisheries targeting shared stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
-In many cases, catch or fishing effort quota on shared stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change in the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+In many cases, catch or fishing effort quota on shared stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change in the fish stocks and fisheries [@Sumby:2021gw, Chapter 2].
 Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by worsening climate change [@Pinsky:2018cb].
@@
 Here, I estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
-I employ a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
-I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks, excluding the high seas in this study).
+I employ a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared (Chapter 2) by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
+I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (Chapter 2) and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks, excluding the high seas in this study).
 Our results suggest that many transboundary stocks have already ‘emerged’ from historical natural variability and overpassed their share threat point in most neighboring EEZs, posing a threat to the sustainability of shared stocks and the resilience of international treaties.
@@
 
-This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks in this analysis.
+This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 (Chapter 2).
+Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd, Chapter  1], resulting in a total of 9,132 transboundary stocks in this analysis.
 Using the transboundary species Atlantic cod (Gadus morhua) as an example, based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of G.
 morhua, but the United States and Greenland do not share a stock of this transboundary species.
-The boundaries of the world’s EEZs were defined using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:toeFigureS1))) including island territories.
-The intersections between polygons were defined using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
+I defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; \@ref(fig:toeFigureS1))) including island territories.
+I determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
 Each EEZ was categorized by the geopolitical regions according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and by natural biome [@Reygondeau:2019uh].
@@
 
-The distribution of marine species were proejcted using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
+I projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
 The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
@@
 [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
-DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
+DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd], Cheung et al.
 *in revision*].
@@
 In the GFDL-ESM2M, atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, I have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
+Because the main approach of this chapter relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
 I do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
@@
 I developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
-This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2)).
+This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2).
 The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
@@
 To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, I included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top 95th percentile (e.g., per = 95%).
-I ran a sensitivity analysis using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
+A sensitivity analysis was ran using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
 The centroid of each EEZ was estimated using the st package in R (Figure \@ref(fig:toeFigureS1)).
-For each ensemble member, neighboring EEZs and transboundary stock, the distance between centroids was computed assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
+For each ensemble member, neighboring EEZs and transboundary stock, I computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
 
@@
 where $\mu_{TI}$ is the yearly mean TI ($\mu$) across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005).
-I set a threshold of TI above which it was consider a shared stock to have emerged from the historical internal variability of the TI.
+I set a threshold of TI above which I consider a shared stock to have emerged from the historical internal variability of the TI.
 The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four), representing a probability of 68% that the transboundary index has emerged from historical variability).
@@
 
-<!-- We created a country-specific vulnerability index (VI) to examine the risk of impacts of species’ range shifts on transboundary fisheries management.
+<!-- I created a country-specific vulnerability index (VI) to examine the risk of impacts of species’ range shifts on transboundary fisheries management.
 The VI reflects both the rate of change of shared stocks’ distribution (indicated by the time-of-emergence of the transboundary index) and the importance of the shifting species to a country’s total catch and revenue from the species.
@@
 However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., "threat point" [@Sumaila:2013hv].
-Here, I define threat point as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Here, I define threat point as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@Sumaila:2020tp, Chapter Four].
 Any proportion below the defined threat point of a country would result in a deflection from the joint management of the shared stock.
@@
 First I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
-I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
+I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2).
 Second, I average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
@@
 skewness = 1.02, kurtosis = 2.83) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
-Specifically, I tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of the transboundary index of transboundary stocks.
+Specifically, I tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of transboundary stocks.
 For both cases, our null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
@@
 
-The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:toeFigure1)).
+The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 \pm 26 years (Figure \@ref(fig:toeFigure1)).
 In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100.
@@
 Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020.
-Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure \@ref(fig:toeFigure1)).
+Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure Figure \@ref(fig:toeFigure1)).
 I based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
 Results from the sensitivity analysis suggest that our method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g.
-the top 90%) of the stock (Figure \@ref(fig:toeFigureS3)).
+the top 90%) of the stock (Figure \@ref(fig:toeFigureS3).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone (EEZ) and transboundary stock."}
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone (EEZ) and transboundary stock: A) results in absolute numbers (green).
+B) results expressed as a percentage of the total emerging stocks and EEZ with emerging stock (purple)."}
 
@@
 In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
-In the case of Brazil, there were four transboundary stocks included, all shared with Uruguay [@PalaciosAbrantes:2020tv]: Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
+In the case of Brazil, four transboundary stocks were included, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
 In all cases, our study suggest that the transboundary index of these stocks will not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A)
@@
 
-On average, stocks that will see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010.
+On average, stocks that will see their transboundary index emerge by 2100 represent 27% \pm 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010.
 However, large variation exists in the proportion that stocks with emerging transboundary index have in the current revenue participation across EEZs.
@@
 I tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence).
-The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
+The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 \pm 28, a difference of about 7 years relative to the 1 s.d.
 threshold (Figure \@ref(fig:toeFigureS2)).
@@
 Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods).
-I found significant differences in the time of emergence of stocks range shifts, indicated by the transbounudary index, by their specie's habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
+Significant differences were found in the time of emergence of stocks range shifts, indicated by the transbounudary index, by their specie's habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
 Specifically, time of emergence of range shifts of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
@@
 I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods - equation \@ref(eq:five)).
-By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A).
-The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4); \@ref(fig:toeFigureS5)).
+By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% \pm 15% (Figure \@ref(fig:toeFigure5)A).
+The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)).
 However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high.
@@
 Our results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
-These EEZs will see one third (33% $\pm$ 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A).
+These EEZs will see one third (33% \pm 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A).
 Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin will see changes in all of the transboundary stocks analyzed while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) will be the only countries with less than 10% of stocks changing beyond the threat point.
@@
 Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point.
-Panel B zooms in to specific areas shown within the grey dashed line boundaries in A."}
+Panel B zooms in to specific areas shown within the grey dashed line boundaries in A.s"}
 
@@
 
-In this paper, I investigate the timing and intensity of changes in the sharing of transboundary stocks between countries.
+In this chapter, I investigate the timing and intensity of changes in the sharing of transboundary stocks between countries.
 Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
@@
 For example, there are 59 commercial fish stocks fished the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas with limited knowledge about the species biology [@Christiansen:2014gg].
-However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@IPCC:2019tn].
+However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@REF].
 As a result, the biomass of Arctic marine animals [@BryndumBuchholz:2019kh] and catch and revenue of associated fisheries [@Tai:2019cq; @Lam:2014fq] are expected to increase in the incoming years, although substantial uncertainty exists on the magnitude of change [@BryndumBuchholz:2019kh].
@@
 The conflict lasted over ten years until Canada and the USA agreed on a mutual conservation fund where the US would contribute the larger proportion to support scientific research, habitat restoration and enhancement of wild stock production in transboundary rivers [@Munro:2002uf].
-Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Pinsky:2018cd].
+Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Chapter5; @Pinsky2018].
 Side payments are mechanisms that provide a solution to the inequality produced by shifts in the stock’s distribution [@Tunca:2019if; @Miller:2004fpa].
@@
 Secondly, due to limited biological and spatially-specific data on stocks sub-population structure of transboundary species, our analysis uses political boundaries (EEZ) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
-On the other hand, in many EEZs, fisheries are often managed at the species level [@PalaciosAbrantes:2020tv] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
+On the other hand, in many EEZs, fisheries are often managed at the species level (Chapter 2) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
 However, reproducing our analysis regionally, where spatially explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
@@
 While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg; @PalaciosAbrantes:2020tv].
-To address the issues raised in this paper, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
+To address the issues raised in this chapter, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
 
@@
 
-```{r FigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
+```{r toeFigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
 

### 2020-12-03 Scripts/Manuscript.Rmd
Changes for dealing with word

@@
   "grid",
-  "knitr",
+  "knitr"#,
  # "kableExtra"
@@
 I tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence).
-The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 \pm 28, a difference of about 7 years relative to the 1 s.d.
+The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
 threshold (Figure \@ref(fig:toeFigureS2)).
@@
 I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods - equation \@ref(eq:five)).
-By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% \pm 15% (Figure \@ref(fig:toeFigure5)A).
+By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A).
 The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)).
@@
 The conflict lasted over ten years until Canada and the USA agreed on a mutual conservation fund where the US would contribute the larger proportion to support scientific research, habitat restoration and enhancement of wild stock production in transboundary rivers [@Munro:2002uf].
-Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Chapter5; @Pinsky2018].
+Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Pinsky:2018cb, Chapter 5].
 Side payments are mechanisms that provide a solution to the inequality produced by shifts in the stock’s distribution [@Tunca:2019if; @Miller:2004fpa].

### 2020-12-03 Scripts/Supplements.Rmd
Changes for dealing with word

@@
   "knitr",
-  "kableExtra",
+  # "kableExtra",
   "png",
@@
           "K-W Difference" = dif.com.difference) %>% 
-  kable("latex",
-        booktabs = T,
-        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  ) %>% 
-  kable_styling(full_width = F,
-                position = "center")
+  kable(caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions")
+  # kable("latex",
+        # booktabs = T,
+        # caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  # ) %>% 
+  # kable_styling(full_width = F,
+                # position = "center")
 
@@
           "K-W Difference" = dif.com.difference) %>% 
-  kable("latex",
+  kable(
         booktabs = T,
         caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  ) %>% 
-  kable_styling(full_width = F,
-                position = "center")
+  )  
+  # kable("latex",
+  #       booktabs = T,
+  #       caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  # ) %>% 
+  # kable_styling(full_width = F,
+  #               position = "center")
 

### 2020-12-08 Scripts/Initial_Analysis.Rmd
figures visuals

@@
              color="black",
-             size=4,
+             size=5,
              angle=0,
@@
              color="black", 
-             size=4, 
+             size=5, 
              angle=0, 
@@
     theme(
-      legend.box.margin = margin(t = -64, l = 249),
+      legend.box.margin = margin(t = -80, l = 40),
       legend.text = element_blank(),
@@
 
-  boxplot <- Eme_Res %>% 
+  boxplot_data <- Eme_Res %>% 
     filter(tresh == "Tresh_one",
@@
            !is.na(un_sub_region)) %>% 
-    mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and the Car.",paste(un_sub_region))) %>% 
-    ggplot() +
+    mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
+                                  ifelse(str_detect(un_sub_region,"Australia"),"Aus. and New Z.",
+                                         ifelse(str_detect(un_sub_region,"Sub-Saharan"),"Sub-Saharan\nAfrica",
+                                         paste(un_sub_region)))
+    )
+    )
+
+    boxplot_data$un_sub_region <- boxplot_data$un_sub_region %>% 
+  gsub("Northern","N.",.) %>% 
+  gsub("Southern","S.",.) %>% 
+  gsub("Western","W.",.) %>% 
+  gsub("Eastern","E.",.) %>% 
+  gsub("South-eastern","S.E.",.)
+
+  boxplot <- ggplot(boxplot_data) +
     geom_boxplot(
@@
     theme(
-      axis.text.x = element_text(size = 10, angle = 45, face = "plain", hjust = 1), 
-      strip.text.x = element_text(size = 14, colour = "black"), 
+      axis.text.x = element_text(size = 16, angle = 45, face = "plain", hjust = 1), 
+      strip.text.x = element_text(size = 18, colour = "black"), 
+      axis.title.y = element_text(size = 18),
+      axis.text.y = element_text(size = 16), 
       # axis.ticks.x = element_blank(),

### 2020-12-14 Scripts/Manuscript.Rmd
changes in the affiliations

--- 2020-11-26 Scripts/Manuscript.Rmd
@@
 subtitle: Challenges to transboundary fisheries under climate change
-author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2^, Gabriel Reygondeau^13^, U. Rashid Sumaila^1^, Colette Wabnitz^14^, and William W.L. Cheung^1^'
-date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern.'
+author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
+date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
+^2^Climate and Environmental Physics, Physics Institute, University of Bern.
+^3^Oeschger Centre for Climate Change Research, University of Bern, Switzerland.
+^4^School of Environmental Sciences, University of Liverpool.
+^5^Center for Ocean Solutions, Stanford University, Stanford, United States'
 output:
@@
 
-<!-- Climate change shifts in marine species distributions is affecting the current frameworks emplaced to manage species shared fisheries.
+Climate change shifts in marine species distributions is affecting the current frameworks emplaced to manage species shared fisheries.
 Here, we rely on a modeling framework with multiple ensemble members to identify the time frame by which shared stocks distribution will change from historical ranges, and the change in the proportion of the distrirbution that each EEZ has under climate change.
@@
 Our results suggest that most changes will happen before 2030 urging policy to be re-formulated to account for the possible implications of these changes.
--->
 
@@
 Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn].
-Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk]).
+Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk].
 As a result of climate change, many marine species have changed their distributions towards higher latitude, deeper water or following local temperature gradients [@Poloczanska:2016kk].
-Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].The projected risks of impacts can be reduced through improving the effectiveness of current fisheries managemen [@Gaines:2018sg; @Cheung:2018dh], including the management of fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
+Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].
+The projected risks of impacts can be reduced through improving the effectiveness of current fisheries managemen [@Gaines:2018sg; @Cheung:2018dh], including the fisheries management of species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
 
 The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
-As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (*i*) transboundary stocks, those that cross neighboring EEZs; (*ii*) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs [@Munro:2004th].
+As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, those that cross neighboring EEZs; (*ii*) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (*iV*) discrete stocks that are only present in the high seas [@Munro:2004th].
+This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
 While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
+A recent study estimates that, globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, and yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
 
-The effectiveness of managing fisheries targeting shared stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
-In many cases, catch or fishing effort quota on shared stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change in the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
-Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by worsening climate change [@Pinsky:2018cb].
-With future shifts in species distributions under climate change, the challenges posed on shared stocks fisheries management will continue to increase in the $21^{st}$ century.
-To prepare ocean governance to meet the challenges in resource management arising from shifting shared stocks, it is important to understand the time when climate change will affect the sharing distribution of these stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
+The effectiveness of fisheries management for transboundary stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
+In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb].
+With future shifts in species distributions under climate change, the challenges for the fisheries management of transboundary stocks will continue to increase during the $21^{st}$ century.
+In order to prepare ocean governance to meet the resource management challenges arising from shifting transboundary stocks, it is important to constrain when climate change will affect the dynamics of transboundary stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
 
 The timing and intensity of climate impacts on managing transboundary fish stocks can be quantitatively examined through the concepts of ‘time of emergence’ (ToE) and ‘threat point’ (TP).
-Time of emergence is defined as the moment in time when an anthropogenic change signal (e.g., future trend) rises above the background noise of natural variability (e.g., historical variation) [@IPCC:2019tn].
-This concept has been applied to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
+Time of emergence is defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (e.g., historical natural variation) [@Hawkins:2012hc] and has been applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
 The premise behind the time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
 The concept of threat point comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
-Game theory has been widely used to investigate and manage shared stocks [@Sumaila:2013hv].
-Specifically, Sumaila et al.
-(*in press*) applied this concept to understand the impacts of climate change to the economics of shared species between Canada and the United States.
+Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary species between Canada and the United States [@Sumaila:2020tp].
 
-Here, I estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
-I employ a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared [@PalaciosAbrantes:2020tv] by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
-I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks, excluding the high seas in this study).
-Our results suggest that many transboundary stocks have already ‘emerged’ from historical natural variability and overpassed their share threat point in most neighboring EEZs, posing a threat to the sustainability of shared stocks and the resilience of international treaties.
+Here, we estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
+We employ a species distribution model driven by output from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species [@PalaciosAbrantes:2020tv] across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
+We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude the high seas).
+We find that many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.
 
@@
 This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks in this analysis.
-Using the transboundary species Atlantic cod (Gadus morhua) as an example, based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of G.
-morhua, but the United States and Greenland do not share a stock of this transboundary species.
-The boundaries of the world’s EEZs were defined using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; Figure \@ref(fig:toeFigureS1))) including island territories.
-The intersections between polygons were defined using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
-Each EEZ was categorized by the geopolitical regions according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and by natural biome [@Reygondeau:2019uh].
-The habitat association of each species was determined based on information available from FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org).
-For each stock and EEZ, I used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
-I report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv]; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks in this analysis.
+Using the transboundary species Atlantic cod (*Gadus morhua*) as an example, and based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species.
+We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1)) including island territories.
+We determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
+Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
+The habitat association of each species was determined following the classification in FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org).
+For each stock and EEZ, we used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
+We report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
 The monthly average US Consumer Price Index (CPI) according to the U.S.
@@
 
-The distribution of marine species were proejcted using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
-The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
-For pelagic species, the model uses environmental variables at the surface (e.g., sea surface temperature) whereas demersal species distribution is driven by ocean bottom variables (e.g., bottom temperature).
-This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (see Cheung et al.
-[-@Cheung:2010dt; -@Cheung:2016jd] for model details).
-DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
+We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
+The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
+For pelagic species, the model uses environmental variables at the surface whereas demersal species distribution is driven by ocean bottom variables .
+DBEM then estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid (see Cheung et al.
+2010 and 2016 for model details).
+Importantly, DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd], Cheung et al.
 *in revision*].
 
-The DBEM was forced with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
-The GFDL-ESM2M is run under historical forcing until 2005 and follows the ‘high’ climate change emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
-In the GFDL-ESM2M, atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, I have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
-I do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
-Thus, for our experiment, each of the ten GFDL-ESM2M-ensembles were started from infinitesimally small differences in Earth system initial conditions resulting in a unique atmosphere and ocean state at each point in time.
-This approach allows us to estimate the natural internal variability (e.g., background noise) and isolate the forced climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
+We forced the DBEM with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+The GFDL-ESM2M is run under historical forcing until 2005 and follows the high greenhouse gas emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand distribution variability during both the historical and the future periods, to infer differences between time frames.
+We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
+Thus, for our experiment, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in year 1950 resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
+By design, variations among ensemble members are then solely due to natural variability.
+This approach allows us to estimate the natural internal variability (e.g., background noise) and isolate the forced anthropogenic climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
 
@@
 where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
-To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, I included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top 95th percentile (e.g., per = 95%).
-I ran a sensitivity analysis using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
+To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the  95th percentile (e.g., per = 95%).
+We ran a sensitivity analysis using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
 The centroid of each EEZ was estimated using the st package in R (Figure \@ref(fig:toeFigureS1)).
-For each ensemble member, neighboring EEZs and transboundary stock, the distance between centroids was computed assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
+For each ensemble member, neighboring EEZs and transboundary stock, the distance between centroids was computed assuming the Earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
 
@@
 where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively and $lon_{eez}$ and $lon_{ts}$ are the respective longitudes.
-Then, for each year I calculated the transboundary index as follows:
+Then, for each year we calculated the transboundary index (TI) as follows:
 
@@
 Knowing the point in time (e.g., year) at which the distribution of a shared stock will diverge from its natural internal variability is important in informing the lead-time to which climate adaptation needs to be implemented [@Link:2010ei].
-Here, I define time of emergence as the time when the stock shared between neighbouring EEZs will exceed the historical natural internal variability as follows:
+Here, we define time of emergence of the tranboundary index (ETI) as the time when the stock shared between neighbouring EEZs will exceed the historical natural internal variability as follows:
 
@@
 \label{eq:four}
-EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}} \\
+ETI = \frac{\mu_{TI[t]}}{SD_{TI_{th}}} \\
 ToE = t when Emergence index[t] > 1 s.d.
@@
 
-where $\mu_{TI}$ is the yearly mean TI ($\mu$) across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005).
-I set a threshold of TI above which it was consider a shared stock to have emerged from the historical internal variability of the TI.
+where $\mu_{TI}$ is the yearly mean TI across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005).
+We set a threshold of TI above which it was consider a shared stock to have emerged from the historical internal variability of the TI.
 The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four), representing a probability of 68% that the transboundary index has emerged from historical variability).
-I tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
+We tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
 
@@
 
-In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than operating on a non-cooperative basis [@Munro:1979dg].
-However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., "threat point" [@Sumaila:2013hv].
-Here, I define threat point as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-Any proportion below the defined threat point of a country would result in a deflection from the joint management of the shared stock.
+In game theory, cooperation over a transboundary resource will more likely result in a better overall outcome for the sharing parties than those operating on a non-cooperative basis [@Munro:1979dg].
+However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., “threat point” [@Sumaila:2013hv].
+Here, we define the threat point as the minimum required proportion of a transboundary stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Any proportion below the defined threat point of a country would result in a deflection from the joint management of the transboundary stock.
 
-First I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
-I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
-Second, I average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
+First we estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
+We did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
+Second, we average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
 The first time period ($SSR_{th}$) spans from 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
 It is assumed as the historic baseline of shared distributions for each stock.
-I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$).
-I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
+We then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$).
+We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
 The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members.
-Third, I defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$.
+Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$.
 This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{et} \geqq (SSR_{th} + \sigma)$ or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
-Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al.
+Finally, we estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al.
 (*in press*),
@@
 skewness = 1.02, kurtosis = 2.83) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
-Specifically, I tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of the transboundary index of transboundary stocks.
+Specifically, we tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of the transboundary index of transboundary stocks.
 For both cases, our null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
@@
 Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure \@ref(fig:toeFigure1)).
-I based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
+We based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
 Results from the sensitivity analysis suggest that our method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g.
@@
 (67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods).
-I tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence).
+We tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence).
 The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
@@
 Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods).
-I found significant differences in the time of emergence of stocks range shifts, indicated by the transbounudary index, by their specie's habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
+We found significant differences in the time of emergence of stocks range shifts, indicated by the transbounudary index, by their specie's habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
 Specifically, time of emergence of range shifts of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
@@
 
-In this paper, I investigate the timing and intensity of changes in the sharing of transboundary stocks between countries.
+In this paper, we investigate the timing and intensity of changes in the sharing of transboundary stocks between countries.
 Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
 Such early emergences of range shifts indicate the urgency of the challenges that climate change is posing on international fisheries management.
-Moreover, I identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management.
-Finally, I show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to level below that conduce cooperative resource management between countries.
+Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management.
+Finally, we show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to level below that conduce cooperative resource management between countries.
 Adaptation measures are needed to reduce the impacts of species range shifts on current international transboundary fisheries treaties.
@@
 Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@@Pinsky:2018cb].
-Here, I have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
-I have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
+Here, we have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
+We have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
 Our results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management.

### 2020-12-14 Scripts/Manuscript.Rmd
Thomas and Alessandro comments

--- 2020-12-03 Scripts/Manuscript.Rmd
@@
 
-<!-- # Instructions  -->
-
-<!-- Research articles should present a major advance and can be short or long (maximum 15,000 words) format.
-Submissions must include an abstract (up to 150 words), an introduction, and sections with brief informative subheadings.
-You may include up to ten figures and/or tables and about 60 references.
-Supplementary materials should be limited to information that is not essential for the general understanding of the research presented in the main text and can include data sets, figures, tables, videos, or audio files.
-You may include up to a total of ten figures and/or tables (combined) throughout the supplemental text.
--->
-
 ```{r setup, eval = T, echo = F, warning = F, message = F, results = 'hide'}
@@
   "knitr"#,
- # "kableExtra"
-  # "here", # for dbem_import `here()`
-  # "data.table", #dbem_import `fread()`
-  # "readxl", # for reading excell files
-  # "janitor", # for clearing names
-  # "geosphere", # estimate distances between points `distm()`
-  # "ggrepel",
-  # "zoo", # for average mean
-  # "parallel", # for mclapply,
-  # "sf", # for mapping
-  # "st", # for mapping
-  # "rgdal", #Spatial analysis
-  # "tools", #Spatial analysis 
-  # "zeallot", # for Juanito's map
-  # "gmt", # for estimating distances between points
-  # "viridis"
 )
@@
 
-<!-- * corresponding author: j.palacios@oceans.ubc.ca -->
-
-<!-- # Abstract (150  words) -->
-
-<!-- An opening sentence that states the question/problem addressed by the research AND Enough background content to give context to the study AND A brief statement of primary results AND A short concluding sentence.
--->
-
-<!-- Climate change shifts in marine species distributions is affecting the current frameworks emplaced to manage species shared fisheries.
-Here, we rely on a modeling framework with multiple ensemble members to identify the time frame by which shared stocks distribution will change from historical ranges, and the change in the proportion of the distrirbution that each EEZ has under climate change.
-Our results suggest that 60% of shared species have emerged since 2006 in 87% of the world EEZs.
-By 2075 all EEZs will have an emerging species and by 2097 and all species will have emerged in at least one EEZ.
-Average changes in the ratio of species-share between neighboring EEZs will fluctuate between 1 and 10% with some exceeding 25% by 2030 relative to today’s ratio.
-Our results suggest that most changes will happen before 2030 urging policy to be re-formulated to account for the possible implications of these changes.
--->
-
 ## Introduction
@@
 As a result of climate change, many marine species have changed their distributions towards higher latitude, deeper water or following local temperature gradients [@Poloczanska:2016kk].
-Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].The projected risks of impacts can be reduced through improving the effectiveness of current fisheries managemen [@Gaines:2018sg; @Cheung:2018dh], including the management of fisheries on species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
+Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].
+The projected risks of impacts can be reduced through improving the effectiveness of current fisheries management [@Gaines:2018sg; @Cheung:2018dh], including the fisheries management of species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
 
 The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
-As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into three non-exclusive categories: (*i*) transboundary stocks, those that cross neighboring EEZs; (*ii*) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; and (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs [@Munro:2004th].
+As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (i) transboundary stocks, those that cross neighboring EEZs; (ii) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; (iii) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (iv) discrete stocks that are only present in the high seas [@Munro:2004th].
+This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
 While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 (Chapter 2).
+A recent study estimates that, globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
 
-The effectiveness of managing fisheries targeting shared stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
-In many cases, catch or fishing effort quota on shared stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change in the fish stocks and fisheries [@Sumby:2021gw, Chapter 2].
-Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by worsening climate change [@Pinsky:2018cb].
-With future shifts in species distributions under climate change, the challenges posed on shared stocks fisheries management will continue to increase in the $21^{st}$ century.
-To prepare ocean governance to meet the challenges in resource management arising from shifting shared stocks, it is important to understand the time when climate change will affect the sharing distribution of these stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
+The effectiveness of fisheries management for transboundary stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
+In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb].
+With future shifts in species distributions under climate change, the challenges for the fisheries management of transboundary stocks will continue to increase during the 21st century.
+In order to prepare ocean governance to meet the resource management challenges arising from shifting transboundary stocks, it is important to constrain when climate change will affect the dynamics of transboundary  stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
 
 The timing and intensity of climate impacts on managing transboundary fish stocks can be quantitatively examined through the concepts of ‘time of emergence’ (ToE) and ‘threat point’ (TP).
-Time of emergence is defined as the moment in time when an anthropogenic change signal (e.g., future trend) rises above the background noise of natural variability (e.g., historical variation) [@IPCC:2019tn].
-This concept has been applied to investigate the emergence of multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
+Time of emergence is defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (e.g., historical natural variation) [@Hawkins:2012hc] and has been applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
 The premise behind the time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
 The concept of threat point comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
-Game theory has been widely used to investigate and manage shared stocks [@Sumaila:2013hv].
-Specifically, Sumaila et al.
-(*in press*) applied this concept to understand the impacts of climate change to the economics of shared species between Canada and the United States.
+Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary species between Canada and the United States (Sumaila et al., *in press*) .
 
-Here, I estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
-I employ a species distribution model driven by output from an Earth system model with ten ensemble members to project the distribution of the 663 selected species shared (Chapter 2) by 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
-I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (Chapter 2) and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks, excluding the high seas in this study).
-Our results suggest that many transboundary stocks have already ‘emerged’ from historical natural variability and overpassed their share threat point in most neighboring EEZs, posing a threat to the sustainability of shared stocks and the resilience of international treaties.
+Here, we estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
+We employ a species distribution model driven by output from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species [@PalaciosAbrantes:2020tv] across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
+We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude  the high seas.
+We find that many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.
 
@@
 This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 (Chapter 2).
-Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd, Chapter  1], resulting in a total of 9,132 transboundary stocks in this analysis.
-Using the transboundary species Atlantic cod (Gadus morhua) as an example, based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of G.
-morhua, but the United States and Greenland do not share a stock of this transboundary species.
-I defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico’s EEZ is divided into Mexico Pacific and Mexico Atlantic; \@ref(fig:toeFigureS1))) including island territories.
-I determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
-Each EEZ was categorized by the geopolitical regions according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and by natural biome [@Reygondeau:2019uh].
-The habitat association of each species was determined based on information available from FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org).
-For each stock and EEZ, I used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
-I report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd, Chapter 2], resulting in a total of 9,132 transboundary stocks in this analysis.
+Using the transboundary species Atlantic cod (Gadus morhua) as an example, and based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species.
+We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1) including island territories.
+We determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
+Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
+The habitat association of each species was determined following the classification in FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org).
+For each stock and EEZ, we used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
+We report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
 The monthly average US Consumer Price Index (CPI) according to the U.S.
@@
 
-I projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
-The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences to sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
-For pelagic species, the model uses environmental variables at the surface (e.g., sea surface temperature) whereas demersal species distribution is driven by ocean bottom variables (e.g., bottom temperature).
-This way, the model estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world (see Cheung et al.
+We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
+The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
+For pelagic species, the model uses environmental variables at the surface whereas demersal species distribution is driven by ocean bottom variables .
+DBEM then estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid (see Cheung et al.
 [-@Cheung:2010dt; -@Cheung:2016jd] for model details).
-DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd], Cheung et al.
+Importantly, DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
 *in revision*].
@@
 The DBEM was forced with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
-The GFDL-ESM2M is run under historical forcing until 2005 and follows the ‘high’ climate change emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
-In the GFDL-ESM2M, atmospheric surface temperature is projected to increase by $3.2^{\circ}C$ by 2081-2100, relative to preindustrial levels.
-Because the main approach of this chapter relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand both the historical as well as the future climate-driven distribution variability, to infer differences between time frames.
+The GFDL-ESM2M is run under historical forcing until 2005 and follows the high greenhouse gas emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand distribution variability during both the historical and the future periods, to infer differences between time frames.
 I do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
-Thus, for our experiment, each of the ten GFDL-ESM2M-ensembles were started from infinitesimally small differences in Earth system initial conditions resulting in a unique atmosphere and ocean state at each point in time.
-This approach allows us to estimate the natural internal variability (e.g., background noise) and isolate the forced climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
+Thus, for my experiment, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in year 1950 resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
+By design, variations among ensemble members are then solely due to natural variability.
+This approach allows us to estimate the natural internal variability (e.g., background noise) and isolate the forced anthropogenic climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
 
@@
 EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}} \\
-ToE = t when Emergence index[t] > 1 s.d.
+ToE = ToE_{t} > 1 s.d.
 \end{split}
@@
 
-In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than operating on a non-cooperative basis [@Munro:1979dg].
-However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., "threat point" [@Sumaila:2013hv].
-Here, I define threat point as the minimum proportion of a shared stock within an EEZ that a country will accept to engage in cooperative management with the country’s sharing neighbor [@Sumaila:2020tp, Chapter Four].
-Any proportion below the defined threat point of a country would result in a deflection from the joint management of the shared stock.
+In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than those operating on a non-cooperative basis [@Munro:1979dg].
+However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., “threat point” [@Sumaila:2013hv].
+Here, we define the threat point as the minimum required proportion of a shared stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Any proportion below the defined threat point of a country would result in a unilateral management of the shared stock.
 
-First I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
-I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2).
-Second, I average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
-The first time period ($SSR_{th}$) spans from 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
-It is assumed as the historic baseline of shared distributions for each stock.
+First, we estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
+We did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2).
+Second, we averaged the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of variability.
+The first time period ($SSR_{th}$) spans 1951 to 2005 and is chosen to match the historical period in the GFDL ESM2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
+It is assumed to represent the historic baseline of shared distributions for each stock.
 I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$).
@@
 The time of emergence results were tested for normality (e.g.
-skewness = 1.02, kurtosis = 2.83) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
+skewness, kurtosis) and performed two non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
 Specifically, I tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of transboundary stocks.
@@
 The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:toeFigure1)).
+This means that on average, the distribution of transboundary stocks will diverge from their natural internal variability by 2029.
 In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100.
 Fish stocks with transboundary index that first emerged from their historical variability in each EEZ was projected to be as early as 2006 and the last one in 2100.
-About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g.
-Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010).
+About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g., Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010).
 The number of EEZs with emerging stocks is projected to increase steeply from 87% in 2006 to 96% in 2020.
-After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.7% per year.
+After the mid-2020s, the rate of increase in the number EEZs slows from a 1% to an average of 0.7% per year.
 By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:toeFigure1)).
 Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020.
-Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure Figure \@ref(fig:toeFigure1)).
-I based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
-Results from the sensitivity analysis suggest that our method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g.
-the top 90%) of the stock (Figure \@ref(fig:toeFigureS3).
+Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure \@ref(fig:toeFigure1)).
+I based the transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
+Results from the sensitivity analysis suggest that this method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g., the top 95%) of the stock (Figure \@ref(fig:toeFigureS3).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone (EEZ) and transboundary stock: A) results in absolute numbers (green).
-B) results expressed as a percentage of the total emerging stocks and EEZ with emerging stock (purple)."}
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence of the transboundary index by Exclusive Economic Zone (EEZ) and transboundary stock"}
 
@@
 
-The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A).
-Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1)) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B).
-In contrast, EEZs located in template regions like northern Europe and eastern Asia will have a significantly later time of emergence (p < 0.05; see Table \@ref(tab:toeTableS1 for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058).
-However, notable exceptions of tropical EEZs with a later average time of emergence across stocks include Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070).
-Moreover, despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) will have relatively early average time of emergence (Figure \@ref(fig:toeFigure2)A).
-Only in a few cases, the transboundary index of a stock will not emerge between 2006 and 2100.
+The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; \@ref(fig:toeFigure2)A).
+Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (\@ref(fig:toeFigure2)B).
+In contrast, EEZs located in temperate regions like northern Europe and eastern Asia have significantly later times of emergence (p < 0.05; see Table @ref(tab:toeTableS1 for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058).
+However, notable exceptions to these broad patterns exist, with  tropical EEZs displaying a later average time of emergence across stocks including the  Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070) and despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) display a relatively early average time of emergence (\@ref(fig:toeFigure2)A).
+There are only a few cases where the transboundary index of a stock does not emerge between 2006 and 2100.
 In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
-In the case of Brazil, four transboundary stocks were included, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
-In all cases, our study suggest that the transboundary index of these stocks will not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A)
+In the case of Brazil, we included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
+However, on all cases, we find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (\@ref(fig:toeFigure2)A).
 
@@
 
-Our analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see their distribution range shifts beyond historical variabilities between 2025 and 2040 (Figure \@ref(fig:toeFigure3)).
-In some cases, like Indonesia (ranked $8^{th}$ in terms of revenue from transboundary stocks), Thailand (ranked $7^{th}$ in terms of revenue from transboundary stocks) and the United States (ranked $4^{th}$ in terms of fishing revenue from transboundary stocks), stocks’ range shifts swill emerge by around 2030.
-The 47 countries that make up the bulk of the global fisheries revenues also have more stocks with range shifts emerging from historical variabilities, being Spain (n = 122), France (n = 85), China and Portugal (both n = 73) and Senegal (n = 66) the top 5 leading countries in number of emerging stocks.
+My analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see the distribution range of their stocks shift beyond historical variabilities between 2025 and 2040 (Figure \@ref(fig:toeFigure3)).
+In some cases, like Indonesia (ranked $8^{th}$ in terms of revenue from transboundary stocks), Thailand (ranked $7^{th}$ in terms of revenue from transboundary stocks) and the United States (ranked $4^{th}$ in terms of fishing revenue from transboundary stocks), the range shift of transboundary stocks will emerge by around 2030.
+The 47 countries that make up the bulk of the global fisheries revenues also have more transboundary stocks with range shifts emerging from historical variability, such as Spain (n = 122), France (n = 85), China and Portugal (both n = 73) and Senegal (n = 66), which are the top 5 leading countries in terms of number of transboundary stocks emerging from historical variability.
 
-On average, stocks that will see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010.
-However, large variation exists in the proportion that stocks with emerging transboundary index have in the current revenue participation across EEZs.
-In some cases e.g., El Salvador and New Zealand, stock’s with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries  in their respective EEZs between 2004 and 2014.
-On the contrary, stocks with emerging transboundary index from the Marshal Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A).
-In addition, some EEZs are projectede to have a few emerging stocks that represent large proportion of revenue from transboundary stocks e.g., in Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively).
-Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks on fishing revenue is less than 20%.
+On average, stocks that see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of their EEZ’s fisheries revenue that targets transboundary species between 2004 and 2010.
+However, large variation exists in the contribution of stocks with emerging transboundary index to the current revenue participation across EEZs.
+In some cases, e.g., El Salvador and New Zealand, stock’s with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries in their respective EEZs between 2004 and 2014.
+On the contrary, stocks with emerging transboundary index from the Marshall Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A).
+In addition, some EEZs are projected to have a few emerging stocks that represent a large proportion of revenue from transboundary stocks e.g., in Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively).
+Overall, Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks to fishing revenue is less than 20%.
+This indicates that that these countries despite having a large proportion of transboundary stocks emerging, the economic contribution of these are relatively small.
 
@@
 (67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods).
-I tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence).
-The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
+We tested the sensitivity to a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence) and the global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
 threshold (Figure \@ref(fig:toeFigureS2)).
-In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric, that is, 40% (n = 1626) fewer emerging stocks than using a 1.s.d threshold.
+In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric,
 
-I compared species time of emergence within EEZs by species’ habitat association (Figure \@ref(fig:toeFigure4)).
+I compared stocks time of emergence within EEZs by the species’ habitat association (Figure \@ref(fig:toeFigure4)).
 Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods).
-Significant differences were found in the time of emergence of stocks range shifts, indicated by the transbounudary index, by their specie's habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
-Specifically, time of emergence of range shifts of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
-On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (toeFigure \@ref(fig:toeFigure4)).
+We found significant differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
+Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table S2).
+On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (Figure \@ref(fig:toeFigure4)).
 Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:toeFigure4)).
@@
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range."}
+Line represents median, and black points represent outliers with values exceeding 1.5 times the interquartile range."}
 
@@
 
-I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods - equation \@ref(eq:five)).
+I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods – \@ref(eq:five)).
 By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A).
@@
 For example, Guatemala shares Panulirus gracilis stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century.
-Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)).
-Our results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
-These EEZs will see one third (33% $\pm$ 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A).
-Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin will see changes in all of the transboundary stocks analyzed while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) will be the only countries with less than 10% of stocks changing beyond the threat point.
+Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4))).
+My results suggest that 87% (n = 238) of the world EEZs experiences changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
+These EEZs see one third (33% $\pm$ 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A).
+Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin see changes beyond the threat point for all of the transboundary stocks analyzed, while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) are the only countries with less than 10% of stocks changing beyond the threat point.
 
@@
 Changes of stock share ratio are largely related to regional changes in biogeography and the geometry of the EEZs (Figure \@ref(fig:toeFigure5)).
-In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa stock share ratio is expected to follow the projected poleward shift of transboundary species under climate change.
-However, central America Pacific and West African coasts are expected to follow an equatorial direction as range shifts of the transboundary species are following local oceanographic gradients with relatively lower temperature waters in lower latitude [@Clarke:2020ep].
-Moreover, some regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no stablished pattern, this is the case of the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).
+In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the stock share ratio is expected to follow the projected poleward shift of transboundary species under climate change.
+However, central America Pacific and West African coasts follow an equatorial direction as range shifts of the transboundary species follow local oceanographic gradients, with relatively lower temperature waters in lower latitude [@Clarke:2020ep].
+Moreover, regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no established pattern, as in the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).
 
@@
 Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
-Such early emergences of range shifts indicate the urgency of the challenges that climate change is posing on international fisheries management.
+Such early emergence of range shifts indicate the urgency of the challenges that climate change is posing on international fisheries management.
 Moreover, I identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management.
-Finally, I show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to level below that conduce cooperative resource management between countries.
-Adaptation measures are needed to reduce the impacts of species range shifts on current international transboundary fisheries treaties.
+Finally, I show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to level below that which encourages cooperative resource management between countries.
+Adaptation measures are therefore needed to reduce the impacts of species range shifts on current international transboundary fisheries treaties.
 
@@
 
-Our projections in the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century corroborate with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk].
+Our projections for the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century concur with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk].
 For example, in the early 2000s, Humboldt squid (Dosidicus gigas) substantially expanded its geographic range poleward, reaching the coast of Washington state (US).
@@
 In the northeast Atlantic, the Atlantic mackerel (Scomber scombrus) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
-However, a range expansion of Atlantic mackerel (Scomber scombrus) into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008 without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij].
-Such changes in quota resulted in disputes between Iceland and Faroes Islands, and the NEAFC member states [@Spijkers:2017ij].
-As transboundary stocks continue to emerge in the incoming years, mal-adapted fisheries management and international disputes like that of Atlantic mackerel are likely to increase in frequency [@Pinsky:2018cb]
+However, a range expansion of Atlantic mackerel (Scomber scombrus) into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008, without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij].
+Such changes resulted in disputes between Iceland and the Faroes Islands, as well as the NEAFC member states [@Spijkers:2017ij].
+As climate change impacts on transboundary stocks continue to emerge from historical variability in the coming years, mal-adapted fisheries management and international disputes like that of Atlantic mackerel are likely to increase in frequency [@Pinsky:2018cb].
 
-The projected range shifts of large number of transboundary stocks emerging since the early 2000s can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
-Marine organisms have a thermal tolerance that limits the temperature range they could live in [@IPCC:2019tn].
-This thermal tolerance explains the sensitivity of the animal to warming and is a main driver of species’ distributional change [@IPCC:2019tn].
-Oxygen plays a similar role to that of temperature [@Clarke:2020ep], especially on tuna species that require large amounts of dissolve oxygen for their metabolism [@IPCC:2014bz].
-It has been estimated that since the early 2000s, increase in sea surface temperature (SST) and decrease in oxygen content have emerged in 79% and 23% of the global ocean, respectively, and by 2080, SST is projected to emerge in 90% of the global ocean while O2 will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq].
+The projected range shifts of large number of transboundary stocks emerging from historical variability since the early 2000s can be partially attributed to the parallel emergence of several environmental variables (e.g., temperature and oxygen) that influence their distribution [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
+Marine organisms have a thermal tolerance that limits their habitable temperature range [@IPCC:2019tn], which then explains the sensitivity of the animal to warming and is a main driver of any distributional change [@IPCC:2019tn].
+Oxygen plays a similar role to that of temperature [@Clarke:2020ep], especially for tuna species that require large amounts of dissolved oxygen for their metabolism [@IPCC:2014bz].
+It has been estimated that since the early 2000s, increased sea surface temperature (SST) and decrease in subsurface oxygen have emerged from historical variability for 79% and 23% of the global ocean, respectively, and that by 2080, SST is projected to emerge in 90% of the global ocean while the area affected by lower O2 will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq].
 Moreover, before 2100, surface acidification, nitrate content and net primary production, also important variables for species distributions, will have emerged from the historical variability of 1861-1900 in 30% and 60% of the ocean under a low (RCP 2.6) and strong (RCP 8.5) climate change scenario, respectively [@IPCC:2019tn].
 These oceanographic changes are important drivers of change in species distributions as represented in the DBEM [@Cheung:2016jd; @Cheung:0ur].
+[I think we can tighten up the end of this paragraph to, e.g.
+from ‘It has been estimated..’ as I am not 100% sure of the point.
+Is it saying that these variables are all included in DBEB?
+If so, does it then imply that we could more formally look at things?
+Or, is the idea here to make the point that SST and O2 changes are most clearly emerging, but that other factors are also at play?]
 
@@
 Specifically, our findings are consistent with results from previous studies that show tropical species are already shifting their distributions in the last few decades [@Poloczanska:2016kk].
-The tropics are among the most species-rich biome [@Reygondeau:2019uh] and have potentially experienced some of the highest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
-In the tropics, marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz].
-Consequently, as a result of climate change, species have to move towards deeper or sub-tropical waters [@Poloczanska:2016kk] although recent studies suggest that in some regions species are shifting towards the equator and to shallow waters to follow local temperature gradients [@Clarke:2020ep.
-Moreover, tropical species are expanding their range in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis shows a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
+The tropics are among the most species-rich biomes [@Reygondeau:2019uh] and have potentially experienced some of the highest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
+In the tropics, marine species also live close to their thermal tolerance, making them highly vulnerable to warming [@IPCC:2014bz].
+Consequently,  species have to move towards deeper or sub-tropical waters as a result of climate change [@Poloczanska:2016kk], although in some regions species follow local temperature gradients towards the equator or shallow waters [@Clarke:2020ep].
+Moreover, tropical species are expanding their range into sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis shows a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
 In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs that border with multiple countries (e.g., The Caribbean).
@@
 
-Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)).
+Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)).
 For example, Polynesia and Micronesia are the only regions where the average time of emergence of the distribution of where all transboundary stocks across all EEZs emerged before 2022 (Figure \@ref(fig:toeFigure5)).
 Moreover, countries in these hotspot regions are highly dependent on pelagic-oceanic (e.g., tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
-On the other hand, “coldspots” of climate risk for transboundary fisheries like the Northeast Atlantic and Northwest Pacific where stocks are expected to emerge later in the 21st century.
+On the other hand, “coldspots” of climate risk exist for transboundary fisheries, like the Northeast Atlantic and Northwest Pacific, where stocks are expected to emerge later in the 21st century.
 Assuming that these countries already have adequate transboundary management to deal with historical and current variations in fish stocks’ distribution e.g., the Pacific Salmon Treaties [@Miller:2004fpa], management agencies are likely to have more time to prepare for emerging stock shifts.
-In cases where stocks are emerging or have already emerged, adaptation policies are urgently needed to address an issue that could be impacting important fisheries [@Burden:2019ij].
-Currently, fisheries targeting tunas shared stocks in the South Pacific Islands region operate under the Vessel Day Scheme - which takes into account shifts in stocks and catches making it a good example where a management system has been set up that is somewhat anticipatory [@Aqorau:2018bh].
-However, even strategies that take into consideration the shift of the species within the Party’s area will likely require to deal with the newcomer issue that NEAFC is currently facing over Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
-As species have been emerging from their historical distribution since 2006, regional management bodies require to revise their management actions as today most of them are not equipped with strategies to cope with shifting species distributions [@Oremus:2020id; @Sumby:2021gw].
+
+In cases where climate driven changes in stocks are emerging, or have already emerged, from historical variability adaptation policies are urgently needed to address the potential impact on important fisheries [@Burden:2019ij].
+Currently, fisheries targeting shared tuna stocks in the South Pacific Islands region operate under the Vessel Day Scheme, which because it takes into account shifts in stocks and catches makes it a good example of a management system that is somewhat anticipatory [@Aqorau:2018bh].
+However, even strategies that take into consideration the shift of the species within the Party’s area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+As the climate trend in species emerges from their historical distribution since 2006, regional management bodies must revise their management actions as most are not currently equipped with strategies to cope with shifting species distributions [@Oremus:2020id; @Sumby:2021gw].
 This is of specific concern for those management bodies located in the emerging “hotspots” here identified.
 
-The presentce of EEZs lacking stocks with emerging transboundary index is a combination of EEZ shape, climate projections, and a reduced number of transbounudary stocks.
-For example, there are 59 commercial fish stocks fished the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas with limited knowledge about the species biology [@Christiansen:2014gg].
-However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@IPCC:2019tn].
+Some EEZs lack stocks with emerging transboundary index because of a combination of their shape, climate projections, and a reduced number of transboundary stocks.
+For example, there are 59 commercial fish stocks fished in the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas, with limited knowledge about the species biology [@Christiansen:2014gg].
+However, the Arctic has been warming faster than many other regions, resulting in a reduction in sea-ice [@Douglas:2010Mb] leading to the  potential for  occupation by species not previously present [@IPCC:2019tn].
 As a result, the biomass of Arctic marine animals [@BryndumBuchholz:2019kh] and catch and revenue of associated fisheries [@Tai:2019cq; @Lam:2014fq] are expected to increase in the incoming years, although substantial uncertainty exists on the magnitude of change [@BryndumBuchholz:2019kh].
 Moreover, projections suggest a larger increment on transboundary stocks in the incoming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further North [@Morley:2018fn; @Frainer:2017ji].
-In the Case of Brazil, in addition to having few transbounudary stocks (only 4 shared with Uruguay in this study), the large diffrence in the EEZs lenghts could also be influencing the results.
+In the specific case of Brazil, in addition to having few transboundary stocks (only four shared with Uruguay in this study), the large difference in the EEZs lengths could also be influencing the results.
 
@@
 Poleward shifts in stock share ratio are expected in regions of the world with relatively long EEZs and few boundaries, like the northeast Pacific and the southeast Atlantic.
-However, despite observations and modelling work indicating that, generally, climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], some regions like the eastern tropical Pacific and eastern tropical Atlantic Oceans will experience a shift in distribution towards the equator.
+However, despite observations and modelling work indicating that climate change is generally shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], some regions like the eastern tropical Pacific and eastern tropical Atlantic Oceans will experience a shift in distribution towards the equator.
 Specifically, in the eastern tropical Pacific, the direction of shift is expected to be from Mexico to Ecuador with EEZs losing stock share ratio to their southern neighbors.
-Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems [@Clarke:2020ep].
-They are consistent with regional modeling studies suggesting that by mid 21st century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
-Moreover, regions characterized by multiple political boundaries, like the Caribbean and the northeast Atlantic, will see a complex network of changes in SSR with no clear pattern (\@ref(fig:toeFigure5)B).
-These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] supporting the implementation of joint management plans [@Burden:2019ij] and the call for new implemented shared policies [@Levin:2018in].
+Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems [@Clarke:2020ep]and are consistent with regional modeling studies suggesting that by the mid 21st century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission scenario [@Clarke:2020ep].
+Moreover, regions characterized by multiple political boundaries, like the Caribbean and the northeast Atlantic, will see a complex network of changes in stock share ratio with no clear pattern (Figure \@ref(fig:toeFigure5)B).
+These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and dispersal of larvae [@Ramesh:2019va] supporting the implementation of joint management plans [@Burden:2019ij] and the call for new implemented shared policies [@Levin:2018in].
 However, game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.
 
-Changes in stock share ratio beyond the threat point of countries sharing a common resource could instigate international conflict and hinder bi-lateral agreements (@Miller:2013iv; @Sumaila:2020tp).
-Lessons learned from historic distributional shifts of shared stocks suggest that treaties that are not prepare to respond to such shifts will be less resilient to changes in transboundary fish stocks share ratio due to climate change [@Miller:2013iv; @Sumaila:2020tp].
+Changes in stock share ratio beyond the threat point of countries sharing a common resource could instigate international conflict and hinder bi-lateral agreements [@Miller:2013iv; @Sumaila:2020tp].
+Lessons learned from historic distributional shifts of shared stocks suggest that treaties that are not prepared to respond to such shifts will be less resilient to changes in transboundary fish stocks share ratio due to climate change [@Miller:2013iv; @Sumaila:2020tp].
 For example, disputes over Pacific salmon between Canada and the USA arose in the 1990s when a climate-related shift in stock abundance favoured Alaska’s salmon fisheries over Canada’s [@Miller:2013iv; @Munro:2002uf; @Song:2017va].
 The conflict lasted over ten years until Canada and the USA agreed on a mutual conservation fund where the US would contribute the larger proportion to support scientific research, habitat restoration and enhancement of wild stock production in transboundary rivers [@Munro:2002uf].
-Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Pinsky:2018cb, Chapter 5].
+Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Pinsky:2018cb].
 Side payments are mechanisms that provide a solution to the inequality produced by shifts in the stock’s distribution [@Tunca:2019if; @Miller:2004fpa].
-These can be monetary (e.g., the stock-“winning” State compensates the stock-“loosing” state for the proportional shift) or other methods like the Pacific salmon case (e.g., a conservation fund payed mainly by one Party) [@Miller:2004fpa; @Miller:2013iv].
-Strengthening current international cooperation will be required as stock distribution will move to neighboring jurisdictions, requiring combine efforts to generate information and set proper management rules [@Miller:2013iv].
-Finally, flexible management rules might be better fitted to capture the stock’s distributional shift.
-For example, quota allocation based on the stock’s current distribution like that for Pacific halibut (*Hippoglossus stenolepis*) managed by Canada and the US [@Sumaila:2020tp; @PalaciosAbrantes:2020vt] might be better able to capture the stock’s shifts than a quota based on a fix-historical proportion like the case of the EU [@Baudron:2020dj].
+These can be monetary (e.g., the stock-“winning” State compensates the stock-“losing” state for the proportional shift) or other methods like the Pacific salmon case (e.g., a conservation fund paid mainly by one Party) [@Miller:2004fpa; @Miller:2013iv].
+Strengthening current international cooperation will be required as stock distribution will move to neighboring jurisdictions, requiring combined efforts to generate information and set proper management rules [@Miller:2013iv].
+Finally, flexible management rules might be more appropriate to capturing t distributional shifts.
+For example, a quota allocation based on the stock’s current distribution like that for Pacific halibut (*Hippoglossus stenolepis*) managed by Canada and the US [@Sumaila:2020tp; @PalaciosAbrantes:2020vt] would be more agile than one based on a fixed-historical proportion like the case of the EU [@Baudron:2020dj].
 

### 2020-12-14 Scripts/Manuscript.Rmd
Addressed Villy's comments

@@
 This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2).
-The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
+The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., ^Pacific halibut (*Hippoglossus stenolepis*) that is transboundary between the United States and Canada.
 Therefore,
@@
 However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high.
-For example, Guatemala shares Panulirus gracilis stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century.
+For example, Guatemala shares *Panulirus gracilis* stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century.
 Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4))).
@@
 Moreover, I identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management.
-Finally, I show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to level below that which encourages cooperative resource management between countries.
+Finally, I show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to levels below of that needed to encourage cooperative resource management between countries.
 Adaptation measures are therefore needed to reduce the impacts of species range shifts on current international transboundary fisheries treaties.
@@
 Our projections for the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century concur with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk].
-For example, in the early 2000s, Humboldt squid (Dosidicus gigas) substantially expanded its geographic range poleward, reaching the coast of Washington state (US).
+For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US).
 A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs].
-In the northeast Atlantic, the Atlantic mackerel (Scomber scombrus) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
-However, a range expansion of Atlantic mackerel (Scomber scombrus) into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008, without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij].
+In the northeast Atlantic, the Atlantic mackerel (*Scomber scombrus*) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
+However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008, without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij].
 Such changes resulted in disputes between Iceland and the Faroes Islands, as well as the NEAFC member states [@Spijkers:2017ij].
@@
 Oxygen plays a similar role to that of temperature [@Clarke:2020ep], especially for tuna species that require large amounts of dissolved oxygen for their metabolism [@IPCC:2014bz].
-It has been estimated that since the early 2000s, increased sea surface temperature (SST) and decrease in subsurface oxygen have emerged from historical variability for 79% and 23% of the global ocean, respectively, and that by 2080, SST is projected to emerge in 90% of the global ocean while the area affected by lower O2 will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq].
+It has been estimated that since the early 2000s, increased sea surface temperature (SST) and decrease in subsurface oxygen have emerged from historical variability for 79% and 23% of the global ocean, respectively, and that by 2080, SST is projected to emerge from historical variability in 90% of the global ocean while the area affected by lower $O_2$ will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq].
 Moreover, before 2100, surface acidification, nitrate content and net primary production, also important variables for species distributions, will have emerged from the historical variability of 1861-1900 in 30% and 60% of the ocean under a low (RCP 2.6) and strong (RCP 8.5) climate change scenario, respectively [@IPCC:2019tn].
 These oceanographic changes are important drivers of change in species distributions as represented in the DBEM [@Cheung:2016jd; @Cheung:0ur].
-[I think we can tighten up the end of this paragraph to, e.g.
-from ‘It has been estimated..’ as I am not 100% sure of the point.
-Is it saying that these variables are all included in DBEB?
-If so, does it then imply that we could more formally look at things?
-Or, is the idea here to make the point that SST and O2 changes are most clearly emerging, but that other factors are also at play?]
 
@@
 Moreover, countries in these hotspot regions are highly dependent on pelagic-oceanic (e.g., tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
-On the other hand, “coldspots” of climate risk exist for transboundary fisheries, like the Northeast Atlantic and Northwest Pacific, where stocks are expected to emerge later in the 21st century.
+On the other hand, “coldspots” of climate risk exist for transboundary fisheries, like the Northeast Atlantic and Northwest Pacific, where stocks are expected to emerge later in the $21^{st}$ century.
 Assuming that these countries already have adequate transboundary management to deal with historical and current variations in fish stocks’ distribution e.g., the Pacific Salmon Treaties [@Miller:2004fpa], management agencies are likely to have more time to prepare for emerging stock shifts.
@@
 Some EEZs lack stocks with emerging transboundary index because of a combination of their shape, climate projections, and a reduced number of transboundary stocks.
-For example, there are 59 commercial fish stocks fished in the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas, with limited knowledge about the species biology [@Christiansen:2014gg].
+For example, there are 59 commercial fish stocks fished in the Arctic with the vast majority on the Barents, Bering and Norwegian seas, with limited knowledge about the species biology [@Christiansen:2014gg].
 However, the Arctic has been warming faster than many other regions, resulting in a reduction in sea-ice [@Douglas:2010Mb] leading to the  potential for  occupation by species not previously present [@IPCC:2019tn].
 As a result, the biomass of Arctic marine animals [@BryndumBuchholz:2019kh] and catch and revenue of associated fisheries [@Tai:2019cq; @Lam:2014fq] are expected to increase in the incoming years, although substantial uncertainty exists on the magnitude of change [@BryndumBuchholz:2019kh].
-Moreover, projections suggest a larger increment on transboundary stocks in the incoming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further North [@Morley:2018fn; @Frainer:2017ji].
+Moreover, projections suggest a larger increment on transboundary stocks in the coming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further north [@Morley:2018fn; @Frainer:2017ji].
 In the specific case of Brazil, in addition to having few transboundary stocks (only four shared with Uruguay in this study), the large difference in the EEZs lengths could also be influencing the results.
@@
 Poleward shifts in stock share ratio are expected in regions of the world with relatively long EEZs and few boundaries, like the northeast Pacific and the southeast Atlantic.
-However, despite observations and modelling work indicating that climate change is generally shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], some regions like the eastern tropical Pacific and eastern tropical Atlantic Oceans will experience a shift in distribution towards the equator.
+However, despite observations and modelling work indicating that climate change is generally shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], some regions like the eastern tropical Pacific and easterntropical Atlantic Oceans will experience a shift in distribution towards the equator.
 Specifically, in the eastern tropical Pacific, the direction of shift is expected to be from Mexico to Ecuador with EEZs losing stock share ratio to their southern neighbors.
@@
 Side payments are mechanisms that provide a solution to the inequality produced by shifts in the stock’s distribution [@Tunca:2019if; @Miller:2004fpa].
-These can be monetary (e.g., the stock-“winning” State compensates the stock-“losing” state for the proportional shift) or other methods like the Pacific salmon case (e.g., a conservation fund paid mainly by one Party) [@Miller:2004fpa; @Miller:2013iv].
+These can be monetary (e.g., the stock-“winning” state compensates the stock-“losing” state for the proportional shift) or other methods like the Pacific salmon case (e.g., a conservation fund paid mainly by one party) [@Miller:2004fpa; @Miller:2013iv].
 Strengthening current international cooperation will be required as stock distribution will move to neighboring jurisdictions, requiring combined efforts to generate information and set proper management rules [@Miller:2013iv].

### 2020-12-14 Scripts/Manuscript.Rmd
standarized stock and species

@@
 While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
+A recent study estimates that, globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 (Chapter 2).
 
 The effectiveness of fisheries management for transboundary stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
-In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
-Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb].
+In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the stocks [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+Misalignment between fisheries resources allocation and stocks’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb].
 With future shifts in species distributions under climate change, the challenges for the fisheries management of transboundary stocks will continue to increase during the 21st century.
@@
 The concept of threat point comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
-Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary species between Canada and the United States (Sumaila et al., *in press*) .
+Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary stocks between Canada and the United States (Sumaila et al., *in press*) .
 
-Here, we estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
-We employ a species distribution model driven by output from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species [@PalaciosAbrantes:2020tv] across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
-We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude  the high seas.
-We find that many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.
+Here, I estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
+I employ a species distribution model driven by output from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species (Chapter 2) across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP 8.5 (see Methods).
+I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (Chapter 2) and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude  the high seas.
+I find that many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.
 
@@
 Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd, Chapter 2], resulting in a total of 9,132 transboundary stocks in this analysis.
-Using the transboundary species Atlantic cod (Gadus morhua) as an example, and based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species.
-We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1) including island territories.
-We determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
+Using the transboundary species Atlantic cod (*Gadus morhua*) as an example, and based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species.
+I defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1) including island territories.
+I determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
 Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
 The habitat association of each species was determined following the classification in FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org).
-For each stock and EEZ, we used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
-We report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+For each stock and EEZ, I used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
+I report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
 The monthly average US Consumer Price Index (CPI) according to the U.S.
@@
 The GFDL-ESM2M is run under historical forcing until 2005 and follows the high greenhouse gas emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand distribution variability during both the historical and the future periods, to infer differences between time frames.
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, I have to understand distribution variability during both the historical and the future periods, to infer differences between time frames.
 I do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
@@
 
-where $D_A$ and $D_B$ represent the distance between the distribution centroids of a species and the centroid of the pair of neighbouring EEZs (A and B) sharing the species for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the pair of neighbouring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
 Thus, a higher TI means that a shared stock becomes more or less abundant relative to its neighbouring EEZs that is beyond what the fisheries experienced historically.
@@
 However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., “threat point” [@Sumaila:2013hv].
-Here, we define the threat point as the minimum required proportion of a shared stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Here, I define the threat point as the minimum required proportion of a shared stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
 Any proportion below the defined threat point of a country would result in a unilateral management of the shared stock.
 
-First, we estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
-We did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2).
-Second, we averaged the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of variability.
+First, I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
+I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2).
+Second, I averaged the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of variability.
 The first time period ($SSR_{th}$) spans 1951 to 2005 and is chosen to match the historical period in the GFDL ESM2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
@@
 The time of emergence results were tested for normality (e.g.
-skewness, kurtosis) and performed two non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
+skewness, kurtosis) and performed two non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary stocks.
 Specifically, I tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of transboundary stocks.
-For both cases, our null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
-<!-- Moreover, we implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
-Specifically, we looked at habitat associations, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
-We hypothesize that all of these factors would have significant influence in the ToE of transboundary species.
+For both cases, my null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
+<!-- Moreover, I implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
+Specifically, I looked at habitat associations, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
+I hypothesize that all of these factors would have significant influence in the ToE of transboundary species.
 -->
@@
 In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
-In the case of Brazil, we included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
-However, on all cases, we find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (\@ref(fig:toeFigure2)A).
+In the case of Brazil, I included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
+However, on all cases, I find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (\@ref(fig:toeFigure2)A).
 
@@
 
-On average, stocks that see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of their EEZ’s fisheries revenue that targets transboundary species between 2004 and 2010.
+On average, stocks that see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of their EEZ’s fisheries revenue that targets transboundary stocks between 2004 and 2010.
 However, large variation exists in the contribution of stocks with emerging transboundary index to the current revenue participation across EEZs.
@@
 (67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods).
-We tested the sensitivity to a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence) and the global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
+I tested the sensitivity to a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence) and the global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
 threshold (Figure \@ref(fig:toeFigureS2)).
@@
 Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods).
-We found significant differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
+I found significant differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
 Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table S2).
@@
 
-<!-- We hypothesized that species environmental preferences (e.g.
+<!-- I hypothesized that species environmental preferences (e.g.
 presence in certain biomes and habitats), climate change effects on water temperature (e.g.
@@
     booktabs = T,
-    captio= "Multiple regression results for time of emergence of transboundary species.
+    captio= "Multiple regression results for time of emergence of transboundary stocks.
 Only showing significant variables.
@@
 In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the stock share ratio is expected to follow the projected poleward shift of transboundary species under climate change.
-However, central America Pacific and West African coasts follow an equatorial direction as range shifts of the transboundary species follow local oceanographic gradients, with relatively lower temperature waters in lower latitude [@Clarke:2020ep].
+However, central America Pacific and West African coasts follow an equatorial direction as range shifts of transboundary stocks follow local oceanographic gradients, with relatively lower temperature waters in lower latitude [@Clarke:2020ep].
 Moreover, regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no established pattern, as in the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).
@@
 In this chapter, I investigate the timing and intensity of changes in the sharing of transboundary stocks between countries.
-Our findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
+My findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
 Such early emergence of range shifts indicate the urgency of the challenges that climate change is posing on international fisheries management.
@@
 Regional differences in time of emergence are attributed to the variations in environmental conditions, species’ climate vulnerability, and the characteristics of political boundaries.
-Specifically, our findings are consistent with results from previous studies that show tropical species are already shifting their distributions in the last few decades [@Poloczanska:2016kk].
+Specifically, my findings are consistent with results from previous studies that show tropical species are already shifting their distributions in the last few decades [@Poloczanska:2016kk].
 The tropics are among the most species-rich biomes [@Reygondeau:2019uh] and have potentially experienced some of the highest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
@@
 Currently, fisheries targeting shared tuna stocks in the South Pacific Islands region operate under the Vessel Day Scheme, which because it takes into account shifts in stocks and catches makes it a good example of a management system that is somewhat anticipatory [@Aqorau:2018bh].
-However, even strategies that take into consideration the shift of the species within the Party’s area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
-As the climate trend in species emerges from their historical distribution since 2006, regional management bodies must revise their management actions as most are not currently equipped with strategies to cope with shifting species distributions [@Oremus:2020id; @Sumby:2021gw].
+However, even strategies that take into consideration the shift of the stock within the party’s area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+As the climate trend in stocks distibution emerges from their historical variabillity since 2006, regional management bodies must revise their management actions as most are not currently equipped with strategies to cope with shifting species distributions [@Oremus:2020id; @Sumby:2021gw].
 This is of specific concern for those management bodies located in the emerging “hotspots” here identified.
@@
 Specifically, in the eastern tropical Pacific, the direction of shift is expected to be from Mexico to Ecuador with EEZs losing stock share ratio to their southern neighbors.
-Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems [@Clarke:2020ep]and are consistent with regional modeling studies suggesting that by the mid 21st century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission scenario [@Clarke:2020ep].
+Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems [@Clarke:2020ep] and are consistent with regional modeling studies suggesting that by the mid $21{st}$ century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission scenario [@Clarke:2020ep].
 Moreover, regions characterized by multiple political boundaries, like the Caribbean and the northeast Atlantic, will see a complex network of changes in stock share ratio with no clear pattern (Figure \@ref(fig:toeFigure5)B).
@@
 
-There are two main sources of uncertainty in our analysis.
-First, our analysis relies on the combination of a single Earth system model (GFDL-ESM2M), fish and fisheries model, and climate change scenario (RCP 8.5).
+There are two main sources of uncertainty in my analysis.
+First, I rely on the combination of a single Earth system model (GFDL-ESM2M), fish and fisheries model, and climate change scenario (RCP 8.5).
 The use of multiple ESMs can capture the structural uncertainty of the models [@Frolicher:2016dg; @Lotze:2019ce], however, exercises looking at changes in fish biomass with multiple models show an overall agreement between models in terms of direction of change but variable in magnitude.
 Further research that includes multiple ESMs and species distribution models could elucidate a broader range structural uncertainty of the climate, fish and fisheries models [@Lotze:2019ce].
-Secondly, due to limited biological and spatially-specific data on stocks sub-population structure of transboundary species, our analysis uses political boundaries (EEZ) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
-On the other hand, in many EEZs, fisheries are often managed at the species level (Chapter 2) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
-However, reproducing our analysis regionally, where spatially explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
+Secondly, due to limited biological and spatially-specific data on stocks sub-population structure of transboundary species, my analysis uses political boundaries (EEZs) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
+On the other hand, in many EEZs, fisheries are often managed at the species level (Chapter 2) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for my analysis [@Popova:2019fs; @Dunn:2019ca].
+However, reproducing my analysis locally, where spatially explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
 Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
@@
 I have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
-Our results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management.
-While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg; @PalaciosAbrantes:2020tv].
+My results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management.
+While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, my results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg; @PalaciosAbrantes:2020tv].
 To address the issues raised in this chapter, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
@@
 
-<!-- The time and number of emerging species decreases as we move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
+<!-- The time and number of emerging species decreases as I move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
 -->

### 2020-12-15 Scripts/Manuscript.Rmd
William final comments

@@
 
-We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
+I projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
 The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
-For pelagic species, the model uses environmental variables at the surface whereas demersal species distribution is driven by ocean bottom variables .
+For pelagic species, the model uses environmental variables at the surface whereas demersal species distribution is driven by ocean bottom variables.
 DBEM then estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid (see Cheung et al.
@@
 I developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
-This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2).
-The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., ^Pacific halibut (*Hippoglossus stenolepis*) that is transboundary between the United States and Canada.
+This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2)).
+The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., Pacific halibut (*Hippoglossus stenolepis*) that is transboundary between the United States and Canada.
 Therefore,
@@
 I set a threshold of TI above which I consider a shared stock to have emerged from the historical internal variability of the TI.
-The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four), representing a probability of 68% that the transboundary index has emerged from historical variability).
+The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four)), representing a probability of 68% that the transboundary index has emerged from historical variability).
 I tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
@@
 
-Estimating the threat point of transboundary stock share
+### Estimating the threat point of transboundary stock share
 
@@
 It is assumed to represent the historic baseline of shared distributions for each stock.
-I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$).
+I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SSR_{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR_{tm}$).
 I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
@@
 Third, I defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$.
-This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{et} \geqq (SSR_{th} + \sigma)$ or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
+This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{te} \geqq (SSR_{th} + \sigma)$ or $SSR[t]_{e,m} \leqq (SSR_{th} - \sigma)$.
 Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al.
@@
 
-Time of emergence of transboundary stocks
+### Time of emergence of transboundary stocks
 
@@
 I based the transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
-Results from the sensitivity analysis suggest that this method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g., the top 95%) of the stock (Figure \@ref(fig:toeFigureS3).
+Results from the sensitivity analysis suggest that this method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g., the top 95%) of the stock (Figure \@ref(fig:toeFigureS3)).
 
@@
 
-The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; \@ref(fig:toeFigure2)A).
-Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (\@ref(fig:toeFigure2)B).
-In contrast, EEZs located in temperate regions like northern Europe and eastern Asia have significantly later times of emergence (p < 0.05; see Table @ref(tab:toeTableS1 for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058).
-However, notable exceptions to these broad patterns exist, with  tropical EEZs displaying a later average time of emergence across stocks including the  Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070) and despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) display a relatively early average time of emergence (\@ref(fig:toeFigure2)A).
+The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A).
+Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B).
+In contrast, EEZs located in temperate regions like northern Europe and eastern Asia have significantly later times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058).
+However, notable exceptions to these broad patterns exist, with  tropical EEZs displaying a later average time of emergence across stocks including the  Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070) and despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) display a relatively early average time of emergence (Figure \@ref(fig:toeFigure2)A).
 There are only a few cases where the transboundary index of a stock does not emerge between 2006 and 2100.
@@
 In the case of Brazil, I included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
-However, on all cases, I find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (\@ref(fig:toeFigure2)A).
+However, on all cases, I find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A).
 
@@
 I found significant differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
-Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table S2).
+Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
 On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (Figure \@ref(fig:toeFigure4)).
@@
 
-I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods – \@ref(eq:five)).
+I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods – Equation \@ref(eq:five)).
 By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A).
@@
 For example, Guatemala shares *Panulirus gracilis* stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century.
-Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4))).
+Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)).
 My results suggest that 87% (n = 238) of the world EEZs experiences changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
@@
 
-#### High present-day climate risk of on transboundary fisheries management
+### High present-day climate risk of on transboundary fisheries management
 
@@
 Global society has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear effects on multiple others [@UnitedNations2018; @Singh:2017ds].
-Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@@Pinsky:2018cb].
+Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@Pinsky:2018cb].
 Here, I have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
@@
 My results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management.
-While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, my results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg; @PalaciosAbrantes:2020tv].
+While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, my results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; @Gaines:2018sg; @PalaciosAbrantes:2020tv].
 To address the issues raised in this chapter, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.

### 2020-12-15 Scripts/Supplements.Rmd
William final comments

@@
 
-```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
+```{r toeFigureS1, eval = T,  echo = F, fig.width=12, fig.height=7, fig.cap =  "World Exclusive Economic Zones used in this study as defined by the Sea Around Us and their centroids (points)."}
 
@@
 
-```{r toeFigureS2, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+```{r toeFigureS2, eval = T,  echo = F, fig.width=12, fig.height=5, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
 The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel).
@@
           "K-W Difference" = dif.com.difference) %>% 
-  kable(caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions")
-  # kable("latex",
-        # booktabs = T,
-        # caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  # ) %>% 
-  # kable_styling(full_width = F,
-                # position = "center")
+  # kable(caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions")
+  kable("latex",
+        booktabs = T,
+        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  ) %>%
+  kable_styling(full_width = T,
+                position = "center")
 
@@
           "K-W Difference" = dif.com.difference) %>% 
-  kable(
+  # kable(
+  #       booktabs = T,
+  #       caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  # )  
+  kable("latex",
         booktabs = T,
         caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  )  
-  # kable("latex",
-  #       booktabs = T,
-  #       caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  # ) %>% 
-  # kable_styling(full_width = F,
-  #               position = "center")
+  ) %>%
+  kable_styling(full_width = F,
+                position = "center")
 

### 2020-12-22 Scripts/Manuscript.Rmd
las changes

@@
 As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (i) transboundary stocks, those that cross neighboring EEZs; (ii) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; (iii) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (iv) discrete stocks that are only present in the high seas [@Munro:2004th].
-This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
+This chapter focuses on transboundary stocks exploited by fisheries operating within EEZs.
 While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 (Chapter 2).
+In the previous chapter, I estimated that globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010.
 
 The effectiveness of fisheries management for transboundary stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
-In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the stocks [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the stocks [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@Sumby:2021gw, Chapter 2].
 Misalignment between fisheries resources allocation and stocks’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb].
 With future shifts in species distributions under climate change, the challenges for the fisheries management of transboundary stocks will continue to increase during the 21st century.
-In order to prepare ocean governance to meet the resource management challenges arising from shifting transboundary stocks, it is important to constrain when climate change will affect the dynamics of transboundary  stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
+In order to prepare ocean governance to meet the resource management challenges arising from shifting transboundary stocks, it is important to constrain when climate change will affect the dynamics of transboundary  stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @Link:2010ei, Chapter 2].
 
@@
 The concept of threat point comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
-Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary stocks between Canada and the United States (Sumaila et al., *in press*) .
+Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary stocks between Canada and the United States [@Sumaila:2020tp].
 
@@
 I employ a species distribution model driven by output from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species (Chapter 2) across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP 8.5 (see Methods).
-I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (Chapter 2) and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude  the high seas.
-I find that many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.
+I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (Chapter 2) and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude the high seas).
+I find that the distribution of many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.
 
@@
 Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd, Chapter 2], resulting in a total of 9,132 transboundary stocks in this analysis.
-Using the transboundary species Atlantic cod (*Gadus morhua*) as an example, and based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species.
-I defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1) including island territories.
-I determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
+Using the transboundary species Atlantic cod (*Gadus morhua*) as an example, and based on the definition of transboundary stock in this chapter, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species.
+I defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1)) including island territories.
+I determined the intersections between polygons using the R statistical software package sf [@PackagesfSimpleF:2018vp].
 Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
@@
 To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, I included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top 95th percentile (e.g., per = 95%).
-A sensitivity analysis was ran using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
+A sensitivity analysis was ran using a subset of species (n = 34) for all EEZs (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
 The centroid of each EEZ was estimated using the st package in R (Figure \@ref(fig:toeFigureS1)).
@@
 However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., “threat point” [@Sumaila:2013hv].
-Here, I define the threat point as the minimum required proportion of a shared stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Here, I define the threat point as the minimum required proportion of a shared stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [Chapter 2, @Sumaila:2020tp].
 Any proportion below the defined threat point of a country would result in a unilateral management of the shared stock.
@@
 I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SSR_{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR_{tm}$).
-I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
+I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
 The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members.
@@
 This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{te} \geqq (SSR_{th} + \sigma)$ or $SSR[t]_{e,m} \leqq (SSR_{th} - \sigma)$.
-Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al.
-(*in press*),
+Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following (See Chapter 4 - Methods),
 
@@
 
-```{r Load_Results, eval = TRUE,  echo = FALSE, include = FALSE}
+```{r Load_Results, eval = F,  echo = FALSE, include = FALSE}
 
@@
 There are only a few cases where the transboundary index of a stock does not emerge between 2006 and 2100.
-In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
+In the Arctic EEZs of Canada and the US, the present chapter included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
 In the case of Brazil, I included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
@@
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range."}
+Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range.", fig.scap="Average time of emergence of stocks’ transboundary index.", out.extra=''}
 
@@
 Fishing revenue on a logarithmic scale.
-Showing country names for the top 90th percentile of annual fishing revenue."}
+Showing country names for the top 90th percentile of annual fishing revenue.", fig.scap="Average time of emergence of stocks’ transboundary index per fishing entity, aggregated by colour and shape according to region", out.extra=''}
 
@@
 threshold (Figure \@ref(fig:toeFigureS2)).
-In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric,
+In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric.
 
@@
 Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods).
-I found significant differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
+I found significant statistical differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
 Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
@@
 
-```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association.
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=6,fig.pos= "H", fig.cap =  "Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association.
 Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Methods).
@@
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers with values exceeding 1.5 times the interquartile range."}
+Line represents median, and black points represent outliers with values exceeding 1.5 times the interquartile range.", fig.scap="Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association.", out.extra=''}
 
@@
 Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point.
-Panel B zooms in to specific areas shown within the grey dashed line boundaries in A."}
+Panel B zooms in to specific areas shown within the grey dashed line boundaries in A.", fig.scap="Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.", out.extra=''}
 
@@
 
-Our projections for the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century concur with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk].
+My projections for the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century concur with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk].
 For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US).
@@
 
-Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)).
+My results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)).
 For example, Polynesia and Micronesia are the only regions where the average time of emergence of the distribution of where all transboundary stocks across all EEZs emerged before 2022 (Figure \@ref(fig:toeFigure5)).
@@
 Moreover, projections suggest a larger increment on transboundary stocks in the coming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further north [@Morley:2018fn; @Frainer:2017ji].
-In the specific case of Brazil, in addition to having few transboundary stocks (only four shared with Uruguay in this study), the large difference in the EEZs lengths could also be influencing the results.
+In the specific case of Brazil, in addition to having few transboundary stocks (only four shared with Uruguay in this chapter), the large difference in the EEZs lengths could also be influencing the results.
 
@@
 Finally, flexible management rules might be more appropriate to capturing t distributional shifts.
-For example, a quota allocation based on the stock’s current distribution like that for Pacific halibut (*Hippoglossus stenolepis*) managed by Canada and the US [@Sumaila:2020tp; @PalaciosAbrantes:2020vt] would be more agile than one based on a fixed-historical proportion like the case of the EU [@Baudron:2020dj].
+For example, a quota allocation based on the stock’s current distribution like that for Pacific halibut (*Hippoglossus stenolepis*) managed by Canada and the US [@Sumaila:2020tp; Chapter 2] would be more agile than one based on a fixed-historical proportion like the case of the EU [@Baudron:2020dj].
 
@@
 My results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management.
-While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, my results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; @Gaines:2018sg; @PalaciosAbrantes:2020tv].
+While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, my results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; @Gaines:2018sg; Chapter 2].
 To address the issues raised in this chapter, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
@@
 
-```{r FigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
+```{r FigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this chapter as devided by the Sea Around Us and their centroids (points)."}
 

### 2020-12-22 Scripts/Supplements.Rmd
las changes

@@
 
-```{r toeFigureS1, eval = T,  echo = F, fig.width=12, fig.height=7, fig.cap =  "World Exclusive Economic Zones used in this study as defined by the Sea Around Us and their centroids (points)."}
+```{r toeFigureS1, eval = T,  echo = F, fig.width=12, fig.height=5, fig.cap =  "World Exclusive Economic Zones used in this study as defined by the Sea Around Us and their centroids (points)."}
 
@@
 The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel).
-Time of emergence is defined as the first year when the TI overpasses historical values (Bottom panel)."}
+Time of emergence is defined as the first year when the TI overpasses historical values (Bottom panel).", fig.scap="Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.",out.extra=''}
 
@@
 ```{r toeFigureS3, eval = T,  echo = F, fig.width=14, fig.height=7, fig.cap ="Sensitivity analysis of the number of grid cells with projected stock abundance within the neighbouring EEZs sharing the stock.
-The level used in this article was the top 95th percentile (e.g., far right column)."}
+The level used in this article was the top 95th percentile (e.g., far right column).", fig.scap="Sensitivity analysis of the number of grid cells with projected stock abundance within the neighbouring EEZs sharing the stock",out.extra=''}
 
@@
 Polygons represent the percentage of initial species that will see a shift.
-Panel B zooms in to specific areas shown in grey boxes of A."}
+Panel B zooms in to specific areas shown in grey boxes of A.", fig.scap="Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.", out.extra=''}
 

### 2020-12-22 Scripts/Initial_Analysis.Rmd
changed treshold to last year rather than first year

@@
       group_by(taxon_key,neighbour_id,n_ens,tresh) %>%
-      summarise(emerging_yr= min(year)) %>%
+      # summarise(emerging_yr= min(year)) %>% # First year that treshold is crossed
+      summarise(emerging_yr= max(year)) %>% # Last year that treshold is crossed
       left_join(EEZ_ids,

### 2021-01-28 Scripts/Initial_Analysis.Rmd
ready to run for p90

@@
 
-```{r Proportion_control_panel, eval = T, echo = F}
+```{r Proportion_control_panel, eval = F, echo = F}
 
@@
                           eez_centroid,
+                          p = NA,
                           Plot = F){
@@
              Neighbours = Neighbours,
-             eez_centroid = eez_centroid
+             eez_centroid = eez_centroid,
+             p = p
              )
@@
 
-```{r Emerging_routine, eval = F, echo = F}
+```{r Emerging_routine, eval = T, echo = F}
 
@@
 ####_______________________ ####
-n_cores <- 24
+# n_cores <- 24
 t_start <- Sys.time()
@@
                eez_centroid = eez_centroid,
+               p = 0.90,
                Plot = F
@@
 
-```{r get_env_data, eval = T, echo = T}
+```{r get_env_data, eval = F, echo = T}
 
@@
 
-```{r ToE_generals,  eval = T, echo = T}
+```{r ToE_generals,  eval = F, echo = T}
 
@@
 
-```{r Time_lapse_data,  eval = T, echo = T}
+```{r Time_lapse_data,  eval = F, echo = T}
 
@@
 
-```{r Time_lapse_plot, eval = T, echo = F}
+```{r Time_lapse_plot, eval = F, echo = F}
 
@@
 
-```{r emergence_map, eval = T}
+```{r emergence_map, eval = F}
 
@@
 
-```{r data_loading, eval = T, echo = F}
+```{r data_loading, eval = F, echo = F}
 
@@
 
-```{r data_analysis, eval = T, echo = F}
+```{r data_analysis, eval = F, echo = F}
 
@@
 
-```{r time_emergence_map,  eval = T, echo = T}
+```{r time_emergence_map,  eval = F, echo = T}
 
@@
 
-```{r histo,  eval = T, echo = T}
+```{r histo,  eval = F, echo = T}
 
@@
 
-```{r temp_plot, eval = T, echo = T}
+```{r temp_plot, eval = F, echo = T}
 
@@
 
-```{r Scatter_plot,  eval = T, echo = T}
+```{r Scatter_plot,  eval = F, echo = T}
 

### 2021-02-01 Scripts/Initial_Analysis.Rmd
Ran analysis for last year over 90 per

@@
   "tidyverse",
-  "here", # for dbem_import `here()`
+  "here", # for dbem_import `here()`da
   "data.table", #dbem_import `fread()`
@@
                          eez_centroid,
-                         p = 0.95 # percentile of data for central distribution
+                         p # percentile of data for central distribution
                          ){
@@
 
-    Final_Result <- mean_sd_index %>%
+    Raw_Result <- mean_sd_index %>%
       filter(year > Hist_y) %>%
@@
       ) %>%
-      gather("tresh","YesNo",tresh_one,tresh_two) %>% 
-      filter(!is.na(YesNo)) %>% 
-      group_by(taxon_key,neighbour_id,n_ens,tresh) %>%
-      # summarise(emerging_yr= min(year)) %>% # First year that treshold is crossed
-      summarise(emerging_yr= max(year)) %>% # Last year that treshold is crossed
+      gather("tresh","emerged",tresh_one,tresh_two)
+
+    taxon <- unique(Raw_Result$taxon_key)
+    name_raw <- paste(my_path("R",extra_path = "/Emergence_last_raw_p90/"),taxon,"_emergence_raw.csv",sep="")
+
+    write_csv(Raw_Result,
+              name_raw)
+
+    ToE_Result <- Raw_Result %>% 
+    # Method the last year when the index overshoots and never goes back
+      filter(is.na(emerged)) %>% 
+      group_by(taxon_key, neighbour_id,n_ens,tresh) %>% 
+      summarize(emerging_yr = max(year)+1) %>% # last year of NA + 1 makes it the first year of overshooting
       left_join(EEZ_ids,
@@
 
-  return(Final_Result)
+  return(ToE_Result)
   }
@@
 
-    name <- paste(my_path("R",extra_path = "/Emergence_2005/"),taxon_key,"_emergence.txt",sep="")
+    name <- paste(my_path("R",extra_path = "/Emergence_last_p90/"),taxon_key,"_emergence.txt",sep="")
 
@@
 
-```{r Emerging_routine, eval = T, echo = F}
+```{r Emerging_routine, eval = F, echo = F}
 
@@
 
-```{r, olda_acumulative}
+```{r, olda_acumulative, eval = F}
 

### 2021-02-16 Scripts/Initial_Analysis.Rmd
SSH Proportion set to over 2sd and figures eddits

@@
     mutate(
-      top_tresh = ensemble_mean+ensemble_sd,
-      low_tresh = ensemble_mean-ensemble_sd
+      top_tresh = ensemble_mean+(2*ensemble_sd), # two s.d.
+      low_tresh = ensemble_mean-(2*ensemble_sd) # two.s.d
     ) %>% 
@@
       summarise(
-        mean_spp = mean(ensemble_mean,na.rm=T) # average spp change from all countries sharing
+        mean_spp = mean(ensemble_mean,na.rm=T) # average spp change from all countries sharing that speceies
       ) %>% 
@@
 
-**Note: Contains step to filter out changes less than +-1 s.d.**
+**Note: Contains step to filter out changes less than +-2 s.d.**
 
@@
 
-    # Number of SSR stocks
+    # Percentage of SSR stocks changing
     n_ssr <- spp %>% 
@@
       summarise(
-        n_spp = length(unique(taxon_key))
-      ) %>% 
-      spread(time_step,n_spp) %>% 
+        # n_spp = length(unique(taxon_key)),
+        n_stock = n(),
+      ) %>% 
+      spread(time_step,n_stock) %>% 
       mutate_at(vars(early,mid),
@@
       summarise(
+        n_spp = length(unique(taxon_key)),
+        n_stock = n(),
         min_eez = min(mean_spp, na.rm=T),
@@
         ) %>% 
-      select(eez_name:max_eez,mean_eez,sd_eez) %>% 
+      select(eez_name:max_eez,mean_eez,sd_eez,n_spp,n_stock) %>% 
       left_join(n_ssr)
@@
         n_spp = length(unique(taxon_key)),
-        # n_spp = n(),
+        n_stock = n(),
         min_eez = min(mean_spp, na.rm=T),
@@
         ) %>% 
-      select(eez_name:max_eez,mean_eez,sd_eez)
+      select(eez_name:max_eez,mean_eez,sd_eez,n_spp,n_stock)
 
@@
   proportion_change <- eez %>% 
-    gather("variable","value",min_eez:n_per) %>% 
+    gather("variable","value",n_spp:n_per) %>% 
     spread(time_step,value) %>% 
@@
     ),
-    funs(round((((.-historical)/historical))*100,2))
+    # funs(round(((.-historical)/(abs(historical)))*100,2)
+    funs(ifelse(variable == "n_per",.,round(((.-historical)/(abs(historical)))*100,2)
+                )
+         )
     )
@@
   filter(!is.na(eez_neighbour)) %>%
-  mutate(per_mid_plot =ifelse(per_mid >= 50, 50,
-                              ifelse(per_mid <= -50,-50,per_mid)
+  mutate(per_mid_plot =ifelse(per_mid >= 100, 100,
+                              ifelse(per_mid <= -100,-100,per_mid)
   ),
-  per_early_plot =ifelse(per_early >= 50, 50,
-                         ifelse(per_early <= -50,-50,per_early)
+  per_early_plot =ifelse(per_early >= 100, 100,
+                         ifelse(per_early <= -100,-100,per_early)
   )
@@
 
-Prop_results <- my_path("R",name="proportion_change_2005.csv", read=T)
+# Prop_resultsb <- my_path("R",name="proportion_change_2005.csv", read=T)
 
@@
             by=  "sf_eez_name") %>% 
-  clean_names()
+  clean_names() %>% 
+  mutate(tresh = str_to_lower(tresh))
 
@@
 path_world <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile")
+path_world <- "/Volumes/Enterprise"
 
@@
 
-World_eez_sf %>% 
-  filter(str_detect(eez_name,"Russia")) %>% 
-  ggplot() +
-  geom_sf() +
-  aes(fill = eez_name)
+# World_eez_sf %>% 
+#   filter(str_detect(eez_name,"Russia")) %>% 
+#   ggplot() +
+#   geom_sf() +
+#   aes(fill = eez_name)
 
@@
 
+### ------------ ##
+# Select a trshold
+selected_tresh <- "tresh_two"
+### ------------ ##
+
 # Total number of emerging stocks
@@
   gather("metric","value",com_sum,per) %>% 
-  filter(tresh == "Tresh_one")
+  filter(tresh == selected_tresh)
 
@@
   filter(taxon_emer <= 2020,
-         tresh == "Tresh_one") %>% 
+         tresh == selected_tresh) %>%  # Presenting results for the second treshold
   group_by(level,metric) %>% 
@@
     min(taxon_emer),
-    max(taxon_emer),
-    min(taxon_emer)
+    mean(taxon_emer),
+    sd(taxon_emer)
   )
@@
   filter(taxon_emer <= 2020,
-         tresh == "Tresh_one",
+         tresh == "tresh_one",
          metric == "per")
@@
   filter(!fishing_entity_id %in% Value_map_data$fishing_entity_id) %>% 
-  filter(!sovereignt %in% c("Antarctica","Brazil"))
-
-grey_eez <- World_eez_sf %>% 
-  filter(!eezid %in% Eez_d_hist$eezid,
-         !eez_name %in% Neighbours$eez_name)
-
-Brazil_land <- World_map %>% 
-  filter(sovereignt == "Brazil")
-
-Brazil_eez <- World_eez_sf %>% 
-  filter(eez_name %in% c("Brazil","USA (Alaska, Arctic)"))
-
+  filter(!sovereignt %in% c("Antarctica"))
+
+# grey_eez <- World_eez_sf %>% 
+#   filter(!eezid %in% Eez_d_hist$eezid,
+#          !eez_name %in% Neighbours$eez_name)
+
+# -------------------------- #
+# Call plot for both tresholds
+# -------------------------- #
 for(i in 1:2){
 
-  tresh_list <- c("Tresh_one","Tresh_two")
+  tresh_list <- c("tresh_one","tresh_two")
   name_save <- paste("./Figures/emer_map_lat_",tresh_list[i],".png",sep = "")
 
-  ## --------------- ##
-  # Latitudinal plot
-  ## --------------- ##
-
-  ### Random plot
-
-# Eez_hist <- Eme_Res %>% 
-#   group_by(eezid,eez_name = sf_eez_name) %>% 
-#   left_join(Eez_centroid) %>% 
-#   filter(tresh == tresh_list[i]) %>% 
-#   group_by(latitude) %>% 
-#   summarise(
-#     number = n()
-#   ) %>% 
-#   group_by() %>% 
-#   mutate(RMean = rollmean(x = number, 
-#                           10, 
-#                           align = "center", 
-#                           fill = number
-#   )
-#   ) %>% 
-#   mutate(lab="lab") %>% 
-#   ggplot() +
-#   geom_area(
-#     aes(
-#       x = latitude,
-#       y = RMean,
-#       fill = lab,
-#       color = lab
-#     )
-#   ) +
-#   coord_flip() +
-#   scale_x_continuous("",
-#     limits = c(-50,80),
-#     breaks = c(-50,0,50,80)
-#   ) +
-#   scale_y_continuous(
-#     "Emerging stocks",
-#     limits = c(0, 124),
-#     breaks = c(0,62,124)
-#     
-#   ) +
-#   scale_fill_viridis(option="magma", discrete = T, begin = 0.7) +
-#   scale_color_viridis(option="magma", discrete = T, begin = 0.7) +
-#   theme_minimal() + 
-#   theme(
-#     panel.background = element_rect(fill = "white", colour = "white"),
-#     panel.grid.major.y = element_blank(),
-#     panel.grid.minor.y = element_blank(),
-#     panel.grid.major.x = element_blank(),
-#     panel.grid.minor.x = element_blank(),
-#     axis.text.y = element_blank(),
-#     axis.ticks.y = element_blank(),
-#     axis.text = element_text(size = 12, colour="black"),
-#     axis.title.x = element_text(size = 12, colour="black", face = "bold"),
-#     legend.position = ""
-#   )
-#   
-  ## --------------- ##
-  # Map
-  ## --------------- ##
-
-  World_map_part <- World_map_nemer# %>% 
-    # filter(
-    #   tresh ==  tresh_list[i], 
-    #   variable == "value"
-    # )
-
-  map <- World_eez_sf %>% 
+  # Value Map
+  World_map_part <- World_map_nemer %>% 
+  filter(
+    tresh ==  tresh_list[i],
+    variable == "value"
+  )
+
+  # ToE Map 
+  toe_map_data <- World_eez_sf %>% 
     left_join(Eez_d_hist,
@@
     filter(tresh == tresh_list[i],
-           metric == "mean_yr") %>% 
-    ggplot() +
+           metric == "mean_yr")
+
+  # Map with countries without ToE
+
+  russia_churchil <- World_eez_sf %>% 
+    filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))
+
+  no_toe_eez <- World_eez_sf %>% 
+    filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
+    rbind(russia_churchil)
+
+  no_toe_land <- World_map %>% 
+    filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !name %in% World_map_part$name) # Filter EEZs with ToE stocks
+
+  map <- ggplot(toe_map_data) +
     geom_sf(
@@
     geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
-    geom_sf(data = grey_eez, aes(), fill = "white", color =  "grey50",size = 0.1) +
-    geom_sf(data = Brazil_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
-    geom_sf(data = Brazil_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    # geom_sf(data = grey_eez, aes(), fill = "white", color =  "grey50",size = 0.1) +
+    geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
     # Set Participation labels
     annotate("text",
-             x =-135,
+             x =-150,
              y = -69,
@@
     annotate("text", 
-             x =-135,
+             x =-150,
              y = -85,
-             label = "Time of Emergence",
+             label = "Year of range shift",
              color="black", 
@@
              y = -85,
-             label = seq(2100,2000,-25), # original
-             # label = c(">2100",seq(2075,2000,-25)), # not inluding the > because we dont know
+             # label = seq(2100,2000,-25), # original
+             label = c(">2100",seq(2075,2000,-25)), # not inluding the > because we dont know
              color="black", 
@@
              # fontface="bold",
+             hjust=0) +
+    # Set No emergence label
+    annotate("text",
+             x =110,
+             y = -77,
+             label = "No shift",
+             color="black",
+             size=5,
+             angle=0,
+             fontface="bold",
              hjust=0) +
@@
     theme(
-      legend.box.margin = margin(t = -80, l = 40),
+      legend.box.margin = margin(t = -62, l = 252),
       legend.text = element_blank(),
@@
   boxplot_data <- Eme_Res %>% 
-    filter(tresh == "Tresh_one",
+    filter(tresh == tresh_list[i],
            !is.na(un_sub_region)) %>%
@@
     left_join(Eme_Res) %>% 
-    filter(tresh == "Tresh_one",
+    filter(tresh == tresh_list[i],
            !is.na(un_sub_region)) %>% 
@@
 
+  # ---------- #
+  # Summary data
+  # ---------- #
+
+  medians_data <- boxplot_data %>% 
+    group_by(un_sub_region,un_geo_region) %>% 
+    summarise(mean = mean(emerging_yr,na.rm = T),
+              sd = sd(emerging_yr,na.rm = T),
+              min = min(emerging_yr,na.rm = T),
+              max = max(emerging_yr,na.rm = T),
+              n_stocks = n()
+    )
+
+  boxplot_data <- boxplot_data %>% 
+    left_join(medians_data,
+              by = c("un_sub_region","un_geo_region")
+              )
+
   boxplot <- ggplot(boxplot_data) +
-    geom_boxplot(
+    geom_violin(
       aes(
@@
         y = emerging_yr#,
-        # fill = un_geo_region
+        # fill = mean,
+        # colour = mean
       ),
@@
                nrow = 1) +
-    scale_fill_viridis("UN Sub region",
-                       discrete = T) +
-    scale_color_grey("") +
+    # scale_fill_viridis("",
+    #                    limits = c(2000,2100),
+    #                    breaks = seq(2000,2100,25),
+    #                    labels = c("","","","","")#,
+    #                    # direction = -1
+    # ) +
+    # scale_colour_viridis("",
+    #                    limits = c(2000,2100),
+    #                    breaks = seq(2000,2100,25),
+    #                    labels = c("","","","","")#,
+    #                    # direction = -1
+    # ) +
     xlab("") +
-    ylab("Time of Emergence") +
+    ylab("Year of range shift") +
+    scale_y_continuous(
+                   labels = c(2000,2025, 2050, 2075, ">2100"), # not inluding the > because we dont know
+    ) +
     my_ggtheme_p() +
@@
       # axis.ticks.x = element_blank(),
-      legend.position = "right"
+      legend.position = ""
     )
@@
             by = c("sau_fishing_entity","tresh")) %>% 
-  filter(tresh == "Tresh_one",
+  filter(tresh == selected_tresh, # tresh hold selection
          variable == "value") %>% 
@@
   y = c(-5.8,-5.8,Percentile_value$x+.1),
-  label = c("Early emergence", "Late emergence","Fisheries > 75th percentile")
+  label = c("Early shift", "Late shift","Fisheries > 75th percentile")
 )
@@
   # right side labels
-  geom_label_repel(data = subset(Buble_data, mean_yr > 2035 & sum_value_bi >= 1),
-                  aes(
-                    x = mean_yr,
-                    y = log_value,
-                    label = sau_fishing_entity
-                  ),
-                  arrow = arrow(length = unit(0.009, "npc"),
-                                type = "closed",
-                                ends = "first"),
-                  force = 2,
-                  box.padding = 0.2,
-                  fill = "NA",
-                  nudge_x = 2050 + subset(Buble_data, mean_yr > 2035 & sum_value_bi >= 1)$mean_yr,
-                  segment.size  = 0.2,
-                  segment.color = "grey50",
-                  direction     = "y",
-                  hjust         = 1
-  ) +
-  # right side labels
-  geom_label_repel(data = subset(Buble_data, mean_yr < 2035 & sum_value_bi >= 1),
+  geom_label_repel(data = subset(Buble_data, mean_yr > 2045 & sum_value_bi >= 0.90),
                   aes(
@@
                   fill = "NA",
-                  nudge_x = 2010 - subset(Buble_data, mean_yr < 2035 & sum_value_bi >= 1)$mean_yr,
-                  segment.size  = 0.2,
+                  nudge_x = 2075 - subset(Buble_data, mean_yr > 2045 & sum_value_bi >= 0.90)$mean_yr,
+                  segment.size  = 0.1,
+                  segment.color = "grey50",
+                  direction     = "y",
+                  hjust         = 1
+  ) +
+  # LEFT side labels
+  geom_label_repel(data = subset(Buble_data, mean_yr < 2045 & sum_value_bi >= 0.90),
+                  aes(
+                    x = mean_yr,
+                    y = log_value,
+                    label = sau_fishing_entity
+                  ),
+                  arrow = arrow(length = unit(0.009, "npc"),
+                                type = "closed",
+                                ends = "first"),
+                  force = 1,
+                  box.padding = 0.2,
+                  fill = "NA",
+                  nudge_x = 2025 - subset(Buble_data, mean_yr < 2045 & sum_value_bi >= 0.90)$mean_yr,
+                  segment.size  = 0.1,
                   segment.color = "grey50",
@@
             ),
-            size = 3,
+            size = 5,
             color = "grey30"
@@
                       discrete = T) +
-  scale_x_continuous("Average time of emergence",
+  scale_x_continuous("Average year of range shift",
                      limits = c(2000,2080),
@@
   ) +
-  scale_size("Number of\nemerging stock") +
+  scale_size("Number of\nshifting stocks") +
   ggsave("./Figures/bubble_plot.png",
@@
 
+```{r participation_text, eval = T, echo=F}
+
+Value_map_data %>% 
+  group_by(tresh,variable) %>% 
+  summarise(m = mean(participation)-2000,
+            sd = sd(participation-2000)
+            )
+
+```
+
 ### Species level analysis
@@
 ```{r deme_pela, eval = F, echo =F}
+
+Dem_pel <- Eme_Res %>% 
+  filter(tresh == selected_tresh,
+         !is.na(un_sub_region)) %>% 
+  left_join(Spp_env_pref) %>% 
+  mutate(DemersPelag = ifelse(is.na(DemersPelag),"other",paste(DemersPelag)
+  #                             ifelse(DemersPelag %in% c("bathypelagic","benthopelagic"), "bathy-bentho-pelagic",paste(DemersPelag)
+  #                             )
+   ),
+    un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and the Car.",paste(un_sub_region)))
 
@@
   mutate(DemersPelag = ifelse(is.na(DemersPelag),"other",DemersPelag)) %>% 
-  group_by(DemersPelag) %>% 
+  group_by(DemersPelag,tresh) %>%
   summarise(mean_toe = median(emerging_yr,na.rm=T),
-            n_spp = length(unique(taxon_key))) %>% 
+            n_spp = length(unique(taxon_key)),
+            n_stock = n()
+  ) %>%
   left_join(Dem_pel) %>%
   mutate(
-    label = paste0(DemersPelag, "\n(n=",n_spp,")")
-  ) %>% 
-  filter(tresh == "Tresh_one",
+    label = paste0(DemersPelag, "\n(n=",n_stock,")")
+  ) %>%
+  filter(tresh == selected_tresh,
          DemersPelag != "pelagic") %>% # just one spp.
   ggplot() +
+  # geom_violin(
   geom_boxplot(
@@
     alpha = 0.8,
-    fill = "grey"
-  ) +
-  # scale_fill_viridis("Habitat preference",
-                     # discrete = T) +
+    fill = "grey",
+    coef = 0, # to have dotted error bars
+    outlier.shape = NA # to have dotted error bars
+  ) +
+  # to have dotted error bars
+  stat_boxplot(
+    aes(x = reorder(label,mean_toe),
+        # aes(x = DemersPelag,
+        y = emerging_yr,
+        # fill = DemersPelag
+    ),
+    geom ='errorbar',
+    width = 0.3,
+    linetype = "dashed") + 
   xlab("") +
-  ylab("Time of emergence") +
+  ylab("Year of range shift") +
   my_ggtheme_p() +
@@
 Stats_data <- Eme_Res %>% 
-  filter(tresh == "Tresh_one",
+  filter(tresh == selected_tresh,
          n_ens == 10)
@@
 Reg_diff_data <- Eme_Res %>% 
-  filter(tresh == "Tresh_one",
+  filter(tresh == selected_tresh,
          !is.na(un_sub_region)
@@
   # head()
+  View()
   write_csv("kw_regional.csv")
@@
 Hab_diff_data <- Eme_Res %>% 
-  filter(tresh == "Tresh_one",
+  filter(tresh == selected_tresh,
          !is.na(un_sub_region)
@@
   # head()
+  View()
   write_csv("kw_habitat.csv")
@@
 
-##### Fig. 1 and S1. (Netowrk)
+##### Fig. 5 and S5. (Proportion + Netowrk)
 
@@
 
-### Median number
+### Median numbef of stocks changing in each time step
 
 Prop_results %>% 
-  filter(variable == "n_spp") %>% 
+  filter(variable == "n_per") %>% 
   group_by() %>% 
-  summarise(
-    median(early,na.rm=T),
-    sd(early,na.rm=T)
+  summarise_at(vars(early,mid),funs(
+    median(.,na.rm=T),
+    sd(.,na.rm=T),
+    length(unique(eez_name)),
+    length(unique(eez_name))/280*100
+    )
     )
@@
   gather("time_step","per_chng",per_early:per_mid) %>% 
-  mutate(per_plot = ifelse(per_chng > 100,100,per_chng),
+  mutate(per_plot = ifelse(per_chng > 50,50,per_chng),
          direction = ifelse(per_chng > 0,"positive","negative")
@@
   gather("time_step","per_chng",per_early:per_mid) %>% 
-  mutate(per_plot = ifelse(per_chng > 100,100,per_chng))
+  mutate(per_plot = ifelse(per_chng > 50,50,per_chng))
 
@@
   summarise(
-    n_trans = unique(length(taxon_key))
+    # n_trans = unique(length(taxon_key))
+    n_trans = n()
   )
@@
   summarise(avrg_spp_per = mean(avrg_spp_per,na.rm=T)) %>% 
-  mutate(time_step = ifelse(time_step == "early","per_early","per_mid"))
-
-# Paper numbers
-mean(n_taxa_data$avrg_spp_per,na.rm=T)
-sd(n_taxa_data$avrg_spp_per,na.rm=T)
+  mutate(time_step = ifelse(time_step == "early","per_early","per_mid")) %>% 
+  mutate(avrg_spp_per = ifelse(avrg_spp_per > 50,50,avrg_spp_per))
 
@@
             )
+
+# Grey land
+grey_land <- World_map %>% 
+  filter(is.na(fishing_entity_id)) %>% 
+  filter(!sovereignt %in% c("Antarctica"))
 
@@
       limits = c(0,max(Map_data$per_plot)),
-      breaks = seq(0,max(Map_data$per_plot),10)#,
+      breaks = seq(0,max(Map_data$per_plot),10),
+      labels = c(seq(0,max(Map_data$per_plot-10),10),">50")
       # option="magma",
@@
       breaks = seq(0,max(Map_data$per_plot),10),
+      labels = c(seq(0,max(Map_data$per_plot-10),10),">50")
       # alpha = 0.8

### 2021-02-16 Scripts/Manuscript.Rmd
Changed discussion and added supplemental fishbase table

--- 2020-12-14 Scripts/Manuscript.Rmd
@@
 ---
+title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change
+subtitle: Challenges to transboundary fisheries under climate change
+author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
+date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
+^2^Climate and Environmental Physics, Physics Institute, University of Bern.
+^3^Oeschger Centre for Climate Change Research, University of Bern, Switzerland.
+^4^School of Environmental Sciences, University of Liverpool.
+^5^Center for Ocean Solutions, Stanford University, Stanford, United States'
 output:
-  bookdown::pdf_book:
-    toc : no
-    keep_tex: no
-    number_sections: yes
-geometry: margin=1in
-editor_options:
-  chunk_output_type: inline
+  # word_document:
+    # reference_docx: ../References/template.docx
+  pdf_document: default
 bibliography: ../References/ToE_Ref.bib
+csl: ../References/science.csl
 header-includes:
@@
 ---
+
+<!-- # Instructions  -->
+
+<!-- Research articles should present a major advance and can be short or long (maximum 15,000 words) format.
+Submissions must include an abstract (up to 150 words), an introduction, and sections with brief informative subheadings.
+You may include up to ten figures and/or tables and about 60 references.
+Supplementary materials should be limited to information that is not essential for the general understanding of the research presented in the main text and can include data sets, figures, tables, videos, or audio files.
+You may include up to a total of ten figures and/or tables (combined) throughout the supplemental text.
+-->
 
@@
   "grid",
-  "knitr"#,
+  "knitr",
+ "kableExtra"
+  # "here", # for dbem_import `here()`
+  # "data.table", #dbem_import `fread()`
+  # "readxl", # for reading excell files
+  # "janitor", # for clearing names
+  # "geosphere", # estimate distances between points `distm()`
+  # "ggrepel",
+  # "zoo", # for average mean
+  # "parallel", # for mclapply,
+  # "sf", # for mapping
+  # "st", # for mapping
+  # "rgdal", #Spatial analysis
+  # "tools", #Spatial analysis 
+  # "zeallot", # for Juanito's map
+  # "gmt", # for estimating distances between points
+  # "viridis"
 )
@@
 
+<!-- * corresponding author: j.palacios@oceans.ubc.ca -->
+
+<!-- # Abstract (150  words) -->
+
+<!-- An opening sentence that states the question/problem addressed by the research AND Enough background content to give context to the study AND A brief statement of primary results AND A short concluding sentence.
+-->
+
+Climate change shifts in marine species distributions is affecting the current frameworks emplaced to manage species shared fisheries.
+Here, we rely on a modeling framework with multiple ensemble members to identify the time frame by which shared stocks distribution will change from historical ranges, and the change in the proportion of the distrirbution that each EEZ has under climate change.
+Our results suggest that 60% of shared species have emerged since 2006 in 87% of the world EEZs.
+By 2075 all EEZs will have an emerging species and by 2097 and all species will have emerged in at least one EEZ.
+Average changes in the ratio of species-share between neighboring EEZs will fluctuate between 1 and 10% with some exceeding 25% by 2030 relative to today’s ratio.
+Our results suggest that most changes will happen before 2030 urging policy to be re-formulated to account for the possible implications of these changes.
+
 ## Introduction
@@
 Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn].
-Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk]).
+Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk].
 As a result of climate change, many marine species have changed their distributions towards higher latitude, deeper water or following local temperature gradients [@Poloczanska:2016kk].
 Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].
-The projected risks of impacts can be reduced through improving the effectiveness of current fisheries management [@Gaines:2018sg; @Cheung:2018dh], including the fisheries management of species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
+The projected risks of impacts can be reduced through improving the effectiveness of current fisheries managemen [@Gaines:2018sg; @Cheung:2018dh], including the fisheries management of species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
 
 The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
-As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (i) transboundary stocks, those that cross neighboring EEZs; (ii) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; (iii) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (iv) discrete stocks that are only present in the high seas [@Munro:2004th].
+As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, those that cross neighboring EEZs; (*ii*) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (*iV*) discrete stocks that are only present in the high seas [@Munro:2004th].
 This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
 While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 (Chapter 2).
+A recent study estimates that, globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, and yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
 
 The effectiveness of fisheries management for transboundary stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
-In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the stocks [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
-Misalignment between fisheries resources allocation and stocks’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb].
-With future shifts in species distributions under climate change, the challenges for the fisheries management of transboundary stocks will continue to increase during the 21st century.
+In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb].
+With future shifts in species distributions under climate change, the challenges for the fisheries management of transboundary stocks will continue to increase during the $21^{st}$ century.
 In order to prepare ocean governance to meet the resource management challenges arising from shifting transboundary stocks, it is important to constrain when climate change will affect the dynamics of transboundary stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
@@
 The concept of threat point comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
-Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary stocks between Canada and the United States (Sumaila et al., *in press*) .
+Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary species between Canada and the United States [@Sumaila:2020tp].
 
-Here, I estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
-I employ a species distribution model driven by output from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species (Chapter 2) across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP 8.5 (see Methods).
-I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (Chapter 2) and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude  the high seas.
-I find that many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.
+Here, we estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
+We employ a species distribution model driven by output from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species [@PalaciosAbrantes:2020tv] across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
+We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude the high seas).
+We find that many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.
 
@@
 
-This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 (Chapter 2).
-Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd, Chapter 2], resulting in a total of 9,132 transboundary stocks in this analysis.
+This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
+Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv]; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks in this analysis.
 Using the transboundary species Atlantic cod (*Gadus morhua*) as an example, and based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species.
-I defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1) including island territories.
-I determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
+We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1)) including island territories.
+We determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
 Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
 The habitat association of each species was determined following the classification in FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org).
-For each stock and EEZ, I used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
-I report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+For each stock and EEZ, we used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
+We report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
 The monthly average US Consumer Price Index (CPI) according to the U.S.
@@
 DBEM then estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid (see Cheung et al.
-[-@Cheung:2010dt; -@Cheung:2016jd] for model details).
-Importantly, DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
+2010 and 2016 for model details).
+Importantly, DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd], Cheung et al.
 *in revision*].
 
-The DBEM was forced with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+We forced the DBEM with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
 The GFDL-ESM2M is run under historical forcing until 2005 and follows the high greenhouse gas emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, I have to understand distribution variability during both the historical and the future periods, to infer differences between time frames.
-I do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
-Thus, for my experiment, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in year 1950 resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand distribution variability during both the historical and the future periods, to infer differences between time frames.
+We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
+Thus, for our experiment, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in year 1950 resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
 By design, variations among ensemble members are then solely due to natural variability.
@@
 I developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
-This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2).
-The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., ^Pacific halibut (*Hippoglossus stenolepis*) that is transboundary between the United States and Canada.
+This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2)).
+The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
 Therefore,
@@
 where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
-To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, I included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top 95th percentile (e.g., per = 95%).
-A sensitivity analysis was ran using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
+To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the  95th percentile (e.g., per = 95%).
+We ran a sensitivity analysis using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
 The centroid of each EEZ was estimated using the st package in R (Figure \@ref(fig:toeFigureS1)).
-For each ensemble member, neighboring EEZs and transboundary stock, I computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
+For each ensemble member, neighboring EEZs and transboundary stock, the distance between centroids was computed assuming the Earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
 
@@
 
-where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the pair of neighbouring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+where $D_A$ and $D_B$ represent the distance between the distribution centroids of a species and the centroid of the pair of neighbouring EEZs (A and B) sharing the species for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
 Thus, a higher TI means that a shared stock becomes more or less abundant relative to its neighbouring EEZs that is beyond what the fisheries experienced historically.
@@
 \label{eq:four}
-EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}} \\
-ToE = ToE_{t} > 1 s.d.
+ETI = \frac{\mu_{TI[t]}}{SD_{TI_{th}}} \\
+ToE = t when Emergence index[t] > 1 s.d.
 \end{split}
@@
 
-where $\mu_{TI}$ is the yearly mean TI ($\mu$) across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005).
-I set a threshold of TI above which I consider a shared stock to have emerged from the historical internal variability of the TI.
+where $\mu_{TI}$ is the yearly mean TI across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005).
+We set a threshold of TI above which it was consider a shared stock to have emerged from the historical internal variability of the TI.
 The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four), representing a probability of 68% that the transboundary index has emerged from historical variability).
-I tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
+We tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
 
@@
 
-<!-- I created a country-specific vulnerability index (VI) to examine the risk of impacts of species’ range shifts on transboundary fisheries management.
+<!-- We created a country-specific vulnerability index (VI) to examine the risk of impacts of species’ range shifts on transboundary fisheries management.
 The VI reflects both the rate of change of shared stocks’ distribution (indicated by the time-of-emergence of the transboundary index) and the importance of the shifting species to a country’s total catch and revenue from the species.
@@
 
-In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than those operating on a non-cooperative basis [@Munro:1979dg].
+In game theory, cooperation over a transboundary resource will more likely result in a better overall outcome for the sharing parties than those operating on a non-cooperative basis [@Munro:1979dg].
 However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., “threat point” [@Sumaila:2013hv].
-Here, I define the threat point as the minimum required proportion of a shared stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-Any proportion below the defined threat point of a country would result in a unilateral management of the shared stock.
+Here, we define the threat point as the minimum required proportion of a transboundary stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Any proportion below the defined threat point of a country would result in a deflection from the joint management of the transboundary stock.
 
-First, I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
-I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2).
-Second, I averaged the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of variability.
-The first time period ($SSR_{th}$) spans 1951 to 2005 and is chosen to match the historical period in the GFDL ESM2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
-It is assumed to represent the historic baseline of shared distributions for each stock.
-I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$).
-I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
+First we estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
+We did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
+Second, we average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
+The first time period ($SSR_{th}$) spans from 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
+It is assumed as the historic baseline of shared distributions for each stock.
+We then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$).
+We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
 The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members.
-Third, I defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$.
+Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$.
 This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{et} \geqq (SSR_{th} + \sigma)$ or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
-Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al.
+Finally, we estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al.
 (*in press*),
@@
 The time of emergence results were tested for normality (e.g.
-skewness, kurtosis) and performed two non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary stocks.
-Specifically, I tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of transboundary stocks.
-For both cases, my null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
-<!-- Moreover, I implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
-Specifically, I looked at habitat associations, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
-I hypothesize that all of these factors would have significant influence in the ToE of transboundary species.
+skewness = 1.02, kurtosis = 2.83) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
+Specifically, we tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of the transboundary index of transboundary stocks.
+For both cases, our null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
+<!-- Moreover, we implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
+Specifically, we looked at habitat associations, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
+We hypothesize that all of these factors would have significant influence in the ToE of transboundary species.
 -->
@@
 The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:toeFigure1)).
-This means that on average, the distribution of transboundary stocks will diverge from their natural internal variability by 2029.
 In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100.
 Fish stocks with transboundary index that first emerged from their historical variability in each EEZ was projected to be as early as 2006 and the last one in 2100.
-About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g., Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010).
+About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g.
+Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010).
 The number of EEZs with emerging stocks is projected to increase steeply from 87% in 2006 to 96% in 2020.
-After the mid-2020s, the rate of increase in the number EEZs slows from a 1% to an average of 0.7% per year.
+After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.7% per year.
 By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:toeFigure1)).
@@
 Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure \@ref(fig:toeFigure1)).
-I based the transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
-Results from the sensitivity analysis suggest that this method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g., the top 95%) of the stock (Figure \@ref(fig:toeFigureS3).
+We based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
+Results from the sensitivity analysis suggest that our method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g.
+the top 90%) of the stock (Figure \@ref(fig:toeFigureS3)).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence of the transboundary index by Exclusive Economic Zone (EEZ) and transboundary stock"}
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone (EEZ) and transboundary stock."}
 
@@
 
-The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; \@ref(fig:toeFigure2)A).
-Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (\@ref(fig:toeFigure2)B).
-In contrast, EEZs located in temperate regions like northern Europe and eastern Asia have significantly later times of emergence (p < 0.05; see Table @ref(tab:toeTableS1 for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058).
-However, notable exceptions to these broad patterns exist, with  tropical EEZs displaying a later average time of emergence across stocks including the  Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070) and despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) display a relatively early average time of emergence (\@ref(fig:toeFigure2)A).
-There are only a few cases where the transboundary index of a stock does not emerge between 2006 and 2100.
+The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A).
+Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1)) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B).
+In contrast, EEZs located in template regions like northern Europe and eastern Asia will have a significantly later time of emergence (p < 0.05; see Table \@ref(tab:toeTableS1 for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058).
+However, notable exceptions of tropical EEZs with a later average time of emergence across stocks include Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070).
+Moreover, despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) will have relatively early average time of emergence (Figure \@ref(fig:toeFigure2)A).
+Only in a few cases, the transboundary index of a stock will not emerge between 2006 and 2100.
 In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
-In the case of Brazil, I included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
-However, on all cases, I find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (\@ref(fig:toeFigure2)A).
+In the case of Brazil, there were four transboundary stocks included, all shared with Uruguay [@PalaciosAbrantes:2020tv]: Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
+In all cases, our study suggest that the transboundary index of these stocks will not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A)
 
@@
 
-My analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see the distribution range of their stocks shift beyond historical variabilities between 2025 and 2040 (Figure \@ref(fig:toeFigure3)).
-In some cases, like Indonesia (ranked $8^{th}$ in terms of revenue from transboundary stocks), Thailand (ranked $7^{th}$ in terms of revenue from transboundary stocks) and the United States (ranked $4^{th}$ in terms of fishing revenue from transboundary stocks), the range shift of transboundary stocks will emerge by around 2030.
-The 47 countries that make up the bulk of the global fisheries revenues also have more transboundary stocks with range shifts emerging from historical variability, such as Spain (n = 122), France (n = 85), China and Portugal (both n = 73) and Senegal (n = 66), which are the top 5 leading countries in terms of number of transboundary stocks emerging from historical variability.
+Our analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see their distribution range shifts beyond historical variabilities between 2025 and 2040 (Figure \@ref(fig:toeFigure3)).
+In some cases, like Indonesia (ranked $8^{th}$ in terms of revenue from transboundary stocks), Thailand (ranked $7^{th}$ in terms of revenue from transboundary stocks) and the United States (ranked $4^{th}$ in terms of fishing revenue from transboundary stocks), stocks’ range shifts swill emerge by around 2030.
+The 47 countries that make up the bulk of the global fisheries revenues also have more stocks with range shifts emerging from historical variabilities, being Spain (n = 122), France (n = 85), China and Portugal (both n = 73) and Senegal (n = 66) the top 5 leading countries in number of emerging stocks.
 
-On average, stocks that see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of their EEZ’s fisheries revenue that targets transboundary stocks between 2004 and 2010.
-However, large variation exists in the contribution of stocks with emerging transboundary index to the current revenue participation across EEZs.
-In some cases, e.g., El Salvador and New Zealand, stock’s with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries in their respective EEZs between 2004 and 2014.
-On the contrary, stocks with emerging transboundary index from the Marshall Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A).
-In addition, some EEZs are projected to have a few emerging stocks that represent a large proportion of revenue from transboundary stocks e.g., in Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively).
-Overall, Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks to fishing revenue is less than 20%.
-This indicates that that these countries despite having a large proportion of transboundary stocks emerging, the economic contribution of these are relatively small.
+On average, stocks that will see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010.
+However, large variation exists in the proportion that stocks with emerging transboundary index have in the current revenue participation across EEZs.
+In some cases e.g., El Salvador and New Zealand, stock’s with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries  in their respective EEZs between 2004 and 2014.
+On the contrary, stocks with emerging transboundary index from the Marshal Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A).
+In addition, some EEZs are projectede to have a few emerging stocks that represent large proportion of revenue from transboundary stocks e.g., in Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively).
+Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks on fishing revenue is less than 20%.
 
@@
 (67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods).
-I tested the sensitivity to a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence) and the global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
+We tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence).
+The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
 threshold (Figure \@ref(fig:toeFigureS2)).
-In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric,
+In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric, that is, 40% (n = 1626) fewer emerging stocks than using a 1.s.d threshold.
 
-I compared stocks time of emergence within EEZs by the species’ habitat association (Figure \@ref(fig:toeFigure4)).
+I compared species time of emergence within EEZs by species’ habitat association (Figure \@ref(fig:toeFigure4)).
 Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods).
-I found significant differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
-Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table S2).
-On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (Figure \@ref(fig:toeFigure4)).
+We found significant differences in the time of emergence of stocks range shifts, indicated by the transbounudary index, by their specie's habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
+Specifically, time of emergence of range shifts of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
+On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (toeFigure \@ref(fig:toeFigure4)).
 Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:toeFigure4)).
@@
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers with values exceeding 1.5 times the interquartile range."}
+Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range."}
 
@@
 
-<!-- I hypothesized that species environmental preferences (e.g.
+<!-- We hypothesized that species environmental preferences (e.g.
 presence in certain biomes and habitats), climate change effects on water temperature (e.g.
@@
     booktabs = T,
-    captio= "Multiple regression results for time of emergence of transboundary stocks.
+    captio= "Multiple regression results for time of emergence of transboundary species.
 Only showing significant variables.
@@
 
-I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods – \@ref(eq:five)).
+I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods - equation \@ref(eq:five)).
 By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A).
-The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)).
+The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4); \@ref(fig:toeFigureS5)).
 However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high.
-For example, Guatemala shares *Panulirus gracilis* stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century.
-Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4))).
-My results suggest that 87% (n = 238) of the world EEZs experiences changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
-These EEZs see one third (33% $\pm$ 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A).
-Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin see changes beyond the threat point for all of the transboundary stocks analyzed, while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) are the only countries with less than 10% of stocks changing beyond the threat point.
+For example, Guatemala shares Panulirus gracilis stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century.
+Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)).
+Our results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
+These EEZs will see one third (33% $\pm$ 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A).
+Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin will see changes in all of the transboundary stocks analyzed while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) will be the only countries with less than 10% of stocks changing beyond the threat point.
 
@@
 Changes of stock share ratio are largely related to regional changes in biogeography and the geometry of the EEZs (Figure \@ref(fig:toeFigure5)).
-In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the stock share ratio is expected to follow the projected poleward shift of transboundary species under climate change.
-However, central America Pacific and West African coasts follow an equatorial direction as range shifts of transboundary stocks follow local oceanographic gradients, with relatively lower temperature waters in lower latitude [@Clarke:2020ep].
-Moreover, regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no established pattern, as in the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).
+In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa stock share ratio is expected to follow the projected poleward shift of transboundary species under climate change.
+However, central America Pacific and West African coasts are expected to follow an equatorial direction as range shifts of the transboundary species are following local oceanographic gradients with relatively lower temperature waters in lower latitude [@Clarke:2020ep].
+Moreover, some regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no stablished pattern, this is the case of the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).
 
@@
 
-In this chapter, I investigate the timing and intensity of changes in the sharing of transboundary stocks between countries.
-My findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
-Such early emergence of range shifts indicate the urgency of the challenges that climate change is posing on international fisheries management.
-Moreover, I identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management.
-Finally, I show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to levels below of that needed to encourage cooperative resource management between countries.
-Adaptation measures are therefore needed to reduce the impacts of species range shifts on current international transboundary fisheries treaties.
+In this paper, we investigate the timing and intensity associated with changes in the share distribution of transboundary stocks between countries.
+Our findings highlight an early distributional shift from historical variations for most transboundary stocks in the world.
+Such early shifts point to the urgency of addressing the challenges that climate change presents for international fisheries management.
+Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries that would benefit from immediate policy responses to support sustainability into the future.
+Finally, we show that for many countries changes in the distribution of transboundary stocks are likely to undermine joint resource management between countries [@Bailey:2010ga; @Munro:1979dg; @PalaciosAbrantes:2020vt; @Bailey:2016eh].
+Thus, our findings emphasize recent calls for the urgent adoption of adaptation measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @CisnerosMata:2019ks; @Oremus:2020id; @Free:2019jq].
 
@@
 
-Our projections for the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century concur with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk].
-For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US).
+Our projections in the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century corroborate with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk].
+For example, in the early 2000s, Humboldt squid (Dosidicus gigas) substantially expanded its geographic range poleward, reaching the coast of Washington state (US).
 A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs].
-In the northeast Atlantic, the Atlantic mackerel (*Scomber scombrus*) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
-However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008, without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij].
-Such changes resulted in disputes between Iceland and the Faroes Islands, as well as the NEAFC member states [@Spijkers:2017ij].
-As climate change impacts on transboundary stocks continue to emerge from historical variability in the coming years, mal-adapted fisheries management and international disputes like that of Atlantic mackerel are likely to increase in frequency [@Pinsky:2018cb].
+In the northeast Atlantic, the Atlantic mackerel (Scomber scombrus) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
+However, a range expansion of Atlantic mackerel (Scomber scombrus) into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008 without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij].
+Such changes in quota resulted in disputes between Iceland and Faroes Islands, and the NEAFC member states [@Spijkers:2017ij].
+As transboundary stocks continue to emerge in the incoming years, mal-adapted fisheries management and international disputes like that of Atlantic mackerel are likely to increase in frequency [@Pinsky:2018cb]
 
-The projected range shifts of large number of transboundary stocks emerging from historical variability since the early 2000s can be partially attributed to the parallel emergence of several environmental variables (e.g., temperature and oxygen) that influence their distribution [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
-Marine organisms have a thermal tolerance that limits their habitable temperature range [@IPCC:2019tn], which then explains the sensitivity of the animal to warming and is a main driver of any distributional change [@IPCC:2019tn].
-Oxygen plays a similar role to that of temperature [@Clarke:2020ep], especially for tuna species that require large amounts of dissolved oxygen for their metabolism [@IPCC:2014bz].
-It has been estimated that since the early 2000s, increased sea surface temperature (SST) and decrease in subsurface oxygen have emerged from historical variability for 79% and 23% of the global ocean, respectively, and that by 2080, SST is projected to emerge from historical variability in 90% of the global ocean while the area affected by lower $O_2$ will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq].
+The projected range shifts of large number of transboundary stocks emerging since the early 2000s can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
+Marine organisms have a thermal tolerance that limits the temperature range they could live in [@IPCC:2019tn].
+This thermal tolerance explains the sensitivity of the animal to warming and is a main driver of species’ distributional change [@IPCC:2019tn].
+Oxygen plays a similar role to that of temperature [@Clarke:2020ep], especially on tuna species that require large amounts of dissolve oxygen for their metabolism [@IPCC:2014bz].
+It has been estimated that since the early 2000s, increase in sea surface temperature (SST) and decrease in oxygen content have emerged in 79% and 23% of the global ocean, respectively, and by 2080, SST is projected to emerge in 90% of the global ocean while O2 will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq].
 Moreover, before 2100, surface acidification, nitrate content and net primary production, also important variables for species distributions, will have emerged from the historical variability of 1861-1900 in 30% and 60% of the ocean under a low (RCP 2.6) and strong (RCP 8.5) climate change scenario, respectively [@IPCC:2019tn].
@@
 Regional differences in time of emergence are attributed to the variations in environmental conditions, species’ climate vulnerability, and the characteristics of political boundaries.
-Specifically, my findings are consistent with results from previous studies that show tropical species are already shifting their distributions in the last few decades [@Poloczanska:2016kk].
-The tropics are among the most species-rich biomes [@Reygondeau:2019uh] and have potentially experienced some of the highest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
-In the tropics, marine species also live close to their thermal tolerance, making them highly vulnerable to warming [@IPCC:2014bz].
-Consequently, species have to move towards deeper or sub-tropical waters as a result of climate change [@Poloczanska:2016kk], although in some regions species follow local temperature gradients towards the equator or shallow waters [@Clarke:2020ep].
-Moreover, tropical species are expanding their range into sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis shows a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
+Specifically, our findings are consistent with results from previous studies that show tropical species are already shifting their distributions in the last few decades [@Poloczanska:2016kk].
+The tropics are among the most species-rich biome [@Reygondeau:2019uh] and have potentially experienced some of the highest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
+In the tropics, marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz].
+Consequently, as a result of climate change, species have to move towards deeper or sub-tropical waters [@Poloczanska:2016kk] although recent studies suggest that in some regions species are shifting towards the equator and to shallow waters to follow local temperature gradients [@Clarke:2020ep.
+Moreover, tropical species are expanding their range in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis shows a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
 In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs that border with multiple countries (e.g., The Caribbean).
@@
 
-Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)).
+Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)).
 For example, Polynesia and Micronesia are the only regions where the average time of emergence of the distribution of where all transboundary stocks across all EEZs emerged before 2022 (Figure \@ref(fig:toeFigure5)).
 Moreover, countries in these hotspot regions are highly dependent on pelagic-oceanic (e.g., tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
-On the other hand, “coldspots” of climate risk exist for transboundary fisheries, like the Northeast Atlantic and Northwest Pacific, where stocks are expected to emerge later in the $21^{st}$ century.
+On the other hand, “coldspots” of climate risk for transboundary fisheries like the Northeast Atlantic and Northwest Pacific where stocks are expected to emerge later in the 21st century.
 Assuming that these countries already have adequate transboundary management to deal with historical and current variations in fish stocks’ distribution e.g., the Pacific Salmon Treaties [@Miller:2004fpa], management agencies are likely to have more time to prepare for emerging stock shifts.
-
-In cases where climate driven changes in stocks are emerging, or have already emerged, from historical variability adaptation policies are urgently needed to address the potential impact on important fisheries [@Burden:2019ij].
-Currently, fisheries targeting shared tuna stocks in the South Pacific Islands region operate under the Vessel Day Scheme, which because it takes into account shifts in stocks and catches makes it a good example of a management system that is somewhat anticipatory [@Aqorau:2018bh].
-However, even strategies that take into consideration the shift of the stock within the party’s area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
-As the climate trend in stocks distibution emerges from their historical variabillity since 2006, regional management bodies must revise their management actions as most are not currently equipped with strategies to cope with shifting species distributions [@Oremus:2020id; @Sumby:2021gw].
+In cases where stocks are emerging or have already emerged, adaptation policies are urgently needed to address an issue that could be impacting important fisheries [@Burden:2019ij].
+Currently, fisheries targeting tunas shared stocks in the South Pacific Islands region operate under the Vessel Day Scheme - which takes into account shifts in stocks and catches making it a good example where a management system has been set up that is somewhat anticipatory [@Aqorau:2018bh].
+However, even strategies that take into consideration the shift of the species within the Party’s area will likely require to deal with the newcomer issue that NEAFC is currently facing over Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+As species have been emerging from their historical distribution since 2006, regional management bodies require to revise their management actions as today most of them are not equipped with strategies to cope with shifting species distributions [@Oremus:2020id; @Sumby:2021gw].
 This is of specific concern for those management bodies located in the emerging “hotspots” here identified.
 
-Some EEZs lack stocks with emerging transboundary index because of a combination of their shape, climate projections, and a reduced number of transboundary stocks.
-For example, there are 59 commercial fish stocks fished in the Arctic with the vast majority on the Barents, Bering and Norwegian seas, with limited knowledge about the species biology [@Christiansen:2014gg].
-However, the Arctic has been warming faster than many other regions, resulting in a reduction in sea-ice [@Douglas:2010Mb] leading to the  potential for occupation by species not previously present [@IPCC:2019tn].
+The presentce of EEZs lacking stocks with emerging transboundary index is a combination of EEZ shape, climate projections, and a reduced number of transbounudary stocks.
+For example, there are 59 commercial fish stocks fished the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas with limited knowledge about the species biology [@Christiansen:2014gg].
+However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@IPCC:2019tn].
 As a result, the biomass of Arctic marine animals [@BryndumBuchholz:2019kh] and catch and revenue of associated fisheries [@Tai:2019cq; @Lam:2014fq] are expected to increase in the incoming years, although substantial uncertainty exists on the magnitude of change [@BryndumBuchholz:2019kh].
-Moreover, projections suggest a larger increment on transboundary stocks in the coming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further north [@Morley:2018fn; @Frainer:2017ji].
-In the specific case of Brazil, in addition to having few transboundary stocks (only four shared with Uruguay in this study), the large difference in the EEZs lengths could also be influencing the results.
+Moreover, projections suggest a larger increment on transboundary stocks in the incoming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further North [@Morley:2018fn; @Frainer:2017ji].
+In the Case of Brazil, in addition to having few transbounudary stocks (only 4 shared with Uruguay in this study), the large diffrence in the EEZs lenghts could also be influencing the results.
 
@@
 Poleward shifts in stock share ratio are expected in regions of the world with relatively long EEZs and few boundaries, like the northeast Pacific and the southeast Atlantic.
-However, despite observations and modelling work indicating that climate change is generally shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], some regions like the eastern tropical Pacific and easterntropical Atlantic Oceans will experience a shift in distribution towards the equator.
+However, despite observations and modelling work indicating that, generally, climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], some regions like the eastern tropical Pacific and eastern tropical Atlantic Oceans will experience a shift in distribution towards the equator.
 Specifically, in the eastern tropical Pacific, the direction of shift is expected to be from Mexico to Ecuador with EEZs losing stock share ratio to their southern neighbors.
-Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems [@Clarke:2020ep] and are consistent with regional modeling studies suggesting that by the mid $21{st}$ century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission scenario [@Clarke:2020ep].
-Moreover, regions characterized by multiple political boundaries, like the Caribbean and the northeast Atlantic, will see a complex network of changes in stock share ratio with no clear pattern (Figure \@ref(fig:toeFigure5)B).
-These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and dispersal of larvae [@Ramesh:2019va] supporting the implementation of joint management plans [@Burden:2019ij] and the call for new implemented shared policies [@Levin:2018in].
+Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems [@Clarke:2020ep].
+They are consistent with regional modeling studies suggesting that by mid 21st century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
+Moreover, regions characterized by multiple political boundaries, like the Caribbean and the northeast Atlantic, will see a complex network of changes in SSR with no clear pattern (\@ref(fig:toeFigure5)B).
+These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] supporting the implementation of joint management plans [@Burden:2019ij] and the call for new implemented shared policies [@Levin:2018in].
 However, game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.
 
-Changes in stock share ratio beyond the threat point of countries sharing a common resource could instigate international conflict and hinder bi-lateral agreements [@Miller:2013iv; @Sumaila:2020tp].
-Lessons learned from historic distributional shifts of shared stocks suggest that treaties that are not prepared to respond to such shifts will be less resilient to changes in transboundary fish stocks share ratio due to climate change [@Miller:2013iv; @Sumaila:2020tp].
+Changes in stock share ratio beyond the threat point of countries sharing a common resource could instigate international conflict and hinder bi-lateral agreements (@Miller:2013iv; @Sumaila:2020tp).
+Lessons learned from historic distributional shifts of shared stocks suggest that treaties that are not prepare to respond to such shifts will be less resilient to changes in transboundary fish stocks share ratio due to climate change [@Miller:2013iv; @Sumaila:2020tp].
 For example, disputes over Pacific salmon between Canada and the USA arose in the 1990s when a climate-related shift in stock abundance favoured Alaska’s salmon fisheries over Canada’s [@Miller:2013iv; @Munro:2002uf; @Song:2017va].
 The conflict lasted over ten years until Canada and the USA agreed on a mutual conservation fund where the US would contribute the larger proportion to support scientific research, habitat restoration and enhancement of wild stock production in transboundary rivers [@Munro:2002uf].
-Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Pinsky:2018cb].
+Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Pinsky:2018cd].
 Side payments are mechanisms that provide a solution to the inequality produced by shifts in the stock’s distribution [@Tunca:2019if; @Miller:2004fpa].
-These can be monetary (e.g., the stock-“winning” state compensates the stock-“losing” state for the proportional shift) or other methods like the Pacific salmon case (e.g., a conservation fund paid mainly by one party) [@Miller:2004fpa; @Miller:2013iv].
-Strengthening current international cooperation will be required as stock distribution will move to neighboring jurisdictions, requiring combined efforts to generate information and set proper management rules [@Miller:2013iv].
-Finally, flexible management rules might be more appropriate to capturing t distributional shifts.
-For example, a quota allocation based on the stock’s current distribution like that for Pacific halibut (*Hippoglossus stenolepis*) managed by Canada and the US [@Sumaila:2020tp; @PalaciosAbrantes:2020vt] would be more agile than one based on a fixed-historical proportion like the case of the EU [@Baudron:2020dj].
+These can be monetary (e.g., the stock-“winning” State compensates the stock-“loosing” state for the proportional shift) or other methods like the Pacific salmon case (e.g., a conservation fund payed mainly by one Party) [@Miller:2004fpa; @Miller:2013iv].
+Strengthening current international cooperation will be required as stock distribution will move to neighboring jurisdictions, requiring combine efforts to generate information and set proper management rules [@Miller:2013iv].
+Finally, flexible management rules might be better fitted to capture the stock’s distributional shift.
+For example, quota allocation based on the stock’s current distribution like that for Pacific halibut (*Hippoglossus stenolepis*) managed by Canada and the US [@Sumaila:2020tp; @PalaciosAbrantes:2020vt] might be better able to capture the stock’s shifts than a quota based on a fix-historical proportion like the case of the EU [@Baudron:2020dj].
 
@@
 
-There are two main sources of uncertainty in my analysis.
-First, I rely on the combination of a single Earth system model (GFDL-ESM2M), fish and fisheries model, and climate change scenario (RCP 8.5).
+There are two main sources of uncertainty in our analysis.
+First, our analysis relies on the combination of a single Earth system model (GFDL-ESM2M), fish and fisheries model, and climate change scenario (RCP 8.5).
 The use of multiple ESMs can capture the structural uncertainty of the models [@Frolicher:2016dg; @Lotze:2019ce], however, exercises looking at changes in fish biomass with multiple models show an overall agreement between models in terms of direction of change but variable in magnitude.
 Further research that includes multiple ESMs and species distribution models could elucidate a broader range structural uncertainty of the climate, fish and fisheries models [@Lotze:2019ce].
-Secondly, due to limited biological and spatially-specific data on stocks sub-population structure of transboundary species, my analysis uses political boundaries (EEZs) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
-On the other hand, in many EEZs, fisheries are often managed at the species level (Chapter 2) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for my analysis [@Popova:2019fs; @Dunn:2019ca].
-However, reproducing my analysis locally, where spatially explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
+Secondly, due to limited biological and spatially-specific data on stocks sub-population structure of transboundary species, our analysis uses political boundaries (EEZ) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
+On the other hand, in many EEZs, fisheries are often managed at the species level [@PalaciosAbrantes:2020tv] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
+However, reproducing our analysis regionally, where spatially explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
 Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
@@
 Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@@Pinsky:2018cb].
-Here, I have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
-I have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
-My results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management.
-While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, my results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg; @PalaciosAbrantes:2020tv].
-To address the issues raised in this chapter, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
+Here, we have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
+We have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
+Our results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management.
+While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg; @PalaciosAbrantes:2020tv].
+To address the issues raised in this paper, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
 
@@
 
+```{r TableS0, eval = T, echo = F}
+
+tibble(
+  "Habitat preference" = unique(Dem_pel$DemersPelag)[1:7],
+  Description = c(
+    "The shallow pelagic zone over the continental shelf; nearshore ocean ecosystems; i.e., those associated with the coasts because the waters are overlying continental shelves and/or the waters are < 200 m deep in areas of coastal submarine slopes",
+    "A largely consolidated wave resistant feature; the upper surface is within 0-20 m of the ocean surface; in the tropics, the upper portion is typically composed mainly of living and non-living remains of coral and coralline algae; rock lying at or near the surface and which may pose a danger to navigation",
+    "Living and feeding on the bottom below 200 m",
+    "Living and feeding near the bottom as well as in midwaters or near the surface.
+Feeding on benthic as well as free swimming organisms.
+Many freshwater fish are opportunistic feeders that forage on the bottom as well as in midwater and near the surface, also pertaining to forms which hover or swim just over the floor of the sea, e.g.
+Halosauridae, Macrouridae, Moridae, Brotulidae; the depth zone about 100 metres off the bottom at all depths below the edge of the continental shelf",
+    "Sinking to or lying on the bottom; living on or near the bottom and feeding on benthic organisms",
+    "Region of the oceanic zone between 1,000 m to 4,000 m; between the mesopelagic layer above and the abyssopelagic layer below.
+Living or feeding in open waters at depths between 1,000 and 4,000 m.
+In FishBase this term is used to include the depth range from 200 m to the bottom and thus the zones mesopelagic, bathypelagic and abyssopelagic",
+
+    "Living and feeding in the open sea; associated with the surface or middle depths of a body of water; free swimming in the seas, oceans or open waters; not in association with the bottom.
+Many pelagic fish feed on plankton.
+In FishBase, referring to surface or mid water from 0 to 200 m depth."
+  )
+) %>% 
+ kable()
+
+```
+
 ```{r TableS1, eval = F, echo = F}
@@
 
-<!-- The time and number of emerging species decreases as I move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
+<!-- The time and number of emerging species decreases as we move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
 -->

### 2021-02-16 Scripts/Initial_Analysis.Rmd
Fixed figure 1

@@
 
-# Manually include rest of EEZs
-
+# Manually include rest of data untill 2100
 cont_eez <- plot_data %>% 
   filter(level == "eez") %>% 
-  slice(1:36) %>%
+  arrange(taxon_emer) %>% 
+  slice(1:24) %>%
   select(-1) %>% 
+  arrange(metric) %>% 
   mutate(
-    value = ifelse(metric == "per",100,232),
-    taxon_emer = c(seq(2082,2099,1),seq(2082,2099,1))
+    value = ifelse(metric == "per",100,228),
+    taxon_emer = c(seq(2088,2099,1),seq(2088,2099,1))
   )
@@
   )
-
-# Percentage increase of EEZ after 2020
-lm_eez <- plot_data %>% 
-  filter(taxon_emer <= 2020,
-         tresh == "tresh_one",
-         metric == "per")
-
-summary(lm(lm_eez$taxon_emer~lm_eez$value))
 

### 2021-03-03 Scripts/Initial_Analysis.Rmd
Included dashed lines and other modifications to figure 1

@@
 path_world <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile")
-path_world <- "/Volumes/Enterprise"
+path_world <- "~/Downloads/SAU_Shapefile"
 
@@
 # Transform sf to df
-Eez_centroid_eck <- as.data.frame(Eez_centroid_sf)
-
-Eez_centroid_eck <- do.call(rbind, st_geometry(Eez_centroid_sf)) %>% 
-  as_tibble() %>% 
-  setNames(c("lon","lat")) %>% 
-  bind_cols(Eez_centroid) %>% 
-  select(eez_name,longitude = lon,latitude = lat)
+# Eez_centroid_eck <- as.data.frame(Eez_centroid_sf)
+
+# Eez_centroid_eck <- do.call(rbind, st_geometry(Eez_centroid_sf)) %>% 
+#   as_tibble() %>% 
+#   setNames(c("lon","lat")) %>% 
+#   bind_cols(Eez_centroid) %>% 
+  # select(eez_name,longitude = lon,latitude = lat)
 
@@
 
+dashed_lines <- tibble(
+  x = c(2030,2050,2080),
+  label = c("SDGs target date","Early range shift  ","Late range shift   ")
+)
+
 # How many between 2006 and 2020?
@@
   )  +
-  scale_y_continuous("Cumulative number of EEZs (n)",
+  # Add dashed lines and labels
+  geom_vline(
+    data = dashed_lines,
+    aes(
+      xintercept = x
+    ),
+     linetype="dashed",
+    color = "grey"
+  ) +
+  geom_text(data = dashed_lines,
+            aes(
+              x = x-4,
+              y = 40,
+              label = label
+            ),
+            size = 4,
+            color = "white",
+            angle = 90
+  ) +
+  scale_y_continuous("Cumulative number of EEZs",
                      expand = c(0,0)
                      ) +
-  scale_x_continuous("",
+  scale_x_continuous("Year of range shift",
                      breaks = seq(2000,2100,25),
@@
   )  +
-  scale_y_continuous("Cumulative number of stocks (n)",
+  # Add dashed lines and labels
+  geom_vline(
+    data = dashed_lines,
+    aes(
+      xintercept = x
+    ),
+     linetype="dashed",
+    color = "grey"
+  ) +
+  geom_text(data = dashed_lines,
+            aes(
+              x = x-4,
+              y = 750,
+              label = label
+            ),
+            size = 4,
+            color = "white",
+            angle = 90
+  ) +
+  scale_y_continuous("Cumulative number of stocks",
                      expand = c(0,0)
                      ) +
-  scale_x_continuous("Time of emergence of transboundary index",
+  scale_x_continuous("",
                      breaks = seq(2000,2100,25),
@@
   ) + 
+  # Add dashed lines and labels
+  geom_vline(
+    data = dashed_lines,
+    aes(
+      xintercept = x
+    ),
+     linetype="dashed",
+    color = "grey"
+  ) +
+  geom_text(data = dashed_lines,
+            aes(
+              x = x-4,
+              y = 18,
+              label = label
+            ),
+            size = 4,
+            color = "white",
+            angle = 90
+  ) +
   scale_y_continuous("Cumulative proportion of the total (%)",
@@
   ) +
-  guides(fill=guide_legend(title= "Level", reverse=FALSE)) +
+  guides(fill=guide_legend(title= "", reverse=FALSE)) +
   scale_fill_viridis(discrete = T,
@@
     # Latin America and the Car.
-    draw_plot(count_eez_plot, x = 0.01, y = 0, width = 0.28, height = 1) +
+    draw_plot(count_stock_plot, x = 0.01, y = 0, width = 0.28, height = 0.90) +
     # Europe
-    draw_plot(count_stock_plot, x = 0.28, y = 0, width = 0.28, height = 1) +
+    draw_plot(count_eez_plot, x = 0.28, y = 0, width = 0.28, height = 0.90) +
     # Percentage plot
-    draw_plot(percentage_plot, x = 0.56, y = 0 , width = 0.28, height = 1) +
+    draw_plot(percentage_plot, x = 0.56, y = 0 , width = 0.28, height = 0.90) +
     # legend
-    draw_plot(legend, x = 0.84, y = 0 , width = 0.28, height = 1)
+    draw_plot(legend, x = 0.84, y = 0 , width = 0.28, height = 1) +
+    # Labels
+    draw_plot_label(label = c("A", "B","C"), 
+                    size = 20,
+                    x = c(0.01, 0.285,0.56), 
+                    y = c(1,1,1)
+    )
 

### 2021-03-03 Scripts/Initial_Analysis.Rmd
removed Index alternative analysis

@@
   mutate(participation = (sum_value/tot_value)*100+2000)
-
-### ----------------- ###
-### Index alternative ###
-### ----------------- ###
-
-# Index_data <- Eez_d_hist %>% 
-#   left_join(Value_map_data) %>% 
-#   filter(
-#     tresh == "Tresh_one",
-#     metric == "mean_yr"
-#   ) %>% 
-#   group_by(sau_fishing_entity,fishing_entity_id,variable) %>% 
-#   summarise(
-#     toe = mean(value,na.rm=T),
-#     # sum_value = sum(sum_value),
-#     mean_part = mean(participation,na.rm=T)
-#   ) %>% 
-#   filter(!is.na(variable)) %>% 
-#   # spread(variable,sum_value) %>% 
-#   spread(variable,mean_part) %>% 
-#   mutate(
-#     imp = catch+value,
-#     index = imp/toe
-#   )
-
-##### Normalize index
-# https://www.datanovia.com/en/blog/how-to-normalize-and-standardize-data-in-r-for-great-heatmap-visualization/
-
-# Get min and maximums
-# min_index <- min(Index_data$index)
-# max_index <- max(Index_data$index)
-
-# Index_data <- Index_data %>% 
-#   mutate(
-    # index_norm =(index-min_index)/(max_index-min_index),
-  #   index_2000 = 2000 + (index-min_index)*(2100-2000)/(max_index-min_index)
-  # )
-
-### Explore index ####
-
-# ggplot(data = Index_data) +
-#   geom_point(
-#     aes(
-#       x = toe,
-#       y = imp,
-#       color = index_2000
-#     ),
-#     size = 4
-#   ) +
-#    scale_color_viridis(option="magma", discrete = F)
-
-#### ------------  ####
-
-# Value_map_data <- Index_data %>%
-  # select(sau_fishing_entity,fishing_entity_id,
-         # tresh,
-         # variable,
-         # participation = index_2000)
-
-### ----------------- ###
 

### 2021-03-03 Scripts/Initial_Analysis.Rmd
Changes i figure 2 and bubble figure

@@
              y = -69,
-             label = "Participation (%)",
+             label = "% Of fishing revenue",
              color="black",
@@
               max = max(emerging_yr,na.rm = T),
+              med = median(emerging_yr,na.rm = T),
               n_stocks = n()
@@
       aes(
-        # x = un_sub_region,
         x = reorder(un_sub_region,order),
-        y = emerging_yr#,
-        # fill = mean,
-        # colour = mean
+        y = emerging_yr
       ),
@@
                nrow = 1) +
-    # scale_fill_viridis("",
-    #                    limits = c(2000,2100),
-    #                    breaks = seq(2000,2100,25),
-    #                    labels = c("","","","","")#,
-    #                    # direction = -1
-    # ) +
-    # scale_colour_viridis("",
-    #                    limits = c(2000,2100),
-    #                    breaks = seq(2000,2100,25),
-    #                    labels = c("","","","","")#,
-    #                    # direction = -1
-    # ) +
+    # Median points
+    geom_point(
+      aes(
+        x = reorder(un_sub_region,order),
+        y = med,
+        colour = med
+      ),
+      size = 5
+    ) + 
+    # h line
+    geom_hline(yintercept = 2030, color = "grey", size = 0.5, linetype="dashed") +
+    scale_colour_viridis("",
+                       limits = c(2000,2100),
+                       breaks = seq(2000,2100,25),
+                       labels = c("","","","",""),
+                       direction = -1
+    ) +
     xlab("") +
@@
 
+dashed_lines <- tibble(
+  x = c(2030,2050,2080),
+  label = c("SDGs target date","Early range shift  ","Late range shift   ")
+)
 # Needs previous chunk
@@
 Buble_data %>% 
-  filter(log_value > Percentile_value$x) %>% 
+  filter(log_value > Percentile_value$x & mean_yr < 2036) %>% 
   View()
@@
 Labels_data <- tibble(
-  x = c(2020,2040,2007),
-  y = c(-5.8,-5.8,Percentile_value$x+.1),
-  label = c("Early shift", "Late shift","Fisheries > 75th percentile")
+  x = c(2030,2045,2055,2007),
+  y = c(-5.8,-5.8,-5.8,Percentile_value$x+.1),
+  label = c("SDGs target", "Early shift","Late shift","Fisheries > 75th percentile")
 )
@@
   ) +
-  # Divisory lines
-  geom_vline(xintercept = 2030, color = "grey90", size = 0.5) +
-  geom_hline(data = Percentile_value,
-             aes(yintercept = x),
-             color = "grey90"
-  )  +
-  # right side labels
-  geom_label_repel(data = subset(Buble_data, mean_yr > 2045 & sum_value_bi >= 0.90),
+  # Right side labels
+  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034),
                   aes(
@@
                   fill = "NA",
-                  nudge_x = 2075 - subset(Buble_data, mean_yr > 2045 & sum_value_bi >= 0.90)$mean_yr,
+                  nudge_x = 2066 - subset(Buble_data,log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034)$mean_yr, # set distance from point
+                  nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
                   segment.size  = 0.1,
@@
   # LEFT side labels
-  geom_label_repel(data = subset(Buble_data, mean_yr < 2045 & sum_value_bi >= 0.90),
+  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034),
                   aes(
@@
                   fill = "NA",
-                  nudge_x = 2025 - subset(Buble_data, mean_yr < 2045 & sum_value_bi >= 0.90)$mean_yr,
+                  nudge_x = 2025 - subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$mean_yr, # set horizontal position
+                  nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
                   segment.size  = 0.1,
@@
                   hjust         = 1
+  ) +
+  # Divisory lines and labels
+  geom_vline(xintercept = c(2030,2050), color = "grey90", size = 0.5, linetype = "dashed") +
+  geom_hline(data = Percentile_value,
+             aes(yintercept = x),
+             color = "grey90"
   )  +

### 2021-03-03 Scripts/Initial_Analysis.Rmd
Updated figures and estimated revenues

@@
 
-## Top fishing nations
+## Top fishing nations close to SDGs
 Buble_data %>% 
   filter(log_value > Percentile_value$x & mean_yr < 2036) %>% 
-  View()
+  # View()
+  group_by() %>% 
+  summarise(sum(sum_value_bi),
+            sum(n_taxa))
+
+## All fishing nations close to SDGs
+Buble_data %>% 
+  filter(mean_yr < 2036) %>% 
+  # nrow() # 76  coutries
+  # View()
+  group_by() %>% 
+  summarise(sum(sum_value_bi),
+            sum(n_taxa))
+
+76/length(unique(Buble_data$sau_fishing_entity))
 

### 2021-03-03 Scripts/Initial_Analysis.Rmd
Included a catch proportion in proportion analysis

@@
 
-  #Determines the amount of grids present in each country
+  #Determines the amount of catch present in each country
   Spp_Grid <- Trans_Spp %>% 
@@
              ensemble) %>% 
-    summarise(n_spp_eez = length(unique(index))) %>% 
+    summarise(n_spp_eez = length(unique(index)),
+              total_mcp = sum(value)
+              ) %>% 
     left_join(Neighbours_List,
@@
       n_spp_eez,
+      total_mcp,
       year,
@@
       n_spp_eez,
+      total_mcp,
       Name=eez_neighbour,
@@
     rowwise() %>%
-    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
+    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T),
+           MCP_Total = sum(total_mcp.x,total_mcp.y,na.rm=T)) %>% # Total gridcelles per Neighbours
     distinct() %>% # Removes false duplicates from `full_join()`
@@
            n_spp_Country = n_spp_eez.x,
-           n_spp_Neighbour = n_spp_eez.y) %>% 
-    mutate(area_index = n_spp_Country/Spp_Total) %>% 
+           n_spp_Neighbour = n_spp_eez.y,
+           mcp_Country = total_mcp.x,
+           mcp_Neighbour = total_mcp.y) %>% 
+    mutate(area_index = n_spp_Country/Spp_Total,
+           catch_proportion = mcp_Country/MCP_Total) %>% 
     # Determine the three time frames for comparrison
@@
     group_by(taxon_key,ensemble,eez_name,eez_neighbour,time_step) %>% 
-    summarise_at(vars("area_index"),
-                 c(temporal_mean=mean,
-                   temporal_sd=sd)
+    summarise_at(vars("area_index","catch_proportion"),
+                 c(temp_mean=mean,
+                   tempo_sd=sd)
     ) %>% 
@@
     group_by(taxon_key,eez_name,eez_neighbour,time_step) %>%
-    summarise(
-      ensemble_mean = mean(temporal_mean, na.rm = T),
-      ensemble_sd = sd(temporal_mean, na.rm = T),
-      ensemble_n = n()
+    summarise_at(vars("area_index_temp_mean","catch_proportion_temp_mean"),
+                 c(ensemble_mean = mean,
+                   ensemble_sd = sd
+                   )
     )
@@
 
-# EstPropChange(taxon_key = 600004,
-#               year = year_df$year,
-#               hist_y = 2000,
-#               data_type = "Catch",
-#               ensemble_list = 102,
+EstPropChange(taxon_key = 600004,
+              year = year_df$year,
+              hist_y = 2000,
+              data_type = "Catch",
+              ensemble_list = c(102,103,104),
               # path = path,
-#               neighbours = neighbours)
-#
+              neighbours = neighbours)
 

### 2021-03-08 Scripts/Initial_Analysis.Rmd
added lines for supp 3 figure and changed code for updated packages ()ughhhhu

--- 2021-03-03 Scripts/Initial_Analysis.Rmd
@@
   select(-nbiom) %>% 
-  distinct(eezid, .keep_all = T) # Croatia is 1/2. We arbritary allocated the first one
+  distinct(eezid, .keep_all = T) # Croatia is 1/2. We arbitrary allocated the first one
 
@@
 
-Catch_trend <- my_path("D","Species","Catch_trend.csv", read = T) %>% 
-  clean_names() %>% 
-  filter(taxon_key %in% Taxon_list_emer) %>% 
-  rename(sau_fishing_entity = territory) %>% 
-  left_join(Fishing_names,
-            by = "sau_fishing_entity") %>% 
-  filter(status != "No Trend") %>% 
-  mutate(category = ifelse(status %in% c("Developing","Rebuilding"),"Category A",
-                         ifelse(status == "Max Exploited","Category B",
-                                ifelse(status %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
-                         )
-         )
-  )
+# Catch_trend <- my_path("D","Species","Catch_trend.csv", read = T) %>% 
+#   clean_names() %>% 
+#   filter(taxon_key %in% Taxon_list_emer) %>% 
+#   rename(sau_fishing_entity = territory) %>% 
+#   left_join(Fishing_names,
+#             by = "sau_fishing_entity") %>% 
+#   filter(status != "No Trend") %>% 
+#   mutate(category = ifelse(status %in% c("Developing","Rebuilding"),"Category A",
+#                          ifelse(status == "Max Exploited","Category B",
+#                                 ifelse(status %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
+#                          )
+#          )
+#   )
 
@@
 
-# For dat availabillity
+# For dat availability
  # write_csv(prop_results,
           # "proportion_change_2005.csv")
-
-# Prop_resultsb <- my_path("R",name="proportion_change_2005.csv", read=T)
 
@@
 
-# warnings() for empty data species
+# warnings() for empty data species # New version of package being a pain in the arse
 
@@
 
-# Get taxon list
-Taxon_list_sens <- list.files(my_path("R",extra_path = "Toe_Sens_per/Emergence_2005_p20"))
-
-# clean names for function
-Taxon_list_sens <- gsub("_.*","",Taxon_list_sens)
-
 # SAU exploited species list
@@
 
+sen <- 95
 read_sen <- function(sen){
@@
 
-  Taxon_list_sens <- list.files(my_path("R",extra_path = "Toe_Sens_per/Emergence_2005_p20"))
-
-  Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_list_emer,"_emergence.txt",sep="")
+  Taxon_list <- gsub("_.*","",Taxon_list_sens)
+
+  Paths_sen <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_list_sens,"_emergence.txt",sep="")
 
@@
   filter(
-    tresh== "tresh_one"
-  )
+    tresh== "tresh_two"
+  )
+
+unique(Sen_data$sen)
 
@@
     # total = ifelse(level == "stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$sf_eez_name))),
-    total = ifelse(level == "stock",max(stock_d$com_sum),max(eez_d$com_sum)),
+    total = ifelse(level == "stock",max(sen_stock_d$com_sum),max(sen_eez_d$com_sum)),
     per = com_sum/total*100
@@
     my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
-    label = ifelse(total == max(stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs"),
+    label = ifelse(total == max(sen_stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs"),
   sen = paste("Percentile above",sen,"%") 
+  )
+
+# Lines data
+dashed_lines <- tibble(
+  x = c(2030,2050,2080),
+  label = c("Agenda 2030","Early range shift  ","Late range shift   ")
 )
@@
   ) +
-  scale_x_continuous("Time of emergence",
+  scale_x_continuous("Year of range shift",
                      breaks = seq(2000,2100,25),
@@
   ) +
+  # Add dashed lines and labels
+  geom_vline(
+    data = dashed_lines,
+    aes(
+      xintercept = x,
+      linetype = label,
+    ),
+    color = "grey"
+  ) + 
   my_ggtheme_p() +
@@
     legend.position = "right",
-    # strip.text.x = element_blank(),
     panel.spacing = unit(3, "lines") # space between plots
   ) +
-  guides(fill=guide_legend(title= "Level", reverse=FALSE)) +
+  guides(fill=guide_legend(title= "", reverse=FALSE),
+         linetype = guide_legend(title = "Time step")) +
   scale_fill_viridis(discrete = T,
@@
   ) +
-  ggsave("./Figures/toe_sen.png",
+  ggsave("./Figures/Supp3.png",
          width = 14,

### 2021-03-09 Scripts/Initial_Analysis.Rmd
changed to catch proportion

@@
   # Reads taxon data
-  proportion_data <- fread(paste(my_path("R",extra_path = "Proportion_2005/"),"proportion_",taxon_key,".csv", sep="")) %>% 
-    filter(!is.na(eez_neighbour))
+  proportion_data <- fread(paste(my_path("R",extra_path = "Proportion_2005c/"),"proportion_",taxon_key,".csv", sep="")) %>% 
+    filter(!is.na(eez_neighbour)) %>% 
+    select(
+      1:4,
+      ensemble_mean = catch_proportion_temp_mean_ensemble_mean,
+      ensemble_sd = catch_proportion_temp_mean_ensemble_sd)
 
@@
 
-    # Min, max and median change per eez_name and neighbour
+    # Min, max and median change per eez_name and neighbor
     eez_summary <- spp %>% 
@@
 # --------------- #
+
+# Read transboundary species database from Palacios-Abrantes et al 2020 (FishForVisa)
+Transboundary_spp <- my_path("D", extra_path = "Species/", name="Transboundary_spp.csv", read = TRUE)
 
@@
 # Get taxon list
-Taxon_list <- list.files(my_path("R",extra_path = "Proportion_2005"))
+Taxon_list <- list.files(my_path("R",extra_path = "Proportion_2005c"))
 
@@
   x = c(2030,2050,2080),
-  label = c("SDGs target date","Early range shift  ","Late range shift   ")
+  label = c("UN 2030 Agenda","Early range shift  ","Late range shift   ")
 )
@@
   ) +
-  my_ggtheme_p() +
-  theme(legend.position = "") +
+  my_ggtheme_p(leg_pos = "") +
   guides(fill=guide_legend(title= "Level", reverse=FALSE)) +
@@
   ) +
-  my_ggtheme_p() +
+  my_ggtheme_p(leg_pos = "right") +
   theme(
-    legend.position = "right",
     strip.text.x = element_blank()
@@
     # legend
-    draw_plot(legend, x = 0.84, y = 0 , width = 0.28, height = 1) +
+    draw_plot(legend, x = 0.77, y = 0 , width = 0.28, height = 1) +
     # Labels
@@
   y = c(-5.8,-5.8,-5.8,Percentile_value$x+.1),
-  label = c("SDGs target", "Early shift","Late shift","Fisheries > 75th percentile")
+  label = c("UN 2030 Agenda", "Early shift","Late shift","Fisheries > 75th percentile")
 )
@@
   # Divisory lines and labels
-  geom_vline(xintercept = c(2030,2050), color = "grey90", size = 0.5, linetype = "dashed") +
+  geom_vline(xintercept = c(2030,2050), color = "grey50", size = 0.5, linetype = "dashed") +
   geom_hline(data = Percentile_value,
              aes(yintercept = x),
-             color = "grey90"
+             color = "grey50"
   )  +
@@
   scale_size("Number of\nshifting stocks") +
+  my_ggtheme_p(legend.position = "right") +
   ggsave("./Figures/bubble_plot.png",
@@
         y = emerging_yr,
-        # fill = DemersPelag
+        fill = mean_toe
     ),
     alpha = 0.8,
-    fill = "grey",
+    # fill = "grey",
     coef = 0, # to have dotted error bars
@@
     linetype = "dashed") + 
+  geom_hline(yintercept = 2030, color = "grey", size = 0.5, linetype="dashed") +
   xlab("") +
   ylab("Year of range shift") +
-  my_ggtheme_p() +
-  theme(
-    axis.text.x = element_text(size = 10, angle = 45, face = "plain", hjust = 1), 
-    # axis.ticks.x = element_blank(),
-    legend.position = "right"
+  my_ggtheme_p(leg_pos = "right", leg_tl_s =11, leg_tx_s = 10, axx_tx_ang = 45, ax_tx_s = 10, hjust = 1) +
+  scale_fill_viridis("Average year of\nrange shift",
+                       limits = c(2000,2100),
+                       breaks = seq(2000,2100,25),
+                       # labels = c("","","","",""),
+                       direction = -1
     ) +
@@
 
-### Median numbef of stocks changing in each time step
+### Median number of stocks changing in each time step
 
@@
   sf::st_set_geometry(NULL) %>% 
-  unnest() %>%
+  unnest(cols = c(coords)) %>%
   mutate_at(vars(X,Y), round,2) %>% 
@@
 Rectangles <- tibble(
-    x1=c(-100,-3,-14), #LatAm, Med, NE
-    x2=c(-55,35,23),
-    y1=c(9,31,50), 
-    y2=c(30,44,67)
+    x1=c(-100,-14), #LatAm, Med, NE
+    x2=c(-55,23),
+    y1=c(9,50), 
+    y2=c(30,67)
   )
@@
   Lat_am_eez <- World_eez_sf %>% 
-    rename(sf_eez_name = eez_name) %>% 
+    select(sf_eez_name = eez_name,everything()) %>% 
     left_join(Matching_names) %>% 
@@
   ## ------------------- ##
-  # The Med
+  # The Med (Removed by Colette)
   ## ------------------- ##
 
-  # Land polygon
-  Med_land <- World_map %>% 
-    filter(region_wb %in% c("Europe & Central Asia","Middle East & North Africa") |
-             sau_fishing_entity %in% c("Spain","France")
-    )
-
-  # EEZ polygon
-  Med_eez <- World_eez_sf %>% 
-    rename(sf_eez_name = eez_name) %>% 
-    left_join(Matching_names) %>% 
-    filter(sau_fishing_entity %in% Med_land$sau_fishing_entity,
-           sf_eez_name != "Spain (Northwest)")
-
-  # Med data
-  Med_data <- Map_data %>% 
-    filter(source %in% Med_eez$sf_eez_name)
-
-  # Med paths
-  Med_paths <- paths %>% 
-    filter(source %in% Med_eez$sf_eez_name,
-           target %in% Med_eez$sf_eez_name)
-
-  # Med plot
-  Med_map <- ggplot() +
-    geom_sf(data = Med_eez, size = 0.5, fill = "white", col = "gray30") +
-    geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
-    geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
-            aes(fill = avrg_spp_per),
-            size = 0.1,col = "white") +
-    geom_point(data = subset(Med_data, time_step == time & per_plot < 0 ),
-               aes(start_lon,
-                   start_lat,
-                   col = per_plot),
-               size = 1,
-               col = "gray30",
-               shape = 20,
-               stroke = 1.5) +
-    geom_path(data = subset(Med_paths, time_step == time & per_plot > 0),
-              aes(X, Y,
-                  group = interaction(line_id,subline_id),
-                  col = per_plot
-              ),
-              size = 1,
-              arrow = arrow(length = unit(0.05, "npc"),
-                            ends = "first"),
-              show.legend = NA) +
-    coord_sf(xlim = c(-3, 35),ylim = c(30, 45)) +
-    labs(x = "", y = "") +
-    scale_x_continuous(breaks = seq(-5,35,10)) +
-    scale_y_continuous(breaks = seq(30,50,5)) +
-    my_ggtheme_m() +
-    theme(legend.position = "",
-          panel.background = element_rect(fill = "white", color = "white"),
-          panel.grid.major =  element_line(color = "white")
-    ) +
-    scale_colour_viridis(
-      name="Average Gain of\nStock-Share Ratio (%)",
-      limits = c(0,max(Med_data$per_plot)),
-      breaks = seq(0,max(Med_data$per_plot),5),
-      # option="magma",
-      # direction = -1
-    )+
-    scale_fill_viridis(
-      name="Average Gain of\nStock-Share Ratio (%)",
-      limits = c(0,max(Med_data$per_plot)),
-      breaks = seq(0,max(Med_data$per_plot),5),
-      # option="magma",
-      # direction = -1
-    )#;Med_map
+  # # Land polygon
+  # Med_land <- World_map %>% 
+  #   filter(region_wb %in% c("Europe & Central Asia","Middle East & North Africa") |
+  #            sau_fishing_entity %in% c("Spain","France")
+  #   )
+  # 
+  # # EEZ polygon
+  # Med_eez <- World_eez_sf %>% 
+  #   select(sf_eez_name = eez_name,everything()) %>% 
+  #   left_join(Matching_names) %>% 
+  #   filter(sau_fishing_entity %in% Med_land$sau_fishing_entity,
+  #          sf_eez_name != "Spain (Northwest)")
+  # 
+  # # Med data
+  # Med_data <- Map_data %>% 
+  #   filter(source %in% Med_eez$sf_eez_name)
+  # 
+  # # Med paths
+  # Med_paths <- paths %>% 
+  #   filter(source %in% Med_eez$sf_eez_name,
+  #          target %in% Med_eez$sf_eez_name)
+  # 
+  # # Med plot
+  # Med_map <- ggplot() +
+  #   geom_sf(data = Med_eez, size = 0.5, fill = "white", col = "gray30") +
+  #   geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
+  #   geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
+  #           aes(fill = avrg_spp_per),
+  #           size = 0.1,col = "white") +
+  #   geom_point(data = subset(Med_data, time_step == time & per_plot < 0 ),
+  #              aes(start_lon,
+  #                  start_lat,
+  #                  col = per_plot),
+  #              size = 1,
+  #              col = "gray30",
+  #              shape = 20,
+  #              stroke = 1.5) +
+  #   geom_path(data = subset(Med_paths, time_step == time & per_plot > 0),
+  #             aes(X, Y,
+  #                 group = interaction(line_id,subline_id),
+  #                 col = per_plot
+  #             ),
+  #             size = 1,
+  #             arrow = arrow(length = unit(0.05, "npc"),
+  #                           ends = "first"),
+  #             show.legend = NA) +
+  #   coord_sf(xlim = c(-3, 35),ylim = c(30, 45)) +
+  #   labs(x = "", y = "") +
+  #   scale_x_continuous(breaks = seq(-5,35,10)) +
+  #   scale_y_continuous(breaks = seq(30,50,5)) +
+  #   my_ggtheme_m() +
+  #   theme(legend.position = "",
+  #         panel.background = element_rect(fill = "white", color = "white"),
+  #         panel.grid.major =  element_line(color = "white")
+  #   ) +
+  #   scale_colour_viridis(
+  #     name="Average Gain of\nStock-Share Ratio (%)",
+  #     limits = c(0,max(Med_data$per_plot)),
+  #     breaks = seq(0,max(Med_data$per_plot),5),
+  #     # option="magma",
+  #     # direction = -1
+  #   )+
+  #   scale_fill_viridis(
+  #     name="Average Gain of\nStock-Share Ratio (%)",
+  #     limits = c(0,max(Med_data$per_plot)),
+  #     breaks = seq(0,max(Med_data$per_plot),5),
+  #     # option="magma",
+  #     # direction = -1
+  #   )#;Med_map
 
@@
   Eu_eez <- World_eez_sf %>% 
-    rename(sf_eez_name = eez_name) %>% 
+    select(sf_eez_name = eez_name,everything()) %>% 
     left_join(Matching_names) %>% 
@@
             size = 0.1,col = "white") +
-    geom_point(data = subset(Med_data, time_step == time & per_plot < 0 ),
+    geom_point(data = subset(Eu_data, time_step == time & per_plot < 0 ),
                aes(start_lon,
@@
       name="Average Gain of\nStock-Share Ratio (%)",
-      limits = c(0,max(Med_data$per_plot)),
-      breaks = seq(0,max(Med_data$per_plot),5),
+      limits = c(0,max(Eu_data$per_plot)),
+      breaks = seq(0,max(Eu_data$per_plot),5),
       # option="magma",
@@
     # Latin America and the Car.
-    draw_plot(Lat_am_map, x = 0.01, y = 0.02, width = 0.33, height = 0.45) +
+    draw_plot(Lat_am_map, x = 0.05, y = 0.00, width = 0.5, height = 0.50) +
     # Europe
-    draw_plot(Eu_map, x = 0.33, y = 0.01, width = 0.33, height = 0.45) +
-    # The Med
-    draw_plot(Med_map, x = 0.66, y = 0.02 , width = 0.33, height = 0.45) +
+    draw_plot(Eu_map, x = 0.52, y = 0.05, width = 0.35, height = 0.35) +
     # Legend
-    draw_plot(map_plot, x = 0.3, y = 0, width = 0.59, height = 0.1) +
+    draw_plot(legend, x = 0.25, y = 0, width = 0.59, height = 0.09) +
     # Labels
-    draw_plot_label(label = c("A", "B", "The Caribbean","Northwestern Europe","The Mediterranean"),
-                    size = c(20,20,15,15,15),
-                    x = c(0.01,0.01,0.058,0.33,0.68),
-                    y = c(1,0.45,0.44,0.44,0.44)
-    )
-
-  name <- paste0("./Figures/prop_",time,".png")
+    draw_plot_label(label = c("A", "B", "The Caribbean","Northwestern Europe"),
+                    size = c(20,20,15,15),
+                    x = c(0.01,0.01,0.14,0.50),
+                    y = c(1,0.45,0.44,0.44)
+    )
+
+  name <- paste0("./Figures/props_",time,".png")
 

### 2021-03-10 Scripts/Manuscript.Rmd
changes to title page

@@
 ---
-title: Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change
-subtitle: Challenges to transboundary fisheries under climate change
-author: 'Authors: Juliano Palacios-Abrantes*^1^, Thomas Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
+title: Early range shifts in transboundary fish stocks challenge their management under climate change
+# subtitle: Changes in sharing of transboundary fisheries under climate change
+author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
 date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
@@
 ^4^School of Environmental Sciences, University of Liverpool.
-^5^Center for Ocean Solutions, Stanford University, Stanford, United States'
+^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States'
 output:
-  # word_document:
-    # reference_docx: ../References/template.docx
-  pdf_document: default
+  # bookdown::word_document2
+    reference_docx: ../References/template.docx
+  # pdf_document: default
 bibliography: ../References/ToE_Ref.bib

### 2021-03-10 Scripts/Supplements.Rmd
Updated supplements for sci submission

--- 2020-12-15 Scripts/Supplements.Rmd
@@
 output:
-  bookdown::pdf_book:
+  bookdown::word_document2:
     toc : no
@@
     number_sections: yes
+reference_docx: ../References/supplementary_template.docx
+bibliography: ../References/ToE_Ref.bib
+csl: ../References/science.csl
 header-includes: #allows you to add in your own Latex packages
@@
 - \usepackage{caption}
-- \captionsetup[figure]{name = Figure A3}
-- \captionsetup[table]{name = Table A3}
+- \captionsetup[figure]{name = Figure S3}
+- \captionsetup[table]{name = Table S3}
 ---
 
-```{r setup, eval = T, echo=F, warning=F,message=F, results='hide'}
+```{r setup_sup, eval = T, echo=F, warning=F,message=F, results='hide'}
 library(MyFunctions)
@@
   "knitr",
-  # "kableExtra",
+  "kableExtra",
   "png",
@@
 
-```{r toeFigureS1, eval = T,  echo = F, fig.width=12, fig.height=7, fig.cap =  "World Exclusive Economic Zones used in this study as defined by the Sea Around Us and their centroids (points)."}
+**Materials and methods**
+
+_**Databases and species selection**_
+
+This analysis includes 633 exploited marine transboundary species that count for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
+Each time a species was shared by a pair of neighboring  EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.
+Using the transboundary species Atlantic cod (*Gadus morhua*) as an example and this study’s definition of transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not.
+We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015; http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
+\@ref(fig:toeFigureS1)) including island territories.
+We determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp].
+Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
+The habitat preference of each species was determined following FishBase (http://www.fishbase.org) for fish species and SeaLifeBase (http://www.fishbase.org) for invertebrates (Fig.
+\@ref(tab:toeTableS1)).
+For each stock and EEZ, we used Sea Around Us data to estimate catch and fishing revenue from fishing activities.
+We report both average catch and revenue for the last available decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+
+We forced the DBEM with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+The GFDL-ESM2M is run under historical forcing until 2005 and follows the high greenhouse gas emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand distribution variability during both the historical and the future periods, to infer differences between time frames.
+We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
+Thus, for our experiment, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in year 1950 resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
+By design, variations among ensemble members are then solely due to natural variability.
+This approach allows us to estimate the natural internal variability (e.g., background noise) and isolate the forced anthropogenic climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
+
+_**Projecting species distributions under climate change**_
+
+We projected the future distribution of transboundary stocks using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
+The DBEM determines the environmental space a species will occupy based on physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
+For pelagic species, the model uses environmental variables at the surface whereas demersal and benthic species’ distributions are driven by ocean bottom variables.
+The DBEM then estimates species’ abundance and maximum catch potential (a proxy for maximum sustainable yield) over a regular spatial grid of 0.5 of latitude per 0.5 of longitude (see Cheung et al.
+2009, 2010 and 2016 for model details).
+Importantly, DBEM is able to project catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
+*in revision*].
+
+We forced the DBEM with simulated ocean conditions from a ten member ensemble simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distribution of transboundary stocks from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+The GFDL-ESM2M is run under historical forcing until 2005 and follows the high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5) from 2006 to2100 [@Riahi:2011dk].
+We chose this time period to match the historical period in the GFDL ESM2M simulations over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Rodgers:2015jq].
+Because our main approach lies in identifying when the future shared distribution of a stock undergoes a substantial change in reference to historical variability, we follow an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
+To this end, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in 1950, resulting in a unique atmosphere and ocean state at each point in time after about three years for surface and eight years for subsurface waters [@Frolicher:2020cc].
+By design, variations among ensemble members are then solely due to natural internal variability (e.g., different phases of the El Niño Southern Oscillation-ENSO).
+
+_**Calculating an index of transboundary range shift**_
+
+We developed a Transboundary Index (TI) to evaluate range shifts in the shared distribution of transboundary stocks under climate change.
+This index represents the shift in the distribution centroid of a transboundary stock relative to the centroid of neighboring EEZs that share this stock (Fig.
+\@ref(fig:toeFigureS2)).
+The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across grid cells within the neighboring EEZs sharing the stock.
+Therefore,
+
+\begin{equation}
+\begin{split}
+\label{eq:one}
+lat_{ts} = \mu (lat_{per}) \\
+lon_{ts} = \mu (lon_{per})
+\end{split}
+\end{equation}
+
+where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
+To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighboring  EEZs sharing the stock was above the 95th percentile.
+A sensitivity analysis using a subset of species (*n* = 34) for all EEZs (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) to examine the effects of different thresholds on the calculated index value showed no apparent difference between percentiles (Fig.
+\@ref(fig:toeFigureS3)).
+The centroid of each EEZ was estimated using the st package in R (Fig.
+\@ref(fig:toeFigureS1)).
+For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids (e.g., EEZ and stock) using the geosphere package in R assuming the Earth is a perfect sphere and ignoring geographic barriers;
+
+\begin{equation}
+\label{eq:two}
+D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
+\end{equation}
+
+where $lat_{eez}$ and lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively, and $lon_{eez}$ and $lon_{ts}$ are associated longitudes.
+Then, for each year and for each ensemble member between 1951 and 2100 we calculated the transboundary index (TI) as follows:
+
+\begin{equation}
+\label{eq:three}
+TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
+\end{equation}
+
+where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring  EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+Thus, a higher TI means that the abundance of a transboundary stock within neighboring  EEZs changes beyond historical levels.
+From the perspective of a state’s fisheries, increasing TI indicates an increase in the challenges to effective management and sustainability of the transboundary stock’s fisheries.
+In contrast, a decreasing TI suggests that the sharing of the stocks between neighboring  EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries.
+The index was smoothed to a 10-year average to reduce interannual variability.
+
+_**Calculating shifts in the shared distribution of transboundary stocks**_
+
+Knowing the point in time at which the distribution of a shared stock will diverge from historical variability is important to inform decision makers when climate adaptation measures will need to be implemented [@Link:2010ei].
+To that end, we adopted the concept of time of emergence, defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (e.g., historical natural variation) [@Hawkins:2012hc], and commonly applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
+The premise behind time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
+The TI’s natural internal variability was estimated from 1951 to 2005 ($TI_\sigma$) across ensemble members, and the TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$).
+We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See (Fig.
+\@ref(fig:toeFigureS2)) for a graphical description):
+
+<!-- $EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$ -->
+<!-- $ToE = t when Emergence index[t] > 1$ -->
+
+\begin{equation}
+\label{eq:four}
+ToE = TI_\mu[t']>2* TI_\sigma 
+\end{equation}
+
+Thus, we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TI “emerges” from historical internal variability (t’).
+This assumes that, from a fisheries’ perspective, the first year a stock’s  distribution shifts from a 10 year average is sufficient to call for caution.
+We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
+We performed two non-parametric Kruskal–Wallis tests by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the range shift of transboundary species (See supplementary information for complete statistical analysis).
+
+_**Quantifying the intensity of transboundary stocks’ range shift**_
+
+Here, we adopt the concept of threat point to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring  EEZs.
+The concept of threat point comes from Game Theory and has been widely used in shared fisheries management [@Bailey:2010ga; @Sumaila:2013hv; @Sumaila:2020tp; @Clark:1975uu; @Munro:1979dg].
+In a game theoretic model, a player’s strategy (e.g., to act cooperatively or not) will have direct consequences on other players, which, in turn, will affect the overall outcome of the game [@Bailey:2010ga].
+Cooperation will often result in the maximization of benefits for the system rather than for individual players.
+However, for a cooperation strategy to work, the benefit a player gets must not be less than the benefit under a non-cooperation strategy.
+Thus, the “threat point” is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
+
+We define the threat point as the minimum required catch proportion of a transboundary stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Any proportion below the defined threat point would result in the unilateral management of the stock.
+First, we estimated the stock share ratio (SSR) of each transboundary stock shared by neighboring  nations between 1951 and 2100.
+We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
+Second, to reduce the effects of variability, we averaged the calculated proportion of stock occupying a given EEZ into three time periods: the historical time period ($SSR_{th}$) from 1951 to 2005, and two future periods, the early 21st century, between 2021 and 2040 ($SSR_{te}$), and the mid 21st century, defined as the average of 2041 to 2060 ($SSR_{tm}$).
+These future time periods were chosen to correspond to the challenges of achieving the fisheries-related UN-SDGs by 2030 (i.e., the 2030 Agenda), such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) and SDG 1.2 (poverty reduction) [@Singh:2017ds; @UnitedNations:1Qe8KIcN].
+The analysis was replicated for projected stock distributions from each of the ten ensemble members and results were averaged across ensemble members.
+Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm 2 \sigma$, where  is the standard deviation of $SSR_{th}$.
+Thus, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds two standard deviations of the historical variations in SSR (i.e., when $SSR_{et} \geqq (SSR_{th} + 2 \sigma$) or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
+Finally, we estimated the percentage change in SSR ($\Delta SSR$) of each future time period ($SSR_{tf}$) relative to the historic time period $SSR_{th}$ for each stock whose share ratio exceeded the threat point following Palacios-Abrantes et al.
+(2020);
+
+<!-- $$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$ -->
+
+\begin{equation}
+\label{eq:five}
+\Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100
+\end{equation}
+
+_**Statistical analysis**_
+
+We tested the results for normality (e.g., skewness, kurtosis) and performed two non-parametric Kruskal–Wallis tests by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the range shift of transboundary species.
+Specifically, we tested if the habitat preference of transboundary species and the geographic location of EEZs would have any effect on the transboundary stocks distributional range shift.
+In both cases, our null hypothesis was that there were no significant differences in the year of shift across habitat association nor EEZs.
+All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages data.table [@Packagedatatable:2019uh], janitor [@PackagejanitorSim:2018ur], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], gmt [@PackagegmtInterfa:2017vh] and zoo [@PackagezooSInfr:2019tm], ggrepel [@PackageggrepelAut:2020to], zeallot [@PackagezeallotMul:2018wy], viridis [@PackageviridisDef:2018us], cowplot [@CowplotStreamlined:2019wt], moments [@PackagemomensMome:2015uh] and pgirmess [@PackagepgirmessSp:2018um].
+Code for all analyses is available at https://github.com/jepa/TransEmergence
+
+\clearpage
+
+_**Figures**_
+
+```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap ="World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their centroids (points)."}
 
@@
 
-```{r toeFigureS2, eval = T,  echo = F, fig.width=12, fig.height=5, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
-The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel).
+\clearpage
+
+```{r toeFigureS2, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap = "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring  Exclusive Economic Zones sharing the stock (Top panel).
 Time of emergence is defined as the first year when the TI overpasses historical values (Bottom panel)."}
@@
 
-```{r toeFigureS3, eval = T,  echo = F, fig.width=14, fig.height=7, fig.cap ="Sensitivity analysis of the number of grid cells with projected stock abundance within the neighbouring EEZs sharing the stock.
-The level used in this article was the top 95th percentile (e.g., far right column)."}
+\clearpage
+
+```{r toeFigureS3, eval = T,  echo = F, fig.width=14, fig.height=7, fig.cap ="Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine the stock’s centroid.
+Top Row: results in absolute numbers.
+Bottom row: results expressed as a percentage of the total emerging stocks and EEZ with emerging stock."}
 
@@
 
-```{r toeFigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
+\clearpage
+
+```{r toeFigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average and standard deviation of distribution of mean year of range shift for a 95% confidence (yellow) and 68% (purple)"}
 
@@
 
-```{r toeFigureS5, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
+\clearpage
+
+```{r toeFigureS5, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Changes in Stock Share Ratio by 2050 (2041-2060) relative to 1951-2005.
+Lines represent the average gain of transboundary stock share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial stocks that will see a shift.
+Panel B zooms into specific areas shown in grey boxes of A."}
 
@@
 <!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005))](../Figures/Supp2.png) -->
+\clearpage
 
-```{r toeFigureS6, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial species that will see a shift.
-Panel B zooms in to specific areas shown in grey boxes of A."}
+_**Tables**_
 
-img <- readPNG("../Figures/Supp6.png")
-grid.raster(img)
+```{r toeTableS1, eval = T, echo = F}
+
+tibble(
+  "Habitat preference" = c("pelagic-neritic","reef-associated","bathydemersal","benthopelagic","demersal","bathypelagic","pelagic-oceanic"),
+  Description = c(
+    "The shallow pelagic zone over the continental shelf; nearshore ocean ecosystems; i.e., those associated with the coasts because the waters are overlying continental shelves and/or the waters are < 200 m deep in areas of coastal submarine slopes",
+    "A largely consolidated wave resistant feature; the upper surface is within 0-20 m of the ocean surface; in the tropics, the upper portion is typically composed mainly of living and non-living remains of coral and coralline algae; rock lying at or near the surface and which may pose a danger to navigation",
+    "Living and feeding on the bottom below 200 m",
+    "Living and feeding near the bottom as well as in midwaters or near the surface.
+Feeding on benthic as well as free swimming organisms.
+Many freshwater fish are opportunistic feeders that forage on the bottom as well as in midwater and near the surface, also pertaining to forms which hover or swim just over the floor of the sea, e.g.
+Halosauridae, Macrouridae, Moridae, Brotulidae; the depth zone about 100 metres off the bottom at all depths below the edge of the continental shelf",
+    "Sinking to or lying on the bottom; living on or near the bottom and feeding on benthic organisms",
+    "Region of the oceanic zone between 1,000 m to 4,000 m; between the mesopelagic layer above and the abyssopelagic layer below.
+Living or feeding in open waters at depths between 1,000 and 4,000 m.
+In FishBase this term is used to include the depth range from 200 m to the bottom and thus the zones mesopelagic, bathypelagic and abyssopelagic",
+
+    "Living and feeding in the open sea; associated with the surface or middle depths of a body of water; free swimming in the seas, oceans or open waters; not in association with the bottom.
+Many pelagic fish feed on plankton.
+In FishBase, referring to surface or mid water from 0 to 200 m depth."
+  )
+) %>% 
+ kable(
+        booktabs = T,
+        caption = "Definition of each habitat preference according to FishBase"
+  ) #%>% 
+  # kable_styling(full_width = F,
+                # position = "center")
 
@@
 
-```{r toeTableS1, eval = T, echo = F}
+\clearpage
+
+```{r toeTableS2, eval = T, echo = F}
 
@@
   select(-statistic,
-         "Habitat preference comparrison" = comparrison,
+         "habitat association comparrison" = comparrison,
          "Significan level" = signif.level,
@@
           "K-W Difference" = dif.com.difference) %>% 
-  # kable(caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions")
-  kable("latex",
+  kable(
         booktabs = T,
         caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  ) %>%
-  kable_styling(full_width = T,
-                position = "center")
+  ) #%>% 
+  # kable_styling(full_width = F,
+                # position = "center")
 
@@
 
-```{r toeTableS2, eval = T, echo = F}
+\clearpage
+
+```{r toeTableS3, eval = T, echo = F}
 
@@
           "K-W Difference" = dif.com.difference) %>% 
-  # kable(
-  #       booktabs = T,
-  #       caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  # )  
-  kable("latex",
+  kable(
         booktabs = T,
         caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  ) %>%
-  kable_styling(full_width = F,
-                position = "center")
+  ) #%>% 
+  # kable_styling(full_width = F,
+                # position = "center")
 

### 2021-03-10 Scripts/Manuscript.Rmd
updated manuscript for submission for gdrive

@@
 ---
-title: Early range shifts in transboundary fish stocks challenge their management under climate change
-# subtitle: Changes in sharing of transboundary fisheries under climate change
+title: Identifying timing of shifts in transboundary fish stocks to support climate resilient fisheries management
+
 author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
@@
 output:
-  # bookdown::word_document2
-    reference_docx: ../References/template.docx
+  bookdown::word_document2
+reference_docx:  ../References/manuscript_template.docx
   # pdf_document: default
@@
 
-<!-- * corresponding author: j.palacios@oceans.ubc.ca -->
+-   corresponding author: [j.palacios\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}
 
-<!-- # Abstract (150  words) -->
+**Abstract**
 
@@
 
-Climate change shifts in marine species distributions is affecting the current frameworks emplaced to manage species shared fisheries.
-Here, we rely on a modeling framework with multiple ensemble members to identify the time frame by which shared stocks distribution will change from historical ranges, and the change in the proportion of the distrirbution that each EEZ has under climate change.
-Our results suggest that 60% of shared species have emerged since 2006 in 87% of the world EEZs.
-By 2075 all EEZs will have an emerging species and by 2097 and all species will have emerged in at least one EEZ.
-Average changes in the ratio of species-share between neighboring EEZs will fluctuate between 1 and 10% with some exceeding 25% by 2030 relative to today’s ratio.
-Our results suggest that most changes will happen before 2030 urging policy to be re-formulated to account for the possible implications of these changes.
+Climate change is shifting the distribution of transboundary fish stocks between neighbouring countries.
+The time-frame of such shifts determine the extent of climate challenges to international fisheries governance.
+Here, using global climate-fisheries simulation modelling, we show that, by 2036, 68% of the world’s transboundary stocks are projected to shift their distributions significantly across international boundaries.
+By 2100, such shifts will occur in all countries under a ‘no mitigation’ scenario.
+Hotspots where transboundary stock shifts are earlier and larger coincide with countries that are highly dependent on fisheries for income and livelihood.
+The early occurrences of transboundary shifts highlight the urgency to re-formulate international fisheries policies to avoid disputes in fish stock sharing.
+Such bi-/multi-lateral discussions should start now as lead-time for policy changes is long.
 
-## Introduction
+**One-Sentence Summary**
 
-Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn].
-Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk].
-As a result of climate change, many marine species have changed their distributions towards higher latitude, deeper water or following local temperature gradients [@Poloczanska:2016kk].
-Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].
-The projected risks of impacts can be reduced through improving the effectiveness of current fisheries managemen [@Gaines:2018sg; @Cheung:2018dh], including the fisheries management of species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
+Early range shifts in transboundary fish stocks challenge their management under climate change
 
-The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
-As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, those that cross neighboring EEZs; (*ii*) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (*iV*) discrete stocks that are only present in the high seas [@Munro:2004th].
+**Main Text.**
+
+Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters, [@IPCC:2019tn].
+Marine species' distributions reflect species' preferences for discrete environmental conditions [@Hutchinson:1957nt; @Nelson:2016kk].
+As a result of climate change, many marine species have shifted their distributions towards higher latitude, deeper water or following local temperature gradients to track and stay within their optimal environmental 'niche'[@Poloczanska:2016kk; @Barange:2014kr].
+The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt; @Barange:2014kr], impacting fisheries production and economies [@Sumaila:2019ew], and compromising our capacity to reach international sustainability goals within the 2030 Agenda such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
+The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management [@Cheung:2018dh], including for species that cross international borders, i.e., shared stocks [@Pinsky:2018cb].
+
+The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by Coastal States [@UnitedNations:1986tl].
+As defined by the United Nations' Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (*iv*) discrete stocks that are only present on the high seas [@Munro:2004th].
 This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
-While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, and yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
+Countries are responsible for the management of stocks within their EEZs, under UNCLOS, and are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
+A recent study estimates that, globally, there are 633 transboundary fish species representing 67% of identified exploited taxa.
+These species yielded an annual average of 48 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
 
-The effectiveness of fisheries management for transboundary stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
-In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the species [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
-Misalignment between fisheries resources allocation and species’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb].
-With future shifts in species distributions under climate change, the challenges for the fisheries management of transboundary stocks will continue to increase during the $21^{st}$ century.
-In order to prepare ocean governance to meet the resource management challenges arising from shifting transboundary stocks, it is important to constrain when climate change will affect the dynamics of transboundary stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].
+The effectiveness of fisheries management for transboundary species, and thus reaching the 2030 Agenda [@Pecl:2017hu], is challenged by shifts in stocks' distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].
+In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017iua; @Spijkers:2017ij; @Crespo:2020cs].
+Such patterns are expected to be exacerbated by continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
+The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the SDGs.
+Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important to developing climate resilient international ocean governance and achieving the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].
 
-The timing and intensity of climate impacts on managing transboundary fish stocks can be quantitatively examined through the concepts of ‘time of emergence’ (ToE) and ‘threat point’ (TP).
-Time of emergence is defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (e.g., historical natural variation) [@Hawkins:2012hc] and has been applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
-The premise behind the time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
-The concept of threat point comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
-Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary species between Canada and the United States [@Sumaila:2020tp].
+Here, we estimate the year in which 663 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variation and the intensity of said shift.
+We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
+We employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks' distribution across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (RCP8.5; Methods).
+We find that the shared distribution of many transboundary stocks has already shifted beyond historical natural variability.
+Moreover, by 2030, the shared catch proportion of such stocks will exceed historical levels in most neighboring EEZs, posing a threat to the sustainability of associated fisheries and the resilience of international treaties that manage their shared exploitation.
 
-Here, we estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
-We employ a species distribution model driven by output from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species [@PalaciosAbrantes:2020tv] across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP8.5 (see Methods).
-We treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude the high seas).
-We find that many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.
+**Identifying range shifts in the shared distribution of transboundary stocks**
 
-## Materials and methods
+We calculated the time of emergence of the transboundary index to identify shifts in the shared distribution of transboundary stocks from 2006 to 2100 under a high emission climate change scenario.
+Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability as early as 2006 and up to 2100 (Fig.
+\@ref(fig:toeFigure1)A).
+This corresponds to 68% of the studied stocks (Fig.
+\@ref(fig:toeFigure1)C).
+On average, the shift in shared distribution of these stocks in all EEZs analyzed is projected to be 2036 $\pm$ 28 years.
+Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% presenting an early distributional shift across stocks (Fig.
+\@ref(fig:toeFigure1)C).
 
-### Databases and species selection
-
-This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv]; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks in this analysis.
-Using the transboundary species Atlantic cod (*Gadus morhua*) as an example, and based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species.
-We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1)) including island territories.
-We determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp].
-Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
-The habitat association of each species was determined following the classification in FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org).
-For each stock and EEZ, we used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
-We report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
-The monthly average US Consumer Price Index (CPI) according to the U.S.
-Bureau of Labor Statistic (https://www.bls.gov/cpi/) was used to standardize all monetary values to 2019 real USD.
-
-### Projecting species distributions under climate change
-
-We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
-The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
-For pelagic species, the model uses environmental variables at the surface whereas demersal species distribution is driven by ocean bottom variables .
-DBEM then estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid (see Cheung et al.
-2010 and 2016 for model details).
-Importantly, DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd], Cheung et al.
-*in revision*].
-
-We forced the DBEM with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
-The GFDL-ESM2M is run under historical forcing until 2005 and follows the high greenhouse gas emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand distribution variability during both the historical and the future periods, to infer differences between time frames.
-We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
-Thus, for our experiment, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in year 1950 resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
-By design, variations among ensemble members are then solely due to natural variability.
-This approach allows us to estimate the natural internal variability (e.g., background noise) and isolate the forced anthropogenic climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
-
-### Calculating an index of transboundary range shift
-
-I developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
-This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2)).
-The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., sockeye salmon (*Oncorhynchus nerka*) that is transboundary between the United States and Canada.
-Therefore,
-
-<!-- $lat_{ts} = \mu (lat_{per})$ -->
-
-<!-- $lon_{ts} = \mu (lon_{per})$ -->
-
-\begin{equation}
-\begin{split}
-\label{eq:one}
-lat_{ts} = \mu (lat_{per}) \\
-lon_{ts} = \mu (lon_{per})
-\end{split}
-\end{equation}
-
-where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
-To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the  95th percentile (e.g., per = 95%).
-We ran a sensitivity analysis using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
-The centroid of each EEZ was estimated using the st package in R (Figure \@ref(fig:toeFigureS1)).
-For each ensemble member, neighboring EEZs and transboundary stock, the distance between centroids was computed assuming the Earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
-
-<!-- $$D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))$$ -->
-
-\begin{equation}
-\label{eq:two}
-D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
-\end{equation}
-
-where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively and $lon_{eez}$ and $lon_{ts}$ are the respective longitudes.
-Then, for each year we calculated the transboundary index (TI) as follows:
-
-<!-- $$TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2$$ -->
-
-\begin{equation}
-\label{eq:three}
-TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
-\end{equation}
-
-where $D_A$ and $D_B$ represent the distance between the distribution centroids of a species and the centroid of the pair of neighbouring EEZs (A and B) sharing the species for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
-Thus, a higher TI means that a shared stock becomes more or less abundant relative to its neighbouring EEZs that is beyond what the fisheries experienced historically.
-From the perspective of a state’s fisheries, higher TI indicates an increase in the challenges to effective management and sustainability of the shared stock’ fisheries.
-In contrast, a smaller TI suggests that the sharing of the stocks between neighbouring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the shared stocks’ fisheries.
-The index was smoothed to a 10-year average to reduce interannual variability.
-
-#### Calculating the time of emergence of the transboundary index
-
-Knowing the point in time (e.g., year) at which the distribution of a shared stock will diverge from its natural internal variability is important in informing the lead-time to which climate adaptation needs to be implemented [@Link:2010ei].
-Here, we define time of emergence of the tranboundary index (ETI) as the time when the stock shared between neighbouring EEZs will exceed the historical natural internal variability as follows:
-
-<!-- $EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$ -->
-<!-- $ToE = t when Emergence index[t] > 1$ -->
-
-\begin{equation}
-\begin{split}
-\label{eq:four}
-ETI = \frac{\mu_{TI[t]}}{SD_{TI_{th}}} \\
-ToE = t when Emergence index[t] > 1 s.d.
-\end{split}
-\end{equation}
-
-where $\mu_{TI}$ is the yearly mean TI across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005).
-We set a threshold of TI above which it was consider a shared stock to have emerged from the historical internal variability of the TI.
-The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four), representing a probability of 68% that the transboundary index has emerged from historical variability).
-We tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
-
-<!-- #### Computing a transboundary fisheries vulnerability index (VI) -->
-
-<!-- We created a country-specific vulnerability index (VI) to examine the risk of impacts of species’ range shifts on transboundary fisheries management.
-The VI reflects both the rate of change of shared stocks’ distribution (indicated by the time-of-emergence of the transboundary index) and the importance of the shifting species to a country’s total catch and revenue from the species.
-The VI is calculated from: -->
-
-<!-- $$ index' = (P_c + P_v)/ToE $$ -->
-
-<!-- \begin{equation} -->
-<!-- \label{eq:five} -->
-<!-- VI' = \frac{(P_c + P_v)}{ToE} -->
-<!-- \end{equation} -->
-
-<!-- where $P_c$ and $P_v$ indicate the proportion accounted for by shared species in total catch and value generated from transboundary fisheries within a country’s EEZ, respectively.
-Here, the proportion of total catch and value are represented as the average of 2005 to 2014 within individual EEZs according to the Sea Around Us.
-The VI was then normalized to a 0 (less vulnerable) to 100 (more vulnerable) scale.
--->
-
-Estimating the threat point of transboundary stock share
-
-In game theory, cooperation over a transboundary resource will more likely result in a better overall outcome for the sharing parties than those operating on a non-cooperative basis [@Munro:1979dg].
-However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., “threat point” [@Sumaila:2013hv].
-Here, we define the threat point as the minimum required proportion of a transboundary stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-Any proportion below the defined threat point of a country would result in a deflection from the joint management of the transboundary stock.
-
-First we estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
-We did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
-Second, we average the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of inter-annual climate variability.
-The first time period ($SSR_{th}$) spans from 1951 to 2005 and matches the historical period in the GFDLESM-2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
-It is assumed as the historic baseline of shared distributions for each stock.
-We then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$).
-We chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
-The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members.
-Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$.
-This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{et} \geqq (SSR_{th} + \sigma)$ or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
-Finally, we estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al.
-(*in press*),
-
-<!-- $$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$ -->
-
-\begin{equation}
-\label{eq:five}
-\Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100
-\end{equation}
-
-### Statistical analysis
-
-The time of emergence results were tested for normality (e.g.
-skewness = 1.02, kurtosis = 2.83) and performed a couple of non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary species.
-Specifically, we tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of the transboundary index of transboundary stocks.
-For both cases, our null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
-<!-- Moreover, we implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
-Specifically, we looked at habitat associations, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
-We hypothesize that all of these factors would have significant influence in the ToE of transboundary species.
--->
-All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages data.table [@Packagedatatable:2019uh], janitor [@PackagejanitorSim:2018ur], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], gmt [@PackagegmtInterfa:2017vh] and zoo [@PackagezooSInfr:2019tm], ggrepel [@PackageggrepelAut:2020to], zeallot [@PackagezeallotMul:2018wy], viridis [@PackageviridisDef:2018us], cowplot [@CowplotStreamlined:2019wt], moments [@PackagemomensMome:2015uh] and pgirmess [@PackagepgirmessSp:2018um].
-All code is available at https://github.com/jepa/TransEmergence
-
-## Results
-
-```{r Load_Results, eval = TRUE,  echo = FALSE, include = FALSE}
-
-prop_r <- read_csv("/Volumes/HALL2000/Data/TransEmergence/Results/proportion_change_2005.csv") %>% 
-  gather("time_step","change",per_early,per_mid)
-
-```
-
-Time of emergence of transboundary stocks
-
-The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:toeFigure1)).
-In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100.
-Fish stocks with transboundary index that first emerged from their historical variability in each EEZ was projected to be as early as 2006 and the last one in 2100.
-About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g.
-Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010).
-The number of EEZs with emerging stocks is projected to increase steeply from 87% in 2006 to 96% in 2020.
-After the mid-2020s, the rate of increase in the number EEZs slow down to an average of 0.7% per year.
-By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:toeFigure1)).
-Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020.
-Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure \@ref(fig:toeFigure1)).
-We based our transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
-Results from the sensitivity analysis suggest that our method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g.
-the top 90%) of the stock (Figure \@ref(fig:toeFigureS3)).
-
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence by Exclusive Economic Zone (EEZ) and transboundary stock."}
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary stock shared distribution.
+Dashed lines represent the year by which countries have committed to reach the full implementation of the 2030 Agenda, as well as an early and late change in range shift, respectively."}
 
@@
 
-The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A).
-Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1)) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B).
-In contrast, EEZs located in template regions like northern Europe and eastern Asia will have a significantly later time of emergence (p < 0.05; see Table \@ref(tab:toeTableS1 for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058).
-However, notable exceptions of tropical EEZs with a later average time of emergence across stocks include Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070).
-Moreover, despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) will have relatively early average time of emergence (Figure \@ref(fig:toeFigure2)A).
-Only in a few cases, the transboundary index of a stock will not emerge between 2006 and 2100.
-In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
-In the case of Brazil, there were four transboundary stocks included, all shared with Uruguay [@PalaciosAbrantes:2020tv]: Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
-In all cases, our study suggest that the transboundary index of these stocks will not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A)
+The median year in which transboundary stocks experience a range shift varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Figure \@ref(fig:toeFigure2)A).
+Overall, most tropical EEZs will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia being significantly earlier (*p* < 0.05; see Table \@ref(tab:toeTableS2) for test statistics) than almost all other regions (Fig.
+\@ref(fig:toeFigure2)B).
+In contrast, EEZs located in temperate regions like northern Europe and eastern Asia are projected to experience significantly later range shifts (Table \@ref(tab:toeTableS2)).
+Stocks in some countries, like Brazil or New Zealand, are not projected to experience range shifts beyond natural variability (e.g., the transboundary index of a stock did not emerge from historical variability under a 95% confidence interval) (Fig.
+\@ref(fig:toeFigure2)A).
+On average, shifting transboundary stocks represented 23% $\pm$ 22% of yearly fishing revenue from fisheries targeting such stocks within global EEZs between 2004 and 2010 (Fig.
+\@ref(fig:toeFigure2)A).
+However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
+Moreover, in some EEZs, while few transboundary stocks are projected to shift above historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).
 
-```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of stocks’ transboundary index.
-A) Land polygon shows the participation of the emerging stocks in the country’s total fishing revenue from transboundary stocks.
-Exclusive Economic Zone polygon represents the average time of emergence of the transboundary index across all stocks.
-Aqua color represents EEZs with no time of emergence between 2006 and 2100.
-B) Time of emergence of the transboundary index according to the United Nations sub-regions.
-Whiskers represent 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range."}
+```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks.
+A) Land polygon shows the contribution of shifting stocks to the country’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygons represent average range shifts per EEZ.
+EEZs with no time of emergence between 2006 and 2100 are represented in aqua color.
+B) Year of shared distribution shifts by region according to United Nations sub-regions.
+Points represent the median year of range shift.
+Horizontal dashed line represents the 2030 Agenda.
+N = North, S = South, W = West and E = East.
+Ltn.
+Ame.
+and the Car.
+= Latin America and the Caribbean.
+Aus and New Z.
+= Australia and New Zealand"}
 
@@
 <!-- ![toeFigure 2.
-Average time of emergence of stocks’ transboundary index.
-A) Land polygon shows the participation of the emerging stocks in the country’s total fishing revenue from transboundary stocks.
-Exclusive Economic Zone polygon represents the average time of emergence of the transboundary index across all stocks.
-Aqua color represents EEZs with no time of emergence between 2006 and 2100.
-B) Time of emergence of the transboundary index according to the United Nations sub-regions.
-Whiskers represent 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent points outside of 1.5*IQR.](../Figures/Fig3.png) -->
+Year of change in the shared distribution of 4,119 transboundary stocks.
+A) Land polygon shows the contribution of shifting stocks to the country’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygons represent average range shifts per EEZ.
+EEZs with no time of emergence between 2006 and 2100 are represented in aqua color.
+B) Year of shared distribution shifts by region according to United Nations sub-regions.
+Points represent the median year of range shift.
+Horizontal dashed line represents the 2030 Agenda.
+N = North, S = South, W = West and E = East.
+Ltn.
+Ame.
+and the Car.
+= Latin America and the Caribbean.
+Aus and New Z.
+= Australia and New Zealand](../Figures/Fig3.png) -->
 
-Our analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see their distribution range shifts beyond historical variabilities between 2025 and 2040 (Figure \@ref(fig:toeFigure3)).
-In some cases, like Indonesia (ranked $8^{th}$ in terms of revenue from transboundary stocks), Thailand (ranked $7^{th}$ in terms of revenue from transboundary stocks) and the United States (ranked $4^{th}$ in terms of fishing revenue from transboundary stocks), stocks’ range shifts swill emerge by around 2030.
-The 47 countries that make up the bulk of the global fisheries revenues also have more stocks with range shifts emerging from historical variabilities, being Spain (n = 122), France (n = 85), China and Portugal (both n = 73) and Senegal (n = 66) the top 5 leading countries in number of emerging stocks.
+We estimated the current (2005 and 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs.
+Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD in fishing revenue are projected to shift within the EEZs of 76 counties (Fig.
+\@ref(fig:toeFigure3)); with 4.2 billion USDs accounted by 321 stocks shifting in just eleven countries (Fig.
+\@ref(fig:toeFigure3)).
+Most of the remaining countries in the top $75^{th}$ revenue percentile will see early shifts (e.g., before 2050) with the exception of Japan, Korea, Peru and Russia, who are projected to have later shifts (e.g., after 2050).
+In terms of the number of transboundary stocks shifting from historical variability, Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66), emerge as the top 5 countries (Fig.
+\@ref(fig:toeFigure3)).
 
-On average, stocks that will see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of their EEZ’s revenue produced by fisheries targeting transboundary species between 2004 and 2010.
-However, large variation exists in the proportion that stocks with emerging transboundary index have in the current revenue participation across EEZs.
-In some cases e.g., El Salvador and New Zealand, stock’s with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries  in their respective EEZs between 2004 and 2014.
-On the contrary, stocks with emerging transboundary index from the Marshal Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A).
-In addition, some EEZs are projectede to have a few emerging stocks that represent large proportion of revenue from transboundary stocks e.g., in Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively).
-Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks on fishing revenue is less than 20%.
-
-```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average time of emergence of stocks’ transboundary index per fishing entity, aggregated by colour and shape according to region.
-Fishing revenue on a logarithmic scale.
-Showing country names for the top 90th percentile of annual fishing revenue."}
+```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks per fishing entity.
+Countries are coded according to different colored symbols representative of 5 distinct regions.
+Horizontal line separating the top 75th percentile of annual fishing revenue.
+Showing country names for those within the top 75th percentile projected to  have an average shift within 5 years of the 2030 Agenda."}
 
@@
 
-I set an arbitrary threshold of 1 s.d.
-(67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods).
-We tested the sensitivity of our methods by considering a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence).
-The global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
-threshold (Figure \@ref(fig:toeFigureS2)).
-In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric, that is, 40% (n = 1626) fewer emerging stocks than using a 1.s.d threshold.
+We compared the year in which transboundary stocks experience a range shift by species’ preferred habitat association (Fig.
+\@ref(fig:toeFigure4)).
+We found significant differences in the range shift year relative to habitat association (Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; See table S3 for full statistics \@ref(tab:TabelS3)).
+The average range shift of bathy-pelagic species (e.g., Blue whiting *Micromesistius poutassou*), pelagic-oceanic (e.g., tunas) and reef species (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to happen before 2030 while bathydemersal species (e.g., Banguela hake, *Merluccius polli*) are expected to experience the latest shift (2040).
 
-I compared species time of emergence within EEZs by species’ habitat association (Figure \@ref(fig:toeFigure4)).
-Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods).
-We found significant differences in the time of emergence of stocks range shifts, indicated by the transbounudary index, by their specie's habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
-Specifically, time of emergence of range shifts of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
-On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (toeFigure \@ref(fig:toeFigure4)).
-Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:toeFigure4)).
-
-```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association.
-Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Methods).
-The number of species included in this analysis for each habitat type is noted in parenthesis.
-Whiskers represent 1.5* interquartile range.
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
+Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
+The category “Other” consists of species that have no “habitat preference” classification.
+The number of stocks included in this analysis for each habitat type is noted in parentheses.
+Dashed whiskers represent the 1.5* interquartile range.
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range."}
+Solid line represents the median time of emergence per habitat preference."}
 
@@
 <!-- ![toeFigure 4.
-Regional time of emergence per habitat association.
-Regions according to the United Nations and habitat association to FishBase.
-Whiskers represent 1.5* interquartile range.
+Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
+Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
+The category “Other” consists of species that have no “habitat preference” classification.
+The number of stocks included in this analysis for each habitat type is noted in parentheses.
+Dashed whiskers represent the 1.5* interquartile range.
 Box represents interquartile range as distance between first and third quartiles.
-Line represents median, and black points represent outliers (outside of 1.5*IQR.](../Figures/Fig5.png) -->
+Solid line represents the median time of emergence per habitat preference.](../Figures/Fig5.png) -->
 
-<!-- We hypothesized that species environmental preferences (e.g.
-presence in certain biomes and habitats), climate change effects on water temperature (e.g.
-ToE of SST), human drivers (e.g.
-catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary species (Table \ref(tab:toeTableS3)).
-Results from the multiple linear regression analysis show a significant relation between ToE of transboundary species and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref(tab:toeTable1)).
--->
+We based the transboundary index on a stock’s centroid defined as the average grid cells that contain the majority of a stock’s abundance (see Methods - Calculating an index of transboundary range shift).
+Results from the sensitivity analysis suggest that our method is robust to the abundance level (e.g., the top 95%) of the stock (Fig.
+\@ref(fig:toeFigureS3)).
+In addition, we set an arbitrary threshold of 2 s.d.
+(95% confidence) to determine the time of emergence of the transboundary index.
+Sensitivity tests to a lower threshold of 1 s.d.
+(for a 68% confidence) resulted in a global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27, a difference of about 7 years (Fig.
+\@ref(fig:toeFigureS4)).
+In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks.
 
-```{r Table1, eval = F,  echo = F, message = F, warning  = F}
+**Changes in the stock share ratio of transboundary stocks**
 
-read_csv("../Tables/mlr_result.csv") %>% 
-  knitr::kable(
-    "latex",
-    booktabs = T,
-    captio= "Multiple regression results for time of emergence of transboundary species.
-Only showing significant variables.
-The model significantly predicts time of emergence (F[18,6155] = 20.68, p < 0.001, R2 = 0.05)"
-  )
-
-```
-
-### Changes in the stock share ratio of transboundary stocks
-
-I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods - equation \@ref(eq:five)).
-By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A).
-The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4); \@ref(fig:toeFigureS5)).
-However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high.
-For example, Guatemala shares Panulirus gracilis stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century.
-Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)).
-Our results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
-These EEZs will see one third (33% $\pm$ 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A).
-Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin will see changes in all of the transboundary stocks analyzed while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) will be the only countries with less than 10% of stocks changing beyond the threat point.
+We identified transboundary stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point by the early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) under a high emission climate change scenario.
+Findings show that by the early $21^{st}$ century, the global average catch proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 59% $\pm$ 17% (Fig.
+\@ref(fig:toeFigure5)A).
+By the early $21^{st}$, 85% (n = 239) of the world EEZs would have experienced such changes.
+Changes in stock share ratio are expected to increase slightly towards the mid 21st century at both the stock and EEZ level (Fig.
+\@ref(fig:toeFigure5)A and \@ref(fig:toeFigureS5)).
 
 ```{r toeFigure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those that are gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point.
-Panel B zooms in to specific areas shown within the grey dashed line boundaries in A."}
+Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors (green to yellow).
+Panel B zooms into specific areas shown within the grey dashed line boundaries in A."}
 
@@
 <!-- ![Figure 1.
-Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average gain of transboundary specie share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial species that will see a shift.
-Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Fig1.png) -->
+Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors (green to yellow).
+Panel B zooms into specific areas shown within the grey dashed line boundaries in A.](../Figures/Fig1.png) -->
 
-Changes of stock share ratio are largely related to regional changes in biogeography and the geometry of the EEZs (Figure \@ref(fig:toeFigure5)).
-In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa stock share ratio is expected to follow the projected poleward shift of transboundary species under climate change.
-However, central America Pacific and West African coasts are expected to follow an equatorial direction as range shifts of the transboundary species are following local oceanographic gradients with relatively lower temperature waters in lower latitude [@Clarke:2020ep].
-Moreover, some regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no stablished pattern, this is the case of the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).
+The direction and intensity of shafts in transboundary stocks are largely related to regional changes in biogeography and the geometry of the EEZs (Fig.
+\@ref(fig:toeFigure5)A).
+In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the stock share ratio is expected to follow a projected poleward shift.
+However, shifts along the coasts of Pacific Central America and West Africa  occur along an equatorial direction.
+Here, transboundary species follow local oceanographic processes like upwelling zones (e.g., Humboldt current) and patterns of oxygen availability, in addition to temperature gradients.
+Regions where EEZs are relatively small and have multiple borders (e.g., Caribbean or Northwestern Europe) are expected to have a particularly complex exchange of stock share ratio with no established pattern (Fig.
+\@ref(fig:toeFigure5)B).
 
-## Discussion
+<!-- ## Discussion -->
 
-In this paper, we investigate the timing and intensity associated with changes in the share distribution of transboundary stocks between countries.
+<!-- In this paper, we investigate the timing and intensity associated with changes in the share distribution of transboundary stocks between countries.
 Our findings highlight an early distributional shift from historical variations for most transboundary stocks in the world.
@@
 Thus, our findings emphasize recent calls for the urgent adoption of adaptation measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @CisnerosMata:2019ks; @Oremus:2020id; @Free:2019jq].
-
-#### High present-day climate risk of on transboundary fisheries management
-
-Our projections in the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century corroborate with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk].
-For example, in the early 2000s, Humboldt squid (Dosidicus gigas) substantially expanded its geographic range poleward, reaching the coast of Washington state (US).
-A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs].
-In the northeast Atlantic, the Atlantic mackerel (Scomber scombrus) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
-However, a range expansion of Atlantic mackerel (Scomber scombrus) into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008 without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij].
-Such changes in quota resulted in disputes between Iceland and Faroes Islands, and the NEAFC member states [@Spijkers:2017ij].
-As transboundary stocks continue to emerge in the incoming years, mal-adapted fisheries management and international disputes like that of Atlantic mackerel are likely to increase in frequency [@Pinsky:2018cb]
-
-The projected range shifts of large number of transboundary stocks emerging since the early 2000s can be partially attributed to the time of emergence of several environmental variables (e.g., temperature, oxygen) that influence the distribution of marine species [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
-Marine organisms have a thermal tolerance that limits the temperature range they could live in [@IPCC:2019tn].
-This thermal tolerance explains the sensitivity of the animal to warming and is a main driver of species’ distributional change [@IPCC:2019tn].
-Oxygen plays a similar role to that of temperature [@Clarke:2020ep], especially on tuna species that require large amounts of dissolve oxygen for their metabolism [@IPCC:2014bz].
-It has been estimated that since the early 2000s, increase in sea surface temperature (SST) and decrease in oxygen content have emerged in 79% and 23% of the global ocean, respectively, and by 2080, SST is projected to emerge in 90% of the global ocean while O2 will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq].
-Moreover, before 2100, surface acidification, nitrate content and net primary production, also important variables for species distributions, will have emerged from the historical variability of 1861-1900 in 30% and 60% of the ocean under a low (RCP 2.6) and strong (RCP 8.5) climate change scenario, respectively [@IPCC:2019tn].
-These oceanographic changes are important drivers of change in species distributions as represented in the DBEM [@Cheung:2016jd; @Cheung:0ur].
-
-Regional differences in time of emergence are attributed to the variations in environmental conditions, species’ climate vulnerability, and the characteristics of political boundaries.
-Specifically, our findings are consistent with results from previous studies that show tropical species are already shifting their distributions in the last few decades [@Poloczanska:2016kk].
-The tropics are among the most species-rich biome [@Reygondeau:2019uh] and have potentially experienced some of the highest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
-In the tropics, marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz].
-Consequently, as a result of climate change, species have to move towards deeper or sub-tropical waters [@Poloczanska:2016kk] although recent studies suggest that in some regions species are shifting towards the equator and to shallow waters to follow local temperature gradients [@Clarke:2020ep.
-Moreover, tropical species are expanding their range in sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis shows a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
-In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs that border with multiple countries (e.g., The Caribbean).
-These tropical areas are projected to have the highest number of transboundary stocks’ distribution shifts emerging in the incoming years.
-
-### Hotspot of climate risk on transboundary fisheries management
-
-Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)).
-For example, Polynesia and Micronesia are the only regions where the average time of emergence of the distribution of where all transboundary stocks across all EEZs emerged before 2022 (Figure \@ref(fig:toeFigure5)).
-Moreover, countries in these hotspot regions are highly dependent on pelagic-oceanic (e.g., tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
-On the other hand, “coldspots” of climate risk for transboundary fisheries like the Northeast Atlantic and Northwest Pacific where stocks are expected to emerge later in the 21st century.
-Assuming that these countries already have adequate transboundary management to deal with historical and current variations in fish stocks’ distribution e.g., the Pacific Salmon Treaties [@Miller:2004fpa], management agencies are likely to have more time to prepare for emerging stock shifts.
-In cases where stocks are emerging or have already emerged, adaptation policies are urgently needed to address an issue that could be impacting important fisheries [@Burden:2019ij].
-Currently, fisheries targeting tunas shared stocks in the South Pacific Islands region operate under the Vessel Day Scheme - which takes into account shifts in stocks and catches making it a good example where a management system has been set up that is somewhat anticipatory [@Aqorau:2018bh].
-However, even strategies that take into consideration the shift of the species within the Party’s area will likely require to deal with the newcomer issue that NEAFC is currently facing over Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
-As species have been emerging from their historical distribution since 2006, regional management bodies require to revise their management actions as today most of them are not equipped with strategies to cope with shifting species distributions [@Oremus:2020id; @Sumby:2021gw].
-This is of specific concern for those management bodies located in the emerging “hotspots” here identified.
-
-The presentce of EEZs lacking stocks with emerging transboundary index is a combination of EEZ shape, climate projections, and a reduced number of transbounudary stocks.
-For example, there are 59 commercial fish stocks fished the Arctic with the vast majority focused on the Barents, Bering and Norwegian seas with limited knowledge about the species biology [@Christiansen:2014gg].
-However, the Arctic has been warming faster than many other regions resulting in a reduction in sea-ice [@Douglas:2010Mb] that could potentially be occupied by species previously not present [@IPCC:2019tn].
-As a result, the biomass of Arctic marine animals [@BryndumBuchholz:2019kh] and catch and revenue of associated fisheries [@Tai:2019cq; @Lam:2014fq] are expected to increase in the incoming years, although substantial uncertainty exists on the magnitude of change [@BryndumBuchholz:2019kh].
-Moreover, projections suggest a larger increment on transboundary stocks in the incoming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further North [@Morley:2018fn; @Frainer:2017ji].
-In the Case of Brazil, in addition to having few transbounudary stocks (only 4 shared with Uruguay in this study), the large diffrence in the EEZs lenghts could also be influencing the results.
-
-<!-- For Brazil and Uruguay, the current and future distribution of the transboundary stocks in combination with the particular characteristics of these EEZs could be influencing the results.
-First, both Argentine menhaden and Argentine shortfin squid have distributions with larger biomass concentration further south, with Brazil serving as the northern limit of the distributional range [@AquaMaps:2019ui; @AquaMaps:2019wt].
-Second, Brazil’s EEZ is substantially larger than Uruguay’s resulting in a long distance between Brazil’s and the stocks’ centroids.
-Third, both stocks are expected to shift further south in the incoming years [@AquaMaps:2019ui; @AquaMaps:2019w].
-Thus, the further southern shift of Argentine menhaden and shortfin squid in combination with the reduced latitudinal gradient of Uruguay’s EEZ results in a small effect in the already long latitudinal distance between Brazil’s EEZ and the stocks’ centroids.
-In the case of the two shrimp species, their destitution range spams along the cost of Brazil and is not expected to have a strong shift in the incoming years, also resulting in a stable transboundary index [@AquaMaps:2019wn; @AquaMaps:2019wj].
 -->
 
-### Changes in stock share ratio will be ubiquitous
+**High present-day climate risk of on transboundary fisheries management**
 
-Changes in stock share ratio beyond a country’s threat point will be seen across the world, suggesting that climate change impacts on transboundary stock management is a global problem (Figure \@ref(fig:toeFigure5)).
-Poleward shifts in stock share ratio are expected in regions of the world with relatively long EEZs and few boundaries, like the northeast Pacific and the southeast Atlantic.
-However, despite observations and modelling work indicating that, generally, climate change is shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], some regions like the eastern tropical Pacific and eastern tropical Atlantic Oceans will experience a shift in distribution towards the equator.
-Specifically, in the eastern tropical Pacific, the direction of shift is expected to be from Mexico to Ecuador with EEZs losing stock share ratio to their southern neighbors.
-Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems [@Clarke:2020ep].
-They are consistent with regional modeling studies suggesting that by mid 21st century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission climate change scenario [@Clarke:2020ep].
-Moreover, regions characterized by multiple political boundaries, like the Caribbean and the northeast Atlantic, will see a complex network of changes in SSR with no clear pattern (\@ref(fig:toeFigure5)B).
-These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and larvae dispersal [@Ramesh:2019va] supporting the implementation of joint management plans [@Burden:2019ij] and the call for new implemented shared policies [@Levin:2018in].
-However, game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.
+Early shifts in the shared distribution of transboundary stocks concur with previous studies that have detected and attributed changes in marine catch composition [e.g., @Cheung:2013gk; @Frainer:2017ji; @Last:2011fd].
+For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US) as a response to climatic, oceanographic and ecological changes [@Zeidberg:2007gs].
+A new fishery targeting Humboldt squid quickly developed on the heels of the species’ range expansion [@Pinsky:2014by].
+In the northeast Atlantic, the Atlantic mackerel (*Scomber scombrus*) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
+However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the fishery’s total allowable catch and a further 18% in 2008, without consultation with NEAFC, threatening the sustainability of the stock [@Spijkers:2017ij].
+These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij].
+As transboundary stocks continue to shift, international disputes like that of Atlantic mackerel are likely to increase in frequency requiring a revision of management plans [@Pinsky:2018cb; @Engler:2020bg].
 
-Changes in stock share ratio beyond the threat point of countries sharing a common resource could instigate international conflict and hinder bi-lateral agreements (@Miller:2013iv; @Sumaila:2020tp).
-Lessons learned from historic distributional shifts of shared stocks suggest that treaties that are not prepare to respond to such shifts will be less resilient to changes in transboundary fish stocks share ratio due to climate change [@Miller:2013iv; @Sumaila:2020tp].
-For example, disputes over Pacific salmon between Canada and the USA arose in the 1990s when a climate-related shift in stock abundance favoured Alaska’s salmon fisheries over Canada’s [@Miller:2013iv; @Munro:2002uf; @Song:2017va].
-The conflict lasted over ten years until Canada and the USA agreed on a mutual conservation fund where the US would contribute the larger proportion to support scientific research, habitat restoration and enhancement of wild stock production in transboundary rivers [@Munro:2002uf].
-Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Pinsky:2018cd].
-Side payments are mechanisms that provide a solution to the inequality produced by shifts in the stock’s distribution [@Tunca:2019if; @Miller:2004fpa].
-These can be monetary (e.g., the stock-“winning” State compensates the stock-“loosing” state for the proportional shift) or other methods like the Pacific salmon case (e.g., a conservation fund payed mainly by one Party) [@Miller:2004fpa; @Miller:2013iv].
-Strengthening current international cooperation will be required as stock distribution will move to neighboring jurisdictions, requiring combine efforts to generate information and set proper management rules [@Miller:2013iv].
-Finally, flexible management rules might be better fitted to capture the stock’s distributional shift.
-For example, quota allocation based on the stock’s current distribution like that for Pacific halibut (*Hippoglossus stenolepis*) managed by Canada and the US [@Sumaila:2020tp; @PalaciosAbrantes:2020vt] might be better able to capture the stock’s shifts than a quota based on a fix-historical proportion like the case of the EU [@Baudron:2020dj].
+**Hotspot of climate risk on transboundary fisheries management**
 
-## Caveats and uncertainties
+Our results underscore regional "hotspots" of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figures \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure5)A).
+Such regions, like the Caribbean, are characterized by high levels of warming relative to natural internal variability [@IPCC:2019tn; @Hawkins:2012hc] and high vulnerability of species to warming waters [@IPCC:2014bz].
+Moreover, they represent relatively small EEZs that border multiple countries (e.g., The Caribbean).
+Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting transboundary stocks.
 
-There are two main sources of uncertainty in our analysis.
-First, our analysis relies on the combination of a single Earth system model (GFDL-ESM2M), fish and fisheries model, and climate change scenario (RCP 8.5).
-The use of multiple ESMs can capture the structural uncertainty of the models [@Frolicher:2016dg; @Lotze:2019ce], however, exercises looking at changes in fish biomass with multiple models show an overall agreement between models in terms of direction of change but variable in magnitude.
-Further research that includes multiple ESMs and species distribution models could elucidate a broader range structural uncertainty of the climate, fish and fisheries models [@Lotze:2019ce].
-Secondly, due to limited biological and spatially-specific data on stocks sub-population structure of transboundary species, our analysis uses political boundaries (EEZ) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
-On the other hand, in many EEZs, fisheries are often managed at the species level [@PalaciosAbrantes:2020tv] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca].
-However, reproducing our analysis regionally, where spatially explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
+Management plans that are not prepared to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
+Thus, focusing on these “hotspots'' is key to anticipate a potential increase in fisheries conflicts in the coming years.
+Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@Pinsky:2018cb].
+For example, countries that are Party to the Nauru Agreement (PNA) manage their fisheries under a  Vessel Day Scheme.
+Because this scheme takes into account shifts in stocks and catches, it is a good example of a management system that is somewhat anticipatory [@Aqorau:2018bh].
+However, even strategies that take into consideration  species’ shift within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+Quota allocation methods based on a stock’s current distribution [@Sumaila:2020tp; @PalaciosAbrantes:2020vt] and/or on fixed-historical proportion [@Baudron:2020dj] need to evolve to be more agile.
+Fisheries based on historic participation could move towards a dynamic method or a combination of both  [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+However, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from the fishery.
+Side payments are mechanisms that can provide a solution to the inequality produced by shifts in a stock’s distribution [@Tunca:2019if].
+These can, but do not have to, be monetary (e.g., the stock-“winning” state compensates the stock-“losing” state for the proportional shift, a conservation fund paid mainly by one Party) [@Miller:2013iv].
+
+**Potential drivers of shift and associated uncertainties**
+
+The early projected range shifts of the large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
+Specifically, Frölicher et al., 2016 estimated the time of emergence of dissolved oxygen ($O_2$ between 100-600m), natural primary production (NPP), and sea surface temperature (SST) relative to historical variability (1986-2005).
+By 2030, the SST signal will have emerged from historical variability for 70% of the global ocean under RCP 8.5 [@Frolicher:2016dg].
+Early emergence of SST will occur mainly in the tropics while the Southern Ocean and the North Atlantic are not expected to experience an emergence on the SST signal before 2100 [@Reygondeau:2020jh; @Frolicher:2016dg].
+For NPP and $O_2$, the signal is projected to emerge in 23% and 10% of the ocean by the end of the $21^st$ century, respectively under RCP 8.5 [@Frolicher:2016dg].
+Results from multiple marine ecosystem models show that SST and NPP are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
+Specifically for the DBEM, SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results [@Cheung:2016tt].
+The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz], while also explaining the later, and sometimes non-existent, shift at higher latitudes (Fig.
+\@ref(fig:toeFigure2)B).
+
+Different levels of uncertainty exist in the projected emergence of environmental variables, mainly related to *i*) climate change scenarios and  *ii*) model structure [@Frolicher:2016dg].
+First, our analysis is based on a high emission scenario (RCP 8.5) representing an extreme case of shifts in the shared distribution of transboundary species.
+Thus, any climate change mitigation pathway (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].
+Second, a substantial source of uncertainty is related to model selection for both climate change and fisheries distribution.
+This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to the emergence of environmental drivers is larger but that to scenarios is lower [@Frolicher:2016dg].
+While such uncertainty could potentially be reduced by incorporating  ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq].
+As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties, and will likely continue to do so in the coming decade.
+This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, and where internal variability (e.g., ENSO for Pacific nations) accounts for most of the uncertainty.
+Model uncertainty also affects the way we project species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh].
+Exercises looking at changes in fish biomass with multiple models, including the DBEM, show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@BryndumBuchholz:2019kh].
+Further research that includes large ensemble simulations of multiple species distribution models could elucidate a broader range in structural uncertainty of fish and fisheries models [@Lotze:2019ce].
+Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks.
+While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca].
+Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
 Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
 
-## Conclusion
+**Conclusion**
 
-Global society has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear effects on multiple others [@UnitedNations2018; @Singh:2017ds].
-Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@@Pinsky:2018cb].
+Global society has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear benefits for a number of other goals [@UnitedNations2018; @Singh:2017ds].
+Preparing anticipatory policy to shifting transboundary stocks is key to achieve the sustainable development goals and effective governance of the world’s fisheries [@Pinsky:2018cb; @Oremus:2020id; @Pecl:2017hu].
 Here, we have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
-We have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
-Our results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management.
-While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg; @PalaciosAbrantes:2020tv].
-To address the issues raised in this paper, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
+First, we have identified the transboundary stocks that will see a significant shift in their shared distribution compared to their historical average and the year in which such shifts will occur.
+Second, we have estimated the intensity of such change.
+Our results have direct implications for ocean governance as the data provided can inform regional management bodies and countries to anticipate the likely climate-driven challenges to international fisheries management.
+While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for shifting transboundary stocks [@Pinsky:2018cb; @PalaciosAbrantes:2020tv].
+Our findings emphasize recent calls for the urgent adoption of measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @CisnerosMata:2019ks; @Oremus:2020id].
 
@@
 
-```{r FigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}
+<!-- Set Figure, table and equations numbering according to the chapter -->
+\renewcommand{\thefigure}{S.\arabic{figure}}
+\setcounter{figure}{0}
+\renewcommand{\thetable}{S.\arabic{table}}
+\setcounter{table}{0}
 
-img <- readPNG("../Figures/Supp1.png")
-grid.raster(img)
+```{r comitee_form, child = '../Scripts/Supplements.Rmd', eval = T}
 
 ```
-
-```{r FigureS2, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
-The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel).
-Time of emergence is defined as the first year when the TI overpasses historical values (Bottom panel)."}
-
-img <- readPNG("../Figures/Supp2.png")
-grid.raster(img)
-
-```
-
-```{r FigureS3, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
-
-img <- readPNG("../Figures/Supp3.png")
-grid.raster(img)
-
-```
-
-```{r FigureS4, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
-
-img <- readPNG("../Figures/Supp4.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005))](../Figures/Supp2.png) -->
-
-```{r FigureS5, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial stocks that will see a shift.
-Panel B zooms in to specific areas shown in grey boxes of A."}
-
-img <- readPNG("../Figures/Supp5.png")
-grid.raster(img)
-
-```
-
-```{r TableS0, eval = T, echo = F}
-
-tibble(
-  "Habitat preference" = unique(Dem_pel$DemersPelag)[1:7],
-  Description = c(
-    "The shallow pelagic zone over the continental shelf; nearshore ocean ecosystems; i.e., those associated with the coasts because the waters are overlying continental shelves and/or the waters are < 200 m deep in areas of coastal submarine slopes",
-    "A largely consolidated wave resistant feature; the upper surface is within 0-20 m of the ocean surface; in the tropics, the upper portion is typically composed mainly of living and non-living remains of coral and coralline algae; rock lying at or near the surface and which may pose a danger to navigation",
-    "Living and feeding on the bottom below 200 m",
-    "Living and feeding near the bottom as well as in midwaters or near the surface.
-Feeding on benthic as well as free swimming organisms.
-Many freshwater fish are opportunistic feeders that forage on the bottom as well as in midwater and near the surface, also pertaining to forms which hover or swim just over the floor of the sea, e.g.
-Halosauridae, Macrouridae, Moridae, Brotulidae; the depth zone about 100 metres off the bottom at all depths below the edge of the continental shelf",
-    "Sinking to or lying on the bottom; living on or near the bottom and feeding on benthic organisms",
-    "Region of the oceanic zone between 1,000 m to 4,000 m; between the mesopelagic layer above and the abyssopelagic layer below.
-Living or feeding in open waters at depths between 1,000 and 4,000 m.
-In FishBase this term is used to include the depth range from 200 m to the bottom and thus the zones mesopelagic, bathypelagic and abyssopelagic",
-
-    "Living and feeding in the open sea; associated with the surface or middle depths of a body of water; free swimming in the seas, oceans or open waters; not in association with the bottom.
-Many pelagic fish feed on plankton.
-In FishBase, referring to surface or mid water from 0 to 200 m depth."
-  )
-) %>% 
- kable()
-
-```
-
-```{r TableS1, eval = F, echo = F}
-
-read.csv("~/GitHub/TransEmergence/Tables/kw_regional.csv") %>% 
-  select(-statistic,
-         "habitat association comparrison" = comparrison,
-         "Significan level" = signif.level,
-         "Observation difference" = dif.com.obs.dif,
-         "Critical difference" = dif.com.critical.dif,
-          "K-W Difference" = dif.com.difference) %>% 
-  kable("latex",
-        booktabs = T,
-        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  ) %>% 
-  kable_styling(full_width = F,
-                position = "center")
-
-```
-
-```{r TableS2, eval = F, echo = F}
-
-read.csv("~/GitHub/TransEmergence/Tables/kw_habitat.csv") %>% 
-  select(-statistic,
-         "Sub region comparrison" = comparrison,
-         "Significan level" = signif.level,
-         "Observation difference" = dif.com.obs.dif,
-         "Critical difference" = dif.com.critical.dif,
-          "K-W Difference" = dif.com.difference) %>% 
-  kable("latex",
-        booktabs = T,
-        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  ) %>% 
-  kable_styling(full_width = F,
-                position = "center")
-
-```
-
-<!-- # Recycled  text -->
-<!-- All of these results have direct implications for managing shared fisheries under climate change that can be partially addressed by the early implementation of joint-management international policies.
--->
-
-<!-- The time and number of emerging species decreases as we move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
--->
-
-<!-- Changes in stock share ratio has contribute to past conflicts between nations with long history of collaboration including Canada and the United States over Pacific salmon and Norway and Iceland over spring spawning herring [@Miller:2013iv].
-In both cases, rules were emplaced and agreed by all parties, however, treaties were not prepared to address shifts in species distributions [@Miller:2013iv; @Song:2017va] and when the benefits felt below the desired level (e.g.
-threat point) the affected country broke off relationships.
-In the case of Canada and the US, the agreement was re-stated when a conservation fund was implemented [@Miller:2013iv].
--->
-
-<!-- In cases where the species is distributed in more than 2 neighboring EEZs (e.g. *xxx*), the index is as follows: -->
-
-<!-- $$I =(\frac{D_{A,y}}{sd(D_{A,h))} - \frac{D_{B,y}}{sd(D_{B,h)))) |  \\  +(\frac{D_{A,y}}{sd(D_{A,h))} - \frac{D_{C,y}}{sd(D_{B,h)))) | \\  +(\frac{D_{A,y}}{sd(D_{A,h))} - \frac{D_{Cy}}{sd(D_{B,h))))/N$$ -->
-
-<!-- In other words, for two neighboring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases (see *TI* equation).
-Consequently, if the centroid shifts are within historical variability (*SNR* < 1), risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability (*SNR* > 1), the risk of dispute is higher.
-Thus, risk is positively related to the frequency of centroid shifts beyond historical variability.
--->

### 2021-03-10 Scripts/Initial_Analysis.Rmd
misc

@@
   ) +
-  my_ggtheme_p() +
-  theme(legend.position = "right") +
+  my_ggtheme_p(leg_pos = "right") +
   scale_fill_viridis(discrete = T, 

### 2021-03-16 Scripts/Manuscript.Rmd
Updated manuscript with title and Alessandro's comments

@@
 ---
-title: Identifying timing of shifts in transboundary fish stocks to support climate resilient fisheries management
-
+title: "Early range shifts in transboundary fish stocks challenge their management under climate change"
 author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
@@
 ^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States'
-output:
-  bookdown::word_document2
-reference_docx:  ../References/manuscript_template.docx
+output: bookdown::word_document2
+reference_docx:  ../References/sci_template.docx
   # pdf_document: default
@@
 csl: ../References/science.csl
-header-includes:
-- \usepackage{amsmath}
-- \usepackage{float} # Use the 'float' package
-- \floatplacement{figure}{H} # Make every figure with caption = h
 ---
@@
 
-Early range shifts in transboundary fish stocks challenge their management under climate change
+Rapid climate-induced shifts in transboundary fish stocks
 
@@
 
-Different levels of uncertainty exist in the projected emergence of environmental variables, mainly related to *i*) climate change scenarios and  *ii*) model structure [@Frolicher:2016dg].
+Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and  *ii*) model structure [@Frolicher:2016dg].
 First, our analysis is based on a high emission scenario (RCP 8.5) representing an extreme case of shifts in the shared distribution of transboundary species.
@@
 This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to the emergence of environmental drivers is larger but that to scenarios is lower [@Frolicher:2016dg].
-While such uncertainty could potentially be reduced by incorporating  ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq].
+While such uncertainty could potentially be reduced by incorporating  ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] and they ignore other sources of structural uncertainty based on the parameterisation of processes.
 As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties, and will likely continue to do so in the coming decade.
-This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, and where internal variability (e.g., ENSO for Pacific nations) accounts for most of the uncertainty.
-Model uncertainty also affects the way we project species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh].
-Exercises looking at changes in fish biomass with multiple models, including the DBEM, show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@BryndumBuchholz:2019kh].
-Further research that includes large ensemble simulations of multiple species distribution models could elucidate a broader range in structural uncertainty of fish and fisheries models [@Lotze:2019ce].
+This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf] and where internal variability (e.g., El Niño Southern Oscillation - ENSO- for Pacific nations) is also important .
+These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh].
+Overall, changes in fish biomass with multiple upper trophic level models, including the DBEM, show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@BryndumBuchholz:2019kh].
+Further research that reduces uncertainty in the response of the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could elucidate a broader range in structural uncertainty of fish and fisheries models [@Lotze:2019ce].
 Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks.

### 2021-03-17 Scripts/Manuscript.Rmd
Ale. Thom. and Rash. Comments

@@
 ---
-title: "Early range shifts in transboundary fish stocks challenge their management under climate change"
-author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
+title: "Imminent range shifts in transboundary fish stocks challenge their management under climate change"
+author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
 date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
@@
 
--   corresponding author: [j.palacios\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}
+corresponding author: [j.palacios\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}
 
@@
 Climate change is shifting the distribution of transboundary fish stocks between neighbouring countries.
-The time-frame of such shifts determine the extent of climate challenges to international fisheries governance.
-Here, using global climate-fisheries simulation modelling, we show that, by 2036, 68% of the world’s transboundary stocks are projected to shift their distributions significantly across international boundaries.
+The time-frame of such shifts determine the extent of climate change challenges to international fisheries governance.
+Using global climate-fisheries model simulations, we show that, within the next 15 years, 68% of the world’s transboundary fish stocks are projected to shift their distributions across international boundaries.
 By 2100, such shifts will occur in all countries under a ‘no mitigation’ scenario.
-Hotspots where transboundary stock shifts are earlier and larger coincide with countries that are highly dependent on fisheries for income and livelihood.
-The early occurrences of transboundary shifts highlight the urgency to re-formulate international fisheries policies to avoid disputes in fish stock sharing.
-Such bi-/multi-lateral discussions should start now as lead-time for policy changes is long.
+Hotspots of transboundary shifts occur in countries that are highly dependent on fisheries for income and livelihood.
+The imminent onset of transboundary shifts highlight the urgency to re-formulate international fisheries policies to avoid disputes over fish stock.
+Bi- and multi-lateral actions discussions should start now as lead-time for policy changes is long.
 
@@
 
-Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters, [@IPCC:2019tn].
-Marine species' distributions reflect species' preferences for discrete environmental conditions [@Hutchinson:1957nt; @Nelson:2016kk].
-As a result of climate change, many marine species have shifted their distributions towards higher latitude, deeper water or following local temperature gradients to track and stay within their optimal environmental 'niche'[@Poloczanska:2016kk; @Barange:2014kr].
+In the past century, human activities have altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters [@IPCC:2019tn].
+Marine species' distributions reflect species' preferences for discrete environmental conditions [@Hutchinson:1957nt].As a result of these past changes, many marine species have shifted their distributions towards higher latitudes, deeper water or to follow local temperature gradients to remain within their optimal environmental "niche" [@Poloczanska:2016kk; @Barange:2014kr].
 The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt; @Barange:2014kr], impacting fisheries production and economies [@Sumaila:2019ew], and compromising our capacity to reach international sustainability goals within the 2030 Agenda such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
-The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management [@Cheung:2018dh], including for species that cross international borders, i.e., shared stocks [@Pinsky:2018cb].
+The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management [@Cheung:2018dh], including for species that cross international borders, which are known as ‘shared stocks’ [@Pinsky:2018cb].
 
 The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by Coastal States [@UnitedNations:1986tl].
-As defined by the United Nations' Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (*iv*) discrete stocks that are only present on the high seas [@Munro:2004th].
+As defined by the United Nations' Food and Agriculture Organization, shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and (*iv*) discrete stocks that are only present in the high seas [@Munro:2004th].
 This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
@@
 In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
-While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017iua; @Spijkers:2017ij; @Crespo:2020cs].
+While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Song:2017iua; @Spijkers:2017ij; @Crespo:2020cs].
 Such patterns are expected to be exacerbated by continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
-The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the SDGs.
-Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important to developing climate resilient international ocean governance and achieving the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].
+The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the  Sustainable Development Goals (SDG) under the United Nations 2030 Agenda.
+Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].
 
-Here, we estimate the year in which 663 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variation and the intensity of said shift.
-We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
-We employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks' distribution across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (RCP8.5; Methods).
-We find that the shared distribution of many transboundary stocks has already shifted beyond historical natural variability.
-Moreover, by 2030, the shared catch proportion of such stocks will exceed historical levels in most neighboring EEZs, posing a threat to the sustainability of associated fisheries and the resilience of international treaties that manage their shared exploitation.
+Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (RCP8.5; Methods).
+We then estimate the year in which the 663 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variation and the intensity of this shift (i.e., time of emergence; see Methods).
+We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (15) and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
 
@@
 
-We calculated the time of emergence of the transboundary index to identify shifts in the shared distribution of transboundary stocks from 2006 to 2100 under a high emission climate change scenario.
 Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability as early as 2006 and up to 2100 (Fig.
@@
 This corresponds to 68% of the studied stocks (Fig.
-\@ref(fig:toeFigure1)C).
-On average, the shift in shared distribution of these stocks in all EEZs analyzed is projected to be 2036 $\pm$ 28 years.
-Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% presenting an early distributional shift across stocks (Fig.
+\@ref(fig:toeFigure1)C).The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be 2036 (s.d.
+of all species, $\pm$ 28 years).
+Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% presenting a distributional shift across stocks by the time the UN 2030 Agenda is intended to be achieved (Fig.
 \@ref(fig:toeFigure1)C).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary stock shared distribution.
-Dashed lines represent the year by which countries have committed to reach the full implementation of the 2030 Agenda, as well as an early and late change in range shift, respectively."}
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary shared stock distribution.
+Dashed lines represent the year by which countries have committed to reach the full implementation of the 2030 Agenda, as well as a mid and late change in range shift, respectively."}
 
@@
 The median year in which transboundary stocks experience a range shift varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Figure \@ref(fig:toeFigure2)A).
-Overall, most tropical EEZs will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia being significantly earlier (*p* < 0.05; see Table \@ref(tab:toeTableS2) for test statistics) than almost all other regions (Fig.
+Overall, most tropical EEZs will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia being significantly earlier (*p* < 0.05; see Table S\@ref(tab:toeTableS2) for test statistics) than almost all other regions (Fig.
 \@ref(fig:toeFigure2)B).
-In contrast, EEZs located in temperate regions like northern Europe and eastern Asia are projected to experience significantly later range shifts (Table \@ref(tab:toeTableS2)).
-Stocks in some countries, like Brazil or New Zealand, are not projected to experience range shifts beyond natural variability (e.g., the transboundary index of a stock did not emerge from historical variability under a 95% confidence interval) (Fig.
+In contrast, EEZs located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts (Table \@ref(tab:toeTableS2)).Stocks in countries such as Brazil and New Zealand are not projected to experience range shifts beyond natural variability (e.g., the transboundary index of a stock did not change beyond historical variability) (Fig.
 \@ref(fig:toeFigure2)A).
-On average, shifting transboundary stocks represented 23% $\pm$ 22% of yearly fishing revenue from fisheries targeting such stocks within global EEZs between 2004 and 2010 (Fig.
+On average, shifting transboundary stocks represented 23% (s.d.
+$\pm$ 22%) of yearly fishing revenue from fisheries targeting such stocks within global EEZs between 2004 and 2010 (Fig.
 \@ref(fig:toeFigure2)A).
 However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
-Moreover, in some EEZs, while few transboundary stocks are projected to shift above historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).
+Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).
 
@@
 A) Land polygon shows the contribution of shifting stocks to the country’s total fishing revenue from transboundary stocks.
-Exclusive Economic Zone polygons represent average range shifts per EEZ.
-EEZs with no time of emergence between 2006 and 2100 are represented in aqua color.
+Exclusive Economic Zone polygons represent average range shifts within them.
+EEZs with no distributional shifts between 2006 and 2100 are represented in aqua color.
 B) Year of shared distribution shifts by region according to United Nations sub-regions.
 Points represent the median year of range shift.
-Horizontal dashed line represents the 2030 Agenda.
+Horizontal dashed line represents the year 2030.
 N = North, S = South, W = West and E = East.
@@
 We estimated the current (2005 and 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs.
-Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD in fishing revenue are projected to shift within the EEZs of 76 counties (Fig.
+Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD in fishing revenue are projected to shift within the EEZs of 76 countries (Fig.
 \@ref(fig:toeFigure3)); with 4.2 billion USDs accounted by 321 stocks shifting in just eleven countries (Fig.
@@
 
-```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks per fishing entity.
+```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks per country.
 Countries are coded according to different colored symbols representative of 5 distinct regions.
@@
 ```
-
-We compared the year in which transboundary stocks experience a range shift by species’ preferred habitat association (Fig.
-\@ref(fig:toeFigure4)).
-We found significant differences in the range shift year relative to habitat association (Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; See table S3 for full statistics \@ref(tab:TabelS3)).
-The average range shift of bathy-pelagic species (e.g., Blue whiting *Micromesistius poutassou*), pelagic-oceanic (e.g., tunas) and reef species (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to happen before 2030 while bathydemersal species (e.g., Banguela hake, *Merluccius polli*) are expected to experience the latest shift (2040).
-
-```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
-Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
-The category “Other” consists of species that have no “habitat preference” classification.
-The number of stocks included in this analysis for each habitat type is noted in parentheses.
-Dashed whiskers represent the 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Solid line represents the median time of emergence per habitat preference."}
-
-img <- readPNG("../Figures/Fig4.png")
-grid.raster(img)
-
-```
-
-<!-- ![toeFigure 4.
-Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
-Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
-The category “Other” consists of species that have no “habitat preference” classification.
-The number of stocks included in this analysis for each habitat type is noted in parentheses.
-Dashed whiskers represent the 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Solid line represents the median time of emergence per habitat preference.](../Figures/Fig5.png) -->
 
@@
 
-We identified transboundary stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point by the early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) under a high emission climate change scenario.
+We identified transboundary stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point by the early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005).
 Findings show that by the early $21^{st}$ century, the global average catch proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 59% $\pm$ 17% (Fig.
@@
 Such regions, like the Caribbean, are characterized by high levels of warming relative to natural internal variability [@IPCC:2019tn; @Hawkins:2012hc] and high vulnerability of species to warming waters [@IPCC:2014bz].
-Moreover, they represent relatively small EEZs that border multiple countries (e.g., The Caribbean).
+Moreover, they represent relatively small EEZs that border multiple countries (e.g., the Caribbean).
 Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting transboundary stocks.
@@
 
-The early projected range shifts of the large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
-Specifically, Frölicher et al., 2016 estimated the time of emergence of dissolved oxygen ($O_2$ between 100-600m), natural primary production (NPP), and sea surface temperature (SST) relative to historical variability (1986-2005).
-By 2030, the SST signal will have emerged from historical variability for 70% of the global ocean under RCP 8.5 [@Frolicher:2016dg].
-Early emergence of SST will occur mainly in the tropics while the Southern Ocean and the North Atlantic are not expected to experience an emergence on the SST signal before 2100 [@Reygondeau:2020jh; @Frolicher:2016dg].
-For NPP and $O_2$, the signal is projected to emerge in 23% and 10% of the ocean by the end of the $21^st$ century, respectively under RCP 8.5 [@Frolicher:2016dg].
+The early projected range shifts of a large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability (e.g., SST which is projected to increase historical variability by 2030 in 70% of the global ocean under RCP 8.5 [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
 Results from multiple marine ecosystem models show that SST and NPP are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
@@
 Second, a substantial source of uncertainty is related to model selection for both climate change and fisheries distribution.
-This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to the emergence of environmental drivers is larger but that to scenarios is lower [@Frolicher:2016dg].
-While such uncertainty could potentially be reduced by incorporating  ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] and they ignore other sources of structural uncertainty based on the parameterisation of processes.
+.
+This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
+While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available **(Deser et al.
+2020)** and cannot address limitations of process level understanding.
 As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties, and will likely continue to do so in the coming decade.
@@
 Overall, changes in fish biomass with multiple upper trophic level models, including the DBEM, show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@BryndumBuchholz:2019kh].
-Further research that reduces uncertainty in the response of the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could elucidate a broader range in structural uncertainty of fish and fisheries models [@Lotze:2019ce].
+Further research that reduces uncertainty in the response of the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to a larger structural uncertainty of fish and fisheries models [@Lotze:2019ce].
 Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks.
@@
 
-Global society has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear benefits for a number of other goals [@UnitedNations2018; @Singh:2017ds].
-Preparing anticipatory policy to shifting transboundary stocks is key to achieve the sustainable development goals and effective governance of the world’s fisheries [@Pinsky:2018cb; @Oremus:2020id; @Pecl:2017hu].
-Here, we have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
-First, we have identified the transboundary stocks that will see a significant shift in their shared distribution compared to their historical average and the year in which such shifts will occur.
-Second, we have estimated the intensity of such change.
-Our results have direct implications for ocean governance as the data provided can inform regional management bodies and countries to anticipate the likely climate-driven challenges to international fisheries management.
+Th global community has set the ambitious goal of managing all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear benefits for a number of other societal goals [@UnitedNations2018; @Singh:2017ds].
+Preparing anticipatory policies to deal with shifting transboundary stocks is key to achieving the SDGs and ensuring effective governance of the world’s fisheries [@Pinsky:2018cb; @Oremus:2020id; @Pecl:2017hu].
+In this contribution, we implemented two steps to demonstrate how to achieve sustainable transboundary fisheries in a changing world [@Link:2010ei].
+First, we identified the transboundary stocks that would likely see shifts in their shared distribution compared to their historical average and the year in which such shifts are predicted to occur.
+Second, we estimated the intensity of such change.
+Our results have direct implications for ocean governance as the data provided can inform regional management bodies and member countries to anticipate the likely climate-driven challenges to transboundary fisheries management.
 While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for shifting transboundary stocks [@Pinsky:2018cb; @PalaciosAbrantes:2020tv].

### 2021-03-17 Scripts/Initial_Analysis.Rmd
FIugres modiff.

@@
 dashed_lines <- tibble(
-  x = c(2030,2050,2080),
-  label = c("UN 2030 Agenda","Early range shift  ","Late range shift   ")
+  x = c(2030,2050,2070),
+  label = c("UN 2030 Agenda","Mid range shift   ","Late range shift   ")
 )
@@
                      limits = c(2005,2100),
-                     # label = c(seq(2000,2075,25),">2100"),
-                     expand = c(0, 0)
+                     label = c(seq(2000,2075,25),">2100")
   ) +
@@
                      limits = c(2005,2100),
-                     # label = c(seq(2000,2075,25),">2100"),
-                     expand = c(0, 0)
+                     label = c(seq(2000,2075,25),">2100")
   ) +
   my_ggtheme_p(leg_pos = "right") +
-  theme(
-    strip.text.x = element_blank()
-  ) +
   guides(fill=guide_legend(title= "", reverse=FALSE)) +
@@
     # Latin America and the Car.
-    draw_plot(count_stock_plot, x = 0.01, y = 0, width = 0.28, height = 0.90) +
+    draw_plot(count_stock_plot, x = 0.01, y = 0, width = 0.29, height = 0.90) +
     # Europe
-    draw_plot(count_eez_plot, x = 0.28, y = 0, width = 0.28, height = 0.90) +
+    draw_plot(count_eez_plot, x = 0.31, y = 0, width = 0.29, height = 0.90) +
     # Percentage plot
-    draw_plot(percentage_plot, x = 0.56, y = 0 , width = 0.28, height = 0.90) +
+    draw_plot(percentage_plot, x = 0.61, y = 0 , width = 0.29, height = 0.90) +
     # legend
-    draw_plot(legend, x = 0.77, y = 0 , width = 0.28, height = 1) +
+    draw_plot(legend, x = 0.84, y = 0 , width = 0.20, height = 1) +
     # Labels
@@
 # -------------------------- #
+tresh_list <- c("tresh_one","tresh_two")
 for(i in 1:2){
 
-  tresh_list <- c("tresh_one","tresh_two")
   name_save <- paste("./Figures/emer_map_lat_",tresh_list[i],".png",sep = "")
@@
 
+  # Weighted verion
   boxplot_data <- Eme_Res %>% 
+    # Get the order of boxes
     filter(tresh == tresh_list[i],
@@
     group_by(un_sub_region) %>%
-    summarise(order = median(emerging_yr)) %>% 
+    summarise(n_eezs = length(unique(sf_eez_name))) %>%
+    # Get the rest of the data
     left_join(Eme_Res) %>%
@@
            !is.na(un_sub_region)) %>%
+    mutate(weight = )
     mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
@@
                                                 paste(un_sub_region)))
-    )
-    )
+    )#,
+    # For number labels
+    # un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
+    )
+
+  # Non-weighted verion
+
+  # boxplot_data <- Eme_Res %>% 
+  #   # Get the order of boxes
+  #   filter(tresh == tresh_list[i],
+  #          !is.na(un_sub_region)) %>%
+  #   group_by(un_sub_region) %>% 
+  #   summarise(order = median(emerging_yr),
+  #             n_eezs = length(unique(sf_eez_name))) %>% 
+  #   # Get the rest of the data
+  #   left_join(Eme_Res) %>% 
+  #   filter(tresh == tresh_list[i],
+  #          !is.na(un_sub_region)) %>% 
+  #   mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
+  #                                 ifelse(str_detect(un_sub_region,"Australia"),"Aus. and New Z.",
+  #                                        ifelse(str_detect(un_sub_region,"Sub-Saharan"),"Sub-Saharan\nAfrica",
+  #                                               paste(un_sub_region)))
+  #   ),
+  #   # For number labels
+  #   un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
+  #   )
 
@@
       legend.position = ""
-    )
+    );boxplot
 

### 2021-03-19 Scripts/Initial_Analysis.Rmd
Changes for Sci. submission

@@
 dashed_lines <- tibble(
-  x = c(2030,2050,2080),
-  label = c("UN 2030 Agenda","Early range shift  ","Late range shift   ")
+  x = c(2030,2050,2070),
+  label = c("UN 2030 Agenda","Mid range shift  ","Late range shift   ")
 )
@@
 Labels_data <- tibble(
-  x = c(2030,2045,2055,2007),
+  x = c(2030,2050,2070,2007),
   y = c(-5.8,-5.8,-5.8,Percentile_value$x+.1),
-  label = c("UN 2030 Agenda", "Early shift","Late shift","Fisheries > 75th percentile")
+  label = c("UN 2030 Agenda", "Mid shift","Late shift","Fisheries > 75th percentile")
 )
@@
   # Divisory lines and labels
-  geom_vline(xintercept = c(2030,2050), color = "grey50", size = 0.5, linetype = "dashed") +
+  geom_vline(xintercept = c(2030,2050,2070), color = "grey50", size = 0.5, linetype = "dashed") +
   geom_hline(data = Percentile_value,
@@
   scale_size("Number of\nshifting stocks") +
-  my_ggtheme_p(legend.position = "right") +
+  my_ggtheme_p(leg_pos = "right") +
   ggsave("./Figures/bubble_plot.png",

### 2021-03-19 Scripts/Manuscript.Rmd
Changes for Sci. submission

@@
 In the past century, human activities have altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters [@IPCC:2019tn].
-Marine species' distributions reflect species' preferences for discrete environmental conditions [@Hutchinson:1957nt].As a result of these past changes, many marine species have shifted their distributions towards higher latitudes, deeper water or to follow local temperature gradients to remain within their optimal environmental "niche" [@Poloczanska:2016kk; @Barange:2014kr].
-The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt; @Barange:2014kr], impacting fisheries production and economies [@Sumaila:2019ew], and compromising our capacity to reach international sustainability goals within the 2030 Agenda such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
-The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management [@Cheung:2018dh], including for species that cross international borders, which are known as ‘shared stocks’ [@Pinsky:2018cb].
+Marine species' distributions reflect species' preferences for discrete environmental conditions [@Hutchinson:1957nt].As a result of these past changes, many marine species have shifted their distributions towards higher latitudes, deeper water or to follow local temperature gradients to remain within their optimal environmental "niche" [@Poloczanska:2016kk].
+The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt], impacting fisheries production and economies [@Sumaila:2019ew], and compromising our capacity to reach international sustainability goals within the 2030 Agenda such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
+The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management, including for species that cross international borders, which are known as ‘shared stocks’ [@Pinsky:2018cb].
 
@@
 The effectiveness of fisheries management for transboundary species, and thus reaching the 2030 Agenda [@Pecl:2017hu], is challenged by shifts in stocks' distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].
-In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
-While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Song:2017iua; @Spijkers:2017ij; @Crespo:2020cs].
+In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs].
 Such patterns are expected to be exacerbated by continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
@@
 
-```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks per country.
+```{r toeFigure3, eval = T,  echo = F, fig.width=10, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks per country.
 Countries are coded according to different colored symbols representative of 5 distinct regions.
@@
 Our results underscore regional "hotspots" of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figures \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure5)A).
-Such regions, like the Caribbean, are characterized by high levels of warming relative to natural internal variability [@IPCC:2019tn; @Hawkins:2012hc] and high vulnerability of species to warming waters [@IPCC:2014bz].
+Such regions, like the Caribbean, are characterized by high levels of warming relative to natural internal variability [@IPCC:2019tn; @Hawkins:2012hc] and high vulnerability of species to warming waters [@IPCC:2019tn].
 Moreover, they represent relatively small EEZs that border multiple countries (e.g., the Caribbean).
@@
 However, even strategies that take into consideration  species’ shift within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
-Quota allocation methods based on a stock’s current distribution [@Sumaila:2020tp; @PalaciosAbrantes:2020vt] and/or on fixed-historical proportion [@Baudron:2020dj] need to evolve to be more agile.
+Quota allocation methods based on a stock’s current distribution and/or on fixed-historical proportion need to evolve to be more agile [@Sumaila:2020tp; @PalaciosAbrantes:2020vt].
 Fisheries based on historic participation could move towards a dynamic method or a combination of both  [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
@@
 Results from multiple marine ecosystem models show that SST and NPP are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
-Specifically for the DBEM, SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results [@Cheung:2016tt].
-The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz], while also explaining the later, and sometimes non-existent, shift at higher latitudes (Fig.
+Specifically for the DBEM, SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results.
+The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2019tn], while also explaining the later, and sometimes non-existent, shift at higher latitudes (Fig.
 \@ref(fig:toeFigure2)B).
@@
 Second, a substantial source of uncertainty is related to model selection for both climate change and fisheries distribution.
-.
 This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
-While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available **(Deser et al.
-2020)** and cannot address limitations of process level understanding.
+While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available [@Deser:2020bj] and cannot address limitations of process level understanding.
 As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties, and will likely continue to do so in the coming decade.
@@
 While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for shifting transboundary stocks [@Pinsky:2018cb; @PalaciosAbrantes:2020tv].
-Our findings emphasize recent calls for the urgent adoption of measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @CisnerosMata:2019ks; @Oremus:2020id].
+Our findings emphasize recent calls for the urgent adoption of measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @Oremus:2020id].
 
@@
 
-```{r comitee_form, child = '../Scripts/Supplements.Rmd', eval = T}
+```{r supplements, child = '../Scripts/Supplements.Rmd', eval = T}
 

### 2021-03-19 Scripts/Supplements.Rmd
Changes for Sci. submission

@@
 
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
+Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
+The category “Other” consists of species that have no “habitat preference” classification.
+The number of stocks included in this analysis for each habitat type is noted in parentheses.
+Dashed whiskers represent the 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Solid line represents the median time of emergence per habitat preference."}
+
+img <- readPNG("../Figures/Fig4.png")
+grid.raster(img)
+
+```
+
+<!-- ![toeFigure 4.
+Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
+Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
+The category “Other” consists of species that have no “habitat preference” classification.
+The number of stocks included in this analysis for each habitat type is noted in parentheses.
+Dashed whiskers represent the 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Solid line represents the median time of emergence per habitat preference.](../Figures/Fig5.png) -->
+
 ```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap ="World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their centroids (points)."}

### 2021-03-19 Scripts/Manuscript.Rmd
grammar comments

@@
 ---
-title: "Imminent range shifts in transboundary fish stocks challenge their management under climate change"
-author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
-date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
-^2^Climate and Environmental Physics, Physics Institute, University of Bern.
-^3^Oeschger Centre for Climate Change Research, University of Bern, Switzerland.
-^4^School of Environmental Sciences, University of Liverpool.
-^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States'
-output: bookdown::word_document2
-reference_docx:  ../References/sci_template.docx
-  # pdf_document: default
+output:
+  bookdown::pdf_book:
+    toc : no
+    keep_tex: no
+    number_sections: yes
+geometry: margin=1in
+editor_options:
+  chunk_output_type: inline
 bibliography: ../References/ToE_Ref.bib
-csl: ../References/science.csl
+header-includes:
+- \usepackage{amsmath}
+- \usepackage{float} # Use the 'float' package
+- \floatplacement{figure}{H} # Make every figure with caption = h
 ---
-
-<!-- # Instructions  -->
-
-<!-- Research articles should present a major advance and can be short or long (maximum 15,000 words) format.
-Submissions must include an abstract (up to 150 words), an introduction, and sections with brief informative subheadings.
-You may include up to ten figures and/or tables and about 60 references.
-Supplementary materials should be limited to information that is not essential for the general understanding of the research presented in the main text and can include data sets, figures, tables, videos, or audio files.
-You may include up to a total of ten figures and/or tables (combined) throughout the supplemental text.
--->
 
@@
   "grid",
-  "knitr",
- "kableExtra"
-  # "here", # for dbem_import `here()`
-  # "data.table", #dbem_import `fread()`
-  # "readxl", # for reading excell files
-  # "janitor", # for clearing names
-  # "geosphere", # estimate distances between points `distm()`
-  # "ggrepel",
-  # "zoo", # for average mean
-  # "parallel", # for mclapply,
-  # "sf", # for mapping
-  # "st", # for mapping
-  # "rgdal", #Spatial analysis
-  # "tools", #Spatial analysis 
-  # "zeallot", # for Juanito's map
-  # "gmt", # for estimating distances between points
-  # "viridis"
+  "knitr"#,
 )
@@
 
-corresponding author: [j.palacios\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}
+## Introduction
 
-**Abstract**
+Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn].
+Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk]).
+As a result of climate change, many marine species have changed their distributions towards higher latitude, deeper water or have followed local temperature gradients [@Poloczanska:2016kk].
+Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds].
+The projected risks of impacts can be reduced through improving the effectiveness of current fisheries management [@Gaines:2018sg; @Cheung:2018dh], including the fisheries management of species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].
 
-<!-- An opening sentence that states the question/problem addressed by the research AND Enough background content to give context to the study AND A brief statement of primary results AND A short concluding sentence.
+The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl].
+As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (i) transboundary stocks, those that cross neighboring EEZs; (ii) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; (iii) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (iv) discrete stocks that are only present in the high seas [@Munro:2004th].
+This chapter focuses on transboundary stocks exploited by fisheries operating within EEZs.
+While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
+In the previous chapter, I estimated that there are 633 transboundary fish species globally, representing 67% or the identified fished taxa, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue (between 2005 and 2010).
+
+The effectiveness of fisheries management for transboundary stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by].
+In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the stocks [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@Sumby:2021gw, Chapter 2].
+Misalignment between fisheries resources allocation and stocks’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb].
+With future shifts in species distributions under climate change, the challenges for the  management of fisheries targeting transboundary stocks will continue to increase during the $21^{st}$ century.
+It is important to constrain when climate change will affect the dynamics of transboundary  stocks and the intensity of the resulting impacts in order to prepare ocean governance to meet the resource management challenges arising from shifting transboundary stocks [@Pinsky:2018cb; @Link:2010ei, Chapter 2].
+
+The timing and intensity of climate impacts on managing transboundary fish stocks can be quantitatively examined through the concepts of ‘time of emergence’ (ToE) and ‘threat point’ (TP).
+Time of emergence is defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (e.g., historical natural variation) [@Hawkins:2012hc] and has been applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
+The premise behind the time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
+The concept of threat point comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
+Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary stocks between Canada and the United States [@Sumaila:2020tp].
+
+Here, I estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide.
+I employ a species distribution model driven by outputs from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species (Chapter 2) across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP 8.5 (see Methods).
+I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (Chapter 2) and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude the high seas).
+I find that the distribution of many transboundary stocks has already shifted from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.
+
+## Materials and methods
+
+### Databases and species selection
+
+This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 (Chapter 2).
+Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd, Chapter 2], resulting in a total of 9,132 transboundary stocks in this analysis.
+Using the transboundary species Atlantic cod (*Gadus morhua*) as an example, and based on the definition of transboundary stock in this chapter, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species.
+I defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1)) including island territories.
+I determined the intersections between polygons using the R statistical software package sf [@PackagesfSimpleF:2018vp].
+Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome location [@Reygondeau:2019uh].
+The habitat association of each species was determined following the classification in FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org).
+For each stock and EEZ, I used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs.
+I report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+The monthly average US Consumer Price Index (CPI) according to the U.S.
+Bureau of Labor Statistic (https://www.bls.gov/cpi/) was used to standardize all monetary values to 2019 real USD.
+
+### Projecting species distributions under climate change
+
+I projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd].
+The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
+For pelagic species, the model uses environmental variables at the surface whereas demersal species distribution is driven by ocean bottom variables.
+The DBEM then estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid (see Cheung et al.
+[-@Cheung:2010dt; -@Cheung:2016jd] for model details).
+Importantly, DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by EEZs that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
+*in revision*].
+
+The DBEM was forced with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+The GFDL-ESM2M was run under historical forcing until 2005 and followed the high greenhouse gas emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
+Because the main approach of this chapter relies on understanding the spatial and temporal variation of a stocks’ distribution, I have to understand distribution variability during both the historical and the future periods, to infer differences between time frames.
+I do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
+Thus, for my experiment, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in year 1950 resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
+By design, variations among ensemble members are then solely due to natural variability.
+This approach allows us to estimate the natural internal variability (e.g., background noise) and isolate the forced anthropogenic climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
+
+### Calculating an index of transboundary range shift
+
+I developed an index to evaluate range shifts in transboundary stocks under climate change (TI).
+This index was based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2)).
+The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., Pacific halibut (*Hippoglossus stenolepis*) that is transboundary between the United States and Canada.
+Therefore,
+
+<!-- $lat_{ts} = \mu (lat_{per})$ -->
+
+<!-- $lon_{ts} = \mu (lon_{per})$ -->
+
+\begin{equation}
+\begin{split}
+\label{eq:one}
+lat_{ts} = \mu (lat_{per}) \\
+lon_{ts} = \mu (lon_{per})
+\end{split}
+\end{equation}
+
+where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
+To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, I included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top 95th percentile (e.g., per = 95%).
+A sensitivity analysis was ran using a subset of species (n = 34) for all EEZs under different abundance levels (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)).
+The centroid of each EEZ was estimated using the st package in R (Figure \@ref(fig:toeFigureS1)).
+For each ensemble member, neighboring EEZs and transboundary stock, I computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;
+
+<!-- $$D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))$$ -->
+
+\begin{equation}
+\label{eq:two}
+D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
+\end{equation}
+
+where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively and $lon_{eez}$ and $lon_{ts}$ are the respective longitudes.
+Then, for each year I calculated the transboundary index as follows:
+
+<!-- $$TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2$$ -->
+
+\begin{equation}
+\label{eq:three}
+TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
+\end{equation}
+
+where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the pair of neighbouring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+Thus, a higher TI means that a shared stock becomes more or less abundant relative to its neighbouring EEZs that is beyond what the fisheries experienced historically.
+From the perspective of a State’s fisheries, higher TI indicates an increase in the challenges to effective management and sustainability of the shared stock’ fisheries.
+In contrast, a smaller TI suggests that the sharing of the stocks between neighbouring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the shared stocks’ fisheries.
+The index was smoothed to a 10-year average to reduce interannual variability.
+
+#### Calculating the time of emergence of the transboundary index
+
+Knowing the point in time (e.g., year) at which the distribution of a shared stock will diverge from its natural internal variability is important for informing the lead-time to which climate adaptation needs to be implemented [@Link:2010ei].
+Here, I defined time of emergence as the time when the stock shared between neighbouring EEZs exceeded the historical natural internal variability as follows:
+
+<!-- $EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$ -->
+<!-- $ToE = t when Emergence index[t] > 1$ -->
+
+\begin{equation}
+\begin{split}
+\label{eq:four}
+EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}} \\
+ToE = ToE_{t} > 1 s.d.
+\end{split}
+\end{equation}
+
+where $\mu_{TI}$ is the yearly mean TI ($\mu$) across all ten ensemble members in year *t* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical (*th*) reference period (1951 - 2005).
+I set a threshold of TI above which I considered a shared stock to have emerged from the historical internal variability of the TI.
+The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four)), representing a probability of 68% that the transboundary index has emerged from historical variability).
+I tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).
+
+<!-- #### Computing a transboundary fisheries vulnerability index (VI) -->
+
+<!-- I created a country-specific vulnerability index (VI) to examine the risk of impacts of species’ range shifts on transboundary fisheries management.
+The VI reflects both the rate of change of shared stocks’ distribution (indicated by the time-of-emergence of the transboundary index) and the importance of the shifting species to a country’s total catch and revenue from the species.
+The VI is calculated from: -->
+
+<!-- $$ index' = (P_c + P_v)/ToE $$ -->
+
+<!-- \begin{equation} -->
+<!-- \label{eq:five} -->
+<!-- VI' = \frac{(P_c + P_v)}{ToE} -->
+<!-- \end{equation} -->
+
+<!-- where $P_c$ and $P_v$ indicate the proportion accounted for by shared species in total catch and value generated from transboundary fisheries within a country’s EEZ, respectively.
+Here, the proportion of total catch and value are represented as the average of 2005 to 2014 within individual EEZs according to the Sea Around Us.
+The VI was then normalized to a 0 (less vulnerable) to 100 (more vulnerable) scale.
 -->
 
-Climate change is shifting the distribution of transboundary fish stocks between neighbouring countries.
-The time-frame of such shifts determine the extent of climate change challenges to international fisheries governance.
-Using global climate-fisheries model simulations, we show that, within the next 15 years, 68% of the world’s transboundary fish stocks are projected to shift their distributions across international boundaries.
-By 2100, such shifts will occur in all countries under a ‘no mitigation’ scenario.
-Hotspots of transboundary shifts occur in countries that are highly dependent on fisheries for income and livelihood.
-The imminent onset of transboundary shifts highlight the urgency to re-formulate international fisheries policies to avoid disputes over fish stock.
-Bi- and multi-lateral actions discussions should start now as lead-time for policy changes is long.
+### Estimating the threat point of transboundary stock share
 
-**One-Sentence Summary**
+In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than those operating on a non-cooperative basis [@Munro:1979dg].
+However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., “threat point” [@Sumaila:2013hv].
+Here, I defined the threat point as the minimum required proportion of a shared stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [Chapter 2, @Sumaila:2020tp].
+Any proportion below the defined threat point of a country would result in a unilateral management of the shared stock.
 
-Rapid climate-induced shifts in transboundary fish stocks
+First, I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100.
+I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2).
+Second, I averaged the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of variability.
+The first time period ($SSR_{th}$) spans 1951 to 2005 and was chosen to match the historical period in the GFDL ESM2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db].
+It was assumed to represent the historic baseline of shared distributions for each stock.
+I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SSR_{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR_{tm}$).
+I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds].
+The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members.
+Third, I defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$.
+This way, a change of SSR beyond an EEZ’s threat point happened when the future SSR exceeded one standard deviation of the historical variations of the SSR i.e when $SSR_{te} \geqq (SSR_{th} + \sigma)$ or $SSR[t]_{e,m} \leqq (SSR_{th} - \sigma)$.
+Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following (See Chapter 4 - Methods),
 
-**Main Text.**
+<!-- $$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$ -->
 
-In the past century, human activities have altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters [@IPCC:2019tn].
-Marine species' distributions reflect species' preferences for discrete environmental conditions [@Hutchinson:1957nt].As a result of these past changes, many marine species have shifted their distributions towards higher latitudes, deeper water or to follow local temperature gradients to remain within their optimal environmental "niche" [@Poloczanska:2016kk].
-The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt], impacting fisheries production and economies [@Sumaila:2019ew], and compromising our capacity to reach international sustainability goals within the 2030 Agenda such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
-The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management, including for species that cross international borders, which are known as ‘shared stocks’ [@Pinsky:2018cb].
+\begin{equation}
+\label{eq:five}
+\Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100
+\end{equation}
 
-The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by Coastal States [@UnitedNations:1986tl].
-As defined by the United Nations' Food and Agriculture Organization, shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and (*iv*) discrete stocks that are only present in the high seas [@Munro:2004th].
-This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
-Countries are responsible for the management of stocks within their EEZs, under UNCLOS, and are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species representing 67% of identified exploited taxa.
-These species yielded an annual average of 48 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
+### Statistical analysis
 
-The effectiveness of fisheries management for transboundary species, and thus reaching the 2030 Agenda [@Pecl:2017hu], is challenged by shifts in stocks' distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].
-In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
-While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs].
-Such patterns are expected to be exacerbated by continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
-The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the  Sustainable Development Goals (SDG) under the United Nations 2030 Agenda.
-Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].
+The time of emergence results were tested for normality (e.g., skewness, kurtosis) and performed two non-parametric Kruskal–Wallis test by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the ToE of transboundary stocks.
+Specifically, I tested if the habitat association of transboundary species and the geographic location of EEZs would have any effect on the time of emergence of transboundary stocks.
+For both cases, my null hypothesis was that there were no significant differences in the time of emergence across habitat association nor EEZs.
+<!-- Moreover, I implemented a multiple linear regression analysis to a series to explain the ToE of each species in each EEZ.
+Specifically, I looked at habitat associations, catch trends, geopolitical and biome location and ToE of both bottom and sea surface temperature (See *Databases and species selection*).
+I hypothesize that all of these factors would have significant influence in the ToE of transboundary species.
+-->
+All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages data.table [@Packagedatatable:2019uh], janitor [@PackagejanitorSim:2018ur], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], gmt [@PackagegmtInterfa:2017vh] and zoo [@PackagezooSInfr:2019tm], ggrepel [@PackageggrepelAut:2020to], zeallot [@PackagezeallotMul:2018wy], viridis [@PackageviridisDef:2018us], cowplot [@CowplotStreamlined:2019wt], moments [@PackagemomensMome:2015uh] and pgirmess [@PackagepgirmessSp:2018um].
+All code is available at https://github.com/jepa/TransEmergence
 
-Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (RCP8.5; Methods).
-We then estimate the year in which the 663 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variation and the intensity of this shift (i.e., time of emergence; see Methods).
-We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (15) and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
+## Results
 
-**Identifying range shifts in the shared distribution of transboundary stocks**
+```{r Load_Results, eval = F,  echo = FALSE, include = FALSE}
 
-Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability as early as 2006 and up to 2100 (Fig.
-\@ref(fig:toeFigure1)A).
-This corresponds to 68% of the studied stocks (Fig.
-\@ref(fig:toeFigure1)C).The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be 2036 (s.d.
-of all species, $\pm$ 28 years).
-Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% presenting a distributional shift across stocks by the time the UN 2030 Agenda is intended to be achieved (Fig.
-\@ref(fig:toeFigure1)C).
+prop_r <- read_csv("/Volumes/HALL2000/Data/TransEmergence/Results/proportion_change_2005.csv") %>% 
+  gather("time_step","change",per_early,per_mid)
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary shared stock distribution.
-Dashed lines represent the year by which countries have committed to reach the full implementation of the 2030 Agenda, as well as a mid and late change in range shift, respectively."}
+```
+
+### Time of emergence of transboundary stocks
+
+The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:toeFigure1)).
+This means that on average, the distribution of transboundary stocks will diverge from their natural internal variability by 2029.
+In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100.
+Fish stocks with transboundary index that first emerged from their historical variability in each EEZ is projected to be as early as 2006 and the last one in 2100.
+About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g., Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010).
+The number of EEZs with emerging stocks is projected to increase steeply from 87% in 2006 to 96% in 2020.
+After the mid-2020s, the rate of increase in the number EEZs slows from a 1% to an average of 0.7% per year.
+By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:toeFigure1)).
+Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020.
+Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure \@ref(fig:toeFigure1)).
+I based the transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift).
+Results from the sensitivity analysis suggest that this method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g., the top 95%) of the stock (Figure \@ref(fig:toeFigureS3)).
+
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Time of emergence of the transboundary index by Exclusive Economic Zone (EEZ) and transboundary stock"}
 
@@
 
-The median year in which transboundary stocks experience a range shift varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Figure \@ref(fig:toeFigure2)A).
-Overall, most tropical EEZs will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia being significantly earlier (*p* < 0.05; see Table S\@ref(tab:toeTableS2) for test statistics) than almost all other regions (Fig.
-\@ref(fig:toeFigure2)B).
-In contrast, EEZs located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts (Table \@ref(tab:toeTableS2)).Stocks in countries such as Brazil and New Zealand are not projected to experience range shifts beyond natural variability (e.g., the transboundary index of a stock did not change beyond historical variability) (Fig.
-\@ref(fig:toeFigure2)A).
-On average, shifting transboundary stocks represented 23% (s.d.
-$\pm$ 22%) of yearly fishing revenue from fisheries targeting such stocks within global EEZs between 2004 and 2010 (Fig.
-\@ref(fig:toeFigure2)A).
-However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
-Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).
+The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A).
+Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B).
+In contrast, EEZs located in temperate regions like northern Europe and eastern Asia have significantly later times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058).
+However, notable exceptions to these broad patterns exist, with  tropical EEZs displaying a later average time of emergence across stocks including the  Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070) and despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) display a relatively early average time of emergence (Figure \@ref(fig:toeFigure2)A).
+There are only a few cases where the transboundary index of a stock does not emerge between 2006 and 2100.
+In the Arctic EEZs of Canada and the US, the present chapter included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*).
+In the case of Brazil, I included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*).
+However, on all cases, I find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A).
 
-```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks.
-A) Land polygon shows the contribution of shifting stocks to the country’s total fishing revenue from transboundary stocks.
-Exclusive Economic Zone polygons represent average range shifts within them.
-EEZs with no distributional shifts between 2006 and 2100 are represented in aqua color.
-B) Year of shared distribution shifts by region according to United Nations sub-regions.
-Points represent the median year of range shift.
-Horizontal dashed line represents the year 2030.
-N = North, S = South, W = West and E = East.
-Ltn.
-Ame.
-and the Car.
-= Latin America and the Caribbean.
-Aus and New Z.
-= Australia and New Zealand"}
+```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of stocks’ transboundary index.
+A) Land polygon shows the participation of the emerging stocks in the country’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygon represents the average time of emergence of the transboundary index across all stocks.
+Aqua color represents EEZs with no time of emergence between 2006 and 2100.
+B) Time of emergence of the transboundary index according to the United Nations sub-regions.
+Whiskers represent 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range.", fig.scap="Average time of emergence of stocks’ transboundary index.", out.extra=''}
 
@@
 <!-- ![toeFigure 2.
-Year of change in the shared distribution of 4,119 transboundary stocks.
-A) Land polygon shows the contribution of shifting stocks to the country’s total fishing revenue from transboundary stocks.
-Exclusive Economic Zone polygons represent average range shifts per EEZ.
-EEZs with no time of emergence between 2006 and 2100 are represented in aqua color.
-B) Year of shared distribution shifts by region according to United Nations sub-regions.
-Points represent the median year of range shift.
-Horizontal dashed line represents the 2030 Agenda.
-N = North, S = South, W = West and E = East.
-Ltn.
-Ame.
-and the Car.
-= Latin America and the Caribbean.
-Aus and New Z.
-= Australia and New Zealand](../Figures/Fig3.png) -->
+Average time of emergence of stocks’ transboundary index.
+A) Land polygon shows the participation of the emerging stocks in the country’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygon represents the average time of emergence of the transboundary index across all stocks.
+Aqua color represents EEZs with no time of emergence between 2006 and 2100.
+B) Time of emergence of the transboundary index according to the United Nations sub-regions.
+Whiskers represent 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Line represents median, and black points represent points outside of 1.5*IQR.](../Figures/Fig3.png) -->
 
-We estimated the current (2005 and 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs.
-Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD in fishing revenue are projected to shift within the EEZs of 76 countries (Fig.
-\@ref(fig:toeFigure3)); with 4.2 billion USDs accounted by 321 stocks shifting in just eleven countries (Fig.
-\@ref(fig:toeFigure3)).
-Most of the remaining countries in the top $75^{th}$ revenue percentile will see early shifts (e.g., before 2050) with the exception of Japan, Korea, Peru and Russia, who are projected to have later shifts (e.g., after 2050).
-In terms of the number of transboundary stocks shifting from historical variability, Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66), emerge as the top 5 countries (Fig.
-\@ref(fig:toeFigure3)).
+My analysis suggests that most (n = 47) countries responsible for 75% of the yearly revenue generated from transboundary stocks between 2005 and 2010 will see the distribution range of their stocks shift beyond historical variabilities between 2025 and 2040 (Figure \@ref(fig:toeFigure3)).
+In some cases, like Indonesia (ranked $8^{th}$ in terms of revenue from transboundary stocks), Thailand (ranked $7^{th}$ in terms of revenue from transboundary stocks) and the United States (ranked $4^{th}$ in terms of fishing revenue from transboundary stocks), the range shift of transboundary stocks will emerge by around 2030.
+The 47 countries that make up the bulk of the global fisheries revenues also have more transboundary stocks with range shifts emerging from historical variability, such as Spain (n = 122), France (n = 85), China and Portugal (both n = 73) and Senegal (n = 66), which are the top 5 leading countries in terms of number of transboundary stocks emerging from historical variability.
 
-```{r toeFigure3, eval = T,  echo = F, fig.width=10, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks per country.
-Countries are coded according to different colored symbols representative of 5 distinct regions.
-Horizontal line separating the top 75th percentile of annual fishing revenue.
-Showing country names for those within the top 75th percentile projected to  have an average shift within 5 years of the 2030 Agenda."}
+On average, stocks that see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of the revenue generated by fisheries that target transboundary stocks within their EEZs between 2004 and 2010.
+However, large variation exists in the contribution of stocks with emerging transboundary index to the current revenue participation across EEZs.
+In some cases, like El Salvador and New Zealand, stocks with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries in their respective EEZs between 2004 and 2014.
+On the contrary, stocks with emerging transboundary index from the Marshall Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A).
+In addition, some EEZs are projected to have a few emerging stocks that represent a large proportion of revenue from transboundary stocks like Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively).
+Overall, Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks to fishing revenue is less than 20%.
+This indicates that, despite having a large proportion of transboundary stocks emerging, the economic contribution of these are relatively small to the overall revenue generated by fisheries targetting transboundary stocks in these countries.
+
+```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average time of emergence of stocks’ transboundary index per fishing entity, aggregated by colour and shape according to region.
+Fishing revenue on a logarithmic scale.
+Showing country names for the top 90th percentile of annual fishing revenue.", fig.scap="Average time of emergence of stocks’ transboundary index per fishing entity, aggregated by colour and shape according to region", out.extra=''}
 
@@
 
-We based the transboundary index on a stock’s centroid defined as the average grid cells that contain the majority of a stock’s abundance (see Methods - Calculating an index of transboundary range shift).
-Results from the sensitivity analysis suggest that our method is robust to the abundance level (e.g., the top 95%) of the stock (Fig.
-\@ref(fig:toeFigureS3)).
-In addition, we set an arbitrary threshold of 2 s.d.
-(95% confidence) to determine the time of emergence of the transboundary index.
-Sensitivity tests to a lower threshold of 1 s.d.
-(for a 68% confidence) resulted in a global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27, a difference of about 7 years (Fig.
-\@ref(fig:toeFigureS4)).
-In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks.
+I set an arbitrary threshold of 1 s.d.
+(67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods).
+I tested the sensitivity to a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence) and the global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d.
+threshold (Figure \@ref(fig:toeFigureS2)).
+In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric.
 
-**Changes in the stock share ratio of transboundary stocks**
+I compared stocks time of emergence within EEZs by the species’ habitat association (Figure \@ref(fig:toeFigure4)).
+Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods).
+I found significant statistical differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001).
+Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)).
+On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (Figure \@ref(fig:toeFigure4)).
+Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:toeFigure4)).
 
-We identified transboundary stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point by the early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005).
-Findings show that by the early $21^{st}$ century, the global average catch proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 59% $\pm$ 17% (Fig.
-\@ref(fig:toeFigure5)A).
-By the early $21^{st}$, 85% (n = 239) of the world EEZs would have experienced such changes.
-Changes in stock share ratio are expected to increase slightly towards the mid 21st century at both the stock and EEZ level (Fig.
-\@ref(fig:toeFigure5)A and \@ref(fig:toeFigureS5)).
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=6,fig.pos= "H", fig.cap =  "Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association.
+Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Methods).
+The number of species included in this analysis for each habitat type is noted in parenthesis.
+Whiskers represent 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Line represents median, and black points represent outliers with values exceeding 1.5 times the interquartile range.", fig.scap="Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association.", out.extra=''}
+
+img <- readPNG("../Figures/Fig4.png")
+grid.raster(img)
+
+```
+
+<!-- ![toeFigure 4.
+Regional time of emergence per habitat association.
+Regions according to the United Nations and habitat association to FishBase.
+Whiskers represent 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Line represents median, and black points represent outliers (outside of 1.5*IQR.](../Figures/Fig5.png) -->
+
+<!-- I hypothesized that species environmental preferences (e.g.
+presence in certain biomes and habitats), climate change effects on water temperature (e.g.
+ToE of SST), human drivers (e.g.
+catch rate), and latitudinal gradient would have significant influence in the ToE of transboundary species (Table \ref(tab:toeTableS3)).
+Results from the multiple linear regression analysis show a significant relation between ToE of transboundary species and latitudinal gradient (*p* > 0.001), tropical (*p* = 0.013) and temperate (*p* = 0.002) biomes, EEZs in the Americas (*p* > 0.001), Asia (*p* > 0.001) and Oceania (*p* = 0.002), and some habitats (*p* > 0.001) (Table \ref(tab:toeTable1)).
+-->
+
+```{r Table1, eval = F,  echo = F, message = F, warning  = F}
+
+read_csv("../Tables/mlr_result.csv") %>% 
+  knitr::kable(
+    "latex",
+    booktabs = T,
+    captio= "Multiple regression results for time of emergence of transboundary stocks.
+Only showing significant variables.
+The model significantly predicts time of emergence (F[18,6155] = 20.68, p < 0.001, R2 = 0.05)"
+  )
+
+```
+
+### Changes in the stock share ratio of transboundary stocks
+
+I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods – Equation \@ref(eq:five)).
+By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A).
+The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)).
+However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high.
+For example, Guatemala shares *Panulirus gracilis* stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century.
+Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)).
+My results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
+These EEZs will changes in the stock share ratio beyond the threat point in one third (33% $\pm$ 21%) of their transboundary stocks (Figure \@ref(fig:toeFigure5)A).
+Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin are projected to have changes beyond the threat point for all of the transboundary stocks analyzed, while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) are the only countries with less than 10% of stocks changing beyond the threat point.
 
 ```{r toeFigure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors (green to yellow).
-Panel B zooms into specific areas shown within the grey dashed line boundaries in A."}
+Lines represent the average gain of transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those that are gaining shares (arrowhead).
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point.
+Panel B zooms in to specific areas shown within the grey dashed line boundaries in A.", fig.scap="Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.", out.extra=''}
 
@@
 <!-- ![Figure 1.
-Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors (green to yellow).
-Panel B zooms into specific areas shown within the grey dashed line boundaries in A.](../Figures/Fig1.png) -->
+Changes in Specie Share Ratio by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average gain of transboundary specie share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial species that will see a shift.
+Panel B zooms in to specific areas shown in grey boxes of A.](../Figures/Fig1.png) -->
 
-The direction and intensity of shafts in transboundary stocks are largely related to regional changes in biogeography and the geometry of the EEZs (Fig.
-\@ref(fig:toeFigure5)A).
-In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the stock share ratio is expected to follow a projected poleward shift.
-However, shifts along the coasts of Pacific Central America and West Africa  occur along an equatorial direction.
-Here, transboundary species follow local oceanographic processes like upwelling zones (e.g., Humboldt current) and patterns of oxygen availability, in addition to temperature gradients.
-Regions where EEZs are relatively small and have multiple borders (e.g., Caribbean or Northwestern Europe) are expected to have a particularly complex exchange of stock share ratio with no established pattern (Fig.
-\@ref(fig:toeFigure5)B).
+Changes of stock share ratio are largely related to regional changes in biogeography and the geometry of the EEZs (Figure \@ref(fig:toeFigure5)).
+In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the stock share ratio is expected to follow the projected poleward shift of transboundary species under climate change.
+However, central America Pacific and West African coasts follow an equatorial direction as range shifts of transboundary stocks follow local oceanographic gradients, with relatively lower temperature waters in lower latitude [@Clarke:2020ep].
+Moreover, regions where EEZs are relatively small and have multiple borders are expected to have a particularly complex exchange of stock share ratio with no established pattern, as in the Caribbean, Mediterranean and northwestern Europe (Figure \@ref(fig:toeFigure5)B).
 
-<!-- ## Discussion -->
+## Discussion
 
-<!-- In this paper, we investigate the timing and intensity associated with changes in the share distribution of transboundary stocks between countries.
-Our findings highlight an early distributional shift from historical variations for most transboundary stocks in the world.
-Such early shifts point to the urgency of addressing the challenges that climate change presents for international fisheries management.
-Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries that would benefit from immediate policy responses to support sustainability into the future.
-Finally, we show that for many countries changes in the distribution of transboundary stocks are likely to undermine joint resource management between countries [@Bailey:2010ga; @Munro:1979dg; @PalaciosAbrantes:2020vt; @Bailey:2016eh].
-Thus, our findings emphasize recent calls for the urgent adoption of adaptation measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @CisnerosMata:2019ks; @Oremus:2020id; @Free:2019jq].
+In this chapter, I investigate the timing and intensity of changes in the sharing of transboundary stocks between countries.
+My findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
+Such early emergence of range shifts indicate the urgency of addressing the challenges that climate change is posing on international fisheries management.
+Moreover, I identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management.
+Finally, I show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to levels below of that needed to encourage cooperative resource management between countries.
+Adaptation measures are therefore needed to reduce the impacts of species range shifts on current international transboundary fisheries treaties.
+
+### High present-day climate risk of on transboundary fisheries management
+
+My projections for the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century concur with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk].
+For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US).
+A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs].
+In the northeast Atlantic, the Atlantic mackerel (*Scomber scombrus*) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
+However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008, without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij].
+Such changes resulted in disputes between Iceland and the Faroes Islands, as well as the NEAFC member states [@Spijkers:2017ij].
+As climate change impacts on transboundary stocks continue to emerge from historical variability in the coming years, mal-adapted fisheries management and international disputes like that of Atlantic mackerel are likely to increase in frequency [@Pinsky:2018cb].
+
+The projected range shifts of large number of transboundary stocks emerging from historical variability since the early 2000s can be partially attributed to the parallel emergence of several environmental variables (e.g., temperature and oxygen) that influence their distribution [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Mahlstein:2011fy].
+Marine organisms have a thermal tolerance that limits their habitable temperature range [@IPCC:2019tn], which then explains the sensitivity of the animal to warming and is a main driver of any distributional change [@IPCC:2019tn].
+Oxygen plays a similar role to that of temperature [@Clarke:2020ep], especially for tuna species that require large amounts of dissolved oxygen for their metabolism [@IPCC:2014bz].
+It has been estimated that since the early 2000s, increased sea surface temperature (SST) and decrease in subsurface oxygen have emerged from historical variability for 79% and 23% of the global ocean, respectively, and that by 2080, SST is projected to emerge from historical variability in 90% of the global ocean while the area affected by lower $O_2$ will have doubled under a high greenhouse gas emissions scenario (RCP 8.5) [@Rodgers:2015jq].
+Moreover, before 2100, surface acidification, nitrate content and net primary production, also important variables for species distributions, will have emerged from the historical variability of 1861-1900 in 30% and 60% of the ocean under a low (RCP 2.6) and strong (RCP 8.5) climate change scenario, respectively [@IPCC:2019tn].
+These oceanographic changes are important drivers of change in species distributions as represented in the DBEM [@Cheung:2016jd; @Cheung:0ur].
+
+Regional differences in the time of emergence of many transboundary stocks are attributed to the variations in environmental conditions, species’ climate vulnerability, and the characteristics of political boundaries.
+Specifically, my findings are consistent with results from previous studies that show tropical species are already shifting their distributions since the last few decades [@Poloczanska:2016kk].
+The tropics are among the most species-rich biomes [@Reygondeau:2019uh] and have potentially experienced some of the highest levels of warming relative to natural internal variability [@IPCC:2019tn; @Roemmich:2015id].
+In the tropics, marine species also live close to their thermal tolerance, making them highly vulnerable to warming [@IPCC:2014bz].
+Consequently, species have to move towards deeper or sub-tropical waters as a result of climate change [@Poloczanska:2016kk], although in some regions species follow local temperature gradients towards the equator or shallow waters [@Clarke:2020ep].
+Moreover, tropical species are expanding their range into sub-tropical biomes [@Fogarty:2017eo] and catch-data analysis shows a reduction in sub-tropical species catches in tropical waters between 1970 and 2006 [@Cheung:2013gk].
+In addition to species natural traits and the environmental pressure from climate change, the tropics comprise areas with relatively small EEZs that border with multiple countries (e.g., the Caribbean).
+These tropical areas are projected to have the highest number of transboundary stocks’ distribution shifts emerging in the incoming years.
+
+### Hotspot of climate risk on transboundary fisheries management
+
+My results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)).
+For example, Polynesia and Micronesia are the only regions where the average time of emergence of transboundary stocks' distribution across all EEZs emerged before 2022 (Figure \@ref(fig:toeFigure5)).
+Moreover, countries in these hotspot regions are highly dependent on pelagic-oceanic (e.g., tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh].
+On the other hand, “coldspots” of climate risk exist for transboundary fisheries, like the Northeast Atlantic and Northwest Pacific, where stocks are expected to emerge later in the $21^{st}$ century.
+Assuming that these countries already have adequate transboundary management to deal with historical and current variations in fish stocks’ distribution e.g., the Pacific Salmon Treaties [@Miller:2004fpa], management agencies are likely to have more time to prepare for emerging stock shifts.
+
+In cases where climate driven changes in stocks are emerging, or have already emerged, from historical variability adaptation policies are urgently needed to address the potential impact on important fisheries [@Burden:2019ij].
+Currently, fisheries targeting shared tuna stocks in the South Pacific Islands region operate under the Vessel Day Scheme, which because it takes into account shifts in stocks and catches makes it a good example of a management system that is somewhat anticipatory [@Aqorau:2018bh].
+However, even strategies that take into consideration the shift of the stock within the party’s area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+As the climate trend in stocks distribution emerges from their historical variability sch since 2006, regional management bodies must revise their management actions as most are not currently equipped with strategies to cope with shifting species distributions [@Oremus:2020id; @Sumby:2021gw].
+This is of specific concern for those management bodies located in the emerging “hotspots” here identified.
+
+Some EEZs lack stocks with emerging transboundary index because of a combination of their shape, climate projections, and a reduced number of transboundary stocks.
+For example, there are 59 commercial fish stocks fished in the Arctic with the vast majority on the Barents, Bering and Norwegian seas, with limited knowledge about the species biology [@Christiansen:2014gg].
+However, the Arctic has been warming faster than many other regions, resulting in a sea-ice reduction [@Douglas:2010Mb] consequently increasing the potential for occupation by species not previously present [@IPCC:2019tn].
+As a result, the biomass of Arctic marine animals [@BryndumBuchholz:2019kh] and catch and revenue of local fisheries [@Tai:2019cq; @Lam:2014fq] are expected to increase in the incoming years, although substantial uncertainty exists on the magnitude of change [@BryndumBuchholz:2019kh].
+Moreover, projections suggest a larger increment on transboundary stocks in the coming years in the Arctic [@Pinsky:2018cb] as climate change pushes sub-polar species further north [@Morley:2018fn; @Frainer:2017ji].
+In the specific case of Brazil, in addition to having few transboundary stocks (only four shared with Uruguay in this chapter), the large difference in the EEZs lengths could also be influencing the results.
+
+<!-- For Brazil and Uruguay, the current and future distribution of the transboundary stocks in combination with the particular characteristics of these EEZs could be influencing the results.
+First, both Argentine menhaden and Argentine shortfin squid have distributions with larger biomass concentration further south, with Brazil serving as the northern limit of the distributional range [@AquaMaps:2019ui; @AquaMaps:2019wt].
+Second, Brazil’s EEZ is substantially larger than Uruguay’s resulting in a long distance between Brazil’s and the stocks’ centroids.
+Third, both stocks are expected to shift further south in the incoming years [@AquaMaps:2019ui; @AquaMaps:2019w].
+Thus, the further southern shift of Argentine menhaden and shortfin squid in combination with the reduced latitudinal gradient of Uruguay’s EEZ results in a small effect in the already long latitudinal distance between Brazil’s EEZ and the stocks’ centroids.
+In the case of the two shrimp species, their destitution range spams along the cost of Brazil and is not expected to have a strong shift in the incoming years, also resulting in a stable transboundary index [@AquaMaps:2019wn; @AquaMaps:2019wj].
 -->
 
-**High present-day climate risk of on transboundary fisheries management**
+### Changes in stock share ratio will be ubiquitous
 
-Early shifts in the shared distribution of transboundary stocks concur with previous studies that have detected and attributed changes in marine catch composition [e.g., @Cheung:2013gk; @Frainer:2017ji; @Last:2011fd].
-For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US) as a response to climatic, oceanographic and ecological changes [@Zeidberg:2007gs].
-A new fishery targeting Humboldt squid quickly developed on the heels of the species’ range expansion [@Pinsky:2014by].
-In the northeast Atlantic, the Atlantic mackerel (*Scomber scombrus*) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
-However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the fishery’s total allowable catch and a further 18% in 2008, without consultation with NEAFC, threatening the sustainability of the stock [@Spijkers:2017ij].
-These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij].
-As transboundary stocks continue to shift, international disputes like that of Atlantic mackerel are likely to increase in frequency requiring a revision of management plans [@Pinsky:2018cb; @Engler:2020bg].
+Changes in stock share ratio beyond a country’s threat point will be seen across the world, suggesting that climate change impacts on transboundary stock management is a global problem (Figure \@ref(fig:toeFigure5)).
+Poleward shifts in stock share ratio are expected in regions of the world with relatively long EEZs and few boundaries, like the northeast Pacific and the southeast Atlantic.
+However, despite observations and modelling work indicating that climate change is generally shifting species distributions poleward [@Cheung:2010dt; @Poloczanska:2016kk; @IPCC:2019tn; @Fogarty:2017eo], some regions like the eastern tropical Pacific and eastern tropical Atlantic Oceans will experience a shift in distribution towards the equator.
+Specifically, in the eastern tropical Pacific, the direction of shift is expected to be from Mexico to Ecuador with northern EEZs losing stock share ratio to their southern neighbors.
+Such range shifts towards the equator reflect movements to cooler habitats characteristic of equatorial upwelling systems [@Clarke:2020ep] and are consistent with regional modeling studies suggesting that by the mid $21{st}$ century, marine species will move between 50 and 100 km southward of Mexico’s EEZ towards Peru under a high emission scenario [@Clarke:2020ep].
+Moreover, regions characterized by multiple political boundaries, like the Caribbean and the northeast Atlantic, will see a complex network of changes in stock share ratio with no clear pattern (Figure \@ref(fig:toeFigure5)B).
+These regions are known for being highly connected by both fish adult migration [@Levin:2018in] and dispersal of larvae [@Ramesh:2019va] supporting the implementation of joint management plans [@Burden:2019ij] and the call for new implemented shared policies [@Levin:2018in].
+However, game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Sumaila:2013hv], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting shared stocks.
 
-**Hotspot of climate risk on transboundary fisheries management**
+Changes in stock share ratio beyond the threat point of countries sharing a common resource could instigate international conflict and hinder bi-lateral agreements [@Miller:2013iv; @Sumaila:2020tp].
+Lessons learned from historic distributional shifts of shared stocks suggest that treaties that are not prepared to respond to such shifts will be less resilient to changes in transboundary fish stocks share ratio due to climate change [@Miller:2013iv; @Sumaila:2020tp].
+For example, disputes over Pacific salmon between Canada and the US arose in the 1990s when a climate-related shift in stock abundance favoured Alaska’s salmon fisheries over Canada’s [@Miller:2013iv; @Munro:2002uf; @Song:2017va].
+The conflict lasted over ten years until Canada and the US agreed on a mutual conservation fund where the US would contribute the larger proportion to support scientific research, habitat restoration and enhancement of wild stock production in transboundary rivers [@Munro:2002uf].
+Identified strategies to cope with changes in the share proportion of transboundary stocks include the incorporation of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and dynamic ocean management [@Pinsky:2018cb].
+Side payments are mechanisms that can provide a solution to the inequality produced by shifts in the stock’s distribution [@Tunca:2019if; @Miller:2004fpa].
+These can be monetary (e.g., the stock-“winning” state compensates the stock-“losing” state for the proportional shift) or other methods like the Pacific salmon case (e.g., a conservation fund paid mainly by one party) [@Miller:2004fpa; @Miller:2013iv].
+Strengthening current international cooperation will be required as stock distribution will move to neighboring jurisdictions, requiring combined efforts to generate information and set proper management rules [@Miller:2013iv].
+Finally, flexible management rules might be more appropriate to capturing t distributional shifts.
+For example, a quota allocation based on the stock’s current distribution like that for Pacific halibut (*Hippoglossus stenolepis*) managed by Canada and the US [@Sumaila:2020tp, Chapter 2] would be more agile than one based on a fixed-historical proportion like the case of the EU [@Baudron:2020dj].
 
-Our results underscore regional "hotspots" of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figures \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure5)A).
-Such regions, like the Caribbean, are characterized by high levels of warming relative to natural internal variability [@IPCC:2019tn; @Hawkins:2012hc] and high vulnerability of species to warming waters [@IPCC:2019tn].
-Moreover, they represent relatively small EEZs that border multiple countries (e.g., the Caribbean).
-Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting transboundary stocks.
+## Caveats and uncertainties
 
-Management plans that are not prepared to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
-Thus, focusing on these “hotspots'' is key to anticipate a potential increase in fisheries conflicts in the coming years.
-Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@Pinsky:2018cb].
-For example, countries that are Party to the Nauru Agreement (PNA) manage their fisheries under a  Vessel Day Scheme.
-Because this scheme takes into account shifts in stocks and catches, it is a good example of a management system that is somewhat anticipatory [@Aqorau:2018bh].
-However, even strategies that take into consideration  species’ shift within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
-Quota allocation methods based on a stock’s current distribution and/or on fixed-historical proportion need to evolve to be more agile [@Sumaila:2020tp; @PalaciosAbrantes:2020vt].
-Fisheries based on historic participation could move towards a dynamic method or a combination of both  [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-However, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from the fishery.
-Side payments are mechanisms that can provide a solution to the inequality produced by shifts in a stock’s distribution [@Tunca:2019if].
-These can, but do not have to, be monetary (e.g., the stock-“winning” state compensates the stock-“losing” state for the proportional shift, a conservation fund paid mainly by one Party) [@Miller:2013iv].
-
-**Potential drivers of shift and associated uncertainties**
-
-The early projected range shifts of a large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability (e.g., SST which is projected to increase historical variability by 2030 in 70% of the global ocean under RCP 8.5 [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
-Results from multiple marine ecosystem models show that SST and NPP are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
-Specifically for the DBEM, SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results.
-The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2019tn], while also explaining the later, and sometimes non-existent, shift at higher latitudes (Fig.
-\@ref(fig:toeFigure2)B).
-
-Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and  *ii*) model structure [@Frolicher:2016dg].
-First, our analysis is based on a high emission scenario (RCP 8.5) representing an extreme case of shifts in the shared distribution of transboundary species.
-Thus, any climate change mitigation pathway (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].
-Second, a substantial source of uncertainty is related to model selection for both climate change and fisheries distribution.
-This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
-While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available [@Deser:2020bj] and cannot address limitations of process level understanding.
-As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties, and will likely continue to do so in the coming decade.
-This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf] and where internal variability (e.g., El Niño Southern Oscillation - ENSO- for Pacific nations) is also important .
-These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh].
-Overall, changes in fish biomass with multiple upper trophic level models, including the DBEM, show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@BryndumBuchholz:2019kh].
-Further research that reduces uncertainty in the response of the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to a larger structural uncertainty of fish and fisheries models [@Lotze:2019ce].
-Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks.
-While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca].
-Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
+There are two main sources of uncertainty in my analysis.
+First, I rely on the combination of a single Earth system model (GFDL-ESM2M), fish and fisheries model, and climate change scenario (RCP 8.5).
+The use of multiple ESMs can capture the structural uncertainty of the models [@Frolicher:2016dg; @Lotze:2019ce], however, exercises looking at changes in fish biomass with multiple models show an overall agreement between models in terms of direction of change but variable in magnitude.
+Further research that includes multiple ESMs and species distribution models could elucidate a broader range structural uncertainty of the climate, fish and fisheries models [@Lotze:2019ce].
+Secondly, due to limited biological and spatially-specific data on stocks sub-population structure of transboundary species, my analysis uses political boundaries (EEZs) to delineate a stock that does not necessarily align with biologically-defined sub-populations within an EEZ.
+On the other hand, in many EEZs, fisheries are often managed at the species level (Chapter 2) and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for my analysis [@Popova:2019fs; @Dunn:2019ca].
+However, reproducing my analysis locally, where spatially explicit stock data is available would allow to produce more certain results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts within meta-populations level [@Link:2010ei; @Kaplan:2016hs; @Archambault:2016ie].
 Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
 
-**Conclusion**
+## Conclusion
 
-Th global community has set the ambitious goal of managing all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear benefits for a number of other societal goals [@UnitedNations2018; @Singh:2017ds].
-Preparing anticipatory policies to deal with shifting transboundary stocks is key to achieving the SDGs and ensuring effective governance of the world’s fisheries [@Pinsky:2018cb; @Oremus:2020id; @Pecl:2017hu].
-In this contribution, we implemented two steps to demonstrate how to achieve sustainable transboundary fisheries in a changing world [@Link:2010ei].
-First, we identified the transboundary stocks that would likely see shifts in their shared distribution compared to their historical average and the year in which such shifts are predicted to occur.
-Second, we estimated the intensity of such change.
-Our results have direct implications for ocean governance as the data provided can inform regional management bodies and member countries to anticipate the likely climate-driven challenges to transboundary fisheries management.
-While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for shifting transboundary stocks [@Pinsky:2018cb; @PalaciosAbrantes:2020tv].
-Our findings emphasize recent calls for the urgent adoption of measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @Oremus:2020id].
+The global community has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear effects on multiple others [@UnitedNations2018; @Singh:2017ds].
+Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@Pinsky:2018cb].
+Here, I have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei].
+I have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change.
+My results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management.
+While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, my results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; @Gaines:2018sg, Chapter 2].
+To address the issues raised in this chapter, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.
 
@@
 
-<!-- Set Figure, table and equations numbering according to the chapter -->
-\renewcommand{\thefigure}{S.\arabic{figure}}
-\setcounter{figure}{0}
-\renewcommand{\thetable}{S.\arabic{table}}
-\setcounter{table}{0}
+```{r FigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this chapter as devided by the Sea Around Us and their centroids (points)."}
 
-```{r supplements, child = '../Scripts/Supplements.Rmd', eval = T}
+img <- readPNG("../Figures/Supp1.png")
+grid.raster(img)
 
 ```
+
+```{r FigureS2, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel).
+Time of emergence is defined as the first year when the TI overpasses historical values (Bottom panel)."}
+
+img <- readPNG("../Figures/Supp2.png")
+grid.raster(img)
+
+```
+
+```{r FigureS3, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
+
+img <- readPNG("../Figures/Supp3.png")
+grid.raster(img)
+
+```
+
+```{r FigureS4, eval = F,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005)"}
+
+img <- readPNG("../Figures/Supp4.png")
+grid.raster(img)
+
+```
+
+<!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005))](../Figures/Supp2.png) -->
+
+```{r FigureS5, eval = F,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
+Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
+Polygons represent the percentage of initial stocks that will see a shift.
+Panel B zooms in to specific areas shown in grey boxes of A."}
+
+img <- readPNG("../Figures/Supp5.png")
+grid.raster(img)
+
+```
+
+```{r TableS1, eval = F, echo = F}
+
+read.csv("~/GitHub/TransEmergence/Tables/kw_regional.csv") %>% 
+  select(-statistic,
+         "habitat association comparrison" = comparrison,
+         "Significan level" = signif.level,
+         "Observation difference" = dif.com.obs.dif,
+         "Critical difference" = dif.com.critical.dif,
+          "K-W Difference" = dif.com.difference) %>% 
+  kable("latex",
+        booktabs = T,
+        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  ) %>% 
+  kable_styling(full_width = F,
+                position = "center")
+
+```
+
+```{r TableS2, eval = F, echo = F}
+
+read.csv("~/GitHub/TransEmergence/Tables/kw_habitat.csv") %>% 
+  select(-statistic,
+         "Sub region comparrison" = comparrison,
+         "Significan level" = signif.level,
+         "Observation difference" = dif.com.obs.dif,
+         "Critical difference" = dif.com.critical.dif,
+          "K-W Difference" = dif.com.difference) %>% 
+  kable("latex",
+        booktabs = T,
+        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  ) %>% 
+  kable_styling(full_width = F,
+                position = "center")
+
+```
+
+<!-- # Recycled  text -->
+<!-- All of these results have direct implications for managing shared fisheries under climate change that can be partially addressed by the early implementation of joint-management international policies.
+-->
+
+<!-- The time and number of emerging species decreases as I move away from the equator and into temperate biomes where species live further away from their temperature tolerance and thus hare less vulnerable to warming [@IPCC] with a smaller peak aligned with the EEZ of the Baltic and North seas.
+-->
+
+<!-- Changes in stock share ratio has contribute to past conflicts between nations with long history of collaboration including Canada and the United States over Pacific salmon and Norway and Iceland over spring spawning herring [@Miller:2013iv].
+In both cases, rules were emplaced and agreed by all parties, however, treaties were not prepared to address shifts in species distributions [@Miller:2013iv; @Song:2017va] and when the benefits felt below the desired level (e.g.
+threat point) the affected country broke off relationships.
+In the case of Canada and the US, the agreement was re-stated when a conservation fund was implemented [@Miller:2013iv].
+-->
+
+<!-- In cases where the species is distributed in more than 2 neighboring EEZs (e.g. *xxx*), the index is as follows: -->
+
+<!-- $$I =(\frac{D_{A,y}}{sd(D_{A,h))} - \frac{D_{B,y}}{sd(D_{B,h)))) |  \\  +(\frac{D_{A,y}}{sd(D_{A,h))} - \frac{D_{C,y}}{sd(D_{B,h)))) | \\  +(\frac{D_{A,y}}{sd(D_{A,h))} - \frac{D_{Cy}}{sd(D_{B,h))))/N$$ -->
+
+<!-- In other words, for two neighboring EEZs sharing a stock, if the centroid shifts towards EEZ A ($D_A$ decreases) but away from EEZ A B ($D_B$ increases), EEZ A may catch a bigger share over B and the risk of dispute increases (see *TI* equation).
+Consequently, if the centroid shifts are within historical variability (*SNR* < 1), risk of dispute between EEZ A and B is lower, finally, if the centroid shifts are beyond historical variability (*SNR* > 1), the risk of dispute is higher.
+Thus, risk is positively related to the frequency of centroid shifts beyond historical variability.
+-->

### 2021-03-19 Scripts/Supplements.Rmd
typos insupplements

--- 2020-12-22 Scripts/Supplements.Rmd
@@
 
-```{r toeFigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time  of Emergence and standar dviation for a 64% confidence treshold  (purple) and a 98% confidence teshold (yellow)"}
+```{r toeFigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average Time of Emergence and standard deviation for a 64% confidence threshold (purple) and a 98% confidence threshold (yellow)"}
 
@@
 
-```{r toeFigureS5, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st centurury relative to today (1951-2005)"}
+```{r toeFigureS5, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) 21st century relative to today (1951-2005)"}
 
@@
 ```{r toeFigureS6, eval = T,  echo = F, fig.width=9, fig.height=12, fig.cap =  "Changes in Stock Share Ration by 2050 (2041-2060) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with srrows going from the losing EEZ (point) to the winner (arrowhead).
+Lines represent the average gain of transboundary stock share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
 Polygons represent the percentage of initial species that will see a shift.

### 2021-03-22 Scripts/Manuscript.Rmd
Addressed william's comments

@@
 Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn].
-Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk]).
+Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk].
 As a result of climate change, many marine species have changed their distributions towards higher latitude, deeper water or have followed local temperature gradients [@Poloczanska:2016kk].
@@
 Overall, Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks to fishing revenue is less than 20%.
-This indicates that, despite having a large proportion of transboundary stocks emerging, the economic contribution of these are relatively small to the overall revenue generated by fisheries targetting transboundary stocks in these countries.
+This indicates that, despite having a large proportion of transboundary stocks emerging, the economic contribution of these stocks are relatively small to the overall revenue generated by fisheries targeting transboundary stocks in these countries.
 
@@
 My results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A).
-These EEZs will changes in the stock share ratio beyond the threat point in one third (33% $\pm$ 21%) of their transboundary stocks (Figure \@ref(fig:toeFigure5)A).
+These EEZs are projected to have changes in the stock share ratio beyond the threat point in one third (33% $\pm$ 21%) of their transboundary stocks (Figure \@ref(fig:toeFigure5)A).
 Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin are projected to have changes beyond the threat point for all of the transboundary stocks analyzed, while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) are the only countries with less than 10% of stocks changing beyond the threat point.
@@
 My findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world.
-Such early emergence of range shifts indicate the urgency of addressing the challenges that climate change is posing on international fisheries management.
+Such early emergence of range shifts indicates the urgency of addressing the challenges that climate change is posing on international fisheries management.
 Moreover, I identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management.

### 2021-03-23 Scripts/Manuscript.Rmd
Included CCW comments untill Fig 2

--- 2021-03-19 Scripts/Manuscript.Rmd
@@
 title: "Imminent range shifts in transboundary fish stocks challenge their management under climate change"
-author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
+author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette C.C. Wabnitz^1,5^, and William W.L. Cheung^1^'
 date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
@@
 Climate change is shifting the distribution of transboundary fish stocks between neighbouring countries.
-The time-frame of such shifts determine the extent of climate change challenges to international fisheries governance.
+The time-frame of such shifts determines the extent of climate change challenges to international fisheries governance.
 Using global climate-fisheries model simulations, we show that, within the next 15 years, 68% of the world’s transboundary fish stocks are projected to shift their distributions across international boundaries.
 By 2100, such shifts will occur in all countries under a ‘no mitigation’ scenario.
-Hotspots of transboundary shifts occur in countries that are highly dependent on fisheries for income and livelihood.
-The imminent onset of transboundary shifts highlight the urgency to re-formulate international fisheries policies to avoid disputes over fish stock.
-Bi- and multi-lateral actions discussions should start now as lead-time for policy changes is long.
+Hotspots of transboundary shifts occur in countries that are highly dependent on fisheries for income, food and livelihoods.
+The imminent onset of transboundary shifts highlights the urgency to re-formulate international fisheries policies to limit conflict over these fish stocks.
+Bi- and multi-lateral actions should start now as lead-time for policy change is long.
 
@@
 
-In the past century, human activities have altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters [@IPCC:2019tn].
-Marine species' distributions reflect species' preferences for discrete environmental conditions [@Hutchinson:1957nt].As a result of these past changes, many marine species have shifted their distributions towards higher latitudes, deeper water or to follow local temperature gradients to remain within their optimal environmental "niche" [@Poloczanska:2016kk].
-The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt], impacting fisheries production and economies [@Sumaila:2019ew], and compromising our capacity to reach international sustainability goals within the 2030 Agenda such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
-The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management, including for species that cross international borders, which are known as ‘shared stocks’ [@Pinsky:2018cb].
+Most fisheries management is poorly adapted to address species’ shifting distributions as a result of human driven warmer, more acidic and less oxygenated ocean waters.
+Over the past century, human activities have dramatically altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters [@IPCC:2019tn].
+Marine species’ distributions reflect species’ preferences for discrete environmental conditions [@Hutchinson:1957nt].
+As a result of changing ocean conditions, many marine species have shifted their distributions towards higher latitudes, deeper water or following local temperature gradients to remain within their optimal environmental “niche” [@Poloczanska:2016kk].
+The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt], impacting fisheries production and economies [@Sumaila:2019ew], and compromising our capacity to reach international sustainability goals within the 2030 Agenda for sustainable development such as Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
+The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management, including for species that cross international borders, also known as ‘shared stocks’ [@Pinsky:2018cb].
 
@@
 This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
-Countries are responsible for the management of stocks within their EEZs, under UNCLOS, and are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
-A recent study estimates that, globally, there are 633 transboundary fish species representing 67% of identified exploited taxa.
-These species yielded an annual average of 48 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].
+Countries are responsible for the management of stocks within their EEZs.
+Under UNCLOS, EEZs are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
+A recent study estimates that there are 633 transboundary fish species globally, representing 67% of identified exploited taxa.
+Between 2005 and 2010, these species yielded an annual average of 48 million tonnes of catch and USD 78 billion in fishing revenue [@PalaciosAbrantes:2020tv].
 
-The effectiveness of fisheries management for transboundary species, and thus reaching the 2030 Agenda [@Pecl:2017hu], is challenged by shifts in stocks' distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].
+The effectiveness of fisheries management for transboundary species is challenged by shifts in stocks’ distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].
 In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
 While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs].
-Such patterns are expected to be exacerbated by continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
+Such patterns are expected to be exacerbated under continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
 The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the  Sustainable Development Goals (SDG) under the United Nations 2030 Agenda.
-Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].
+Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the goals set out under the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].
 
@@
 of all species, $\pm$ 28 years).
-Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% presenting a distributional shift across stocks by the time the UN 2030 Agenda is intended to be achieved (Fig.
+Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% having recorded a distributional shift in stocks by the UN 2030 Agenda’s deadline (Fig.
 \@ref(fig:toeFigure1)C).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary shared stock distribution.
-Dashed lines represent the year by which countries have committed to reach the full implementation of the 2030 Agenda, as well as a mid and late change in range shift, respectively."}
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "The timeline is presented according to cumulative A.
+Number of stocks (cumulative), B.
+Number of EEZs and C.
+proportion of total stocks and EEZs.
+Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change  in range shift (Methods), respectively."}
 
@@
 
-The median year in which transboundary stocks experience a range shift varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Figure \@ref(fig:toeFigure2)A).
-Overall, most tropical EEZs will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia being significantly earlier (*p* < 0.05; see Table S\@ref(tab:toeTableS2) for test statistics) than almost all other regions (Fig.
-\@ref(fig:toeFigure2)B).
-In contrast, EEZs located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts (Table \@ref(tab:toeTableS2)).Stocks in countries such as Brazil and New Zealand are not projected to experience range shifts beyond natural variability (e.g., the transboundary index of a stock did not change beyond historical variability) (Fig.
+The median year in which transboundary stocks experience a range shift varied significantly according to the geographic region of the neighboring EEZs (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Figure \@ref(fig:toeFigure2)A)) and a species’ habitat association (Methods; *p* < 0.05; Table S\@ref(tab:toeTableS2)).
+Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; see Tables S\@ref(tab:toeTableS1, S\@ref(tab:toeTableS2 for full statistics) than almost any other region (Fig.
+\@ref(fig:toeFigure2)B, \@ref(fig:toeFigureSX)).
+In contrast, EEZs and stocks located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts (Table \@ref(tab:toeTableS2)).Stocks in countries such as Brazil and New Zealand are not projected to experience range shifts beyond natural variability (Fig.
 \@ref(fig:toeFigure2)A).

### 2021-03-29 Scripts/Manuscript.Rmd
Included first round of co-authors final comments

@@
 
-Most fisheries management is poorly adapted to address species’ shifting distributions as a result of human driven warmer, more acidic and less oxygenated ocean waters.
+Most fisheries management is poorly adapted to address species’ shifting distributions as a result of changing ocean conditions.
 Over the past century, human activities have dramatically altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters [@IPCC:2019tn].
@@
 As a result of changing ocean conditions, many marine species have shifted their distributions towards higher latitudes, deeper water or following local temperature gradients to remain within their optimal environmental “niche” [@Poloczanska:2016kk].
-The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt], impacting fisheries production and economies [@Sumaila:2019ew], and compromising our capacity to reach international sustainability goals within the 2030 Agenda for sustainable development such as Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
+The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt], impacting fish stocks, fisheries production and economies [@Sumaila:2019ew].
+These impacts are compromising our capacity to reach international sustainability goals under the 2030 Agenda for Sustainable Development such as Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
 The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management, including for species that cross international borders, also known as ‘shared stocks’ [@Pinsky:2018cb].
@@
 The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by Coastal States [@UnitedNations:1986tl].
-As defined by the United Nations' Food and Agriculture Organization, shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and (*iv*) discrete stocks that are only present in the high seas [@Munro:2004th].
+As defined by the United Nations’ Food and Agriculture Organization, shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and (*iv*) discrete stocks that are only present in the high seas [@Munro:2004th].
 This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
-Countries are responsible for the management of stocks within their EEZs.
-Under UNCLOS, EEZs are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl].
+Under UNCLOS, countries are responsible for the management of stocks within their EEZs and encouraged to cooperate when stocks are shared  [@UnitedNations:1986tl].
 A recent study estimates that there are 633 transboundary fish species globally, representing 67% of identified exploited taxa.
@@
 
-The effectiveness of fisheries management for transboundary species is challenged by shifts in stocks’ distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].
-In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+The effectiveness of fisheries management for transboundary species is challenged by shifts in stocks’ distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
 While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs].
@@
 Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (RCP8.5; Methods).
-We then estimate the year in which the 663 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variation and the intensity of this shift (i.e., time of emergence; see Methods).
-We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (15) and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
+We then estimate the year in which the 663 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variation and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods).
+We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
 
@@
 This corresponds to 68% of the studied stocks (Fig.
-\@ref(fig:toeFigure1)C).The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be 2036 (s.d.
+\@ref(fig:toeFigure1)C).
+The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be 2036 (s.d.
 of all species, $\pm$ 28 years).
-Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% having recorded a distributional shift in stocks by the UN 2030 Agenda’s deadline (Fig.
+Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% having recorded a distributional shift in stocks by the UN 2030 Agenda’s deadline (i.e., year 2030) (Fig.
 \@ref(fig:toeFigure1)C).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "The timeline is presented according to cumulative A.
-Number of stocks (cumulative), B.
-Number of EEZs and C.
-proportion of total stocks and EEZs.
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary shared stock distribution.
+(A) Number of stocks (cumulative) with their distribution centroids that have shifted beyond natural variations, (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one shifting stock, beyond natural variability.
 Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change  in range shift (Methods), respectively."}
@@
 
-<!-- ![Figure 1. Cumulative increase in time of emergence by Exclusive Economic Zone and transboundary species in absolute numbers and percentage.](../Figures/Fig2.png) -->
-
-The median year in which transboundary stocks experience a range shift varied significantly according to the geographic region of the neighboring EEZs (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Figure \@ref(fig:toeFigure2)A)) and a species’ habitat association (Methods; *p* < 0.05; Table S\@ref(tab:toeTableS2)).
-Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; see Tables S\@ref(tab:toeTableS1, S\@ref(tab:toeTableS2 for full statistics) than almost any other region (Fig.
-\@ref(fig:toeFigure2)B, \@ref(fig:toeFigureSX)).
-In contrast, EEZs and stocks located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts (Table \@ref(tab:toeTableS2)).Stocks in countries such as Brazil and New Zealand are not projected to experience range shifts beyond natural variability (Fig.
+The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; Tables S\@ref(tab:toeTableS1), S\@ref(tab:toeTableS2) and Fig.
+S\@ref(fig:toeFigureS1)) and geographic region of the neighboring EEZs (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Table S3 and Fig.
+2A).
+Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables S\@ref(tab:toeTableS2 and S\@ref(tab:toeTableS3) than almost any other region (Figs.
+\@ref(fig:toeFigure2)B, S\@ref(fig:toeFigureS1) and S\@ref(fig:toeFigureS2)).
+In contrast, EEZs and stocks located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts.
+Stocks in countries such as Brazil and New Zealand are not projected to experience range shifts beyond natural variability (Fig.
 \@ref(fig:toeFigure2)A).
@@
 ```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks.
-A) Land polygon shows the contribution of shifting stocks to the country’s total fishing revenue from transboundary stocks.
-Exclusive Economic Zone polygons represent average range shifts within them.
-EEZs with no distributional shifts between 2006 and 2100 are represented in aqua color.
-B) Year of shared distribution shifts by region according to United Nations sub-regions.
-Points represent the median year of range shift.
-Horizontal dashed line represents the year 2030.
+A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygons display average year of range shifts within them.
+EEZs with no distributional shift between 2006 and 2100 are represented in aqua color.
+B) Year of shared distribution shifts by UN sub-regions.
+Points color coded by the median year of range shift.
+Horizontal dashed line represents the year 2030 – the Sustainable Development Agenda deadline.
 N = North, S = South, W = West and E = East.
@@
 Aus and New Z.
-= Australia and New Zealand"}
+= Australia and New Zealand."}
 
@@
 
-We estimated the current (2005 and 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs.
-Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD in fishing revenue are projected to shift within the EEZs of 76 countries (Fig.
-\@ref(fig:toeFigure3)); with 4.2 billion USDs accounted by 321 stocks shifting in just eleven countries (Fig.
+We estimated the current (2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs.
+Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD in fishing revenue are projected to shift in the EEZs of 76 countries (Fig.
+\@ref(fig:toeFigure3)); with 4.2 billion USD accounted by 321 stocks shifting in just eleven countries (Fig.
 \@ref(fig:toeFigure3)).
-Most of the remaining countries in the top $75^{th}$ revenue percentile will see early shifts (e.g., before 2050) with the exception of Japan, Korea, Peru and Russia, who are projected to have later shifts (e.g., after 2050).
-In terms of the number of transboundary stocks shifting from historical variability, Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66), emerge as the top 5 countries (Fig.
+In terms of the number of transboundary stocks shifting from historical variability, Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66), emerge as the top five countries (Fig.
 \@ref(fig:toeFigure3)).
 
-```{r toeFigure3, eval = T,  echo = F, fig.width=10, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks per country.
+```{r toeFigure3, eval = T,  echo = F, fig.width=10, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks and corresponding fishing revenue per country.
 Countries are coded according to different colored symbols representative of 5 distinct regions.
-Horizontal line separating the top 75th percentile of annual fishing revenue.
-Showing country names for those within the top 75th percentile projected to  have an average shift within 5 years of the 2030 Agenda."}
+Horizontal line separates countries above and below the 75th percentile of annual fishing revenue.
+Showing country names for those within the top 75th percentile projected to have an average shift within 5 years after the 2030 deadline of the Sustainable Development Agenda.}
 
@@
 
-We based the transboundary index on a stock’s centroid defined as the average grid cells that contain the majority of a stock’s abundance (see Methods - Calculating an index of transboundary range shift).
+**Changes in the stock share ratio of transboundary stocks**
+
+<!-- We based the transboundary index on a stock’s centroid defined as the average grid cells that contain the majority of a stock’s abundance (see Methods - Calculating an index of transboundary range shift).
 Results from the sensitivity analysis suggest that our method is robust to the abundance level (e.g., the top 95%) of the stock (Fig.
@@
 In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks.
+SENT TO METHODS!!!!
+-->
 
-**Changes in the stock share ratio of transboundary stocks**
+We estimated the intensity of the distributional shift of transboundary stocks in terms of changes in catch proportion by 2030 (2020-2040) and 2050 (2040-2060), relative to the recent past (1951-2005).
+Findings show that by 2030, the global average catch proportion of transboundary stocks that are projected to change beyond the EEZ’s historic proportion is 59% ± 17% (Fig.
+\@ref(fig:toeFigure4)A).
+Moreover,  85% (n = 239) of the world EEZs would have experienced changes in catch proportion of transboundary stocks by 2030.
+Changes in catch proportion are expected to become more common and intense slightly towards the mid -21st century at both the stock and EEZ level (Figs.
+S\@ref(fig:toeFigureS3) and S\@ref(fig:toeFigureS4)).
 
-We identified transboundary stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point by the early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005).
-Findings show that by the early $21^{st}$ century, the global average catch proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 59% $\pm$ 17% (Fig.
-\@ref(fig:toeFigure5)A).
-By the early $21^{st}$, 85% (n = 239) of the world EEZs would have experienced such changes.
-Changes in stock share ratio are expected to increase slightly towards the mid 21st century at both the stock and EEZ level (Fig.
-\@ref(fig:toeFigure5)A and \@ref(fig:toeFigureS5)).
-
-```{r toeFigure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock proportion of neighboring Exclusive Economic Zones by 2030 (2021-2040) relative to 1951-2005.
 Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors (green to yellow).
-Panel B zooms into specific areas shown within the grey dashed line boundaries in A."}
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors.
+Panel B highlights changes for the Caribbean and Northwestern Europe."}
 
@@
 Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors (green to yellow).
-Panel B zooms into specific areas shown within the grey dashed line boundaries in A.](../Figures/Fig1.png) -->
+Panel B zooms into specific areas shown within the gray dashed line boundaries in A.](../Figures/Fig1.png) -->
 
-The direction and intensity of shafts in transboundary stocks are largely related to regional changes in biogeography and the geometry of the EEZs (Fig.
-\@ref(fig:toeFigure5)A).
-In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the stock share ratio is expected to follow a projected poleward shift.
+The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of the EEZs (Fig.
+\@ref(fig:toeFigure4)A).
+In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the proportion of the  stock is expected to follow a projected poleward shift.
 However, shifts along the coasts of Pacific Central America and West Africa occur along an equatorial direction.
@@
 Regions where EEZs are relatively small and have multiple borders (e.g., Caribbean or Northwestern Europe) are expected to have a particularly complex exchange of stock share ratio with no established pattern (Fig.
-\@ref(fig:toeFigure5)B).
+\@ref(fig:toeFigure4)B).
 
@@
 
-Early shifts in the shared distribution of transboundary stocks concur with previous studies that have detected and attributed changes in marine catch composition [e.g., @Cheung:2013gk; @Frainer:2017ji; @Last:2011fd].
+Our findings highlighting early shifts in the shared distribution of transboundary stocks concur with previous studies that have detected and attributed changes in marine catch composition [e.g., @Cheung:2013gk; @Frainer:2017ji; @Last:2011fd].
 For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US) as a response to climatic, oceanographic and ecological changes [@Zeidberg:2007gs].
 A new fishery targeting Humboldt squid quickly developed on the heels of the species’ range expansion [@Pinsky:2014by].
-In the northeast Atlantic, the Atlantic mackerel (*Scomber scombrus*) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
+In the northeast Atlantic, fisheries on Atlantic mackerel (*Scomber scombrus*) are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
 However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the fishery’s total allowable catch and a further 18% in 2008, without consultation with NEAFC, threatening the sustainability of the stock [@Spijkers:2017ij].
 These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij].
-As transboundary stocks continue to shift, international disputes like that of Atlantic mackerel are likely to increase in frequency requiring a revision of management plans [@Pinsky:2018cb; @Engler:2020bg].
+As transboundary stocks continue to shift, such international disputes are likely to increase in frequency and requiring a revision of management plans [@Pinsky:2018cb; @Engler:2020bg].
 
@@
 
-Our results underscore regional "hotspots" of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figures \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure5)A).
-Such regions, like the Caribbean, are characterized by high levels of warming relative to natural internal variability [@IPCC:2019tn; @Hawkins:2012hc] and high vulnerability of species to warming waters [@IPCC:2019tn].
-Moreover, they represent relatively small EEZs that border multiple countries (e.g., the Caribbean).
+Our results identify regional “hotspots” of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figures \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure4)A).
+Such regions, like the Caribbean, are characterized by high levels of warming relative to natural internal variability [@IPCC:2019tn; @Hawkins:2012hc] and species’ high vulnerability to warming waters [@IPCC:2019tn].
+Moreover, they represent relatively small EEZs that border multiple countries.
 Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting transboundary stocks.
 
-Management plans that are not prepared to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
-Thus, focusing on these “hotspots'' is key to anticipate a potential increase in fisheries conflicts in the coming years.
-Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments [@Tunca:2019if], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@Pinsky:2018cb].
-For example, countries that are Party to the Nauru Agreement (PNA) manage their fisheries under a  Vessel Day Scheme.
-Because this scheme takes into account shifts in stocks and catches, it is a good example of a management system that is somewhat anticipatory [@Aqorau:2018bh].
-However, even strategies that take into consideration  species’ shift within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+Management plans that are not designed to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
+Thus, focusing on these “hotspots” is key to anticipate a potential increase in fisheries conflicts in coming years.
+Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments, which do not have to be monetary [@Tunca:2019if; @Miller:2013iv], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@PalaciosAbrantes:2020vt; @Pinsky:2018cb].
+For example, countries that are Party to the Nauru Agreement (PNA) manage their fisheries under a so-called Vessel Day Scheme, which accounts for shifts in stocks and catches, and is therefore somewhat anticipatory [@Aqorau:2018bh].
+However, even strategies that take into consideration species’ shifts within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
 Quota allocation methods based on a stock’s current distribution and/or on fixed-historical proportion need to evolve to be more agile [@Sumaila:2020tp; @PalaciosAbrantes:2020vt].
 Fisheries based on historic participation could move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-However, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from the fishery.
-Side payments are mechanisms that can provide a solution to the inequality produced by shifts in a stock’s distribution [@Tunca:2019if].
-These can, but do not have to, be monetary (e.g., the stock-“winning” state compensates the stock-“losing” state for the proportional shift, a conservation fund paid mainly by one Party) [@Miller:2013iv].
+However, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery.
 
@@
 
-The early projected range shifts of a large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability (e.g., SST which is projected to increase historical variability by 2030 in 70% of the global ocean under RCP 8.5 [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
-Results from multiple marine ecosystem models show that SST and NPP are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
-Specifically for the DBEM, SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results.
-The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2019tn], while also explaining the later, and sometimes non-existent, shift at higher latitudes (Fig.
+The early projected range shifts of a large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability e.g., SST is projected to increase beyond historical variability by 2030 in 70% of the global ocean under RCP 8.5 [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
+Results from multiple marine ecosystem models show that sea surface temperature (SST) and primary production (NPP) are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].Specifically, in the species distribution model used here (see methods), SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results.
+The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2019tn], while also explaining the later, and sometimes non-existent, shifts at higher latitudes (Fig.
 \@ref(fig:toeFigure2)B).
@@
 First, our analysis is based on a high emission scenario (RCP 8.5) representing an extreme case of shifts in the shared distribution of transboundary species.
-Thus, any climate change mitigation pathway (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].
-Second, a substantial source of uncertainty is related to model selection for both climate change and fisheries distribution.
+Thus, any mitigation efforts (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the 21^st century [@Frolicher:2016dg].
+Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
 This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
 While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available [@Deser:2020bj] and cannot address limitations of process level understanding.
-As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties, and will likely continue to do so in the coming decade.
-This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf] and where internal variability (e.g., El Niño Southern Oscillation - ENSO- for Pacific nations) is also important .
+As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties and will likely continue to do so in the coming decade.
+This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf] and where internal variability (e.g., El Niño Southern Oscillation - ENSO- for Pacific nations) is also important.
 These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh].
-Overall, changes in fish biomass with multiple upper trophic level models, including the DBEM, show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@BryndumBuchholz:2019kh].
-Further research that reduces uncertainty in the response of the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to a larger structural uncertainty of fish and fisheries models [@Lotze:2019ce].
+Overall, changes in fish biomass with multiple upper trophic level models show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@Lotze:2019ce].
+Further research that reduces uncertainty in the response of the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to a smaller structural uncertainty of fish and fisheries models [@BryndumBuchholz:2019kh].
 Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks.
 While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca].
-Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
+Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identity different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
 Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
@@
 
-Th global community has set the ambitious goal of managing all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear benefits for a number of other societal goals [@UnitedNations2018; @Singh:2017ds].
+The global community has set the ambitious goal of managing all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear benefits for a number of other societal goals [@UnitedNations2018; @Singh:2017ds].
 Preparing anticipatory policies to deal with shifting transboundary stocks is key to achieving the SDGs and ensuring effective governance of the world’s fisheries [@Pinsky:2018cb; @Oremus:2020id; @Pecl:2017hu].

### 2021-03-29 Scripts/Manuscript.Rmd
fixed bugs

@@
 Horizontal line separates countries above and below the 75th percentile of annual fishing revenue.
-Showing country names for those within the top 75th percentile projected to have an average shift within 5 years after the 2030 deadline of the Sustainable Development Agenda.}
+Showing country names for those within the top 75th percentile projected to have an average shift within 5 years after the 2030 deadline of the Sustainable Development Agenda."}
 
@@
 
-img <- readPNG("../Figures/Fig5.png")
+img <- readPNG("../Figures/Fig4.png")
 grid.raster(img)
@@
 
-```{r supplements, child = '../Scripts/Supplements.Rmd', eval = T}
+```{r supplements, child = '../Scripts/Supplements.Rmd', eval = F}
 

### 2021-03-29 Scripts/Supplements.Rmd
William comments from first round of co-authors

--- 2021-03-19 Scripts/Supplements.Rmd
@@
 ---
+title: "Supplementary Materials for Imminent range shifts in transboundary fish stocks challenge their management under climate change"
+author: 'Juliano Palacios-Abrantes, Thomas L. Frölicher, Gabriel Reygondeau, U. Rashid Sumaila, Alessandro Tagliabue, Colette C.C. Wabnitz, and William W.L. Cheung'
 output:
@@
 csl: ../References/science.csl
-header-includes: #allows you to add in your own Latex packages
+header-includes:
 - \usepackage{float} #use the 'float' package
@@
 - \usepackage{caption}
-- \captionsetup[figure]{name = Figure S3}
-- \captionsetup[table]{name = Table S3}
+- \captionsetup[figure]{name = Figure S}
+- \captionsetup[table]{name = Table S}
 ---
@@
 
-This analysis includes 633 exploited marine transboundary species that count for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-Each time a species was shared by a pair of neighboring  EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.
-Using the transboundary species Atlantic cod (*Gadus morhua*) as an example and this study’s definition of transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not.
+The analyses detailed herein are based on 633 exploited marine transboundary species that account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
+When a species was shared by a pair of neighboring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.
+Using the transboundary species Atlantic cod (*Gadus morhua*) as an example and this study’s definition of transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not as they do not have juxtaposing EEZs.
 We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015; http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
-\@ref(fig:toeFigureS1)) including island territories.
+\@ref(fig:toeFigureS5)) including island territories.
+.
 We determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp].
@@
 
-We forced the DBEM with simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
-The GFDL-ESM2M is run under historical forcing until 2005 and follows the high greenhouse gas emissions scenario, the Representative Concentration Pathway 8.5 (RCP 8.5) over the 2006-2100 period [@Riahi:2011dk].
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a stocks’ distribution, we have to understand distribution variability during both the historical and the future periods, to infer differences between time frames.
-We do this by following an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
-Thus, for our experiment, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in year 1950 resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
-By design, variations among ensemble members are then solely due to natural variability.
-This approach allows us to estimate the natural internal variability (e.g., background noise) and isolate the forced anthropogenic climate signal of a stock’s distribution by averaging the secular trend over all ten ensemble members.
-
 _**Projecting species distributions under climate change**_
@@
 For pelagic species, the model uses environmental variables at the surface whereas demersal and benthic species’ distributions are driven by ocean bottom variables.
-The DBEM then estimates species’ abundance and maximum catch potential (a proxy for maximum sustainable yield) over a regular spatial grid of 0.5 of latitude per 0.5 of longitude (see Cheung et al.
-2009, 2010 and 2016 for model details).
+The DBEM then estimates species’ abundance and maximum catch potential (a proxy for maximum sustainable yield) over a regular spatial grid of 0.5 of latitude per 0.5 of longitude (see @Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd for model details).
 Importantly, DBEM is able to project catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
@@
 
-We forced the DBEM with simulated ocean conditions from a ten member ensemble simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distribution of transboundary stocks from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
-The GFDL-ESM2M is run under historical forcing until 2005 and follows the high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5) from 2006 to2100 [@Riahi:2011dk].
+ DBEM was driven by simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+The GFDL-ESM2M was run under historical forcing until 2005 and follows a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5), over the 2006-2100 period [@Riahi:2011dk].
 We chose this time period to match the historical period in the GFDL ESM2M simulations over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Rodgers:2015jq].
-Because our main approach lies in identifying when the future shared distribution of a stock undergoes a substantial change in reference to historical variability, we follow an ensemble member approach where each member constitutes a different realization of the Earth system variability condition [@Rodgers:2015jq; @Frolicher:2009ge].
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a stock’s distribution, we have to understand distribution variability during both historical and future periods, to infer differences between time frames.
 To this end, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in 1950, resulting in a unique atmosphere and ocean state at each point in time after about three years for surface and eight years for subsurface waters [@Frolicher:2020cc].
 By design, variations among ensemble members are then solely due to natural internal variability (e.g., different phases of the El Niño Southern Oscillation-ENSO).
+Thus, for our experiment, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in the year 1950, resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
 
@@
 This index represents the shift in the distribution centroid of a transboundary stock relative to the centroid of neighboring EEZs that share this stock (Fig.
-\@ref(fig:toeFigureS2)).
-The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across grid cells within the neighboring EEZs sharing the stock.
+\@ref(fig:toeFigureS6)).
+The centroid of a transboundary stock was determined by the average ($\mu$) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across grid cells within the neighboring EEZs sharing the stock.
 Therefore,
@@
 
-where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance.
-To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighboring  EEZs sharing the stock was above the 95th percentile.
-A sensitivity analysis using a subset of species (*n* = 34) for all EEZs (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) to examine the effects of different thresholds on the calculated index value showed no apparent difference between percentiles (Fig.
-\@ref(fig:toeFigureS3)).
+where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the specified percentile ($per^{th}$) of the projected transboundary stock abundance.
+To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighboring EEZs sharing the stock was above the $95^{th}$ percentile.
+A sensitivity analysis using a subset of species (n = 34) for all EEZs  to examine the effects of different thresholds (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) on the calculated index value showed no apparent difference between percentiles (Fig.
+\@ref(fig:toeFigureS7)).
 The centroid of each EEZ was estimated using the st package in R (Fig.
-\@ref(fig:toeFigureS1)).
-For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids (e.g., EEZ and stock) using the geosphere package in R assuming the Earth is a perfect sphere and ignoring geographic barriers;
+\@ref(fig:toeFigureS5)).
+For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids (i.e., EEZ and stock) using the geosphere package in R assuming the Earth is a perfect sphere and ignoring geographic barriers;
 
@@
 
-where $lat_{eez}$ and lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively, and $lon_{eez}$ and $lon_{ts}$ are associated longitudes.
+where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively, and $lon_{eez}$ and $lon_{ts}$ are associated longitudes.
 Then, for each year and for each ensemble member between 1951 and 2100 we calculated the transboundary index (TI) as follows:
@@
 
-where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring  EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (*th*, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (th, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
 Thus, a higher TI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels.
@@
 In contrast, a decreasing TI suggests that the sharing of the stocks between neighboring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries.
-The index was smoothed to a 10-year average to reduce interannual variability.
+The index was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg].
 
@@
 
-Knowing the point in time at which the distribution of a shared stock will diverge from historical variability is important to inform decision makers when climate adaptation measures will need to be implemented [@Link:2010ei].
-To that end, we adopted the concept of time of emergence, defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (e.g., historical natural variation) [@Hawkins:2012hc], and commonly applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev].
+Knowing the point in time at which the distribution of a shared stock will diverge from historical variability is important to inform decision makers by when, ideally, climate adaptation plans will need to have been implemented [@Link:2010ei].
+To that end, we adopted the concept of time of emergence, commonly applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev] and defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (i.e., historical natural variation) [@Hawkins:2012hc].
 The premise behind time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
 The TI’s natural internal variability was estimated from 1951 to 2005 ($TI_\sigma$) across ensemble members, and the TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$).
-We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See (Fig.
-\@ref(fig:toeFigureS2)) for a graphical description):
+We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig.
+\@ref(fig:toeFigureS6) for a graphical description):
 
@@
 
-Thus, we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TI “emerges” from historical internal variability (t’).
-This assumes that, from a fisheries’ perspective, the first year a stock’s  distribution shifts from a 10 year average is sufficient to call for caution.
+Thus, we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TI “emerges” from historical internal variability (*t’*).
+It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution.
 We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
-We performed two non-parametric Kruskal–Wallis tests by ranks [@Hollander:2013wz] to investigate geopolitical and ecological differences in the range shift of transboundary species (See supplementary information for complete statistical analysis).
+Sensitivity tests to a lower threshold of 1 s.d.
+(or 68% confidence) resulted in a global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
+\@ref(fig:toeFigureS8)).
+In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
+\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
 
@@
 The concept of threat point comes from Game Theory and has been widely used in shared fisheries management [@Bailey:2010ga; @Sumaila:2013hv; @Sumaila:2020tp; @Clark:1975uu; @Munro:1979dg].
-In a game theoretic model, a player’s strategy (e.g., to act cooperatively or not) will have direct consequences on other players, which, in turn, will affect the overall outcome of the game [@Bailey:2010ga].
+In a game theoretic model, a player’s strategy (e.g., to act cooperatively or not) will have direct consequences for other players, which, in turn, will affect the overall outcome of the game [@Bailey:2010ga].
 Cooperation will often result in the maximization of benefits for the system rather than for individual players.
-However, for a cooperation strategy to work, the benefit a player gets must not be less than the benefit under a non-cooperation strategy.
-Thus, the “threat point” is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr].
-
-We define the threat point as the minimum required catch proportion of a transboundary stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+However, for a cooperative strategy to work, the benefit a player gets must not be less than the benefit under a non-cooperative strategy.
+Thus, the “threat point” is defined as the minimum payoff a player is willing to receive to cooperate in a game-theoretic model [@Nash:1953vr].
+We define the threat point as the minimum required catch proportion of a transboundary stock within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
 Any proportion below the defined threat point would result in the unilateral management of the stock.
 First, we estimated the stock share ratio (SSR) of each transboundary stock shared by neighboring nations between 1951 and 2100.
-We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ [@PalaciosAbrantes:2020tv].
-Second, to reduce the effects of variability, we averaged the calculated proportion of stock occupying a given EEZ into three time periods: the historical time period ($SSR_{th}$) from 1951 to 2005, and two future periods, the early 21st century, between 2021 and 2040 ($SSR_{te}$), and the mid 21st century, defined as the average of 2041 to 2060 ($SSR_{tm}$).
-These future time periods were chosen to correspond to the challenges of achieving the fisheries-related UN-SDGs by 2030 (i.e., the 2030 Agenda), such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) and SDG 1.2 (poverty reduction) [@Singh:2017ds; @UnitedNations:1Qe8KIcN].
+We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock's catch held within each EEZ [@PalaciosAbrantes:2020tv].
+Second, to reduce the effects of variability, we averaged the calculated proportion of stock occupying a given EEZ into three time periods: the historical time period ($SSR_{th}$) from 1951 to 2005, and two future periods, the early $21^{st}$ century, between 2021 and 2040 ($SSR_{te}$), and the mid $21^{st}$ century, defined as the average of 2041 to 2060 ($SSR_{tm}$).
+These future time periods were chosen to correspond to the challenges of achieving fisheries-relevant UN-SDGs by 2030 (i.e., the 2030 Agenda), such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) and SDG 1.2 (poverty reduction) [@Singh:2017ds; @UnitedNations:1Qe8KIcN].
 The analysis was replicated for projected stock distributions from each of the ten ensemble members and results were averaged across ensemble members.
@@
 Thus, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds two standard deviations of the historical variations in SSR (i.e., when $SSR_{et} \geqq (SSR_{th} + 2 \sigma$) or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
-Finally, we estimated the percentage change in SSR ($\Delta SSR$) of each future time period ($SSR_{tf}$) relative to the historic time period $SSR_{th}$ for each stock whose share ratio exceeded the threat point following Palacios-Abrantes et al.
-(2020);
+Finally, we estimated the percentage change in SSR ($\Delta SSR$) of each future time period ($SSR_{tf}$) relative to the historic time period $SSR_{th}$ for each stock whose share ratio exceeded the threat point following [@PalaciosAbrantes:2020tv];
 
@@
 Specifically, we tested if the habitat preference of transboundary species and the geographic location of EEZs would have any effect on the transboundary stocks distributional range shift.
-In both cases, our null hypothesis was that there were no significant differences in the year of shift across habitat association nor EEZs.
+In both cases, our null hypothesis was that there were no significant differences in the year of shift across habitat association or EEZs.
 All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages data.table [@Packagedatatable:2019uh], janitor [@PackagejanitorSim:2018ur], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], gmt [@PackagegmtInterfa:2017vh] and zoo [@PackagezooSInfr:2019tm], ggrepel [@PackageggrepelAut:2020to], zeallot [@PackagezeallotMul:2018wy], viridis [@PackageviridisDef:2018us], cowplot [@CowplotStreamlined:2019wt], moments [@PackagemomensMome:2015uh] and pgirmess [@PackagepgirmessSp:2018um].
@@
 
-```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
+```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
 Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
@@
 
-img <- readPNG("../Figures/Fig4.png")
+img <- readPNG("../Figures/FigS1.png")
 grid.raster(img)
@@
 
-```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap ="World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their centroids (points)."}
+\clearpage
 
-img <- readPNG("../Figures/Supp1.png")
+```{r toeFigureS2, eval = T,  echo = F, fig.width=12, fig.height=7, fig.cap =".
+Year of change in the shared distribution of 5,745 transboundary stocks using a one standard deviation threshold.
+A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygons display average year of range shifts within them.
+EEZs with no distributional shift between 2006 and 2100 are represented in aqua color.
+B) Year of shared distribution shifts by UN sub-regions.
+Points color coded by the median year of range shift.
+Horizontal dashed line represents the year 2030 – the Sustainable Development Agenda deadline.
+N = North, S = South, W = West and E = East.
+Ltn.
+Ame.
+and the Car.
+= Latin America and the Caribbean.
+Aus and New Z.
+= Australia and New Zealand."}
+
+img <- readPNG("../Figures/FigS2.png")
 grid.raster(img)
@@
 
-```{r toeFigureS2, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap = "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
-The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring  Exclusive Economic Zones sharing the stock (Top panel).
-Time of emergence is defined as the first year when the TI overpasses historical values (Bottom panel)."}
+```{r toeFigureS3, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Distribution of changes in transboundary stock proportion per stock and neighboring Exclusive Economic Zones.
+Yellow color represents changes by early, 2030 (2021-2040) and purple by mid 2050 (2041-60) 21st century relative to 1951-2005."}
 
-img <- readPNG("../Figures/Supp2.png")
+img <- readPNG("../Figures/FigS3.png")
 grid.raster(img)
@@
 
-```{r toeFigureS3, eval = T,  echo = F, fig.width=14, fig.height=7, fig.cap ="Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine the stock’s centroid.
-Top Row: results in absolute numbers.
-Bottom row: results expressed as a percentage of the total emerging stocks and EEZ with emerging stock."}
+```{r toeFigureS4, eval = T,  echo = F, fig.width=14, fig.height=7, fig.cap ="Changes in stock proportion of neighboring Exclusive Economic Zones by 2050 (2041-2060) relative to 1951-2005.
+Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors.
+Panel B highlights changes for the Caribbean and Northwestern Europe."}
 
-img <- readPNG("../Figures/Supp3.png")
+img <- readPNG("../Figures/FigS4.png")
 grid.raster(img)
@@
 
-```{r toeFigureS4, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Average and standard deviation of distribution of mean year of range shift for a 95% confidence (yellow) and 68% (purple)"}
+```{r toeFigureS5, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their estimated centroids (black points)."}
 
-img <- readPNG("../Figures/Supp4.png")
+img <- readPNG("../Figures/FigS5.png")
 grid.raster(img)
@@
 
-img <- readPNG("../Figures/Supp5.png")
+img <- readPNG("../Figures/FigS5.png")
 grid.raster(img)
@@
 
-<!-- ![Figure S2. Distribution of Stock Share Ratio percentage change by the early (2021-2040) and middle (2040-2060) $21^{st}$ centurury relative to today (1951-2005))](../Figures/Supp2.png) -->
 \clearpage
+
+```{r toeFigureS6, eval = T,  echo = F, fig.width=14 fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring Exclusive Economic Zones sharing the stock (Top panel).
+Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel)."}
+
+img <- readPNG("../Figures/FigS6.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigureS7, eval = T,  echo = F, fig.width=14 fig.height=10, fig.cap =  "Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine the stock’s centroid.
+Top Row: results in absolute numbers.
+Bottom row: results expressed as a percentage of the total emerging stocks and EEZs with emerging stock."}
+
+img <- readPNG("../Figures/FigS7.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigureS7, eval = T,  echo = F, fig.width=6 fig.height=6, fig.cap =  "Average distribution of transboundary stock’s year of change in their shared distribution.
+Left panel is the average year across stocks and right panel the standard deviation.
+Yellow represents a one standard deviation distance from the historical average and purple represents a two standard deviation distance from the historical average."}
+
+img <- readPNG("../Figures/FigS8.png")
+grid.raster(img)
+
+```
 
@@
 
+```{r toeTableS3, eval = T, echo = F}
+
+read.csv("~/GitHub/TransEmergence/Tables/kw_habitat.csv") %>% 
+  select(-statistic,
+         "Sub region comparrison" = comparrison,
+         "Significan level" = signif.level,
+         "Observation difference" = dif.com.obs.dif,
+         "Critical difference" = dif.com.critical.dif,
+          "K-W Difference" = dif.com.difference) %>% 
+  kable(
+        booktabs = T,
+        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+  ) #%>% 
+  # kable_styling(full_width = F,
+                # position = "center")
+
+```
+
+\clearpage
+
 ```{r toeTableS2, eval = T, echo = F}
@@
 ```
-
-\clearpage
-
-```{r toeTableS3, eval = T, echo = F}
-
-read.csv("~/GitHub/TransEmergence/Tables/kw_habitat.csv") %>% 
-  select(-statistic,
-         "Sub region comparrison" = comparrison,
-         "Significan level" = signif.level,
-         "Observation difference" = dif.com.obs.dif,
-         "Critical difference" = dif.com.critical.dif,
-          "K-W Difference" = dif.com.difference) %>% 
-  kable(
-        booktabs = T,
-        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  ) #%>% 
-  # kable_styling(full_width = F,
-                # position = "center")
-
-```

### 2021-03-29 Scripts/Supplements.Rmd
William comments for co-authors

@@
 Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
-The habitat preference of each species was determined following FishBase (http://www.fishbase.org) for fish species and SeaLifeBase (http://www.fishbase.org) for invertebrates (Fig.
-\@ref(tab:toeTableS1)).
+The habitat preference of each species was determined following FishBase (http://www.fishbase.org) for fish species and SeaLifeBase (http://www.fishbase.org) for invertebrates (Table S\@ref(tab:toeTableS1)).
 For each stock and EEZ, we used Sea Around Us data to estimate catch and fishing revenue from fishing activities.
@@
 
-<!-- ![toeFigure 4.
-Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
-Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
-The category “Other” consists of species that have no “habitat preference” classification.
-The number of stocks included in this analysis for each habitat type is noted in parentheses.
-Dashed whiskers represent the 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Solid line represents the median time of emergence per habitat preference.](../Figures/Fig5.png) -->
-
 \clearpage
@@
 
-```{r toeFigureS5, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "Changes in Stock Share Ratio by 2050 (2041-2060) relative to 1951-2005.
-Lines represent the average gain of transboundary stock share ratio with arrows going from the losing EEZ (point) to the winner (arrowhead).
-Polygons represent the percentage of initial stocks that will see a shift.
-Panel B zooms into specific areas shown in grey boxes of A."}
-
-img <- readPNG("../Figures/FigS5.png")
-grid.raster(img)
-
-```
-
-\clearpage
-
-```{r toeFigureS6, eval = T,  echo = F, fig.width=14 fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+```{r toeFigureS6, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
 The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring Exclusive Economic Zones sharing the stock (Top panel).
@@
 
-```{r toeFigureS7, eval = T,  echo = F, fig.width=14 fig.height=10, fig.cap =  "Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine the stock’s centroid.
+```{r toeFigureS7, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine the stock’s centroid.
 Top Row: results in absolute numbers.
@@
 
-```{r toeFigureS7, eval = T,  echo = F, fig.width=6 fig.height=6, fig.cap =  "Average distribution of transboundary stock’s year of change in their shared distribution.
+```{r toeFigureS8, eval = T,  echo = F, fig.width=6, fig.height=6, fig.cap ="Average distribution of transboundary stock’s year of change in their shared distribution.
 Left panel is the average year across stocks and right panel the standard deviation.

### 2021-04-05 Scripts/Initial_Analysis.Rmd
Changes for submission

@@
 
-Transboundary_spp %>% 
+n_stocks <- Transboundary_spp %>% 
   left_join(Neighbours,
@@
   group_by() %>% 
-  summarise(t_trans_stocks = sum(n_stock)) # 9132
+  summarise(t_trans_stocks = sum(n_stock)) %>% 
+  pull(t_trans_stocks) # 9132
 
@@
 
-# Join all data for analysis
-
-Main_data <- Eme_Res %>% # 19728
-  filter(tresh == "Tresh_one") %>% # 11490
-  left_join(Spp_depth) %>% 
-  select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel,un_region= un_geo_region) %>% 
-  left_join(Toe_temp,
-            by = c("eezid","tresh")
-            ) %>% 
-  left_join(Catch_trend,
-            by = c("taxon_key","sau_fishing_entity","fishing_entity_id")
-            ) %>% # 11492
-  left_join(Eez_biogeo,
-            by = "eezid") %>% 
-  left_join(Spp_env_pref,
-            by = "taxon_key") %>% 
-  left_join(Eez_centroids,
-            by = "eezid") %>% 
-  mutate(latitude = abs(latitude))
+#
+
+# See chunk Emerging_control_panel
+n_stock <- 9132
+n_eez <- length(unique(Eez_centroid_sf$eez_name))
 
@@
     emerging_stocks = sum(emerging_stocks),
-    prop_total = (emerging_stocks/9132)*100,
+    prop_total = (emerging_stocks/n_stock)*100,
     )
@@
 
+# By 2030
+Eme_Res %>% 
+filter(emerging_yr >= 2030) %>% 
+  group_by(tresh,neighbour_id) %>% 
+  summarise(
+    emerging_stocks = length(unique(taxon_key))
+    ) %>% 
+  group_by(tresh) %>% 
+  summarise(
+    emerging_stocks = sum(emerging_stocks),
+    prop_total = (emerging_stocks/n_stock)*100
+    )
+
 ```
@@
 
-max(stock_d$com_sum)
+# max(stock_d$com_sum)
 
@@
 
-max(eez_d$com_sum)
+# max(eez_d$com_sum)
 
@@
     # total = ifelse(level == "stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$sf_eez_name))),
-    total = ifelse(level == "stock",max(stock_d$com_sum),max(eez_d$com_sum)),
+    total = ifelse(level == "stock",n_stock,n_eez),
     per = com_sum/total*100
@@
 
-# Manually include rest of data untill 2100
+# Manually include rest of data until 2100
 cont_eez <- plot_data %>% 
@@
   mutate(
-    value = ifelse(metric == "per",100,228),
+    value = ifelse(metric == "per",81,228),
     taxon_emer = c(seq(2088,2099,1),seq(2088,2099,1))
@@
   ) +
-  my_ggtheme_p() +
-  theme(legend.position = "") +
-  guides(fill=guide_legend(title= "Level", reverse=FALSE)) +
+  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
   scale_fill_viridis(discrete = T,
@@
   ) +
-  my_ggtheme_p(leg_pos = "") +
-  guides(fill=guide_legend(title= "Level", reverse=FALSE)) +
+  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
   scale_fill_viridis(discrete = T,
@@
               x = x-4,
-              y = 18,
+              y = 14,
               label = label
@@
   ) +
-  my_ggtheme_p(leg_pos = "right") +
-  guides(fill=guide_legend(title= "", reverse=FALSE)) +
+  my_ggtheme_p(leg_pos = "right", hjust = 0.5) +
   scale_fill_viridis(discrete = T,
@@
   plot_all <- ggdraw() +
-    # Latin America and the Car.
+    # Count stock plot
     draw_plot(count_stock_plot, x = 0.01, y = 0, width = 0.29, height = 0.90) +
-    # Europe
+    # Count EEZ plot
     draw_plot(count_eez_plot, x = 0.31, y = 0, width = 0.29, height = 0.90) +
@@
                     size = 20,
-                    x = c(0.01, 0.285,0.56), 
+                    x = c(0.01, 0.285,0.59), 
                     y = c(1,1,1)
@@
 # coord_sf(xlim = c(-3, 35),ylim = c(30, 45)) +
-Rectangles <- tibble(
-    x1=c(-100,-14), #LatAm, Med, NE
-    x2=c(-55,23),
-    y1=c(9,50), 
-    y2=c(30,67)
-  )
+# Rectangles <- tibble(
+#     x1=c(-100,-14), #LatAm, Med, NE
+#     x2=c(-55,23),
+#     y1=c(9,50), 
+#     y2=c(30,67)
+#   )
 
@@
     # Closeup Rectangles 
-    geom_rect(data=Rectangles,
-              aes(xmin=x1, 
-                  xmax=x2,
-                  ymin=y1,
-                  ymax=y2),
-              color="grey50", 
-              fill = NA,
-              linetype = "dashed"
-    ) +
+    # geom_rect(data=Rectangles,
+    #           aes(xmin=x1, 
+    #               xmax=x2,
+    #               ymin=y1,
+    #               ymax=y2),
+    #           color="grey50", 
+    #           fill = NA,
+    #           linetype = "dashed"
+    # ) +
     my_ggtheme_m() +

### 2021-05-03 Scripts/Manuscript.Rmd
misc

--- 2021-03-29 Scripts/Manuscript.Rmd
@@
 title: "Imminent range shifts in transboundary fish stocks challenge their management under climate change"
-author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette C.C. Wabnitz^1,5^, and William W.L. Cheung^1^'
-date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
+author: "SUBMITTED MANUSCRIPT. DO NOT SHARE."
+date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
 ^2^Climate and Environmental Physics, Physics Institute, University of Bern.
@@
 ^4^School of Environmental Sciences, University of Liverpool.
-^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States'
-output: bookdown::word_document2
+^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States"
+subtitle: "Juliano Palacios-Abrantes*^1^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette C.C. Wabnitz^1,5^, and William W.L. Cheung^1^"
+output: bookdown::pdf_document2
 reference_docx:  ../References/sci_template.docx
-  # pdf_document: default
 bibliography: ../References/ToE_Ref.bib
 csl: ../References/science.csl
+header-includes:
+  - \usepackage{setspace}
+  - \usepackage{caption}
+  - \usepackage{float} # Use the 'float' package
+  - \floatplacement{figure}{H} # Make every figure with caption = h
+  - \captionsetup[figure]{name = Fig}
 ---
+
+\setlength{\parindent}{4em} 
+\linespread{1}
+\doublespacing
 
@@
 
-corresponding author: [j.palacios\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}
+corresponding author: [j.palacio\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}
 
@@
 
-Climate change is shifting the distribution of transboundary fish stocks between neighbouring countries.
+Climate change is shifting the distribution of transboundary fish stocks between neighboring countries.
 The time-frame of such shifts determines the extent of climate change challenges to international fisheries governance.
-Using global climate-fisheries model simulations, we show that, within the next 15 years, 68% of the world’s transboundary fish stocks are projected to shift their distributions across international boundaries.
-By 2100, such shifts will occur in all countries under a ‘no mitigation’ scenario.
-Hotspots of transboundary shifts occur in countries that are highly dependent on fisheries for income, food and livelihoods.
+Using global climate-fisheries model simulations, we show that, by 2030, transboundary fish stocks are projected to shift their distributions across 78% of the world’s international boundaries.
+By 2100, 45% of stocks will have shifted for 81% of countries under a ‘no mitigation’ scenario.
+Hotspots of transboundary shifts occur for countries that are highly dependent on fisheries for income, food and livelihoods.
 The imminent onset of transboundary shifts highlights the urgency to re-formulate international fisheries policies to limit conflict over these fish stocks.
-Bi- and multi-lateral actions should start now as lead-time for policy change is long.
+Bi- and multi-lateral actions are urgently needed as lead-time for policy change is long.
 
-**One-Sentence Summary**
+<!-- **One-Sentence Summary** -->
 
-Rapid climate-induced shifts in transboundary fish stocks
+<!-- Rapid climate-induced shifts in transboundary fish stocks challenges international fisheries management -->
+
+\clearpage
 
@@
 The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt], impacting fish stocks, fisheries production and economies [@Sumaila:2019ew].
-These impacts are compromising our capacity to reach international sustainability goals under the 2030 Agenda for Sustainable Development such as Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
+These impacts are compromising our capacity to reach international sustainability goals under the United Nations (UN)  2030 Agenda for Sustainable Development such as Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
 The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management, including for species that cross international borders, also known as ‘shared stocks’ [@Pinsky:2018cb].
@@
 The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by Coastal States [@UnitedNations:1986tl].
-As defined by the United Nations’ Food and Agriculture Organization, shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and (*iv*) discrete stocks that are only present in the high seas [@Munro:2004th].
+As defined by the UN's Food and Agriculture Organization, shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and (*iv*) discrete stocks that are only present in the high seas [@Munro:2004th].
 This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
@@
 
-The effectiveness of fisheries management for transboundary species is challenged by shifts in stocks’ distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+The effectiveness of fisheries management for transboundary species is challenged by shifts in stocks’ distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].
+In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
 While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs].
 Such patterns are expected to be exacerbated under continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
-The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the Sustainable Development Goals (SDG) under the United Nations 2030 Agenda.
+The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the Sustainable Development Goals (SDG) under the UN 2030 Agenda.
 Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the goals set out under the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].
 
-Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (RCP8.5; Methods).
-We then estimate the year in which the 663 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variation and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods).
+Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (see Methods).
+We then estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods).
 We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
@@
 \@ref(fig:toeFigure1)A).
-This corresponds to 68% of the studied stocks (Fig.
+This corresponds to 45% of the studied stocks (Fig.
 \@ref(fig:toeFigure1)C).
@@
 of all species, $\pm$ 28 years).
-Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% having recorded a distributional shift in stocks by the UN 2030 Agenda’s deadline (i.e., year 2030) (Fig.
+Furthermore, 81% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with 78% having recorded a distributional shift in stocks by the UN 2030 Agenda’s deadline (i.e., year 2030) (Fig.
 \@ref(fig:toeFigure1)C).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary shared stock distribution.
-(A) Number of stocks (cumulative) with their distribution centroids that have shifted beyond natural variations, (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one shifting stock, beyond natural variability.
-Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change  in range shift (Methods), respectively."}
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "**Timeline of shifts in transboundary shared stock distribution.** (A) Number of stocks (cumulative) with their distribution centroids that have shifted beyond natural variations, (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one shifting stock, beyond natural variability.
+Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change in range shift (see Methods), respectively."}
 
@@
 
-The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; Tables S\@ref(tab:toeTableS1), S\@ref(tab:toeTableS2) and Fig.
-S\@ref(fig:toeFigureS1)) and geographic region of the neighboring EEZs (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Table S3 and Fig.
+The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig.
+\@ref(fig:toeFigureS1)) and geographic region of the neighboring EEZs (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Table S3 and Fig.
 2A).
-Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables S\@ref(tab:toeTableS2 and S\@ref(tab:toeTableS3) than almost any other region (Figs.
-\@ref(fig:toeFigure2)B, S\@ref(fig:toeFigureS1) and S\@ref(fig:toeFigureS2)).
+Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2 and \@ref(tab:toeTableS3) than almost any other region (Figs.
+\@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS1) and \@ref(fig:toeFigureS2)).
 In contrast, EEZs and stocks located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts.
@@
 
-```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks.
-A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
+```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "**Year of change in the shared distribution of 4,119 transboundary stocks.** A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
 Exclusive Economic Zone polygons display average year of range shifts within them.
@@
 Points color coded by the median year of range shift.
-Horizontal dashed line represents the year 2030 – the Sustainable Development Agenda deadline.
+Horizontal dashed line represent the year by which countries have committed to reach full implementation of the 2030 Agenda.
 N = North, S = South, W = West and E = East.
@@
 
-```{r toeFigure3, eval = T,  echo = F, fig.width=10, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks and corresponding fishing revenue per country.
-Countries are coded according to different colored symbols representative of 5 distinct regions.
+```{r toeFigure3, eval = T,  echo = F, fig.width=8, fig.height=6, fig.pos= "H", fig.cap =  "**Average range shift of transboundary stocks and corresponding fishing revenue per country.** Countries are coded according to different colored symbols representative of 5 distinct regions.
 Horizontal line separates countries above and below the 75th percentile of annual fishing revenue.
-Showing country names for those within the top 75th percentile projected to have an average shift within 5 years after the 2030 deadline of the Sustainable Development Agenda.}
+Showing country names for those within the top 75th percentile projected to have an average shift within 5 years after the 2030 Agenda."}
 
@@
 
-**Changes in the stock share ratio of transboundary stocks**
+**Shifting intensity of transboundary stocks**
 
@@
 
-We estimated the intensity of the distributional shift of transboundary stocks in terms of changes in catch proportion by 2030 (2020-2040) and 2050 (2040-2060), relative to the recent past (1951-2005).
+We estimated the intensity of the distributional shift of transboundary stocks in terms of changes in catch proportion by 2030 (2020-2040) and 2050 (2040-2060), relative to the recent past (1951-2005) (see Methods).
 Findings show that by 2030, the global average catch proportion of transboundary stocks that are projected to change beyond the EEZ’s historic proportion is 59% ± 17% (Fig.
@@
 Changes in catch proportion are expected to become more common and intense slightly towards the mid -21st century at both the stock and EEZ level (Figs.
-S\@ref(fig:toeFigureS3) and S\@ref(fig:toeFigureS4)).
+\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).
 
-```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock proportion of neighboring Exclusive Economic Zones by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "**Changes in stock proportion of neighboring Exclusive Economic Zones by 2030 (2021-2040) relative to 1951-2005.** Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
 Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors.
@@
 
-**Potential drivers of shift and associated uncertainties**
+**Potential drivers of shift and managing an uncertain future**
 
-The early projected range shifts of a large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability e.g., SST is projected to increase beyond historical variability by 2030 in 70% of the global ocean under RCP 8.5 [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
-Results from multiple marine ecosystem models show that sea surface temperature (SST) and primary production (NPP) are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].Specifically, in the species distribution model used here (see methods), SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results.
+The early projected range shifts of a large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability.
+For example, sea surface temperature (SST) is projected to increase beyond historical variability by 2030 in 50 to 70% of the global ocean [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
+Results from multiple marine ecosystem models show that SST and primary production (NPP) are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
+Specifically, in the species distribution model used here (see methods), SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results.
 The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2019tn], while also explaining the later, and sometimes non-existent, shifts at higher latitudes (Fig.
@@
 Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and *ii*) model structure [@Frolicher:2016dg].
-First, our analysis is based on a high emission scenario (RCP 8.5) representing an extreme case of shifts in the shared distribution of transboundary species.
-Thus, any mitigation efforts (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the 21^st century [@Frolicher:2016dg].
+First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.
+Thus, any mitigation efforts (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].
 Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
-This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
+This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is often larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
 While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available [@Deser:2020bj] and cannot address limitations of process level understanding.
 As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties and will likely continue to do so in the coming decade.
-This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf] and where internal variability (e.g., El Niño Southern Oscillation - ENSO- for Pacific nations) is also important.
+This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf].
 These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh].
 Overall, changes in fish biomass with multiple upper trophic level models show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@Lotze:2019ce].
-Further research that reduces uncertainty in the response of the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to a smaller structural uncertainty of fish and fisheries models [@BryndumBuchholz:2019kh].
+Further research that reduces uncertainty in the NPP response at the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to a smaller structural uncertainty of fish and fisheries models [@BryndumBuchholz:2019kh].
 Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks.
@@
 Preparing anticipatory policies to deal with shifting transboundary stocks is key to achieving the SDGs and ensuring effective governance of the world’s fisheries [@Pinsky:2018cb; @Oremus:2020id; @Pecl:2017hu].
-In this contribution, we implemented two steps to demonstrate how to achieve sustainable transboundary fisheries in a changing world [@Link:2010ei].
+Here, we implemented two steps to demonstrate how to achieve sustainable transboundary fisheries in a changing world [@Link:2010ei].
 First, we identified the transboundary stocks that would likely see shifts in their shared distribution compared to their historical average and the year in which such shifts are predicted to occur.
@@
 
-<!-- ## Supplements -->
+\clearpage
+
+**References**
+
+<div id="refs"></div>
+
+\clearpage
+
+**Supplements**
 
 <!-- Set Figure, table and equations numbering according to the chapter -->
-\renewcommand{\thefigure}{S.\arabic{figure}}
+\renewcommand{\thefigure}{S\arabic{figure}}
 \setcounter{figure}{0}
-\renewcommand{\thetable}{S.\arabic{table}}
+\renewcommand{\thetable}{S\arabic{table}}
 \setcounter{table}{0}

### 2021-05-04 Scripts/Reviews/reviewer_one.Rmd
Started reviewing process

---
title: "Response to Reviewer One"
output: pdf_document
---

- Response: [Not responded] 
- *Response*: [Draft response here] 
- **Response**: [Response compleded her and in manuscript]

# Main comment

I appreciated the opportunity to review “Imminent range shifts in transboundary fish stocks challenge their management under climate change.” The figures were quite informative and visually pleasing so I comment the authors on this effect. I had only two large scale questions or concerns the first of which was more related to the policy-side raised in the abstract. **Specifically thinking how many bilateral agreements exist, and which ones should be considered “ready” for addressing transboundary shifts ** In addition, **the issue of model-specific variance is not represented in the manuscript**. Variance as a topic is well discussed in the discussion but I am left wanting to know the error bounds around some of the numbers presented (# of stocks shifting, net gain / loss of stocks by EEZ, years of response as well). I think if these two concerns can be addressed, the manuscript would be of great interest to a wide audience. I have some finer scale points raised below as well.

- Response:

P1L12 – “between neighbouring countries” – Is this not redundant with “transboundary?” Are you only examining stocks that currently cross jurisdictional boundaries? After further reading / review, it seems like this delineation would be useful to make explicit in the abstract.

- **Response:** Yes, "transboundary between neighboring countries” is indeed redundant. We have modified the manuscript to say "*Climate change is shifting the distribution of shared fish stocks between neighboring countries (transboundary).*" We believe it is important to keep both parts of the sentence for clarification.

P1L17-18 – which countries? Include specifics.

- **Response:** We are very limited in terms of word counting in the abstract so unfortunately, we cannot include specifics here. However, we have included some in-text examples, please see line *XXX* of page *XX*.

P2L1 – “Bi and multi” Also, some examples already exist and would be good to discuss as those that are particularly successful vs. those that aren’t.

- *Response:* Similarly, we have included some discussion of regarding adapted management programs, please see line *XXX* of page *XX*.

P2L5 – I would say preferences for discrete environmental conditions are a proxy for factors that influence species distribution not as you’ve laid out here.

- *Response:* We have changed this sentence to *Marine species’ distributions mainly reflect species’ preferences for discrete environmental conditions (2).*

P4L3 – “633 transboundary fish species globally, representing 67% of identified exploited taxa.” – how many of these have bi-lateral agreements as well?

- Response:

P5L7 – “Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability as early as 2006 and up to 2100 (Fig. 1A).” This is confusing as written. I would suggest “Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig. 1A), with the first anomaly occurring as early as 2006.” if meaning is correct.

- **Response:** Thank you for the input, we have made the suggested change.

P5L8 – “as early as 2006” is there a way to validate the historical period of these predictions (e.g. 2006-2020) to see if the model is actually predicting observed changes in transboundary stocks? This would greatly improve the validity of the manuscript.

- Response:

P8L6 – can you include information on direction here as well? Are these top five countries gaining stocks, losing stocks or just shifting stocks? It would be good to know which countries are predicted to be winners vs. losers.

- Response:

P9L7-8 – “catch proportion are expected to become more common and intense” – this sentence is a bit too jargon-y for the non-fisheries reader. Please specify whether you’re referring to increased specialization? Or?

- Response:

P10L6-7 – “with higher gains identified in warmer colors” Please be specific here on the scale as well. Which colors represent which # of stocks transitioning? This could be absolute number (stocks gained – stocks leaving) or relative (stocks gained – stocks leaving) / historical EEZ stocks. Both of these metrics could be quite useful for future planning.

- Response:

P12L2-3 – “As transboundary stocks continue to shift, such international disputes are likely to increase in frequency and requiring a revision of management plans (9, 28).” I do not disagree with this statement but it rings a bit hollow without knowing which country management plans need revision. I am not asking the authors to fully assess this literature but instead provide examples of successful management plans in response to transboundary shifts and examples those that are existing that would need to be revised.

- Response:

Pgs 12-13 deal with this a bit but this could use more context.

- Response:

P13L10-12 – “Results from multiple marine ecosystem models show that SST and primary production (NPP) are the main drivers of changes in species distributions across oceanic basins (37, 38).” This statement is misleading in my opinion. They are often important drivers in our models but this may be in part because we don’t measure other metrics at the appropriate scale for global or basin-scale modeling. I would revise to say “are the main indicators of species distributional changes across ocean basins.”

- **Response:** Thank you for highlighting this. We have changed the sentence to reflect this comment.

P14L14-16 - “Overall, changes in fish biomass with multiple upper trophic level models show overall agreement across models in terms of direction of change, but vary in the magnitude of that change (38).” I really appreciate the discussion of uncertainty but it seems important to present uncertainty for the models you have presented as well. Is there an upper and lower bounds (based on RCP8.5) due to model uncertainty from the predictions you have included? This could be incorporated in a number of the presented data points, e.g. 4,119 stocks by 2100, 239 of the world EEZs, etc.

- Response:

P14L20-23 – “While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level (43) and sub-populations are potentially interconnected (44), thus, providing additional ecological ground for our analysis (45).” A counterpoint would argue that subpopulations may have different ecological responses to climate variability and change thus species-specific models may miss this nuance. You get to this topic a bit on the next page though.

- Response:

### 2021-05-04 Scripts/Reviews/reviewer_three.Rmd
Started reviewing process

---
title: "Responses to Reviewer Three"
author: "Juliano Palacios"
date: "03/05/2021"
output: pdf_document
---

- Response: [Not responded] 
- *Response*: [Draft response here] 
- **Response**: [Response compleded her and in manuscript]

Overview
This paper covers an interesting and important question: how soon and how much will fisheries that span country boundaries shift with climate change? The authors make use of global climate models, predicted species responses to multiple changing environmental factors, and global catch data to provide a global synthesis of what bilateral and multilateral fisheries agreements will need to contend with in the coming decades. I really like the authors’ approach to dealing with natural temporal variability of fish stocks by accounting for that variability using historical model hindcasts and examining when stock distributions are predicted to move outside of historical bounds of variation. However, I had a couple of major methodological suggestions that I think should be addressed, and a number of smaller points to help improve clarity and interpretation. **My largest concern is the failure to consider national stocks that become transboundary stocks (and vice versa), which seem like cases central to the point of the study and yet overlooked in the analysis.** I elaborate on these points below.

# Major comments

- Informing global fisheries sustainability and meeting the SDG 2030 Agenda is a useful framing, in my opinion. However, given that framing and the research questions posed in this paper, I think the authors have overlooked two very important cases: 1) stocks that are currently confined to a single EEZ (“national stocks”) that will shift into transboundary stocks by 2100, and 2) stocks that are currently transboundary stocks that will shift entirel into a single EEZ by 2100, becoming national stocks. From a sustainability perspective, I am more concerned about case #1 than stocks that are already transboundary that are going to shift their distributions but maintain their transboundary status (the only cases currently examined in this paper). Some of those stocks might already have cooperative management agreements in place. And if they don’t, they are already plagued by the overharvesting that is likely to occur in the absence of international cooperation. Stocks that are currently national, however, have a higher likelihood of being managed sustainably now, and that likelihood will decline when they become transboundary. This seems like the most important case to analyze, and one that seems totally feasible to examine within the authors’ analysis framework, and yet this seems to have been completely overlooked. I also think it would be interesting to determine what proportion of current transboundary stocks become national stocks – the authors should have been able to detect these cases and I would assume there are some, but it is not mentioned anywhere in the paper. This seems like a missed opportunity for greater insights and is a mismatch between the stated goals of the paper and the results that are presented.

- The methods are pretty dense, so I would recommend making sure that all of the results
are really clearly described. In a few places, I found them confusing despite having read
the methods very carefully. Specifically:

 - Pg 6, line 16: Is 23% the proportion of revenue from transboundary stocks? Or the proportion of total fishing revenue? It is not totally clear to me as written. I think it’s the former, in which case line 19 should say “…less than 1% of transboundary fishing revenue” and the legend in Fig. 2A should read “% of transboundary fishing revenue.”

 - Pg 9, Line 6: is 59% the proportion of stocks that will shift their catch distribution? Or is it the average proportion of the catch that shifts between countries?

- The methods are ENTIRELY in the supplementary materials. I have no problem with more abbreviated methods in the main paper and more details in the supplement, but there needs to be some methods in the main paper. A reader needs to be able to understand the basic approach without going to the supplementary materials.

- *Response:* We have now included a summarized version of the methods in the main article.

- Pg 31-32: Does the TI capture the equality of sharing of the stock between two countries, or just how stable it is over time? The text claims an increasing TI means management is getting more challenging and a decreasing TI that the sharing between countries is becoming more stable (but not necessarily more equal)….is that the right interpretation? And how does that align with the example in Fig S6, where the TI seems to decline over the entire time series…..does that mean the sharing is becoming more stable?

- I like the time of emergence concept and how it was applied in this analysis.

**Response:** We appreciate the acknowledgement.

- I also like the idea of the threat point analysis to identify when a stock is sufficiently shared with another country to motivate cooperative management. However, I was confused by how the threat point was implemented using the stock share ratio (SSR) – why is the threat point determined relative to the historic SSR? I assume the idea of a threat point from game theory is based on the proportion of catch or value that is shared, and I don’t see how the starting point SSR would be relevant? Maybe I misunderstood what the authors did here, but conceptually/theoretically I’m not understanding the rationale for their approach so at the very least it needs more explanation and justification. But I’m concerned it actually needs to be based on a theoretical proportion, not a quantity interpreted relative to a historic proportion.

- I think a version of Fig. S6 would be useful for the main paper. (But a few small comments about this figure: in panel a, I don’t understand why the EEZ centroids aren’t shown in the middle of the EEZ boxes? Also, for panel b, I see a red open circle in the legend, but I don’t see that anywhere on the actual figure.)

**Response:** THank you for noticing the EEZ centroids. The figure has been corrected and included the figure in the main text.

# Minor comments

- I know this is nitpicky, but please don’t have line numbers restart on each page. It just makes it more cumbersome on a reviewer trying to give detailed comments.

**Response:** We understand (and agree!) with this comment. Unfortunately, this is how the *Science* template is currently distributed.

- Page1, Line 15: I suggest reporting the proportion of stocks (which looks like it would be between 20-25%), rather than the 78% reported here. Saying that transboundary stocks will shift their distributions across 78% of “the world’s international boundaries” feels a little misleading…..the results show that 78% of EEZs will experience a shifting stock. The proportion of stocks is a much easier to understand figure (particularly in the abstract) an is more directly comparable to the 45% figure given in the next sentence. The 78% feels like an unnecessary attempt to be flashy in the abstract.

- Pg 3, Line 2: First sentence doesn’t reflect the content in the rest of the paragraph. It would work better as a topic sentence for the third paragraph.

- Pg 3, Line 7: change “have shifted” to “are shifting” – this paper shows the shifts aren’t done happening yet

- **Response:** Thank you for noticing this. We have changed the wording to reflect this comment.

- Pg 4, Line 11-12: I don’t understand what is meant by “misalignment between fisheries resources allocation and distribution shifts” – language is confusing. Do the authors mean misalignment between fishing effort and stock biomass? Also in this paragraph, I think it would be helpful for the authors to add a couple of sentences of explanation for why transboundary stocks are more likely to be over-exploited.

- Pg 29: I realize the DBEM model is described in other papers, but a little more detail here would be helpful so that a reader doesn’t have to consult other papers to have a better sense of what the inputs are, how it was parameterized for each stock, etc. I suggest adding another paragraph to this effect.

- Pg 30, lines 8-12: this text is repeated from earlier in the paragraph

- *Response:* Thank you for noticing. We have removed the repeated text.

- Fig. S4 needs a color bar legend

- *Response:* Thank you for noticing. Fig. S4 has now a bar legend.

### 2021-05-04 Scripts/Manuscript.Rmd
Easy Responses to rev 1

@@
 
-Climate change is shifting the distribution of transboundary fish stocks between neighboring countries.
-The time-frame of such shifts determines the extent of climate change challenges to international fisheries governance.
+**Climate change is shifting the distribution of shared fish stocks between neighboring countries (transboundary).** The time-frame of such shifts determines the extent of climate change challenges to international fisheries governance.
 Using global climate-fisheries model simulations, we show that, by 2030, transboundary fish stocks are projected to shift their distributions across 78% of the world’s international boundaries.
@@
 Marine species’ distributions reflect species’ preferences for discrete environmental conditions [@Hutchinson:1957nt].
-As a result of changing ocean conditions, many marine species have shifted their distributions towards higher latitudes, deeper water or following local temperature gradients to remain within their optimal environmental “niche” [@Poloczanska:2016kk].
+As a result of changing ocean conditions, many marine species **are shifting** their distributions towards higher latitudes, deeper water or following local temperature gradients to remain within their optimal environmental “niche” [@Poloczanska:2016kk].
 The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt], impacting fish stocks, fisheries production and economies [@Sumaila:2019ew].
@@
 
-Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability as early as 2006 and up to 2100 (Fig.
-\@ref(fig:toeFigure1)A).
-This corresponds to 45% of the studied stocks (Fig.
+**Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
+\@ref(fig:toeFigure1)A), with the first shift occurring as early as 2006.** This corresponds to 45% of the studied stocks (Fig.
 \@ref(fig:toeFigure1)C).
@@
 For example, sea surface temperature (SST) is projected to increase beyond historical variability by 2030 in 50 to 70% of the global ocean [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
-Results from multiple marine ecosystem models show that SST and primary production (NPP) are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
+Results from multiple marine ecosystem models show that SST and primary production (NPP) are the main indicators of species distributional changes across ocean basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
 Specifically, in the species distribution model used here (see methods), SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results.

### 2021-05-04 Scripts/Initial_Analysis.Rmd
Included legend of figures 4 and S4 that were missing

--- 2021-04-05 Scripts/Initial_Analysis.Rmd
@@
     scale_colour_viridis(
-      name="Stock share ratio gain \nand changing stocks (%)",
+      name= "Changes in catch proportion \nand changing stocks (%)",
       limits = c(0,max(Map_data$per_plot)),
@@
     scale_fill_viridis(
-      name="Stock share ratio gain \nand changing stocks (%)",
+      name= "Changes in catch proportion \nand changing stocks (%)",
       limits = c(0,max(Map_data$per_plot)),
@@
   # Legend
-  map_plot <- get_legend(Prop_map)
+  map_legend <- get_legend(Prop_map)
 
@@
     # Europe
-    draw_plot(Eu_map, x = 0.52, y = 0.05, width = 0.35, height = 0.35) +
+    draw_plot(Eu_map, x = 0.53, y = 0.05, width = 0.35, height = 0.35) +
     # Legend
-    draw_plot(legend, x = 0.25, y = 0, width = 0.59, height = 0.09) +
+    draw_plot(map_legend, x = 0.25, y = 0, width = 0.59, height = 0.08) +
     # Labels

### 2021-05-04 Scripts/Manuscript.Rmd
Continue addressing reviewer 1

@@
 Moreover,  85% (n = 239) of the world EEZs would have experienced changes in catch proportion of transboundary stocks by 2030.
-Changes in catch proportion are expected to become more common and intense slightly towards the mid -21st century at both the stock and EEZ level (Figs.
-\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).
+**There will be an increase in the number of stocks changing their catch proportion and the number of EEZs with changing stocks towards the mid-21st century.
+Such increase will be followed by an also increase in the intensity of the proportion change (Figs.
+\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).**
 

### 2021-05-04 Scripts/Reviews/reviewer_one.Rmd
Continue addressing reviewer 1

@@
 - *Response*: [Draft response here] 
-- **Response**: [Response compleded her and in manuscript]
+- **Response**: [Response completed her and in manuscript]
 
@@
 
-- Response:
+- **Response:** Thank you for your comment.
+We provide direction of change in the next figure (Figure 5 of the revised manuscript) showing winners and losers according to neighboring EEZs.
+Here we show how each neighboring EEZ will stand to win or lose against its neighbors.
+**NOTE FOR CO-AUTHORS:** *I am hesitant of actually pointing out winners and losers (e.g., stating that Spain will loos most of its catch to France).
+Hence, I rather leave the results and discussion more general.
+THoughts?*
 
@@
 
-- Response:
+- **Response:** We have modified this sentence in the revised manuscript with the hopes to make it less jargon-y.
+It now reeds on line *7* of *P9* the revised manuscript: "*There will be an increase in the number of stocks changing their catch proportion and the number of EEZs with changing stocks towards the mid-21st century.
+Such increase will be followed by an also increase in the intensity of the proportion change (Figs.
+S3 and S4).*"
 
@@
 
-- Response:
+- **Response:** We realize the figure was missing the legend and color bar making it confusing for the reader. We have now included the legend explaining the legend and color bar.
 
@@
 
-- Response:
+- *Response:* Thank you for this comment. We have included a couple of examples of management plans that are somehow prepared or have respond to the issues we highlight.
 
@@
 
-- Response:
+- **Response:** Please see previous comment.
 
@@
 
-- Response:
+- **Response:** Thank you for your comment.
+For this paper we used one climate model (the GFDL) forced following one climate change scenario (RCP 8.5) to couple with one species distribution model.
+For this reason, providing quantitative upper and lower bounds of our results is not possible and would require substantial extra computation.
+However, because we have 10 different runs of the GFDL we are capable of providing a a less conservative estimate of the numbers you mention (i.e., the number of shifting stocks and EEZs).
+This was originally presented in the methods section as a sensitivity analyses and has now been moved to the Methods section of the main paper.
+It reads on line *XXX* of page *XX*: *We also computed a less conservative estimation of shifting transboundary stocks by including a threshold of one standard deviation (i.e., representing a probability of 68% that the stock had shifted from natural variation; see methods).
+In this case, the global average shift in the shared distribution of transboundary stocks is 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
+S8).
+In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
+S2 and S8).*
 
@@
 
-- Response:
+- **Response:** We agree with this comment. However, based on the data available we are unable to increase the detail of our study.

### 2021-05-06 Scripts/Reviews/reviewer_one.Rmd
Addressed all but one comment from Rev. 1

@@
 
-- Response:
+- Response: Thank you very much for commenting on our figures.
+Here, we provide responses to the two large scale questions you provide and below we respond the finer scale points.
+We hope that this clarifies all of your comments.
+
+ - **Regarding the bi-lateral agreements**, this is a great question that, unfortunately, we cannot respond at this moment.
+Currently, there is only one database on bi-lateral agreements produce by Oremus *et al* (2020).
+On their paper, the authors state "we analysed 27 publicly available international fisheries agreements in the South Pacific, Southeast Asia, South Asia, Northwest Africa, and Central America and the Caribbean" and " For each agreement, we recorded the following information...
+If relevant, to which specific stock or species the agreement pertained".
+However, the authors do not provide any information at the stock level and, unfortunately, they never respond/ did not shared their database with us.
+
+We hope that the reviewer understands the complexity of building such database and that this exercise falls outside the scope and stage of our current analysis.
+We agree that having such information would increase the manuscript impact, but, would not necessarily directly impact our results nor the message we want to covein (Ultimately oremus *et at* state it very clearly in their paper, "None of the 127 agreements contains language that is directly related to climate change, range shifts or stock exits").
+
+Oremus KL, Bone J, Costello C, Molinos JG, Lee A, Mangin T, Salzman J.
+2020.
+Governance challenges for tropical nations losing fish species due to climate change.
+Nature Sustainability 6:1–4.
+Nature Publishing Group.
+Available from https://www.nature.com/articles/s41893-020-0476-y.
+
+ - **Regarding model-specific variance.** For this paper we used one climate model (the GFDL) forced following one climate change scenario (RCP 8.5) to couple with one species distribution model.
+For this reason, providing quantitative upper and lower bounds of our results is not possible and would require substantial extra computation.
+However, because we have 10 different runs of the GFDL and employ the time of emergence concept, we are capable of providing a less conservative estimate of the numbers you mention (i.e., the number of shifting stocks and EEZs).
+This was originally presented in the methods section as a sensitivity analysis and has now been moved to the Methods section of the main paper (**NOTE TO CO-AUTHORS**: Should we include this in the results?).
+It reads on line *XXX* of page *XX*: *We also computed a less conservative estimation of shifting transboundary stocks by including a threshold of one standard deviation (i.e., representing a probability of 68% that the stock had shifted from natural variation; see methods).
+In this case, the global average shift in the shared distribution of transboundary stocks is 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
+S8).
+In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
+S2 and S8).*
 
@@
 - **Response:** We are very limited in terms of word counting in the abstract so unfortunately, we cannot include specifics here.
-However, we have included some in-text examples, please see line *XXX* of page *XX*.
+Please note that we do provide some in-text examples in the results section, specifically the revised Figure 4.
 
@@
 
-- *Response:* Similarly, we have included some discussion of regarding adapted management programs, please see line *XXX* of page *XX*.
+- **Response:** Similarly, we do not have space to discuss this here. Please note that we discuss management plans on the paragraph starting on line *13* of page *12*.
 
@@
 
-- *Response:* We have changed this sentence to *Marine species’ distributions mainly reflect species’ preferences for discrete environmental conditions (2).*
+- **Response:** We have changed this sentence to *Marine species’ distributions mainly reflect species’ preferences for discrete environmental conditions (2).*
 
@@
 
-- Response:
+- **Response:** Please see the main comment response above.
 
@@
 
-- Response:
+- *Response:* In the section "High present-day climate risk of on transboundary fisheries management" we describe early shifts in the shared distribution of transboundary stocks including the cases of Humboldt squid (*Dosidicus gigas*) and Atlantic mackerel (*Scomber scombrus*) both in the early 2000s.
+We have included some more punctual examples of documented shifts between 2006 and 2021.
+It reads on line **xx** of page **XX**: "Similar shifts have been documented for multiple European stocks [@Baudron:2020dj] like European hake (*Merluccius merluccius*) [@Baudron:2015il] as well as...
 
@@
 Hence, I rather leave the results and discussion more general.
-THoughts?*
+Thoughts?*
 
@@
 
-- *Response:* Thank you for this comment. We have included a couple of examples of management plans that are somehow prepared or have respond to the issues we highlight.
+- **Response:** Thank you for this comment. We have eddited the paragraph starting at line **13** of page **12** to reflect this comment.
 
@@
 
-- **Response:** Thank you for your comment.
-For this paper we used one climate model (the GFDL) forced following one climate change scenario (RCP 8.5) to couple with one species distribution model.
-For this reason, providing quantitative upper and lower bounds of our results is not possible and would require substantial extra computation.
-However, because we have 10 different runs of the GFDL we are capable of providing a a less conservative estimate of the numbers you mention (i.e., the number of shifting stocks and EEZs).
-This was originally presented in the methods section as a sensitivity analyses and has now been moved to the Methods section of the main paper.
-It reads on line *XXX* of page *XX*: *We also computed a less conservative estimation of shifting transboundary stocks by including a threshold of one standard deviation (i.e., representing a probability of 68% that the stock had shifted from natural variation; see methods).
-In this case, the global average shift in the shared distribution of transboundary stocks is 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
-S8).
-In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
-S2 and S8).*
+- **Response:**  Please see the response to the main comment above.
 

### 2021-05-06 Scripts/Manuscript.Rmd
Marked changes for Rev 1

@@
 
-Management plans that are not designed to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
+**Management plans that are not designed to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
 Thus, focusing on these “hotspots” is key to anticipate a potential increase in fisheries conflicts in coming years.
 Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments, which do not have to be monetary [@Tunca:2019if; @Miller:2013iv], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@PalaciosAbrantes:2020vt; @Pinsky:2018cb].
-For example, countries that are Party to the Nauru Agreement (PNA) manage their fisheries under a so-called Vessel Day Scheme, which accounts for shifts in stocks and catches, and is therefore somewhat anticipatory [@Aqorau:2018bh].
-However, even strategies that take into consideration species’ shifts within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
-Quota allocation methods based on a stock’s current distribution and/or on fixed-historical proportion need to evolve to be more agile [@Sumaila:2020tp; @PalaciosAbrantes:2020vt].
-Fisheries based on historic participation could move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-However, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery.
+Some examples of current treaties that have adapted some of these strategies include the Pacific Salmon Treaty between Canada and the U.S.
+to manage Pacific salmon (Oncorhynchus sp.) that has a conservation fund that works as a side payment [@Miller:2013iv], the International Pacific Halibut Commission between Canada and the United States that manages Pacific halibut (*Hippoglossus stenolepis*) and allocates quota based on the stock’s yearly distribution [@PalaciosAbrantes:2020vt], and the Parties to the Nauru Agreement (PNA) in the Pacific Island States that manage skipjack tuna (*Katsuwonus pelamis*).
+The PNA manage their fisheries under a so-called Vessel Day Scheme, which accounts for shifts in stocks and catches, and is therefore somewhat anticipatory [@Aqorau:2018bh].
+Quota allocation methods based on a stock’s current distribution and/or on fixed-historical proportion like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+However, even management plans that take into consideration such strategies might not be fully prepared for the consequences of shifting transboundary stocks [@Pinsky:2018cb; @Koubrak:2020hw; @Engler:2020bg].
+For example species’ shifts within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+Moreover, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery they have historically been entitled.
+Finally, changes in policy can take years (e.g., it took the PST about 10 years to agree on a conservation fund) and sometimes, the time of political change is longer than that of species shifts [@Pinsky:2012kq].**
 

### 2021-05-07 Scripts/Reviews/reviewer_one.Rmd
Responses to reviewer one

@@
 
-- *Response:* In the section "High present-day climate risk of on transboundary fisheries management" we describe early shifts in the shared distribution of transboundary stocks including the cases of Humboldt squid (*Dosidicus gigas*) and Atlantic mackerel (*Scomber scombrus*) both in the early 2000s.
-We have included some more punctual examples of documented shifts between 2006 and 2021.
-It reads on line **xx** of page **XX**: "Similar shifts have been documented for multiple European stocks [@Baudron:2020dj] like European hake (*Merluccius merluccius*) [@Baudron:2015il] as well as...
+- **Response:** In the section "High present-day climate risk of on transboundary fisheries management" we describe early shifts in the shared distribution of transboundary stocks including the cases of Humboldt squid (*Dosidicus gigas*) and Atlantic mackerel (*Scomber scombrus*) both in the early 2000s and have included a couple more punctual examples of documented shifts between 2006 and 2021.
+However, there is not a vast documentation on shifts in the distribution of transboundary species (i.e., specifically from one EEZ to another).
+Mapping these shifts is out of the scope of the current project as there is an ongoing effort to document the global hifts of transboundary species.
 

### 2021-05-07 Scripts/Initial_Analysis.Rmd
fixed point in fig4

@@
     mutate(
-      over_tresh = ifelse(ensemble_mean > top_tresh | ensemble_mean < low_tresh,"keep",
+      over_tresh = ifelse(time_step != "historical" & ensemble_mean > top_tresh | ensemble_mean < low_tresh,"keep",
                           ifelse(time_step == "historical","keep","drop")
@@
     ) %>% 
+    # filter(time_step !="historical" | over_tresh == "keep") %>%
     filter(over_tresh == "keep") %>%
@@
 }
+
+```
+
+#### Fun `EEZarrive_exit`
+
+```{r fun_EEZexit, eval = F, echo = F}
+
+# taxon_list <- c(600004,600006)
+
+# EEZProp <- function(taxon_list,links="NA"){
+
+  spp <- bind_rows(
+    lapply(Taxon_list,SummaryProp,links)
+  ) %>% 
+    spread(time_step,mean_spp) %>% 
+    # mutate_if(is.numeric,replace_na,0) %>% # If NA 0 meaning wins and looses
+    mutate_at(vars(
+      "per_early" = early,
+      "per_mid" = mid
+    ),
+    funs(round(((.-historical)/(abs(historical)))*100,2))) %>%  # original
+         # funs(round(((historical-.)/(abs((historical+.)/2)))*100,2)
+    # funs((historical-.)*100)) %>%  
+    # funs(ifelse(variable == "n_per",.,round(((.-historical)/(abs(historical)))*100,2)
+                # )
+         # )
+    # ) %>%
+  gather("time_step","percentage_change",per_early:per_mid)
 
@@
     # funs(round(((.-historical)/(abs(historical)))*100,2)
-    funs(ifelse(variable == "n_per",.,round(((.-historical)/(abs(historical)))*100,2)
+    funs(ifelse(variable == "n_per",.,round(((.-historical)/(abs(historical)))*100,2) # original
+    #             )
+    # funs(ifelse(variable == "n_per",.,(historical-.)*100)
                 )
@@
     geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
-    geom_point(data = subset(Map_data, time_step == time & per_plot < 0),
+    geom_point(data = subset(Map_data, variable =="mean_eez" & time_step == time & per_plot < 0), 
                aes(start_lon,
@@
                stroke = 1.5) +
-    geom_path(data = subset(paths, time_step == time & per_plot > 0),
+    geom_path(data = subset(paths, variable =="mean_eez" & time_step == time & per_plot > 0), # filter data 
               aes(X,

### 2021-05-10 Scripts/Initial_Analysis.Rmd
Cleaned script

@@
     ),
-    funs(round(((.-historical)/(abs(historical)))*100,2))) %>%  # original
-         # funs(round(((historical-.)/(abs((historical+.)/2)))*100,2)
-    # funs((historical-.)*100)) %>%  
-    # funs(ifelse(variable == "n_per",.,round(((.-historical)/(abs(historical)))*100,2)
-                # )
-         # )
-    # ) %>%
+    funs(round(((.-historical)/(abs(historical)))*100,2))
+    ) %>%  # original
   gather("time_step","percentage_change",per_early:per_mid)
@@
     ),
-    # funs(round(((.-historical)/(abs(historical)))*100,2)
     funs(ifelse(variable == "n_per",.,round(((.-historical)/(abs(historical)))*100,2) # original
-    #             )
-    # funs(ifelse(variable == "n_per",.,(historical-.)*100)
                 )

### 2021-05-10 Scripts/Manuscript.Rmd
included most of rev. 3 changes

@@
 **Climate change is shifting the distribution of shared fish stocks between neighboring countries (transboundary).** The time-frame of such shifts determines the extent of climate change challenges to international fisheries governance.
-Using global climate-fisheries model simulations, we show that, by 2030, transboundary fish stocks are projected to shift their distributions across 78% of the world’s international boundaries.
+Using global climate-fisheries model simulations, we show that 78% the world’s international boundaries will experience a shifting transboundary stock by 2030.
 By 2100, 45% of stocks will have shifted for 81% of countries under a ‘no mitigation’ scenario.
@@
 
-Most fisheries management is poorly adapted to address species’ shifting distributions as a result of changing ocean conditions.
 Over the past century, human activities have dramatically altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters [@IPCC:2019tn].
@@
 
+Most fisheries management is poorly adapted to address species’ shifting distributions as a result of changing ocean conditions.
 The effectiveness of fisheries management for transboundary species is challenged by shifts in stocks’ distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].
@@
 While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs].
+For example, Pacific salmon (Oncorhynchus sp.) jointly managed by Canada and the U.S.
+was heavily fished in the early 1990s after a stock shift favored the Alaskan fleet over the Canadian fleet [@Miller:2013iv].
 Such patterns are expected to be exacerbated under continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
@@
 \@ref(fig:toeFigure2)A).
+
 On average, shifting transboundary stocks represented 23% (s.d.
-$\pm$  22%) of yearly fishing revenue from fisheries targeting such stocks within global EEZs between 2004 and 2010 (Fig.
+$\pm$  22%) of yearly fishing revenue from fisheries targeting **transboundary** stocks within global EEZs between 2004 and 2010 (Fig.
 \@ref(fig:toeFigure2)A).
-However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
+**From all fisheries targeting transboundary stocks between 2004 and 2010 in the global EEZs, shifting stocks represented on average 23% (s.d.
+$\pm$  22%) of the yearly revenue (Fig.
+\@ref(fig:toeFigure2)A).** However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
 Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).
@@
 We estimated the intensity of the distributional shift of transboundary stocks in terms of changes in catch proportion by 2030 (2020-2040) and 2050 (2040-2060), relative to the recent past (1951-2005) (see Methods).
-Findings show that by 2030, the global average catch proportion of transboundary stocks that are projected to change beyond the EEZ’s historic proportion is 59% ± 17% (Fig.
+**Findings show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks is projected to change beyond an EEZ’s historic proportion (i.e., between 2004 and 2010;** Fig.
 \@ref(fig:toeFigure4)A).

### 2021-05-10 Scripts/Manuscript.Rmd
Included a summary of the methods in the main article

@@
 
-Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (see Methods).
-We then estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods).
+Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (see Methods **and Supplemental Materials**).
+We then estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig.
+\@ref(fig:toeFigure1)A) and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods **and Supplemental Materials**).
 We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
@@
 
+**Methods **
+
+We estimate the year in which 663 transboundary species previously identified by [@PalaciosAbrantes:2020tv] (i.e., 9,132 stocks) would undergo changes in their shared distribution beyond historical variation and the intensity of said changes (see the Supplementary Materials - *Databases and species selection*).
+We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
+We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division, noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
+\@ref(fig:toeFigureS5)) including island territories.
+
+We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks’ under climate change (see the Supplementary Materials).
+Roughly, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and  spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
+The DBEM was driven for ten different versions of the GFDL (i.e., a ten-ensemble member simulation) each of them representing an unique atmosphere and ocean state at the start of the simulation [@Dunne:2012bo; @Dunne:2013db; @Frolicher:2020cc].
+The simulations were run under historical forcing until 2005, afterwards and until 2100, simulations followed a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5) [@Riahi:2011dk].
+
+Furthermore, we developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
+The index is based on changes in the distance of the stock's abundance centroid and that of the EEZs that share such stock (Fig.
+\@ref(fig:toeFigure1)A)).
+We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution (see the Supplementary Materials - *Calculating shifts in the shared distribution of transboundary stocks*).
+The time of emergence was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ.
+\@ref(equ:four)).
+We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
+In this case, the global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
+\@ref(fig:toeFigureS8)).
+In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
+\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
+
+Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Clark:1975uu; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (See Supplemental Material - *Quantifying the intensity of transboundary stocks’ range shift*).
+A threat point is defined as the minimum payoff a player would be willing to receive to cooperate in a game-theoretic model.
+Here, we define the minimum payoff as two standard deviations from the historical catch proportion (average of 1951 to 2005) that each EEZ had for each stock.
+This assumes that no EEZ will be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Finally, we computed changes in the catch proportion that were below the threat point of the neighbouring EEZs sharing a stock for two time periods; 2030 (2021 and 2040) matching the UN 2030 Agenda and a late date (2040 to 2050), relative to 1951 to 2005.
+
 <div id="refs"></div>

### 2021-05-10 Scripts/Supplements.Rmd
Moved some of the methods to the main article and icnluded a more detailed version of the DBEM

@@
 \@ref(fig:toeFigureS5)) including island territories.
-.
 We determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp].
@@
 
-We projected the future distribution of transboundary stocks using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
+We projected the future distribution of transboundary stocks using a dynamic bioclimatic envelope model (hereafter called DBEM).
 The DBEM determines the environmental space a species will occupy based on physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
-For pelagic species, the model uses environmental variables at the surface whereas demersal and benthic species’ distributions are driven by ocean bottom variables.
-The DBEM then estimates species’ abundance and maximum catch potential (a proxy for maximum sustainable yield) over a regular spatial grid of 0.5 of latitude per 0.5 of longitude (see @Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd for model details).
-Importantly, DBEM is able to project catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
+The DBEM then estimates species’ abundance and maximum catch potential (a proxy for maximum sustainable yield) over a regular spatial grid of 0.5 of latitude per 0.5 of longitude.
+Below we provide an overall description of the model.
+Please see @Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd for a more detailed version.
+Importantly, the DBEM is able to project catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
 *in revision*].
 
- DBEM was driven by simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+For each species, the DBEM uses the Sea Around Us (www.searaundus.org) distributional data and environmental variables from 1970 to 2000 to estimate the current distribution and environmental profile of each species.
+Moreover, it projects the future habitat suitability in each grid-cell (*i*) based on the following components;
+
+\begin{equation}
+\label{eq:one}
+Hs = P(T_{i},TPP) *  P(Ba_{i}, Min_{D}, Max_{D}) * P(Hab_{i,j}, Hab_{Assoc}) * (Sal_{i}, Sal_{Assoc}) * P(Ice_{i}, Ice_{P})
+\end{equation}
+
+where *T* is seawater temperature, *TPP* is the species temperature profile, *Ba* is bathymetry, $Min_{D}$ and $Max{D}$ are the species minimum and maximum depth limits, *Hab* is the proportion of area of the habitat type *j* relative to the total seawater area of the cell *i*, $Hab_{Assoc}$ is an habitat association index, *Sal* is the salinity class of cell *i* and $Sal_{Assoc}$ is a salinity class association index.
+Finally, for polar regions and species, *Ice* is sea ice extent, and $Ice_{P}$ is association to sea ice.
+For pelagic species, the model uses environmental variables at the surface whereas demersal and benthic species’ distributions are driven by ocean bottom variables.
+
+The DBEM also integrates each species preference profile with physiology and population dynamics to project relative biomass assuming that spatio-temporal dynamics are determined by intrinsic growth rate, larval dispersion, and adult migration;
+
+\begin{equation}
+\label{eq:two}
+\frac{dA}{dt} = \sum_{j=1}^{n} (L_{ji} + Ad_{ji} + G_i)
+\end{equation}
+
+where $A_i$ is the relative abundance of the gri-cell *i*, *G* is the intrinsic population growth and $L_{ji}$ and $Ad_{ji}$ are settled larvae and net migrated adults from surrounding cells (*j*), respectively.
+Larval dispersal is modeled through ocean current with an advection-diffusion-reaction model.
+Finally, $G_i$ estimates the intrinsic growth following a logistic equation and the intrinsic rate of population increase (*r*):
+
+\begin{equation}
+\label{eq:three}
+G_i = r * A_i * (1-\frac{A_i}{A_{\infty,i}})
+\end{equation}
+
+Finally, the DBEM projects maximum catch potential by setting fishing mortality to half of the intrinsic population growth rate of the species (*f = r/2*).
+
+The DBEM was driven by simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
 The GFDL-ESM2M was run under historical forcing until 2005 and follows a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5), over the 2006-2100 period [@Riahi:2011dk].
@@
 It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution.
-We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
-Sensitivity tests to a lower threshold of 1 s.d.
-(or 68% confidence) resulted in a global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
-\@ref(fig:toeFigureS8)).
-In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
-\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
 

### 2021-05-10 Scripts/Reviews/reviewer_three.Rmd
Almost done!

@@
 title: "Responses to Reviewer Three"
-author: "Juliano Palacios"
-date: "03/05/2021"
 output: pdf_document
@@
 
-- The methods are pretty dense, so I would recommend making sure that all of the results
-are really clearly described. In a few places, I found them confusing despite having read
-the methods very carefully. Specifically:
+- **Response:** We appreciate the reviewer's comment regarding our method and understand the concern on consider national stocks that become transboundary stocks and *vice versa*, however, these questions have already been responded by Pinsky *et al* (2018) and Oremus *et al.* (2020) and consequently, follow outside the scope of the current paper.
 
- - Pg 6, line 16: Is 23% the proportion of revenue from transboundary stocks?
+In their 2018 paper, Pinsky *et al* look at how many *new* transboundary species each EEZ of the world will have towards the end of the century under climate change (i.e., national stocks that become transboundary stocks).
+Indeed, the authors find that "*many of the world’s EEZs are likely to receive one to 12 new, climate-driven transboundary stocks by the end of the century*." On the other hand, Oremus *et al* (2020) "*estimates the number of species exiting national jurisdictions under climate change up to 2100*" concluding that *tropical nations stand to lose the most species under these scenarios.* However, both studies fail to identify changes in the *current* transboundary stock, gap that we look to fill int his paper.
+Here, we used the species (n = 633) identified by Palacios-Abrantes *et al* (2020) as transboundary and consequently do not have national stocks identified.
+This was on purpose as the objective of our paper is precisely to look at the impacts of climate change on *current* transboundary species.
+Addressing the question of entering/exciting stocks would require us to change our initial species list, change our otherwise already complex and long methodology, and likely give the paper a different perspective.
+
+ - Oremus KL, Bone J, Costello C, Molinos JG, Lee A, Mangin T, Salzman J. 2020. Governance challenges for tropical nations losing fish species due to climate change. *Nature Sustainability* 6:1–4.
+
+ - Palacios-Abrantes J, Reygondeau G, Wabnitz CCC, Cheung WWL. 2020. The transboundary nature of the world’s exploited marine species. *Scientific Reports* 10:415–12.
+
+ - Pinsky ML, Reygondeau G, Caddell R, Palacios-Abrantes J, Spijkers J, Cheung WWL. 2018. Preparing ocean governance for species on the move. *Science* 360:1189–1191.
+
+The methods are pretty dense, so I would recommend making sure that all of the results are really clearly described.
+In a few places, I found them confusing despite having read the methods very carefully.
+Specifically:
+
+Pg 6, line 16: Is 23% the proportion of revenue from transboundary stocks?
 Or the proportion of total fishing revenue?
@@
 
+ - **Response:** We agree that this was confusing and have modified the initial sentence and adopted the suggestions provided in the revised manuscript.
+It reads in line **XX** of page **X**: "*From all fisheries targeting transboundary stocks between 2004 and 2010 in the global EEZs, shifting stocks represented on average 23% (s.d.
+± 22%) of the yearly revenue (Fig.
+2A).*"
+
  - Pg 9, Line 6: is 59% the proportion of stocks that will shift their catch distribution? Or is it the average proportion of the catch that shifts between countries?
+
+ - **Response:** We meant that 59% is the average proportion of the catch that shifts between countries, or more specific, the stocks reponsable for that catch.
+We have revised the manuscript to clarify this, it reads on line **XX** of page *X*: "*Findings show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks is projected to change beyond an EEZ’s historic proportion (i.e., between 2004 and 2010; Fig.
+4A)*"
 
@@
 
-- *Response:* We have now included a summarized version of the methods in the main article.
+- **Response:** The main structure was following the journal's guideline. We have now included a summarized version of the methods in the main article.
 
@@
 
+*Response:* William, thoughts?
+
 - I like the time of emergence concept and how it was applied in this analysis.
 
-**Response:** We appreciate the acknowledgement.
+**Response:** We appreciate the acknowledgment on our methods.
 
@@
 
+- Response: Missing
+
 - I think a version of Fig.
@@
 
-**Response:** THank you for noticing the EEZ centroids. The figure has been corrected and included the figure in the main text.
+**Response:** Thank you for noticing the misplace of the EEZ centroids.
+The figure has been corrected and included the figure in the main text (It is now Fig.
+1).
+The open red circles are located in the bottom panel (B) in each time step where the future distribution overpasses the natural variability.
 
@@
 
-**Response:** We understand (and agree!) with this comment. Unfortunately, this is how the *Science* template is currently distributed.
+**Response:** We understand (and agree!) with this comment. Unfortunately, this is how the *Science* template was distributed to us.
 
@@
 
+**Response:** Thank you, we have adopted the suggested approach.
+You will read in line 14 of page 1: "*Using global climate-fisheries model simulations, we show that 78% the world’s international boundaries will experience a shifting transboundary stock by 2030*".
+
 - Pg 3, Line 2: First sentence doesn’t reflect the content in the rest of the paragraph. It would work better as a topic sentence for the third paragraph.
+
+- **Response**: We agree with this observation and have modified the revised manuscript to reflect it.
 
@@
 Do the authors mean misalignment between fishing effort and stock biomass?
+
+- **Response:** We meant misalignment between fishing allocation rules and the stock's biomass this has been clarified in the revised manuscript.
+
 Also in this paragraph, I think it would be helpful for the authors to add a couple of sentences of explanation for why transboundary stocks are more likely to be over-exploited.
+
+- **Response:** We have included one of the examples we referenced, it now reads in line *XX* of page *xxx in the revised manuscript:* *"For example, Pacific salmon (Oncorhynchus sp.) jointly managed by Canada and the U.S.
+was heavily fished in the early 1990s after a stock shift favored the Alaskan fleet over the Canadian fleet (18).*"
 
@@
 
+- **Response:** We have included an overall description of the DBEM (including the main equations) in line **XX** of page **X** of the supplemental material.
+
 - Pg 30, lines 8-12: this text is repeated from earlier in the paragraph
 
-- *Response:* Thank you for noticing. We have removed the repeated text.
+- **Response:** Thank you for noticing. We have removed the repeated text.
 
@@
 
-- *Response:* Thank you for noticing. Fig. S4 has now a bar legend.
+- **Response:** Thank you for noticing. Both figures 4 (now Fig. 5) and S4 have a bar color legend now.

### 2021-05-11 Scripts/Manuscript.Rmd
Included a summary of methods and re-organized figures

@@
 We then estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig.
-\@ref(fig:toeFigure1)A) and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods **and Supplemental Materials**).
+\@ref(fig:toeFigure1N)) and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods **and Supplemental Materials**).
 We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
+
+```{r toeFigure1N, eval = F,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring Exclusive Economic Zones sharing the stock (Top panel).
+Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel)."}
+
+img <- readPNG("../Figures/Fig1.png")
+grid.raster(img)
+
+```
+
+\clearpage
 
@@
 
-img <- readPNG("../Figures/Fig1.png")
+img <- readPNG("../Figures/Fig2.png")
 grid.raster(img)
@@
 
-img <- readPNG("../Figures/Fig2.png")
+img <- readPNG("../Figures/Fig3.png")
 grid.raster(img)
@@
 
-img <- readPNG("../Figures/Fig3.png")
+img <- readPNG("../Figures/Fig4.png")
 grid.raster(img)
@@
 
-img <- readPNG("../Figures/Fig4.png")
+img <- readPNG("../Figures/Fig5.png")
 grid.raster(img)
@@
 
-We estimate the year in which 663 transboundary species previously identified by [@PalaciosAbrantes:2020tv] (i.e., 9,132 stocks) would undergo changes in their shared distribution beyond historical variation and the intensity of said changes (see the Supplementary Materials - *Databases and species selection*).
+We estimate the year in which 663 transboundary species (i.e., 9,132 stocks) previously identified by Palacios-Abrantes *et al.* (2020) [@PalaciosAbrantes:2020tv] would undergo changes in their shared distribution beyond historical variation and the intensity of said changes (see the Supplementary Materials).
 We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
-We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division, noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
-\@ref(fig:toeFigureS5)) including island territories.
+We defined the boundaries of the world’s EEZs using the Sea Around Us (http://www.seaaroundus.org/) spatial division, noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
+\@ref(fig:toeFigureS5)) including island territories (see the Supplementary Materials - *Databases and species selection*).
 
@@
 We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution (see the Supplementary Materials - *Calculating shifts in the shared distribution of transboundary stocks*).
-The time of emergence was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ.
+The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ.
 \@ref(equ:four)).
@@
 Here, we define the minimum payoff as two standard deviations from the historical catch proportion (average of 1951 to 2005) that each EEZ had for each stock.
-This assumes that no EEZ will be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-Finally, we computed changes in the catch proportion that were below the threat point of the neighbouring EEZs sharing a stock for two time periods; 2030 (2021 and 2040) matching the UN 2030 Agenda and a late date (2040 to 2050), relative to 1951 to 2005.
+This assumes that countries are adapted to the historical sharing but no country will be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Finally, we computed changes in the catch proportion that were below the threat point of the neighbouring EEZs sharing a stock for two time periods relative to the historical proportion; 2030 (2021 and 2040) matching the UN 2030 Agenda and a late date (2040 to 2050).
 

### 2021-05-11 Scripts/Supplements.Rmd
Removed TI direcction statement since it is obsolete

@@
 
+<!-- Included more details of the DBEM in response to reviewer 3 -->
 For each species, the DBEM uses the Sea Around Us (www.searaundus.org) distributional data and environmental variables from 1970 to 2000 to estimate the current distribution and environmental profile of each species.
@@
 \begin{split}
-\label{eq:one}
+\label{eq:four}
 lat_{ts} = \mu (lat_{per}) \\
@@
 \begin{equation}
-\label{eq:two}
+\label{eq:five}
 D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
@@
 \begin{equation}
-\label{eq:three}
+\label{eq:six}
 TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
@@
 where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (th, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
-Thus, a higher TI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels.
+
+<!-- Removed in response to reviewer 3 -->
+<!-- Thus, a higher TI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels.
 From the perspective of a state’s fisheries, increasing TI indicates an increase in the challenges to effective management and sustainability of the transboundary stock’s fisheries.
 In contrast, a decreasing TI suggests that the sharing of the stocks between neighboring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries.
+-->
+
 The index was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg].
@@
 \begin{equation}
-\label{eq:four}
+\label{eq:seven}
 ToE = TI_\mu[t']>2* TI_\sigma 
@@
 \begin{equation}
-\label{eq:five}
+\label{eq:eight}
 \Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100
@@
 img <- readPNG("../Figures/FigS5.png")
-grid.raster(img)
-
-```
-
-\clearpage
-
-```{r toeFigureS6, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
-The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring Exclusive Economic Zones sharing the stock (Top panel).
-Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel)."}
-
-img <- readPNG("../Figures/FigS6.png")
 grid.raster(img)

### 2021-05-11 Scripts/Reviews/reviewer_three.Rmd
Initial round of revisions done

@@
 
-*Response:* William, thoughts?
+*Response:* Thank you for your comment.
+The transboundary index only captures the sharing stability over time.
+We had originally explored the directionality of the TI per stock, however, this was removed from the analysis as we rather focused on the "time of emergence" of the future TI from historical variability.
+Thus, since we do not talk about the directionallity of the TI (nor the TI *per se*) we have removed this sentence to avoid misunderstanding.
+
+The equality of sharing is measured by the Stock Share Ratio described under *Quantifying the intensity of transboundary stocks’ range shift.*
 

### 2021-05-11 Scripts/Supplements.Rmd
Initial round of revisions done

@@
 We define the threat point as the minimum required catch proportion of a transboundary stock within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-Any proportion below the defined threat point would result in the unilateral management of the stock.
-First, we estimated the stock share ratio (SSR) of each transboundary stock shared by neighboring nations between 1951 and 2100.
+**Any proportion below the defined threat point would result in the unilateral management of the stock assuming that no country will be willing to engage in cooperative management if the future catch proportion of the transboundary stock is lower than it has ever been.**
+
+First, we estimated **the catch proportion as the** stock share ratio (SSR) of each transboundary stock shared by neighboring nations between 1951 and 2100.
 We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock's catch held within each EEZ [@PalaciosAbrantes:2020tv].

### 2021-05-11 Scripts/Reviews/reviewer_one.Rmd
All responses done
Initial round of revisions done

@@
 title: "Response to Reviewer One"
-output: pdf_document
+# output: pdf_document
+output: word_document
 ---
-
-- Response: [Not responded] 
-- *Response*: [Draft response here] 
-- **Response**: [Response completed her and in manuscript]
 
@@
 
-- Response: Thank you very much for commenting on our figures.
+- **Response:** Thank you very much for commenting on our figures.
 Here, we provide responses to the two large scale questions you provide and below we respond the finer scale points.
@@
 Currently, there is only one database on bi-lateral agreements produce by Oremus *et al* (2020).
-On their paper, the authors state "we analysed 27 publicly available international fisheries agreements in the South Pacific, Southeast Asia, South Asia, Northwest Africa, and Central America and the Caribbean" and " For each agreement, we recorded the following information...
+On their paper, the authors state "we analysed 127 publicly available international fisheries agreements in the South Pacific, Southeast Asia, South Asia, Northwest Africa, and Central America and the Caribbean" and " For each agreement, we recorded the following information...
 If relevant, to which specific stock or species the agreement pertained".
-However, the authors do not provide any information at the stock level and, unfortunately, they never respond/ did not shared their database with us.
+We contacted Dr.
+(Kimberly) Oremus who provided the provided the mentioned dataset.
+However, non of the 127 treaties have identified subjects at the species level (except for bigeye, skipjack and yellowtail tunas).
 
 We hope that the reviewer understands the complexity of building such database and that this exercise falls outside the scope and stage of our current analysis.
-We agree that having such information would increase the manuscript impact, but, would not necessarily directly impact our results nor the message we want to covein (Ultimately oremus *et at* state it very clearly in their paper, "None of the 127 agreements contains language that is directly related to climate change, range shifts or stock exits").
+We agree that having such information would increase the manuscript impact, but, would not necessarily directly impact our results nor the message we want to covein (Ultimately Oremus *et at* state it very clearly in their paper, "None of the 127 agreements contains language that is directly related to climate change, range shifts or stock exits").
 
-Oremus KL, Bone J, Costello C, Molinos JG, Lee A, Mangin T, Salzman J.
+ - Oremus KL, Bone J, Costello C, Molinos JG, Lee A, Mangin T, Salzman J.
 2020.
@@
 Nature Publishing Group.
-Available from https://www.nature.com/articles/s41893-020-0476-y.
 
@@
 However, because we have 10 different runs of the GFDL and employ the time of emergence concept, we are capable of providing a less conservative estimate of the numbers you mention (i.e., the number of shifting stocks and EEZs).
-This was originally presented in the methods section as a sensitivity analysis and has now been moved to the Methods section of the main paper (**NOTE TO CO-AUTHORS**: Should we include this in the results?).
+This was originally presented in the methods section as a sensitivity analysis and has now been moved to the Methods section of the main paper (**NOTE TO CO-AUTHORS**: Should we include this in the results or leave it in the methods?).
 It reads on line *XXX* of page *XX*: *We also computed a less conservative estimation of shifting transboundary stocks by including a threshold of one standard deviation (i.e., representing a probability of 68% that the stock had shifted from natural variation; see methods).
@@
 
-- **Response:** Thank you for this comment. We have eddited the paragraph starting at line **13** of page **12** to reflect this comment.
+- **Response:** Thank you for this comment. We have edited the paragraph starting at line **13** of page **12** to reflect this comment.
 

### 2021-05-11 Scripts/Reviews/reviewer_three.Rmd
Initial round of revisions done

@@
 title: "Responses to Reviewer Three"
-output: pdf_document
+# output: pdf_document
+output: word_document
 ---
-
-- Response: [Not responded] 
-- *Response*: [Draft response here] 
-- **Response**: [Response compleded her and in manuscript]
 
@@
 
-- Informing global fisheries sustainability and meeting the SDG 2030 Agenda is a useful framing, in my opinion.
+Informing global fisheries sustainability and meeting the SDG 2030 Agenda is a useful framing, in my opinion.
 However, given that framing and the research questions posed in this paper, I think the authors have overlooked two very important cases: 1) stocks that are currently confined to a single EEZ (“national stocks”) that will shift into transboundary stocks by 2100, and 2) stocks that are currently transboundary stocks that will shift entirel into a single EEZ by 2100, becoming national stocks.
@@
 
- - Pg 9, Line 6: is 59% the proportion of stocks that will shift their catch distribution? Or is it the average proportion of the catch that shifts between countries?
+Pg 9, Line 6: is 59% the proportion of stocks that will shift their catch distribution? Or is it the average proportion of the catch that shifts between countries?
 
@@
 
-- The methods are ENTIRELY in the supplementary materials.
+The methods are ENTIRELY in the supplementary materials.
 I have no problem with more abbreviated methods in the main paper and more details in the supplement, but there needs to be some methods in the main paper.
@@
 
-- Pg 31-32: Does the TI capture the equality of sharing of the stock between two countries, or just how stable it is over time?
+Pg 31-32: Does the TI capture the equality of sharing of the stock between two countries, or just how stable it is over time?
 The text claims an increasing TI means management is getting more challenging and a decreasing TI that the sharing between countries is becoming more stable (but not necessarily more equal)….is that the right interpretation?
@@
 
-*Response:* Thank you for your comment.
+- **Response:** Thank you for your comment.
 The transboundary index only captures the sharing stability over time.
@@
 Thus, since we do not talk about the directionallity of the TI (nor the TI *per se*) we have removed this sentence to avoid misunderstanding.
+Please note that the equality of sharing is measured by the Stock Share Ratio described under *Quantifying the intensity of transboundary stocks’ range shift.*
 
-The equality of sharing is measured by the Stock Share Ratio described under *Quantifying the intensity of transboundary stocks’ range shift.*
+I like the time of emergence concept and how it was applied in this analysis.
 
-- I like the time of emergence concept and how it was applied in this analysis.
+- **Response:** We appreciate the acknowledgment on our methods.
 
-**Response:** We appreciate the acknowledgment on our methods.
-
-- I also like the idea of the threat point analysis to identify when a stock is sufficiently shared with another country to motivate cooperative management.
+I also like the idea of the threat point analysis to identify when a stock is sufficiently shared with another country to motivate cooperative management.
 However, I was confused by how the threat point was implemented using the stock share ratio (SSR) – why is the threat point determined relative to the historic SSR?
@@
 
-- Response: Missing
+- **Response:** We have revised the text to clarify this point as we believe the reviewer misunderstood our approach.
+We ofer and explanation below, and inlcuded some more detailed in the methods of the main manuscript (see page *XXX*, line *XX*) and in the Supplemental Material (see page *XXX* line *XX*).
 
-- I think a version of Fig.
+  The reviewer is correct to "*assume the idea of a threat point from game theory is based on the proportion of catch or value that is shared"* and that is exactly what we mean here with the stock share ratio (SSR).
+Here, the SSR represents the proportion of a transboundary stock's catch that each country has per year.
+We do realize we loosly used the concepts of SSR and catch proportion when they mean the same thing.
+This has been clarified in the text.
+
+  If we think on two countries managing a transboundary stock from 1951 to 2000, we can assume that both countries are adapted to the historical stock share ratio.
+This is not a single value but a series of combinations (e.g., year 1 40/50, year 2 30/60, year 3 50/50...
+n).
+Since management continues over the historical time period, we can assume both countries are happy with the historic SSRs.
+However, as game theory states, for a cooperative strategy to work, the benefit a player gets (in this case the SSR) must not be less than the benefit under a non-cooperative strategy.
+While we could set the threat point as a theoretical proportion as suggested by the reviewer, but if we assume countries are collaborating, then the threat point can be the minimum historic SSR that each country had (in our case SSR $\pm$ 2 s.d for a more conservative scenario; see equation **X**).
+This assumes that no country will be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been.
+This method was previously applied by Palacios-Abrantes *et al.* (2020) and Sumaila *et al* (2020).
+
+ - Palacios-Abrantes J, Sumaila UR, Cheung WWL. 2020. Challenges to transboundary fisheries management in North America under climate change. Ecology and Society 25:art41–17.
+
+ - Sumaila UR, Palacios-Abrantes J, Cheung WWL. 2020. Climate change, shifting threat points, and the management of transboundary fish stocks. Ecology and Society 25:art40–9. The Resilience Alliance.
+
+I think a version of Fig.
 S6 would be useful for the main paper.
@@
 
-**Response:** Thank you for noticing the misplace of the EEZ centroids.
+- **Response:** Thank you for noticing the misplace of the EEZ centroids.
 The figure has been corrected and included the figure in the main text (It is now Fig.
@@
 
-- I know this is nitpicky, but please don’t have line numbers restart on each page. It just makes it more cumbersome on a reviewer trying to give detailed comments.
+I know this is nitpicky, but please don’t have line numbers restart on each page. It just makes it more cumbersome on a reviewer trying to give detailed comments.
 
-**Response:** We understand (and agree!) with this comment. Unfortunately, this is how the *Science* template was distributed to us.
+- **Response:** We understand (and agree!) with this comment. Unfortunately, this is how the *Science* template was distributed to us.
 
-- Page1, Line 15: I suggest reporting the proportion of stocks (which looks like it would be between 20-25%), rather than the 78% reported here.
+Page1, Line 15: I suggest reporting the proportion of stocks (which looks like it would be between 20-25%), rather than the 78% reported here.
 Saying that transboundary stocks will shift their distributions across 78% of “the world’s international boundaries” feels a little misleading…..the results show that 78% of EEZs will experience a shifting stock.
@@
 
-**Response:** Thank you, we have adopted the suggested approach.
+- **Response:** Thank you, we have adopted the suggested approach.
 You will read in line 14 of page 1: "*Using global climate-fisheries model simulations, we show that 78% the world’s international boundaries will experience a shifting transboundary stock by 2030*".
 
-- Pg 3, Line 2: First sentence doesn’t reflect the content in the rest of the paragraph. It would work better as a topic sentence for the third paragraph.
+Pg 3, Line 2: First sentence doesn’t reflect the content in the rest of the paragraph. It would work better as a topic sentence for the third paragraph.
 
@@
 
-- Pg 3, Line 7: change “have shifted” to “are shifting” – this paper shows the shifts aren’t done happening yet
+Pg 3, Line 7: change “have shifted” to “are shifting” – this paper shows the shifts aren’t done happening yet
 
@@
 
-- Pg 4, Line 11-12: I don’t understand what is meant by “misalignment between fisheries resources allocation and distribution shifts” – language is confusing.
+Pg 4, Line 11-12: I don’t understand what is meant by “misalignment between fisheries resources allocation and distribution shifts” – language is confusing.
 Do the authors mean misalignment between fishing effort and stock biomass?
@@
 
-- Pg 29: I realize the DBEM model is described in other papers, but a little more detail here would be helpful so that a reader doesn’t have to consult other papers to have a better sense of what the inputs are, how it was parameterized for each stock, etc.
+Pg 29: I realize the DBEM model is described in other papers, but a little more detail here would be helpful so that a reader doesn’t have to consult other papers to have a better sense of what the inputs are, how it was parameterized for each stock, etc.
 I suggest adding another paragraph to this effect.
@@
 
-- Pg 30, lines 8-12: this text is repeated from earlier in the paragraph
+Pg 30, lines 8-12: this text is repeated from earlier in the paragraph
 
@@
 
-- Fig. S4 needs a color bar legend
+Fig. S4 needs a color bar legend
 

### 2021-05-11 Scripts/Manuscript.Rmd
Fixed images links

@@
 subtitle: "Juliano Palacios-Abrantes*^1^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette C.C. Wabnitz^1,5^, and William W.L. Cheung^1^"
-output: bookdown::pdf_document2
+# output: bookdown::pdf_document2
+output: bookdown::word_document2
 reference_docx:  ../References/sci_template.docx
@@
 
-```{r toeFigure1N, eval = F,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
 The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring Exclusive Economic Zones sharing the stock (Top panel).
@@
 2A).
-Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2 and \@ref(tab:toeTableS3) than almost any other region (Figs.
+Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2) and \@ref(tab:toeTableS3) than almost any other region (Figs.
 \@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS1) and \@ref(fig:toeFigureS2)).
@@
 
-\clearpage
-
-**References**
-
 **Methods **
@@
 The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ.
-\@ref(equ:four)).
+\@ref(eq:four)).
 We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
@@
 Finally, we computed changes in the catch proportion that were below the threat point of the neighbouring EEZs sharing a stock for two time periods relative to the historical proportion; 2030 (2021 and 2040) matching the UN 2030 Agenda and a late date (2040 to 2050).
+
+\clearpage
+
+**References**
 

### 2021-05-11 Scripts/Supplements.Rmd
FIxede images links

@@
 This index represents the shift in the distribution centroid of a transboundary stock relative to the centroid of neighboring EEZs that share this stock (Fig.
-\@ref(fig:toeFigureS6)).
+\@ref(fig:toeFigure1)).
 The centroid of a transboundary stock was determined by the average ($\mu$) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across grid cells within the neighboring EEZs sharing the stock.
@@
 We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig.
-\@ref(fig:toeFigureS6) for a graphical description):
+\@ref(fig:toeFigure1) for a graphical description):
 
@@
 
-First, we estimated **the catch proportion as the** stock share ratio (SSR) of each transboundary stock shared by neighboring nations between 1951 and 2100.
+First, we estimated **the catch proportion as the** stock share ratio (SSR) catch of each transboundary stock shared by neighboring nations between 1951 and 2100.
 We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock's catch held within each EEZ [@PalaciosAbrantes:2020tv].
@@
 Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm 2 \sigma$, where  is the standard deviation of $SSR_{th}$.
-Thus, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds two standard deviations of the historical variations in SSR (i.e., when $SSR_{et} \geqq (SSR_{th} + 2 \sigma$) or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
+Thus, a change of SSR catch beyond an EEZ’s threat point happens when the future SSR catch exceeds two standard deviations of the historical variations in SSR (i.e., when $SSR_{et} \geqq (SSR_{th} + 2 \sigma$) or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
 Finally, we estimated the percentage change in SSR ($\Delta SSR$) of each future time period ($SSR_{tf}$) relative to the historic time period $SSR_{th}$ for each stock whose share ratio exceeded the threat point following [@PalaciosAbrantes:2020tv];

### 2021-05-11 Scripts/Reviews/reviewer_three.Rmd
Updates to reviewers responses

@@
 
-- **Response:** We have revised the text to clarify this point as we believe the reviewer misunderstood our approach.
-We ofer and explanation below, and inlcuded some more detailed in the methods of the main manuscript (see page *XXX*, line *XX*) and in the Supplemental Material (see page *XXX* line *XX*).
+- **Response:** We have revised the text to clarify this point as we realize from the reviewer comment that we have not been clear enough.
+We offer an explanation below, and included some more details in the methods section of the main text (see page *XXX*, line *XX*) and in the Supplemental Materials (see page *XXX* line *XX*).
 
 The reviewer is correct to "*assume the idea of a threat point from game theory is based on the proportion of catch or value that is shared"* and that is exactly what we mean here with the stock share ratio (SSR).
-Here, the SSR represents the proportion of a transboundary stock's catch that each country has per year.
-We do realize we loosly used the concepts of SSR and catch proportion when they mean the same thing.
-This has been clarified in the text.
+Here, the SSR represents the proportion of a transboundary stock's catch that each country takes each year.
+We do realize we loosely used the concepts of SSR and catch proportion interchangeably when they mean the same thing.
+This has been now clarified in the text.
 
 If we think on two countries managing a transboundary stock from 1951 to 2000, we can assume that both countries are adapted to the historical stock share ratio.
-This is not a single value but a series of combinations (e.g., year 1 40/50, year 2 30/60, year 3 50/50...
+This is not a single value but a series of combinations of catch proportions (e.g., year 1 40/50, year 2 30/60, year 3 50/50...
 n).
-Since management continues over the historical time period, we can assume both countries are happy with the historic SSRs.
-However, as game theory states, for a cooperative strategy to work, the benefit a player gets (in this case the SSR) must not be less than the benefit under a non-cooperative strategy.
-While we could set the threat point as a theoretical proportion as suggested by the reviewer, but if we assume countries are collaborating, then the threat point can be the minimum historic SSR that each country had (in our case SSR $\pm$ 2 s.d for a more conservative scenario; see equation **X**).
-This assumes that no country will be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been.
+Since management continues over the historical time period, we can assume that both countries are happy with the historic SSRs catch.
+However, as game theory states, for a cooperative strategy to work, the benefit a player gets (in this case the SSR catch) must not be less than the benefit under a non-cooperative strategy.
+While we could set the threat point as a theoretical proportion as suggested by the reviewer, but if we assume countries are collaborating, then the threat point can be the minimum historic SSR catch that each country had (in our case SSR catch $\pm$ 2 s.d for a more conservative scenario; see equation **X**).
+This assumes that no country will be willing to collaborate if the future catch proportion of the transboundary stock it will recieve is lower than it has ever been.
 This method was previously applied by Palacios-Abrantes *et al.* (2020) and Sumaila *et al* (2020).

### 2021-05-11 Scripts/Manuscript.Rmd
Ended first revision

@@
 These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij].
-As transboundary stocks continue to shift, such international disputes are likely to increase in frequency and requiring a revision of management plans [@Pinsky:2018cb; @Engler:2020bg].
+**Early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].**As transboundary stocks continue to shift, international disputes are likely to increase in frequency and stocks' health decrease requiring international collaboration and revision of management plans [@Pinsky:2018cb; @Engler:2020bg; @Maureaud:2021gj].
 
@@
 Moreover, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery they have historically been entitled.
-Finally, changes in policy can take years (e.g., it took the PST about 10 years to agree on a conservation fund) and sometimes, the time of political change is longer than that of species shifts [@Pinsky:2012kq].**
+Finally, changes in policy can take years (e.g., it took the PST about 10 years to agree on a conservation fund) and sometimes, the time of political change is longer than that of species shifts [@Pinsky:2012kq].
+This is specifically true as in many cases we lack standardized data to track shifts of species across international jurisdictions [@Maureaud:2021gj] as these might already be happening.**
 

### 2021-05-11 Scripts/Reviews/reviewer_one.Rmd
Ended first revision

@@
  - **Regarding the bi-lateral agreements**, this is a great question that, unfortunately, we cannot respond at this moment.
-Currently, there is only one database on bi-lateral agreements produce by Oremus *et al* (2020).
-On their paper, the authors state "we analysed 127 publicly available international fisheries agreements in the South Pacific, Southeast Asia, South Asia, Northwest Africa, and Central America and the Caribbean" and " For each agreement, we recorded the following information...
+Currently, there is only one database on bi/multi-lateral agreements produced by Oremus *et al* (2020).
+On their paper, the authors "analysed 127 publicly available international fisheries agreements in the South Pacific, Southeast Asia, South Asia, Northwest Africa, and Central America and the Caribbean" and "For each agreement, we [the authors] recorded the following information...
 If relevant, to which specific stock or species the agreement pertained".
 We contacted Dr.
-(Kimberly) Oremus who provided the provided the mentioned dataset.
+(Kimberly) Oremus who provided  the mentioned data-set.
 However, non of the 127 treaties have identified subjects at the species level (except for bigeye, skipjack and yellowtail tunas).
+Thus, we were unable to match the stocks with existing agreements.
 
 We hope that the reviewer understands the complexity of building such database and that this exercise falls outside the scope and stage of our current analysis.
-We agree that having such information would increase the manuscript impact, but, would not necessarily directly impact our results nor the message we want to covein (Ultimately Oremus *et at* state it very clearly in their paper, "None of the 127 agreements contains language that is directly related to climate change, range shifts or stock exits").
+We agree that having such information would increase the manuscript impact, but, would not necessarily affect our results, nor the message we want to covey (Ultimately Oremus *et at* state it very clearly in their paper, "None of the 127 agreements contains language that is directly related to climate change, range shifts or stock exits").
 
@@
 However, because we have 10 different runs of the GFDL and employ the time of emergence concept, we are capable of providing a less conservative estimate of the numbers you mention (i.e., the number of shifting stocks and EEZs).
-This was originally presented in the methods section as a sensitivity analysis and has now been moved to the Methods section of the main paper (**NOTE TO CO-AUTHORS**: Should we include this in the results or leave it in the methods?).
+This was originally presented in the Supplemental Material section as a sensitivity analysis and has now been moved to the Methods section of the revised paper (**NOTE TO CO-AUTHORS**: Should we include this in the results or leave it in the methods?).
 It reads on line *XXX* of page *XX*: *We also computed a less conservative estimation of shifting transboundary stocks by including a threshold of one standard deviation (i.e., representing a probability of 68% that the stock had shifted from natural variation; see methods).
@@
 
-- **Response:** Yes, "transboundary between neighboring countries” is indeed redundant.
+- **Response:** Yes, "transboundary between neighboring countries” is indeed redundant, thank you for noticing.
 We have modified the manuscript to say "*Climate change is shifting the distribution of shared fish stocks between neighboring countries (transboundary).*" We believe it is important to keep both parts of the sentence for clarification.
@@
 
-- **Response:** Similarly, we do not have space to discuss this here. Please note that we discuss management plans on the paragraph starting on line *13* of page *12*.
+- **Response:** Similarly, we do not have space to discuss this here. Please note that we discuss management plans on the paragraph starting on line *XXX* of page *XX*.
 
@@
 
-- **Response:** We have changed this sentence to *Marine species’ distributions mainly reflect species’ preferences for discrete environmental conditions (2).*
+- **Response:** We have changed this sentence to reflect the suggestion, it now reads: *Marine species’ distributions mainly reflect species’ preferences for discrete environmental conditions (2).*
 
@@
 
-- **Response:** In the section "High present-day climate risk of on transboundary fisheries management" we describe early shifts in the shared distribution of transboundary stocks including the cases of Humboldt squid (*Dosidicus gigas*) and Atlantic mackerel (*Scomber scombrus*) both in the early 2000s and have included a couple more punctual examples of documented shifts between 2006 and 2021.
-However, there is not a vast documentation on shifts in the distribution of transboundary species (i.e., specifically from one EEZ to another).
-Mapping these shifts is out of the scope of the current project as there is an ongoing effort to document the global hifts of transboundary species.
+- **Response:** There is not a vast documentation on shifts in the distribution of transboundary species (i.e., specifically from one EEZ to another).
+This is partially because, from a scientific perspective, there is not enough standardized data to track changes across countries (See Maureaud *et all*, 2021 below).
+Mapping these shifts is out of the scope of the current project as there is an ongoing effort to document the global hotspots of transboundary species.
+That being said, in the section "High present-day climate risk of on transboundary fisheries management" we describe early shifts in the shared distribution of transboundary stocks including the cases of Humboldt squid (*Dosidicus gigas*) and Atlantic mackerel (*Scomber scombrus*) both in the early 2000s.
+Moreover, we have included a couple of other examples of shifts reported in neighbouring countries.
+It reads on line **XX** of page **X** of the revised manuscript: *Early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].*
 
@@
 We provide direction of change in the next figure (Figure 5 of the revised manuscript) showing winners and losers according to neighboring EEZs.
-Here we show how each neighboring EEZ will stand to win or lose against its neighbors.
-**NOTE FOR CO-AUTHORS:** *I am hesitant of actually pointing out winners and losers (e.g., stating that Spain will loos most of its catch to France).
-Hence, I rather leave the results and discussion more general.
-Thoughts?*
+This analysis is based on the time of emergence result, that has no direction, we show how each neighboring EEZ will stand to win or lose against its neighbors.
 
@@
 
-- **Response:** We have modified this sentence in the revised manuscript with the hopes to make it less jargon-y.
-It now reeds on line *7* of *P9* the revised manuscript: "*There will be an increase in the number of stocks changing their catch proportion and the number of EEZs with changing stocks towards the mid-21st century.
+- **Response:** Thank you for noticing this.
+We have modified this sentence in the revised manuscript with the hopes to make it less jargon-y.
+It now reeds on line *XX* of *XXX* the revised manuscript: "*There will be an increase in the number of stocks changing their catch proportion and the number of EEZs with changing stocks towards the mid-21st century.
 Such increase will be followed by an also increase in the intensity of the proportion change (Figs.
@@
 
-- **Response:** Thank you for this comment. We have edited the paragraph starting at line **13** of page **12** to reflect this comment.
+- **Response:** Thank you for this comment. We have edited the paragraph starting at line **XXX** of page **XX** to reflect this comment.
 
@@
 
-- **Response:** Thank you for highlighting this. We have changed the sentence to reflect this comment.
+- **Response:** We agre and have changed the sentence to reflect this comment.
 

### 2021-05-11 Scripts/Reviews/reviewer_three.Rmd
Ended first revision

@@
 
-- **Response:** We appreciate the reviewer's comment regarding our method and understand the concern on consider national stocks that become transboundary stocks and *vice versa*, however, these questions have already been responded by Pinsky *et al* (2018) and Oremus *et al.* (2020) and consequently, follow outside the scope of the current paper.
+- **Response:** We appreciate the reviewer's comment regarding our method and understand the concern on consider national stocks that become transboundary stocks and *vice versa*, however, these questions follow outside the scope of the current paper as they have already been responded by Pinsky *et al* (2018) and Oremus *et al.* (2020).
 
 In their 2018 paper, Pinsky *et al* look at how many *new* transboundary species each EEZ of the world will have towards the end of the century under climate change (i.e., national stocks that become transboundary stocks).
-Indeed, the authors find that "*many of the world’s EEZs are likely to receive one to 12 new, climate-driven transboundary stocks by the end of the century*." On the other hand, Oremus *et al* (2020) "*estimates the number of species exiting national jurisdictions under climate change up to 2100*" concluding that *tropical nations stand to lose the most species under these scenarios.* However, both studies fail to identify changes in the *current* transboundary stock, gap that we look to fill int his paper.
+Indeed, the authors find that "*many of the world’s EEZs are likely to receive one to 12 new, climate-driven transboundary stocks by the end of the century*." On the other hand, Oremus *et al* (2020) "*estimates the number of species exiting national jurisdictions under climate change up to 2100*" concluding that *tropical nations stand to lose the most species under these scenarios.* However, both studies fail to identify changes in *current* transboundary stocks, gap that we look to fill int his paper.
 Here, we used the species (n = 633) identified by Palacios-Abrantes *et al* (2020) as transboundary and consequently do not have national stocks identified.
 This was on purpose as the objective of our paper is precisely to look at the impacts of climate change on *current* transboundary species.
-Addressing the question of entering/exciting stocks would require us to change our initial species list, change our otherwise already complex and long methodology, and likely give the paper a different perspective.
+Addressing the question of entering/exciting stocks would require us to change our initial species list, change our otherwise already dense methodology as you rightfully point out, and likely give the paper a different perspective.
 
@@
 
- - **Response:** We meant that 59% is the average proportion of the catch that shifts between countries, or more specific, the stocks reponsable for that catch.
+ - **Response:** We meant that 59% is the average proportion of the catch that shifts between countries, or more specific, the stocks responsible for that catch.
 We have revised the manuscript to clarify this, it reads on line **XX** of page *X*: "*Findings show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks is projected to change beyond an EEZ’s historic proportion (i.e., between 2004 and 2010; Fig.
@@
 
-- **Response:** The main structure was following the journal's guideline. We have now included a summarized version of the methods in the main article.
+- **Response:** The main structure was following the journal's guideline. We have now included a summarized version of the methods in the main article. Please see line **XXX** of page **XX*.*
 
@@
 - **Response:** Thank you, we have adopted the suggested approach.
-You will read in line 14 of page 1: "*Using global climate-fisheries model simulations, we show that 78% the world’s international boundaries will experience a shifting transboundary stock by 2030*".
+It reads now in line 14 of page 1 of the revised manuscript: "*Using global climate-fisheries model simulations, we show that 78% the world’s international boundaries will experience a shifting transboundary stock by 2030*".
 
@@
 
-- **Response:** Thank you for noticing this. We have changed the wording to reflect this comment.
+- **Response:** Thank you for noticing this. We have changed the wording to reflect this comment in the revised manuscript.
 
@@
 
-- **Response:** We have included one of the examples we referenced, it now reads in line *XX* of page *xxx in the revised manuscript:* *"For example, Pacific salmon (Oncorhynchus sp.) jointly managed by Canada and the U.S.
+- **Response:** We have included one of the examples we referenced, it now reads in line *XX* of page *xxx *in the revised manuscript:* *"For example, Pacific salmon (Oncorhynchus sp.) jointly managed by Canada and the U.S.
 was heavily fished in the early 1990s after a stock shift favored the Alaskan fleet over the Canadian fleet (18).*"

### 2021-05-17 Scripts/Supplements.Rmd
Incorporate THomas' comments

@@
 
-The DBEM was driven by simulated ocean conditions from a ten-ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+The DBEM was driven by simulated ocean conditions from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
 The GFDL-ESM2M was run under historical forcing until 2005 and follows a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5), over the 2006-2100 period [@Riahi:2011dk].
-We chose this time period to match the historical period in the GFDL ESM2M simulations over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Rodgers:2015jq].
+We chose the 1951-2100 time period to match the historical period in the GFDL ESM2M simulations over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Rodgers:2015jq].
 Because the main approach of this paper relies on understanding the spatial and temporal variation of a stock’s distribution, we have to understand distribution variability during both historical and future periods, to infer differences between time frames.
@@
 By design, variations among ensemble members are then solely due to natural internal variability (e.g., different phases of the El Niño Southern Oscillation-ENSO).
-Thus, for our experiment, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in the year 1950, resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
+Thus, for our experiment, each of the ten GFDL-ESM2M ensemble simulations were started from infinitesimally small differences in Earth system initial conditions in the year 1950, resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
 
@@
 It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution.
+The ToE was also computed for a less conservative estimation using a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
+In this case, the global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
+\@ref(fig:toeFigureS8)).
+In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
+\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
 

### 2021-05-17 Scripts/Manuscript.Rmd
incorporating changes from thomas

@@
 The imminent onset of transboundary shifts highlights the urgency to re-formulate international fisheries policies to limit conflict over these fish stocks.
-Bi- and multi-lateral actions are urgently needed as lead-time for policy change is long.
+**A strong expansion of bi- and multi-lateral actions** are urgently needed as lead-time for policy change is long.
 
@@
 We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
-In this case, the global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
-\@ref(fig:toeFigureS8)).
-In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
-\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
 

### 2021-05-19 Scripts/Initial_Analysis.Rmd
misc

@@
 
-       ### Code to produce graphs for Brazil and ruguay
+       ### Code to produce graphs for Brazil and Uruguay
 
@@
 
-# GetCentroids(taxon_key = 600076,
+# GetCentroids(taxon_key = 600004,
 #              year = year_df$year[1:5],
@@
 #              Neighbours = Neighbours,
-#              eez_centroid = eez_centroid
+#              eez_centroid = eez_centroid,
+#              p = 0.95
 #              )
@@
     filter(year <= Hist_y) %>% 
-    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
+    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first change to ensembe
     summarise_at(vars("trans_index"),
@@
                  x = year,
-                 y = trend,
+                 y = trend
+               ),
                colour = "red",
-               ),
                shape = 1,
@@
       } +
+        geom_line(data = Trans_index,
+              aes(
+                x = year,
+                y = RMean,
+                color = as.factor(ensemble)
+              ),
+              alpha = 0.3
+    ) +
     geom_vline(xintercept = Hist_y,colour="blue",alpha = 0.5) +
-    scale_x_continuous(breaks = seq(1950,2100,5)) +
+    scale_x_continuous(breaks = seq(1960,2100,5),limits = c(1960,2100)) +
+    scale_color_brewer("Ensembles", palette = "Greens") +
     theme_classic()
@@
 
-  }
-
-  # If no historic data then emergin year is the first year of data
+  } #End test plot
+
+  # If no historic data then emerging year is the first year of data
   if(nrow(historic_sd) == 0){
@@
 # Last year of "historic" data
-Historic_year = 2005
+Historic_year <- 2005
 

### 2021-05-20 Scripts/Initial_Analysis.Rmd
Sensitivity of alternative analysis

--- 2021-05-10 Scripts/Initial_Analysis.Rmd
@@
     filter(year <= Hist_y) %>% 
+    # Alternative method sensitivity analysis
     group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
+    # group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
     summarise_at(vars("trans_index"),
@@
 ### ------------ ##
-# Select a trshold
+# Select a treshold
 selected_tresh <- "tresh_two"
@@
            !is.na(un_sub_region)) %>%
-    mutate(weight = )
+    # mutate(weight = )
     mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
@@
 
-Testing different Per values for the centroids
+### Testing different Per values for the centroids
 
@@
   )
+
+```
+
+### Noise for all ensemble members
+
+```{r read_raw_data}
+
+# --------------- #
+# Load year of emergence results
+# --------------- #
+
+# Taxon_list_emer <- list.files(my_path("R",extra_path = "Emergence_2005"))
+Taxon_list_emer <- list.files("~/Desktop/Emergence_p95_alt")
+
+# clean names for function
+Taxon_list_emer <- gsub("_.*","",Taxon_list_emer)
+
+# Paths <- paste(my_path("R",extra_path = "Emergence_raw_p95"),Taxon_list_emer,"_emergence_raw.csv",sep="")[1:10]
+
+Paths <- list.files("~/Desktop/Emergence_p95_alt",full.names = T)
+
+# Load all sp in one table
+Eme_Alt_Res <- bind_rows(lapply(Paths, fread)) %>% 
+  left_join(Neighbours,
+            # by = c("eez_name","eez_neighbour")
+  ) %>%
+  rename(sf_eez_name=eez_name) %>%
+  left_join(Matching_names,
+            by=  "sf_eez_name") %>%
+  clean_names() %>%
+  mutate(tresh = str_to_lower(tresh)) %>% 
+  rename(emerging_yr = mean_emer)
 

### 2021-05-20 Scripts/Initial_Analysis.Rmd
Ran sensitivity for alternative method

@@
     # Alternative method sensitivity analysis
-    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
+    group_by(taxon_key,neighbour_id,year) %>% #Average historic data per year first
     # group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
@@
     taxon <- unique(Raw_Result$taxon_key)
-    name_raw <- paste(my_path("R",extra_path = "/Emergence_last_raw_p90/"),taxon,"_emergence_raw.csv",sep="")
+    name_raw <- paste(my_path("R",extra_path = "/Emergence_alt_sens_raw/"),taxon,"_emergence_raw.csv",sep="")
 
@@
 
-    name <- paste(my_path("R",extra_path = "/Emergence_last_p90/"),taxon_key,"_emergence.txt",sep="")
+    name <- paste(my_path("R",extra_path = "/Emergence_alt_sens/"),taxon_key,"_emergence.txt",sep="")
 
@@
 # Already modeled species
-spp_done <- list.files(my_path("R", extra_path = "Emergence"))
+spp_done <- list.files(my_path("R", extra_path = "Emergence_alt_sens"))
 
@@
 
+# For plotting test
 rand_plot <- sample(spp_reman,5)
+
+# For alternative method
+alt_sens <- sample(spp_reman,30)
 
@@
     Emerging_result <- bind_rows(
-      mclapply(spp_list,
+      mclapply(alt_sens,
                GetTransIndex,
@@
                eez_centroid = eez_centroid,
-               p = 0.90,
+               p = 0.95, # main is 95
                Plot = F

### 2021-05-21 Scripts/Manuscript.Rmd
Included co-authors comments

@@
 
-**Climate change is shifting the distribution of shared fish stocks between neighboring countries (transboundary).** The time-frame of such shifts determines the extent of climate change challenges to international fisheries governance.
-Using global climate-fisheries model simulations, we show that 78% the world’s international boundaries will experience a shifting transboundary stock by 2030.
-By 2100, 45% of stocks will have shifted for 81% of countries under a ‘no mitigation’ scenario.
-Hotspots of transboundary shifts occur for countries that are highly dependent on fisheries for income, food and livelihoods.
-The imminent onset of transboundary shifts highlights the urgency to re-formulate international fisheries policies to limit conflict over these fish stocks.
-**A strong expansion of bi- and multi-lateral actions** are urgently needed as lead-time for policy change is long.
+**Climate change is shifting the distribution of shared fish stocks between neighboring countries and the high seas.
+The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.** By coupling a large ensemble simulation of an Earth system model to a species distribution model, we show that 78% the world’s Exclusive Economic Zones will experience at least one shifting transboundary stock by 2030.
+By 2100, 45% of stocks would have shifted out of 81% of countries/political entities waters under a high emissions scenario.
+Importantly, hotspots for transboundary shifts occur for countries that are highly dependent on fisheries for income, food and livelihoods.
+The imminent onset of transboundary shifts in fish stocks highlights the urgent need to re-formulate international fisheries policies to limit conflict over these fish stocks.
+**A strong expansion of bi- and multi-lateral actions**
 
@@
 The effectiveness of fisheries management for transboundary species is challenged by shifts in stocks’ distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].
-In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
+In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of a changing climate on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
 While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs].
-For example, Pacific salmon (Oncorhynchus sp.) jointly managed by Canada and the U.S.
+For example, Pacific salmon (*Oncorhynchus sp.*) jointly managed by Canada and the U.S.
 was heavily fished in the early 1990s after a stock shift favored the Alaskan fleet over the Canadian fleet [@Miller:2013iv].
-Such patterns are expected to be exacerbated under continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
+Such patterns are expected to be exacerbated under increasing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
 The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the Sustainable Development Goals (SDG) under the UN 2030 Agenda.
@@
 
-Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (see Methods **and Supplemental Materials**).
+Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (see Methods **and Supplemental Materials**).
 We then estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig.
 \@ref(fig:toeFigure1N)) and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods **and Supplemental Materials**).
-We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
+We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
 
-```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks.
+```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone to another.
 The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring Exclusive Economic Zones sharing the stock (Top panel).
-Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel)."}
+Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel).
+Only showing a subset of ensemble members for clarification."}
 
@@
 **Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
-\@ref(fig:toeFigure1)A), with the first shift occurring as early as 2006.** This corresponds to 45% of the studied stocks (Fig.
+\@ref(fig:toeFigure1)A), with the first shift having occurred back in 2006.** This corresponds to 45% of the studied stocks (Fig.
 \@ref(fig:toeFigure1)C).
@@
 of all species, $\pm$ 28 years).
-Furthermore, 81% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with 78% having recorded a distributional shift in stocks by the UN 2030 Agenda’s deadline (i.e., year 2030) (Fig.
+Furthermore, 81% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with 78% recording a distributional shift in stocks by the UN 2030 Agenda’s deadline (i.e., year 2030) (Fig.
 \@ref(fig:toeFigure1)C).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "**Timeline of shifts in transboundary shared stock distribution.** (A) Number of stocks (cumulative) with their distribution centroids that have shifted beyond natural variations, (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one shifting stock, beyond natural variability.
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "**Timeline of shifts in transboundary shared stock distribution.** (A) Number of stocks (cumulative) with their distribution centroids that have shifted beyond natural variations, (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one stock shifting beyond natural variability.
 Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change in range shift (see Methods), respectively."}
@@
 In contrast, EEZs and stocks located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts.
-Stocks in countries such as Brazil and New Zealand are not projected to experience range shifts beyond natural variability (Fig.
+Stocks in the EEZ of countries such as Brazil and New Zealand are projected to experience no range shifts beyond natural variability by the end of this century (Fig.
 \@ref(fig:toeFigure2)A).
@@
 
-We estimated the current (2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs.
-Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD in fishing revenue are projected to shift in the EEZs of 76 countries (Fig.
+We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs.
+Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD generated in fishing revenue are projected to shift in the EEZs of 76 countries (Fig.
 \@ref(fig:toeFigure3)); with 4.2 billion USD accounted by 321 stocks shifting in just eleven countries (Fig.
 \@ref(fig:toeFigure3)).
-In terms of the number of transboundary stocks shifting from historical variability, Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66), emerge as the top five countries (Fig.
+In terms of the number of transboundary stocks shifting from historical variability, the EEZs of Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66), emerge as the top five (Fig.
 \@ref(fig:toeFigure3)).
@@
 **There will be an increase in the number of stocks changing their catch proportion and the number of EEZs with changing stocks towards the mid-21st century.
-Such increase will be followed by an also increase in the intensity of the proportion change (Figs.
+Such increase will be followed by an increase in the intensity of the proportion that changes (Figs.
 \@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).**
@@
 ```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "**Changes in stock proportion of neighboring Exclusive Economic Zones by 2030 (2021-2040) relative to 1951-2005.** Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors.
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors.
 Panel B highlights changes for the Caribbean and Northwestern Europe."}
@@
 
-The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of the EEZs (Fig.
+The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the shape of the EEZs (Fig.
 \@ref(fig:toeFigure4)A).
-In some cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the proportion of the  stock is expected to follow a projected poleward shift.
-However, shifts along the coasts of Pacific Central America and West Africa occur along an equatorial direction.
-Here, transboundary species follow local oceanographic processes like upwelling zones (e.g., Humboldt current) and patterns of oxygen availability, in addition to temperature gradients.
-Regions where EEZs are relatively small and have multiple borders (e.g., Caribbean or Northwestern Europe) are expected to have a particularly complex exchange of stock share ratio with no established pattern (Fig.
+In cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the proportion of the stock is expected to follow a projected poleward shift.
+However, shifts in the coasts of Pacific Central America and West Africa occur along an equatorial direction.
+Here, transboundary species follow specific oceanographic regions (*sensu* Longhurst) like upwelling zones (e.g., Humboldt current) and local patterns of oxygen availability, in addition to temperature gradients.
+Regions where EEZs are relatively small and have multiple borders (e.g., Caribbean or Northwestern Europe) are expected to have a particularly complex exchange of stock share ratio with no clear established pattern (Fig.
 \@ref(fig:toeFigure4)B).
@@
 
-**High present-day climate risk of on transboundary fisheries management**
+**High present-day climate risk of transboundary fisheries management**
 
@@
 These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij].
-**Early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].**As transboundary stocks continue to shift, international disputes are likely to increase in frequency and stocks' health decrease requiring international collaboration and revision of management plans [@Pinsky:2018cb; @Engler:2020bg; @Maureaud:2021gj].
+**Early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].
+While these represent some of the documented cases of species shifting across jurisdictions, to date is not very clear which and how many transboundary stocks are shifting, where are they shifting and whether they are jointly managed.
+Monitoring present and future shifts in the distribution of transboundary stocks is key to achieve sustainable goals as shifts in distributions are likely to increase the frequency of international disputes and decrease stocks’ health [@Pinsky:2018cb; @Engler:2020bg; @Oremus:2020id].
+This is particularly true as climate change continues to change the distribution of transboundary stocks to a point that some EEZs will see new transboundary stocks [@Oremus:2020id] while others will stand to lose them completely [@Pinsky:2018cb].**
 
@@
 Our results identify regional “hotspots” of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figures \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure4)A).
-Such regions, like the Caribbean, are characterized by high levels of warming relative to natural internal variability [@IPCC:2019tn; @Hawkins:2012hc] and species’ high vulnerability to warming waters [@IPCC:2019tn].
-Moreover, they represent relatively small EEZs that border multiple countries.
+Regions such as the Caribbean, are characterized by high levels of warming relative to historical natural variability [@IPCC:2019tn; @Hawkins:2012hc] and species’ high vulnerability to warming waters [@IPCC:2019tn].
+Moreover, they represent relatively low area EEZs that border multiple countries.
 Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting transboundary stocks.
@@
 **Management plans that are not designed to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
-Thus, focusing on these “hotspots” is key to anticipate a potential increase in fisheries conflicts in coming years.
+Thus, focusing on these “hotspots” is key to anticipating a potential increase in fisheries conflicts in coming years.
 Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments, which do not have to be monetary [@Tunca:2019if; @Miller:2013iv], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@PalaciosAbrantes:2020vt; @Pinsky:2018cb].
@@
 The PNA manage their fisheries under a so-called Vessel Day Scheme, which accounts for shifts in stocks and catches, and is therefore somewhat anticipatory [@Aqorau:2018bh].
-Quota allocation methods based on a stock’s current distribution and/or on fixed-historical proportion like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-However, even management plans that take into consideration such strategies might not be fully prepared for the consequences of shifting transboundary stocks [@Pinsky:2018cb; @Koubrak:2020hw; @Engler:2020bg].
-For example species’ shifts within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+Quota allocation methods based on a stock’s current distribution and/or on fixed-historical proportion like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].**
+
+**Even management plans that take into consideration such strategies might not be fully prepared for the consequences of shifting transboundary stocks [@Pinsky:2018cb; @Koubrak:2020hw; @Engler:2020bg].
+For example, species’ shifts within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
 Moreover, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery they have historically been entitled.
 Finally, changes in policy can take years (e.g., it took the PST about 10 years to agree on a conservation fund) and sometimes, the time of political change is longer than that of species shifts [@Pinsky:2012kq].
-This is specifically true as in many cases we lack standardized data to track shifts of species across international jurisdictions [@Maureaud:2021gj] as these might already be happening.**
+This is specifically true we could be missing shifting stocks as we lack standardized data to track shifts across international jurisdictions [@Maureaud:2021gj].
+Moreover, recent efforts looking at international fisheries plans found that all 127 plans lacked direct actions to address topics of climate change or species range shifts across jurisdictions.
+Furthermore, authors were rarely able to identify the stock in question, and when they did, it was as broad as skipjack tuna or billfishes [@Oremus:2020id].
+Having a better understanding of what stocks are managed as transboundary and under what rules would be an important complement to our study and an important step towards the sustainability of international stocks.**
 
@@
 Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and *ii*) model structure [@Frolicher:2016dg].
-First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.
-Thus, any mitigation efforts (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].
-Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
+First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.**Thus, any mitigation efforts could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].** Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
 This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is often larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
@@
 Roughly, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
-The DBEM was driven for ten different versions of the GFDL (i.e., a ten-ensemble member simulation) each of them representing an unique atmosphere and ocean state at the start of the simulation [@Dunne:2012bo; @Dunne:2013db; @Frolicher:2020cc].
+The DBEM was driven with ten different simulations of the GFDL ESM2M (i.e., a ten-ensemble member simulation) each of them representing a unique atmosphere and ocean state at the start of the simulation [@Dunne:2012bo; @Dunne:2013db; @Frolicher:2020cc].
 The simulations were run under historical forcing until 2005, afterwards and until 2100, simulations followed a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5) [@Riahi:2011dk].
@@
 
-Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Clark:1975uu; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (See Supplemental Material - *Quantifying the intensity of transboundary stocks’ range shift*).
+Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (See Supplemental Material - *Quantifying the intensity of transboundary stocks’ range shift*).
 A threat point is defined as the minimum payoff a player would be willing to receive to cooperate in a game-theoretic model.
 Here, we define the minimum payoff as two standard deviations from the historical catch proportion (average of 1951 to 2005) that each EEZ had for each stock.
-This assumes that countries are adapted to the historical sharing but no country will be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+This assumes that countries are adapted to the historical sharing, but no country will be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
 Finally, we computed changes in the catch proportion that were below the threat point of the neighbouring EEZs sharing a stock for two time periods relative to the historical proportion; 2030 (2021 and 2040) matching the UN 2030 Agenda and a late date (2040 to 2050).

### 2021-05-21 Scripts/Manuscript.Rmd
Included ticks on Fig2

@@
 These impacts are compromising our capacity to reach international sustainability goals under the United Nations (UN)  2030 Agenda for Sustainable Development such as Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
-The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management, including for species that cross international borders, also known as ‘shared stocks’ [@Pinsky:2018cb].
+The projected risks and impacts can be reduced by improving the effectiveness of current governance and fisheries management, including for species that cross international borders, also known as ‘shared stocks’ [@Pinsky:2018cb].
 
-The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by Coastal States [@UnitedNations:1986tl].
-As defined by the UN's Food and Agriculture Organization, shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and (*iv*) discrete stocks that are only present in the high seas [@Munro:2004th].
+The concept of shared stocks was developed following the ratification of the UN on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by Coastal States [@UnitedNations:1986tl].
+As defined by the UN's Food and Agriculture Organization, shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and (*iv*) discrete stocks that are only present on the high seas [@Munro:2004th].
 This study focuses on transboundary stocks exploited by fisheries operating within EEZs.
@@
 
-Most fisheries management is poorly adapted to address species’ shifting distributions as a result of changing ocean conditions.
+Most fisheries management is poorly adapted to address species’ shifting distributions because of changing ocean conditions.
 The effectiveness of fisheries management for transboundary species is challenged by shifts in stocks’ distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id].
 In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of a changing climate on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw].
-While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs].
-For example, Pacific salmon (*Oncorhynchus sp.*) jointly managed by Canada and the U.S.
-was heavily fished in the early 1990s after a stock shift favored the Alaskan fleet over the Canadian fleet [@Miller:2013iv].
-Such patterns are expected to be exacerbated under increasing climate change [@Pinsky:2018cb; @Sumaila:2011bq].
-The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the Sustainable Development Goals (SDG) under the UN 2030 Agenda.
+While many transboundary stocks experience natural seasonal variation in their distribution (e.g., migration, reproductive cycle), misalignment between fisheries resources' allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Spijkers:2017ij; @Crespo:2020cs].
+For example, Pacific salmon species (*Oncorhynchus sp.*), jointly managed by Canada and the U.S., were heavily fished in the early 1990s after a stock shift favored the Alaskan fleet over the Canadian fleet [@Miller:2013iv].
+Continuing climate change is expected to exacerbate such patterns [@Pinsky:2018cb; @Sumaila:2011bq].
 Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the goals set out under the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].
@@
 
-```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone to another.
-The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring Exclusive Economic Zones sharing the stock (Top panel).
+```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone (EEZ) to another.
+The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring EEZ sharing the stock (Top panel).
 Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel).
-Only showing a subset of ensemble members for clarification."}
+Only showing a subset of ensemble members for clarity."}
 
@@
 **Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
-\@ref(fig:toeFigure1)A), with the first shift having occurred back in 2006.** This corresponds to 45% of the studied stocks (Fig.
+\@ref(fig:toeFigure1)A), with the first shift modelled to have occurred back in 2006.** This corresponds to 45% of the studied stocks (Fig.
 \@ref(fig:toeFigure1)C).
@@
 of all species, $\pm$ 28 years).
-Furthermore, 81% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with 78% recording a distributional shift in stocks by the UN 2030 Agenda’s deadline (i.e., year 2030) (Fig.
+Furthermore, 81% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with 78% recording a distributional shift in stocks by the UN 2030 Agenda’s deadline (Fig.
 \@ref(fig:toeFigure1)C).

### 2021-05-21 Scripts/Manuscript.Rmd
Addressed Colettes eddits up to results

@@
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "**Timeline of shifts in transboundary shared stock distribution.** (A) Number of stocks (cumulative) with their distribution centroids that have shifted beyond natural variations, (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one stock shifting beyond natural variability.
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "**Timeline of shifts in transboundary shared stock distribution.** (A) Number of stocks (cumulative) whose distribution centroids that have shifted beyond natural variations over time (from 2006 to >2100), (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one stock shifting beyond natural variability.
 Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change in range shift (see Methods), respectively."}
@@
 The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig.
-\@ref(fig:toeFigureS1)) and geographic region of the neighboring EEZs (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Table S3 and Fig.
+\@ref(fig:toeFigureS1)) and EEZs' geographic regions (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Table S3 and Fig.
 2A).
@@
 \@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS1) and \@ref(fig:toeFigureS2)).
-In contrast, EEZs and stocks located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts.
-Stocks in the EEZ of countries such as Brazil and New Zealand are projected to experience no range shifts beyond natural variability by the end of this century (Fig.
+In contrast, EEZs and stocks located in temperate regions, like northern Europe and eastern Asia, are projected to experience much later range shifts.
+Stocks in Brazil and New Zealand's EEZ are projected to experience no range shifts beyond natural variability by the end of this century (Fig.
 \@ref(fig:toeFigure2)A).
@@
 On average, shifting transboundary stocks represented 23% (s.d.
-$\pm$  22%) of yearly fishing revenue from fisheries targeting **transboundary** stocks within global EEZs between 2004 and 2010 (Fig.
+$\pm$  22%) of annual revenue from fisheries targeting **transboundary** stocks across the world's EEZs between 2004 and 2010 (Fig.
 \@ref(fig:toeFigure2)A).
@@
 
-```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "**Year of change in the shared distribution of 4,119 transboundary stocks.** A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
+```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "**Year of change in the shared distribution of 4,119 transboundary stocks.
+A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
 Exclusive Economic Zone polygons display average year of range shifts within them.
-EEZs with no distributional shift between 2006 and 2100 are represented in aqua color.
+Warm colours are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cool colours represent a late shift/low fishing revenue.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
 B) Year of shared distribution shifts by UN sub-regions.
-Points color coded by the median year of range shift.
-Horizontal dashed line represent the year by which countries have committed to reach full implementation of the 2030 Agenda.
+Points color coded by the average year of range shift.
+Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda.
 N = North, S = South, W = West and E = East.
@@
 \@ref(fig:toeFigure3)).
-In terms of the number of transboundary stocks shifting from historical variability, the EEZs of Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66), emerge as the top five (Fig.
+The number of transboundary stocks shifting from historical variability was highest in the EEZs of Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66) (Fig.
 \@ref(fig:toeFigure3)).
 
-```{r toeFigure3, eval = T,  echo = F, fig.width=8, fig.height=6, fig.pos= "H", fig.cap =  "**Average range shift of transboundary stocks and corresponding fishing revenue per country.** Countries are coded according to different colored symbols representative of 5 distinct regions.
-Horizontal line separates countries above and below the 75th percentile of annual fishing revenue.
-Showing country names for those within the top 75th percentile projected to have an average shift within 5 years after the 2030 Agenda."}
+```{r toeFigure3, eval = T,  echo = F, fig.width=8, fig.height=6, fig.pos= "H", fig.cap =  "**Average range shift of transboundary stocks and corresponding fishing revenue per country.
+Countries are coded according to different colored symbols representative of 5 distinct regions.
+The horizontal line separates countries above and below the 75th percentile of annual fishing revenue.
+Showing country names for the eleven nations above the 75th percentile and projected to have an average shift across transboundary stocks within its EEZ within 5 years of the 2030 Agenda deadline."}
 
@@
 \@ref(fig:toeFigure4)A).
-Moreover,  85% (n = 239) of the world EEZs would have experienced changes in catch proportion of transboundary stocks by 2030.
+Moreover, 85% (n = 239) of the world's EEZs will have experienced changes in catch proportion of transboundary stocks by 2030.
 **There will be an increase in the number of stocks changing their catch proportion and the number of EEZs with changing stocks towards the mid-21st century.
@@
 
-```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "**Changes in stock proportion of neighboring Exclusive Economic Zones by 2030 (2021-2040) relative to 1951-2005.** Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "**Changes in stock proportion of neighboring Exclusive Economic Zones by 2030 (2021-2040) relative to 1951-2005.** Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
 Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors.
@@
 
-The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the shape of the EEZs (Fig.
+The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the shape of EEZs (Fig.
 \@ref(fig:toeFigure4)A).
-In cases, like the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the proportion of the stock is expected to follow a projected poleward shift.
+Along the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the proportion of the stock is expected to follow a projected poleward shift.
 However, shifts in the coasts of Pacific Central America and West Africa occur along an equatorial direction.

### 2021-05-21 Scripts/Manuscript.Rmd
Inlcuded Colette's main eddits. Missing some specific comments

@@
 In contrast, EEZs and stocks located in temperate regions, like northern Europe and eastern Asia, are projected to experience much later range shifts.
-Stocks in Brazil and New Zealand's EEZ are projected to experience no range shifts beyond natural variability by the end of this century (Fig.
+Stocks in a number of EEZs including Brazil and New Zealand’s EEZs are projected to experience no range shifts beyond natural variability by the end of this century (Fig.
 \@ref(fig:toeFigure2)A).
@@
 We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs.
-Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD generated in fishing revenue are projected to shift in the EEZs of 76 countries (Fig.
+Within five years of the UN 2030 Agenda, 979 transboundary stocks accounting for USD 5.3 billion in fishing revenue are projected to shift within 76 EEZs (Fig.
 \@ref(fig:toeFigure3)); with 4.2 billion USD accounted by 321 stocks shifting in just eleven countries (Fig.
@@
 
-The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the shape of EEZs (Fig.
+The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
 \@ref(fig:toeFigure4)A).
@@
 
-Our findings highlighting early shifts in the shared distribution of transboundary stocks concur with previous studies that have detected and attributed changes in marine catch composition [e.g., @Cheung:2013gk; @Frainer:2017ji; @Last:2011fd].
-For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US) as a response to climatic, oceanographic and ecological changes [@Zeidberg:2007gs].
+Our findings highlighting early shifts in the shared distribution of transboundary stocks concur with previous studies that have detected changes in marine catch composition and attributed these to climate change [e.g., @Cheung:2013gk; @Frainer:2017ji; @Last:2011fd].
+For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US), in response to climatic, and associated oceanographic and ecological changes [@Zeidberg:2007gs].
 A new fishery targeting Humboldt squid quickly developed on the heels of the species’ range expansion [@Pinsky:2014by].
 In the northeast Atlantic, fisheries on Atlantic mackerel (*Scomber scombrus*) are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
-However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the fishery’s total allowable catch and a further 18% in 2008, without consultation with NEAFC, threatening the sustainability of the stock [@Spijkers:2017ij].
+However, a range expansion of Atlantic mackerel due to environmental variation into Icelandic waters in 2007 resulted in Iceland capturing 6% of the fishery’s total allowable catch and a further 18% in 2008, without consultation with NEAFC, threatening the sustainability of the stock [@Spijkers:2017ij].
 These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij].
@@
 
-**Hotspot of climate risk on transboundary fisheries management**
+**Climate risk hotspots for transboundary fisheries management**
 
 Our results identify regional “hotspots” of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figures \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure4)A).
-Regions such as the Caribbean, are characterized by high levels of warming relative to historical natural variability [@IPCC:2019tn; @Hawkins:2012hc] and species’ high vulnerability to warming waters [@IPCC:2019tn].
+Regions such as the Caribbean are characterized by high levels of warming relative to historical natural variability [@IPCC:2019tn; @Hawkins:2012hc] and species’ high vulnerability to warming waters [@IPCC:2019tn].
 Moreover, they represent relatively low area EEZs that border multiple countries.
-Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting transboundary stocks.
+Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei].
+Thus, coordinating the management of shifting transboundary stocks for countries in these regions will be particularly challenging.
 
-**Management plans that are not designed to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
-Thus, focusing on these “hotspots” is key to anticipating a potential increase in fisheries conflicts in coming years.
+**Most management plans that are not designed or prepared to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
+Focusing on identified “hotspots” is key to anticipating a potential increase in and averting fisheries conflicts in coming years.
 Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments, which do not have to be monetary [@Tunca:2019if; @Miller:2013iv], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@PalaciosAbrantes:2020vt; @Pinsky:2018cb].
-Some examples of current treaties that have adapted some of these strategies include the Pacific Salmon Treaty between Canada and the U.S.
-to manage Pacific salmon (Oncorhynchus sp.) that has a conservation fund that works as a side payment [@Miller:2013iv], the International Pacific Halibut Commission between Canada and the United States that manages Pacific halibut (*Hippoglossus stenolepis*) and allocates quota based on the stock’s yearly distribution [@PalaciosAbrantes:2020vt], and the Parties to the Nauru Agreement (PNA) in the Pacific Island States that manage skipjack tuna (*Katsuwonus pelamis*).
-The PNA manage their fisheries under a so-called Vessel Day Scheme, which accounts for shifts in stocks and catches, and is therefore somewhat anticipatory [@Aqorau:2018bh].
-Quota allocation methods based on a stock’s current distribution and/or on fixed-historical proportion like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].**
+Some examples of current treaties that have adapted some of these strategies include: the Pacific Salmon Treaty between Canada and the U.S.
+to manage Pacific salmon (Oncorhynchus sp.) and which has a conservation fund functioning as a side payment [@Miller:2013iv], the International Pacific Halibut Commission between Canada and the United States, which as the name implies manages Pacific halibut (*Hippoglossus stenolepis*) and allocates quota based on the stock’s yearly distribution [@PalaciosAbrantes:2020vt]; and the Parties to the Nauru Agreement (PNA) an Oceania subregional agreement among 8 Pacific Island countries and Tokelau, which collectively manage the largest tuna fishery in the world, controlling around 50% of the global supply of skipjack tuna (*Katsuwonus pelamis*).
+The PNA manage fishing effort under a so-called Vessel Day Scheme, an adaptive and equitable system which accounts for shifts in stocks and catches as a result of climate variability [@Aqorau:2018bh].
+Quota allocation methods based on a stock’s current distribution and/or according to fixed-historical proportions like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].**
 
 **Even management plans that take into consideration such strategies might not be fully prepared for the consequences of shifting transboundary stocks [@Pinsky:2018cb; @Koubrak:2020hw; @Engler:2020bg].
-For example, species’ shifts within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
-Moreover, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery they have historically been entitled.
+For example, species’ shifts within the PNA area will likely also have to deal with stocks expanding to new jurisdictions, issue that the NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+Moreover, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery they have historically been entitled to.
 Finally, changes in policy can take years (e.g., it took the PST about 10 years to agree on a conservation fund) and sometimes, the time of political change is longer than that of species shifts [@Pinsky:2012kq].
@@
 
-The early projected range shifts of a large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability.
+The early projected range shifts of large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability.
 For example, sea surface temperature (SST) is projected to increase beyond historical variability by 2030 in 50 to 70% of the global ocean [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
 Results from multiple marine ecosystem models show that SST and primary production (NPP) are the main indicators of species distributional changes across ocean basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
-Specifically, in the species distribution model used here (see methods), SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results.
-The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2019tn], while also explaining the later, and sometimes non-existent, shifts at higher latitudes (Fig.
+Specifically, in the species distribution model used here (see methods), SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems, while NPP drives temperate regions’ results.
+The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics, where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2019tn], while also explaining the later, and sometimes non-existent, shifts at higher latitudes (Fig.
 \@ref(fig:toeFigure2)B).
-
 Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and *ii*) model structure [@Frolicher:2016dg].
@@
 This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is often larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
-While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available [@Deser:2020bj] and cannot address limitations of process level understanding.
-As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties and will likely continue to do so in the coming decade.
-This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf].
+While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available [@Deser:2020bj] and cannot address limitations of process-level understanding.
+It is unlikely that these complex uncertainties will adequately be addressed in the coming decade.
+This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, and the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf].
 These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh].
-Overall, changes in fish biomass with multiple upper trophic level models show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@Lotze:2019ce].
-Further research that reduces uncertainty in the NPP response at the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to a smaller structural uncertainty of fish and fisheries models [@BryndumBuchholz:2019kh].
+Overall, multiple upper trophic level models present overall agreement in terms of directional change of fish biomass, but variability in the magnitude of that change [@Lotze:2019ce].
+Further research that reduces uncertainty in the NPP response at the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to smaller structural uncertainty of fish and fisheries models [@BryndumBuchholz:2019kh].
 Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks.
 While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca].
-Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identity different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
+Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identify different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
 Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
@@
 
-The global community has set the ambitious goal of managing all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear benefits for a number of other societal goals [@UnitedNations2018; @Singh:2017ds].
-Preparing anticipatory policies to deal with shifting transboundary stocks is key to achieving the SDGs and ensuring effective governance of the world’s fisheries [@Pinsky:2018cb; @Oremus:2020id; @Pecl:2017hu].
-Here, we implemented two steps to demonstrate how to achieve sustainable transboundary fisheries in a changing world [@Link:2010ei].
-First, we identified the transboundary stocks that would likely see shifts in their shared distribution compared to their historical average and the year in which such shifts are predicted to occur.
+The global community has set the ambitious goal of managing all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear benefits for several other societal goals [@UnitedNations2018; @Singh:2017ds].
+Developing anticipatory policies to deal with shifting transboundary stocks is key to achieving the SDGs and ensuring effective governance of the world’s fisheries [@Pinsky:2018cb; @Oremus:2020id; @Pecl:2017hu].
+Here, we developed an approach to inform the sustainable management and governance of transboundary fisheries in a changing world [@Link:2010ei].
+First, we identified the transboundary stocks that would likely see shifts in their shared distribution compared to their historical average and the year in which such shifts would occur.
 Second, we estimated the intensity of such change.
-Our results have direct implications for ocean governance as the data provided can inform regional management bodies and member countries to anticipate the likely climate-driven challenges to transboundary fisheries management.
 While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for shifting transboundary stocks [@Pinsky:2018cb; @PalaciosAbrantes:2020tv].
@@
 
-We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks’ under climate change (see the Supplementary Materials).
-Roughly, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
+We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks under climate change (see the Supplementary Materials).
+In brief, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
 The DBEM was driven with ten different simulations of the GFDL ESM2M (i.e., a ten-ensemble member simulation) each of them representing a unique atmosphere and ocean state at the start of the simulation [@Dunne:2012bo; @Dunne:2013db; @Frolicher:2020cc].
@@
 
-Furthermore, we developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
-The index is based on changes in the distance of the stock's abundance centroid and that of the EEZs that share such stock (Fig.
-\@ref(fig:toeFigure1)A)).
+We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
+The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.
+\@ref(fig:toeFigure1)A).
 We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution (see the Supplementary Materials - *Calculating shifts in the shared distribution of transboundary stocks*).
@@
 Here, we define the minimum payoff as two standard deviations from the historical catch proportion (average of 1951 to 2005) that each EEZ had for each stock.
-This assumes that countries are adapted to the historical sharing, but no country will be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-Finally, we computed changes in the catch proportion that were below the threat point of the neighbouring EEZs sharing a stock for two time periods relative to the historical proportion; 2030 (2021 and 2040) matching the UN 2030 Agenda and a late date (2040 to 2050).
+This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Finally, we computed changes in catch proportion that were below the threat point of neighboring EEZs sharing a stock for two time periods relative to the historical proportion: 2030 (2021 and 2040) matching the UN 2030 Agenda and a mid-century timeline (2040 to 2050).
 

### 2021-05-21 Scripts/Manuscript.Rmd
I cluded William's comments

@@
 **Most management plans that are not designed or prepared to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
-Focusing on identified “hotspots” is key to anticipating a potential increase in and averting fisheries conflicts in coming years.
-Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments, which do not have to be monetary [@Tunca:2019if; @Miller:2013iv], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@PalaciosAbrantes:2020vt; @Pinsky:2018cb].
+Focusing on identified “hotspots” will help anticipate any potential increases in and averting fisheries conflicts in coming years.
+Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments (including non-monetary arrangements) [@Tunca:2019if; @Miller:2013iv], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@PalaciosAbrantes:2020vt; @Pinsky:2018cb].
 Some examples of current treaties that have adapted some of these strategies include: the Pacific Salmon Treaty between Canada and the U.S.

### 2021-05-21 Scripts/Initial_Analysis.Rmd
Ran alternative analysis for revissions

@@
       select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr) %>% 
-      mutate(tresh= NA)
+      mutate(tresh= "NA")
 
@@
     # Second Step
+
+    ### TEST it 
+    # for (i in 1:length(Ids)){
+    #   print(Ids[i])
+    #   print(TransIndex(Step_one,Ids[i]))
+    # }
+
     # Get the transboudary index based on all centroids
@@
 
-    Emerging_result <- bind_rows(
-      mclapply(alt_sens,
+      mclapply(spp_reman,
                GetTransIndex,
@@
                Plot = F
-      )
       )

### 2021-05-22 Scripts/Initial_Analysis.Rmd
Included sensibility analysis

--- 2021-05-20 Scripts/Initial_Analysis.Rmd
@@
   my_ggtheme_p(leg_pos = "", hjust = 0.5) +
+  theme(
+    axis.ticks.x = element_line()
+  ) +
   scale_fill_viridis(discrete = T,
@@
                      end = 0.7
+  ) + theme(
+    axis.ticks.x = element_line()
   )
@@
                      labels = c("EEZs","Stocks")
+  )+theme(
+    axis.ticks.x = element_line()
   )
@@
         x = reorder(un_sub_region,order),
-        y = med,
-        colour = med
+        y = mean,
+        colour = mean
       ),
@@
 
-```{r read_raw_data}
+```{r read_raw_data, eval = T, echo = F}
 
@@
 
-# Taxon_list_emer <- list.files(my_path("R",extra_path = "Emergence_2005"))
-Taxon_list_emer <- list.files("~/Desktop/Emergence_p95_alt")
-
-# clean names for function
-Taxon_list_emer <- gsub("_.*","",Taxon_list_emer)
-
-# Paths <- paste(my_path("R",extra_path = "Emergence_raw_p95"),Taxon_list_emer,"_emergence_raw.csv",sep="")[1:10]
-
-Paths <- list.files("~/Desktop/Emergence_p95_alt",full.names = T)
+sens_paths <- list.files("~/Desktop/Sens_analysis/Emergence_alt_sens",full.names = T)
 
 # Load all sp in one table
-Eme_Alt_Res <- bind_rows(lapply(Paths, fread)) %>% 
+Eme_Alt_Res <- bind_rows(lapply(sens_paths, fread)) %>% 
+  mutate(method= "alt") %>% 
   left_join(Neighbours,
-            # by = c("eez_name","eez_neighbour")
+            by = c("eez_name","eez_neighbour")
   ) %>% 
@@
   mutate(tresh = str_to_lower(tresh)) %>% 
-  rename(emerging_yr = mean_emer)
+  filter(emerging_yr != 2100)
+
+# Combine both datas
+sens_data <- Eme_Res %>% 
+  filter(taxon_key %in% Eme_Alt_Res$taxon_key) %>% 
+  mutate(method = "org") %>% 
+  bind_rows(Eme_Alt_Res)
+
+n_stock <- sens_data %>% 
+  filter(method == "org") %>% 
+  group_by(neighbour_id) %>% 
+  summarise(n_stock = length(unique(taxon_key))) %>% 
+  # View()
+  group_by() %>% 
+  summarise(t_trans_stocks = sum(n_stock)) %>% 
+  pull(t_trans_stocks)
+
+n_eez <- length(unique(sens_data$sf_eez_name))
+
+```
+
+## General numbers
+
+```{r sens_numbers, eval = T, echo = F}
+
+# Total number of emerging stocks
+
+# Global estimates
+sens_data %>% 
+  group_by(tresh,neighbour_id,method) %>% 
+  summarise(
+    emerging_stocks = length(unique(taxon_key))
+    ) %>% 
+  group_by(tresh,method) %>% 
+  summarise(
+    emerging_stocks = sum(emerging_stocks),
+    prop_total = (emerging_stocks/n_stock)*100,
+    )
+
+# Looks like there is no difference in the total number of emerging stocks
+# method emerging_stocks prop_total
+# alt                158       85.9
+# org                147       79.9
+
+# Global average ToE
+sens_data %>% 
+  group_by(tresh,method) %>% 
+  summarise(
+    # emerging_stocks = length(unique(taxon_key)),
+    # prop_total = (emerging_stocks/9132)*100,
+    mean= mean(emerging_yr),
+    sd = sd(emerging_yr),
+    min = min(emerging_yr),
+    max = max(emerging_yr)
+           )
+
+# Looks like the alternative method sets a later mean ToE but no real difference in s.d. min nor max
+# method  mean    sd   min   max
+# alt    2054.  29.2  2007  2099
+# org    2029.  25.4  2006  2098
+
+```
+
+```{r read_raw_data, eval = T, echo = F}
+
+  # Per Stock
+stock_d <-  
+  sens_data %>% 
+  # left_join(Neighbours) %>% 
+  # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
+  group_by(taxon_key,neighbour_id,tresh,method) %>% 
+  summarise(
+    taxon_emer = min(emerging_yr)
+  ) %>% 
+  group_by(taxon_emer,tresh,method) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  group_by(tresh,method) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "stock"
+  )
+
+# max(stock_d$com_sum)
+
+#  EEZ data
+eez_d <-  sens_data %>% 
+  # left_join(Neighbours) %>% 
+  filter(n_ens == 10) %>% 
+  # distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
+  group_by(sf_eez_name,tresh,method) %>% 
+  summarise(
+    taxon_emer = min(emerging_yr)
+  ) %>% 
+  group_by(taxon_emer,tresh,method) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  group_by(tresh,method) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "eez"
+  )
+
+# max(eez_d$com_sum)
+
+# Plot Data
+plot_data <- stock_d %>% 
+  bind_rows(eez_d) %>% 
+  filter(tresh != "NA") %>% 
+  mutate(
+    total = ifelse(level == "stock",n_stock,n_eez),
+    per = com_sum/total*100
+  ) %>% 
+  gather("metric","value",com_sum,per) %>% 
+  filter(tresh == "tresh_two")
+
+# 
+# # Manually include rest of data until 2100
+# cont_eez <- plot_data %>% 
+#   filter(level == "eez") %>% 
+#   arrange(taxon_emer) %>% 
+#   slice(1:24) %>%
+#   select(-1) %>% 
+#   arrange(metric) %>% 
+#   mutate(
+#     value = ifelse(metric == "per",81,228),
+#     taxon_emer = c(seq(2088,2099,1),seq(2088,2099,1))
+#   )
+
+plot_data <- plot_data %>% 
+  # bind_rows(cont_eez) %>% 
+  mutate(
+    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
+    label = ifelse(total == max(stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs")
+  )
+
+```
+
+```{r Time_lapse_plot, eval = F, echo = F}
+
+### Three plots
+
+count_eez_plot <- ggplot() +
+  geom_area(data = subset(plot_data, metric == "com_sum" & level == "eez"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  )  +
+  scale_y_continuous("Cumulative number of EEZs",
+                     expand = c(0,0)
+                     ) +
+  scale_x_continuous("Year of range shift",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     label = c(seq(2000,2075,25),">2100")
+  ) +
+  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
+  scale_fill_viridis(discrete = T,
+                     begin = 0.1, 
+                     end = 0.1
+  ) +
+  facet_wrap(~ method);count_eez_plot
+
+#### Count stock plot 
+count_stock_plot <- ggplot() +
+  geom_area(data = subset(plot_data, metric == "com_sum" & level == "stock"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  )  +
+  scale_y_continuous("Cumulative number of stocks",
+                     expand = c(0,0)
+                     ) +
+  scale_x_continuous("",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     label = c(seq(2000,2075,25),">2100")
+  ) +
+  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
+  scale_fill_viridis(discrete = T,
+                     begin = 0.7, 
+                     end = 0.7
+  ) +
+  facet_wrap(~ method);count_stock_plot
+
+# Percentage plot
+percentage_plot <- ggplot() +
+  geom_area(data = subset(plot_data, level == "eez" & my_facet == "Percentage (%)"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  ) + 
+  geom_area(data = subset(plot_data, level == "stock" & my_facet == "Percentage (%)"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  ) + 
+  scale_y_continuous("Cumulative proportion of the total (%)",
+                     expand = c(0,0)
+  ) +
+  scale_x_continuous("",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     label = c(seq(2000,2075,25),">2100")
+  ) +
+  my_ggtheme_p(leg_pos = "right", hjust = 0.5) +
+  scale_fill_viridis(discrete = T,
+                     begin = 0.1,
+                     end = 0.7,
+                     labels = c("EEZs","Stocks")
+  )+
+  facet_wrap(~ method);percentage_plot
+
+# Legend
+  legend <- get_legend(percentage_plot)
+
+  percentage_plot <- percentage_plot + 
+    theme(legend.position = "")
+
+# Draw figure
+  plot_all <- ggdraw() +
+    # Count stock plot
+    draw_plot(count_stock_plot, x = 0.01, y = 0, width = 0.29, height = 0.90) +
+    # Count EEZ plot
+    draw_plot(count_eez_plot, x = 0.31, y = 0, width = 0.29, height = 0.90) +
+    # Percentage plot
+    draw_plot(percentage_plot, x = 0.61, y = 0 , width = 0.29, height = 0.90) +
+    # legend
+    draw_plot(legend, x = 0.84, y = 0 , width = 0.20, height = 1) +
+    # Labels
+    draw_plot_label(label = c("A", "B","C"), 
+                    size = 20,
+                    x = c(0.01, 0.285,0.59), 
+                    y = c(1,1,1)
+    )
+
+  # Save plot
+  save_plot(
+    "./Figures/cumu_per_freq.png",
+    plot_all,
+    base_height = 5,
+    base_width = 12
+  )
 

### 2021-05-22 Scripts/Initial_Analysis.Rmd
Finished analysis for sensitivity of historical s.d.

@@
 
-sens_paths <- list.files("~/Desktop/Sens_analysis/Emergence_alt_sens",full.names = T)
+fread2 <- function(x){
+
+  tb <- fread(x)
+
+  if(nrow(tb) >0){
+    return(tb)
+  }
+
+}
+
+sens_paths <- list.files(my_path("R","Emergence_alt_sens"),full.names = T)
 
 # Load all sp in one table
-Eme_Alt_Res <- bind_rows(lapply(sens_paths, fread)) %>% 
+Eme_Alt_Res <- bind_rows(lapply(sens_paths, fread2)) %>% 
   mutate(method= "alt") %>% 
@@
 
-## General numbers
+#### General numbers
 
@@
 # Looks like there is no difference in the total number of emerging stocks
-# method emerging_stocks prop_total
-# alt                158       85.9
-# org                147       79.9
+ # tresh     method emerging_stocks prop_total
+# 1 tresh_one alt               6665      116. 
+# 2 tresh_one org               5729      100  
+# 3 tresh_two alt               5258       91.8
+# 4 tresh_two org               4108       71.7
 
@@
 # Looks like the alternative method sets a later mean ToE but no real difference in s.d. min nor max
-# method  mean    sd   min   max
-# alt    2054.  29.2  2007  2099
-# org    2029.  25.4  2006  2098
+#  tresh     method  mean    sd   min   max
+# 1 tresh_one alt    2055.  28.7  2007  2099
+# 2 tresh_one org    2029.  26.7  2006  2099
+# 3 tresh_two alt    2057.  28.6  2007  2099
+# 4 tresh_two org    2036.  28.1  2006  2099
 
@@
 
+#### Area plots 
 ```{r read_raw_data, eval = T, echo = F}
@@
   save_plot(
-    "./Figures/cumu_per_freq.png",
+    "./Figures/cumu_per_freq_sens.png",
     plot_all,
@@
     base_width = 12
+  )
+
+```
+
+#### Comparative map
+
+```{r}
+
+# Per EEZs
+Eez_d_hist <- sens_data %>% 
+  group_by(eezid,sau_fishing_entity,tresh,method) %>% 
+  summarise(
+    mean_yr = mean(emerging_yr,na.rm=T),
+    sd_yr = sd(emerging_yr,na.rm=T),
+    n_taxa = length(unique(taxon_key))
+    ) %>% 
+  gather("metric","value",mean_yr:n_taxa) %>% 
+  mutate(
+    reverse = abs(value-2100)  + 2000
+  )
+
+# Estimate the value (catch and USD) of emerging species
+Eme_value_data <- sens_data %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,tresh,taxon_key,method) %>% 
+  summarise(n()) %>% 
+  left_join(Sau_data,
+            by = c("fishing_entity_id","taxon_key")) %>% 
+  filter(!is.na(variable)) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,tresh,variable,method) %>% 
+  summarise(
+   sum_value = sum(value,na.rm=T),
+    n_taxa = length(unique(taxon_key)) 
+  )
+
+# Total value of each fishing nation
+
+Total_value <- Sau_data %>% 
+  group_by(fishing_entity_id,variable) %>% 
+  summarise(tot_value= sum(value,na.rm=T))
+
+Value_map_data <- Eme_value_data %>% 
+  left_join(Total_value,
+            by = c("fishing_entity_id","variable")) %>% 
+  mutate(participation = (sum_value/tot_value)*100+2000)
+
+# Per fishing entity
+
+fish_d_hist <- sens_data %>% 
+  group_by(sau_fishing_entity,tresh,method) %>% 
+  summarise(
+    value = length(unique(taxon_key))
+    ) %>% 
+  filter(!is.na(sau_fishing_entity))
+
+World_map_nemer <- World_map %>%
+  full_join(Value_map_data,
+            by = c("sau_fishing_entity","fishing_entity_id")
+  ) %>% 
+  filter(!is.na(participation))
+
+# These are regions without transboundary species
+grey_land <- World_map %>% 
+  filter(!fishing_entity_id %in% Value_map_data$fishing_entity_id) %>% 
+  filter(!sovereignt %in% c("Antarctica"))
+
+# -------------------------- #
+# Call plot
+# -------------------------- #
+tresh_list <- "tresh_two"
+
+  name_save <- paste("./Figures/emer_map_sens_",tresh_list,".png",sep = "")
+
+  # Value Map
+  World_map_part <- World_map_nemer %>% 
+  filter(
+    tresh ==  tresh_list,
+    variable == "value"
+  )
+
+  # ToE Map 
+  toe_map_data <- World_eez_sf %>% 
+    left_join(Eez_d_hist,
+              by = "eezid") %>% 
+    filter(tresh == tresh_list,
+           metric == "mean_yr")
+
+  # Map with countries without ToE
+
+  russia_churchil <- World_eez_sf %>% 
+    filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))
+
+  no_toe_eez <- World_eez_sf %>% 
+    filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
+    rbind(russia_churchil)
+
+  no_toe_land <- World_map %>% 
+    filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !name %in% World_map_part$name) # Filter EEZs with ToE stocks
+
+  map <- ggplot(toe_map_data) +
+    geom_sf(
+      aes(
+        fill = reverse
+      ),
+      color =  "white",
+      size = 0.1
+    ) +
+    geom_sf(data = World_map_part,
+            aes(fill = participation),
+            color =  "white",
+            size = 0.1
+    ) +
+    geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
+    geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    scale_fill_viridis("",
+                       limits = c(2000,2100),
+                       breaks = seq(2000,2100,25),
+                       labels = c(">2100",seq(2075,2000,-25)) # not inluding the > because we
+    ) +
+    my_ggtheme_m() +
+    facet_wrap(~method,
+               nrow = 2)
+
+  ggsave(name_save,
+         width = 10,
+         height = 6,
+         units = "in"
   )

### 2021-05-24 Scripts/Initial_Analysis.Rmd
sd, not mean.-

@@
     # Alternative method sensitivity analysis
-    group_by(taxon_key,neighbour_id,year) %>% #Average historic data per year first
+    group_by(taxon_key,neighbour_id,year) %>% #s.d. historic data per year first
     # group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
     summarise_at(vars("trans_index"),
-                 c(trend= mean),
+                 c(trend = sd),
                  na.rm=T
@@
 
-    taxon <- unique(Raw_Result$taxon_key)
-    name_raw <- paste(my_path("R",extra_path = "/Emergence_alt_sens_raw/"),taxon,"_emergence_raw.csv",sep="")
-
-    write_csv(Raw_Result,
-              name_raw)
+    # taxon <- unique(Raw_Result$taxon_key)
+    # name_raw <- paste(my_path("R",extra_path = "/Emergence_alt_sens_raw/"),taxon,"_emergence_raw.csv",sep="")
+    # 
+    # write_csv(Raw_Result,
+    #           name_raw)
 
@@
 
-    name <- paste(my_path("R",extra_path = "/Emergence_alt_sens/"),taxon_key,"_emergence.txt",sep="")
+    name <- paste(my_path("R",extra_path = "/Emergence_alt_sens_sd/"),taxon_key,"_emergence.txt",sep="")
 

### 2021-05-24 Scripts/Supplements.Rmd
Inlcuded text for alternative method

@@
 The premise behind time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
-The TI’s natural internal variability was estimated from 1951 to 2005 ($TI_\sigma$) across ensemble members, and the TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$).
+**The TI’s natural internal variability was estimated from the ensemble members by firstly averaging the TI from 1951 to 2005 and then estimating the variation ($TI_\sigma$) across ensemble members.**
+
+The TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$).
 We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig.
@@
 It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution.
+
+**Since the ToE method is sensitive to the noise estimate (i.e., natural variability) [@Reference?], we conducted sensitivity analysis where we first calculated the standard deviation across the 10 ensemble members for each year between 1951 to 2005 and the calculating the average over 1950-2005 [@Reference?].
+Results show that the alternative method estimates 1139 more emerging stocks than the original (5,258 stocks vs 4,119 of the original method), an it also estimates an overall latter ToE (2057 vs 2036 of the original method).
+The earlier and latter ToE remains the same, and the number of EEZs with emerging stocks slightly increases.**
+
 The ToE was also computed for a less conservative estimation using a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).

### 2021-05-25 Scripts/Manuscript.Rmd
Done with Colette's comments

@@
 
-**Climate change is shifting the distribution of shared fish stocks between neighboring countries and the high seas.
-The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.** By coupling a large ensemble simulation of an Earth system model to a species distribution model, we show that 78% the world’s Exclusive Economic Zones will experience at least one shifting transboundary stock by 2030.
-By 2100, 45% of stocks would have shifted out of 81% of countries/political entities waters under a high emissions scenario.
-Importantly, hotspots for transboundary shifts occur for countries that are highly dependent on fisheries for income, food and livelihoods.
+**Climate change is shifting the distribution of shared fish stocks between neighboring countries’ Exclusive Economic Zones (EEZ) and the high seas.
+The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.** Coupling a large ensemble simulation of an Earth system model to a species distribution model, we show that 78% of the world will experience at least one shifting transboundary stock by 2030.
+By 2100, 45% of stocks would have shifted out of 81% of EEZ waters under a high emissions scenario.
+Importantly, many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts.
 The imminent onset of transboundary shifts in fish stocks highlights the urgent need to re-formulate international fisheries policies to limit conflict over these fish stocks.
-**A strong expansion of bi- and multi-lateral actions**
+**A strong expansion of bi- and multi-lateral actions.**
 
@@
 of all species, $\pm$ 28 years).
-Furthermore, 81% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with 78% recording a distributional shift in stocks by the UN 2030 Agenda’s deadline (Fig.
+Furthermore, 81% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 78% recording a distributional shift in stocks by the UN 2030 Agenda’s deadline (Fig.
 \@ref(fig:toeFigure1)C).
@@
 Exclusive Economic Zone polygons display average year of range shifts within them.
-Warm colours are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cool colours represent a late shift/low fishing revenue.
+Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cool colors represent a late shift/low fishing revenue.
 EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
 B) Year of shared distribution shifts by UN sub-regions.
-Points color coded by the average year of range shift.
+Points color coded by the median year of range shift.
 Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda.
@@
 We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs.
-Within five years of the UN 2030 Agenda, 979 transboundary stocks accounting for USD 5.3 billion in fishing revenue are projected to shift within 76 EEZs (Fig.
-\@ref(fig:toeFigure3)); with 4.2 billion USD accounted by 321 stocks shifting in just eleven countries (Fig.
+Within five years of the UN 2030 Agenda timeline, 979 transboundary stocks accounting for USD 5.3 billion in fishing revenue are projected to shift within 76 EEZs (Fig.
+\@ref(fig:toeFigure3); above Fisheries >75th percentile); with 4.2 billion USD accounted by 321 stocks shifting in just eleven countries (Fig.
 \@ref(fig:toeFigure3)).
@@
 We estimated the intensity of the distributional shift of transboundary stocks in terms of changes in catch proportion by 2030 (2020-2040) and 2050 (2040-2060), relative to the recent past (1951-2005) (see Methods).
-**Findings show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks is projected to change beyond an EEZ’s historic proportion (i.e., between 2004 and 2010;** Fig.
+**Projections show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks will change beyond historic variability experinced within an EEZ’s (Fig.
+\@ref(fig:toeFigure4)A).** Moreover, 85% (n = 239) of the world’s EEZs will have experienced changes in catch proportion of transboundary stocks by 2030.
+**By mid-21st century, the number of EEZs with changing stocks, as well as the number of stocks experiencing changes in catch proportion and the magnitude of that change will increase (Figs.
+\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).**The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
+5A).
+Along the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, shifts in stocks are expected to benefit poleward EEZs.
+However, shifts along the coasts of Pacific Central America and West Africa occur in an equatorial direction.
+The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
 \@ref(fig:toeFigure4)A).
-Moreover, 85% (n = 239) of the world's EEZs will have experienced changes in catch proportion of transboundary stocks by 2030.
-**There will be an increase in the number of stocks changing their catch proportion and the number of EEZs with changing stocks towards the mid-21st century.
-Such increase will be followed by an increase in the intensity of the proportion that changes (Figs.
-\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).**
+Along the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the proportion of the stock is expected to follow a projected poleward shift.
+However, shifts in the coasts of Pacific Central America and West Africa occur along an equatorial direction.
 
@@
 
-The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
+<!-- The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
 \@ref(fig:toeFigure4)A).
@@
 \@ref(fig:toeFigure4)B).
+-->
 
@@
 
-**High present-day climate risk of transboundary fisheries management**
+**High present-day climate risk for transboundary fisheries**
 
@@
 These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij].
-**Early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].
-While these represent some of the documented cases of species shifting across jurisdictions, to date is not very clear which and how many transboundary stocks are shifting, where are they shifting and whether they are jointly managed.
-Monitoring present and future shifts in the distribution of transboundary stocks is key to achieve sustainable goals as shifts in distributions are likely to increase the frequency of international disputes and decrease stocks’ health [@Pinsky:2018cb; @Engler:2020bg; @Oremus:2020id].
-This is particularly true as climate change continues to change the distribution of transboundary stocks to a point that some EEZs will see new transboundary stocks [@Oremus:2020id] while others will stand to lose them completely [@Pinsky:2018cb].**
+**Other documented cases of early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].
+Despite these documented cases, broadly speaking, there is a lack of information and data regarding which and how many transboundary stocks are shifting, where they are shifting to and whether they are jointly managed.
+This is particularly important as modeling exercises project that climate change will continue to change the distribution of transboundary stocks to the point that some tropical EEZs will lose them completely [@Oremus:2020id] while other EEZs, mainly in higher latitudes, will stand to win new stocks [@Pinsky:2018cb].**
 
@@
 to manage Pacific salmon (Oncorhynchus sp.) and which has a conservation fund functioning as a side payment [@Miller:2013iv], the International Pacific Halibut Commission between Canada and the United States, which as the name implies manages Pacific halibut (*Hippoglossus stenolepis*) and allocates quota based on the stock’s yearly distribution [@PalaciosAbrantes:2020vt]; and the Parties to the Nauru Agreement (PNA) an Oceania subregional agreement among 8 Pacific Island countries and Tokelau, which collectively manage the largest tuna fishery in the world, controlling around 50% of the global supply of skipjack tuna (*Katsuwonus pelamis*).
-The PNA manage fishing effort under a so-called Vessel Day Scheme, an adaptive and equitable system which accounts for shifts in stocks and catches as a result of climate variability [@Aqorau:2018bh].
+The PNA manage fishing effort under a so-called Vessel Day Scheme, an adaptive and equitable system which accounts for shifts in stocks and catches as a result of climate variability [@Bahri:2021wz].
 Quota allocation methods based on a stock’s current distribution and/or according to fixed-historical proportions like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].**
@@
 Moreover, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery they have historically been entitled to.
-Finally, changes in policy can take years (e.g., it took the PST about 10 years to agree on a conservation fund) and sometimes, the time of political change is longer than that of species shifts [@Pinsky:2012kq].
-This is specifically true we could be missing shifting stocks as we lack standardized data to track shifts across international jurisdictions [@Maureaud:2021gj].
-Moreover, recent efforts looking at international fisheries plans found that all 127 plans lacked direct actions to address topics of climate change or species range shifts across jurisdictions.
-Furthermore, authors were rarely able to identify the stock in question, and when they did, it was as broad as skipjack tuna or billfishes [@Oremus:2020id].
-Having a better understanding of what stocks are managed as transboundary and under what rules would be an important complement to our study and an important step towards the sustainability of international stocks.**
+Finally, changes in policy can be sluggish (e.g., it took the PST about 10 years to agree on a conservation fund) compared to the timeframe over which species shift [@Pinsky:2012kq].
+This is specifically true as in many cases we lack standardized data to track shifts of species across international jurisdictions [@Maureaud:2021gj] as these might already be happening.
+Addressing shifting transboundary stocks is also challenged by an overall lack of knowledge of how many international agreements exist or what transboundary stocks are jointly managed.
+Recent efforts looking at 127 international fisheries plans found that they all lacked direct actions to address topics of climate change or species range shifts across jurisdictions [@Oremus:2020id].
+Having a better understanding of what stocks are managed as transboundary and under what rules would be an important complement to our study and an important step towards the sustainability of international stocks.
+[@Oremus:2020id].**
 
@@
 Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and *ii*) model structure [@Frolicher:2016dg].
-First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.**Thus, any mitigation efforts could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].** Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
+First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.**Thus, any mitigation efforts could result in a substantial delay in distribution shifts as environmental signals are responsive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].** Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
 This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is often larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].

### 2021-05-25 Scripts/Initial_Analysis.Rmd
Changes for Colette

@@
 
-This results only shows the average percentage change in proportion of all speecies within an EEZ without relationship with neighbours
+This results only shows the average percentage change in proportion of all speecies within an EEZ without relationship with neighbors
 
@@
 
-This function estimates the total change in MCP proportion of each species for each EEZ. It looses the connection with neighbours as you only know for x species houw much is the proportion changing.
+This function estimates the total change in MCP proportion of each species for each EEZ. It looses the connection with neighbors as you only know for x species how much is the proportion changing.
 
@@
 # Read proportion results
+suppressMessages(
 Prop_results <- PropChange(Taxon_list, links="y") %>%
@@
   )
-
+)
 # For dat availability
@@
 dashed_lines <- tibble(
-  x = c(2030,2050,2070),
-  label = c("UN 2030 Agenda","Mid range shift   ","Late range shift   ")
+  x = c(2030
+        # ,2050,2070),
+  ),
+  label = c("UN 2030 Agenda"
+            # ,"Mid range shift   ","Late range shift   "
+            )
 )
@@
 
-dashed_lines <- tibble(
-  x = c(2030,2050,2070),
-  label = c("UN 2030 Agenda","Mid range shift  ","Late range shift   ")
-)
 # Needs previous chunk
@@
 Labels_data <- tibble(
-  x = c(2030,2050,2070,2007),
-  y = c(-5.8,-5.8,-5.8,Percentile_value$x+.1),
-  label = c("UN 2030 Agenda", "Mid shift","Late shift","Fisheries > 75th percentile")
+  x = c(2029,
+        # 2050,2070,
+        2007),
+  y = c(-5.2,
+        # -5.8,-5.8,
+        Percentile_value$x+.1),
+  label = c("UN 2030 Agenda", 
+            # "Mid shift","Late shift",
+            "Fisheries > 75th percentile")
 )
@@
   # Divisory lines and labels
-  geom_vline(xintercept = c(2030,2050,2070), color = "grey50", size = 0.5, linetype = "dashed") +
-  geom_hline(data = Percentile_value,
-             aes(yintercept = x),
-             color = "grey50"
-  )  +
-  geom_text(data = Labels_data,
+  geom_vline(xintercept = 2030, color = "grey50", size = 0.5, linetype = "dashed") +
+  geom_text(data = subset(Labels_data, x == 2029),
             aes(
@@
             ),
-            size = 5,
+            size = 4,
+            color = "grey30",
+            angle = 90
+  ) +
+  geom_hline(data = Percentile_value,
+             aes(yintercept = x),
+             color = "grey50"
+  )  +
+  geom_text(data = subset(Labels_data, x == 2007),
+            aes(
+              x = x,
+              y = y,
+              label = label
+            ),
+            size = 4,
             color = "grey30"

### 2021-05-26 Scripts/Initial_Analysis.Rmd
Cleaned code to represent paper estimates

--- 2021-05-22 Scripts/Initial_Analysis.Rmd
@@
                             align = "right", 
-                            fill = trans_index,
+                            fill = NA,
                             na.rm=T)
     ) %>% 
-    # filter(year > 1960) %>%
+    # filter(year > 1960) %>% # drop first 10 years of running mean
     select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)
@@
     # Valid ensembles
-  # Only perform analysis on ensembles containing all years between 1951 and 2040
+  # Only perform analysis on ensembles containing all years between 1961 and 2040
     n_years <- mean_sd_index %>% 
@@
     # Alternative method sensitivity analysis
-    group_by(taxon_key,neighbour_id,year) %>% #Average historic data per year first
-    # group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
-    summarise_at(vars("trans_index"),
-                 c(trend= mean),
+    # group_by(taxon_key,neighbour_id,year) %>% #s.d. historic data per year first
+    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
+    summarise_at(vars("RMean"),
+                 c(trend = mean),
                  na.rm=T
@@
 
-    Raw_Result <- mean_sd_index %>%
+Final_Result <- mean_sd_index %>%
   filter(year > Hist_y) %>%
@@
   ) %>%
-      gather("tresh","emerged",tresh_one,tresh_two)
-
-    taxon <- unique(Raw_Result$taxon_key)
-    name_raw <- paste(my_path("R",extra_path = "/Emergence_alt_sens_raw/"),taxon,"_emergence_raw.csv",sep="")
-
-    write_csv(Raw_Result,
-              name_raw)
-
-    ToE_Result <- Raw_Result %>% 
-    # Method the last year when the index overshoots and never goes back
-      filter(is.na(emerged)) %>% 
-      group_by(taxon_key, neighbour_id,n_ens,tresh) %>% 
-      summarize(emerging_yr = max(year)+1) %>% # last year of NA + 1 makes it the first year of overshooting
+  gather("tresh","YesNo",tresh_one,tresh_two) %>% 
+  filter(!is.na(YesNo)) %>% 
+  group_by(taxon_key,neighbour_id,n_ens,tresh) %>%
+  summarise(emerging_yr= min(year)) %>%
+  # summarise(emerging_yr= max(year)) %>% # Last year that treshold is crossed
   left_join(EEZ_ids,
@@
 
-  return(ToE_Result)
+  return(Final_Result)
   }
@@
 
-    name <- paste(my_path("R",extra_path = "/Emergence_alt_sens/"),taxon_key,"_emergence.txt",sep="")
+    name <- paste(my_path("R",extra_path = "/Emergence_95/"),taxon_key,"_emergence.txt",sep="")
 

### 2021-05-26 Scripts/Initial_Analysis.Rmd

--- 2021-05-25 Scripts/Initial_Analysis.Rmd
@@
     # Alternative method sensitivity analysis
-    group_by(taxon_key,neighbour_id,year) %>% #s.d. historic data per year first
-    # group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
-    summarise_at(vars("trans_index"),
-                 c(trend = sd),
+    # group_by(taxon_key,neighbour_id,year) %>% #s.d. historic data per year first
+    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
+    summarise_at(vars("RMean"),
+                 c(trend = mean),
                  na.rm=T
@@
   ) %>%
-      gather("tresh","emerged",tresh_one,tresh_two)
-
-    # taxon <- unique(Raw_Result$taxon_key)
-    # name_raw <- paste(my_path("R",extra_path = "/Emergence_alt_sens_raw/"),taxon,"_emergence_raw.csv",sep="")
-    # 
-    # write_csv(Raw_Result,
-    #           name_raw)
-
-    ToE_Result <- Raw_Result %>% 
-    # Method the last year when the index overshoots and never goes back
-      filter(is.na(emerged)) %>% 
-      group_by(taxon_key, neighbour_id,n_ens,tresh) %>% 
-      summarize(emerging_yr = max(year)+1) %>% # last year of NA + 1 makes it the first year of overshooting
+  gather("tresh","YesNo",tresh_one,tresh_two) %>% 
+  filter(!is.na(YesNo)) %>% 
+  group_by(taxon_key,neighbour_id,n_ens,tresh) %>%
+  summarise(emerging_yr= min(year)) %>%
+  # summarise(emerging_yr= max(year)) %>% # Last year that treshold is crossed
   left_join(EEZ_ids,
@@
 
-    name <- paste(my_path("R",extra_path = "/Emergence_alt_sens_sd/"),taxon_key,"_emergence.txt",sep="")
+    name <- paste(my_path("R",extra_path = "/Emergence_95/"),taxon_key,"_emergence.txt",sep="")
 

### 2021-05-26 Scripts/Supplements.Rmd
Updated methods to relfect sensitivity analysis of natural variabillity

--- 2021-05-17 Scripts/Supplements.Rmd
@@
 where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (th, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+The TI was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg] and to match transboundary fisheries management that often integrates a longer time-frame in their consideration.
 
@@
 
-The index was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg].
-
 _**Calculating shifts in the shared distribution of transboundary stocks**_
@@
 The premise behind time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
-The TI’s natural internal variability was estimated from 1951 to 2005 ($TI_\sigma$) across ensemble members, and the TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$).
+The TI’s natural internal variability was estimated from the ensemble members by firstly averaging the TI from 1951 to 2005 and then estimating the variation ($TI_\sigma$) across ensemble members.
+The TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$).
 We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig.
@@
 It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution.
+
+Since the ToE method is sensitive to the noise estimate (i.e., historical natural variability) [@Reference?], we conducted sensitivity analysis where we first calculated the average across the 10 ensemble members for each year between 1951 to 2005 and the calculating the standard deviation over 1950-2005 [@Reference?].
+Under this method, the natural variability is slightly larger resulting in a latter overall time of emergence (average ToE = 2057 vs 2036 of the original method).
+However, this method projects 1,139 more emerging stocks than the original (5,258 stocks vs 4,119 of the original method).
+The minimum and maximum ToE remains the same, and the number of EEZs with emerging stocks slightly increases with the alternative method.
+Further research looking at different approaches of estimating the natural variability component could help reduce the uncertainty in the estimation of the time of emergence of transboundary stocks.
 The ToE was also computed for a less conservative estimation using a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
@@
 In both cases, our null hypothesis was that there were no significant differences in the year of shift across habitat association or EEZs.
-All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages data.table [@Packagedatatable:2019uh], janitor [@PackagejanitorSim:2018ur], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], gmt [@PackagegmtInterfa:2017vh] and zoo [@PackagezooSInfr:2019tm], ggrepel [@PackageggrepelAut:2020to], zeallot [@PackagezeallotMul:2018wy], viridis [@PackageviridisDef:2018us], cowplot [@CowplotStreamlined:2019wt], moments [@PackagemomensMome:2015uh] and pgirmess [@PackagepgirmessSp:2018um].
+All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages cowplot [@CowplotStreamlined:2019wt], data.table [@Packagedatatable:2019uh], geosphere [@], ggrepel [@PackageggrepelAut:2020to], gmt [@PackagegmtInterfa:2017vh], janitor [@PackagejanitorSim:2018ur], moments [@PackagemomensMome:2015uh], pgirmess [@PackagepgirmessSp:2018um], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], viridis [@PackageviridisDef:2018us], zeallot [@PackagezeallotMul:2018wy], and zoo [@PackagezooSInfr:2019tm].
 Code for all analyses is available at https://github.com/jepa/TransEmergence

### 2021-05-26 Scripts/Supplements.Rmd
Changed title

@@
 ---
-title: "Supplementary Materials for Imminent range shifts in transboundary fish stocks challenge their management under climate change"
+title: "Timing and magnitude of climate driven range shifts in transboundary fish stocks challenge their management"
 author: 'Juliano Palacios-Abrantes, Thomas L. Frölicher, Gabriel Reygondeau, U. Rashid Sumaila, Alessandro Tagliabue, Colette C.C. Wabnitz, and William W.L. Cheung'

### 2021-05-27 Scripts/Manuscript.Rmd
Changed title
Changed my affiliation and adressed final details before re-submission

@@
 ---
-title: "Imminent range shifts in transboundary fish stocks challenge their management under climate change"
-author: "SUBMITTED MANUSCRIPT. DO NOT SHARE."
+title: "Timing and magnitude of climate driven range shifts in transboundary fish stocks challenge their management"
+author: "REVISED MANUSCRIPT. DO NOT SHARE."
 date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
@@
 ^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States"
-subtitle: "Juliano Palacios-Abrantes*^1^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette C.C. Wabnitz^1,5^, and William W.L. Cheung^1^"
+subtitle: "Juliano Palacios-Abrantes*^1,2^, Thomas L. Frölicher^3,4^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^5^, Colette C.C. Wabnitz^1,6^, and William W.L. Cheung^1^"
 # output: bookdown::pdf_document2
@@
 **Climate change is shifting the distribution of shared fish stocks between neighboring countries’ Exclusive Economic Zones (EEZ) and the high seas.
-The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.** Coupling a large ensemble simulation of an Earth system model to a species distribution model, we show that 78% of the world will experience at least one shifting transboundary stock by 2030.
+The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.** Coupling a large ensemble simulation of an Earth system model to a species distribution model, we show that by 2030 23% of transboundary stocks would have shifted and 78% of the world will experience at least one shifting stock.
 By 2100, 45% of stocks would have shifted out of 81% of EEZ waters under a high emissions scenario.
 Importantly, many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts.
-The imminent onset of transboundary shifts in fish stocks highlights the urgent need to re-formulate international fisheries policies to limit conflict over these fish stocks.
-**A strong expansion of bi- and multi-lateral actions.**
+**The imminent onset of transboundary shifts in fish stocks highlights the urgent need to create international fisheries policies to limit conflict over these fish stocks and re-formulate those existing.**
 
@@
 ```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "**Timeline of shifts in transboundary shared stock distribution.** (A) Number of stocks (cumulative) whose distribution centroids that have shifted beyond natural variations over time (from 2006 to >2100), (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one stock shifting beyond natural variability.
-Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change in range shift (see Methods), respectively."}
+Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change in range shift (Methods), respectively."}
 
@@
 
-On average, shifting transboundary stocks represented 23% (s.d.
-$\pm$  22%) of annual revenue from fisheries targeting **transboundary** stocks across the world's EEZs between 2004 and 2010 (Fig.
-\@ref(fig:toeFigure2)A).
-**From all fisheries targeting transboundary stocks between 2004 and 2010 in the global EEZs, shifting stocks represented on average 23% (s.d.
-$\pm$  22%) of the yearly revenue (Fig.
+**On average, shifting transboundary stocks represented 23% (s.d.
+$\pm$  22%) of annual revenue from fisheries targeting transboundary stocks across the world's EEZs between 2004 and 2010 (Fig.
 \@ref(fig:toeFigure2)A).** However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
@@
 We estimated the intensity of the distributional shift of transboundary stocks in terms of changes in catch proportion by 2030 (2020-2040) and 2050 (2040-2060), relative to the recent past (1951-2005) (see Methods).
-**Projections show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks will change beyond historic variability experinced within an EEZ’s (Fig.
+**Projections show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks will change beyond the historic variability experienced within an EEZ’s (Fig.
 \@ref(fig:toeFigure4)A).** Moreover, 85% (n = 239) of the world’s EEZs will have experienced changes in catch proportion of transboundary stocks by 2030.
@@
 \@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).**The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
-5A).
+\@ref(fig:toeFigure4)A).
 Along the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, shifts in stocks are expected to benefit poleward EEZs.
 However, shifts along the coasts of Pacific Central America and West Africa occur in an equatorial direction.
-The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
-\@ref(fig:toeFigure4)A).
-Along the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the proportion of the stock is expected to follow a projected poleward shift.
-However, shifts in the coasts of Pacific Central America and West Africa occur along an equatorial direction.
 
@@
 Despite these documented cases, broadly speaking, there is a lack of information and data regarding which and how many transboundary stocks are shifting, where they are shifting to and whether they are jointly managed.
-This is particularly important as modeling exercises project that climate change will continue to change the distribution of transboundary stocks to the point that some tropical EEZs will lose them completely [@Oremus:2020id] while other EEZs, mainly in higher latitudes, will stand to win new stocks [@Pinsky:2018cb].**
+This is particularly important for international fisheries management as modeling exercises project that climate change will continue to change the distribution of transboundary stocks to the point that some tropical EEZs will lose them completely [@Oremus:2020id] while other EEZs, mainly in higher latitudes, will stand to win new stocks [@Pinsky:2018cb].**
 
@@
 
-Our results identify regional “hotspots” of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figures \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure4)A).
+Our results identify regional “hotspots” of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figs.
+\@ref(fig:toeFigure2)A and \@ref(fig:toeFigure4)A).
 Regions such as the Caribbean are characterized by high levels of warming relative to historical natural variability [@IPCC:2019tn; @Hawkins:2012hc] and species’ high vulnerability to warming waters [@IPCC:2019tn].

### 2021-05-27 Scripts/Supplements.Rmd
Inlcuded sensitivity analysis text

@@
 
-Since the ToE method is sensitive to the noise estimate (i.e., historical natural variability) [@Reference?], we conducted sensitivity analysis where we first calculated the average across the 10 ensemble members for each year between 1951 to 2005 and the calculating the standard deviation over 1950-2005 [@Reference?].
+Since the ToE method is sensitive to the noise estimate (i.e., historical natural variability), we conducted sensitivity analysis where we first calculated the average across the 10 ensemble members for each year between 1951 to 2005 and the calculating the standard deviation over 1950-2005.
 Under this method, the natural variability is slightly larger resulting in a latter overall time of emergence (average ToE = 2057 vs 2036 of the original method).

### 2021-07-19 Scripts/Initial_Analysis.Rmd
Sci adv revissions (#4)
* Started reviewing process

* Easy Responses to rev 1

* Included legend of figures 4 and S4 that were missing

* Continue addressing reviewer 1

* Addressed all but one comment from Rev. 1

* Marked changes for Rev 1

* Responses to reviewer one

* fixed point in fig4

* Cleaned script

* Updated figure order including S2 as 1

* included most of rev. 3 changes

* All responses done

* Included a summary of the methods in the main article

* Moved some of the methods to the main article and icnluded a more detailed version of the DBEM

* Almost done!

* misc.

* Inclu

@@
 
-# How many between 2006 and 2020?
+# How many between 2006 and 2030?
 plot_data %>% 
-  filter(taxon_emer <= 2020,
+  filter(taxon_emer <= 2030,
          tresh == selected_tresh) %>%  # Presenting results for the second treshold
@@
              y = -69,
-             label = "% Of fishing revenue",
+             label = "Transboundary fishing revenue (%)",
              color="black",
@@
              y = -85,
-             label = "Year of range shift",
+             label = "Average year of range shift",
              color="black", 

### 2021-07-19 Scripts/Manuscript.Rmd
updates to pnas

@@
 bibliography: ../References/ToE_Ref.bib
-csl: ../References/science.csl
+csl: ../References/pnas.csl
 header-includes:
@@
 
-**Abstract**
+# Abstract
 
@@
 
-**Climate change is shifting the distribution of shared fish stocks between neighboring countries’ Exclusive Economic Zones (EEZ) and the high seas.
-The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.** Coupling a large ensemble simulation of an Earth system model to a species distribution model, we show that by 2030 23% of transboundary stocks would have shifted and 78% of the world will experience at least one shifting stock.
-By 2100, 45% of stocks would have shifted out of 81% of EEZ waters under a high emissions scenario.
-Importantly, many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts.
-**The imminent onset of transboundary shifts in fish stocks highlights the urgent need to create international fisheries policies to limit conflict over these fish stocks and re-formulate those existing.**
-
-<!-- **One-Sentence Summary** -->
-
-<!-- Rapid climate-induced shifts in transboundary fish stocks challenges international fisheries management -->
+Climate change is shifting the distribution of shared fish stocks between neighboring countries’ Exclusive Economic Zones (EEZ) and the high seas.
+The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.
+Coupling a large ensemble simulation of an Earth system model to a species distribution model, we show that by 2030 23% of transboundary stocks would have shifted and 78% of the world will experience at least one shifting stock.
+**By 2100, 81% of EEZ waters will see at least one shifting stock with a total of 45% of stocks shifting globally, under a high emissions scenario.** Importantly, many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts.
+**Existing fisheries agreements need to be assessed for their capacity in addressing transboundary shifts, and strengthened where necessary to limit conflict over these fish stocks.**
 
@@
 
-**Main Text.**
+# Introduction
 
@@
 Marine species’ distributions reflect species’ preferences for discrete environmental conditions [@Hutchinson:1957nt].
-As a result of changing ocean conditions, many marine species **are shifting** their distributions towards higher latitudes, deeper water or following local temperature gradients to remain within their optimal environmental “niche” [@Poloczanska:2016kk].
+As a result of changing ocean conditions, many marine species are shifting their distributions towards higher latitudes, deeper water or following local temperature gradients to remain within their optimal environmental “niche” [@Poloczanska:2016kk].
 The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt], impacting fish stocks, fisheries production and economies [@Sumaila:2019ew].
@@
 
-Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (see Methods **and Supplemental Materials**).
+Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (see Methods and Supplemental Materials).
 We then estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig.
-\@ref(fig:toeFigure1N)) and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods **and Supplemental Materials**).
+\@ref(fig:toeFigure1N)) and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods and Supplemental Materials).
 We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
@@
 
-**Identifying range shifts in the shared distribution of transboundary stocks**
+<!-- # Methods -->
 
-**Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
-\@ref(fig:toeFigure1)A), with the first shift modelled to have occurred back in 2006.** This corresponds to 45% of the studied stocks (Fig.
+<!-- We estimate the year in which 663 transboundary species (i.e., 9,132 stocks) previously identified by Palacios-Abrantes *et al.* (2020) [@PalaciosAbrantes:2020tv] would undergo changes in their shared distribution beyond historical variation and the intensity of said changes (see the Supplementary Materials).
+We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
+We defined the boundaries of the world’s EEZs using the Sea Around Us (http://www.seaaroundus.org/) spatial division, noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
+\@ref(fig:toeFigureS5)) including island territories (see the Supplementary Materials - *Databases and species selection*).
+-->
+
+<!-- We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks under climate change (see the Supplementary Materials).
+In brief, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
+The DBEM was driven with ten different simulations of the GFDL ESM2M (i.e., a ten-ensemble member simulation) each of them representing a unique atmosphere and ocean state at the start of the simulation [@Dunne:2012bo; @Dunne:2013db; @Frolicher:2020cc].
+The simulations were run under historical forcing until 2005, afterwards and until 2100, simulations followed a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5) [@Riahi:2011dk].
+-->
+
+<!-- We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
+The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.
+\@ref(fig:toeFigure1)A).
+We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution (see the Supplementary Materials - *Calculating shifts in the shared distribution of transboundary stocks*).
+The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ.
+\@ref(eq:four)).
+We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
+-->
+
+<!-- Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (See Supplemental Material - *Quantifying the intensity of transboundary stocks’ range shift*).
+A threat point is defined as the minimum payoff a player would be willing to receive to cooperate in a game-theoretic model.
+Here, we define the minimum payoff as two standard deviations from the historical catch proportion (average of 1951 to 2005) that each EEZ had for each stock.
+This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Finally, we computed changes in catch proportion that were below the threat point of neighboring EEZs sharing a stock for two time periods relative to the historical proportion: 2030 (2021 and 2040) matching the UN 2030 Agenda and a mid-century timeline (2040 to 2050).
+-->
+
+# Results
+
+### Identifying range shifts in the shared distribution of transboundary stocks
+
+Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
+\@ref(fig:toeFigure1)A), with the first shift modelled to have occurred back in 2006.
+This corresponds to 45% of the studied stocks (Fig.
 \@ref(fig:toeFigure1)C).
@@
 
-**On average, shifting transboundary stocks represented 23% (s.d.
+On average, shifting transboundary stocks represented 23% (s.d.
 $\pm$  22%) of annual revenue from fisheries targeting transboundary stocks across the world's EEZs between 2004 and 2010 (Fig.
-\@ref(fig:toeFigure2)A).** However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
+\@ref(fig:toeFigure2)A).
+However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
 Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).
@@
 
-**Shifting intensity of transboundary stocks**
+### Shifting intensity of transboundary stocks
 
@@
 We estimated the intensity of the distributional shift of transboundary stocks in terms of changes in catch proportion by 2030 (2020-2040) and 2050 (2040-2060), relative to the recent past (1951-2005) (see Methods).
-**Projections show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks will change beyond the historic variability experienced within an EEZ’s (Fig.
-\@ref(fig:toeFigure4)A).** Moreover, 85% (n = 239) of the world’s EEZs will have experienced changes in catch proportion of transboundary stocks by 2030.
-**By mid-21st century, the number of EEZs with changing stocks, as well as the number of stocks experiencing changes in catch proportion and the magnitude of that change will increase (Figs.
-\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).**The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
+Projections show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks will change beyond the historic variability experienced within an EEZ’s (Fig.
+\@ref(fig:toeFigure4)A).
+Moreover, 85% (n = 239) of the world’s EEZs will have experienced changes in catch proportion of transboundary stocks by 2030.
+By mid-21st century, the number of EEZs with changing stocks, as well as the number of stocks experiencing changes in catch proportion and the magnitude of that change will increase (Figs.
+\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).
+The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
 \@ref(fig:toeFigure4)A).
@@
 
-<!-- ## Discussion -->
+# Discussion
 
@@
 
-**High present-day climate risk for transboundary fisheries**
+### High present-day climate risk for transboundary fisheries
 
@@
 These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij].
-**Other documented cases of early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].
+Other documented cases of early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].
 Despite these documented cases, broadly speaking, there is a lack of information and data regarding which and how many transboundary stocks are shifting, where they are shifting to and whether they are jointly managed.
-This is particularly important for international fisheries management as modeling exercises project that climate change will continue to change the distribution of transboundary stocks to the point that some tropical EEZs will lose them completely [@Oremus:2020id] while other EEZs, mainly in higher latitudes, will stand to win new stocks [@Pinsky:2018cb].**
+This is particularly important for international fisheries management as modeling exercises project that climate change will continue to change the distribution of transboundary stocks to the point that some tropical EEZs will lose them completely [@Oremus:2020id] while other EEZs, mainly in higher latitudes, will stand to win new stocks [@Pinsky:2018cb].
 
-**Climate risk hotspots for transboundary fisheries management**
+###mClimate risk hotspots for transboundary fisheries management
 
@@
 
-**Most management plans that are not designed or prepared to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
+Most management plans that are not designed or prepared to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
 Focusing on identified “hotspots” will help anticipate any potential increases in and averting fisheries conflicts in coming years.
@@
 The PNA manage fishing effort under a so-called Vessel Day Scheme, an adaptive and equitable system which accounts for shifts in stocks and catches as a result of climate variability [@Bahri:2021wz].
-Quota allocation methods based on a stock’s current distribution and/or according to fixed-historical proportions like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].**
+Quota allocation methods based on a stock’s current distribution and/or according to fixed-historical proportions like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
 
-**Even management plans that take into consideration such strategies might not be fully prepared for the consequences of shifting transboundary stocks [@Pinsky:2018cb; @Koubrak:2020hw; @Engler:2020bg].
+Even management plans that take into consideration such strategies might not be fully prepared for the consequences of shifting transboundary stocks [@Pinsky:2018cb; @Koubrak:2020hw; @Engler:2020bg].
 For example, species’ shifts within the PNA area will likely also have to deal with stocks expanding to new jurisdictions, issue that the NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
@@
 Having a better understanding of what stocks are managed as transboundary and under what rules would be an important complement to our study and an important step towards the sustainability of international stocks.
-[@Oremus:2020id].**
+[@Oremus:2020id].
 
-**Potential drivers of shift and managing an uncertain future**
+### Potential drivers of shift and managing an uncertain future
 
@@
 Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and *ii*) model structure [@Frolicher:2016dg].
-First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.**Thus, any mitigation efforts could result in a substantial delay in distribution shifts as environmental signals are responsive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].** Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
+First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.Thus, any mitigation efforts could result in a substantial delay in distribution shifts as environmental signals are responsive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].
+Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
 This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is often larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
@@
 
-**Conclusion**
+# Conclusion
 
@@
 
-**Methods **
-
-We estimate the year in which 663 transboundary species (i.e., 9,132 stocks) previously identified by Palacios-Abrantes *et al.* (2020) [@PalaciosAbrantes:2020tv] would undergo changes in their shared distribution beyond historical variation and the intensity of said changes (see the Supplementary Materials).
-We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
-We defined the boundaries of the world’s EEZs using the Sea Around Us (http://www.seaaroundus.org/) spatial division, noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
-\@ref(fig:toeFigureS5)) including island territories (see the Supplementary Materials - *Databases and species selection*).
-
-We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks under climate change (see the Supplementary Materials).
-In brief, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
-The DBEM was driven with ten different simulations of the GFDL ESM2M (i.e., a ten-ensemble member simulation) each of them representing a unique atmosphere and ocean state at the start of the simulation [@Dunne:2012bo; @Dunne:2013db; @Frolicher:2020cc].
-The simulations were run under historical forcing until 2005, afterwards and until 2100, simulations followed a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5) [@Riahi:2011dk].
-
-We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
-The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.
-\@ref(fig:toeFigure1)A).
-We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution (see the Supplementary Materials - *Calculating shifts in the shared distribution of transboundary stocks*).
-The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ.
-\@ref(eq:four)).
-We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
-
-Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (See Supplemental Material - *Quantifying the intensity of transboundary stocks’ range shift*).
-A threat point is defined as the minimum payoff a player would be willing to receive to cooperate in a game-theoretic model.
-Here, we define the minimum payoff as two standard deviations from the historical catch proportion (average of 1951 to 2005) that each EEZ had for each stock.
-This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-Finally, we computed changes in catch proportion that were below the threat point of neighboring EEZs sharing a stock for two time periods relative to the historical proportion: 2030 (2021 and 2040) matching the UN 2030 Agenda and a mid-century timeline (2040 to 2050).
-
 \clearpage
 
-**References**
+# References
 
@@
 
-**Supplements**
+# Supplements
 

### 2021-07-26 Scripts/Manuscript.Rmd
duplicated paragraph removed

@@
 
-Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (see Methods and Supplemental Materials).
+Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods).
 We then estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig.
-\@ref(fig:toeFigure1N)) and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods and Supplemental Materials).
+\@ref(fig:toeFigure1N)) and the intensity of this shift (i.e., time of emergence and stock share ratio; Methods).
 We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
@@
 
-<!-- We estimate the year in which 663 transboundary species (i.e., 9,132 stocks) previously identified by Palacios-Abrantes *et al.* (2020) [@PalaciosAbrantes:2020tv] would undergo changes in their shared distribution beyond historical variation and the intensity of said changes (see the Supplementary Materials).
-We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
-We defined the boundaries of the world’s EEZs using the Sea Around Us (http://www.seaaroundus.org/) spatial division, noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
-\@ref(fig:toeFigureS5)) including island territories (see the Supplementary Materials - *Databases and species selection*).
--->
-
-<!-- We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks under climate change (see the Supplementary Materials).
+We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks under climate change (see the Supplementary Materials).
 In brief, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
@@
 The simulations were run under historical forcing until 2005, afterwards and until 2100, simulations followed a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5) [@Riahi:2011dk].
--->
 
-<!-- We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
+We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
 The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.
@@
 We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
--->
 
-<!-- Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (See Supplemental Material - *Quantifying the intensity of transboundary stocks’ range shift*).
+Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (See Supplemental Material - *Quantifying the intensity of transboundary stocks’ range shift*).
 A threat point is defined as the minimum payoff a player would be willing to receive to cooperate in a game-theoretic model.
@@
 Finally, we computed changes in catch proportion that were below the threat point of neighboring EEZs sharing a stock for two time periods relative to the historical proportion: 2030 (2021 and 2040) matching the UN 2030 Agenda and a mid-century timeline (2040 to 2050).
--->
 
@@
 
-###mClimate risk hotspots for transboundary fisheries management
+### Climate risk hotspots for transboundary fisheries management
 

### 2021-07-26 Scripts/Initial_Analysis.Rmd
No need for arrive/excit analysis

@@
 }
-
-```
-
-#### Fun `EEZarrive_exit`
-
-```{r fun_EEZexit, eval = F, echo = F}
-
-# taxon_list <- c(600004,600006)
-
-# EEZProp <- function(taxon_list,links="NA"){
-
-  spp <- bind_rows(
-    lapply(Taxon_list,SummaryProp,links)
-  ) %>% 
-    spread(time_step,mean_spp) %>% 
-    # mutate_if(is.numeric,replace_na,0) %>% # If NA 0 meaning wins and looses
-    mutate_at(vars(
-      "per_early" = early,
-      "per_mid" = mid
-    ),
-    funs(round(((.-historical)/(abs(historical)))*100,2))
-    ) %>%  # original
-  gather("time_step","percentage_change",per_early:per_mid)
 

### 2021-07-26 Scripts/Manuscript.Rmd
last paragraph changed

@@
 
-We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks under climate change (see the Supplementary Materials).
+We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks under climate change (Methods).
 In brief, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
@@
 \@ref(fig:toeFigure1)A).
-We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution (see the Supplementary Materials - *Calculating shifts in the shared distribution of transboundary stocks*).
-The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ.
-\@ref(eq:four)).
+We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution.
+The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence".
 We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
 
-Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (See Supplemental Material - *Quantifying the intensity of transboundary stocks’ range shift*).
+Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (Methods).
 A threat point is defined as the minimum payoff a player would be willing to receive to cooperate in a game-theoretic model.
@@
 
-Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
-\@ref(fig:toeFigure1)A), with the first shift modelled to have occurred back in 2006.
+**Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
+\@ref(fig:toeFigure1)A), using a two s.d.
+threshold (i.e., representing a probability of 95% that the stock has shifted).
 This corresponds to 45% of the studied stocks (Fig.
 \@ref(fig:toeFigure1)C).
-The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be 2036 (s.d.
-of all species, $\pm$ 28 years).
-Furthermore, 81% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 78% recording a distributional shift in stocks by the UN 2030 Agenda’s deadline (Fig.
-\@ref(fig:toeFigure1)C).
+Using a less conservative threshold (i.e., 1 s.d.
+representing a probability of 68% that the stock has shifted) results in 33% (n = 5,745) more shifting stocks, that is 63% of all studied stocks (Figs.
+\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
+For both cases, the first shift is modelled to have occurred back in 2006.
+The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be between 2029 $\pm$ 27 years (1 s.d.
+threshold) and 2036 $\pm$ 28 years (2 s.d.
+threshold).
+Furthermore, between 81% and 83% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 77% to 80% recording a distributional shift in stocks by the UN 2030 Agenda’s deadline (Fig.
+\@ref(fig:toeFigure1)C).**
 
@@
 
-The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig.
-\@ref(fig:toeFigureS1)) and EEZs' geographic regions (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Table S3 and Fig.
+The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; using the 2 s.d.
+threshold; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig.
+\@ref(fig:toeFigureS1)) and EEZs' geographic regions (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; using the 2 s.d.
+threshold; Table S3 and Fig.
 2A).
@@
 
-On average, shifting transboundary stocks represented 23% (s.d.
-$\pm$  22%) of annual revenue from fisheries targeting transboundary stocks across the world's EEZs between 2004 and 2010 (Fig.
+**On average, shifting transboundary stocks represented 27% to 23% (1 and 2 s.d.
+thresholds) of annual revenue from fisheries targeting transboundary stocks across the world's EEZs between 2004 and 2010 (Fig.
 \@ref(fig:toeFigure2)A).
 However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
-Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).
+Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).**
 
@@
 
-We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs.
-Within five years of the UN 2030 Agenda timeline, 979 transboundary stocks accounting for USD 5.3 billion in fishing revenue are projected to shift within 76 EEZs (Fig.
-\@ref(fig:toeFigure3); above Fisheries >75th percentile); with 4.2 billion USD accounted by 321 stocks shifting in just eleven countries (Fig.
+We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs for both thresholds (one and two s.d.).
+Within five years of the UN 2030 Agenda timeline, 979 transboundary stocks accounting for USD 5.3 billion in fishing revenue are projected to shift within 76 EEZs considering a two s.d.
+threshold (Fig.
+\@ref(fig:toeFigure3); above Fisheries >75th percentile).
+In this case, 321 stocks shifting in just eleven countries account for 4.2 billion USD (Fig.
 \@ref(fig:toeFigure3)).
@@
 \@ref(fig:toeFigure3)).
+The values are substantially higher when considering a more relaxed threshold.
+In this case, 2,136 stocks accounting for USD 24 billion in fishing revenue are projected to shift within five years of the UN 2030 Agenda timeline in 134 EEZs.
 

### 2021-07-26 Scripts/Supplements.Rmd
updates to pnas
changed methods to reflect thresholds

@@
 The TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$).
-We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig.
+We then set a threshold (*tresh*) of $TI_\sigma$ to define the time of emergence of transboundary stocks (See Fig.
 \@ref(fig:toeFigure1) for a graphical description):
+
+<!-- We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig.
+\@ref(fig:toeFigure1) for a graphical description): -->
 
@@
 \label{eq:seven}
-ToE = TI_\mu[t']>2* TI_\sigma 
+ToE = TI_\mu[t']> tresh
 \end{equation}
 
-Thus, we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TI “emerges” from historical internal variability (*t’*).
+**This way we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TI “emerges” from historical internal variability (*tresh*).
 It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution.
+Two different arbitrary values of *tresh* were used.
+A conservative value ($tresh = 2* TI_\sigma$) representing a 95% probability that the index has emerged from historical variability and a more relaxed value ($tresh=TI_\sigma$) for a 68% probability.
+**
 
@@
 Further research looking at different approaches of estimating the natural variability component could help reduce the uncertainty in the estimation of the time of emergence of transboundary stocks.
-The ToE was also computed for a less conservative estimation using a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
-In this case, the global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
+
+<!-- In this case, the global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
 \@ref(fig:toeFigureS8)).
@@
 \@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
+-->
 
@@
 We define the threat point as the minimum required catch proportion of a transboundary stock within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-**Any proportion below the defined threat point would result in the unilateral management of the stock assuming that no country will be willing to engage in cooperative management if the future catch proportion of the transboundary stock is lower than it has ever been.**
+Any proportion below the defined threat point would result in the unilateral management of the stock assuming that no country will be willing to engage in cooperative management if the future catch proportion of the transboundary stock is lower than it has ever been.
 
@@
 
-img <- readPNG("../Figures/FigS7.png")
+img <- readPNG("../Figures/FigS6.png")
 grid.raster(img)
@@
 
-img <- readPNG("../Figures/FigS8.png")
+img <- readPNG("../Figures/FigS7.png")
 grid.raster(img)
@@
 
-```{r toeTableS3, eval = T, echo = F}
+```{r toeTableS3, eval = F, echo = F}
 
@@
 
-```{r toeTableS2, eval = T, echo = F}
+```{r toeTableS2, eval = F, echo = F}
 

### 2021-07-26 Scripts/Manuscript.Rmd
flaged changes

@@
 \@ref(fig:toeFigure1)C).
-Using a less conservative threshold (i.e., 1 s.d.
-representing a probability of 68% that the stock has shifted) results in 33% (n = 5,745) more shifting stocks, that is 63% of all studied stocks (Figs.
+Using a less conservative threshold (i.e., one s.d.
+representing a probability of 68% that the stock has shifted) results in 33% (n = 5,745) more shifting stocks, that is, 63% of all studied stocks (Figs.
 \@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
 For both cases, the first shift is modelled to have occurred back in 2006.
-The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be between 2029 $\pm$ 27 years (1 s.d.
-threshold) and 2036 $\pm$ 28 years (2 s.d.
+The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be between 2029 $\pm$ 27 years (one s.d.
+threshold) and 2036 $\pm$ 28 years (two s.d.
 threshold).
-Furthermore, between 81% and 83% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 77% to 80% recording a distributional shift in stocks by the UN 2030 Agenda’s deadline (Fig.
+Furthermore, between 81% and 83% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 77% to 80% of EEZs recording a distributional stocks shift by the UN 2030 Agenda’s deadline (Fig.
 \@ref(fig:toeFigure1)C).**

### 2021-07-26 Scripts/Initial_Analysis.Rmd
misc

@@
 
-# Neighbour list and their respective id (see chunk 3)
+# Neighbor list and their respective id (see chunk 3)
 Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)
@@
 
-# Species depth clasification (SAU-fishbase)
-Spp_depth <-fread(paste(my_path("D"),"SAU/SppTaxonName.csv", sep ="")) %>% 
+# Species depth classification (SAU-fishbase)
+Spp_depth <- fread(paste(my_path("D"),"SAU/SppTaxonName.csv", sep ="")) %>% 
   clean_names()
@@
 
-# The path
-path_world <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile")
-path_world <- "~/Downloads/SAU_Shapefile"
-
-# The File
-fnam_world <- "SAUEEZ_July2015.shp"
-
-# Load it!
-World_eez_sf <- st_read(dsn = path_world,
-                        layer =file_path_sans_ext(fnam_world)) %>% 
+World_eez_sf <- st_read(my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile", "SAUEEZ_July2015.shp")) %>% 
   rename(eez_name = Name) %>% 
@@
   ) %>% 
-  gather("metric","value",com_sum,per) %>% 
-  filter(tresh == selected_tresh)
+  gather("metric","value",com_sum,per) #%>% 
+  # filter(tresh == selected_tresh)
 
@@
             by = c("sau_fishing_entity","tresh")) %>% 
-  filter(tresh == selected_tresh, # tresh hold selection
+  filter(#tresh == selected_tresh, # tresh hold selection
          variable == "value") %>% 
@@
             by = c("sau_fishing_entity","fishing_entity_id")) %>% 
-  distinct(sau_fishing_entity,.keep_all = TRUE) %>% 
+  distinct(sau_fishing_entity,tresh,.keep_all = TRUE) %>%
   filter(!is.na(un_geo_region)) %>% 
@@
 
-### Fucntion for estimating percentile
+### Function for estimating percentile
 
@@
 Percentile_value <- Buble_data %>% 
-  group_by() %>%  # for whole distribution
+  group_by(tresh) %>%  # for whole distribution
         do(quibble(.$log_value, 0.75))
@@
   filter(log_value > Percentile_value$x & mean_yr < 2036) %>% 
-  # View()
-  group_by() %>% 
+  View()
+  group_by(tresh) %>% 
   summarise(sum(sum_value_bi),
@@
   filter(mean_yr < 2036) %>% 
-  # nrow() # 76  coutries
-  # View()
-  group_by() %>% 
+  # nrow() # 76  countries
+  View()
+  group_by(tresh) %>% 
   summarise(sum(sum_value_bi),
@@
 76/length(unique(Buble_data$sau_fishing_entity))
+134/length(unique(Buble_data$sau_fishing_entity))
 

### 2021-07-26 Scripts/Supplements.Rmd
Sent to co-authors

@@
 
-**Materials and methods**
+# Materials and methods
 
-_**Databases and species selection**_
+### Databases and species selection
 
@@
 
-_**Projecting species distributions under climate change**_
+### Projecting species distributions under climate change
 
@@
 
-_**Calculating an index of transboundary range shift**_
+### Calculating an index of transboundary range shift
 
@@
 
-_**Calculating shifts in the shared distribution of transboundary stocks**_
+### Calculating shifts in the shared distribution of transboundary stocks
 
@@
 
-_**Quantifying the intensity of transboundary stocks’ range shift**_
+### Quantifying the intensity of transboundary stocks’ range shift
 
@@
 
-First, we estimated **the catch proportion as the** stock share ratio (SSR) catch of each transboundary stock shared by neighboring nations between 1951 and 2100.
+First, we estimated the catch proportion as the stock share ratio (SSR) catch of each transboundary stock shared by neighboring nations between 1951 and 2100.
 We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock's catch held within each EEZ [@PalaciosAbrantes:2020tv].
@@
 
-_**Statistical analysis**_
+### Statistical analysis
 
@@
 All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages cowplot [@CowplotStreamlined:2019wt], data.table [@Packagedatatable:2019uh], geosphere [@], ggrepel [@PackageggrepelAut:2020to], gmt [@PackagegmtInterfa:2017vh], janitor [@PackagejanitorSim:2018ur], moments [@PackagemomensMome:2015uh], pgirmess [@PackagepgirmessSp:2018um], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], viridis [@PackageviridisDef:2018us], zeallot [@PackagezeallotMul:2018wy], and zoo [@PackagezooSInfr:2019tm].
-Code for all analyses is available at https://github.com/jepa/TransEmergence
+Code for all analyses is available at https://github.com/jepa/EmergingFish
 
@@
 
-_**Figures**_
+# Supplemental Figures
 

### 2021-07-26 Scripts/Manuscript.Rmd
Updated manuscript and read me for submission

@@
 
-# Introduction
-
 Over the past century, human activities have dramatically altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters [@IPCC:2019tn].
@@
 This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-Finally, we computed changes in catch proportion that were below the threat point of neighboring EEZs sharing a stock for two time periods relative to the historical proportion: 2030 (2021 and 2040) matching the UN 2030 Agenda and a mid-century timeline (2040 to 2050).
+<!-- Finally, we computed changes in catch proportion that were below the threat point of neighboring EEZs sharing a stock for two time periods relative to the historical proportion: 2030 (2021 and 2040) matching the UN 2030 Agenda and a mid-century timeline (2040 to 2050).
+-->
 
@@
 threshold).
-Furthermore, between 81% and 83% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 77% to 80% of EEZs recording a distributional stocks shift by the UN 2030 Agenda’s deadline (Fig.
+Furthermore, between 81% and 83% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 77% to 80% of EEZs recording a distributional stock shift by the UN 2030 Agenda’s deadline (Fig.
 \@ref(fig:toeFigure1)C).**
@@
 
-\clearpage
-
-# References
-
-<div id="refs"></div>
-
-\clearpage
-
-# Supplements
+# Methods
 
@@
 ```
+
+# References
+
+<div id="refs"></div>

### 2021-07-30 Scripts/Manuscript.Rmd
moved methods to main article

@@
 # output: bookdown::pdf_document2
-output: bookdown::word_document2
+output: 
+  bookdown::word_document2:
+    number_section: no
 reference_docx:  ../References/sci_template.docx
@@
   - \usepackage{float} # Use the 'float' package
-  - \floatplacement{figure}{H} # Make every figure with caption = h
   - \captionsetup[figure]{name = Fig}
@@
 
-corresponding author: [j.palacio\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}
+**Email**: [j.palacio\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}
 
-# Abstract
+**Author Contributions:**  
+Conceptualization: JPA, WWLC, TF, AT
+Data curation; TF; WWLC; GR
+Formal Analysis: JPA
+Methodology: JPA, WWLC, TF, AT, GR, CCCW
+Visualization: JPA
+Writing – original draft: JPA; WWLC; CCCW; TF
+Writing – review & editing: JPA; WWLC; CCCW; TF; AT: URS; GR
+
+**Competing Interest Statement:** Authors declare that they have no competing interests.
+
+Classification: Biological Science, Environmental Sciences
+
+This PDF file includes:
+Main Text
+Figures 1 to 5
+
+\clearpage
+
+# Abstract {-}
 
@@
 Coupling a large ensemble simulation of an Earth system model to a species distribution model, we show that by 2030 23% of transboundary stocks would have shifted and 78% of the world will experience at least one shifting stock.
-**By 2100, 81% of EEZ waters will see at least one shifting stock with a total of 45% of stocks shifting globally, under a high emissions scenario.** Importantly, many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts.
-**Existing fisheries agreements need to be assessed for their capacity in addressing transboundary shifts, and strengthened where necessary to limit conflict over these fish stocks.**
+By 2100, 81% of EEZ waters will see at least one shifting stock with a total of 45% of stocks shifting globally, under a high emissions scenario.
+Importantly, many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts.
+Existing fisheries agreements need to be assessed for their capacity in addressing transboundary shifts, and strengthened where necessary to limit conflict over these fish stocks.
+
+# Significant statement {-}
+
+Marine fish stocks are shifting their distribution in response to climate change.
+This is of particular importance for the management of stocks shared between neighboring Exclusive Economic Zones, also known as transboundary stocks.
+Here, we find that imminent climate change-induced shifts in transboundary fish stocks are challenging international management and that urging fisheries agreements to be assessed for their capacity in addressing transboundary shifts, and strengthened where necessary to limit conflict over these fish stocks.
 
 \clearpage
+
+# Main text {-}
 
@@
 Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods).
-We then estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig.
-\@ref(fig:toeFigure1N)) and the intensity of this shift (i.e., time of emergence and stock share ratio; Methods).
+We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
+The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.
+\@ref(fig:toeFigure1N)).
+By employing the concept of *time of emergence*, we estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig.
+1).
+Such estimation is based on two thresholds representing a 67 and a 95% probability of shift.
+Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (Methods).
+This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
 We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
@@
 
-We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks under climate change (Methods).
-In brief, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
-The DBEM was driven with ten different simulations of the GFDL ESM2M (i.e., a ten-ensemble member simulation) each of them representing a unique atmosphere and ocean state at the start of the simulation [@Dunne:2012bo; @Dunne:2013db; @Frolicher:2020cc].
+We coupled a mechanistic population dynamic model with a ten-ensemble member simulation of an Earth System Model to project the future global distribution of transboundary stocks under climate change (Methods).
+
+<!-- In brief, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd].
+-->
+<!-- The DBEM was driven with ten different simulations of the GFDL ESM2M (i.e., a ten-ensemble member simulation) each of them representing a unique atmosphere and ocean state at the start of the simulation [@Dunne:2012bo; @Dunne:2013db; @Frolicher:2020cc].
 The simulations were run under historical forcing until 2005, afterwards and until 2100, simulations followed a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5) [@Riahi:2011dk].
+-->
 
-We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
+<!-- We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
 The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.
@@
 The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence".
-We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).
+We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 67% that the index has emerged).
+-->
 
-Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (Methods).
+<!-- Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (Methods).
 A threat point is defined as the minimum payoff a player would be willing to receive to cooperate in a game-theoretic model.
@@
 This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+-->
 <!-- Finally, we computed changes in catch proportion that were below the threat point of neighboring EEZs sharing a stock for two time periods relative to the historical proportion: 2030 (2021 and 2040) matching the UN 2030 Agenda and a mid-century timeline (2040 to 2050).
@@
 
-# Results
+# Results {-}
 
-### Identifying range shifts in the shared distribution of transboundary stocks
+### Identifying range shifts in the shared distribution of transboundary stocks {-}
 
-**Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
+Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
 \@ref(fig:toeFigure1)A), using a two s.d.
@@
 Using a less conservative threshold (i.e., one s.d.
-representing a probability of 68% that the stock has shifted) results in 33% (n = 5,745) more shifting stocks, that is, 63% of all studied stocks (Figs.
+representing a probability of 67% that the stock has shifted) results in 33% (n = 5,745) more shifting stocks, that is, 67% of all studied stocks (Figs.
 \@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
@@
 Furthermore, between 81% and 83% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 77% to 80% of EEZs recording a distributional stock shift by the UN 2030 Agenda’s deadline (Fig.
-\@ref(fig:toeFigure1)C).**
+\@ref(fig:toeFigure1)C).
 
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "**Timeline of shifts in transboundary shared stock distribution.** (A) Number of stocks (cumulative) whose distribution centroids that have shifted beyond natural variations over time (from 2006 to >2100), (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one stock shifting beyond natural variability.
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary shared stock distribution.
+(A) Number of stocks (cumulative) whose distribution centroids that have shifted beyond natural variations over time (from 2006 to >2100), (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one stock shifting beyond natural variability.
 Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change in range shift (Methods), respectively."}
@@
 
-**On average, shifting transboundary stocks represented 27% to 23% (1 and 2 s.d.
+On average, shifting transboundary stocks represented 27% to 23% (1 and 2 s.d.
 thresholds) of annual revenue from fisheries targeting transboundary stocks across the world's EEZs between 2004 and 2010 (Fig.
@@
 However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
-Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).**
+Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).
 
-```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "**Year of change in the shared distribution of 4,119 transboundary stocks.
+```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks.
 A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
@@
 
-```{r toeFigure3, eval = T,  echo = F, fig.width=8, fig.height=6, fig.pos= "H", fig.cap =  "**Average range shift of transboundary stocks and corresponding fishing revenue per country.
+```{r toeFigure3, eval = T,  echo = F, fig.width=8, fig.height=6, fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks and corresponding fishing revenue per country.
 Countries are coded according to different colored symbols representative of 5 distinct regions.
@@
 
-### Shifting intensity of transboundary stocks
+### Shifting intensity of transboundary stocks {-}
 
@@
 Sensitivity tests to a lower threshold of 1 s.d.
-(for a 68% confidence) resulted in a global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27, a difference of about 7 years (Fig.
+(for a 67% confidence) resulted in a global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27, a difference of about 7 years (Fig.
 \@ref(fig:toeFigureS4)).
@@
 
-```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "**Changes in stock proportion of neighboring Exclusive Economic Zones by 2030 (2021-2040) relative to 1951-2005.** Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock proportion of neighboring Exclusive Economic Zones by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
 Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors.
@@
 
-# Discussion
+# Discussion {-}
 
@@
 
-### High present-day climate risk for transboundary fisheries
+### High present-day climate risk for transboundary fisheries {-}
 
@@
 
-### Climate risk hotspots for transboundary fisheries management
+### Climate risk hotspots for transboundary fisheries management {-}
 
@@
 
-### Potential drivers of shift and managing an uncertain future
+### Potential drivers of shift and managing an uncertain future {-}
 
@@
 
-# Conclusion
+<!-- # Conclusion {-} -->
 
@@
 
-# Methods
+# Materials and Methods {-}
 
-<!-- Set Figure, table and equations numbering according to the chapter -->
-\renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{figure}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{table}{0}
+## Databases and species selection
 
-```{r supplements, child = '../Scripts/Supplements.Rmd', eval = T}
+The analyses detailed herein are based on 633 exploited marine transboundary species that account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
+When a species was shared by a pair of neighboring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.
+Using the transboundary species Atlantic cod (*Gadus morhua*) as an example and this study’s definition of transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not as they do not have juxtaposing EEZs.
+We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015; http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
+\@ref(fig:toeFigureS5)) including island territories.
+We determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp].
+Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
+The habitat preference of each species was determined following FishBase (http://www.fishbase.org) for fish species and SeaLifeBase (http://www.fishbase.org) for invertebrates (Table S\@ref(tab:toeTableS1)).
+For each stock and EEZ, we used Sea Around Us data to estimate catch and fishing revenue from fishing activities.
+We report both average catch and revenue for the last available decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
+
+## Projecting species distributions under climate change
+
+We projected the future distribution of transboundary stocks using a dynamic bioclimatic envelope model (hereafter called DBEM).
+The DBEM determines the environmental space a species will occupy based on physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
+The DBEM then estimates species’ abundance and maximum catch potential (a proxy for maximum sustainable yield) over a regular spatial grid of 0.5 of latitude per 0.5 of longitude.
+Below we provide an overall description of the model.
+Please see @Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd for a more detailed version.
+Importantly, the DBEM is able to project catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
+*in revision*].
+
+<!-- Included more details of the DBEM in response to reviewer 3 -->
+For each species, the DBEM uses the Sea Around Us (www.searaundus.org) distributional data and environmental variables from 1970 to 2000 to estimate the current distribution and environmental profile of each species.
+Moreover, it projects the future habitat suitability in each grid-cell (*i*) based on the following components;
+
+\begin{equation}
+\label{eq:one}
+Hs = P(T_{i},TPP) *  P(Ba_{i}, Min_{D}, Max_{D}) * P(Hab_{i,j}, Hab_{Assoc}) * (Sal_{i}, Sal_{Assoc}) * P(Ice_{i}, Ice_{P})
+\end{equation}
+
+where *T* is seawater temperature, *TPP* is the species temperature profile, *Ba* is bathymetry, $Min_{D}$ and $Max{D}$ are the species minimum and maximum depth limits, *Hab* is the proportion of area of the habitat type *j* relative to the total seawater area of the cell *i*, $Hab_{Assoc}$ is an habitat association index, *Sal* is the salinity class of cell *i* and $Sal_{Assoc}$ is a salinity class association index.
+Finally, for polar regions and species, *Ice* is sea ice extent, and $Ice_{P}$ is association to sea ice.
+For pelagic species, the model uses environmental variables at the surface whereas demersal and benthic species’ distributions are driven by ocean bottom variables.
+
+The DBEM also integrates each species preference profile with physiology and population dynamics to project relative biomass assuming that spatio-temporal dynamics are determined by intrinsic growth rate, larval dispersion, and adult migration;
+
+\begin{equation}
+\label{eq:two}
+\frac{dA}{dt} = \sum_{j=1}^{n} (L_{ji} + Ad_{ji} + G_i)
+\end{equation}
+
+where $A_i$ is the relative abundance of the gri-cell *i*, *G* is the intrinsic population growth and $L_{ji}$ and $Ad_{ji}$ are settled larvae and net migrated adults from surrounding cells (*j*), respectively.
+Larval dispersal is modeled through ocean current with an advection-diffusion-reaction model.
+Finally, $G_i$ estimates the intrinsic growth following a logistic equation and the intrinsic rate of population increase (*r*):
+
+\begin{equation}
+\label{eq:three}
+G_i = r * A_i * (1-\frac{A_i}{A_{\infty,i}})
+\end{equation}
+
+Finally, the DBEM projects maximum catch potential by setting fishing mortality to half of the intrinsic population growth rate of the species (*f = r/2*).
+
+The DBEM was driven by simulated ocean conditions from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
+The GFDL-ESM2M was run under historical forcing until 2005 and follows a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5), over the 2006-2100 period [@Riahi:2011dk].
+We chose the 1951-2100 time period to match the historical period in the GFDL ESM2M simulations over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Rodgers:2015jq].
+Because the main approach of this paper relies on understanding the spatial and temporal variation of a stock’s distribution, we have to understand distribution variability during both historical and future periods, to infer differences between time frames.
+To this end, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in 1950, resulting in a unique atmosphere and ocean state at each point in time after about three years for surface and eight years for subsurface waters [@Frolicher:2020cc].
+By design, variations among ensemble members are then solely due to natural internal variability (e.g., different phases of the El Niño Southern Oscillation-ENSO).
+Thus, for our experiment, each of the ten GFDL-ESM2M ensemble simulations were started from infinitesimally small differences in Earth system initial conditions in the year 1950, resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
+
+## Calculating an index of transboundary range shift
+
+We developed a Transboundary Index (TI) to evaluate range shifts in the shared distribution of transboundary stocks under climate change.
+This index represents the shift in the distribution centroid of a transboundary stock relative to the centroid of neighboring EEZs that share this stock (Fig.
+\@ref(fig:toeFigure1)).
+The centroid of a transboundary stock was determined by the average ($\mu$) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across grid cells within the neighboring EEZs sharing the stock.
+Therefore,
+
+\begin{equation}
+\begin{split}
+\label{eq:four}
+lat_{ts} = \mu (lat_{per}) \\
+lon_{ts} = \mu (lon_{per})
+\end{split}
+\end{equation}
+
+where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the specified percentile ($per^{th}$) of the projected transboundary stock abundance.
+To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighboring EEZs sharing the stock was above the $95^{th}$ percentile.
+A sensitivity analysis using a subset of species (n = 34) for all EEZs  to examine the effects of different thresholds (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) on the calculated index value showed no apparent difference between percentiles (Fig.
+\@ref(fig:toeFigureS7)).
+The centroid of each EEZ was estimated using the st package in R (Fig.
+\@ref(fig:toeFigureS5)).
+For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids (i.e., EEZ and stock) using the geosphere package in R assuming the Earth is a perfect sphere and ignoring geographic barriers;
+
+\begin{equation}
+\label{eq:five}
+D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
+\end{equation}
+
+where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively, and $lon_{eez}$ and $lon_{ts}$ are associated longitudes.
+Then, for each year and for each ensemble member between 1951 and 2100 we calculated the transboundary index (TI) as follows:
+
+\begin{equation}
+\label{eq:six}
+TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
+\end{equation}
+
+where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (th, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
+The TI was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg] and to match transboundary fisheries management that often integrates a longer time-frame in their consideration.
+
+<!-- Removed in response to reviewer 3 -->
+<!-- Thus, a higher TI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels.
+From the perspective of a state’s fisheries, increasing TI indicates an increase in the challenges to effective management and sustainability of the transboundary stock’s fisheries.
+In contrast, a decreasing TI suggests that the sharing of the stocks between neighboring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries.
+-->
+
+## Calculating shifts in the shared distribution of transboundary stocks
+
+Knowing the point in time at which the distribution of a shared stock will diverge from historical variability is important to inform decision makers by when, ideally, climate adaptation plans will need to have been implemented [@Link:2010ei].
+To that end, we adopted the concept of time of emergence, commonly applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev] and defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (i.e., historical natural variation) [@Hawkins:2012hc].
+The premise behind time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
+The TI’s natural internal variability was estimated from the ensemble members by firstly averaging the TI from 1951 to 2005 and then estimating the variation ($TI_\sigma$) across ensemble members.
+The TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$).
+We then set a threshold (*tresh*) of $TI_\sigma$ to define the time of emergence of transboundary stocks (See Fig.
+\@ref(fig:toeFigure1) for a graphical description):
+
+<!-- We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig.
+\@ref(fig:toeFigure1) for a graphical description): -->
+
+<!-- $EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$ -->
+<!-- $ToE = t when Emergence index[t] > 1$ -->
+
+\begin{equation}
+\label{eq:seven}
+ToE = TI_\mu[t']> tresh
+\end{equation}
+
+This way we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TI “emerges” from historical internal variability (*tresh*).
+It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution.
+Two different arbitrary values of *tresh* were used.
+A conservative value ($tresh = 2* TI_\sigma$) representing a 95% probability that the index has emerged from historical variability and a more relaxed value ($tresh=TI_\sigma$) for a 67% probability.
+
+Since the ToE method is sensitive to the noise estimate (i.e., historical natural variability), we conducted sensitivity analysis where we first calculated the average across the 10 ensemble members for each year between 1951 to 2005 and the calculating the standard deviation over 1950-2005.
+Under this method, the natural variability is slightly larger resulting in a latter overall time of emergence (average ToE = 2057 vs 2036 of the original method).
+However, this method projects 1,139 more emerging stocks than the original (5,258 stocks vs 4,119 of the original method).
+The minimum and maximum ToE remains the same, and the number of EEZs with emerging stocks slightly increases with the alternative method.
+Further research looking at different approaches of estimating the natural variability component could help reduce the uncertainty in the estimation of the time of emergence of transboundary stocks.
+
+<!-- In this case, the global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
+\@ref(fig:toeFigureS8)).
+In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
+\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
+-->
+
+## Quantifying the intensity of transboundary stocks’ range shift
+
+Here, we adopt the concept of threat point to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs.
+The concept of threat point comes from Game Theory and has been widely used in shared fisheries management [@Bailey:2010ga; @Sumaila:2013hv; @Sumaila:2020tp; @Clark:1975uu; @Munro:1979dg].
+In a game theoretic model, a player’s strategy (e.g., to act cooperatively or not) will have direct consequences for other players, which, in turn, will affect the overall outcome of the game [@Bailey:2010ga].
+Cooperation will often result in the maximization of benefits for the system rather than for individual players.
+However, for a cooperative strategy to work, the benefit a player gets must not be less than the benefit under a non-cooperative strategy.
+Thus, the “threat point” is defined as the minimum payoff a player is willing to receive to cooperate in a game-theoretic model [@Nash:1953vr].
+We define the threat point as the minimum required catch proportion of a transboundary stock within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+Any proportion below the defined threat point would result in the unilateral management of the stock assuming that no country will be willing to engage in cooperative management if the future catch proportion of the transboundary stock is lower than it has ever been.
+
+First, we estimated the catch proportion as the stock share ratio (SSR) catch of each transboundary stock shared by neighboring nations between 1951 and 2100.
+We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock's catch held within each EEZ [@PalaciosAbrantes:2020tv].
+Second, to reduce the effects of variability, we averaged the calculated proportion of stock occupying a given EEZ into three time periods: the historical time period ($SSR_{th}$) from 1951 to 2005, and two future periods, the early $21^{st}$ century, between 2021 and 2040 ($SSR_{te}$), and the mid $21^{st}$ century, defined as the average of 2041 to 2060 ($SSR_{tm}$).
+These future time periods were chosen to correspond to the challenges of achieving fisheries-relevant UN-SDGs by 2030 (i.e., the 2030 Agenda), such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) and SDG 1.2 (poverty reduction) [@Singh:2017ds; @UnitedNations:1Qe8KIcN].
+The analysis was replicated for projected stock distributions from each of the ten ensemble members and results were averaged across ensemble members.
+Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm 2 \sigma$, where  is the standard deviation of $SSR_{th}$.
+Thus, a change of SSR catch beyond an EEZ’s threat point happens when the future SSR catch exceeds two standard deviations of the historical variations in SSR (i.e., when $SSR_{et} \geqq (SSR_{th} + 2 \sigma$) or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
+Finally, we estimated the percentage change in SSR ($\Delta SSR$) of each future time period ($SSR_{tf}$) relative to the historic time period $SSR_{th}$ for each stock whose share ratio exceeded the threat point following [@PalaciosAbrantes:2020tv];
+
+<!-- $$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$ -->
+
+\begin{equation}
+\label{eq:eight}
+\Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100
+\end{equation}
+
+## Statistical analysis
+
+We tested the results for normality (e.g., skewness, kurtosis) and performed two non-parametric Kruskal–Wallis tests by ranks [@Hollander:2013wz]  to investigate geopolitical and ecological differences in the range shift of transboundary species.
+Specifically, we tested if the habitat preference of transboundary species and the geographic location of EEZs would have any effect on the transboundary stocks distributional range shift.
+In both cases, our null hypothesis was that there were no significant differences in the year of shift across habitat association or EEZs.
+All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages cowplot [@CowplotStreamlined:2019wt], data.table [@Packagedatatable:2019uh], geosphere [@], ggrepel [@PackageggrepelAut:2020to], gmt [@PackagegmtInterfa:2017vh], janitor [@PackagejanitorSim:2018ur], moments [@PackagemomensMome:2015uh], pgirmess [@PackagepgirmessSp:2018um], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], viridis [@PackageviridisDef:2018us], zeallot [@PackagezeallotMul:2018wy], and zoo [@PackagezooSInfr:2019tm].
+Code for all analyses is available at https://github.com/jepa/EmergingFish
+
+# Acknowledgments {-}
+We would like to acknowledge inputs from T.
+Clarke and L.
+Kwong on previous iterations of the manuscript and figures.
+A.T.
+acknowledges  funding from the Horizon 2020 Framework Programme (724289).
+C.C.C.W acknowledges funding from the David and Lucile Packard Foundation (2019-68336), the Gordon and Betty Moore Foundation (GBMF5668.02), and the Walton Family Foundation (2018-1371).
+J.P.A and W.W.L.C.
+acknowledge funding from the Social Sciences and Humanities Research Council of Canada, Partnership Grant and the Natural Sciences and Engineering Research Council of Canada, Partnership Grant.
+T.F.
+acknowledges funding from the Horizon 2020 Framework Programme (820989), the Swiss National Science Foundation (PP00P2_170687), and the Swiss National Supercomputing Centre.
+U.R.S.
+acknowledges funding from the Social Sciences and Humanities Research Council of Canada, Partnership Grant.
+
+# References {-}
+
+<div id="refs"></div>
+
+# Figures and Tables
+
+```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone (EEZ) to another.
+The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring EEZ sharing the stock (Top panel).
+Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel).
+Only showing a subset of ensemble members for clarity."}
+
+img <- readPNG("../Figures/Fig1.png")
+grid.raster(img)
 
@@
 
-# References
+\clearpage
 
-<div id="refs"></div>
+```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary shared stock distribution.
+(A) Number of stocks (cumulative) whose distribution centroids that have shifted beyond natural variations over time (from 2006 to >2100), (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one stock shifting beyond natural variability.
+Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change in range shift (Methods), respectively."}
+
+img <- readPNG("../Figures/Fig2.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks.
+A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygons display average year of range shifts within them.
+Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cool colors represent a late shift/low fishing revenue.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
+B) Year of shared distribution shifts by UN sub-regions.
+Points color coded by the median year of range shift.
+Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda.
+N = North, S = South, W = West and E = East.
+Ltn.
+Ame.
+and the Car.
+= Latin America and the Caribbean.
+Aus and New Z.
+= Australia and New Zealand."}
+
+img <- readPNG("../Figures/Fig3.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigure3, eval = T,  echo = F, fig.width=8, fig.height=6, fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks and corresponding fishing revenue per country.
+Countries are coded according to different colored symbols representative of 5 distinct regions.
+The horizontal line separates countries above and below the 75th percentile of annual fishing revenue.
+Showing country names for the eleven nations above the 75th percentile and projected to have an average shift across transboundary stocks within its EEZ within 5 years of the 2030 Agenda deadline."}
+
+img <- readPNG("../Figures/Fig4.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock proportion of neighboring Exclusive Economic Zones by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors.
+Panel B highlights changes for the Caribbean and Northwestern Europe."}
+
+img <- readPNG("../Figures/Fig5.png")
+grid.raster(img)
+
+```

### 2021-07-30 Scripts/Supplements.Rmd
moved methods to main article

@@
 output:
-  bookdown::word_document2:
+  # bookdown::word_document2
+    bookdown::pdf_document2:
     toc : no
@@
     number_sections: yes
-reference_docx: ../References/supplementary_template.docx
-bibliography: ../References/ToE_Ref.bib
-csl: ../References/science.csl
 header-includes:
@@
 
-# Materials and methods
-
-### Databases and species selection
-
-The analyses detailed herein are based on 633 exploited marine transboundary species that account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-When a species was shared by a pair of neighboring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.
-Using the transboundary species Atlantic cod (*Gadus morhua*) as an example and this study’s definition of transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not as they do not have juxtaposing EEZs.
-We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015; http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
-\@ref(fig:toeFigureS5)) including island territories.
-We determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp].
-Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
-The habitat preference of each species was determined following FishBase (http://www.fishbase.org) for fish species and SeaLifeBase (http://www.fishbase.org) for invertebrates (Table S\@ref(tab:toeTableS1)).
-For each stock and EEZ, we used Sea Around Us data to estimate catch and fishing revenue from fishing activities.
-We report both average catch and revenue for the last available decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc].
-
-### Projecting species distributions under climate change
-
-We projected the future distribution of transboundary stocks using a dynamic bioclimatic envelope model (hereafter called DBEM).
-The DBEM determines the environmental space a species will occupy based on physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry.
-The DBEM then estimates species’ abundance and maximum catch potential (a proxy for maximum sustainable yield) over a regular spatial grid of 0.5 of latitude per 0.5 of longitude.
-Below we provide an overall description of the model.
-Please see @Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd for a more detailed version.
-Importantly, the DBEM is able to project catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
-*in revision*].
-
-<!-- Included more details of the DBEM in response to reviewer 3 -->
-For each species, the DBEM uses the Sea Around Us (www.searaundus.org) distributional data and environmental variables from 1970 to 2000 to estimate the current distribution and environmental profile of each species.
-Moreover, it projects the future habitat suitability in each grid-cell (*i*) based on the following components;
-
-\begin{equation}
-\label{eq:one}
-Hs = P(T_{i},TPP) *  P(Ba_{i}, Min_{D}, Max_{D}) * P(Hab_{i,j}, Hab_{Assoc}) * (Sal_{i}, Sal_{Assoc}) * P(Ice_{i}, Ice_{P})
-\end{equation}
-
-where *T* is seawater temperature, *TPP* is the species temperature profile, *Ba* is bathymetry, $Min_{D}$ and $Max{D}$ are the species minimum and maximum depth limits, *Hab* is the proportion of area of the habitat type *j* relative to the total seawater area of the cell *i*, $Hab_{Assoc}$ is an habitat association index, *Sal* is the salinity class of cell *i* and $Sal_{Assoc}$ is a salinity class association index.
-Finally, for polar regions and species, *Ice* is sea ice extent, and $Ice_{P}$ is association to sea ice.
-For pelagic species, the model uses environmental variables at the surface whereas demersal and benthic species’ distributions are driven by ocean bottom variables.
-
-The DBEM also integrates each species preference profile with physiology and population dynamics to project relative biomass assuming that spatio-temporal dynamics are determined by intrinsic growth rate, larval dispersion, and adult migration;
-
-\begin{equation}
-\label{eq:two}
-\frac{dA}{dt} = \sum_{j=1}^{n} (L_{ji} + Ad_{ji} + G_i)
-\end{equation}
-
-where $A_i$ is the relative abundance of the gri-cell *i*, *G* is the intrinsic population growth and $L_{ji}$ and $Ad_{ji}$ are settled larvae and net migrated adults from surrounding cells (*j*), respectively.
-Larval dispersal is modeled through ocean current with an advection-diffusion-reaction model.
-Finally, $G_i$ estimates the intrinsic growth following a logistic equation and the intrinsic rate of population increase (*r*):
-
-\begin{equation}
-\label{eq:three}
-G_i = r * A_i * (1-\frac{A_i}{A_{\infty,i}})
-\end{equation}
-
-Finally, the DBEM projects maximum catch potential by setting fishing mortality to half of the intrinsic population growth rate of the species (*f = r/2*).
-
-The DBEM was driven by simulated ocean conditions from a ten ensemble member simulation of the Geophysical Fluid Dynamics Laboratory Earth system model (GFDL-ESM2M) to project the distributions of the 633 species from 1951 to 2100 [@Dunne:2012bo; @Dunne:2013db; @Rodgers:2015jq].
-The GFDL-ESM2M was run under historical forcing until 2005 and follows a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5), over the 2006-2100 period [@Riahi:2011dk].
-We chose the 1951-2100 time period to match the historical period in the GFDL ESM2M simulations over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Rodgers:2015jq].
-Because the main approach of this paper relies on understanding the spatial and temporal variation of a stock’s distribution, we have to understand distribution variability during both historical and future periods, to infer differences between time frames.
-To this end, each of the ten GFDL-ESM2M ensembles were started from infinitesimally small differences in Earth system initial conditions in 1950, resulting in a unique atmosphere and ocean state at each point in time after about three years for surface and eight years for subsurface waters [@Frolicher:2020cc].
-By design, variations among ensemble members are then solely due to natural internal variability (e.g., different phases of the El Niño Southern Oscillation-ENSO).
-Thus, for our experiment, each of the ten GFDL-ESM2M ensemble simulations were started from infinitesimally small differences in Earth system initial conditions in the year 1950, resulting in a unique atmosphere and ocean state at each point in time after about three years for surface waters and eight years for subsurface waters [@Frolicher:2020cc].
-
-### Calculating an index of transboundary range shift
-
-We developed a Transboundary Index (TI) to evaluate range shifts in the shared distribution of transboundary stocks under climate change.
-This index represents the shift in the distribution centroid of a transboundary stock relative to the centroid of neighboring EEZs that share this stock (Fig.
-\@ref(fig:toeFigure1)).
-The centroid of a transboundary stock was determined by the average ($\mu$) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across grid cells within the neighboring EEZs sharing the stock.
-Therefore,
-
-\begin{equation}
-\begin{split}
-\label{eq:four}
-lat_{ts} = \mu (lat_{per}) \\
-lon_{ts} = \mu (lon_{per})
-\end{split}
-\end{equation}
-
-where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the specified percentile ($per^{th}$) of the projected transboundary stock abundance.
-To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighboring EEZs sharing the stock was above the $95^{th}$ percentile.
-A sensitivity analysis using a subset of species (n = 34) for all EEZs  to examine the effects of different thresholds (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) on the calculated index value showed no apparent difference between percentiles (Fig.
-\@ref(fig:toeFigureS7)).
-The centroid of each EEZ was estimated using the st package in R (Fig.
-\@ref(fig:toeFigureS5)).
-For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids (i.e., EEZ and stock) using the geosphere package in R assuming the Earth is a perfect sphere and ignoring geographic barriers;
-
-\begin{equation}
-\label{eq:five}
-D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
-\end{equation}
-
-where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively, and $lon_{eez}$ and $lon_{ts}$ are associated longitudes.
-Then, for each year and for each ensemble member between 1951 and 2100 we calculated the transboundary index (TI) as follows:
-
-\begin{equation}
-\label{eq:six}
-TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
-\end{equation}
-
-where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (th, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
-The TI was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg] and to match transboundary fisheries management that often integrates a longer time-frame in their consideration.
-
-<!-- Removed in response to reviewer 3 -->
-<!-- Thus, a higher TI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels.
-From the perspective of a state’s fisheries, increasing TI indicates an increase in the challenges to effective management and sustainability of the transboundary stock’s fisheries.
-In contrast, a decreasing TI suggests that the sharing of the stocks between neighboring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries.
--->
-
-### Calculating shifts in the shared distribution of transboundary stocks
-
-Knowing the point in time at which the distribution of a shared stock will diverge from historical variability is important to inform decision makers by when, ideally, climate adaptation plans will need to have been implemented [@Link:2010ei].
-To that end, we adopted the concept of time of emergence, commonly applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev] and defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (i.e., historical natural variation) [@Hawkins:2012hc].
-The premise behind time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
-The TI’s natural internal variability was estimated from the ensemble members by firstly averaging the TI from 1951 to 2005 and then estimating the variation ($TI_\sigma$) across ensemble members.
-The TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$).
-We then set a threshold (*tresh*) of $TI_\sigma$ to define the time of emergence of transboundary stocks (See Fig.
-\@ref(fig:toeFigure1) for a graphical description):
-
-<!-- We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig.
-\@ref(fig:toeFigure1) for a graphical description): -->
-
-<!-- $EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$ -->
-<!-- $ToE = t when Emergence index[t] > 1$ -->
-
-\begin{equation}
-\label{eq:seven}
-ToE = TI_\mu[t']> tresh
-\end{equation}
-
-**This way we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TI “emerges” from historical internal variability (*tresh*).
-It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution.
-Two different arbitrary values of *tresh* were used.
-A conservative value ($tresh = 2* TI_\sigma$) representing a 95% probability that the index has emerged from historical variability and a more relaxed value ($tresh=TI_\sigma$) for a 68% probability.
-**
-
-Since the ToE method is sensitive to the noise estimate (i.e., historical natural variability), we conducted sensitivity analysis where we first calculated the average across the 10 ensemble members for each year between 1951 to 2005 and the calculating the standard deviation over 1950-2005.
-Under this method, the natural variability is slightly larger resulting in a latter overall time of emergence (average ToE = 2057 vs 2036 of the original method).
-However, this method projects 1,139 more emerging stocks than the original (5,258 stocks vs 4,119 of the original method).
-The minimum and maximum ToE remains the same, and the number of EEZs with emerging stocks slightly increases with the alternative method.
-Further research looking at different approaches of estimating the natural variability component could help reduce the uncertainty in the estimation of the time of emergence of transboundary stocks.
-
-<!-- In this case, the global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig.
-\@ref(fig:toeFigureS8)).
-In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
-\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
--->
-
-### Quantifying the intensity of transboundary stocks’ range shift
-
-Here, we adopt the concept of threat point to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs.
-The concept of threat point comes from Game Theory and has been widely used in shared fisheries management [@Bailey:2010ga; @Sumaila:2013hv; @Sumaila:2020tp; @Clark:1975uu; @Munro:1979dg].
-In a game theoretic model, a player’s strategy (e.g., to act cooperatively or not) will have direct consequences for other players, which, in turn, will affect the overall outcome of the game [@Bailey:2010ga].
-Cooperation will often result in the maximization of benefits for the system rather than for individual players.
-However, for a cooperative strategy to work, the benefit a player gets must not be less than the benefit under a non-cooperative strategy.
-Thus, the “threat point” is defined as the minimum payoff a player is willing to receive to cooperate in a game-theoretic model [@Nash:1953vr].
-We define the threat point as the minimum required catch proportion of a transboundary stock within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-Any proportion below the defined threat point would result in the unilateral management of the stock assuming that no country will be willing to engage in cooperative management if the future catch proportion of the transboundary stock is lower than it has ever been.
-
-First, we estimated the catch proportion as the stock share ratio (SSR) catch of each transboundary stock shared by neighboring nations between 1951 and 2100.
-We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock's catch held within each EEZ [@PalaciosAbrantes:2020tv].
-Second, to reduce the effects of variability, we averaged the calculated proportion of stock occupying a given EEZ into three time periods: the historical time period ($SSR_{th}$) from 1951 to 2005, and two future periods, the early $21^{st}$ century, between 2021 and 2040 ($SSR_{te}$), and the mid $21^{st}$ century, defined as the average of 2041 to 2060 ($SSR_{tm}$).
-These future time periods were chosen to correspond to the challenges of achieving fisheries-relevant UN-SDGs by 2030 (i.e., the 2030 Agenda), such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) and SDG 1.2 (poverty reduction) [@Singh:2017ds; @UnitedNations:1Qe8KIcN].
-The analysis was replicated for projected stock distributions from each of the ten ensemble members and results were averaged across ensemble members.
-Third, we defined a threat point for each EEZ’s stock as $SSR_{th} \pm 2 \sigma$, where  is the standard deviation of $SSR_{th}$.
-Thus, a change of SSR catch beyond an EEZ’s threat point happens when the future SSR catch exceeds two standard deviations of the historical variations in SSR (i.e., when $SSR_{et} \geqq (SSR_{th} + 2 \sigma$) or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$.
-Finally, we estimated the percentage change in SSR ($\Delta SSR$) of each future time period ($SSR_{tf}$) relative to the historic time period $SSR_{th}$ for each stock whose share ratio exceeded the threat point following [@PalaciosAbrantes:2020tv];
-
-<!-- $$ \Delta{SSR_{e,m}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100 $$ -->
-
-\begin{equation}
-\label{eq:eight}
-\Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100
-\end{equation}
-
-### Statistical analysis
-
-We tested the results for normality (e.g., skewness, kurtosis) and performed two non-parametric Kruskal–Wallis tests by ranks [@Hollander:2013wz]  to investigate geopolitical and ecological differences in the range shift of transboundary species.
-Specifically, we tested if the habitat preference of transboundary species and the geographic location of EEZs would have any effect on the transboundary stocks distributional range shift.
-In both cases, our null hypothesis was that there were no significant differences in the year of shift across habitat association or EEZs.
-All analyses were run using the statistical software R version 3.5.2 (2018-12-20; Eggshell Igloo) with the packages cowplot [@CowplotStreamlined:2019wt], data.table [@Packagedatatable:2019uh], geosphere [@], ggrepel [@PackageggrepelAut:2020to], gmt [@PackagegmtInterfa:2017vh], janitor [@PackagejanitorSim:2018ur], moments [@PackagemomensMome:2015uh], pgirmess [@PackagepgirmessSp:2018um], rfishbase [@PackagerfishbaseR:2019th], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], viridis [@PackageviridisDef:2018us], zeallot [@PackagezeallotMul:2018wy], and zoo [@PackagezooSInfr:2019tm].
-Code for all analyses is available at https://github.com/jepa/EmergingFish
-
-\clearpage
-
 # Supplemental Figures
 
-```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
-Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
-The category “Other” consists of species that have no “habitat preference” classification.
-The number of stocks included in this analysis for each habitat type is noted in parentheses.
-Dashed whiskers represent the 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Solid line represents the median time of emergence per habitat preference."}
+S1
 
-img <- readPNG("../Figures/FigS1.png")
-grid.raster(img)
-
-```
-
-\clearpage
-
-```{r toeFigureS2, eval = T,  echo = F, fig.width=12, fig.height=7, fig.cap =".
-Year of change in the shared distribution of 5,745 transboundary stocks using a one standard deviation threshold.
+```{r toeFigureS2, eval = T,  echo = F, fig.width=12, fig.height=7, fig.cap ="Year of change in the shared distribution of 5,745 transboundary stocks using a one standard deviation threshold.
 A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
@@
 
+S2
+
+```{r toeFigureS8, eval = T,  echo = F, fig.width=6, fig.height=6, fig.cap ="Average distribution of transboundary stock’s year of change in their shared distribution.
+Left panel is the average year across stocks and right panel the standard deviation.
+Yellow represents a one standard deviation distance from the historical average and purple represents a two standard deviation distance from the historical average."}
+
+img <- readPNG("../Figures/FigS7.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+S3
+
+```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
+Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
+The category “Other” consists of species that have no “habitat preference” classification.
+The number of stocks included in this analysis for each habitat type is noted in parentheses.
+Dashed whiskers represent the 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Solid line represents the median time of emergence per habitat preference."}
+
+img <- readPNG("../Figures/FigS1.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+S4
+
 ```{r toeFigureS3, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Distribution of changes in transboundary stock proportion per stock and neighboring Exclusive Economic Zones.
@@
 \clearpage
+
+S5
 
@@
 
-```{r toeFigureS5, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their estimated centroids (black points)."}
-
-img <- readPNG("../Figures/FigS5.png")
-grid.raster(img)
-
-```
-
-\clearpage
+S6
 
@@
 
-```{r toeFigureS8, eval = T,  echo = F, fig.width=6, fig.height=6, fig.cap ="Average distribution of transboundary stock’s year of change in their shared distribution.
-Left panel is the average year across stocks and right panel the standard deviation.
-Yellow represents a one standard deviation distance from the historical average and purple represents a two standard deviation distance from the historical average."}
+S7
 
-img <- readPNG("../Figures/FigS7.png")
+```{r toeFigureS5, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their estimated centroids (black points)."}
+
+img <- readPNG("../Figures/FigS5.png")
 grid.raster(img)
@@
 ```
+
+\clearpage
 
@@
 ) %>% 
+  write.csv(.,"habitat_preference.csv")
  kable(

### 2021-08-24 Scripts/Manuscript.Rmd
updates for submission to gcb

@@
 title: "Timing and magnitude of climate driven range shifts in transboundary fish stocks challenge their management"
-author: "REVISED MANUSCRIPT. DO NOT SHARE."
-date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.
-^2^Climate and Environmental Physics, Physics Institute, University of Bern.
-^3^Oeschger Centre for Climate Change Research, University of Bern, Switzerland.
-^4^School of Environmental Sciences, University of Liverpool.
-^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States"
+date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada.^2^Center for Limnology, University of Wisconsin, Madison.
+^3^Climate and Environmental Physics, Physics Institute, University of Bern.
+^4^Oeschger Centre for Climate Change Research, University of Bern, Switzerland.
+^5^School of Environmental Sciences, University of Liverpool.
+^6^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States"
 subtitle: "Juliano Palacios-Abrantes*^1,2^, Thomas L. Frölicher^3,4^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^5^, Colette C.C. Wabnitz^1,6^, and William W.L. Cheung^1^"
-# output: bookdown::pdf_document2
 output: 
-  bookdown::word_document2:
+  bookdown::pdf_document2:
     number_section: no
+    toc: no
+    latex_engine: xelatex
 reference_docx:  ../References/sci_template.docx
 bibliography: ../References/ToE_Ref.bib
-csl: ../References/pnas.csl
+csl: ../References/gcb.csl
 header-includes:
@@
   - \usepackage{caption}
+  - \usepackage[left]{lineno} # for numbers in lines
   - \usepackage{float} # Use the 'float' package
   - \captionsetup[figure]{name = Fig}
+  - \floatplacement{figure}{H}
 ---
+
+<!-- Running title: Imminent shifts of transboundary fish stocks -->
 
@@
 \doublespacing
-
-<!-- # Instructions  -->
-
-<!-- Research articles should present a major advance and can be short or long (maximum 15,000 words) format.
-Submissions must include an abstract (up to 150 words), an introduction, and sections with brief informative subheadings.
-You may include up to ten figures and/or tables and about 60 references.
-Supplementary materials should be limited to information that is not essential for the general understanding of the research presented in the main text and can include data sets, figures, tables, videos, or audio files.
-You may include up to a total of ten figures and/or tables (combined) throughout the supplemental text.
--->
 
@@
 
-**Email**: [j.palacio\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}
+<!-- **Email**: [j.palacio\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email} -->
 
-**Author Contributions:**  
-Conceptualization: JPA, WWLC, TF, AT
-Data curation; TF; WWLC; GR
-Formal Analysis: JPA
-Methodology: JPA, WWLC, TF, AT, GR, CCCW
-Visualization: JPA
-Writing – original draft: JPA; WWLC; CCCW; TF
-Writing – review & editing: JPA; WWLC; CCCW; TF; AT: URS; GR
+<!-- **Author Contributions:**   -->
+<!-- Conceptualization: JPA, WWLC, TF, AT -->
+<!-- Data curation; TF; WWLC; GR -->
+<!-- Formal Analysis: JPA -->
+<!-- Methodology: JPA, WWLC, TF, AT, GR, CCCW -->
+<!-- Visualization: JPA -->
+<!-- Writing – original draft: JPA; WWLC; CCCW; TF -->
+<!-- Writing – review & editing: JPA; WWLC; CCCW; TF; AT: URS; GR -->
 
-**Competing Interest Statement:** Authors declare that they have no competing interests.
+<!-- **Competing Interest Statement:** Authors declare that they have no competing interests. -->
 
-Classification: Biological Science, Environmental Sciences
+<!-- Thomas Frölicher -->
+<!-- Horizon 2020 Framework Programme 820989 -->
+<!-- Swiss National Science Foundation PP00P2_170687 -->
+<!-- Swiss National Supercomputing Centre -->
 
-This PDF file includes:
-Main Text
-Figures 1 to 5
+<!-- Alessandro Tagliabue -->
+<!-- Horizon 2020 Framework Programme  724289 -->
+
+<!-- Juliano Palacios-Abrantes -->
+<!-- Natural Sciences and Engineering Research Council of Canada  Discovery Grant -->
+<!-- Social Sciences and Humanities Research Council of Canada Partnership Grant -->
+
+<!-- William Cheung -->
+<!-- Natural Sciences and Engineering Research Council of Canada - Discovery Grant -->
+<!-- Social Sciences and Humanities Research Council of Canada - Partnership Grant -->
+
+<!-- U. Rashid Sumaila -->
+<!-- Social Sciences and Humanities Research Council of Canada --> <!-- Partnership Grant -->
+
+<!-- Colette C.C. Wabnitz -->
+<!-- David and Lucile Packard Foundation 2019-68336 -->
+<!-- Gordon and Betty Moore Foundation GBMF5668.02 -->
+<!-- Walton Family Foundation 2018-1371 -->
+
+\linenumbers
+
+# Abstract
+
+Climate change is shifting the distribution of shared fish stocks between neighboring countries’ Exclusive Economic Zones (EEZ) and the high seas.
+The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.
+Coupling a large ensemble simulation of an Earth system model to a species distribution model, we show that by 2030, 23% of transboundary stocks would have shifted and 78% of the world EEZs will experience at least one shifting stock under a high emission climate change scenario.
+By the end of this century, 81% of EEZ waters will see at least one shifting stock with a total of 45% of stocks shifting globally, under a high emissions scenario.
+Importantly, many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts showing early, and sometimes past shifts.
+Existing fisheries agreements need to be assessed for their capacity in addressing transboundary shifts, and strengthened where necessary to limit conflict over these fish stocks while new agreements are urged to considere this problematic in order to be resilient to global change.
+
+<!-- # Significant statement {-} -->
+
+<!-- Marine fish stocks are shifting their distribution in response to climate change.
+This is of particular importance for the management of stocks shared between neighboring Exclusive Economic Zones, also known as transboundary stocks.
+Here, we find that imminent climate change-induced shifts in transboundary fish stocks are challenging international management and that urging fisheries agreements to be assessed for their capacity in addressing transboundary shifts, and strengthened where necessary to limit conflict over these fish stocks.
+-->
 
@@
 
-# Abstract {-}
-
-<!-- An opening sentence that states the question/problem addressed by the research AND Enough background content to give context to the study AND A brief statement of primary results AND A short concluding sentence.
--->
-
-Climate change is shifting the distribution of shared fish stocks between neighboring countries’ Exclusive Economic Zones (EEZ) and the high seas.
-The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.
-Coupling a large ensemble simulation of an Earth system model to a species distribution model, we show that by 2030 23% of transboundary stocks would have shifted and 78% of the world will experience at least one shifting stock.
-By 2100, 81% of EEZ waters will see at least one shifting stock with a total of 45% of stocks shifting globally, under a high emissions scenario.
-Importantly, many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts.
-Existing fisheries agreements need to be assessed for their capacity in addressing transboundary shifts, and strengthened where necessary to limit conflict over these fish stocks.
-
-# Significant statement {-}
-
-Marine fish stocks are shifting their distribution in response to climate change.
-This is of particular importance for the management of stocks shared between neighboring Exclusive Economic Zones, also known as transboundary stocks.
-Here, we find that imminent climate change-induced shifts in transboundary fish stocks are challenging international management and that urging fisheries agreements to be assessed for their capacity in addressing transboundary shifts, and strengthened where necessary to limit conflict over these fish stocks.
-
-\clearpage
-
-# Main text {-}
+# Introduction
 
@@
 
-```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone (EEZ) to another.
+```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.pos="H", fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone (EEZ) to another.
 The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring EEZ sharing the stock (Top panel).
@@
 
-\clearpage
-
 <!-- # Methods -->
 
-We coupled a mechanistic population dynamic model with a ten-ensemble member simulation of an Earth System Model to project the future global distribution of transboundary stocks under climate change (Methods).
+<!-- We coupled a mechanistic population dynamic model with a ten-ensemble member simulation of an Earth System Model to project the future global distribution of transboundary stocks under climate change (Methods).
+-->
 
@@
 
-# Results {-}
-
-### Identifying range shifts in the shared distribution of transboundary stocks {-}
-
-Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
-\@ref(fig:toeFigure1)A), using a two s.d.
-threshold (i.e., representing a probability of 95% that the stock has shifted).
-This corresponds to 45% of the studied stocks (Fig.
-\@ref(fig:toeFigure1)C).
-Using a less conservative threshold (i.e., one s.d.
-representing a probability of 67% that the stock has shifted) results in 33% (n = 5,745) more shifting stocks, that is, 67% of all studied stocks (Figs.
-\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
-For both cases, the first shift is modelled to have occurred back in 2006.
-The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be between 2029 $\pm$ 27 years (one s.d.
-threshold) and 2036 $\pm$ 28 years (two s.d.
-threshold).
-Furthermore, between 81% and 83% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 77% to 80% of EEZs recording a distributional stock shift by the UN 2030 Agenda’s deadline (Fig.
-\@ref(fig:toeFigure1)C).
-
-```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary shared stock distribution.
-(A) Number of stocks (cumulative) whose distribution centroids that have shifted beyond natural variations over time (from 2006 to >2100), (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one stock shifting beyond natural variability.
-Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change in range shift (Methods), respectively."}
-
-img <- readPNG("../Figures/Fig2.png")
-grid.raster(img)
-
-```
-
-The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; using the 2 s.d.
-threshold; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig.
-\@ref(fig:toeFigureS1)) and EEZs' geographic regions (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; using the 2 s.d.
-threshold; Table S3 and Fig.
-2A).
-Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2) and \@ref(tab:toeTableS3) than almost any other region (Figs.
-\@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS1) and \@ref(fig:toeFigureS2)).
-In contrast, EEZs and stocks located in temperate regions, like northern Europe and eastern Asia, are projected to experience much later range shifts.
-Stocks in a number of EEZs including Brazil and New Zealand’s EEZs are projected to experience no range shifts beyond natural variability by the end of this century (Fig.
-\@ref(fig:toeFigure2)A).
-
-On average, shifting transboundary stocks represented 27% to 23% (1 and 2 s.d.
-thresholds) of annual revenue from fisheries targeting transboundary stocks across the world's EEZs between 2004 and 2010 (Fig.
-\@ref(fig:toeFigure2)A).
-However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
-Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).
-
-```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks.
-A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
-Exclusive Economic Zone polygons display average year of range shifts within them.
-Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cool colors represent a late shift/low fishing revenue.
-EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
-B) Year of shared distribution shifts by UN sub-regions.
-Points color coded by the median year of range shift.
-Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda.
-N = North, S = South, W = West and E = East.
-Ltn.
-Ame.
-and the Car.
-= Latin America and the Caribbean.
-Aus and New Z.
-= Australia and New Zealand."}
-
-img <- readPNG("../Figures/Fig3.png")
-grid.raster(img)
-
-```
-
-<!-- ![toeFigure 2.
-Year of change in the shared distribution of 4,119 transboundary stocks.
-A) Land polygon shows the contribution of shifting stocks to the country’s total fishing revenue from transboundary stocks.
-Exclusive Economic Zone polygons represent average range shifts per EEZ.
-EEZs with no time of emergence between 2006 and 2100 are represented in aqua color.
-B) Year of shared distribution shifts by region according to United Nations sub-regions.
-Points represent the median year of range shift.
-Horizontal dashed line represents the 2030 Agenda.
-N = North, S = South, W = West and E = East.
-Ltn.
-Ame.
-and the Car.
-= Latin America and the Caribbean.
-Aus and New Z.
-= Australia and New Zealand](../Figures/Fig3.png) -->
-
-We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs for both thresholds (one and two s.d.).
-Within five years of the UN 2030 Agenda timeline, 979 transboundary stocks accounting for USD 5.3 billion in fishing revenue are projected to shift within 76 EEZs considering a two s.d.
-threshold (Fig.
-\@ref(fig:toeFigure3); above Fisheries >75th percentile).
-In this case, 321 stocks shifting in just eleven countries account for 4.2 billion USD (Fig.
-\@ref(fig:toeFigure3)).
-The number of transboundary stocks shifting from historical variability was highest in the EEZs of Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66) (Fig.
-\@ref(fig:toeFigure3)).
-The values are substantially higher when considering a more relaxed threshold.
-In this case, 2,136 stocks accounting for USD 24 billion in fishing revenue are projected to shift within five years of the UN 2030 Agenda timeline in 134 EEZs.
-
-```{r toeFigure3, eval = T,  echo = F, fig.width=8, fig.height=6, fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks and corresponding fishing revenue per country.
-Countries are coded according to different colored symbols representative of 5 distinct regions.
-The horizontal line separates countries above and below the 75th percentile of annual fishing revenue.
-Showing country names for the eleven nations above the 75th percentile and projected to have an average shift across transboundary stocks within its EEZ within 5 years of the 2030 Agenda deadline."}
-
-img <- readPNG("../Figures/Fig4.png")
-grid.raster(img)
-
-```
-
-### Shifting intensity of transboundary stocks {-}
-
-<!-- We based the transboundary index on a stock’s centroid defined as the average grid cells that contain the majority of a stock’s abundance (see Methods - Calculating an index of transboundary range shift).
-Results from the sensitivity analysis suggest that our method is robust to the abundance level (e.g., the top 95%) of the stock (Fig.
-\@ref(fig:toeFigureS3)).
-In addition, we set an arbitrary threshold of 2 s.d.
-(95% confidence) to determine the time of emergence of the transboundary index.
-Sensitivity tests to a lower threshold of 1 s.d.
-(for a 67% confidence) resulted in a global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27, a difference of about 7 years (Fig.
-\@ref(fig:toeFigureS4)).
-In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks.
-SENT TO METHODS!!!!
--->
-
-We estimated the intensity of the distributional shift of transboundary stocks in terms of changes in catch proportion by 2030 (2020-2040) and 2050 (2040-2060), relative to the recent past (1951-2005) (see Methods).
-Projections show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks will change beyond the historic variability experienced within an EEZ’s (Fig.
-\@ref(fig:toeFigure4)A).
-Moreover, 85% (n = 239) of the world’s EEZs will have experienced changes in catch proportion of transboundary stocks by 2030.
-By mid-21st century, the number of EEZs with changing stocks, as well as the number of stocks experiencing changes in catch proportion and the magnitude of that change will increase (Figs.
-\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).
-The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
-\@ref(fig:toeFigure4)A).
-Along the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, shifts in stocks are expected to benefit poleward EEZs.
-However, shifts along the coasts of Pacific Central America and West Africa occur in an equatorial direction.
-
-```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock proportion of neighboring Exclusive Economic Zones by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors.
-Panel B highlights changes for the Caribbean and Northwestern Europe."}
-
-img <- readPNG("../Figures/Fig5.png")
-grid.raster(img)
-
-```
-
-<!-- ![Figure 1.
-Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors (green to yellow).
-Panel B zooms into specific areas shown within the gray dashed line boundaries in A.](../Figures/Fig1.png) -->
-
-<!-- The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
-\@ref(fig:toeFigure4)A).
-Along the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the proportion of the stock is expected to follow a projected poleward shift.
-However, shifts in the coasts of Pacific Central America and West Africa occur along an equatorial direction.
-Here, transboundary species follow specific oceanographic regions (*sensu* Longhurst) like upwelling zones (e.g., Humboldt current) and local patterns of oxygen availability, in addition to temperature gradients.
-Regions where EEZs are relatively small and have multiple borders (e.g., Caribbean or Northwestern Europe) are expected to have a particularly complex exchange of stock share ratio with no clear established pattern (Fig.
-\@ref(fig:toeFigure4)B).
--->
-
-# Discussion {-}
-
-<!-- In this paper, we investigate the timing and intensity associated with changes in the share distribution of transboundary stocks between countries.
-Our findings highlight an early distributional shift from historical variations for most transboundary stocks in the world.
-Such early shifts point to the urgency of addressing the challenges that climate change presents for international fisheries management.
-Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries that would benefit from immediate policy responses to support sustainability into the future.
-Finally, we show that for many countries changes in the distribution of transboundary stocks are likely to undermine joint resource management between countries [@Bailey:2010ga; @Munro:1979dg; @PalaciosAbrantes:2020vt; @Bailey:2016eh].
-Thus, our findings emphasize recent calls for the urgent adoption of adaptation measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @CisnerosMata:2019ks; @Oremus:2020id; @Free:2019jq].
--->
-
-### High present-day climate risk for transboundary fisheries {-}
-
-Our findings highlighting early shifts in the shared distribution of transboundary stocks concur with previous studies that have detected changes in marine catch composition and attributed these to climate change [e.g., @Cheung:2013gk; @Frainer:2017ji; @Last:2011fd].
-For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US), in response to climatic, and associated oceanographic and ecological changes [@Zeidberg:2007gs].
-A new fishery targeting Humboldt squid quickly developed on the heels of the species’ range expansion [@Pinsky:2014by].
-In the northeast Atlantic, fisheries on Atlantic mackerel (*Scomber scombrus*) are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
-However, a range expansion of Atlantic mackerel due to environmental variation into Icelandic waters in 2007 resulted in Iceland capturing 6% of the fishery’s total allowable catch and a further 18% in 2008, without consultation with NEAFC, threatening the sustainability of the stock [@Spijkers:2017ij].
-These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij].
-Other documented cases of early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].
-Despite these documented cases, broadly speaking, there is a lack of information and data regarding which and how many transboundary stocks are shifting, where they are shifting to and whether they are jointly managed.
-This is particularly important for international fisheries management as modeling exercises project that climate change will continue to change the distribution of transboundary stocks to the point that some tropical EEZs will lose them completely [@Oremus:2020id] while other EEZs, mainly in higher latitudes, will stand to win new stocks [@Pinsky:2018cb].
-
-### Climate risk hotspots for transboundary fisheries management {-}
-
-Our results identify regional “hotspots” of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figs.
-\@ref(fig:toeFigure2)A and \@ref(fig:toeFigure4)A).
-Regions such as the Caribbean are characterized by high levels of warming relative to historical natural variability [@IPCC:2019tn; @Hawkins:2012hc] and species’ high vulnerability to warming waters [@IPCC:2019tn].
-Moreover, they represent relatively low area EEZs that border multiple countries.
-Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei].
-Thus, coordinating the management of shifting transboundary stocks for countries in these regions will be particularly challenging.
-
-Most management plans that are not designed or prepared to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
-Focusing on identified “hotspots” will help anticipate any potential increases in and averting fisheries conflicts in coming years.
-Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments (including non-monetary arrangements) [@Tunca:2019if; @Miller:2013iv], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@PalaciosAbrantes:2020vt; @Pinsky:2018cb].
-Some examples of current treaties that have adapted some of these strategies include: the Pacific Salmon Treaty between Canada and the U.S.
-to manage Pacific salmon (Oncorhynchus sp.) and which has a conservation fund functioning as a side payment [@Miller:2013iv], the International Pacific Halibut Commission between Canada and the United States, which as the name implies manages Pacific halibut (*Hippoglossus stenolepis*) and allocates quota based on the stock’s yearly distribution [@PalaciosAbrantes:2020vt]; and the Parties to the Nauru Agreement (PNA) an Oceania subregional agreement among 8 Pacific Island countries and Tokelau, which collectively manage the largest tuna fishery in the world, controlling around 50% of the global supply of skipjack tuna (*Katsuwonus pelamis*).
-The PNA manage fishing effort under a so-called Vessel Day Scheme, an adaptive and equitable system which accounts for shifts in stocks and catches as a result of climate variability [@Bahri:2021wz].
-Quota allocation methods based on a stock’s current distribution and/or according to fixed-historical proportions like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-
-Even management plans that take into consideration such strategies might not be fully prepared for the consequences of shifting transboundary stocks [@Pinsky:2018cb; @Koubrak:2020hw; @Engler:2020bg].
-For example, species’ shifts within the PNA area will likely also have to deal with stocks expanding to new jurisdictions, issue that the NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
-Moreover, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery they have historically been entitled to.
-Finally, changes in policy can be sluggish (e.g., it took the PST about 10 years to agree on a conservation fund) compared to the timeframe over which species shift [@Pinsky:2012kq].
-This is specifically true as in many cases we lack standardized data to track shifts of species across international jurisdictions [@Maureaud:2021gj] as these might already be happening.
-Addressing shifting transboundary stocks is also challenged by an overall lack of knowledge of how many international agreements exist or what transboundary stocks are jointly managed.
-Recent efforts looking at 127 international fisheries plans found that they all lacked direct actions to address topics of climate change or species range shifts across jurisdictions [@Oremus:2020id].
-Having a better understanding of what stocks are managed as transboundary and under what rules would be an important complement to our study and an important step towards the sustainability of international stocks.
-[@Oremus:2020id].
-
-### Potential drivers of shift and managing an uncertain future {-}
-
-The early projected range shifts of large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability.
-For example, sea surface temperature (SST) is projected to increase beyond historical variability by 2030 in 50 to 70% of the global ocean [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
-Results from multiple marine ecosystem models show that SST and primary production (NPP) are the main indicators of species distributional changes across ocean basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
-Specifically, in the species distribution model used here (see methods), SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems, while NPP drives temperate regions’ results.
-The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics, where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2019tn], while also explaining the later, and sometimes non-existent, shifts at higher latitudes (Fig.
-\@ref(fig:toeFigure2)B).
-Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and *ii*) model structure [@Frolicher:2016dg].
-First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.Thus, any mitigation efforts could result in a substantial delay in distribution shifts as environmental signals are responsive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].
-Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
-This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is often larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
-While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available [@Deser:2020bj] and cannot address limitations of process-level understanding.
-It is unlikely that these complex uncertainties will adequately be addressed in the coming decade.
-This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, and the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf].
-These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh].
-Overall, multiple upper trophic level models present overall agreement in terms of directional change of fish biomass, but variability in the magnitude of that change [@Lotze:2019ce].
-Further research that reduces uncertainty in the NPP response at the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to smaller structural uncertainty of fish and fisheries models [@BryndumBuchholz:2019kh].
-Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks.
-While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca].
-Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identify different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
-Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
-
-<!-- # Conclusion {-} -->
-
-The global community has set the ambitious goal of managing all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear benefits for several other societal goals [@UnitedNations2018; @Singh:2017ds].
-Developing anticipatory policies to deal with shifting transboundary stocks is key to achieving the SDGs and ensuring effective governance of the world’s fisheries [@Pinsky:2018cb; @Oremus:2020id; @Pecl:2017hu].
-Here, we developed an approach to inform the sustainable management and governance of transboundary fisheries in a changing world [@Link:2010ei].
-First, we identified the transboundary stocks that would likely see shifts in their shared distribution compared to their historical average and the year in which such shifts would occur.
-Second, we estimated the intensity of such change.
-While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for shifting transboundary stocks [@Pinsky:2018cb; @PalaciosAbrantes:2020tv].
-Our findings emphasize recent calls for the urgent adoption of measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @Oremus:2020id].
-
-# Materials and Methods {-}
+# Materials and Methods
 
@@
 Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh].
-The habitat preference of each species was determined following FishBase (http://www.fishbase.org) for fish species and SeaLifeBase (http://www.fishbase.org) for invertebrates (Table S\@ref(tab:toeTableS1)).
+The habitat preference of each species was determined following FishBase (http://www.fishbase.org) for fish species and SeaLifeBase (http://www.fishbase.org) for invertebrates (Table \@ref(tab:toeTableS1)).
 For each stock and EEZ, we used Sea Around Us data to estimate catch and fishing revenue from fishing activities.
@@
 
-# Acknowledgments {-}
-We would like to acknowledge inputs from T.
-Clarke and L.
-Kwong on previous iterations of the manuscript and figures.
-A.T.
-acknowledges  funding from the Horizon 2020 Framework Programme (724289).
-C.C.C.W acknowledges funding from the David and Lucile Packard Foundation (2019-68336), the Gordon and Betty Moore Foundation (GBMF5668.02), and the Walton Family Foundation (2018-1371).
-J.P.A and W.W.L.C.
-acknowledge funding from the Social Sciences and Humanities Research Council of Canada, Partnership Grant and the Natural Sciences and Engineering Research Council of Canada, Partnership Grant.
-T.F.
-acknowledges funding from the Horizon 2020 Framework Programme (820989), the Swiss National Science Foundation (PP00P2_170687), and the Swiss National Supercomputing Centre.
-U.R.S.
-acknowledges funding from the Social Sciences and Humanities Research Council of Canada, Partnership Grant.
+# Results
 
-# References {-}
+## Identifying range shifts in the shared distribution of transboundary stocks {-}
 
-<div id="refs"></div>
-
-# Figures and Tables
-
-```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone (EEZ) to another.
-The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring EEZ sharing the stock (Top panel).
-Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel).
-Only showing a subset of ensemble members for clarity."}
-
-img <- readPNG("../Figures/Fig1.png")
-grid.raster(img)
-
-```
-
-\clearpage
+Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig.
+\@ref(fig:toeFigure1)A), using a two s.d.
+threshold (i.e., representing a probability of 95% that the stock has shifted).
+This corresponds to 45% of the studied stocks (Fig.
+\@ref(fig:toeFigure1)C).
+Using a less conservative threshold (i.e., one s.d.
+representing a probability of 67% that the stock has shifted) results in 33% (n = 5,745) more shifting stocks, that is, 67% of all studied stocks (Figs.
+\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
+For both cases, the first shift is modelled to have occurred back in 2006.
+The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be between 2029 $\pm$ 27 years (one s.d.
+threshold) and 2036 $\pm$ 28 years (two s.d.
+threshold).
+Furthermore, between 81% and 83% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 77% to 80% of EEZs recording a distributional stock shift by the UN 2030 Agenda’s deadline (Fig.
+\@ref(fig:toeFigure1)C).
 
@@
 
-\clearpage
+The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; using the 2 s.d.
+threshold; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig.
+\@ref(fig:toeFigureS1)) and EEZs' geographic regions (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; using the 2 s.d.
+threshold; Table S3 and Fig.
+2A).
+Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2) and \@ref(tab:toeTableS3) than almost any other region (Figs.
+\@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS1) and \@ref(fig:toeFigureS2)).
+In contrast, EEZs and stocks located in temperate regions, like northern Europe and eastern Asia, are projected to experience much later range shifts.
+Stocks in a number of EEZs including Brazil and New Zealand’s EEZs are projected to experience no range shifts beyond natural variability by the end of this century (Fig.
+\@ref(fig:toeFigure2)A).
+
+<!-- ![toeFigure 2.
+Year of change in the shared distribution of 4,119 transboundary stocks.
+A) Land polygon shows the contribution of shifting stocks to the country’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygons represent average range shifts per EEZ.
+EEZs with no time of emergence between 2006 and 2100 are represented in aqua color.
+B) Year of shared distribution shifts by region according to United Nations sub-regions.
+Points represent the median year of range shift.
+Horizontal dashed line represents the 2030 Agenda.
+N = North, S = South, W = West and E = East.
+Ltn.
+Ame.
+and the Car.
+= Latin America and the Caribbean.
+Aus and New Z.
+= Australia and New Zealand](../Figures/Fig3.png) -->
 
@@
 
-\clearpage
+On average, shifting transboundary stocks represented 27% to 23% (1 and 2 s.d.
+thresholds) of annual revenue from fisheries targeting transboundary stocks across the world's EEZs between 2004 and 2010 (Fig.
+\@ref(fig:toeFigure2)A).
+However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands).
+Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).
+
+We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs for both thresholds (one and two s.d.).
+Within five years of the UN 2030 Agenda timeline, 979 transboundary stocks accounting for USD 5.3 billion in fishing revenue are projected to shift within 76 EEZs considering a two s.d.
+threshold (Fig.
+\@ref(fig:toeFigure3); above Fisheries >75th percentile).
+In this case, 321 stocks shifting in just eleven countries account for 4.2 billion USD (Fig.
+\@ref(fig:toeFigure3)).
+The number of transboundary stocks shifting from historical variability was highest in the EEZs of Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66) (Fig.
+\@ref(fig:toeFigure3)).
+The values are substantially higher when considering a more relaxed threshold.
+In this case, 2,136 stocks accounting for USD 24 billion in fishing revenue are projected to shift within five years of the UN 2030 Agenda timeline in 134 EEZs.
 
@@
 
-\clearpage
+## Shifting intensity of transboundary stocks
+
+<!-- We based the transboundary index on a stock’s centroid defined as the average grid cells that contain the majority of a stock’s abundance (see Methods - Calculating an index of transboundary range shift).
+Results from the sensitivity analysis suggest that our method is robust to the abundance level (e.g., the top 95%) of the stock (Fig.
+\@ref(fig:toeFigureS3)).
+In addition, we set an arbitrary threshold of 2 s.d.
+(95% confidence) to determine the time of emergence of the transboundary index.
+Sensitivity tests to a lower threshold of 1 s.d.
+(for a 67% confidence) resulted in a global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27, a difference of about 7 years (Fig.
+\@ref(fig:toeFigureS4)).
+In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks.
+SENT TO METHODS!!!!
+-->
+
+We estimated the intensity of the distributional shift of transboundary stocks in terms of changes in catch proportion by 2030 (2020-2040) and 2050 (2040-2060), relative to the recent past (1951-2005) (see Methods).
+Projections show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks will change beyond the historic variability experienced within an EEZ’s (Fig.
+\@ref(fig:toeFigure4)A).
+Moreover, 85% (n = 239) of the world’s EEZs will have experienced changes in catch proportion of transboundary stocks by 2030.
+By mid-21st century, the number of EEZs with changing stocks, as well as the number of stocks experiencing changes in catch proportion and the magnitude of that change will increase (Figs.
+\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).
+The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
+\@ref(fig:toeFigure4)A).
+Along the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, shifts in stocks are expected to benefit poleward EEZs.
+However, shifts along the coasts of Pacific Central America and West Africa occur in an equatorial direction.
 
@@
 ```
+
+<!-- ![Figure 1.
+Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors (green to yellow).
+Panel B zooms into specific areas shown within the gray dashed line boundaries in A.](../Figures/Fig1.png) -->
+
+<!-- The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
+\@ref(fig:toeFigure4)A).
+Along the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, the proportion of the stock is expected to follow a projected poleward shift.
+However, shifts in the coasts of Pacific Central America and West Africa occur along an equatorial direction.
+Here, transboundary species follow specific oceanographic regions (*sensu* Longhurst) like upwelling zones (e.g., Humboldt current) and local patterns of oxygen availability, in addition to temperature gradients.
+Regions where EEZs are relatively small and have multiple borders (e.g., Caribbean or Northwestern Europe) are expected to have a particularly complex exchange of stock share ratio with no clear established pattern (Fig.
+\@ref(fig:toeFigure4)B).
+-->
+
+# Discussion
+
+<!-- In this paper, we investigate the timing and intensity associated with changes in the share distribution of transboundary stocks between countries.
+Our findings highlight an early distributional shift from historical variations for most transboundary stocks in the world.
+Such early shifts point to the urgency of addressing the challenges that climate change presents for international fisheries management.
+Moreover, we identify regional patterns of high climate risk to important fish stocks and associated fisheries that would benefit from immediate policy responses to support sustainability into the future.
+Finally, we show that for many countries changes in the distribution of transboundary stocks are likely to undermine joint resource management between countries [@Bailey:2010ga; @Munro:1979dg; @PalaciosAbrantes:2020vt; @Bailey:2016eh].
+Thus, our findings emphasize recent calls for the urgent adoption of adaptation measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @CisnerosMata:2019ks; @Oremus:2020id; @Free:2019jq].
+-->
+
+## High present-day climate risk for transboundary fisheries
+
+Our findings highlighting early shifts in the shared distribution of transboundary stocks concur with previous studies that have detected changes in marine catch composition and attributed these to climate change [e.g., @Cheung:2013gk; @Frainer:2017ji; @Last:2011fd].
+For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US), in response to climatic, and associated oceanographic and ecological changes [@Zeidberg:2007gs].
+A new fishery targeting Humboldt squid quickly developed on the heels of the species’ range expansion [@Pinsky:2014by].
+In the northeast Atlantic, fisheries on Atlantic mackerel (*Scomber scombrus*) are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC).
+However, a range expansion of Atlantic mackerel due to environmental variation into Icelandic waters in 2007 resulted in Iceland capturing 6% of the fishery’s total allowable catch and a further 18% in 2008, without consultation with NEAFC, threatening the sustainability of the stock [@Spijkers:2017ij].
+These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij].
+Other documented cases of early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].
+Despite these documented cases, broadly speaking, there is a lack of information and data regarding which and how many transboundary stocks are shifting, where they are shifting to and whether they are jointly managed.
+This is particularly important for international fisheries management as modeling exercises project that climate change will continue to change the distribution of transboundary stocks to the point that some tropical EEZs will lose them completely [@Oremus:2020id] while other EEZs, mainly in higher latitudes, will stand to win new stocks [@Pinsky:2018cb].
+
+## Climate risk hotspots for transboundary fisheries management
+
+Our results identify regional “hotspots” of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figs.
+\@ref(fig:toeFigure2)A and \@ref(fig:toeFigure4)A).
+Regions such as the Caribbean are characterized by high levels of warming relative to historical natural variability [@IPCC:2019tn; @Hawkins:2012hc] and species’ high vulnerability to warming waters [@IPCC:2019tn].
+Moreover, they represent relatively low area EEZs that border multiple countries.
+Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei].
+Thus, coordinating the management of shifting transboundary stocks for countries in these regions will be particularly challenging.
+
+Most management plans that are not designed or prepared to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp].
+Focusing on identified “hotspots” will help anticipate any potential increases in and averting fisheries conflicts in coming years.
+Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments (including non-monetary arrangements) [@Tunca:2019if; @Miller:2013iv], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@PalaciosAbrantes:2020vt; @Pinsky:2018cb].
+Some examples of current treaties that have adapted some of these strategies include: the Pacific Salmon Treaty between Canada and the U.S.
+to manage Pacific salmon (Oncorhynchus sp.) and which has a conservation fund functioning as a side payment [@Miller:2013iv], the International Pacific Halibut Commission between Canada and the United States, which as the name implies manages Pacific halibut (*Hippoglossus stenolepis*) and allocates quota based on the stock’s yearly distribution [@PalaciosAbrantes:2020vt]; and the Parties to the Nauru Agreement (PNA) an Oceania subregional agreement among 8 Pacific Island countries and Tokelau, which collectively manage the largest tuna fishery in the world, controlling around 50% of the global supply of skipjack tuna (*Katsuwonus pelamis*).
+The PNA manage fishing effort under a so-called Vessel Day Scheme, an adaptive and equitable system which accounts for shifts in stocks and catches as a result of climate variability [@Bahri:2021wz].
+Quota allocation methods based on a stock’s current distribution and/or according to fixed-historical proportions like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
+
+Even management plans that take into consideration such strategies might not be fully prepared for the consequences of shifting transboundary stocks [@Pinsky:2018cb; @Koubrak:2020hw; @Engler:2020bg].
+For example, species’ shifts within the PNA area will likely also have to deal with stocks expanding to new jurisdictions, issue that the NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb].
+Moreover, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery they have historically been entitled to.
+Finally, changes in policy can be sluggish (e.g., it took the PST about 10 years to agree on a conservation fund) compared to the timeframe over which species shift [@Pinsky:2012kq].
+This is specifically true as in many cases we lack standardized data to track shifts of species across international jurisdictions [@Maureaud:2021gj] as these might already be happening.
+Addressing shifting transboundary stocks is also challenged by an overall lack of knowledge of how many international agreements exist or what transboundary stocks are jointly managed.
+Recent efforts looking at 127 international fisheries plans found that they all lacked direct actions to address topics of climate change or species range shifts across jurisdictions [@Oremus:2020id].
+Having a better understanding of what stocks are managed as transboundary and under what rules would be an important complement to our study and an important step towards the sustainability of international stocks.
+[@Oremus:2020id].
+
+## Potential drivers of shift and managing an uncertain future
+
+The early projected range shifts of large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability.
+For example, sea surface temperature (SST) is projected to increase beyond historical variability by 2030 in 50 to 70% of the global ocean [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy].
+Results from multiple marine ecosystem models show that SST and primary production (NPP) are the main indicators of species distributional changes across ocean basins [@BryndumBuchholz:2019kh; @Lotze:2019ce].
+Specifically, in the species distribution model used here (see methods), SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems, while NPP drives temperate regions’ results.
+The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics, where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2019tn], while also explaining the later, and sometimes non-existent, shifts at higher latitudes (Fig.
+\@ref(fig:toeFigure2)B).
+Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and *ii*) model structure [@Frolicher:2016dg].
+First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.Thus, any mitigation efforts could result in a substantial delay in distribution shifts as environmental signals are responsive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].
+Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
+This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is often larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg].
+While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available [@Deser:2020bj] and cannot address limitations of process-level understanding.
+It is unlikely that these complex uncertainties will adequately be addressed in the coming decade.
+This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, and the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf].
+These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh].
+Overall, multiple upper trophic level models present overall agreement in terms of directional change of fish biomass, but variability in the magnitude of that change [@Lotze:2019ce].
+Further research that reduces uncertainty in the NPP response at the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to smaller structural uncertainty of fish and fisheries models [@BryndumBuchholz:2019kh].
+Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks.
+While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca].
+Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identify different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
+Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
+
+<!-- # Conclusion {-} -->
+
+The global community has set the ambitious goal of managing all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear benefits for several other societal goals [@UnitedNations2018; @Singh:2017ds].
+Developing anticipatory policies to deal with shifting transboundary stocks is key to achieving the SDGs and ensuring effective governance of the world’s fisheries [@Pinsky:2018cb; @Oremus:2020id; @Pecl:2017hu].
+Here, we developed an approach to inform the sustainable management and governance of transboundary fisheries in a changing world [@Link:2010ei].
+First, we identified the transboundary stocks that would likely see shifts in their shared distribution compared to their historical average and the year in which such shifts would occur.
+Second, we estimated the intensity of such change.
+While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, our results provide an important baseline on which to build when preparing ocean governance for shifting transboundary stocks [@Pinsky:2018cb; @PalaciosAbrantes:2020tv].
+Our findings emphasize recent calls for the urgent adoption of measures in support of more adaptive, flexible fisheries management and governance to support resilient fisheries and durable management systems [@Pinsky:2018cb; @Pinsky:2020jx; @Oremus:2020id].
+
+<!-- # Acknowledgments {-} -->
+<!-- We would like to acknowledge inputs from T.
+Clarke and L.
+Kwong on previous iterations of the manuscript and figures.
+A.T.
+acknowledges  funding from the Horizon 2020 Framework Programme (724289).
+C.C.C.W acknowledges funding from the David and Lucile Packard Foundation (2019-68336), the Gordon and Betty Moore Foundation (GBMF5668.02), and the Walton Family Foundation (2018-1371).
+J.P.A and W.W.L.C.
+acknowledge funding from the Social Sciences and Humanities Research Council of Canada, Partnership Grant and the Natural Sciences and Engineering Research Council of Canada, Partnership Grant.
+T.F.
+acknowledges funding from the Horizon 2020 Framework Programme (820989), the Swiss National Science Foundation (PP00P2_170687), and the Swiss National Supercomputing Centre.
+U.R.S.
+acknowledges funding from the Social Sciences and Humanities Research Council of Canada, Partnership Grant.
+-->
+
+# References
+
+<div id="refs"></div>
+
+<!-- # Supplemental material -->
+
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{table}{0}
+
+```{r supplements, child = 'Supplements.Rmd', eval = T}
+```

### 2021-08-24 Scripts/Supplements.Rmd
updates for submission to gcb

@@
 ---
-title: "Timing and magnitude of climate driven range shifts in transboundary fish stocks challenge their management"
-author: 'Juliano Palacios-Abrantes, Thomas L. Frölicher, Gabriel Reygondeau, U. Rashid Sumaila, Alessandro Tagliabue, Colette C.C. Wabnitz, and William W.L. Cheung'
+title: "Supplemental material for: Timing and magnitude of climate driven range shifts in transboundary fish stocks challenge their management"
+date: 'Juliano Palacios-Abrantes, Thomas L. Frölicher, Gabriel Reygondeau, U. Rashid Sumaila, Alessandro Tagliabue, Colette C.C. Wabnitz, and William W.L. Cheung'
 output: 
-  # bookdown::word_document2
   bookdown::pdf_document2:
-    toc : no
-    keep_tex: no
-    number_sections: yes
+    number_section: no
+    toc: no
+    latex_engine: xelatex
+reference_docx:  ../References/sci_template.docx
 header-includes:
@@
 - \captionsetup[table]{name = Table S}
+- \usepackage{lscape}
 ---
@@
 
-# Supplemental Figures
+# Figures
 
-S1
+```{r toeFigureS5, eval = T,  echo = F, fig.width = 10, fig.height= 8, fig.pos="H", fig.cap ="World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their estimated centroids (black points)."}
 
-```{r toeFigureS2, eval = T,  echo = F, fig.width=12, fig.height=7, fig.cap ="Year of change in the shared distribution of 5,745 transboundary stocks using a one standard deviation threshold.
+img <- readPNG("../Figures/FigS7.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigureS7, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine the stock’s centroid.
+Top Row: results in absolute numbers.
+Bottom row: results expressed as a percentage of the total emerging stocks and EEZs with emerging stock."}
+
+img <- readPNG("../Figures/FigS6.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigureS2, eval = T,  echo = F, fig.width=13, fig.height=13, fig.cap ="Year of change in the shared distribution of 5,745 transboundary stocks using a one standard deviation threshold.
 A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
@@
 
+img <- readPNG("../Figures/FigS1.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigureS8, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap ="Average distribution of transboundary stock’s year of change in their shared distribution.
+Left panel is the average year across stocks and right panel the standard deviation.
+Yellow represents a two standard deviation distance from the historical average and purple represents a one standard deviation distance from the historical average."}
+
 img <- readPNG("../Figures/FigS2.png")
@@
 
-S2
-
-```{r toeFigureS8, eval = T,  echo = F, fig.width=6, fig.height=6, fig.cap ="Average distribution of transboundary stock’s year of change in their shared distribution.
-Left panel is the average year across stocks and right panel the standard deviation.
-Yellow represents a one standard deviation distance from the historical average and purple represents a two standard deviation distance from the historical average."}
-
-img <- readPNG("../Figures/FigS7.png")
-grid.raster(img)
-
-```
-
-\clearpage
-
-S3
-
-```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
+```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=6,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
 Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
@@
 Solid line represents the median time of emergence per habitat preference."}
-
-img <- readPNG("../Figures/FigS1.png")
-grid.raster(img)
-
-```
-
-\clearpage
-
-S4
-
-```{r toeFigureS3, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Distribution of changes in transboundary stock proportion per stock and neighboring Exclusive Economic Zones.
-Yellow color represents changes by early, 2030 (2021-2040) and purple by mid 2050 (2041-60) 21st century relative to 1951-2005."}
 
@@
 
-S5
-
-```{r toeFigureS4, eval = T,  echo = F, fig.width=14, fig.height=7, fig.cap ="Changes in stock proportion of neighboring Exclusive Economic Zones by 2050 (2041-2060) relative to 1951-2005.
-Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors.
-Panel B highlights changes for the Caribbean and Northwestern Europe."}
+```{r toeFigureS3, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Distribution of changes in transboundary stock proportion per stock and neighboring Exclusive Economic Zones.
+Yellow color represents changes by early, 2030 (2021-2040) and purple by mid 2050 (2041-60) 21st century relative to 1951-2005."}
 
@@
 
-S6
-
-```{r toeFigureS7, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine the stock’s centroid.
-Top Row: results in absolute numbers.
-Bottom row: results expressed as a percentage of the total emerging stocks and EEZs with emerging stock."}
-
-img <- readPNG("../Figures/FigS6.png")
-grid.raster(img)
-
-```
-
-\clearpage
-
-S7
-
-```{r toeFigureS5, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their estimated centroids (black points)."}
+```{r toeFigureS4, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap ="Changes in stock proportion of neighboring Exclusive Economic Zones by 2050 (2041-2060) relative to 1951-2005.
+Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors.
+Panel B highlights changes for the Caribbean and Northwestern Europe."}
 
@@
 
-_**Tables**_
+# Tables
 
@@
 ) %>% 
-  write.csv(.,"habitat_preference.csv")
+  # write.csv(.,"habitat_preference.csv")
  kable(
         booktabs = T,
+        # longtable = T,
         caption = "Definition of each habitat preference according to FishBase"
-  ) #%>% 
-  # kable_styling(full_width = F,
-                # position = "center")
+  ) %>%
+  kable_styling(full_width = T,
+                position = "center")
 
@@
 
-```{r toeTableS3, eval = F, echo = F}
+```{r toeTableS3, eval = T, echo = F}
 
-read.csv("~/GitHub/TransEmergence/Tables/kw_habitat.csv") %>% 
+read.csv("../Tables/kw_habitat.csv") %>% 
   select(-statistic,
@@
         booktabs = T,
+        longtable = T,
         caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  ) #%>% 
-  # kable_styling(full_width = F,
-                # position = "center")
+  ) %>% 
+  kable_styling(#full_width = T,
+                position = "center")
 
@@
 
-```{r toeTableS2, eval = F, echo = F}
+\begin{landscape}
 
-read.csv("~/GitHub/TransEmergence/Tables/kw_regional.csv") %>% 
+```{r toeTableS2, eval = T, echo = F}
+
+read.csv("../Tables/kw_regional.csv") %>% 
   select(-statistic,
@@
         booktabs = T,
+        longtable = T,
         caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
-  ) #%>% 
-  # kable_styling(full_width = F,
-                # position = "center")
+  ) %>% 
+  kable_styling(#full_width = T,
+                # position = "center",
+                latex_options = c("repeat_header")
+                )
 
 ```
+
+\end{landscape}

### 2021-10-28 Scripts/Manuscript.Rmd
misc.

@@
 output: 
-  bookdown::pdf_document2:
+  bookdown::word_document2:
     number_section: no
@@
     latex_engine: xelatex
+always_allow_html: true
 reference_docx:  ../References/sci_template.docx

### 2021-11-01 Scripts/Manuscript.Rmd
Addressed some of comments

@@
 By the end of this century, 81% of EEZ waters will see at least one shifting stock with a total of 45% of stocks shifting globally, under a high emissions scenario.
-Importantly, many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts showing early, and sometimes past shifts.
-Existing fisheries agreements need to be assessed for their capacity in addressing transboundary shifts, and strengthened where necessary to limit conflict over these fish stocks while new agreements are urged to considere this problematic in order to be resilient to global change.
+**Many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts.
+These hotspots are characterized for showing early shifts in the distribution of important transboundary stocks.** Existing fisheries agreements need to be assessed for their capacity in addressing transboundary shifts, and strengthened where necessary to limit conflict over these fish stocks while new agreements are urged to **consider** this problematic in order to be resilient to global change.
 
@@
 The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt], impacting fish stocks, fisheries production and economies [@Sumaila:2019ew].
-These impacts are compromising our capacity to reach international sustainability goals under the United Nations (UN)  2030 Agenda for Sustainable Development such as Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
+These impacts are compromising our capacity to reach international sustainability goals under the United Nations (UN)  2030 Agenda for Sustainable Development such as **Goals 14 - life below water and 17 - partnerships for the Goals** [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu].
 The projected risks and impacts can be reduced by improving the effectiveness of current governance and fisheries management, including for species that cross international borders, also known as ‘shared stocks’ [@Pinsky:2018cb].
@@
 Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods).
-We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
+**We define a ‘stock’ unit as a species shared by neighboring EEZs.** Firthermore, we  developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
 The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.
@@
 The analyses detailed herein are based on 633 exploited marine transboundary species that account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-When a species was shared by a pair of neighboring EEZs it was considered a transboundary stock [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.
-Using the transboundary species Atlantic cod (*Gadus morhua*) as an example and this study’s definition of transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not as they do not have juxtaposing EEZs.
+**We define a (transboundary) "stock" unit when a species was shared by a pair of neighboring EEZs [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.** Using the transboundary species Atlantic cod (*Gadus morhua*) as an example and this study’s definition of transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not as they do not have juxtaposing EEZs.
 We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015; http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
@@
 Please see @Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd for a more detailed version.
-Importantly, the DBEM is able to project catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al.
-*in revision*].
+Importantly, the DBEM is able to project catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, @cheung_marine_2021].
 
@@
 
-where *T* is seawater temperature, *TPP* is the species temperature profile, *Ba* is bathymetry, $Min_{D}$ and $Max{D}$ are the species minimum and maximum depth limits, *Hab* is the proportion of area of the habitat type *j* relative to the total seawater area of the cell *i*, $Hab_{Assoc}$ is an habitat association index, *Sal* is the salinity class of cell *i* and $Sal_{Assoc}$ is a salinity class association index.
+where *T* is seawater temperature, *TPP* is the species temperature profile, *Ba* is bathymetry, $Min_{D}$ and $Max{D}$ are the species minimum and maximum depth limits, *Hab* is the proportion of area of the habitat type *j* relative to the total seawater area of the cell *i*, $Hab_{Assoc}$ is **a habitat** association index, *Sal* is the salinity class of cell *i* and $Sal_{Assoc}$ is a salinity class association index.
 Finally, for polar regions and species, *Ice* is sea ice extent, and $Ice_{P}$ is association to sea ice.
@@
 where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the specified percentile ($per^{th}$) of the projected transboundary stock abundance.
-To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, we included grid cells where the projected stock abundance within the neighboring EEZs sharing the stock was above the $95^{th}$ percentile.
-A sensitivity analysis using a subset of species (n = 34) for all EEZs  to examine the effects of different thresholds (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) on the calculated index value showed no apparent difference between percentiles (Fig.
+**Fish stocks are not evenly distributed across their entire range, thus we focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place.
+Thus, we estimated the centroid of a stock only including grid cells where the projected stock abundance within the neighboring EEZs sharing the stock was above the $95^{th}$ percentile.** A sensitivity analysis using a subset of species (n = 34) for all EEZs  to examine the effects of different thresholds (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) on the calculated index value showed no apparent difference between percentiles (Fig.
 \@ref(fig:toeFigureS7)).
@@
 
+We estimated the participation of the transboundary stocks that are shifting between neighboring EEZs relative to the total revenue from all transboundary stocks that each EEZ has.
 On average, shifting transboundary stocks represented 27% to 23% (1 and 2 s.d.

### 2021-11-01 Scripts/responses_rev_gcb.Rmd
Addressed some of comments

---
title: "Response to Reviewers"
output: pdf_document
---

# Edditor

Overall, the two reviewers agree that this work tackles an important, but collectively have raised a considerable number of points that will need to be addressed. Of most concern, is that it is not currently clear how this current MS differs from other recent papers that have established methods used on this paper and implications of climate change in the context of transboundary fisheries - including the authors' one works (Sumaila et al. 2020 Ecology and Society, Palacios-Abrantes et al. 2020 Ecology and Society, Pinsky et al. 2018 Science, and Palacios-Abrantes et al. 2020 Scientific Reports). It is important that GCB papers deliver the step change in knowledge that has not been realised elsewhere; therefore' we ask the authors to ensure the paper wholly demonstrates and articulates the transformation achieved in order to progress with the submission.

- **Response:**

# Reviewer 1

Comments to the Author
In this work, the authors analyze potential shifts in harvested marine species under climate change, in relation to their distribution across neighboring exclusive economic zones. They find that many countries will experience shifting species, and investigate how the number and intensity of transboundary shifts will vary by species type. Overall, I find the authors take an interesting approach to the analysis of transboundary fish populations under climate change, and the Discussion of potential management implications is strong, in my opinion. However, I think the manuscript needs some significant clarifications and edits before it can be published. In the space below, I outline my major comments, followed by more specific and/or minor comments, accompanied by line numbers from the draft pdf.

- **Response:** We are glad that the reviewer found our manuscript to be interesting. We have revised the manuscript paying special attention to clarifying the parts the reviewer flagged. We thank the reviewer for his/her comments as they have strengthen our paper. Please find below direct responses to the reviewer's comments.

Major comments:
I am concerned about the lack of a clear definition of “fish stock” in this manuscript, since it is absolutely central to the analysis. I cannot find where in the manuscript the definition of stocks is laid out. In most literature (and especially in fisheries science), stocks have a specific (if flexible) definition as populations of a species with some degree of demographic isolation from other stocks, such that the population dynamics of separate stocks may be considered independent. In this manuscript, it seems that instead, a transboundary stock is simply defined as any pair of EEZs that share a species distribution. I would recommend that the authors define the precise definition of fish stock explicitly within the manuscript, including how their definition is different from the definition in other, related literature.

- **Response:** Thank you for this comment. We acknowledge that we did not defined clearly pur concept of stock in the previous version of the manuscript. The definition of stock has now been included in the revised manuscript in both the introduction and the methods as follows;

  - Introduction Line XX, *We define a ‘stock’ unit as a species shared by neighboring EEZs.*

  - Methods Line XX, *We define a (transboundary) "stock" unit when a species was shared by a pair of neighboring EEZs [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks*

I have a number of questions and concerns about this approach to stock definition, including: how are neighboring stocks of the same species considered? If, for example, a species’ distribution exists within British Columbia, the United States, and Mexico, is this considered one stock for the purpose of the analyses in the paper (e.g. TI, ToE, SSR), or is this two stocks (one between Canada and the US and one between the US and Mexico)? How does the choice of stock definition influence the calculation of SSR? Given the same sample as above, can the US lose SSR of one stock of a species to Canada, while gaining SSR of the same species from Mexico?

- **Response:** We acknowledged that this was not fully clear in the previous version of the manuscript and have work towards improving clarity in the revise manuscript. For example, pacific sardine is a species present along the coast of Mexico, the United States (US) and, sometimes, British Columbia (BC) [@cisneros-montemayor_ecosystem-based_2020]. For the purpose of our study, we define two stocks: One that is shared by Mexico and the US and one shared by the US and BC. We have made this explicit in the following section of the methods, hopefully clarifying our definition;

    - Methods Line XX, *Let us consider the transboundary species Atlantic cod (*Gadus morhua*) as an example. Atlantic cod's distribution in the Northwest Atlantic ranges from the United States to Greenland. Based on the previous definition of a transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not share a stock as they do not have juxtaposing EEZs.*

    Thus, this definition will influence the calculation of SSR *between* neighboring EEZs (i.e., Mexico and the US SSR for sardine and US and BC SSR for sardine, independently one of another). As the reviewer suggests, in this scenario Mexico could be loosing sardine SSR to the US contiguous states (this can also be seen as the US gaining SSR from Mexico), while the US contiguous states looses sardine SSR to BC (again, this could also be seen as Canada winning SSR from the US contiguous states ). This effect is appreciable in the average SSR gains of Figure 5 where Mexico losses to the US contiguous states, who in turn looses to BC, who actually gains from both US continuous states and Alaska. This is not explain on the methods as follows **NOT SURE WE NEED THIS**;

    - Line XX, *While this means that a shift in the stocks could happened from the US to Canada and from Canada to Greenland, such shift could not (analytically) go directly from the US to Greenland*

    This way, our definition of a shared stock allows for stocks to go to or leave an EEZ from all borders. Continuing using Mexico as an example, Figure 5 shows how in average Mexico will loose SSR of the stocks it shares with the US up north probably do to a poleward shift on the Californian current species [@cisneros-montemayor_ecosystem-based_2020] but also to the south to Guatemala as stocks follow local environmental gradients [@Clarke:2020ep].

The above are just example situations. I do not mean to be nitpicky about definitions, but I am concerned that the definition of stocks used in this manuscript fundamentally determine the top-line results (e.g., number and proportion of shifting stocks, calculation of the transboundary index and time of emergence, loss or gain of SSR, etc. etc.).

  - **Response:** We understand the reviewer's concern and hope that our response have clarified this issue. However, it is almost certain that our definition of stock is influencing the results of our paper, just by tha nature of the analysis. While a more fine scale definition (e.g., biologically-defined populations of stock will probably arose other results, not limiting the species to, at least, an EEZ could result in an overestimation of our results (See Ovando *et al.* https://www.pnas.org/content/118/23/e2025958118).

Further, I am concerned about the averaging and weighting of species that is done across species in this analysis. As one example, why is the geographic centroid (used in the calculation of TI) not weighted by probability of occurrence? Fish stocks are not evenly distributed across their entire range, and unweighted centroids can severely skew the actual distribution of the species.

- **Response:** We have revised the manuscript to clarify that we did account for the heterogeneous distribution of fish stocks across their entire range. The geographic centroid of each stock is indeed weighted by the abundance of the stock. The DBEM uses the species probability of occurrence to estimate the biomass of the species at each 0.5 x 0.5 gridcell as indicated from lines XX to XX of the methods. The centroid of a transboundary stock was determined by the average latitude and longitude ($lon_{ts}$) of the grid cells holding the specified percentile ($per^{th}$) of the projected transboundary stock abundance. This is, we determined the centroid of the stock using only those regions where abundace is higher (i.e., using the probability of occurrence of the DBEM). It reads as follows:

  - Line, *Fish stocks are not evenly distributed across their entire range, thus we focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place. Thus, we estimated the centroid of a stock only including grid cells where the projected stock abundance within the neighboring EEZs sharing the stock was above the $95^{th}$ percentile.*

    We agree that and unweighted centroids can severely skew the actual distribution of the species and influence our results. We estimated the sensitivity of our results to the (arbitrarily) $95^{th}$ percentile threshold as described in lines XX and shown in Figure S6 of the supplements.

Weighting comes up again in the analyses that result in Figures 3, 4, and 5. Having one major, valuable species shift rapidly is a much bigger concern for many countries than having many small or low-value stocks shift slowly. These sorts of dynamics, while interesting, are difficult if not impossible to parse from the current Figures. For example, would the arrows look different in Figure 5 if the averages were weighted by each fish stock’s value?

**NOT SURE I UNDERSTAND THIS COMMENT**

- **Response (INCOMPLETE):** We were not clear in our message and have worked on clarifying all of these results. Specifically, in figure three we are showing the participation of the transboundary stocks that are shifting between neighbouring EEZs relative to the total revenue from all transboundary stocks that each EEZ has. So, for example, the transboundary stocks that are shifting in the Mexican Pacific coast to/from US and Guatemala represent about 15% of Mexico's total revenue from transboundary stocks. They average year of shift is ~2025. Indeed as the reviewer suggests, Figure 4 does not include any weight value, but rather absolute revenue amounts. The figure's objective is not only to inform the reader about the economic importance, in terms of absolute USD, of these shifting transboundary stocks (rather than their weighted value that is presented in Figure 3) but also highlight that these shifts are happening close to the UN 2030 Agenda. Finally, Figure 5 does indeed show weighted values. Here, we are showing the...

We have made the following modifications to the manuscript:

  - Line XX, *We estimated the participation of the transboundary stocks that are shifting between neighboring EEZs relative to the total revenue from all transboundary stocks that each EEZ has. On average, shifting transboundary stocks represented 27% to 23% (1 and 2 s.d. thresholds) of annual revenue from fisheries targeting transboundary stocks across the world's EEZs between 2004 and 2010 (Fig. \@ref(fig:toeFigure2)A).*

Specific comments:

Abstract:
Line 10: what does this line mean: “showing early, and sometimes past shifts”?

- **Response:** We have modify this sentence to now read:

  - Abstract Line XX; *Many countries that are highly dependent on fisheries for income, food and nutrition security, as well as livelihoods emerge as hotspots for transboundary shifts. These hotspots are characterized for showing early shifts in the distribution of important transboundary stocks.*

Introduction:
Be careful of verb tense. The Introduction mixes past with present tense. For example, the first two sentences starting at Line 57.

- **Response: (Incomplete)** Many thanks for this observation. We have revised the whole manuscript to be sure verb tense is maintained through the document.

Methods:

Line 78-79: Is this the definition of a stock? What if a species spans more than two EEZs? What if it is a straddling stock? Please be more explicit about what kinds of stocks are included (and not) in your analysis.

- **Response:** This is indeed the definition of our stock. Please refer to our responses to the main comment.

Line 79: where is “this study's definition of a transboundary stock?”

- **Response:** Please see reposes to the reviewer's main comments. We have defined a trnaboundary stock as follows;

  - Methods Line XX, *We define a (transboundary) "stock" unit when a species was shared by a pair of neighboring EEZs [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks*

Lines 220-226: How does the threat point analysis align with my concerns about the definition of stocks described above? Can an EEZ both lose and gain SSR that exceeds a threshold, but of the same species? If so, this seems misleading.

- **Response:** Thank you for your concern. In short the answer is yes, an EEZ could both lose and gain SSR of the same *species* but not of the same *stock.* This is because, according to our definition of stock, a species who spans multiple EEZs has multiple stocks (see response to the previus question and the main comment). let us use the Peruvian anchoveta (*Engraulis ringens*) as an example.

Peru has identified two main stocks of Peruvian anchoveta. one in the central/norther part of the EEZ and one in the south, shared with Chile. Although the northern stocks is not considered transboundary [@Cashion:2018cg; @PalaciosAbrantes:2020tv] and Equador hardly ever reports anchoveta landings (wwww.searaundus.org), in some few years, environmental variability has pushed the central/north stock towards the EEZ of Ecuador. In those years one could argue that, the Peruvian anchoveta SSR of Peru-Ecuador favors Ecuador.

At the same time, there are regional concerns that the Peruvian southern stock of anchoveta, shared with Chile, has been slowly shifting south into Chilean waters to the point that these two nations are now pushing towards cooperative management of such stock. In this case, the SSR of Peru-Chile, favors Chile.

This is only and example as our study is not capable of capturing this specific nuisance since we are not able to differentiate stocks within an EEZ nor more detailed, local, environmental factors (in this case, for example, the Humboldt current). However, we believe this

Results:

Lines 244 and 246- check your math here. The percentages seem off. For example, if 4119 stock is 45% of the total, how is 5745 stocks 33% more stocks? If, as stated in the Methods, there are 9132 stocks, then 5745 is ~63% (not 67%) of all stocks, and it is <20% more than 4119 stocks. Please check these numbers and clarify

- **Response:**

Fig. 3 - While I like the idea of Figure 3, I feel that there is too much conveyed in this figure. It is confusing to have two different scales (for land polygons and EEZs, respectively) on the same color ramp, even though they show completely different things. Moreover, see my comments on averaging above. Why is average year used for coloring EEZs, but median year used in Fig 3B?

- **Response:**

Year of range shift- is this the median emergence year for all stocks that experience a shift? A shift beyond the threat point? Does it include stocks that are transboundary but don’t shift? Please clarify how this median is calculated.

- **Response:**

Lines 261-266: I think this is an interesting result, and is lost in the averaging of the range shifts. It would be interesting to see a figure or Table with some of the biggest “winners and losers” in terms of high-value species that are shifting out of or into particular EEZs. Or, could there be a figure displaying the 10-20 biggest/highest-value stocks that are expected to shift? Which countries are expected to lose/gain these species?

- **Response:**

Fig. 4- This figure is a bit messy, and seems to be relating two dubiously-related statistics (that is, the projected average year of range shift with the current value of transboundary fisheries) by world region. I am just not sure what the main message of this figure is. If the point is to show the value of species that may shift over time, I would suggest a time series (by country and/or region), simply showing the total (present) value of transboundary stocks over time.

- **Response:**

Fig. 5- this figure is hard to read, and is another figure that may be misleading if not weighted by abundance or value. It is strange to “average” across a unitless ratio (SSR). For example, a country could “gain” a small share of an extremely high-value species, perhaps enough to offset economically a number of losses of smaller or less valuable stocks. However, in the figure, the “average” SSR for that country would be negative (arrow going out). The authors should distinguish (in the text and in this figure) different possible situations like this.

- **Response:**

Can you determine from your analyses where the biggest positive or negative changes might be in stock distributions, by nation/EEZ? In other words, are there nations with a greater number of stocks entering than exiting, and if so, where are they? This would be an interesting addition to the manuscript

- **Response:**

Discussion

The Discussion is well-organized and lays out the major implications of the work. I particularly commend the authors on their discussion of specific fisheries and cases where the transboundary issue is already a concern, as well as the description of examples of management approaches to deal with transboundary species. I do not have any major comments.

- **Response:** Many thanks for your kind words.

# Reviewer 2

This MS estimates both the extent and expected timing of range shifts of transboundary species due to climate change. Overall, the MS is well-written and methods are clearly explained (except Fig. 1 could be slightly modified- see comments below). The authors provided good examples and appropriately place their results into both a global sustainability framework as well as detail the potential application in regional management. From my perspective, this is a great contribution to the field of adaptation planning, and my comments below are to strengthen the connection to adaptation planning.

- **Response:**

Comments:

Line 12. Typo: considere>consider

- **Response:** Changed in the revised manuscript, thank you.

Line 22: After reading the paper—it’s clear that results are also relevant to the SDG 17: https://sdgs.un.org/goals/goal17

- **Response:** We agree with the reviewer and have included the SDG in the paper.

For example: Your results are the type of evidence that would be used by nations to develop adaptation plans, in hotspots, . See 17.16 Enhance the global partnership for sustainable development, complemented by multi-stakeholder partnerships that mobilize and share knowledge, expertise, technology and financial resources, to support the achievement of the sustainable development goals in all countries, in particular developing countries

17.16.1
Number of countries reporting progress in multi-stakeholder development effectiveness monitoring frameworks that support the achievement of the sustainable development goals

- **Response:**

Line 60: I suggest re-naming this Transboundary Shift and Emergence Index or Transboundary Shift Index. Something a little more descriptive. TI is so vague.

- **Response:**
Fig. 1: This is a wonderful schematic, my concerns: 1) stocks that do not occur as ovals but occur in patches or blobs instead-- the centroid may be too crude to detect movement of non-centric distributions- I don;t think your approach should change but re-cast as a proof of concept, researchers interested in applying the technique would be able to apply it at a regional level, and perhaps adapt it to patch distributions (separate centroids).
2) I don;t know what the empty red circle means in bottom panel, I can’t find it. do you mean the white area since the red dot?
3) what is diff btwn distributional and geographic centroid? (suggest using one term or other).
4) WHat if EEZ Ca moves at the same rate in the same direction as EEZ Cb. Is the TI only useful if one (more northerly) stock is static? What about adding and index of each Ca and EEZb distance to border? Given how great this MS is, I imagine a lot of thought went through what constitutes a “shift”, so If I have misunderstood index. OH! Now I think I get it, after reading equations 5 and 6, is it possible to include more details in Fig 1?

- **Response:**

Line 109- type- an habitat> a habitat

- **Response:** Changed in the revised manuscript, thank you.

Line 140- Nice real-world application of the DBEM and ensemble ocean models.

- **Response:** Thank you!

Line 180- Time of emergence is excellent -- will allow nations to triage which species have to be jointly managed. Bi-lateral adaptation plans and governance will be a priority for those stocks with earlier times of emergence – really useful information.

- **Response:** Thank you!

Line 185: Re ToE and noise. In my mind, there is no right answer, you may reduce the uncertainty but the natural variability contributing to that uncertainty is inherent in these projections. Instead, the various ToE estimates, produced using different assumptions, can be considered by decision-makers as risk in a risk assessment process. For example, assuming historically high natural variability, the ToE would be n years. Assuming lower natural variability, the ToE would be n years. In other words, having the array of options, some decision-makers may choose to adopt that risk while in other instances, they could choose to be precautionary and act sooner. This information could be used in scenario-management planning. See the following link for ideas.. https://www.fisheries.noaa.gov/resource/document/scenario-planning- introduction-fishery-managers

- **Response:**

Line 249- From an adaptation planning perspective, the sds mean that the average year of shifts are the same regardless of threshold== happening now or soon—again, information that could be used in risk assessment or scenario management planning.

- **Response:**

Line 256: See previous comment on SDG 17

- **Response:**

~Line 274 To re-iterate previous statement, this information can be used in a risk assessment framework/scenario planning to estimate risk/loss of revenue if adaptation is not developed cooperatively.

- **Response:**

Fig.5, can you make the vectors a different color? Currently they seem to be the same color as the low end of the changes in catch proportion scale

- **Response:**

Line 341- This last sentence will be of most interest to decision-makers. As Sumaila is a co-author, add a sentence to expand on the "how-to".

- **Response:**

Line 401-I really like this closing sentence and following paragraphs. Your work provides context and a starting point for regional application. The global scale study provides information to identify global hotspots, and this para discusses how estimates could be "down-scaled", so it is the approach or framework that is a useful starting point.

- **Response:**

### 2021-11-02 Scripts/Initial_Analysis.Rmd
Included the sf fix. and commented the weighted method of figure 3

@@
 my_lib(packages)
+
+# Fix new updates of sf package
+sf::sf_use_s2(use_s2 = FALSE)
 
@@
 
-##### Fig. 1. Accumulative plots
+##### Fig. 2. Accumulative plots
 
@@
   summarise(
-    mean_yr = mean(emerging_yr,na.rm=T),
+    mean_yr = median(emerging_yr,na.rm=T),
     sd_yr = sd(emerging_yr,na.rm=T),
@@
     # geom_hline(yintercept = c(-50,0,50,80)) +
-    my_ggtheme_m() +
+    my_ggtheme_m(leg_pos = "bottom") +
     theme(
@@
   # Weighted verion
+  # boxplot_data <- Eme_Res %>% 
+  #   # Get the order of boxes
+  #   filter(tresh == tresh_list[i],
+  #          !is.na(un_sub_region)) %>%
+  #   group_by(un_sub_region) %>%
+  #   summarise(n_eezs = length(unique(sf_eez_name))) %>%
+  #   # Get the rest of the data
+  #   left_join(Eme_Res) %>%
+  #   filter(tresh == tresh_list[i],
+  #          !is.na(un_sub_region)) %>%
+  #   # mutate(weight = )
+  #   mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
+  #                                 ifelse(str_detect(un_sub_region,"Australia"),"Aus. and New Z.",
+  #                                        ifelse(str_detect(un_sub_region,"Sub-Saharan"),"Sub-Saharan\nAfrica",
+  #                                               paste(un_sub_region)))
+  #   )#,
+  #   # For number labels
+  #   # un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
+  #   )
+
+  # Non-weighted verion
+
   boxplot_data <- Eme_Res %>%
@@
     group_by(un_sub_region) %>%
-    summarise(n_eezs = length(unique(sf_eez_name))) %>%
+    summarise(order = median(emerging_yr),
+              n_eezs = length(unique(sf_eez_name))) %>%
     # Get the rest of the data
@@
            !is.na(un_sub_region)) %>%
-    # mutate(weight = )
     mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
@@
                                                 paste(un_sub_region)))
-    )#,
+    ),
     # For number labels
-    # un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
-    )
-
-  # Non-weighted verion
-
-  # boxplot_data <- Eme_Res %>% 
-  #   # Get the order of boxes
-  #   filter(tresh == tresh_list[i],
-  #          !is.na(un_sub_region)) %>%
-  #   group_by(un_sub_region) %>% 
-  #   summarise(order = median(emerging_yr),
-  #             n_eezs = length(unique(sf_eez_name))) %>% 
-  #   # Get the rest of the data
-  #   left_join(Eme_Res) %>% 
-  #   filter(tresh == tresh_list[i],
-  #          !is.na(un_sub_region)) %>% 
-  #   mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
-  #                                 ifelse(str_detect(un_sub_region,"Australia"),"Aus. and New Z.",
-  #                                        ifelse(str_detect(un_sub_region,"Sub-Saharan"),"Sub-Saharan\nAfrica",
-  #                                               paste(un_sub_region)))
-  #   ),
-  #   # For number labels
-  #   un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
-  #   )
+    un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
+    )
 

### 2021-11-02 Scripts/Initial_Analysis.Rmd
back to mean

@@
   summarise(
-    mean_yr = median(emerging_yr,na.rm=T),
+    mean_yr = mean(emerging_yr,na.rm=T),
     sd_yr = sd(emerging_yr,na.rm=T),

### 2021-11-09 Scripts/Manuscript.Rmd
Addressed most comments

@@
 Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods).
-**We define a ‘stock’ unit as a species shared by neighboring EEZs.** Firthermore, we  developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
+**We define a ‘stock’ unit as a species shared by neighboring EEZs.** Firthermore, we  developed a Transboundary Shift Index to evaluate range shifts in the shared distribution of transboundary stocks.
 The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.
@@
 
-```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.pos="H", fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone (EEZ) to another.
-The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring EEZ sharing the stock (Top panel).
-Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel).
+```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.pos="H", fig.cap =  "Schematic diagram of the transboundary index (TSI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone (EEZ) to another.
+**Hypothetical distribution range and shape of a shared stock.** The index is based on the distance between the distributional centroid of the transboundary stock and the centroid of the neighboring EEZ sharing the stock (Top panel).
+Time of emergence is defined as the first year when the TSI overshoots historical values (Bottom panel).
 Only showing a subset of ensemble members for clarity."}
@@
 
-<!-- We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks.
+<!-- We then developed a Transboundary Shift Index to evaluate range shifts in the shared distribution of transboundary stocks.
 The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.
@@
 
-We developed a Transboundary Index (TI) to evaluate range shifts in the shared distribution of transboundary stocks under climate change.
+We developed a Transboundary Shift Index (TSI) to evaluate range shifts in the shared distribution of transboundary stocks under climate change.
 This index represents the shift in the distribution centroid of a transboundary stock relative to the centroid of neighboring EEZs that share this stock (Fig.
@@
 where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively, and $lon_{eez}$ and $lon_{ts}$ are associated longitudes.
-Then, for each year and for each ensemble member between 1951 and 2100 we calculated the transboundary index (TI) as follows:
+Then, for each year and for each ensemble member between 1951 and 2100 we calculated the transboundary index (TSI) as follows:
 
@@
 \label{eq:six}
-TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
+TSI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
 \end{equation}
@@
 where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (th, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.
-The TI was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg] and to match transboundary fisheries management that often integrates a longer time-frame in their consideration.
+The TSI was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg] and to match transboundary fisheries management that often integrates a longer time-frame in their consideration.
 
 <!-- Removed in response to reviewer 3 -->
-<!-- Thus, a higher TI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels.
-From the perspective of a state’s fisheries, increasing TI indicates an increase in the challenges to effective management and sustainability of the transboundary stock’s fisheries.
-In contrast, a decreasing TI suggests that the sharing of the stocks between neighboring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries.
+<!-- Thus, a higher TSI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels.
+From the perspective of a state’s fisheries, increasing TSI indicates an increase in the challenges to effective management and sustainability of the transboundary stock’s fisheries.
+In contrast, a decreasing TSI suggests that the sharing of the stocks between neighboring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries.
 -->
@@
 The premise behind time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc].
-The TI’s natural internal variability was estimated from the ensemble members by firstly averaging the TI from 1951 to 2005 and then estimating the variation ($TI_\sigma$) across ensemble members.
-The TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$).
-We then set a threshold (*tresh*) of $TI_\sigma$ to define the time of emergence of transboundary stocks (See Fig.
+The TSI’s natural internal variability was estimated from the ensemble members by firstly averaging the TSI from 1951 to 2005 and then estimating the variation ($TSI_\sigma$) across ensemble members.
+The TSI signal represents the average TSI across ensemble members from 2006 to 2100 ($TSI_\mu$).
+We then set a threshold (*tresh*) of $TSI_\sigma$ to define the time of emergence of transboundary stocks (See Fig.
 \@ref(fig:toeFigure1) for a graphical description):
 
-<!-- We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig.
+<!-- We then set a threshold of two times $TSI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig.
 \@ref(fig:toeFigure1) for a graphical description): -->
 
-<!-- $EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$ -->
+<!-- $EmergenceTSIndex = \frac{\mu_{TSI[t]}}{SD_{TSI_{th}}}$ -->
 <!-- $ToE = t when Emergence index[t] > 1$ -->
@@
 \label{eq:seven}
-ToE = TI_\mu[t']> tresh
+ToE = TSI_\mu[t']> tresh
 \end{equation}
 
-This way we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TI “emerges” from historical internal variability (*tresh*).
+This way we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TSI “emerges” from historical internal variability (*tresh*).
 It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution.
 Two different arbitrary values of *tresh* were used.
-A conservative value ($tresh = 2* TI_\sigma$) representing a 95% probability that the index has emerged from historical variability and a more relaxed value ($tresh=TI_\sigma$) for a 67% probability.
+A conservative value ($tresh = 2* TSI_\sigma$) representing a 95% probability that the index has emerged from historical variability and a more relaxed value ($tresh=TSI_\sigma$) for a 67% probability.
 
@@
 Using a less conservative threshold (i.e., one s.d.
-representing a probability of 67% that the stock has shifted) results in 33% (n = 5,745) more shifting stocks, that is, 67% of all studied stocks (Figs.
+representing a probability of 67% that the stock has shifted) **results in 18% (n = 5,745) more shifting stocks, that is, 63% of all studied stocks** (Figs.
 \@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
@@
 B) Year of shared distribution shifts by UN sub-regions.
-Points color coded by the median year of range shift.
+Points color coded by the median year for all shifting stocks in the region.
 Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda.

### 2021-11-09 Scripts/responses_rev_gcb.Rmd
Addressed most comments

@@
 
-- **Response:**
+- **Response:** Thank you for pointing this out.
+In the original manuscript we were showing the percentage difference between the two numbers (i.e., $x = \frac{|45-63|}{|\frac{45+63}{2}|}*100$), hence the 33.3% difference.
+However, now we realize this can be misleading and have adopted a more straightforward computation.
+We also corrected the 63% typo (not 67%).
+It now reads:
+
+  - - Line XX; *Using a less conservative threshold (i.e., one s.d.
+representing a probability of 67% that the stock has shifted) results in 18% (n = 5,745) more shifting stocks, that is, 63% of all studied stocks*
 
@@
 
-- **Response:**
+- **Response:** Thank you for your commnets on our fiugre 3. We have devided the responses by each point you make. Please see below
+
+  While I like the idea of Figure 3, I feel that there is too much conveyed in this figure.
+It is confusing to have two different scales (for land polygons and EEZs, respectively) on the same color ramp, even though they show completely different things.
+
+  - **Response:**  **NOTE: I do not (fully) agree with this comment.
+yes, the figure is complex, but I do not think having different color scales will make it any less complex.
+I actually think it will make it even harder to read.
+Do you think we can counter-argue this comment?**
+
+  Moreover, see my comments on averaging above.
+
+  - **Response:** Please see our response to the comment above.
+
+  Why is average year used for coloring EEZs, but median year used in Fig 3B?
+
+  - **Response:** In the case of Fig.3B we used the median because this figure talks directly to the statistical non-parametric analysis (Kruskal-Wallis) looking at differences in the time of emergence by geographic region.
+**NOTE: Shall we change Fig.3A to show the median instead of the mean, for the sake of consistency (and make the reviewer happy)?**
 
@@
 
-- **Response:**
+- **Response:** Yes, this refers to the median emergence year for all shifting stocks in each region. This is now included in the figure legend.
+
+  - - Fig. 3 legend, *B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region.*
 
@@
 
-- **Response:**
+- **Response:** We intended to show this information in Figure 4.
+However, the message in the figure is not clear, as the reviewer pointed out in the next comment.
+We have modified figure 4 to show this as a figure.
 
@@
 
-- **Response:**
+- **Response:** Thank you so much for taking the time to revise our manuscript and for your words of encouragement.
 
@@
 
-- **Response:**
+- **Response:** Thank you for this comment. We have now changed the term to "*Transboundary Shift Index (TSI)*"
+
 Fig.
 1: This is a wonderful schematic, my concerns: 1) stocks that do not occur as ovals but occur in patches or blobs instead-- the centroid may be too crude to detect movement of non-centric distributions- I don;t think your approach should change but re-cast as a proof of concept, researchers interested in applying the technique would be able to apply it at a regional level, and perhaps adapt it to patch distributions (separate centroids).
+
+- **Response:** We agree with the reviewer that the distribution of fish stocks in the ocean is not necessarily oval.
+In fact, the distribution of some of our stocks does not follow this patter.
+Nevertheless, we believe our method responds well to other patterns (as long as they are considered one single stock unit) since the centroid of the stock's distribution is based on the "high biomass-level" grid cells.
+We have included a clarification regarding the stocks distribution in the figure legend of the revised manuscript.
+
+  - - *Hypothetical distribution range and shape of a shared stock.*
+
 2) I don;t know what the empty red circle means in bottom panel, I can’t find it. do you mean the white area since the red dot?
+
+- **Response:** Red circles represent the years the TI is larger than the historical variation. These were quite faded in the original diagram and have made them clearer in the revised figure.
+
 3) what is diff btwn distributional and geographic centroid? (suggest using one term or other).
+
+- **Response:** The distributional centroid represent's the stocks centroid while "geographical" centroid is the EEZs centroid.
+We have changed the second term as suggested by the reviewer, it is now called simply "EEZ centroid".
+
 4) WHat if EEZ Ca moves at the same rate in the same direction as EEZ Cb.
@@
 
-- **Response:**
+- **Response:** Ca and Cb represent the centroid of the EEZs, and thus, they do not move.
+What moves is the centroid of the stock (the smaller blue circle) along the whole distribution of the stock that spans both EEZs.
+We realize this was not clear in the legend and have now change it.
+
+  - - : *say EEZ $C_A$: Centroid of the Exclusive Economic Zone A and EEZ $C_B$: Centroid of the Exclusive Economic Zone B*
 
@@
 
-- **Response:**
+- **Response:** See our previous response on SDG 17
 
@@
 
-- **Response:**
+- **Response:** Thank you very much for your kind words

### 2021-11-10 Scripts/responses_rev_gcb.Rmd
Response to the edditor

@@
 
-- **Response:**
+- **Response: PLEASE PAY SPECIAL ATTENTION TO THIS POINT. LOOKS LIKE WE REALLY NEED TO CONVINCE THE EDITOR OUR RESEARCH IS NOVEL**
+
+Below you will find the response to both Reviewers 1 and 2.
+In addition, we would like to offer here an explanation of how our paper differs from recent published papers, including those authored by some of us.
+
+In their 2018 paper, Pinsky *et al.* look at how many *new* transboundary species each EEZ of the world will have towards the end of the century under climate change (i.e., national stocks that become transboundary stocks).
+Indeed, the authors find that “many of the world’s EEZs are likely to receive one to 12 new, climate-driven transboundary stocks by the end of the century.” On the other hand, Oremus *et al.* (2020) “estimates the number of species *exiting* national jurisdictions under climate change up to 2100” concluding that tropical nations stand to lose the most species under these scenarios.
+However, both studies fail to identify changes in *current* transboundary stocks including the intensity of such shift, gap that we look to fill in this paper.
+Here, we build from Palacios-Abrantes *et al.* (2020) and use the transboundary species (n = 633) identified in such paper.
+This means that, unlike Pinsky *et al.* (2018) and Oremus *et al.* (2020) who look at national stocks shifting towards new or away from EEZs, our paper focuses on transboundary stocks shifting within EEZs.
+Moreover, as far as we are concerned no other study, including the research done by us, identifies the actual timeline of shifts in shared stocks, nor the intensity of such shifts.
+
+To identify the timeline we adopt the concept of *Time of Emergence* commonly used in oceanic physical and chemical studies (See Rodgers *et al.* 2015, Keller *et al.* 2014 and has just recently been used in ecological studies by Cheung *et al.* 2020 looking at ecological responses to marine heatwaves.
+As far as we are concerned, no other study has adopted this methodology to looking at the tyming of shifts in transbounary fisheries.
+On the other hand, to identify the intensity of the shift, we (partially) used methods developed by Palacios *et al.* and Sumaila *et al* (2020), published in *Ecology and society*.
+These papers adopt the concept of threat point to evaluate the resilience of specific treaties celebrated by Canada-US tranboundary to manage four transboundary stocks.
+Here, we take this idea of threat point and expand the analysis globally to estimate the intensity that the shifts in transboundary stocks will have.
+
+In short, the current manuscript builds upon Palacios-Abrantes *et al.* 2020 *Scientific Reports* combining methods developed in Sumaila *et al.* 2020 and Palacios-Abrantes *et al.* 2020 *Ecology and Society* to address a question that neither Pinsky *et al.* 2018 *Science* nor Oremus *et al.* (2020) *Nature Sust.* addressed.
+We hope to have clarified the novelty of our research and how it differs from previous studies.
+The new information that we present here can be critical for fisheries management to adapt to shifts in the distribution of transboundary stocks under climate change as highlighted by both reviewers.
 

### 2021-11-10 Scripts/Manuscript.Rmd
typos

@@
 Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods).
-**We define a ‘stock’ unit as a species shared by neighboring EEZs.** Firthermore, we  developed a Transboundary Shift Index to evaluate range shifts in the shared distribution of transboundary stocks.
+**We define a ‘stock’ unit as a species shared by neighboring EEZs.** Furthermore, we  developed a Transboundary Shift Index to evaluate range shifts in the shared distribution of transboundary stocks.
 The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.

### 2021-11-12 Scripts/responses_rev_gcb.Rmd
REsponded to edditor and to stock question

@@
 - **Response:** Thank you for this comment.
-We acknowledge that we did not defined clearly pur concept of stock in the previous version of the manuscript.
+We acknowledge that we did not defined clearly our concept of stock in the previous version of the manuscript.
 The definition of stock has now been included in the revised manuscript in both the introduction and the methods as follows;
@@
 
+- **Response:** Please see below direct responses to each sub-part of the question.
+
+  How are neighboring stocks of the same species considered?
+If, for example, a species’ distribution exists within British Columbia, the United States, and Mexico, is this considered one stock for the purpose of the analyses in the paper (e.g.
+TI, ToE, SSR), or is this two stocks (one between Canada and the US and one between the US and Mexico)?
+
 - **Response:** We acknowledged that this was not fully clear in the previous version of the manuscript and have work towards improving clarity in the revise manuscript.
@@
 
-    Thus, this definition will influence the calculation of SSR *between* neighboring EEZs (i.e., Mexico and the US SSR for sardine and US and BC SSR for sardine, independently one of another).
-As the reviewer suggests, in this scenario Mexico could be loosing sardine SSR to the US contiguous states (this can also be seen as the US gaining SSR from Mexico), while the US contiguous states looses sardine SSR to BC (again, this could also be seen as Canada winning SSR from the US contiguous states ).
+    How does the choice of stock definition influence the calculation of SSR?
+Given the same sample as above, can the US lose SSR of one stock of a species to Canada, while gaining SSR of the same species from Mexico?
+
+- **Response:** Our definition will influence the calculation of SSR *between* neighboring EEZs (i.e., Mexico and the US SSR for sardine and US and BC SSR for sardine, independently one of another).
+As the reviewer suggests, in this scenario Mexico could be loosing sardine SSR to the US contiguous states (this can also be seen as the US gaining SSR from Mexico), while the US contiguous states looses sardine SSR to BC (again, this could also be seen as Canada winning SSR from the US contiguous states).
 This effect is appreciable in the average SSR gains of Figure 5 where Mexico losses to the US contiguous states, who in turn looses to BC, who actually gains from both US continuous states and Alaska.
-This is not explain on the methods as follows **NOT SURE WE NEED THIS**;
+This is now explained on the methods as follows **NOT SURE WE NEED THIS**;
 
@@
 
-    This way, our definition of a shared stock allows for stocks to go to or leave an EEZ from all borders.
-Continuing using Mexico as an example, Figure 5 shows how in average Mexico will loose SSR of the stocks it shares with the US up north probably do to a poleward shift on the Californian current species [@cisneros-montemayor_ecosystem-based_2020] but also to the south to Guatemala as stocks follow local environmental gradients [@Clarke:2020ep].
+  How does the choice of stock definition influence the calculation of SSR?
+
+- **Response:** The definition of "stock" will likely influence the results as it is a fundamental part of the analysis.
+Replicating this analysis at the species level will undoubtedly result in multiple countries sharing each species.
+For example, in the case of Atlantic Cod, this would be shared by US, Canada and Greenland (and maybe enven northern Europe).
+Choosing this approach invariably comes with the assumption that fishing Cod in, e.g., the Gulf of Maine, will have direct impact on e.g., Greenland fisheries.
+While this might be true for some few highly migratory species (e.g., tunas) we believe is a stretch for all other species in this analysis.
+
+  However, even when a stock of non highly-migratory species does span along the coast of more than two countries, our results seem to be similar (this relates to the part regarding Mexico US and Canada).
+For example, Pacific halibut (*Hippoglossus stenolepis*) is considered to be one single stock that extends along the Pacific coast of the US contiguous states, Canada, and Alaska [@iphc_iphc_2016_report_2016].
+Palacios-Abrantes *et al.* (2020) analyzed the shift in SSR between the three parties and found that the US contiguous states will likely loose SSR to Canada while Canada will loose to Alaska.
+A similar patter shown in the present study.
+
+    Ideally, one could define a stock as a biologically-defined population.
+This could be the case of Peruvian anchoveta *(Engraulis ringens)* off the coast of Chile and Peru.
+There are two identified stocks of anchoveta; one in the central/norther part of the Peruvian EEZ and one in the south, shared with Chile.
+This differentiation will likely impact both the ToE and SSR results.
+However, as far as we are concern, such analysis is challenged by a lack of stock-level data around the world.
 
@@
   - **Response:** We understand the reviewer's concern and hope that our response have clarified this issue.
-However, it is almost certain that our definition of stock is influencing the results of our paper, just by tha nature of the analysis.
+However, it is almost certain that our definition of stock is influencing the results of our paper, just by that nature of the analysis.
 While a more fine scale definition (e.g., biologically-defined populations of stock will probably arose other results, not limiting the species to, at least, an EEZ could result in an overestimation of our results (See Ovando *et al.* https://www.pnas.org/content/118/23/e2025958118).
+This is a main caveat of our study and have made an effort to be very clear and upfront about it in line XX of the revised manuscript:
+
+  - - **Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks, rather than biologically-defined populations.
+While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca].
+Moreover, not limiting the species to, at least, an EEZ could result in an overestimation of our results (**NOTE:** See Ovando *et al.* https://www.pnas.org/content/118/23/e2025958118).**
 
@@
 
-- **Response:** We have revised the manuscript to clarify that we did account for the heterogeneous distribution of fish stocks across their entire range.
+- **Response:** We agree with the reviewer that and unweighted centroid can severely skew the actual distribution of the species and influence our results.
+We have revised the manuscript to clarify that we did account for the heterogeneous distribution of fish stocks across their entire range.
 The geographic centroid of each stock is indeed weighted by the abundance of the stock.
 The DBEM uses the species probability of occurrence to estimate the biomass of the species at each 0.5 x 0.5 gridcell as indicated from lines XX to XX of the methods.
-The centroid of a transboundary stock was determined by the average latitude and longitude ($lon_{ts}$) of the grid cells holding the specified percentile ($per^{th}$) of the projected transboundary stock abundance.
-This is, we determined the centroid of the stock using only those regions where abundace is higher (i.e., using the probability of occurrence of the DBEM).
-It reads as follows:
+The centroid of a transboundary stock was determined by the average latitude and longitude ($lon_{ts}$) of the grid cells holding the specified percentile ($per^{th}$) of the projected transboundary stock abundance (see below).
+This is, we determined the centroid of the stock using only those regions where abundance was higher (i.e., using the probability of occurrence of the DBEM).
+Moreover, we estimated the sensitivity of our results to the (arbitrarily) $per^{th} = 95^{th}$ percentile threshold as described in lines XX and shown in Figure S6 of the supplements.
+It reads as follows in line XX:
 
-  - Line, *Fish stocks are not evenly distributed across their entire range, thus we focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place.
+  - Line XX, *Fish stocks are not evenly distributed across their entire range, thus we focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place.
 Thus, we estimated the centroid of a stock only including grid cells where the projected stock abundance within the neighboring EEZs sharing the stock was above the $95^{th}$ percentile.*
-
-    We agree that and unweighted centroids can severely skew the actual distribution of the species and influence our results.
-We estimated the sensitivity of our results to the (arbitrarily) $95^{th}$ percentile threshold as described in lines XX and shown in Figure S6 of the supplements.
 
@@
 For example, would the arrows look different in Figure 5 if the averages were weighted by each fish stock’s value?
-
-**NOT SURE I UNDERSTAND THIS COMMENT**
 
@@
 - **Response:** Thank you for your concern.
-In short the answer is yes, an EEZ could both lose and gain SSR of the same *species* but not of the same *stock.* This is because, according to our definition of stock, a species who spans multiple EEZs has multiple stocks (see response to the previus question and the main comment).
+In short the answer is yes, an EEZ could both lose and gain SSR of the same *species* but not of the same *stock.* This is because, according to our definition of stock, a species who spans multiple EEZs has multiple stocks (see response to the previous question and the main comment).
 let us use the Peruvian anchoveta (*Engraulis ringens*) as an example.

### 2021-11-12 Scripts/Manuscript.Rmd
responded to edditor and stock question

@@
 Further research that reduces uncertainty in the NPP response at the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to smaller structural uncertainty of fish and fisheries models [@BryndumBuchholz:2019kh].
-Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks.
+**Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks, rather than biologically-defined populations.
 While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca].
-Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identify different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
+Moreover, not limiting the species to, at least, an EEZ could result in an overestimation of our results (NOTE: See Ovando *et al.* https://www.pnas.org/content/118/23/e2025958118).** Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identify different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
 Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.

### 2021-11-12 Scripts/Initial_Analysis.Rmd
Created top 5 stocks plot

@@
 }
+
+```
+
+##### Fig. N3 Mapping the time of emergence Of Top stocks
+
+```{r emergence_map, eval = F}
+
+top_stocks_per_country <- Eme_Res %>% 
+  rename(geo_region = un_sub_region) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,geo_region,tresh,taxon_key) %>% 
+  summarise(n()) %>% 
+  left_join(Sau_data,
+            by = c("fishing_entity_id","taxon_key")) %>% 
+  filter(!is.na(variable)) %>% 
+  # group_by(sau_fishing_entity,fishing_entity_id,tresh,variable) %>% 
+  # summarise(sum_value = sum(value,na.rm=T)) %>% 
+  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
+  # Select top 10 most valuable stocks
+  top_n(5,value)
+
+# Per EEZs
+top_toe <- Eme_Res %>% 
+  rename(geo_region = un_sub_region) %>% 
+  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
+  # summarise(n_tax = length(unique(taxon_key))) %>% 
+  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
+  right_join(top_stocks_per_country,
+             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
+             ) %>% 
+  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
+  summarise_at(vars(mean_yr),
+               c("mean"=mean,
+                 "sd" = sd)) %>% 
+  mutate(
+    geo_region = ifelse(is.na(geo_region),"Other",geo_region)
+  )
+
+```
+
+###### The bar option by UN sub region
+
+```{r}
+
+# Create data
+Species_Data <- top_toe %>% 
+  filter(tresh == selected_tresh,
+         variable == "value") %>% 
+  mutate(mean = mean-2000,
+         group = geo_region)
+
+# Set a number of 'empty bar' to add at the end of each group
+Char <- as.factor(Species_Data$group)
+empty_bar <- 5
+to_add <- data.frame( matrix(NA, empty_bar*nlevels(Char), ncol(Species_Data)) )
+colnames(to_add) <- colnames(Species_Data)
+to_add$group <- rep(levels(Char), each=empty_bar)
+Species_Data <- rbind(Species_Data, to_add)
+Species_Data <- Species_Data %>% arrange(group,mean)
+Species_Data$id <- seq(1, nrow(Species_Data))
+
+# Species_Data$id <- seq(1, nrow(Species_Data))
+
+# Get the name and the y position of each label
+label_data <- Species_Data 
+number_of_bar <- nrow(label_data)
+angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar
+label_data$hjust <- ifelse( angle < -90, 1, 0)
+label_data$angle <- ifelse(angle < -90, angle+180, angle)
+
+# prepare a data frame for base lines
+base_data <- Species_Data %>% 
+  group_by(group) %>% 
+  summarize(start=min(id), 
+            end=max(id) - empty_bar, 
+            median = median(mean, na.rm = T)
+  ) %>% 
+  rowwise() %>% 
+  mutate(title=mean(c(start, end)))
+
+# prepare a data frame for grid (scales)
+grid_data <- base_data
+grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
+grid_data$start <- grid_data$start - 1
+# grid_data <- grid_data[-1,]
+# grid_data[1,2] <- 5 # Manually include bathydemerssal
+# grid_data[1,3] <- 1
+
+un_regions <- grid_data %>% 
+  select(group)
+
+ggplot(Species_Data, 
+       aes(x=as.factor(id), 
+           y=mean,
+           fill=mean)
+) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
+  geom_bar(aes(x=as.factor(id),
+               y=mean,
+               fill=mean),
+           stat="identity",
+           alpha=0.5
+  ) +
+  geom_segment(data=grid_data,
+               aes(x = end, y = 30, xend = start, yend = 30),
+               colour = "grey90", alpha=1, size=0.3, inherit.aes = FALSE) +
+  geom_bar(aes(x=as.factor(id),
+               y=mean,
+               fill=mean
+  ),
+  stat="identity",
+  alpha=0.5) +
+  annotate("text", 
+           x = 0.5,
+           y = 37,
+           label = "UN Agenda",
+           color= "grey60", 
+           size=3, 
+           angle=-5
+  ) +
+  # Country names
+  geom_text(data=label_data,
+            aes(x=id, 
+                y= 100,
+                label=sau_fishing_entity,
+                hjust=hjust),
+            color="black",
+            fontface="italic",
+            alpha = 1,
+            size = 4 ,
+            angle = label_data$angle,
+            inherit.aes = FALSE) +
+  # Continents labels
+  # geom_text(data=base_data,
+  #           aes(x = title,
+  #               y = 85, 
+  #               label=group), 
+  #           # hjust=c(1,1,1,0,0,0),
+  #           colour = "black",
+  #           # angle = myAng,
+  #           size=5,
+  #           fontface="bold",
+  #           inherit.aes = FALSE) +
+  ylim(-100,100) + # sets the circle the first value the central the last value how wide
+  theme_minimal() +
+  theme(
+    legend.position = c(0.5,0.5),
+    legend.text = element_text(size = 13),
+    legend.title = element_text(size = 14, face = "bold"),
+    legend.title.align = 0.5,
+    axis.text = element_blank(),
+    axis.title = element_blank(),
+    panel.grid = element_blank(),
+    panel.background = element_rect(colour = "white"),
+    legend.direction="horizontal",
+    legend.key.width = unit(3, 
+                            "line")
+  ) +
+  coord_polar(start = 0) +
+  scale_fill_viridis("Average year of range shift\nfor top 5 most valuable stocks",
+                     limits = c(0,100),
+                     breaks = seq(0,100,20),
+                     labels = seq(2000,2100,20),
+                     direction = -1
+  ) +
+  guides(
+    fill=guide_colorbar(title.position="top", 
+                        title.hjust =0.5)
+  )
+
+ggsave(
+  plot = last_plot(),
+  width = 13,
+  height = 13,
+  units = "in",
+  filename = "./Figures/top_stocks_circular_trend.png"
+)
+
+```
+
+###### The Map option with tops
+
+```{r the map}
+
+# Create data
+Top_map_data <- top_toe %>% 
+  filter(tresh == selected_tresh,
+         variable == "value") %>% 
+  mutate(mean = mean-2000) %>%
+  arrange(mean)
+
+# These are regions without transboundary species
+grey_land <- World_map %>% 
+  filter(!fishing_entity_id %in% Value_map_data$fishing_entity_id) %>% 
+  filter(!sovereignt %in% c("Antarctica"))
+
+# grey_eez <- World_eez_sf %>% 
+#   filter(!eezid %in% Eez_d_hist$eezid,
+#          !eez_name %in% Neighbours$eez_name)
+
+# -------------------------- #
+# Call plot for both tresholds
+# -------------------------- #
+# tresh_list <- c("tresh_one","tresh_two")
+
+  # Value Map
+  World_map_part <- World_map_nemer %>% 
+  filter(
+    tresh ==  ,
+    variable == "value"
+  )
+
+  # ToE Map 
+  toe_map_data <- World_eez_sf %>% 
+    left_join(Eez_d_hist,
+              by = "eezid") %>% 
+    filter(tresh == tresh_list[i],
+           metric == "mean_yr")
+
+  # Map with countries without ToE
+
+  russia_churchil <- World_eez_sf %>% 
+    filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))
+
+  no_toe_eez <- World_eez_sf %>% 
+    filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
+    rbind(russia_churchil)
+
+  no_toe_land <- World_map %>% 
+    filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !name %in% World_map_part$name) # Filter EEZs with ToE stocks
+
+  map <- ggplot(toe_map_data) +
+    geom_sf(
+      aes(
+        fill = reverse #original map
+        # fill = value
+      ),
+      color =  "white",
+      size = 0.1
+    ) +
+    # Inlcude the participation of species
+    geom_sf(data = World_map_part,
+            # aes(fill = participation+2000), #original map
+            aes(fill = participation),
+            color =  "white",
+            size = 0.1
+    ) +
+    # grey maps
+    # ggplot() +
+    geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
+    # geom_sf(data = grey_eez, aes(), fill = "white", color =  "grey50",size = 0.1) +
+    geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    # Set Participation labels
+    annotate("text",
+             x =-150,
+             y = -69,
+             label = "Transboundary fishing revenue (%)",
+             color="black",
+             size=5,
+             angle=0,
+             fontface="bold",
+             hjust=0) +
+    annotate("text", 
+             x = c(-72,-35,5,46,82),
+             y = -69,
+             # x =(c(-8500000,-3000000,2500000,7500000,12800000)),
+             # y = -9050000,
+             label = c(0,25,50,75,100), 
+             color="black", 
+             size=4, 
+             angle=0, 
+             # fontface="bold",
+             hjust=0) +
+    # Set ToE labels
+    annotate("text", 
+             x =-150,
+             y = -85,
+             label = "Average year of range shift",
+             color="black", 
+             size=5, 
+             angle=0, 
+             fontface="bold",
+             hjust=0) +
+    annotate("text", 
+             x = c(-79,-39,2,42,81),
+             y = -85,
+             # label = seq(2100,2000,-25), # original
+             label = c(">2100",seq(2075,2000,-25)), # not inluding the > because we dont know
+             color="black", 
+             size=4, 
+             angle=0, 
+             # fontface="bold",
+             hjust=0) +
+    # Set No emergence label
+    annotate("text",
+             x =110,
+             y = -77,
+             label = "No shift",
+             color="black",
+             size=5,
+             angle=0,
+             fontface="bold",
+             hjust=0) +
+    scale_fill_viridis("",
+                       limits = c(2000,2100),
+                       breaks = seq(2000,2100,25),
+                       labels = c("","","","","")#,
+                       # direction = -1
+    ) +
+    # scale_fill_gradientn("",
+    #                      limits = c(2000,2100),
+    #                      breaks = seq(2000,2100,25),
+    #                       labels = c("","","","",""),
+    #                      colours = pal
+    # ) +
+    # geom_hline(yintercept = c(-50,0,50,80)) +
+    my_ggtheme_m(leg_pos = "bottom") +
+    theme(
+      legend.box.margin = margin(t = -62, l = 252),
+      legend.text = element_blank(),
+      legend.key.width = unit(5, "line")
+    )
+
+  ## --------------- ##
+  # Regional Boxplot
+  ## --------------- ##
+
+  # Weighted verion
+  # boxplot_data <- Eme_Res %>% 
+  #   # Get the order of boxes
+  #   filter(tresh == tresh_list[i],
+  #          !is.na(un_sub_region)) %>%
+  #   group_by(un_sub_region) %>%
+  #   summarise(n_eezs = length(unique(sf_eez_name))) %>%
+  #   # Get the rest of the data
+  #   left_join(Eme_Res) %>%
+  #   filter(tresh == tresh_list[i],
+  #          !is.na(un_sub_region)) %>%
+  #   # mutate(weight = )
+  #   mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
+  #                                 ifelse(str_detect(un_sub_region,"Australia"),"Aus. and New Z.",
+  #                                        ifelse(str_detect(un_sub_region,"Sub-Saharan"),"Sub-Saharan\nAfrica",
+  #                                               paste(un_sub_region)))
+  #   )#,
+  #   # For number labels
+  #   # un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
+  #   )
+
+  # Non-weighted verion
+
+  boxplot_data <- Eme_Res %>%
+    # Get the order of boxes
+    filter(tresh == tresh_list[i],
+           !is.na(un_sub_region)) %>%
+    group_by(un_sub_region) %>%
+    summarise(order = median(emerging_yr),
+              n_eezs = length(unique(sf_eez_name))) %>%
+    # Get the rest of the data
+    left_join(Eme_Res) %>%
+    filter(tresh == tresh_list[i],
+           !is.na(un_sub_region)) %>%
+    mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
+                                  ifelse(str_detect(un_sub_region,"Australia"),"Aus. and New Z.",
+                                         ifelse(str_detect(un_sub_region,"Sub-Saharan"),"Sub-Saharan\nAfrica",
+                                                paste(un_sub_region)))
+    ),
+    # For number labels
+    un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
+    )
+
+  boxplot_data$un_sub_region <- boxplot_data$un_sub_region %>% 
+    gsub("Northern","N.",.) %>% 
+    gsub("Southern","S.",.) %>% 
+    gsub("Western","W.",.) %>% 
+    gsub("Eastern","E.",.) %>% 
+    gsub("South-eastern","S.E.",.)
+
+  # ---------- #
+  # Summary data
+  # ---------- #
+
+  medians_data <- boxplot_data %>% 
+    group_by(un_sub_region,un_geo_region) %>% 
+    summarise(mean = mean(emerging_yr,na.rm = T),
+              sd = sd(emerging_yr,na.rm = T),
+              min = min(emerging_yr,na.rm = T),
+              max = max(emerging_yr,na.rm = T),
+              med = median(emerging_yr,na.rm = T),
+              n_stocks = n()
+    )
+
+  boxplot_data <- boxplot_data %>% 
+    left_join(medians_data,
+              by = c("un_sub_region","un_geo_region")
+              )
+
+  boxplot <- ggplot(boxplot_data) +
+    geom_violin(
+      aes(
+        x = reorder(un_sub_region,order),
+        y = emerging_yr
+      ),
+      fill = "grey"
+    ) + 
+    facet_wrap(~un_geo_region,
+               scales = "free_x",
+               nrow = 1) +
+    # Median points
+    geom_point(
+      aes(
+        x = reorder(un_sub_region,order),
+        y = mean,
+        colour = mean
+      ),
+      size = 5
+    ) + 
+    # h line
+    geom_hline(yintercept = 2030, color = "grey", size = 0.5, linetype="dashed") +
+    scale_colour_viridis("",
+                       limits = c(2000,2100),
+                       breaks = seq(2000,2100,25),
+                       labels = c("","","","",""),
+                       direction = -1
+    ) +
+    xlab("") +
+    ylab("Year of range shift") +
+    scale_y_continuous(
+                   labels = c(2000,2025, 2050, 2075, ">2100"), # not inluding the > because we dont know
+    ) +
+    my_ggtheme_p() +
+    theme(
+      axis.text.x = element_text(size = 16, angle = 45, face = "plain", hjust = 1), 
+      strip.text.x = element_text(size = 18, colour = "black"), 
+      axis.title.y = element_text(size = 18),
+      axis.text.y = element_text(size = 16), 
+      # axis.ticks.x = element_blank(),
+      legend.position = ""
+    );boxplot
+
+  ## --------------- ##
+  # Combine plots and save them
+  ## --------------- ##
+
+  p <- ggdraw() +
+    # ToE Latitudinal plot
+    # draw_plot(Eez_hist, x = 0.873, y = 0.53, width = 0.124, height = 0.37) +
+    # Time of Emergence Map
+    draw_plot(map, x = 0, y = 0.4, width = 1, height = 0.6) +
+    # Regional boxplot
+    draw_plot(boxplot, x = 0, y = -0.03, width = 1, height = 0.44) +
+    # Labels
+    draw_plot_label(label = c("A", "B"), 
+                    size = 20,
+                    x = c(0, 0), 
+                    y = c(0.95, 0.46)
+    )
+
+  save_plot(
+    name_save,
+    p,
+    base_height = 11,
+    base_width = 12
+  )
+
+}
+
+```
+
+##### New Fig. 4. Top shifting stocks/EEZs weighted (proportion)
+
+```{r, top_stocks_eez, eval = F, echo = F}
+
+Value_map_data %>% 
+  group_by(variable,tresh) %>% 
+  top_frac(0.10,participation) %>% 
+  filter(tresh == "tresh_two",
+         variable == "value") %>% 
+  # View()
+  left_join(Sau_data) %>% 
+  group_by(variable,tresh) %>% 
+  top_frac(0.10,value) %>%
+  left_join(my_path("G","SAU","exploited_species_list.csv",read = T)) %>% 
+  ggplot() +
+  geom_bar(
+    aes(
+      x = common_name,
+      y = value
+    ),
+    stat = "identity"
+  ) +
+  facet_wrap(~sau_fishing_entity,
+             scale = "free") +
+  theme(axis.text.x = element_text(angle = 90))
 

### 2021-11-12 Scripts/Initial_Analysis.Rmd
Created top 5 map and circular plot. I think the map is better

@@
   filter(tresh == selected_tresh,
-         variable == "value") %>% 
-  mutate(mean = mean-2000) %>%
-  arrange(mean)
+         variable == "value") # tried with catch but is quite the same
+
+World_map_top <- World_map %>%
+  full_join(Top_map_data,
+            by = c("sau_fishing_entity")
+  ) %>% 
+  filter(!is.na(variable))
 
@@
 grey_land <- World_map %>% 
-  filter(!fishing_entity_id %in% Value_map_data$fishing_entity_id) %>% 
+  filter(!sau_fishing_entity %in% Top_map_data$sau_fishing_entity) %>% 
   filter(!sovereignt %in% c("Antarctica"))
-
-# grey_eez <- World_eez_sf %>% 
-#   filter(!eezid %in% Eez_d_hist$eezid,
-#          !eez_name %in% Neighbours$eez_name)
-
-# -------------------------- #
-# Call plot for both tresholds
-# -------------------------- #
-# tresh_list <- c("tresh_one","tresh_two")
-
-  # Value Map
-  World_map_part <- World_map_nemer %>% 
-  filter(
-    tresh ==  ,
-    variable == "value"
-  )
-
-  # ToE Map 
-  toe_map_data <- World_eez_sf %>% 
-    left_join(Eez_d_hist,
-              by = "eezid") %>% 
-    filter(tresh == tresh_list[i],
-           metric == "mean_yr")
 
@@
 
-  map <- ggplot(toe_map_data) +
+  map <- ggplot(World_map_top) +
     geom_sf(
       aes(
-        fill = reverse #original map
+        fill = mean #original map
         # fill = value
       ),
-      color =  "white",
-      size = 0.1
-    ) +
-    # Inlcude the participation of species
-    geom_sf(data = World_map_part,
-            # aes(fill = participation+2000), #original map
-            aes(fill = participation),
       color =  "white",
@@
     geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
-    # geom_sf(data = grey_eez, aes(), fill = "white", color =  "grey50",size = 0.1) +
-    geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    # geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
     geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
     # Set Participation labels
-    annotate("text",
-             x =-150,
-             y = -69,
-             label = "Transboundary fishing revenue (%)",
-             color="black",
-             size=5,
-             angle=0,
-             fontface="bold",
-             hjust=0) +
-    annotate("text", 
-             x = c(-72,-35,5,46,82),
-             y = -69,
-             # x =(c(-8500000,-3000000,2500000,7500000,12800000)),
-             # y = -9050000,
-             label = c(0,25,50,75,100), 
-             color="black", 
-             size=4, 
-             angle=0, 
-             # fontface="bold",
-             hjust=0) +
-    # Set ToE labels
-    annotate("text", 
-             x =-150,
-             y = -85,
-             label = "Average year of range shift",
-             color="black", 
-             size=5, 
-             angle=0, 
-             fontface="bold",
-             hjust=0) +
-    annotate("text", 
-             x = c(-79,-39,2,42,81),
-             y = -85,
-             # label = seq(2100,2000,-25), # original
-             label = c(">2100",seq(2075,2000,-25)), # not inluding the > because we dont know
-             color="black", 
-             size=4, 
-             angle=0, 
-             # fontface="bold",
-             hjust=0) +
-    # Set No emergence label
-    annotate("text",
-             x =110,
-             y = -77,
-             label = "No shift",
-             color="black",
-             size=5,
-             angle=0,
-             fontface="bold",
-             hjust=0) +
-    scale_fill_viridis("",
+    scale_fill_viridis("Average year of range shift",
                        limits = c(2000,2100),
                        breaks = seq(2000,2100,25),
-                       labels = c("","","","","")#,
-                       # direction = -1
-    ) +
-    # scale_fill_gradientn("",
-    #                      limits = c(2000,2100),
-    #                      breaks = seq(2000,2100,25),
-    #                       labels = c("","","","",""),
-    #                      colours = pal
-    # ) +
-    # geom_hline(yintercept = c(-50,0,50,80)) +
+                       direction = -1
+    ) +
     my_ggtheme_m(leg_pos = "bottom") +
-    theme(
-      legend.box.margin = margin(t = -62, l = 252),
-      legend.text = element_blank(),
-      legend.key.width = unit(5, "line")
-    )
-
-  ## --------------- ##
-  # Regional Boxplot
-  ## --------------- ##
-
-  # Weighted verion
-  # boxplot_data <- Eme_Res %>% 
-  #   # Get the order of boxes
-  #   filter(tresh == tresh_list[i],
-  #          !is.na(un_sub_region)) %>%
-  #   group_by(un_sub_region) %>%
-  #   summarise(n_eezs = length(unique(sf_eez_name))) %>%
-  #   # Get the rest of the data
-  #   left_join(Eme_Res) %>%
-  #   filter(tresh == tresh_list[i],
-  #          !is.na(un_sub_region)) %>%
-  #   # mutate(weight = )
-  #   mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
-  #                                 ifelse(str_detect(un_sub_region,"Australia"),"Aus. and New Z.",
-  #                                        ifelse(str_detect(un_sub_region,"Sub-Saharan"),"Sub-Saharan\nAfrica",
-  #                                               paste(un_sub_region)))
-  #   )#,
-  #   # For number labels
-  #   # un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
-  #   )
-
-  # Non-weighted verion
-
-  boxplot_data <- Eme_Res %>%
-    # Get the order of boxes
-    filter(tresh == tresh_list[i],
-           !is.na(un_sub_region)) %>%
-    group_by(un_sub_region) %>%
-    summarise(order = median(emerging_yr),
-              n_eezs = length(unique(sf_eez_name))) %>%
-    # Get the rest of the data
-    left_join(Eme_Res) %>%
-    filter(tresh == tresh_list[i],
-           !is.na(un_sub_region)) %>%
-    mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
-                                  ifelse(str_detect(un_sub_region,"Australia"),"Aus. and New Z.",
-                                         ifelse(str_detect(un_sub_region,"Sub-Saharan"),"Sub-Saharan\nAfrica",
-                                                paste(un_sub_region)))
-    ),
-    # For number labels
-    un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
-    )
-
-  boxplot_data$un_sub_region <- boxplot_data$un_sub_region %>% 
-    gsub("Northern","N.",.) %>% 
-    gsub("Southern","S.",.) %>% 
-    gsub("Western","W.",.) %>% 
-    gsub("Eastern","E.",.) %>% 
-    gsub("South-eastern","S.E.",.)
-
-  # ---------- #
-  # Summary data
-  # ---------- #
-
-  medians_data <- boxplot_data %>% 
-    group_by(un_sub_region,un_geo_region) %>% 
-    summarise(mean = mean(emerging_yr,na.rm = T),
-              sd = sd(emerging_yr,na.rm = T),
-              min = min(emerging_yr,na.rm = T),
-              max = max(emerging_yr,na.rm = T),
-              med = median(emerging_yr,na.rm = T),
-              n_stocks = n()
-    )
-
-  boxplot_data <- boxplot_data %>% 
-    left_join(medians_data,
-              by = c("un_sub_region","un_geo_region")
-              )
-
-  boxplot <- ggplot(boxplot_data) +
-    geom_violin(
-      aes(
-        x = reorder(un_sub_region,order),
-        y = emerging_yr
-      ),
-      fill = "grey"
-    ) + 
-    facet_wrap(~un_geo_region,
-               scales = "free_x",
-               nrow = 1) +
-    # Median points
-    geom_point(
-      aes(
-        x = reorder(un_sub_region,order),
-        y = mean,
-        colour = mean
-      ),
-      size = 5
-    ) + 
-    # h line
-    geom_hline(yintercept = 2030, color = "grey", size = 0.5, linetype="dashed") +
-    scale_colour_viridis("",
-                       limits = c(2000,2100),
-                       breaks = seq(2000,2100,25),
-                       labels = c("","","","",""),
-                       direction = -1
-    ) +
-    xlab("") +
-    ylab("Year of range shift") +
-    scale_y_continuous(
-                   labels = c(2000,2025, 2050, 2075, ">2100"), # not inluding the > because we dont know
-    ) +
-    my_ggtheme_p() +
-    theme(
-      axis.text.x = element_text(size = 16, angle = 45, face = "plain", hjust = 1), 
-      strip.text.x = element_text(size = 18, colour = "black"), 
-      axis.title.y = element_text(size = 18),
-      axis.text.y = element_text(size = 16), 
-      # axis.ticks.x = element_blank(),
-      legend.position = ""
-    );boxplot
-
-  ## --------------- ##
-  # Combine plots and save them
-  ## --------------- ##
-
-  p <- ggdraw() +
-    # ToE Latitudinal plot
-    # draw_plot(Eez_hist, x = 0.873, y = 0.53, width = 0.124, height = 0.37) +
-    # Time of Emergence Map
-    draw_plot(map, x = 0, y = 0.4, width = 1, height = 0.6) +
-    # Regional boxplot
-    draw_plot(boxplot, x = 0, y = -0.03, width = 1, height = 0.44) +
-    # Labels
-    draw_plot_label(label = c("A", "B"), 
-                    size = 20,
-                    x = c(0, 0), 
-                    y = c(0.95, 0.46)
-    )
-
-  save_plot(
-    name_save,
-    p,
-    base_height = 11,
-    base_width = 12
-  )
-
-}
+    theme(legend.key.width = unit(5, "line")) +
+
+    guides(
+      fill=guide_colorbar(title.position="top", 
+                        title.hjust =0.5)
+  )
+
+  ggsave(
+    plot = map,
+    "./Figures/top_toe_map.png",
+    width = 10,
+    height = 5,
+    units = "in"
+  )
 

### 2021-11-12 Scripts/Initial_Analysis.Rmd
Created new figuere on top stocks and wrote text about it

@@
   # Select top 10 most valuable stocks
-  top_n(5,value)
 
@@
   )
+
+# Numbers for paper
+x <- top_toe %>% 
+  filter(mean_yr <= 2021,
+         variable == "value",
+         tresh == selected_tresh)
+
+length(unique(x$sau_fishing_entity))/length(unique(Sau_data$fishing_entity_id))
+
+# Number of shifting stocks per geo-region
+x %>% 
+  group_by(geo_region) %>% 
+  tally() %>% 
+  arrange(n) %>% 
+  top_frac(0.2,n) # top 20%
 
@@
          variable == "value") # tried with catch but is quite the same
+
+mean(Top_map_data$mean)
+sd(Top_map_data$mean)
 

### 2021-11-12 Scripts/Manuscript.Rmd
Created new figuere on top stocks and wrote text about it

@@
 
+**Having high-valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting.
+On average, countries will see a range shift of the top 5 most valuable stocks by 2038 $\pm$ 16 years (Figure S-NEW).
+Moreover, our projections suggest that these high value stocks are already shifting in 66% of coastal nations, with the regions of Polynesia (n = 22), Sub-Saharan Africa (n = 36) and Latin America and the Caribbean (n = 62) presenting the larger number of shifting stocks.**
+
+(**New Figuree**).
+
 We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs for both thresholds (one and two s.d.).

### 2021-11-12 Scripts/responses_rev_gcb.Rmd
Created new figuere on top stocks and wrote text about it

@@
 
-- **Response (INCOMPLETE):** We were not clear in our message and have worked on clarifying all of these results.
-Specifically, in figure three we are showing the participation of the transboundary stocks that are shifting between neighbouring EEZs relative to the total revenue from all transboundary stocks that each EEZ has.
-So, for example, the transboundary stocks that are shifting in the Mexican Pacific coast to/from US and Guatemala represent about 15% of Mexico's total revenue from transboundary stocks.
-They average year of shift is ~2025.
-Indeed as the reviewer suggests, Figure 4 does not include any weight value, but rather absolute revenue amounts.
-The figure's objective is not only to inform the reader about the economic importance, in terms of absolute USD, of these shifting transboundary stocks (rather than their weighted value that is presented in Figure 3) but also highlight that these shifts are happening close to the UN 2030 Agenda.
-Finally, Figure 5 does indeed show weighted values.
-Here, we are showing the...
+- **Response:** This is a very good observation and we believe it has increased the impact of our manuscript.
+We have included two new figure (**Supplement Figure XX**) showing the average range shift of the top 5 most valuable transboundary stocks per country in addition to the following text on line XX:
+
+  - Line XX: *Having high-valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting.
+On average, countries will see a range shift of the top 5 most valuable stocks by 2038 $\pm$ 16 years (Figure S-NEW).
+Moreover, our projections suggest that these high value stocks are already shifting in 66% of coastal nations, with the regions of Polynesia (n = 22), Sub-Saharan Africa (n = 36) and Latin America and the Caribbean (n = 62) presenting the larger number of shifting stocks (20% of total high value stocks).*
+
+![Top 5](../Figures/top_toe_map.png)
 

### 2021-11-12 Scripts/Initial_Analysis.Rmd
Updates of the top stocks for SSR

@@
   ) +
-  my_ggtheme_m() +
+  my_ggtheme_m(leg_pos = "bottom") +
   ggsave("botTemp_ToE.png",
@@
 }
+
+SummaryProp(taxon_key)
 
@@
   group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
-  # Select top 10 most valuable stocks
+  # Select top 5 most valuable stocks
+  top_n(5,value)
 
@@
     theme(legend.key.width = unit(5, "line")) +
-
     guides(
@@
   )
-
-```
-
-##### New Fig. 4. Top shifting stocks/EEZs weighted (proportion)
-
-```{r, top_stocks_eez, eval = F, echo = F}
-
-Value_map_data %>% 
-  group_by(variable,tresh) %>% 
-  top_frac(0.10,participation) %>% 
-  filter(tresh == "tresh_two",
-         variable == "value") %>% 
-  # View()
-  left_join(Sau_data) %>% 
-  group_by(variable,tresh) %>% 
-  top_frac(0.10,value) %>%
-  left_join(my_path("G","SAU","exploited_species_list.csv",read = T)) %>% 
-  ggplot() +
-  geom_bar(
-    aes(
-      x = common_name,
-      y = value
-    ),
-    stat = "identity"
-  ) +
-  facet_wrap(~sau_fishing_entity,
-             scale = "free") +
-  theme(axis.text.x = element_text(angle = 90))
 
@@
     # ) +
-    my_ggtheme_m() +
+    my_ggtheme_m(leg_pos = "bottom") +
     scale_colour_viridis(
@@
     scale_y_continuous(breaks = seq(10,30,10)) +
-    my_ggtheme_m() +
+    my_ggtheme_m(leg_pos = "bottom") +
     theme(legend.position = "",
@@
     # scale_x_continuous(breaks = seq(45,65,5)) +
-    my_ggtheme_m() +
+    my_ggtheme_m(leg pos = "bottom") +
     theme(legend.position = "",
@@
 
+##### Fig. TOP 5
+
+```{r fig_map_Juan, eval = F, echo =F}
+
+top_stocks_per_country <- Eme_Res %>% 
+  rename(geo_region = un_sub_region) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,geo_region,tresh,taxon_key) %>% 
+  summarise(n()) %>% 
+  left_join(Sau_data,
+            by = c("fishing_entity_id","taxon_key")) %>% 
+  filter(!is.na(variable)) %>% 
+  # group_by(sau_fishing_entity,fishing_entity_id,tresh,variable) %>% 
+  # summarise(sum_value = sum(value,na.rm=T)) %>% 
+  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
+  # Select top 5 most valuable stocks
+  top_n(5,value)
+
+top_taxon_list <-  unique(top_stocks_per_country$taxon_key)
+# Read proportion results
+suppressMessages(
+top_prop_results <- PropChange(top_taxon_list, links="y") %>%
+  filter(!is.na(eez_neighbour)) %>%
+  mutate(per_mid_plot =ifelse(per_mid >= 100, 100,
+                              ifelse(per_mid <= -100,-100,per_mid)
+  ),
+  per_early_plot =ifelse(per_early >= 100, 100,
+                         ifelse(per_early <= -100,-100,per_early)
+  )
+  )
+)
+
+# Prepare data from results
+top_data <- top_prop_results %>% 
+  # get coords for sources
+  left_join(Eez_centroid,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  left_join(Eez_centroid,
+            by = "eez_name") %>% 
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  mutate(
+    per_early = ifelse(is.na(per_early),0,per_early),
+    per_mid = ifelse(is.na(per_mid),0,per_mid)
+  ) %>% 
+  drop_na() %>%
+  # gather("variable","value",per_early:per_early_plot) %>% 
+  filter(
+    variable %in% c("mean_eez","n_spp")#,
+    # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
+  )
+
+# Convert to data to spatial
+top_network_data_sf <- top_data %>%
+  select(start_lon, start_lat, end_lon, end_lat) %>% 
+  purrr::transpose() %>% 
+  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
+  purrr::map(st_linestring) %>%
+  st_sfc(crs = 4326) %>%
+  # st_sfc(crs = "+proj=eck4") %>% # for tbular plot
+  st_sf(geometry = .)%>%
+  bind_cols(top_data) %>%
+  select(everything(), geometry) %>% 
+  # Set circles ('cus the world is not flat!)
+  st_segmentize(units::set_units(100, km)) %>% 
+  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>%
+  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>%
+  sf::st_sf(crs = 4326)
+  # sf::st_sf(crs = "+proj=eck4") # for tbular plot
+
+#  unpack the coordinates of each feature 
+coord_data <- as.data.frame( st_coordinates(st_cast(top_network_data_sf,"MULTILINESTRING")$geometry))
+
+c(top_network_data_sf$line_id, top_network_data_sf$coords) %<-% (
+  coord_data %>% 
+    rename(subline_id = L1) %>%   
+    nest(-L2)
+)
+
+# Create data path for line connections
+top_paths <- network_data_sf %>% 
+  sf::st_set_geometry(NULL) %>% 
+  unnest(cols = c(coords)) %>%
+  mutate_at(vars(X,Y), round,2) %>% 
+  gather("time_step","per_chng",per_early:per_mid) %>% 
+  mutate(per_plot = ifelse(per_chng > 50,50,per_chng),
+         direction = ifelse(per_chng > 0,"positive","negative")
+         )
+
+# Plot it!
+
+# Proportion map data 
+top_map_data <- top_data %>% 
+  gather("time_step","per_chng",per_early:per_mid) %>% 
+  mutate(per_plot = ifelse(per_chng > 50,50,per_chng))
+
+time <- "per_early"
+
+top_rop_map <- ggplot() +
+  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
+  geom_sf(data = World_map, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
+  geom_point(data = subset(top_map_data, variable =="mean_eez" & time_step == time & per_plot < 0), 
+             aes(start_lon,
+                 start_lat
+             ),
+             size = 0.05,
+             col = "gray80",
+             shape = 20,
+             stroke = 1.5) +
+  geom_path(data = subset(top_paths, variable =="mean_eez" & time_step == time & per_plot > 0), # filter data 
+            aes(X,
+                Y,
+                group = interaction(line_id,subline_id),
+                col = per_plot
+            ),
+            size = 0.9,
+            lineend = "round",
+            arrow = arrow(length = unit(0.008, "npc"),
+                          ends = "first"),
+            show.legend = NA) +
+  my_ggtheme_m(leg_pos = "bottom") +
+  scale_colour_viridis(
+    name= "Changes in catch proportion",
+    limits = c(0,max(Map_data$per_plot)),
+    breaks = seq(0,max(Map_data$per_plot),10),
+    labels = c(seq(0,max(Map_data$per_plot-10),10),">50")
+  ) + guides(
+      color=guide_colorbar(title.position="top", 
+                        title.hjust =0.5)
+  )
+
+ggsave(
+  plot = top_rop_map,
+  filename = "./Figures/top_ssr.png",
+  width = 12,
+  height = 8,
+  units = "in"
+)
+
+```
+
 ##### Fig S2 (Histogram of change)
@@
     ) +
-    my_ggtheme_m() +
+    my_ggtheme_m(leg_pos = "bottom") +
     facet_wrap(~method,

### 2021-11-12 Scripts/Manuscript.Rmd
Updates of the top stocks for SSR

--- 2021-11-12 Scripts/Manuscript.Rmd
@@
 
+**Having high-valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting.
+On average, countries will see a range shift of the top 5 most valuable stocks by 2038 $\pm$ 16 years (Figure S-NEW).
+Moreover, our projections suggest that these high value stocks are already shifting in 66% of coastal nations, with the regions of Polynesia (n = 22), Sub-Saharan Africa (n = 36) and Latin America and the Caribbean (n = 62) presenting the larger number of shifting stocks (20% of total high value stocks).**
+
 We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs for both thresholds (one and two s.d.).

### 2021-11-12 Scripts/responses_rev_gcb.Rmd
Updates of the top stocks for SSR

@@
 As one example, why is the geographic centroid (used in the calculation of TI) not weighted by probability of occurrence?
-Fish stocks are not evenly distributed across their entire range, and unweighted centroids can severely skew the actual distribution of the species.
+Fish stocks are not evenly distributed across their entire range, and un-weighted centroids can severely skew the actual distribution of the species.
 
-- **Response:** We agree with the reviewer that and unweighted centroid can severely skew the actual distribution of the species and influence our results.
+- **Response:** We agree with the reviewer that and un-weighted centroid can severely skew the actual distribution of the species and influence our results.
 We have revised the manuscript to clarify that we did account for the heterogeneous distribution of fish stocks across their entire range.
@@
 
-- **Response:** This is a very good observation and we believe it has increased the impact of our manuscript.
-We have included two new figure (**Supplement Figure XX**) showing the average range shift of the top 5 most valuable transboundary stocks per country in addition to the following text on line XX:
+- **Response:   (INCOMPLETE)** This is a very good observation and we believe it has increased the impact of our manuscript.
+We have included two new figures (**Supplement Figures XX and XX**) showing the average range shift and SSR of the top 5 most valuable transboundary stocks per country (Please note that figure 4 has been removed from the revised manuscript).
 
-  - Line XX: *Having high-valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting.
+  - Range shift, line XX: *Having high-valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting.
 On average, countries will see a range shift of the top 5 most valuable stocks by 2038 $\pm$ 16 years (Figure S-NEW).
@@
 
-We have made the following modifications to the manuscript:
+- SSR , line XX: **
 
-  - Line XX, *We estimated the participation of the transboundary stocks that are shifting between neighboring EEZs relative to the total revenue from all transboundary stocks that each EEZ has.
-On average, shifting transboundary stocks represented 27% to 23% (1 and 2 s.d.
-thresholds) of annual revenue from fisheries targeting transboundary stocks across the world's EEZs between 2004 and 2010 (Fig.
-\@ref(fig:toeFigure2)A).*
+![Top 5](../Figures/top_ssr.png)
 
@@
 
-- **Response: (Incomplete)** Many thanks for this observation. We have revised the whole manuscript to be sure verb tense is maintained through the document.
+- **Response: (INCOMPLETE)** Many thanks for this observation. We have revised the whole manuscript to be sure verb tense is maintained through the document.
 
@@
 
-- **Response:** This is indeed the definition of our stock. Please refer to our responses to the main comment.
+- **Response:** This is indeed the definition of our stock.
+We do not deferenciate between transboundary and straddling stocks as per defined in Palacios-Abanes *et al.* (2020).
+Please refer to our responses to the main comment.
 
@@
 
-- **Response:** Please see reposes to the reviewer's main comments. We have defined a trnaboundary stock as follows;
+- **Response:** Please see responses to the reviewer's main comments. We have defined a transboundary stock as follows;
 

### 2021-11-12 Scripts/Initial_Analysis.Rmd
Finalized new figure 4

@@
 
-```
-
-###  Year of Eergence Analysis
-
-```{r ToE_generals,  eval = F, echo = T}
-
 ### ------------ ##
@@
 ### ------------ ##
+
+```
+
+###  Year of Eergence Analysis
+
+```{r ToE_generals,  eval = F, echo = T}
 
@@
   filter(!is.na(variable)) %>% 
-  # group_by(sau_fishing_entity,fishing_entity_id,tresh,variable) %>% 
-  # summarise(sum_value = sum(value,na.rm=T)) %>% 
   group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
@@
   group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
-  # summarise(n_tax = length(unique(taxon_key))) %>% 
   summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
@@
 # Numbers for paper
-x <- top_toe %>% 
-  filter(mean_yr <= 2021,
-         variable == "value",
-         tresh == selected_tresh)
-
-length(unique(x$sau_fishing_entity))/length(unique(Sau_data$fishing_entity_id))
+Eme_Res %>% 
+  rename(geo_region = un_sub_region) %>% 
+  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
+  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
+  right_join(top_stocks_per_country,
+             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
+             ) %>% 
+  group_by(variable,tresh) %>% 
+  summarise_at(vars(mean_yr),
+               c("mean"=mean,
+                 "sd" = sd))
+
+# Number of coastal nations
+Eme_Res %>% 
+  rename(geo_region = un_sub_region) %>% 
+  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
+  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
+  right_join(top_stocks_per_country,
+             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
+             ) %>% 
+  filter(mean_yr <= 2021) %>% 
+  group_by(variable,tresh) %>% 
+  summarise(
+    n_countries = length(unique(sau_fishing_entity))
+  ) %>% 
+  mutate(per = n_countries/length(unique(Sau_data$fishing_entity_id)))
 
 # Number of shifting stocks per geo-region
-x %>% 
+Eme_Res %>% 
+  rename(geo_region = un_sub_region) %>% 
+  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
+  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
+  right_join(top_stocks_per_country,
+             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
+             ) %>% 
+  filter(mean_yr <= 2021) %>% 
   group_by(geo_region) %>% 
@@
   arrange(n) %>% 
-  top_frac(0.2,n) # top 20%
+  top_frac(0.25,n) # top 25%
 
@@
 
-mean(Top_map_data$mean)
-sd(Top_map_data$mean)
-
-World_map_top <- World_map %>%
+World_map_top <- World_eez_sf %>% 
+  left_join(Matching_names) %>% 
   full_join(Top_map_data,
@@
 
-# These are regions without transboundary species
-grey_land <- World_map %>% 
-  filter(!sau_fishing_entity %in% Top_map_data$sau_fishing_entity) %>% 
-  filter(!sovereignt %in% c("Antarctica"))
-
 # Map with countries without ToE
@@
     filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
-           !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
+           !eez_name %in% World_map_top$eez_name) %>%  # Filter EEZs with ToE stocks
     rbind(russia_churchil)
@@
     filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
-           !name %in% World_map_part$name) # Filter EEZs with ToE stocks
+           !name %in% World_map_top$name) # Filter EEZs with ToE stocks
 
@@
     # ggplot() +
-    geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
-    # geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
-    geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    geom_sf(data = World_map, aes(), fill = "grey50", color =  "white",size = 0.1) +
+    geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    # geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
     # Set Participation labels
-    scale_fill_viridis("Average year of range shift",
+    scale_fill_viridis("Average year of range shift of top valued transboundary stocks",
                        limits = c(2000,2100),

### 2021-11-12 Scripts/Manuscript.Rmd
Finalized new figure 4

--- 2021-11-12 Scripts/Manuscript.Rmd
@@
 
-We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs for both thresholds (one and two s.d.).
+**While the total number of shifting stocks is important for fisheries management, having high-valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting.
+Thus, we looked at the year of range shift of the top five most valuable transboundary stocks of each country (New Fig 4).
+On average, countries are projected to see a range shift of the top five most valuable stocks by 2038 $\pm$ 25 years under threshold two( 2028 $\pm$ 22 years under threshold one).
+Moreover, our projections suggest that these high value stocks are already shifting in 52 (27%) coastal nations under threshold two (n = 85 44% under threshold one), with the regions of Latin America and the Caribbean (n = 63), Sub-Saharan Africa (n = 59), South-eastern Asia  (n = 28), and Polynesia (n = 19) presenting the larger number of shifting stocks (25% of total high value stocks).**
+
+<!-- We estimated the current (i.e., 2005 - 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs for both thresholds (one and two s.d.).
 Within five years of the UN 2030 Agenda timeline, 979 transboundary stocks accounting for USD 5.3 billion in fishing revenue are projected to shift within 76 EEZs considering a two s.d.
@@
 In this case, 2,136 stocks accounting for USD 24 billion in fishing revenue are projected to shift within five years of the UN 2030 Agenda timeline in 134 EEZs.
+-->
 

### 2021-11-12 Scripts/responses_rev_gcb.Rmd
Finalized new figure 4

@@
 Replicating this analysis at the species level will undoubtedly result in multiple countries sharing each species.
-For example, in the case of Atlantic Cod, this would be shared by US, Canada and Greenland (and maybe enven northern Europe).
+For example, in the case of Atlantic Cod, this would be shared by US, Canada and Greenland (and maybe even northern Europe).
 Choosing this approach invariably comes with the assumption that fishing Cod in, e.g., the Gulf of Maine, will have direct impact on e.g., Greenland fisheries.
@@
 
-- **Response:** Please see responses to the reviewer's main comments. We have defined a transboundary stock as follows;
+- **Response:** The definition is now clearly stated in the revised manuscript as follows;
 
@@
 In the original manuscript we were showing the percentage difference between the two numbers (i.e., $x = \frac{|45-63|}{|\frac{45+63}{2}|}*100$), hence the 33.3% difference.
-However, now we realize this can be misleading and have adopted a more straightforward computation.
+However, now we realize this can be misleading and have adopted a more straightforward computation that aligns with the reviewer's math.
 We also corrected the 63% typo (not 67%).
@@
 
-- **Response:** Thank you for your commnets on our fiugre 3. We have devided the responses by each point you make. Please see below
+- **Response:** Thank you for your comments on our figure 3 and see below the responses to each point you make.
 
@@
 
-  - **Response:**  **NOTE: I do not (fully) agree with this comment.
-yes, the figure is complex, but I do not think having different color scales will make it any less complex.
-I actually think it will make it even harder to read.
-Do you think we can counter-argue this comment?**
+  - **Response:**  **NOTE: I do not (fully) agree with this comment, Do you think we can counter-argue it?
+(We had the same type of plot in the Sci.
+Rep.
+paper).
+Yes, the figure is complex, but I do not think having different color scales will make it any less complex.
+I actually think it will make it even harder to read.**
 
@@
 
-  - **Response:** In the case of Fig.3B we used the median because this figure talks directly to the statistical non-parametric analysis (Kruskal-Wallis) looking at differences in the time of emergence by geographic region.
-**NOTE: Shall we change Fig.3A to show the median instead of the mean, for the sake of consistency (and make the reviewer happy)?**
+  - **Response:** In the case of Fig.3B we used the median because this figure talks directly to the statistical non-parametric analysis (Kruskal-Wallis) looking at differences in the time of emergence by geographic region .
+**NOTE: Shall we change Fig.3A to show the median instead of the mean, for the sake of consistency or the inverse?
+(and make the reviewer happy)?**
 
@@
 
-- **Response:** We intended to show this information in Figure 4.
-However, the message in the figure is not clear, as the reviewer pointed out in the next comment.
-We have modified figure 4 to show this as a figure.
+- **Response:** We agree with the reviewer that showing the biggest “winners and losers” in terms of high-value species would be interesting.
+Thus, we have included a new figure and text in this section showing the year of range shift of the 5 most valuable stocks per country (see response to main comments).
 
@@
 
-- **Response:**
+- **Response:NOTE THINKING ON CHANGEING FIGURES** Our original intention was to show the importance in terms of value, of shifting stocks.
+However, we agree that this figure is confusing and it's message was ambiguous.
+Moreover, we think that the reviewers suggestion of the most valuable species brings home a mor important message and have decided to remove this figure from the revised manuscript and include the "top 5 most valuable stocks" figure instead.
 

### 2021-11-13 Scripts/Initial_Analysis.Rmd
First draft of responses to Rev 1

@@
   # Select top 5 most valuable stocks
-  top_n(5,value)
+  top_n(5,value) %>% 
+  left_join(Exploited_spp_list)
 
@@
 ### Median number of stocks changing in each time step
-
 Prop_results %>% 
@@
   mutate(time_step = ifelse(time_step == "early","per_early","per_mid")) %>%
-  mutate(avrg_spp_per = ifelse(avrg_spp_per > 50,50,avrg_spp_per))
+  mutate(avrg_spp_per_100 = avrg_spp_per,
+    avrg_spp_per = ifelse(avrg_spp_per > 50,50,avrg_spp_per))
+
+# Paper numbers for Rev. 1
+
+n_taxa_data %>% 
+  group_by(time_step) %>% 
+  summarise_at(vars(avrg_spp_per),
+               c(mean,sd,min,max),
+               na.rm=T)
 
@@
 time_sets <- c("per_early","per_mid")
-# i = 1
+i = 1
 for(i in 1:2){
@@
               ),
-              size = 0.3,
+              size = 1,
               lineend = "round",
@@
               show.legend = NA) +
-    # Closeup Rectangles 
-    # geom_rect(data=Rectangles,
-    #           aes(xmin=x1, 
-    #               xmax=x2,
-    #               ymin=y1,
-    #               ymax=y2),
-    #           color="grey50", 
-    #           fill = NA,
-    #           linetype = "dashed"
-    # ) +
     my_ggtheme_m(leg_pos = "bottom") +
@@
     ) #+
-  # ggsave("./Figures/fig_sup_1.png",
-  # width = 12,
-  # height = 8,
-  # units = "in"
-  # )
+
+  ggsave(
+    plot = Prop_map,
+    filename = "./Figures/fig_5n.png",
+    width = 10,
+    height = 5,
+    units = "in"
+  )
 
@@
 
+### Median number of stocks changing in each time step
+
+top_prop_results %>% 
+  filter(variable == "n_per") %>% 
+  group_by() %>% 
+  summarise_at(vars(early,mid),funs(
+    median(.,na.rm=T),
+    sd(.,na.rm=T),
+    length(unique(eez_name)),
+    length(unique(eez_name))/280*100
+  )
+  )
+
 # Prepare data from results
@@
 
+View(top_data)
+
 # Convert to data to spatial
@@
   geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
-  geom_sf(data = World_map, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
+  geom_sf(data = subset(World_map,sovereignt != "Antarctica"), size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
+
   geom_point(data = subset(top_map_data, variable =="mean_eez" & time_step == time & per_plot < 0), 
@@
   filename = "./Figures/top_ssr.png",
-  width = 12,
-  height = 8,
+  width = 10,
+  height = 5,
   units = "in"

### 2021-11-13 Scripts/Manuscript.Rmd
First draft of responses to Rev 1

@@
 Thus, we looked at the year of range shift of the top five most valuable transboundary stocks of each country (New Fig 4).
-On average, countries are projected to see a range shift of the top five most valuable stocks by 2038 $\pm$ 25 years under threshold two( 2028 $\pm$ 22 years under threshold one).
+On average, countries are projected to see a range shift of the top five most valuable stocks by 2038 $\pm$ 25 years under threshold two (2028 $\pm$ 22 years under threshold one).
 Moreover, our projections suggest that these high value stocks are already shifting in 52 (27%) coastal nations under threshold two (n = 85 44% under threshold one), with the regions of Latin America and the Caribbean (n = 63), Sub-Saharan Africa (n = 59), South-eastern Asia  (n = 28), and Polynesia (n = 19) presenting the larger number of shifting stocks (25% of total high value stocks).**
@@
 
-```{r toeFigure3, eval = T,  echo = F, fig.width=8, fig.height=6, fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks and corresponding fishing revenue per country.
-Countries are coded according to different colored symbols representative of 5 distinct regions.
-The horizontal line separates countries above and below the 75th percentile of annual fishing revenue.
-Showing country names for the eleven nations above the 75th percentile and projected to have an average shift across transboundary stocks within its EEZ within 5 years of the 2030 Agenda deadline."}
+```{r toeFigure3, eval = T,  echo = F, fig.width=8, fig.height=6, fig.pos= "H", fig.cap =  "Year of change in the shared distribution of the top five most valuable transboundary stocks of each country.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue."}
 
@@
 Panel B zooms into specific areas shown within the gray dashed line boundaries in A.](../Figures/Fig1.png) -->
+
+**Most EEZs are expected to see a relative small number of stocks shifting their SSR (Fig.
+\@ref(fig:toeFigure4)A).
+On average, EEZs are expected to see changes in SSR on 16% $\pm$ 10 of transboundary stocks increasing to 20% $\pm$ 11 by mid of the $21^{st}$ century.
+However, this result masks mask large variation in changes across EEZs.
+By early century, the EEZs of French Guiana (54%), South Africa (59%), Falkland Isl.
+(61%), Brazil (75%) and Kerguelen Isl., and Pitcairn (both 100%) are projected to see changes in SSR of over 50% in their transboundary stocks.
+On the other hand, some island territories such as Anguilla are projected to see changes in as low as 2% of their transboundary stocks.**
+
+**Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 61 $\pm$ 18% by 2030 and 72 $\pm$ 17% by 2050 in 83% of the world's EEZs (Figure SX).
+In some cases, the SSR of one EEZ is projected increase in more than double relative to their neighbors (e.g., from 10% to 30% SSR).
+Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Cyprus from Greece and Turkey in the Mediterranean Sea, and Russia from Norway and Japan in the Barents Sea and Western Pacific, respectively**
 

### 2021-11-13 Scripts/responses_rev_gcb.Rmd
First draft of responses to Rev 1

@@
 
-Major comments:
+## Major comments:
 I am concerned about the lack of a clear definition of “fish stock” in this manuscript, since it is absolutely central to the analysis.
@@
 
-    - Methods Line XX, *Let us consider the transboundary species Atlantic cod (*Gadus morhua*) as an example.
+  - Methods Line XX, *Let us consider the transboundary species Atlantic cod (Gadus morhua) as an example.
 Atlantic cod's distribution in the Northwest Atlantic ranges from the United States to Greenland.
@@
 We have revised the manuscript to clarify that we did account for the heterogeneous distribution of fish stocks across their entire range.
-The geographic centroid of each stock is indeed weighted by the abundance of the stock.
+Specifically, the geographic centroid of each stock is indeed weighted by the abundance of the stock.
 The DBEM uses the species probability of occurrence to estimate the biomass of the species at each 0.5 x 0.5 gridcell as indicated from lines XX to XX of the methods.
@@
 
-- **Response:   (INCOMPLETE)** This is a very good observation and we believe it has increased the impact of our manuscript.
-We have included two new figures (**Supplement Figures XX and XX**) showing the average range shift and SSR of the top 5 most valuable transboundary stocks per country (Please note that figure 4 has been removed from the revised manuscript).
+- **Response:** This is a very good observation and we believe it has increased the impact of our manuscript.
+We have included two new figures (**Figure 4** and **Figure SXX**) showing the average range shift of the top 5 most valuable transboundary stocks per country (Please note that the old figure 4 has been removed from the revised manuscript, see response to the Fig.
+4 comment, below).
 
   - Range shift, line XX: *Having high-valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting.
-On average, countries will see a range shift of the top 5 most valuable stocks by 2038 $\pm$ 16 years (Figure S-NEW).
+On average, countries will see a range shift of the top 5 most valuable stocks by 2038 $\pm$ 16 years (Figure SXX).
 Moreover, our projections suggest that these high value stocks are already shifting in 66% of coastal nations, with the regions of Polynesia (n = 22), Sub-Saharan Africa (n = 36) and Latin America and the Caribbean (n = 62) presenting the larger number of shifting stocks (20% of total high value stocks).*
@@
 
-- SSR , line XX: **
+## Specific comments:
 
-![Top 5](../Figures/top_ssr.png)
-
-Specific comments:
-
-Abstract:
+### Abstract:
 Line 10: what does this line mean: “showing early, and sometimes past shifts”?
@@
 
-Introduction:
+### Introduction:
 Be careful of verb tense. The Introduction mixes past with present tense. For example, the first two sentences starting at Line 57.
@@
 - **Response: (INCOMPLETE)** Many thanks for this observation. We have revised the whole manuscript to be sure verb tense is maintained through the document. 
-
-Methods:
+### Methods:
 
@@
 - **Response:** This is indeed the definition of our stock.
-We do not deferenciate between transboundary and straddling stocks as per defined in Palacios-Abanes *et al.* (2020).
+We do not differentiate between transboundary and straddling stocks as per defined in Palacios-Abanes *et al.* (2020).
 Please refer to our responses to the main comment.
@@
 - **Response:** Thank you for your concern.
-In short the answer is yes, an EEZ could both lose and gain SSR of the same *species* but not of the same *stock.* This is because, according to our definition of stock, a species who spans multiple EEZs has multiple stocks (see response to the previous question and the main comment).
-let us use the Peruvian anchoveta (*Engraulis ringens*) as an example.
+In short the answer is yes, an EEZ *could* both lose and gain SSR of the same *species* but not of the same *stock.* This is because, according to our definition of stock, a species who spans multiple EEZs has multiple stocks (see response to the previous question and the main comment).
+Please refer above to our response to the question *How does the choice of stock definition influence the calculation of SSR?*
 
-Peru has identified two main stocks of Peruvian anchoveta.
-one in the central/norther part of the EEZ and one in the south, shared with Chile.
-Although the northern stocks is not considered transboundary [@Cashion:2018cg; @PalaciosAbrantes:2020tv] and Equador hardly ever reports anchoveta landings (wwww.searaundus.org), in some few years, environmental variability has pushed the central/north stock towards the EEZ of Ecuador.
-In those years one could argue that, the Peruvian anchoveta SSR of Peru-Ecuador favors Ecuador.
-
-At the same time, there are regional concerns that the Peruvian southern stock of anchoveta, shared with Chile, has been slowly shifting south into Chilean waters to the point that these two nations are now pushing towards cooperative management of such stock.
-In this case, the SSR of Peru-Chile, favors Chile.
-
-This is only and example as our study is not capable of capturing this specific nuisance since we are not able to differentiate stocks within an EEZ nor more detailed, local, environmental factors (in this case, for example, the Humboldt current).
-However, we believe this
-
-Results:
+### Results:
 
@@
 
-- **Response:NOTE THINKING ON CHANGEING FIGURES** Our original intention was to show the importance in terms of value, of shifting stocks.
+- **Response:NOTE THINKING ON CHANGING FIGURES** Our original intention was to show the importance in terms of value, of shifting stocks.
 However, we agree that this figure is confusing and it's message was ambiguous.
-Moreover, we think that the reviewers suggestion of the most valuable species brings home a mor important message and have decided to remove this figure from the revised manuscript and include the "top 5 most valuable stocks" figure instead.
+Moreover, we think that the reviewer's suggestion of the most valuable species brings home a more important message and have decided to remove this figure from the revised manuscript and include the "top 5 most valuable stocks" figure instead.
 
@@
 
-- **Response:**
+- **Response:**  Thank you for your comment. See below responses to both parts of the question.
+
+Fig. 5- this figure is hard to read,
+
+- **Response:** We agree with the reviewer that this figure is quite busy and have divided the figure in two.
+Figure 5 is now just panel A while the regional zooms were sent to the supplemental material.
+We believe that the arrows are now easier to appreciate.
+
+![New Figure 5](../Figures/Fig5n.png)
+
+![New Figure SX for regional](../Figures/Fig5SX_ssr)
+
+and is another figure that may be misleading if not weighted by abundance or value. It is strange to “average” across a unitless ratio (SSR).
+
+- **Response:** We would like to clarify here that that the Stock Share Ratio (SSR) is not unit less.
+It represents the percentage of the shared distribution of a stock that each EEZ has.
+So for example, Canada and the US share Pacific halibut the SSR for Canada is 20% and that of the US is 80%, meaning that Canada holds 20% of the distribution of the stock within its waters and the US the other 80%.
+We have modified the methods to make sure this is clearer now:
+
+  - - *This way, if, Canada and the US share a specific stock the SSR for Canada could be 20% and that of the US 80%, meaning that Canada holds 20% of the distribution of the stock within its waters and the US the other 80%.*
+
+For example, a country could “gain” a small share of an extremely high-value species, perhaps enough to offset economically a number of losses of smaller or less valuable stocks.
+However, in the figure, the “average” SSR for that country would be negative (arrow going out).
+The authors should distinguish (in the text and in this figure) different possible situations like this.
+
+- **Response:** Yes, we agree with the reviewer and just like for Fig.
+3, believe including a figure of high-value stocks will increased the impact of our manuscript.
+We have included such figure in the supplements and the following text in the results section:
+
+- line XX: **Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 61 $\pm$ 18% by 2030 and 72 $\pm$ 17% by 2050 in 83% of the world's EEZs (Figure SX).
+In some cases, the SSR of one EEZ is projected increase in more than double relative to their neighbors (e.g., from 10% to 30% SSR).
+Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Cyprus from Greece and Turkey in the Mediterranean Sea, and Russia from Norway and Japan in the Barents Sea and Western Pacific, respectively**
+
+![Top 5](../Figures/Fig5SXX_top.png)
 
@@
 
-- **Response:**
+- **Response:** We show that result in Figure 5 in the land polygon as the percentage of stocks that are changing but realize we did not include an in-text description of the result.
+It now reads on line XX:
 
-Discussion
+  - - **Most EEZs are expected to see a relative small number of stocks shifting their SSR (Fig.
+\@ref(fig:toeFigure4)A).
+On average, EEZs are expected to see changes in SSR on 16% $\pm$ 10 of transboundary stocks increasing to 20% $\pm$ 11 by mid of the $21^{st}$ century.
+However, this result masks large changes, specially on island territories.
+By early century, the EEZs of French Guiana (54%), South Africa (59%), Falkland Isl.
+(61%), Brazil (75%) and Kerguelen Isl., and Pitcairn (both 100%) are projected to see changes in SSR of over 50% in their transboundary stocks.
+On the other hand, some island territories such as Anguilla are projected to see changes in as low as 2% of their transboundary stocks.**
+
+### Discussion
 

### 2021-11-15 Scripts/Initial_Analysis.Rmd
First draft towards co-authors

@@
   ) %>% 
-  filter(!is.na(participation))
+  filter(!is.na(participation)) %>% 
+sf::st_transform(crs = "+proj=eck4")
 
@@
   filter(!fishing_entity_id %in% Value_map_data$fishing_entity_id) %>% 
-  filter(!sovereignt %in% c("Antarctica"))
+  filter(!sovereignt %in% c("Antarctica"))%>% 
+sf::st_transform(crs = "+proj=eck4")
 
@@
     filter(tresh == tresh_list[i],
-           metric == "mean_yr")
+           metric == "mean_yr")%>% 
+sf::st_transform(crs = "+proj=eck4")
 
@@
   russia_churchil <- World_eez_sf %>% 
-    filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))
+    filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))%>% 
+sf::st_transform(crs = "+proj=eck4")
 
@@
            !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
-    rbind(russia_churchil)
+    rbind(russia_churchil)%>% 
+sf::st_transform(crs = "+proj=eck4")
 
@@
     filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
-           !name %in% World_map_part$name) # Filter EEZs with ToE stocks
+           !name %in% World_map_part$name)%>% 
+sf::st_transform(crs = "+proj=eck4")  # Filter EEZs with ToE stocks
 
@@
 
-##### Fig. N3 Mapping the time of emergence Of Top stocks
+##### Fig. 4N Mapping the time of emergence Of Top stocks
 
@@
   ) %>% 
-  filter(!is.na(variable))
+  filter(!is.na(variable)) %>% 
+  sf::st_transform(crs = "+proj=eck4")
 
@@
   russia_churchil <- World_eez_sf %>% 
-    filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))
+    filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))%>% 
+    sf::st_transform(crs = "+proj=eck4")
 
   no_toe_eez <- World_eez_sf %>% 
+    sf::st_transform(crs = "+proj=eck4") %>% 
     filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
            !eez_name %in% World_map_top$eez_name) %>%  # Filter EEZs with ToE stocks
-    rbind(russia_churchil)
+    rbind(russia_churchil)%>% 
+    sf::st_transform(crs = "+proj=eck4")
 
@@
     filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
-           !name %in% World_map_top$name) # Filter EEZs with ToE stocks
+           !name %in% World_map_top$name
+           ) %>% 
+    sf::st_transform(crs = "+proj=eck4")
 
@@
     # ggplot() +
-    geom_sf(data = World_map, aes(), fill = "grey50", color =  "white",size = 0.1) +
+    geom_sf(data = subset(World_map, sovereignt != "Antarctica"), aes(), fill = "grey50", color =  "white",size = 0.1) +
     geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
@@
     "./Figures/top_toe_map.png",
-    width = 10,
+    width = 8,
     height = 5,
@@
   st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>%
-  sf::st_sf(crs = 4326)
-  # sf::st_sf(crs = "+proj=eck4") # for tbular plot
+  # sf::st_sf(crs = 4326)
+  sf::st_sf(crs = "+proj=eck4") # for tbular plot
 
@@
 time <- "per_early"
+
+World_eez_sf <- World_eez_sf %>% 
+sf::st_transform(crs = "+proj=eck4")
+
+World_map <- World_map %>% 
+sf::st_transform(crs = "+proj=eck4")
 

### 2021-11-15 Scripts/Manuscript.Rmd
First draft towards co-authors

@@
 We did this by aggregating the projected catch within the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock's catch held within each EEZ [@PalaciosAbrantes:2020tv].
-Second, to reduce the effects of variability, we averaged the calculated proportion of stock occupying a given EEZ into three time periods: the historical time period ($SSR_{th}$) from 1951 to 2005, and two future periods, the early $21^{st}$ century, between 2021 and 2040 ($SSR_{te}$), and the mid $21^{st}$ century, defined as the average of 2041 to 2060 ($SSR_{tm}$).
+*This way, if, Canada and the US share a specific stock the SSR for Canada could be 20% and that of the US 80%, meaning that Canada holds 20% of the distribution of the stock within its waters and the US the other 80%.*Second, to reduce the effects of variability, we averaged the calculated proportion of stock occupying a given EEZ into three time periods: the historical time period ($SSR_{th}$) from 1951 to 2005, and two future periods, the early $21^{st}$ century, between 2021 and 2040 ($SSR_{te}$), and the mid $21^{st}$ century, defined as the average of 2041 to 2060 ($SSR_{tm}$).
 These future time periods were chosen to correspond to the challenges of achieving fisheries-relevant UN-SDGs by 2030 (i.e., the 2030 Agenda), such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) and SDG 1.2 (poverty reduction) [@Singh:2017ds; @UnitedNations:1Qe8KIcN].
@@
 
-**Most EEZs are expected to see a relative small number of stocks shifting their SSR (Fig.
+**Most EEZs are expected to see a relatively small number of stocks shifting their SSR (Fig.
 \@ref(fig:toeFigure4)A).
 On average, EEZs are expected to see changes in SSR on 16% $\pm$ 10 of transboundary stocks increasing to 20% $\pm$ 11 by mid of the $21^{st}$ century.
-However, this result masks mask large variation in changes across EEZs.
-By early century, the EEZs of French Guiana (54%), South Africa (59%), Falkland Isl.
-(61%), Brazil (75%) and Kerguelen Isl., and Pitcairn (both 100%) are projected to see changes in SSR of over 50% in their transboundary stocks.
+However, this result masks large variation in changes across EEZs.
+By early century, the EEZs of French Guiana (54%), South Africa (59%), Falkland Island (61%), Brazil (75%) and Kerguelen Island, and Pitcairn (both 100%) are projected to see changes in SSR of over 50% in their transboundary stocks.
 On the other hand, some island territories such as Anguilla are projected to see changes in as low as 2% of their transboundary stocks.**
@@
 **Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 61 $\pm$ 18% by 2030 and 72 $\pm$ 17% by 2050 in 83% of the world's EEZs (Figure SX).
-In some cases, the SSR of one EEZ is projected increase in more than double relative to their neighbors (e.g., from 10% to 30% SSR).
+In some cases, the SSR of one EEZ is projected to increase more than double relative to their neighbors (e.g., from 10% to 30% SSR).
 Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Cyprus from Greece and Turkey in the Mediterranean Sea, and Russia from Norway and Japan in the Barents Sea and Western Pacific, respectively**
@@
 
-Our results identify regional “hotspots” of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figs.
+**The SDG target 17.16 (Enhance the global partnership for sustainable development) calls for collaborations between stakeholders to support the achievement of the SDGs in all countries, in particular developing countries [@SDGREF].
+Such collaborations will be critical if we are to sustainably manage transboundary fisheries under climate change [@Pinsky:2018cb; @Oremus:2020id].** Our results identify regional “hotspots” of climate risk for transboundary fisheries management that will require the prompt adaptation of collaborative management plans (Figs.
 \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure4)A).
@@
 \@ref(fig:toeFigure2)B).
+
 Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and *ii*) model structure [@Frolicher:2016dg].
-First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.Thus, any mitigation efforts could result in a substantial delay in distribution shifts as environmental signals are responsive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].
+First, our analysis is based on a high emission scenario (RCP 8.5) representing an "extreme" case of shifts in the shared distribution of transboundary species.
+Thus, any mitigation efforts could result in a substantial delay in distribution shifts as environmental signals are responsive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg].
 Second, a substantial source of uncertainty is related to model selection for both climate change and fish distribution.
@@
 It is unlikely that these complex uncertainties will adequately be addressed in the coming decade.
-This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, and the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf].
+**NOTE OPTION A.
+Therefore, decision-makers are facing the challenge of taking action under an uncertain future.
+One strategy for that is adapting scenario planning into fisheries management [@NOAA_scen_plan_2020].
+This will allow for managers to integrate the uncertainty into planning for managing fisheries.
+For example, if stocks are expected to shift their distribution by 2036 $\pm$ 28 years (Fig.
+2), managers could prepare for a scenario where stocks are currently changing and another where they will shift in 20 years.
+Scenario planning can also be used to estimate potential looses in revenue when adaptation to climate change is not developed cooperatively [@Sumaila:2020tp; @Miller:2013iv].** This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, and the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf].
+
 These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh].
@@
 Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
+
+**NOTE: OPTION B, From a decision-making point of view, these uncertainties hinder the management of natural resources forcing managers to taking action under an uncertain future.
+Rather than wait for more certain methods to be developed (in some cases is unlikely that we could develop them in the next decade or so, Forlicher *et al.* 2016), manager could build scenario planning into fisheries management [@NOAA_scen_plan_2020].
+This will allow for managers to integrate the uncertainty into planning for managing fisheries.
+For example, if stocks are expected to shift their distribution by 2036 $\pm$ 28 years (Fig.
+2), managers could prepare for a scenario where stocks are currently changing and another where they will shift in 20 years.
+Scenario planning can also be used to estimate potential looses in revenue when adaptation to climate change is not developed cooperatively [@Sumaila:2020tp; @Miller:2013iv].**
 

### 2021-11-15 Scripts/Initial_Analysis.Rmd
Cylindrical maps

@@
     annotate("text",
-             x =-150,
-             y = -69,
-             label = "Transboundary fishing revenue (%)",
+             x = -14012580,
+             y = -8357450,
+             label = "Transboundary fish revenue (%)",
              color="black",
@@
     annotate("text",
-             x = c(-72,-35,5,46,82),
-             y = -69,
+             x = c(-3812580,-4500,3800000,7700000,11512580),
+             y = -8400000,
              # x =(c(-8500000,-3000000,2500000,7500000,12800000)),
@@
              hjust=0) +
-    # Set ToE labels
+    # # Set ToE labels
     annotate("text",
-             x =-150,
-             y = -85,
+             # x =-150,
+             # y = -85,
+             x = -13012580,
+             y = -9657450,
              label = "Average year of range shift",
@@
     annotate("text",
-             x = c(-79,-39,2,42,81),
-             y = -85,
+             # x = c(-79,-39,2,42,81),
+             # y = -85,
+             x = c(-4412580,-430000,3410000,7300000,11312580),
+             y = -9657450,
              # label = seq(2100,2000,-25), # original
@@
              hjust=0) +
-    # Set No emergence label
+    # # Set No emergence label
     annotate("text",
-             x =110,
-             y = -77,
+             # x =110,
+             # y = -77,
+             x = 13000000,
+             y = -9000000,
              label = "No shift",
@@
              hjust=0) +
+     annotate("rect", 
+             xmin = 15500000, 
+             xmax = 16500000, 
+             ymin = -8600000, 
+             ymax = -9300000,
+           fill = "lightblue") +
     scale_fill_viridis("",
@@
     ) +
-    # scale_fill_gradientn("",
-    #                      limits = c(2000,2100),
-    #                      breaks = seq(2000,2100,25),
-    #                       labels = c("","","","",""),
-    #                      colours = pal
-    # ) +
-    # geom_hline(yintercept = c(-50,0,50,80)) +
     my_ggtheme_m(leg_pos = "bottom") +
     theme(
-      legend.box.margin = margin(t = -62, l = 252),
+      legend.box.margin = margin(t = -61,l = 180),
       legend.text = element_blank(),
+      legend.background = element_blank(),
       legend.key.width = unit(5, "line")
@@
       axis.text.y = element_text(size = 16), 
+      panel.background = element_rect(colour = "white", fill = "white"),
       # axis.ticks.x = element_blank(),
@@
     # Time of Emergence Map
-    draw_plot(map, x = 0, y = 0.4, width = 1, height = 0.6) +
+    draw_plot(map, x = 0, y = 0.39, width = 1, height = 0.70) +
     # Regional boxplot
@@
                     x = c(0, 0), 
-                    y = c(0.95, 0.46)
-    )
+                    y = c(0.98, 0.46)
+    ) + 
+  theme(plot.background = element_rect(fill="white", color = NA))
 
@@
     name_save,
+      # "./Figures/test.png",
     p,
@@
   )
-
   }

### 2021-11-15 Scripts/Initial_Analysis.Rmd
Finally created cilindrical SSR maps!!!!!

@@
             ) %>% 
-  st_set_crs(4326) #%>% # to match shapefiles
-  # st_transform(crs = "+proj=eck4")
+  st_set_crs(4326) %>% # to match shapefiles
+  st_transform(crs = "+proj=eck4")
+
+ggplot(Eez_centroid_sf) +
+  geom_sf()
 
 # Transform sf to df
-# Eez_centroid_eck <- as.data.frame(Eez_centroid_sf)
-
-# Eez_centroid_eck <- do.call(rbind, st_geometry(Eez_centroid_sf)) %>% 
-#   as_tibble() %>% 
-#   setNames(c("lon","lat")) %>% 
-#   bind_cols(Eez_centroid) %>% 
-  # select(eez_name,longitude = lon,latitude = lat)
+Eez_centroid_eck <- as.data.frame(Eez_centroid_sf)
+
+Eez_centroid_eck <- do.call(rbind, st_geometry(Eez_centroid_sf)) %>%
+  as_tibble() %>%
+  setNames(c("lon","lat")) %>%
+  bind_cols(Eez_centroid) %>%
+select(eez_name,longitude = lon,latitude = lat)
 
@@
 
-```{r fig_map_Juan, eval = F, echo =F}
-# Figure made thanks to Juanito!
-# http://jsmayorga.com/post/mapping-the-global-network-of-transnational-fisheries/
-
-### Median number of stocks changing in each time step
-Prop_results %>% 
-  filter(variable == "n_per") %>% 
-  group_by() %>% 
-  summarise_at(vars(early,mid),funs(
-    median(.,na.rm=T),
-    sd(.,na.rm=T),
-    length(unique(eez_name)),
-    length(unique(eez_name))/280*100
-  )
-  )
+```{r}
+
+# Done with the help of https://www.data-to-viz.com/story/MapConnection.html
+
+# A function that makes a dateframe per connection (we will use these connections to plot each lines)
+data_for_connection=function( dep_lon, dep_lat, arr_lon, arr_lat, group){
+  inter <- gcIntermediate(c(dep_lon, dep_lat), c(arr_lon, arr_lat), n=50, addStartEnd=TRUE, breakAtDateLine=F)             
+  inter=data.frame(inter)
+  inter$group=NA
+  diff_of_lon=abs(dep_lon) + abs(arr_lon)
+  if(diff_of_lon > 180){
+    inter$group[ which(inter$lon>=0)]=paste(group, "A",sep="")
+    inter$group[ which(inter$lon<0)]=paste(group, "B",sep="")
+  }else{
+    inter$group=group
+  }
+  return(inter)
+}
+pal <- wesanderson::wes_palette("Zissou1",100,type = "continuous")
 
@@
   left_join(Eez_centroid,
+  # left_join(Eez_centroid_eck,
             by = "eez_name") %>%
@@
   # get coords for targets
+  # left_join(Eez_centroid_eck,
   left_join(Eez_centroid,
@@
   ) %>% 
-  drop_na() %>%
-  # gather("variable","value",per_early:per_early_plot) %>% 
+  drop_na()
+
+# Get paths data
+path_data <- Data %>% 
   filter(
-    variable %in% c("mean_eez","n_spp")#,
+    variable %in% c("mean_eez")#,
     # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
@@
 
-# Convert to data to spatial
-network_data_sf <- Data %>%
-  select(start_lon, start_lat, end_lon, end_lat) %>% 
-  purrr::transpose() %>% 
-  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
-  purrr::map(st_linestring) %>%
-  st_sfc(crs = 4326) %>%
-  # st_sfc(crs = "+proj=eck4") %>% # for tbular plot
-  st_sf(geometry = .)%>%
-  bind_cols(Data) %>%
-  select(everything(), geometry) %>% 
-  # Set circles ('cus the world is not flat!)
-  st_segmentize(units::set_units(100, km)) %>% 
-  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>%
-  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>%
-  sf::st_sf(crs = 4326)
-  # sf::st_sf(crs = "+proj=eck4") # for tbular plot
-
-#  unpack the coordinates of each feature 
-coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
-
-c(network_data_sf$line_id, network_data_sf$coords) %<-% (
-  coord_data %>% 
-    rename(subline_id = L1) %>%   
-    nest(-L2)
-)
-
-# Create data path for line connections
-paths <- network_data_sf %>% 
-  sf::st_set_geometry(NULL) %>% 
-  unnest(cols = c(coords)) %>%
-  mutate_at(vars(X,Y), round,2) %>% 
-  gather("time_step","per_chng",per_early:per_mid) %>% 
-  mutate(per_plot = ifelse(per_chng > 50,50,per_chng),
-         direction = ifelse(per_chng > 0,"positive","negative")
-         )
-
-# Plot it!
-
-# Proportion map data 
-Map_data <- Data %>% 
-  gather("time_step","per_chng",per_early:per_mid) %>% 
-  mutate(per_plot = ifelse(per_chng > 50,50,per_chng))
-
-# Average number of taxa to change per eez
-
-# Total number of transboundary stocks per EEZ
+data_ready_plot=data.frame()
+
+for(i in c(1:nrow(path_data))){
+  tmp=data_for_connection(path_data$start_lon[i], path_data$start_lat[i], path_data$end_lon[i], path_data$end_lat[i] , i)
+  # tmp$homecontinent=summary$homecontinent[i]
+  tmp$per_early_plot=path_data$per_early_plot[i]
+  tmp$per_mid_plot=path_data$per_mid_plot[i]
+  tmp$source=path_data$source[i]
+  tmp$objective=path_data$target[i]
+  data_ready_plot=rbind(data_ready_plot,tmp)
+}
+
+data_ready_plot <- data_ready_plot %>% 
+  mutate(
+    per_early_plot = ifelse(per_early_plot >= 50,50,per_early_plot),
+    per_mid_plot = ifelse(per_mid_plot >= 50,50,per_mid_plot))
+
+sf_ready_plot <- st_as_sf(data_ready_plot,
+            coords = c("lon", "lat")
+            ) %>% 
+  st_set_crs(4326) %>% # to match shapefiles
+  st_transform(crs = "+proj=eck4")
+
+# Get arrow data
+arrows_data <- as.data.frame(st_coordinates(st_cast(sf_ready_plot,"POINT")$geometry)) %>% 
+  bind_cols(data_ready_plot)
+
+# Get point data
+point_data <- path_data %>% 
+  filter(per_early <= 0) %>% 
+  st_as_sf(
+    coords = c("start_lon", "start_lat")
+  ) %>% 
+  select(source,target,geometry) %>% 
+  st_set_crs(4326) %>% 
+  st_transform(crs = "+proj=eck4")
+
+#### Get the change of species
+
 n_trans_eez <- Transboundary_spp %>% 
@@
 
+# Included in map
+n_taxa_map <- World_map %>% 
+  full_join(n_taxa_data,
+            by = c("sau_fishing_entity","fishing_entity_id")
+            )
+
+# Grey land
+grey_land <- World_map %>% 
+  filter(is.na(fishing_entity_id)) %>% 
+  filter(!sovereignt %in% c("Antarctica"))
+
+# Plot
+p <-
+  ggplot() +
+  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
+  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
+            aes(fill = avrg_spp_per),
+            size = 0.1,
+            col = "white") +
+  geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
+  geom_sf(data = subset(sf_ready_plot, per_early_plot >= 0), aes(color = per_early_plot), size = 0.1, fill = "white") +
+  geom_sf(data = point_data, aes(), color = "grey90") +
+  my_ggtheme_m(leg_pos = "bottom") +
+    geom_path(data = subset(arrows_data, per_early_plot >= 0),
+              aes(X, Y,
+                  group = group,
+                  col = per_early_plot
+              ),
+              size = 1,
+              arrow = arrow(length = unit(0.01, "npc"),
+                            ends = "first"),
+              show.legend = NA) +
+  my_ggtheme_m(leg_pos = "bottom") +
+  scale_color_gradientn(name= "Changes in catch proportion \nand changing stocks (%)",
+                        colours = pal,
+                        limits = c(0,max(data_ready_plot$per_early_plot,na.rm=T)),
+                        breaks = seq(0,max(data_ready_plot$per_early_plot,na.rm=T),10),
+                        labels = c(seq(0,max(data_ready_plot$per_early_plot-10,na.rm=T),10),">50")
+  ) +
+  scale_fill_gradientn(name= "Changes in catch proportion \nand changing stocks (%)",
+                        colours = pal,
+                        limits = c(0,max(data_ready_plot$per_early_plot,na.rm=T)),
+                        breaks = seq(0,max(data_ready_plot$per_early_plot,na.rm=T),10),
+                        labels = c(seq(0,max(data_ready_plot$per_early_plot-10,na.rm=T),10),">50")
+  )
+
+ggsave(plot =  p,
+       "./Figures/ssr_early.png",
+       width = 10,
+       height = 5,
+       units = "in")
+
+```
+
+```{r fig_map_Juan, eval = F, echo =F}
+# Figure made thanks to Juanito!
+# http://jsmayorga.com/post/mapping-the-global-network-of-transnational-fisheries/
+
+### Median number of stocks changing in each time step
+Prop_results %>% 
+  filter(variable == "n_per") %>% 
+  group_by() %>% 
+  summarise_at(vars(early,mid),funs(
+    median(.,na.rm=T),
+    sd(.,na.rm=T),
+    length(unique(eez_name)),
+    length(unique(eez_name))/280*100
+  )
+  )
+
+# Prepare data from results
+Data <- Prop_results %>% 
+  # get coords for sources
+  left_join(Eez_centroid,
+  # left_join(Eez_centroid_eck,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  # left_join(Eez_centroid_eck,
+  left_join(Eez_centroid,
+            by = "eez_name") %>%
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  mutate(
+    per_early = ifelse(is.na(per_early),0,per_early),
+    per_mid = ifelse(is.na(per_mid),0,per_mid)
+  ) %>% 
+  drop_na() %>%
+  # gather("variable","value",per_early:per_early_plot) %>% 
+  filter(
+    variable %in% c("mean_eez","n_spp")#,
+    # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
+  )
+
+# Convert to data to spatial
+network_data_sf <- Data %>%
+  select(start_lon, start_lat, end_lon, end_lat) %>% 
+  purrr::transpose() %>% 
+  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
+  purrr::map(st_linestring) %>%
+  # st_sfc(crs = 4326) %>%
+  st_sfc(crs = "+proj=eck4") %>% # for tubular plot
+  st_sf(geometry = .) %>%
+  bind_cols(Data) %>%
+  select(everything(), geometry) %>% 
+  # Set circles ('cus the world is not flat!)
+  # st_segmentize(units::set_units(100, km)) %>%
+  # mutate(geometry = (geometry + c(18000000,90000000)) %% c(18000000) - c(18000000,90000000)) %>%
+  # mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>%
+  # st_wrap_dateline(options = "WRAPDATELINE=YES", quiet = TRUE); network_data_sf # for tubular plot
+  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=60"), quiet = TRUE)
+
+network_data_sf
+
+#  unpack the coordinates of each feature 
+coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
+
+c(network_data_sf$line_id, network_data_sf$coords) %<-% (
+  coord_data %>% 
+    rename(subline_id = L1) %>%   
+    nest(-L2)
+)
+
+# Create data path for line connections
+paths <- network_data_sf %>% 
+  sf::st_set_geometry(NULL) %>% 
+  unnest(cols = c(coords)) %>%
+  mutate_at(vars(X,Y), round,2) %>% 
+  gather("time_step","per_chng",per_early:per_mid) %>% 
+  mutate(per_plot = ifelse(per_chng > 50,50,per_chng),
+         direction = ifelse(per_chng > 0,"positive","negative")
+         )
+
+# Plot it!
+
+# Proportion map data 
+Map_data <- Data %>% 
+  gather("time_step","per_chng",per_early:per_mid) %>% 
+  mutate(per_plot = ifelse(per_chng > 50,50,per_chng))
+
+# Average number of taxa to change per eez
+
+# Total number of transboundary stocks per EEZ
+n_trans_eez <- Transboundary_spp %>% 
+  group_by(eez_name) %>% 
+  summarise(
+    # n_trans = unique(length(taxon_key))
+    n_trans = n()
+  )
+
+# Estimate the percentage changing
+n_taxa_data <- Prop_results %>% 
+  filter(variable == "n_spp") %>%
+  select(eez_name,early,mid) %>% 
+  left_join(n_trans_eez) %>% 
+  # mutate(per = (early/n_trans)*100) %>% 
+  mutate_at(vars(early,mid),
+            .funs = ~(./n_trans)*100) %>% 
+  rename(sf_eez_name = eez_name) %>% 
+  left_join(Matching_names) %>% 
+  select(sau_fishing_entity,fishing_entity_id,early,mid) %>% 
+  gather("time_step","avrg_spp_per",early:mid) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,time_step) %>%
+  summarise(avrg_spp_per = mean(avrg_spp_per,na.rm=T)) %>% 
+  mutate(time_step = ifelse(time_step == "early","per_early","per_mid")) %>%
+  mutate(avrg_spp_per_100 = avrg_spp_per,
+    avrg_spp_per = ifelse(avrg_spp_per > 50,50,avrg_spp_per))
+
 # Paper numbers for Rev. 1
@@
 i = 1
-for(i in 1:2){
+# for(i in 1:2){
   gc()# clean ram
@@
     plot = Prop_map,
-    filename = "./Figures/fig_5n.png",
+    filename = "./Figures/test.png",
     width = 10,

### 2021-11-15 Scripts/responses_rev_gcb.Rmd
misc

@@
 title: "Response to Reviewers"
-output: pdf_document
+output:
+  word_document: default
+  pdf_document: default
 ---
 
-# Edditor
+# Editor
 
@@
 
-- **Response: PLEASE PAY SPECIAL ATTENTION TO THIS POINT. LOOKS LIKE WE REALLY NEED TO CONVINCE THE EDITOR OUR RESEARCH IS NOVEL**
-
-Below you will find the response to both Reviewers 1 and 2.
+- **Response: PLEASE PAY SPECIAL ATTENTION TO THIS POINT.
+LOOKS LIKE WE REALLY NEED TO CONVINCE THE EDITOR OUR RESEARCH IS NOVEL** Below you will find the response to both Reviewers 1 and 2.
 In addition, we would like to offer here an explanation of how our paper differs from recent published papers, including those authored by some of us.
@@
   To identify the timeline we adopt the concept of *Time of Emergence* commonly used in oceanic physical and chemical studies (See Rodgers *et al.* 2015, Keller *et al.* 2014 and has just recently been used in ecological studies by Cheung *et al.* 2020 looking at ecological responses to marine heatwaves.
-As far as we are concerned, no other study has adopted this methodology to looking at the tyming of shifts in transbounary fisheries.
+As far as we are concerned, no other study has adopted this methodology to look at the timing of shifts in transboundary fisheries.
 On the other hand, to identify the intensity of the shift, we (partially) used methods developed by Palacios *et al.* and Sumaila *et al* (2020), published in *Ecology and society*.
-These papers adopt the concept of threat point to evaluate the resilience of specific treaties celebrated by Canada-US tranboundary to manage four transboundary stocks.
+These papers adopt the concept of threat point to evaluate the resilience of specific treaties celebrated by Canada-US transboundary to manage four transboundary stocks.
 Here, we take this idea of threat point and expand the analysis globally to estimate the intensity that the shifts in transboundary stocks will have.
 
-In short, the current manuscript builds upon Palacios-Abrantes *et al.* 2020 *Scientific Reports* combining methods developed in Sumaila *et al.* 2020 and Palacios-Abrantes *et al.* 2020 *Ecology and Society* to address a question that neither Pinsky *et al.* 2018 *Science* nor Oremus *et al.* (2020) *Nature Sust.* addressed.
+  In short, our manuscript builds upon Palacios-Abrantes *et al.* 2020 *Scientific Reports* combining methods developed in Sumaila *et al.* 2020 and Palacios-Abrantes *et al.* 2020 *Ecology and Society* to address a question that neither Pinsky *et al.* 2018 *Science* nor Oremus *et al.* (2020) *Nature Sust.* addressed.
 We hope to have clarified the novelty of our research and how it differs from previous studies.
@@
 We have revised the manuscript paying special attention to clarifying the parts the reviewer flagged.
-We thank the reviewer for his/her comments as they have strengthen our paper.
+We thank the reviewer for his/her comments as they have strengthened our paper.
 Please find below direct responses to the reviewer's comments.
@@
 - **Response:** Thank you for this comment.
-We acknowledge that we did not defined clearly our concept of stock in the previous version of the manuscript.
+We acknowledge that we did not clearly define our concept of stock in the previous version of the manuscript.
 The definition of stock has now been included in the revised manuscript in both the introduction and the methods as follows;
@@
 
-  - **Response:** We understand the reviewer's concern and hope that our response have clarified this issue.
+  - **Response:** We understand the reviewer's concern and hope that our responses have clarified this issue.
 However, it is almost certain that our definition of stock is influencing the results of our paper, just by that nature of the analysis.
@@
 
-- **Response:** We agree with the reviewer that and un-weighted centroid can severely skew the actual distribution of the species and influence our results.
+- **Response:** We agree with the reviewer that an un-weighted centroid can severely skew the actual distribution of the species and influence our results.
 We have revised the manuscript to clarify that we did account for the heterogeneous distribution of fish stocks across their entire range.
@@
 
-![Top 5](../Figures/top_toe_map.png)
+![Fig 4.
+Average year of range shift of the top five most valuable transboundary stocks of each EEZ between 2005 and 2014.
+Warm colors are indicative of an early shift  while cool colors represent a late shift.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.](../Figures/Fig4.png)
 
@@
 - **Response:** This is indeed the definition of our stock.
-We do not differentiate between transboundary and straddling stocks as per defined in Palacios-Abanes *et al.* (2020).
+We do not differentiate between transboundary, straddling nor highly-migratory stocks as per defined in Palacios-Abanes *et al.* (2020).
 Please refer to our responses to the main comment.
+Moreover, all of the catch and value data used represent fishing activities within countries’ EEZs.
 
@@
 
-  - - Line XX; *Using a less conservative threshold (i.e., one s.d.
+  - Line XX; *Using a less conservative threshold (i.e., one s.d.
 representing a probability of 67% that the stock has shifted) results in 18% (n = 5,745) more shifting stocks, that is, 63% of all studied stocks*
@@
 
-  - - Fig. 3 legend, *B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region.*
+  - Fig. 3 legend, *B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region.*
 
@@
 
-![New Figure 5](../Figures/Fig5n.png)
+![New Figure 5](../Figures/Fig5.png)
 
@@
 
-  - - *This way, if, Canada and the US share a specific stock the SSR for Canada could be 20% and that of the US 80%, meaning that Canada holds 20% of the distribution of the stock within its waters and the US the other 80%.*
+  - *This way, if, Canada and the US share a specific stock the SSR for Canada could be 20% and that of the US 80%, meaning that Canada holds 20% of the distribution of the stock within its waters and the US the other 80%.*
 
@@
 
-- line XX: **Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 61 $\pm$ 18% by 2030 and 72 $\pm$ 17% by 2050 in 83% of the world's EEZs (Figure SX).
-In some cases, the SSR of one EEZ is projected increase in more than double relative to their neighbors (e.g., from 10% to 30% SSR).
-Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Cyprus from Greece and Turkey in the Mediterranean Sea, and Russia from Norway and Japan in the Barents Sea and Western Pacific, respectively**
+- line XX: *Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 61 $\pm$ 18% by 2030 and 72 $\pm$ 17% by 2050 in 83% of the world's EEZs (Figure SX).
+In some cases, the SSR of one EEZ is projected to increase more than double relative to their neighbors (e.g., from 10% to 30% SSR).
+Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Cyprus from Greece and Turkey in the Mediterranean Sea, and Russia from Norway and Japan in the Barents Sea and Western Pacific, respectively*
 
@@
 
-  - - **Most EEZs are expected to see a relative small number of stocks shifting their SSR (Fig.
+  - *Most EEZs are expected to see a relatively small number of stocks shifting their SSR (Fig.
 \@ref(fig:toeFigure4)A).
 On average, EEZs are expected to see changes in SSR on 16% $\pm$ 10 of transboundary stocks increasing to 20% $\pm$ 11 by mid of the $21^{st}$ century.
-However, this result masks large changes, specially on island territories.
-By early century, the EEZs of French Guiana (54%), South Africa (59%), Falkland Isl.
-(61%), Brazil (75%) and Kerguelen Isl., and Pitcairn (both 100%) are projected to see changes in SSR of over 50% in their transboundary stocks.
-On the other hand, some island territories such as Anguilla are projected to see changes in as low as 2% of their transboundary stocks.**
+However, this result masks large variation in changes across EEZs.
+By early century, the EEZs of French Guiana (54%), South Africa (59%), Falkland Island (61%), Brazil (75%) and Kerguelen Island, and Pitcairn (both 100%) are projected to see changes in SSR of over 50% in their transboundary stocks.
+On the other hand, some island territories such as Anguilla are projected to see changes in as low as 2% of their transboundary stocks.*
 
@@
 
-Comments:
+## Comments:
 
@@
 
-- **Response:** We agree with the reviewer and have included the SDG in the paper.
+- **Response:** We agree with the reviewer and have included the SDG 17 in the instrucution and discussion of the paper.
 
@@
 
-17.16.1
-Number of countries reporting progress in multi-stakeholder development effectiveness monitoring frameworks that support the achievement of the sustainable development goals
+- **Response: (NOTE: Help from COLETE here)** We have included the following in the discussion as a response to this comment:
 
-- **Response:**
+  -- Line xx **The SDG target 17.16 (Enhance the global partnership for sustainable development) calls for collaborations between stakeholders to support the achievement of the SDGs in all countries, in particular developing countries [@SDGREF].
+Such collaborations will be critical if we are to sustainably manage transboundary fisheries under climate change [@Pinsky:2018cb; @Oremus:2020id].**
 
@@
 
-- **Response:** Thank you for this comment. We have now changed the term to "*Transboundary Shift Index (TSI)*"
+- **Response:** Thank you for this comment. We have now changed the term to *Transboundary Shift Index (TSI)*
 
@@
 
-- **Response:** We agree with the reviewer that the distribution of fish stocks in the ocean is not necessarily oval.
-In fact, the distribution of some of our stocks does not follow this patter.
+- **Response: (NOTE: Is the reviewer referring to multiple stocks within an EEZ??)** We agree with the reviewer that the distribution of fish stocks in the ocean is not necessarily oval.
+In fact, the distribution of some of our stocks does not follow this pattern.
 Nevertheless, we believe our method responds well to other patterns (as long as they are considered one single stock unit) since the centroid of the stock's distribution is based on the "high biomass-level" grid cells.
@@
 
-- **Response:** Red circles represent the years the TI is larger than the historical variation. These were quite faded in the original diagram and have made them clearer in the revised figure.
+- **Response:** Red circles represent the years the TI is larger than the historical variation. They are all located in the graph along the black line.
 
@@
 
-- **Response:** The distributional centroid represent's the stocks centroid while "geographical" centroid is the EEZs centroid.
-We have changed the second term as suggested by the reviewer, it is now called simply "EEZ centroid".
+- **Response:** The distributional centroid represents the stocks centroid while "geographical" centroid is the EEZs centroid.
+We notice that this could be confusing and have changed the second term as suggested by the reviewer, it is now called simply "EEZ centroid".
 
@@
 
-- **Response:**
+- **Response: (NOTE: I AM UNSURE HOW TO RESPOND TO THIS COMMENT, ARE WE ASKED TO LINK THE UNCERTAINTY TO THE DECISION MAKING PROCESS?)**: Indeed uncertainty will be an inherit aspect of climate change and decision making.
+We have included a couple of limes in the discussion making a more evident linkage between the historical variability and the decision making process:
+
+  - Line XX, *Therefore, decision-makers are facing the challenge of taking action under an uncertain future.
+One strategy for that is adapting scenario planning into fisheries management [@NOAA_scen_plan_2020].
+This will allow for managers to integrate the uncertainty into planning for managing fisheries.
+For example, if stocks are expected to shift their distribution by 2036 $\pm$ 28 years (Fig.
+2), managers could prepare for a scenario where stocks are currently changing and another where they will shift in 20 years.*
 
@@
 
-- **Response:**
+- **Response:** Please see response to the previous comment
 
@@
 
-- **Response:**
+- **Response:** Please see response to the Scenario planning comment.
 
@@
 
-- **Response:**
+- **Response: (NOTE: Don't understand, what vectors?)**
 

### 2021-11-17 Scripts/Initial_Analysis.Rmd
Created splitted figure 3 for rev.2

@@
 
-SummaryProp(taxon_key)
+# SummaryProp(taxon_key)
 
@@
   rename(eez_name = Name) %>% 
-  st_transform(crs = 4326) %>% # 4326
-  # st_transform(crs = "+proj=eck4") %>% # for tbular plot
+  # st_transform(crs = 4326) %>% # 4326
+  st_transform(crs = "+proj=eck4") %>% # for tbular plot
   st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
@@
   select(iso_a3,sovereignt,sov_a3,name,name_long,region_wb,continent,geometry) %>% 
-  # st_transform(crs = "+proj=eck4") %>%
-  st_transform(crs = 4326) %>% # 4326
+  st_transform(crs = "+proj=eck4") %>%
+  # st_transform(crs = 4326) %>% # 4326
   mutate(iso_a3 = ifelse(sovereignt == "Norway","NOR",
@@
 
-ggplot(Eez_centroid_sf) +
-  geom_sf()
+# ggplot(Eez_centroid_sf) +
+#   geom_sf() +
+#   geom_sf(data = World_map, aes(), fill = "grey") +
+#   geom_sf(data = World_eez_sf, aes(), fill = "white")
 
@@
 
+### Splitted Figure 3 for supplements
+
+Note this chunk needs the previous one to be ran up to the forloop
+
+```{r supplement_figure_splitted}
+
+  # Value Map
+  World_map_part <- World_map_nemer %>% 
+  filter(
+    tresh ==  "tresh_two",
+    variable == "value"
+  )
+
+  # ToE Map 
+  toe_map_data <- World_eez_sf %>% 
+    left_join(Eez_d_hist,
+              by = "eezid") %>% 
+    filter(tresh == "tresh_two",
+           metric == "mean_yr")%>% 
+sf::st_transform(crs = "+proj=eck4")
+
+  # Map with countries without ToE
+
+  russia_churchil <- World_eez_sf %>% 
+    filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))%>% 
+sf::st_transform(crs = "+proj=eck4")
+
+  no_toe_eez <- World_eez_sf %>% 
+    filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
+    rbind(russia_churchil)%>% 
+sf::st_transform(crs = "+proj=eck4") %>% 
+    mutate(value = NA)
+
+  no_toe_land <- World_map %>% 
+    filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !name %in% World_map_part$name)%>% 
+sf::st_transform(crs = "+proj=eck4") %>%   # Filter EEZs with ToE stocks 
+    mutate(value = NA)
+
+  # Grey land
+grey_land <- World_map %>% 
+  # filter(is.na(fishing_entity_id)) %>% 
+  filter(!sovereignt %in% c("Antarctica"))
+
+  # Inlcude the participation of species
+
+    #   geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    #   geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
+
+eez_map <- ggplot(toe_map_data) +
+  geom_sf(aes(fill = value), # for 2000 to 2100),
+          color =  "white",
+          size = 0.1
+  ) +
+  geom_sf(data = grey_land, aes(),
+          fill = "grey50", 
+          color =  "white",
+          size = 0.1) +
+  geom_sf(data = no_toe_eez, aes(fill = as.numeric(value)),
+          color =  "white",
+          size = 0.1) +
+  scale_fill_viridis("Average year of range shift",
+                     limits = c(2000,2100),
+                     breaks = seq(2000,2100,25),
+                     labels = seq(2000,2100,25),
+                     direction = -1,
+                     na.value = "lightblue"
+  ) +
+  my_ggtheme_m(leg_pos = "bottom") +
+  theme(legend.key.width = unit(3, "line")) +
+  guides(
+    fill=guide_colorbar(title.position="top", 
+                        title.hjust =0.5)
+  )
+
+land_map <- ggplot(World_map_part) +
+  geom_sf(aes(fill = participation-2000),
+          color =  "white",
+          size = 0.1
+  ) +
+  geom_sf(data = no_toe_land, aes(fill = as.numeric(value)),
+          color =  "white",
+          size = 0.1) +
+  scale_fill_viridis("Transboundary fish revenue (%)",
+                     limits = c(0,100),
+                     breaks = seq(0,100,25),
+                     labels = seq(0,100,25),
+                     direction = 1,
+                     na.value = "lightblue",
+                     option = "A"
+  ) +
+  my_ggtheme_m(leg_pos = "bottom") +
+  theme(legend.key.width = unit(3, "line")) +
+  guides(
+    fill=guide_colorbar(title.position="top", 
+                        title.hjust =0.5)
+  )
+
+# ggsave(
+#   plot = land_map,
+#   width = 8,
+#   height = 6,
+#   units = "in",
+#   filename = "./Figures/eez_map_supp.png"
+# )
+
+## --------------- ##
+  # Combine plots and save them
+  ## --------------- ##
+
+  p <- ggdraw() +
+    draw_plot(eez_map, x = 0, y = 0.5, width = 1, height = 0.5) +
+    # Regional boxplot
+    draw_plot(land_map, x = 0, y = 0, width = 1, height = 0.5) +
+    # Labels
+    draw_plot_label(label = c("A", "B"), 
+                    size = 20,
+                    x = c(0, 0), 
+                    y = c(0.98, 0.46)
+    ) + 
+  theme(plot.background = element_rect(fill="white", color = NA))
+
+  save_plot(
+      "./Figures/FigSplitSX.png",
+    p,
+    base_height = 10,
+    base_width = 10
+  )
+
+```
+
 ##### Fig. 4N Mapping the time of emergence Of Top stocks
@@
   mutate(
-    per_early_plot = ifelse(per_early_plot >= 50,50,per_early_plot),
-    per_mid_plot = ifelse(per_mid_plot >= 50,50,per_mid_plot))
-
-sf_ready_plot <- st_as_sf(data_ready_plot,
-            coords = c("lon", "lat")
-            ) %>% 
+    per_early = ifelse(per_early_plot >= 50,50,per_early_plot),
+    per_mid = ifelse(per_mid_plot >= 50,50,per_mid_plot)) %>% 
+  gather("time_step","percentage",per_early,per_mid)
+
+# SELECT A TIME STEP
+# time_toplot <- "per_mid"
+time_toplot <- "per_early"
+
+# ----------------------------- #
+
+sf_ready_plot <- data_ready_plot %>% 
+  filter(time_step == time_toplot) %>% 
+  st_as_sf(coords = c("lon", "lat")) %>% 
   st_set_crs(4326) %>% # to match shapefiles
@@
 arrows_data <- as.data.frame(st_coordinates(st_cast(sf_ready_plot,"POINT")$geometry)) %>% 
-  bind_cols(data_ready_plot)
+  bind_cols(data_ready_plot %>% 
+              filter(time_step == time_toplot)
+            )
 
@@
 point_data <- path_data %>% 
-  filter(per_early <= 0) %>% 
+  gather("time_step","percentage",per_early:per_mid) %>% 
+  filter(time_step == time_toplot,
+    percentage <= 0) %>% 
   st_as_sf(
@@
   mutate(avrg_spp_per_100 = avrg_spp_per,
-    avrg_spp_per = ifelse(avrg_spp_per > 50,50,avrg_spp_per))
+    avrg_spp_per = ifelse(avrg_spp_per > 50,50,avrg_spp_per)) %>% 
+  filter(time_step == time_toplot)
 
@@
   geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
-  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
+  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
             aes(fill = avrg_spp_per),
@@
   geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
-  geom_sf(data = subset(sf_ready_plot, per_early_plot >= 0), aes(color = per_early_plot), size = 0.1, fill = "white") +
+  geom_sf(data = subset(sf_ready_plot, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
   geom_sf(data = point_data, aes(), color = "grey90") +
   my_ggtheme_m(leg_pos = "bottom") +
-    geom_path(data = subset(arrows_data, per_early_plot >= 0),
+    geom_path(data = subset(arrows_data, percentage >= 0),
               aes(X, Y,
                   group = group,
-                  col = per_early_plot
+                  col = percentage
               ),
@@
                         colours = pal,
-                        limits = c(0,max(data_ready_plot$per_early_plot,na.rm=T)),
-                        breaks = seq(0,max(data_ready_plot$per_early_plot,na.rm=T),10),
-                        labels = c(seq(0,max(data_ready_plot$per_early_plot-10,na.rm=T),10),">50")
+                        limits = c(0,max(sf_ready_plot$percentage,na.rm=T)),
+                        breaks = seq(0,max(sf_ready_plot$percentage,na.rm=T),10),
+                        labels = c(seq(0,max(sf_ready_plot$percentage-10,na.rm=T),10),">50")
   ) +
@@
                         colours = pal,
-                        limits = c(0,max(data_ready_plot$per_early_plot,na.rm=T)),
-                        breaks = seq(0,max(data_ready_plot$per_early_plot,na.rm=T),10),
-                        labels = c(seq(0,max(data_ready_plot$per_early_plot-10,na.rm=T),10),">50")
-  )
-
+                        limits = c(0,max(sf_ready_plot$percentage,na.rm=T)),
+                        breaks = seq(0,max(sf_ready_plot$percentage,na.rm=T),10),
+                        labels = c(seq(0,max(sf_ready_plot$percentage-10,na.rm=T),10),">50")
+  )
+
+name <- paste0("./Figures/ssr_",time_toplot,".png")
+# Save figure
 ggsave(plot =  p,
-       "./Figures/ssr_early.png",
+       name,
+       width = 10,
+       height = 5,
+       units = "in")
+
+## ------------------- ##
+  # Regional Plots
+  ## ------------------- ##
+
+  Lat_am_land <- World_map %>% 
+    filter(region_wb %in% c("Latin America & Caribbean") |
+             sau_fishing_entity %in% c("USA","Mexico")
+    )
+
+  Lat_am_eez <- World_eez_sf %>% 
+    select(sf_eez_name = eez_name,everything()) %>% 
+    left_join(Matching_names) %>% 
+    filter(
+      !str_detect(sf_eez_name,"Pacific"),
+      !sf_eez_name %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador")
+    )
+
+  Lat_am_data <- sf_ready_plot %>% 
+    filter(source %in% Lat_am_eez$sf_eez_name)
+
+  Lat_am_paths <- sf_ready_plot %>% 
+    filter(source %in% Lat_am_eez$sf_eez_name,
+           objective %in% Lat_am_eez$sf_eez_name)
+
+  Lat_am_points <- point_data %>% 
+    filter(source %in% Lat_am_eez$sf_eez_name,
+           target %in% Lat_am_eez$sf_eez_name)
+
+  Lat_am_map <- ggplot() +
+    geom_sf(data = Lat_am_eez, size = 0.1, fill = "white", col = "gray30") +
+    geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
+    geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time_toplot),
+            aes(fill = avrg_spp_per),
+            size = 0.1,col = "white") +
+    geom_sf(data = Lat_am_points, aes(), color = "grey90") +
+  my_ggtheme_m(leg_pos = "bottom") +
+    geom_path(data = subset(Lat_am_paths, percentage > 0),
+              aes(X, Y,
+                  group = interaction(line_id,subline_id),
+                  col = per_plot
+              ),
+              size = 1,
+              arrow = arrow(length = unit(0.05, "npc"),
+                            ends = "first"),
+              show.legend = NA) +
+    coord_sf(xlim = c(-100, -55),ylim = c(9, 30)) + 
+    labs(x = "", y = "") +
+    scale_x_continuous(breaks = seq(-100,60,10)) +
+    scale_y_continuous(breaks = seq(10,30,10)) +
+    my_ggtheme_m(leg_pos = "bottom") +
+    theme(legend.position = "",
+          panel.background = element_rect(fill = "white", color = "white"),
+          panel.grid.major =  element_line(color = "white")
+    ) +
+    scale_colour_viridis(
+      name="Average Gain of\nStock-Share Ratio (%)",
+      limits = c(0,max(Lat_am_data$percentage)),
+      breaks = seq(0,max(Lat_am_data$percentage),10),
+      # option="magma",
+      # direction = -1
+    )+
+    scale_fill_viridis(
+      name="Average Gain of\nStock-Share Ratio (%)",
+      limits = c(0,max(Lat_am_data$percentage)),
+      breaks = seq(0,max(Lat_am_data$percentage),10),
+      # option="magma",
+      # direction = -1
+    )
+
+  ggsave(plot =  Lat_am_map,
+       "test.png",
        width = 10,

### 2021-11-17 Scripts/responses_rev_gcb.Rmd
Created splitted figure 3 for rev.2

@@
 
-  - **Response:**  **NOTE: I do not (fully) agree with this comment, Do you think we can counter-argue it?
-(We had the same type of plot in the Sci.
-Rep.
-paper).
-Yes, the figure is complex, but I do not think having different color scales will make it any less complex.
-I actually think it will make it even harder to read.**
+  - **Response:** We understand the reviewer’s concern with the complexity of our figure.
+However, after some tryouts, we concluded that having different color scales does not make it any less complex.
+Thus, We have made an extra effort to describe it as clearly as possible in the figure legend (see below).
+In addition, we have split up the land and polygons into two maps and include them in the supplemental figures (See supplement FX).
 
@@
 
-  - Fig. 3 legend, *B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region.*
+  - Fig.
+3 legend, - *Year of change in the shared distribution of 4,119 transboundary stocks.
+A) Land polygons show the contribution of shifting stocks to a country or territory's total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygons display average year of range shifts within them (Note that the year scale goes from end of the century to recently).
+Warm colors are indicative of an early shift and high fishing revenue contribution from transboundary stocks while cold colors represent a later shift and low fishing revenue.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
+See Figure SXX for a detached version.
+B) Year of shared distribution shifts by UN sub-regions.
+Points color coded by the median year for all shifting stocks in the region reflecting the Kruskal-Wallis test.
+Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda.
+N = North, S = South, W = West and E = East.
+Ltn.
+Ame.
+and the Car.
+= Latin America and the Caribbean.
+Aus and New Z.
+= Australia and New Zealand.*
 

### 2021-11-17 Scripts/Initial_Analysis.Rmd
changes to legend

@@
 
-### Splitted Figure 3 for supplements
-
-Note this chunk needs the previous one to be ran up to the forloop
+###### Splitted Figure 3 for supplements
+
+Note this chunk needs the previous one to be ran up to the for-loop
+
+Year of change in the shared distribution of 4,119 transboundary stocks and their contribution to fishing revenue.
+A) Contribution of shifting stocks to a country or territory's total fishing revenue from transboundary stocks.
+B) Average year of range shift across an EEZ’s transboundary stocks.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
+This figure is a detached version of Figure 3
 
@@
   p <- ggdraw() +
-    draw_plot(eez_map, x = 0, y = 0.5, width = 1, height = 0.5) +
+    draw_plot(eez_map, x = 0, y = 0, width = 1, height = 0.5) +
     # Regional boxplot
-    draw_plot(land_map, x = 0, y = 0, width = 1, height = 0.5) +
+    draw_plot(land_map, x = 0, y = 0.5, width = 1, height = 0.5) +
     # Labels

### 2021-11-17 Scripts/Manuscript.Rmd
changes to legend

@@
 Exclusive Economic Zone polygons display average year of range shifts within them.
-Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cool colors represent a late shift/low fishing revenue.
+Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cold colors represent a late shift/low fishing revenue.
 EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.

### 2021-11-23 Scripts/Initial_Analysis.Rmd
Modified F5 to match F4
Created Supplmental zoom figures of Fig5

@@
 
-##### Fig. 4N Mapping the time of emergence Of Top stocks
+##### Fig. 4 Mapping the time of emergence Of Top stocks
 
@@
 
-###### The bar option by UN sub region
-
-```{r}
-
-# Create data
-Species_Data <- top_toe %>% 
-  filter(tresh == selected_tresh,
-         variable == "value") %>% 
-  mutate(mean = mean-2000,
-         group = geo_region)
-
-# Set a number of 'empty bar' to add at the end of each group
-Char <- as.factor(Species_Data$group)
-empty_bar <- 5
-to_add <- data.frame( matrix(NA, empty_bar*nlevels(Char), ncol(Species_Data)) )
-colnames(to_add) <- colnames(Species_Data)
-to_add$group <- rep(levels(Char), each=empty_bar)
-Species_Data <- rbind(Species_Data, to_add)
-Species_Data <- Species_Data %>% arrange(group,mean)
-Species_Data$id <- seq(1, nrow(Species_Data))
-
-# Species_Data$id <- seq(1, nrow(Species_Data))
-
-# Get the name and the y position of each label
-label_data <- Species_Data 
-number_of_bar <- nrow(label_data)
-angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar
-label_data$hjust <- ifelse( angle < -90, 1, 0)
-label_data$angle <- ifelse(angle < -90, angle+180, angle)
-
-# prepare a data frame for base lines
-base_data <- Species_Data %>% 
-  group_by(group) %>% 
-  summarize(start=min(id), 
-            end=max(id) - empty_bar, 
-            median = median(mean, na.rm = T)
-  ) %>% 
-  rowwise() %>% 
-  mutate(title=mean(c(start, end)))
-
-# prepare a data frame for grid (scales)
-grid_data <- base_data
-grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
-grid_data$start <- grid_data$start - 1
-# grid_data <- grid_data[-1,]
-# grid_data[1,2] <- 5 # Manually include bathydemerssal
-# grid_data[1,3] <- 1
-
-un_regions <- grid_data %>% 
-  select(group)
-
-ggplot(Species_Data, 
-       aes(x=as.factor(id), 
-           y=mean,
-           fill=mean)
-) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
-  geom_bar(aes(x=as.factor(id),
-               y=mean,
-               fill=mean),
-           stat="identity",
-           alpha=0.5
-  ) +
-  geom_segment(data=grid_data,
-               aes(x = end, y = 30, xend = start, yend = 30),
-               colour = "grey90", alpha=1, size=0.3, inherit.aes = FALSE) +
-  geom_bar(aes(x=as.factor(id),
-               y=mean,
-               fill=mean
-  ),
-  stat="identity",
-  alpha=0.5) +
-  annotate("text", 
-           x = 0.5,
-           y = 37,
-           label = "UN Agenda",
-           color= "grey60", 
-           size=3, 
-           angle=-5
-  ) +
-  # Country names
-  geom_text(data=label_data,
-            aes(x=id, 
-                y= 100,
-                label=sau_fishing_entity,
-                hjust=hjust),
-            color="black",
-            fontface="italic",
-            alpha = 1,
-            size = 4 ,
-            angle = label_data$angle,
-            inherit.aes = FALSE) +
-  # Continents labels
-  # geom_text(data=base_data,
-  #           aes(x = title,
-  #               y = 85, 
-  #               label=group), 
-  #           # hjust=c(1,1,1,0,0,0),
-  #           colour = "black",
-  #           # angle = myAng,
-  #           size=5,
-  #           fontface="bold",
-  #           inherit.aes = FALSE) +
-  ylim(-100,100) + # sets the circle the first value the central the last value how wide
-  theme_minimal() +
-  theme(
-    legend.position = c(0.5,0.5),
-    legend.text = element_text(size = 13),
-    legend.title = element_text(size = 14, face = "bold"),
-    legend.title.align = 0.5,
-    axis.text = element_blank(),
-    axis.title = element_blank(),
-    panel.grid = element_blank(),
-    panel.background = element_rect(colour = "white"),
-    legend.direction="horizontal",
-    legend.key.width = unit(3, 
-                            "line")
-  ) +
-  coord_polar(start = 0) +
-  scale_fill_viridis("Average year of range shift\nfor top 5 most valuable stocks",
-                     limits = c(0,100),
-                     breaks = seq(0,100,20),
-                     labels = seq(2000,2100,20),
-                     direction = -1
-  ) +
-  guides(
-    fill=guide_colorbar(title.position="top", 
-                        title.hjust =0.5)
-  )
-
-ggsave(
-  plot = last_plot(),
-  width = 13,
-  height = 13,
-  units = "in",
-  filename = "./Figures/top_stocks_circular_trend.png"
-)
-
-```
-
 ###### The Map option with tops
@@
     units = "in"
-  )
-
-```
-
-##### Fig. 4. Bubble plot
-
-```{r, buble_plot, eval = F, echo = F}
-
-# Needs previous chunk
-
-Buble_data <- Eez_d_hist %>% 
-  filter(metric == "mean_yr") %>% 
-  group_by(sau_fishing_entity,tresh) %>% 
-  summarise(mean_yr = mean(value)) %>% 
-  left_join(Eme_value_data,
-            by = c("sau_fishing_entity","tresh")) %>% 
-  filter(#tresh == selected_tresh, # tresh hold selection
-         variable == "value") %>% 
-  mutate(
-    weight_toe = sum_value/mean_yr,
-    sum_value_bi = sum_value/1000000000,
-    log_value = log10(sum_value_bi)
-    ) %>%   # Billion usd
-  left_join(Matching_names,
-            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
-  distinct(sau_fishing_entity,tresh,.keep_all = TRUE) %>%
-  filter(!is.na(un_geo_region)) %>% 
-  ungroup()# %>%  # no more asterisk
-  # mutate(
-  #   sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity))
-
-### Function for estimating percentile
-
-quibble <- function(x, q = p) {
-        tibble(x = quantile(x, q), q = q)
-}
-
-Percentile_value <- Buble_data %>% 
-  group_by(tresh) %>%  # for whole distribution
-        do(quibble(.$log_value, 0.75))
-
-## Top fishing nations close to SDGs
-Buble_data %>% 
-  filter(log_value > Percentile_value$x & mean_yr < 2036) %>% 
-  View()
-  group_by(tresh) %>% 
-  summarise(sum(sum_value_bi),
-            sum(n_taxa))
-
-## All fishing nations close to SDGs
-Buble_data %>% 
-  filter(mean_yr < 2036) %>% 
-  # nrow() # 76  countries
-  View()
-  group_by(tresh) %>% 
-  summarise(sum(sum_value_bi),
-            sum(n_taxa))
-
-76/length(unique(Buble_data$sau_fishing_entity))
-134/length(unique(Buble_data$sau_fishing_entity))
-
-# Labels Data
-
-Labels_data <- tibble(
-  x = c(2029,
-        # 2050,2070,
-        2007),
-  y = c(-5.2,
-        # -5.8,-5.8,
-        Percentile_value$x+.1),
-  label = c("UN 2030 Agenda", 
-            # "Mid shift","Late shift",
-            "Fisheries > 75th percentile")
-)
-
-# Plot the bubbles!
-ggplot(Buble_data) +
-  geom_point(
-    aes(
-      x = mean_yr,
-      y = log_value,
-      size = n_taxa,
-      colour = un_geo_region,
-      shape = un_geo_region
-    ),
-    alpha = 0.7
-  ) +
-  # Right side labels
-  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034),
-                  aes(
-                    x = mean_yr,
-                    y = log_value,
-                    label = sau_fishing_entity
-                  ),
-                  arrow = arrow(length = unit(0.009, "npc"),
-                                type = "closed",
-                                ends = "first"),
-                  force = 1,
-                  box.padding = 0.2,
-                  fill = "NA",
-                  nudge_x = 2066 - subset(Buble_data,log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034)$mean_yr, # set distance from point
-                  nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
-                  segment.size  = 0.1,
-                  segment.color = "grey50",
-                  direction     = "y",
-                  hjust         = 1
-  ) +
-  # LEFT side labels
-  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034),
-                  aes(
-                    x = mean_yr,
-                    y = log_value,
-                    label = sau_fishing_entity
-                  ),
-                  arrow = arrow(length = unit(0.009, "npc"),
-                                type = "closed",
-                                ends = "first"),
-                  force = 1,
-                  box.padding = 0.2,
-                  fill = "NA",
-                  nudge_x = 2025 - subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$mean_yr, # set horizontal position
-                  nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
-                  segment.size  = 0.1,
-                  segment.color = "grey50",
-                  direction     = "y",
-                  hjust         = 1
-  ) +
-  # Divisory lines and labels
-  geom_vline(xintercept = 2030, color = "grey50", size = 0.5, linetype = "dashed") +
-  geom_text(data = subset(Labels_data, x == 2029),
-            aes(
-              x = x,
-              y = y,
-              label = label
-            ),
-            size = 4,
-            color = "grey30",
-            angle = 90
-  ) +
-  geom_hline(data = Percentile_value,
-             aes(yintercept = x),
-             color = "grey50"
-  )  +
-  geom_text(data = subset(Labels_data, x == 2007),
-            aes(
-              x = x,
-              y = y,
-              label = label
-            ),
-            size = 4,
-            color = "grey30"
-  ) + 
-  scale_shape("Region") +
-  scale_color_viridis(name="Region",
-                      discrete = T) +
-  scale_x_continuous("Average year of range shift",
-                     limits = c(2000,2080),
-                     breaks = seq(2000,2080,10)
-  ) +
-  scale_y_continuous("Fishing revenue (log10)",
-                     limits = c(-6,2),
-                     breaks = seq(-6,2,2),
-                     expand = c(0,0)
-  ) +
-  scale_size("Number of\nshifting stocks") +
-  my_ggtheme_p(leg_pos = "right") +
-  ggsave("./Figures/bubble_plot.png",
-         width = 12,
-         height = 8,
-         units = "in"
-  )
-
-```
-
-```{r participation_text, eval = T, echo=F}
-
-Value_map_data %>% 
-  group_by(tresh,variable) %>% 
-  summarise(m = mean(participation)-2000,
-            sd = sd(participation-2000)
   )
@@
 
-## Regional differences
+#### Regional differences
 
@@
 
-## Habitat preference differences
+#### Habitat preference differences
 
@@
 
-####  Prooportion Analysis
-
-##### Fig. 5 and S5. (Proportion + Netowrk)
-
-```{r}
-
-# Done with the help of https://www.data-to-viz.com/story/MapConnection.html
+##  Prooportion Analysis
+
+### Fig. 5 and S5 Proportion + Netowrk
+
+Note that this map is based on the data viz site: https://www.data-to-viz.com/story/MapConnection.html
+
+```{r fig5_paths, eval = T, echo = T}
 
@@
   geom_sf(data = point_data, aes(), color = "grey90") +
-  my_ggtheme_m(leg_pos = "bottom") +
     geom_path(data = subset(arrows_data, percentage >= 0),
@@
   my_ggtheme_m(leg_pos = "bottom") +
-  scale_color_gradientn(name= "Changes in catch proportion \nand changing stocks (%)",
+  scale_color_gradientn(name= "Changes in catch proportion and changing stocks (%)",
                         colours = pal,
@@
   ) +
-  scale_fill_gradientn(name= "Changes in catch proportion \nand changing stocks (%)",
+  scale_fill_gradientn(name= "Changes in catch proportion and changing stocks (%)",
                         colours = pal,
@@
                         labels = c(seq(0,max(sf_ready_plot$percentage-10,na.rm=T),10),">50")
-  )
-
-name <- paste0("./Figures/ssr_",time_toplot,".png")
+  ) + guides(
+      color=guide_colorbar(title.position="top")
+  ) +
+  theme(legend.key.width = unit(4,"line"))
+
+# name <- paste0("./Figures/ssr_",time_toplot,".png")
 # Save figure
-ggsave(plot =  p,
-       name,
-       width = 10,
-       height = 5,
-       units = "in")
+# ggsave(plot =  p,
+#        name,
+#        width = 8,
+#        height = 5,
+#        units = "in")
+
+```
+
+### Figure S6. Regional plots
+
+Note that the following chunk needs the previous one to be ran
+
+```{r}
 
@@
 
-  Lat_am_land <- World_map %>% 
-    filter(region_wb %in% c("Latin America & Caribbean") |
-             sau_fishing_entity %in% c("USA","Mexico")
-    )
+  # Lat_am_land <- World_map %>% 
+    # filter(region_wb %in% c("Latin America & Caribbean") |
+             # sau_fishing_entity %in% c("USA","Mexico")
+    # )
 
@@
     filter(
+      un_sub_region == "Latin America and the Caribbean",
       !str_detect(sf_eez_name,"Pacific"),
@@
   Lat_am_data <- sf_ready_plot %>% 
-    filter(source %in% Lat_am_eez$sf_eez_name)
-
-  Lat_am_paths <- sf_ready_plot %>% 
+    filter(source %in% Lat_am_eez$sf_eez_name,
+           objective %in% Lat_am_eez$sf_eez_name)
+
+  Lat_am_paths <- arrows_data %>% 
     filter(source %in% Lat_am_eez$sf_eez_name,
@@
 
-  Lat_am_map <- ggplot() +
+  Lat_am_n_taxa <- n_taxa_map %>% 
+    filter(
+      region_wb == "Latin America & Caribbean",
+      # !str_detect(sf_eez_name,"Pacific"),
+      # !sf_eez_name %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador")
+    )
+
+  lat_am_map <- ggplot() +
   geom_sf(data = Lat_am_eez, size = 0.1, fill = "white", col = "gray30") +
-    geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
-    geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time_toplot),
+  geom_sf(data = subset(Lat_am_n_taxa, sovereignt != "Antarctica"),
             aes(fill = avrg_spp_per),
-            size = 0.1,col = "white") +
+            size = 0.1,
+            col = "white") +
+  # geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
+  geom_sf(data = subset(Lat_am_data, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
   geom_sf(data = Lat_am_points, aes(), color = "grey90") +
-  my_ggtheme_m(leg_pos = "bottom") +
-    geom_path(data = subset(Lat_am_paths, percentage > 0),
+    geom_path(data = subset(Lat_am_paths, percentage >= 0),
               aes(X, Y,
-                  group = interaction(line_id,subline_id),
-                  col = per_plot
+                  group = group,
+                  col = percentage
               ),
               size = 1,
-              arrow = arrow(length = unit(0.05, "npc"),
+              arrow = arrow(length = unit(0.01, "npc"),
                             ends = "first"),
               show.legend = NA) +
-    coord_sf(xlim = c(-100, -55),ylim = c(9, 30)) + 
+  my_ggtheme_m(leg_pos = "bottom") +
+    coord_sf(xlim = c(-9000000,-5000000),ylim = c(1000000, 4000000)) + 
+    labs(x = "", y = "") +
+    # scale_x_continuous(breaks = seq(-100,60,10)) +
+    # scale_y_continuous(breaks = seq(10,30,10)) +
+    my_ggtheme_m(leg_pos = "bottom") +
+    theme(
+          panel.background = element_rect(fill = "white", color = "white"),
+          panel.grid.major =  element_line(color = "white")
+    ) +
+    scale_color_gradientn(name= "Changes in catch proportion and changing stocks (%)",
+                        colours = pal,
+                        limits = c(0,max(Lat_am_data$percentage)),
+                        breaks = seq(0,max(Lat_am_data$percentage),10)
+  ) +
+  scale_fill_gradientn(name= "Changes in catch proportion and changing stocks (%)",
+                        colours = pal,
+                        limits = c(0,max(Lat_am_data$percentage)),
+                       breaks = seq(0,max(Lat_am_data$percentage),10)
+  ) + guides(
+      color=guide_colorbar(title.position="top", 
+                        title.hjust =0.5)
+  ) +
+  theme(legend.key.width = unit(4,"line"))
+
+  ggsave(filename = "lat_am_mid.png",
+         lat_am_map,
+         width = 6,
+         height = 6)
+
+  ## ------------------- ##
+  # Northern Europe
+  ## ------------------- ##
+
+  EU_data <- sf_ready_plot %>% 
+    filter(source %in% EU_eez$sf_eez_name)
+
+EU_map <- ggplot() +
+  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
+  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+            aes(fill = avrg_spp_per),
+            size = 0.1,
+            col = "white") +
+  # geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
+  geom_sf(data = subset(sf_ready_plot, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
+  geom_sf(data = point_data, aes(), color = "grey90") +
+    geom_path(data = subset(arrows_data, percentage >= 0),
+              aes(X, Y,
+                  group = group,
+                  col = percentage
+              ),
+              size = 1,
+              arrow = arrow(length = unit(0.03, "npc"),
+                            ends = "first"),
+              show.legend = NA) +
+  # my_ggtheme_m(leg_pos = "bottom") +
+    coord_sf(xlim = c(-1300000,2000000),ylim = c(5500000, 7500000))  +
   labs(x = "", y = "") +
@@
     my_ggtheme_m(leg_pos = "bottom") +
-    theme(legend.position = "",
+    theme(
           panel.background = element_rect(fill = "white", color = "white"),
@@
     ) +
-    scale_colour_viridis(
-      name="Average Gain of\nStock-Share Ratio (%)",
-      limits = c(0,max(Lat_am_data$percentage)),
-      breaks = seq(0,max(Lat_am_data$percentage),10),
-      # option="magma",
-      # direction = -1
-    )+
-    scale_fill_viridis(
-      name="Average Gain of\nStock-Share Ratio (%)",
-      limits = c(0,max(Lat_am_data$percentage)),
-      breaks = seq(0,max(Lat_am_data$percentage),10),
-      # option="magma",
-      # direction = -1
-    )
-
-  ggsave(plot =  Lat_am_map,
-       "test.png",
-       width = 10,
-       height = 5,
-       units = "in")
+    scale_color_gradientn(name= "Changes in catch proportion and changing stocks (%)",
+                        colours = pal,
+                        limits = c(0,max(EU_data$percentage)),
+                        breaks = seq(0,max(EU_data$percentage),10)
+  ) +
+  scale_fill_gradientn(name= "Changes in catch proportion and changing stocks (%)",
+                        colours = pal,
+                        limits = c(0,max(EU_data$percentage)),
+                       breaks = seq(0,max(EU_data$percentage),10)
+  ) + guides(
+      color=guide_colorbar(title.position="top",
+                        title.hjust =0.5)
+  ) +
+  theme(legend.key.width = unit(4,"line"))
+
+    ggsave(filename = "EU_early.png",
+         EU_map,
+         width = 6,
+         height = 6)
 
@@
 
-```{r multiple_reg, eval = F, echo = F}
-
 Arctic <- c("Canada (Arctic)","USA (Alaska, Arctic)", "Russia (Laptev to Chukchi Sea)")
@@
 ```
+
+# Old code
+
+## Fig. 4. Bubble plot
+
+```{r, buble_plot, eval = F, echo = F}
+
+# Needs previous chunk
+
+Buble_data <- Eez_d_hist %>% 
+  filter(metric == "mean_yr") %>% 
+  group_by(sau_fishing_entity,tresh) %>% 
+  summarise(mean_yr = mean(value)) %>% 
+  left_join(Eme_value_data,
+            by = c("sau_fishing_entity","tresh")) %>% 
+  filter(#tresh == selected_tresh, # tresh hold selection
+         variable == "value") %>% 
+  mutate(
+    weight_toe = sum_value/mean_yr,
+    sum_value_bi = sum_value/1000000000,
+    log_value = log10(sum_value_bi)
+    ) %>%   # Billion usd
+  left_join(Matching_names,
+            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
+  distinct(sau_fishing_entity,tresh,.keep_all = TRUE) %>%
+  filter(!is.na(un_geo_region)) %>% 
+  ungroup()# %>%  # no more asterisk
+  # mutate(
+  #   sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity))
+
+### Function for estimating percentile
+
+quibble <- function(x, q = p) {
+        tibble(x = quantile(x, q), q = q)
+}
+
+Percentile_value <- Buble_data %>% 
+  group_by(tresh) %>%  # for whole distribution
+        do(quibble(.$log_value, 0.75))
+
+## Top fishing nations close to SDGs
+Buble_data %>% 
+  filter(log_value > Percentile_value$x & mean_yr < 2036) %>% 
+  View()
+  group_by(tresh) %>% 
+  summarise(sum(sum_value_bi),
+            sum(n_taxa))
+
+## All fishing nations close to SDGs
+Buble_data %>% 
+  filter(mean_yr < 2036) %>% 
+  # nrow() # 76  countries
+  View()
+  group_by(tresh) %>% 
+  summarise(sum(sum_value_bi),
+            sum(n_taxa))
+
+76/length(unique(Buble_data$sau_fishing_entity))
+134/length(unique(Buble_data$sau_fishing_entity))
+
+# Labels Data
+
+Labels_data <- tibble(
+  x = c(2029,
+        # 2050,2070,
+        2007),
+  y = c(-5.2,
+        # -5.8,-5.8,
+        Percentile_value$x+.1),
+  label = c("UN 2030 Agenda", 
+            # "Mid shift","Late shift",
+            "Fisheries > 75th percentile")
+)
+
+# Plot the bubbles!
+ggplot(Buble_data) +
+  geom_point(
+    aes(
+      x = mean_yr,
+      y = log_value,
+      size = n_taxa,
+      colour = un_geo_region,
+      shape = un_geo_region
+    ),
+    alpha = 0.7
+  ) +
+  # Right side labels
+  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034),
+                  aes(
+                    x = mean_yr,
+                    y = log_value,
+                    label = sau_fishing_entity
+                  ),
+                  arrow = arrow(length = unit(0.009, "npc"),
+                                type = "closed",
+                                ends = "first"),
+                  force = 1,
+                  box.padding = 0.2,
+                  fill = "NA",
+                  nudge_x = 2066 - subset(Buble_data,log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034)$mean_yr, # set distance from point
+                  nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
+                  segment.size  = 0.1,
+                  segment.color = "grey50",
+                  direction     = "y",
+                  hjust         = 1
+  ) +
+  # LEFT side labels
+  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034),
+                  aes(
+                    x = mean_yr,
+                    y = log_value,
+                    label = sau_fishing_entity
+                  ),
+                  arrow = arrow(length = unit(0.009, "npc"),
+                                type = "closed",
+                                ends = "first"),
+                  force = 1,
+                  box.padding = 0.2,
+                  fill = "NA",
+                  nudge_x = 2025 - subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$mean_yr, # set horizontal position
+                  nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
+                  segment.size  = 0.1,
+                  segment.color = "grey50",
+                  direction     = "y",
+                  hjust         = 1
+  ) +
+  # Divisory lines and labels
+  geom_vline(xintercept = 2030, color = "grey50", size = 0.5, linetype = "dashed") +
+  geom_text(data = subset(Labels_data, x == 2029),
+            aes(
+              x = x,
+              y = y,
+              label = label
+            ),
+            size = 4,
+            color = "grey30",
+            angle = 90
+  ) +
+  geom_hline(data = Percentile_value,
+             aes(yintercept = x),
+             color = "grey50"
+  )  +
+  geom_text(data = subset(Labels_data, x == 2007),
+            aes(
+              x = x,
+              y = y,
+              label = label
+            ),
+            size = 4,
+            color = "grey30"
+  ) + 
+  scale_shape("Region") +
+  scale_color_viridis(name="Region",
+                      discrete = T) +
+  scale_x_continuous("Average year of range shift",
+                     limits = c(2000,2080),
+                     breaks = seq(2000,2080,10)
+  ) +
+  scale_y_continuous("Fishing revenue (log10)",
+                     limits = c(-6,2),
+                     breaks = seq(-6,2,2),
+                     expand = c(0,0)
+  ) +
+  scale_size("Number of\nshifting stocks") +
+  my_ggtheme_p(leg_pos = "right") +
+  ggsave("./Figures/bubble_plot.png",
+         width = 12,
+         height = 8,
+         units = "in"
+  )
+
+```
+
+```{r participation_text, eval = T, echo=F}
+
+Value_map_data %>% 
+  group_by(tresh,variable) %>% 
+  summarise(m = mean(participation)-2000,
+            sd = sd(participation-2000)
+            )
+
+```
+
+## Fig. 4N Top fishers on bars
+
+The bar option by UN sub region of Figure 4N of top fishers
+
+```{r}
+
+# Create data
+Species_Data <- top_toe %>% 
+  filter(tresh == selected_tresh,
+         variable == "value") %>% 
+  mutate(mean = mean-2000,
+         group = geo_region)
+
+# Set a number of 'empty bar' to add at the end of each group
+Char <- as.factor(Species_Data$group)
+empty_bar <- 5
+to_add <- data.frame( matrix(NA, empty_bar*nlevels(Char), ncol(Species_Data)) )
+colnames(to_add) <- colnames(Species_Data)
+to_add$group <- rep(levels(Char), each=empty_bar)
+Species_Data <- rbind(Species_Data, to_add)
+Species_Data <- Species_Data %>% arrange(group,mean)
+Species_Data$id <- seq(1, nrow(Species_Data))
+
+# Species_Data$id <- seq(1, nrow(Species_Data))
+
+# Get the name and the y position of each label
+label_data <- Species_Data 
+number_of_bar <- nrow(label_data)
+angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar
+label_data$hjust <- ifelse( angle < -90, 1, 0)
+label_data$angle <- ifelse(angle < -90, angle+180, angle)
+
+# prepare a data frame for base lines
+base_data <- Species_Data %>% 
+  group_by(group) %>% 
+  summarize(start=min(id), 
+            end=max(id) - empty_bar, 
+            median = median(mean, na.rm = T)
+  ) %>% 
+  rowwise() %>% 
+  mutate(title=mean(c(start, end)))
+
+# prepare a data frame for grid (scales)
+grid_data <- base_data
+grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
+grid_data$start <- grid_data$start - 1
+# grid_data <- grid_data[-1,]
+# grid_data[1,2] <- 5 # Manually include bathydemerssal
+# grid_data[1,3] <- 1
+
+un_regions <- grid_data %>% 
+  select(group)
+
+ggplot(Species_Data, 
+       aes(x=as.factor(id), 
+           y=mean,
+           fill=mean)
+) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
+  geom_bar(aes(x=as.factor(id),
+               y=mean,
+               fill=mean),
+           stat="identity",
+           alpha=0.5
+  ) +
+  geom_segment(data=grid_data,
+               aes(x = end, y = 30, xend = start, yend = 30),
+               colour = "grey90", alpha=1, size=0.3, inherit.aes = FALSE) +
+  geom_bar(aes(x=as.factor(id),
+               y=mean,
+               fill=mean
+  ),
+  stat="identity",
+  alpha=0.5) +
+  annotate("text", 
+           x = 0.5,
+           y = 37,
+           label = "UN Agenda",
+           color= "grey60", 
+           size=3, 
+           angle=-5
+  ) +
+  # Country names
+  geom_text(data=label_data,
+            aes(x=id, 
+                y= 100,
+                label=sau_fishing_entity,
+                hjust=hjust),
+            color="black",
+            fontface="italic",
+            alpha = 1,
+            size = 4 ,
+            angle = label_data$angle,
+            inherit.aes = FALSE) +
+  # Continents labels
+  # geom_text(data=base_data,
+  #           aes(x = title,
+  #               y = 85, 
+  #               label=group), 
+  #           # hjust=c(1,1,1,0,0,0),
+  #           colour = "black",
+  #           # angle = myAng,
+  #           size=5,
+  #           fontface="bold",
+  #           inherit.aes = FALSE) +
+  ylim(-100,100) + # sets the circle the first value the central the last value how wide
+  theme_minimal() +
+  theme(
+    legend.position = c(0.5,0.5),
+    legend.text = element_text(size = 13),
+    legend.title = element_text(size = 14, face = "bold"),
+    legend.title.align = 0.5,
+    axis.text = element_blank(),
+    axis.title = element_blank(),
+    panel.grid = element_blank(),
+    panel.background = element_rect(colour = "white"),
+    legend.direction="horizontal",
+    legend.key.width = unit(3, 
+                            "line")
+  ) +
+  coord_polar(start = 0) +
+  scale_fill_viridis("Average year of range shift\nfor top 5 most valuable stocks",
+                     limits = c(0,100),
+                     breaks = seq(0,100,20),
+                     labels = seq(2000,2100,20),
+                     direction = -1
+  ) +
+  guides(
+    fill=guide_colorbar(title.position="top", 
+                        title.hjust =0.5)
+  )
+
+ggsave(
+  plot = last_plot(),
+  width = 13,
+  height = 13,
+  units = "in",
+  filename = "./Figures/top_stocks_circular_trend.png"
+)
+
+```

### 2021-11-23 Scripts/Supplements.Rmd
Fixed supplemental figures order and legends

@@
 
-```{r toeFigureS5, eval = T,  echo = F, fig.width = 10, fig.height= 8, fig.pos="H", fig.cap ="World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their estimated centroids (black points)."}
+```{r toeFigureS1, eval = T,  echo = F, fig.width = 10, fig.height= 8, fig.pos="H", fig.cap ="World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their estimated centroids (black points)."}
 
-img <- readPNG("../Figures/FigS7.png")
+img <- readPNG("../Figures/FigS1.png")
 grid.raster(img)
@@
 
-```{r toeFigureS7, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine the stock’s centroid.
-Top Row: results in absolute numbers.
-Bottom row: results expressed as a percentage of the total emerging stocks and EEZs with emerging stock."}
+```{r toeFigureSA, eval =F,  echo = F, fig.width=4, fig.height=4, fig.cap ="Average distribution of transboundary stock’s year of change in their shared distribution.
+Left panel is the average year across stocks and right panel the standard deviation.
+Yellow represents a two standard deviation distance from the historical average and purple represents a one standard deviation distance from the historical average."}
 
-img <- readPNG("../Figures/FigS6.png")
+img <- readPNG("../Figures/FigSA.png")
 grid.raster(img)
@@
 
-```{r toeFigureS2, eval = T,  echo = F, fig.width=13, fig.height=13, fig.cap ="Year of change in the shared distribution of 5,745 transboundary stocks using a one standard deviation threshold.
+```{r toeFigureS2, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine the stock’s centroid.
+Top Row: results in absolute numbers.
+Bottom row: results expressed as a percentage of the total emerging stocks and EEZs with emerging stock."}
+
+img <- readPNG("../Figures/FigS2.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigureS3, eval = T,  echo = F, fig.width=13, fig.height=13, fig.cap ="Year of change in the shared distribution of 5,745 transboundary stocks using a one standard deviation threshold.
 A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
@@
 
-img <- readPNG("../Figures/FigS1.png")
-grid.raster(img)
-
-```
-
-\clearpage
-
-```{r toeFigureS8, eval = T,  echo = F, fig.width=4, fig.height=4, fig.cap ="Average distribution of transboundary stock’s year of change in their shared distribution.
-Left panel is the average year across stocks and right panel the standard deviation.
-Yellow represents a two standard deviation distance from the historical average and purple represents a one standard deviation distance from the historical average."}
-
-img <- readPNG("../Figures/FigS2.png")
-grid.raster(img)
-
-```
-
-\clearpage
-
-```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=6,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
-Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
-The category “Other” consists of species that have no “habitat preference” classification.
-The number of stocks included in this analysis for each habitat type is noted in parentheses.
-Dashed whiskers represent the 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Solid line represents the median time of emergence per habitat preference."}
-
 img <- readPNG("../Figures/FigS3.png")
@@
 
-```{r toeFigureS3, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Distribution of changes in transboundary stock proportion per stock and neighboring Exclusive Economic Zones.
-Yellow color represents changes by early, 2030 (2021-2040) and purple by mid 2050 (2041-60) 21st century relative to 1951-2005."}
+```{r toeFigureS4,  eval = T,  echo = F, fig.width=13, fig.height=13, fig.cap ="Year of change in the shared distribution of 4,119 transboundary stocks and their contribution to fishing revenue.
+A) Contribution of shifting stocks to a country or territory's total fishing revenue from transboundary stocks.
+B) Average year of range shift across an EEZ’s transboundary stocks.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
+This figure is identical to Figure 3A but shows the information in two different maps and color scales; A shows the land polygons in 3A and B shows the EEZ polygons in 3B."}
 
@@
 
-```{r toeFigureS4, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap ="Changes in stock proportion of neighboring Exclusive Economic Zones by 2050 (2041-2060) relative to 1951-2005.
+```{r toeFigureS6, eval = T,  echo = F, fig.width=9, fig.height=6,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
+Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
+The category “Other” consists of species that have no “habitat preference” classification.
+The number of stocks included in this analysis for each habitat type is noted in parentheses.
+Dashed whiskers represent the 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Solid line represents the median time of emergence per habitat preference."}
+
+img <- readPNG("../Figures/FigS6.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigureS7, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap ="Changes in stock proportion of neighboring Exclusive Economic Zones by 2050 (2041-2060) relative to 1951-2005.
 Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors.
-Panel B highlights changes for the Caribbean and Northwestern Europe."}
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors."}
 
-img <- readPNG("../Figures/FigS5.png")
+img <- readPNG("../Figures/FigS7.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigureS8, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Distribution of changes in transboundary stock proportion per stock and neighboring Exclusive Economic Zones.
+Yellow color represents changes by early, 2030 (2021-2040) and purple by mid 2050 (2041-60) 21st century relative to 1951-2005."}
+
+img <- readPNG("../Figures/FigS8.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
+```{r toeFigureS9, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Changes in stock proportion of neighboring Exclusive Economic Zones of Latin America and the Caribeen and Northwestern Europe by 2030 (2021-2040) and by 2060 (2041-2060) relative to 1951-2005.
+Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors.}
+
+img <- readPNG("../Figures/FigS9.png")
 grid.raster(img)

### 2021-11-23 Scripts/Manuscript.Rmd
updated suppplemental figures order

@@
 We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015; http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
-\@ref(fig:toeFigureS5)) including island territories.
+\@ref(fig:toeFigureS1)) including island territories.
 We determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp].
@@
 Thus, we estimated the centroid of a stock only including grid cells where the projected stock abundance within the neighboring EEZs sharing the stock was above the $95^{th}$ percentile.** A sensitivity analysis using a subset of species (n = 34) for all EEZs  to examine the effects of different thresholds (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) on the calculated index value showed no apparent difference between percentiles (Fig.
-\@ref(fig:toeFigureS7)).
+\@ref(fig:toeFigureS2)).
 The centroid of each EEZ was estimated using the st package in R (Fig.
-\@ref(fig:toeFigureS5)).
+\@ref(fig:toeFigureS1)).
 For each ensemble member, neighboring EEZs and transboundary stock, we computed the distance between centroids (i.e., EEZ and stock) using the geosphere package in R assuming the Earth is a perfect sphere and ignoring geographic barriers;
@@
 In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs.
-\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
+\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS8)).
 -->
@@
 representing a probability of 67% that the stock has shifted) **results in 18% (n = 5,745) more shifting stocks, that is, 63% of all studied stocks** (Figs.
-\@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
+\@ref(fig:toeFigureS3), \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)).
 For both cases, the first shift is modelled to have occurred back in 2006.
@@
 threshold; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig.
-\@ref(fig:toeFigureS1)) and EEZs' geographic regions (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; using the 2 s.d.
+\@ref(fig:toeFigureS6)) and EEZs' geographic regions (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; using the 2 s.d.
 threshold; Table S3 and Fig.
@@
 Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2) and \@ref(tab:toeTableS3) than almost any other region (Figs.
-\@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS1) and \@ref(fig:toeFigureS2)).
+\@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).
 In contrast, EEZs and stocks located in temperate regions, like northern Europe and eastern Asia, are projected to experience much later range shifts.
@@
 Results from the sensitivity analysis suggest that our method is robust to the abundance level (e.g., the top 95%) of the stock (Fig.
-\@ref(fig:toeFigureS3)).
+\@ref(fig:toeFigureS6)).
 In addition, we set an arbitrary threshold of 2 s.d.
@@
 Projections show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks will change beyond the historic variability experienced within an EEZ’s (Fig.
-\@ref(fig:toeFigure4)A).
+\@ref(fig:toeFigure4)).
 Moreover, 85% (n = 239) of the world’s EEZs will have experienced changes in catch proportion of transboundary stocks by 2030.
 By mid-21st century, the number of EEZs with changing stocks, as well as the number of stocks experiencing changes in catch proportion and the magnitude of that change will increase (Figs.
-\@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).
+\@ref(fig:toeFigureS7) and \@ref(fig:toeFigureS8)).
 The direction and intensity of shifts in transboundary stocks are largely related to regional changes in biogeography and the geometry of EEZs (Fig.
-\@ref(fig:toeFigure4)A).
+\@ref(fig:toeFigure4).
 Along the Atlantic and Pacific coasts of Northern and Southern America and the Atlantic coast of Southern Africa, shifts in stocks are expected to benefit poleward EEZs.
 However, shifts along the coasts of Pacific Central America and West Africa occur in an equatorial direction.
+Regions where EEZs are relatively small and have multiple borders (e.g., Caribbean or Northwestern Europe) are expected to have a particularly complex exchange of stock share ratio with no clear established pattern (Fig.
+\@ref(fig:toeFigureS9)).
 
@@
 Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors.
-Panel B highlights changes for the Caribbean and Northwestern Europe."}
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors."}
 
@@
 **Most EEZs are expected to see a relatively small number of stocks shifting their SSR (Fig.
-\@ref(fig:toeFigure4)A).
+\@ref(fig:toeFigure4)).
 On average, EEZs are expected to see changes in SSR on 16% $\pm$ 10 of transboundary stocks increasing to 20% $\pm$ 11 by mid of the $21^{st}$ century.

### 2021-11-23 Scripts/Initial_Analysis.Rmd
FIxed missing siland nations of Fig S4A

@@
 
-World_map <- rnaturalearth::ne_countries(scale = 'small', returnclass = c("sf")) %>%
+World_map <- rnaturalearth::ne_countries(scale = 'large', returnclass = c("sf")) %>%
   clean_names() %>% 
@@
 
-# ggplot(Eez_centroid_sf) +
-#   geom_sf() +
-#   geom_sf(data = World_map, aes(), fill = "grey") +
-#   geom_sf(data = World_eez_sf, aes(), fill = "white")
-
 # Transform sf to df
@@
 ### ------------ ##
+
+```
+
+```{r}
+
+centroid_map <- ggplot() +
+  geom_sf(data = World_eez_sf, aes()) +
+  geom_sf(data = subset(World_map, !sovereignt %in% c("Antarctica")), aes(),fill = "grey90") +
+  geom_sf(data = Eez_centroid_sf, aes()) +
+  my_ggtheme_m(leg_pos = "")
+
+ggsave(
+  plot = centroid_map,
+  "./Figures/centroid_map.png",
+  width = 8,
+  height = 5,
+  units = "in"
+)
 
@@
 
-land_map <- ggplot(World_map_part) +
-  geom_sf(aes(fill = participation-2000),
-          color =  "white",
+land_map <- ggplot() +
+  geom_sf(data = World_map_part,
+          aes(fill = participation-2000, 
+              color = participation-2000),
           size = 0.1
   )  +
-  geom_sf(data = no_toe_land, aes(fill = as.numeric(value)),
-          color =  "white",
+  geom_sf(data = no_toe_land, aes(fill = as.numeric(value),
+                                  color = as.numeric(value)),
           size = 0.1) +
   scale_fill_viridis("Transboundary fish revenue (%)",
+                     limits = c(0,100),
+                     breaks = seq(0,100,25),
+                     labels = seq(0,100,25),
+                     direction = 1,
+                     na.value = "lightblue",
+                     option = "A"
+  ) +
+  scale_color_viridis("Transboundary fish revenue (%)",
                       limits = c(0,100),
@@
 )
+
+ggplot() +
+  geom_sf(data = rnaturalearth::ne_countries(scale = 'large', returnclass = c("sf")),aes())
 

### 2021-11-24 Scripts/Initial_Analysis.Rmd
finalized supplemental plots

@@
 
-This results only shows the average percentage change in proportion of all speecies within an EEZ without relationship with neighbors
+This results only shows the average percentage change in proportion of all species within an EEZ without relationship with neighbours
 
@@
 
-This function estimates the total change in MCP proportion of each species for each EEZ. It looses the connection with neighbors as you only know for x species how much is the proportion changing.
+This function estimates the total change in MCP proportion of each species for each EEZ. It looses the connection with neighbours as you only know for x species how much is the proportion changing.
 
@@
 
-The next step is to summarize all species per eez so our results shows average change of all species within an EEZ towards a neighbour.
+The next step is to summarize all species per EEZ so our results shows average change of all species within an EEZ towards a neighbour.
 
@@
 
-# Neighbor list and their respective id (see chunk 3)
+# Neighbor list and their respective id from Palacios-Abrantes et al 2020 (FishForVisa)
 Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)
@@
 
-# Species catch trend (from FishForVisa)
-
-# Catch_trend <- my_path("D","Species","Catch_trend.csv", read = T) %>% 
-#   clean_names() %>% 
-#   filter(taxon_key %in% Taxon_list_emer) %>% 
-#   rename(sau_fishing_entity = territory) %>% 
-#   left_join(Fishing_names,
-#             by = "sau_fishing_entity") %>% 
-#   filter(status != "No Trend") %>% 
-#   mutate(category = ifelse(status %in% c("Developing","Rebuilding"),"Category A",
-#                          ifelse(status == "Max Exploited","Category B",
-#                                 ifelse(status %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
-#                          )
-#          )
-#   )
-
 # --------------- #
@@
 # For dat availability
-# write_csv(prop_results,
-# "proportion_change_2005.csv")
+# write_csv(Prop_results,
+# "Data/ssr_data.csv")
 
@@
 # warnings() for empty data species # New version of package being a pain in the arse
+
+# Write data file for Github
+
+toe_data <- bind_rows(lapply(Paths, fread)) %>% 
+  filter(
+    n_ens == 10
+  ) %>% 
+  left_join(Exploited_spp_list) %>% 
+            select(taxon_key,taxon_name,common_name,everything(),-genus,-species)
+
+write_csv(toe_data, "./Data/toe_data.csv")
 
@@
             # aes(fill = participation+2000), #original map
-            aes(fill = participation),
-            color =  "white",
+            aes(fill = participation, 
+                color = participation),
+            # color =  "white",
             size = 0.1
@@
     # ggplot() +
-    geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
+    geom_sf(data = grey_land, aes(), fill = "grey50", 
+            # color =  "white",
+            size = 0.1) +
     # geom_sf(data = grey_eez, aes(), fill = "white", color =  "grey50",size = 0.1) +
@@
     ) +
+    scale_color_viridis("",
+                       limits = c(2000,2100),
+                       breaks = seq(2000,2100,25),
+                       labels = c("","","","","")#,
+                       # direction = -1
+    ) +
     my_ggtheme_m(leg_pos = "bottom") +

### 2021-11-24 Scripts/Initial_Analysis.Rmd
Removed old unecessary data

@@
   select(eezid,latitude,longitude)
-
-# Species depth classification (SAU-fishbase)
-Spp_depth <- fread(paste(my_path("D"),"SAU/SppTaxonName.csv", sep ="")) %>% 
-  clean_names()
-
-# Temperature Time of Emergence
-Toe_temp <- left_join(my_path("D","Climate", "ToE_sst.csv", read = T),
-                      my_path("D","Climate", "ToE_bottemp.csv", read = T)
-) %>% 
-  clean_names()
-
-# Biogeographic regions (from Rey 2020)
-Biogeo_reg <-fread(paste(my_path("D"),"Climate/SPATIAL_PARTITION_OCEAN.csv", sep ="")) %>% 
-  clean_names()
-
-# Fix coordinate system and re-name biomes
-Biogeo_reg <- suppressWarnings(Biogeo_reg[order(Biogeo_reg$lat, rev(Biogeo_reg$lon),decreasing=TRUE), ]) %>% 
-  mutate(index = seq(1,259200,1)) %>% 
-  mutate(
-    biome = ifelse(biome2 %in% c(1,4), "tropical",
-                   ifelse(biome2  %in% c(2,5),"temperate",
-                          ifelse(biome2  %in% c(3,6), "polar",
-                                 biome2)
-                   )
-    )
-  ) %>% 
-  filter(biome2 != "NaN")
-
-# biome2 1 = Coastal tropical 
-# biome2 2 = Coastal sub-tropical
-# biome2 3 = Coastal polar
-# biome2 4 = Oceanic tropical
-# biome2 5 = Oceanic sub-tropical
-# biome2 6 = Oceanic polar
-
-# Biogeo_reg %>%
-# gather("biome_type","value",biome:biome2) %>%
-# ggplot() +
-# geom_tile(
-#   aes(
-#     x = lon,
-#     y = lat,
-#     fill = biome
-#   )
-# )
-#   facet_wrap(~biome_type,
-#              ncol=1)
-
-# Get eez to index relational data
-EEZ_CellID <- my_path("G", extra_path = "Spatial/DBEM", name="EEZ_CellID.xlsx", read = TRUE) %>%
-  clean_names()
-colnames(EEZ_CellID) <- c("eezid","index")
-
-Eez_biogeo <- EEZ_CellID %>% 
-  # filter(eezid == 8) %>%
-  left_join(Biogeo_reg,
-            by = "index") %>% 
-  # Deal with NaNs in data
-  mutate(biome = ifelse(biome == "NaN",0,biome)) %>% 
-  filter(biome > 0) %>% 
-  # Get frequency of biome per index
-  group_by(eezid,biome) %>% 
-  summarise(nbiom=n()) %>% 
-  # Select the more frequent biome per eez
-  group_by(eezid) %>% 
-  top_n(1,nbiom) %>% 
-  ungroup() %>% 
-  select(-nbiom) %>% 
-  distinct(eezid, .keep_all = T) # Croatia is 1/2. We arbitrary allocated the first one
 

### 2021-11-24 Scripts/Initial_Analysis.Rmd
Updated repository for data availabillity

@@
 
-toe_data <- bind_rows(lapply(Paths, fread)) %>% 
-  filter(
-    n_ens == 10
-  ) %>% 
-  left_join(Exploited_spp_list) %>% 
-            select(taxon_key,taxon_name,common_name,everything(),-genus,-species)
-
-write_csv(toe_data, "./Data/toe_data.csv")
+# toe_data <- bind_rows(lapply(Paths, fread)) %>% 
+#   filter(
+#     n_ens == 10
+#   ) %>% 
+#   left_join(Exploited_spp_list) %>% 
+#             select(taxon_key,taxon_name,common_name,everything(),-genus,-species)
+# 
+# write_csv(toe_data, "./Data/toe_data.csv")
 
@@
 
+### Centroid Supplemental Figure
 ```{r}

### 2021-11-24 Scripts/Manuscript.Rmd
updated info on top 5 SSR plot

@@
 
-**Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 61 $\pm$ 18% by 2030 and 72 $\pm$ 17% by 2050 in 83% of the world's EEZs (Figure SX).
+**Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 65 $\pm$ 21% by 2030 and 78 $\pm$ 18% by 2050 in 77% of the world's EEZs (Figure SX).
 In some cases, the SSR of one EEZ is projected to increase more than double relative to their neighbors (e.g., from 10% to 30% SSR).
-Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Cyprus from Greece and Turkey in the Mediterranean Sea, and Russia from Norway and Japan in the Barents Sea and Western Pacific, respectively**
+Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Mozambique from Madagascar, and Russia from Norway and Japan in the Barents Sea**
 

### 2021-11-24 Scripts/responses_rev_gcb.Rmd
updated info on top 5 SSR plot

@@
 
-- line XX: *Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 61 $\pm$ 18% by 2030 and 72 $\pm$ 17% by 2050 in 83% of the world's EEZs (Figure SX).
+- line XX: *Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 65 $\pm$ 21% by 2030 and 78 $\pm$ 18% by 2050 in 77% of the world's EEZs (Figure SX).
 In some cases, the SSR of one EEZ is projected to increase more than double relative to their neighbors (e.g., from 10% to 30% SSR).
-Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Cyprus from Greece and Turkey in the Mediterranean Sea, and Russia from Norway and Japan in the Barents Sea and Western Pacific, respectively*
+Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Mozambique from Madagascar, and Russia from Norway and Japan in the Barents Sea*
 

### 2021-11-24 Scripts/Initial_Analysis.Rmd
FIxed top 5 ssr map

@@
 
-SummaryProp <- function(taxon_key,links="NA"){
+# taxon_key = "600147"
+# top_stocks_per_country
+
+SummaryProp <- function(taxon_key,links="NA", top = F){
+
+  if(top == T){
+    # Reads taxon data
+    proportion_data <- fread(paste(my_path("R",extra_path = "Proportion_2005c/"),"proportion_",taxon_key,".csv", sep="")) %>% 
+      filter(!is.na(eez_neighbour)) %>% 
+      rename(sf_eez_name = eez_name) %>% 
+      left_join(Matching_names) %>% 
+      semi_join(top_stocks_per_country) %>% 
+      select(
+      1:4,
+      ensemble_mean = catch_proportion_temp_mean_ensemble_mean,
+      ensemble_sd = catch_proportion_temp_mean_ensemble_sd) %>% 
+      rename(eez_name = sf_eez_name)
+
+  }else{
 
@@
 
+  }
+
   # Identify SSR treshold for natural spatial variability of the stock
@@
 
-EEZProp <- function(taxon_list,links="NA"){
+EEZProp <- function(taxon_list,links="NA",top = F){
 
   spp <- bind_rows(
-    lapply(taxon_list,SummaryProp,links)
+    lapply(taxon_list,SummaryProp,links,top = top)
   )
@@
 
-PropChange <- function(taxon_list,links="NA"){
-
-  eez <- EEZProp(taxon_list,links)
+PropChange <- function(taxon_list,links="NA", top = F){
+
+  eez <- EEZProp(taxon_list,links, top = top)
 
@@
 
-###  Year of Eergence Analysis
+###  Time of Emergence Analysis
 
@@
 
+Timeline of shifts in transboundary shared stock distribution.
+(A) Number of stocks (cumulative) whose distribution centroids that have shifted beyond natural variations over time (from 2006 to >2100), (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one stock shifting beyond natural variability.
+Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change in range shift (Methods), respectively.
+
+###### Generate data needed
+
 ```{r Time_lapse_data,  eval = F, echo = T}
@@
 ```
+
+###### Plot histograms
 
@@
 
-##### Fig. 3 Mapping the time of emergence
-
-###### Mean ToE and vulnerabillity Index
-
-```{r emergence_map, eval = F}
-
+##### Fig. 3 Mapping the time of emergence and the participation
+
+Year of change in the shared distribution of 4,119 transboundary stocks.
+A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygons display average year of range shifts within them.
+Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cold colors represent a late shift/low fishing revenue.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
+B) Year of shared distribution shifts by UN sub-regions.
+Points color coded by the median year for all shifting stocks in the region.
+Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda.
+N = North, S = South, W = West and E = East.
+Ltn.
+Ame.
+and the Car.
+= Latin America and the Caribbean.
+Aus and New Z.
+= Australia and New Zealand.
+
+###### Generate mean ToE across EEZs' stocks
+
+```{r emergence_map_data, eval = F}
 # Per EEZs
@@
 
-# grey_eez <- World_eez_sf %>% 
-#   filter(!eezid %in% Eez_d_hist$eezid,
-#          !eez_name %in% Neighbours$eez_name)
-
+```
+
+###### Plot ToE map
+
+```{r emergence_map_plot, eval = F}
 # -------------------------- #
@@
 
-Note this chunk needs the previous one to be ran up to the for-loop
+Note this chunk needs `emergence_map_plot` to be ran.
 
@@
 
-# Grey land
-grey_land <- World_map %>% 
-  # filter(is.na(fishing_entity_id)) %>% 
-  filter(!sovereignt %in% c("Antarctica"))
-
-# Inlcude the participation of species
-
-#   geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
-#   geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
-
+# EEZ map
 eez_map <- ggplot(toe_map_data) +
@@
 
+# Land Map
 land_map <- ggplot() +
@@
 
-# ggsave(
-#   plot = land_map,
-#   width = 8,
-#   height = 6,
-#   units = "in",
-#   filename = "./Figures/eez_map_supp.png"
-# )
-
  ## --------------- ##
@@
 
-ggplot() +
-  geom_sf(data = rnaturalearth::ne_countries(scale = 'large', returnclass = c("sf")),aes())
-
 ```
 
-##### Fig. 4 Mapping the time of emergence Of Top stocks
+##### Fig. 4 Mapping the time of emergence of Top stocks
 
@@
   # Select top 5 most valuable stocks
-  top_n(5,value) %>% 
-  left_join(Exploited_spp_list)
+  top_n(5,value)
 
@@
 
-### Species level analysis
-
-#### Fig. 5 Species Boxplot
+#### Species level analysis
+
+Supplemental box picture by habitat preference
 
@@
   left_join(Spp_env_pref) %>% 
-  mutate(DemersPelag = ifelse(is.na(DemersPelag),"other",paste(DemersPelag)
-                              #                             ifelse(DemersPelag %in% c("bathypelagic","benthopelagic"), "bathy-bentho-pelagic",paste(DemersPelag)
-                              #                             )
-  ),
+  mutate(DemersPelag = ifelse(is.na(DemersPelag),"other",paste(DemersPelag)),
   un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and the Car.",paste(un_sub_region)))
@@
 # ------------------- #
-# Test normal asumptions 
+# Test normal assumptions 
 # ------------------- #
 
-hist(Stats_data$emerging_yr) # Positevely skewed
+hist(Stats_data$emerging_yr) # Positively skewed
 qqnorm(Stats_data$emerging_yr) # not normally distributed
@@
 
-# Multiple differences but Eastern Asia, Lat. Ame. Car. and Polinesia are the nes who differe the most (e.g. differe with more revions).
+# Multiple differences but Eastern Asia, Lat. Ame. Car. and Polynesia are the ones who differ the most (e.g. differ with more regions).
 
@@
 
-##  Prooportion Analysis
+##  Stock Share Ratio Analysis
 
@@
 
-```{r fig5_paths, eval = T, echo = T}
-
+#### Prepare data
+
+```{r ssr_map_data, eval = T, echo = T}
 # A function that makes a dateframe per connection (we will use these connections to plot each lines)
@@
 }
+
+# Use a different palet between SSR and ToE
 pal <- wesanderson::wes_palette("Zissou1",100,type = "continuous")
@@
   left_join(Eez_centroid,
-            # left_join(Eez_centroid_eck,
             by = "eez_name") %>%
@@
   # get coords for targets
-  # left_join(Eez_centroid_eck,
   left_join(Eez_centroid,
@@
   filter(
-    variable %in% c("mean_eez")#,
-    # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
+    variable %in% c("mean_eez")
   )
@@
   tmp=data_for_connection(path_data$start_lon[i], path_data$start_lat[i], path_data$end_lon[i], path_data$end_lat[i] , i)
-  # tmp$homecontinent=summary$homecontinent[i]
   tmp$per_early_plot=path_data$per_early_plot[i]
@@
 
+```
+
+#### Select a time step
+
+```{r}
 # SELECT A TIME STEP
@@
 time_toplot <- "per_early"
-
-# ----------------------------- #
+```
+
+#### Get data ready for time step
+
+```{r ssr_map_data, eval = T, echo = T}
 
@@
   filter(!sovereignt %in% c("Antarctica"))
-
+```
+
+#### Plot SSR map
+
+```{r ssr_map_plot, eval = T, echo = T}
 # Plot
@@
 
-### Figure S6. Regional plots
-
-Note that the following chunk needs the previous one to be ran
+#### Regional maps for supplements
+
+Note that the following chunk needs `ssr_map_data` to work.
 
@@
 ## ------------------- ##
-# Regional Plots
+# Latin america Map
 ## ------------------- ##
 
-# Lat_am_land <- World_map %>% 
-# filter(region_wb %in% c("Latin America & Caribbean") |
-# sau_fishing_entity %in% c("USA","Mexico")
-# )
-
-Lat_am_eez <- World_eez_sf %>% 
-  select(sf_eez_name = eez_name,everything()) %>% 
-  left_join(Matching_names) %>% 
-  filter(
-    un_sub_region == "Latin America and the Caribbean",
-    !str_detect(sf_eez_name,"Pacific"),
-    !sf_eez_name %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador")
-  )
-
-Lat_am_data <- sf_ready_plot %>% 
-  filter(source %in% Lat_am_eez$sf_eez_name,
-         objective %in% Lat_am_eez$sf_eez_name)
-
-Lat_am_paths <- arrows_data %>% 
-  filter(source %in% Lat_am_eez$sf_eez_name,
-         objective %in% Lat_am_eez$sf_eez_name)
-
-Lat_am_points <- point_data %>% 
-  filter(source %in% Lat_am_eez$sf_eez_name,
-         target %in% Lat_am_eez$sf_eez_name)
-
-Lat_am_n_taxa <- n_taxa_map %>% 
-  filter(
-    region_wb == "Latin America & Caribbean",
-    # !str_detect(sf_eez_name,"Pacific"),
-    # !sf_eez_name %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador")
-  )
-
 lat_am_map <- ggplot() +
-  geom_sf(data = Lat_am_eez, size = 0.1, fill = "white", col = "gray30") +
-  geom_sf(data = subset(Lat_am_n_taxa, sovereignt != "Antarctica"),
-          aes(fill = avrg_spp_per),
-          size = 0.1,
-          col = "white") +
-  # geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
-  geom_sf(data = subset(Lat_am_data, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
-  geom_sf(data = Lat_am_points, aes(), color = "grey90") +
-  geom_path(data = subset(Lat_am_paths, percentage >= 0),
-            aes(X, Y,
-                group = group,
-                col = percentage
-            ),
-            size = 1,
-            arrow = arrow(length = unit(0.01, "npc"),
-                          ends = "first"),
-            show.legend = NA) +
-  my_ggtheme_m(leg_pos = "bottom") +
-  coord_sf(xlim = c(-9000000,-5000000),ylim = c(1000000, 4000000)) + 
-  labs(x = "", y = "") +
-  # scale_x_continuous(breaks = seq(-100,60,10)) +
-  # scale_y_continuous(breaks = seq(10,30,10)) +
-  my_ggtheme_m(leg_pos = "bottom") +
-  theme(
-    panel.background = element_rect(fill = "white", color = "white"),
-    panel.grid.major =  element_line(color = "white")
-  ) +
-  scale_color_gradientn(name= "Changes in catch proportion and changing stocks (%)",
-                        colours = pal,
-                        limits = c(0,max(Lat_am_data$percentage)),
-                        breaks = seq(0,max(Lat_am_data$percentage),10)
-  ) +
-  scale_fill_gradientn(name= "Changes in catch proportion and changing stocks (%)",
-                       colours = pal,
-                       limits = c(0,max(Lat_am_data$percentage)),
-                       breaks = seq(0,max(Lat_am_data$percentage),10)
-  ) + guides(
-    color=guide_colorbar(title.position="top", 
-                         title.hjust =0.5)
-  ) +
-  theme(legend.key.width = unit(4,"line"))
-
-ggsave(filename = "lat_am_mid.png",
-       lat_am_map,
-       width = 6,
-       height = 6)
-
-## ------------------- ##
-# Northern Europe
-## ------------------- ##
-
-EU_data <- sf_ready_plot %>% 
-  filter(source %in% EU_eez$sf_eez_name)
-
-EU_map <- ggplot() +
   geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
@@
             show.legend = NA) +
-  # my_ggtheme_m(leg_pos = "bottom") +
+  coord_sf(xlim = c(-9000000,-5000000),ylim = c(1000000, 4000000)) + 
+  labs(x = "", y = "") +
+  # scale_x_continuous(breaks = seq(-100,60,10)) +
+  # scale_y_continuous(breaks = seq(10,30,10)) +
+  my_ggtheme_m(leg_pos = "bottom") +
+  theme(
+    panel.background = element_rect(fill = "white", color = "white"),
+    panel.grid.major =  element_line(color = "white")
+  ) +
+  scale_color_gradientn(name= "Changes in catch proportion and changing stocks (%)",
+                        colours = pal,
+                        limits = c(0,max(Lat_am_data$percentage)),
+                        breaks = seq(0,max(Lat_am_data$percentage),10)
+  ) +
+  scale_fill_gradientn(name= "Changes in catch proportion and changing stocks (%)",
+                       colours = pal,
+                       limits = c(0,max(Lat_am_data$percentage)),
+                       breaks = seq(0,max(Lat_am_data$percentage),10)
+  ) + guides(
+    color=guide_colorbar(title.position="top", 
+                         title.hjust =0.5)
+  ) +
+  theme(legend.key.width = unit(4,"line"))
+
+ggsave(filename = "lat_am_mid.png",
+       lat_am_map,
+       width = 6,
+       height = 6)
+
+## ------------------- ##
+# Northern Europe Map
+## ------------------- ##
+
+EU_map <- ggplot() +
+  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
+  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+          aes(fill = avrg_spp_per),
+          size = 0.1,
+          col = "white") +
+  geom_sf(data = subset(sf_ready_plot, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
+  geom_sf(data = point_data, aes(), color = "grey90") +
+  geom_path(data = subset(arrows_data, percentage >= 0),
+            aes(X, Y,
+                group = group,
+                col = percentage
+            ),
+            size = 1,
+            arrow = arrow(length = unit(0.03, "npc"),
+                          ends = "first"),
+            show.legend = NA) +
   coord_sf(xlim = c(-1300000,2000000),ylim = c(5500000, 7500000))  +
@@
 ```
+
+#### SSR TOP 5 figure
+
+```{r fig_map_Juan, eval = F, echo =F}
+
+top_stocks_per_country <- Eme_Res %>% 
+  rename(geo_region = un_sub_region) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,geo_region,tresh,taxon_key) %>% 
+  summarise(n()) %>% 
+  left_join(Sau_data,
+            by = c("fishing_entity_id","taxon_key")) %>% 
+  filter(!is.na(variable)) %>% 
+  # group_by(sau_fishing_entity,fishing_entity_id,tresh,variable) %>% 
+  # summarise(sum_value = sum(value,na.rm=T)) %>% 
+  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
+  # Select top 5 most valuable stocks
+  top_n(5,value)
+
+top_taxon_list <-  unique(top_stocks_per_country$taxon_key)
+# Read proportion results
+suppressMessages(
+  top_prop_results <- PropChange(top_taxon_list, links="y",top = T) %>%
+    filter(!is.na(eez_neighbour)) %>%
+    mutate(per_mid_plot =ifelse(per_mid >= 100, 100,
+                                ifelse(per_mid <= -100,-100,per_mid)
+    ),
+    per_early_plot =ifelse(per_early >= 100, 100,
+                           ifelse(per_early <= -100,-100,per_early)
+    )
+    )
+)
+
+### Median number of stocks changing in each time step
+
+top_prop_results %>% 
+  filter(variable == "n_per") %>% 
+  group_by() %>% 
+  summarise_at(vars(early,mid),funs(
+    median(.,na.rm=T),
+    sd(.,na.rm=T),
+    length(unique(eez_name)),
+    length(unique(eez_name))/280*100
+  )
+  )
+
+# Prepare data from results
+top_data <- top_prop_results %>% 
+  # get coords for sources
+  left_join(Eez_centroid,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  left_join(Eez_centroid,
+            by = "eez_name") %>% 
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  mutate(
+    per_early = ifelse(is.na(per_early),0,per_early),
+    per_mid = ifelse(is.na(per_mid),0,per_mid)
+  ) %>% 
+  drop_na() %>%
+  # gather("variable","value",per_early:per_early_plot) %>% 
+  filter(variable %in% c("mean_eez"))
+
+top_data_ready_plot=data.frame()
+
+for(i in c(1:nrow(top_data))){
+  tmp=data_for_connection(top_data$start_lon[i], top_data$start_lat[i], top_data$end_lon[i], top_data$end_lat[i] , i)
+  tmp$per_early_plot=top_data$per_early_plot[i]
+  tmp$per_mid_plot=top_data$per_mid_plot[i]
+  tmp$source=top_data$source[i]
+  tmp$objective=top_data$target[i]
+  top_data_ready_plot=rbind(top_data_ready_plot,tmp)
+}
+
+top_data_ready_plot <- top_data_ready_plot %>% 
+  mutate(
+    per_early = ifelse(per_early_plot >= 50,50,per_early_plot),
+    per_mid = ifelse(per_mid_plot >= 50,50,per_mid_plot)) %>% 
+  gather("time_step","percentage",per_early,per_mid)
+
+top_sf_ready_plot <- top_data_ready_plot %>% 
+  filter(time_step == "per_early") %>% 
+  st_as_sf(coords = c("lon", "lat")) %>% 
+  st_set_crs(4326) %>% # to match shapefiles
+  st_transform(crs = "+proj=eck4")
+
+# Get arrow data
+top_arrows_data <- as.data.frame(st_coordinates(st_cast(top_sf_ready_plot,"POINT")$geometry)) %>% 
+  bind_cols(top_data_ready_plot %>% 
+              filter(time_step == "per_early")
+  )
+
+# Get point data
+top_point_data <- top_data %>% 
+  gather("time_step","percentage",per_early:per_mid) %>% 
+  filter(time_step == "per_early",
+         percentage <= 0) %>% 
+  st_as_sf(
+    coords = c("start_lon", "start_lat")
+  ) %>% 
+  select(source,target,geometry) %>% 
+  st_set_crs(4326) %>% 
+  st_transform(crs = "+proj=eck4")
+
+# # # Estimate the percentage changing
+# n_taxa_data <- top_prop_results %>%
+#   filter(variable == "n_spp") %>%
+#   select(eez_name,early,mid) %>%
+#   left_join(n_trans_eez) %>%
+#   # mutate(per = (early/n_trans)*100) %>%
+#   mutate_at(vars(early,mid),
+#             .funs = ~(./n_trans)*100) %>%
+#   rename(sf_eez_name = eez_name) %>%
+#   left_join(Matching_names) %>%
+#   select(sau_fishing_entity,fishing_entity_id,early,mid) %>%
+#   gather("time_step","avrg_spp_per",early:mid) %>%
+#   group_by(sau_fishing_entity,fishing_entity_id,time_step) %>%
+#   summarise(avrg_spp_per = mean(avrg_spp_per,na.rm=T)) %>%
+#   mutate(time_step = ifelse(time_step == "early","per_early","per_mid")) %>%
+#   mutate(avrg_spp_per_100 = avrg_spp_per,
+#          avrg_spp_per = ifelse(avrg_spp_per > 50,50,avrg_spp_per)) %>%
+#   filter(time_step == time_toplot)
+
+# Included in map
+# n_taxa_map <- World_map %>% 
+#   full_join(n_taxa_data,
+#             by = c("sau_fishing_entity","fishing_entity_id")
+#   )
+
+```
+
+```{r}
+top_ssr_map <-
+  ggplot() +
+  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
+  # geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
+  #         aes(fill = avrg_spp_per),
+  #         size = 0.1,
+  #         col = "white") +
+  geom_sf(data = subset(World_map,sovereignt != "Antarctica"),
+          size = 0.1, fill = "grey90", col = "gray30") +
+  geom_sf(data = subset(top_sf_ready_plot, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
+  geom_sf(data = top_point_data, aes(), color = "grey90") +
+  geom_path(data = subset(top_arrows_data, percentage >= 0),
+            aes(X, Y,
+                group = group,
+                col = percentage
+            ),
+            size = 1,
+            arrow = arrow(length = unit(0.01, "npc"),
+                          ends = "first"),
+            show.legend = NA) +
+  my_ggtheme_m(leg_pos = "bottom") +
+  scale_color_gradientn(name= "Changes in catch proportion and changing stocks (%)",
+                        colours = pal,
+                        limits = c(0,max(top_sf_ready_plot$percentage,na.rm=T)),
+                        breaks = seq(0,max(top_sf_ready_plot$percentage,na.rm=T),10),
+                        labels = c(seq(0,max(top_sf_ready_plot$percentage-10,na.rm=T),10),">50")
+  ) +
+  scale_fill_gradientn(name= "Changes in catch proportion and changing stocks (%)",
+                       colours = pal,
+                       limits = c(0,max(top_sf_ready_plot$percentage,na.rm=T)),
+                       breaks = seq(0,max(top_sf_ready_plot$percentage,na.rm=T),10),
+                       labels = c(seq(0,max(top_sf_ready_plot$percentage-10,na.rm=T),10),">50")
+  ) + guides(
+    color=guide_colorbar(title.position="top")
+  ) +
+  theme(legend.key.width = unit(4,"line"))
+
+# Save figure
+ggsave(plot =  top_ssr_map,
+       "./Figures/ssr_top.png",
+       width = 8,
+       height = 5,
+       units = "in")
+
+```
+
+##### Fig S2 (Histogram of change)
+
+```{r hist_change}
+
+Prop_results %>% 
+  gather("time_step","change",per_early,per_mid) %>% 
+  filter(variable == "mean_eez" & 
+           !is.na(change)
+  ) %>%
+  ggplot() +
+  geom_density(
+    aes(
+      x = ifelse(change >= 100,100,change),
+      # x = change,
+      fill = ifelse(time_step == "per_early","q",time_step)
+      # fill = time_step,
+
+    ),
+    # stat = "count",
+    # alpha= 0.8,
+    size=0.1,
+    # position = "stack"
+  ) +
+  scale_y_continuous("Density",
+                     # breaks = seq(0,90,10),
+                     # limits = c(0,100)
+  ) +
+  scale_x_continuous("Percentage Change",
+                     breaks = seq(-75,500,25)
+  ) +
+  my_ggtheme_p(leg_pos = "right") +
+  scale_fill_viridis(discrete = T, 
+                     option = "plasma",
+                     alpha = 0.5,
+                     name = "Time Step", labels = c("Middle", "Early")
+  ) +
+  coord_flip() +
+  ggsave("./Figures/hist_change.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
+```
+
+## Sensitivity analysis
+
+### Testing different Per values for the centroids
+
+```{r data_loading, eval = F, echo = F}
+
+# SAU exploited species list
+Exploited_spp_list <- my_path("G","SAU","exploited_species_list.csv",read = T) %>%
+  clean_names() %>% 
+  pull(taxon_name)
+
+# Load all sp in one table
+sen_list <- c(20,50,90,95)
+
+sen <- 95
+read_sen <- function(sen){
+
+  if(sen != 95){
+
+    level <- paste0("Emergence_2005_p",sen)
+    extra_path <- paste0("Toe_Sens_per/",level)
+
+    # Get taxon list
+    Taxon_list_sens <- list.files(my_path("R",extra_path = extra_path))
+
+    Paths_sen <- paste(my_path("R",extra_path = extra_path),Taxon_list_sens,sep="")
+
+  }else{
+
+    Taxon_list <- gsub("_.*","",Taxon_list_sens)
+
+    Paths_sen <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_list_sens,"_emergence.txt",sep="")
+
+  }
+
+  Eme_sen <- bind_rows(lapply(Paths_sen, fread)) %>% 
+    mutate(sen = sen) %>% 
+    filter(
+      n_ens == 10
+    ) %>% 
+    left_join(Neighbours,
+              by = c("eez_name","eez_neighbour")
+    ) %>% 
+    rename(sf_eez_name=eez_name) %>% 
+    left_join(Matching_names,
+              by=  "sf_eez_name") %>% 
+    clean_names()
+
+  return(Eme_sen)
+
+}
+
+Sen_data <- bind_rows(
+  lapply(sen_list,read_sen)
+) %>% 
+  filter(
+    tresh== "tresh_two"
+  )
+
+unique(Sen_data$sen)
+
+```
+
+```{r data_analysis, eval = F, echo = F}
+
+# Per Stock
+sen_stock_d <-  
+  Sen_data %>% 
+  left_join(Neighbours) %>%
+  group_by(taxon_key,neighbour_id,sen) %>% 
+  summarise(
+    taxon_emer = min(emerging_yr)
+  ) %>% 
+  group_by(taxon_emer,sen) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  group_by(sen) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "stock"
+  )
+
+#  EEZ data
+sen_eez_d <-  Sen_data %>% 
+  left_join(Neighbours) %>% 
+  filter(n_ens == 10) %>% 
+  group_by(sf_eez_name,sen) %>% 
+  summarise(
+    taxon_emer = min(emerging_yr)
+  ) %>% 
+  group_by(taxon_emer,sen) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  group_by(sen) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "eez"
+  )
+
+# Plot Data
+sen_plot_data <- sen_stock_d %>% 
+  bind_rows(sen_eez_d) %>% 
+  mutate(
+    # total = ifelse(level == "stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$sf_eez_name))),
+    total = ifelse(level == "stock",max(sen_stock_d$com_sum),max(sen_eez_d$com_sum)),
+    per = com_sum/total*100
+  ) %>% 
+  gather("metric","value",com_sum,per) %>% 
+  ungroup() %>% 
+  mutate(
+    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
+    label = ifelse(total == max(sen_stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs"),
+    sen = paste("Percentile above",sen,"%") 
+  )
+
+# Lines data
+dashed_lines <- tibble(
+  x = c(2030,2050,2080),
+  label = c("Agenda 2030","Early range shift  ","Late range shift   ")
+)
+
+# Plot me
+ggplot() +
+  geom_area(data = subset(sen_plot_data, level == "stock"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  )  +
+  geom_area(data = subset(sen_plot_data, level == "eez"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  )  +
+  # Shameslly plot it again to fix colors
+  geom_area(data = subset(sen_plot_data, level == "stock" & my_facet == "Percentage (%)"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  ) +
+  facet_wrap(~my_facet + sen,
+             scales = "free",
+             nrow=2) + 
+  scale_y_continuous("",
+                     expand = c(0,0)
+  ) +
+  scale_x_continuous("Year of range shift",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     label = c(seq(2000,2075,25),">2100"),
+                     expand = c(0,0)
+  ) +
+  # Add dashed lines and labels
+  geom_vline(
+    data = dashed_lines,
+    aes(
+      xintercept = x,
+      linetype = label,
+    ),
+    color = "grey"
+  ) + 
+  my_ggtheme_p() +
+  theme(
+    legend.position = "right",
+    panel.spacing = unit(3, "lines") # space between plots
+  ) +
+  guides(fill=guide_legend(title= "", reverse=FALSE),
+         linetype = guide_legend(title = "Time step")) +
+  scale_fill_viridis(discrete = T,
+                     begin = 0.1, 
+                     end = 0.7,
+                     labels = c("EEZs","Stocks")
+  ) +
+  ggsave("./Figures/Supp3.png",
+         width = 14,
+         height = 10,
+         units = "in"
+  )
+
+```
+
+### Noise for all ensemble members
+
+```{r read_raw_data, eval = T, echo = F}
+
+# --------------- #
+# Load year of emergence results
+# --------------- #
+
+fread2 <- function(x){
+
+  tb <- fread(x)
+
+  if(nrow(tb) >0){
+    return(tb)
+  }
+
+}
+
+sens_paths <- list.files(my_path("R","Emergence_alt_sens"),full.names = T)
+
+# Load all sp in one table
+Eme_Alt_Res <- bind_rows(lapply(sens_paths, fread2)) %>% 
+  mutate(method= "alt") %>% 
+  left_join(Neighbours,
+            by = c("eez_name","eez_neighbour")
+  ) %>% 
+  rename(sf_eez_name=eez_name) %>% 
+  left_join(Matching_names,
+            by=  "sf_eez_name") %>% 
+  clean_names() %>% 
+  mutate(tresh = str_to_lower(tresh)) %>% 
+  filter(emerging_yr != 2100)
+
+# Combine both datas
+sens_data <- Eme_Res %>% 
+  filter(taxon_key %in% Eme_Alt_Res$taxon_key) %>% 
+  mutate(method = "org") %>% 
+  bind_rows(Eme_Alt_Res)
+
+n_stock <- sens_data %>% 
+  filter(method == "org") %>% 
+  group_by(neighbour_id) %>% 
+  summarise(n_stock = length(unique(taxon_key))) %>% 
+  # View()
+  group_by() %>% 
+  summarise(t_trans_stocks = sum(n_stock)) %>% 
+  pull(t_trans_stocks)
+
+n_eez <- length(unique(sens_data$sf_eez_name))
+
+```
+
+#### General numbers
+
+```{r sens_numbers, eval = T, echo = F}
+
+# Total number of emerging stocks
+
+# Global estimates
+sens_data %>% 
+  group_by(tresh,neighbour_id,method) %>% 
+  summarise(
+    emerging_stocks = length(unique(taxon_key))
+  ) %>% 
+  group_by(tresh,method) %>% 
+  summarise(
+    emerging_stocks = sum(emerging_stocks),
+    prop_total = (emerging_stocks/n_stock)*100,
+  )
+
+# Looks like there is no difference in the total number of emerging stocks
+# tresh     method emerging_stocks prop_total
+# 1 tresh_one alt               6665      116. 
+# 2 tresh_one org               5729      100  
+# 3 tresh_two alt               5258       91.8
+# 4 tresh_two org               4108       71.7
+
+# Global average ToE
+sens_data %>% 
+  group_by(tresh,method) %>% 
+  summarise(
+    # emerging_stocks = length(unique(taxon_key)),
+    # prop_total = (emerging_stocks/9132)*100,
+    mean= mean(emerging_yr),
+    sd = sd(emerging_yr),
+    min = min(emerging_yr),
+    max = max(emerging_yr)
+  )
+
+# Looks like the alternative method sets a later mean ToE but no real difference in s.d. min nor max
+#  tresh     method  mean    sd   min   max
+# 1 tresh_one alt    2055.  28.7  2007  2099
+# 2 tresh_one org    2029.  26.7  2006  2099
+# 3 tresh_two alt    2057.  28.6  2007  2099
+# 4 tresh_two org    2036.  28.1  2006  2099
+
+```
+
+#### Area plots 
+```{r read_raw_data, eval = T, echo = F}
+
+# Per Stock
+stock_d <-  
+  sens_data %>% 
+  # left_join(Neighbours) %>% 
+  # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
+  group_by(taxon_key,neighbour_id,tresh,method) %>% 
+  summarise(
+    taxon_emer = min(emerging_yr)
+  ) %>% 
+  group_by(taxon_emer,tresh,method) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  group_by(tresh,method) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "stock"
+  )
+
+# max(stock_d$com_sum)
+
+#  EEZ data
+eez_d <-  sens_data %>% 
+  # left_join(Neighbours) %>% 
+  filter(n_ens == 10) %>% 
+  # distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
+  group_by(sf_eez_name,tresh,method) %>% 
+  summarise(
+    taxon_emer = min(emerging_yr)
+  ) %>% 
+  group_by(taxon_emer,tresh,method) %>% 
+  summarise(
+    n_tax = n()
+  ) %>% 
+  group_by(tresh,method) %>% 
+  mutate(
+    com_sum = cumsum(n_tax),
+    level = "eez"
+  )
+
+# max(eez_d$com_sum)
+
+# Plot Data
+plot_data <- stock_d %>% 
+  bind_rows(eez_d) %>% 
+  filter(tresh != "NA") %>% 
+  mutate(
+    total = ifelse(level == "stock",n_stock,n_eez),
+    per = com_sum/total*100
+  ) %>% 
+  gather("metric","value",com_sum,per) %>% 
+  filter(tresh == "tresh_two")
+
+# 
+# # Manually include rest of data until 2100
+# cont_eez <- plot_data %>% 
+#   filter(level == "eez") %>% 
+#   arrange(taxon_emer) %>% 
+#   slice(1:24) %>%
+#   select(-1) %>% 
+#   arrange(metric) %>% 
+#   mutate(
+#     value = ifelse(metric == "per",81,228),
+#     taxon_emer = c(seq(2088,2099,1),seq(2088,2099,1))
+#   )
+
+plot_data <- plot_data %>% 
+  # bind_rows(cont_eez) %>% 
+  mutate(
+    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
+    label = ifelse(total == max(stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs")
+  )
+
+```
+
+```{r Time_lapse_plot, eval = F, echo = F}
+
+### Three plots
+
+count_eez_plot <- ggplot() +
+  geom_area(data = subset(plot_data, metric == "com_sum" & level == "eez"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  )  +
+  scale_y_continuous("Cumulative number of EEZs",
+                     expand = c(0,0)
+  ) +
+  scale_x_continuous("Year of range shift",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     label = c(seq(2000,2075,25),">2100")
+  ) +
+  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
+  scale_fill_viridis(discrete = T,
+                     begin = 0.1, 
+                     end = 0.1
+  ) +
+  facet_wrap(~ method);count_eez_plot
+
+#### Count stock plot 
+count_stock_plot <- ggplot() +
+  geom_area(data = subset(plot_data, metric == "com_sum" & level == "stock"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  )  +
+  scale_y_continuous("Cumulative number of stocks",
+                     expand = c(0,0)
+  ) +
+  scale_x_continuous("",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     label = c(seq(2000,2075,25),">2100")
+  ) +
+  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
+  scale_fill_viridis(discrete = T,
+                     begin = 0.7, 
+                     end = 0.7
+  ) +
+  facet_wrap(~ method);count_stock_plot
+
+# Percentage plot
+percentage_plot <- ggplot() +
+  geom_area(data = subset(plot_data, level == "eez" & my_facet == "Percentage (%)"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  ) + 
+  geom_area(data = subset(plot_data, level == "stock" & my_facet == "Percentage (%)"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = value,
+              fill = level
+            ),
+            size=0.5,
+            colour="black"
+  ) + 
+  scale_y_continuous("Cumulative proportion of the total (%)",
+                     expand = c(0,0)
+  ) +
+  scale_x_continuous("",
+                     breaks = seq(2000,2100,25),
+                     limits = c(2005,2100),
+                     label = c(seq(2000,2075,25),">2100")
+  ) +
+  my_ggtheme_p(leg_pos = "right", hjust = 0.5) +
+  scale_fill_viridis(discrete = T,
+                     begin = 0.1,
+                     end = 0.7,
+                     labels = c("EEZs","Stocks")
+  )+
+  facet_wrap(~ method);percentage_plot
+
+# Legend
+legend <- get_legend(percentage_plot)
+
+percentage_plot <- percentage_plot + 
+  theme(legend.position = "")
+
+# Draw figure
+plot_all <- ggdraw() +
+  # Count stock plot
+  draw_plot(count_stock_plot, x = 0.01, y = 0, width = 0.29, height = 0.90) +
+  # Count EEZ plot
+  draw_plot(count_eez_plot, x = 0.31, y = 0, width = 0.29, height = 0.90) +
+  # Percentage plot
+  draw_plot(percentage_plot, x = 0.61, y = 0 , width = 0.29, height = 0.90) +
+  # legend
+  draw_plot(legend, x = 0.84, y = 0 , width = 0.20, height = 1) +
+  # Labels
+  draw_plot_label(label = c("A", "B","C"), 
+                  size = 20,
+                  x = c(0.01, 0.285,0.59), 
+                  y = c(1,1,1)
+  )
+
+# Save plot
+save_plot(
+  "./Figures/cumu_per_freq_sens.png",
+  plot_all,
+  base_height = 5,
+  base_width = 12
+)
+
+```
+
+#### Comparative map
+
+```{r}
+
+# Per EEZs
+Eez_d_hist <- sens_data %>% 
+  group_by(eezid,sau_fishing_entity,tresh,method) %>% 
+  summarise(
+    mean_yr = mean(emerging_yr,na.rm=T),
+    sd_yr = sd(emerging_yr,na.rm=T),
+    n_taxa = length(unique(taxon_key))
+  ) %>% 
+  gather("metric","value",mean_yr:n_taxa) %>% 
+  mutate(
+    reverse = abs(value-2100)  + 2000
+  )
+
+# Estimate the value (catch and USD) of emerging species
+Eme_value_data <- sens_data %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,tresh,taxon_key,method) %>% 
+  summarise(n()) %>% 
+  left_join(Sau_data,
+            by = c("fishing_entity_id","taxon_key")) %>% 
+  filter(!is.na(variable)) %>% 
+  group_by(sau_fishing_entity,fishing_entity_id,tresh,variable,method) %>% 
+  summarise(
+    sum_value = sum(value,na.rm=T),
+    n_taxa = length(unique(taxon_key)) 
+  )
+
+# Total value of each fishing nation
+
+Total_value <- Sau_data %>% 
+  group_by(fishing_entity_id,variable) %>% 
+  summarise(tot_value= sum(value,na.rm=T))
+
+Value_map_data <- Eme_value_data %>% 
+  left_join(Total_value,
+            by = c("fishing_entity_id","variable")) %>% 
+  mutate(participation = (sum_value/tot_value)*100+2000)
+
+# Per fishing entity
+
+fish_d_hist <- sens_data %>% 
+  group_by(sau_fishing_entity,tresh,method) %>% 
+  summarise(
+    value = length(unique(taxon_key))
+  ) %>% 
+  filter(!is.na(sau_fishing_entity))
+
+World_map_nemer <- World_map %>%
+  full_join(Value_map_data,
+            by = c("sau_fishing_entity","fishing_entity_id")
+  ) %>% 
+  filter(!is.na(participation))
+
+# These are regions without transboundary species
+grey_land <- World_map %>% 
+  filter(!fishing_entity_id %in% Value_map_data$fishing_entity_id) %>% 
+  filter(!sovereignt %in% c("Antarctica"))
+
+# -------------------------- #
+# Call plot
+# -------------------------- #
+tresh_list <- "tresh_two"
+
+name_save <- paste("./Figures/emer_map_sens_",tresh_list,".png",sep = "")
+
+# Value Map
+World_map_part <- World_map_nemer %>% 
+  filter(
+    tresh ==  tresh_list,
+    variable == "value"
+  )
+
+# ToE Map 
+toe_map_data <- World_eez_sf %>% 
+  left_join(Eez_d_hist,
+            by = "eezid") %>% 
+  filter(tresh == tresh_list,
+         metric == "mean_yr")
+
+# Map with countries without ToE
+
+russia_churchil <- World_eez_sf %>% 
+  filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))
+
+no_toe_eez <- World_eez_sf %>% 
+  filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+         !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
+  rbind(russia_churchil)
+
+no_toe_land <- World_map %>% 
+  filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+         !name %in% World_map_part$name) # Filter EEZs with ToE stocks
+
+map <- ggplot(toe_map_data) +
+  geom_sf(
+    aes(
+      fill = reverse
+    ),
+    color =  "white",
+    size = 0.1
+  ) +
+  geom_sf(data = World_map_part,
+          aes(fill = participation),
+          color =  "white",
+          size = 0.1
+  ) +
+  geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
+  geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+  geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+  scale_fill_viridis("",
+                     limits = c(2000,2100),
+                     breaks = seq(2000,2100,25),
+                     labels = c(">2100",seq(2075,2000,-25)) # not inluding the > because we
+  ) +
+  my_ggtheme_m(leg_pos = "bottom") +
+  facet_wrap(~method,
+             nrow = 2)
+
+ggsave(name_save,
+       width = 10,
+       height = 6,
+       units = "in"
+)
+
+```
+
+# END PAPER
+
+# Troubleshooting
+
+### Weird connections from Australia
+
+```{r fig_map_Juan, eval = F, echo =F}
+
+##_----------------------- #
+# Checking on China and Indonesia
+# Probably due to dispute territories
+
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
+  clean_names()
+
+# 
+# Neighbours %>% 
+#   filter(eez_name == "China") %>% 
+#   pull(eez_neighbour) %>% 
+#   unique()
+
+##_----------------------- #
+
+# Get the center poligon of each EEZ for map source/target
+coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
+  ungroup() %>% 
+  select(eez_name,geometry) %>% 
+  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
+  # Manually fix issues
+  mutate(
+    longitude = ifelse(
+      eez_name == "Fiji", 178.065033,
+      ifelse(eez_name == "Tuvalu",177.6493,
+             ifelse( eez_name == "New Zealand",174.8860,
+                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
+                     )
+             )
+      )
+    )
+  )
+
+# Remove "c()"
+coords$longitude <- gsub("\\(","",coords$longitude)
+coords$longitude <- gsub("c","",coords$longitude)
+coords$latitude <- gsub(")","",coords$latitude)
+
+coords <- coords %>% 
+  mutate_at(vars(longitude,latitude),as.numeric)
+
+# Prepare data from results
+Test_Data <- results %>% 
+  # get coords for sources
+  left_join(coords,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  left_join(coords,
+            by = "eez_name") %>% 
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  mutate(
+    per_mid = ifelse(is.na(per_mid),0,per_mid)
+  ) %>% 
+  drop_na() %>% 
+  filter(
+    # per_mid >= 20,
+    variable == "mean_spp",
+  )
+
+# Convert to data to spatial
+network_data_sf <- Test_Data %>%
+  select(start_lon, start_lat, end_lon, end_lat) %>% 
+  purrr::transpose() %>% 
+  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
+  purrr::map(st_linestring) %>%
+  st_sfc(crs = 4326) %>%
+  st_sf(geometry = .)%>%
+  bind_cols(Test_Data) %>%
+  select(everything(), geometry) %>% 
+  # Set circles ('cus the world is not flat!)
+  st_segmentize(units::set_units(100, km)) %>% 
+  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
+  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
+  sf::st_sf(crs = 4326)
+
+#  unpack the coordinates of each feature 
+coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
+
+c(network_data_sf$line_id, network_data_sf$coords) %<-% (
+  coord_data %>% 
+    rename(subline_id = L1) %>%   
+    nest(-L2)
+)
+
+# Create data path for line connections
+paths <- network_data_sf %>% 
+  sf::st_set_geometry(NULL) %>% 
+  unnest() %>% 
+  mutate_at(vars(X,Y), round,2)
+
+# Plot it!
+
+country <- c("Fiji","Solomon Isl.","Tuvalu")
+
+ggplot() +
+  geom_sf(
+    # data = subset(world_map,name %in% country),
+    data = world_map,
+    size = 0.1, fill = "gray90", col = "gray90") +
+  geom_sf(
+    # data = subset(World_eez_sf, eez_name %in% country),
+    data = World_eez_sf,
+    size = 0.1, fill = "white", col = "gray70") +
+  geom_point(data = Test_Data,
+             # data = subset(Test_Data, source %in% country & target %in% source),
+             aes(start_lon,
+                 start_lat,
+                 col = per_mid),
+             size = 0.1,
+             col = "black",
+             # shape = 21,
+             stroke = 1.5) +
+  geom_path(data = paths,
+            # data = subset(paths, source %in% country),
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = per_mid
+            ),
+            size = 0.3,
+            arrow = arrow(length = unit(0.009, "npc")),
+            show.legend = NA) +
+  # geom_sf_text(data = subset(World_eez_sf, eez_name %in% country),
+  #              aes(label =eez_name),
+  #              size = 2) +
+  # geom_text(data = subset(Test_Data, source %in% country),
+  #              aes(start_lon,
+  #                start_lat,
+  #                label = paste(source,"to",target)
+  #              ),
+  #              size = 2) +
+  # geom_text_repel(data = Test_Data,
+  #   aes(start_lon,
+# start_lat,
+# label = paste(source,"to",target)
+# )
+# )+
+ggtheme_map() +
+  scale_colour_gradientn("Gain of Stock-Share Ratio (%)",
+                         colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
+                         limits = c(0,25),
+                         breaks = seq(0,25,5)
+  )
+
+```
+
+### Arctic and Brasil missing species
+
+```{r arctic, eval = F, echo =F}
+
+# Old Code
+
+###### Table.1 Multy linear model
+
+Arctic <- c("Canada (Arctic)","USA (Alaska, Arctic)", "Russia (Laptev to Chukchi Sea)")
+
+can_us_arctic <- Transboundary_spp %>% 
+  filter(eez_name == "Canada (Arctic)",
+         eez_neighbour == "USA (Alaska, Arctic)")
+
+arctic_spp <- Transboundary_spp %>% 
+  filter(eez_name %in% Arctic,
+         eez_neighbour %in% Arctic)
+
+arctic_spp <- Spp_depth %>% 
+  filter(taxon_key %in% can_us_arctic$taxon_key)
+
+# Transbounudary species that will see no change
+# 600252 Mallotus villosus  (Capelin)
+# 600315 Eleginus gracilis (Safforn cod)
+# 601520 Clupea pallasii  (Pacific herring)
+
+Eme_Res %>% 
+  filter(taxon_key %in% can_us_arctic$taxon_key) %>% 
+  # filter(sau_eez_name %in% arctic_spp$eez_name) %>% 
+  View(.)
+
+```
+
+```{r Brasa, eval = F, echo =F}
+
+brasil <- Transboundary_spp %>% 
+  filter(eez_name == "Brazil")
+
+brasil_spp <- Spp_depth %>% 
+  filter(taxon_key %in% brasil$taxon_key)
+
+# Transbounudary species that will see no change
+# 601590  Brevoortia pectinata  
+# 690022  Pleoticus muelleri 
+# 690023  Illex argentinus  
+# 690338  Penaeus paulensis
+
+Eme_Res %>% 
+  filter(taxon_key %in% brasil_spp$taxon_key) %>% 
+  View(.)
+
+```
+
+```{r TransIndex, eval = F, echo = F}
+
+TransIndex <- function(Spp,year, data_type = "Catch",ensemble_list,path,Neighbours,eez_centroid){
+
+  # ----------------- #
+  # Get Species Centrodis
+  #  ---------------- #
+
+  # Get model data from spps
+  Spp_Dist <- bind_rows(
+    lapply(
+      ensemble_list,
+      dbem_import,
+      taxon_key = Spp,
+      year = year,
+      data_type = data_type,
+      path = path
+    )
+  )
+
+  #### Control for Spp id so I can fix problems
+
+  if(file.exists("~/Desktop/spp_df.csv") == FALSE){
+    spp_df <- tibble() %>% 
+      write_csv(.,
+                "~/Desktop/spp_df.csv")
+  }
+
+  suppressMessages(
+    spp_df <- read_csv("~/Desktop/spp_df.csv")
+  )
+  spp_df <- tibble(Spp) %>% 
+    bind_rows(spp_df) %>% 
+    write_csv(.,
+              "~/Desktop/spp_df.csv")
+
+  #____________ Selecting only the transboundary nature of the species _________ #
+  Trans_Spp <- Spp_Dist %>%
+    left_join(index_code,
+              by= "index") %>% 
+    filter(!is.na(value)) %>% 
+    semi_join(Transboundary_spp, # filter only transboundary cases
+              by = c("taxon_key","eez_name")
+    )
+
+  # Step
+  nr <- nrow(Trans_Spp)
+  if(nr > 0){
+
+    # Filter eez by those transboundary
+    unique_eez <- unique(Trans_Spp$eez_name)
+
+    Neighbours_combo <- Neighbours %>% 
+      filter(eez_name %in% unique_eez,
+             eez_neighbour %in% unique_eez)
+
+    # Get the centroid of each country
+    EEZ_centroid <- eez_centroid %>% 
+      filter(eez_name %in% Trans_Spp$eez_name) %>% 
+      mutate(taxon_key = Spp)
+
+    Ids <- unique(Neighbours_combo$neighbour_id)
+
+    corefx <- function(Ids){
+
+      EEZ_ids <- Neighbours_combo %>% 
+        filter(neighbour_id == Ids)
+
+      Sub_Trans_Spp <- Trans_Spp %>% 
+        filter(eez_name %in% EEZ_ids$eez_name)
+
+      Sub_EEZ_centroid <- EEZ_centroid %>% 
+        filter(eez_name %in% EEZ_ids$eez_name)
+
+      # Centroid is defined as the tile with the higest species value
+      Centroids <- Sub_Trans_Spp %>% 
+        left_join(EEZ_ids, by = "eez_name") %>% 
+        group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
+        top_n(1,value) %>% 
+        distinct(value,year, .keep_all = TRUE) %>%  # removes duplicates when tile is in between EEZs
+        left_join(Sub_EEZ_centroid, by = "taxon_key") %>%
+        select(eez_name =eez_name.y, everything() ,-eez_name.x)
+
+      # Estimate distances between spp centroid and EEZ centroid
+      Distance_Data <- Centroids %>%
+        rename(
+          catch_lon = lon,
+          catch_lat = lat,
+          eez_lon = longitude,
+          eez_lat= latitude
+        ) %>% 
+        # Estimate the distance between centriods
+        mutate(
+          distance = geodist(eez_lon, eez_lat, catch_lon, catch_lat, units="km")
+        ) %>% 
+        select(taxon_key,eez_name,neighbour_id,ensemble,year,distance)
+
+      # Historical mean and sd of distance
+      historical <- Distance_Data %>%
+        filter(year <= 2000) %>%
+        ungroup() %>%
+        group_by(taxon_key,ensemble,eez_name,neighbour_id) %>%
+        summarise_at(vars("distance"),
+                     c(hd_temp_mean=mean,
+                       hd_temp_sd=sd)
+        )
+
+      # Future results
+      future <- Distance_Data %>% 
+        filter(year > 2000)
+
+      # Transboundary Index
+      Trans_index <- Distance_Data %>% 
+        left_join(historical,
+                  by = c("taxon_key","eez_name","ensemble","neighbour_id")
+        ) %>%
+        # Re arrenging table for mutate
+        gather("variable","value",distance:hd_temp_sd) %>%
+        ungroup()
+
+      Partial <- Trans_index %>% 
+        mutate(
+          names = ifelse(eez_name ==Sub_EEZ_centroid$eez_name[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_"))
+        ) %>%
+        select(-eez_name,-variable) %>%
+        spread(names,value) %>% 
+        # mutate index
+        mutate(
+          trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
+        ) %>% 
+        group_by(taxon_key,ensemble,neighbour_id) %>% 
+        mutate(RMean = rollmean(x = trans_index, 
+                                10, 
+                                align = "right", 
+                                fill = trans_index)
+        ) %>% 
+        filter(year > 1960) %>% 
+        select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)
+
+      # Average of 10 years mean of ensemble members results
+      mean_sd_index <- Partial %>% 
+        group_by(taxon_key,year,neighbour_id) %>% 
+        summarise_at(vars("RMean"),
+                     c(trend=mean,
+                       ensemble_sd=sd)
+        )
+
+      # Estimate noice from ensemble members
+      historic_sd <- mean_sd_index %>% 
+        filter(year < 2000) %>% 
+        group_by(taxon_key) %>% 
+        summarise_at(vars("trend"),
+                     c(hist_mean=mean,
+                       noise=sd)
+        )
+
+      # Plot the result, bruh
+      # ggplot() +
+      #   geom_line(data = mean_sd_index,
+      #             aes(
+      #               x = year,
+      #               y = trend
+      #             )
+      #   ) +
+      #   geom_ribbon(data = mean_sd_index,
+      #               aes(x=year,
+      #                   ymin=historic_sd$hist_mean-historic_sd$noise,
+      #                   ymax=historic_sd$hist_mean+historic_sd$noise),
+      #               alpha = 0.5,
+      #               fill = "grey50"
+      #   ) +
+      #   geom_line(data = mean_sd_index,
+      #             aes(
+      #               x = year,
+      #               y = historic_sd$hist_mean
+      #             ),
+      #             colour = "grey50"
+      #   ) +
+      #   geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
+      #              aes(
+      #                x = year,
+      #                y = trend,
+      #                colour = "red",
+      #              ),
+      #              shape = 1,
+      #              size = 3
+      #   ) +
+      #   geom_vline(xintercept = 2000,colour="blue",alpha = 0.5) +
+      #   theme_classic()
+
+      # #     
+      Final_Result <- mean_sd_index %>%
+        filter(year > 2000,
+               trend > (historic_sd$hist_mean+historic_sd$noise) |
+                 trend < (historic_sd$hist_mean-historic_sd$noise)
+        ) %>%
+        group_by(taxon_key,neighbour_id) %>%
+        summarise(emerging_yr= min(year)) %>%
+        left_join(EEZ_ids) %>%
+        ungroup() %>%
+        select(taxon_key,eez_name,eez_neighbour,emerging_yr)
+
+      # if(i == 1){
+      #   f_df <- Final_Result
+      # }else{
+      #   f_df <- bind_rows(f_df,Final_Result)
+      # }
+      return(Final_Result)
+    }
+
+    x <- bind_rows(
+      mclapply(Ids,corefx)
+    )
+    # return(f_df)
+
+  }else{
+
+    print("no share data")
+
+    x <- tibble(
+      taxon_key = Spp,
+      eez_name = NA,
+      emerging_yr = NA
+    )
+
+  }
+
+  return(x)
+
+}
+
+#  #One spp works 
+# suppressMessages(
+# TransIndex(Spp = 600006,
+#              year,
+#              data_type = "Catch",
+#              ensemble_list,
+#              path,
+#              Neighbours = Neighbours,
+#              eez_centroid)
+# Multiple spp
+
+# spp_list <- c(600004,
+#               600005,
+#               600245,
+#               600006
+#               )
+# 
+# suppressMessages(
+#   suppressWarnings(
+#     Test <- bind_rows(
+#       lapply(spp_list,
+#              TransIndex,
+#              year,
+#              data_type = "Catch",
+#              ensemble_list,
+#              path,
+#              Neighbours,
+#              eez_centroid
+#       )
+#     )
+#   )
+# )
+
+```
+
+###### With relation to neighbours
+
+```{r time_emergence_map,  eval = F, echo = T}
+
+##_----------------------- #
+# Checking on China and Indonesia
+# Probably due to dispute territories
+
+Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
+  clean_names()
+
+# 
+# Neighbours %>% 
+#   filter(eez_name == "China") %>% 
+#   pull(eez_neighbour) %>% 
+#   unique()
+
+##_----------------------- #
+
+# Get the center poligon of each EEZ for map source/target
+coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
+  ungroup() %>% 
+  select(eez_name,geometry) %>% 
+  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
+  # Manually fix issues
+  mutate(
+    longitude = ifelse(
+      eez_name == "Fiji", 178.065033,
+      ifelse(eez_name == "Tuvalu",177.6493,
+             ifelse( eez_name == "New Zealand",174.8860,
+                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
+                     )
+             )
+      )
+    )
+  )
+
+# Remove "c()"
+coords$longitude <- gsub("\\(","",coords$longitude)
+coords$longitude <- gsub("c","",coords$longitude)
+coords$latitude <- gsub(")","",coords$latitude)
+
+coords <- coords %>% 
+  mutate_at(vars(longitude,latitude),as.numeric)
+
+# Prepare data from results
+Test_Data <- Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # emerging_yr >= 2020
+  ) %>% 
+  group_by(eez_name,eez_neighbour,emerging_yr) %>% 
+  summarise(
+    n_tax = length(unique(taxon_key))
+  ) %>% 
+  filter(
+    emerging_yr <= 2050
+  ) %>%
+  group_by(eez_name,eez_neighbour) %>% 
+  summarise(
+    n_tax_emer = sum(n_tax)
+  ) %>% 
+  # get coords for sources
+  left_join(coords,
+            by = "eez_name") %>%
+  rename(source = eez_name,
+         start_lon = longitude,
+         start_lat = latitude,
+         eez_name=eez_neighbour) %>% 
+  # get coords for targets
+  left_join(coords,
+            by = "eez_name") %>% 
+  rename(target = eez_name,
+         end_lon = longitude,
+         end_lat = latitude) %>% 
+  ungroup() %>% 
+  # Plot miscs
+  # mutate(
+  #   per_mid = ifelse(is.na(per_mid),0,per_mid)
+  # ) %>% 
+  drop_na() #%>% 
+# filter(
+# per_mid >= 20,
+# variable == "mean_spp",
+# )
+
+# Convert to data to spatial
+network_data_sf <- Test_Data %>%
+  select(start_lon, start_lat, end_lon, end_lat) %>% 
+  purrr::transpose() %>% 
+  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
+  purrr::map(st_linestring) %>%
+  st_sfc(crs = 4326) %>%
+  st_sf(geometry = .)%>%
+  bind_cols(Test_Data) %>%
+  select(everything(), geometry) %>% 
+  # Set circles ('cus the world is not flat!)
+  st_segmentize(units::set_units(100, km)) %>% 
+  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
+  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
+  sf::st_sf(crs = 4326)
+
+#  unpack the coordinates of each feature 
+coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
+
+c(network_data_sf$line_id, network_data_sf$coords) %<-% (
+  coord_data %>% 
+    rename(subline_id = L1) %>%   
+    nest(-L2)
+)
+
+# Create data path for line connections
+paths <- network_data_sf %>% 
+  sf::st_set_geometry(NULL) %>% 
+  unnest() %>% 
+  mutate_at(vars(X,Y), round,2)
+
+# Plot it!
+
+country <- c("Fiji","Solomon Isl.","Tuvalu")
+
+ggplot() +
+  geom_sf(
+    # data = subset(world_map,name %in% country),
+    data = World_map,
+    size = 0.1, fill = "gray90", col = "gray90") +
+  geom_sf(
+    # data = subset(World_eez_sf, eez_name %in% country),
+    data = World_eez_sf,
+    size = 0.1, fill = "white", col = "gray70") +
+  geom_point(data = Test_Data,
+             # data = subset(Test_Data, source %in% country & target %in% source),
+             aes(start_lon,
+                 start_lat,
+                 col = n_tax_emer
+             ),
+             size = 0.1,
+             col = "grey50",
+             # shape = 21,
+             stroke = 1.5) +
+  geom_path(data = paths,
+            # data = subset(paths, source %in% country),
+            aes(X, Y,
+                group = interaction(line_id,subline_id),
+                col = n_tax_emer
+            ),
+            size = 0.3,
+            # arrow = arrow(length = unit(0.009, "npc")),
+            show.legend = NA) +
+  ggtheme_map() +
+  scale_colour_viridis(
+    name="Number of emerging stocks\nby 2050",
+    limits = c(0,40),
+    breaks = seq(0,40,10),
+    # option="magma",
+    direction = 1
+  ) +
+  ggsave("./Figures/figure_emerging_2050.png",
+         width = 10,
+         height = 6,
+         units = "in"
+  )
+
+```
+
+### Non selected Results
+
+```{r, olda_acumulative, eval = F}
+
+# filter(emerging_yr >= 2020) %>% 
+ggplot() +
+  geom_area(data = subset(plot_data, level == "stock"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = com_sum,
+              fill = level
+            ),
+            alpha= 0.8,
+            size=0.5,
+            colour="black"
+  )  +
+  geom_area(data = subset(plot_data, level == "eez"),
+            aes(
+              x = taxon_emer,
+              y = com_sum,
+              fill = level
+            ),
+            alpha= 0.8,
+            size=0.5,
+            colour="black"
+  ) +
+  facet_wrap(~tresh) + 
+  # Comment out for option A (Change plot name!!)
+  #   geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
+  # geom_text(aes(label="Total number of EEZs", x = 2010, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
+  #   geom_hline(yintercept = length(unique(Eme_Res$taxon_key)), color = "grey",linetype = "dashed") +
+  # geom_text(aes(label="Total number of Taxa", x = 2010, y = length(unique(Eme_Res$taxon_key))+10), color = "grey70",size =3) +
+  scale_y_continuous("Cumulatuve Curve",
+                     breaks = seq(0,600,50),
+                     limits = c(0,600)
+  ) +
+  scale_x_continuous("Year of emergence",
+                     breaks = seq(2005,2100,10),
+                     limits = c(2005,2100)
+  ) +
+  my_ggtheme_p() +
+  scale_fill_viridis(discrete = T) +
+  ggsave("./Figures/figure_accumulative_curve_ab.png",
+         width = 6,
+         height = 6,
+         units = "in" 
+  )
+
+### Percentage accumulation
+
+ggplot() +
+  geom_area(data = subset(plot_data, level == "stock"),
+            aes(
+              x = as.numeric(taxon_emer),
+              y = per,
+              fill = level
+            ),
+            alpha= 0.8,
+            size=0.5,
+            colour="black"
+  )  +
+  geom_area(data = subset(plot_data, level == "eez"),
+            aes(
+              x = taxon_emer,
+              y = per,
+              fill = level
+            ),
+            alpha= 0.8,
+            size=0.5,
+            colour="black"
+  ) +
+  scale_y_continuous("Cumulatuve Curve",
+                     breaks = seq(0,100,25),
+                     limits = c(0,100)
+  ) +
+  scale_x_continuous("Year of emergence",
+                     breaks = seq(2005,2100,15),
+                     limits = c(2005,2100)
+  ) +
+  my_ggtheme_p() +
+  scale_fill_viridis(discrete = T) +
+  # facet_wrap(~tresh) + 
+  ggsave("./Figures/Fig4.png",
+         width = 8,
+         height = 8,
+         units = "in"
+  )
+
+```
+
+```{r histo,  eval = F, echo = T}
+
+# First time a Species becomes emergent
+Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # tresh == "tresh_one"
+  ) %>% 
+  group_by(tresh,taxon_key,emerging_yr) %>% 
+  summarise(
+    n_trans = length(unique(sf_eez_name))
+  ) %>% 
+  group_by(tresh,taxon_key) %>% 
+  summarise(
+    ntrans = min(emerging_yr)
+  ) %>% 
+  group_by(tresh,ntrans) %>% 
+  summarise(
+    n_taxa = length(unique(taxon_key))
+  ) %>% 
+  arrange(-n_taxa) %>% 
+  # View()
+  # group_by(tresh) %>% 
+  # summarise(sum(n_taxa))
+  ggplot() +
+  geom_line(
+    aes(
+      x = ntrans,
+      y = n_taxa,
+      color = tresh
+    )
+  )
+
+# EEZ, first time an EEZ has a emergent stock
+Eme_Res %>% 
+  filter(
+    n_ens == 10#,
+    # tresh == "tresh_one"
+  ) %>% 
+  group_by(tresh,sf_eez_name,emerging_yr) %>% 
+  summarise(
+    n_trans = length(unique(taxon_key))
+  ) %>% 
+  group_by(tresh,sf_eez_name) %>% 
+  summarise(
+    ntrans = mean(emerging_yr)
+  ) %>% 
+  group_by(tresh,ntrans) %>% 
+  summarise(
+    n_taxa = length(unique(sf_eez_name))
+  ) %>% 
+  arrange(-n_taxa) %>% 
+  View()
+# group_by(tresh) %>%
+# summarise(sum(n_taxa))
+# ggplot() +
+# geom_line(
+#   aes(
+#     x = ntrans,
+#     y = n_taxa,
+#     color = tresh
+#   )
+# )
+
+# Numer of EEZs where stocks will emerge
+stock_d_hist <- Eme_Res %>% 
+  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
+  group_by(emerging_yr,taxon_key,tresh) %>%
+  summarise(n_tax = length(unique(sf_eez_name))) %>%
+  # summarise(n()) %>% 
+  mutate(
+    level = "EEZs per Spp"
+  )  %>% 
+  ungroup() %>% 
+  select(-taxon_key)
+
+# Per EEZs
+eez_d_hist <- Eme_Res %>% 
+  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
+  group_by(emerging_yr,sf_eez_name,tresh,neighbour_id) %>%  
+  # summarise(n_tax = length(unique(taxon_key))) %>% 
+  summarise(n()) %>% 
+  mutate(
+    level = "Spp per EEZ"
+  ) %>% 
+  ungroup() %>% 
+  select(-sf_eez_name)
+
+stock_d_hist %>% 
+  bind_rows(eez_d_hist) %>% 
+  filter(tresh != "NA") %>% 
+  ggplot() +
+  geom_density(
+    aes(
+      x = emerging_yr,
+      fill = level
+    ),
+    stat = "count",
+    alpha= 0.8,
+    size=0.5,
+  ) +
+  scale_y_continuous("Frequency",
+                     # breaks = seq(0,300,50),
+                     # limits = c(0,300)
+  ) +
+  scale_x_continuous("Year of emergence",
+                     # breaks = seq(2006,2100,25),
+                     # limits = c(2000,2100)
+  ) +
+  my_ggtheme_p() +
+  scale_fill_viridis(discrete = T) +
+  facet_wrap(~tresh) +
+  ggsave("./Figures/hist_freq.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
+```
+
+##### Temperature gradient
+
+##### Fun `GetEnvData`
+
+```{r get_env_data, eval = F, echo = T}
+
+getEnvVar=function(yr,ModelRCP,RCP,Variable){
+
+  if(ModelRCP == "IPSL"){
+    ModelRCP = "IPSL-CM5A-MR"
+    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
+  }
+
+  if(ModelRCP == "MPI"){
+    ModelRCP = "MPI-ESM-MR"
+    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
+  }
+
+  if(ModelRCP == "GFDL"){
+
+    if(RCP == "rcp26"){
+      ModelRCP <- "GFDL26"
+    }else{
+      ModelRCP <- "GFDL85"
+    }
+    Filepath <- paste("Z:/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")
+  }
+
+  if(Variable == "Ice"){
+
+    Bottom <- paste("IceExt_",yr,".txt",sep="")
+
+  }
+
+  if(Variable == "Temperature"){
+
+    Bottom = paste("SST_",yr,".txt",sep="")
+
+  }
+
+  if(Variable == "Salinity"){
+    # Bottom
+    Bottom = paste("Salinity_btm_",yr,".txt",sep="")
+
+  }
+
+  Read_me <- paste(Filepath,Bottom,sep="")[1]
+  Bottom_Data <- read.csv(Read_me,header=F)
+
+  # Retrun
+
+  getTemp <- Bottom_Data 
+}
+
+Years <-seq(1951,2100,1)
+
+for(y in 1:length(Years)){
+  Data <- getEnvVar(yr = Years[y],
+                    Model = "GFDL",
+                    Variable = "Temperature",
+                    RCP = "rcp85"
+  )
+
+  if(y == 1){
+
+    F_BData <- Data[1]
+    colnames(F_BData) <- Years[1]
+
+  }else{
+
+    F_BData <- cbind(F_BData,Data[1])
+    colnames(F_BData)[y] <- Years[y]
+  }
+}
+
+F_BData <- F_BData %>% 
+  rowid_to_column("index") %>% 
+  gather("year","temp",-index) %>% 
+  mutate(temp = ifelse(temp == -9999,NA,temp)) %>% 
+  group_by(year) %>% 
+  summarise(
+    mean_temp = mean(temp,na.rm=T)
+  ) %>% 
+  write_csv(.,
+            "SST_data.csv")
+
+```
+
+```{r temp_plot, eval = F, echo = T}
+
+### NEED TO RUN TIME LAPSE CHUNK FIRST TO GET ACCUMULATIVE DATA
+
+# Get sst_data
+sst_data <- my_path("D",T,"Climate","SST_data.csv",read = T)
+
+sst_hist <- sst_data %>% 
+  filter(year < 2000) %>% 
+  group_by() %>% 
+  summarise(hist_sst =mean(mean_temp)) %>% 
+  pull()
+
+sst_change <- sst_data %>%
+  filter(year > 2020) %>% 
+  mutate(chng = round(mean_temp-sst_hist,2))
+
+# Plot Data
+stock_d %>% 
+  bind_rows(eez_d) %>% 
+  filter(taxon_emer >= 2020) %>% 
+  rename(year = taxon_emer) %>% 
+  left_join(sst_change) %>% 
+  ggplot() +
+  geom_line(
+    aes(
+      x = chng,
+      y = com_sum,
+      colour = level
+    )
+  ) +
+  geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
+  geom_text(aes(label="Total number of EEZs", x = 0.5, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
+  scale_y_continuous("Cumulatuve Curve",
+                     breaks = seq(0,450,50),
+                     limits = c(0,500)
+  ) +
+  scale_x_continuous("Change in SST (C)",
+                     breaks = seq(0,2,0.2),
+                     limits = c(0.3,2.3)
+  ) +
+  my_ggtheme_p() +
+  scale_colour_viridis(discrete = T) +
+  ggsave("./Figures/figure_temp.png",
+         width = 6,
+         height = 6,
+         units = "in"
+  )
+
+```
+
+```{r Scatter_plot,  eval = F, echo = T}
+
+data_stock <- Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr,eez_name) %>% 
+  summarise(
+    tax_n = length(unique(taxon_key))
+  ) %>% 
+  group_by(emerging_yr) %>% 
+  summarise(emerging_tax = sum(tax_n))
+
+data_eez <- Eme_Res %>% 
+  filter(
+    n_ens == 10,
+    emerging_yr >= 2020
+  ) %>% 
+  group_by(emerging_yr,taxon_key) %>% 
+  summarise(
+    eez_n = length(unique(eez_name))
+  ) %>% 
+  group_by(emerging_yr) %>% 
+  summarise(emerging_eez = sum(eez_n))
+
+data_stock %>% 
+  left_join(data_eez,
+            by = "emerging_yr") %>% 
+  rename(year = emerging_yr) %>% 
+  left_join(sst_data) %>% 
+  ggplot() +
+  geom_point(
+    aes(
+      x = mean_temp,
+      y = emerging_eez,
+      fill = year,
+      colour = year
+    )
+  )
+
+```
+
+# Old code
+
+## Fig. 4. Bubble plot
+
+```{r, buble_plot, eval = F, echo = F}
+
+# Needs previous chunk
+
+Buble_data <- Eez_d_hist %>% 
+  filter(metric == "mean_yr") %>% 
+  group_by(sau_fishing_entity,tresh) %>% 
+  summarise(mean_yr = mean(value)) %>% 
+  left_join(Eme_value_data,
+            by = c("sau_fishing_entity","tresh")) %>% 
+  filter(#tresh == selected_tresh, # tresh hold selection
+    variable == "value") %>% 
+  mutate(
+    weight_toe = sum_value/mean_yr,
+    sum_value_bi = sum_value/1000000000,
+    log_value = log10(sum_value_bi)
+  ) %>%   # Billion usd
+  left_join(Matching_names,
+            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
+  distinct(sau_fishing_entity,tresh,.keep_all = TRUE) %>%
+  filter(!is.na(un_geo_region)) %>% 
+  ungroup()# %>%  # no more asterisk
+# mutate(
+#   sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity))
+
+### Function for estimating percentile
+
+quibble <- function(x, q = p) {
+  tibble(x = quantile(x, q), q = q)
+}
+
+Percentile_value <- Buble_data %>% 
+  group_by(tresh) %>%  # for whole distribution
+  do(quibble(.$log_value, 0.75))
+
+## Top fishing nations close to SDGs
+Buble_data %>% 
+  filter(log_value > Percentile_value$x & mean_yr < 2036) %>% 
+  View()
+group_by(tresh) %>% 
+  summarise(sum(sum_value_bi),
+            sum(n_taxa))
+
+## All fishing nations close to SDGs
+Buble_data %>% 
+  filter(mean_yr < 2036) %>% 
+  # nrow() # 76  countries
+  View()
+group_by(tresh) %>% 
+  summarise(sum(sum_value_bi),
+            sum(n_taxa))
+
+76/length(unique(Buble_data$sau_fishing_entity))
+134/length(unique(Buble_data$sau_fishing_entity))
+
+# Labels Data
+
+Labels_data <- tibble(
+  x = c(2029,
+        # 2050,2070,
+        2007),
+  y = c(-5.2,
+        # -5.8,-5.8,
+        Percentile_value$x+.1),
+  label = c("UN 2030 Agenda", 
+            # "Mid shift","Late shift",
+            "Fisheries > 75th percentile")
+)
+
+# Plot the bubbles!
+ggplot(Buble_data) +
+  geom_point(
+    aes(
+      x = mean_yr,
+      y = log_value,
+      size = n_taxa,
+      colour = un_geo_region,
+      shape = un_geo_region
+    ),
+    alpha = 0.7
+  ) +
+  # Right side labels
+  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034),
+                   aes(
+                     x = mean_yr,
+                     y = log_value,
+                     label = sau_fishing_entity
+                   ),
+                   arrow = arrow(length = unit(0.009, "npc"),
+                                 type = "closed",
+                                 ends = "first"),
+                   force = 1,
+                   box.padding = 0.2,
+                   fill = "NA",
+                   nudge_x = 2066 - subset(Buble_data,log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034)$mean_yr, # set distance from point
+                   nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
+                   segment.size  = 0.1,
+                   segment.color = "grey50",
+                   direction     = "y",
+                   hjust         = 1
+  ) +
+  # LEFT side labels
+  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034),
+                   aes(
+                     x = mean_yr,
+                     y = log_value,
+                     label = sau_fishing_entity
+                   ),
+                   arrow = arrow(length = unit(0.009, "npc"),
+                                 type = "closed",
+                                 ends = "first"),
+                   force = 1,
+                   box.padding = 0.2,
+                   fill = "NA",
+                   nudge_x = 2025 - subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$mean_yr, # set horizontal position
+                   nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
+                   segment.size  = 0.1,
+                   segment.color = "grey50",
+                   direction     = "y",
+                   hjust         = 1
+  ) +
+  # Divisory lines and labels
+  geom_vline(xintercept = 2030, color = "grey50", size = 0.5, linetype = "dashed") +
+  geom_text(data = subset(Labels_data, x == 2029),
+            aes(
+              x = x,
+              y = y,
+              label = label
+            ),
+            size = 4,
+            color = "grey30",
+            angle = 90
+  ) +
+  geom_hline(data = Percentile_value,
+             aes(yintercept = x),
+             color = "grey50"
+  )  +
+  geom_text(data = subset(Labels_data, x == 2007),
+            aes(
+              x = x,
+              y = y,
+              label = label
+            ),
+            size = 4,
+            color = "grey30"
+  ) + 
+  scale_shape("Region") +
+  scale_color_viridis(name="Region",
+                      discrete = T) +
+  scale_x_continuous("Average year of range shift",
+                     limits = c(2000,2080),
+                     breaks = seq(2000,2080,10)
+  ) +
+  scale_y_continuous("Fishing revenue (log10)",
+                     limits = c(-6,2),
+                     breaks = seq(-6,2,2),
+                     expand = c(0,0)
+  ) +
+  scale_size("Number of\nshifting stocks") +
+  my_ggtheme_p(leg_pos = "right") +
+  ggsave("./Figures/bubble_plot.png",
+         width = 12,
+         height = 8,
+         units = "in"
+  )
+
+```
+
+```{r participation_text, eval = T, echo=F}
+
+Value_map_data %>% 
+  group_by(tresh,variable) %>% 
+  summarise(m = mean(participation)-2000,
+            sd = sd(participation-2000)
+  )
+
+```
+
+## Fig. 4N Top fishers on bars
+
+The bar option by UN sub region of Figure 4N of top fishers
+
+```{r}
+
+# Create data
+Species_Data <- top_toe %>% 
+  filter(tresh == selected_tresh,
+         variable == "value") %>% 
+  mutate(mean = mean-2000,
+         group = geo_region)
+
+# Set a number of 'empty bar' to add at the end of each group
+Char <- as.factor(Species_Data$group)
+empty_bar <- 5
+to_add <- data.frame( matrix(NA, empty_bar*nlevels(Char), ncol(Species_Data)) )
+colnames(to_add) <- colnames(Species_Data)
+to_add$group <- rep(levels(Char), each=empty_bar)
+Species_Data <- rbind(Species_Data, to_add)
+Species_Data <- Species_Data %>% arrange(group,mean)
+Species_Data$id <- seq(1, nrow(Species_Data))
+
+# Species_Data$id <- seq(1, nrow(Species_Data))
+
+# Get the name and the y position of each label
+label_data <- Species_Data 
+number_of_bar <- nrow(label_data)
+angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar
+label_data$hjust <- ifelse( angle < -90, 1, 0)
+label_data$angle <- ifelse(angle < -90, angle+180, angle)
+
+# prepare a data frame for base lines
+base_data <- Species_Data %>% 
+  group_by(group) %>% 
+  summarize(start=min(id), 
+            end=max(id) - empty_bar, 
+            median = median(mean, na.rm = T)
+  ) %>% 
+  rowwise() %>% 
+  mutate(title=mean(c(start, end)))
+
+# prepare a data frame for grid (scales)
+grid_data <- base_data
+grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
+grid_data$start <- grid_data$start - 1
+# grid_data <- grid_data[-1,]
+# grid_data[1,2] <- 5 # Manually include bathydemerssal
+# grid_data[1,3] <- 1
+
+un_regions <- grid_data %>% 
+  select(group)
+
+ggplot(Species_Data, 
+       aes(x=as.factor(id), 
+           y=mean,
+           fill=mean)
+) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
+  geom_bar(aes(x=as.factor(id),
+               y=mean,
+               fill=mean),
+           stat="identity",
+           alpha=0.5
+  ) +
+  geom_segment(data=grid_data,
+               aes(x = end, y = 30, xend = start, yend = 30),
+               colour = "grey90", alpha=1, size=0.3, inherit.aes = FALSE) +
+  geom_bar(aes(x=as.factor(id),
+               y=mean,
+               fill=mean
+  ),
+  stat="identity",
+  alpha=0.5) +
+  annotate("text", 
+           x = 0.5,
+           y = 37,
+           label = "UN Agenda",
+           color= "grey60", 
+           size=3, 
+           angle=-5
+  ) +
+  # Country names
+  geom_text(data=label_data,
+            aes(x=id, 
+                y= 100,
+                label=sau_fishing_entity,
+                hjust=hjust),
+            color="black",
+            fontface="italic",
+            alpha = 1,
+            size = 4 ,
+            angle = label_data$angle,
+            inherit.aes = FALSE) +
+  # Continents labels
+  # geom_text(data=base_data,
+  #           aes(x = title,
+  #               y = 85, 
+  #               label=group), 
+  #           # hjust=c(1,1,1,0,0,0),
+  #           colour = "black",
+  #           # angle = myAng,
+  #           size=5,
+  #           fontface="bold",
+  #           inherit.aes = FALSE) +
+ylim(-100,100) + # sets the circle the first value the central the last value how wide
+  theme_minimal() +
+  theme(
+    legend.position = c(0.5,0.5),
+    legend.text = element_text(size = 13),
+    legend.title = element_text(size = 14, face = "bold"),
+    legend.title.align = 0.5,
+    axis.text = element_blank(),
+    axis.title = element_blank(),
+    panel.grid = element_blank(),
+    panel.background = element_rect(colour = "white"),
+    legend.direction="horizontal",
+    legend.key.width = unit(3, 
+                            "line")
+  ) +
+  coord_polar(start = 0) +
+  scale_fill_viridis("Average year of range shift\nfor top 5 most valuable stocks",
+                     limits = c(0,100),
+                     breaks = seq(0,100,20),
+                     labels = seq(2000,2100,20),
+                     direction = -1
+  ) +
+  guides(
+    fill=guide_colorbar(title.position="top", 
+                        title.hjust =0.5)
+  )
+
+ggsave(
+  plot = last_plot(),
+  width = 13,
+  height = 13,
+  units = "in",
+  filename = "./Figures/top_stocks_circular_trend.png"
+)
+
+```
+
+## Figure 5 code from Juan
 
@@
 ```
-
-##### Fig. TOP 5
-
-```{r fig_map_Juan, eval = F, echo =F}
-
-top_stocks_per_country <- Eme_Res %>% 
-  rename(geo_region = un_sub_region) %>% 
-  group_by(sau_fishing_entity,fishing_entity_id,geo_region,tresh,taxon_key) %>% 
-  summarise(n()) %>% 
-  left_join(Sau_data,
-            by = c("fishing_entity_id","taxon_key")) %>% 
-  filter(!is.na(variable)) %>% 
-  # group_by(sau_fishing_entity,fishing_entity_id,tresh,variable) %>% 
-  # summarise(sum_value = sum(value,na.rm=T)) %>% 
-  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
-  # Select top 5 most valuable stocks
-  top_n(5,value)
-
-top_taxon_list <-  unique(top_stocks_per_country$taxon_key)
-# Read proportion results
-suppressMessages(
-  top_prop_results <- PropChange(top_taxon_list, links="y") %>%
-    filter(!is.na(eez_neighbour)) %>%
-    mutate(per_mid_plot =ifelse(per_mid >= 100, 100,
-                                ifelse(per_mid <= -100,-100,per_mid)
-    ),
-    per_early_plot =ifelse(per_early >= 100, 100,
-                           ifelse(per_early <= -100,-100,per_early)
-    )
-    )
-)
-
-### Median number of stocks changing in each time step
-
-top_prop_results %>% 
-  filter(variable == "n_per") %>% 
-  group_by() %>% 
-  summarise_at(vars(early,mid),funs(
-    median(.,na.rm=T),
-    sd(.,na.rm=T),
-    length(unique(eez_name)),
-    length(unique(eez_name))/280*100
-  )
-  )
-
-# Prepare data from results
-top_data <- top_prop_results %>% 
-  # get coords for sources
-  left_join(Eez_centroid,
-            by = "eez_name") %>%
-  rename(source = eez_name,
-         start_lon = longitude,
-         start_lat = latitude,
-         eez_name=eez_neighbour) %>% 
-  # get coords for targets
-  left_join(Eez_centroid,
-            by = "eez_name") %>% 
-  rename(target = eez_name,
-         end_lon = longitude,
-         end_lat = latitude) %>% 
-  ungroup() %>% 
-  # Plot miscs
-  mutate(
-    per_early = ifelse(is.na(per_early),0,per_early),
-    per_mid = ifelse(is.na(per_mid),0,per_mid)
-  ) %>% 
-  drop_na() %>%
-  # gather("variable","value",per_early:per_early_plot) %>% 
-  filter(
-    variable %in% c("mean_eez","n_spp")#,
-    # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
-  )
-
-View(top_data)
-
-# Convert to data to spatial
-top_network_data_sf <- top_data %>%
-  select(start_lon, start_lat, end_lon, end_lat) %>% 
-  purrr::transpose() %>% 
-  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
-  purrr::map(st_linestring) %>%
-  st_sfc(crs = 4326) %>%
-  # st_sfc(crs = "+proj=eck4") %>% # for tbular plot
-  st_sf(geometry = .)%>%
-  bind_cols(top_data) %>%
-  select(everything(), geometry) %>% 
-  # Set circles ('cus the world is not flat!)
-  st_segmentize(units::set_units(100, km)) %>% 
-  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>%
-  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>%
-  # sf::st_sf(crs = 4326)
-  sf::st_sf(crs = "+proj=eck4") # for tbular plot
-
-#  unpack the coordinates of each feature 
-coord_data <- as.data.frame( st_coordinates(st_cast(top_network_data_sf,"MULTILINESTRING")$geometry))
-
-c(top_network_data_sf$line_id, top_network_data_sf$coords) %<-% (
-  coord_data %>% 
-    rename(subline_id = L1) %>%   
-    nest(-L2)
-)
-
-# Create data path for line connections
-top_paths <- network_data_sf %>% 
-  sf::st_set_geometry(NULL) %>% 
-  unnest(cols = c(coords)) %>%
-  mutate_at(vars(X,Y), round,2) %>% 
-  gather("time_step","per_chng",per_early:per_mid) %>% 
-  mutate(per_plot = ifelse(per_chng > 50,50,per_chng),
-         direction = ifelse(per_chng > 0,"positive","negative")
-  )
-
-# Plot it!
-
-# Proportion map data 
-top_map_data <- top_data %>% 
-  gather("time_step","per_chng",per_early:per_mid) %>% 
-  mutate(per_plot = ifelse(per_chng > 50,50,per_chng))
-
-time <- "per_early"
-
-World_eez_sf <- World_eez_sf %>% 
-  sf::st_transform(crs = "+proj=eck4")
-
-World_map <- World_map %>% 
-  sf::st_transform(crs = "+proj=eck4")
-
-top_rop_map <- ggplot() +
-  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
-  geom_sf(data = subset(World_map,sovereignt != "Antarctica"), size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
-
-  geom_point(data = subset(top_map_data, variable =="mean_eez" & time_step == time & per_plot < 0), 
-             aes(start_lon,
-                 start_lat
-             ),
-             size = 0.05,
-             col = "gray80",
-             shape = 20,
-             stroke = 1.5) +
-  geom_path(data = subset(top_paths, variable =="mean_eez" & time_step == time & per_plot > 0), # filter data 
-            aes(X,
-                Y,
-                group = interaction(line_id,subline_id),
-                col = per_plot
-            ),
-            size = 0.9,
-            lineend = "round",
-            arrow = arrow(length = unit(0.008, "npc"),
-                          ends = "first"),
-            show.legend = NA) +
-  my_ggtheme_m(leg_pos = "bottom") +
-  scale_colour_viridis(
-    name= "Changes in catch proportion",
-    limits = c(0,max(Map_data$per_plot)),
-    breaks = seq(0,max(Map_data$per_plot),10),
-    labels = c(seq(0,max(Map_data$per_plot-10),10),">50")
-  ) + guides(
-    color=guide_colorbar(title.position="top", 
-                         title.hjust =0.5)
-  )
-
-ggsave(
-  plot = top_rop_map,
-  filename = "./Figures/top_ssr.png",
-  width = 10,
-  height = 5,
-  units = "in"
-)
-
-```
-
-##### Fig S2 (Histogram of change)
-
-```{r hist_change}
-
-Prop_results %>% 
-  gather("time_step","change",per_early,per_mid) %>% 
-  filter(variable == "mean_eez" & 
-           !is.na(change)
-  ) %>%
-  ggplot() +
-  geom_density(
-    aes(
-      x = ifelse(change >= 100,100,change),
-      # x = change,
-      fill = ifelse(time_step == "per_early","q",time_step)
-      # fill = time_step,
-
-    ),
-    # stat = "count",
-    # alpha= 0.8,
-    size=0.1,
-    # position = "stack"
-  ) +
-  scale_y_continuous("Density",
-                     # breaks = seq(0,90,10),
-                     # limits = c(0,100)
-  ) +
-  scale_x_continuous("Percentage Change",
-                     breaks = seq(-75,500,25)
-  ) +
-  my_ggtheme_p(leg_pos = "right") +
-  scale_fill_viridis(discrete = T, 
-                     option = "plasma",
-                     alpha = 0.5,
-                     name = "Time Step", labels = c("Middle", "Early")
-  ) +
-  coord_flip() +
-  ggsave("./Figures/hist_change.png",
-         width = 6,
-         height = 6,
-         units = "in"
-  )
-
-```
-
-## Sensitivity analysis
-
-### Testing different Per values for the centroids
-
-```{r data_loading, eval = F, echo = F}
-
-# SAU exploited species list
-Exploited_spp_list <- my_path("G","SAU","exploited_species_list.csv",read = T) %>%
-  clean_names() %>% 
-  pull(taxon_name)
-
-# Load all sp in one table
-sen_list <- c(20,50,90,95)
-
-sen <- 95
-read_sen <- function(sen){
-
-  if(sen != 95){
-
-    level <- paste0("Emergence_2005_p",sen)
-    extra_path <- paste0("Toe_Sens_per/",level)
-
-    # Get taxon list
-    Taxon_list_sens <- list.files(my_path("R",extra_path = extra_path))
-
-    Paths_sen <- paste(my_path("R",extra_path = extra_path),Taxon_list_sens,sep="")
-
-  }else{
-
-    Taxon_list <- gsub("_.*","",Taxon_list_sens)
-
-    Paths_sen <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_list_sens,"_emergence.txt",sep="")
-
-  }
-
-  Eme_sen <- bind_rows(lapply(Paths_sen, fread)) %>% 
-    mutate(sen = sen) %>% 
-    filter(
-      n_ens == 10
-    ) %>% 
-    left_join(Neighbours,
-              by = c("eez_name","eez_neighbour")
-    ) %>% 
-    rename(sf_eez_name=eez_name) %>% 
-    left_join(Matching_names,
-              by=  "sf_eez_name") %>% 
-    clean_names()
-
-  return(Eme_sen)
-
-}
-
-Sen_data <- bind_rows(
-  lapply(sen_list,read_sen)
-) %>% 
-  filter(
-    tresh== "tresh_two"
-  )
-
-unique(Sen_data$sen)
-
-```
-
-```{r data_analysis, eval = F, echo = F}
-
-# Per Stock
-sen_stock_d <-  
-  Sen_data %>% 
-  left_join(Neighbours) %>%
-  group_by(taxon_key,neighbour_id,sen) %>% 
-  summarise(
-    taxon_emer = min(emerging_yr)
-  ) %>% 
-  group_by(taxon_emer,sen) %>% 
-  summarise(
-    n_tax = n()
-  ) %>% 
-  group_by(sen) %>% 
-  mutate(
-    com_sum = cumsum(n_tax),
-    level = "stock"
-  )
-
-#  EEZ data
-sen_eez_d <-  Sen_data %>% 
-  left_join(Neighbours) %>% 
-  filter(n_ens == 10) %>% 
-  group_by(sf_eez_name,sen) %>% 
-  summarise(
-    taxon_emer = min(emerging_yr)
-  ) %>% 
-  group_by(taxon_emer,sen) %>% 
-  summarise(
-    n_tax = n()
-  ) %>% 
-  group_by(sen) %>% 
-  mutate(
-    com_sum = cumsum(n_tax),
-    level = "eez"
-  )
-
-# Plot Data
-sen_plot_data <- sen_stock_d %>% 
-  bind_rows(sen_eez_d) %>% 
-  mutate(
-    # total = ifelse(level == "stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$sf_eez_name))),
-    total = ifelse(level == "stock",max(sen_stock_d$com_sum),max(sen_eez_d$com_sum)),
-    per = com_sum/total*100
-  ) %>% 
-  gather("metric","value",com_sum,per) %>% 
-  ungroup() %>% 
-  mutate(
-    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
-    label = ifelse(total == max(sen_stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs"),
-    sen = paste("Percentile above",sen,"%") 
-  )
-
-# Lines data
-dashed_lines <- tibble(
-  x = c(2030,2050,2080),
-  label = c("Agenda 2030","Early range shift  ","Late range shift   ")
-)
-
-# Plot me
-ggplot() +
-  geom_area(data = subset(sen_plot_data, level == "stock"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = value,
-              fill = level
-            ),
-            size=0.5,
-            colour="black"
-  )  +
-  geom_area(data = subset(sen_plot_data, level == "eez"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = value,
-              fill = level
-            ),
-            size=0.5,
-            colour="black"
-  )  +
-  # Shameslly plot it again to fix colors
-  geom_area(data = subset(sen_plot_data, level == "stock" & my_facet == "Percentage (%)"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = value,
-              fill = level
-            ),
-            size=0.5,
-            colour="black"
-  ) +
-  facet_wrap(~my_facet + sen,
-             scales = "free",
-             nrow=2) + 
-  scale_y_continuous("",
-                     expand = c(0,0)
-  ) +
-  scale_x_continuous("Year of range shift",
-                     breaks = seq(2000,2100,25),
-                     limits = c(2005,2100),
-                     label = c(seq(2000,2075,25),">2100"),
-                     expand = c(0,0)
-  ) +
-  # Add dashed lines and labels
-  geom_vline(
-    data = dashed_lines,
-    aes(
-      xintercept = x,
-      linetype = label,
-    ),
-    color = "grey"
-  ) + 
-  my_ggtheme_p() +
-  theme(
-    legend.position = "right",
-    panel.spacing = unit(3, "lines") # space between plots
-  ) +
-  guides(fill=guide_legend(title= "", reverse=FALSE),
-         linetype = guide_legend(title = "Time step")) +
-  scale_fill_viridis(discrete = T,
-                     begin = 0.1, 
-                     end = 0.7,
-                     labels = c("EEZs","Stocks")
-  ) +
-  ggsave("./Figures/Supp3.png",
-         width = 14,
-         height = 10,
-         units = "in"
-  )
-
-```
-
-### Noise for all ensemble members
-
-```{r read_raw_data, eval = T, echo = F}
-
-# --------------- #
-# Load year of emergence results
-# --------------- #
-
-fread2 <- function(x){
-
-  tb <- fread(x)
-
-  if(nrow(tb) >0){
-    return(tb)
-  }
-
-}
-
-sens_paths <- list.files(my_path("R","Emergence_alt_sens"),full.names = T)
-
-# Load all sp in one table
-Eme_Alt_Res <- bind_rows(lapply(sens_paths, fread2)) %>% 
-  mutate(method= "alt") %>% 
-  left_join(Neighbours,
-            by = c("eez_name","eez_neighbour")
-  ) %>% 
-  rename(sf_eez_name=eez_name) %>% 
-  left_join(Matching_names,
-            by=  "sf_eez_name") %>% 
-  clean_names() %>% 
-  mutate(tresh = str_to_lower(tresh)) %>% 
-  filter(emerging_yr != 2100)
-
-# Combine both datas
-sens_data <- Eme_Res %>% 
-  filter(taxon_key %in% Eme_Alt_Res$taxon_key) %>% 
-  mutate(method = "org") %>% 
-  bind_rows(Eme_Alt_Res)
-
-n_stock <- sens_data %>% 
-  filter(method == "org") %>% 
-  group_by(neighbour_id) %>% 
-  summarise(n_stock = length(unique(taxon_key))) %>% 
-  # View()
-  group_by() %>% 
-  summarise(t_trans_stocks = sum(n_stock)) %>% 
-  pull(t_trans_stocks)
-
-n_eez <- length(unique(sens_data$sf_eez_name))
-
-```
-
-#### General numbers
-
-```{r sens_numbers, eval = T, echo = F}
-
-# Total number of emerging stocks
-
-# Global estimates
-sens_data %>% 
-  group_by(tresh,neighbour_id,method) %>% 
-  summarise(
-    emerging_stocks = length(unique(taxon_key))
-  ) %>% 
-  group_by(tresh,method) %>% 
-  summarise(
-    emerging_stocks = sum(emerging_stocks),
-    prop_total = (emerging_stocks/n_stock)*100,
-  )
-
-# Looks like there is no difference in the total number of emerging stocks
-# tresh     method emerging_stocks prop_total
-# 1 tresh_one alt               6665      116. 
-# 2 tresh_one org               5729      100  
-# 3 tresh_two alt               5258       91.8
-# 4 tresh_two org               4108       71.7
-
-# Global average ToE
-sens_data %>% 
-  group_by(tresh,method) %>% 
-  summarise(
-    # emerging_stocks = length(unique(taxon_key)),
-    # prop_total = (emerging_stocks/9132)*100,
-    mean= mean(emerging_yr),
-    sd = sd(emerging_yr),
-    min = min(emerging_yr),
-    max = max(emerging_yr)
-  )
-
-# Looks like the alternative method sets a later mean ToE but no real difference in s.d. min nor max
-#  tresh     method  mean    sd   min   max
-# 1 tresh_one alt    2055.  28.7  2007  2099
-# 2 tresh_one org    2029.  26.7  2006  2099
-# 3 tresh_two alt    2057.  28.6  2007  2099
-# 4 tresh_two org    2036.  28.1  2006  2099
-
-```
-
-#### Area plots 
-```{r read_raw_data, eval = T, echo = F}
-
-# Per Stock
-stock_d <-  
-  sens_data %>% 
-  # left_join(Neighbours) %>% 
-  # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
-  group_by(taxon_key,neighbour_id,tresh,method) %>% 
-  summarise(
-    taxon_emer = min(emerging_yr)
-  ) %>% 
-  group_by(taxon_emer,tresh,method) %>% 
-  summarise(
-    n_tax = n()
-  ) %>% 
-  group_by(tresh,method) %>% 
-  mutate(
-    com_sum = cumsum(n_tax),
-    level = "stock"
-  )
-
-# max(stock_d$com_sum)
-
-#  EEZ data
-eez_d <-  sens_data %>% 
-  # left_join(Neighbours) %>% 
-  filter(n_ens == 10) %>% 
-  # distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
-  group_by(sf_eez_name,tresh,method) %>% 
-  summarise(
-    taxon_emer = min(emerging_yr)
-  ) %>% 
-  group_by(taxon_emer,tresh,method) %>% 
-  summarise(
-    n_tax = n()
-  ) %>% 
-  group_by(tresh,method) %>% 
-  mutate(
-    com_sum = cumsum(n_tax),
-    level = "eez"
-  )
-
-# max(eez_d$com_sum)
-
-# Plot Data
-plot_data <- stock_d %>% 
-  bind_rows(eez_d) %>% 
-  filter(tresh != "NA") %>% 
-  mutate(
-    total = ifelse(level == "stock",n_stock,n_eez),
-    per = com_sum/total*100
-  ) %>% 
-  gather("metric","value",com_sum,per) %>% 
-  filter(tresh == "tresh_two")
-
-# 
-# # Manually include rest of data until 2100
-# cont_eez <- plot_data %>% 
-#   filter(level == "eez") %>% 
-#   arrange(taxon_emer) %>% 
-#   slice(1:24) %>%
-#   select(-1) %>% 
-#   arrange(metric) %>% 
-#   mutate(
-#     value = ifelse(metric == "per",81,228),
-#     taxon_emer = c(seq(2088,2099,1),seq(2088,2099,1))
-#   )
-
-plot_data <- plot_data %>% 
-  # bind_rows(cont_eez) %>% 
-  mutate(
-    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
-    label = ifelse(total == max(stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs")
-  )
-
-```
-
-```{r Time_lapse_plot, eval = F, echo = F}
-
-### Three plots
-
-count_eez_plot <- ggplot() +
-  geom_area(data = subset(plot_data, metric == "com_sum" & level == "eez"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = value,
-              fill = level
-            ),
-            size=0.5,
-            colour="black"
-  )  +
-  scale_y_continuous("Cumulative number of EEZs",
-                     expand = c(0,0)
-  ) +
-  scale_x_continuous("Year of range shift",
-                     breaks = seq(2000,2100,25),
-                     limits = c(2005,2100),
-                     label = c(seq(2000,2075,25),">2100")
-  ) +
-  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
-  scale_fill_viridis(discrete = T,
-                     begin = 0.1, 
-                     end = 0.1
-  ) +
-  facet_wrap(~ method);count_eez_plot
-
-#### Count stock plot 
-count_stock_plot <- ggplot() +
-  geom_area(data = subset(plot_data, metric == "com_sum" & level == "stock"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = value,
-              fill = level
-            ),
-            size=0.5,
-            colour="black"
-  )  +
-  scale_y_continuous("Cumulative number of stocks",
-                     expand = c(0,0)
-  ) +
-  scale_x_continuous("",
-                     breaks = seq(2000,2100,25),
-                     limits = c(2005,2100),
-                     label = c(seq(2000,2075,25),">2100")
-  ) +
-  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
-  scale_fill_viridis(discrete = T,
-                     begin = 0.7, 
-                     end = 0.7
-  ) +
-  facet_wrap(~ method);count_stock_plot
-
-# Percentage plot
-percentage_plot <- ggplot() +
-  geom_area(data = subset(plot_data, level == "eez" & my_facet == "Percentage (%)"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = value,
-              fill = level
-            ),
-            size=0.5,
-            colour="black"
-  ) + 
-  geom_area(data = subset(plot_data, level == "stock" & my_facet == "Percentage (%)"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = value,
-              fill = level
-            ),
-            size=0.5,
-            colour="black"
-  ) + 
-  scale_y_continuous("Cumulative proportion of the total (%)",
-                     expand = c(0,0)
-  ) +
-  scale_x_continuous("",
-                     breaks = seq(2000,2100,25),
-                     limits = c(2005,2100),
-                     label = c(seq(2000,2075,25),">2100")
-  ) +
-  my_ggtheme_p(leg_pos = "right", hjust = 0.5) +
-  scale_fill_viridis(discrete = T,
-                     begin = 0.1,
-                     end = 0.7,
-                     labels = c("EEZs","Stocks")
-  )+
-  facet_wrap(~ method);percentage_plot
-
-# Legend
-legend <- get_legend(percentage_plot)
-
-percentage_plot <- percentage_plot + 
-  theme(legend.position = "")
-
-# Draw figure
-plot_all <- ggdraw() +
-  # Count stock plot
-  draw_plot(count_stock_plot, x = 0.01, y = 0, width = 0.29, height = 0.90) +
-  # Count EEZ plot
-  draw_plot(count_eez_plot, x = 0.31, y = 0, width = 0.29, height = 0.90) +
-  # Percentage plot
-  draw_plot(percentage_plot, x = 0.61, y = 0 , width = 0.29, height = 0.90) +
-  # legend
-  draw_plot(legend, x = 0.84, y = 0 , width = 0.20, height = 1) +
-  # Labels
-  draw_plot_label(label = c("A", "B","C"), 
-                  size = 20,
-                  x = c(0.01, 0.285,0.59), 
-                  y = c(1,1,1)
-  )
-
-# Save plot
-save_plot(
-  "./Figures/cumu_per_freq_sens.png",
-  plot_all,
-  base_height = 5,
-  base_width = 12
-)
-
-```
-
-#### Comparative map
-
-```{r}
-
-# Per EEZs
-Eez_d_hist <- sens_data %>% 
-  group_by(eezid,sau_fishing_entity,tresh,method) %>% 
-  summarise(
-    mean_yr = mean(emerging_yr,na.rm=T),
-    sd_yr = sd(emerging_yr,na.rm=T),
-    n_taxa = length(unique(taxon_key))
-  ) %>% 
-  gather("metric","value",mean_yr:n_taxa) %>% 
-  mutate(
-    reverse = abs(value-2100)  + 2000
-  )
-
-# Estimate the value (catch and USD) of emerging species
-Eme_value_data <- sens_data %>% 
-  group_by(sau_fishing_entity,fishing_entity_id,tresh,taxon_key,method) %>% 
-  summarise(n()) %>% 
-  left_join(Sau_data,
-            by = c("fishing_entity_id","taxon_key")) %>% 
-  filter(!is.na(variable)) %>% 
-  group_by(sau_fishing_entity,fishing_entity_id,tresh,variable,method) %>% 
-  summarise(
-    sum_value = sum(value,na.rm=T),
-    n_taxa = length(unique(taxon_key)) 
-  )
-
-# Total value of each fishing nation
-
-Total_value <- Sau_data %>% 
-  group_by(fishing_entity_id,variable) %>% 
-  summarise(tot_value= sum(value,na.rm=T))
-
-Value_map_data <- Eme_value_data %>% 
-  left_join(Total_value,
-            by = c("fishing_entity_id","variable")) %>% 
-  mutate(participation = (sum_value/tot_value)*100+2000)
-
-# Per fishing entity
-
-fish_d_hist <- sens_data %>% 
-  group_by(sau_fishing_entity,tresh,method) %>% 
-  summarise(
-    value = length(unique(taxon_key))
-  ) %>% 
-  filter(!is.na(sau_fishing_entity))
-
-World_map_nemer <- World_map %>%
-  full_join(Value_map_data,
-            by = c("sau_fishing_entity","fishing_entity_id")
-  ) %>% 
-  filter(!is.na(participation))
-
-# These are regions without transboundary species
-grey_land <- World_map %>% 
-  filter(!fishing_entity_id %in% Value_map_data$fishing_entity_id) %>% 
-  filter(!sovereignt %in% c("Antarctica"))
-
-# -------------------------- #
-# Call plot
-# -------------------------- #
-tresh_list <- "tresh_two"
-
-name_save <- paste("./Figures/emer_map_sens_",tresh_list,".png",sep = "")
-
-# Value Map
-World_map_part <- World_map_nemer %>% 
-  filter(
-    tresh ==  tresh_list,
-    variable == "value"
-  )
-
-# ToE Map 
-toe_map_data <- World_eez_sf %>% 
-  left_join(Eez_d_hist,
-            by = "eezid") %>% 
-  filter(tresh == tresh_list,
-         metric == "mean_yr")
-
-# Map with countries without ToE
-
-russia_churchil <- World_eez_sf %>% 
-  filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))
-
-no_toe_eez <- World_eez_sf %>% 
-  filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
-         !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
-  rbind(russia_churchil)
-
-no_toe_land <- World_map %>% 
-  filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
-         !name %in% World_map_part$name) # Filter EEZs with ToE stocks
-
-map <- ggplot(toe_map_data) +
-  geom_sf(
-    aes(
-      fill = reverse
-    ),
-    color =  "white",
-    size = 0.1
-  ) +
-  geom_sf(data = World_map_part,
-          aes(fill = participation),
-          color =  "white",
-          size = 0.1
-  ) +
-  geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
-  geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
-  geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
-  scale_fill_viridis("",
-                     limits = c(2000,2100),
-                     breaks = seq(2000,2100,25),
-                     labels = c(">2100",seq(2075,2000,-25)) # not inluding the > because we
-  ) +
-  my_ggtheme_m(leg_pos = "bottom") +
-  facet_wrap(~method,
-             nrow = 2)
-
-ggsave(name_save,
-       width = 10,
-       height = 6,
-       units = "in"
-)
-
-```
-
-# END PAPER
-
-# Troubleshooting
-
-### Weird connections from Australia
-
-```{r fig_map_Juan, eval = F, echo =F}
-
-##_----------------------- #
-# Checking on China and Indonesia
-# Probably due to dispute territories
-
-Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
-  clean_names()
-
-# 
-# Neighbours %>% 
-#   filter(eez_name == "China") %>% 
-#   pull(eez_neighbour) %>% 
-#   unique()
-
-##_----------------------- #
-
-# Get the center poligon of each EEZ for map source/target
-coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
-  ungroup() %>% 
-  select(eez_name,geometry) %>% 
-  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
-  # Manually fix issues
-  mutate(
-    longitude = ifelse(
-      eez_name == "Fiji", 178.065033,
-      ifelse(eez_name == "Tuvalu",177.6493,
-             ifelse( eez_name == "New Zealand",174.8860,
-                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
-                     )
-             )
-      )
-    )
-  )
-
-# Remove "c()"
-coords$longitude <- gsub("\\(","",coords$longitude)
-coords$longitude <- gsub("c","",coords$longitude)
-coords$latitude <- gsub(")","",coords$latitude)
-
-coords <- coords %>% 
-  mutate_at(vars(longitude,latitude),as.numeric)
-
-# Prepare data from results
-Test_Data <- results %>% 
-  # get coords for sources
-  left_join(coords,
-            by = "eez_name") %>%
-  rename(source = eez_name,
-         start_lon = longitude,
-         start_lat = latitude,
-         eez_name=eez_neighbour) %>% 
-  # get coords for targets
-  left_join(coords,
-            by = "eez_name") %>% 
-  rename(target = eez_name,
-         end_lon = longitude,
-         end_lat = latitude) %>% 
-  ungroup() %>% 
-  # Plot miscs
-  mutate(
-    per_mid = ifelse(is.na(per_mid),0,per_mid)
-  ) %>% 
-  drop_na() %>% 
-  filter(
-    # per_mid >= 20,
-    variable == "mean_spp",
-  )
-
-# Convert to data to spatial
-network_data_sf <- Test_Data %>%
-  select(start_lon, start_lat, end_lon, end_lat) %>% 
-  purrr::transpose() %>% 
-  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
-  purrr::map(st_linestring) %>%
-  st_sfc(crs = 4326) %>%
-  st_sf(geometry = .)%>%
-  bind_cols(Test_Data) %>%
-  select(everything(), geometry) %>% 
-  # Set circles ('cus the world is not flat!)
-  st_segmentize(units::set_units(100, km)) %>% 
-  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
-  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
-  sf::st_sf(crs = 4326)
-
-#  unpack the coordinates of each feature 
-coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
-
-c(network_data_sf$line_id, network_data_sf$coords) %<-% (
-  coord_data %>% 
-    rename(subline_id = L1) %>%   
-    nest(-L2)
-)
-
-# Create data path for line connections
-paths <- network_data_sf %>% 
-  sf::st_set_geometry(NULL) %>% 
-  unnest() %>% 
-  mutate_at(vars(X,Y), round,2)
-
-# Plot it!
-
-country <- c("Fiji","Solomon Isl.","Tuvalu")
-
-ggplot() +
-  geom_sf(
-    # data = subset(world_map,name %in% country),
-    data = world_map,
-    size = 0.1, fill = "gray90", col = "gray90") +
-  geom_sf(
-    # data = subset(World_eez_sf, eez_name %in% country),
-    data = World_eez_sf,
-    size = 0.1, fill = "white", col = "gray70") +
-  geom_point(data = Test_Data,
-             # data = subset(Test_Data, source %in% country & target %in% source),
-             aes(start_lon,
-                 start_lat,
-                 col = per_mid),
-             size = 0.1,
-             col = "black",
-             # shape = 21,
-             stroke = 1.5) +
-  geom_path(data = paths,
-            # data = subset(paths, source %in% country),
-            aes(X, Y,
-                group = interaction(line_id,subline_id),
-                col = per_mid
-            ),
-            size = 0.3,
-            arrow = arrow(length = unit(0.009, "npc")),
-            show.legend = NA) +
-  # geom_sf_text(data = subset(World_eez_sf, eez_name %in% country),
-  #              aes(label =eez_name),
-  #              size = 2) +
-  # geom_text(data = subset(Test_Data, source %in% country),
-  #              aes(start_lon,
-  #                start_lat,
-  #                label = paste(source,"to",target)
-  #              ),
-  #              size = 2) +
-  # geom_text_repel(data = Test_Data,
-  #   aes(start_lon,
-# start_lat,
-# label = paste(source,"to",target)
-# )
-# )+
-ggtheme_map() +
-  scale_colour_gradientn("Gain of Stock-Share Ratio (%)",
-                         colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
-                         limits = c(0,25),
-                         breaks = seq(0,25,5)
-  )
-
-```
-
-### Arctic and Brasil missing species
-
-```{r arctic, eval = F, echo =F}
-
-# Old Code
-
-###### Table.1 Multy linear model
-
-Arctic <- c("Canada (Arctic)","USA (Alaska, Arctic)", "Russia (Laptev to Chukchi Sea)")
-
-can_us_arctic <- Transboundary_spp %>% 
-  filter(eez_name == "Canada (Arctic)",
-         eez_neighbour == "USA (Alaska, Arctic)")
-
-arctic_spp <- Transboundary_spp %>% 
-  filter(eez_name %in% Arctic,
-         eez_neighbour %in% Arctic)
-
-arctic_spp <- Spp_depth %>% 
-  filter(taxon_key %in% can_us_arctic$taxon_key)
-
-# Transbounudary species that will see no change
-# 600252 Mallotus villosus  (Capelin)
-# 600315 Eleginus gracilis (Safforn cod)
-# 601520 Clupea pallasii  (Pacific herring)
-
-Eme_Res %>% 
-  filter(taxon_key %in% can_us_arctic$taxon_key) %>% 
-  # filter(sau_eez_name %in% arctic_spp$eez_name) %>% 
-  View(.)
-
-```
-
-```{r Brasa, eval = F, echo =F}
-
-brasil <- Transboundary_spp %>% 
-  filter(eez_name == "Brazil")
-
-brasil_spp <- Spp_depth %>% 
-  filter(taxon_key %in% brasil$taxon_key)
-
-# Transbounudary species that will see no change
-# 601590  Brevoortia pectinata  
-# 690022  Pleoticus muelleri 
-# 690023  Illex argentinus  
-# 690338  Penaeus paulensis
-
-Eme_Res %>% 
-  filter(taxon_key %in% brasil_spp$taxon_key) %>% 
-  View(.)
-
-```
-
-```{r TransIndex, eval = F, echo = F}
-
-TransIndex <- function(Spp,year, data_type = "Catch",ensemble_list,path,Neighbours,eez_centroid){
-
-  # ----------------- #
-  # Get Species Centrodis
-  #  ---------------- #
-
-  # Get model data from spps
-  Spp_Dist <- bind_rows(
-    lapply(
-      ensemble_list,
-      dbem_import,
-      taxon_key = Spp,
-      year = year,
-      data_type = data_type,
-      path = path
-    )
-  )
-
-  #### Control for Spp id so I can fix problems
-
-  if(file.exists("~/Desktop/spp_df.csv") == FALSE){
-    spp_df <- tibble() %>% 
-      write_csv(.,
-                "~/Desktop/spp_df.csv")
-  }
-
-  suppressMessages(
-    spp_df <- read_csv("~/Desktop/spp_df.csv")
-  )
-  spp_df <- tibble(Spp) %>% 
-    bind_rows(spp_df) %>% 
-    write_csv(.,
-              "~/Desktop/spp_df.csv")
-
-  #____________ Selecting only the transboundary nature of the species _________ #
-  Trans_Spp <- Spp_Dist %>%
-    left_join(index_code,
-              by= "index") %>% 
-    filter(!is.na(value)) %>% 
-    semi_join(Transboundary_spp, # filter only transboundary cases
-              by = c("taxon_key","eez_name")
-    )
-
-  # Step
-  nr <- nrow(Trans_Spp)
-  if(nr > 0){
-
-    # Filter eez by those transboundary
-    unique_eez <- unique(Trans_Spp$eez_name)
-
-    Neighbours_combo <- Neighbours %>% 
-      filter(eez_name %in% unique_eez,
-             eez_neighbour %in% unique_eez)
-
-    # Get the centroid of each country
-    EEZ_centroid <- eez_centroid %>% 
-      filter(eez_name %in% Trans_Spp$eez_name) %>% 
-      mutate(taxon_key = Spp)
-
-    Ids <- unique(Neighbours_combo$neighbour_id)
-
-    corefx <- function(Ids){
-
-      EEZ_ids <- Neighbours_combo %>% 
-        filter(neighbour_id == Ids)
-
-      Sub_Trans_Spp <- Trans_Spp %>% 
-        filter(eez_name %in% EEZ_ids$eez_name)
-
-      Sub_EEZ_centroid <- EEZ_centroid %>% 
-        filter(eez_name %in% EEZ_ids$eez_name)
-
-      # Centroid is defined as the tile with the higest species value
-      Centroids <- Sub_Trans_Spp %>% 
-        left_join(EEZ_ids, by = "eez_name") %>% 
-        group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
-        top_n(1,value) %>% 
-        distinct(value,year, .keep_all = TRUE) %>%  # removes duplicates when tile is in between EEZs
-        left_join(Sub_EEZ_centroid, by = "taxon_key") %>%
-        select(eez_name =eez_name.y, everything() ,-eez_name.x)
-
-      # Estimate distances between spp centroid and EEZ centroid
-      Distance_Data <- Centroids %>%
-        rename(
-          catch_lon = lon,
-          catch_lat = lat,
-          eez_lon = longitude,
-          eez_lat= latitude
-        ) %>% 
-        # Estimate the distance between centriods
-        mutate(
-          distance = geodist(eez_lon, eez_lat, catch_lon, catch_lat, units="km")
-        ) %>% 
-        select(taxon_key,eez_name,neighbour_id,ensemble,year,distance)
-
-      # Historical mean and sd of distance
-      historical <- Distance_Data %>%
-        filter(year <= 2000) %>%
-        ungroup() %>%
-        group_by(taxon_key,ensemble,eez_name,neighbour_id) %>%
-        summarise_at(vars("distance"),
-                     c(hd_temp_mean=mean,
-                       hd_temp_sd=sd)
-        )
-
-      # Future results
-      future <- Distance_Data %>% 
-        filter(year > 2000)
-
-      # Transboundary Index
-      Trans_index <- Distance_Data %>% 
-        left_join(historical,
-                  by = c("taxon_key","eez_name","ensemble","neighbour_id")
-        ) %>%
-        # Re arrenging table for mutate
-        gather("variable","value",distance:hd_temp_sd) %>%
-        ungroup()
-
-      Partial <- Trans_index %>% 
-        mutate(
-          names = ifelse(eez_name ==Sub_EEZ_centroid$eez_name[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_"))
-        ) %>%
-        select(-eez_name,-variable) %>%
-        spread(names,value) %>% 
-        # mutate index
-        mutate(
-          trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
-        ) %>% 
-        group_by(taxon_key,ensemble,neighbour_id) %>% 
-        mutate(RMean = rollmean(x = trans_index, 
-                                10, 
-                                align = "right", 
-                                fill = trans_index)
-        ) %>% 
-        filter(year > 1960) %>% 
-        select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)
-
-      # Average of 10 years mean of ensemble members results
-      mean_sd_index <- Partial %>% 
-        group_by(taxon_key,year,neighbour_id) %>% 
-        summarise_at(vars("RMean"),
-                     c(trend=mean,
-                       ensemble_sd=sd)
-        )
-
-      # Estimate noice from ensemble members
-      historic_sd <- mean_sd_index %>% 
-        filter(year < 2000) %>% 
-        group_by(taxon_key) %>% 
-        summarise_at(vars("trend"),
-                     c(hist_mean=mean,
-                       noise=sd)
-        )
-
-      # Plot the result, bruh
-      # ggplot() +
-      #   geom_line(data = mean_sd_index,
-      #             aes(
-      #               x = year,
-      #               y = trend
-      #             )
-      #   ) +
-      #   geom_ribbon(data = mean_sd_index,
-      #               aes(x=year,
-      #                   ymin=historic_sd$hist_mean-historic_sd$noise,
-      #                   ymax=historic_sd$hist_mean+historic_sd$noise),
-      #               alpha = 0.5,
-      #               fill = "grey50"
-      #   ) +
-      #   geom_line(data = mean_sd_index,
-      #             aes(
-      #               x = year,
-      #               y = historic_sd$hist_mean
-      #             ),
-      #             colour = "grey50"
-      #   ) +
-      #   geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
-      #              aes(
-      #                x = year,
-      #                y = trend,
-      #                colour = "red",
-      #              ),
-      #              shape = 1,
-      #              size = 3
-      #   ) +
-      #   geom_vline(xintercept = 2000,colour="blue",alpha = 0.5) +
-      #   theme_classic()
-
-      # #     
-      Final_Result <- mean_sd_index %>%
-        filter(year > 2000,
-               trend > (historic_sd$hist_mean+historic_sd$noise) |
-                 trend < (historic_sd$hist_mean-historic_sd$noise)
-        ) %>%
-        group_by(taxon_key,neighbour_id) %>%
-        summarise(emerging_yr= min(year)) %>%
-        left_join(EEZ_ids) %>%
-        ungroup() %>%
-        select(taxon_key,eez_name,eez_neighbour,emerging_yr)
-
-      # if(i == 1){
-      #   f_df <- Final_Result
-      # }else{
-      #   f_df <- bind_rows(f_df,Final_Result)
-      # }
-      return(Final_Result)
-    }
-
-    x <- bind_rows(
-      mclapply(Ids,corefx)
-    )
-    # return(f_df)
-
-  }else{
-
-    print("no share data")
-
-    x <- tibble(
-      taxon_key = Spp,
-      eez_name = NA,
-      emerging_yr = NA
-    )
-
-  }
-
-  return(x)
-
-}
-
-#  #One spp works 
-# suppressMessages(
-# TransIndex(Spp = 600006,
-#              year,
-#              data_type = "Catch",
-#              ensemble_list,
-#              path,
-#              Neighbours = Neighbours,
-#              eez_centroid)
-# Multiple spp
-
-# spp_list <- c(600004,
-#               600005,
-#               600245,
-#               600006
-#               )
-# 
-# suppressMessages(
-#   suppressWarnings(
-#     Test <- bind_rows(
-#       lapply(spp_list,
-#              TransIndex,
-#              year,
-#              data_type = "Catch",
-#              ensemble_list,
-#              path,
-#              Neighbours,
-#              eez_centroid
-#       )
-#     )
-#   )
-# )
-
-```
-
-###### With relation to neighbours
-
-```{r time_emergence_map,  eval = F, echo = T}
-
-##_----------------------- #
-# Checking on China and Indonesia
-# Probably due to dispute territories
-
-Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
-  clean_names()
-
-# 
-# Neighbours %>% 
-#   filter(eez_name == "China") %>% 
-#   pull(eez_neighbour) %>% 
-#   unique()
-
-##_----------------------- #
-
-# Get the center poligon of each EEZ for map source/target
-coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
-  ungroup() %>% 
-  select(eez_name,geometry) %>% 
-  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
-  # Manually fix issues
-  mutate(
-    longitude = ifelse(
-      eez_name == "Fiji", 178.065033,
-      ifelse(eez_name == "Tuvalu",177.6493,
-             ifelse( eez_name == "New Zealand",174.8860,
-                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
-                     )
-             )
-      )
-    )
-  )
-
-# Remove "c()"
-coords$longitude <- gsub("\\(","",coords$longitude)
-coords$longitude <- gsub("c","",coords$longitude)
-coords$latitude <- gsub(")","",coords$latitude)
-
-coords <- coords %>% 
-  mutate_at(vars(longitude,latitude),as.numeric)
-
-# Prepare data from results
-Test_Data <- Eme_Res %>% 
-  filter(
-    n_ens == 10#,
-    # emerging_yr >= 2020
-  ) %>% 
-  group_by(eez_name,eez_neighbour,emerging_yr) %>% 
-  summarise(
-    n_tax = length(unique(taxon_key))
-  ) %>% 
-  filter(
-    emerging_yr <= 2050
-  ) %>%
-  group_by(eez_name,eez_neighbour) %>% 
-  summarise(
-    n_tax_emer = sum(n_tax)
-  ) %>% 
-  # get coords for sources
-  left_join(coords,
-            by = "eez_name") %>%
-  rename(source = eez_name,
-         start_lon = longitude,
-         start_lat = latitude,
-         eez_name=eez_neighbour) %>% 
-  # get coords for targets
-  left_join(coords,
-            by = "eez_name") %>% 
-  rename(target = eez_name,
-         end_lon = longitude,
-         end_lat = latitude) %>% 
-  ungroup() %>% 
-  # Plot miscs
-  # mutate(
-  #   per_mid = ifelse(is.na(per_mid),0,per_mid)
-  # ) %>% 
-  drop_na() #%>% 
-# filter(
-# per_mid >= 20,
-# variable == "mean_spp",
-# )
-
-# Convert to data to spatial
-network_data_sf <- Test_Data %>%
-  select(start_lon, start_lat, end_lon, end_lat) %>% 
-  purrr::transpose() %>% 
-  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
-  purrr::map(st_linestring) %>%
-  st_sfc(crs = 4326) %>%
-  st_sf(geometry = .)%>%
-  bind_cols(Test_Data) %>%
-  select(everything(), geometry) %>% 
-  # Set circles ('cus the world is not flat!)
-  st_segmentize(units::set_units(100, km)) %>% 
-  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
-  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
-  sf::st_sf(crs = 4326)
-
-#  unpack the coordinates of each feature 
-coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))
-
-c(network_data_sf$line_id, network_data_sf$coords) %<-% (
-  coord_data %>% 
-    rename(subline_id = L1) %>%   
-    nest(-L2)
-)
-
-# Create data path for line connections
-paths <- network_data_sf %>% 
-  sf::st_set_geometry(NULL) %>% 
-  unnest() %>% 
-  mutate_at(vars(X,Y), round,2)
-
-# Plot it!
-
-country <- c("Fiji","Solomon Isl.","Tuvalu")
-
-ggplot() +
-  geom_sf(
-    # data = subset(world_map,name %in% country),
-    data = World_map,
-    size = 0.1, fill = "gray90", col = "gray90") +
-  geom_sf(
-    # data = subset(World_eez_sf, eez_name %in% country),
-    data = World_eez_sf,
-    size = 0.1, fill = "white", col = "gray70") +
-  geom_point(data = Test_Data,
-             # data = subset(Test_Data, source %in% country & target %in% source),
-             aes(start_lon,
-                 start_lat,
-                 col = n_tax_emer
-             ),
-             size = 0.1,
-             col = "grey50",
-             # shape = 21,
-             stroke = 1.5) +
-  geom_path(data = paths,
-            # data = subset(paths, source %in% country),
-            aes(X, Y,
-                group = interaction(line_id,subline_id),
-                col = n_tax_emer
-            ),
-            size = 0.3,
-            # arrow = arrow(length = unit(0.009, "npc")),
-            show.legend = NA) +
-  ggtheme_map() +
-  scale_colour_viridis(
-    name="Number of emerging stocks\nby 2050",
-    limits = c(0,40),
-    breaks = seq(0,40,10),
-    # option="magma",
-    direction = 1
-  ) +
-  ggsave("./Figures/figure_emerging_2050.png",
-         width = 10,
-         height = 6,
-         units = "in"
-  )
-
-```
-
-### Non selected Results
-
-```{r, olda_acumulative, eval = F}
-
-# filter(emerging_yr >= 2020) %>% 
-ggplot() +
-  geom_area(data = subset(plot_data, level == "stock"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = com_sum,
-              fill = level
-            ),
-            alpha= 0.8,
-            size=0.5,
-            colour="black"
-  )  +
-  geom_area(data = subset(plot_data, level == "eez"),
-            aes(
-              x = taxon_emer,
-              y = com_sum,
-              fill = level
-            ),
-            alpha= 0.8,
-            size=0.5,
-            colour="black"
-  ) +
-  facet_wrap(~tresh) + 
-  # Comment out for option A (Change plot name!!)
-  #   geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
-  # geom_text(aes(label="Total number of EEZs", x = 2010, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
-  #   geom_hline(yintercept = length(unique(Eme_Res$taxon_key)), color = "grey",linetype = "dashed") +
-  # geom_text(aes(label="Total number of Taxa", x = 2010, y = length(unique(Eme_Res$taxon_key))+10), color = "grey70",size =3) +
-  scale_y_continuous("Cumulatuve Curve",
-                     breaks = seq(0,600,50),
-                     limits = c(0,600)
-  ) +
-  scale_x_continuous("Year of emergence",
-                     breaks = seq(2005,2100,10),
-                     limits = c(2005,2100)
-  ) +
-  my_ggtheme_p() +
-  scale_fill_viridis(discrete = T) +
-  ggsave("./Figures/figure_accumulative_curve_ab.png",
-         width = 6,
-         height = 6,
-         units = "in" 
-  )
-
-### Percentage accumulation
-
-ggplot() +
-  geom_area(data = subset(plot_data, level == "stock"),
-            aes(
-              x = as.numeric(taxon_emer),
-              y = per,
-              fill = level
-            ),
-            alpha= 0.8,
-            size=0.5,
-            colour="black"
-  )  +
-  geom_area(data = subset(plot_data, level == "eez"),
-            aes(
-              x = taxon_emer,
-              y = per,
-              fill = level
-            ),
-            alpha= 0.8,
-            size=0.5,
-            colour="black"
-  ) +
-  scale_y_continuous("Cumulatuve Curve",
-                     breaks = seq(0,100,25),
-                     limits = c(0,100)
-  ) +
-  scale_x_continuous("Year of emergence",
-                     breaks = seq(2005,2100,15),
-                     limits = c(2005,2100)
-  ) +
-  my_ggtheme_p() +
-  scale_fill_viridis(discrete = T) +
-  # facet_wrap(~tresh) + 
-  ggsave("./Figures/Fig4.png",
-         width = 8,
-         height = 8,
-         units = "in"
-  )
-
-```
-
-```{r histo,  eval = F, echo = T}
-
-# First time a Species becomes emergent
-Eme_Res %>% 
-  filter(
-    n_ens == 10#,
-    # tresh == "tresh_one"
-  ) %>% 
-  group_by(tresh,taxon_key,emerging_yr) %>% 
-  summarise(
-    n_trans = length(unique(sf_eez_name))
-  ) %>% 
-  group_by(tresh,taxon_key) %>% 
-  summarise(
-    ntrans = min(emerging_yr)
-  ) %>% 
-  group_by(tresh,ntrans) %>% 
-  summarise(
-    n_taxa = length(unique(taxon_key))
-  ) %>% 
-  arrange(-n_taxa) %>% 
-  # View()
-  # group_by(tresh) %>% 
-  # summarise(sum(n_taxa))
-  ggplot() +
-  geom_line(
-    aes(
-      x = ntrans,
-      y = n_taxa,
-      color = tresh
-    )
-  )
-
-# EEZ, first time an EEZ has a emergent stock
-Eme_Res %>% 
-  filter(
-    n_ens == 10#,
-    # tresh == "tresh_one"
-  ) %>% 
-  group_by(tresh,sf_eez_name,emerging_yr) %>% 
-  summarise(
-    n_trans = length(unique(taxon_key))
-  ) %>% 
-  group_by(tresh,sf_eez_name) %>% 
-  summarise(
-    ntrans = mean(emerging_yr)
-  ) %>% 
-  group_by(tresh,ntrans) %>% 
-  summarise(
-    n_taxa = length(unique(sf_eez_name))
-  ) %>% 
-  arrange(-n_taxa) %>% 
-  View()
-# group_by(tresh) %>%
-# summarise(sum(n_taxa))
-# ggplot() +
-# geom_line(
-#   aes(
-#     x = ntrans,
-#     y = n_taxa,
-#     color = tresh
-#   )
-# )
-
-# Numer of EEZs where stocks will emerge
-stock_d_hist <- Eme_Res %>% 
-  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
-  group_by(emerging_yr,taxon_key,tresh) %>%
-  summarise(n_tax = length(unique(sf_eez_name))) %>%
-  # summarise(n()) %>% 
-  mutate(
-    level = "EEZs per Spp"
-  )  %>% 
-  ungroup() %>% 
-  select(-taxon_key)
-
-# Per EEZs
-eez_d_hist <- Eme_Res %>% 
-  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
-  group_by(emerging_yr,sf_eez_name,tresh,neighbour_id) %>%  
-  # summarise(n_tax = length(unique(taxon_key))) %>% 
-  summarise(n()) %>% 
-  mutate(
-    level = "Spp per EEZ"
-  ) %>% 
-  ungroup() %>% 
-  select(-sf_eez_name)
-
-stock_d_hist %>% 
-  bind_rows(eez_d_hist) %>% 
-  filter(tresh != "NA") %>% 
-  ggplot() +
-  geom_density(
-    aes(
-      x = emerging_yr,
-      fill = level
-    ),
-    stat = "count",
-    alpha= 0.8,
-    size=0.5,
-  ) +
-  scale_y_continuous("Frequency",
-                     # breaks = seq(0,300,50),
-                     # limits = c(0,300)
-  ) +
-  scale_x_continuous("Year of emergence",
-                     # breaks = seq(2006,2100,25),
-                     # limits = c(2000,2100)
-  ) +
-  my_ggtheme_p() +
-  scale_fill_viridis(discrete = T) +
-  facet_wrap(~tresh) +
-  ggsave("./Figures/hist_freq.png",
-         width = 6,
-         height = 6,
-         units = "in"
-  )
-
-```
-
-##### Temperature gradient
-
-##### Fun `GetEnvData`
-
-```{r get_env_data, eval = F, echo = T}
-
-getEnvVar=function(yr,ModelRCP,RCP,Variable){
-
-  if(ModelRCP == "IPSL"){
-    ModelRCP = "IPSL-CM5A-MR"
-    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
-  }
-
-  if(ModelRCP == "MPI"){
-    ModelRCP = "MPI-ESM-MR"
-    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
-  }
-
-  if(ModelRCP == "GFDL"){
-
-    if(RCP == "rcp26"){
-      ModelRCP <- "GFDL26"
-    }else{
-      ModelRCP <- "GFDL85"
-    }
-    Filepath <- paste("Z:/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")
-  }
-
-  if(Variable == "Ice"){
-
-    Bottom <- paste("IceExt_",yr,".txt",sep="")
-
-  }
-
-  if(Variable == "Temperature"){
-
-    Bottom = paste("SST_",yr,".txt",sep="")
-
-  }
-
-  if(Variable == "Salinity"){
-    # Bottom
-    Bottom = paste("Salinity_btm_",yr,".txt",sep="")
-
-  }
-
-  Read_me <- paste(Filepath,Bottom,sep="")[1]
-  Bottom_Data <- read.csv(Read_me,header=F)
-
-  # Retrun
-
-  getTemp <- Bottom_Data 
-}
-
-Years <-seq(1951,2100,1)
-
-for(y in 1:length(Years)){
-  Data <- getEnvVar(yr = Years[y],
-                    Model = "GFDL",
-                    Variable = "Temperature",
-                    RCP = "rcp85"
-  )
-
-  if(y == 1){
-
-    F_BData <- Data[1]
-    colnames(F_BData) <- Years[1]
-
-  }else{
-
-    F_BData <- cbind(F_BData,Data[1])
-    colnames(F_BData)[y] <- Years[y]
-  }
-}
-
-F_BData <- F_BData %>% 
-  rowid_to_column("index") %>% 
-  gather("year","temp",-index) %>% 
-  mutate(temp = ifelse(temp == -9999,NA,temp)) %>% 
-  group_by(year) %>% 
-  summarise(
-    mean_temp = mean(temp,na.rm=T)
-  ) %>% 
-  write_csv(.,
-            "SST_data.csv")
-
-```
-
-```{r temp_plot, eval = F, echo = T}
-
-### NEED TO RUN TIME LAPSE CHUNK FIRST TO GET ACCUMULATIVE DATA
-
-# Get sst_data
-sst_data <- my_path("D",T,"Climate","SST_data.csv",read = T)
-
-sst_hist <- sst_data %>% 
-  filter(year < 2000) %>% 
-  group_by() %>% 
-  summarise(hist_sst =mean(mean_temp)) %>% 
-  pull()
-
-sst_change <- sst_data %>%
-  filter(year > 2020) %>% 
-  mutate(chng = round(mean_temp-sst_hist,2))
-
-# Plot Data
-stock_d %>% 
-  bind_rows(eez_d) %>% 
-  filter(taxon_emer >= 2020) %>% 
-  rename(year = taxon_emer) %>% 
-  left_join(sst_change) %>% 
-  ggplot() +
-  geom_line(
-    aes(
-      x = chng,
-      y = com_sum,
-      colour = level
-    )
-  ) +
-  geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
-  geom_text(aes(label="Total number of EEZs", x = 0.5, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
-  scale_y_continuous("Cumulatuve Curve",
-                     breaks = seq(0,450,50),
-                     limits = c(0,500)
-  ) +
-  scale_x_continuous("Change in SST (C)",
-                     breaks = seq(0,2,0.2),
-                     limits = c(0.3,2.3)
-  ) +
-  my_ggtheme_p() +
-  scale_colour_viridis(discrete = T) +
-  ggsave("./Figures/figure_temp.png",
-         width = 6,
-         height = 6,
-         units = "in"
-  )
-
-```
-
-```{r Scatter_plot,  eval = F, echo = T}
-
-data_stock <- Eme_Res %>% 
-  filter(
-    n_ens == 10,
-    emerging_yr >= 2020
-  ) %>% 
-  group_by(emerging_yr,eez_name) %>% 
-  summarise(
-    tax_n = length(unique(taxon_key))
-  ) %>% 
-  group_by(emerging_yr) %>% 
-  summarise(emerging_tax = sum(tax_n))
-
-data_eez <- Eme_Res %>% 
-  filter(
-    n_ens == 10,
-    emerging_yr >= 2020
-  ) %>% 
-  group_by(emerging_yr,taxon_key) %>% 
-  summarise(
-    eez_n = length(unique(eez_name))
-  ) %>% 
-  group_by(emerging_yr) %>% 
-  summarise(emerging_eez = sum(eez_n))
-
-data_stock %>% 
-  left_join(data_eez,
-            by = "emerging_yr") %>% 
-  rename(year = emerging_yr) %>% 
-  left_join(sst_data) %>% 
-  ggplot() +
-  geom_point(
-    aes(
-      x = mean_temp,
-      y = emerging_eez,
-      fill = year,
-      colour = year
-    )
-  )
-
-```
-
-# Old code
-
-## Fig. 4. Bubble plot
-
-```{r, buble_plot, eval = F, echo = F}
-
-# Needs previous chunk
-
-Buble_data <- Eez_d_hist %>% 
-  filter(metric == "mean_yr") %>% 
-  group_by(sau_fishing_entity,tresh) %>% 
-  summarise(mean_yr = mean(value)) %>% 
-  left_join(Eme_value_data,
-            by = c("sau_fishing_entity","tresh")) %>% 
-  filter(#tresh == selected_tresh, # tresh hold selection
-    variable == "value") %>% 
-  mutate(
-    weight_toe = sum_value/mean_yr,
-    sum_value_bi = sum_value/1000000000,
-    log_value = log10(sum_value_bi)
-  ) %>%   # Billion usd
-  left_join(Matching_names,
-            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
-  distinct(sau_fishing_entity,tresh,.keep_all = TRUE) %>%
-  filter(!is.na(un_geo_region)) %>% 
-  ungroup()# %>%  # no more asterisk
-# mutate(
-#   sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity))
-
-### Function for estimating percentile
-
-quibble <- function(x, q = p) {
-  tibble(x = quantile(x, q), q = q)
-}
-
-Percentile_value <- Buble_data %>% 
-  group_by(tresh) %>%  # for whole distribution
-  do(quibble(.$log_value, 0.75))
-
-## Top fishing nations close to SDGs
-Buble_data %>% 
-  filter(log_value > Percentile_value$x & mean_yr < 2036) %>% 
-  View()
-group_by(tresh) %>% 
-  summarise(sum(sum_value_bi),
-            sum(n_taxa))
-
-## All fishing nations close to SDGs
-Buble_data %>% 
-  filter(mean_yr < 2036) %>% 
-  # nrow() # 76  countries
-  View()
-group_by(tresh) %>% 
-  summarise(sum(sum_value_bi),
-            sum(n_taxa))
-
-76/length(unique(Buble_data$sau_fishing_entity))
-134/length(unique(Buble_data$sau_fishing_entity))
-
-# Labels Data
-
-Labels_data <- tibble(
-  x = c(2029,
-        # 2050,2070,
-        2007),
-  y = c(-5.2,
-        # -5.8,-5.8,
-        Percentile_value$x+.1),
-  label = c("UN 2030 Agenda", 
-            # "Mid shift","Late shift",
-            "Fisheries > 75th percentile")
-)
-
-# Plot the bubbles!
-ggplot(Buble_data) +
-  geom_point(
-    aes(
-      x = mean_yr,
-      y = log_value,
-      size = n_taxa,
-      colour = un_geo_region,
-      shape = un_geo_region
-    ),
-    alpha = 0.7
-  ) +
-  # Right side labels
-  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034),
-                   aes(
-                     x = mean_yr,
-                     y = log_value,
-                     label = sau_fishing_entity
-                   ),
-                   arrow = arrow(length = unit(0.009, "npc"),
-                                 type = "closed",
-                                 ends = "first"),
-                   force = 1,
-                   box.padding = 0.2,
-                   fill = "NA",
-                   nudge_x = 2066 - subset(Buble_data,log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034)$mean_yr, # set distance from point
-                   nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
-                   segment.size  = 0.1,
-                   segment.color = "grey50",
-                   direction     = "y",
-                   hjust         = 1
-  ) +
-  # LEFT side labels
-  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034),
-                   aes(
-                     x = mean_yr,
-                     y = log_value,
-                     label = sau_fishing_entity
-                   ),
-                   arrow = arrow(length = unit(0.009, "npc"),
-                                 type = "closed",
-                                 ends = "first"),
-                   force = 1,
-                   box.padding = 0.2,
-                   fill = "NA",
-                   nudge_x = 2025 - subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$mean_yr, # set horizontal position
-                   nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
-                   segment.size  = 0.1,
-                   segment.color = "grey50",
-                   direction     = "y",
-                   hjust         = 1
-  ) +
-  # Divisory lines and labels
-  geom_vline(xintercept = 2030, color = "grey50", size = 0.5, linetype = "dashed") +
-  geom_text(data = subset(Labels_data, x == 2029),
-            aes(
-              x = x,
-              y = y,
-              label = label
-            ),
-            size = 4,
-            color = "grey30",
-            angle = 90
-  ) +
-  geom_hline(data = Percentile_value,
-             aes(yintercept = x),
-             color = "grey50"
-  )  +
-  geom_text(data = subset(Labels_data, x == 2007),
-            aes(
-              x = x,
-              y = y,
-              label = label
-            ),
-            size = 4,
-            color = "grey30"
-  ) + 
-  scale_shape("Region") +
-  scale_color_viridis(name="Region",
-                      discrete = T) +
-  scale_x_continuous("Average year of range shift",
-                     limits = c(2000,2080),
-                     breaks = seq(2000,2080,10)
-  ) +
-  scale_y_continuous("Fishing revenue (log10)",
-                     limits = c(-6,2),
-                     breaks = seq(-6,2,2),
-                     expand = c(0,0)
-  ) +
-  scale_size("Number of\nshifting stocks") +
-  my_ggtheme_p(leg_pos = "right") +
-  ggsave("./Figures/bubble_plot.png",
-         width = 12,
-         height = 8,
-         units = "in"
-  )
-
-```
-
-```{r participation_text, eval = T, echo=F}
-
-Value_map_data %>% 
-  group_by(tresh,variable) %>% 
-  summarise(m = mean(participation)-2000,
-            sd = sd(participation-2000)
-  )
-
-```
-
-## Fig. 4N Top fishers on bars
-
-The bar option by UN sub region of Figure 4N of top fishers
-
-```{r}
-
-# Create data
-Species_Data <- top_toe %>% 
-  filter(tresh == selected_tresh,
-         variable == "value") %>% 
-  mutate(mean = mean-2000,
-         group = geo_region)
-
-# Set a number of 'empty bar' to add at the end of each group
-Char <- as.factor(Species_Data$group)
-empty_bar <- 5
-to_add <- data.frame( matrix(NA, empty_bar*nlevels(Char), ncol(Species_Data)) )
-colnames(to_add) <- colnames(Species_Data)
-to_add$group <- rep(levels(Char), each=empty_bar)
-Species_Data <- rbind(Species_Data, to_add)
-Species_Data <- Species_Data %>% arrange(group,mean)
-Species_Data$id <- seq(1, nrow(Species_Data))
-
-# Species_Data$id <- seq(1, nrow(Species_Data))
-
-# Get the name and the y position of each label
-label_data <- Species_Data 
-number_of_bar <- nrow(label_data)
-angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar
-label_data$hjust <- ifelse( angle < -90, 1, 0)
-label_data$angle <- ifelse(angle < -90, angle+180, angle)
-
-# prepare a data frame for base lines
-base_data <- Species_Data %>% 
-  group_by(group) %>% 
-  summarize(start=min(id), 
-            end=max(id) - empty_bar, 
-            median = median(mean, na.rm = T)
-  ) %>% 
-  rowwise() %>% 
-  mutate(title=mean(c(start, end)))
-
-# prepare a data frame for grid (scales)
-grid_data <- base_data
-grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
-grid_data$start <- grid_data$start - 1
-# grid_data <- grid_data[-1,]
-# grid_data[1,2] <- 5 # Manually include bathydemerssal
-# grid_data[1,3] <- 1
-
-un_regions <- grid_data %>% 
-  select(group)
-
-ggplot(Species_Data, 
-       aes(x=as.factor(id), 
-           y=mean,
-           fill=mean)
-) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
-  geom_bar(aes(x=as.factor(id),
-               y=mean,
-               fill=mean),
-           stat="identity",
-           alpha=0.5
-  ) +
-  geom_segment(data=grid_data,
-               aes(x = end, y = 30, xend = start, yend = 30),
-               colour = "grey90", alpha=1, size=0.3, inherit.aes = FALSE) +
-  geom_bar(aes(x=as.factor(id),
-               y=mean,
-               fill=mean
-  ),
-  stat="identity",
-  alpha=0.5) +
-  annotate("text", 
-           x = 0.5,
-           y = 37,
-           label = "UN Agenda",
-           color= "grey60", 
-           size=3, 
-           angle=-5
-  ) +
-  # Country names
-  geom_text(data=label_data,
-            aes(x=id, 
-                y= 100,
-                label=sau_fishing_entity,
-                hjust=hjust),
-            color="black",
-            fontface="italic",
-            alpha = 1,
-            size = 4 ,
-            angle = label_data$angle,
-            inherit.aes = FALSE) +
-  # Continents labels
-  # geom_text(data=base_data,
-  #           aes(x = title,
-  #               y = 85, 
-  #               label=group), 
-  #           # hjust=c(1,1,1,0,0,0),
-  #           colour = "black",
-  #           # angle = myAng,
-  #           size=5,
-  #           fontface="bold",
-  #           inherit.aes = FALSE) +
-ylim(-100,100) + # sets the circle the first value the central the last value how wide
-  theme_minimal() +
-  theme(
-    legend.position = c(0.5,0.5),
-    legend.text = element_text(size = 13),
-    legend.title = element_text(size = 14, face = "bold"),
-    legend.title.align = 0.5,
-    axis.text = element_blank(),
-    axis.title = element_blank(),
-    panel.grid = element_blank(),
-    panel.background = element_rect(colour = "white"),
-    legend.direction="horizontal",
-    legend.key.width = unit(3, 
-                            "line")
-  ) +
-  coord_polar(start = 0) +
-  scale_fill_viridis("Average year of range shift\nfor top 5 most valuable stocks",
-                     limits = c(0,100),
-                     breaks = seq(0,100,20),
-                     labels = seq(2000,2100,20),
-                     direction = -1
-  ) +
-  guides(
-    fill=guide_colorbar(title.position="top", 
-                        title.hjust =0.5)
-  )
-
-ggsave(
-  plot = last_plot(),
-  width = 13,
-  height = 13,
-  units = "in",
-  filename = "./Figures/top_stocks_circular_trend.png"
-)
-
-```

### 2021-11-24 Scripts/Supplements.Rmd
Included Figure S10 top

@@
 
+```{r toeFigureS10, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Changes in the stock share proportion of top 5 most valuable stocks by 2030 (2021-2040) relative to 1951-2005.
+Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).}
+
+img <- readPNG("../Figures/FigS10.png")
+grid.raster(img)
+
+```
+
 # Tables

### 2021-11-24 Scripts/Manuscript.Rmd
Misc

@@
 
-**Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 65 $\pm$ 21% by 2030 and 78 $\pm$ 18% by 2050 in 77% of the world's EEZs (Figure SX).
+**Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 65 $\pm$ 21% by 2030 and 78 $\pm$ 18% by 2050 in 77% of the world's EEZs (Fig.
+\@ref(fig:toeFigureS10)).
 In some cases, the SSR of one EEZ is projected to increase more than double relative to their neighbors (e.g., from 10% to 30% SSR).

### 2021-11-25 Scripts/Initial_Analysis.Rmd
Updated stats results with tresh_two

@@
 tresh_list <- c("tresh_one","tresh_two")
-for(i in 1:2){
+i = 2
+for(i in 2){
 
@@
 
+  ggsave("mean_boxplot.png",boxplot,width = 10)
+
   ## --------------- ##
@@
   ) %>% 
+  # View()
   filter(mean_yr <= 2021) %>% 
   group_by(variable,tresh) %>% 
+  # View()
   summarise(
@@
   ) %>% 
-  mutate(per = n_countries/length(unique(Sau_data$fishing_entity_id)))
+  mutate(per = n_countries/length(unique(Sau_data$fishing_entity_id))*100)
 
@@
   ) %>% 
-  filter(mean_yr <= 2021) %>% 
-  group_by(geo_region) %>% 
+  filter(mean_yr <= 2021,
+         variable == "value",
+         tresh == "tresh_two",
+         !is.na(geo_region)) %>% 
+  group_by(geo_region,tresh) %>% 
   tally() %>% 
-  arrange(n) %>% 
-  top_frac(0.25,n) # top 25%
+  arrange(desc(n)) %>% 
+  # pull(n) %>% 
+  # sum()
+
+139/263
 
@@
 qqnorm(Stats_data$emerging_yr) # not normally distributed
-skewness(Stats_data$emerging_yr) #1.02
-kurtosis(Stats_data$emerging_yr) #2.82
+skewness(Stats_data$emerging_yr) #0.61
+kurtosis(Stats_data$emerging_yr) #2.11
 
@@
 # data:  un_sub_region by emerging_yr
-# Kruskal-Wallis chi-squared = 287.23, df = 93, p-value < 2.2e-16
+# Kruskal-Wallis chi-squared = 242.11, df = 93, p-value = 3.128e-15
 
@@
   # head()
-  View()
+  # View()
 write_csv("kw_regional.csv")

### 2021-11-25 Scripts/Manuscript.Rmd
Updated stats results with tresh_two

@@
 Continuing climate change is expected to exacerbate such patterns [@Pinsky:2018cb; @Sumaila:2011bq].
-Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the goals set out under the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].
+**Recent studies have looked at global gains (Pinsky et.
+al., 2018) and losses (Oremus et al., 2020) of transboundary stocks within the world’s EEZs as well as regional sharing dynamics between neighbouring countries (Palacios-Abrantes et al., 2020, Sumaila et al., 2020) towards the end of the century because of climate change.
+However, previous studies fail to identify changes in the distribution of current transboundary stocks at the global level including the actual timeline of shifts in shared stocks, nor the intensity of such shifts.** Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the goals set out under the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].
 
 Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods).
-**We define a ‘stock’ unit as a species shared by neighboring EEZs.** Furthermore, we  developed a Transboundary Shift Index to evaluate range shifts in the shared distribution of transboundary stocks.
+**We adopt the definition of a ‘stock’ unit developed and applied in Palacios-Abrantes *et al.* 2020; Teh & Sumaila, 2015, which considers a stock unit as a species shared by neighbouring EEZs.** Furthermore, we  developed a Transboundary Shift Index to evaluate range shifts in the shared distribution of transboundary stocks.
 The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig.
 \@ref(fig:toeFigure1N)).
-By employing the concept of *time of emergence*, we estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig.
-1).
-Such estimation is based on two thresholds representing a 67 and a 95% probability of shift.
+**We base our analysis on the 633 transboundary species identified by Palacios-Abrantes (2020) defining a ‘stock’ unit as each time a transboundary species is shared by neighbouring EEZs.
+By employing the concept of time of emergence, we estimate the year in which the stocks are projected to shift their shared distribution beyond historical variations (Fig.
+\@ref(fig:toeFigure1N)).** Such estimation is based on two thresholds representing a 67 and a 95% probability of shift.
 Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (Methods).
 This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp].
-We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).
+**We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).**
 
@@
 The analyses detailed herein are based on 633 exploited marine transboundary species that account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 [@PalaciosAbrantes:2020tv].
-**We define a (transboundary) "stock" unit when a species was shared by a pair of neighboring EEZs [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.** Using the transboundary species Atlantic cod (*Gadus morhua*) as an example and this study’s definition of transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not as they do not have juxtaposing EEZs.
-We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015; http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
+**We define a (transboundary) "stock" unit when a species was shared by a pair of neighboring EEZs [@PalaciosAbrantes:2020tv; @Teh:2015gd], resulting in a total of 9,132 transboundary stocks.** **Let us consider the transboundary species Atlantic cod (Gadus morhua) as an example.
+Atlantic cod’s distribution in the Northwest Atlantic ranges from the United States to Greenland.
+Based on the previous definition of a transboundary stock, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share a separate stock of Atlantic cod, but the United States and Greenland do not share a stock as they do not have juxtaposing EEZs.** We defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015; http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig.
 \@ref(fig:toeFigureS1)) including island territories.
@@
 
-**While the total number of shifting stocks is important for fisheries management, having high-valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting.
-Thus, we looked at the year of range shift of the top five most valuable transboundary stocks of each country (New Fig 4).
+**While the total number of shifting stocks is important for fisheries management, having more valuable stocks shifting sooner could be a much bigger concern than having many low-value stocks shifting.
+Thus, we looked at the year of range shift of the top five most valuable transboundary stocks of each country (Fig.
+\@ref(fig:toeFigure3)).
 On average, countries are projected to see a range shift of the top five most valuable stocks by 2038 $\pm$ 25 years under threshold two (2028 $\pm$ 22 years under threshold one).
-Moreover, our projections suggest that these high value stocks are already shifting in 52 (27%) coastal nations under threshold two (n = 85 44% under threshold one), with the regions of Latin America and the Caribbean (n = 63), Sub-Saharan Africa (n = 59), South-eastern Asia  (n = 28), and Polynesia (n = 19) presenting the larger number of shifting stocks (25% of total high value stocks).**
+Moreover, our projections suggest that these high value stocks are already shifting in 129 (66%) coastal nations under threshold two (n = 154, 79% under threshold one), with the regions of Latin America and the Caribbean (n = 62), Sub-Saharan Africa (n = 36), Polynesia (n = 22), and Southern Europe (n = 19) presenting the larger number of high value shifting stocks (50% of shifting high value stocks).**
 
@@
 While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca].
-Moreover, not limiting the species to, at least, an EEZ could result in an overestimation of our results (NOTE: See Ovando *et al.* https://www.pnas.org/content/118/23/e2025958118).** Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identify different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
+Moreover, the population structure of most exploited species in the world are poorly- or un- defined [@ref].
+Future studies that focus on particular shared stocks could use more specific unit stock definition that is available for the cases.** Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results and potentially identify different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie].
 Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.

### 2021-11-25 Scripts/Initial_Analysis.Rmd
Showing median in figure 3, not mean

@@
 tresh_list <- c("tresh_one","tresh_two")
-i = 2
-for(i in 2){
+
+for(i in 1:2){
 
@@
         x = reorder(un_sub_region,order),
-        y = mean,
-        colour = mean
+        y = med,
+        colour = med
       ),
@@
 
-  ggsave("mean_boxplot.png",boxplot,width = 10)
+  # ggsave("mean_boxplot.png",boxplot,width = 10)
 

### 2021-11-25 Scripts/Initial_Analysis.Rmd
Nvm.. Back to showing the mean

--- 2021-11-25 Scripts/Initial_Analysis.Rmd
@@
     group_by(un_sub_region) %>%
-    summarise(order = median(emerging_yr),
+    summarise(order = mean(emerging_yr),
               n_eezs = length(unique(sf_eez_name))) %>%
@@
 }
+
+```
+
+###### Lat Am Figure 3 for supplements
+
+```{r}
+
+World_map_part <- World_map_nemer %>% 
+    filter(
+      variable == "value"
+    )
+
+  # ToE Map 
+  toe_map_data <- World_eez_sf %>% 
+    left_join(Eez_d_hist,
+              by = "eezid") %>% 
+    filter(metric == "mean_yr")%>% 
+    sf::st_transform(crs = "+proj=eck4")
+
+  no_toe_eez <- World_eez_sf %>% 
+    filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
+    sf::st_transform(crs = "+proj=eck4")
+
+  no_toe_land <- World_map %>% 
+    filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !name %in% World_map_part$name)%>% 
+    sf::st_transform(crs = "+proj=eck4")
+
+lat_ammap <- ggplot(toe_map_data) +
+    geom_sf(
+      aes(
+        fill = reverse #original map
+        # fill = value
+      ),
+      color =  "white",
+      size = 0.1
+    ) +
+    # Inlcude the participation of species
+    geom_sf(data = World_map_part,
+            # aes(fill = participation+2000), #original map
+            aes(fill = participation, 
+                color = participation),
+            # color =  "white",
+            size = 0.1
+    ) +
+    # grey maps
+    # ggplot() +
+    geom_sf(data = grey_land, aes(), fill = "grey50", 
+            # color =  "white",
+            size = 0.1) +
+    # geom_sf(data = grey_eez, aes(), fill = "white", color =  "grey50",size = 0.1) +
+    geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+  coord_sf(xlim = c(-9000000,-5000000),ylim = c(1000000, 4000000)) + 
+  facet_wrap(~tresh) +
+  scale_colour_viridis("",
+                         limits = c(2000,2100),
+                         breaks = seq(2000,2100,25),
+                         # labels = c("","","","",""),
+                         direction = 1
+    ) +
+    scale_fill_viridis("",
+                         limits = c(2000,2100),
+                         breaks = seq(2000,2100,25),
+                         # labels = c("","","","",""),
+                         direction = 1
+    ) +
+    my_ggtheme_m(map_type = "regional",leg_pos = "")
+
+ggsave("lat_am_map.png",
+       lat_ammap,
+       height = 5,
+       width = 6)
+
+```
+
+###### SP Islands Figure 3 for supplements
+
+```{r}
+
+  # Value Map
+  spi_map_part <- World_map_nemer %>% 
+    filter(
+      # tresh ==  tresh_list[i],
+      variable == "value"
+    ) %>% 
+  sf::st_transform(crs = 3832)
+
+  # ToE Map 
+  spi_toe_map_data <- World_eez_sf %>% 
+    left_join(Eez_d_hist,
+              by = "eezid") %>% 
+    filter(#tresh == tresh_list[i],
+           metric == "mean_yr")%>% 
+    sf::st_transform(crs = 3832)
+
+  # Map with countries without ToE
+  spi_no_toe_eez <- World_eez_sf %>% 
+    filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
+    sf::st_transform(crs = 3832)
+
+  spi_no_toe_land <- World_map %>% 
+    filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
+           !name %in% World_map_part$name)%>% 
+    sf::st_transform(crs = 3832) # Filter EEZs with ToE stocks
+
+  sp_map <-  ggplot(spi_toe_map_data) +
+    geom_sf(
+      aes(
+        fill = reverse #original map
+        # fill = value
+      ),
+      color =  "white",
+      size = 0.1
+    ) +
+    # Inlcude the participation of species
+    geom_sf(data = spi_map_part,
+            # aes(fill = participation+2000), #original map
+            aes(fill = participation, 
+                color = participation),
+            # color =  "white",
+            size = 0.1
+    ) +
+    # grey maps
+    geom_sf(data = spi_no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    geom_sf(data = spi_no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
+    coord_sf(xlim = c(9500000,1000000),ylim = c(-3500000, 1900000)) + 
+    facet_wrap(~tresh) +
+    scale_colour_viridis("",
+                         limits = c(2000,2100),
+                         breaks = seq(2000,2100,25),
+                         # labels = c("","","","",""),
+                         direction = 1
+    ) +
+    scale_fill_viridis("",
+                         limits = c(2000,2100),
+                         breaks = seq(2000,2100,25),
+                         # labels = c("","","","",""),
+                         direction = 1
+    ) +
+    my_ggtheme_m(map_type = "regional",leg_pos = "bottom") +
+    theme(legend.key.width = unit(3, "line"))
 

### 2021-11-25 Scripts/Initial_Analysis.Rmd
Updated manuscript and analysis to inlcude zooms to latin america and south pacific islands as supplements

@@
 
-lat_ammap <- ggplot(toe_map_data) +
+lat_am_map <- ggplot(toe_map_data) +
     geom_sf(
@@
 ggsave("lat_am_map.png",
-       lat_ammap,
-       height = 5,
-       width = 6)
+       lat_am_map,
+       width = 8,
+       height = 6)
 
@@
 
-  sp_map <-  ggplot(spi_toe_map_data) +
+spi_map <-  ggplot(spi_toe_map_data) +
   geom_sf(
@@
   ) +
+  scale_x_continuous(limits = c(-120,180),
+                     breaks = c(-120,180)
+                     # labels =c("a","b")
+                     ) + 
   my_ggtheme_m(map_type = "regional",leg_pos = "bottom") +
@@
 
+ggsave("spi_map.png",
+       spi_map,
+       width = 8,
+       height = 6)
 ```

### 2021-11-25 Scripts/Manuscript.Rmd
Updated manuscript and analysis to inlcude zooms to latin america and south pacific islands as supplements

@@
 2A).
-Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2) and \@ref(tab:toeTableS3) than almost any other region (Figs.
+Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean, Melanesia and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2) and \@ref(tab:toeTableS3) than almost any other region (Figs.
 \@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)).
@@
 EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
+See Fig.
+S11 for zoom into the Caribbean and South Pacific Islands.
 B) Year of shared distribution shifts by UN sub-regions.

### 2021-11-25 Scripts/Initial_Analysis.Rmd
misc

@@
           # color =  "white",
-          size = 0.1
+          size = 1
   ) +
@@
 
-ggsave("spi_map.png",
+ggsave(".Figures/spi_map_bf.png",
        spi_map,

### 2021-11-29 Scripts/Supplements.Rmd
Updates supplemental figures

@@
 
+```{r toeFigureS5, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Year of change in the shared distribution of transboundary stocks in the Caribbean and South Pacific Islands.
+Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygons display average year of range shifts within them.
+Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cold colors represent a late shift/low fishing revenue.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue."}
+
+img <- readPNG("../Figures/FigS11.png")
+grid.raster(img)
+
+```
+
+\clearpage
+
 ```{r toeFigureS6, eval = T,  echo = F, fig.width=9, fig.height=6,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
@@
 Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
-Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors.}
+Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors."}
 
@@
 ```{r toeFigureS10, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Changes in the stock share proportion of top 5 most valuable stocks by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).}
+Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead)."}
 

### 2021-11-30 Scripts/Initial_Analysis.Rmd
Included dummy data for F1 toe

@@
 ggsave(map,filename="/Users/william/Google Drive/Projects/MHWs Global/MHW_SST_plot_map.png",height=8,width=11,dpi=300) (edited)
+
+```
+
+## Figures for F1
+
+This is a hypothetical figure to show the idea behind the ToE. It does not follow the actual method but only shows the trend.
+
+```{r tsi_figure, eval = F}
+
+# Create dummy data
+dummy_data <- data.frame(
+  year = seq(1951,2100,1),
+  tsi = c(runif(80,-10,10),
+          runif(20,-10,4),
+          runif(30,-12,0),
+          runif(20,-14,-4)),
+  tsia = c(runif(80,-10,10),
+          runif(20,-10,4),
+          runif(30,-12,0),
+          runif(20,-14,-4)),
+  tsib = c(runif(80,-10,10),
+           runif(20,-10,4),
+           runif(30,-12,0),
+           runif(20,-14,-4)),
+  tsic = c(runif(80,-10,10),
+           runif(20,-10,4),
+           runif(30,-12,0),
+           runif(20,-14,-4))
+) %>% 
+    rowwise() %>% 
+    mutate(
+      mean_tsi = mean(c(tsi,tsia,tsib,tsic),na.rm=T)
+    ) %>% 
+    ungroup() %>% 
+  group_by() %>% 
+  mutate_all(funs(zoo::rollmean(., 
+                          10, 
+                          align = "left", 
+                          fill = NA,
+                          na.rm=T)
+  )) %>% 
+  gather("eme","value",tsi:tsic)
+
+# Get min, max and toe for box/noise
+min <- dummy_data %>% 
+  filter(year < 2006) %>% 
+  pull(value) %>% 
+  min()
+
+max <- dummy_data %>% 
+  filter(year < 2006) %>% 
+  pull(value) %>% 
+  max()
+
+# ToE
+toe <- dummy_data %>% 
+  filter(year > 2005 & mean_tsi < min+0.5) %>% 
+  slice(1)
+
+# Make hypotetical figure
+ggplot(dummy_data) +
+  # Historical variation box (noise)
+  geom_rect(
+    aes(xmin=1951,
+        xmax=2100,
+        ymin=min+0.5,
+        ymax=max-1),
+    fill="grey80",
+    alpha = 0.5
+  ) +
+  # Historical mean
+  geom_rect(
+    aes(xmin=1951,
+            xmax=2100,
+            ymin = mean(c(min,max)),
+            ymax = mean(c(min,max)),
+            ),
+    color = "grey60",
+    size = 0.1
+    ) +
+  # Each ensemle
+  geom_line(
+    aes(
+      x = year,
+      y = value,
+      fill = eme
+    ),
+    alpha = 0.8,
+    color = "#F8AFA8"
+  ) +
+  # Average of ensembles (trend)
+  geom_line(
+    aes(
+      x = year,
+      y = mean_tsi
+    ),
+    color = "black"
+  ) +
+  # Mark ToE
+  geom_point( data = toe,
+    aes(
+      x = year,
+      y = mean_tsi
+    ),
+    color = "#FF0000",
+    size = 5,
+    alpha = 0.8
+  ) +
+  # Historical / Future division line
+  geom_vline(xintercept = 2005, color = "#046C9A") +
+  # Aesthetics
+  scale_color_brewer(palette = "Oranges") +
+  theme_classic()+
+  theme(axis.text = element_blank(),
+        axis.ticks = element_blank())+
+  labs(x = "Transboundary shift index (TSI)",
+       y = "Year")
+
+# save image
+ggsave(
+  filename = "./Figures/tsi_fig.png",
+  plot = last_plot(),
+  width = 8,
+  height = 4
+)
 

### 2021-11-30 Scripts/Initial_Analysis.Rmd
Updated figures S

@@
   coord_sf(xlim = c(-9000000,-5000000),ylim = c(1000000, 4000000)) + 
-  facet_wrap(~tresh) +
+  facet_wrap(~tresh, ncol = 1) +
   scale_colour_viridis("",
@@
 
-ggsave("lat_am_map.png",
+ggsave("Figures/lat_am_map.png",
        lat_am_map,
-       width = 8,
-       height = 6)
+       width = 6,
+       height = 8)
 
@@
   coord_sf(xlim = c(9500000,1000000),ylim = c(-3500000, 1900000)) + 
-  facet_wrap(~tresh) +
+  facet_wrap(~tresh, ncol = 1) +
   scale_colour_viridis("",
@@
 
-ggsave(".Figures/spi_map_bf.png",
+ggsave("Figures/spi_map_bf.png",
        spi_map,
-       width = 8,
-       height = 6)
+       width = 6,
+       height = 8)
 ```
@@
 ```{r ssr_map_data, eval = T, echo = T}
-# A function that makes a dateframe per connection (we will use these connections to plot each lines)
+# A function that makes a dataframe per connection (we will use these connections to plot each lines)
 data_for_connection=function( dep_lon, dep_lat, arr_lon, arr_lat, group){
@@
 # SELECT A TIME STEP
-# time_toplot <- "per_mid"
-time_toplot <- "per_early"
+time_toplot <- "per_mid"
+# time_toplot <- "per_early"
 ```
@@
 lat_am_map <- ggplot() +
-  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
+  # PLot EEZs
+  geom_sf(data = subset(World_eez_sf,
+                        !str_detect(eez_name,"Pacific") &
+                        !str_detect(eez_name,"El Salvador") &
+                        !eez_name %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil")
+  ), size = 0.1, fill = "white", col = "gray30") +
+  # Plot Land information
   geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
@@
           col = "white") +
-  # geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
-  geom_sf(data = subset(sf_ready_plot, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
-  geom_sf(data = point_data, aes(), color = "grey90") +
-  geom_path(data = subset(arrows_data, percentage >= 0),
+  geom_sf(data = subset(sf_ready_plot, percentage >= 0 & 
+                          !str_detect(source,"Pacific") & 
+                        !source %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador") &
+                        !str_detect(objective,"Pacific") &
+                        !objective %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador")
+                        ), aes(color = percentage), size = 0.1, fill = "white") +
+  geom_sf(data = subset(point_data,
+                        !str_detect(source,"Pacific") & 
+                        !source %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador") &
+                        !str_detect(target,"Pacific") &
+                        !target %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador")
+                        ),
+          aes(), color = "grey90") +
+  geom_path(data = subset(arrows_data, percentage >= 0 &
+                            !str_detect(objective,"Pacific") &
+                        !objective %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador")
+                        ),
             aes(X, Y,
@@
   # scale_y_continuous(breaks = seq(10,30,10)) +
-  my_ggtheme_m(leg_pos = "bottom") +
+  my_ggtheme_m(map_type = "NA",leg_pos = "bottom") +
   theme(
@@
                         colours = pal,
-                        limits = c(0,max(Lat_am_data$percentage)),
-                        breaks = seq(0,max(Lat_am_data$percentage),10)
+                        limits = c(0,50),
+                        breaks = seq(0,50,10)
   ) +
@@
                        colours = pal,
-                       limits = c(0,max(Lat_am_data$percentage)),
-                       breaks = seq(0,max(Lat_am_data$percentage),10)
+                       limits = c(0,50),
+                        breaks = seq(0,50,10)
   ) + guides(
@@
 
-ggsave(filename = "lat_am_mid.png",
+name <- paste0("Figures/lat_am_",time_toplot,".png")
+
+ggsave(filename = name,
        lat_am_map,
@@
 
+# Name to save
+name_eu <- paste0("Figures/EU_",time_toplot,".png")
+
+# PLot EU map
 EU_map <- ggplot() +
@@
   scale_x_continuous(breaks = seq(-100,60,10)) +
-  scale_y_continuous(breaks = seq(10,30,10)) +
-  my_ggtheme_m(leg_pos = "bottom") +
+  # scale_y_continuous(breaks = seq(10,30,10)) +
+  my_ggtheme_m(map_type = "NA",leg_pos = "bottom") +
   theme(
@@
                         colours = pal,
-                        limits = c(0,max(EU_data$percentage)),
-                        breaks = seq(0,max(EU_data$percentage),10)
+                        limits = c(0,50),
+                        breaks = seq(0,50,10)
   ) +
@@
                        colours = pal,
-                       limits = c(0,max(EU_data$percentage)),
-                       breaks = seq(0,max(EU_data$percentage),10)
+                       limits = c(0,50),
+                        breaks = seq(0,50,10)
   ) + guides(
@@
 
-ggsave(filename = "EU_early.png",
+# Save map
+ggsave(filename = name_eu,
        EU_map,

### 2021-12-01 Scripts/responses_rev_gcb.Rmd
misc.

@@
 output:
-  word_document: default
+  # word_document: default
   pdf_document: default
@@
 
-- **Response: PLEASE PAY SPECIAL ATTENTION TO THIS POINT.
-LOOKS LIKE WE REALLY NEED TO CONVINCE THE EDITOR OUR RESEARCH IS NOVEL** Below you will find the response to both Reviewers 1 and 2.
+- **Response**: Below you will find the response to both Reviewers 1 and 2.
 In addition, we would like to offer here an explanation of how our paper differs from recent published papers, including those authored by some of us.
 
-  In their 2018 paper, Pinsky *et al.* look at how many *new* transboundary species each EEZ of the world will have towards the end of the century under climate change (i.e., national stocks that become transboundary stocks).
-Indeed, the authors find that “many of the world’s EEZs are likely to receive one to 12 new, climate-driven transboundary stocks by the end of the century.” On the other hand, Oremus *et al.* (2020) “estimates the number of species *exiting* national jurisdictions under climate change up to 2100” concluding that tropical nations stand to lose the most species under these scenarios.
-However, both studies fail to identify changes in *current* transboundary stocks including the intensity of such shift, gap that we look to fill in this paper.
-Here, we build from Palacios-Abrantes *et al.* (2020) and use the transboundary species (n = 633) identified in such paper.
-This means that, unlike Pinsky *et al.* (2018) and Oremus *et al.* (2020) who look at national stocks shifting towards new or away from EEZs, our paper focuses on transboundary stocks shifting within EEZs.
-Moreover, as far as we are concerned no other study, including the research done by us, identifies the actual timeline of shifts in shared stocks, nor the intensity of such shifts.
+  We understand the Editor’s concern as all of these papers look at the consequences of climate change for shared stocks.
+We have stated the differences between our study and those previously made in the introduction (see below).
+In short, our manuscript critically builds on the previous studies highlighted, using the species identified by Palacios-Abrantes et al.
+(2020) *Scientific Reports* and combining methods developed in Sumaila et al.
+(2020) and Palacios-Abrantes et al.
+(2020) *Ecology and Society* to address the questions of what is the timeline and intensity of changes in the distribution of current transboundary stocks at the global level?
+A question that neither Pinsky et al.
+(2018) *Science* nor Oremus et al.
+(2020) *Nature Sust*.
+addressed.
+Please note that we offer a longer explanation of these differences after the in-text statements.
+We hope to have clarified the novelty of our research and how it differs from previous studies.
+The new information that we present here can be critical for fisheries management as it offers tangible reference points at country level around which to consider future climate change impacts and to inform discussions, for instance, around agreements that are relevant to shared stocks.
 
-  To identify the timeline we adopt the concept of *Time of Emergence* commonly used in oceanic physical and chemical studies (See Rodgers *et al.* 2015, Keller *et al.* 2014 and has just recently been used in ecological studies by Cheung *et al.* 2020 looking at ecological responses to marine heatwaves.
-As far as we are concerned, no other study has adopted this methodology to look at the timing of shifts in transboundary fisheries.
-On the other hand, to identify the intensity of the shift, we (partially) used methods developed by Palacios *et al.* and Sumaila *et al* (2020), published in *Ecology and society*.
+   - Line XX, introduction:  *Recent studies have looked at global gains (Pinsky et.
+al., 2018) and losses (Oremus et al., 2020) of transboundary stocks within the world’s EEZs as well as regional sharing dynamics between neighbouring countries (Palacios-Abrantes et al., 2020, Sumaila et al., 2020) by the end of the century due to climate change.
+However, previous studies failed to identify both changes in the distribution of current transboundary stocks at the global level including the actual timeline of shifts in shared stocks and the intensity of such shifts.*
+
+  - Line XX, introduction: *We base our analysis on the 633 transboundary species identified by Palacios-Abrantes (2020), defining a ‘stock’ unit as each time a transboundary species is shared by neighboring EEZs.*
+
+  In their 2018 paper, Pinsky et al.
+look at how many new transboundary species will enter each EEZ of the world towards the end of the century under climate change (i.e., national stocks that become transboundary stocks).
+The authors find that “many of the world’s EEZs are likely to receive one to 12 new, climate-driven transboundary stocks by the end of the century.” On the other hand, Oremus et al.
+(2020) “estimate the number of species exiting national jurisdictions under climate change up to 2100” concluding that tropical nations stand to lose the most species.
+Both studies do not identify changes in current transboundary stocks, including the intensity of such shifts, an important knowledge gap that we aim to close in this paper.
+Here, we build from Palacios-Abrantes et al.
+(2020) and use the existing no.
+of transboundary species (n = 633) identified in that earlier paper.
+This means that, unlike Pinsky et al.
+(2018) and Oremus et al.
+(2020) that focus their study on the gaining of new stocks in a EEZ from expansion of distribution of national stocks from other EEZs, or losses of existing stocks from range contraction,  our paper examines the shifts of transboundary stocks  between EEZs that are already sharing the stocks now.
+These different types of shifts in transboundary stocks are causing different types of impacts on fisheries and would require different sets of adaptation responses [@Pinsky:2014by].
+We also identify the timeline of shifts in shared stocks and the intensity of such shifts, which is not explored in previous studies.
+
+To identify the timeline of shifts we adopt the concept of “Time of Emergence” commonly used in oceanic physical and chemical studies (see for example in Rodgers et al.
+2015 or Keller et al.
+2014), and has just recently been used in ecological studies by Cheung et al.
+-[@Cheung:2020ja] looking at ecological responses to marine heatwaves and Trisos et al [-@trisos_projected_2020] looking at the timing of species’ exposure to potentially dangerous climate conditions.
+The “Time of Emergence” is generally defined in the literature as the time when the increase in intensity of the signal exceeds the noise (Bindoff et al.
+2019).
+To our knowledge, no other study has adopted this methodology to look at the time of emergence of shifts in transboundary fisheries.
+Moreover, to identify the intensity of the shift, we (partially) used methods developed by Palacios et al.
+(2020) Scientific Reports and Sumaila et al (2020), published in Ecology and society.
 These papers adopt the concept of threat point to evaluate the resilience of specific treaties celebrated by Canada-US transboundary to manage four transboundary stocks.
-Here, we take this idea of threat point and expand the analysis globally to estimate the intensity that the shifts in transboundary stocks will have.
-
-  In short, our manuscript builds upon Palacios-Abrantes *et al.* 2020 *Scientific Reports* combining methods developed in Sumaila *et al.* 2020 and Palacios-Abrantes *et al.* 2020 *Ecology and Society* to address a question that neither Pinsky *et al.* 2018 *Science* nor Oremus *et al.* (2020) *Nature Sust.* addressed.
-We hope to have clarified the novelty of our research and how it differs from previous studies.
-The new information that we present here can be critical for fisheries management to adapt to shifts in the distribution of transboundary stocks under climate change as highlighted by both reviewers.
+Here, we further develop this method to apply the concept of threat point to the Stock Share Ratio estimation while expanding the analysis globally to estimate the intensity that the shifts in transboundary stocks will have as well as their time of emergence.
 
@@
 
-- **Response:** We are glad that the reviewer found our manuscript to be interesting.
+- **Response:** We are glad that the reviewer found our manuscript interesting.
 We have revised the manuscript paying special attention to clarifying the parts the reviewer flagged.
 We thank the reviewer for his/her comments as they have strengthened our paper.
-Please find below direct responses to the reviewer's comments.
+Please find below direct responses to the reviewer’s comments.
 
@@
 
-- **Response:** We acknowledged that this was not fully clear in the previous version of the manuscript and have work towards improving clarity in the revise manuscript.
-For example, pacific sardine is a species present along the coast of Mexico, the United States (US) and, sometimes, British Columbia (BC) [@cisneros-montemayor_ecosystem-based_2020].
+- **Response:** We acknowledged that this was not fully clear in the previous version of the manuscript.
+We have improved the clarity of the definition of the stock unit in the revised manuscript.
+Specifically, in considering “neighboring stocks of the same species”, for example, Pacific sardine is a species that is present along the coast of Mexico, the United States (US) and, sometimes, British Columbia (BC) [@cisneros-montemayor_ecosystem-based_2020].
 For the purpose of our study, we define two stocks: One that is shared by Mexico and the US and one shared by the US and BC.
-We have made this explicit in the following section of the methods, hopefully clarifying our definition;
+We have made this explicit in the following section of the methods to clarify our definition;
 
@@
 
-- **Response:** Our definition will influence the calculation of SSR *between* neighboring EEZs (i.e., Mexico and the US SSR for sardine and US and BC SSR for sardine, independently one of another).
-As the reviewer suggests, in this scenario Mexico could be loosing sardine SSR to the US contiguous states (this can also be seen as the US gaining SSR from Mexico), while the US contiguous states looses sardine SSR to BC (again, this could also be seen as Canada winning SSR from the US contiguous states).
-This effect is appreciable in the average SSR gains of Figure 5 where Mexico losses to the US contiguous states, who in turn looses to BC, who actually gains from both US continuous states and Alaska.
-This is now explained on the methods as follows **NOT SURE WE NEED THIS**;
+- **Response:** Our definition will influence the calculation of SSR between neighboring EEZs (i.e., Mexico and the US SSR for sardine and US and BC SSR for sardine, independently one of another).
+As the reviewer suggests, in this scenario Mexico could be losing sardine SSR to the US contiguous states (this can also be seen as the US gaining SSR from Mexico), while the US contiguous states loses sardine SSR to BC (again, this could also be seen as Canada winning SSR from the US contiguous states).
+This is now explained on the methods as follows:
 
@@
 
-  However, even when a stock of non highly-migratory species does span along the coast of more than two countries, our results seem to be similar (this relates to the part regarding Mexico US and Canada).
-For example, Pacific halibut (*Hippoglossus stenolepis*) is considered to be one single stock that extends along the Pacific coast of the US contiguous states, Canada, and Alaska [@iphc_iphc_2016_report_2016].
-Palacios-Abrantes *et al.* (2020) analyzed the shift in SSR between the three parties and found that the US contiguous states will likely loose SSR to Canada while Canada will loose to Alaska.
-A similar patter shown in the present study.
-
-    Ideally, one could define a stock as a biologically-defined population.
-This could be the case of Peruvian anchoveta *(Engraulis ringens)* off the coast of Chile and Peru.
-There are two identified stocks of anchoveta; one in the central/norther part of the Peruvian EEZ and one in the south, shared with Chile.
-This differentiation will likely impact both the ToE and SSR results.
-However, as far as we are concern, such analysis is challenged by a lack of stock-level data around the world.
-
 The above are just example situations.
@@
 
-  - **Response:** We understand the reviewer's concern and hope that our responses have clarified this issue.
-However, it is almost certain that our definition of stock is influencing the results of our paper, just by that nature of the analysis.
-While a more fine scale definition (e.g., biologically-defined populations of stock will probably arose other results, not limiting the species to, at least, an EEZ could result in an overestimation of our results (See Ovando *et al.* https://www.pnas.org/content/118/23/e2025958118).
-This is a main caveat of our study and have made an effort to be very clear and upfront about it in line XX of the revised manuscript:
+  - **Response:** We understand the reviewer’s concern and hope that our responses have clarified our choice of the definition.
+In addition, in the revised manuscript, we have explained the implications of our choice of the definition of stock unit on our findings and conclusions.
+(line XX of the revised manuscript):
 
-  - - **Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks, rather than biologically-defined populations.
+    - **Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks, rather than biologically-defined populations.
 While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca].
@@
 - **Response:** We agree with the reviewer that an un-weighted centroid can severely skew the actual distribution of the species and influence our results.
-We have revised the manuscript to clarify that we did account for the heterogeneous distribution of fish stocks across their entire range.
+We clarified in the manuscript that we did account for the heterogeneous distribution of fish stocks across their entire range.
 Specifically, the geographic centroid of each stock is indeed weighted by the abundance of the stock.
@@
 The centroid of a transboundary stock was determined by the average latitude and longitude ($lon_{ts}$) of the grid cells holding the specified percentile  ($per^{th}$)  of the projected transboundary stock abundance (see below).
-This is, we determined the centroid of the stock using only those regions where abundance was higher (i.e., using the probability of occurrence of the DBEM).
-Moreover, we estimated the sensitivity of our results to the (arbitrarily) $per^{th} = 95^{th}$ percentile threshold as described in lines XX and shown in Figure S6 of the supplements.
+Thus, we determined the centroid of the stock using only those regions where abundance was higher (i.e., using the probability of occurrence of the DBEM).
+Moreover, we estimated the sensitivity of our results to the (arbitrarily) percentile threshold $per^{th} = 95^{th}$ as described in lines XX and shown in Figure S6 of the supplements.
 It reads as follows in line XX:

### 2021-12-01 Scripts/Supplements.Rmd
Created equation table

@@
 
+```{r}
+library(tidyverse)
+library(knitr)
+
+data.frame(
+   "Acronym" = c(
+     "Hs",
+     "_i_",
+     "_T_",
+     "TPP",
+     "B$\\alpha$~i~",
+     "Min~D~",
+     "Max~D~",
+     "Hab~ij~",
+     "_u_",
+     "Hab~Assoc~",
+     "Sal",
+     "Sal~Assoc~",
+     "Ice",
+     "Ice~p~",
+     "A~i~",
+     "_G_",
+     "L~ji~",
+     "Ad~ji~",
+     "_j_",
+     "_r_",
+     "_K_",
+     "_f_",
+     "lat~ts~",
+     "lat~per~",
+     "lon~ts~",
+     "lon~per~",
+     "$\\mu$",
+     "per^th^",
+     "D~ens~",
+     "lat~EEZ~",
+     "lat~ts~",
+     "lon~EEZ~",
+     "lon~ts~",
+     "TSI",
+     "D~A~",
+     "D~B~",
+     "sd",
+     "t",
+     "th",
+     "ToE",
+     "TSI~$\\mu$~",
+     "TSI~f$\\mu$~",
+     "TSI~$\\omega$~",
+     "_t'_",
+     "tresh",
+     "ToE",
+     "SSR~th~",
+     "SSR~te~",
+     "SSR~tm~",
+     "SSR~tf~",
+     "_s_"
+
+     ),
+   "Meaning" = c(
+     "Habitat suitability",
+     "Grid-cell of 0.5^o^ latitude x 0.5^p^ longitude",
+     "Seawater temperature",
+     "Species temperature profile",
+     "Bathymetry",
+     "Species' minimum depth limit",
+     "Species' maximum depth limit",
+     "Proportion of area of the habitat type",
+     "Habitat type",
+     "Habitat association index",
+     "Salinity",
+     "Salinity association index",
+     "Sea ice extent",
+     "Association to sea ice",
+     "Relative abundance of a given species in each grid-cell _i_",
+     "Intrinsic population growth",
+     "Number of settled larvae",
+     "Net migrated adults from surrounding cells _j_",
+     "Surrounding grid-cell",
+     "Intrinsic rate of population increase",
+     "Carrying capacity of grid-cell _i_",
+     "Fishing mortality",
+     "Average latitude across the high-abundance grid-cells",
+     "Latitude of grid-cells holding a given percentile of projected abundance",
+     "Average longitude across the high-abundance grid-cells",
+     "Longitude of grid-cells holding a given percentile of projected abundance",
+     "Mean",
+     "Percentile of projected transboundary stock abundance",
+     "Distance between the centroid of the EEZs and the centroid of the transboundary stock",
+     "Latitude of the EEZ's centroid",
+     "Latitude of the transboundary stock's centroid",
+     "Longitude of the EEZ's centroid",
+     "Longitude of the transboundary stock's centroid",
+     "Transboundary Stock Index",
+     "Distance between the distribution centroids of a stock and the centroid of the neighboring EEZs A",
+     "Distance between the distribution centroids of a stock and the centroid of the neighboring EEZs B",
+     "Standard deviation",
+     "Time step from 1951 to 2006",
+     "Historical time period (1951 to 2005)",
+     "Time of emergence",
+     "Average TSI from 1951 to 2005",
+     "Average TSI from 2006 to 2100",
+     "Standard deviation of TSI from 1951 to 2005",
+     "Years from 2006 to 2100",
+     "Threshold to determine the time of emergence",
+     "Year when TSI~f$\\mu$~ overpasses the _tresh_",
+     "Stocks share ratio over the historical time period (1951 to 2005)",
+     "Stocks share ratio at the early time period (2020 to 2040)",
+     "Stocks share ratio at the late time period (2050 to 2060)",
+     "Stocks share ratio at the early or late time period",
+     "Stock whose share ratio exceeded the threat point of an EEZ"
+   )
+) %>% 
+  kable()
+```
+
 \clearpage

### 2021-12-01 Scripts/Supplements.Rmd
re-rganized figures

@@
 
-```{r toeFigureS1, eval = T,  echo = F, fig.width = 10, fig.height= 8, fig.pos="H", fig.cap ="World Exclusive Economic Zones used in this study as divided by the Sea Around Us and their estimated centroids (black points)."}
+```{r toeFigureS1, eval = T,  echo = F, fig.width = 10, fig.height= 8, fig.pos="H", fig.cap ="World Exclusive Economic Zones (EEZs) used in this study as divided by the Sea Around Us and their estimated centroids (black points)."}
 
@@
 
-```{r toeFigureSA, eval =F,  echo = F, fig.width=4, fig.height=4, fig.cap ="Average distribution of transboundary stock’s year of change in their shared distribution.
-Left panel is the average year across stocks and right panel the standard deviation.
-Yellow represents a two standard deviation distance from the historical average and purple represents a one standard deviation distance from the historical average."}
-
-img <- readPNG("../Figures/FigSA.png")
-grid.raster(img)
-
-```
-
-\clearpage
-
-```{r toeFigureS2, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine the stock’s centroid.
+```{r toeFigureS2, eval =F,  echo = F, fig.width=4, fig.height=4, fig.cap ="Average year of range shift by Exclusive Economic Zone (EEZ) and transboundary stock using different levels of abundance percentile to determine a stock’s centroid.
 Top Row: results in absolute numbers.
-Bottom row: results expressed as a percentage of the total emerging stocks and EEZs with emerging stock."}
+Bottom row: results expressed as a percentage of the total number of emerging stocks and EEZs with at least one emerging stock."}
 
@@
 A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
-Exclusive Economic Zone polygons display average year of range shifts within them.
-EEZs with no distributional shift between 2006 and 2100 are represented in aqua color.
+Exclusive Economic Zone polygons display the average year of range shifts within them.
+EEZs with no distributional shifts between 2006 and 2100 are represented in pale blue.
 B) Year of shared distribution shifts by UN sub-regions.
-Points color coded by the median year of range shift.
+Points are color coded by the mean year of range shift.
 Horizontal dashed line represents the year 2030 – the Sustainable Development Agenda deadline.
@@
 
-```{r toeFigureS4,  eval = T,  echo = F, fig.width=13, fig.height=13, fig.cap ="Year of change in the shared distribution of 4,119 transboundary stocks and their contribution to fishing revenue.
-A) Contribution of shifting stocks to a country or territory's total fishing revenue from transboundary stocks.
-B) Average year of range shift across an EEZ’s transboundary stocks.
-EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
-This figure is identical to Figure 3A but shows the information in two different maps and color scales; A shows the land polygons in 3A and B shows the EEZ polygons in 3B."}
+```{r toeFigureS4, eval = T,  echo = F, fig.width=9, fig.height=6,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
+Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
+The category “Other” consists of species that have no “habitat preference” classification.
+The number of stocks included in this analysis for each habitat type is noted in parentheses.
+Dashed whiskers represent the 1.5* interquartile range.
+Box represents interquartile range as distance between first and third quartiles.
+Solid line represents the median time of emergence per habitat preference."}
 
@@
 
-```{r toeFigureS5, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Year of change in the shared distribution of transboundary stocks in the Caribbean and South Pacific Islands.
-Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
-Exclusive Economic Zone polygons display average year of range shifts within them.
-Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cold colors represent a late shift/low fishing revenue.
-EEZs with no distributional shift between 2006 and 2100 are represented in pale blue."}
+```{r toeFigureS5,  eval = T,  echo = F, fig.width=13, fig.height=13, fig.cap ="Year of change in the shared distribution of 4,119 transboundary stocks and their contribution to fishing revenue.
+A) Contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
+B) Average year of range shift across an EEZ’s transboundary stocks.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.
+This figure is identical to Fig.
+3A but shows the information in two different maps and color scales; A shows the land polygons in 3A and B shows the EEZ polygons in 3B."}
 
-img <- readPNG("../Figures/FigS11.png")
+img <- readPNG("../Figures/FigS5.png")
 grid.raster(img)
@@
 
-```{r toeFigureS6, eval = T,  echo = F, fig.width=9, fig.height=6,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association.
-Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1).
-The category “Other” consists of species that have no “habitat preference” classification.
-The number of stocks included in this analysis for each habitat type is noted in parentheses.
-Dashed whiskers represent the 1.5* interquartile range.
-Box represents interquartile range as distance between first and third quartiles.
-Solid line represents the median time of emergence per habitat preference."}
+```{r toeFigureS6, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Year of change in the shared distribution of transboundary stocks in the Caribbean and South Pacific Islands using a one s.d.
+(top row) two s.d.
+(bottom row) thresholds.
+Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks.
+Exclusive Economic Zone polygons display the average year of range shifts within them.
+Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks, while cooler colors represent a late shift/low fishing revenue.
+EEZs with no distributional shift between 2006 and 2100 are represented in pale blue."}
 
@@
 
-```{r toeFigureS7, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap ="Changes in stock proportion of neighboring Exclusive Economic Zones by 2050 (2041-2060) relative to 1951-2005.
-Lines represent the average change in transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+```{r toeFigureS7, eval = T,  echo = F, fig.width=14, fig.height=14, fig.cap ="Changes in stock proportion between neighboring Exclusive Economic Zones by 2050 (2041-2060) relative to 1951-2005.
+Lines represent the average change in transboundary stock share ratio (SSR) with arrows going from EEZs with a decrease in stock share (point) to those gaining shares (arrowhead).
 Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point with higher gains identified in warmer colors."}
@@
 ```{r toeFigureS8, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Distribution of changes in transboundary stock proportion per stock and neighboring Exclusive Economic Zones.
-Yellow color represents changes by early, 2030 (2021-2040) and purple by mid 2050 (2041-60) 21st century relative to 1951-2005."}
+Yellow color represents changes by 2030 (2021-2040) and purple by mid-century (2041- 2060) relative to the historical baseline (1951-2005)."}
 
@@
 
-```{r toeFigureS9, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Changes in stock proportion of neighboring Exclusive Economic Zones of Latin America and the Caribeen and Northwestern Europe by 2030 (2021-2040) and by 2060 (2041-2060) relative to 1951-2005.
-Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead).
+```{r toeFigureS9, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Changes in stock proportion of neighboring Exclusive Economic Zones of Latin America and the Caribbean and Northwestern Europe by 2030 (2021-2040) and by 2060 (2041-2060) relative to 1951-2005.
+Lines represent the average change in transboundary stock share ratio (SSR; Methods) with arrows going from EEZ with a decrease in stock shares (point) to those gaining shares (arrowhead).
 Land polygons represent the percentage of stocks that are projected to change their stock share ratio (i.e., over two standard deviations from the historical catch proportion) beyond the identified threat point with higher gains identified in warmer colors."}
@@
 
-```{r toeFigureS10, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Changes in the stock share proportion of top 5 most valuable stocks by 2030 (2021-2040) relative to 1951-2005.
-Lines represent the average change in transboundary stock share ratio (Methods) with arrows going from EEZ with decreasing stock share (point) to those gaining shares (arrowhead)."}
+```{r toeFigureS10, eval = T,  echo = F, fig.width=7, fig.height=5, fig.cap = "Changes in the stock share proportion of the top 5 most valuable stocks by 2030 (2021-2040) relative to the historical baseline (1951-2005).
+Lines represent the average change in transboundary stock share ratio (SSR; Methods) with arrows going from EEZ with decreasing stock shares (point) to those gaining shares (arrowhead)."}
 

### 2021-12-03 Scripts/Supplements.Rmd

@@
 - \usepackage{caption}
-- \captionsetup[figure]{name = Figure S}
-- \captionsetup[table]{name = Table S}
 - \usepackage{lscape}
@@
 
-```{r setup_sup, eval = T, echo=F, warning=F,message=F, results='hide'}
+<!-- - \captionsetup[figure]{name = Figure S} -->
+<!-- - \captionsetup[table]{name = Table S} -->
+
+```{r setup_sup, echo=F, warning=F,message=F, results='hide'}
 library(MyFunctions)
@@
 ```
+
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 
@@
 
-```{r}
-library(tidyverse)
-library(knitr)
+\clearpage
+
+```{r toeTableS2, eval = T, echo = F}
 
@@
      "Hs",
-     "_i_",
-     "_T_",
+     "i",
+     "T",
      "TPP",
-     "B$\\alpha$~i~",
-     "Min~D~",
-     "Max~D~",
-     "Hab~ij~",
-     "_u_",
-     "Hab~Assoc~",
+     "B$\\alpha_i$",
+     "$Min_D$",
+     "$Max_D$",
+     "$Hab_{ij}$",
+     "u",
+     "$Hab_{Assoc}$",
      "Sal",
-     "Sal~Assoc~",
+     "$Sal_{Assoc}$",
      "Ice",
-     "Ice~p~",
-     "A~i~",
-     "_G_",
-     "L~ji~",
-     "Ad~ji~",
-     "_j_",
-     "_r_",
-     "_K_",
-     "_f_",
-     "lat~ts~",
-     "lat~per~",
-     "lon~ts~",
-     "lon~per~",
+     "$Ice_{i}$",
+     "$A_i$",
+     "G",
+     "$L_{ji}$",
+     "$Ad_{ji}$",
+     "j",
+     "r",
+     "K",
+     "f",
+     "$lat_{ts}$",
+     "$lat_{per}$",
+     "$lon_{ts}$",
+     "$lon_{per}$",
      "$\\mu$",
-     "per^th^",
-     "D~ens~",
-     "lat~EEZ~",
-     "lat~ts~",
-     "lon~EEZ~",
-     "lon~ts~",
+     "$per^{th}$",
+     "$D_{ens}$",
+     "$lat_{EEZ}$",
+     "$lon_{EEZ}$",
      "TSI",
-     "D~A~",
-     "D~B~",
+     "$D_A$",
+     "$D_B$",
      "sd",
@@
      "ToE",
-     "TSI~$\\mu$~",
-     "TSI~f$\\mu$~",
-     "TSI~$\\omega$~",
-     "_t'_",
+     "$TSI_\\mu$",
+     "$TSI_{f\\mu}$",
+     "$TSI_\\omega$",
+     "t'",
      "tresh",
-     "ToE",
-     "SSR~th~",
-     "SSR~te~",
-     "SSR~tm~",
-     "SSR~tf~",
-     "_s_"
-
+     # "ToE",
+     "$SSR_{th}$",
+     "$SSR_{te}$",
+     "$SSR_{tm}$",
+     "$SSR_{tf}$",
+     "s"
      ),
-   "Meaning" = c(
+   "Definition" = c(
      "Habitat suitability",
-     "Grid-cell of 0.5^o^ latitude x 0.5^p^ longitude",
+     "Grid-cell of $0.5^o$ latitude x $0.5^o$ longitude",
      "Seawater temperature",
@@
      "Association to sea ice",
-     "Relative abundance of a given species in each grid-cell _i_",
+     "Relative abundance of a given species in each grid-cell i",
      "Intrinsic population growth",
      "Number of settled larvae",
-     "Net migrated adults from surrounding cells _j_",
+     "Net migrated adults from surrounding cells j",
      "Surrounding grid-cell",
      "Intrinsic rate of population increase",
-     "Carrying capacity of grid-cell _i_",
+     "Carrying capacity of grid-cell i",
      "Fishing mortality",
@@
      "Distance between the centroid of the EEZs and the centroid of the transboundary stock",
-     "Latitude of the EEZ's centroid",
      "Latitude of the transboundary stock's centroid",
-     "Longitude of the EEZ's centroid",
      "Longitude of the transboundary stock's centroid",
      "Transboundary Stock Index",
-     "Distance between the distribution centroids of a stock and the centroid of the neighboring EEZs A",
-     "Distance between the distribution centroids of a stock and the centroid of the neighboring EEZs B",
+     "Distance between the centroids of a stock and the neighboring EEZs A",
+     "Distance between the centroids of a stock and the neighboring EEZs B",
      "Standard deviation",
@@
      "Historical time period (1951 to 2005)",
-     "Time of emergence",
+     "Time of emergence, year when $TSI_{f\\mu}$ overpasses the tresh",
      "Average TSI from 1951 to 2005",
@@
      "Threshold to determine the time of emergence",
-     "Year when TSI~f$\\mu$~ overpasses the _tresh_",
      "Stocks share ratio over the historical time period (1951 to 2005)",
@@
 ) %>% 
-  kable()
+  arrange(Acronym) %>%
+  kable(format = "latex",
+        booktabs = T,
+        longtable = T,
+        escape = F,
+        caption = "Term abbreviations used in equations throughout the methods. For each parameter we provide a short definition."
+  ) %>%
+  kable_styling(#full_width = T,
+                # position = "center",
+                latex_options = c("repeat_header")
+                )
 ```
@@
 
-```{r toeTableS2, eval = T, echo = F}
+```{r toeTableS4, eval = T, echo = F}
 

### 2021-12-15 Scripts/Supplements.Rmd
Included abstracts in spanish and portuguese

@@
         longtable = T,
-        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
+        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across species habitat association according to FishBase and SeaLifeBase"
   ) %>% 
@@
 \end{landscape}
+
+# Translated Abstracts
+
+Climate change is shifting the distribution of shared fish stocks between neighboring countries’ Exclusive Economic Zones (EEZ) and the high seas.
+The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.
+Coupling a large ensemble simulation of an Earth system model to a dynamic population model, we show that by 2030, under a high emission climate change scenario, 23% of transboundary stocks will have shifted and 78% of the world’s EEZs will have experienced at least one shifting stock.
+By the end of this century, projections under a high emissions scenario show 81% of EEZs waters with at least one shifting stock and a total of 45% of stocks shifting globally.
+Many countries that are highly dependent on fisheries for livelihood and food security emerge as hotspots for transboundary shifts.
+These hotspots are characterized by early shifts in the distribution of a important number of transboundary stocks.
+Existing fisheries agreements need to be assessed for their capacity to address the social-ecological implications of climate change driven transboundary shifts and strengthened where necessary to limit potential conflict between the parties of interest, while new agreements need to be anticipatory and consider these concerns to be resilient to global change.
+
+## Español
+
+El cambio climático está afectando la distribución de poblaciones transfronterizas de fauna marina que se encuentran en Zonas Económicas Exclusivas (ZEEs) de países vecinos y en el alta mar.
+Los efectos del cambio climático en el manejo pesquero internacional estarán determinados por la escala temporal de dichos desplazamientos.
+Para determinar esa escala temporal, el presente estudio combinó un modelo dinámico poblacional, con una serie de simulaciones de un modelo del sistema terrestre, bajo un escenario de cambio climático de altas emisiones.
+Los resultados siguieren que para 2030, el 23% de las poblaciones transfronterizas se habrán desplazado y en el 78% de las ZEEs del mundo habrán experimentado cambios en la distribución de al menos una población transfronteriza.
+Para fines de este siglo, las proyecciones muestran que el 45% de las poblaciones transfronterizas globales habrán cambiado su distribución y 81% de las ZEEs tendrán al menos una población en movimiento.
+La magnitud de tal desplazamiento se reflejará en un cambio promedio del 59% de la proporción de captura de poblaciones transfronterizas entre ZEEs vecinas para el 2030.
+Muchos países que dependen de la pesca para sustento económico y seguridad alimentaria emergen como zonas críticas de cambios transfronterizos.
+Estas zonas se caracterizan por cambios tempranos en la distribución de un número importante de poblaciones transfronterizas.
+Por lo tanto, los acuerdos pesqueros internacionales deben evaluarse por su capacidad para responder a los impactos socio-ecológicos del desplazamiento de poblaciones transfronterizas debido al cambio climático.
+Dichos acuerdos deberán fortalecerse, cuando sea necesario, para limitar los posibles conflictos entre las partes de interés y evitar amenazar la sustentabilidad del recurso.
+Así mismo, los nuevos acuerdos que vayan a establecerse deberán considerar los posibles cambios en la distribución de poblaciones compartidas (y la incertidumbre asociada) para anticiparse a dichos conflictos y aumentar la resiliencia frente al cambio climático.
+
+## François
+
+## Português
+
+As mudanças climáticas vêm promovendo alterações na distribuição dos estoques de peixes compartilhados por países vizinhos, tanto nas suas Zonas Econômicas Exclusivas (ZEE) como em águas oceânicas internacionais.
+A escala de tempo desse deslocamento transfronteiriço vai determinar como as mudanças climáticas afetarão o manejo pesqueiro internacional.
+Diante disso, no presente trabalho teve o objetivo de analisar essa escala de tempo, combinando um amplo conjunto de simulações de um modelo do sistema terrestre sob um cenário de mudanças climáticas de altas emissões a um modelo de dinâmica populacional.
+Foi observado que, para 2030, 23% dos estoques transfronteiriços terão suas distribuições alteradas e 78% das ZEEs do mundo terão experimentado deslocamentos em pelo menos um estoque transfronteiriço.
+No final deste século, as projeções mostram que 45% dos estoques transfronteiriços do mundo sofrerão alterações, e que 81% das ZEEs apresentarão alterações em pelo menos um estoque.
+A magnitude de tal deslocamento será refletida por uma mudança média de 59% na proporção de capturas de estoques transfronteiriços entre ZEEs vizinhas no ano de 2030.
+Muitos países que são altamente dependentes da pesca para subsistência e segurança alimentar surgem como pontos críticos para mudanças transfronteiriças.
+Estes são caracterizados por mudanças iniciais na distribuição de um número importante de estoques.
+Os acordos internacionais de pesca precisam ser avaliados quanto à sua capacidade de abordar as implicações socioecológicas de deslocamentos transfronteiriços impulsionados pelas mudanças climáticas, e reforçados quando necessário no sentido de minimizar possíveis conflitos entre as partes de interesse.
+Da mesma forma, novos acordos devem considerar possíveis mudanças na distribuição de populações transfronteiriças, a fim de antecipar tais conflitos e construir resiliência em face das mudanças climáticas e das incertezas que as acompanha.

### 2021-12-21 Scripts/Supplements.Rmd
Updated trnslated abstracts

@@
 
-Climate change is shifting the distribution of shared fish stocks between neighboring countries’ Exclusive Economic Zones (EEZ) and the high seas.
-The timescale of these transboundary shifts determines how climate change will affect international fisheries governance.
-Coupling a large ensemble simulation of an Earth system model to a dynamic population model, we show that by 2030, under a high emission climate change scenario, 23% of transboundary stocks will have shifted and 78% of the world’s EEZs will have experienced at least one shifting stock.
-By the end of this century, projections under a high emissions scenario show 81% of EEZs waters with at least one shifting stock and a total of 45% of stocks shifting globally.
-Many countries that are highly dependent on fisheries for livelihood and food security emerge as hotspots for transboundary shifts.
-These hotspots are characterized by early shifts in the distribution of a important number of transboundary stocks.
-Existing fisheries agreements need to be assessed for their capacity to address the social-ecological implications of climate change driven transboundary shifts and strengthened where necessary to limit potential conflict between the parties of interest, while new agreements need to be anticipatory and consider these concerns to be resilient to global change.
-
 ## Español
@@
 
-## François
+## Français
+
+Le changement climatique altère la distribution des stocks de poissons exploités posant de sérieux problèmes de juridiction et gestion des espèces partagées entre pays voisins, et/ou avec la haute mer.
+C’est en analysant l’échelle de temps de ces migrations transfrontalières que l’impact du changement climatique sur la gouvernance mondiale des pêches peut être évalué.
+Dans cette étude, nous explorons cette échelle de temps à l'aide d’un modèle de dynamique des populations marines exploitées couplé à des simulations dérivées d’un ensemble de modèles globaux océan-atmosphère.
+Les résultats montrent que d’ici 2030, pour le scénario à hautes émissions, 23% des stocks transfrontaliers auront changé de distribution et que 78% des zones économiques exclusives (ZEE) expérimenteront au moins une nouvelle espèce transfrontalière.
+A la fin du siècle, et pour ce même scénario, 81% des ZEE auront au moins une espèce transfrontalière et 45% des stocks transfrontaliers auront changé de distribution.
+La magnitude de tels changements de distribution est ici quantifiée par la variation dans la proportion de capture entre ZEE partageant ce stock transfrontalier.
+D’ici 2030, de tels changements entre ZEE seront de l’ordre de 59% à l'échelle globale, avec de nombreux  pays dont la qualité de vie et la sécurité alimentaire dépendent de la pêche émergeant comme zones à haut risque.
+Ces zones se caractérisent par le déplacement précoce d’un grand nombre de stocks transfrontaliers .
+A la lumière de ces résultats, les traités et accords de pêche internationaux doivent être évalués pour leur capacité à répondre aux implications socio-écologiques du changement climatique et renégocier afin d’éviter tout conflit entre pays voisins.
+En anticipant des changements potentiels de distribution entre stocks transfrontaliers, tout nouvel accord de pêche se voudra plus résilient aux effets du changement climatique
 

### 2022-01-03 Scripts/Initial_Analysis.Rmd
Misc.

@@
       summarise(
-        n_spp = length(unique(taxon_key)),
-        n_stock = n(),
-        min_eez = min(mean_spp, na.rm=T),
-        max_eez = max(mean_spp, na.rm=T),
+        # n_spp = length(unique(taxon_key)),
+        n_stock = n(), # Number of shifting stocks
+        min_eez = min(mean_spp, na.rm=T), #minimum proportion across all shifting stocks
+        max_eez = max(mean_spp, na.rm=T), # maximum proportion across all shifting stocks
         sd_eez = sd(mean_spp, na.rm=T), # for some weird reason it wont work if the order is fliped
-        mean_eez = median(mean_spp,na.rm=T)
-      ) %>% 
-      select(eez_name:max_eez,mean_eez,sd_eez,n_spp,n_stock) %>% 
+        mean_eez = median(mean_spp,na.rm=T) # mean proportion across all shifting stocks
+      ) %>% 
+      select(eez_name:max_eez,mean_eez,sd_eez,n_stock) %>% 
       left_join(n_ssr)
@@
       summarise(
-        n_spp = length(unique(taxon_key)),
+        # n_spp = length(unique(taxon_key)),
         n_stock = n(),
@@
       ) %>% 
-      select(eez_name:max_eez,mean_eez,sd_eez,n_spp,n_stock)
+      select(eez_name:max_eez,mean_eez,sd_eez,n_stock)
 
@@
   proportion_change <- eez %>% 
-    gather("variable","value",n_spp:n_per) %>% 
+    gather("variable","value",n_stock:n_per) %>% 
     spread(time_step,value) %>% 
@@
 n_taxa_data <- Prop_results %>% 
-  filter(variable == "n_spp") %>%
+  filter(variable == "n_stock") %>%
   select(eez_name,early,mid) %>% 
@@
   filter(
-    variable %in% c("mean_eez","n_spp")#,
+    variable %in% c("mean_eez","n_stock")#,
     # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
